### 2014-11-17 tex/codatadp.tex
copied template plus skeleton paper

%envcountsect,
\documentclass[a4paper,oribibl,envcountsame,draft]{llncs}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{arydshln}
\usepackage[scaled=.82]{beramono}
\usepackage{booktabs}
\usepackage{bussproofs}
\usepackage{calc}
\usepackage{cite}
\usepackage{mathptmx}
%\usepackage{txfonts}
%\usepackage{mathrsfs}
%\usepackage{pifont}
%\usepackage{smallcap}
\usepackage{stmaryrd}
\usepackage{subfigure}
\usepackage[usenames]{color}
%\usepackage{graphicx}
%\usepackage{newcent}
\usepackage{textcomp}
%\usepackage{tipa}
\usepackage{units}
\usepackage{url}
\usepackage{version}
\usepackage[all]{xy}

\DeclareFontFamily{OT1}{pzc}{}
\DeclareFontShape{OT1}{pzc}{m}{it}{<-> s * [1.10] pzcmi7t}{}
\DeclareMathAlphabet{\mathscr}{OT1}{pzc}{m}{it}

\DeclareMathAlphabet{\mathcal}{OT1}{pzc}{m}{it}

%%% For final version as well?
\usepackage[
   a4paper,
   pdftex,
   pdftitle={A Decision Procedure for Codatatypes in SMT Solvers},
   pdfauthor={Andrew J. Reynolds and Jasmin Christian Blanchette},
   pdfkeywords={},
   pdfborder={0 0 0},
   draft=false,
   bookmarksnumbered,
   bookmarks,
   bookmarksdepth=2,
   bookmarksopenlevel=2,
   bookmarksopen]{hyperref}

\urlstyle{ttstyle}

\global\def\figurename{Figure}

\DeclareSymbolFont{letters}{OML}{txmi}{m}{it}

%%% REMOVE BEFORE SUBMITTING ABSOLUTELY FINAL VERSION
\makeatletter
\ps@myheadings
\makeatother

\include{defs}

\begin{document}

%\title{Mechanising the Metatheory of %(Many-Sorted)
%  First-Order Logic}
%\subtitle{Infrastructure and Application to Sort Encodings}
%\title{A Mechanised Metatheory of First-Order Logic \\
%  and Its Application to Sort Erasure}
\title{A Decision Procedure for Codatatypes in SMT Solvers}

\author {Andrew J. Reynolds\inst{1} \and Jasmin Christian Blanchette\inst{2,3}}
\authorrunning {A. J. Reynolds \and J. C. Blanchette}
\institute{
\'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL), Switzerland
\and
Inria \& LORIA, Villers-l`es-Nancy, France \\
Max-Planck-Institut f\"ur Informatik, Saarbr\"ucken, Germany
}

\maketitle

\begin{abstract}
Blah.
\end{abstract}

%% The institutions above shouldn't count as footnotes
\setcounter{footnote}{0}

\section{Introduction}
\label{sec:introduction}

Blah.

\section{Conclusion}
\label{sec:conclusion}

\def\ackname{Acknowledgement}
\ourparagraph{\ackname.}
We thank Viktor Kuncak and Tobias Nipkow for making this work possible.
The research was partially supported by the Deutsche
Forschungs\-gemein\-schaft (DFG) project
\relax{Hardening the Hammer} (grant Ni\,491\slash 14-1).
% The authors are listed in alphabetical order.
% regardless of individual contributions or seniority.

\bibliographystyle{splncs03}
\bibliography{bib}{}

\end{document}

### 2014-11-18 tex/codatadp.tex
more ideas / examples

@@
    pdftex,
-   pdftitle={A Decision Procedure for Codatatypes in SMT Solvers},
+   pdftitle={A Decision Procedure for (Co)datatypes in SMT Solvers},
    pdfauthor={Andrew J. Reynolds and Jasmin Christian Blanchette},
@@
 %  and Its Application to Sort Erasure}
-\title{A Decision Procedure for Codatatypes in SMT Solvers}
+\title{A Decision Procedure for (Co)datatypes in SMT Solvers}
 
@@
 \and
-Inria \& LORIA, Villers-l`es-Nancy, France \\
+Inria \& LORIA, Villers-l\`es-Nancy, France \\
 Max-Planck-Institut f\"ur Informatik, Saarbr\"ucken, Germany
@@
 
+  * briefly: datatypes and why they are useful
+
+  * less briefly: codatatypes and why they are useful
+    * in Agda, Coq, Matita -- recently also in Isabelle/HOL
+    * but not in SMT-LIB (2.5?)
+
+  * single decision procedure for datatypes and codatatypes
+  * datatypes are implementation following the lines of Barrett et al.
+  * codatatypes were added later, motivated by CVC4 as Isabelle backend
+
+  * setting (Section~\ref{sec:theory-of-co-datatypes}):
+    * universal formulas
+    * many-sorted logic
+    * mutually (co)recursive types with constructors, selectors, and
+      discriminators
+
+  * codatatypes: from a theoretical and implementational point of view, like
+      datatypes but:
+    * codatatypes are never empty (e.g. finite streams are rejected)
+    * no acyclicity (e.g. xs = lcons(0, xs) is satisfiable)
+    * instead: bisimilarity
+
+  * motivation:
+    * acyclicity and bisimilarity cannot be axiomatized, so they really belong in
+      a decision procedure
+    * other properties can be done more efficiently by a decision procedure
+
+  * decision procedure is described abstractly as a calculus
+    (Section~\ref{sec:the-calculus})
+    * includes rewriting
+    * and inferences
+
+  * integrated with Nelson-Oppen
+
+  * implemented in CVC4 as rewriting and a theory solver
+    (\ref{sec:the-theory-solver})
+    * precise strategy for applying inferences
+    * about 2000 lines of code, among which 1600 are shared between datatypes
+      and codatatypes
+
+  * useful both for proving and for model finding
+    * in particular, acyclicity and bisimilarity are necessary for some proofs
+      (and cannot be axiomatized finitely)
+    * and for model finding, without them we quickly get spurious models
+
+%\ref{sec:examples}
+
+  * benchmarking is often an issue -- esp. codatatypes
+  * the decision procedure is evaluated on two sets of benchmarks
+    (Section~\ref{sec:experimental-results}):
+    * first set: hand-crafted examples
+    * second set: generated from Isabelle theories using the Sledgehammer tool
+
+\paragraph{Related Work.}
+
+    * point to Barrett et al. for SMT datatypes
+    * additional ones since then (e.g. strings?)
+    * anything about codatatypes?
+      * proof assistants like Agda, Coq, etc. have them
+      * Dafny, CoALP
+      * also a lot of theoretical research, some of which is loosely connected,
+        e.g. decision procedure for corecursive functions (Henning in Nijmegen)
+
+\section{Theory of (Co)datatypes}
+\label{sec:theory-of-co-datatypes}
+
+  * unlike datatypes, codatatypes are never empty
+
+\section{The Calculus}
+\label{sec:the-calculus}
+
+  * SMT background
+
+  * Rewriter
+  * Actual calculus
+
+  * Correctness
+
+  * no "internal search"
+
+\section{The Theory Solver}
+\label{sec:the-theory-solver}
+
+  * Strategies
+  * Implementation
+    * what is an SMT theory solver
+    * cooperation with Nelson-Oppen / integration into SMT solver
+    * "mu" stuff
+
+  * Enumerator?
+    * for model generation
+
+%\section{Examples}
+%\label{sec:examples}
+
+\section{Experimental Results}
+\label{sec:experimental-results}
+
+  * Constructed examples
+  * Isabelle benchmarks
+
 \section{Conclusion}
@@
 
-\def\ackname{Acknowledgement}
-\ourparagraph{\ackname.}
-We thank Viktor Kuncak and Tobias Nipkow for making this work possible.
+Future work:
+
+  * more evaluation
+  * recursive and corecursive functions
+    (and rewriting)
+
+\def\ackname{Acknowledgment}
+\paragraph{\ackname.}
+
+Clark Barrett and Cesare Tinelli.
+
+Bosses: Viktor Kuncak, Tobias Nipkow, and Stephan Merz.
+
 The research was partially supported by the Deutsche
@@
 \relax{Hardening the Hammer} (grant Ni\,491\slash 14-1).
-% The authors are listed in alphabetical order.
-% regardless of individual contributions or seniority.
 

### 2014-11-18 tex/codatadp.tex
more

@@
   * briefly: datatypes and why they are useful
+    (terminology: freely-generated, inductive, algebraic, ..., sometimes
+    with different meanings; we'll clarify below)
+    * Hoare's recursive data structures
+    * comparatively easy to reason about and automate
+    * common in practice, esp. in computer science applications
 
@@
   * datatypes are implementation following the lines of Barrett et al.
-  * codatatypes were added later, motivated by CVC4 as Isabelle backend
+  * codatatypes were added later, motivated by the use of SMT solvers as
+    backends to proof assistants (more specifically, CVC4 to Isabelle/HOL)
 
@@
     * instead: bisimilarity
+    * (also: ``enumeration'', e.g. of streams)
 
@@
     * point to Barrett et al. for SMT datatypes
+      * about their own work, they say: "our focus is on generality and
+        efficiency rather than immediacy of implementation"
+      * (deal more directly with finite sorts than Barrett et al., Section 6.1)
     * additional ones since then (e.g. strings?)
@@
 
+    * ODDITY: Oppen 1980: single-constructor, recursive -- infinite values?
+
 \section{Theory of (Co)datatypes}
@@
   * unlike datatypes, codatatypes are never empty
+
+  * how to deal with "wrong" selectors, e.g.
+      hd(nil1) = hd(nil2)?
+  * ODDITY: Barrett et al. overspecify things (not good w.r.t. SMT-LIB 2, or
+    other solvers, e.g., Z3)
 

### 2014-11-19 tex/codatadp.tex
more

@@
     * no acyclicity (e.g. xs = lcons(0, xs) is satisfiable)
-    * instead: bisimilarity
+    * instead: bisimilarity check
     * (also: ``enumeration'', e.g. of streams)
@@
 
-  * unlike datatypes, codatatypes are never empty
+  * (co)datatype specifications:
+    * N mutually recursive types (datatypes or codatatypes -- no mixture)
+    * list of n constructors (n >= 1)
+    * each constructor has m arguments (m >= 0)
+    * each argument has a sort, which may be
+
+  * don't consider mixtures (co)datatypes (cf. Agda), nor polymorphic types, nor
+    is it possible to do any nesting; e.g. this is outside our scope:
+
+      datatype tree = Node Int (Array tree)
+
+  * initial algebra
+    * ``no junk, no confusion''
+    * initial algebra ``exists'' only if well-founded
+      otherwise it would be empty, which is disallowed by most formulations of
+      first-order logic (including CVC4 and other SMT solvers)
+      and higher-order logic (including Isabelle/HOL)
+
+  * final coalgebra
+    * this time, no condition: always exists
+    e.g. the specification
+
+      codatatype x = X x
+
+    gives rise to a singleton type {X (X (X ...))}.
+
+  * discriminators/selectors for initial algebras
+  * constructors for final coalgebras
 
@@
         hd(nil1) = hd(nil2)?
+      * leave them unspecified; hence hd(nil1) = hd(nil2) in some models,
+        and not in other models
+
   * ODDITY: Barrett et al. overspecify things (not good w.r.t. SMT-LIB 2, or
     other solvers, e.g., Z3)
+
+  * axiomatically
+    * injectivity
+      distinctness
+      exhaustiveness
+    * datatypes:
+      * induction axiom (schema)
+        * consequence: no infinite objects (in particular, no cyclic objects)
+
+      * when looking at the universal ground (?) theory:
+        enough to consider acyclicity -- no way to specify infinite objects
+        otherwise
+
+    * codatatypes:
+      * coinduction axiom (schema)
+        * consequence: infinite objects are allowed, but
+          two objects yielding the same observations must be equal
+
+      * again: when looking at the universal ground (?) theory, enough to
+        consider bisimilarity/bisimulation
 

### 2014-11-26 tex/codatadp.tex
Initial rough draft of calculus (incomplete).

@@
 %\usepackage{smallcap}
+\usepackage{mathpartir} 
 \usepackage{stmaryrd}
@@
 \DeclareMathAlphabet{\mathcal}{OT1}{pzc}{m}{it}
+
+\renewcommand{\vec}[1]{\boldsymbol{#1}}
+\newcommand{\Ec}{\mathsf{E}}
+\newcommand{\tEc}{\mathcal{T}(\Ec)}
+\newcommand{\rn}[1]{\textsf{\small #1}}
+\newcommand{\cvc}{\textsc{cvc}{\small 4}\xspace}
+\newcommand{\teq}{\approx}
+\newcommand{\tneq}{\not\teq}
+\newcommand{\rem}[1]{\textcolor{red}{[#1]}}
+\newcommand{\tester}[1]{is\text{-} #1}
+\newcommand{\ror}{\quad \parallel \quad}
+\newcommand{\tpath}[2]{\mathcal{P}_{ #2 \rightarrow \_ }( #1 )}
+\newcommand{\ttpath}[3]{\mathcal{P}_{ #2 \rightarrow #3 }( #1 )}
+\newcommand{\ec}[1]{ [ #1 ] }
+
+% change?
+\newcommand{\thD}{T_{D}}
 
@@
 
-  * no "internal search"
+\rem{Discuss signature $\Sigma$, theory $\thD$}
+This section presents a calculus for determining the satisfiability of conjunctions of $\thD$-constraints.
+
+Our calculus consists of derivation rules that operate on a set $\Ec$ of equalities and disequalities between $\Sigma$-terms.
+We will write $\tEc$ to denote the set of $\Sigma$-terms occurring in $\Ec$.
+We will commonly denote tuples of terms $( t_1, \ldots, t_n )$ in bold font, as $\vec t$.
+We assume that all $\Sigma$-terms in $\Ec$ are \emph{normalized}, meaning that all subterms of the form $s^i_j( C_j( \vec t ) )$
+are replaced by $t_i$, and moreover assume that all additional constraints added to $\Ec$ are normalized in this manner.
+
+\begin{figure}[t]
+\centering
+\begin{tabular}{c}
+\rn{Refl}
+\(
+\inferrule{
+  t \in \tEc
+}{
+  \Ec := \Ec, t \teq t
+}
+\)
+\qquad
+\rn{Symm}
+\(
+\inferrule{
+ t_1 \teq t_2 \in \Ec
+}{
+ \Ec := \Ec, t_2 \teq t_1
+}
+\)
+\qquad
+\rn{Conflict}
+\(
+\inferrule{
+  t \tneq t \in \Ec
+}{
+  \bot
+}
+\)
+\\[3.7ex]
+\rn{Trans}
+\(
+\inferrule{
+  t_1 \teq t_2, t_2 \teq t_3 \in \Ec
+}{
+  \Ec := \Ec, t_1 \teq t_3
+}
+\)
+\qquad
+\rn{Cong} 
+\(
+\inferrule{
+  \vec t \teq \vec u \in \Ec \quad f( \vec t ), f( \vec u ) \in \tEc
+}{
+  \Ec := \Ec, f( \vec t ) \teq f( \vec u )
+}
+\)
+\\[3.7ex]
+\rn{Unify$_1$} 
+\(
+\inferrule{
+  C_1( \vec t ) \teq C_1( \vec u ) \in \Ec
+}{
+  \Ec := \Ec, \vec t \teq \vec u
+}
+\)
+\qquad
+\rn{Unify$_2$} 
+\(
+\inferrule{
+  C_1( \vec t ) \teq C_2( \vec u ) \in \Ec
+}{
+  \bot
+}
+\)
+\end{tabular}
+\caption{Rules for bidirectional congruence closure.
+}
+\label{fig:cc-rules}
+\end{figure}
+
+Following the conventions from~\cite{}, 
+the derivation rules of our calculus are given in Figures~\ref{fig:cc-rules},~\ref{fig:ab-rules}, and~\ref{fig:split-rule} in \emph{guarded assignment form},
+where a rule can be applied to $\Ec$ if it meets all of the specified preconditions for $\Ec$.
+The conclusion of a rule either describes equalities to be added to $\Ec$ (in which case we will call an application of it to be \emph{non-terminal}),
+or is $\bot$ (in which case we will call an application of it to be \emph{terminal}).
+A rule may have multiple conclusions separated by $\parallel$, which denotes a non-deterministic branching.
+An application of a rule is \emph{redundant} if it is non-terminal and at least one branch in its conclusion does not add a new equality to $\Ec$.
+
+Figure~\ref{fig:cc-rules} gives the rules for constructing the (bidirectional) congruence closure of $\Ec$. \rem{more?}
+
+Assume that $\Ec$ is saturated with respect to the rules in Figure~\ref{fig:cc-rules}, that is,
+no non-redudant application of a rule can be applied to $\Ec$.
+In this case, it is easy to see that $\Ec$ induces an equivalence relation over $\tEc$ such that two terms $t_1$ and $t_2$ are equivalent if and only if $t_1 \teq t_2 \in \Ec$.
+Thus, we will in the following consider $\Ec$ as a set of equivalence classes of terms. 
+For term $t \in \tEc$, we will write $\ec{t}$ to denote the equivalence class in $\Ec$ that containing $t$.
+
+To explain the rules in Figure~\ref{fig:ab-rules}, we introduce the following notions.
+First, an edge $\ec{t_1} \rightarrow^{C}_n \ec{t_2}$ is a transition between two equivalence classes $\ec{t_1}$ and $\ec{t_2}$ (which we will call its source and destination respectively),
+labelled by a constructor $C$ and integer $n$.
+We say that $\Ec$ \emph{induces edge} $\ec{t_1} \rightarrow^{C}_n \ec{t_2}$ if and only if $C( \vec u ) \in \ec{t_1}$ for some $\vec u$ and $\ec{t_2} = \ec{u_n}$.
+A \emph{path} $p$ is a non-empty ordered list of edges $e_1, \ldots e_n$ such that the destination of $e_i$ is the source of $e_{i+1}$ for $i = 1, \ldots (n-1)$,
+and the sources of $e_1, \ldots, e_n$ are distinct.
+We say that $\Ec$ \emph{induces path} $p$ if it induces each of its edges.
+Notice that since $\Ec$ contains a finite set of equivalence classes, all paths induced by $\Ec$ are finite.
+The set $\ttpath{\Ec}{\ec{t_1}}{\ec{t_2}}$ is the smallest set containing all induced paths starting with source $\ec{t_1}$ and destination $\ec{t_2}$.
+The set $\tpath{\Ec}{\ec{t_1}}$ is the union of $\ttpath{\Ec}{\ec{t_1}}{\ec{t_2}}$ for all $t_2 \in \tEc$.
+
+\begin{figure}[t]
+\centering
+\begin{tabular}{c}
+\rn{Acyclic}
+\(
+\inferrule{
+  t : \tau \text{ inductive } 
+  \quad
+  p \in \ttpath{\Ec}{\ec{t}}{\ec{t}}
+}{
+  \bot
+}
+\)
+\\[3.7ex]
+\rn{Bisimilar}
+\(
+\inferrule{
+ t_1, t_2 : \tau \text{ co-inductive }
+  \quad
+ \exists \sigma. p \sigma \in \tpath{\Ec}{\ec{t_1}} \text{ for each } p \in \tpath{\Ec}{\ec{t_2}}
+}{
+ \Ec := \Ec, t_1 \teq t_2
+}
+\)
+\end{tabular}
+\caption{Rules for acyclicity and bisimilarity.
+}
+\label{fig:ab-rules}
+\end{figure}
+
+\rem{Change notation for inductive/co-inductive.}
+
+\begin{figure}[t]
+\centering
+\begin{tabular}{c}
+\rn{Split} 
+\(
+\inferrule{
+  s^i_j( t ) \in \tEc, \text{ or } t : \tau \text{ finite }
+}{
+  \Ec := \Ec, \tester{C_1}( t ) \ror \ldots \ror \Ec := \Ec, \tester{C_n}( t ) 
+}
+\)
+\end{tabular}
+\caption{Rule for splitting.  We write $\tester{C_i}( t )$ as shorthand for the equality $t \teq C_i( s^1_i( t ), \ldots, s^n_i( t ) )$.
+}
+\label{fig:split-rule}
+\end{figure}
+
+A \emph{derivation tree} is a tree whose nodes are sets of equalities, where non-root nodes are obtained by a non-redundant application of a derivation rule to its parent node.
+A derivation tree is \emph{closed} if all of its leaf nodes are $\bot$.
+A node is \emph{saturated} if no non-redundant application of a rule can be applied to it.
+
+\subsection{Correctness}
+
+\begin{lemma}[Refutation Soundness]
+If there exists a closed derivation tree with root node $\Ec$, then $\Ec$ is unsatisfiable in $\thD$.
+\end{lemma}
+
+\begin{lemma}[Solution Soundness]
+If there exists a derivation tree containing a saturated node with root node $\Ec$, then $\Ec$ is satisfiable in $\thD$.
+\end{lemma}
+
+\begin{lemma}[Termination]
+All derivation trees are finite.
+\end{lemma}
+
+\rem{Thus, sound and complete.}
 
@@
     * "mu" stuff
-
+    * no "internal search"
   * Enumerator?

### 2014-11-27 tex/codatadp.tex
Minor updates to text.

@@
 \inferrule{
-  t \tneq t \in \Ec
+  t \teq u, t \tneq u \in \Ec
 }{
@@
 \end{tabular}
-\caption{Rules for bidirectional congruence closure.
+\caption{Rules for bidirectional closure.
 }
@@
 
+We present the calculus in three phases, shown in Figures~\ref{fig:cc-rules},~\ref{fig:ab-rules}, and~\ref{fig:split-rule}.
 Following the conventions from~\cite{}, 
-the derivation rules of our calculus are given in Figures~\ref{fig:cc-rules},~\ref{fig:ab-rules}, and~\ref{fig:split-rule} in \emph{guarded assignment form},
+the derivation rules of our calculus are given in \emph{guarded assignment form},
 where a rule can be applied to $\Ec$ if it meets all of the specified preconditions for $\Ec$.
@@
 
-Figure~\ref{fig:cc-rules} gives the rules for constructing the (bidirectional) congruence closure of $\Ec$. \rem{more?}
+First, Figure~\ref{fig:cc-rules} gives the rules for constructing the bidirectional closure of $\Ec$.
+The rules \rn{Refl}, \rn{Symm}, \rn{Trans}, and \rn{Cong} together compute the (upwards) congruence closure,
+after which a conflict may be recognized by \rn{Conflict} when an inconsistency is found.
+The rules \rn{Unify$_1$} and \rn{Unify$_2$} together compute (downwards) unification,
+where additional equalities are inferred based on the injectivity of constructors (by \rn{Unify$_1$}),
+and failures to unify equivalent terms are recognized (by \rn{Unify$_2$}).
 
-Assume that $\Ec$ is saturated with respect to the rules in Figure~\ref{fig:cc-rules}, that is,
+Second, when applying the rules in Figure~\ref{fig:ab-rules}, 
+we assume that $\Ec$ is saturated with respect to the rules in Figure~\ref{fig:cc-rules}, that is,
 no non-redudant application of a rule can be applied to $\Ec$.
@@
 To explain the rules in Figure~\ref{fig:ab-rules}, we introduce the following notions.
-First, an edge $\ec{t_1} \rightarrow^{C}_n \ec{t_2}$ is a transition between two equivalence classes $\ec{t_1}$ and $\ec{t_2}$ (which we will call its source and destination respectively),
+First, an edge $\ec{t_1} \rightarrow^{C}_n \ec{t_2}$ is a transition between two equivalence classes $\ec{t_1}$ and $\ec{t_2}$ 
+(which we will call its \emph{source} and \emph{destination} respectively),
 labelled by a constructor $C$ and integer $n$.
-We say that $\Ec$ \emph{induces edge} $\ec{t_1} \rightarrow^{C}_n \ec{t_2}$ if and only if $C( \vec u ) \in \ec{t_1}$ for some $\vec u$ and $\ec{t_2} = \ec{u_n}$.
+We say that edge $\ec{t_1} \rightarrow^{C}_n \ec{t_2}$ is \emph{induced by $\Ec$} if and only if $C( \vec u ) \in \ec{t_1}$ for some $\vec u$ and $\ec{t_2} = \ec{u_n}$.
 A \emph{path} $p$ is a non-empty ordered list of edges $e_1, \ldots e_n$ such that the destination of $e_i$ is the source of $e_{i+1}$ for $i = 1, \ldots (n-1)$,
 and the sources of $e_1, \ldots, e_n$ are distinct.
-We say that $\Ec$ \emph{induces path} $p$ if it induces each of its edges.
-Notice that since $\Ec$ contains a finite set of equivalence classes, all paths induced by $\Ec$ are finite.
-The set $\ttpath{\Ec}{\ec{t_1}}{\ec{t_2}}$ is the smallest set containing all induced paths starting with source $\ec{t_1}$ and destination $\ec{t_2}$.
+We say that path $p$ is \emph{induced by $\Ec$} if and only if each of its edges are induced by $\Ec$.
+Notice that since $\Ec$ contains a finite set of equivalence classes, all paths induced by $\Ec$ have finite length.
+We call the source of a path is the source of its first edge, and the destination of a path is the destination of its last edge.
+The set $\ttpath{\Ec}{\ec{t_1}}{\ec{t_2}}$ is the smallest set containing all paths with source $\ec{t_1}$ and destination $\ec{t_2}$ that are induced by $\Ec$.
 The set $\tpath{\Ec}{\ec{t_1}}$ is the union of $\ttpath{\Ec}{\ec{t_1}}{\ec{t_2}}$ for all $t_2 \in \tEc$.

### 2014-12-01 tex/codatadp.tex
minor
more theory

@@
 
-%\title{Mechanising the Metatheory of %(Many-Sorted)
-%  First-Order Logic}
-%\subtitle{Infrastructure and Application to Sort Encodings}
-%\title{A Mechanised Metatheory of First-Order Logic \\
-%  and Its Application to Sort Erasure}
 \title{A Decision Procedure for (Co)datatypes in SMT Solvers}
@@
 
-  * (co)datatype specifications:
-    * N mutually recursive types (datatypes or codatatypes -- no mixture)
+\newcommand\keyw[1]{\textbf{#1}}
+
+For the purposes of this paper, a (co)datatype specification consists of
+$k$~mutually recursive types which are either all datatypes or all codatatypes.
+Each datatype $\delta$ is equipped by $m \ge 1$ constructors, and each
+constructor for $\delta$ takes $n \ge 0$ arguments, each associated with a
+selector, and returns a $\delta$ value. The argument types must be either
+already existing types or be one of the newly introduced (co)datatypes. The
+introduced names for the (co)datatypes, the constructors, and the selectors must
+be distinct with already introduced ones and among themselves. Schematically:
+%
+\[
+\begin{aligned}[t]
+(\keyw{co})\keyw{datatype}\;\,
+  \delta_1 & {}=\, C_{11}\;(s_{111}{:}\; \sigma_{111}) \,\ldots\, (s_{11n_{11}}{:}\; \sigma_{11n_{11}}) & \\
+           & {}\;\phantom{=\,}\;{} \smash{\,\,\vdots} \\ % {}\;\phantom{=}\llap{\ensuremath{\mid}}\; \cdots \\
+           & {}\;\phantom{=\,}\llap{\ensuremath{\mid}}\; C_{1m}\;(s_{1m1}{:}\; \sigma_{1m1}) \,\ldots\, (s_{1mn_{1m}}{:}\; \sigma_{1mn_{1m}}) \\[-.5\jot]
+   \smash{\vdots\,\,\,} \\[-.5\jot]
+  \keyw{and}\; \,\delta_k & =\, \ldots
+\end{aligned}
+\]
+
+(datatypes or codatatypes -- no mixture)
     * list of n constructors (n >= 1)

### 2014-12-01 tex/codatadp.tex
more

@@
 
+  * consider a natural fragment---datatypes as supported in modern SMT solvers
+    and the SMT-LIB 2 standard, and codatatypes as their duals
+    * mutual recursion, but no polymorphism, nested recursion
+
   * decision procedure is described abstractly as a calculus
@@
 
+Polymorphic types, nested (co)recursion, and datatype--codatatype mixtures fall
+
+\paragraph{Notations.}
+Blah
+
 \paragraph{Related Work.}
@@
 \newcommand\keyw[1]{\textbf{#1}}
+\newcommand\const[1]{\textsf{#1}}
+\newcommand\ty[1]{\mathit{#1}}
 
 For the purposes of this paper, a (co)datatype specification consists of
-$k$~mutually recursive types which are either all datatypes or all codatatypes.
+$l$~mutually recursive types which are either all datatypes or all codatatypes.
 Each datatype $\delta$ is equipped by $m \ge 1$ constructors, and each
@@
            & {}\;\phantom{=\,}\;{} \smash{\,\,\vdots} \\ % {}\;\phantom{=}\llap{\ensuremath{\mid}}\; \cdots \\
-           & {}\;\phantom{=\,}\llap{\ensuremath{\mid}}\; C_{1m}\;(s_{1m1}{:}\; \sigma_{1m1}) \,\ldots\, (s_{1mn_{1m}}{:}\; \sigma_{1mn_{1m}}) \\[-.5\jot]
-   \smash{\vdots\,\,\,} \\[-.5\jot]
-  \keyw{and}\; \,\delta_k & =\, \ldots
+           & {}\;\phantom{=\,}\llap{\ensuremath{\mid}}\; C_{1m}\;(s_{1m1}{:}\; \sigma_{1m1}) \,\ldots\, (s_{1mn_{1m}}{:}\; \sigma_{1mn_{1m}}) \\[-1\jot]
+   \smash{\vdots\,\,\,} \\[-1\jot]
+  \keyw{and}\; \,\delta_l & =\, \ldots
 \end{aligned}
 \]
+%
+with
+$\const{C}_{ij} : \sigma_{ij1}\times\cdots\times\sigma_{ijk_{ij}} \to \delta_i$
+and $\const{s}_{ijk} : \delta_i \to \sigma_{ijk}$.
 
-(datatypes or codatatypes -- no mixture)
-    * list of n constructors (n >= 1)
-    * each constructor has m arguments (m >= 0)
-    * each argument has a sort, which may be
-
-  * don't consider mixtures (co)datatypes (cf. Agda), nor polymorphic types, nor
-    is it possible to do any nesting; e.g. this is outside our scope:
-
-      datatype tree = Node Int (Array tree)
+Polymorphic types, nested (co)recursion, and datatype--codatatype mixtures fall
+outside this fragment. For example, the specification
+\[\begin{aligned}[t]
+      \keyw{datatype}\;\, \ty{tree} \,=\, \const{Node}\;\ty{int}\;\ty{forest} \\[-\jot]
+      \keyw{and}\;\, \ty{forest} \,=\, \const{FNil} \mid \const{FCons}\; \ty{tree}\; \ty{forest}
+\end{aligned}
+\]
+is supported, whereas
+\[\begin{aligned}[t]
+      \keyw{codatatype}\; \,\ty{llist}(\alpha) & \,=\, \const{LNil} \mid \const{LCons}\; \alpha\; \ty{llist}(\alpha) \\
+      \keyw{datatype}\;\, \ty{tree} & \,=\, \const{Node}\; \ty{int}\; \ty{llist}(\ty{tree})
+\end{aligned}
+\]
+is beyond the fragment considered here.
+% (although it makes sense and is supported by Isabelle \cite{xxx}).
 

### 2014-12-01 tex/codatadp.tex
minor

@@
 \and
-Inria \& LORIA, Villers-l\`es-Nancy, France \\
+Inria Nancy \& LORIA, Villers-l\`es-Nancy, France
+\and
 Max-Planck-Institut f\"ur Informatik, Saarbr\"ucken, Germany
@@
 (\keyw{co})\keyw{datatype}\;\,
-  \delta_1 & {}=\, C_{11}\;(s_{111}{:}\; \sigma_{111}) \,\ldots\, (s_{11n_{11}}{:}\; \sigma_{11n_{11}}) & \\
+  \delta_1 & {}=\, \const{C}_{11}(\bigl[\const{s}_{111}{:}\bigr]\; \sigma_{111}, \,\ldots,\, \bigl[\const{s}_{11n_{11}}{:}\bigr]\; \sigma_{11n_{11}}) & \\
            & {}\;\phantom{=\,}\;{} \smash{\,\,\vdots} \\ % {}\;\phantom{=}\llap{\ensuremath{\mid}}\; \cdots \\
-           & {}\;\phantom{=\,}\llap{\ensuremath{\mid}}\; C_{1m}\;(s_{1m1}{:}\; \sigma_{1m1}) \,\ldots\, (s_{1mn_{1m}}{:}\; \sigma_{1mn_{1m}}) \\[-1\jot]
+           & {}\;\phantom{=\,}\llap{\ensuremath{\mid}}\; \const{C}_{1m}(\bigl[\const{s}_{1m1}{:}\bigr]\; \sigma_{1m1}, \,\ldots,\, \big[\const{s}_{1mn_{1m}}{:}\big]\; \sigma_{1mn_{1m}}) \\[-1\jot]
    \smash{\vdots\,\,\,} \\[-1\jot]
@@
 $\const{C}_{ij} : \sigma_{ij1}\times\cdots\times\sigma_{ijk_{ij}} \to \delta_i$
-and $\const{s}_{ijk} : \delta_i \to \sigma_{ijk}$.
+and $\const{s}_{ijk} : \delta_i \to \sigma_{ijk}$. The selector names are often
+omitted in our examples, in which case default names are generated.
 
@@
 \[\begin{aligned}[t]
-      \keyw{datatype}\;\, \ty{tree} \,=\, \const{Node}\;\ty{int}\;\ty{forest} \\[-\jot]
-      \keyw{and}\;\, \ty{forest} \,=\, \const{FNil} \mid \const{FCons}\; \ty{tree}\; \ty{forest}
+      \keyw{datatype}\;\, \ty{tree} & \,=\, \const{Node}(\ty{int},\, \ty{forest}) \\[-\jot]
+      \keyw{and}\;\, \ty{forest} & \,=\, \const{FNil} \mid \const{FCons}(\ty{tree},\, \ty{forest})
 \end{aligned}
@@
 \[\begin{aligned}[t]
-      \keyw{codatatype}\; \,\ty{llist}(\alpha) & \,=\, \const{LNil} \mid \const{LCons}\; \alpha\; \ty{llist}(\alpha) \\
-      \keyw{datatype}\;\, \ty{tree} & \,=\, \const{Node}\; \ty{int}\; \ty{llist}(\ty{tree})
+      \keyw{codatatype}\; \,\ty{llist}(\alpha) & \,=\, \const{LNil} \mid \const{LCons}(\alpha,\, \ty{llist}(\alpha)) \\
+      \keyw{datatype}\;\, \ty{tree} & \,=\, \const{Node}(\ty{int},\, \ty{llist}(\ty{tree}))
 \end{aligned}
@@
   * constructors for final coalgebras
+
+  * Introduce the discriminators already here.
 
@@
 We will commonly denote tuples of terms $( t_1, \ldots, t_n )$ in bold font, as $\vec t$.
-We assume that all $\Sigma$-terms in $\Ec$ are \emph{normalized}, meaning that all subterms of the form $s^i_j( C_j( \vec t ) )$
+We assume that all $\Sigma$-terms in $\Ec$ are \emph{normalized}, meaning that all subterms of the form $s^i_j( \const{C}_j( \vec t ) )$
 are replaced by $t_i$, and moreover assume that all additional constraints added to $\Ec$ are normalized in this manner.
@@
 \inferrule{
-  C_1( \vec t ) \teq C_1( \vec u ) \in \Ec
+  \const{C}_1( \vec t ) \teq \const{C}_1( \vec u ) \in \Ec
 }{
@@
 \inferrule{
-  C_1( \vec t ) \teq C_2( \vec u ) \in \Ec
+  \const{C}_1( \vec t ) \teq \const{C}_2( \vec u ) \in \Ec
 }{
@@
 (which we will call its \emph{source} and \emph{destination} respectively),
-labelled by a constructor $C$ and integer $n$.
+labeled by a constructor $C$ and integer $n$.
 We say that edge $\ec{t_1} \rightarrow^{C}_n \ec{t_2}$ is \emph{induced by $\Ec$} if and only if $C( \vec u ) \in \ec{t_1}$ for some $\vec u$ and $\ec{t_2} = \ec{u_n}$.
@@
 }{
-  \Ec := \Ec, \tester{C_1}( t ) \ror \ldots \ror \Ec := \Ec, \tester{C_n}( t ) 
+  \Ec := \Ec, \tester{\const{C}_1}( t ) \ror \ldots \ror \Ec := \Ec, \tester{\const{C}_n}( t ) 
 }
@@
 \end{tabular}
-\caption{Rule for splitting.  We write $\tester{C_i}( t )$ as shorthand for the equality $t \teq C_i( s^1_i( t ), \ldots, s^n_i( t ) )$.
+\caption{Rule for splitting.  We write $\tester{\const{C}_i}( t )$ as shorthand for the equality $t \teq \const{C}_i( s^1_i( t ), \ldots, s^n_i( t ) )$.
 }

### 2014-12-01 tex/codatadp.tex
sketched an abstract

@@
 
+\hyphenation{data-type data-types co-data-type co-data-types}
+
 \begin{document}
@@
 \begin{abstract}
-Blah.
+Codatatypes naturally capture potentially infinite data structures and
+processes. We introduce a decision procedure that combines reasoning about
+datatypes and codatatypes. The dual of the acyclicity rule of datatypes is a
+bisimilarity rule that identifies observationally equal values. The procedure is
+compatible with the Nelson--Oppen combination method and is useful for both
+proving and model finding. It has been implemented in the latest version of
+CVC4, a modern SMT solver. An evaluation based on problems generated from
+Isabelle/HOL theories demonstrates the potential of the procedure in a proving
+context.
 \end{abstract}
@@
     * and for model finding, without them we quickly get spurious models
+    * explain how finite model finding works
 

### 2014-12-01 tex/codatadp.tex
tuned abstract

@@
 processes. We introduce a decision procedure that combines reasoning about
-datatypes and codatatypes. The dual of the acyclicity rule of datatypes is a
-bisimilarity rule that identifies observationally equal values. The procedure is
-compatible with the Nelson--Oppen combination method and is useful for both
-proving and model finding. It has been implemented in the latest version of
-CVC4, a modern SMT solver. An evaluation based on problems generated from
-Isabelle/HOL theories demonstrates the potential of the procedure in a proving
-context.
+datatypes and codatatypes. The dual of the acyclicity rule for datatypes is a
+bisimilarity rule that identifies observationally equal, possibly cyclic
+values. The procedure is compatible with the Nelson--Oppen method
+and is useful for both proving and model finding. It has been implemented in the
+latest version of CVC4, a modern SMT solver. An evaluation based on problems
+generated from Isabelle/HOL theories demonstrates the potential of the procedure
+in a proving context.
 \end{abstract}

### 2014-12-02 tex/codatadp.tex
more work on intro etc.

@@
 datatypes and codatatypes. The dual of the acyclicity rule for datatypes is a
-bisimilarity rule that identifies observationally equal, possibly cyclic
-values. The procedure is compatible with the Nelson--Oppen method
-and is useful for both proving and model finding. It has been implemented in the
-latest version of CVC4, a modern SMT solver. An evaluation based on problems
-generated from Isabelle/HOL theories demonstrates the potential of the procedure
-in a proving context.
+bisimilarity rule that identifies observationally equal, possibly cyclic values.
+The procedure is complete for the universal fragment and is compatible with the
+Nelson--Oppen method. It has been implemented in the latest version of CVC4, a
+modern SMT solver. An evaluation based on problems generated from Isabelle
+theories demonstrates the potential of the procedure for both proving and model
+finding.
 \end{abstract}
@@
   * briefly: datatypes and why they are useful
-    (terminology: freely-generated, inductive, algebraic, ..., sometimes
-    with different meanings; we'll clarify below)
     * Hoare's recursive data structures
@@
     * common in practice, esp. in computer science applications
+
+    (terminology: freely-generated, inductive, algebraic, ..., sometimes
+    with different meanings; we'll clarify below what we use)
 
@@
 
-\paragraph{Notations.}
-Blah
-
 \paragraph{Related Work.}
@@
 
+\paragraph{Conventions.}
+
+* types are simply atomic sorts, with no structure, and interpreted by nonempty
+  domains
+
+* although nothing prevents composing the decision procedure with theories
+  providing polymorphic types (parametric sorts), such as the theory of arrays
+
+* $\sigma, \tau$ range over types---whether (co)datatypes or uninterpreted types
+* $\delta$ range over (co)datatypes
+
+* use sans-serif for function symbols (e.g., $\const{f}$) to distinguish them
+from variables (e.g., $x$).
+
+* function symbols are accompanied by a type signature, e.g. $\const{f} : \sigma_1 \times \cdots \times \sigma_n \to
+\sigma$ indicates
+
+* don't distinguish between constants and function symbols
+  a constant is merely the case $n = 0$ of a function $f :
+
 \section{Theory of (Co)datatypes}
@@
 \newcommand\const[1]{\textsf{#1}}
-\newcommand\ty[1]{\mathit{#1}}
+\newcommand\ty[1]{\textit{#1}}
 
-For the purposes of this paper, a (co)datatype specification consists of
-$l$~mutually recursive types which are either all datatypes or all codatatypes.
-Each datatype $\delta$ is equipped by $m \ge 1$ constructors, and each
-constructor for $\delta$ takes $n \ge 0$ arguments, each associated with a
-selector, and returns a $\delta$ value. The argument types must be either
-already existing types or be one of the newly introduced (co)datatypes. The
-introduced names for the (co)datatypes, the constructors, and the selectors must
-be distinct with already introduced ones and among themselves. Schematically:
+A signature consists of:
+  * set of plain types, which can be interpreted or uninterpreted
+    * must contain the distinguished \ty{bool} type of Booleans
+  * set of function symbols with argument and return types
+  * list of codatatype specifications, described below
+
+A (co)datatype specification consists of $l$~mutually recursive types which are
+either all datatypes or all codatatypes. Each datatype $\delta$ is equipped by
+$m \ge 1$ constructors, and each constructor for $\delta$ takes zero or more
+arguments and returns a $\delta$ value. The argument types must be either
+plain, be or be one of the newly introduced (co)datatypes. To every
+argument corresponds a selector. The introduced names for the (co)datatypes, the
+constructors, and the selectors must be distinct and different from already
+introduced ones. Schematically:
 %
@@
   \delta_1 & {}=\, \const{C}_{11}(\bigl[\const{s}_{111}{:}\bigr]\; \sigma_{111}, \,\ldots,\, \bigl[\const{s}_{11n_{11}}{:}\bigr]\; \sigma_{11n_{11}}) & \\
-           & {}\;\phantom{=\,}\;{} \smash{\,\,\vdots} \\ % {}\;\phantom{=}\llap{\ensuremath{\mid}}\; \cdots \\
-           & {}\;\phantom{=\,}\llap{\ensuremath{\mid}}\; \const{C}_{1m}(\bigl[\const{s}_{1m1}{:}\bigr]\; \sigma_{1m1}, \,\ldots,\, \big[\const{s}_{1mn_{1m}}{:}\big]\; \sigma_{1mn_{1m}}) \\[-1\jot]
+           & {}\;\phantom{=\,}\;{} \smash{\,\,\,\vdots} \\[-\jot] % {}\;\phantom{=}\llap{\ensuremath{\mid}}\; \cdots \\
+           & {}\;\phantom{=\,}\llap{\ensuremath{\,\mid\,}}\; \const{C}_{1m}(\bigl[\const{s}_{1m1}{:}\bigr]\; \sigma_{1m1}, \,\ldots,\, \big[\const{s}_{1mn_{1m}}{:}\big]\; \sigma_{1mn_{1m}}) \\[-1\jot]
    \smash{\vdots\,\,\,} \\[-1\jot]
@@
 $\const{C}_{ij} : \sigma_{ij1}\times\cdots\times\sigma_{ijk_{ij}} \to \delta_i$
-and $\const{s}_{ijk} : \delta_i \to \sigma_{ijk}$. The selector names are often
-omitted in our examples, in which case default names are generated.
+and $\const{s}_{ijk} : \delta_i \to \sigma_{ijk}$. Defaults are assumed for
+the selector names if they are omitted.
 
@@
       \keyw{datatype}\;\, \ty{tree} & \,=\, \const{Node}(\ty{int},\, \ty{forest}) \\[-\jot]
-      \keyw{and}\;\, \ty{forest} & \,=\, \const{FNil} \mid \const{FCons}(\ty{tree},\, \ty{forest})
+      \keyw{and}\;\, \ty{forest} & \,=\, \const{FNil} \,\mid\, \const{FCons}(\ty{tree},\, \ty{forest})
 \end{aligned}
@@
 \[\begin{aligned}[t]
-      \keyw{codatatype}\; \,\ty{llist}(\alpha) & \,=\, \const{LNil} \mid \const{LCons}(\alpha,\, \ty{llist}(\alpha)) \\
+      \keyw{codatatype}\; \,\ty{llist}(\alpha) & \,=\, \const{LNil} \,\mid\, \const{LCons}(\alpha,\, \ty{llist}(\alpha)) \\[-.5\jot]
       \keyw{datatype}\;\, \ty{tree} & \,=\, \const{Node}(\ty{int},\, \ty{llist}(\ty{tree}))
@@
 \]
-is beyond the fragment considered here.
-% (although it makes sense and is supported by Isabelle \cite{xxx}).
+is beyond the fragment considered here. In principle, rank-1 (top-level)
+polymorphism would be easy to support; nesting datatypes inside datatypes,
+and likewise for codatatypes, can be reduced to the mutual case. So the only
+genuinely interesting cases missing are mixed nested (co)recursion
+(as in the \ty{tree} example above) as well as (co)recursion through a
+non-(co)datatype (both of which do make sense \cite{blanchette-et-al-2014-codata}).
 
@@
   * Constructed examples
+    * proving
+    * model finding?
+
   * Isabelle benchmarks

### 2014-12-02 tex/codatadp.tex
fix

@@
 \DeclareMathAlphabet{\mathcal}{OT1}{pzc}{m}{it}
+
+\newcommand\keyw[1]{\textbf{#1}}
+\newcommand\const[1]{\textsf{#1}}
+\newcommand\ty[1]{\textit{#1}}
 
@@
 
-* $\sigma, \tau$ range over types---whether (co)datatypes or uninterpreted types
+* $\tau$ range over types---whether (co)datatypes or uninterpreted types
 * $\delta$ range over (co)datatypes
@@
 
-* function symbols are accompanied by a type signature, e.g. $\const{f} : \sigma_1 \times \cdots \times \sigma_n \to
-\sigma$ indicates
+* function symbols are accompanied by a type signature, e.g. $\const{f} : \tau_1 \times \cdots \times \tau_n \to
+\tau$ indicates
 
 * don't distinguish between constants and function symbols
-  a constant is merely the case $n = 0$ of a function $f :
+  a constant is merely the case $n = 0$ of a function $f : $
 
@@
 \label{sec:theory-of-co-datatypes}
-
-\newcommand\keyw[1]{\textbf{#1}}
-\newcommand\const[1]{\textsf{#1}}
-\newcommand\ty[1]{\textit{#1}}
 
@@
 (\keyw{co})\keyw{datatype}\;\,
-  \delta_1 & {}=\, \const{C}_{11}(\bigl[\const{s}_{111}{:}\bigr]\; \sigma_{111}, \,\ldots,\, \bigl[\const{s}_{11n_{11}}{:}\bigr]\; \sigma_{11n_{11}}) & \\
+  \delta_1 & {}=\, \const{C}_{11}(\bigl[\const{s}_{111}{:}\bigr]\; \tau_{111}, \,\ldots,\, \bigl[\const{s}_{11n_{11}}{:}\bigr]\; \tau_{11n_{11}}) & \\
            & {}\;\phantom{=\,}\;{} \smash{\,\,\,\vdots} \\[-\jot] % {}\;\phantom{=}\llap{\ensuremath{\mid}}\; \cdots \\
-           & {}\;\phantom{=\,}\llap{\ensuremath{\,\mid\,}}\; \const{C}_{1m}(\bigl[\const{s}_{1m1}{:}\bigr]\; \sigma_{1m1}, \,\ldots,\, \big[\const{s}_{1mn_{1m}}{:}\big]\; \sigma_{1mn_{1m}}) \\[-1\jot]
+           & {}\;\phantom{=\,}\llap{\ensuremath{\,\mid\,}}\; \const{C}_{1m}(\bigl[\const{s}_{1m1}{:}\bigr]\; \tau_{1m1}, \,\ldots,\, \big[\const{s}_{1mn_{1m}}{:}\big]\; \tau_{1mn_{1m}}) \\[-1\jot]
    \smash{\vdots\,\,\,} \\[-1\jot]
@@
 with
-$\const{C}_{ij} : \sigma_{ij1}\times\cdots\times\sigma_{ijk_{ij}} \to \delta_i$
-and $\const{s}_{ijk} : \delta_i \to \sigma_{ijk}$. Defaults are assumed for
+$\const{C}_{ij} : \tau_{ij1}\times\cdots\times\tau_{ijk_{ij}} \to \delta_i$
+and $\const{s}_{ijk} : \delta_i \to \tau_{ijk}$. Defaults are assumed for
 the selector names if they are omitted.

### 2014-12-02 tex/codatadp.tex
more

@@
 * function symbols are accompanied by a type signature, e.g. $\const{f} : \tau_1 \times \cdots \times \tau_n \to
-\tau$ indicates
+\tau$ indicates takes $n \ge 0$ arguments of types $\tau_1,\ldots,\tau_n$ and returns a value of type $\tau$;
+* applied as $\const{f}(t_1,\ldots,t_n)$ to well-typed arguments $t_1 : \tau_1$, \ldots,
+$t_n : \tau_n$.
 
-* don't distinguish between constants and function symbols
-  a constant is merely the case $n = 0$ of a function $f : $
+* nullary function symbols are called constants;
+  following tradition, we simply write $\const{c} : \tau$ for the type signature 
+  and $\const{c}$ rather than $\const{c}()$.
+
+* in keeping with SMT conventions, predicate symbols are simply functions that
+  return $\ty{Bool}$
 
@@
 (as in the \ty{tree} example above) as well as (co)recursion through a
-non-(co)datatype (both of which do make sense \cite{blanchette-et-al-2014-codata}).
+non-(co)datatype (both of which make sense \cite{blanchette-et-al-2014-codata}).
 
@@
   * more evaluation
+    * use it in a new Nitpick-like tool
   * recursive and corecursive functions
     (and rewriting)
+  * extensions, perhaps straightforward, to polymorphism and nested
+    (co)recursion
 

### 2014-12-02 tex/codatadp.tex
more

@@
    pdftitle={A Decision Procedure for (Co)datatypes in SMT Solvers},
-   pdfauthor={Andrew J. Reynolds and Jasmin Christian Blanchette},
+   pdfauthor={Andrew Reynolds and Jasmin Christian Blanchette},
    pdfkeywords={},
@@
 
-\author {Andrew J. Reynolds\inst{1} \and Jasmin Christian Blanchette\inst{2,3}}
-\authorrunning {A. J. Reynolds \and J. C. Blanchette}
+\author {Andrew Reynolds\inst{1} \and Jasmin Christian Blanchette\inst{2,3}}
+\authorrunning {A. Reynolds \and J. C. Blanchette}
 \institute{
@@
 
+* monomorphic (or many-sorted) first-order logic
+* Signature: $\Sigma = (\mathcal{T}, \mathcal{F})$
+
 * types are simply atomic sorts, with no structure, and interpreted by nonempty
   domains
+  * set of types must contain the distinguished \ty{bool} type of Booleans, interpreted
+    as the set of truth values, and may contain other interpreted types (e.g., \ty{int}, \ty{real}).
 
 * although nothing prevents composing the decision procedure with theories
-  providing polymorphic types (parametric sorts), such as the theory of arrays
+  providing polymorphic types (parametric sorts), such as for arrays (e.g., \ty{array}(\alpha,\beta))
+
+* in keeping with SMT conventions, there are no separate predicate symbols;
+  predicates are modeled as functions that return $\ty{Bool}$
 
@@
 
-* in keeping with SMT conventions, predicate symbols are simply functions that
-  return $\ty{Bool}$
-
 \section{Theory of (Co)datatypes}
@@
 
-A signature consists of:
-  * set of plain types, which can be interpreted or uninterpreted
-    * must contain the distinguished \ty{bool} type of Booleans
-  * set of function symbols with argument and return types
-  * list of codatatype specifications, described below
+We assume a signature $\Sigma = (\mathcal{T}, \mathcal{F})$.
+Furthermore, a partition of the types:
+  $\mathcal{T} = \mathcal{I} \uplus \mathcal{C} \uplus \mathcal{N}$
+with
+  * $\mathcal{I}$: datatypes
+  * $\mathcal{C}$: codatatypes
+  * $\mathcal{N}$: non-(co)datatypes
+
+In practice, e.g. in an SMT problem, we'd have:
+  * declaration of uninterpreted types (in any order)
+  * specification of mutual groups of (co)datatypes, in a correct order
 

### 2014-12-02 tex/codatadp.tex
more

@@
 * although nothing prevents composing the decision procedure with theories
-  providing polymorphic types (parametric sorts), such as for arrays (e.g., \ty{array}(\alpha,\beta))
+  providing polymorphic types (parametric sorts), such as for arrays (e.g., $\ty{array}(\alpha,\beta)$)
 
@@
 We assume a signature $\Sigma = (\mathcal{T}, \mathcal{F})$.
+
 Furthermore, a partition of the types:
@@
 with
-  * $\mathcal{I}$: datatypes
+  * $\mathcal{D}$: datatypes
   * $\mathcal{C}$: codatatypes
   * $\mathcal{N}$: non-(co)datatypes
+
+And a partition of the functions:
+  * $\mathcal{C}$
+
+for each datatype or codatatype $\delta$:
+  *
 

### 2014-12-02 tex/codatadp.tex
Add example for bisimilar.  Add explanation.

@@
 processes. We introduce a decision procedure that combines reasoning about
-datatypes and codatatypes. The dual of the acyclicity rule for datatypes is a
+datatypes and codatatypes. The dual of the cyclicity rule for datatypes is a
 bisimilarity rule that identifies observationally equal, possibly cyclic values.
@@
 
-We present the calculus in three phases, shown in Figures~\ref{fig:cc-rules},~\ref{fig:ab-rules}, and~\ref{fig:split-rule}.
+We present the calculus in three sections, shown in Figures~\ref{fig:cc-rules},~\ref{fig:ab-rules}, and~\ref{fig:split-rule}.
 Following the conventions from~\cite{}, 
@@
 The rules \rn{Unify$_1$} and \rn{Unify$_2$} together compute (downwards) unification,
-where additional equalities are inferred based on the injectivity of constructors (by \rn{Unify$_1$}),
-and failures to unify equivalent terms are recognized (by \rn{Unify$_2$}).
+where additional equalities are inferred based on the injectivity of constructors by \rn{Unify$_1$},
+and failures to unify equivalent terms are recognized by \rn{Unify$_2$}.
 
-Second, when applying the rules in Figure~\ref{fig:ab-rules}, 
+When applying the rules in Figure~\ref{fig:ab-rules}, 
 we assume that $\Ec$ is saturated with respect to the rules in Figure~\ref{fig:cc-rules}, that is,
@@
 We say that edge $\ec{t_1} \rightarrow^{C}_n \ec{t_2}$ is \emph{induced by $\Ec$} if and only if $C( \vec u ) \in \ec{t_1}$ for some $\vec u$ and $\ec{t_2} = \ec{u_n}$.
+In other words, $t_1$ is equivalent to a $C$-application whose $n^{th}$ argument is equivalent to $t_2$.
 A \emph{path} $p$ is a non-empty ordered list of edges $e_1, \ldots e_n$ such that the destination of $e_i$ is the source of $e_{i+1}$ for $i = 1, \ldots (n-1)$,
@@
 Notice that since $\Ec$ contains a finite set of equivalence classes, all paths induced by $\Ec$ have finite length.
-We call the source of a path is the source of its first edge, and the destination of a path is the destination of its last edge.
+We call the \emph{source} of a path is the source of its first edge, and the \emph{destination} of a path is the destination of its last edge.
 The set $\ttpath{\Ec}{\ec{t_1}}{\ec{t_2}}$ is the smallest set containing all paths with source $\ec{t_1}$ and destination $\ec{t_2}$ that are induced by $\Ec$.
 The set $\tpath{\Ec}{\ec{t_1}}$ is the union of $\ttpath{\Ec}{\ec{t_1}}{\ec{t_2}}$ for all $t_2 \in \tEc$.
+We will write, for instance, $\ec{t_1} \rightarrow^{C_1}_{n_1} \ec{t_2} \rightarrow^{C_2}_{n_2} \ec{t_3}$ 
+as shorthand for the path containing an edge from $\ec{t_1}$ to $\ec{t_2}$, followed by one from $\ec{t_2}$ to $\ec{t_3}$.
+
+The rule \rn{Cyclic} recognizes when a path exists from an equivalence class $\ec{t}$ to itself, as indicated
+by the assumption $p \in \ttpath{\Ec}{\ec{t}}{\ec{t}}$.
+In this case, $\Ec$ entails that $t$ contains itself as a (strict) subterm,
+and thus if $t$ is a datatype, $\Ec$ must be unsatisfiable.
+The rule \rn{Bisimilar} recognizes when two equivalence classes $\ec{t_1}$ and $\ec{t_2}$ having co-datatype type are observationally equivalent.
+In particular, if there is a mapping $\sigma$ over equivalence classes such that all paths $p$ from $t_2$
+have a corresponding path $p \sigma$ from $t_1$,
+then $t_1$ must be equal to $t_2$.
+We demonstrate this rule concretely with the following example.
+
+\begin{example}
+Let $\Ec$ be the set $\{ x \teq \const{LCons}(y,\const{LCons}(y,x)), x \tneq \const{LCons}(y,x) \}$.  \rem{ensure that \const{LCons} understood}
+The equivalence classes of $\Ec$ are $\{ x, \const{LCons}(y,\const{LCons}(y,x)) \}$, $\{ \const{LCons}(y,x) \}$, and $\{ y \}$.
+The set $\tpath{\Ec}{\ec{x}}$ is:
+\[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
+\{ & \ec{x} \rightarrow^{\const{LCons}}_1 & \ec{y}, & & \\ 
+   & \ec{x} \rightarrow^{\const{LCons}}_2 & \ec{\const{LCons}(y,x)} \rightarrow^{\const{LCons}}_1 & \ec{y},& \\
+   & \ec{x} \rightarrow^{\const{LCons}}_2 & \ec{\const{LCons}(y,x)} \rightarrow^{\const{LCons}}_2 & \ec{x} & \}.
+\end{array}\]
+The set $\tpath{\Ec}{\ec{\const{LCons}(y,x)}}$ is:
+\[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
+\{ & \ec{\const{LCons}(y,x)} \rightarrow^{\const{LCons}}_1 & \ec{y}, & & \\ 
+   & \ec{\const{LCons}(y,x)} \rightarrow^{\const{LCons}}_2 & \ec{x} \rightarrow^{\const{LCons}}_1 & \ec{y},& \\
+   & \ec{\const{LCons}(y,x)} \rightarrow^{\const{LCons}}_2 & \ec{x} \rightarrow^{\const{LCons}}_2 & \ec{\const{LCons}(y,x)} & \}.
+\end{array}\]
+Thus, we may conclude $x \teq \const{LCons}(y,x)$ by the rule \rn{Bisimilar}
+for the mapping $\sigma :=$ $\{ \ec{x} \mapsto \ec{\const{LCons}(y,x)}, \ec{\const{LCons}(y,x)} \mapsto \ec{x} \}$,
+after which the rule \rn{Conflict} may be applied.
+\end{example}
 
@@
 \begin{tabular}{c}
-\rn{Acyclic}
+\rn{Cyclic}
 \(
@@
 \end{tabular}
-\caption{Rules for acyclicity and bisimilarity.
+\caption{Rules for cyclicity and bisimilarity.
 }
@@
 }{
-  \Ec := \Ec, \tester{\const{C}_1}( t ) \ror \ldots \ror \Ec := \Ec, \tester{\const{C}_n}( t ) 
+  \Ec := \Ec, t \teq \const{C}_1( s^1_1( t ), \ldots, s^{a_1}_1( t ) ) \ror \ldots \ror \Ec := \Ec, t \teq \const{C}_n( s^1_n( t ), \ldots, s^{a_n}_n( t ) ) 
 }
@@
 \end{tabular}
-\caption{Rule for splitting.  We write $\tester{\const{C}_i}( t )$ as shorthand for the equality $t \teq \const{C}_i( s^1_i( t ), \ldots, s^n_i( t ) )$.
+\caption{Rule for splitting.  We split on the constructor for $t$ if there is a selector applied to $t$ or if its type is finite.
 }

### 2014-12-02 tex/codatadp.tex
more stuff

--- 2014-12-02 tex/codatadp.tex
@@
 \newcommand{\thD}{T_{D}}
+
+%\newcommand\Types{\mathcal{T}}
+\newcommand\Types{\mathcal{Y}}
+\newcommand\Funcs{\mathcal{F}}
+
+\newcommand\Data{\Types_{\mathrm{dt}}}
+\newcommand\Codata{\Types_{\mathrm{codt}}}
+\newcommand\Nondata{\Types_{\mathrm{other}}}
+
+\newcommand\Ctr{\Funcs_{\mathrm{ctr}}}
+\newcommand\Sel{\Funcs_{\mathrm{sel}}}
+%\newcommand\Plainfuncs{\Funcs_{\mathrm{other}}}
 
@@
 * monomorphic (or many-sorted) first-order logic
-* Signature: $\Sigma = (\mathcal{T}, \mathcal{F})$
+* Signature: $\Sigma = (\Types, \Funcs)$
 
@@
 
-We assume a signature $\Sigma = (\mathcal{T}, \mathcal{F})$.
+We assume a signature $\Sigma = (\Types, \Funcs)$.
 
 Furthermore, a partition of the types:
-  $\mathcal{T} = \mathcal{I} \uplus \mathcal{C} \uplus \mathcal{N}$
+  $\Types = \Data \uplus \Codata \uplus \Nondata$
 with
-  * $\mathcal{D}$: datatypes
-  * $\mathcal{C}$: codatatypes
-  * $\mathcal{N}$: non-(co)datatypes
+  * $\Data$: datatypes
+  * $\Codata$: codatatypes
+  * $\Nondata$: non-(co)datatypes
 
 And a partition of the functions:
-  * $\mathcal{C}$
+  $\Funcs = \Ctr \uplus \Sel$
+with
+  * $\Ctr$: constructors
+  * $\Sel$: selectors
 
-for each datatype or codatatype $\delta$:
-  *
+(Do not need to consider other function symbols, because they would be
+abstracted away as variables in the context of theory combination.)
 
@@
   * declaration of uninterpreted types (in any order)
-  * specification of mutual groups of (co)datatypes, in a correct order
+  * list of specifications of mutually (co)recursive groups of (co)datatypes
 

### 2014-12-02 tex/codatadp.tex
more

@@
 \DeclareMathAlphabet{\mathcal}{OT1}{pzc}{m}{it}
+
+\newcommand\Sig{\mathrm{\Sigma}}
 
@@
     * codatatypes are never empty (e.g. finite streams are rejected)
+
+\[
+      \keyw{codatatype}\; \,\ty{llist} \,=\, \const{LNil} \,\mid\, \const{LCons}(\ty{int},\, \ty{llist})
+\]
+
     * no acyclicity (e.g. xs = lcons(0, xs) is satisfiable)
@@
 * monomorphic (or many-sorted) first-order logic
-* Signature: $\Sigma = (\Types, \Funcs)$
+* Signature: $\Sig = (\Types, \Funcs)$
 
@@
 
-We assume a signature $\Sigma = (\Types, \Funcs)$.
+We assume a signature $\Sig = (\Types, \Funcs)$.
 
@@
 (\keyw{co})\keyw{datatype}\;\,
-  \delta_1 & {}=\, \const{C}_{11}(\bigl[\const{s}_{111}{:}\bigr]\; \tau_{111}, \,\ldots,\, \bigl[\const{s}_{11n_{11}}{:}\bigr]\; \tau_{11n_{11}}) & \\
+  \delta_1 & {}=\, \const{C}_{1}\rlap{${}_{1}$}\phantom{{}_{m}}(\bigl[\const{s}_{111}{:}\bigr]\; \tau_{111}, \,\ldots,\, \bigl[\const{s}_{11n_{11}}{:}\bigr]\; \tau_{11n_{11}}) & \\
            & {}\;\phantom{=\,}\;{} \smash{\,\,\,\vdots} \\[-\jot] % {}\;\phantom{=}\llap{\ensuremath{\mid}}\; \cdots \\
@@
 \[\begin{aligned}[t]
-      \keyw{codatatype}\; \,\ty{llist}(\alpha) & \,=\, \const{LNil} \,\mid\, \const{LCons}(\alpha,\, \ty{llist}(\alpha)) \\[-.5\jot]
-      \keyw{datatype}\;\, \ty{tree} & \,=\, \const{Node}(\ty{int},\, \ty{llist}(\ty{tree}))
+      \keyw{codatatype}\; \,\ty{pllist}(\alpha) & \,=\, \const{PLNil} \,\mid\, \const{PLCons}(\alpha,\, \ty{pllist}(\alpha)) \\[-.5\jot]
+      \keyw{datatype}\;\, \ty{tree} & \,=\, \const{Node}(\ty{int},\, \ty{pllist}(\ty{tree}))
 \end{aligned}
@@
 
-\rem{Discuss signature $\Sigma$, theory $\thD$}
+\rem{Discuss signature $\Sig$, theory $\thD$}
 This section presents a calculus for determining the satisfiability of conjunctions of $\thD$-constraints.
 
-Our calculus consists of derivation rules that operate on a set $\Ec$ of equalities and disequalities between $\Sigma$-terms.
-We will write $\tEc$ to denote the set of $\Sigma$-terms occurring in $\Ec$.
+Our calculus consists of derivation rules that operate on a set $\Ec$ of equalities and disequalities between $\Sig$-terms.
+We will write $\tEc$ to denote the set of $\Sig$-terms occurring in $\Ec$.
 We will commonly denote tuples of terms $( t_1, \ldots, t_n )$ in bold font, as $\vec t$.
-We assume that all $\Sigma$-terms in $\Ec$ are \emph{normalized}, meaning that all subterms of the form $s^i_j( \const{C}_j( \vec t ) )$
+We assume that all $\Sig$-terms in $\Ec$ are \emph{normalized}, meaning that all subterms of the form $s^i_j( \const{C}_j( \vec t ) )$
 are replaced by $t_i$, and moreover assume that all additional constraints added to $\Ec$ are normalized in this manner.

### 2014-12-02 tex/codatadp.tex
Termination proof (draft).

@@
 \newcommand{\tEc}{\mathcal{T}(\Ec)}
+\newcommand{\tcEc}{\mathcal{T}^\ast(\Ec)}
 \newcommand{\rn}[1]{\textsf{\small #1}}
@@
 
-When applying the rules in Figure~\ref{fig:ab-rules}, 
+\begin{figure}[t]
+\centering
+\begin{tabular}{c}
+\rn{Cyclic}
+\(
+\inferrule{
+  t : \tau \text{ inductive } 
+  \quad
+  p \in \ttpath{\Ec}{\ec{t}}{\ec{t}}
+}{
+  \bot
+}
+\)
+\\[3.7ex]
+\rn{Bisimilar}
+\(
+\inferrule{
+ t_1, t_2 : \tau \text{ co-inductive }
+  \quad
+ \exists \sigma. p \sigma \in \tpath{\Ec}{\ec{t_1}} \text{ for each } p \in \tpath{\Ec}{\ec{t_2}}
+}{
+ \Ec := \Ec, t_1 \teq t_2
+}
+\)
+\end{tabular}
+\caption{Rules for cyclicity and bisimilarity.
+}
+\label{fig:ab-rules}
+\end{figure}
+
+Before applying the rules in Figure~\ref{fig:ab-rules}, 
 we assume that $\Ec$ is saturated with respect to the rules in Figure~\ref{fig:cc-rules}, that is,
-no non-redudant application of a rule can be applied to $\Ec$.
-In this case, it is easy to see that $\Ec$ induces an equivalence relation over $\tEc$ such that two terms $t_1$ and $t_2$ are equivalent if and only if $t_1 \teq t_2 \in \Ec$.
+no non-redudant application of one of its rules can be applied to $\Ec$.
+When this is the case, it is easy to see that $\Ec$ induces an equivalence relation over $\tEc$ such that two terms $t_1$ and $t_2$ are equivalent if and only if $t_1 \teq t_2 \in \Ec$.
 Thus, we will in the following consider $\Ec$ as a set of equivalence classes of terms. 
@@
 
-\begin{figure}[t]
-\centering
-\begin{tabular}{c}
-\rn{Cyclic}
-\(
-\inferrule{
-  t : \tau \text{ inductive } 
-  \quad
-  p \in \ttpath{\Ec}{\ec{t}}{\ec{t}}
-}{
-  \bot
-}
-\)
-\\[3.7ex]
-\rn{Bisimilar}
-\(
-\inferrule{
- t_1, t_2 : \tau \text{ co-inductive }
-  \quad
- \exists \sigma. p \sigma \in \tpath{\Ec}{\ec{t_1}} \text{ for each } p \in \tpath{\Ec}{\ec{t_2}}
-}{
- \Ec := \Ec, t_1 \teq t_2
-}
-\)
-\end{tabular}
-\caption{Rules for cyclicity and bisimilarity.
-}
-\label{fig:ab-rules}
-\end{figure}
-
 \rem{Change notation for inductive/co-inductive.}
@@
 
-A \emph{derivation tree} is a tree whose nodes are sets of equalities, where non-root nodes are obtained by a non-redundant application of a derivation rule to its parent node.
+Finally, if neither a rule from Figure~\ref{fig:cc-rules} or Figure~\ref{fig:ab-rules} applies to $\Ec$, 
+we split on the constructor type for certain terms $t$.
+In particular, if there is a selector term applied to $t$, or if $t$ has a finite type, 
+then we require the equivalence class of $t$ to contain a constructor term, which is enforced by the rule \rn{Split}.
+
+\subsection{Correctness}
+
+A \emph{derivation tree} is a tree whose nodes are sets of equalities, where non-root nodes are obtained by 
+a non-redundant application of a derivation rule to its parent node.
 A derivation tree is \emph{closed} if all of its leaf nodes are $\bot$.
@@
 
-\subsection{Correctness}
+\begin{lemma}[Termination]
+All derivation trees are finite.
+\end{lemma}
+\begin{proof}
+Consider a derivation tree with root node $\Ec$.
+Let $S^0$ be the set of terms that are arguments of a selector application in $\tEc$.
+For $i \geq 0$, let $S^{i+1}$ $=$ $S_i \cup \{ s( t ) \mid t \in S^i, s \in sel( t ), t : \tau \text{ finite} \}$.
+Let $S^\ast$ be the fix point of this sequence, which is guarenteed to be finite 
+since $S^\ast \setminus S^0$ only includes selectors applied to terms having finite type.
+Now, let $\tcEc$ be the set all subterms of $\Ec \cup \{ C_i( s^1_i( t ), \ldots, s^{a_j}_i( t ) ) \mid t \in S^\ast, C_i \in cons( t ) \}$.
+In a derivation tree with root node $\Ec$, 
+it can be shown by case analysis on the rules of the calculus that each non-root node is 
+obtained as a result of adding an equality between two terms from $\tcEc$ to its parent node.
+Thus, the maximum depth of a branch in a derivation tree with root node $\Ec$ is $\mid \tcEc \mid^2$,
+which is finite since $\tcEc$ is finite.
+\end{proof}
 
@@
 \end{lemma}
+\begin{proof}
+TODO
+\end{proof}
 
@@
 \end{lemma}
-
-\begin{lemma}[Termination]
-All derivation trees are finite.
-\end{lemma}
+\begin{proof}
+TODO
+\end{proof}
 

### 2014-12-02 tex/codatadp.tex
conventions

--- 2014-12-02 tex/codatadp.tex
@@
 \newcommand\Data{\Types_{\mathrm{dt}}}
-\newcommand\Codata{\Types_{\mathrm{codt}}}
-\newcommand\Nondata{\Types_{\mathrm{other}}}
+\newcommand\Codata{\Types_{\mathrm{cdt}}}
+\newcommand\Nondata{\Types_{\mathrm{oth}}}
 
@@
     * in Agda, Coq, Matita -- recently also in Isabelle/HOL
-    * but not in SMT-LIB (2.5?)
+    * but not in SMT-LIB 2 \cite{xxx}
 
@@
 \paragraph{Conventions.}
+This paper assumes a monomorphic (or many-sorted) first-order logic throughout.
+A signature $\Sig = (\Types, \Funcs)$ consists of a set of types $\Types$ and a
+set of function symbols $\Funcs$. Types are simply atomic sorts, with no
+structure, and interpreted by nonempty domains. The set $\Types$ must contain a
+distinguished type \ty{bool} interpreted as the set of truth
+values. %, and may contain other interpreted types (e.g., \ty{int}, \ty{real}).
+Predicates are simply functions that return $\ty{bool}$.
+The metavariables $\delta, \tau$ range over types, with $\delta$ reserved
+for (co)datatypes.
 
-* monomorphic (or many-sorted) first-order logic
-* Signature: $\Sig = (\Types, \Funcs)$
+Function symbols are written in a sans-serif font (e.g., $\const{f}$, $\const{g}$) to
+distinguish them from variables (e.g., $x$, $y$). With each function symbol \const{f}
+is associated a list of argument types $\tau_1,\ldots,\tau_n$ (with $n \ge 0$)
+and a return type $\tau$. This connection can be expressed compactly as
+$\const{f} : \tau_1 \times \cdots \times \tau_n \to \tau$
+(which collapses to $\const{f} : \tau$ if $n = 0$).
+For a term $t$, the notation $t : \tau$ indicates that $t$ has type $\tau$.
+Functions are applied the standard way, with $\const{f}(t_1,\ldots,t_n)$
+applying \const{f} to $n$ well-typed arguments $t_1 : \tau_1$, \ldots, $t_n :
+\tau_n$. Nullary function symbols are called constants and can appear as is in
+terms, without parentheses.
 
-* types are simply atomic sorts, with no structure, and interpreted by nonempty
-  domains
-  * set of types must contain the distinguished \ty{bool} type of Booleans, interpreted
-    as the set of truth values, and may contain other interpreted types (e.g., \ty{int}, \ty{real}).
-
-* although nothing prevents composing the decision procedure with theories
-  providing polymorphic types (parametric sorts), such as for arrays (e.g., $\ty{array}(\alpha,\beta)$)
-
-* in keeping with SMT conventions, there are no separate predicate symbols;
-  predicates are modeled as functions that return $\ty{Bool}$
-
-* $\tau$ range over types---whether (co)datatypes or uninterpreted types
-* $\delta$ range over (co)datatypes
-
-* use sans-serif for function symbols (e.g., $\const{f}$) to distinguish them
-from variables (e.g., $x$).
-
-* function symbols are accompanied by a type signature, e.g. $\const{f} : \tau_1 \times \cdots \times \tau_n \to
-\tau$ indicates takes $n \ge 0$ arguments of types $\tau_1,\ldots,\tau_n$ and returns a value of type $\tau$;
-* applied as $\const{f}(t_1,\ldots,t_n)$ to well-typed arguments $t_1 : \tau_1$, \ldots,
-$t_n : \tau_n$.
-
-* nullary function symbols are called constants;
-  following tradition, we simply write $\const{c} : \tau$ for the type signature 
-  and $\const{c}$ rather than $\const{c}()$.
+%* although nothing prevents composing the decision procedure with theories
+%  providing polymorphic types (parametric sorts), such as for arrays (e.g., $\ty{array}(\alpha,\beta)$)
 
@@
 
-We assume a signature $\Sig = (\Types, \Funcs)$.
+We assume a signature $\Sig = (\Types, \Funcs)$. The types are partioned into
+$\Types = \Data \mathrel{\uplus} \Codata \mathrel{\uplus} \Nondata$, where $\Data$ are the
+datatypes, $\Codata$ are the codatatypes, and $\Nondata$ are the remaining
+types. The functions are partitioned into $\Funcs = \Ctr \mathrel{\uplus} \Sel$, where
+$\Ctr$ are the constructors and $\Sel$ are the selectors. (There is no need to
+consider other function symbols, because they can be abstracted away as
+variables when combining theories.)
 
-Furthermore, a partition of the types:
-  $\Types = \Data \uplus \Codata \uplus \Nondata$
-with
-  * $\Data$: datatypes
-  * $\Codata$: codatatypes
-  * $\Nondata$: non-(co)datatypes
-
-And a partition of the functions:
-  $\Funcs = \Ctr \uplus \Sel$
-with
-  * $\Ctr$: constructors
-  * $\Sel$: selectors
-
-(Do not need to consider other function symbols, because they would be
-abstracted away as variables in the context of theory combination.)
-
-In practice, e.g. in an SMT problem, we'd have:
-  * declaration of uninterpreted types (in any order)
-  * list of specifications of mutually (co)recursive groups of (co)datatypes
+In practice (e.g., in an SMT problem), the signature is given by specifying the
+uninterpreted types first (in any order), then the (co)datatypes with their
+constructors and selectors in groups of mutually (co)recursive groups of
+(co)datatypes, and finally any other function symbols.
 
@@
 constructors, and the selectors must be distinct and different from already
-introduced ones. Schematically:
+introduced ones.%
+\footnote{In practice, it can be useful to specify the same selector
+for several constructors of the same (co)datatype,
+as long as the argument types coincide \cite[Section~3]{blanchette-et-al-2014-codata}.
+However, this is not allowed by SMT-LIB 2.}
+Schematically, using a syntax similar to Standard ML, Haskell,
+or SMT-LIB:
 %

### 2014-12-02 tex/codatadp.tex
minor
more

@@
   * less briefly: codatatypes and why they are useful
-    * in Agda, Coq, Matita -- recently also in Isabelle/HOL
+    * in Agda, Coq, Matita -- recently also in Isabelle/HOL \cite{nipkow-et-al-2002}
     * but not in SMT-LIB 2 \cite{xxx}
@@
 Functions are applied the standard way, with $\const{f}(t_1,\ldots,t_n)$
-applying \const{f} to $n$ well-typed arguments $t_1 : \tau_1$, \ldots, $t_n :
+applying the $n$-ary function symbol
+\const{f} to $n$ well-typed arguments $t_1 : \tau_1$, \ldots, $t_n :
 \tau_n$. Nullary function symbols are called constants and can appear as is in

### 2014-12-02 tex/codatadp.tex
more

@@
     * in Agda, Coq, Matita -- recently also in Isabelle/HOL \cite{nipkow-et-al-2002}
-    * but not in SMT-LIB 2 \cite{xxx}
+    * but not in SMT-LIB 2 \cite{barrett-et-al-2010}
 
@@
 \const{f} to $n$ well-typed arguments $t_1 : \tau_1$, \ldots, $t_n :
-\tau_n$. Nullary function symbols are called constants and can appear as is in
-terms, without parentheses.
+\tau_n$. Nullary function symbols are called constants and can appear without
+parentheses in terms.
 

### 2014-12-02 tex/codatadp.tex
Refactoring

@@
 
+We present the calculus in three sections, shown in Figures~\ref{fig:cc-rules},~\ref{fig:ab-rules}, and~\ref{fig:split-rule}.
+Following the conventions from~\cite{}, 
+the derivation rules of our calculus are given in \emph{guarded assignment form},
+where a rule can be applied to $\Ec$ if it meets all of the specified preconditions for $\Ec$.
+The conclusion of a rule either describes equalities to be added to $\Ec$ (in which case we will call an application of it to be \emph{non-terminal}),
+or is $\bot$ (in which case we will call an application of it to be \emph{terminal}).
+A rule may have multiple conclusions separated by $\parallel$, which denotes a non-deterministic branching.
+An application of a rule is \emph{redundant} if it is non-terminal and at least one branch in its conclusion does not add a new equality to $\Ec$.
+
 \begin{figure}[t]
@@
 
-We present the calculus in three sections, shown in Figures~\ref{fig:cc-rules},~\ref{fig:ab-rules}, and~\ref{fig:split-rule}.
-Following the conventions from~\cite{}, 
-the derivation rules of our calculus are given in \emph{guarded assignment form},
-where a rule can be applied to $\Ec$ if it meets all of the specified preconditions for $\Ec$.
-The conclusion of a rule either describes equalities to be added to $\Ec$ (in which case we will call an application of it to be \emph{non-terminal}),
-or is $\bot$ (in which case we will call an application of it to be \emph{terminal}).
-A rule may have multiple conclusions separated by $\parallel$, which denotes a non-deterministic branching.
-An application of a rule is \emph{redundant} if it is non-terminal and at least one branch in its conclusion does not add a new equality to $\Ec$.
-
-First, Figure~\ref{fig:cc-rules} gives the rules for constructing the bidirectional closure of $\Ec$.
+Figure~\ref{fig:cc-rules} gives the rules for constructing the bidirectional closure of $\Ec$.
 The rules \rn{Refl}, \rn{Symm}, \rn{Trans}, and \rn{Cong} together compute the (upwards) congruence closure,
@@
 \inferrule{
-  t : \tau \text{ inductive } 
+  t : \tau, \tau \in \Data
   \quad
@@
 \inferrule{
- t_1, t_2 : \tau \text{ co-inductive }
+ t_1, t_2 : \tau, \tau \in \Codata
   \quad
@@
 
-\rem{Change notation for inductive/co-inductive.}
-
 \begin{figure}[t]
@@
 Let $S^\ast$ be the fix point of this sequence, which is guarenteed to be finite 
-since $S^\ast \setminus S^0$ only includes selectors applied to terms having finite type.
-Now, let $\tcEc$ be the set all subterms of $\Ec \cup \{ C_i( s^1_i( t ), \ldots, s^{a_j}_i( t ) ) \mid t \in S^\ast, C_i \in cons( t ) \}$.
+since $S^i \setminus S^0$ only includes selectors applied to terms having finite type.
+Now, let $\tcEc$ be the set of subterms of $\Ec \cup \{ C_i( s^1_i( t ), \ldots, s^{a_j}_i( t ) ) \mid t \in S^\ast, C_i \in cons( t ) \}$.
 In a derivation tree with root node $\Ec$, 
@@
 \begin{proof}
-TODO
+%We show that if a derivation tree with $\Ec$ has a model $\M$, 
 \end{proof}

### 2014-12-02 tex/codatadp.tex
more

--- 2014-12-02 tex/codatadp.tex
@@
 
-  * initial algebra
+Axiomatization:
+
+discriminators go here
+
+specifies the
+
     * ``no junk, no confusion''
-    * initial algebra ``exists'' only if well-founded
-      otherwise it would be empty, which is disallowed by most formulations of
-      first-order logic (including CVC4 and other SMT solvers)
-      and higher-order logic (including Isabelle/HOL)
+
+Because the types are inhabited (nonempty), a datatype specification is
+admissible only if a ground term can be exhibited that belongs to that datatype.
+This rules out ill-founded specifications such as
+%
+\[\keyw{datatype}\;\, \ty{fstream} \,=\, \const{FSCons}(\ty{int},\,\ty{fstream})\]
+%
+For codatatypes, no such check is necessary. Given the specification
+%
+\[\keyw{codatatype}\;\, \ty{stream} \,=\, \const{SCons}(\ty{int},\,\ty{stream})\]
+%
+the infinite value $\const{SCons}(0, \const{SCons}(0, \ldots))$ witnesses
+nonemptiness.
+
+Datatypes can be defined semantically as the initial model of an equational
+specification of the selector--constructor equations. This is the view followed
+by Barrett et al.\ \cite{barrett-et-al-2010}. A disadvantage with this approach
+is that it does not naturally account for selectors applied to wrong
+constructors. Barrett et al.\ address this by parameterizing the construction by
+default values, but these give rise to spurious equalities between unrelated
+terms (e.g., $hd(nil1) \teq hd(nil2)$). This could be corrected, but the added
+compexity seems to suggest that selectors are better specified axiomatically.
+
+A related semantic view of datatypes is as initial algebra. Codatatypes are then
+defined dually as a final coalgebra. The datatypes are generated by their
+constructors, whereas the codatatypes are viewed through their selectors
+\cite{xxx}. By uniformly focusing on the constructors, the axiomatic approach
+emphasizes the commonality between datatypes and codatatypes, while sacrificing
+a theoretically fruitful notion of duality.
 
@@
 
-Bosses: Viktor Kuncak, Tobias Nipkow, and Stephan Merz.
+Bosses: Viktor Kuncak, Stephan Merz, Tobias Nipkow, and Christoph Weidenbach.
 

### 2014-12-02 tex/codatadp.tex
Refutation soundness proof (draft).

@@
 Let $S^0$ be the set of terms that are arguments of a selector application in $\tEc$.
-For $i \geq 0$, let $S^{i+1}$ $=$ $S_i \cup \{ s( t ) \mid t \in S^i, s \in sel( t ), t : \tau \text{ finite} \}$.
-Let $S^\ast$ be the fix point of this sequence, which is guarenteed to be finite 
+For $i \geq 0$, let $S^{i+1}$ $=$ $S_i \cup \{ s( t ) \mid t \in S^i, t : \tau \text{ finite}, s \in \Sel^\tau  \}$.
+Let $S^\ast$ be the fixed point of this sequence, which is guarenteed to be a finite set 
 since $S^i \setminus S^0$ only includes selectors applied to terms having finite type.
-Now, let $\tcEc$ be the set of subterms of $\Ec \cup \{ C_i( s^1_i( t ), \ldots, s^{a_j}_i( t ) ) \mid t \in S^\ast, C_i \in cons( t ) \}$.
+Now, let $\tcEc$ be the set of subterms of $\Ec \cup \{ C_i( s^1_i( t ), \ldots, s^{a_j}_i( t ) ) \mid t \in S^\ast, t : \tau, C_i \in \Ctr^\tau \}$.
 In a derivation tree with root node $\Ec$, 
@@
 which is finite since $\tcEc$ is finite.
+\qed
 \end{proof}
@@
 \begin{proof}
-%We show that if a derivation tree with $\Ec$ has a model $\M$, 
+We prove this by structural induction on the (closed) derivation tree with root node $\Ec$.
+First, if the derivation tree is an application of \rn{Conflict}, \rn{Unify$_2$}, or \rn{Cyclic},
+then $\Ec$ is unsatisfiable in $\thD$.
+\rem{Elaborate on cyclic}
+If the children of $\Ec$ are closed derivation trees 
+whose root node is the result of applying the rule \rn{Split} on term $t$ of type $\tau$,
+then by the induction hypothesis $\Ec \cup t \teq C_i( s^1_i( t ), \ldots, s^{a_i}_i( t ) )$ is unsatisfiable
+for each $C_i \in \Ctr^\tau$.
+Since all models of $\thD$ entail exactly one equality $t \teq C_i( s^1_i( t ), \ldots, s^{a_i}_i( t ) )$,
+we know $\Ec$ is unsatisfiable in $\thD$.
+Otherwise, the child of $\Ec$ is a closed derivation tree 
+whose root node is $\Ec \cup t_1 \teq t_2$ obtained by applying one of the rules \rn{Refl}, \rn{Symm}, \rn{Trans}, \rn{Cong}, \rn{Unify$_1$}, or \rn{Bisimilar}.
+In all cases, we have that $\Ec \models_{\thD} t_1 \teq t_2$.
+\rem{Elaborate on Bisimilar}
+By the induction hypothesis, we have $\Ec \cup t_1 \teq t_2$ is unsatisfiable in $\thD$, 
+and since $\Ec \models_{\thD} t_1 \teq t_2$,
+we have that $\Ec$ is unsatisfiable in $\thD$.
+\qed
 \end{proof}

### 2014-12-02 tex/codatadp.tex
more

--- 2014-12-02 tex/codatadp.tex
@@
 %\newcommand\Plainfuncs{\Funcs_{\mathrm{other}}}
+
+\newcommand\vthinspace{\kern+0.083333em}
+\newcommand\negvthinspace{\kern-0.083333em}
 
@@
 (\keyw{co})\keyw{datatype}\;\,
-  \delta_1 & {}=\, \const{C}_{1}\rlap{${}_{1}$}\phantom{{}_{m}}(\bigl[\const{s}_{111}{:}\bigr]\; \tau_{111}, \,\ldots,\, \bigl[\const{s}_{11n_{11}}{:}\bigr]\; \tau_{11n_{11}}) & \\
+  \delta_1 & {}=\, \const{C}_{1}\rlap{${}_{1}$}\phantom{{}_{m}}(\bigl[\const{s}_{11}^1{:}\bigr]\; \tau_{11}^1, \,\ldots,\, \bigl[\const{s}_{11}^{n_{11}}{:}\bigr]\; \tau_{11}^{n_{11}}) & \\
            & {}\;\phantom{=\,}\;{} \smash{\,\,\,\vdots} \\[-\jot] % {}\;\phantom{=}\llap{\ensuremath{\mid}}\; \cdots \\
-           & {}\;\phantom{=\,}\llap{\ensuremath{\,\mid\,}}\; \const{C}_{1m}(\bigl[\const{s}_{1m1}{:}\bigr]\; \tau_{1m1}, \,\ldots,\, \big[\const{s}_{1mn_{1m}}{:}\big]\; \tau_{1mn_{1m}}) \\[-1\jot]
+           & {}\;\phantom{=\,}\llap{\ensuremath{\,\mid\,}}\; \const{C}_{1m}(\bigl[\const{s}_{1m}^1{:}\bigr]\; \tau_{1m}^1, \,\ldots,\, \big[\const{s}_{1m}^{n_{1m}}{:}\big]\; \tau_{1m}^{n_{1m}}) \\[-1\jot]
    \smash{\vdots\,\,\,} \\[-1\jot]
@@
 with
-$\const{C}_{ij} : \tau_{ij1}\times\cdots\times\tau_{ijk_{ij}} \to \delta_i$
-and $\const{s}_{ijk} : \delta_i \to \tau_{ijk}$. Defaults are assumed for
+$\const{C}_{i\negvthinspace j} : \tau_{i\negvthinspace j}^1\times\cdots\times\tau_{i\negvthinspace j}^{\,k_{\smash{i\negvthinspace j}}} \to \delta_i$
+and $\const{s}_{i\negvthinspace j}^{\,k} : \delta_i \to \tau_{i\negvthinspace j}^{\,k}$. Defaults are assumed for
 the selector names if they are omitted.
@@
 \[\begin{aligned}[t]
-      \keyw{codatatype}\; \,\ty{pllist}(\alpha) & \,=\, \const{PLNil} \,\mid\, \const{PLCons}(\alpha,\, \ty{pllist}(\alpha)) \\[-.5\jot]
-      \keyw{datatype}\;\, \ty{tree} & \,=\, \const{Node}(\ty{int},\, \ty{pllist}(\ty{tree}))
+      \keyw{codatatype}\; \,\ty{gllist}(\alpha) & \,=\, \const{GLNil} \,\mid\, \const{GLCons}(\alpha,\, \ty{gllist}(\alpha)) \\[-.5\jot]
+      \keyw{datatype}\;\, \ty{tree} & \,=\, \const{Node}(\ty{int},\, \ty{gllist}(\ty{tree}))
 \end{aligned}
@@
 
+  * axiomatically
+    * injectivity
+      distinctness
+      exhaustiveness
+    * datatypes:
+      * induction axiom (schema)
+        * consequence: no infinite objects (in particular, no cyclic objects)
+
+      * when looking at the universal ground (?) theory:
+        enough to consider acyclicity -- no way to specify infinite objects
+        otherwise
+
+    * codatatypes:
+      * coinduction axiom (schema)
+        * consequence: infinite objects are allowed, but
+          two objects yielding the same observations must be equal
+
+      * again: when looking at the universal ground (?) theory, enough to
+        consider bisimilarity/bisimulation
+
 discriminators go here
@@
     * ``no junk, no confusion''
+
+  * how to deal with "wrong" selectors, e.g.
+        hd(nil1) = hd(nil2)?
+      * leave them unspecified; hence hd(nil1) = hd(nil2) in some models,
+        and not in other models
 
@@
 default values, but these give rise to spurious equalities between unrelated
-terms (e.g., $hd(nil1) \teq hd(nil2)$). This could be corrected, but the added
+terms---for example, $\const{s}_{11}(\const{C}_2) \teq \const{s}_{11}(\const{C}_3)$.
+This could be corrected, but the added
 compexity seems to suggest that selectors are better specified axiomatically.
@@
 a theoretically fruitful notion of duality.
-
-  * final coalgebra
-    * this time, no condition: always exists
-    e.g. the specification
-
-      codatatype x = X x
-
-    gives rise to a singleton type {X (X (X ...))}.
-
-  * discriminators/selectors for initial algebras
-  * constructors for final coalgebras
-
-  * Introduce the discriminators already here.
-
-  * how to deal with "wrong" selectors, e.g.
-        hd(nil1) = hd(nil2)?
-      * leave them unspecified; hence hd(nil1) = hd(nil2) in some models,
-        and not in other models
-
-  * ODDITY: Barrett et al. overspecify things (not good w.r.t. SMT-LIB 2, or
-    other solvers, e.g., Z3)
-
-  * axiomatically
-    * injectivity
-      distinctness
-      exhaustiveness
-    * datatypes:
-      * induction axiom (schema)
-        * consequence: no infinite objects (in particular, no cyclic objects)
-
-      * when looking at the universal ground (?) theory:
-        enough to consider acyclicity -- no way to specify infinite objects
-        otherwise
-
-    * codatatypes:
-      * coinduction axiom (schema)
-        * consequence: infinite objects are allowed, but
-          two objects yielding the same observations must be equal
-
-      * again: when looking at the universal ground (?) theory, enough to
-        consider bisimilarity/bisimulation
 

### 2014-12-02 tex/codatadp.tex
more

@@
 \DeclareMathAlphabet{\mathcal}{OT1}{pzc}{m}{it}
+
+\newcommand\jPrime{j\vthinspace'\negvthinspace}
 
@@
 values. %, and may contain other interpreted types (e.g., \ty{int}, \ty{real}).
-Predicates are simply functions that return $\ty{bool}$.
+The only predicate is $\teq$, and it is treated as a logical symbol,
+just like the connectives and quantifiers of first-order logic.
+Other predicates can be represented as functions that return $\ty{bool}$.
 The metavariables $\delta, \tau$ range over types, with $\delta$ reserved
-for (co)datatypes.
+for datatypes and codatatypes.
 
 Function symbols are written in a sans-serif font (e.g., $\const{f}$, $\const{g}$) to
-distinguish them from variables (e.g., $x$, $y$). With each function symbol \const{f}
+distinguish them from variables (e.g., $x$, $y$). Upper case names (e.g., $\const{C}$,
+$\const{Nil}$) are reserved for constructors. With each function symbol \const{f}
 is associated a list of argument types $\tau_1,\ldots,\tau_n$ (with $n \ge 0$)
@@
 parentheses in terms.
+%
+Finally, the notation $\bar x$ abbreviates a list or tuple $x_1,\ldots,x_n$.
 
@@
 
-Axiomatization:
+Because the types are inhabited (nonempty), a datatype specification is
+admissible only if a ground term can be exhibited that belongs to that datatype.
+This rules out ill-founded specifications such as
+%
+\[\keyw{datatype}\;\, \ty{fstream} \,=\, \const{FSCons}(\ty{int},\,\ty{fstream})\]
+%
+For codatatypes, no such check is necessary. Given the specification
+%
+\[\keyw{codatatype}\;\, \ty{stream} \,=\, \const{SCons}(\ty{int},\,\ty{stream})\]
+%
+the infinite value $\const{SCons}(0, \const{SCons}(0, \ldots))$ witnesses
+nonemptiness.
 
-  * axiomatically
-    * injectivity
-      distinctness
-      exhaustiveness
+Datatypes and codatatypes can be characterized axiomatically. They share basic
+properties of constructors and selectors. All properties below range over all
+$i$, $j$, $\jPrime$, and $k$ within bounds and to all possible splits of the
+$n$-ary constructor $\const{C}_{i\negvthinspace j}$'s argument list into $\bar x,
+y, \bar z$:
+%
+\[
+\begin{aligned}[t]
+\text{Distinctness:}\quad
+  & \forall \bar x, \bar y.\;\, \const{C}_{i\negvthinspace j}(\bar x) \tneq \const{C}_{i\negvthinspace \jPrime}(\bar y) \qquad\text{if $j \not= \jPrime$}
+  \\
+\text{Injectivity:}\quad
+  & \forall \bar x, y, \bar z.\;\, \const{C}_{i\negvthinspace j}(\bar x, y, \bar z) \teq \const{C}_{i\negvthinspace j}(\bar x, y', \bar z) \longrightarrow y \teq y'
+  \\
+\text{Exhaustiveness:}\quad
+  &
+  \\
+\text{Selection:}\quad
+  &
+\end{aligned}
+\]
+
     * datatypes:
@@
 
-Because the types are inhabited (nonempty), a datatype specification is
-admissible only if a ground term can be exhibited that belongs to that datatype.
-This rules out ill-founded specifications such as
-%
-\[\keyw{datatype}\;\, \ty{fstream} \,=\, \const{FSCons}(\ty{int},\,\ty{fstream})\]
-%
-For codatatypes, no such check is necessary. Given the specification
-%
-\[\keyw{codatatype}\;\, \ty{stream} \,=\, \const{SCons}(\ty{int},\,\ty{stream})\]
-%
-the infinite value $\const{SCons}(0, \const{SCons}(0, \ldots))$ witnesses
-nonemptiness.
-
 Datatypes can be defined semantically as the initial model of an equational
@@
 This could be corrected, but the added
-compexity seems to suggest that selectors are better specified axiomatically.
+complexity seems to suggest that selectors are better specified axiomatically.
 

### 2014-12-02 tex/codatadp.tex
Restate cyclicity/bisimilarity rules, cleanup.

@@
 \inferrule{
-  t : \tau, \tau \in \Data
+  t : \tau
   \quad
-  p \in \ttpath{\Ec}{\ec{t}}{\ec{t}}
+  \tau \in \Data
+  \quad
+  \ttpath{\Ec}{\ec{t}}{\ec{t}} \neq \emptyset
 }{
@@
 \inferrule{
- t_1, t_2 : \tau, \tau \in \Codata
+ t_1, t_2 : \tau
   \quad
- \exists \sigma. p \sigma \in \tpath{\Ec}{\ec{t_1}} \text{ for each } p \in \tpath{\Ec}{\ec{t_2}}
+ \tau \in \Codata
+  \quad
+ \tpath{\Ec}{\ec{t_1}} = \tpath{\Ec}{\ec{t_2}}\sigma \neq \emptyset
 }{
@@
 We call the \emph{source} of a path is the source of its first edge, and the \emph{destination} of a path is the destination of its last edge.
-The set $\ttpath{\Ec}{\ec{t_1}}{\ec{t_2}}$ is the smallest set containing all paths with source $\ec{t_1}$ and destination $\ec{t_2}$ that are induced by $\Ec$.
+We say a path $p$ is \emph{maximal} if its destination $\ec{t}$ is either
+the source of an edge in $p$ or does not contain an application of a constructor (including the case when $t$ is not of datatype/codatatype type).
+The set $\ttpath{\Ec}{\ec{t_1}}{\ec{t_2}}$ is the smallest set containing all maximal paths with source $\ec{t_1}$ and destination $\ec{t_2}$ that are induced by $\Ec$.
 The set $\tpath{\Ec}{\ec{t_1}}$ is the union of $\ttpath{\Ec}{\ec{t_1}}{\ec{t_2}}$ for all $t_2 \in \tEc$.
@@
 The rule \rn{Cyclic} recognizes when a path exists from an equivalence class $\ec{t}$ to itself, as indicated
-by the assumption $p \in \ttpath{\Ec}{\ec{t}}{\ec{t}}$.
+by the assumption $\ttpath{\Ec}{\ec{t}}{\ec{t}} \neq \emptyset$.
 In this case, $\Ec$ entails that $t$ contains itself as a (strict) subterm,
@@
 We demonstrate this rule concretely with the following example.
+\rem{ensure that \const{LCons} understood}
 
 \begin{example}
-Let $\Ec$ be the set $\{ x \teq \const{LCons}(y,\const{LCons}(y,x)), x \tneq \const{LCons}(y,x) \}$.  \rem{ensure that \const{LCons} understood}
-The equivalence classes of $\Ec$ are $\{ x, \const{LCons}(y,\const{LCons}(y,x)) \}$, $\{ \const{LCons}(y,x) \}$, and $\{ y \}$.
+Let $\Ec$ be the set $\{ x \teq \const{LCons}(y,\const{LCons}(y,x)) \}$.
+After taking the bidirectional closure of $\Ec$, we obtain equivalence classes 
+$\{ x, \const{LCons}(y,\const{LCons}(y,x)) \}$, $\{ \const{LCons}(y,x) \}$, and $\{ y \}$.
 The set $\tpath{\Ec}{\ec{x}}$ is:
@@
 \end{array}\]
-Thus, we may conclude $x \teq \const{LCons}(y,x)$ by the rule \rn{Bisimilar}
-for the mapping $\sigma :=$ $\{ \ec{x} \mapsto \ec{\const{LCons}(y,x)}, \ec{\const{LCons}(y,x)} \mapsto \ec{x} \}$,
-after which the rule \rn{Conflict} may be applied.
+Thus, we may infer $x \teq \const{LCons}(y,x)$ by the rule \rn{Bisimilar}
+for the mapping $\sigma :=$ $\{ \ec{x} \mapsto \ec{\const{LCons}(y,x)}, \ec{\const{LCons}(y,x)} \mapsto \ec{x} \}$.
+\qed
 \end{example}
@@
 since $S^i \setminus S^0$ only includes selectors applied to terms having finite type.
-Now, let $\tcEc$ be the set of subterms of $\Ec \cup \{ C_i( s^1_i( t ), \ldots, s^{a_j}_i( t ) ) \mid t \in S^\ast, t : \tau, C_i \in \Ctr^\tau \}$.
+Now, let $\tcEc$ be the set of subterms of $\Ec \cup \{ C_i( s^1_i( t ), \ldots, s^{a_i}_i( t ) ) \mid t \in S^\ast, t : \tau, C_i \in \Ctr^\tau \}$.
 In a derivation tree with root node $\Ec$, 
 it can be shown by case analysis on the rules of the calculus that each non-root node is 
-obtained as a result of adding an equality between two terms from $\tcEc$ to its parent node.
-Thus, the maximum depth of a branch in a derivation tree with root node $\Ec$ is $\mid \tcEc \mid^2$,
+the result of adding an equality between two terms from $\tcEc$ to its parent node.
+Thus, the depth of a branch in a derivation tree with root node $\Ec$ is at most $\mid \tcEc \mid^2$,
 which is finite since $\tcEc$ is finite.
@@
 By the induction hypothesis, we have $\Ec \cup t_1 \teq t_2$ is unsatisfiable in $\thD$, 
-and since $\Ec \models_{\thD} t_1 \teq t_2$,
-we have that $\Ec$ is unsatisfiable in $\thD$.
+and thus $\Ec$ is unsatisfiable in $\thD$.
 \qed

### 2014-12-03 tex/codatadp.tex
minor

--- 2014-12-02 tex/codatadp.tex
@@
 \newcommand{\rem}[1]{\textcolor{red}{[#1]}}
-\newcommand{\tester}[1]{is\text{-} #1}
+%\newcommand{\is}[1]{is\text{-} #1}
+\newcommand{\is}[1]{\const{is#1}}
 \newcommand{\ror}{\quad \parallel \quad}
@@
 
+discriminators go here
+
 Datatypes and codatatypes can be characterized axiomatically. They share basic
-properties of constructors and selectors. All properties below range over all
-$i$, $j$, $\jPrime$, and $k$ within bounds and to all possible splits of the
-$n$-ary constructor $\const{C}_{i\negvthinspace j}$'s argument list into $\bar x,
-y, \bar z$:
+properties of constructors and selectors. All properties below are implicitly
+universally quantified and range over all
+$i$, $j$, $\jPrime$, and $k$ within bounds:
+% and to all possible splits of the
+%$n$-ary constructor $\const{C}_{i\negvthinspace j}$'s argument list into $\bar x,
+%y, \bar z$:
 %
@@
 \text{Distinctness:}\quad
-  & \forall \bar x, \bar y.\;\, \const{C}_{i\negvthinspace j}(\bar x) \tneq \const{C}_{i\negvthinspace \jPrime}(\bar y) \qquad\text{if $j \not= \jPrime$}
+  & %\forall \bar x, \bar y.\;\,
+    \const{C}_{i\negvthinspace j}(\bar x) \tneq \const{C}_{i\negvthinspace \jPrime}(\bar y) \qquad\text{if $j \not= \jPrime$}
   \\
 \text{Injectivity:}\quad
-  & \forall \bar x, y, \bar z.\;\, \const{C}_{i\negvthinspace j}(\bar x, y, \bar z) \teq \const{C}_{i\negvthinspace j}(\bar x, y', \bar z) \longrightarrow y \teq y'
+  & %\forall x_1,\ldots,x_{n_{i\negvthinspace j}} y.\;\,
+    \const{C}_{i\negvthinspace j}(x_1,\ldots,x_{n_{i\negvthinspace j}}) \teq \const{C}_{i\negvthinspace j}(x_1,\ldots,x_{k-1},y,x_{k+1},\ldots,x_{n_{i\negvthinspace j}}) \longrightarrow x_k \teq y
   \\
 \text{Exhaustiveness:}\quad
-  &
+  & \is{C}_{i1}(x) \mathrel\lor \cdots \mathrel\lor \is{C}_{im_i}(x)
   \\
 \text{Selection:}\quad
-  &
+  & s_{ij}^k(\const{C}_{i\negvthinspace j}(x_1,\ldots,x_{n_{i\negvthinspace j}})) = x_k
 \end{aligned}
 \]
+%
+Expressed in the algebraic jargon,
+exhaustiveness ensures that ``no junk'' arises, whereas
+distinctness and injectivity ensure that ``no confusion'' occur,
+whereas
+
+%Sometimes, the requirements on constructors are stipulated as ``no junk, no
+%confusion.''
 
@@
 
-discriminators go here
-
 specifies the
-
-    * ``no junk, no confusion''
 

### 2014-12-04 tex/codatadp.tex
more ideas

@@
       datatypes but:
+    * infinite values (infinitely many nested constructors)
     * codatatypes are never empty (e.g. finite streams are rejected)

### 2014-12-04 tex/codatadp.tex
Add singleton rule.

--- 2014-12-03 tex/codatadp.tex
@@
 \inferrule{
-  t \teq u, t \tneq u \in \Ec
+  t_1 \teq t_2, t_1 \tneq t_2 \in \Ec
 }{
@@
 
-Figure~\ref{fig:cc-rules} gives the rules for constructing the bidirectional closure of $\Ec$.
-The rules \rn{Refl}, \rn{Symm}, \rn{Trans}, and \rn{Cong} together compute the (upwards) congruence closure,
-after which a conflict may be recognized by \rn{Conflict} when an inconsistency is found.
-The rules \rn{Unify$_1$} and \rn{Unify$_2$} together compute (downwards) unification,
-where additional equalities are inferred based on the injectivity of constructors by \rn{Unify$_1$},
-and failures to unify equivalent terms are recognized by \rn{Unify$_2$}.
+Figure~\ref{fig:cc-rules} gives the basic rules of our calculus.
+Together with the rules \rn{Refl}, \rn{Symm}, \rn{Trans}, the rule \rn{Cong} computes the (upwards) congruence closure,
+while the rules \rn{Unify$_1$} and \rn{Unify$_2$} together compute (downwards) unification.
+In the unification rules, additional equalities are inferred based on the injectivity of constructors by \rn{Unify$_1$},
+and failures to unify equated terms are recognized by \rn{Unify$_2$}.
+The rule \rn{Conflict} recognizes when an equality and its negation both occur in $\Ec$, in which case $\Ec$ has no model.
 
@@
 We say a path $p$ is \emph{maximal} if its destination $\ec{t}$ is either
-the source of an edge in $p$ or does not contain an application of a constructor (including the case when $t$ is not of datatype/codatatype type).
+the source of an edge in $p$, 
+or does not contain an application of a constructor (including the case when $t$ is not of datatype/codatatype type).
 The set $\ttpath{\Ec}{\ec{t_1}}{\ec{t_2}}$ is the smallest set containing all maximal paths with source $\ec{t_1}$ and destination $\ec{t_2}$ that are induced by $\Ec$.
@@
 \begin{tabular}{c}
+\rn{Singleton}
+\(
+\inferrule{
+  t, u \in \tEc
+  \quad
+  t, u : \tau,
+  \tau \in \Codata,
+  \mid \tau \mid = 1
+}{
+  \Ec := \Ec, t \teq u
+}
+\)
+\\[3.7ex]
 \rn{Split} 
@@
 \inferrule{
-  s^i_j( t ) \in \tEc, \text{ or } t : \tau \text{ finite }
+  s^i_j( t ) \in \tEc \quad \text{ or } \quad t : \tau, \tau \in \Data, \tau \text{ finite }
 }{
@@
 \end{tabular}
-\caption{Rule for splitting.  We split on the constructor for $t$ if there is a selector applied to $t$ or if its type is finite.
+\caption{Rules for splitting.  
+%All pairs of terms whose type has cardinality 1 are entailed to be equal (the degenerate case).
+%Constructors must be assigned for all terms $t$ if has a selector is applied to it, or if $t$ has finite datatype type.
 }
@@
 
+\rem{Require : all co-datatypes either have infinite cardinality, or have cardinality 1.}
+
 Finally, if neither a rule from Figure~\ref{fig:cc-rules} or Figure~\ref{fig:ab-rules} applies to $\Ec$, 
-we split on the constructor type for certain terms $t$.
-In particular, if there is a selector term applied to $t$, or if $t$ has a finite type, 
+the rules from Figure~\ref{fig:split-rule} are applied.
+The rule \rn{Singleton} handles the degenerate case when there are two terms from $\tEc$ having a singleton codatatype type,
+in which case we can infer they are equal.
+Otherwise, in the case there is a selector term applied to $t$, or if $t$ has a finite datatype type, 
 then we require the equivalence class of $t$ to contain a constructor term, which is enforced by the rule \rn{Split}.
@@
 Let $S^0$ be the set of terms that are arguments of a selector application in $\tEc$.
-For $i \geq 0$, let $S^{i+1}$ $=$ $S_i \cup \{ s( t ) \mid t \in S^i, t : \tau \text{ finite}, s \in \Sel^\tau  \}$.
+For $i \geq 0$, let $S^{i+1}$ $=$ $S_i \cup \{ s( t ) \mid t \in S^i, t : \tau, \tau \in \Data, \tau \in  \text{ finite}, s \in \Sel^\tau  \}$.
 Let $S^\ast$ be the fixed point of this sequence, which is guarenteed to be a finite set 
-since $S^i \setminus S^0$ only includes selectors applied to terms having finite type.
+since $S^i \setminus S^0$ only includes selectors applied to terms having finite datatype type.
 Now, let $\tcEc$ be the set of subterms of $\Ec \cup \{ C_i( s^1_i( t ), \ldots, s^{a_i}_i( t ) ) \mid t \in S^\ast, t : \tau, C_i \in \Ctr^\tau \}$.
@@
 then $\Ec$ is unsatisfiable in $\thD$.
-\rem{Elaborate on cyclic}
+\rem{Elaborate on cyclic?}
 If the children of $\Ec$ are closed derivation trees 
@@
 Otherwise, the child of $\Ec$ is a closed derivation tree 
-whose root node is $\Ec \cup t_1 \teq t_2$ obtained by applying one of the rules \rn{Refl}, \rn{Symm}, \rn{Trans}, \rn{Cong}, \rn{Unify$_1$}, or \rn{Bisimilar}.
+whose root node is $\Ec \cup t_1 \teq t_2$ obtained by applying one of the rules \rn{Refl}, \rn{Symm}, \rn{Trans}, \rn{Cong}, \rn{Unify$_1$}, \rn{Bisimilar}, or \rn{Singleton}.
 In all cases, we have that $\Ec \models_{\thD} t_1 \teq t_2$.
-\rem{Elaborate on Bisimilar}
+\rem{Elaborate on Bisimilar?}
 By the induction hypothesis, we have $\Ec \cup t_1 \teq t_2$ is unsatisfiable in $\thD$, 

### 2014-12-04 tex/codatadp.tex
copied stuff into Sect. 2

@@
 
+    semantically, $k$ types are mutually (co)recursive if their dependency graph
+    (x -> y if x is the type of an argument to one of y's constructors) is
+    strongly connected
+
+    (co)datatypes can be declared together as mutually (co)recursive even if
+    they are not actually (co)recursive
+
+    the semantic notion is more fruitful and the one we will focus on in the
+    sequel
+
+    Also, nothing forbids specifications with no (co)recursion, e.g.
+
+      datatype enat = Nat(nat) | Infty
+      codatatype complex = Complex(re: real, im: real)
+
+    In this case, it makes no difference whether the type is introduced as a
+    datatype or as a codatatype. Without loss of generality, we consider that
+    these types are datatypes instead of codatatypes.
+
+Degenerate Codatatypes
+
+    Consider
+
+      codatatype a = A(a)
+
+    This is a finite codatatype, with a single value, which is infinite:
+    $\mu a. A(a)$
+
+    We need to detect those finite codatatypes. They are easy to
+    characterize semantically.
+
+    Lemma D.1. The only possible finite cardinality for the domain interpreting
+    a (corecursive) codatatype is 1.
+
+    Proof.
+      Assume you have at least 2 different values of a codatatype.
+      They differ at one point.
+      * 
+    Qed
+
+    An immediate consequence by the distinctness axiom is that these types have
+    necessarly one constructor, and that all arguments are of types of
+    cardinality 1.
+
+    Easy case: $k$ mutually recursive codatatypes, each equipped with one
+    constructor, all of which take each other in arguments in various
+    combinations, with potential other types that are known to have cardinality
+    1. e.g.
+
+      codatatype b = B b c b
+             and c = C c b c
+
+    Medium case: like above, but some of the arguments are other types not
+    participating in the corecursion that are known to have cardinality 1, e.g.
+
+      codatatype d = D unit d d unit
+
+    where "unit" is a datatype with the single constructor "Unity : unit".
+
+    Though case: like above, but with any other types where it's not clear
+    whether it could have cardinality one or not.
+
+      type e
+      codatatype f = F e f f unit
+
 \section{The Calculus}

### 2014-12-04 tex/codatadp.tex
Improve termination proof.

@@
 \inferrule{
-  t, u \in \tEc
+  t, u : \tau \in \tEc
   \quad
-  t, u : \tau,
   \tau \in \Codata,
@@
 \inferrule{
-  s^i_j( t ) \in \tEc \quad \text{ or } \quad t : \tau, \tau \in \Data, \tau \text{ finite }
+  t : \tau \in \tEc 
+  \quad 
+  ( s( t ) \in \tEc, s \in \Sel^\tau ) 
+  \text{ or } 
+  ( \tau \in \Data, \tau \text{ finite } )
 }{
@@
 The rule \rn{Singleton} handles the degenerate case when there are two terms from $\tEc$ having a singleton codatatype type,
-in which case we can infer they are equal.
+in which case they must be equivalent.
 Otherwise, in the case there is a selector term applied to $t$, or if $t$ has a finite datatype type, 
-then we require the equivalence class of $t$ to contain a constructor term, which is enforced by the rule \rn{Split}.
+then the equivalence class of $t$ must contain an application of one of the constructors of $\tau$ (where $\Ctr^\tau = \{ C_1, \ldots C_n \}$), 
+which is enforced by the rule \rn{Split}.
 
@@
 Consider a derivation tree with root node $\Ec$.
-Let $S^0$ be the set of terms that are arguments of a selector application in $\tEc$.
-For $i \geq 0$, let $S^{i+1}$ $=$ $S_i \cup \{ s( t ) \mid t \in S^i, t : \tau, \tau \in \Data, \tau \in  \text{ finite}, s \in \Sel^\tau  \}$.
-Let $S^\ast$ be the fixed point of this sequence, which is guarenteed to be a finite set 
-since $S^i \setminus S^0$ only includes selectors applied to terms having finite datatype type.
-Now, let $\tcEc$ be the set of subterms of $\Ec \cup \{ C_i( s^1_i( t ), \ldots, s^{a_i}_i( t ) ) \mid t \in S^\ast, t : \tau, C_i \in \Ctr^\tau \}$.
+Let $S$ be the set of terms occurring as the argument of a selector in $\tEc$.
+For each term $t \in \tEc$ of finite datatype type $\tau$,
+let $S^0_t$ be the set $\{ t \}$,
+and for $i \geq 0$ let $S^{i+1}_t =$ $S^i_t \cup \{ s( u ) \mid u : \upsilon \in S^i, \upsilon \in \Data, \upsilon \text{ finite}, s \in \Sel^{\upsilon}  \}$.
+Let $S^\ast_t$ be the fixed point of this sequence for all $t$, which we know is finite since the values of $\tau$ are of finite size.
+Let $S^\ast$ be the union of $S^\ast_t$ for all such $t$,
+and let $\tcEc$ be the set of subterms of $\Ec \cup \{ C_i( s^1_i( t ), \ldots, s^{a_i}_i( t ) ) \mid t : \tau \in S \cup S^\ast, C_i \in \Ctr^\tau \}$.
 In a derivation tree with root node $\Ec$, 
-it can be shown by case analysis on the rules of the calculus that each non-root node is 
-the result of adding an equality between two terms from $\tcEc$ to its parent node.
+it can be shown by case analysis on the rules of the calculus that each non-root node $\Ec'$ is such that 
+$\mathcal{T}(\Ec') \subseteq \tcEc $, and moreover thus is the result of adding an equality between two terms from $\tcEc$ to its parent node.
 Thus, the depth of a branch in a derivation tree with root node $\Ec$ is at most $\mid \tcEc \mid^2$,
@@
 for each $C_i \in \Ctr^\tau$.
-Since all models of $\thD$ entail exactly one equality $t \teq C_i( s^1_i( t ), \ldots, s^{a_i}_i( t ) )$,
+Since all models of $\thD$ entail (exactly) one equality $t \teq C_i( s^1_i( t ), \ldots, s^{a_i}_i( t ) )$,
 we know $\Ec$ is unsatisfiable in $\thD$.

### 2014-12-04 tex/codatadp.tex
more

--- 2014-12-04 tex/codatadp.tex
@@
 \newcommand\Codata{\Types_{\mathrm{cdt}}}
-\newcommand\Nondata{\Types_{\mathrm{oth}}}
+\newcommand\Nondata{\Types_{\mathrm{pln}}}
 
@@
 values. %, and may contain other interpreted types (e.g., \ty{int}, \ty{real}).
-The only predicate is $\teq$, and it is treated as a logical symbol,
+The only predicate is equality ($\teq$), and it is treated as a logical symbol,
 just like the connectives and quantifiers of first-order logic.
@@
 
-We assume a signature $\Sig = (\Types, \Funcs)$. The types are partioned into
+We fix a signature $\Sig = (\Types, \Funcs)$. The types are partioned into
 $\Types = \Data \mathrel{\uplus} \Codata \mathrel{\uplus} \Nondata$, where $\Data$ are the
-datatypes, $\Codata$ are the codatatypes, and $\Nondata$ are the remaining
-types. The functions are partitioned into $\Funcs = \Ctr \mathrel{\uplus} \Sel$, where
+datatypes, $\Codata$ are the codatatypes, and $\Nondata$ are the remaining,
+plain types. The functions are partitioned into $\Funcs = \Ctr \mathrel{\uplus} \Sel$, where
 $\Ctr$ are the constructors and $\Sel$ are the selectors. (There is no need to
@@
 
-In practice (e.g., in an SMT problem), the signature is given by specifying the
-uninterpreted types first (in any order), then the (co)datatypes with their
-constructors and selectors in groups of mutually (co)recursive groups of
-(co)datatypes, and finally any other function symbols.
+In an SMT problem, the signature is given by specifying the uninterpreted types
+first (in any order), then the (co)datatypes with their constructors and
+selectors in groups of mutually (co)recursive groups of (co)datatypes, and
+finally any other function symbols.
 
@@
 introduced ones.%
-\footnote{In practice, it can be useful to specify the same selector
-for several constructors of the same (co)datatype,
-as long as the argument types coincide \cite[Section~3]{blanchette-et-al-2014-codata}.
-However, this is not allowed by SMT-LIB 2.}
+\footnote{It can be convenient to specify the same selector
+for several constructors associated with the same (co)data\-type,
+as long as the argument types coincide. % \cite[Section~3]{blanchette-et-al-2014-codata}.
+However, this is disallowed by SMT-LIB 2, so we do not consider it here.}
 Schematically, using a syntax similar to Standard ML, Haskell,
@@
 
+For types with several constructors, it is customary to provide discriminators
+$\const{d} : \const{s}_{i\negvthinspace j} : \delta_i \to \ty{bool}$. However,
+they are not strictly necessary:
+To simplify the presentation of the decision procedure, we treat $\const{d}_{i\negvthinspace
+j}(x)$ as an abbreviation for
+$x = \const{C}_{i\negvthinspace j}(\const{s}_{i\negvthinspace j}^1(x), \ldots, \const{s}_{i\negvthinspace j}^{\,n_{\smash{{i\negvthinspace j}}}}(x))$
+instead of extending the signature.
+
 Polymorphic types, nested (co)recursion, and datatype--codatatype mixtures fall
@@
 is beyond the fragment considered here. In principle, rank-1 (top-level)
-polymorphism would be easy to support; nesting datatypes inside datatypes,
-and likewise for codatatypes, can be reduced to the mutual case. So the only
-genuinely interesting cases missing are mixed nested (co)recursion
-(as in the \ty{tree} example above) as well as (co)recursion through a
-non-(co)datatype (both of which make sense \cite{blanchette-et-al-2014-codata}).
+polymorphism would be easy to support; nesting datatypes inside datatypes, and
+likewise for codatatypes, can be reduced to the mutual case
+\cite{gunter-1993-not}. So the only genuinely interesting cases missing are
+mixed nested (co)recursion (as in the \ty{tree} example above) as well as
+(co)recursion through a non-(co)datatype (both of which make sense
+\cite{blanchette-et-al-2014-codata}).
 
@@
 nonemptiness.
-
-discriminators go here
 
@@
 %
-Expressed in the algebraic jargon,
-exhaustiveness ensures that ``no junk'' arises, whereas
-distinctness and injectivity ensure that ``no confusion'' occur,
-whereas
+Expressed in the algebraic jargon, exhaustiveness ensures that ``no junk''
+exists, whereas distinctness and injectivity ensure that ``no confusion''
+arises.
 
-%Sometimes, the requirements on constructors are stipulated as ``no junk, no
-%confusion.''
+Datatypes are additionally characterized by an induction axiom schema:
+%
+\[
+\begin{aligned}[t]
+\text{Induction:}\quad
+for each i, j: (\forall \bar x. IH_{ij}[\bar x] --> P_i(C_{ij}(\bar x)))
+---------------------------
+P_1[x_1] \& ... \& P_l[x_l]
+\end{aligned}
+\]
+%
+Ensures that we get the smallest model. E.g. for natural numbers, this rules out
+nonstandard models, which could contain cyclic elements (e.g., n such that
+n = Suc n) or even infinite acyclic elements $Suc (Suc \ldots)$.
+
+For codatatype, the dual notion is that of a coinduction principle that
+depends on witnesses $R$ that are required to be bisimulations:
+%
+\[
+\begin{aligned}[t]
+\text{Conduction:}\quad
+for each i: R_i[x_i, y_i]
+for each i: (\forall x y. R_i[x, y] --> \bigwedge_k s_{ij}^k(x) \sim s_{ij}^k(y))
+---------------------------------
+x_1 \teq y_1 & ... & x_l \teq y_l
+\end{aligned}
+\]
+
+For codatatypes
 
@@
   * Implementation
+    * mention that discriminators are their own function symbols
     * what is an SMT theory solver

### 2014-12-04 tex/codatadp.tex
more

@@
 \usepackage[all]{xy}
+
+% for "bussproofs" package
+\EnableBpAbbreviations
+\def\ScoreOverhang{1.5pt}
+\def\proofSkipAmount{\vskip 0pt}
+\def\defaultHypSeparation{\hskip0.75em}
 
@@
 
+The proofs are part of a report.
+
 \paragraph{Related Work.}
@@
 distinguished type \ty{bool} interpreted as the set of truth
-values. %, and may contain other interpreted types (e.g., \ty{int}, \ty{real}).
+values $\{\bot, \top\}$. %, and may contain other interpreted types (e.g., \ty{int}, \ty{real}).
 The only predicate is equality ($\teq$), and it is treated as a logical symbol,
@@
 \text{Induction:}\quad
-for each i, j: (\forall \bar x. IH_{ij}[\bar x] --> P_i(C_{ij}(\bar x)))
----------------------------
-P_1[x_1] \& ... \& P_l[x_l]
+\AXC{$\forall x_1 \ldots \vthinspace x_{n_{ij}}.\; \bigl(\bigwedge_{\,k\,} \mathit{IH}_{ij}^k[\bar x_k]\bigr) \longrightarrow P_i[C_{ij}(\bar x)]$\quad\text{for each $i,j$}}
+\UIC{$P_1[v_1] \mathrel\land \cdots \mathrel\land P_l[v_l]$}
+\DP
 \end{aligned}
 \]
+where the induction hypothesis $\mathit{IH}_{ij}^k(x)$
+denotes either $P_{i'}(x)$ for some $i$ such that this is type
+correct or else $\top$.
 %
@@
 \text{Conduction:}\quad
-for each i: R_i[x_i, y_i]
-for each i: (\forall x y. R_i[x, y] --> \bigwedge_k s_{ij}^k(x) \sim s_{ij}^k(y))
----------------------------------
-x_1 \teq y_1 & ... & x_l \teq y_l
+\AXC{for each i: $R_i[x_i, y_i]$}
+\AXC{for each i: $(\forall x y. R_i[x, y] \longrightarrow \bigwedge_k s_{ij}^k(x) \sim s_{ij}^k(y))$}
+\BIC{$x_1 \teq y_1 \& ... \& x_l \teq y_l$}
+\DP
 \end{aligned}

### 2014-12-04 tex/codatadp.tex
more

@@
 
+\newcommand\iPrime{i\vthinspace'\negvthinspace}
 \newcommand\jPrime{j\vthinspace'\negvthinspace}
@@
 where the induction hypothesis $\mathit{IH}_{ij}^k(x)$
-denotes either $P_{i'}(x)$ for some $i$ such that this is type
+denotes either $P_{\iPrime}(x)$ if there exists some $i$ such that this is type
 correct or else $\top$.
 %
-Ensures that we get the smallest model. E.g. for natural numbers, this rules out
-nonstandard models, which could contain cyclic elements (e.g., n such that
-n = Suc n) or even infinite acyclic elements $Suc (Suc \ldots)$.
+The axiom schema ensures that we get a smallest model. For example,
+for a datatype of natural numbers constructed from $\const{Z}$ and $\const{S}$,
+this rules out nonstandard models, which could contain cyclic values (e.g.,
+an $n$ such that $n \teq \const{S}(n)$) or even infinite acyclic values
+$\const{S}(\const{S}(\ldots))$.
 

### 2014-12-04 tex/codatadp.tex
more

@@
 \text{Selection:}\quad
-  & s_{ij}^k(\const{C}_{i\negvthinspace j}(x_1,\ldots,x_{n_{i\negvthinspace j}})) = x_k
+  & s_{i\negvthinspace j}^k(\const{C}_{i\negvthinspace j}(x_1,\ldots,x_{n_{i\negvthinspace j}})) = x_k
 \end{aligned}
@@
 \text{Induction:}\quad
-\AXC{$\forall x_1 \ldots \vthinspace x_{n_{ij}}.\; \bigl(\bigwedge_{\,k\,} \mathit{IH}_{ij}^k[\bar x_k]\bigr) \longrightarrow P_i[C_{ij}(\bar x)]$\quad\text{for each $i,j$}}
+\AXC{$\forall x_1 \ldots \vthinspace x_{n_{i\negvthinspace j}}.\; \bigl(\bigwedge_{\,k\,} \mathit{IH}_{i\negvthinspace j}^k[\bar x_k]\bigr) \longrightarrow P_i[C_{i\negvthinspace j}(\bar x)]$\quad\text{for each $i,j$}}
 \UIC{$P_1[v_1] \mathrel\land \cdots \mathrel\land P_l[v_l]$}
@@
 \]
-where the induction hypothesis $\mathit{IH}_{ij}^k(x)$
+where the induction hypothesis $\mathit{IH}_{i\negvthinspace j}^k(x)$
 denotes either $P_{\iPrime}(x)$ if there exists some $i$ such that this is type
@@
 \AXC{for each i: $R_i[x_i, y_i]$}
-\AXC{for each i: $(\forall x y. R_i[x, y] \longrightarrow \bigwedge_k s_{ij}^k(x) \sim s_{ij}^k(y))$}
+\AXC{for each i: $(\forall x y. R_i[x, y] \longrightarrow \bigwedge_k s_{i\negvthinspace j}^k(x) \sim s_{i\negvthinspace j}^k(y))$}
 \BIC{$x_1 \teq y_1 \& ... \& x_l \teq y_l$}
@@
     * what is an SMT theory solver
+    * sharing is caring (and other "refinements" on calculus)
     * cooperation with Nelson-Oppen / integration into SMT solver

### 2014-12-04 tex/codatadp.tex
more
git push

@@
 %
-Finally, the notation $\bar x$ abbreviates a list or tuple $x_1,\ldots,x_n$.
+Finally, the notation $\bar x$ abbreviates a list or tuple $x_1,\ldots,x_n$,
+and $\bigwedge_i \varphi_i$ abbeviates a conjunction $\varphi_1 \mathrel\land \cdots \mathrel\land \varphi_n$.
 
@@
   & %\forall \bar x, \bar y.\;\,
-    \const{C}_{i\negvthinspace j}(\bar x) \tneq \const{C}_{i\negvthinspace \jPrime}(\bar y) \qquad\text{if $j \not= \jPrime$}
+    \const{C}_{i\negvthinspace j}(\bar x) \tneq \const{C}_{i\negvthinspace \jPrime}(\bar y) \quad\text{if $j \not= \jPrime$}
   \\
@@
 \text{Induction:}\quad
-\AXC{$\forall x_1 \ldots \vthinspace x_{n_{i\negvthinspace j}}.\; \bigl(\bigwedge_{\,k\,} \mathit{IH}_{i\negvthinspace j}^k[\bar x_k]\bigr) \longrightarrow P_i[C_{i\negvthinspace j}(\bar x)]$\quad\text{for each $i,j$}}
-\UIC{$P_1[v_1] \mathrel\land \cdots \mathrel\land P_l[v_l]$}
+\AXC{\strut$\bigwedge_{i,j} \forall x_1 \ldots \vthinspace x_{n_{i\negvthinspace j}}.\; \bigl(\bigwedge_{\,k\,} \mathit{IH}_{i\negvthinspace j}^k[x_k]\bigr) \longrightarrow P_i[C_{i\negvthinspace j}(x_1,\ldots,x_{n_{ij}})]$}
+\UIC{\strut$\bigwedge_{i} P_i[v_i]$}
 \DP
@@
 \begin{aligned}[t]
-\text{Conduction:}\quad
-\AXC{for each i: $R_i[x_i, y_i]$}
-\AXC{for each i: $(\forall x y. R_i[x, y] \longrightarrow \bigwedge_k s_{i\negvthinspace j}^k(x) \sim s_{i\negvthinspace j}^k(y))$}
-\BIC{$x_1 \teq y_1 \& ... \& x_l \teq y_l$}
+\text{Coinduction:}\quad
+\AXC{\strut$\begin{gathered}\textstyle \bigwedge_i R_i[v_i, w_i] \\[-\jot]\textstyle
+\bigwedge_i \forall v\; w.\;\, R_i[v, w] \longrightarrow
+  \bigwedge_j \const{d}_j(v) \teq \const{d}_j(w)
+  \mathrel\land
+  \const{d}_j(v) \longrightarrow \bigwedge_k \const{s}_{i\negvthinspace j}^{\,k}(v) \sim \const{s}_{i\negvthinspace j}^{\,k}(w)
+\end{gathered}$}
+\UIC{\strut$\bigwedge_i v_i \teq w_i$}
 \DP
@@
 \]
+where
+
+This axiom schema may look intimidating, but what it
 
@@
     * sharing is caring (and other "refinements" on calculus)
+    * theory combination more complicated (and Theorem Completeness too weak)
+      * need something like the string paper
     * cooperation with Nelson-Oppen / integration into SMT solver

### 2014-12-05 tex/codatadp.tex
more

@@
 \]
-where
+where $x \sim y$ denotes either $R_i[x, y]$ if there exists some $i$ such that
+this is type correct or $x \teq y$ otherwise.
 
-This axiom schema may look intimidating, but what it
+FIXME: Find out what ensures that codatatypes have infinite values.
 
-For codatatypes
+This axiom schema captures the idea that two values that yield the same
+observations must be equal, where the observations are made by using the
+selectors and discriminators.
 
-    * datatypes:
-      * induction axiom (schema)
-        * consequence: no infinite objects (in particular, no cyclic objects)
+% TODO: where does this go?
+%      * when looking at the universal ground (?) theory:
+%        enough to consider acyclicity -- no way to specify infinite objects
+%        otherwise
 
-      * when looking at the universal ground (?) theory:
-        enough to consider acyclicity -- no way to specify infinite objects
-        otherwise
-
-    * codatatypes:
-      * coinduction axiom (schema)
-        * consequence: infinite objects are allowed, but
-          two objects yielding the same observations must be equal
+%    * codatatypes:
+%      * coinduction axiom (schema)
+%        * consequence: infinite objects are allowed, but
+%          two objects yielding the same observations must be equal
 
@@
     (co)recursion
+  * quantifiers?
 

### 2014-12-05 tex/codatadp.tex
more

@@
 
-Blah.
+Freely generated datatypes are widely used both in functional programs and in
+logical specifications. They can be implemented efficiently and enjoy nice
+properties that can be exploited in automated reasoners. Although they have
+their uses in mathematics, they are especially prominent in computer science
+applications.
 
-  * briefly: datatypes and why they are useful
-    * Hoare's recursive data structures
-    * comparatively easy to reason about and automate
-    * common in practice, esp. in computer science applications
+However, because datatype values correspond to ground terms, they can only
+represent finite objects. For example, the datatype of natural numbers
+constructed by $\const{Z} : \ty{nat}$ and $\const{S} : \ty{nat} \to \ty{nat}$
+only allow values of the form $\const{S}(\ldots(\const{S}(\const{Z}))\ldots)$.
+To represent infinite values such as $\const{S}(\const{S}(\ldots))$, the natural
+choice is to turn to codatatypes, the coinductive dual of datatypes.
 
-    (terminology: freely-generated, inductive, algebraic, ..., sometimes
-    with different meanings; we'll clarify below what we use)
+  * are sometimes perceived as esoteric
+  * but they have their uses where they are nearly irreplacable
+  * for example, CompCert and JinjaThreads
+
+  * codatatype approach unifies the finite and infinite case; it is often more
+    convenient to have
+      codatatype enat = EZ | ES enat
+    than
+      datatype enat = EZ | ES enat | EI
+    with $\infty$ represented by $ES(ES(\ldots))$. Notice that adding
+    $\infty$, as expected
+
+  * also can be understood very easily:
+    intuitively, they consist of all ground constructor terms, finite and infinite
+    and only those
+  * one implication is that if $m$ equals $ES(n)$ and $n$ equals $ES(m)$, necessarily
+    $m$ and $n$ must be equal.
+
+  * perhaps the most commonly used codatatype is that of lazy lists or sequences.
+    using a syntax similar to Standard ML, Haskell, or SMT-LIB
+
+$\mu$ notation
 
@@
 
+%    (terminology: freely-generated, inductive, algebraic, ..., sometimes
+%    with different meanings; we'll clarify below what we use)
+
 We fix a signature $\Sig = (\Types, \Funcs)$. The types are partioned into
@@
 However, this is disallowed by SMT-LIB 2, so we do not consider it here.}
-Schematically, using a syntax similar to Standard ML, Haskell,
-or SMT-LIB:
+Schematically:
 %

### 2014-12-05 tex/codatadp.tex
more

@@
 bisimilarity rule that identifies observationally equal, possibly cyclic values.
-The procedure is complete for the universal fragment and is compatible with the
+The procedure is complete for ground problems and is compatible with the
 Nelson--Oppen method. It has been implemented in the latest version of CVC4, a
@@
 applications.
-
+%
 However, because datatype values correspond to ground terms, they can only
@@
 only allow values of the form $\const{S}(\ldots(\const{S}(\const{Z}))\ldots)$.
-To represent infinite values such as $\const{S}(\const{S}(\ldots))$, the natural
-choice is to turn to codatatypes, the coinductive dual of datatypes.
 
-  * are sometimes perceived as esoteric
-  * but they have their uses where they are nearly irreplacable
-  * for example, CompCert and JinjaThreads
+To represent infinite values such as $\const{S}(\const{S}(\ldots))$, a natural
+choice is to turn to codatatypes (or coinductive datatypes), the
+non-well-founded dual of (inductive) datatypes.
+%
+Although they are sometimes perceived as esoteric, codatatypes have a role to
+play in computer science. The verified C compiler CompCert \cite{leroy-2009},
+the verified Java compiler JinjaThreads \cite{lochbihler-2010-jinja}, and the
+formalized Java memory model \cite{lochbihler-2014-jmm} all rely on
+codatatypes to capture infinite processes.
 
-  * codatatype approach unifies the finite and infinite case; it is often more
-    convenient to have
-      codatatype enat = EZ | ES enat
-    than
-      datatype enat = EZ | ES enat | EI
-    with $\infty$ represented by $ES(ES(\ldots))$. Notice that adding
-    $\infty$, as expected
+Codatatypes, like datatypes, are freely generated by their constructors. But
+whereas datatype values are finite, for codatatypes infinite constructor terms
+also denote values. Intuitively, a codatatype consists of all finite and
+infinite ground constructor terms, and only those. For example, the
+specification
+%
+\[\keyw{codatatype}~\,\ty{enat} \,=\, \const{EZ} \,\mid\, \const{ES}(\ty{enat})\]
+%
+(expressed in a notation inspired by typed functional programming languages)
+introduces a type that aptly models the natural numbers extended with an
+infinite value $\infty = \const{ES}(\const{ES}(\const{ES}(\ldots)))$. Compared
+with the more conventional, inductive definition
+\[\keyw{datatype}~\,\ty{enat} \,=\, \const{EZ} \,\mid\, \const{ES}(\ty{enat}) \,\mid\, \const{EInf}\]
+%
+the codatatype eliminates one case by unifying the finite and infinite cases.
+Equations such as $\const{ES}(\infty) \teq \infty$ hold without any additional
+work, because both sides of the equality share the same infinite expansion
+$\const{ES}(\const{ES}(\const{ES}(\ldots)))$.
 
-  * also can be understood very easily:
-    intuitively, they consist of all ground constructor terms, finite and infinite
-    and only those
-  * one implication is that if $m$ equals $ES(n)$ and $n$ equals $ES(m)$, necessarily
-    $m$ and $n$ must be equal.
+%  * one implication is that if $m$ equals $ES(n)$ and $n$ equals $ES(m)$, necessarily
+%    $m$ and $n$ must be equal.
 

### 2014-12-05 tex/codatadp.tex
more

@@
 Although they are sometimes perceived as esoteric, codatatypes have a role to
-play in computer science. The verified C compiler CompCert \cite{leroy-2009},
-the verified Java compiler JinjaThreads \cite{lochbihler-2010-jinja}, and the
-formalized Java memory model \cite{lochbihler-2014-jmm} all rely on
-codatatypes to capture infinite processes.
+play in computer science. The verified C compiler CompCert
+\cite{leroy-2009}, the verified Java compiler JinjaThreads
+\cite{lochbihler-2010-jinja}, and the formalized Java memory model
+\cite{lochbihler-2014-jmm} all rely on codatatypes to capture infinite
+processes. Many proof assistants, including Agda, Coq, Isabelle, Matita, and
+PVS, explicitly support them, as do the CoALP dialect of Prolog and the Dafny
+specification language.
 
@@
 %
-(expressed in a notation inspired by typed functional programming languages)
-introduces a type that aptly models the natural numbers extended with an
+(using an ML-like syntax) introduces a type that aptly models the natural numbers extended with an
 infinite value $\infty = \const{ES}(\const{ES}(\const{ES}(\ldots)))$. Compared

### 2014-12-05 tex/codatadp.tex
more

@@
 
-Freely generated datatypes are widely used both in functional programs and in
-logical specifications. They can be implemented efficiently and enjoy nice
-properties that can be exploited in automated reasoners. Although they have
-their uses in mathematics, they are especially prominent in computer science
-applications.
+Freely generated datatypes are ubiquitous in functional programs and
+logical specifications. They can be implemented efficiently and enjoy
+properties that can be exploited in automated reasoners. They are especially
+useful to represent finite data structures in computer science applications but
+also arise in mathematical formalizations.
 %
-However, because datatype values correspond to ground terms, they can only
-represent finite objects. For example, the datatype of natural numbers
-constructed by $\const{Z} : \ty{nat}$ and $\const{S} : \ty{nat} \to \ty{nat}$
-only allow values of the form $\const{S}(\ldots(\const{S}(\const{Z}))\ldots)$.
+%However, because datatype values correspond to finite ground terms, they
+%are generally not adequate to represent infinite objects.
+%For example, the datatype of natural numbers
+%constructed by $\const{Z} : \ty{nat}$ and $\const{S} : \ty{nat} \to \ty{nat}$
+%only allow values of the form $\const{S}(\ldots(\const{S}(\const{Z}))\ldots)$.
 
-To represent infinite values such as $\const{S}(\const{S}(\ldots))$, a natural
-choice is to turn to codatatypes (or coinductive datatypes), the
+To represent infinite values, % such as $\const{S}(\const{S}(\const{S}(\ldots))$,
+a natural choice is to turn to codatatypes (or coinductive datatypes), the
 non-well-founded dual of (inductive) datatypes.
@@
 Although they are sometimes perceived as esoteric, codatatypes have a role to
-play in computer science. The verified C compiler CompCert
+play in computer science. For example, the verified C compiler CompCert
 \cite{leroy-2009}, the verified Java compiler JinjaThreads
 \cite{lochbihler-2010-jinja}, and the formalized Java memory model
-\cite{lochbihler-2014-jmm} all rely on codatatypes to capture infinite
-processes. Many proof assistants, including Agda, Coq, Isabelle, Matita, and
-PVS, explicitly support them, as do the CoALP dialect of Prolog and the Dafny
-specification language.
+\cite{lochbihler-2014-jmm} all rely crucially on codatatypes to capture infinite
+processes.
+% as do the CoALP dialect of Prolog and the Dafny specification language.
 
-Codatatypes, like datatypes, are freely generated by their constructors. But
-whereas datatype values are finite, for codatatypes infinite constructor terms
-also denote values. Intuitively, a codatatype consists of all finite and
-infinite ground constructor terms, and only those. For example, the
-specification
+Codatatypes, like datatypes, are freely generated by their constructors;
+but for codatatypes, infinite constructor terms also denote values
+(Section~\ref{sec:theory-of-co-datatypes}). Intuitively, a
+codatatype consists of all finite and infinite ground constructor terms, and
+only those. For example, the specification
 %
-\[\keyw{codatatype}~\,\ty{enat} \,=\, \const{EZ} \,\mid\, \const{ES}(\ty{enat})\]
+\[\keyw{codatatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat})\]
 %
 (using an ML-like syntax) introduces a type that aptly models the natural numbers extended with an
-infinite value $\infty = \const{ES}(\const{ES}(\const{ES}(\ldots)))$. Compared
+infinite value $\infty = \const{S}(\const{S}(\const{S}(\ldots)))$. Compared
 with the more conventional, inductive definition
-\[\keyw{datatype}~\,\ty{enat} \,=\, \const{EZ} \,\mid\, \const{ES}(\ty{enat}) \,\mid\, \const{EInf}\]
+\[\keyw{datatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat}) \,\mid\, \const{Inf}\]
 %
 the codatatype eliminates one case by unifying the finite and infinite cases.
-Equations such as $\const{ES}(\infty) \teq \infty$ hold without any additional
+Equations such as $\const{S}(\infty) \teq \infty$ hold without any additional
 work, because both sides of the equality share the same infinite expansion
-$\const{ES}(\const{ES}(\const{ES}(\ldots)))$.
+$\const{S}(\const{S}(\const{S}(\ldots)))$.
 
-%  * one implication is that if $m$ equals $ES(n)$ and $n$ equals $ES(m)$, necessarily
-%    $m$ and $n$ must be equal.
+Datatypes and codatatypes are an integral part of most popular proof assistants,
+including Agda, Coq, Isabelle, Matita, and PVS. In recent years, more and more
+automatic theorem provers have been extended with dedicated reasoning facilities
+for datatypes. The SMT-LIB~2 \cite{barrett-et-al-2010} syntax, implemented by
+most SMT (satisfiability modulo theories) solvers, includes a theory of
+datatypes.
 
-  * perhaps the most commonly used codatatype is that of lazy lists or sequences.
-    using a syntax similar to Standard ML, Haskell, or SMT-LIB
+In this paper, we introduce a unified decision procedure for both datatypes and
+codatatypes. The procedure is described abstractly as a calculus
+(Section~\ref{sec:the-calculus}). To our knowledge, this is the first decision
+procedure for codatatypes. For the datatype case, it roughly follows the lines
+of Barrett et al.\ \cite{barrett-et-al-2007}. Because datatypes and codatatypes
+share many of the same properties, it is sensible to consider them together.
 
-$\mu$ notation
+%  * codatatypes were added later, motivated by the use of SMT solvers as
+%    backends to proof assistants (more specifically, CVC4 to Isabelle/HOL)
 
-  * less briefly: codatatypes and why they are useful
-    * in Agda, Coq, Matita -- recently also in Isabelle/HOL \cite{nipkow-et-al-2002}
-    * but not in SMT-LIB 2 \cite{barrett-et-al-2010}
-
-  * single decision procedure for datatypes and codatatypes
-  * datatypes are implementation following the lines of Barrett et al.
-  * codatatypes were added later, motivated by the use of SMT solvers as
-    backends to proof assistants (more specifically, CVC4 to Isabelle/HOL)
-
-  * setting (Section~\ref{sec:theory-of-co-datatypes}):
+%  * setting :
     * universal formulas
@@
 
-  * decision procedure is described abstractly as a calculus
-    (Section~\ref{sec:the-calculus})
-    * includes rewriting
-    * and inferences
-
   * integrated with Nelson-Oppen
@@
 
+\newpage
+
+* SMT-Lib
+
+%  * one implication is that if $m$ equals $ES(n)$ and $n$ equals $ES(m)$, necessarily
+%    $m$ and $n$ must be equal.
+
+  * perhaps the most commonly used codatatype is that of lazy lists or sequences.
+    using a syntax similar to Standard ML, Haskell, or SMT-LIB
+
+$\mu$ notation
+
+  * less briefly: codatatypes and why they are useful
+    * in Agda, Coq, Matita -- recently also in Isabelle/HOL \cite{nipkow-et-al-2002}
+    * but not in SMT-LIB 2 \cite{barrett-et-al-2010}
+
 \paragraph{Related Work.}
@@
       * (deal more directly with finite sorts than Barrett et al., Section 6.1)
+    * SPASS-Pirate
     * additional ones since then (e.g. strings?)
@@
 
+  * presented a decision procedure for datatypes and codatatypes
+
+  * part of our program to reduce the gap between automatic and interactive
+    theorem proving
+  * and to enrich provers with more high-level features that can be implemented
+    efficiently
+
 Future work:

### 2014-12-05 tex/codatadp.tex
more

@@
 
+\newcommand\cpp{C\nobreak\raisebox{.05ex}{+}\nobreak\raisebox{.05ex}{+}}
+
 \newcommand\iPrime{i\vthinspace'\negvthinspace}
@@
 \newcommand\Codata{\Types_{\mathrm{cdt}}}
-\newcommand\Nondata{\Types_{\mathrm{pln}}}
+\newcommand\Nondata{\Types_{\mathrm{oth}}}
 
@@
 Codatatypes, like datatypes, are freely generated by their constructors;
-but for codatatypes, infinite constructor terms also denote values
+but for them, infinite constructor terms are also legitimate values
 (Section~\ref{sec:theory-of-co-datatypes}). Intuitively, a
@@
 Equations such as $\const{S}(\infty) \teq \infty$ hold without any additional
-work, because both sides of the equality share the same infinite expansion
-$\const{S}(\const{S}(\const{S}(\ldots)))$.
+work, because both sides share the same infinite expansion
+$\const{S}(\const{S}(\const{S}(\ldots)))$, which uniquely identifies a
+codatatype value.
 
@@
 
-In this paper, we introduce a unified decision procedure for both datatypes and
-codatatypes. The procedure is described abstractly as a calculus
-(Section~\ref{sec:the-calculus}). To our knowledge, this is the first decision
-procedure for codatatypes. For the datatype case, it roughly follows the lines
-of Barrett et al.\ \cite{barrett-et-al-2007}. Because datatypes and codatatypes
-share many of the same properties, it is sensible to consider them together.
+In this paper, we introduce a unified decision procedure for ground problems
+involving both datatypes and codatatypes. The procedure is described abstractly
+as a calculus (Section~\ref{sec:the-calculus}) and is design to be composable
+using the Nelson--Oppen method \cite{nelson-oppen-1979}. To our knowledge, this
+is the first such decision procedure for codatatypes. For the datatype case, it
+roughly follows the lines of Barrett et al.\ \cite{barrett-et-al-2007}.
+
+\newcommand\cheat{\vskip.75\smallskipamount} %% TYPESETTING
+
+Datatypes and codatatypes share many of the same properties, so it makes sense
+to consider them together. The main differences are as follows:
+%
+\begin{itemize}
+\item
+No well-foundedness check; e.g. the type of infinite sequences, or streams, over an
+element type $\tau$
+%
+\[\keyw{codatatype}~\;\ty{stream}_{\,\tau} \,=\, \const{SCons}(\tau,\, \ty{stream}_{\,\tau})\]
+is allowed, whereas the corresponding datatype would be rejected, because no
+finite ground constructor terms exist.
+
+\cheat
+\item
+The cyclicity check must be dropped. A constraint such as
+$\const{x} \teq \const{C}(\const{x})$, where $\const{C}$ is a constructor,
+is unsatisfiable for datatypes but satisfiable for codatatypes.
+Instead: bisimilarity check, which states that two values having the same (possibly
+infinite) expansion must be equal. From $\const{x} \teq \const{C}(\const{y})$ and
+$\const{y} \teq \const{C}(\const{x})$, deduce $\const{x} = \const{y}$.
+
+\cheat
+\item
+Normal forms are needed and must be enumeratable. For finite model finding, requires a $\mu$ notation,
+e.g., $\const{SCons}(0,\, \mu s.\; \const{SCons}(1,\, \const{SCons}(2,\, s)))$
+represents the lasso-shaped sequence $0, 1, 2, 1, 2, \ldots$.
+\end{itemize}
+
+Both cyclicity and bisimilarity checks cannot be replaced by finite
+axiomatizations, so they really belong in a decision procedure. They are
+necessary for completeness of ground problems and for soundness of found models.
+%    * in particular, acyclicity and bisimilarity are necessary for some proofs
+%%      (and cannot be axiomatized finitely)
+%    * and for model finding, without them we quickly get spurious models
+%    * explain how finite model finding works
+ And the other
+(co)datatype properties, such as distinctness of constructors and
+exhaustiveness, can be done more efficiently by a decision procedure than by an
+axiom.
+%
+The proofs or soundness and completeness are included in the technical report
+associated with this paper.
+
+The procedure is implemented in CVC4 as rewriting and a theory solver
+(Section~\ref{sec:the-theory-solver}). It implements a precise strategy for applying
+inferences. It consists of about 2000 lines of \cpp{} code, among which 1600 are
+shared between datatypes and codatatypes. The code is integrated in the
+development version of the solver and is expected to be part of the 1.5 release.
+%
+The decision procedure is evaluated on %hand-crafted examples and on
+problems generated from Isabelle theories using Isabelle's Sledgehammer tool
+(Section~\ref{sec:experimental-results}).
+
+%  * useful both for proving and for model finding
+
+%\ref{sec:examples}
+
+%  * benchmarking is often an issue -- esp. codatatypes
+%
+%Polymorphic types, nested (co)recursion, and datatype--codatatype mixtures fall
+
+%  * setting: FOL
+%    * core procedure is restricted to ground
+%      * but theory solver cooperates
+%    * still, universal conjectures fall into the ground fragment via negation
+%      and skolemization (or, alternatively, herbrandization)
 
@@
 %  * setting :
-    * universal formulas
-    * many-sorted logic
-    * mutually (co)recursive types with constructors, selectors, and
-      discriminators
+%    * universal formulas
+%    * many-sorted logic
+%   * mutually (co)recursive types with constructors, selectors, and
+%      discriminators
 
-  * codatatypes: from a theoretical and implementational point of view, like
-      datatypes but:
-    * infinite values (infinitely many nested constructors)
-    * codatatypes are never empty (e.g. finite streams are rejected)
+%  * codatatypes: from a theoretical and implementational point of view, like
+%      datatypes but:
+%    * infinite values (infinitely many nested constructors)
+%    * codatatypes are never empty (e.g. finite streams are rejected)
 
-\[
-      \keyw{codatatype}\; \,\ty{llist} \,=\, \const{LNil} \,\mid\, \const{LCons}(\ty{int},\, \ty{llist})
-\]
+%\[
+%      \keyw{codatatype}\; \,\ty{llist} \,=\, \const{LNil} \,\mid\, \const{LCons}(\ty{int},\, \ty{llist})
+%\]
 
-    * no acyclicity (e.g. xs = lcons(0, xs) is satisfiable)
-    * instead: bisimilarity check
-    * (also: ``enumeration'', e.g. of streams)
+%  * consider a natural fragment---datatypes as supported in modern SMT solvers
+%    and the SMT-LIB 2 standard, and codatatypes as their duals
+%    * mutual recursion, but no polymorphism, nested recursion
+%  * integrated with Nelson-Oppen
 
-  * motivation:
-    * acyclicity and bisimilarity cannot be axiomatized, so they really belong in
-      a decision procedure
-    * other properties can be done more efficiently by a decision procedure
-
-  * consider a natural fragment---datatypes as supported in modern SMT solvers
-    and the SMT-LIB 2 standard, and codatatypes as their duals
-    * mutual recursion, but no polymorphism, nested recursion
-
-  * integrated with Nelson-Oppen
-
-  * implemented in CVC4 as rewriting and a theory solver
-    (\ref{sec:the-theory-solver})
-    * precise strategy for applying inferences
-    * about 2000 lines of code, among which 1600 are shared between datatypes
-      and codatatypes
-
-  * useful both for proving and for model finding
-    * in particular, acyclicity and bisimilarity are necessary for some proofs
-      (and cannot be axiomatized finitely)
-    * and for model finding, without them we quickly get spurious models
-    * explain how finite model finding works
-
-%\ref{sec:examples}
-
-  * benchmarking is often an issue -- esp. codatatypes
-  * the decision procedure is evaluated on two sets of benchmarks
-    (Section~\ref{sec:experimental-results}):
-    * first set: hand-crafted examples
-    * second set: generated from Isabelle theories using the Sledgehammer tool
-
-Polymorphic types, nested (co)recursion, and datatype--codatatype mixtures fall
-
-The proofs are part of a report.
-
-\newpage
-
-* SMT-Lib
+% * SMT-Lib
 
@@
 
-  * perhaps the most commonly used codatatype is that of lazy lists or sequences.
-    using a syntax similar to Standard ML, Haskell, or SMT-LIB
+%  * perhaps the most commonly used codatatype is that of lazy lists or sequences.
+%    using a syntax similar to Standard ML, Haskell, or SMT-LIB
 
-$\mu$ notation
+%  * less briefly: codatatypes and why they are useful
+%    * in Agda, Coq, Matita -- recently also in Isabelle/HOL \cite{nipkow-et-al-2002}
+%    * but not in SMT-LIB 2 \cite{barrett-et-al-2010}
 
-  * less briefly: codatatypes and why they are useful
-    * in Agda, Coq, Matita -- recently also in Isabelle/HOL \cite{nipkow-et-al-2002}
-    * but not in SMT-LIB 2 \cite{barrett-et-al-2010}
-
-\paragraph{Related Work.}
-
-    * point to Barrett et al. for SMT datatypes
-      * about their own work, they say: "our focus is on generality and
-        efficiency rather than immediacy of implementation"
-      * (deal more directly with finite sorts than Barrett et al., Section 6.1)
-    * SPASS-Pirate
-    * additional ones since then (e.g. strings?)
-    * anything about codatatypes?
-      * proof assistants like Agda, Coq, etc. have them
-      * Dafny, CoALP
-      * also a lot of theoretical research, some of which is loosely connected,
-        e.g. decision procedure for corecursive functions (Henning in Nijmegen)
-
-    * ODDITY: Oppen 1980: single-constructor, recursive -- infinite values?
+% \paragraph{Related Work.}
+% 
+%     * point to Barrett et al. for SMT datatypes
+%       * about their own work, they say: "our focus is on generality and
+%         efficiency rather than immediacy of implementation"
+%       * (deal more directly with finite sorts than Barrett et al., Section 6.1)
+%     * SPASS-Pirate
+%     * additional ones since then (e.g. strings?)
+%     * anything about codatatypes?
+%       * proof assistants like Agda, Coq, etc. have them
+%       * Dafny, CoALP
+%       * also a lot of theoretical research, some of which is loosely connected,
+%         e.g. decision procedure for corecursive functions (Henning in Nijmegen)
+% 
+%     * ODDITY: Oppen 1980: single-constructor, recursive -- infinite values?
 
@@
 A (co)datatype specification consists of $l$~mutually recursive types which are
-either all datatypes or all codatatypes. Each datatype $\delta$ is equipped by
+either all datatypes or all codatatypes. Polymorphic types, nested
+(co)recursion, and datatype--codatatype mixtures fall outside this fragment. In
+the presentation, we allow ourselves some metalevel parameterization
+% at the metalevel
+by using subscripts---for example, $\ty{stream}_{\,\tau}$ denotes a
+family of ground types $\ty{stream}_{\,\ty{int}}$, $\ty{stream}_{\,\ty{bool}}$,
+$\ty{stream}_{\,\ty{stream}_{\,\ty{real}}}$, etc.
+
+Each datatype $\delta$ is equipped by
 $m \ge 1$ constructors, and each constructor for $\delta$ takes zero or more
 arguments and returns a $\delta$ value. The argument types must be either
-plain, be or be one of the newly introduced (co)datatypes. To every
-argument corresponds a selector. The introduced names for the (co)datatypes, the
+plain, be or be one of the newly introduced (co)datatypes.
+%
+To every argument corresponds a selector. The introduced names for the (co)data\-types, the
 constructors, and the selectors must be distinct and different from already
@@
 
-Polymorphic types, nested (co)recursion, and datatype--codatatype mixtures fall
-outside this fragment. For example, the specification
+Here are a few examples of legal specifications of (co)datatype families:
 \[\begin{aligned}[t]
-      \keyw{datatype}\;\, \ty{tree} & \,=\, \const{Node}(\ty{int},\, \ty{forest}) \\[-\jot]
-      \keyw{and}\;\, \ty{forest} & \,=\, \const{FNil} \,\mid\, \const{FCons}(\ty{tree},\, \ty{forest})
+      \keyw{datatype}\;\, \ty{tree}_{\,\tau} & \,=\, \const{Node}(\tau,\, \ty{forest}_{\,\tau}) \\[-\jot]
+      \keyw{and}\;\, \ty{forest}_{\,\tau} & \,=\, \const{FNil} \,\mid\, \const{FCons}(\ty{tree}_{\,\tau},\, \ty{forest}_{\,\tau}) \\%[-.5\jot]
+      \keyw{codatatype}\; \,\ty{llist}_{\,\tau} & \,=\, \const{LNil} \,\mid\, \const{LCons}(\tau,\, \ty{llist}_{\,\tau})
 \end{aligned}
 \]
-is supported, whereas
-\[\begin{aligned}[t]
-      \keyw{codatatype}\; \,\ty{gllist}(\alpha) & \,=\, \const{GLNil} \,\mid\, \const{GLCons}(\alpha,\, \ty{gllist}(\alpha)) \\[-.5\jot]
-      \keyw{datatype}\;\, \ty{tree} & \,=\, \const{Node}(\ty{int},\, \ty{gllist}(\ty{tree}))
-\end{aligned}
-\]
-is beyond the fragment considered here. In principle, rank-1 (top-level)
-polymorphism would be easy to support; nesting datatypes inside datatypes, and
-likewise for codatatypes, can be reduced to the mutual case
-\cite{gunter-1993-not}. So the only genuinely interesting cases missing are
-mixed nested (co)recursion (as in the \ty{tree} example above) as well as
-(co)recursion through a non-(co)datatype (both of which make sense
-\cite{blanchette-et-al-2014-codata}).
 
@@
 admissible only if a ground term can be exhibited that belongs to that datatype.
-This rules out ill-founded specifications such as
+%This rules out ill-founded specifications such as
 %
-\[\keyw{datatype}\;\, \ty{fstream} \,=\, \const{FSCons}(\ty{int},\,\ty{fstream})\]
+%\[\keyw{datatype}\;\, \ty{fstream} \,=\, \const{FSCons}(\ty{int},\,\ty{fstream})\]
 %
-For codatatypes, no such check is necessary. Given the specification
+For codatatypes, no such check is necessary because there is always a term,
+finite or infinite, that witnesses nonemptiness.
+%Given the specification
 %
@@
 %
-the infinite value $\const{SCons}(0, \const{SCons}(0, \ldots))$ witnesses
-nonemptiness.
+%the infinite value $\const{SCons}(0, \const{SCons}(0, \ldots))$ witnesses
+%nonemptiness.
 
@@
   * SMT background
+
+\subsection{The Ground Case}
 
@@
 
-\subsection{Correctness}
+\paragraph{Correctness.}
 
@@
 
+\subsection{Extension to Quantified Formulas}
+
+Blah.
+
 \section{The Theory Solver}
@@
     (co)recursion
+
+In principle, rank-1 (top-level)
+polymorphism would be easy to support; nesting datatypes inside datatypes, and
+likewise for codatatypes, can be reduced to the mutual case
+\cite{gunter-1993-not}. So the only genuinely interesting cases missing are
+mixed nested (co)recursion (as in the \ty{tree} example above) as well as
+(co)recursion through a non-(co)datatype (both of which make sense
+\cite{blanchette-et-al-2014-codata}).
+
   * quantifiers?

### 2014-12-05 tex/codatadp.tex
more

@@
 
+\let\labelitemi=\labelitemii %% CHEAT!
+
 \newcommand\cpp{C\nobreak\raisebox{.05ex}{+}\nobreak\raisebox{.05ex}{+}}
@@
 Codatatypes naturally capture potentially infinite data structures and
-processes. We introduce a decision procedure that combines reasoning about
+processes. We present a decision procedure that combines reasoning about
 datatypes and codatatypes. The dual of the cyclicity rule for datatypes is a
-bisimilarity rule that identifies observationally equal, possibly cyclic values.
+bisimilarity rule that identifies equal values, also in the presence of cycles.
 The procedure is complete for ground problems and is compatible with the
@@
 Freely generated datatypes are ubiquitous in functional programs and
-logical specifications. They can be implemented efficiently and enjoy
-properties that can be exploited in automated reasoners. They are especially
+logical specifications. They are especially
 useful to represent finite data structures in computer science applications but
 also arise in mathematical formalizations.
+They can be implemented efficiently and enjoy
+properties that can be exploited in automated reasoners. 
 %
@@
 play in computer science. For example, the verified C compiler CompCert
-\cite{leroy-2009}, the verified Java compiler JinjaThreads
+\cite{leroy-2009}, the verified Java compiler Jinja\-Threads
 \cite{lochbihler-2010-jinja}, and the formalized Java memory model
@@
 
-Codatatypes, like datatypes, are freely generated by their constructors;
-but for them, infinite constructor terms are also legitimate values
+Codatatypes are freely generated by their constructors, but unlike for datatypes
+infinite constructor terms are also legitimate values
 (Section~\ref{sec:theory-of-co-datatypes}). Intuitively, a
@@
 %
-(using an ML-like syntax) introduces a type that aptly models the natural numbers extended with an
-infinite value $\infty = \const{S}(\const{S}(\const{S}(\ldots)))$. Compared
+(using an ML-like syntax) introduces a type that faithfully
+models the natural numbers extended with an
+infinity $\infty = \const{S}(\const{S}(\const{S}(\ldots)))$. Compared
 with the more conventional, inductive definition
@@
 the codatatype eliminates one case by unifying the finite and infinite cases.
-Equations such as $\const{S}(\infty) \teq \infty$ hold without any additional
+Moreover, equations such as $\const{S}(\infty) \teq \infty$ hold without additional
 work, because both sides share the same infinite expansion
-$\const{S}(\const{S}(\const{S}(\ldots)))$, which uniquely identifies a
+$\const{S}(\const{S}(\const{S}(\ldots)))$, which uniquely identifies the
 codatatype value.
@@
 \item
-No well-foundedness check; e.g. the type of infinite sequences, or streams, over an
+\emph{No well-foundedness check is necessary for codatatypes.}
+For example, the type of infinite sequences, or streams, over an
 element type $\tau$
@@
 \[\keyw{codatatype}~\;\ty{stream}_{\,\tau} \,=\, \const{SCons}(\tau,\, \ty{stream}_{\,\tau})\]
-is allowed, whereas the corresponding datatype would be rejected, because no
-finite ground constructor terms exist.
+is allowed; the corresponding datatype would be rejected as
+non-well-founded.
 
@@
 \item
-The cyclicity check must be dropped. A constraint such as
-$\const{x} \teq \const{C}(\const{x})$, where $\const{C}$ is a constructor,
-is unsatisfiable for datatypes but satisfiable for codatatypes.
-Instead: bisimilarity check, which states that two values having the same (possibly
-infinite) expansion must be equal. From $\const{x} \teq \const{C}(\const{y})$ and
-$\const{y} \teq \const{C}(\const{x})$, deduce $\const{x} = \const{y}$.
+\emph{A bisimilarity rule takes the place of the cyclicity rule of datatypes.}
+A cyclic constraint such as
+$\const{x} \teq \const{S}(\const{x})$ %, where $\const{C}$ is a constructor,
+is unsatisfiable for a datatype but satisfiable for a codatatype.
+The bisimilarity principle states that two values having the same (possibly
+infinite) expansion must be equal. From $\const{x} \teq \const{S}(\const{y})$ and
+$\const{y} \teq \const{S}(\const{x})$, it deduces $\const{x} \teq \const{y}$.
 
@@
 Normal forms are needed and must be enumeratable. For finite model finding, requires a $\mu$ notation,
-e.g., $\const{SCons}(0,\, \mu s.\; \const{SCons}(1,\, \const{SCons}(2,\, s)))$
-represents the lasso-shaped sequence $0, 1, 2, 1, 2, \ldots$.
+e.g., $\const{SCons}(3,\, \mu s.\; \const{SCons}(1,\, \const{SCons}(4,\, s)))$
+represents the lasso-shaped sequence $3, 1, 4, 1, 4, \ldots$\vthinspace.
 \end{itemize}
@@
 
-The procedure is implemented in CVC4 as rewriting and a theory solver
+The procedure is implemented in the SMT solver CVC4 as rewriting and a theory solver
 (Section~\ref{sec:the-theory-solver}). It implements a precise strategy for applying

### 2014-12-05 tex/codatadp.tex
more

@@
 e.g., $\const{SCons}(3,\, \mu s.\; \const{SCons}(1,\, \const{SCons}(4,\, s)))$
-represents the lasso-shaped sequence $3, 1, 4, 1, 4, \ldots$\vthinspace.
+represents the lasso-shaped sequence $3, 1, 4, 1, 4, \ldots$\kern0.1em. %%% TYPESETTING: kern
 \end{itemize}

### 2014-12-05 tex/codatadp.tex
more

@@
 
-\newcommand\cheat{\vskip.75\smallskipamount} %% TYPESETTING
+\newcommand\cheat{\vskip\smallskipamount} %% TYPESETTING
 
@@
 \item
-Normal forms are needed and must be enumeratable. For finite model finding, requires a $\mu$ notation,
+\emph{It must be possible to express cyclic values as closed terms and
+to enumerate them.} For finite model finding, requires a $\mu$ notation,
 e.g., $\const{SCons}(3,\, \mu s.\; \const{SCons}(1,\, \const{SCons}(4,\, s)))$

### 2014-12-05 tex/codatadp.tex
more

@@
 \emph{It must be possible to express cyclic values as closed terms and
-to enumerate them.} For finite model finding, requires a $\mu$ notation,
-e.g., $\const{SCons}(3,\, \mu s.\; \const{SCons}(1,\, \const{SCons}(4,\, s)))$
+to enumerate them.} This is necessary both for finite model finding (modulo theories)
+and for theory combinations. The $\mu$ notation is used to represent cyclic ($\omega$-regular)
+values. For example,
+$\const{SCons}(3,\, \mu s.\; \const{SCons}(1,\, \const{SCons}(4,\, s)))$
 represents the lasso-shaped sequence $3, 1, 4, 1, 4, \ldots$\kern0.1em. %%% TYPESETTING: kern
@@
 The procedure is implemented in the SMT solver CVC4 as rewriting and a theory solver
-(Section~\ref{sec:the-theory-solver}). It implements a precise strategy for applying
-inferences. It consists of about 2000 lines of \cpp{} code, among which 1600 are
+(Section~\ref{sec:the-theory-solver}).
+It consists of about 2000 lines of \cpp{} code, among which 1600 are
 shared between datatypes and codatatypes. The code is integrated in the
@@
 %
-The decision procedure is evaluated on %hand-crafted examples and on
+An evaluation on %hand-crafted examples and on
 problems generated from Isabelle theories using Isabelle's Sledgehammer tool
-(Section~\ref{sec:experimental-results}).
+demonstrates the usefulness of the approach (Section~\ref{sec:experimental-results}).
 

### 2014-12-05 tex/codatadp.tex
intro stuff

@@
 bisimilarity rule that identifies equal values, also in the presence of cycles.
-The procedure is complete for ground problems and is compatible with the
+The procedure is complete for universal problems and is compatible with the
 Nelson--Oppen method. It has been implemented in the latest version of CVC4, a
-modern SMT solver. An evaluation based on problems generated from Isabelle
-theories demonstrates the potential of the procedure for both proving and model
-finding.
+state-of-the-art SMT solver. An evaluation based on problems generated from
+Isabelle theories demonstrates the potential of the procedure for both proving
+and model finding.
 \end{abstract}
@@
 %
-Although they are sometimes perceived as esoteric, codatatypes have a role to
-play in computer science. For example, the verified C compiler CompCert
+Although they are sometimes perceived as esoteric, codatatypes have an important
+role to play in computer science. For example, the verified C compiler CompCert
 \cite{leroy-2009}, the verified Java compiler Jinja\-Threads
@@
 
-Codatatypes are freely generated by their constructors, but unlike for datatypes
-infinite constructor terms are also legitimate values
+Codatatypes are freely generated by their constructors, but in contrast with datatypes,
+infinit\-e constructor terms are also legitimate values of codatatypes
 (Section~\ref{sec:theory-of-co-datatypes}). Intuitively, a
-codatatype consists of all finite and infinite ground constructor terms, and
-only those. For example, the specification
+codatatype consists of all well-typed finite and infinite ground constructor
+terms, and only those. For example, the specification
 %
@@
 Moreover, equations such as $\const{S}(\infty) \teq \infty$ hold without additional
-work, because both sides share the same infinite expansion
+work, because both sides expand to the infinite term
 $\const{S}(\const{S}(\const{S}(\ldots)))$, which uniquely identifies the
@@
 involving both datatypes and codatatypes. The procedure is described abstractly
-as a calculus (Section~\ref{sec:the-calculus}) and is design to be composable
-using the Nelson--Oppen method \cite{nelson-oppen-1979}. To our knowledge, this
-is the first such decision procedure for codatatypes. For the datatype case, it
+as a calculus (Section~\ref{sec:the-calculus}) and is designed to be composable
+via the Nelson--Oppen method \cite{nelson-oppen-1979}.
+For the datatype case, it
 roughly follows the lines of Barrett et al.\ \cite{barrett-et-al-2007}.
+To our knowledge, ours is the first decision procedure for codatatypes.
 
@@
 Datatypes and codatatypes share many of the same properties, so it makes sense
-to consider them together. The main differences are as follows:
+to consider them together. There are, however, a number of important
+differences:
 %
@@
 \emph{A bisimilarity rule takes the place of the cyclicity rule of datatypes.}
-A cyclic constraint such as
+Cyclic constraints such as
 $\const{x} \teq \const{S}(\const{x})$ %, where $\const{C}$ is a constructor,
-is unsatisfiable for a datatype but satisfiable for a codatatype.
+are unsatisfiable for datatypes but satisfiable for codatatypes.
 The bisimilarity principle states that two values having the same (possibly
-infinite) expansion must be equal. From $\const{x} \teq \const{S}(\const{y})$ and
+infinite) expansion must be equal; from $\const{x} \teq \const{S}(\const{y})$ and
 $\const{y} \teq \const{S}(\const{x})$, it deduces $\const{x} \teq \const{y}$.
@@
 to enumerate them.} This is necessary both for finite model finding (modulo theories)
-and for theory combinations. The $\mu$ notation is used to represent cyclic ($\omega$-regular)
-values. For example,
-$\const{SCons}(3,\, \mu s.\; \const{SCons}(1,\, \const{SCons}(4,\, s)))$
-represents the lasso-shaped sequence $3, 1, 4, 1, 4, \ldots$\kern0.1em. %%% TYPESETTING: kern
+and for theory combinations. The $\mu$-binder notation is used to represent cyclic ($\omega$-regular)
+values internally. For example,
+the term $\const{SCons}(3,\, \mu s.\; \const{SCons}(1,\, \const{SCons}(4,\, s)))$
+denotes the lasso-shaped sequence $3, 1, 4, 1, 4, \ldots$\kern0.1em. %%% TYPESETTING: kern
 \end{itemize}
 
-Both cyclicity and bisimilarity checks cannot be replaced by finite
-axiomatizations, so they really belong in a decision procedure. They are
-necessary for completeness of ground problems and for soundness of found models.
+The cyclicity and bisimilarity rules are
+needed to ensure completeness on ground problems and the absence of spurious
+models in the context of finite model finding.
+They cannot be replaced by finite
+axiomatizations, so they naturally belong in a decision procedure. 
 %    * in particular, acyclicity and bisimilarity are necessary for some proofs
@@
 %    * explain how finite model finding works
- And the other
-(co)datatype properties, such as distinctness of constructors and
-exhaustiveness, can be done more efficiently by a decision procedure than by an
-axiom.
-%
+For the other (co)datatype properties---the injectivity, distinctness, and
+exhaustiveness of constructors and the selector laws---it is
+widely recognized that decision procedures are more efficient than
+axiomatizations.
+%axiom.
+
 The proofs or soundness and completeness are included in the technical report
-associated with this paper.
-
+associated with this paper \cite{our-report}.
 The procedure is implemented in the SMT solver CVC4 as rewriting and a theory solver
@@
 shared between datatypes and codatatypes. The code is integrated in the
-development version of the solver and is expected to be part of the 1.5 release.
+development version of the solver and is expected to be part of the CVC4~1.5 release.
 %
 An evaluation on %hand-crafted examples and on
-problems generated from Isabelle theories using Isabelle's Sledgehammer tool
+problems generated from Isabelle theories using the Sledgehammer tool
 demonstrates the usefulness of the approach (Section~\ref{sec:experimental-results}).
@@
 \paragraph{Conventions.}
-This paper assumes a monomorphic (or many-sorted) first-order logic throughout.
+The setting is a monomorphic (or many-sorted) first-order logic.
 A signature $\Sig = (\Types, \Funcs)$ consists of a set of types $\Types$ and a
 set of function symbols $\Funcs$. Types are simply atomic sorts, with no
-structure, and interpreted by nonempty domains. The set $\Types$ must contain a
+structure, and interpreted by nonempty domains. The set~$\Types$ must contain a
 distinguished type \ty{bool} interpreted as the set of truth
 values $\{\bot, \top\}$. %, and may contain other interpreted types (e.g., \ty{int}, \ty{real}).
-The only predicate is equality ($\teq$), and it is treated as a logical symbol,
-just like the connectives and quantifiers of first-order logic.
+The only predicate is equality ($\teq$) and is considered a logical symbol.
 Other predicates can be represented as functions that return $\ty{bool}$.
@@
 Function symbols are written in a sans-serif font (e.g., $\const{f}$, $\const{g}$) to
-distinguish them from variables (e.g., $x$, $y$). Upper case names (e.g., $\const{C}$,
-$\const{Nil}$) are reserved for constructors. With each function symbol \const{f}
+distinguish them from variables (e.g., $x$, $y$). Names starting with an uppercase
+letter (e.g.,
+$\const{S}$, $\const{SCons}$) are reserved for constructors. With each function symbol \const{f}
 is associated a list of argument types $\tau_1,\ldots,\tau_n$ (with $n \ge 0$)
 and a return type $\tau$. This connection can be expressed compactly as
-$\const{f} : \tau_1 \times \cdots \times \tau_n \to \tau$
-(which collapses to $\const{f} : \tau$ if $n = 0$).
+$\const{f} : \tau_1 \times \cdots \times \tau_n \to \tau$,
+which collapses to $\const{f} : \tau$ if $n = 0$.
 For a term $t$, the notation $t : \tau$ indicates that $t$ has type $\tau$.
-Functions are applied the standard way, with $\const{f}(t_1,\ldots,t_n)$
+Functions are invoked in the standard way, with $\const{f}(t_1,\ldots,t_n)$
 applying the $n$-ary function symbol
@@
 %
-Finally, the notation $\bar x$ abbreviates a list or tuple $x_1,\ldots,x_n$,
-and $\bigwedge_i \varphi_i$ abbeviates a conjunction $\varphi_1 \mathrel\land \cdots \mathrel\land \varphi_n$.
+Finally, the notation $\bar x$ abbreviates a list or tuple $x_1,\ldots,x_n$.
+% and $\bigwedge_i \varphi_i$ abbeviates a conjunction $\varphi_1 \mathrel\land \cdots \mathrel\land \varphi_n$.
 

### 2014-12-05 tex/codatadp.tex
more

@@
 
+\excludeversion{report}
+\includeversion{paper}
+
+\def\thewordpaper{paper}
+
+\begin{report}
+\def\thewordpaper{report}
+\end{report}
+
 % for "bussproofs" package
@@
 state-of-the-art SMT solver. An evaluation based on problems generated from
-Isabelle theories demonstrates the potential of the procedure for both proving
-and model finding.
+Isabelle theories demonstrates the potential of the procedure.
 \end{abstract}
@@
 
-In this paper, we introduce a unified decision procedure for ground problems
-involving both datatypes and codatatypes. The procedure is described abstractly
-as a calculus (Section~\ref{sec:the-calculus}) and is designed to be composable
-via the Nelson--Oppen method \cite{nelson-oppen-1979}.
+In this \thewordpaper, we introduce a unified decision procedure for ground and
+universal problems involving both datatypes and codatatypes. The procedure is
+described abstractly as a calculus (Section~\ref{sec:the-calculus}) and is
+designed to be composable via the Nelson--Oppen method \cite{nelson-oppen-1979}.
 For the datatype case, it
@@
 Datatypes and codatatypes share many of the same properties, so it makes sense
-to consider them together. There are, however, a number of important
+to consider them together. There are, however, at least three important
 differences:
@@
 
-The proofs or soundness and completeness are included in the technical report
+\begin{paper}
+The proofs of soundness and completeness are included in the technical report
 associated with this paper \cite{our-report}.
+\end{paper}%
 The procedure is implemented in the SMT solver CVC4 as rewriting and a theory solver

### 2014-12-05 tex/codatadp.tex
minor

@@
 \cite{lochbihler-2014-jmm} all rely crucially on codatatypes to capture infinite
-processes.
+execution traces.
 % as do the CoALP dialect of Prolog and the Dafny specification language.
@@
 values internally. For example,
-the term $\const{SCons}(3,\, \mu s.\; \const{SCons}(1,\, \const{SCons}(4,\, s)))$
-denotes the lasso-shaped sequence $3, 1, 4, 1, 4, \ldots$\kern0.1em. %%% TYPESETTING: kern
+the term $\const{SCons}(1,\, \mu s.\; \const{SCons}(0,\, \const{SCons}(9,\, s)))$
+denotes the lasso-shaped sequence $1, 0, 9, 0, 9, 0, 9, \ldots$\kern0.1em. %%% TYPESETTING: kern
 \end{itemize}
+
+%%% 1090909 is a prime number
 

### 2014-12-05 tex/codatadp.tex
more

@@
 either all datatypes or all codatatypes. Polymorphic types, nested
-(co)recursion, and datatype--codatatype mixtures fall outside this fragment. In
-the presentation, we allow ourselves some metalevel parameterization
+(co)recursion, and datatype--codatatype mixtures fall outside this fragment.%
+\footnote{%
+In principle, rank-1 (top-level) polymorphism \cite{blanchette-paskevich-2013}
+would be easy to support; nesting datatypes inside datatypes, and
+likewise for codatatypes, can be reduced to the mutual case
+\cite{gunter-1993-not}. So the only genuinely interesting cases missing are
+mixed nested (co)recursion (as in the \ty{tree} example above) as well as
+(co)recursion through a non-(co)datatype (both of which make sense
+\cite{blanchette-et-al-2014-codata}).}
+In the presentation, we allow ourselves some metalevel parameterization
 % at the metalevel
@@
 
-  * Constructed examples
-    * proving
-    * model finding?
-
-  * Isabelle benchmarks
-
 \section{Conclusion}
@@
 
-  * presented a decision procedure for datatypes and codatatypes
-
-  * part of our program to reduce the gap between automatic and interactive
-    theorem proving
-  * and to enrich provers with more high-level features that can be implemented
-    efficiently
-
-Future work:
-
-  * more evaluation
-    * use it in a new Nitpick-like tool
-  * recursive and corecursive functions
-    (and rewriting)
-  * extensions, perhaps straightforward, to polymorphism and nested
-    (co)recursion
-
-In principle, rank-1 (top-level)
-polymorphism would be easy to support; nesting datatypes inside datatypes, and
-likewise for codatatypes, can be reduced to the mutual case
-\cite{gunter-1993-not}. So the only genuinely interesting cases missing are
-mixed nested (co)recursion (as in the \ty{tree} example above) as well as
-(co)recursion through a non-(co)datatype (both of which make sense
-\cite{blanchette-et-al-2014-codata}).
-
-  * quantifiers?
+We presented a decision procedure for datatypes and codatatypes and implemented
+it in the SMT solver CVC4. Our main theoretical contribution has been the
+extension of the procedure with codatatypes. Codatatypes are not very difficult
+per se, but there are tricky corner cases to take care of if we care about
+completeness (which we do). On the practical side, we obtained interesting
+results on Isabelle benchmarks, including some rather
+%mind-boggling
+puzzling bisimulation proofs.
+
+This work is part of a larger program that aims at enriching automatic
+provers with more high-level features and reducing the gap between automatic and
+interactive theorem proving. As future work, we want to try out CVC4 and its
+(co)datatypes for higher-order model finding in Isabelle, as an
+alternative to the counterexample generator Nitpick
+\cite{blanchette-nipkow-2010}. We also see some opportunities to enrich CVC4
+with recursive and corecursive functions. Finally, it might be possible to go
+further in terms of supporting nested and mixed (co)recursion and quantified
+formulas over (co)datatypes in SMT solvers.
 
@@
 \paragraph{\ackname.}
-
-Clark Barrett and Cesare Tinelli.
-
-Bosses: Viktor Kuncak, Stephan Merz, Tobias Nipkow, and Christoph Weidenbach.
-
-The research was partially supported by the Deutsche
+We owe a great debt to Clark Barrett and Cesare Tinelli for datatype case blah
+blah blah.
+%
+Our present and former bosses, Viktor Kuncak, Stephan Merz, Tobias Nipkow,
+Cesare Tinelli, and Christoph Weidenbach, have either encouraged the research on
+codatatype or at least tolerated it, both of which we are thankful for.
+%
+Blanchette's research was partially supported by the Deutsche
 Forschungs\-gemein\-schaft (DFG) project

### 2014-12-05 tex/codatadp.tex
more

@@
 
-We presented a decision procedure for datatypes and codatatypes and implemented
+We introduced a decision procedure for datatypes and codatatypes and implemented
 it in the SMT solver CVC4. Our main theoretical contribution has been the
-extension of the procedure with codatatypes. Codatatypes are not very difficult
+addition of codatatypes. These are not very difficult
 per se, but there are tricky corner cases to take care of if we care about
 completeness (which we do). On the practical side, we obtained interesting
-results on Isabelle benchmarks, including some rather
+results on benchmarks generated from Isabelle theories, including some rather
 %mind-boggling
@@
 %
+Andreas Lochbihler and Peter Gammie shared their private codatatype-based
+theories with us so we could include them in the benchmarks.
+%
 Blanchette's research was partially supported by the Deutsche

### 2014-12-05 tex/codatadp.tex
Minor cleanup, starting solution soundness.

@@
 \newcommand{\Ec}{\mathsf{E}}
+\newcommand{\Fc}{\mathsf{F}}
 \newcommand{\tEc}{\mathcal{T}(\Ec)}
@@
 \newcommand{\ec}[1]{ [ #1 ] }
+\newcommand{\M}{\mathcal{M}}
+\newcommand{\Val}{\mathcal{A}}
 
 % change?
-\newcommand{\thD}{T_{D}}
+\newcommand{\thO}{T_{o}}
+\newcommand{\thD}{T_{d}}
 
@@
 \rem{Discuss signature $\Sig$, theory $\thD$}
+\rem{DPLL(T) : we can restrict ourselves to conjunctions.}
 This section presents a calculus for determining the satisfiability of conjunctions of $\thD$-constraints.
@@
 the source of an edge in $p$, 
-or does not contain an application of a constructor (including the case when $t$ is not of datatype/codatatype type).
+or does not contain an application of a non-nullary constructor (including the case when $t$ is not of datatype/codatatype type).
 The set $\ttpath{\Ec}{\ec{t_1}}{\ec{t_2}}$ is the smallest set containing all maximal paths with source $\ec{t_1}$ and destination $\ec{t_2}$ that are induced by $\Ec$.
@@
 Let $S$ be the set of terms occurring as the argument of a selector in $\tEc$.
-For each term $t \in \tEc$ of finite datatype type $\tau$,
+Let $T$ be the set of terms in $\tEc$ of finite datatype type.
+For each term $t \in T$,
 let $S^0_t$ be the set $\{ t \}$,
-and for $i \geq 0$ let $S^{i+1}_t =$ $S^i_t \cup \{ s( u ) \mid u : \upsilon \in S^i, \upsilon \in \Data, \upsilon \text{ finite}, s \in \Sel^{\upsilon}  \}$.
-Let $S^\ast_t$ be the fixed point of this sequence for all $t$, which we know is finite since the values of $\tau$ are of finite size.
-Let $S^\ast$ be the union of $S^\ast_t$ for all such $t$,
+for $i \geq 0$ let $S^{i+1}_t =$ $S^i_t \cup \{ s( u ) \mid u : \upsilon \in S^i, \upsilon \in \Data, \upsilon \text{ finite}, s \in \Sel^{\upsilon}  \}$,
+and let $S^\ast_t$ be the fixed point of this sequence.
+We know this is a finite set for each $t$ since the values of $t$'s type are of finite size.
+Let $S^\ast$ be the union of $S^\ast_t$ for all $t \in T$,
 and let $\tcEc$ be the set of subterms of $\Ec \cup \{ C_i( s^1_i( t ), \ldots, s^{a_i}_i( t ) ) \mid t : \tau \in S \cup S^\ast, C_i \in \Ctr^\tau \}$.
 In a derivation tree with root node $\Ec$, 
-it can be shown by case analysis on the rules of the calculus that each non-root node $\Ec'$ is such that 
-$\mathcal{T}(\Ec') \subseteq \tcEc $, and moreover thus is the result of adding an equality between two terms from $\tcEc$ to its parent node.
+it can be shown by case analysis on the rules of the calculus that each non-root node $\Fc$ is such that 
+$\mathcal{T}(\Fc) \subseteq \tcEc $, and moreover thus contains an equality between two terms from $\tcEc$ not occurring in its parent node.
 Thus, the depth of a branch in a derivation tree with root node $\Ec$ is at most $\mid \tcEc \mid^2$,
@@
 \begin{lemma}[Solution Soundness]
-If there exists a derivation tree containing a saturated node with root node $\Ec$, then $\Ec$ is satisfiable in $\thD$.
+If there exists a derivation tree with root node $\Ec$ containing a saturated node, then $\Ec$ is satisfiable in $\thD$.
 \end{lemma}
 \begin{proof}
-TODO
+Let $\Fc$ be a saturated node in a derivation tree with root node $\Ec$.
+%First, we construct a valuation $\Val$ as follows.
+Let $C$ be the set of variables occurring in the same equivalence class as a constructor term,
+and let $X$ be the set of variables occurring in $\mathcal{T}(\Fc)$ but not in $C$.
+
+\qed
 \end{proof}

### 2014-12-08 tex/codatadp.tex
tuned intro

@@
 
+\newcommand\afterLdots{\kern.1em} %% TYPESETTING
+
 % for "bussproofs" package
@@
 \cite{lochbihler-2010-jinja}, and the formalized Java memory model
-\cite{lochbihler-2014-jmm} all rely crucially on codatatypes to capture infinite
-execution traces.
+\cite{lochbihler-2014-jmm} all depend on codatatypes to capture infinite
+processes.
 % as do the CoALP dialect of Prolog and the Dafny specification language.
@@
 Codatatypes are freely generated by their constructors, but in contrast with datatypes,
-infinit\-e constructor terms are also legitimate values of codatatypes
-(Section~\ref{sec:theory-of-co-datatypes}). Intuitively, a
-codatatype consists of all well-typed finite and infinite ground constructor
+infinit\-e constructor terms are also legitimate values for codatatypes
+(Section~\ref{sec:theory-of-co-datatypes}). Intuitively, the
+values of a codatatype consist of all well-typed finite and infinite ground constructor
 terms, and only those. For example, the specification
@@
 %
-(using an ML-like syntax) introduces a type that faithfully
-models the natural numbers extended with an
-infinity $\infty = \const{S}(\const{S}(\const{S}(\ldots)))$. Compared
+(using an ML-like syntax) introduces a type that
+models the natural numbers $\const{Z}$, $\const{S}(\const{Z})$, $\const{S}(\const{S}(\const{Z}))$, $\ldots$\afterLdots{},
+using Peano notation but extended with an
+infinity value $\infty = \const{S}(\const{S}(\const{S}(\ldots)))$. Compared
 with the more conventional, inductive definition
@@
 $\const{S}(\const{S}(\const{S}(\ldots)))$, which uniquely identifies the
-codatatype value.
+codatatype value $\infty$.
 
 Datatypes and codatatypes are an integral part of most popular proof assistants,
-including Agda, Coq, Isabelle, Matita, and PVS. In recent years, more and more
+including Agda, Coq, Isabelle, Matita, and PVS. In recent years, some
 automatic theorem provers have been extended with dedicated reasoning facilities
@@
 designed to be composable via the Nelson--Oppen method \cite{nelson-oppen-1979}.
-For the datatype case, it
-roughly follows the lines of Barrett et al.\ \cite{barrett-et-al-2007}.
-To our knowledge, ours is the first decision procedure for codatatypes.
+For the datatype case, it follows the lines of Barrett et al.\ \cite{barrett-et-al-2007}.
+To our knowledge, our procedure is the first of its kind for the theory of
+codatatypes.
 
@@
 \item
-\emph{No well-foundedness check is necessary for codatatypes.}
-For example, the type of infinite sequences, or streams, over an
-element type $\tau$
+\emph{Codatatypes need not be well-founded.}
+For example, the type
 %
 \[\keyw{codatatype}~\;\ty{stream}_{\,\tau} \,=\, \const{SCons}(\tau,\, \ty{stream}_{\,\tau})\]
-is allowed; the corresponding datatype would be rejected as
-non-well-founded.
+%
+of infinite sequences, or streams, over an element type $\tau$ is allowed; the
+corresponding datatype would be rejected as non-well-founded.
 
@@
 \item
-\emph{It must be possible to express cyclic values as closed terms and
+\emph{It must be possible to express cyclic }(\emph{$\omega$-regular}) \emph{values as closed terms and
 to enumerate them.} This is necessary both for finite model finding (modulo theories)
-and for theory combinations. The $\mu$-binder notation is used to represent cyclic ($\omega$-regular)
+and for theory combinations. The $\mu$-binder notation is used to represent cyclic
 values internally. For example,
 the term $\const{SCons}(1,\, \mu s.\; \const{SCons}(0,\, \const{SCons}(9,\, s)))$
-denotes the lasso-shaped sequence $1, 0, 9, 0, 9, 0, 9, \ldots$\kern0.1em. %%% TYPESETTING: kern
+denotes the lasso-shaped sequence $1, 0, 9, 0, 9, 0, 9, \ldots$\afterLdots.
 \end{itemize}
@@
 exhaustiveness of constructors and the selector laws---it is
-widely recognized that decision procedures are more efficient than
+widely recognized that decision procedures can be more efficient than
 axiomatizations.
@@
 \end{paper}%
-The procedure is implemented in the SMT solver CVC4 as rewriting and a theory solver
+The procedure is implemented in the SMT solver CVC4 as a combination
+of rewriting and a theory solver
 (Section~\ref{sec:the-theory-solver}).
@@
 %
+Andrei Popescu helped clarify our thoughts about axiomatizing codatatypes.
 Andreas Lochbihler and Peter Gammie shared their private codatatype-based

### 2014-12-08 tex/codatadp.tex
different setup for conference and report
more

@@
-%envcountsect,
-\documentclass[a4paper,oribibl,envcountsame,draft]{llncs}
 \usepackage[T1]{fontenc}
@@
 \usepackage{url}
-\usepackage{version}
 \usepackage[all]{xy}
@@
 \newcommand\Codata{\Types_{\mathrm{cdt}}}
-\newcommand\Nondata{\Types_{\mathrm{oth}}}
+\newcommand\Nondata{\Types_{\mathrm{ord}}}
 
@@
 $\Types = \Data \mathrel{\uplus} \Codata \mathrel{\uplus} \Nondata$, where $\Data$ are the
-datatypes, $\Codata$ are the codatatypes, and $\Nondata$ are the remaining,
-plain types. The functions are partitioned into $\Funcs = \Ctr \mathrel{\uplus} \Sel$, where
-$\Ctr$ are the constructors and $\Sel$ are the selectors. (There is no need to
-consider other function symbols, because they can be abstracted away as
-variables when combining theories.)
-
-In an SMT problem, the signature is given by specifying the uninterpreted types
-first (in any order), then the (co)datatypes with their constructors and
-selectors in groups of mutually (co)recursive groups of (co)datatypes, and
+datatypes, $\Codata$ are the codatatypes, and $\Nondata$ are the remaining
+\emph{ordinary} types. The functions are partitioned into $\Funcs = \Ctr
+\mathrel{\uplus} \Sel$, where $\Ctr$ are the constructors and $\Sel$ are the
+selectors. There is no need to consider other function symbols, because they can
+be abstracted away as variables when combining theories.
+%
+In an SMT problem, the signature is normally given by specifying the
+uninterpreted types in any order, the (co)datatypes with their constructors
+and selectors in groups of mutually (co)recursive groups of (co)datatypes, and
 finally any other function symbols.
@@
 In principle, rank-1 (top-level) polymorphism \cite{blanchette-paskevich-2013}
-would be easy to support; nesting datatypes inside datatypes, and
-likewise for codatatypes, can be reduced to the mutual case
+should not raise any special difficulties. Nesting datatypes inside datatypes,
+and likewise for codatatypes, can be reduced to the mutual case
 \cite{gunter-1993-not}. So the only genuinely interesting cases missing are
-mixed nested (co)recursion (as in the \ty{tree} example above) as well as
+mixed nested (co)recursion as well as
 (co)recursion through a non-(co)datatype (both of which make sense
@@
 by using subscripts---for example, $\ty{stream}_{\,\tau}$ denotes a
-family of ground types $\ty{stream}_{\,\ty{int}}$, $\ty{stream}_{\,\ty{bool}}$,
-$\ty{stream}_{\,\ty{stream}_{\,\ty{real}}}$, etc.
-
-Each datatype $\delta$ is equipped by
+family of ground types including
+$\ty{stream}_{\,\ty{int}}$, $\ty{stream}_{\,\ty{bool}}$,
+and \smash{$\ty{stream}_{\,\ty{stream}_{\,\ty{real}}}$}.
+
+Each datatype $\delta$ is equipped with
 $m \ge 1$ constructors, and each constructor for $\delta$ takes zero or more
 arguments and returns a $\delta$ value. The argument types must be either
-plain, be or be one of the newly introduced (co)datatypes.
+ordinary, among the already know (co)datatypes, or among the (co)datatypes being
+introduced.
 %

### 2014-12-08 tex/codatadp.tex
more

@@
 
-\newcommand\cheat{\vskip\smallskipamount} %% TYPESETTING
-
 Datatypes and codatatypes share many of the same properties, so it makes sense
 to consider them together. There are, however, at least three important
-differences:
-%
-\begin{itemize}
-\item
-\emph{Codatatypes need not be well-founded.}
+differences.
+
+First, \emph{codatatypes need not be well-founded.}
 For example, the type
@@
 
-\cheat
-\item
-\emph{A bisimilarity rule takes the place of the cyclicity rule of datatypes.}
+Second, \emph{a bisimilarity rule takes the place of the cyclicity rule of datatypes.}
 Cyclic constraints such as
@@
 
-\cheat
-\item
-\emph{It must be possible to express cyclic }(\emph{$\omega$-regular}) \emph{values as closed terms and
+Third, \emph{it must be possible to express cyclic }(\emph{$\omega$-regular}) \emph{values as closed terms and
 to enumerate them.} This is necessary both for finite model finding (modulo theories)
@@
 denotes the lasso-shaped sequence $1, 0, 9, 0, 9, 0, 9, \ldots$\afterLdots.
-\end{itemize}
 

### 2014-12-08 tex/codatadp.tex
removed footnotes from conf version

@@
 \def\thewordpaper{paper}
+\newcommand\dotReportFootnote[1]{.}
 
@@
 \def\thewordpaper{report}
+\newcommand\dotReportFootnote[1]{.\footnote{#1}}
 \end{report}
@@
 
-% \paragraph{Related Work.}
-% 
+\paragraph{Related Work.}
+Although it was written several years ago, the account of related work in
+Barrett et al.\ \cite{barrett-et-al-2007} is still largely up to date.
+Since then, datatypes have been added to Z3 in unpublished work by Leonardo de
+Moura and to a SPASS-like prototype prover called Pirate by Daniel Wand.
+%\cite{wand-weidenbach-201x}.
+Closely related is the support for structural
+induction in CVC4 by Reynolds and Kuncak \cite{reynolds-kuncak-2015} and in
+Pirate by Wand and Weidenbach \cite{wand-weidenbach-201x}, which depends on a
+notion of datatype.
+
 %     * point to Barrett et al. for SMT datatypes
@@
 (co)recursion, and datatype--codatatype mixtures fall outside this fragment.%
-\footnote{%
+\dotReportFootnote{%
 In principle, rank-1 (top-level) polymorphism \cite{blanchette-paskevich-2013}
@@
 constructors, and the selectors must be distinct and different from already
-introduced ones.%
-\footnote{It can be convenient to specify the same selector
+introduced ones%
+\dotReportFootnote{It can be convenient to specify the same selector
 for several constructors associated with the same (co)data\-type,

### 2014-12-08 tex/codatadp.tex
more edits

@@
 $\Types = \Data \mathrel{\uplus} \Codata \mathrel{\uplus} \Nondata$, where $\Data$ are the
-datatypes, $\Codata$ are the codatatypes, and $\Nondata$ are the remaining
-\emph{ordinary} types. The functions are partitioned into $\Funcs = \Ctr
-\mathrel{\uplus} \Sel$, where $\Ctr$ are the constructors and $\Sel$ are the
-selectors. There is no need to consider other function symbols, because they can
+\emph{datatypes}, $\Codata$ are the \emph{codatatypes}, and $\Nondata$ are the remaining
+\emph{ordinary types}. The functions are partitioned into $\Funcs = \Ctr
+\mathrel{\uplus} \Sel$, where $\Ctr$ are the \emph{constructors} and $\Sel$ are the
+\emph{selectors}; there is no need to consider other function symbols, because they can
 be abstracted away as variables when combining theories.
@@
 either all datatypes or all codatatypes. Polymorphic types, nested
-(co)recursion, and datatype--codatatype mixtures fall outside this fragment.%
+(co)recursion, and datatype--codatatype mixtures fall outside this fragment%
 \dotReportFootnote{%
@@
 % at the metalevel
-by using subscripts---for example, $\ty{stream}_{\,\tau}$ denotes a
+through subscripts---for example, $\ty{stream}_{\,\tau}$ denotes a
 family of ground types including
@@
 For types with several constructors, it is customary to provide discriminators
-$\const{d} : \const{s}_{i\negvthinspace j} : \delta_i \to \ty{bool}$. However,
-they are not strictly necessary:
-To simplify the presentation of the decision procedure, we treat $\const{d}_{i\negvthinspace
-j}(x)$ as an abbreviation for
-$x = \const{C}_{i\negvthinspace j}(\const{s}_{i\negvthinspace j}^1(x), \ldots, \const{s}_{i\negvthinspace j}^{\,n_{\smash{{i\negvthinspace j}}}}(x))$
-instead of extending the signature.
+$\const{d}_{i\negvthinspace j} : \delta_i \to \ty{bool}$. However,
+it is not necessary to extend the signature:
+$\const{d}_{i\negvthinspace j}(x)$ can be seen as an abbreviation for
+$x = \const{C}_{i\negvthinspace j}\bigl(\const{s}_{i\negvthinspace j}^1(x), \ldots, \const{s}_{i\negvthinspace j}^{\,n_{\smash{{i\negvthinspace j}}}}(x)\bigr)$.
+This will simplify the presentation of the decision procedure.
 
@@
 Because the types are inhabited (nonempty), a datatype specification is
-admissible only if a ground term can be exhibited that belongs to that datatype.
-%This rules out ill-founded specifications such as
-%
-%\[\keyw{datatype}\;\, \ty{fstream} \,=\, \const{FSCons}(\ty{int},\,\ty{fstream})\]
-%
+admissible only if a ground constructor term can be exhibited that belongs to
+that datatype. This rules out ill-founded specifications such as
+\[\keyw{datatype}\;\, \ty{fstream} \,=\, \const{FSCons}(\ty{int},\,\ty{fstream})\]
 For codatatypes, no such check is necessary because there is always a term,
@@
 %
-\[\keyw{codatatype}\;\, \ty{stream} \,=\, \const{SCons}(\ty{int},\,\ty{stream})\]
+%\[\keyw{codatatype}\;\, \ty{stream} \,=\, \const{SCons}(\ty{int},\,\ty{stream})\]
 %

### 2014-12-08 tex/codatadp.tex
Work on solution soundness.

--- 2014-12-08 tex/codatadp.tex
@@
+%envcountsect,
+\documentclass[a4paper,oribibl,envcountsame,draft]{llncs}
 \usepackage[T1]{fontenc}
@@
 \usepackage{url}
+\usepackage{version}
 \usepackage[all]{xy}
@@
 \end{report}
-
-\newcommand\afterLdots{\kern.1em} %% TYPESETTING
 
@@
 
-\renewcommand{\vec}[1]{\boldsymbol{#1}}
+\renewcommand{\vec}[1]{\mathbf{#1}}
 \newcommand{\Ec}{\mathsf{E}}
 \newcommand{\Fc}{\mathsf{F}}
+\newcommand{\Gc}{\mathsf{G}}
+\newcommand{\Ac}{\mathsf{A}}
+\newcommand{\Dc}{\mathsf{D}}
+\newcommand{\Rc}{\mathsf{R}}
 \newcommand{\tEc}{\mathcal{T}(\Ec)}
@@
 \newcommand{\ttpath}[3]{\mathcal{P}_{ #2 \rightarrow #3 }( #1 )}
+\newcommand{\dpath}[3]{\delta^{#1}_{#2}( #3 )}
 \newcommand{\ec}[1]{ [ #1 ] }
@@
 
+\def\AIF{\qtab\keyword{if}\ }
+\def\ATHEN{\ \keyword{then}\ }
+\def\AELSE{\untab\qtab\keyword{else}\ }
+\def\AFI{\untab}
+\def\ARETURN{\keyword{return}\ }
+\def\AENDPROC{\untab}
+
 % change?
@@
 \newcommand\Codata{\Types_{\mathrm{cdt}}}
-\newcommand\Nondata{\Types_{\mathrm{ord}}}
+\newcommand\Nondata{\Types_{\mathrm{oth}}}
 
@@
    bookmarksopen]{hyperref}
+%\usepackage{program}
 
@@
 \cite{lochbihler-2010-jinja}, and the formalized Java memory model
-\cite{lochbihler-2014-jmm} all depend on codatatypes to capture infinite
-processes.
+\cite{lochbihler-2014-jmm} all rely crucially on codatatypes to capture infinite
+execution traces.
 % as do the CoALP dialect of Prolog and the Dafny specification language.
@@
 Codatatypes are freely generated by their constructors, but in contrast with datatypes,
-infinit\-e constructor terms are also legitimate values for codatatypes
-(Section~\ref{sec:theory-of-co-datatypes}). Intuitively, the
-values of a codatatype consist of all well-typed finite and infinite ground constructor
+infinit\-e constructor terms are also legitimate values of codatatypes
+(Section~\ref{sec:theory-of-co-datatypes}). Intuitively, a
+codatatype consists of all well-typed finite and infinite ground constructor
 terms, and only those. For example, the specification
@@
 %
-(using an ML-like syntax) introduces a type that
-models the natural numbers $\const{Z}$, $\const{S}(\const{Z})$, $\const{S}(\const{S}(\const{Z}))$, $\ldots$\afterLdots{},
-using Peano notation but extended with an
-infinity value $\infty = \const{S}(\const{S}(\const{S}(\ldots)))$. Compared
+(using an ML-like syntax) introduces a type that faithfully
+models the natural numbers extended with an
+infinity $\infty = \const{S}(\const{S}(\const{S}(\ldots)))$. Compared
 with the more conventional, inductive definition
@@
 $\const{S}(\const{S}(\const{S}(\ldots)))$, which uniquely identifies the
-codatatype value $\infty$.
+codatatype value.
 
 Datatypes and codatatypes are an integral part of most popular proof assistants,
-including Agda, Coq, Isabelle, Matita, and PVS. In recent years, some
+including Agda, Coq, Isabelle, Matita, and PVS. In recent years, more and more
 automatic theorem provers have been extended with dedicated reasoning facilities
@@
 designed to be composable via the Nelson--Oppen method \cite{nelson-oppen-1979}.
-For the datatype case, it follows the lines of Barrett et al.\ \cite{barrett-et-al-2007}.
-To our knowledge, our procedure is the first of its kind for the theory of
-codatatypes.
+For the datatype case, it
+roughly follows the lines of Barrett et al.\ \cite{barrett-et-al-2007}.
+To our knowledge, ours is the first decision procedure for codatatypes.
+
+\newcommand\cheat{\vskip\smallskipamount} %% TYPESETTING
 
@@
 to consider them together. There are, however, at least three important
-differences.
-
-First, \emph{codatatypes need not be well-founded.}
-For example, the type
+differences:
+%
+\begin{itemize}
+\item
+\emph{No well-foundedness check is necessary for codatatypes.}
+For example, the type of infinite sequences, or streams, over an
+element type $\tau$
 %
 \[\keyw{codatatype}~\;\ty{stream}_{\,\tau} \,=\, \const{SCons}(\tau,\, \ty{stream}_{\,\tau})\]
-%
-of infinite sequences, or streams, over an element type $\tau$ is allowed; the
-corresponding datatype would be rejected as non-well-founded.
-
-Second, \emph{a bisimilarity rule takes the place of the cyclicity rule of datatypes.}
+is allowed; the corresponding datatype would be rejected as
+non-well-founded.
+
+\cheat
+\item
+\emph{A bisimilarity rule takes the place of the cyclicity rule of datatypes.}
 Cyclic constraints such as
@@
 
-Third, \emph{it must be possible to express cyclic }(\emph{$\omega$-regular}) \emph{values as closed terms and
+\cheat
+\item
+\emph{It must be possible to express cyclic values as closed terms and
 to enumerate them.} This is necessary both for finite model finding (modulo theories)
-and for theory combinations. The $\mu$-binder notation is used to represent cyclic
+and for theory combinations. The $\mu$-binder notation is used to represent cyclic ($\omega$-regular)
 values internally. For example,
 the term $\const{SCons}(1,\, \mu s.\; \const{SCons}(0,\, \const{SCons}(9,\, s)))$
-denotes the lasso-shaped sequence $1, 0, 9, 0, 9, 0, 9, \ldots$\afterLdots.
+denotes the lasso-shaped sequence $1, 0, 9, 0, 9, 0, 9, \ldots$\kern0.1em. %%% TYPESETTING: kern
+\end{itemize}
 
@@
 exhaustiveness of constructors and the selector laws---it is
-widely recognized that decision procedures can be more efficient than
+widely recognized that decision procedures are more efficient than
 axiomatizations.
@@
 \end{paper}%
-The procedure is implemented in the SMT solver CVC4 as a combination
-of rewriting and a theory solver
+The procedure is implemented in the SMT solver CVC4 as rewriting and a theory solver
 (Section~\ref{sec:the-theory-solver}).
@@
 $\Types = \Data \mathrel{\uplus} \Codata \mathrel{\uplus} \Nondata$, where $\Data$ are the
-datatypes, $\Codata$ are the codatatypes, and $\Nondata$ are the remaining
-\emph{ordinary} types. The functions are partitioned into $\Funcs = \Ctr
-\mathrel{\uplus} \Sel$, where $\Ctr$ are the constructors and $\Sel$ are the
-selectors. There is no need to consider other function symbols, because they can
-be abstracted away as variables when combining theories.
-%
-In an SMT problem, the signature is normally given by specifying the
-uninterpreted types in any order, the (co)datatypes with their constructors
-and selectors in groups of mutually (co)recursive groups of (co)datatypes, and
+datatypes, $\Codata$ are the codatatypes, and $\Nondata$ are the remaining,
+plain types. The functions are partitioned into $\Funcs = \Ctr \mathrel{\uplus} \Sel$, where
+$\Ctr$ are the constructors and $\Sel$ are the selectors. (There is no need to
+consider other function symbols, because they can be abstracted away as
+variables when combining theories.)
+
+In an SMT problem, the signature is given by specifying the uninterpreted types
+first (in any order), then the (co)datatypes with their constructors and
+selectors in groups of mutually (co)recursive groups of (co)datatypes, and
 finally any other function symbols.
@@
 In principle, rank-1 (top-level) polymorphism \cite{blanchette-paskevich-2013}
-should not raise any special difficulties. Nesting datatypes inside datatypes,
-and likewise for codatatypes, can be reduced to the mutual case
+would be easy to support; nesting datatypes inside datatypes, and
+likewise for codatatypes, can be reduced to the mutual case
 \cite{gunter-1993-not}. So the only genuinely interesting cases missing are
-mixed nested (co)recursion as well as
+mixed nested (co)recursion (as in the \ty{tree} example above) as well as
 (co)recursion through a non-(co)datatype (both of which make sense
@@
 by using subscripts---for example, $\ty{stream}_{\,\tau}$ denotes a
-family of ground types including
-$\ty{stream}_{\,\ty{int}}$, $\ty{stream}_{\,\ty{bool}}$,
-and \smash{$\ty{stream}_{\,\ty{stream}_{\,\ty{real}}}$}.
-
-Each datatype $\delta$ is equipped with
+family of ground types $\ty{stream}_{\,\ty{int}}$, $\ty{stream}_{\,\ty{bool}}$,
+$\ty{stream}_{\,\ty{stream}_{\,\ty{real}}}$, etc.
+
+Each datatype $\delta$ is equipped by
 $m \ge 1$ constructors, and each constructor for $\delta$ takes zero or more
 arguments and returns a $\delta$ value. The argument types must be either
-ordinary, among the already know (co)datatypes, or among the (co)datatypes being
-introduced.
+plain, be or be one of the newly introduced (co)datatypes.
 %
@@
 as shorthand for the path containing an edge from $\ec{t_1}$ to $\ec{t_2}$, followed by one from $\ec{t_2}$ to $\ec{t_3}$.
+%A set of paths is \emph{disjoint} if it does not contain two paths $p_1$ and $p_2$ such that $p_1 = p_2 \sigma$ for some $\sigma$.
 
@@
 it can be shown by case analysis on the rules of the calculus that each non-root node $\Fc$ is such that 
-$\mathcal{T}(\Fc) \subseteq \tcEc $, and moreover thus contains an equality between two terms from $\tcEc$ not occurring in its parent node.
+$\mathcal{T}(\Fc) \subseteq \tcEc$, and moreover thus contains an equality between two terms from $\tcEc$ not occurring in its parent node.
 Thus, the depth of a branch in a derivation tree with root node $\Ec$ is at most $\mid \tcEc \mid^2$,
@@
 
+It remains to show the inverse of the previous lemma, that when a derivation tree with root node $\Ec$ contains a saturated node $\Fc$,
+then $\Ec$ is satisfiable $\thD$.
+
+%We first show that for any saturated node $\Fc$, 
+%additional equalities $\Gc$ can be added to $\Fc$ such that each codatatype/datatype equivalence class of $\Fc \cup \Gc$ has a constructor term.
+%For this purpose, we extend the calculus in this section.
+%For a subset of equivalence classes $S$ of $\Ec$, write $\dpath{\Ec}{C,n}{S}$ to denote the set
+%such that $e' \in \dpath{\Ec}{C,n}{S}$ if and only if there exists an $e \in S$ such that $e \rightarrow^{C}_n e'$ is induced by $\Ec$.
+
+\begin{comment}
+\begin{figure}[t]
+\centering
+\begin{tabular}{c}
+\rn{O-Finish}
+\(
+\inferrule{
+  \Ac := x
+  \quad
+  x : \tau
+  \quad
+  \tau \in \Nondata
+  \quad
+  c \not\in \tEc
+}{
+  \Ac := \emptyset \quad \Ec := \Ec \cup \{ x \teq c \}
+}
+\)
+\\[3.7ex]
+\rn{C-Finish}
+\(
+\inferrule{
+  \Ac := x
+  \quad
+  x : \tau
+  \quad
+  \tau \in \Codata
+  \quad
+  C_i \in \Ctr^\tau
+  \quad 
+  C_i \not\in \{ C_j \mid d \in \Dc, C_j( \vec t ) \in d \}
+}{
+  \Ac := \emptyset \quad \Ec := \Ec \cup \{ x \teq C_i( \vec y ) \} \quad \Rc := \Rc \cup \vec y
+}
+\)
+\\[3.7ex]
+\rn{C-Rec}
+\(
+\inferrule{
+  \Ac := x
+  \quad
+  x : \tau
+  \quad
+  \tau \in \Codata
+  \quad
+  C_i \in \Ctr^\tau
+  \quad
+  C_i \text{ infinite arg } j
+}{
+  \Ac := y_j \quad 
+  \Ec := \Ec \cup \{ x \teq C_i( \vec y ) \} \quad 
+  \Rc := \Rc \cup ( \vec y \setminus y_j ) \quad 
+  \Dc := \dpath{\Ec}{C_i,j}{\Dc}
+}
+\)
+\end{tabular}
+\caption{Rules for completion.
+}
+\label{fig:completion-rules}
+\end{figure}
+\end{comment}
+
+%and second show how to construct a valuation for the variables of $\Fc \cup \Gc$.
+
+%\begin{lemma}
+%\end{lemma}
+
+%Given a set of disjoint paths $P$ where each $p \in P$
+%we write $\interp{P}$ to denote the
+
 \begin{lemma}[Solution Soundness]
-If there exists a derivation tree with root node $\Ec$ containing a saturated node, then $\Ec$ is satisfiable in $\thD$.
+If there exists a derivation tree with root node $\Ec$ containing a saturated node $\Fc$, then $\Ec$ is satisfiable in $\thD$.
 \end{lemma}
@@
 %First, we construct a valuation $\Val$ as follows.
-Let $C$ be the set of variables occurring in the same equivalence class as a constructor term,
-and let $X$ be the set of variables occurring in $\mathcal{T}(\Fc)$ but not in $C$.
-
+%Let $C$ be the set of variables occurring in the same equivalence class as a constructor term,
+%and let $X$ be the set of variables occurring in $\mathcal{T}(\Fc)$ but not in $C$.
+
+We assume that:
+\begin{enumerate}
+\item Each equivalence class of $\Fc$ contains at least one variable,
+\item Each equivalence class $\ec{x}$ of $\Fc$ where $x : \tau$, $\tau \in \Nondata$ contains a constant,
+\item Each equivalence class $\ec{x}$ of $\Fc$ where $x : \tau$, $\tau \in \Data \cup \Codata$ contains a constructor term.
+\end{enumerate}
+The first statement comes with no loss of generality since new equalities of the form $y \teq t$ for fresh variable $y$ can be added to $\Fc$ without affecting its satisfiability.
+The second statement can be assumed for $x : \tau$, since we know that $\tau$ is infinite.
+The third statement can be assumed by \rem{TODO}.
+
+We construct a model $\M$ of $\thD$ that interprets constructors and correctly-applied selectors in the obvious way,
+interprets wrongly-applied selectors based on information in $\Fc$,
+interprets the variables and $x$ of $\Fc$ based on a valuation $\Val$ which construct as follows.
+For each equivalence class $e : \tau$, where $\tau \in \Nondata$ containing constant $c$,
+$\Val(x) = c$ for each $x \in e$.
+For each $e : \tau$, where $\tau \in \Data$ containing the constructor term $C( \vec u )$ where the equivalence classes $\ec{\vec u}$ contains the variables $\vec y$,
+$\Val(x) = C( \Val( \vec y ) )$ for each $x \in e$.
+Since \rn{Cyclic} does not apply, this definition is well-defined, and since \rn{Cong} does not apply, we have that $\Val{x} = \Val{y}$ if and only if $\ec{x} = \ec{y}$.
+For each $e : \tau$, where $\tau \in \Codata$,
+\rem{TODO}
+
+We show that all equalities and disequalities in $\Fc$ are satisfied by $\M$.
+Say $t_1 \teq t_2 \in \Fc$ where $t_1$ and $t_2$ have type $\tau$ where $\tau \not\in \Codata$,
+and assume that $\ec{t_1}$ contains the variable $x$.
+We show that $\M( t ) = \Val( x )$ for each term $t \in \ec{x}$.
+In the case where $t$ is a variable, then $\M( t ) = \Val( x )$ since $t$ and $x$ are assigned identical values by construction.
+In the case where $t$ is a constructor term $C( \vec u )$, say our construction built a valuation for $x$ based on the constructor term $C'( \vec u' ) \in \ec{x}$.
+Since \rn{Unify$_2$} does not apply to $\Fc$, we have that $C = C'$.
+Since \rn{Unify$_1$} does not apply to $\Fc$, we have that $\ec{\vec u} = \ec{\vec u'}$.
+By structural induction, we have $\M( \vec u ) = \M( \vec u' )$, and thus $\M( t ) = \M( C( \vec u ) ) = \M( C'( \vec u ' ) ) = \Val( x )$.
+In the case where $t$ is a selector term $s^j_k( u )$,
+then since \rn{Split} does not apply, 
+$\ec{u}$ must contain a term of the form $C_i( s^1_i( u ), \ldots, s^{a_i}_i( u ) )$ for some $i$.
+Let $x_1, \ldots, x_n$ be the variables in equivalence classes $\ec{s^1_i( u )}, \ldots, \ec{s^{a_i}_i( u )}$.
+By our construction, $\M( u ) = C_i( \Val( x_1 ), \ldots, \Val( x_{a_i} ) )$.
+If $i = k$, then $\M( t ) = \M( s^j_k( u ) ) = \Val( x_j )$, which is $\Val( x )$ since $x_j \in \ec{s^j_i( u )} = \ec{x}$.
+If $i \neq k$, then $\thD$ does not constraint the value $\M( t )$, and thus we interpret $s^j_k( \M( u ) )$ as $\Val( x )$.
+
+%If $t$ and $s$ are both variables, then this follows since $\ec{t} = \ec{s}$.
+
+Let $t \tneq s \in \Fc$.
+Since \rn{Conflict} does not apply, we know $t \teq s \not\in \Fc$ and thus $\ec{t}$ and $\ec{s}$ are distinct.
+Since by above, $\M( t ) = \M( s )$ if and only if $\ec{t}=\ec{s}$, we have $\M( t ) \neq \M( s )$, and thus $\M \models t \tneq s$.
+
+By these two, 
+%$\M \models_{\thD} \Fc$, and since $\Fc$ is a superset of $\Ec$, we have that $\Ec$ is satisfied by $\M$.
 \qed
@@
 %
-Andrei Popescu helped clarify our thoughts about axiomatizing codatatypes.
 Andreas Lochbihler and Peter Gammie shared their private codatatype-based

### 2014-12-08 tex/codatadp.tex
Cleanup, more detail in proof.

@@
 \newcommand{\ttpath}[3]{\mathcal{P}_{ #2 \rightarrow #3 }( #1 )}
+\newcommand{\interp}[1]{[\![ #1 ]\!]}
 \newcommand{\dpath}[3]{\delta^{#1}_{#2}( #3 )}
@@
 
-%Given a set of disjoint paths $P$ where each $p \in P$
-%we write $\interp{P}$ to denote the
+Given a set of disjoint paths $P$ where each $p \in P$
+we write $\interp{P}$ to denote the value
 
@@
 Let $\Fc$ be a saturated node in a derivation tree with root node $\Ec$.
-%First, we construct a valuation $\Val$ as follows.
-%Let $C$ be the set of variables occurring in the same equivalence class as a constructor term,
-%and let $X$ be the set of variables occurring in $\mathcal{T}(\Fc)$ but not in $C$.
-
 We assume that:
@@
 interprets wrongly-applied selectors based on information in $\Fc$,
-interprets the variables and $x$ of $\Fc$ based on a valuation $\Val$ which construct as follows.
+interprets the variables and $x$ of $\Fc$ based on a valuation $\Val$, which is constructed as follows.
 For each equivalence class $e : \tau$, where $\tau \in \Nondata$ containing constant $c$,
@@
 $\Val(x) = C( \Val( \vec y ) )$ for each $x \in e$.
-Since \rn{Cyclic} does not apply, this definition is well-defined, and since \rn{Cong} does not apply, we have that $\Val{x} = \Val{y}$ if and only if $\ec{x} = \ec{y}$.
+Since \rn{Cyclic} does not apply, this definition is well-defined. 
 For each $e : \tau$, where $\tau \in \Codata$, 
-\rem{TODO}
-
-We show that all equalities and disequalities in $\Fc$ are satisfied by $\M$.
-Say $t_1 \teq t_2 \in \Fc$ where $t_1$ and $t_2$ have type $\tau$ where $\tau \not\in \Codata$,
-and assume that $\ec{t_1}$ contains the variable $x$.
-We show that $\M( t ) = \Val( x )$ for each term $t \in \ec{x}$.
+$\Val( x ) = \interp{ \tpath{\Fc}{e} }$ for each $x \in e$.
+
+Since \rn{Cong} does not apply, we have that:
+\begin{eqnarray} \label{eqn:m-bij-dt}
+\Val(x) = \Val(y) \text{ if and only if } \ec{x} = \ec{y}
+\end{eqnarray}
+
+We first show $\M$ satisfies all equalities in $t_1 \teq t_2 \in \Fc$, where $t_1$ and $t_2$ have type $\tau$.
+
+If $\tau \in \Nondata$, then $\M( t_1 ) = \M( t_2 )$ by construction.
+
+If $\tau \in \Data$, then assume the equivalence class of $t_1$ and $t_2$ contains variable $x$.
+We show that $\M( t ) = \Val( x )$ for each term $t \in \ec{x}$, which implies $\M \models t_1 \teq t_2$.
 In the case where $t$ is a variable, then $\M( t ) = \Val( x )$ since $t$ and $x$ are assigned identical values by construction.
@@
 By our construction, $\M( u ) = C_i( \Val( x_1 ), \ldots, \Val( x_{a_i} ) )$.
-If $i = k$, then $\M( t ) = \M( s^j_k( u ) ) = \Val( x_j )$, which is $\Val( x )$ since $x_j \in \ec{s^j_i( u )} = \ec{x}$.
-If $i \neq k$, then $\thD$ does not constraint the value $\M( t )$, and thus we interpret $s^j_k( \M( u ) )$ as $\Val( x )$.
-
-%If $t$ and $s$ are both variables, then this follows since $\ec{t} = \ec{s}$.
-
-Let $t \tneq s \in \Fc$.
-Since \rn{Conflict} does not apply, we know $t \teq s \not\in \Fc$ and thus $\ec{t}$ and $\ec{s}$ are distinct.
-Since by above, $\M( t ) = \M( s )$ if and only if $\ec{t}=\ec{s}$, we have $\M( t ) \neq \M( s )$, and thus $\M \models t \tneq s$.
-
-By these two, 
-%$\M \models_{\thD} \Fc$, and since $\Fc$ is a superset of $\Ec$, we have that $\Ec$ is satisfied by $\M$.
+If $i = k$, then $\M( t ) = \M( s^j_k( u ) ) = s^j_k( \M( u ) ) = \Val( x_j )$, which is $\Val( x )$ since $x_j \in \ec{s^j_i( u )} = \ec{x}$.
+If $i \neq k$, then $\thD$ does not constraint the value of $\M( t )$ and since $s^j_k( \M( u ) )$ is otherwise unconstrained in our model construction since \rn{Cong} does not apply, 
+we interpret $s^j_k( \M( u ) )$ as $\Val( x )$.
+
+If $\tau \in \Codata$, \rem{TODO}
+
+We show that all disequalities in $\Fc$ are satisfied by $\M$.
+Say $t_1 \tneq t_2 \in \Fc$.
+Since \rn{Conflict} does not apply, we know $t_1 \teq t_2 \not\in \Fc$ and thus $\ec{t_1}$ and $\ec{t_2}$ are distinct.
+By the previous paragraph and statement~\ref{eqn:m-bij-dt}, $\M( t_1 ) \neq \M( t_2 )$, and thus $\M \models t_1 \tneq t_2$.
+
+Thus, $\M \models \Fc$
+and since $\Fc$ is a superset of $\Ec$, we have that $\Ec$ is satisfied by $\M$.
 \qed

### 2014-12-08 tex/codatadp.tex
Clarify a few points.

@@
 
-\paragraph{Correctness.}
-
 A \emph{derivation tree} is a tree whose nodes are sets of equalities, where non-root nodes are obtained by 
@@
 A node is \emph{saturated} if no non-redundant application of a rule can be applied to it.
-
-\begin{lemma}[Termination]
+If there exists a closed derivation tree with root node $\Ec$, then we conclude that $\Ec$ is $\thD$-unsatisfiable.
+If there exists a derivation tree with root node $\Ec$ that contains a saturated node, then we conclude that $\Ec$ is $\thD$-satisfiable.
+
+\paragraph{Correctness.}
+
+\begin{lemma}[Termination] \label{lem:t}
 All derivation trees are finite.
@@
 
-\begin{lemma}[Refutation Soundness]
+\begin{lemma}[Refutation Soundness] \label{lem:rs}
 If there exists a closed derivation tree with root node $\Ec$, then $\Ec$ is unsatisfiable in $\thD$.
@@
 
-\begin{lemma}[Solution Soundness]
+\begin{lemma}[Solution Soundness] \label{lem:ss}
 If there exists a derivation tree with root node $\Ec$ containing a saturated node $\Fc$, then $\Ec$ is satisfiable in $\thD$.
@@
 The first statement comes with no loss of generality since new equalities of the form $y \teq t$ for fresh variable $y$ can be added to $\Fc$ without affecting its satisfiability.
-The second statement can be assumed for $x : \tau$, since we know that $\tau$ is infinite.
+The second statement can be assumed for $x : \tau$, since $\tau$ is infinite.
+Hence, we can assign constants for each such equivalence class.
+Moreover, since $\Fc$ contains $\thD$-constraints only, these constants can be chosen arbitrarily.
 The third statement can be assumed by \rem{TODO}.
@@
 We construct a model $\M$ of $\thD$ that interprets constructors and correctly-applied selectors in the obvious way,
-interprets wrongly-applied selectors based on information in $\Fc$,
+interprets wrongly-applied selectors based on equality information in $\Fc$,
 interprets the variables and $x$ of $\Fc$ based on a valuation $\Val$, which is constructed as follows.
@@
 $\Val( x ) = \interp{ \tpath{\Fc}{e} }$ for each $x \in e$.
-
-Since \rn{Cong} does not apply, we have that:
+Since neither \rn{Cong} or \rn{Bisimilar} applies to $\Fc$, we have that for all variables $x$ and $y$:
 \begin{eqnarray} \label{eqn:m-bij-dt}
@@
 If $\tau \in \Data$, then assume the equivalence class of $t_1$ and $t_2$ contains variable $x$.
-We show that $\M( t ) = \Val( x )$ for each term $t \in \ec{x}$, which implies $\M \models t_1 \teq t_2$.
+By induction on the size of $\Val( x )$, we show that $\M( t ) = \Val( x )$ for each term $t \in \ec{x}$, which implies $\M \models t_1 \teq t_2$.
 In the case where $t$ is a variable, then $\M( t ) = \Val( x )$ since $t$ and $x$ are assigned identical values by construction.
@@
 Since \rn{Unify$_1$} does not apply to $\Fc$, we have that $\ec{\vec u} = \ec{\vec u'}$.
-By structural induction, we have $\M( \vec u ) = \M( \vec u' )$, and thus $\M( t ) = \M( C( \vec u ) ) = \M( C'( \vec u ' ) ) = \Val( x )$.
+By the induction hypothesis, we have $\M( \vec u ) = \M( \vec u' )$, and thus $\M( t ) = C( \M( \vec u ) ) = C'( \M( \vec u ' ) ) = \Val( x )$.
 In the case where $t$ is a selector term $s^j_k( u )$,
@@
 If $i \neq k$, then $\thD$ does not constraint the value of $\M( t )$ and since $s^j_k( \M( u ) )$ is otherwise unconstrained in our model construction since \rn{Cong} does not apply, 
-we interpret $s^j_k( \M( u ) )$ as $\Val( x )$.
+we may interpret $s^j_k( \M( u ) )$ as $\Val( x )$.
 
@@
 
-\rem{Thus, sound and complete.}
+By Lemmas~\ref{lem:t},~\ref{lem:rs}, and~\ref{lem:ss}, our calculus is sound and complete for $\thD$.
 

### 2014-12-09 tex/codatadp.tex
minor

--- 2014-12-08 tex/codatadp.tex
@@
 \def\thewordpaper{report}
+\newcommand\dotReportFootnote[1]{.\footnote{#1}}
 \end{report}
+
+\newcommand\afterLdots{\kern.1em} %% TYPESETTING
 
@@
 
-\renewcommand{\vec}[1]{\mathbf{#1}}
+\renewcommand{\vec}[1]{\boldsymbol{#1}}
 \newcommand{\Ec}{\mathsf{E}}
 \newcommand{\Fc}{\mathsf{F}}
-\newcommand{\Gc}{\mathsf{G}}
-\newcommand{\Ac}{\mathsf{A}}
-\newcommand{\Dc}{\mathsf{D}}
-\newcommand{\Rc}{\mathsf{R}}
 \newcommand{\tEc}{\mathcal{T}(\Ec)}
@@
 \newcommand{\ttpath}[3]{\mathcal{P}_{ #2 \rightarrow #3 }( #1 )}
-\newcommand{\dpath}[3]{\delta^{#1}_{#2}( #3 )}
 \newcommand{\ec}[1]{ [ #1 ] }
@@
 
-\def\AIF{\qtab\keyword{if}\ }
-\def\ATHEN{\ \keyword{then}\ }
-\def\AELSE{\untab\qtab\keyword{else}\ }
-\def\AFI{\untab}
-\def\ARETURN{\keyword{return}\ }
-\def\AENDPROC{\untab}
-
 % change?
@@
    bookmarksopen]{hyperref}
-%\usepackage{program}
 
@@
 designed to be composable via the Nelson--Oppen method \cite{nelson-oppen-1979}.
-For the datatype case, it
-roughly follows the lines of Barrett et al.\ \cite{barrett-et-al-2007}.
-To our knowledge, ours is the first decision procedure for codatatypes.
-
-\newcommand\cheat{\vskip\smallskipamount} %% TYPESETTING
+For the datatype case, it follows the lines of Barrett et al.\ \cite{barrett-et-al-2007}.
+To our knowledge, our procedure is the first of its kind for the theory of
+codatatypes.
 
@@
 to consider them together. There are, however, at least three important
-differences:
-%
-\begin{itemize}
-\item
-\emph{No well-foundedness check is necessary for codatatypes.}
-For example, the type of infinite sequences, or streams, over an
-element type $\tau$
+differences.
+
+First, \emph{codatatypes need not be well-founded.}
+For example, the type
 %
 \[\keyw{codatatype}~\;\ty{stream}_{\,\tau} \,=\, \const{SCons}(\tau,\, \ty{stream}_{\,\tau})\]
-is allowed; the corresponding datatype would be rejected as
-non-well-founded.
-
-\cheat
-\item
-\emph{A bisimilarity rule takes the place of the cyclicity rule of datatypes.}
+%
+of infinite sequences, or streams, over an element type $\tau$ is allowed; the
+corresponding datatype would be rejected as non-well-founded.
+
+Second, \emph{a bisimilarity rule takes the place of the cyclicity rule of datatypes.}
 Cyclic constraints such as
@@
 
-\cheat
-\item
-\emph{It must be possible to express cyclic values as closed terms and
+Third, \emph{it must be possible to express cyclic }(\emph{$\omega$-regular}) \emph{values as closed terms and
 to enumerate them.} This is necessary both for finite model finding (modulo theories)
-and for theory combinations. The $\mu$-binder notation is used to represent cyclic ($\omega$-regular)
+and for theory combinations. The $\mu$-binder notation is used to represent cyclic
 values internally. For example,
 the term $\const{SCons}(1,\, \mu s.\; \const{SCons}(0,\, \const{SCons}(9,\, s)))$
-denotes the lasso-shaped sequence $1, 0, 9, 0, 9, 0, 9, \ldots$\kern0.1em. %%% TYPESETTING: kern
-\end{itemize}
+denotes the lasso-shaped sequence $1, 0, 9, 0, 9, 0, 9, \ldots$\afterLdots.
 
@@
 $\Types = \Data \mathrel{\uplus} \Codata \mathrel{\uplus} \Nondata$, where $\Data$ are the
-datatypes, $\Codata$ are the codatatypes, and $\Nondata$ are the remaining,
-plain types. The functions are partitioned into $\Funcs = \Ctr \mathrel{\uplus} \Sel$, where
-$\Ctr$ are the constructors and $\Sel$ are the selectors. (There is no need to
-consider other function symbols, because they can be abstracted away as
-variables when combining theories.)
-
-In an SMT problem, the signature is given by specifying the uninterpreted types
-first (in any order), then the (co)datatypes with their constructors and
-selectors in groups of mutually (co)recursive groups of (co)datatypes, and
+\emph{datatypes}, $\Codata$ are the \emph{codatatypes}, and $\Nondata$ are the remaining
+\emph{ordinary types}. The functions are partitioned into $\Funcs = \Ctr
+\mathrel{\uplus} \Sel$, where $\Ctr$ are the \emph{constructors} and $\Sel$ are the
+\emph{selectors}; there is no need to consider other function symbols, because they can
+be abstracted away as variables when combining theories.
+%
+In an SMT problem, the signature is normally given by specifying the
+uninterpreted types in any order, the (co)datatypes with their constructors
+and selectors in groups of mutually (co)recursive groups of (co)datatypes, and
 finally any other function symbols.
@@
 either all datatypes or all codatatypes. Polymorphic types, nested
-(co)recursion, and datatype--codatatype mixtures fall outside this fragment.%
-\footnote{%
+(co)recursion, and datatype--codatatype mixtures fall outside this fragment%
+\dotReportFootnote{%
 In principle, rank-1 (top-level) polymorphism \cite{blanchette-paskevich-2013}
-would be easy to support; nesting datatypes inside datatypes, and
-likewise for codatatypes, can be reduced to the mutual case
+should not raise any special difficulties. Nesting datatypes inside datatypes,
+and likewise for codatatypes, can be reduced to the mutual case
 \cite{gunter-1993-not}. So the only genuinely interesting cases missing are
-mixed nested (co)recursion (as in the \ty{tree} example above) as well as
+mixed nested (co)recursion as well as
 (co)recursion through a non-(co)datatype (both of which make sense
@@
 % at the metalevel
-by using subscripts---for example, $\ty{stream}_{\,\tau}$ denotes a
-family of ground types $\ty{stream}_{\,\ty{int}}$, $\ty{stream}_{\,\ty{bool}}$,
-$\ty{stream}_{\,\ty{stream}_{\,\ty{real}}}$, etc.
-
-Each datatype $\delta$ is equipped by
+through subscripts---for example, $\ty{stream}_{\,\tau}$ denotes a
+family of ground types including
+$\ty{stream}_{\,\ty{int}}$, $\ty{stream}_{\,\ty{bool}}$,
+and \smash{$\ty{stream}_{\,\ty{stream}_{\,\ty{real}}}$}.
+
+Each datatype $\delta$ is equipped with
 $m \ge 1$ constructors, and each constructor for $\delta$ takes zero or more
 arguments and returns a $\delta$ value. The argument types must be either
-plain, be or be one of the newly introduced (co)datatypes.
+ordinary, among the already know (co)datatypes, or among the (co)datatypes being
+introduced.
 %
@@
 
-Datatypes and codatatypes can be characterized axiomatically. They share basic
-properties of constructors and selectors. All properties below are implicitly
-universally quantified and range over all
-$i$, $j$, $\jPrime$, and $k$ within bounds:
+Datatypes and codatatypes share basic properties pertaining to constructors and
+selectors. All properties below are implicitly universally quantified and range
+over all $i$, $j$, $\jPrime$, and $k$ within bounds:
 % and to all possible splits of the
@@
 as shorthand for the path containing an edge from $\ec{t_1}$ to $\ec{t_2}$, followed by one from $\ec{t_2}$ to $\ec{t_3}$.
-%A set of paths is \emph{disjoint} if it does not contain two paths $p_1$ and $p_2$ such that $p_1 = p_2 \sigma$ for some $\sigma$.
 
@@
 it can be shown by case analysis on the rules of the calculus that each non-root node $\Fc$ is such that 
-$\mathcal{T}(\Fc) \subseteq \tcEc$, and moreover thus contains an equality between two terms from $\tcEc$ not occurring in its parent node.
+$\mathcal{T}(\Fc) \subseteq \tcEc $, and moreover thus contains an equality between two terms from $\tcEc$ not occurring in its parent node.
 Thus, the depth of a branch in a derivation tree with root node $\Ec$ is at most $\mid \tcEc \mid^2$,
@@
 
-It remains to show the inverse of the previous lemma, that when a derivation tree with root node $\Ec$ contains a saturated node $\Fc$,
-then $\Ec$ is satisfiable $\thD$.
-
-%We first show that for any saturated node $\Fc$, 
-%additional equalities $\Gc$ can be added to $\Fc$ such that each codatatype/datatype equivalence class of $\Fc \cup \Gc$ has a constructor term.
-%For this purpose, we extend the calculus in this section.
-%For a subset of equivalence classes $S$ of $\Ec$, write $\dpath{\Ec}{C,n}{S}$ to denote the set
-%such that $e' \in \dpath{\Ec}{C,n}{S}$ if and only if there exists an $e \in S$ such that $e \rightarrow^{C}_n e'$ is induced by $\Ec$.
-
-\begin{comment}
-\begin{figure}[t]
-\centering
-\begin{tabular}{c}
-\rn{O-Finish}
-\(
-\inferrule{
-  \Ac := x
-  \quad
-  x : \tau
-  \quad
-  \tau \in \Nondata
-  \quad
-  c \not\in \tEc
-}{
-  \Ac := \emptyset \quad \Ec := \Ec \cup \{ x \teq c \}
-}
-\)
-\\[3.7ex]
-\rn{C-Finish}
-\(
-\inferrule{
-  \Ac := x
-  \quad
-  x : \tau
-  \quad
-  \tau \in \Codata
-  \quad
-  C_i \in \Ctr^\tau
-  \quad 
-  C_i \not\in \{ C_j \mid d \in \Dc, C_j( \vec t ) \in d \}
-}{
-  \Ac := \emptyset \quad \Ec := \Ec \cup \{ x \teq C_i( \vec y ) \} \quad \Rc := \Rc \cup \vec y
-}
-\)
-\\[3.7ex]
-\rn{C-Rec}
-\(
-\inferrule{
-  \Ac := x
-  \quad
-  x : \tau
-  \quad
-  \tau \in \Codata
-  \quad
-  C_i \in \Ctr^\tau
-  \quad
-  C_i \text{ infinite arg } j
-}{
-  \Ac := y_j \quad 
-  \Ec := \Ec \cup \{ x \teq C_i( \vec y ) \} \quad 
-  \Rc := \Rc \cup ( \vec y \setminus y_j ) \quad 
-  \Dc := \dpath{\Ec}{C_i,j}{\Dc}
-}
-\)
-\end{tabular}
-\caption{Rules for completion.
-}
-\label{fig:completion-rules}
-\end{figure}
-\end{comment}
-
-%and second show how to construct a valuation for the variables of $\Fc \cup \Gc$.
-
-%\begin{lemma}
-%\end{lemma}
-
-%Given a set of disjoint paths $P$ where each $p \in P$
-%we write $\interp{P}$ to denote the
-
 \begin{lemma}[Solution Soundness]
-If there exists a derivation tree with root node $\Ec$ containing a saturated node $\Fc$, then $\Ec$ is satisfiable in $\thD$.
+If there exists a derivation tree with root node $\Ec$ containing a saturated node, then $\Ec$ is satisfiable in $\thD$.
 \end{lemma}
@@
 %First, we construct a valuation $\Val$ as follows.
-%Let $C$ be the set of variables occurring in the same equivalence class as a constructor term,
-%and let $X$ be the set of variables occurring in $\mathcal{T}(\Fc)$ but not in $C$.
-
-We assume that:
-\begin{enumerate}
-\item Each equivalence class of $\Fc$ contains at least one variable,
-\item Each equivalence class $\ec{x}$ of $\Fc$ where $x : \tau$, $\tau \in \Nondata$ contains a constant,
-\item Each equivalence class $\ec{x}$ of $\Fc$ where $x : \tau$, $\tau \in \Data \cup \Codata$ contains a constructor term.
-\end{enumerate}
-The first statement comes with no loss of generality since new equalities of the form $y \teq t$ for fresh variable $y$ can be added to $\Fc$ without affecting its satisfiability.
-The second statement can be assumed for $x : \tau$, since we know that $\tau$ is infinite.
-The third statement can be assumed by \rem{TODO}.
-
-We construct a model $\M$ of $\thD$ that interprets constructors and correctly-applied selectors in the obvious way,
-interprets wrongly-applied selectors based on information in $\Fc$,
-interprets the variables and $x$ of $\Fc$ based on a valuation $\Val$ which construct as follows.
-For each equivalence class $e : \tau$, where $\tau \in \Nondata$ containing constant $c$,
-$\Val(x) = c$ for each $x \in e$.
-For each $e : \tau$, where $\tau \in \Data$ containing the constructor term $C( \vec u )$ where the equivalence classes $\ec{\vec u}$ contains the variables $\vec y$,
-$\Val(x) = C( \Val( \vec y ) )$ for each $x \in e$.
-Since \rn{Cyclic} does not apply, this definition is well-defined, and since \rn{Cong} does not apply, we have that $\Val{x} = \Val{y}$ if and only if $\ec{x} = \ec{y}$.
-For each $e : \tau$, where $\tau \in \Codata$,
-\rem{TODO}
-
-We show that all equalities and disequalities in $\Fc$ are satisfied by $\M$.
-Say $t_1 \teq t_2 \in \Fc$ where $t_1$ and $t_2$ have type $\tau$ where $\tau \not\in \Codata$,
-and assume that $\ec{t_1}$ contains the variable $x$.
-We show that $\M( t ) = \Val( x )$ for each term $t \in \ec{x}$.
-In the case where $t$ is a variable, then $\M( t ) = \Val( x )$ since $t$ and $x$ are assigned identical values by construction.
-In the case where $t$ is a constructor term $C( \vec u )$, say our construction built a valuation for $x$ based on the constructor term $C'( \vec u' ) \in \ec{x}$.
-Since \rn{Unify$_2$} does not apply to $\Fc$, we have that $C = C'$.
-Since \rn{Unify$_1$} does not apply to $\Fc$, we have that $\ec{\vec u} = \ec{\vec u'}$.
-By structural induction, we have $\M( \vec u ) = \M( \vec u' )$, and thus $\M( t ) = \M( C( \vec u ) ) = \M( C'( \vec u ' ) ) = \Val( x )$.
-In the case where $t$ is a selector term $s^j_k( u )$,
-then since \rn{Split} does not apply, 
-$\ec{u}$ must contain a term of the form $C_i( s^1_i( u ), \ldots, s^{a_i}_i( u ) )$ for some $i$.
-Let $x_1, \ldots, x_n$ be the variables in equivalence classes $\ec{s^1_i( u )}, \ldots, \ec{s^{a_i}_i( u )}$.
-By our construction, $\M( u ) = C_i( \Val( x_1 ), \ldots, \Val( x_{a_i} ) )$.
-If $i = k$, then $\M( t ) = \M( s^j_k( u ) ) = \Val( x_j )$, which is $\Val( x )$ since $x_j \in \ec{s^j_i( u )} = \ec{x}$.
-If $i \neq k$, then $\thD$ does not constraint the value $\M( t )$, and thus we interpret $s^j_k( \M( u ) )$ as $\Val( x )$.
-
-%If $t$ and $s$ are both variables, then this follows since $\ec{t} = \ec{s}$.
-
-Let $t \tneq s \in \Fc$.
-Since \rn{Conflict} does not apply, we know $t \teq s \not\in \Fc$ and thus $\ec{t}$ and $\ec{s}$ are distinct.
-Since by above, $\M( t ) = \M( s )$ if and only if $\ec{t}=\ec{s}$, we have $\M( t ) \neq \M( s )$, and thus $\M \models t \tneq s$.
-
-By these two, 
-%$\M \models_{\thD} \Fc$, and since $\Fc$ is a superset of $\Ec$, we have that $\Ec$ is satisfied by $\M$.
+Let $C$ be the set of variables occurring in the same equivalence class as a constructor term,
+and let $X$ be the set of variables occurring in $\mathcal{T}(\Fc)$ but not in $C$.
+
 \qed

### 2014-12-09 tex/codatadp.tex

@@
 \newcommand{\ttpath}[3]{\mathcal{P}_{ #2 \rightarrow #3 }( #1 )}
+\newcommand{\interp}[1]{[\![ #1 ]\!]}
+\newcommand{\dpath}[3]{\delta^{#1}_{#2}( #3 )}
 \newcommand{\ec}[1]{ [ #1 ] }
@@
 \newcommand{\Val}{\mathcal{A}}
+
+\def\AIF{\qtab\keyword{if}\ }
+\def\ATHEN{\ \keyword{then}\ }
+\def\AELSE{\untab\qtab\keyword{else}\ }
+\def\AFI{\untab}
+\def\ARETURN{\keyword{return}\ }
+\def\AENDPROC{\untab}
 
@@
 
-\begin{lemma}[Solution Soundness]
-If there exists a derivation tree with root node $\Ec$ containing a saturated node, then $\Ec$ is satisfiable in $\thD$.
+It remains to show the inverse of the previous lemma, that when a derivation tree with root node $\Ec$ contains a saturated node $\Fc$,
+then $\Ec$ is satisfiable $\thD$.
+
+%We first show that for any saturated node $\Fc$, 
+%additional equalities $\Gc$ can be added to $\Fc$ such that each codatatype/datatype equivalence class of $\Fc \cup \Gc$ has a constructor term.
+%For this purpose, we extend the calculus in this section.
+%For a subset of equivalence classes $S$ of $\Ec$, write $\dpath{\Ec}{C,n}{S}$ to denote the set
+%such that $e' \in \dpath{\Ec}{C,n}{S}$ if and only if there exists an $e \in S$ such that $e \rightarrow^{C}_n e'$ is induced by $\Ec$.
+
+\begin{comment}
+\begin{figure}[t]
+\centering
+\begin{tabular}{c}
+\rn{O-Finish}
+\(
+\inferrule{
+  \Ac := x
+  \quad
+  x : \tau
+  \quad
+  \tau \in \Nondata
+  \quad
+  c \not\in \tEc
+}{
+  \Ac := \emptyset \quad \Ec := \Ec \cup \{ x \teq c \}
+}
+\)
+\\[3.7ex]
+\rn{C-Finish}
+\(
+\inferrule{
+  \Ac := x
+  \quad
+  x : \tau
+  \quad
+  \tau \in \Codata
+  \quad
+  C_i \in \Ctr^\tau
+  \quad 
+  C_i \not\in \{ C_j \mid d \in \Dc, C_j( \vec t ) \in d \}
+}{
+  \Ac := \emptyset \quad \Ec := \Ec \cup \{ x \teq C_i( \vec y ) \} \quad \Rc := \Rc \cup \vec y
+}
+\)
+\\[3.7ex]
+\rn{C-Rec}
+\(
+\inferrule{
+  \Ac := x
+  \quad
+  x : \tau
+  \quad
+  \tau \in \Codata
+  \quad
+  C_i \in \Ctr^\tau
+  \quad
+  C_i \text{ infinite arg } j
+}{
+  \Ac := y_j \quad 
+  \Ec := \Ec \cup \{ x \teq C_i( \vec y ) \} \quad 
+  \Rc := \Rc \cup ( \vec y \setminus y_j ) \quad 
+  \Dc := \dpath{\Ec}{C_i,j}{\Dc}
+}
+\)
+\end{tabular}
+\caption{Rules for completion.
+}
+\label{fig:completion-rules}
+\end{figure}
+\end{comment}
+
+%and second show how to construct a valuation for the variables of $\Fc \cup \Gc$.
+
+%\begin{lemma}
+%\end{lemma}
+
+Given a set of disjoint paths $P$ where each $p \in P$
+we write $\interp{P}$ to denote the value
+
+\begin{lemma}[Solution Soundness] \label{lem:ss}
+If there exists a derivation tree with root node $\Ec$ containing a saturated node $\Fc$, then $\Ec$ is satisfiable in $\thD$.
 \end{lemma}
@@
 Let $\Fc$ be a saturated node in a derivation tree with root node $\Ec$.
-%First, we construct a valuation $\Val$ as follows.
-Let $C$ be the set of variables occurring in the same equivalence class as a constructor term,
-and let $X$ be the set of variables occurring in $\mathcal{T}(\Fc)$ but not in $C$.
-
+We assume that:
+\begin{enumerate}
+\item Each equivalence class of $\Fc$ contains at least one variable,
+\item Each equivalence class $\ec{x}$ of $\Fc$ where $x : \tau$, $\tau \in \Nondata$ contains a constant,
+\item Each equivalence class $\ec{x}$ of $\Fc$ where $x : \tau$, $\tau \in \Data \cup \Codata$ contains a constructor term.
+\end{enumerate}
+The first statement comes with no loss of generality since new equalities of the form $y \teq t$ for fresh variable $y$ can be added to $\Fc$ without affecting its satisfiability.
+The second statement can be assumed for $x : \tau$, since $\tau$ is infinite.
+Hence, we can assign constants for each such equivalence class.
+Moreover, since $\Fc$ contains $\thD$-constraints only, these constants can be chosen arbitrarily.
+The third statement can be assumed by \rem{TODO}.
+
+We construct a model $\M$ of $\thD$ that interprets constructors and correctly-applied selectors in the obvious way,
+interprets wrongly-applied selectors based on equality information in $\Fc$,
+interprets the variables and $x$ of $\Fc$ based on a valuation $\Val$, which is constructed as follows.
+For each equivalence class $e : \tau$, where $\tau \in \Nondata$ containing constant $c$,
+$\Val(x) = c$ for each $x \in e$.
+For each $e : \tau$, where $\tau \in \Data$ containing the constructor term $C( \vec u )$ where the equivalence classes $\ec{\vec u}$ contains the variables $\vec y$,
+$\Val(x) = C( \Val( \vec y ) )$ for each $x \in e$.
+Since \rn{Cyclic} does not apply, this definition is well-defined. 
+For each $e : \tau$, where $\tau \in \Codata$, 
+$\Val( x ) = \interp{ \tpath{\Fc}{e} }$ for each $x \in e$.
+Since neither \rn{Cong} or \rn{Bisimilar} applies to $\Fc$, we have that for all variables $x$ and $y$:
+\begin{eqnarray} \label{eqn:m-bij-dt}
+\Val(x) = \Val(y) \text{ if and only if } \ec{x} = \ec{y}
+\end{eqnarray}
+
+We first show $\M$ satisfies all equalities in $t_1 \teq t_2 \in \Fc$, where $t_1$ and $t_2$ have type $\tau$.
+
+If $\tau \in \Nondata$, then $\M( t_1 ) = \M( t_2 )$ by construction.
+
+If $\tau \in \Data$, then assume the equivalence class of $t_1$ and $t_2$ contains variable $x$.
+By induction on the size of $\Val( x )$, we show that $\M( t ) = \Val( x )$ for each term $t \in \ec{x}$, which implies $\M \models t_1 \teq t_2$.
+In the case where $t$ is a variable, then $\M( t ) = \Val( x )$ since $t$ and $x$ are assigned identical values by construction.
+In the case where $t$ is a constructor term $C( \vec u )$, say our construction built a valuation for $x$ based on the constructor term $C'( \vec u' ) \in \ec{x}$.
+Since \rn{Unify$_2$} does not apply to $\Fc$, we have that $C = C'$.
+Since \rn{Unify$_1$} does not apply to $\Fc$, we have that $\ec{\vec u} = \ec{\vec u'}$.
+By the induction hypothesis, we have $\M( \vec u ) = \M( \vec u' )$, and thus $\M( t ) = C( \M( \vec u ) ) = C'( \M( \vec u ' ) ) = \Val( x )$.
+In the case where $t$ is a selector term $s^j_k( u )$,
+then since \rn{Split} does not apply, 
+$\ec{u}$ must contain a term of the form $C_i( s^1_i( u ), \ldots, s^{a_i}_i( u ) )$ for some $i$.
+Let $x_1, \ldots, x_n$ be the variables in equivalence classes $\ec{s^1_i( u )}, \ldots, \ec{s^{a_i}_i( u )}$.
+By our construction, $\M( u ) = C_i( \Val( x_1 ), \ldots, \Val( x_{a_i} ) )$.
+If $i = k$, then $\M( t ) = \M( s^j_k( u ) ) = s^j_k( \M( u ) ) = \Val( x_j )$, which is $\Val( x )$ since $x_j \in \ec{s^j_i( u )} = \ec{x}$.
+If $i \neq k$, then $\thD$ does not constraint the value of $\M( t )$ and since $s^j_k( \M( u ) )$ is otherwise unconstrained in our model construction since \rn{Cong} does not apply, 
+we may interpret $s^j_k( \M( u ) )$ as $\Val( x )$.
+
+If $\tau \in \Codata$, \rem{TODO}
+
+We show that all disequalities in $\Fc$ are satisfied by $\M$.
+Say $t_1 \tneq t_2 \in \Fc$.
+Since \rn{Conflict} does not apply, we know $t_1 \teq t_2 \not\in \Fc$ and thus $\ec{t_1}$ and $\ec{t_2}$ are distinct.
+By the previous paragraph and statement~\ref{eqn:m-bij-dt}, $\M( t_1 ) \neq \M( t_2 )$, and thus $\M \models t_1 \tneq t_2$.
+
+Thus, $\M \models \Fc$
+and since $\Fc$ is a superset of $\Ec$, we have that $\Ec$ is satisfied by $\M$.
 \qed

### 2014-12-09 tex/codatadp.tex
Elaborate on refutation soundness.  Minor fix to intro.

@@
 
-\renewcommand{\vec}[1]{\mathbf{#1}}
+%\renewcommand{\vec}[1]{\mathbf{#1}}
+\renewcommand{\vec}[1]{\bar #1}
 \newcommand{\Ec}{\mathsf{E}}
@@
 
-In this \thewordpaper, we introduce a unified decision procedure for ground and
-universal problems involving both datatypes and codatatypes. The procedure is
+In this \thewordpaper, we introduce a unified decision procedure for ground
+problems involving both datatypes and codatatypes, and discuss extensions 
+of the procedure when handling universal (non-ground) problems.
+The procedure is
 described abstractly as a calculus (Section~\ref{sec:the-calculus}) and is
@@
 We will write $\tEc$ to denote the set of $\Sig$-terms occurring in $\Ec$.
-We will commonly denote tuples of terms $( t_1, \ldots, t_n )$ in bold font, as $\vec t$.
+%We will commonly denote tuples of terms $( t_1, \ldots, t_n )$ in bold font, as $\vec t$.
 We assume that all $\Sig$-terms in $\Ec$ are \emph{normalized}, meaning that all subterms of the form $s^i_j( \const{C}_j( \vec t ) )$
@@
 Let $\Ec$ be the set $\{ x \teq \const{LCons}(y,\const{LCons}(y,x)) \}$.
-After taking the bidirectional closure of $\Ec$, we obtain equivalence classes 
+After applying the rules in Figure~\ref{fig:cc-rules} to completion, the equivalence classes of $\Ec$ are
 $\{ x, \const{LCons}(y,\const{LCons}(y,x)) \}$, $\{ \const{LCons}(y,x) \}$, and $\{ y \}$.
@@
 then $\Ec$ is unsatisfiable in $\thD$.
-\rem{Elaborate on cyclic?}
+For \rn{Conflict}, this is a consequence of equality reasoning.
+For \rn{Unify$_2$}, this is a consequence of distinctness.
+For \rn{Cyclic}, this is a consequence of induction. \rem{elaborate?}
 If the children of $\Ec$ are closed derivation trees 
@@
 for each $C_i \in \Ctr^\tau$.
-Since all models of $\thD$ entail (exactly) one equality $t \teq C_i( s^1_i( t ), \ldots, s^{a_i}_i( t ) )$,
+Since by exhaustiveness, all models of $\thD$ entail (exactly) one equality $t \teq C_i( s^1_i( t ), \ldots, s^{a_i}_i( t ) )$,
 we know $\Ec$ is unsatisfiable in $\thD$.
@@
 In all cases, we have that $\Ec \models_{\thD} t_1 \teq t_2$.
-\rem{Elaborate on Bisimilar?}
+For \rn{Refl}, \rn{Symm}, \rn{Trans}, \rn{Cong}, this is a consequence of equality reasoning.
+For \rn{Unify$_1$}, this is a consequence of injectivity.
+For \rn{Bisimilar}, this is a consequence of co-induction. \rem{elaborate?}
+For \rn{Singleton}, clearly $t_1$ and $t_2$ must have the same value since the cardinality of their type is one.
 By the induction hypothesis, we have $\Ec \cup t_1 \teq t_2$ is unsatisfiable in $\thD$, 
@@
 If $\tau \in \Data$, then assume the equivalence class of $t_1$ and $t_2$ contains variable $x$.
-By induction on the size of $\Val( x )$, we show that $\M( t ) = \Val( x )$ for each term $t \in \ec{x}$, which implies $\M \models t_1 \teq t_2$.
+We show by induction on the size of $\Val( x )$ that $\M( t ) = \Val( x )$ for each term $t \in \ec{x}$, which implies $\M \models t_1 \teq t_2$.
 In the case where $t$ is a variable, then $\M( t ) = \Val( x )$ since $t$ and $x$ are assigned identical values by construction.
@@
 
-By Lemmas~\ref{lem:t},~\ref{lem:rs}, and~\ref{lem:ss}, our calculus is sound and complete for $\thD$.
+By Lemmas~\ref{lem:t},~\ref{lem:rs}, and~\ref{lem:ss}, the calculus described in this section is sound and complete for $\thD$.
 

### 2014-12-09 tex/codatadp.tex
Reorganize approach for solution soundness proof.  Cleanup.

@@
 \newcommand{\Rc}{\mathsf{R}}
+\newcommand{\vrange}{\mathsf{range}}
+\newcommand{\vdom}{\mathsf{dom}}
 \newcommand{\tEc}{\mathcal{T}(\Ec)}
@@
 The rule \rn{Bisimilar} recognizes when two equivalence classes $\ec{t_1}$ and $\ec{t_2}$ having co-datatype type are observationally equivalent.
-In particular, if there is a mapping $\sigma$ over equivalence classes such that all paths $p$ from $t_2$
-have a corresponding path $p \sigma$ from $t_1$,
-then $t_1$ must be equal to $t_2$.
+In this rule, $\sigma$ is a complete mapping over the equivalence classes that occur as sources in the edges of paths in $\tpath{\Ec}{\ec{t_2}}$.
+The rule applies if $\tpath{\Ec}{\ec{t_1}}$ is comprised of the paths $p \sigma$ for each (and only) the paths $p \in \tpath{\Ec}{\ec{t_2}}$,
+in which case $t_1$ must be equal to $t_2$.
 We demonstrate this rule concretely with the following example.
@@
 Let $\Fc$ be a saturated node in a derivation tree with root node $\Ec$.
+We construct a model $\M$ of $\thD$ that interprets constructors and correctly-applied selectors in the obvious way,
+interprets wrongly-applied selectors based on equality information in $\Fc$,
+interprets the variables and $x$ of $\Fc$ based on a mapping $\Val$ from the equivalence classes of $\Fc$ to values, 
+which is constructed based on the following rules.
+
+\rn{Ord}
+\(
+\inferrule{
+  \ec{x} \not\in \vdom(\Val)
+  \quad
+  x : \tau
+  \quad
+  \tau \in \Nondata
+  \quad
+  c \not\in \vrange(\Val)
+}{
+  \Val := \Val \cup \ec{x} \mapsto c
+}
+\)
+
+\rn{D-Ind}
+\(
+\inferrule{
+  \ec{x} \not\in \vdom(\Val)
+  \quad
+  x : \tau
+  \quad
+  \tau \in \Data
+  \quad
+  C( \vec t ) \in \ec{x}
+  \quad
+  \ec{\vec t} \in \vdom(\Val)
+}{
+  \Val := \Val \cup \ec{x} \mapsto C( \Val \ec{\vec t}  )
+}
+\)
+
+\rn{D-Base}
+\(
+\inferrule{
+  \ec{x} \not\in \vdom(\Val)
+  \quad
+  x : \tau
+  \quad
+  \tau \in \Data
+  \quad
+  \ec{x} \text{ ctr-free}
+  \quad
+  c \not\in \vrange(\Val)
+}{
+  \Val := \Val \cup \ec{x} \mapsto c
+}
+\)
+
+We assume these rules are applied in the order in which they are written, for instance, 
+the rule \rn{D-Ind} is applied to completion before applying \rn{D-Base}.
+
+\begin{eqnarray} \label{eqn:m-bij-dt}
+\Val \ec{x} = \Val \ec{y} \text{ if and only if } \ec{x} = \ec{y}
+\end{eqnarray}
+
+\begin{comment}
 We assume that:
@@
 
-We construct a model $\M$ of $\thD$ that interprets constructors and correctly-applied selectors in the obvious way,
-interprets wrongly-applied selectors based on equality information in $\Fc$,
-interprets the variables and $x$ of $\Fc$ based on a valuation $\Val$, which is constructed as follows.
 For each equivalence class $e : \tau$, where $\tau \in \Nondata$ containing constant $c$,
@@
 \end{eqnarray}
-
+\end{comment}
+
+Without loss of generality, we will assume in the following that all equivalence class of $\Fc$ contain at least one variable,
+which comes with no loss of generality since new equalities of the form $y \teq t$ for fresh variable $y$ can be added to $\Fc$ without affecting its satisfiability.
+We let $\M( x ) = \Val \ec{x}$ for each variable $x$ in $\Fc$.
 We first show $\M$ satisfies all equalities in $t_1 \teq t_2 \in \Fc$, where $t_1$ and $t_2$ have type $\tau$.
 
-If $\tau \in \Nondata$, then $\M( t_1 ) = \M( t_2 )$ by construction.
+If $\tau \in \Nondata$, $t_1$ and $t_2$ must be variables, and thus $\M( t_1 ) = \M( t_2 )$ by construction.
 
 If $\tau \in \Data$, then assume the equivalence class of $t_1$ and $t_2$ contains variable $x$.
-We show by induction on the size of $\Val( x )$ that $\M( t ) = \Val( x )$ for each term $t \in \ec{x}$, which implies $\M \models t_1 \teq t_2$.
-In the case where $t$ is a variable, then $\M( t ) = \Val( x )$ since $t$ and $x$ are assigned identical values by construction.
-In the case where $t$ is a constructor term $C( \vec u )$, say our construction built a valuation for $x$ based on the constructor term $C'( \vec u' ) \in \ec{x}$.
+We show by induction on the size of $\Val \ec{ x }$ that $\M( t ) = \Val \ec{x}$ for each term $t \in \ec{x}$, which implies $\M \models t_1 \teq t_2$.
+In the case where $t$ is a variable, then $\M( t ) = \Val \ec{ t } = \Val \ec{ x }$.
+In the case where $t$ is a constructor term $C( \vec u )$, 
+then $\Val \ec{x}$ must have been assigned by \rn{D-Ind} for some constructor term $C'( \vec u' ) \in \ec{x}$.
 Since \rn{Unify$_2$} does not apply to $\Fc$, we have that $C = C'$.
 Since \rn{Unify$_1$} does not apply to $\Fc$, we have that $\ec{\vec u} = \ec{\vec u'}$.
-By the induction hypothesis, we have $\M( \vec u ) = \M( \vec u' )$, and thus $\M( t ) = C( \M( \vec u ) ) = C'( \M( \vec u ' ) ) = \Val( x )$.
+By the induction hypothesis, we have $\M( \vec u ) = \M( \vec u' )$, and thus $\M( t ) = C( \M( \vec u ) ) = C'( \M( \vec u ' ) ) = \Val \ec{ x }$.
 In the case where $t$ is a selector term $s^j_k( u )$,
@@
 Let $x_1, \ldots, x_n$ be the variables in equivalence classes $\ec{s^1_i( u )}, \ldots, \ec{s^{a_i}_i( u )}$.
-By our construction, $\M( u ) = C_i( \Val( x_1 ), \ldots, \Val( x_{a_i} ) )$.
-If $i = k$, then $\M( t ) = \M( s^j_k( u ) ) = s^j_k( \M( u ) ) = \Val( x_j )$, which is $\Val( x )$ since $x_j \in \ec{s^j_i( u )} = \ec{x}$.
+By our construction, $\M( u ) = C_i( \Val \ec{ x_1 }, \ldots, \Val \ec{ x_{a_i} } )$.
+If $i = k$, then $\M( t ) = \M( s^j_k( u ) ) = s^j_k( \M( u ) ) = \Val \ec{x_j}$, which is $\Val( x )$ since $x_j \in \ec{s^j_i( u )} = \ec{x}$.
 If $i \neq k$, then $\thD$ does not constraint the value of $\M( t )$ and since $s^j_k( \M( u ) )$ is otherwise unconstrained in our model construction since \rn{Cong} does not apply, 
-we may interpret $s^j_k( \M( u ) )$ as $\Val( x )$.
+we may interpret $s^j_k( \M( u ) )$ as $\Val \ec{ x }$.
 

### 2014-12-09 tex/codatadp.tex
More work on solution soundness.

@@
 \begin{lemma}[Solution Soundness] \label{lem:ss}
-If there exists a derivation tree with root node $\Ec$ containing a saturated node $\Fc$, then $\Ec$ is satisfiable in $\thD$.
+If there exists a derivation tree with root node $\Ec$ containing a saturated node, then $\Ec$ is satisfiable in $\thD$.
 \end{lemma}
 \begin{proof}
-Let $\Fc$ be a saturated node in a derivation tree with root node $\Ec$.
+Let $\Fc_0$ be a saturated node in a derivation tree with root node $\Ec$.
+We will construct a model $\M$ for a set of equalities $\Fc$ that is equivalent to $\Fc_0$, and where 
+all equivalence classes of $\Fc$ contain at least one variable, and all selectors in $\Fc$ are applied to variables only.
+The former comes with no loss of generality since new equalities of the form $y \teq t$ for fresh variable $y$ can be added to $\Fc$ without affecting its satisfiability,
+The latter also comes with no loss of generality since nested applications of selectors in $\Fc_0$ can be replaced by fresh variables while adding additional equalities to $\Fc$.
+
 We construct a model $\M$ of $\thD$ that interprets constructors and correctly-applied selectors in the obvious way,
@@
 
-Without loss of generality, we will assume in the following that all equivalence class of $\Fc$ contain at least one variable,
-which comes with no loss of generality since new equalities of the form $y \teq t$ for fresh variable $y$ can be added to $\Fc$ without affecting its satisfiability.
-We let $\M( x ) = \Val \ec{x}$ for each variable $x$ in $\Fc$.
+Now, we let $\M( x ) = \Val \ec{x}$ for each variable $x$ in $\Fc$.
 We first show $\M$ satisfies all equalities in $t_1 \teq t_2 \in \Fc$, where $t_1$ and $t_2$ have type $\tau$.
@@
 
-If $\tau \in \Data$, then assume the equivalence class of $t_1$ and $t_2$ contains variable $x$.
-We show by induction on the size of $\Val \ec{ x }$ that $\M( t ) = \Val \ec{x}$ for each term $t \in \ec{x}$, which implies $\M \models t_1 \teq t_2$.
+If $\tau \in \Data$,
+we first show that $\M( t ) = \Val \ec{x}$ for all non-selector terms $t$ that occur in datatype equivalence classes $\ec{x}$.
+We show this by induction on the size of $\Val \ec{x}$.
 In the case where $t$ is a variable, then $\M( t ) = \Val \ec{ t } = \Val \ec{ x }$.
@@
 By the induction hypothesis, we have $\M( \vec u ) = \M( \vec u' )$, and thus $\M( t ) = C( \M( \vec u ) ) = C'( \M( \vec u ' ) ) = \Val \ec{ x }$.
-In the case where $t$ is a selector term $s^j_k( u )$,
-then since \rn{Split} does not apply, 
-$\ec{u}$ must contain a term of the form $C_i( s^1_i( u ), \ldots, s^{a_i}_i( u ) )$ for some $i$.
-Let $x_1, \ldots, x_n$ be the variables in equivalence classes $\ec{s^1_i( u )}, \ldots, \ec{s^{a_i}_i( u )}$.
-By our construction, $\M( u ) = C_i( \Val \ec{ x_1 }, \ldots, \Val \ec{ x_{a_i} } )$.
-If $i = k$, then $\M( t ) = \M( s^j_k( u ) ) = s^j_k( \M( u ) ) = \Val \ec{x_j}$, which is $\Val( x )$ since $x_j \in \ec{s^j_i( u )} = \ec{x}$.
-If $i \neq k$, then $\thD$ does not constraint the value of $\M( t )$ and since $s^j_k( \M( u ) )$ is otherwise unconstrained in our model construction since \rn{Cong} does not apply, 
-we may interpret $s^j_k( \M( u ) )$ as $\Val \ec{ x }$.
+
+We now show that $\M( t ) = \Val \ec{x}$ for all selector terms $s^j_k( y )$ that occur in datatype equivalence classes $\ec{x}$.
+Since \rn{Split} does not apply, 
+$\ec{y}$ must contain a term of the form $C_i( s^1_i( y ), \ldots, s^{a_i}_i( y ) )$ for some $i$.
+Let $x_1, \ldots, x_n$ be variables in equivalence classes $\ec{s^1_i( y )}, \ldots, \ec{s^{a_i}_i( y )}$.
+By the previous paragraph, we have $\M( y ) = \M( C_i( s^1_i( y ), \ldots, s^{a_i}_i( y ) ) )$,
+which is $C_i( \Val \ec{ x_1 }, \ldots, \Val \ec{ x_{a_i} } )$.
+If $i = k$, then $\M( t ) = \M( s^j_k( y ) ) = s^j_k( \M( y ) ) = \Val \ec{x_j}$, which is $\Val \ec{ x }$ since $\ec{x_j} = \ec{s^j_i( y )} = \ec{x}$.
+If $i \neq k$, then $\thD$ does not constraint the value of $\M( t )$,
+and since $s^j_k( \M( y ) )$ is otherwise unconstrained in our model construction since \rn{Cong} does not apply, 
+we may interpret $s^j_k( \M( y ) )$ as $\Val \ec{ x }$. 
+Thus, for all datatype equivalence classes $\ec{x}$, we have $\M( t ) = \Val \ec{x}$ for each term $t \in \ec{x}$, which implies $\M \models t_1 \teq t_2$.
 
@@
 
-We show that all disequalities in $\Fc$ are satisfied by $\M$.
+We now show that all disequalities in $\Fc$ are satisfied by $\M$.
 Say $t_1 \tneq t_2 \in \Fc$.
@@
 
-Thus, $\M \models \Fc$
+Since by assumption $\Fc$ contains only equalities and disequalities, we have $\M \models \Fc$,
 and since $\Fc$ is a superset of $\Ec$, we have that $\Ec$ is satisfied by $\M$.

### 2014-12-09 tex/codatadp.tex
Another refactor to solution soundness, now work with lazy construction of valuation.

@@
 interprets wrongly-applied selectors based on equality information in $\Fc$,
-interprets the variables and $x$ of $\Fc$ based on a mapping $\Val$ from the equivalence classes of $\Fc$ to values, 
-which is constructed based on the following rules.
-
+interprets the variables and $x$ of $\Fc$ based on a mapping $\Val$ from the equivalence classes of $\Fc$ to values.
+Initially, $\Val$ contains $\ec{x} \mapsto v_{\ec{x}}$ for each equivalence class $\ec{x}$ of $\Fc$,
+where $v_{\ec{x}}$ is a fresh variable of the same type as $x$ not occurring in $\Fc$.
+Then, $\Val$ is updated according to the following rules:
+
+\begin{comment}
 \rn{Ord}
@@
 \)
-
-We assume these rules are applied in the order in which they are written, for instance, 
-the rule \rn{D-Ind} is applied to completion before applying \rn{D-Base}.
+\end{comment}
+
+\rn{Ind}
+\(
+\inferrule{
+  %x : \tau
+  %\quad
+  %\tau \in \Codata
+  v_{\ec{x}} \in FV( \Val )
+  \quad
+  C( t_1, \ldots, t_n ) \in \ec{x}
+  \quad
+  C \in \Ctr
+}{
+  \Val := \Val \{ v_{\ec{x}} \mapsto \mu v_{\ec{x}}. C( v_{\ec{t_1}}, \ldots, v_{\ec{t_n}} ) \}
+}
+\)
+
+\rn{Base}
+\(
+\inferrule{
+  %x : \tau
+  %\quad
+  %\tau \in \Codata
+  %\quad
+  \ec{x} \mapsto v_{\ec{x}} \in \Val
+  %\quad
+  %\ec{x} \text{ ctr-free}
+  \quad
+  c \not\sim c' \text{ for all } c' \in \vrange(\Val)
+}{
+  \Val := \Val \{ v_{\ec{x}} \mapsto c \}
+}
+\)
+
+\rem{define $\sim$}
+We assume these rules are applied to completion in the order in which they are written,
+that is, \rn{Ind} is applied exhaustively before applying \rn{Base}.
+%the rule \rn{D-Ind} is applied to completion before applying \rn{D-Base}.
+In the rule \rn{Ind}, we write $FV( \Val )$ to denote the free variables occurring in $\Val$.
+Notice that this excludes variables beneath $\mu$-binders.
+It is clear that after applying these rules to completion, $\Val$ is a mapping from equivalence classes to ground values.
 
@@
 In the case where $t$ is a constructor term $C( \vec u )$, 
-then $\Val \ec{x}$ must have been assigned by \rn{D-Ind} for some constructor term $C'( \vec u' ) \in \ec{x}$.
+then $\Val \ec{x}$ must have been assigned by \rn{Ind} for some constructor term $C'( \vec u' ) \in \ec{x}$.
 Since \rn{Unify$_2$} does not apply to $\Fc$, we have that $C = C'$.

### 2014-12-09 tex/codatadp.tex
Example run of model generation.

@@
   %\quad
-  \ec{x} \mapsto v_{\ec{x}} \in \Val
+  v_{\ec{x}} \in FV( \Val )
   %\quad
@@
 \rem{define $\sim$}
-We assume these rules are applied to completion in the order in which they are written,
-that is, \rn{Ind} is applied exhaustively before applying \rn{Base}.
+We assume that first \rn{Ind} is applied exhaustively, after which \rn{Base} is applied exhaustively.
 %the rule \rn{D-Ind} is applied to completion before applying \rn{D-Base}.
@@
 Notice that this excludes variables beneath $\mu$-binders.
+
+As an example,
+say that $\Fc$ contains equivalence classes $\ec{x}$ and $\ec{y}$ where $\const{C}(z,x) \in \ec{y}$ and $\const{C}(w,y) \in \ec{x}$, where $z$ and $w$ are variables of integer type.
+A possible application of the above rules, showing the state of $\Val$ after each application, is as follows:
+\[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
+\hline
+\text{Rule} & \multicolumn{4}{c}{\Val} \ 
+\\
+\hline
+& \{ & \ec{x} \mapsto v_{\ec{x}}, & \ec{z} \mapsto v_{\ec{z}}, \\
+&    & \ec{y} \mapsto v_{\ec{y}}, & \ec{w} \mapsto v_{\ec{w}} & \} \\
+\rn{Ind} \ec{x}
+& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ), & \ec{z} \mapsto v_{\ec{z}}, \\
+&    & \ec{y} \mapsto v_{\ec{y}}, & \ec{w} \mapsto v_{\ec{w}} & \} \\
+\rn{Ind} \ec{y}
+& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ) ), & \ec{z} \mapsto v_{\ec{z}}, \\
+&    & \ec{y} \mapsto \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
+\rn{Ind} \ec{x}
+& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ) ), & \ec{z} \mapsto v_{\ec{z}}, \\
+&    & \ec{y} \mapsto \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ) ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
+\rn{Base} \ec{z}
+& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu v_{\ec{y}}. \const{C}( \const{0}, v_{\ec{x}} ) ), & \ec{z} \mapsto \const{0}, \\
+&    & \ec{y} \mapsto \mu v_{\ec{y}}. \const{C}( \const{0}, \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ) ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
+\rn{Base} \ec{w}
+& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( \const{1}, \mu v_{\ec{y}}. \const{C}( \const{0}, v_{\ec{x}} ) ), & \ec{z} \mapsto \const{0}, \\
+&    & \ec{y} \mapsto \mu v_{\ec{y}}. \const{C}( \const{0}, \mu v_{\ec{x}}. \const{C}( \const{1}, v_{\ec{y}} ) ), & \ec{w} \mapsto \const{1} & \} \\
+\end{array}\]
+
 It is clear that after applying these rules to completion, $\Val$ is a mapping from equivalence classes to ground values.

### 2014-12-10 tools/undata.c
started converter (in C!)

@@ -1,0 +1,179 @@
+#include <stdio.h>
+#include <string.h>
+
+typedef enum {
+    Tok_Left_Paren,
+    Tok_Right_Paren,
+    Tok_Declare_Datatypes,
+    Tok_Declare_Codatatypes,
+    Tok_Ident,
+    Tok_End
+} token;
+
+const int BIG_NUM = 65536;
+
+char tok_str[BIG_NUM];
+
+void print_token(FILE *out, token tok)
+{
+    switch (tok) {
+    case Tok_Left_Paren:
+        fputc('(', out);
+        break;
+    case Tok_Right_Paren:
+        fputc(')', out);
+        break;
+    case Tok_Declare_Datatypes:
+        fprintf(out, "declare-datatypes ");
+        break;
+    case Tok_Declare_Codatatypes:
+        fprintf(out, "declare-codatatypes ");
+        break;
+    case Tok_Ident:
+        fprintf(out, "%s", tok_str);
+        fputc(' ', out);
+        break;
+    case Tok_End:
+        ;
+    }
+}
+
+int get_token(FILE *in, int *next_ch)
+{
+    for (;;) {
+        switch (*next_ch) {
+        case EOF:
+            return Tok_End;
+        case ';':
+            while (*next_ch != EOF &&
+                    *next_ch != ' ') {
+                *next_ch = fgetc(in);
+            }
+            break;
+        case '(':
+            *next_ch = fgetc(in);
+            return Tok_Left_Paren;
+        case ')':
+            *next_ch = fgetc(in);
+            return Tok_Right_Paren;
+        case ' ':
+        case '\a':
+        case '\n':
+        case '\r':
+        case '\t':
+        case '\v':
+            *next_ch = fgetc(in);
+            break;
+        default:
+            {
+                int i = 0;
+
+                while (*next_ch != EOF &&
+                        *next_ch != '(' &&
+                        *next_ch != ')' &&
+                        *next_ch != ';' &&
+                        *next_ch != ' ' &&
+                        *next_ch != '\a' &&
+                        *next_ch != '\n' &&
+                        *next_ch != '\r' &&
+                        *next_ch != '\t' &&
+                        *next_ch != '\v') {
+                    tok_str[i++] = *next_ch;
+                    *next_ch = fgetc(in);
+                }
+                tok_str[i] = '\0';
+                if (strcmp(tok_str, "declare-datatypes") == 0) {
[…]

### 2014-12-10 tools/undata.c
more

@@ -46,7 +46,7 @@ int get_token(FILE *in, int *next_ch)
             return Tok_End;
         case ';':
             while (*next_ch != EOF &&
-                    *next_ch != ' ') {
+                    *next_ch != '\n') {
                 *next_ch = fgetc(in);
             }
             break;
@@ -132,10 +132,10 @@ int main(int argc, char **argv)
         return 1;
     }
 
-    token tok;
     int next_ch = fgetc(in);
+    token tok = get_token(in, &next_ch);
 
-    while ((tok = get_token(in, &next_ch)) != Tok_End) {
+    while (tok != Tok_End) {
         if (tok == Tok_Left_Paren) {
             tok = get_token(in, &next_ch);
 
@@ -162,16 +162,19 @@ int main(int argc, char **argv)
                     } else if (tok == Tok_Right_Paren) {
                         --depth;
                     }
+                    tok = get_token(in, &next_ch);
                 }
 
                 fputc('\n', outNix);
                 fputc('\n', outData);
                 fputc('\n', outCodata);
             }
-        } else if (tok == Tok_Right_Paren) {
+        } else {
             print_token(outNix, tok);
             print_token(outData, tok);
             print_token(outCodata, tok);
+
+            tok = get_token(in, &next_ch);
         }
     }
 

### 2014-12-10 tex/codatadp.tex
Save space on example, begin discussion of model values.

@@
 in which case $t_1$ must be equal to $t_2$.
-We demonstrate this rule concretely with the following example.
-\rem{ensure that \const{LCons} understood}
+We demonstrate an application of this rule concretely with the following example.
 
 \begin{example}
-Let $\Ec$ be the set $\{ x \teq \const{LCons}(y,\const{LCons}(y,x)) \}$.
+Let $\const{C}$ be a constructor of type $\tau_1 \times \tau_2 \rightarrow \tau_2$,
+and let $\Ec$ be the set $\{ x \teq \const{C}(y,\const{C}(y,x)) \}$.
 After applying the rules in Figure~\ref{fig:cc-rules} to completion, the equivalence classes of $\Ec$ are
-$\{ x, \const{LCons}(y,\const{LCons}(y,x)) \}$, $\{ \const{LCons}(y,x) \}$, and $\{ y \}$.
-The set $\tpath{\Ec}{\ec{x}}$ is:
-\[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
-\{ & \ec{x} \rightarrow^{\const{LCons}}_1 & \ec{y}, & & \\ 
-   & \ec{x} \rightarrow^{\const{LCons}}_2 & \ec{\const{LCons}(y,x)} \rightarrow^{\const{LCons}}_1 & \ec{y},& \\
-   & \ec{x} \rightarrow^{\const{LCons}}_2 & \ec{\const{LCons}(y,x)} \rightarrow^{\const{LCons}}_2 & \ec{x} & \}.
+$\{ x, \const{C}(y,\const{C}(y,x)) \}$, $\{ \const{C}(y,x) \}$, and $\{ y \}$.
+The sets $\tpath{\Ec}{\ec{x}}$ and $\tpath{\Ec}{\ec{\const{C}(y,x)}}$ are:
+\[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
+\{ & \ec{x} \rightarrow^{\const{C}}_1 & \ec{y}, & & & 
+\{ & \ec{\const{C}(y,x)} \rightarrow^{\const{C}}_1 & \ec{y}, & & \\ 
+   & \ec{x} \rightarrow^{\const{C}}_2 & \ec{\const{C}(y,x)} \rightarrow^{\const{C}}_1 & \ec{y},&  &
+   & \ec{\const{C}(y,x)} \rightarrow^{\const{C}}_2 & \ec{x} \rightarrow^{\const{C}}_1 & \ec{y},& \\
+   & \ec{x} \rightarrow^{\const{C}}_2 & \ec{\const{C}(y,x)} \rightarrow^{\const{C}}_2 & \ec{x} & \} &
+   & \ec{\const{C}(y,x)} \rightarrow^{\const{C}}_2 & \ec{x} \rightarrow^{\const{C}}_2 & \ec{\const{C}(y,x)} & \}.
 \end{array}\]
-The set $\tpath{\Ec}{\ec{\const{LCons}(y,x)}}$ is:
-\[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
-\{ & \ec{\const{LCons}(y,x)} \rightarrow^{\const{LCons}}_1 & \ec{y}, & & \\ 
-   & \ec{\const{LCons}(y,x)} \rightarrow^{\const{LCons}}_2 & \ec{x} \rightarrow^{\const{LCons}}_1 & \ec{y},& \\
-   & \ec{\const{LCons}(y,x)} \rightarrow^{\const{LCons}}_2 & \ec{x} \rightarrow^{\const{LCons}}_2 & \ec{\const{LCons}(y,x)} & \}.
-\end{array}\]
-Thus, we may infer $x \teq \const{LCons}(y,x)$ by the rule \rn{Bisimilar}
-for the mapping $\sigma :=$ $\{ \ec{x} \mapsto \ec{\const{LCons}(y,x)}, \ec{\const{LCons}(y,x)} \mapsto \ec{x} \}$.
+Thus, by the rule \rn{Bisimilar}
+for the mapping $\sigma :=$ $\{ \ec{x} \mapsto \ec{\const{C}(y,x)}, \ec{\const{C}(y,x)} \mapsto \ec{x} \}$,
+we may infer $x \teq \const{C}(y,x)$.
 \qed
@@
 
-Given a set of disjoint paths $P$ where each $p \in P$
-we write $\interp{P}$ to denote the value
+%Given a set of disjoint paths $P$ where each $p \in P$
+%we write $\interp{P}$ to denote the value
+
+Recall that the $\mu$-notation can be used to represent (possibly cyclic) values taken from an initial algebra, \rem{elaborate?} 
+e.g. $\mu x. \const{C}( \const{0}, x )$ represents the infinite sequence $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
+In this setting, a value can be viewed as a term-like structure where all subterms 
+(which we will refer to as \emph{subvalues}) are either 
+$\mu x. \const{C}( \vec v )$ for some constructor $\const{C}$ and subvalues $\vec v$
+\footnote{We assume all such values are well-typed.}, or a variable $x$.
+We will refer to values of this form as \emph{$\mu$-values}.
+We say a $\mu$-value is \emph{closed} if it contains no free variables.
+For convienience, we will represent the values of both datatypes and co-datatypes as closed $\mu$-values,
+where the case of datatypes adds the restriction that none of its subvalues are variables.
+
+To reason about the uniqueness of $\mu$-values, we introduce the following notions.
+First, the \emph{interpretation of subvalue $u$ in $v$} is the value $v\interp{u}^\emptyset$ returned by the following,
+where $x \mapsto \mu x. C( \vec t ) \in v$ indicates that $\mu x. C( \vec t )$ is the subvalue of $v$ that binds this occurrence of variable $x$.
+
+\[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
+v\interp{x}^M & = & x & x \in M \\
+v\interp{x}^M & = & v\interp{\mu x. C( \vec t )}^M & x \not\in M, x \mapsto \mu x. C( \vec t ) \in v  \\
+v\interp{\mu x. C( \vec t )}^M & = & x & x \in M \\
+v\interp{\mu x. C( \vec t )}^M & = & \mu x. C( v\interp{ \vec t }^{M \cup \{ x \}} ) & x \not\in M
+\end{array}\]
+
+We say a (closed) $\mu$-value $v$ of the form $\mu x. \const{C}( \vec w )$ is \emph{normal} if and only if 
+$v\interp{u}^\emptyset$ is not equivalent to $v$ modulo renaming for all of its strict subvalues $u$,
+and all of its strict closed subvalues are normal.
+For instance, the value $v$ of the form $\mu x. \const{S}( \mu y. \const{S}( x ) )$ is not normal, 
+since 
+$v\interp{\mu y. \const{S}( x )}^\emptyset =$ 
+$\mu y. \const{S}( v\interp{x}^{\{y\}} ) =$ 
+$\mu y. \const{S}( v\interp{\mu x. \const{S}( \mu y. \const{S}( x ) ) }^{\{y\}} ) =$ 
+$\mu y. \const{S}( \mu x. \const{S}( v\interp{\mu y. \const{S}( x )}^{\{x,y\}} ) ) =$ 
+$\mu y. \const{S}( \mu x. \const{S}( y ) )$,
+which is equivalent to $v$ modulo the renaming $\{ x \mapsto y, y \mapsto x \}$.
 
@@
 
-\begin{comment}
-\rn{Ord}
-\(
-\inferrule{
-  \ec{x} \not\in \vdom(\Val)
-  \quad
-  x : \tau
-  \quad
-  \tau \in \Nondata
-  \quad
-  c \not\in \vrange(\Val)
-}{
-  \Val := \Val \cup \ec{x} \mapsto c
-}
-\)
-
-\rn{D-Ind}
-\(
-\inferrule{
-  \ec{x} \not\in \vdom(\Val)
-  \quad
-  x : \tau
-  \quad
-  \tau \in \Data
-  \quad
-  C( \vec t ) \in \ec{x}
-  \quad
-  \ec{\vec t} \in \vdom(\Val)
-}{
-  \Val := \Val \cup \ec{x} \mapsto C( \Val \ec{\vec t}  )
-}
-\)
-
-\rn{D-Base}
-\(
-\inferrule{
-  \ec{x} \not\in \vdom(\Val)
-  \quad
-  x : \tau
-  \quad
-  \tau \in \Data
-  \quad
-  \ec{x} \text{ ctr-free}
-  \quad
-  c \not\in \vrange(\Val)
-}{
-  \Val := \Val \cup \ec{x} \mapsto c
-}
-\)
-\end{comment}
-
 \rn{Ind}
@@
 \inferrule{
-  %x : \tau
-  %\quad
-  %\tau \in \Codata
   v_{\ec{x}} \in FV( \Val )
@@
 \inferrule{
-  %x : \tau
-  %\quad
-  %\tau \in \Codata
-  %\quad
   v_{\ec{x}} \in FV( \Val )
-  %\quad
-  %\ec{x} \text{ ctr-free}
   \quad
@@
 \rem{define $\sim$}
-We assume that first \rn{Ind} is applied exhaustively, after which \rn{Base} is applied exhaustively.
-%the rule \rn{D-Ind} is applied to completion before applying \rn{D-Base}.
+
 In the rule \rn{Ind}, we write $FV( \Val )$ to denote the free variables occurring in $\Val$.
 Notice that this excludes variables beneath $\mu$-binders.
-
 As an example,
@@
 
-It is clear that after applying these rules to completion, $\Val$ is a mapping from equivalence classes to ground values.
+In our construction, the rule \rn{Ind} is first applied exhaustively.
+Since the height of any value mapped to by $\Val$ is bounded by the number of equivalence classes of $\Fc$,
+the rule \rn{Ind} can be applied only a finite number of times.
+Second, \rn{Base} is applied exhaustively.
+Since \rn{Base} removes one variable from the set $FV( \Val )$ and does not add any additional variables to $FV( \Val )$, 
+the rule \rn{Base} can only be applied a finite number of times and does not trigger any new applications of \rn{Ind}.
+
+We first argue that $\Val$ contains only ground values.
+First, since \rn{Split} does not apply to $\Fc$,
+all equivalence classes of finite type contain constructor terms.
+Thus, after applying \rn{Ind} exhaustively, the free variables of $\Val$ all have infinite type.
+
+%It is clear that after applying these rules to completion, $\Val$ is a mapping from equivalence classes to ground values.
 

### 2014-12-10 tools/undata.c
more

@@ -12,9 +12,11 @@ typedef enum {
 
 const int BIG_NUM = 65536;
 
+int next_ch;
+token tok;
 char tok_str[BIG_NUM];
 
-void print_token(FILE *out, token tok)
+void print_token(FILE *out)
 {
     switch (tok) {
     case Tok_Left_Paren:
@@ -38,23 +40,23 @@ void print_token(FILE *out, token tok)
     }
 }
 
-int get_token(FILE *in, int *next_ch)
+int get_token(FILE *in)
 {
     for (;;) {
-        switch (*next_ch) {
+        switch (next_ch) {
         case EOF:
             return Tok_End;
         case ';':
-            while (*next_ch != EOF &&
-                    *next_ch != '\n') {
-                *next_ch = fgetc(in);
+            while (next_ch != EOF &&
+                    next_ch != '\n') {
+                next_ch = fgetc(in);
             }
             break;
         case '(':
-            *next_ch = fgetc(in);
+            next_ch = fgetc(in);
             return Tok_Left_Paren;
         case ')':
-            *next_ch = fgetc(in);
+            next_ch = fgetc(in);
             return Tok_Right_Paren;
         case ' ':
         case '\a':
@@ -62,24 +64,24 @@ int get_token(FILE *in, int *next_ch)
         case '\r':
         case '\t':
         case '\v':
-            *next_ch = fgetc(in);
+            next_ch = fgetc(in);
             break;
         default:
             {
                 int i = 0;
 
-                while (*next_ch != EOF &&
-                        *next_ch != '(' &&
-                        *next_ch != ')' &&
-                        *next_ch != ';' &&
-                        *next_ch != ' ' &&
-                        *next_ch != '\a' &&
-                        *next_ch != '\n' &&
-                        *next_ch != '\r' &&
-                        *next_ch != '\t' &&
-                        *next_ch != '\v') {
-                    tok_str[i++] = *next_ch;
-                    *next_ch = fgetc(in);
+                while (next_ch != EOF &&
+                        next_ch != '(' &&
+                        next_ch != ')' &&
+                        next_ch != ';' &&
+                        next_ch != ' ' &&
+                        next_ch != '\a' &&
+                        next_ch != '\n' &&
+                        next_ch != '\r' &&
[…]

### 2014-12-10 tex/codatadp.tex
More work on mu-values.

@@
 \newcommand{\Val}{\mathcal{A}}
+\newcommand{\nf}[1]{{#1}\!\!\downarrow}
+\newcommand{\vsim}{\sim}
 
@@
 $\mu x. \const{C}( \vec v )$ for some constructor $\const{C}$ and subvalues $\vec v$
-\footnote{We assume all such values are well-typed.}, or a variable $x$.
+\footnote{We assume all values are well-typed. 
+Here, the type of $\mu x. \const{C}( \vec v )$ and $x$ is the return type of $\const{C}$.}, or a variable $x$.
 We will refer to values of this form as \emph{$\mu$-values}.
@@
 
-To reason about the uniqueness of $\mu$-values, we introduce the following notions.
-First, the \emph{interpretation of subvalue $u$ in $v$} is the value $v\interp{u}^\emptyset$ returned by the following,
+We introduce the following notions concerning $\mu$-values.
+First, we write $v \vsim u$ if $\mu$-values $v$ and $u$ are equivalent modulo renaming of $\mu$-bound variables.
+The \emph{interpretation of subvalue $u$ in $v$} is the value $v\interp{u}^\emptyset$ returned by the following recursive procedure,
 where $x \mapsto \mu x. C( \vec t ) \in v$ indicates that $\mu x. C( \vec t )$ is the subvalue of $v$ that binds this occurrence of variable $x$.
-
 \[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
-v\interp{x}^M & = & x & x \in M \\
 v\interp{x}^M & = & v\interp{\mu x. C( \vec t )}^M & x \not\in M, x \mapsto \mu x. C( \vec t ) \in v  \\
-v\interp{\mu x. C( \vec t )}^M & = & x & x \in M \\
-v\interp{\mu x. C( \vec t )}^M & = & \mu x. C( v\interp{ \vec t }^{M \cup \{ x \}} ) & x \not\in M
+v\interp{x}^M & = & x & \text{otherwise} \\
+v\interp{\mu x. C( \vec t )}^M & = & \mu x. C( v\interp{ \vec t }^{M \cup \{ x \}} ) & x \not\in M \\
+v\interp{\mu x. C( \vec t )}^M & = & x & \text{otherwise}
 \end{array}\]
-
-We say a (closed) $\mu$-value $v$ of the form $\mu x. \const{C}( \vec w )$ is \emph{normal} if and only if 
-$v\interp{u}^\emptyset$ is not equivalent to $v$ modulo renaming for all of its strict subvalues $u$,
-and all of its strict closed subvalues are normal.
+Given a $\mu$-value $v$ of the form $\mu x. C( \vec w )$, 
+a subvalue $u$ of $v$ is a \emph{bisimilar subvalue} if and only if $v\interp{u}^\emptyset \sim v$.
+We say $v$ is \emph{normal} if and only if all of its bisimilar subvalues are $x$,
+and all of its strict subvalues that do not contain $v$ are also normal.
 For instance, the value $v$ of the form $\mu x. \const{S}( \mu y. \const{S}( x ) )$ is not normal, 
@@
 which is equivalent to $v$ modulo the renaming $\{ x \mapsto y, y \mapsto x \}$.
+For any $\mu$-value $v$ of the form $\mu x. C( \vec w )$, we may recursively construct a normal value $\nf{v}$
+(which we will call the \emph{normal form for $v$})
+that is observationally equivalent to $v$.
+We construct $\nf{v}$ by replacing all of the bisimilar subvalues of $v$ with $x$,
+and by replacing all of the non-normal closed subvalues of $v$ with their corresponding normal forms.
+For instance, $\nf{(\mu x. \const{S}( \mu y. \const{S}( x ) ))} = \mu x. \const{S}( x )$.
 

### 2014-12-10 tex/codatadp.tex
Refactor cyclic/bisimilar based on new definitions.

@@
 
-We present the calculus in three sections, shown in Figures~\ref{fig:cc-rules},~\ref{fig:ab-rules}, and~\ref{fig:split-rule}.
+We present the calculus in three parts.
+In the first part we compute the bidirectional closure of $\Ec$,
+in the second part we make inferences based on cyclicity and bisimilarity,
+and in the third part we (when necessary) split on the constructor type for various terms in $\tEc$.
 Following the conventions from~\cite{}, 
@@
 
+\paragraph{Step 1: Compute Birectional Closure}
 Figure~\ref{fig:cc-rules} gives the basic rules of our calculus.
@@
   \quad
-  \ttpath{\Ec}{\ec{t}}{\ec{t}} \neq \emptyset
+  %\ttpath{\Ec}{\ec{t}}{\ec{t}} \neq \emptyset
+  \Val \ec{t} = \mu x. t'
+  \quad
+  x \in FV( t' )
 }{
@@
  \quad
- \tpath{\Ec}{\ec{t_1}} = \tpath{\Ec}{\ec{t_2}}\sigma \neq \emptyset
+ %\tpath{\Ec}{\ec{t_1}} = \tpath{\Ec}{\ec{t_2}}\sigma \neq \emptyset
+ \Val \ec{t_1} \vsim \Val \ec{t_2}
 }{
@@
 
-Before applying the rules in Figure~\ref{fig:ab-rules}, 
+\paragraph{Step 2: Apply cyclicity and bisimilarity.}
+In this step,
 we assume that $\Ec$ is saturated with respect to the rules in Figure~\ref{fig:cc-rules}, that is,
@@
 
+Recall that the $\mu$-notation can be used to represent (possibly cyclic) terms and values,
+for instance, $\mu x. \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
+Formally, we recursively define a $\mu$-term is either of the form
+$\mu x. \const{C}( \vec t )$ for some constructor $\const{C} \in \Ctr$ and $\mu$-terms $\vec t$, 
+or a variable $x$.
+We assume all $\mu$-terms are \emph{well-typed}, which we defined recursively:
+if the type of $\const{C}$ is $\vec \tau \rightarrow \upsilon$, 
+then if $\vec v$ are well-type $\mu$-terms under the assumption that $x$ has type $\upsilon$,
+then $\mu x. \const{C}( \vec v )$ is a well-typed term of type $\upsilon$.
+We say a $\mu$-term is \emph{closed} if it contains no free variables.
+We write $t \vsim s$ if $\mu$-terms $t$ and $s$ are syntactically equivalent modulo renaming of $\mu$-bound variables.
+For convienience, we will use $\mu$-terms to refer to (classes of) values for both co-datatype and datatype terms,
+where in the latter case, a $\mu$-term with a bound variable denotes an invalid term.
+%where the latter case adds the restriction on $\mu$-terms $t$ that no subterms of $t$ contain are bound variables.
+
+At the beginning of this step, 
+we compute a mapping $\Val$ from equivalence classes to $\mu$-terms.
+Informally, the terms mapped to by $\Val$ describe the (class of) values that terms in each equivalence class can take in the models of $\Ec$. 
+For each equivalence class $\ec{x}$ of $\Ec$, we associate a fresh variable $v_{\ec{x}}$ of the same type as $x$ not occurring in $\tEc$,
+and initially say that $\Val$ contains $\ec{x} \mapsto v_{\ec{x}}$ for each equivalence class $\ec{x}$ of $\Ec$ 
+(in other words, there are initially no constraints on the values for terms).
+Then, $\Val$ is updated by applying the following rule exhaustively:
+
+\(
+\inferrule{
+  v_{\ec{x}} \in FV( \Val )
+  \quad
+  C( t_1, \ldots, t_n ) \in \ec{x}
+  \quad
+  C \in \Ctr
+}{
+  \Val := \Val \{ v_{\ec{x}} \mapsto \mu v_{\ec{x}}. C( v_{\ec{t_1}}, \ldots, v_{\ec{t_n}} ) \}
+}
+\)
+
+Here, we write $FV( \Val )$ to denote the free variables occurring in $\Val$, which
+excludes variables that occur beneath $\mu$-binders.
+It is straightforward to show that the height of terms produced as a result of this rule
+is bounded by the number of equivalence classes of $\Ec$,
+our construction of $\Val$ is terminating.
+
+\begin{example}
+Say that $\Ec$ contains equivalence classes $\ec{x}$ and $\ec{y}$ where $\const{C}(z,x) \in \ec{y}$ and $\const{C}(w,y) \in \ec{x}$.
+A possible sequence of applications of the above rule is shown below.
+In the first column, we show the equivalence class that was considered,
+and in the second column, we show the state of $\Val$.
+\[\begin{array}{|l|@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l|}
+\hline
+\text{Eqc} & \multicolumn{3}{c}{\Val} & \ 
+\\
+\hline
+& \{ & \ec{x} \mapsto v_{\ec{x}}, & \ec{z} \mapsto v_{\ec{z}}, & \\
+&    & \ec{y} \mapsto v_{\ec{y}}, & \ec{w} \mapsto v_{\ec{w}} & \} \\
+\ec{x}
+& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ), & \ec{z} \mapsto v_{\ec{z}}, & \\
+&    & \ec{y} \mapsto v_{\ec{y}}, & \ec{w} \mapsto v_{\ec{w}} & \} \\
+\ec{y}
+& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ) ), & \ec{z} \mapsto v_{\ec{z}}, & \\
+&    & \ec{y} \mapsto \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
+\ec{x}
+& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ) ), & \ec{z} \mapsto v_{\ec{z}}, & \\
+&    & \ec{y} \mapsto \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ) ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
+\hline
+\end{array}\]
+In other words, the (class of) values for $v_{\ec{x}}$ in models of $\Ec$ 
+is $\const{C}( v_{\ec{w}}, \const{C}( v_{\ec{z}}, \ldots ))$, 
+and the class of values for $v_{\ec{y}}$ is $\const{C}( v_{\ec{z}}, \const{C}( v_{\ec{w}}, \ldots ))$,
+where $v_{\ec{z}}$ and $v_{\ec{w}}$ are free variables.
+$\Box$
+\end{example}
+
+Given $\Val$, the rules for cyclicity and bisimilar can be stated as follows, and are shown in Figure~\ref{fig:ab-rules}.
+For cyclicity, if there is an equivalence class $\ec{t}$ of datatype type whose possible values $\mu x. t'$ are cyclic 
+(as indicated by the assumption $x \in FV( t' )$),
+then we can conclude that $\Ec$ is unsatisfiable.
+For bisimilarity, if there are two equivalence classes $\ec{t_1}$ and $\ec{t_2}$ 
+of co-datatype type whose possible values $\Val \ec{t_1}$ and $\Val \ec{t_2}$ are equivalent modulo renaming of $\mu$-bound variables,
+then we can conclude $t_1 \teq t_2$.
+We demonstrate an application of this rule concretely with the following example.
+
+\begin{example}
+Let $\const{C}$ be a constructor of type $\tau_1 \times \tau_2 \rightarrow \tau_2$,
+and let $\Ec$ be the set $\{ x \teq \const{C}(y,\const{C}(y,x)) \}$.
+After applying the rules in Figure~\ref{fig:cc-rules} to completion, the equivalence classes of $\Ec$ are
+$\{ x, \const{C}(y,\const{C}(y,x)) \}$, $\{ \const{C}(y,x) \}$, and $\{ y \}$.
+Our construction of $\Val$ gives us:
+\[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l}
+\Val \ec{x} & = & 
+\mu v_{\ec{x}}. \const{C}( v_{\ec{y}}, \mu v_{\ec{\const{C}(y,x)}}. \const{C}( v_{\ec{y}}, v_{\ec{x}} ) ) \\
+\Val \ec{\const{C}(y,x)} & = & 
+\mu v_{\ec{\const{C}(y,x)}}. \const{C}( v_{\ec{y}}, \mu v_{\ec{x}}. \const{C}( v_{\ec{y}}, v_{\ec{\const{C}(y,x)}} ) )
+\end{array}\]
+Notice that $\Val \ec{x}$ and $\Val \ec{\const{C}(y,x)}$ are syntactically equivalent for the renaming
+$\{ v_{\ec{x}} \mapsto v_{\ec{\const{C}(y,x)}}, v_{\ec{\const{C}(y,x)}} \mapsto v_{\ec{x}} \}$,
+and thus our calculus may infer the equality $x \teq \const{C}(y,x)$.
+\qed
+\end{example}
+
+\begin{comment}
 To explain the rules in Figure~\ref{fig:ab-rules}, we introduce the following notions.
@@
 We demonstrate an application of this rule concretely with the following example.
-
-\begin{example}
-Let $\const{C}$ be a constructor of type $\tau_1 \times \tau_2 \rightarrow \tau_2$,
-and let $\Ec$ be the set $\{ x \teq \const{C}(y,\const{C}(y,x)) \}$.
-After applying the rules in Figure~\ref{fig:cc-rules} to completion, the equivalence classes of $\Ec$ are
-$\{ x, \const{C}(y,\const{C}(y,x)) \}$, $\{ \const{C}(y,x) \}$, and $\{ y \}$.
-The sets $\tpath{\Ec}{\ec{x}}$ and $\tpath{\Ec}{\ec{\const{C}(y,x)}}$ are:
-\[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
-\{ & \ec{x} \rightarrow^{\const{C}}_1 & \ec{y}, & & & 
-\{ & \ec{\const{C}(y,x)} \rightarrow^{\const{C}}_1 & \ec{y}, & & \\ 
-   & \ec{x} \rightarrow^{\const{C}}_2 & \ec{\const{C}(y,x)} \rightarrow^{\const{C}}_1 & \ec{y},&  &
-   & \ec{\const{C}(y,x)} \rightarrow^{\const{C}}_2 & \ec{x} \rightarrow^{\const{C}}_1 & \ec{y},& \\
-   & \ec{x} \rightarrow^{\const{C}}_2 & \ec{\const{C}(y,x)} \rightarrow^{\const{C}}_2 & \ec{x} & \} &
-   & \ec{\const{C}(y,x)} \rightarrow^{\const{C}}_2 & \ec{x} \rightarrow^{\const{C}}_2 & \ec{\const{C}(y,x)} & \}.
-\end{array}\]
-Thus, by the rule \rn{Bisimilar}
-for the mapping $\sigma :=$ $\{ \ec{x} \mapsto \ec{\const{C}(y,x)}, \ec{\const{C}(y,x)} \mapsto \ec{x} \}$,
-we may infer $x \teq \const{C}(y,x)$.
-\qed
-\end{example}
+\end{comment}
 
@@
 
+\paragraph{Step 3: Branching}
 \rem{Require : all co-datatypes either have infinite cardinality, or have cardinality 1.}
@@
 
+\rem{Revise : now $\Val$ is constructed by calculus.}
 Recall that the $\mu$-notation can be used to represent (possibly cyclic) values taken from an initial algebra, \rem{elaborate?} 

### 2014-12-10 tex/codatadp.tex
More refactor.  Now somewhat clean.

@@
 in the second part we make inferences based on cyclicity and bisimilarity,
-and in the third part we (when necessary) split on the constructor type for various terms in $\tEc$.
+and in the third part (when necessary) we branch on constructor types for various terms in $\tEc$.
 Following the conventions from~\cite{}, 
@@
 or a variable $x$.
-We assume all $\mu$-terms are \emph{well-typed}, which we defined recursively:
+We consider only $\mu$-terms that are \emph{well-typed}, which we define recursively:
 if the type of $\const{C}$ is $\vec \tau \rightarrow \upsilon$, 
@@
 and initially say that $\Val$ contains $\ec{x} \mapsto v_{\ec{x}}$ for each equivalence class $\ec{x}$ of $\Ec$ 
-(in other words, there are initially no constraints on the values for terms).
+(in other words, there are initially no constraints on the values for any equivalence class).
 Then, $\Val$ is updated by applying the following rule exhaustively:
@@
 is bounded by the number of equivalence classes of $\Ec$,
-our construction of $\Val$ is terminating.
+and thus our construction of $\Val$ is terminating.
 
@@
 
-Given $\Val$, the rules for cyclicity and bisimilar can be stated as follows, and are shown in Figure~\ref{fig:ab-rules}.
-For cyclicity, if there is an equivalence class $\ec{t}$ of datatype type whose possible values $\mu x. t'$ are cyclic 
-(as indicated by the assumption $x \in FV( t' )$),
+Given $\Val$, the rules for cyclicity and bisimilar (shown in Figure~\ref{fig:ab-rules}) can be stated as follows.
+For cyclicity, if there is an equivalence class $\ec{t}$ of datatype type whose class of values $\Val \ec{t} = \mu x. t'$ are cyclic,
+as indicated by the assumption $x \in FV( t' )$,
 then we can conclude that $\Ec$ is unsatisfiable.
 For bisimilarity, if there are two equivalence classes $\ec{t_1}$ and $\ec{t_2}$ 
-of co-datatype type whose possible values $\Val \ec{t_1}$ and $\Val \ec{t_2}$ are equivalent modulo renaming of $\mu$-bound variables,
+of co-datatype type whose classes of values $\Val \ec{t_1}$ and $\Val \ec{t_2}$ are equivalent modulo renaming of $\mu$-bound variables,
 then we can conclude $t_1 \teq t_2$.
-We demonstrate an application of this rule concretely with the following example.
+We demonstrate an application of the rule for bisimilarity concretely with the following example.
 
@@
 $\{ x, \const{C}(y,\const{C}(y,x)) \}$, $\{ \const{C}(y,x) \}$, and $\{ y \}$.
-Our construction of $\Val$ gives us:
-\[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l}
+We then construct $\Val$ using the aforementioned steps, after which we obtain:
+\[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}r@{\hspace{.5em}}l@{\hspace{.5em}}r@{\hspace{.5em}}l}
 \Val \ec{x} & = & 
-\mu v_{\ec{x}}. \const{C}( v_{\ec{y}}, \mu v_{\ec{\const{C}(y,x)}}. \const{C}( v_{\ec{y}}, v_{\ec{x}} ) ) \\
+\mu v_{\ec{x}}. & \const{C}( v_{\ec{y}}, & \mu v_{\ec{\const{C}(y,x)}}. & \const{C}( v_{\ec{y}}, v_{\ec{x}} ) ) \\
 \Val \ec{\const{C}(y,x)} & = & 
-\mu v_{\ec{\const{C}(y,x)}}. \const{C}( v_{\ec{y}}, \mu v_{\ec{x}}. \const{C}( v_{\ec{y}}, v_{\ec{\const{C}(y,x)}} ) )
+\mu v_{\ec{\const{C}(y,x)}}. & \const{C}( v_{\ec{y}}, & \mu v_{\ec{x}}. & \const{C}( v_{\ec{y}}, v_{\ec{\const{C}(y,x)}} ) )
 \end{array}\]
@@
 $\{ v_{\ec{x}} \mapsto v_{\ec{\const{C}(y,x)}}, v_{\ec{\const{C}(y,x)}} \mapsto v_{\ec{x}} \}$,
-and thus our calculus may infer the equality $x \teq \const{C}(y,x)$.
+and thus our calculus may infer $x \teq \const{C}(y,x)$.
+Intuitively, this equality holds, since the values of $x$ and $\const{C}( y, x )$ are observationally equivalent for any value of $y$.
 \qed
@@
 
-%We first show that for any saturated node $\Fc$, 
-%additional equalities $\Gc$ can be added to $\Fc$ such that each codatatype/datatype equivalence class of $\Fc \cup \Gc$ has a constructor term.
-%For this purpose, we extend the calculus in this section.
-%For a subset of equivalence classes $S$ of $\Ec$, write $\dpath{\Ec}{C,n}{S}$ to denote the set
-%such that $e' \in \dpath{\Ec}{C,n}{S}$ if and only if there exists an $e \in S$ such that $e \rightarrow^{C}_n e'$ is induced by $\Ec$.
-
-\begin{comment}
-\begin{figure}[t]
-\centering
-\begin{tabular}{c}
-\rn{O-Finish}
-\(
-\inferrule{
-  \Ac := x
-  \quad
-  x : \tau
-  \quad
-  \tau \in \Nondata
-  \quad
-  c \not\in \tEc
-}{
-  \Ac := \emptyset \quad \Ec := \Ec \cup \{ x \teq c \}
-}
-\)
-\\[3.7ex]
-\rn{C-Finish}
-\(
-\inferrule{
-  \Ac := x
-  \quad
-  x : \tau
-  \quad
-  \tau \in \Codata
-  \quad
-  C_i \in \Ctr^\tau
-  \quad 
-  C_i \not\in \{ C_j \mid d \in \Dc, C_j( \vec t ) \in d \}
-}{
-  \Ac := \emptyset \quad \Ec := \Ec \cup \{ x \teq C_i( \vec y ) \} \quad \Rc := \Rc \cup \vec y
-}
-\)
-\\[3.7ex]
-\rn{C-Rec}
-\(
-\inferrule{
-  \Ac := x
-  \quad
-  x : \tau
-  \quad
-  \tau \in \Codata
-  \quad
-  C_i \in \Ctr^\tau
-  \quad
-  C_i \text{ infinite arg } j
-}{
-  \Ac := y_j \quad 
-  \Ec := \Ec \cup \{ x \teq C_i( \vec y ) \} \quad 
-  \Rc := \Rc \cup ( \vec y \setminus y_j ) \quad 
-  \Dc := \dpath{\Ec}{C_i,j}{\Dc}
-}
-\)
-\end{tabular}
-\caption{Rules for completion.
-}
-\label{fig:completion-rules}
-\end{figure}
-\end{comment}
-
 %and second show how to construct a valuation for the variables of $\Fc \cup \Gc$.
@@
 \rem{Revise : now $\Val$ is constructed by calculus.}
-Recall that the $\mu$-notation can be used to represent (possibly cyclic) values taken from an initial algebra, \rem{elaborate?} 
-e.g. $\mu x. \const{C}( \const{0}, x )$ represents the infinite sequence $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
-In this setting, a value can be viewed as a term-like structure where all subterms 
-(which we will refer to as \emph{subvalues}) are either 
-$\mu x. \const{C}( \vec v )$ for some constructor $\const{C}$ and subvalues $\vec v$
-\footnote{We assume all values are well-typed. 
-Here, the type of $\mu x. \const{C}( \vec v )$ and $x$ is the return type of $\const{C}$.}, or a variable $x$.
-We will refer to values of this form as \emph{$\mu$-values}.
-We say a $\mu$-value is \emph{closed} if it contains no free variables.
-For convienience, we will represent the values of both datatypes and co-datatypes as closed $\mu$-values,
-where the case of datatypes adds the restriction that none of its subvalues are variables.
-
-We introduce the following notions concerning $\mu$-values.
-First, we write $v \vsim u$ if $\mu$-values $v$ and $u$ are equivalent modulo renaming of $\mu$-bound variables.
+
+First, we require additional terminology concerning $\mu$-terms.
 The \emph{interpretation of subvalue $u$ in $v$} is the value $v\interp{u}^\emptyset$ returned by the following recursive procedure,
@@
 interprets wrongly-applied selectors based on equality information in $\Fc$,
-interprets the variables and $x$ of $\Fc$ based on a mapping $\Val$ from the equivalence classes of $\Fc$ to values.
-Initially, $\Val$ contains $\ec{x} \mapsto v_{\ec{x}}$ for each equivalence class $\ec{x}$ of $\Fc$,
-where $v_{\ec{x}}$ is a fresh variable of the same type as $x$ not occurring in $\Fc$.
-Then, $\Val$ is updated according to the following rules:
-
-\rn{Ind}
-\(
-\inferrule{
-  v_{\ec{x}} \in FV( \Val )
-  \quad
-  C( t_1, \ldots, t_n ) \in \ec{x}
-  \quad
-  C \in \Ctr
-}{
-  \Val := \Val \{ v_{\ec{x}} \mapsto \mu v_{\ec{x}}. C( v_{\ec{t_1}}, \ldots, v_{\ec{t_n}} ) \}
-}
-\)
-
-\rn{Base}
+interprets the variables and $x$ of $\Fc$ based on the valuation $\Val$ which we constructed in step two of our calculus.
+
+Recall that $\Val$ was constructed as a mapping from the equivalence classes of $\Fc$ to $\mu$-terms.
+We must first \emph{complete} $\Val$ so that it maps equivalence classes only to closed $\mu$-terms, that is, terms containing no free variables.
+This is done by exhaustively applying of the following rule:
+
 \(
@@
 \)
-
-\rem{define $\sim$}
-
-In the rule \rn{Ind}, we write $FV( \Val )$ to denote the free variables occurring in $\Val$.
-Notice that this excludes variables beneath $\mu$-binders.
-As an example,
-say that $\Fc$ contains equivalence classes $\ec{x}$ and $\ec{y}$ where $\const{C}(z,x) \in \ec{y}$ and $\const{C}(w,y) \in \ec{x}$, where $z$ and $w$ are variables of integer type.
-A possible application of the above rules, showing the state of $\Val$ after each application, is as follows:
-\[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
-\hline
-\text{Rule} & \multicolumn{4}{c}{\Val} \ 
-\\
-\hline
-& \{ & \ec{x} \mapsto v_{\ec{x}}, & \ec{z} \mapsto v_{\ec{z}}, \\
-&    & \ec{y} \mapsto v_{\ec{y}}, & \ec{w} \mapsto v_{\ec{w}} & \} \\
-\rn{Ind} \ec{x}
-& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ), & \ec{z} \mapsto v_{\ec{z}}, \\
-&    & \ec{y} \mapsto v_{\ec{y}}, & \ec{w} \mapsto v_{\ec{w}} & \} \\
-\rn{Ind} \ec{y}
-& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ) ), & \ec{z} \mapsto v_{\ec{z}}, \\
-&    & \ec{y} \mapsto \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
-\rn{Ind} \ec{x}
-& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ) ), & \ec{z} \mapsto v_{\ec{z}}, \\
-&    & \ec{y} \mapsto \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ) ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
-\rn{Base} \ec{z}
-& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu v_{\ec{y}}. \const{C}( \const{0}, v_{\ec{x}} ) ), & \ec{z} \mapsto \const{0}, \\
-&    & \ec{y} \mapsto \mu v_{\ec{y}}. \const{C}( \const{0}, \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ) ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
-\rn{Base} \ec{w}
-& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( \const{1}, \mu v_{\ec{y}}. \const{C}( \const{0}, v_{\ec{x}} ) ), & \ec{z} \mapsto \const{0}, \\
-&    & \ec{y} \mapsto \mu v_{\ec{y}}. \const{C}( \const{0}, \mu v_{\ec{x}}. \const{C}( \const{1}, v_{\ec{y}} ) ), & \ec{w} \mapsto \const{1} & \} \\
-\end{array}\]
-
-In our construction, the rule \rn{Ind} is first applied exhaustively.
-Since the height of any value mapped to by $\Val$ is bounded by the number of equivalence classes of $\Fc$,
-the rule \rn{Ind} can be applied only a finite number of times.
-Second, \rn{Base} is applied exhaustively.
-Since \rn{Base} removes one variable from the set $FV( \Val )$ and does not add any additional variables to $FV( \Val )$, 
-the rule \rn{Base} can only be applied a finite number of times and does not trigger any new applications of \rn{Ind}.
-
-We first argue that $\Val$ contains only ground values.
+Since this rule removes one variable from the set $FV( \Val )$ and does not add any additional variables to $FV( \Val )$, 
+it can only be applied a finite number of times.
+
+We first argue that $\Val$ contains only to ground values (which include closed $\mu$-terms), in other words, $FV( \Val ) = \emptyset$.
 First, since \rn{Split} does not apply to $\Fc$,
 all equivalence classes of finite type contain constructor terms.
-Thus, after applying \rn{Ind} exhaustively, the free variables of $\Val$ all have infinite type.
+Thus, all variables in $FV( \Val )$ have infinite type.
 

### 2014-12-10 tex/codatadp.tex
More refactor.

@@
 \begin{example}
-Say that $\Ec$ contains equivalence classes $\ec{x}$ and $\ec{y}$ where $\const{C}(z,x) \in \ec{y}$ and $\const{C}(w,y) \in \ec{x}$.
+Say that $\Ec$ contains (distinct) equivalence classes $\ec{x}$, $\ec{y}$, $\ec{z}$ and $\ec{w}$ where $\const{C}(z,x) \in \ec{y}$ and $\const{C}(w,y) \in \ec{x}$.
 A possible sequence of applications of the above rule is shown below.
@@
 \end{array}\]
-In other words, the (class of) values for $v_{\ec{x}}$ in models of $\Ec$ 
-is $\const{C}( v_{\ec{w}}, \const{C}( v_{\ec{z}}, \ldots ))$, 
-and the class of values for $v_{\ec{y}}$ is $\const{C}( v_{\ec{z}}, \const{C}( v_{\ec{w}}, \ldots ))$,
-where $v_{\ec{z}}$ and $v_{\ec{w}}$ are free variables.
+Informally, this indicates that the values for $x$ (resp. $y$) in models of $\Ec$
+are of the form $\const{C}( w, \const{C}( z, \ldots ))$ (resp. $\const{C}( z, \const{C}( w, \ldots ))$) for some value of $w$ and $z$.
 $\Box$
@@
 
-Given $\Val$, the rules for cyclicity and bisimilar (shown in Figure~\ref{fig:ab-rules}) can be stated as follows.
+Given $\Val$, the rules for cyclicity and bisimilarity (shown in Figure~\ref{fig:ab-rules}) can be stated as follows.
 For cyclicity, if there is an equivalence class $\ec{t}$ of datatype type whose class of values $\Val \ec{t} = \mu x. t'$ are cyclic,
@@
 then we can conclude $t_1 \teq t_2$.
-We demonstrate an application of the rule for bisimilarity concretely with the following example.
+We demonstrate the bisimilarity rule concretely with the following example.
 
@@
 \paragraph{Step 3: Branching}
-\rem{Require : all co-datatypes either have infinite cardinality, or have cardinality 1.}
-
-Finally, if neither a rule from Figure~\ref{fig:cc-rules} or Figure~\ref{fig:ab-rules} applies to $\Ec$, 
+If neither a rule from Figure~\ref{fig:cc-rules} or Figure~\ref{fig:ab-rules} applies to $\Ec$, 
 the rules from Figure~\ref{fig:split-rule} are applied.
@@
 Thus, all variables in $FV( \Val )$ have infinite type.
+%\rem{Require : all co-datatypes either have infinite cardinality, or have cardinality 1.}
 

### 2014-12-11 tools/undata.c
more

@@ -10,7 +10,7 @@ typedef enum {
     Tok_End
 } token;
 
-const int BIG_NUM = 65536;
+const int BIG_NUM = 1024;
 
 int next_ch;
 token tok;
@@ -96,7 +96,7 @@ int get_token(FILE *in)
     }
 }
 
-int eliminate_co_datatype(FILE *in)
+int eliminate_co_datatype(FILE *in, FILE *out)
 {
     char typeName[BIG_NUM];
 
@@ -111,17 +111,23 @@ int eliminate_co_datatype(FILE *in)
         return 1;
     }
 
-    while (tok != Tok_Right_Paren) {
+    int depth = 1;
+
+    while (depth != 0 && tok != Tok_End) {
+        if (tok == Tok_Left_Paren) {
+            ++depth;
+        } else if (tok == Tok_Right_Paren) {
+            --depth;
+        }
         tok = get_token(in);
     }
-    tok = get_token(in);
 
     return 0;
 }
 
-int eliminate_co_datatypes(FILE *in)
+int eliminate_co_datatypes(FILE *in, FILE *out)
 {
-    token tok = get_token(in);
+    tok = get_token(in);
     if (tok != Tok_Left_Paren) {
         return 1;
     }
@@ -136,7 +142,7 @@ int eliminate_co_datatypes(FILE *in)
         return 1;
     }
 
-    eliminate_co_datatype(in);
+    eliminate_co_datatype(in, out);
 
     while (tok != Tok_Right_Paren) {
         tok = get_token(in);
@@ -186,9 +192,9 @@ int generate_one_file(const char *inName, int keep_data, int keep_codata)
             tok = get_token(in);
 
             if (tok == Tok_Declare_Datatypes && !keep_data) {
-                eliminate_co_datatypes(in);
+                eliminate_co_datatypes(in, out);
             } else if (tok == Tok_Declare_Codatatypes && !keep_codata) {
-                eliminate_co_datatypes(in);
+                eliminate_co_datatypes(in, out);
             } else {
                 int depth = 1;
 
@@ -225,6 +231,6 @@ int main(int argc, char **argv)
     }
 
     return generate_one_file(argv[1], 0, 0)
-            && generate_one_file(argv[1], 0, 1)
-            && generate_one_file(argv[1], 1, 0);
+            + generate_one_file(argv[1], 0, 1)
+            + generate_one_file(argv[1], 1, 0);
 }

### 2014-12-11 tools/undata.c
more

@@ -96,6 +96,56 @@ int get_token(FILE *in)
     }
 }
 
+int eliminate_constructor_argument(const char *typeName, const char *ctorName,
+        FILE *in, FILE *out)
+{
+    char selName[BIG_NUM];
+    char selSort[BIG_NUM];
+
+    if (tok != Tok_Left_Paren) {
+        return 1;
+    }
+
+    tok = get_token(in);
+    strcpy(selName, tok_str);
+
+    tok = get_token(in);
+    strcpy(selSort, tok_str);
+
+    tok = get_token(in);
+    if (tok != Tok_Right_Paren) {
+        return 1;
+    }
+
+    tok = get_token(in);
+    return 0;
+}
+
+int eliminate_constructor(const char *typeName, FILE *in, FILE *out)
+{
+    char ctorName[BIG_NUM];
+
+    if (tok != Tok_Left_Paren) {
+        return 1;
+    }
+
+    tok = get_token(in);
+    strcpy(ctorName, tok_str);
+
+    tok = get_token(in);
+
+    while (tok == Tok_Left_Paren) {
+        eliminate_constructor_argument(typeName, ctorName, in, out);
+    }
+
+    if (tok != Tok_Right_Paren) {
+        return 1;
+    }
+
+    tok = get_token(in);
+    return 0;
+}
+
 int eliminate_co_datatype(FILE *in, FILE *out)
 {
     char typeName[BIG_NUM];
@@ -107,21 +157,16 @@ int eliminate_co_datatype(FILE *in, FILE *out)
     }
 
     tok = get_token(in);
-    if (tok != Tok_Left_Paren) {
+
+    while (tok == Tok_Left_Paren) {
+        eliminate_constructor(typeName, in, out);
+    }
+
+    if (tok != Tok_Right_Paren) {
         return 1;
     }
 
-    int depth = 1;
-
-    while (depth != 0 && tok != Tok_End) {
-        if (tok == Tok_Left_Paren) {
-            ++depth;
-        } else if (tok == Tok_Right_Paren) {
-            --depth;
-        }
-        tok = get_token(in);
-    }
-
+    tok = get_token(in);
     return 0;
 }
 
@@ -144,7 +189,7 @@ int eliminate_co_datatypes(FILE *in, FILE *out)
 
     eliminate_co_datatype(in, out);
 
-    while (tok != Tok_Right_Paren) {
+    while (tok != Tok_End && tok != Tok_Right_Paren) {
         tok = get_token(in);
     }
     tok = get_token(in);
[…]

### 2014-12-11 tools/undata.c
more

@@ -10,11 +10,11 @@ typedef enum {
     Tok_End
 } token;
 
-const int BIG_NUM = 1024;
+const int Big_Num = 1024;
 
 int next_ch;
 token tok;
-char tok_str[BIG_NUM];
+char tok_str[Big_Num];
 
 void print_token(FILE *out)
 {
@@ -96,21 +96,30 @@ int get_token(FILE *in)
     }
 }
 
-int eliminate_constructor_argument(const char *typeName, const char *ctorName,
-        FILE *in, FILE *out)
+int eliminate_constructor_argument(const char *sort_name,
+        const char *ctor_name, char (*funs)[Big_Num], int *num_funs,
+        char (*args)[Big_Num], int *num_args, FILE *in, FILE *out)
 {
-    char selName[BIG_NUM];
-    char selSort[BIG_NUM];
+    char sel_name[Big_Num];
+    char arg_sort[Big_Num];
 
     if (tok != Tok_Left_Paren) {
         return 1;
     }
 
     tok = get_token(in);
-    strcpy(selName, tok_str);
+    strcpy(sel_name, tok_str);
 
     tok = get_token(in);
-    strcpy(selSort, tok_str);
+    strcpy(arg_sort, tok_str);
+    fprintf(stderr, "%d %s %s %s %s\n", (int)tok, sort_name, ctor_name,
+            sel_name, arg_sort);
+
+    strcpy(*args++, arg_sort);
+
+    char sel_sig[Big_Num];
+    sprintf(sel_sig, "%s (%s)%s", sel_name, sort_name, arg_sort);
+    strcpy(funs[(*num_funs)++], sel_sig);
 
     tok = get_token(in);
     if (tok != Tok_Right_Paren) {
@@ -121,21 +130,38 @@ int eliminate_constructor_argument(const char *typeName, const char *ctorName,
     return 0;
 }
 
-int eliminate_constructor(const char *typeName, FILE *in, FILE *out)
+int eliminate_constructor(const char *sort_name, char (*funs)[Big_Num],
+        int *num_funs, FILE *in, FILE *out)
 {
-    char ctorName[BIG_NUM];
+    char ctor_name[Big_Num];
+    char args[Big_Num][Big_Num];
+    int num_args = 0;
 
     if (tok != Tok_Left_Paren) {
         return 1;
     }
 
     tok = get_token(in);
-    strcpy(ctorName, tok_str);
[…]

### 2014-12-11 tools/undata.c
more

@@ -112,8 +112,8 @@ int eliminate_constructor_argument(const char *sort_name,
 
     tok = get_token(in);
     strcpy(arg_sort, tok_str);
-    fprintf(stderr, "%d %s %s %s %s\n", (int)tok, sort_name, ctor_name,
-            sel_name, arg_sort);
+//    fprintf(stderr, "%d %s %s %s %s\n", (int)tok, sort_name, ctor_name,
+//            sel_name, arg_sort);
 
     strcpy(*args++, arg_sort);
 
@@ -143,7 +143,7 @@ int eliminate_constructor(const char *sort_name, char (*funs)[Big_Num],
 
     tok = get_token(in);
     strcpy(ctor_name, tok_str);
-    fprintf(stderr, "%d %s %s\n", (int)tok, sort_name, tok_str);
+//    fprintf(stderr, "%d %s %s\n", (int)tok, sort_name, tok_str);
 
     char arg_sorts[Big_Num];
     for (int i = 0; i < num_args; i++) {
@@ -184,7 +184,7 @@ int eliminate_co_datatype(char (*funs)[Big_Num], FILE *in, FILE *out)
     tok = get_token(in);
 
     strcpy(sort_name, tok_str);
-    fprintf(stderr, "%d %s\n", (int)tok, sort_name);
+//    fprintf(stderr, "%d %s\n", (int)tok, sort_name);
     tok = get_token(in);
 
     fprintf(out, "(declare-sort %s 0)\n", sort_name);

### 2014-12-11 tools/undata.c
seems to work

@@ -115,7 +115,7 @@ int eliminate_constructor_argument(const char *sort_name,
 //    fprintf(stderr, "%d %s %s %s %s\n", (int)tok, sort_name, ctor_name,
 //            sel_name, arg_sort);
 
-    strcpy(*args++, arg_sort);
+    strcpy(args[(*num_args)++], arg_sort);
 
     char sel_sig[Big_Num];
     sprintf(sel_sig, "%s (%s)%s", sel_name, sort_name, arg_sort);
@@ -145,24 +145,23 @@ int eliminate_constructor(const char *sort_name, char (*funs)[Big_Num],
     strcpy(ctor_name, tok_str);
 //    fprintf(stderr, "%d %s %s\n", (int)tok, sort_name, tok_str);
 
-    char arg_sorts[Big_Num];
-    for (int i = 0; i < num_args; i++) {
-        strcat(arg_sorts, args[i]);
-        if (i > 0) {
-            strcat(arg_sorts, " ");
-        }
-    }
-
-    char ctor_sig[Big_Num];
-    sprintf(ctor_sig, "%s (%s)%s", ctor_name, arg_sorts, sort_name);
-    strcpy(funs[(*num_funs)++], ctor_sig);
-
     tok = get_token(in);
 
     while (tok == Tok_Left_Paren) {
         eliminate_constructor_argument(sort_name, ctor_name, funs, num_funs,
             args, &num_args, in, out);
     }
+
+    char arg_sorts[Big_Num];
+    arg_sorts[0] = '\0';
+    for (int i = 0; i < num_args; i++) {
+        strcat(arg_sorts, args[i]);
+        strcat(arg_sorts, " ");
+    }
+
+    char ctor_sig[Big_Num];
+    sprintf(ctor_sig, "%s (%s)%s", ctor_name, arg_sorts, sort_name);
+    strcpy(funs[(*num_funs)++], ctor_sig);
 
     if (tok != Tok_Right_Paren) {
         return 1;
@@ -172,10 +171,10 @@ int eliminate_constructor(const char *sort_name, char (*funs)[Big_Num],
     return 0;
 }
 
-int eliminate_co_datatype(char (*funs)[Big_Num], FILE *in, FILE *out)
+int eliminate_co_datatype(char (*funs)[Big_Num], int *num_funs, FILE *in,
+        FILE *out)
 {
     char sort_name[Big_Num];
-    int num_funs = 0;
 
     if (tok != Tok_Left_Paren) {
[…]

### 2014-12-11 tools/undata.c
more

@@ -10,7 +10,7 @@ typedef enum {
     Tok_End
 } token;
 
-const int Big_Num = 1024;
+const int Big_Num = 16384;
 
 int next_ch;
 token tok;
@@ -80,6 +80,10 @@ int get_token(FILE *in)
                         next_ch != '\r' &&
                         next_ch != '\t' &&
                         next_ch != '\v') {
+                    if (i >= Big_Num - 1) {
+                        fprintf(stderr, "Token too long\n");
+                        return Tok_End;
+                    }
                     tok_str[i++] = next_ch;
                     next_ch = fgetc(in);
                 }
@@ -134,7 +138,7 @@ int eliminate_constructor(const char *sort_name, char (*funs)[Big_Num],
         int *num_funs, FILE *in, FILE *out)
 {
     char ctor_name[Big_Num];
-    char args[Big_Num][Big_Num];
+    static char args[Big_Num][Big_Num];
     int num_args = 0;
 
     if (tok != Tok_Left_Paren) {
@@ -219,7 +223,7 @@ int eliminate_co_datatypes(FILE *in, FILE *out)
 
     tok = get_token(in);
 
-    char funs[Big_Num][Big_Num];
+    static char funs[Big_Num][Big_Num];
     int num_funs = 0;
 
     for (int i = 0; i < Big_Num; i++) {
@@ -324,12 +328,16 @@ int generate_one_file(const char *in_name, int keep_data, int keep_codata)
 
 int main(int argc, char **argv)
 {
-    if (argc != 2) {
-        fprintf(stderr, "Usage: undata FILE\n");
+    if (argc < 2) {
+        fprintf(stderr, "Usage: undata FILE...\n");
         return 1;
     }
 
-    return generate_one_file(argv[1], 0, 0)
-            + generate_one_file(argv[1], 0, 1)
-            + generate_one_file(argv[1], 1, 0);
+    for (int i = 1; i < argc; i++) {
+        fprintf(stderr, "%s\n", argv[i]);
+        generate_one_file(argv[i], 0, 0);
+        generate_one_file(argv[i], 0, 1);
+        generate_one_file(argv[i], 1, 0);
+    }
+    return 0;
 }

### 2014-12-11 tools/undata.c
release resources

@@ -323,6 +323,8 @@ int generate_one_file(const char *in_name, int keep_data, int keep_codata)
         }
     }
 
+    fclose(in);
+    fclose(out);
     return 0;
 }
 

### 2014-12-11 tex/codatadp.tex
More work on solution soundness, cleanup description of bisimilarity/cyclicity.

@@
 \inferrule{
-  t : \tau
+  t_1 : \tau
   \quad
@@
   %\ttpath{\Ec}{\ec{t}}{\ec{t}} \neq \emptyset
-  \Val \ec{t} = \mu x. t'
+  \Val \ec{t_1} = \mu x. t_2
   \quad
-  x \in FV( t' )
+  x \in FV( t_2 )
 }{
@@
 
-Recall that the $\mu$-notation can be used to represent (possibly cyclic) terms and values,
+%For presentation of the rules in this step, we rely on a representation of terms in the $\mu$-notation.
+The rules of this step contain premises that involve terms containing $\mu$-bindings.
+Recall that these bindings can be used for representing (possibly cyclic) values,
 for instance, $\mu x. \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
-Formally, we recursively define a $\mu$-term is either of the form
+Formally, we recursively define a $\mu$-term as a term-like structure either of the form
 $\mu x. \const{C}( \vec t )$ for some constructor $\const{C} \in \Ctr$ and $\mu$-terms $\vec t$, 
 or a variable $x$.
+For consistency, we will also consider $\mu x. \const{c}$ to be a $\mu$-term
+in the case where $c$ is a constant of type $\tau \in \Nondata$.
 We consider only $\mu$-terms that are \emph{well-typed}, which we define recursively:
-if the type of $\const{C}$ is $\vec \tau \rightarrow \upsilon$, 
-then if $\vec v$ are well-type $\mu$-terms under the assumption that $x$ has type $\upsilon$,
+if the type of $\const{C}$ is $\vec \tau \rightarrow \upsilon$
+and $\vec v$ are well-typed $\mu$-terms of type $\vec \tau$ under the assumption that $x$ has type $\upsilon$,
 then $\mu x. \const{C}( \vec v )$ is a well-typed term of type $\upsilon$.
 We say a $\mu$-term is \emph{closed} if it contains no free variables.
-We write $t \vsim s$ if $\mu$-terms $t$ and $s$ are syntactically equivalent modulo renaming of $\mu$-bound variables.
-For convienience, we will use $\mu$-terms to refer to (classes of) values for both co-datatype and datatype terms,
-where in the latter case, a $\mu$-term with a bound variable denotes an invalid term.
+We write $t \vsim s$ if $\mu$-terms $t$ and $s$ are syntactically equivalent for some renaming of $\mu$-bound variables.
+%For convienience, we will use $\mu$-terms to refer to (classes of) values for both co-datatype and datatype terms,
+%where in the latter case, a $\mu$-term with a bound variable denotes an infeasible term.
 %where the latter case adds the restriction on $\mu$-terms $t$ that no subterms of $t$ contain are bound variables.
 
+Intuitively, a $\mu$-term $t$ of type $\tau$ describes a class of values of $\tau$, 
+either in the case when $\tau$ is a co-datatype or datatype type,
+where in the latter case, such a $t$ containing a bound variable describes an infeasible class of values.
 At the beginning of this step, 
-we compute a mapping $\Val$ from equivalence classes to $\mu$-terms.
-Informally, the terms mapped to by $\Val$ describe the (class of) values that terms in each equivalence class can take in the models of $\Ec$. 
+we compute a mapping $\Val$ from equivalence classes to $\mu$-terms 
+that describe the class of values that terms in that equivalence class can take in the models of $\Ec$. 
 For each equivalence class $\ec{x}$ of $\Ec$, we associate a fresh variable $v_{\ec{x}}$ of the same type as $x$ not occurring in $\tEc$,
@@
 
-Here, we write $FV( \Val )$ to denote the free variables occurring in $\Val$, which
+Here, we write $FV( \Val )$ to denote the \emph{free variables} occurring in $\Val$, which
 excludes variables that occur beneath $\mu$-binders.
-It is straightforward to show that the height of terms produced as a result of this rule
+It is straightforward to see that the height of terms produced as a result of this rule
 is bounded by the number of equivalence classes of $\Ec$,
@@
 In the first column, we show the equivalence class that was considered,
-and in the second column, we show the state of $\Val$.
+and in the second column, we show the content of $\Val$.
 \[\begin{array}{|l|@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l|}
@@
 Informally, this indicates that the values for $x$ (resp. $y$) in models of $\Ec$
-are of the form $\const{C}( w, \const{C}( z, \ldots ))$ (resp. $\const{C}( z, \const{C}( w, \ldots ))$) for some value of $w$ and $z$.
+are of the form $\const{C}( v_{\ec{w}}, \const{C}( v_{\ec{z}}, \ldots ))$ (resp. $\const{C}( v_{\ec{z}}, \const{C}( v_{\ec{w}}, \ldots ))$) 
+for some value of $v_{\ec{w}}$ and $v_{\ec{z}}$.
 $\Box$
@@
 Given $\Val$, the rules for cyclicity and bisimilarity (shown in Figure~\ref{fig:ab-rules}) can be stated as follows.
-For cyclicity, if there is an equivalence class $\ec{t}$ of datatype type whose class of values $\Val \ec{t} = \mu x. t'$ are cyclic,
-as indicated by the assumption $x \in FV( t' )$,
+For cyclicity, if there is an equivalence class $\ec{t_1}$ of datatype type whose class of values $\Val \ec{t_1} = \mu x. t_2$ are cyclic,
+as indicated by the assumption $x \in FV( t_2 )$,
 then we can conclude that $\Ec$ is unsatisfiable.
@@
 and thus our calculus may infer $x \teq \const{C}(y,x)$.
-Intuitively, this equality holds, since the values of $x$ and $\const{C}( y, x )$ are observationally equivalent for any value of $y$.
+Intuitively, this equality holds, since the values of $x$ and $\const{C}( y, x )$ are observationally equivalent for any value of $v_{\ec{y}}$.
 \qed
@@
 For \rn{Unify$_2$}, this is a consequence of distinctness.
-For \rn{Cyclic}, this is a consequence of induction. \rem{elaborate?}
+For \rn{Cyclic}, our construction of $\Val$ indicates that the class of values that $t_1$ can take in models of $\Ec$ is infeasible,
+and thus $\Ec$ is unsatisfiable.
 If the children of $\Ec$ are closed derivation trees 
@@
 For \rn{Unify$_1$}, this is a consequence of injectivity.
-For \rn{Bisimilar}, this is a consequence of co-induction. \rem{elaborate?}
+For \rn{Bisimilar}, our construction of $\Val$ indicates that the value of $t_1$ and $t_2$ are observationally equivalent in all models of $\Ec$.
 For \rn{Singleton}, clearly $t_1$ and $t_2$ must have the same value since the cardinality of their type is one.
@@
 then $\Ec$ is satisfiable $\thD$.
-
-%and second show how to construct a valuation for the variables of $\Fc \cup \Gc$.
-
-%\begin{lemma}
-%\end{lemma}
-
-%Given a set of disjoint paths $P$ where each $p \in P$
-%we write $\interp{P}$ to denote the value
-
-\rem{Revise : now $\Val$ is constructed by calculus.}
-
 First, we require additional terminology concerning $\mu$-terms.
-The \emph{interpretation of subvalue $u$ in $v$} is the value $v\interp{u}^\emptyset$ returned by the following recursive procedure,
-where $x \mapsto \mu x. C( \vec t ) \in v$ indicates that $\mu x. C( \vec t )$ is the subvalue of $v$ that binds this occurrence of variable $x$.
+If $u$ is a subterm of $v$, then the \emph{interpretation of $u$ in $v$} is the term $v\interp{u}^\emptyset$ returned by the following recursive procedure,
+where $x \mapsto \mu x. C( \vec t ) \in v$ indicates that $\mu x. C( \vec t )$ is the subterm of $v$ that binds this occurrence of variable $x$.
 \[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
@@
 \end{array}\]
-Given a $\mu$-value $v$ of the form $\mu x. C( \vec w )$, 
-a subvalue $u$ of $v$ is a \emph{bisimilar subvalue} if and only if $v\interp{u}^\emptyset \sim v$.
-We say $v$ is \emph{normal} if and only if all of its bisimilar subvalues are $x$,
-and all of its strict subvalues that do not contain $v$ are also normal.
-For instance, the value $v$ of the form $\mu x. \const{S}( \mu y. \const{S}( x ) )$ is not normal, 
+Given a $\mu$-term $v$ of the form $\mu x. C( \vec w )$, 
+a subterm $u$ of $v$ is a \emph{bisimilar subterm} if and only if $v\interp{u}^\emptyset \vsim v$.
+We say $v$ is \emph{normal} if and only if all of its bisimilar subterms are $x$,
+and all of its strict subterms that do not contain $v$ are also normal.
+For instance, the $\mu$-term $v$ of the form $\mu x. \const{S}( \mu y. \const{S}( x ) )$ is not normal, 
 since 
@@
 $\mu y. \const{S}( \mu x. \const{S}( y ) )$,
-which is equivalent to $v$ modulo the renaming $\{ x \mapsto y, y \mapsto x \}$.
-For any $\mu$-value $v$ of the form $\mu x. C( \vec w )$, we may recursively construct a normal value $\nf{v}$
-(which we will call the \emph{normal form for $v$})
-that is observationally equivalent to $v$.
-We construct $\nf{v}$ by replacing all of the bisimilar subvalues of $v$ with $x$,
-and by replacing all of the non-normal closed subvalues of $v$ with their corresponding normal forms.
-For instance, $\nf{(\mu x. \const{S}( \mu y. \const{S}( x ) ))} = \mu x. \const{S}( x )$.
+which is syntactically equivalent to $v$ for the renaming $\{ x \mapsto y, y \mapsto x \}$.
+
+%\begin{lemma}
+%\end{lemma}
+
+%For any $\mu$-value $v$ of the form $\mu x. C( \vec w )$, we may recursively construct a normal value $\nf{v}$
+%(which we will call the \emph{normal form for $v$})
+%that is observationally equivalent to $v$.
+%We construct $\nf{v}$ by replacing all of the bisimilar subvalues of $v$ with $x$,
+%and by replacing all of the non-normal closed subvalues of $v$ with their corresponding normal forms.
+%For instance, $\nf{(\mu x. \const{S}( \mu y. \const{S}( x ) ))} = \mu x. \const{S}( x )$.
 
@@
   \quad
+  c \text{ normal}  
+  \quad
   c \not\sim c' \text{ for all } c' \in \vrange(\Val)
@@
 \)
+
 Since this rule removes one variable from the set $FV( \Val )$ and does not add any additional variables to $FV( \Val )$, 
@@
 
-We first argue that $\Val$ contains only to ground values (which include closed $\mu$-terms), in other words, $FV( \Val ) = \emptyset$.
-First, since \rn{Split} does not apply to $\Fc$,
-all equivalence classes of finite type contain constructor terms.
+Let $\Val^\ast$ be the result of applying the above rule to completion on $\Val$.
+We first argue that $FV( \Val^\ast ) = \emptyset$.
+Since \rn{Split} does not apply to $\Fc$,
+all equivalence classes of $\Fc$ finite datatype type contain constructor terms, 
+and thus $FV( \Val )$ contains no variables of finite datatype type.
+For each equivalence classes of $\Fc$ with finite co-datatype $\tau$, 
+we know $\tau$ must have cardinality one, due to \rem{reference}.
+Since \rn{Singleton} does not apply to $\Fc$, 
+there is only one equivalence class of type $\tau$, call it $\ec{x}$.
+If $\Val \ec{x} = v_{\ec{x}}$, then we may apply the above rule for the (single) value $c$ of type $\tau$, 
+for which the second premise holds since $\ec{x}$ is the only equivalence class of type $\tau$.
 Thus, all variables in $FV( \Val )$ have infinite type.
-%\rem{Require : all co-datatypes either have infinite cardinality, or have cardinality 1.}
 

### 2014-12-11 tex/codatadp.tex
More progress on model/proof.

@@
 then $\mu x. \const{C}( \vec v )$ is a well-typed term of type $\upsilon$.
-We say a $\mu$-term is \emph{closed} if it contains no free variables.
-We write $t \vsim s$ if $\mu$-terms $t$ and $s$ are syntactically equivalent for some renaming of $\mu$-bound variables.
+We say a $\mu$-term is \emph{closed} if and only if it contains no free variables.
+A $\mu$-term is \emph{cyclic} if and only if it contains a bound variable.
+We write $t \vsim s$ if $\mu$-terms $t$ and $s$ are syntactically equivalent for some renaming of $\mu$-bound variables,
+in which case we will commonly say $t$ and $s$ are \emph{bisimilar}.
 %For convienience, we will use $\mu$-terms to refer to (classes of) values for both co-datatype and datatype terms,
@@
 First, we require additional terminology concerning $\mu$-terms.
-If $u$ is a subterm of $v$, then the \emph{interpretation of $u$ in $v$} is the term $v\interp{u}^\emptyset$ returned by the following recursive procedure,
+If $u$ is a subterm of $\mu$-term $v$, then the \emph{interpretation of $u$ in $v$} is the $\mu$-term $v\interp{u}^\emptyset$ as returned by the following recursive procedure,
 where $x \mapsto \mu x. C( \vec t ) \in v$ indicates that $\mu x. C( \vec t )$ is the subterm of $v$ that binds this occurrence of variable $x$.
@@
 \end{array}\]
-Given a $\mu$-term $v$ of the form $\mu x. C( \vec w )$, 
-a subterm $u$ of $v$ is a \emph{bisimilar subterm} if and only if $v\interp{u}^\emptyset \vsim v$.
+We will write $v\interp{u}$ as shorthand for $v\interp{u}^\emptyset$.
+Given a $\mu$-term $v$ of the form $\mu x. C( \vec t )$, 
+we say $u$ is a \emph{bisimilar subterm} of $v$ if and only if $v\interp{u} \vsim v$.
 We say $v$ is \emph{normal} if and only if all of its bisimilar subterms are $x$,
-and all of its strict subterms that do not contain $v$ are also normal.
+and all of its strict subterms that do not contain $x$ are normal.
 For instance, the $\mu$-term $v$ of the form $\mu x. \const{S}( \mu y. \const{S}( x ) )$ is not normal, 
@@
 
-We construct a model $\M$ of $\thD$ that interprets constructors and correctly-applied selectors in the obvious way,
+We construct a model $\M$ of $\thD$ that interprets each type $\tau$
+as a maximal set of closed normal $\mu$-terms of that type that are 
+(1) unique up to renaming of $\mu$-bound variables,
+(2) acyclic when $\tau \not\in \Codata$.
+We assume $\M$ interprets constructors and correctly-applied selectors in the obvious way
+\footnote{
+To interpret a correctly-applied selector, if $v = \mu x. C_i( \vec t )$, then $\M( s^j_i( v ) )$ is the value in $\M( \tau )$ that is a renaming of $v\interp{t_j}$,
+where $t_j$ has type $\tau$.
+},
 interprets wrongly-applied selectors based on equality information in $\Fc$,
-interprets the variables and $x$ of $\Fc$ based on the valuation $\Val$ which we constructed in step two of our calculus.
-
-Recall that $\Val$ was constructed as a mapping from the equivalence classes of $\Fc$ to $\mu$-terms.
-We must first \emph{complete} $\Val$ so that it maps equivalence classes only to closed $\mu$-terms, that is, terms containing no free variables.
-This is done by exhaustively applying of the following rule:
-
+interprets the variables of $\Fc$ based on the valuation $\Val$ which was constructed in step two of our calculus.
+
+Recall that $\Val$ is a mapping from the equivalence classes of $\Fc$ to $\mu$-terms.
+We first \emph{complete} $\Val$ so that it maps equivalence classes only to closed $\mu$-terms, that is, terms containing no free variables.
+This is done by exhaustively applying the following rule:
+
+\rn{Complete}
 \(
@@
   \quad
-  c \text{ normal}  
+  c_1 \text{ closed normal}  
   \quad
-  c \not\sim c' \text{ for all } c' \in \vrange(\Val)
+  c_1 \not\sim c_2 \text{ for all } c_2 \in \mathcal{T}(\Val)
 }{
-  \Val := \Val \{ v_{\ec{x}} \mapsto c \}
+  \Val := \Val \{ v_{\ec{x}} \mapsto c_1 \}
 }
@@
 
+We write $\mathcal{T}(\Val)$ to denote the smallest set containing $v\interp{u}$ for each subterm $u$ of $v$ in the range of $\Val$.
 Since this rule removes one variable from the set $FV( \Val )$ and does not add any additional variables to $FV( \Val )$, 
@@
 
-Let $\Val^\ast$ be the result of applying the above rule to completion on $\Val$.
+Let $\Val^\ast$ be the result of applying the above rule to completion on the $\Val$ we constructed during step 2 of our calculus.
 We first argue that $FV( \Val^\ast ) = \emptyset$.
@@
 there is only one equivalence class of type $\tau$, call it $\ec{x}$.
-If $\Val \ec{x} = v_{\ec{x}}$, then we may apply the above rule for the (single) value $c$ of type $\tau$, 
+%It cannot be the case that $v_{\ec{x}} \in FV(\Val^\ast \ec{x})$, then we may apply \rn{Complete} for the (single) value $c$ of type $\tau$, 
+%for which the second premise holds since $\ec{x}$ is the only equivalence class of type $\tau$.
+If it were the case that $v_{\ec{x}} \in FV(\Val^\ast )$,
+then \rn{Complete} would apply for the (single) value $c$ of type $\tau$, 
 for which the second premise holds since $\ec{x}$ is the only equivalence class of type $\tau$.
-Thus, all variables in $FV( \Val )$ have infinite type.
-
-%It is clear that after applying these rules to completion, $\Val$ is a mapping from equivalence classes to ground values.
-
-\begin{eqnarray} \label{eqn:m-bij-dt}
-\Val \ec{x} = \Val \ec{y} \text{ if and only if } \ec{x} = \ec{y}
+Since all $\tau \in \Nondata$ are infinite, all variables in $FV( \Val^\ast )$ have infinite type.
+For any infinite type $\tau$,
+there exists a closed normal $\mu$-term $c_1$ that is not bisimilar to any $c_2 \in \mathcal{T}(\Val^\ast)$, since $\mathcal{T}(\Val^\ast)$ is a finite set. \rem{elaborate?}
+Thus, $FV( \Val^\ast )$ contains no variables of infinite type,
+and consequently $\Val^\ast$ is a mapping from the equivalence classes of $\Fc$ to closed $\mu$-terms.
+We now argue that:
+\begin{eqnarray}
+\label{eqn:m-normal} 
+\Val^\ast \text{ maps to normal $\mu$-terms only},  \\
+\label{eqn:m-bij} 
+\Val^\ast \ec{x} \sim \Val^\ast \ec{y} \text{ if and only if } \ec{x} = \ec{y}
 \end{eqnarray}
-
-\begin{comment}
-We assume that:
-\begin{enumerate}
-\item Each equivalence class of $\Fc$ contains at least one variable,
-\item Each equivalence class $\ec{x}$ of $\Fc$ where $x : \tau$, $\tau \in \Nondata$ contains a constant,
-\item Each equivalence class $\ec{x}$ of $\Fc$ where $x : \tau$, $\tau \in \Data \cup \Codata$ contains a constructor term.
-\end{enumerate}
-The first statement comes with no loss of generality since new equalities of the form $y \teq t$ for fresh variable $y$ can be added to $\Fc$ without affecting its satisfiability.
-The second statement can be assumed for $x : \tau$, since $\tau$ is infinite.
-Hence, we can assign constants for each such equivalence class.
-Moreover, since $\Fc$ contains $\thD$-constraints only, these constants can be chosen arbitrarily.
-The third statement can be assumed by \rem{TODO}.
-
-For each equivalence class $e : \tau$, where $\tau \in \Nondata$ containing constant $c$,
-$\Val(x) = c$ for each $x \in e$.
-For each $e : \tau$, where $\tau \in \Data$ containing the constructor term $C( \vec u )$ where the equivalence classes $\ec{\vec u}$ contains the variables $\vec y$,
-$\Val(x) = C( \Val( \vec y ) )$ for each $x \in e$.
-Since \rn{Cyclic} does not apply, this definition is well-defined. 
-For each $e : \tau$, where $\tau \in \Codata$, 
-$\Val( x ) = \interp{ \tpath{\Fc}{e} }$ for each $x \in e$.
-Since neither \rn{Cong} or \rn{Bisimilar} applies to $\Fc$, we have that for all variables $x$ and $y$:
-\begin{eqnarray} \label{eqn:m-bij-dt}
-\Val(x) = \Val(y) \text{ if and only if } \ec{x} = \ec{y}
-\end{eqnarray}
-\end{comment}
-
-Now, we let $\M( x ) = \Val \ec{x}$ for each variable $x$ in $\Fc$.
+We show this by induction on the number of applications of the rule \rn{Complete}.
+For the base case, note that since \rn{Bisimilar} does not apply to $\Fc$,
+we have that $\Val$ satisfies both~\ref{eqn:m-normal} and~\ref{eqn:m-bij}.
+It satisfies the former since if a $\mu$-term $\mu v_{\ec{x}}. t_1$ had a bisimilar subterm $\mu v_{\ec{y}}. t_2$,
+then the rule \rn{Bisimilar} would have applied to $x$ and $y$. \rem{elaborate?}
+For the inductive case, say that we apply \rn{Complete} to $\Val$ for the substitution $\sigma := \{ v_{\ec{x}} \mapsto c_1 \}$.
+\rem{TODO}
+
+Now, we let $\M( x ) = \Val^\ast \ec{x}$ for each variable $x$ in $\mathcal{T}(\Fc)$.
 We first show $\M$ satisfies all equalities in $t_1 \teq t_2 \in \Fc$, where $t_1$ and $t_2$ have type $\tau$.
@@
 If $\tau \in \Data$,
-we first show that $\M( t ) = \Val \ec{x}$ for all non-selector terms $t$ that occur in datatype equivalence classes $\ec{x}$.
-We show this by induction on the size of $\Val \ec{x}$.
-In the case where $t$ is a variable, then $\M( t ) = \Val \ec{ t } = \Val \ec{ x }$.
+we first show that $\M( t ) = \Val^\ast \ec{x}$ for all non-selector terms $t$ that occur in datatype equivalence classes $\ec{x}$.
+We show this by induction on the size of $\Val^\ast \ec{x}$.
+In the case where $t$ is a variable, then $\M( t ) = \Val^\ast \ec{ t } = \Val^\ast \ec{ x }$.
 In the case where $t$ is a constructor term $C( \vec u )$, 
-then $\Val \ec{x}$ must have been assigned by \rn{Ind} for some constructor term $C'( \vec u' ) \in \ec{x}$.
+then $\Val^\ast \ec{x}$ must have been assigned by \rn{Ind} for some constructor term $C'( \vec u' ) \in \ec{x}$.
 Since \rn{Unify$_2$} does not apply to $\Fc$, we have that $C = C'$.
 Since \rn{Unify$_1$} does not apply to $\Fc$, we have that $\ec{\vec u} = \ec{\vec u'}$.
-By the induction hypothesis, we have $\M( \vec u ) = \M( \vec u' )$, and thus $\M( t ) = C( \M( \vec u ) ) = C'( \M( \vec u ' ) ) = \Val \ec{ x }$.
-
-We now show that $\M( t ) = \Val \ec{x}$ for all selector terms $s^j_k( y )$ that occur in datatype equivalence classes $\ec{x}$.
+By the induction hypothesis, we have $\M( \vec u ) = \M( \vec u' )$, and thus $\M( t ) = C( \M( \vec u ) ) = C'( \M( \vec u ' ) ) = \Val^\ast \ec{ x }$.
+
+We now show that $\M( t ) = \Val^\ast \ec{x}$ for all selector terms $s^j_k( y )$ that occur in datatype equivalence classes $\ec{x}$.
 Since \rn{Split} does not apply, 
@@
 By the previous paragraph, we have $\M( y ) = \M( C_i( s^1_i( y ), \ldots, s^{a_i}_i( y ) ) )$,
-which is $C_i( \Val \ec{ x_1 }, \ldots, \Val \ec{ x_{a_i} } )$.
-If $i = k$, then $\M( t ) = \M( s^j_k( y ) ) = s^j_k( \M( y ) ) = \Val \ec{x_j}$, which is $\Val \ec{ x }$ since $\ec{x_j} = \ec{s^j_i( y )} = \ec{x}$.
+which is $C_i( \Val^\ast \ec{ x_1 }, \ldots, \Val^\ast \ec{ x_{a_i} } )$.
+If $i = k$, then $\M( t ) = \M( s^j_k( y ) ) = s^j_k( \M( y ) ) = \Val^\ast \ec{x_j}$, which is $\Val^\ast \ec{ x }$ since $\ec{x_j} = \ec{s^j_i( y )} = \ec{x}$.
 If $i \neq k$, then $\thD$ does not constraint the value of $\M( t )$,
 and since $s^j_k( \M( y ) )$ is otherwise unconstrained in our model construction since \rn{Cong} does not apply, 
-we may interpret $s^j_k( \M( y ) )$ as $\Val \ec{ x }$. 
-Thus, for all datatype equivalence classes $\ec{x}$, we have $\M( t ) = \Val \ec{x}$ for each term $t \in \ec{x}$, which implies $\M \models t_1 \teq t_2$.
+we may interpret $s^j_k( \M( y ) )$ as $\Val^\ast \ec{ x }$. 
+Thus, for all datatype equivalence classes $\ec{x}$, we have $\M( t ) = \Val^\ast \ec{x}$ for each term $t \in \ec{x}$, which implies $\M \models t_1 \teq t_2$.
 
@@
 Since \rn{Conflict} does not apply, we know $t_1 \teq t_2 \not\in \Fc$ and thus $\ec{t_1}$ and $\ec{t_2}$ are distinct.
-By the previous paragraph and statement~\ref{eqn:m-bij-dt}, $\M( t_1 ) \neq \M( t_2 )$, and thus $\M \models t_1 \tneq t_2$.
+By the previous paragraph and statement~\ref{eqn:m-bij}, $\M( t_1 ) \neq \M( t_2 )$, and thus $\M \models t_1 \tneq t_2$.
 

### 2014-12-11 tex/codatadp.tex
more text in Sect 2

--- 2014-12-10 tex/codatadp.tex
@@
 \newcommand{\Val}{\mathcal{A}}
-\newcommand{\nf}[1]{{#1}\!\!\downarrow}
-\newcommand{\vsim}{\sim}
 
@@
 \newcommand\Data{\Types_{\mathrm{dt}}}
-\newcommand\Codata{\Types_{\mathrm{cdt}}}
+\newcommand\Codata{\Types_{\mathrm{codt}}}
 \newcommand\Nondata{\Types_{\mathrm{ord}}}
@@
 
-Freely generated datatypes are ubiquitous in functional programs and
+Freely generated inductive datatypes are ubiquitous in functional programs and
 logical specifications. They are especially
@@
 To represent infinite values, % such as $\const{S}(\const{S}(\const{S}(\ldots))$,
-a natural choice is to turn to codatatypes (or coinductive datatypes), the
-non-well-founded dual of (inductive) datatypes.
+a natural choice is to turn to \emph{codatatypes} (or coinductive datatypes), the
+non-well-founded dual of (inductive) \relax{datatypes}.
 %
 Although they are sometimes perceived as esoteric, codatatypes have an important
-role to play in computer science. For example, the verified C compiler CompCert
+role to play in computer science. The verified C compiler CompCert
 \cite{leroy-2009}, the verified Java compiler Jinja\-Threads
@@
 with the more conventional, inductive definition
-\[\keyw{datatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat}) \,\mid\, \const{Inf}\]
-%
-the codatatype eliminates one case by unifying the finite and infinite cases.
+\[\keyw{datatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat}) \,\mid\, \const{Infty}\]
+%
+the codatatype avoids one case by unifying the finite and infinite nonzero cases.
 Moreover, equations such as $\const{S}(\infty) \teq \infty$ hold without additional
@@
 $\const{S}(\const{S}(\const{S}(\ldots)))$, which uniquely identifies the
-codatatype value $\infty$.
+value $\infty$.
+
+\nopagebreak
 
@@
 In this \thewordpaper, we introduce a unified decision procedure for ground
-problems involving both datatypes and codatatypes, and discuss extensions 
-of the procedure when handling universal (non-ground) problems.
+problems involving datatypes and codatatypes in combination and discuss extensions 
+of the procedure for handling universal (non-ground) problems.
 The procedure is
@@
 %
-\[\keyw{codatatype}~\;\ty{stream}_{\,\tau} \,=\, \const{SCons}(\tau,\, \ty{stream}_{\,\tau})\]
+\[\keyw{codatatype}~\;\ty{stream}_{\,\tau} \,=\, \const{SCons}(\tau,\: \ty{stream}_{\,\tau})\]
 %
@@
 values internally. For example,
-the term $\const{SCons}(1,\, \mu s.\; \const{SCons}(0,\, \const{SCons}(9,\, s)))$
+the term $\const{SCons}(1,\: \mu s.\; \const{SCons}(0,\: \const{SCons}(9,\: s)))$
 denotes the lasso-shaped sequence $1, 0, 9, 0, 9, 0, 9, \ldots$\afterLdots.
@@
 \begin{paper}
-The proofs of soundness and completeness are included in the technical report
+Detailed proofs of soundness and completeness are included in the technical report
 associated with this paper \cite{our-report}.
@@
 %\[
-%      \keyw{codatatype}\; \,\ty{llist} \,=\, \const{LNil} \,\mid\, \const{LCons}(\ty{int},\, \ty{llist})
+%      \keyw{codatatype}\; \,\ty{llist} \,=\, \const{LNil} \,\mid\, \const{LCons}(\ty{int},\: \ty{llist})
 %\]
@@
 Although it was written several years ago, the account of related work in
-Barrett et al.\ \cite{barrett-et-al-2007} is still largely up to date.
+Barrett et al.\ \cite{barrett-et-al-2007} is still a good reference.
 Since then, datatypes have been added to Z3 in unpublished work by Leonardo de
@@
 values $\{\bot, \top\}$. %, and may contain other interpreted types (e.g., \ty{int}, \ty{real}).
-The only predicate is equality ($\teq$) and is considered a logical symbol.
+The only predicate is equality ($\teq$) and belongs to the logical symbols.
 Other predicates can be represented as functions that return $\ty{bool}$.
-The metavariables $\delta, \tau$ range over types, with $\delta$ reserved
+The metavariables $\delta,\: \tau$ range over types, with $\delta$ reserved
 for datatypes and codatatypes.
@@
 which collapses to $\const{f} : \tau$ if $n = 0$.
-For a term $t$, the notation $t : \tau$ indicates that $t$ has type $\tau$.
+For a term $t$, the notation $t : \tau$ indicates that it has type $\tau$.
 Functions are invoked in the standard way, with $\const{f}(t_1,\ldots,t_n)$
 applying the $n$-ary function symbol
-\const{f} to $n$ well-typed arguments $t_1 : \tau_1$, \ldots, $t_n :
+\const{f} to $n$ well-typed arguments $t_1 :\nobreak \tau_1$, \ldots, $t_n :
 \tau_n$. Nullary function symbols are called constants and can appear without
@@
 \mathrel{\uplus} \Sel$, where $\Ctr$ are the \emph{constructors} and $\Sel$ are the
-\emph{selectors}; there is no need to consider other function symbols, because they can
+\emph{selectors}; there is no need to consider other function symbols because they can
 be abstracted away as variables when combining theories.
@@
 arguments and returns a $\delta$ value. The argument types must be either
-ordinary, among the already know (co)datatypes, or among the (co)datatypes being
-introduced.
-%
-To every argument corresponds a selector. The introduced names for the (co)data\-types, the
+ordinary, among the already known (co)datatypes, or among the (co)datatypes
+being introduced.
+%
+To every argument corresponds a selector. The names for the (co)data\-types, the
 constructors, and the selectors must be distinct and different from already
@@
 (\keyw{co})\keyw{datatype}\;\,
-  \delta_1 & {}=\, \const{C}_{1}\rlap{${}_{1}$}\phantom{{}_{m}}(\bigl[\const{s}_{11}^1{:}\bigr]\; \tau_{11}^1, \,\ldots,\, \bigl[\const{s}_{11}^{n_{11}}{:}\bigr]\; \tau_{11}^{n_{11}}) & \\
+  \delta_1 & {}=\, \smash{\const{C}_{11}(\bigl[\const{s}_{11}^1{:}\bigr]\; \tau_{11}^1, \,\ldots,\, \bigl[\const{s}_{11}^{n_{11}}{:}\bigr]\; \tau_{11}^{n_{11}})} & \\
            & {}\;\phantom{=\,}\;{} \smash{\,\,\,\vdots} \\[-\jot] % {}\;\phantom{=}\llap{\ensuremath{\mid}}\; \cdots \\
-           & {}\;\phantom{=\,}\llap{\ensuremath{\,\mid\,}}\; \const{C}_{1m}(\bigl[\const{s}_{1m}^1{:}\bigr]\; \tau_{1m}^1, \,\ldots,\, \big[\const{s}_{1m}^{n_{1m}}{:}\big]\; \tau_{1m}^{n_{1m}}) \\[-1\jot]
+           & {}\;\phantom{=\,}\llap{\ensuremath{\,\mid\,}}\; \smash{\const{C}_{1m}(\bigl[\const{s}_{1m}^1{:}\bigr]\; \tau_{1m}^1, \,\ldots,\, \big[\const{s}_{1m}^{n_{1m}}{:}\big]\; \tau_{1m}^{n_{1m}})} \\[-1\jot]
    \smash{\vdots\,\,\,} \\[-1\jot]
@@
 $x = \const{C}_{i\negvthinspace j}\bigl(\const{s}_{i\negvthinspace j}^1(x), \ldots, \const{s}_{i\negvthinspace j}^{\,n_{\smash{{i\negvthinspace j}}}}(x)\bigr)$.
-This will simplify the presentation of the decision procedure.
+This will simplify the presentation. % of the decision procedure.
 
@@
 \[\begin{aligned}[t]
-      \keyw{datatype}\;\, \ty{tree}_{\,\tau} & \,=\, \const{Node}(\tau,\, \ty{forest}_{\,\tau}) \\[-\jot]
-      \keyw{and}\;\, \ty{forest}_{\,\tau} & \,=\, \const{FNil} \,\mid\, \const{FCons}(\ty{tree}_{\,\tau},\, \ty{forest}_{\,\tau}) \\%[-.5\jot]
-      \keyw{codatatype}\; \,\ty{llist}_{\,\tau} & \,=\, \const{LNil} \,\mid\, \const{LCons}(\tau,\, \ty{llist}_{\,\tau})
+      \keyw{datatype}\;\, \ty{tree}_{\,\tau} & \,=\, \const{Node}(\tau,\: \ty{forest}_{\,\tau}) \\[-\jot]
+      \keyw{and}\;\, \ty{forest}_{\,\tau} & \,=\, \const{FNil} \,\mid\, \const{FCons}(\ty{tree}_{\,\tau},\:\, \ty{forest}_{\,\tau}) \\%[-.5\jot]
+      \keyw{codatatype}\; \,\ty{llist}_{\,\tau} & \,=\, \const{LNil} \,\mid\, \const{LCons}(\const{lhead}{:}\; \tau,\; \const{ltail}{:}\;\ty{llist}_{\,\tau})
 \end{aligned}
@@
 that datatype. This rules out ill-founded specifications such as
-\[\keyw{datatype}\;\, \ty{fstream} \,=\, \const{FSCons}(\ty{int},\,\ty{fstream})\]
+\[\keyw{datatype}\;\, \ty{fstream} \,=\, \const{FSCons}(\ty{int},\:\, \ty{fstream})\]
 For codatatypes, no such check is necessary because there is always a term,
@@
 %
-%\[\keyw{codatatype}\;\, \ty{stream} \,=\, \const{SCons}(\ty{int},\,\ty{stream})\]
+%\[\keyw{codatatype}\;\, \ty{stream} \,=\, \const{SCons}(\ty{int},\:\ty{stream})\]
 %
@@
 \text{Selection:}\quad
-  & s_{i\negvthinspace j}^k(\const{C}_{i\negvthinspace j}(x_1,\ldots,x_{n_{i\negvthinspace j}})) = x_k
+  & s_{i\negvthinspace j}^{\,k}(\const{C}_{i\negvthinspace j}(x_1,\ldots,x_{n_{i\negvthinspace j}})) = x_k
 \end{aligned}
@@
 exists, whereas distinctness and injectivity ensure that ``no confusion''
-arises.
+arises. The result of selectors applied to the wrong constructor is left
+completely unspecified.
+
+%  * how to deal with "wrong" selectors, e.g.
+%        hd(nil1) = hd(nil2)?
+%      * leave them unspecified; hence hd(nil1) = hd(nil2) in some models,
+%        and not in other models
 
@@
 for a datatype of natural numbers constructed from $\const{Z}$ and $\const{S}$,
-this rules out nonstandard models, which could contain cyclic values (e.g.,
-an $n$ such that $n \teq \const{S}(n)$) or even infinite acyclic values
+this rules out nonstandard models, which could contain cyclic values---e.g.,
+an $n$ such that $n \teq \const{S}(n)$---or even infinite acyclic values
 $\const{S}(\const{S}(\ldots))$.
@@
 this is type correct or $x \teq y$ otherwise.
-
-FIXME: Find out what ensures that codatatypes have infinite values.
-
 This axiom schema captures the idea that two values that yield the same
@@
 selectors and discriminators.
+
+In addition, codatatypes are guaranteed to contain all values corresponding to
+infinite ground constructor terms. This cannot be capture by a first-order
+axiomatization, since there may be uncountably many of them. For example,
+$\ty{stream}_{\,\ty{int}}$ is isomorphic to the uncountable function space
+$\ty{nat} \to \ty{int}$.
 
@@
 
-      * again: when looking at the universal ground (?) theory, enough to
-        consider bisimilarity/bisimulation
-
-specifies the
-
-  * how to deal with "wrong" selectors, e.g.
-        hd(nil1) = hd(nil2)?
-      * leave them unspecified; hence hd(nil1) = hd(nil2) in some models,
-        and not in other models
+%      * again: when looking at the universal ground (?) theory, enough to
+%        consider bisimilarity/bisimulation
+
+% specifies the
 
@@
 
-We present the calculus in three parts.
-In the first part we compute the bidirectional closure of $\Ec$,
-in the second part we make inferences based on cyclicity and bisimilarity,
-and in the third part (when necessary) we branch on constructor types for various terms in $\tEc$.
+We present the calculus in three sections, shown in Figures~\ref{fig:cc-rules},~\ref{fig:ab-rules}, and~\ref{fig:split-rule}.
 Following the conventions from~\cite{}, 
@@
 
-\paragraph{Step 1: Compute Birectional Closure}
 Figure~\ref{fig:cc-rules} gives the basic rules of our calculus.
@@
   \quad
-  %\ttpath{\Ec}{\ec{t}}{\ec{t}} \neq \emptyset
-  \Val \ec{t} = \mu x. t'
-  \quad
-  x \in FV( t' )
+  \ttpath{\Ec}{\ec{t}}{\ec{t}} \neq \emptyset
 }{
@@
   \quad
- %\tpath{\Ec}{\ec{t_1}} = \tpath{\Ec}{\ec{t_2}}\sigma \neq \emptyset
- \Val \ec{t_1} \vsim \Val \ec{t_2}
+ \tpath{\Ec}{\ec{t_1}} = \tpath{\Ec}{\ec{t_2}}\sigma \neq \emptyset
 }{
@@
 
-\paragraph{Step 2: Apply cyclicity and bisimilarity.}
-In this step,
+Before applying the rules in Figure~\ref{fig:ab-rules}, 
 we assume that $\Ec$ is saturated with respect to the rules in Figure~\ref{fig:cc-rules}, that is,
@@
 
-Recall that the $\mu$-notation can be used to represent (possibly cyclic) terms and values,
-for instance, $\mu x. \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
-Formally, we recursively define a $\mu$-term is either of the form
-$\mu x. \const{C}( \vec t )$ for some constructor $\const{C} \in \Ctr$ and $\mu$-terms $\vec t$, 
-or a variable $x$.
-We consider only $\mu$-terms that are \emph{well-typed}, which we define recursively:
-if the type of $\const{C}$ is $\vec \tau \rightarrow \upsilon$, 
-then if $\vec v$ are well-type $\mu$-terms under the assumption that $x$ has type $\upsilon$,
-then $\mu x. \const{C}( \vec v )$ is a well-typed term of type $\upsilon$.
-We say a $\mu$-term is \emph{closed} if it contains no free variables.
-We write $t \vsim s$ if $\mu$-terms $t$ and $s$ are syntactically equivalent modulo renaming of $\mu$-bound variables.
-For convienience, we will use $\mu$-terms to refer to (classes of) values for both co-datatype and datatype terms,
-where in the latter case, a $\mu$-term with a bound variable denotes an invalid term.
-%where the latter case adds the restriction on $\mu$-terms $t$ that no subterms of $t$ contain are bound variables.
-
-At the beginning of this step, 
-we compute a mapping $\Val$ from equivalence classes to $\mu$-terms.
-Informally, the terms mapped to by $\Val$ describe the (class of) values that terms in each equivalence class can take in the models of $\Ec$. 
-For each equivalence class $\ec{x}$ of $\Ec$, we associate a fresh variable $v_{\ec{x}}$ of the same type as $x$ not occurring in $\tEc$,
-and initially say that $\Val$ contains $\ec{x} \mapsto v_{\ec{x}}$ for each equivalence class $\ec{x}$ of $\Ec$ 
-(in other words, there are initially no constraints on the values for any equivalence class).
-Then, $\Val$ is updated by applying the following rule exhaustively:
-
-\(
-\inferrule{
-  v_{\ec{x}} \in FV( \Val )
-  \quad
-  C( t_1, \ldots, t_n ) \in \ec{x}
-  \quad
-  C \in \Ctr
-}{
-  \Val := \Val \{ v_{\ec{x}} \mapsto \mu v_{\ec{x}}. C( v_{\ec{t_1}}, \ldots, v_{\ec{t_n}} ) \}
-}
-\)
-
-Here, we write $FV( \Val )$ to denote the free variables occurring in $\Val$, which
-excludes variables that occur beneath $\mu$-binders.
-It is straightforward to show that the height of terms produced as a result of this rule
-is bounded by the number of equivalence classes of $\Ec$,
-and thus our construction of $\Val$ is terminating.
-
-\begin{example}
-Say that $\Ec$ contains (distinct) equivalence classes $\ec{x}$, $\ec{y}$, $\ec{z}$ and $\ec{w}$ where $\const{C}(z,x) \in \ec{y}$ and $\const{C}(w,y) \in \ec{x}$.
-A possible sequence of applications of the above rule is shown below.
-In the first column, we show the equivalence class that was considered,
-and in the second column, we show the state of $\Val$.
-\[\begin{array}{|l|@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l|}
-\hline
-\text{Eqc} & \multicolumn{3}{c}{\Val} & \ 
-\\
-\hline
-& \{ & \ec{x} \mapsto v_{\ec{x}}, & \ec{z} \mapsto v_{\ec{z}}, & \\
-&    & \ec{y} \mapsto v_{\ec{y}}, & \ec{w} \mapsto v_{\ec{w}} & \} \\
-\ec{x}
-& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ), & \ec{z} \mapsto v_{\ec{z}}, & \\
-&    & \ec{y} \mapsto v_{\ec{y}}, & \ec{w} \mapsto v_{\ec{w}} & \} \\
-\ec{y}
-& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ) ), & \ec{z} \mapsto v_{\ec{z}}, & \\
-&    & \ec{y} \mapsto \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
-\ec{x}
-& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ) ), & \ec{z} \mapsto v_{\ec{z}}, & \\
-&    & \ec{y} \mapsto \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ) ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
-\hline
-\end{array}\]
-Informally, this indicates that the values for $x$ (resp. $y$) in models of $\Ec$
-are of the form $\const{C}( w, \const{C}( z, \ldots ))$ (resp. $\const{C}( z, \const{C}( w, \ldots ))$) for some value of $w$ and $z$.
-$\Box$
-\end{example}
-
-Given $\Val$, the rules for cyclicity and bisimilarity (shown in Figure~\ref{fig:ab-rules}) can be stated as follows.
-For cyclicity, if there is an equivalence class $\ec{t}$ of datatype type whose class of values $\Val \ec{t} = \mu x. t'$ are cyclic,
-as indicated by the assumption $x \in FV( t' )$,
-then we can conclude that $\Ec$ is unsatisfiable.
-For bisimilarity, if there are two equivalence classes $\ec{t_1}$ and $\ec{t_2}$ 
-of co-datatype type whose classes of values $\Val \ec{t_1}$ and $\Val \ec{t_2}$ are equivalent modulo renaming of $\mu$-bound variables,
-then we can conclude $t_1 \teq t_2$.
-We demonstrate the bisimilarity rule concretely with the following example.
-
-\begin{example}
-Let $\const{C}$ be a constructor of type $\tau_1 \times \tau_2 \rightarrow \tau_2$,
-and let $\Ec$ be the set $\{ x \teq \const{C}(y,\const{C}(y,x)) \}$.
-After applying the rules in Figure~\ref{fig:cc-rules} to completion, the equivalence classes of $\Ec$ are
-$\{ x, \const{C}(y,\const{C}(y,x)) \}$, $\{ \const{C}(y,x) \}$, and $\{ y \}$.
-We then construct $\Val$ using the aforementioned steps, after which we obtain:
-\[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}r@{\hspace{.5em}}l@{\hspace{.5em}}r@{\hspace{.5em}}l}
-\Val \ec{x} & = & 
-\mu v_{\ec{x}}. & \const{C}( v_{\ec{y}}, & \mu v_{\ec{\const{C}(y,x)}}. & \const{C}( v_{\ec{y}}, v_{\ec{x}} ) ) \\
-\Val \ec{\const{C}(y,x)} & = & 
-\mu v_{\ec{\const{C}(y,x)}}. & \const{C}( v_{\ec{y}}, & \mu v_{\ec{x}}. & \const{C}( v_{\ec{y}}, v_{\ec{\const{C}(y,x)}} ) )
-\end{array}\]
-Notice that $\Val \ec{x}$ and $\Val \ec{\const{C}(y,x)}$ are syntactically equivalent for the renaming
-$\{ v_{\ec{x}} \mapsto v_{\ec{\const{C}(y,x)}}, v_{\ec{\const{C}(y,x)}} \mapsto v_{\ec{x}} \}$,
-and thus our calculus may infer $x \teq \const{C}(y,x)$.
-Intuitively, this equality holds, since the values of $x$ and $\const{C}( y, x )$ are observationally equivalent for any value of $y$.
-\qed
-\end{example}
-
-\begin{comment}
 To explain the rules in Figure~\ref{fig:ab-rules}, we introduce the following notions.
@@
 in which case $t_1$ must be equal to $t_2$.
-We demonstrate an application of this rule concretely with the following example.
-\end{comment}
+We demonstrate this rule concretely with the following example.
+\rem{ensure that \const{LCons} understood}
+
+\begin{example}
+Let $\Ec$ be the set $\{ x \teq \const{LCons}(y,\const{LCons}(y,x)) \}$.
+After applying the rules in Figure~\ref{fig:cc-rules} to completion, the equivalence classes of $\Ec$ are
+$\{ x, \const{LCons}(y,\const{LCons}(y,x)) \}$, $\{ \const{LCons}(y,x) \}$, and $\{ y \}$.
+The set $\tpath{\Ec}{\ec{x}}$ is:
+\[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
+\{ & \ec{x} \rightarrow^{\const{LCons}}_1 & \ec{y}, & & \\ 
+   & \ec{x} \rightarrow^{\const{LCons}}_2 & \ec{\const{LCons}(y,x)} \rightarrow^{\const{LCons}}_1 & \ec{y},& \\
+   & \ec{x} \rightarrow^{\const{LCons}}_2 & \ec{\const{LCons}(y,x)} \rightarrow^{\const{LCons}}_2 & \ec{x} & \}.
+\end{array}\]
+The set $\tpath{\Ec}{\ec{\const{LCons}(y,x)}}$ is:
+\[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
+\{ & \ec{\const{LCons}(y,x)} \rightarrow^{\const{LCons}}_1 & \ec{y}, & & \\ 
+   & \ec{\const{LCons}(y,x)} \rightarrow^{\const{LCons}}_2 & \ec{x} \rightarrow^{\const{LCons}}_1 & \ec{y},& \\
+   & \ec{\const{LCons}(y,x)} \rightarrow^{\const{LCons}}_2 & \ec{x} \rightarrow^{\const{LCons}}_2 & \ec{\const{LCons}(y,x)} & \}.
+\end{array}\]
+Thus, we may infer $x \teq \const{LCons}(y,x)$ by the rule \rn{Bisimilar}
+for the mapping $\sigma :=$ $\{ \ec{x} \mapsto \ec{\const{LCons}(y,x)}, \ec{\const{LCons}(y,x)} \mapsto \ec{x} \}$.
+\qed
+\end{example}
 
@@
 
-\paragraph{Step 3: Branching}
-If neither a rule from Figure~\ref{fig:cc-rules} or Figure~\ref{fig:ab-rules} applies to $\Ec$, 
+\rem{Require : all co-datatypes either have infinite cardinality, or have cardinality 1.}
+
+Finally, if neither a rule from Figure~\ref{fig:cc-rules} or Figure~\ref{fig:ab-rules} applies to $\Ec$, 
 the rules from Figure~\ref{fig:split-rule} are applied.
@@
 
+%We first show that for any saturated node $\Fc$, 
+%additional equalities $\Gc$ can be added to $\Fc$ such that each codatatype/datatype equivalence class of $\Fc \cup \Gc$ has a constructor term.
+%For this purpose, we extend the calculus in this section.
+%For a subset of equivalence classes $S$ of $\Ec$, write $\dpath{\Ec}{C,n}{S}$ to denote the set
+%such that $e' \in \dpath{\Ec}{C,n}{S}$ if and only if there exists an $e \in S$ such that $e \rightarrow^{C}_n e'$ is induced by $\Ec$.
+
+\begin{comment}
+\begin{figure}[t]
+\centering
+\begin{tabular}{c}
+\rn{O-Finish}
+\(
+\inferrule{
+  \Ac := x
+  \quad
+  x : \tau
+  \quad
+  \tau \in \Nondata
+  \quad
+  c \not\in \tEc
+}{
+  \Ac := \emptyset \quad \Ec := \Ec \cup \{ x \teq c \}
+}
+\)
+\\[3.7ex]
+\rn{C-Finish}
+\(
+\inferrule{
+  \Ac := x
+  \quad
+  x : \tau
+  \quad
+  \tau \in \Codata
+  \quad
+  C_i \in \Ctr^\tau
+  \quad 
+  C_i \not\in \{ C_j \mid d \in \Dc, C_j( \vec t ) \in d \}
+}{
+  \Ac := \emptyset \quad \Ec := \Ec \cup \{ x \teq C_i( \vec y ) \} \quad \Rc := \Rc \cup \vec y
+}
+\)
+\\[3.7ex]
+\rn{C-Rec}
+\(
+\inferrule{
+  \Ac := x
+  \quad
+  x : \tau
+  \quad
+  \tau \in \Codata
+  \quad
+  C_i \in \Ctr^\tau
+  \quad
+  C_i \text{ infinite arg } j
+}{
+  \Ac := y_j \quad 
+  \Ec := \Ec \cup \{ x \teq C_i( \vec y ) \} \quad 
+  \Rc := \Rc \cup ( \vec y \setminus y_j ) \quad 
+  \Dc := \dpath{\Ec}{C_i,j}{\Dc}
+}
+\)
+\end{tabular}
+\caption{Rules for completion.
+}
+\label{fig:completion-rules}
+\end{figure}
+\end{comment}
+
 %and second show how to construct a valuation for the variables of $\Fc \cup \Gc$.
@@
 
-%Given a set of disjoint paths $P$ where each $p \in P$
-%we write $\interp{P}$ to denote the value
-
-\rem{Revise : now $\Val$ is constructed by calculus.}
-
-First, we require additional terminology concerning $\mu$-terms.
-The \emph{interpretation of subvalue $u$ in $v$} is the value $v\interp{u}^\emptyset$ returned by the following recursive procedure,
-where $x \mapsto \mu x. C( \vec t ) \in v$ indicates that $\mu x. C( \vec t )$ is the subvalue of $v$ that binds this occurrence of variable $x$.
-\[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
-v\interp{x}^M & = & v\interp{\mu x. C( \vec t )}^M & x \not\in M, x \mapsto \mu x. C( \vec t ) \in v  \\
-v\interp{x}^M & = & x & \text{otherwise} \\
-v\interp{\mu x. C( \vec t )}^M & = & \mu x. C( v\interp{ \vec t }^{M \cup \{ x \}} ) & x \not\in M \\
-v\interp{\mu x. C( \vec t )}^M & = & x & \text{otherwise}
-\end{array}\]
-Given a $\mu$-value $v$ of the form $\mu x. C( \vec w )$, 
-a subvalue $u$ of $v$ is a \emph{bisimilar subvalue} if and only if $v\interp{u}^\emptyset \sim v$.
-We say $v$ is \emph{normal} if and only if all of its bisimilar subvalues are $x$,
-and all of its strict subvalues that do not contain $v$ are also normal.
-For instance, the value $v$ of the form $\mu x. \const{S}( \mu y. \const{S}( x ) )$ is not normal, 
-since 
-$v\interp{\mu y. \const{S}( x )}^\emptyset =$ 
-$\mu y. \const{S}( v\interp{x}^{\{y\}} ) =$ 
-$\mu y. \const{S}( v\interp{\mu x. \const{S}( \mu y. \const{S}( x ) ) }^{\{y\}} ) =$ 
-$\mu y. \const{S}( \mu x. \const{S}( v\interp{\mu y. \const{S}( x )}^{\{x,y\}} ) ) =$ 
-$\mu y. \const{S}( \mu x. \const{S}( y ) )$,
-which is equivalent to $v$ modulo the renaming $\{ x \mapsto y, y \mapsto x \}$.
-For any $\mu$-value $v$ of the form $\mu x. C( \vec w )$, we may recursively construct a normal value $\nf{v}$
-(which we will call the \emph{normal form for $v$})
-that is observationally equivalent to $v$.
-We construct $\nf{v}$ by replacing all of the bisimilar subvalues of $v$ with $x$,
-and by replacing all of the non-normal closed subvalues of $v$ with their corresponding normal forms.
-For instance, $\nf{(\mu x. \const{S}( \mu y. \const{S}( x ) ))} = \mu x. \const{S}( x )$.
+Given a set of disjoint paths $P$ where each $p \in P$
+we write $\interp{P}$ to denote the value
 
@@
 interprets wrongly-applied selectors based on equality information in $\Fc$,
-interprets the variables and $x$ of $\Fc$ based on the valuation $\Val$ which we constructed in step two of our calculus.
-
-Recall that $\Val$ was constructed as a mapping from the equivalence classes of $\Fc$ to $\mu$-terms.
-We must first \emph{complete} $\Val$ so that it maps equivalence classes only to closed $\mu$-terms, that is, terms containing no free variables.
-This is done by exhaustively applying of the following rule:
-
-\(
-\inferrule{
+interprets the variables and $x$ of $\Fc$ based on a mapping $\Val$ from the equivalence classes of $\Fc$ to values.
+Initially, $\Val$ contains $\ec{x} \mapsto v_{\ec{x}}$ for each equivalence class $\ec{x}$ of $\Fc$,
+where $v_{\ec{x}}$ is a fresh variable of the same type as $x$ not occurring in $\Fc$.
+Then, $\Val$ is updated according to the following rules:
+
+\begin{comment}
+\rn{Ord}
+\(
+\inferrule{
+  \ec{x} \not\in \vdom(\Val)
+  \quad
+  x : \tau
+  \quad
+  \tau \in \Nondata
+  \quad
+  c \not\in \vrange(\Val)
+}{
+  \Val := \Val \cup \ec{x} \mapsto c
+}
+\)
+
+\rn{D-Ind}
+\(
+\inferrule{
+  \ec{x} \not\in \vdom(\Val)
+  \quad
+  x : \tau
+  \quad
+  \tau \in \Data
+  \quad
+  C( \vec t ) \in \ec{x}
+  \quad
+  \ec{\vec t} \in \vdom(\Val)
+}{
+  \Val := \Val \cup \ec{x} \mapsto C( \Val \ec{\vec t}  )
+}
+\)
+
+\rn{D-Base}
+\(
+\inferrule{
+  \ec{x} \not\in \vdom(\Val)
+  \quad
+  x : \tau
+  \quad
+  \tau \in \Data
+  \quad
+  \ec{x} \text{ ctr-free}
+  \quad
+  c \not\in \vrange(\Val)
+}{
+  \Val := \Val \cup \ec{x} \mapsto c
+}
+\)
+\end{comment}
+
+\rn{Ind}
+\(
+\inferrule{
+  %x : \tau
+  %\quad
+  %\tau \in \Codata
   v_{\ec{x}} \in FV( \Val )
   \quad
+  C( t_1, \ldots, t_n ) \in \ec{x}
+  \quad
+  C \in \Ctr
+}{
+  \Val := \Val \{ v_{\ec{x}} \mapsto \mu v_{\ec{x}}. C( v_{\ec{t_1}}, \ldots, v_{\ec{t_n}} ) \}
+}
+\)
+
+\rn{Base}
+\(
+\inferrule{
+  %x : \tau
+  %\quad
+  %\tau \in \Codata
+  %\quad
+  v_{\ec{x}} \in FV( \Val )
+  %\quad
+  %\ec{x} \text{ ctr-free}
+  \quad
   c \not\sim c' \text{ for all } c' \in \vrange(\Val)
@@
 \)
-Since this rule removes one variable from the set $FV( \Val )$ and does not add any additional variables to $FV( \Val )$, 
-it can only be applied a finite number of times.
-
-We first argue that $\Val$ contains only to ground values (which include closed $\mu$-terms), in other words, $FV( \Val ) = \emptyset$.
-First, since \rn{Split} does not apply to $\Fc$,
-all equivalence classes of finite type contain constructor terms.
-Thus, all variables in $FV( \Val )$ have infinite type.
-%\rem{Require : all co-datatypes either have infinite cardinality, or have cardinality 1.}
-
-%It is clear that after applying these rules to completion, $\Val$ is a mapping from equivalence classes to ground values.
+
+\rem{define $\sim$}
+We assume that first \rn{Ind} is applied exhaustively, after which \rn{Base} is applied exhaustively.
+%the rule \rn{D-Ind} is applied to completion before applying \rn{D-Base}.
+In the rule \rn{Ind}, we write $FV( \Val )$ to denote the free variables occurring in $\Val$.
+Notice that this excludes variables beneath $\mu$-binders.
+
+As an example,
+say that $\Fc$ contains equivalence classes $\ec{x}$ and $\ec{y}$ where $\const{C}(z,x) \in \ec{y}$ and $\const{C}(w,y) \in \ec{x}$, where $z$ and $w$ are variables of integer type.
+A possible application of the above rules, showing the state of $\Val$ after each application, is as follows:
+\[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
+\hline
+\text{Rule} & \multicolumn{4}{c}{\Val} \ 
+\\
+\hline
+& \{ & \ec{x} \mapsto v_{\ec{x}}, & \ec{z} \mapsto v_{\ec{z}}, \\
+&    & \ec{y} \mapsto v_{\ec{y}}, & \ec{w} \mapsto v_{\ec{w}} & \} \\
+\rn{Ind} \ec{x}
+& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ), & \ec{z} \mapsto v_{\ec{z}}, \\
+&    & \ec{y} \mapsto v_{\ec{y}}, & \ec{w} \mapsto v_{\ec{w}} & \} \\
+\rn{Ind} \ec{y}
+& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ) ), & \ec{z} \mapsto v_{\ec{z}}, \\
+&    & \ec{y} \mapsto \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
+\rn{Ind} \ec{x}
+& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ) ), & \ec{z} \mapsto v_{\ec{z}}, \\
+&    & \ec{y} \mapsto \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ) ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
+\rn{Base} \ec{z}
+& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu v_{\ec{y}}. \const{C}( \const{0}, v_{\ec{x}} ) ), & \ec{z} \mapsto \const{0}, \\
+&    & \ec{y} \mapsto \mu v_{\ec{y}}. \const{C}( \const{0}, \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ) ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
+\rn{Base} \ec{w}
+& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( \const{1}, \mu v_{\ec{y}}. \const{C}( \const{0}, v_{\ec{x}} ) ), & \ec{z} \mapsto \const{0}, \\
+&    & \ec{y} \mapsto \mu v_{\ec{y}}. \const{C}( \const{0}, \mu v_{\ec{x}}. \const{C}( \const{1}, v_{\ec{y}} ) ), & \ec{w} \mapsto \const{1} & \} \\
+\end{array}\]
+
+It is clear that after applying these rules to completion, $\Val$ is a mapping from equivalence classes to ground values.
 

### 2014-12-11 tex/codatadp.tex
More clarification on model.

--- 2014-12-11 tex/codatadp.tex
@@
 in which case we will commonly say $t$ and $s$ are \emph{bisimilar}.
+%, or informally are \emph{observationally equivalent}.
 %For convienience, we will use $\mu$-terms to refer to (classes of) values for both co-datatype and datatype terms,
@@
 We say $v$ is \emph{normal} if and only if all of its bisimilar subterms are $x$,
-and all of its strict subterms that do not contain $x$ are normal.
+and all of its strict subterms are also normal.
 For instance, the $\mu$-term $v$ of the form $\mu x. \const{S}( \mu y. \const{S}( x ) )$ is not normal, 
@@
 which is syntactically equivalent to $v$ for the renaming $\{ x \mapsto y, y \mapsto x \}$.
-
-%\begin{lemma}
-%\end{lemma}
-
-%For any $\mu$-value $v$ of the form $\mu x. C( \vec w )$, we may recursively construct a normal value $\nf{v}$
-%(which we will call the \emph{normal form for $v$})
-%that is observationally equivalent to $v$.
-%We construct $\nf{v}$ by replacing all of the bisimilar subvalues of $v$ with $x$,
-%and by replacing all of the non-normal closed subvalues of $v$ with their corresponding normal forms.
-%For instance, $\nf{(\mu x. \const{S}( \mu y. \const{S}( x ) ))} = \mu x. \const{S}( x )$.
+For any $\mu$-value $v$ of the form $\mu x. C( \vec w )$, we may recursively construct a normal value $\nf{v}$
+(which we will call the \emph{normal form} of $v$)
+that is a renaming of $v$ by replacing all of the bisimilar subvalues of $v$ with $x$,
+and by replacing all of the non-normal subterms of $v$ with their corresponding normal forms.
+For instance, $\nf{(\mu x. \const{S}( \mu y. \const{S}( x ) ))} = \mu x. \const{S}( x )$.
 
@@
 as a maximal set of closed normal $\mu$-terms of that type that are 
-(1) unique up to renaming of $\mu$-bound variables,
-(2) acyclic when $\tau \not\in \Codata$.
-We assume $\M$ interprets constructors and correctly-applied selectors in the obvious way
-\footnote{
-To interpret a correctly-applied selector, if $v = \mu x. C_i( \vec t )$, then $\M( s^j_i( v ) )$ is the value in $\M( \tau )$ that is a renaming of $v\interp{t_j}$,
-where $t_j$ has type $\tau$.
-},
-interprets wrongly-applied selectors based on equality information in $\Fc$,
-interprets the variables of $\Fc$ based on the valuation $\Val$ which was constructed in step two of our calculus.
+unique up to renaming of $\mu$-bound variables,
+and acyclic when $\tau \not\in \Codata$.
+The interpretation of constructor terms
+$\M( C( \vec t ) )$ for $C \in \Ctr$ is the value in $\M( \tau )$ that is a renaming of $\nf{(\mu x. C( \M( \vec t ) ))}$ where $x \not\in FV( \vec t )$.
+The interpretation of (correctly-applied) selector terms
+$\M( s^j_i( t ) )$ for $s^j_i \in \Sel$ is the value in $\M( \tau )$ that is a renaming of $\M( t )\interp{u_j}$
+where $\M( t )$ is of the form $\mu x. C_i( \vec u )$ for some $\vec u$.
+For instance, if $\M( t ) = \mu x. \const{C}_1( x )$ (call this term $v$), 
+then $\M( s_1^1( t ) )$ is a renaming of
+$v\interp{x} =$ 
+$v\interp{\mu x. \const{C}_1( x )}^\emptyset =$
+$\mu x. \const{C}_1( v\interp{x}^{\{x\}} ) =$
+$\mu x. \const{C}_1( x )$.
+%To interpret a correctly-applied selector in $\M$, if $v = \mu x. C_i( \vec t )$, then $\M( s^j_i( v ) )$ is the value in $\M( \tau )$ that is a renaming of $v\interp{t_j}$,
+%where $t_j$ has type $\tau$.
+%We assume $\M$ interprets constructors and correctly-applied selectors in the obvious way
+%\footnote{
+%The interpretation of $\M( C( \vec t ) )$ is the value in $\M( \tau )$ that is a renaming of $\nf{\mu x. C( \M( \vec t ) ) }$.
+%To interpret a correctly-applied selector, if $v = \mu x. C_i( \vec t )$, then $\M( s^j_i( v ) )$ is the value in $\M( \tau )$ that is a renaming of $v\interp{t_j}$,
+%where $t_j$ has type $\tau$.
+%},
+Additionally, $\M$ will interpret wrongly-applied selectors based on equality information in $\Fc$,
+and will interpret the variables of $\Fc$ based on the valuation $\Val$ which was constructed in step two of our calculus.
 
@@
 For any infinite type $\tau$,
-there exists a closed normal $\mu$-term $c_1$ that is not bisimilar to any $c_2 \in \mathcal{T}(\Val^\ast)$, since $\mathcal{T}(\Val^\ast)$ is a finite set. \rem{elaborate?}
+there exists a closed normal $\mu$-term $c_1$ that is not bisimilar to any $c_2 \in \mathcal{T}(\Val^\ast)$, 
+since $\mathcal{T}(\Val^\ast)$ is a finite set (noting that a closed $\mu$-term is only bisimilar to other closed $\mu$-terms).
 Thus, $FV( \Val^\ast )$ contains no variables of infinite type,
@@
 then the rule \rn{Bisimilar} would have applied to $x$ and $y$. \rem{elaborate?}
-For the inductive case, say that we apply \rn{Complete} to $\Val$ for the substitution $\sigma := \{ v_{\ec{x}} \mapsto c_1 \}$.
+For the inductive case, we assume $\Val$ satisfies ~\ref{eqn:m-normal} and~\ref{eqn:m-bij},
+and show that $\Val \sigma$ does as well where $\sigma := \{ v_{\ec{x}} \mapsto c_1 \}$ is a substitution applied by \rn{Complete}.
+By contradiction, assume that $\Val \sigma \ec{y}$ is a non-normal $\mu$-term for some $\ec{y}$.
+%Since $c_1$ is normal, we have $\ec{x} \neq \ec{y}$, 
+%and thus $\Val \ec{y}$ must have been of the form $\mu v_{\ec{y}}. t$ where $\ec{x} \in FV( t )$.
+Let $u$ be the (strict) subterm of $\Val \ec{y}$ such that $(\Val \ec{y} \sigma ) \interp{u \sigma } \sim (\Val \ec{y} \sigma)$. 
+Since $\Val$ maps only to normal $\mu$-terms, we have $x \in FV( u )$.
 \rem{TODO}
@@
 
-If $\tau \in \Data$,
-we first show that $\M( t ) = \Val^\ast \ec{x}$ for all non-selector terms $t$ that occur in datatype equivalence classes $\ec{x}$.
+If $\tau \in \Data \cup \Codata$,
+we first show that $\M( t ) = \Val^\ast \ec{x}$ for all non-selector terms $t$ that occur in equivalence classes $\ec{x}$.
 We show this by induction on the size of $\Val^\ast \ec{x}$.
@@
 Thus, for all datatype equivalence classes $\ec{x}$, we have $\M( t ) = \Val^\ast \ec{x}$ for each term $t \in \ec{x}$, which implies $\M \models t_1 \teq t_2$.
-
-If $\tau \in \Codata$, \rem{TODO}
 

### 2014-12-11 tex/codatadp.tex
special cases

--- 2014-12-11 tex/codatadp.tex
@@
 Other predicates can be represented as functions that return $\ty{bool}$.
-The metavariables $\delta,\: \tau$ range over types, with $\delta$ reserved
+The metavariables $\delta,\:\varepsilon,\:\tau$ range over types, with $\delta$ reserved
 for datatypes and codatatypes.
@@
 
+\paragraph{Specifications.}
 A (co)datatype specification consists of $l$~mutually recursive types which are
@@
 
+\paragraph{Properties.}
 Datatypes and codatatypes share basic properties pertaining to constructors and
@@
 
+\paragraph{Semantics.}
 Datatypes can be defined semantically as the initial model of an equational
@@
 
-A related semantic view of datatypes is as initial algebra. Codatatypes are then
-defined dually as a final coalgebra. The datatypes are generated by their
+A related semantic view of datatypes is as an initial algebra. Codatatypes are
+then defined dually as a final coalgebra. The datatypes are generated by their
 constructors, whereas the codatatypes are viewed through their selectors
 \cite{xxx}. By uniformly focusing on the constructors, the axiomatic approach
-emphasizes the commonality between datatypes and codatatypes, while sacrificing
-a theoretically fruitful notion of duality.
-
-    semantically, $k$ types are mutually (co)recursive if their dependency graph
-    (x -> y if x is the type of an argument to one of y's constructors) is
-    strongly connected
-
-    (co)datatypes can be declared together as mutually (co)recursive even if
-    they are not actually (co)recursive
-
-    the semantic notion is more fruitful and the one we will focus on in the
-    sequel
-
-    Also, nothing forbids specifications with no (co)recursion, e.g.
-
-      datatype enat = Nat(nat) | Infty
-      codatatype complex = Complex(re: real, im: real)
-
-    In this case, it makes no difference whether the type is introduced as a
-    datatype or as a codatatype. Without loss of generality, we consider that
-    these types are datatypes instead of codatatypes.
+emphasizes the commonality between datatypes and codatatypes, while
+sacrificing a theoretically fruitful notion of duality.
+
+\paragraph{Special Cases.}
+A type $\delta$ depends on another type $\varepsilon$ if $\varepsilon$ is the
+type of an argument to one of $\delta$'s constructors. Semantically, a set of
+types is mutually (co)recursive if and only if the associated dependency graph
+is strongly connected. Types can be declared together as mutually
+(co)recursive even if they are not actually (co)recursive; however, the
+semantic notion is more fruitful and is the one we focus on.
+
+Furthermore, nothing forbids specifications with no (co)recursion, such as
+\[\begin{aligned}[t]
+      \keyw{datatype}\;\, \ty{enat} & \,=\, \const{Nat}(\ty{nat}) \,\mid\, \const{Infty} \\[-.5\jot]
+      \keyw{codatatype}\; \,\ty{complex} & \,=\, \const{Complex}(\const{re}{:}\; \ty{real},\; \const{im}{:}\;\ty{real})
+\end{aligned}
+\]
+Semantically, it makes no difference whether the types are introduced as
+datatypes or as codatatypes.
+%Without loss of generality, we consider that these types are datatypes
+%instead of codatatypes.
 

### 2014-12-11 tex/codatadp.tex
more sect 2

@@
 values internally. For example,
-the term $\const{SCons}(1,\: \mu s.\; \const{SCons}(0,\: \const{SCons}(9,\: s)))$
+the term $\const{SCons}(1,\: \mu\vthinspace s.\; \const{SCons}(0,\: \const{SCons}(9,\: s)))$
 denotes the lasso-shaped sequence $1, 0, 9, 0, 9, 0, 9, \ldots$\afterLdots.
@@
 (co)recursive even if they are not actually (co)recursive; however, the
-semantic notion is more fruitful and is the one we focus on.
+semantic notion is more fruitful and is the one that interests us.
 
@@
 
-Degenerate Codatatypes
-
-    Consider
-
-      codatatype a = A(a)
-
-    This is a finite codatatype, with a single value, which is infinite:
-    $\mu a. A(a)$
-
-    We need to detect those finite codatatypes. They are easy to
-    characterize semantically.
-
-    Lemma D.1. The only possible finite cardinality for the domain interpreting
-    a (corecursive) codatatype is 1.
-
-    Proof.
-      Assume you have at least 2 different values of a codatatype.
-      They differ at one point.
-      * 
-    Qed
+Finally, some codatatypes are so degenerate that they have infinite values
+even though they are finite. The simplest example is
+\[\keyw{datatype}\;\, \ty{a} \,=\, \const{A}(\ty{a})\]
+This is a finite codatatype, of cardinality one, whose single value is
+infinite: $\mu\vthinspace a.\; A(a)$. Other specimens are
+\[\begin{aligned}[t]
+      \keyw{codatatype}\;\, \ty{b} & \,=\, \const{B}(\ty{b},\: \ty{c},\: \ty{b},\: \ty{unit}) \\[-\jot]
+             \keyw{and}\;\, \ty{c} & \,=\, \const{C}(\ty{unit},\: \ty{c},\: \ty{b},\: \ty{c})
+\end{aligned}
+\]
+where \ty{unit} is a datatype with the single constructor $\const{Unity} :
+\ty{unit}$. We call such types \emph{corecursive singletons}.
+For the decision procedure, it will be important to detect those finite
+codatatypes.
+
+% In other words, all finite corecursive codatatypes are corecursive singletons.
+
+\begin{lemma}
+If the domain interpreting a corecursive codatatype is finite, it must be $1$.
+\end{lemma}
+
+\begin{proof}
+TODO.
+\qed
+\end{proof}
 
 An immediate consequence by the distinctness axiom is that these types have
-    necessarly one constructor, and that all arguments are of types of
-    cardinality 1.
-
-    Easy case: $k$ mutually recursive codatatypes, each equipped with one
-    constructor, all of which take each other in arguments in various
-    combinations, with potential other types that are known to have cardinality
-    1. e.g.
-
-      codatatype b = B b c b
-             and c = C c b c
-
-    Medium case: like above, but some of the arguments are other types not
-    participating in the corecursion that are known to have cardinality 1, e.g.
-
-      codatatype d = D unit d d unit
-
-    where "unit" is a datatype with the single constructor "Unity : unit".
-
-    Though case: like above, but with any other types where it's not clear
-    whether it could have cardinality one or not.
-
-      type e
-      codatatype f = F e f f unit
+necessarly one constructor and that all arguments have types of cardinality 1.
+%
+Sometimes, a type might be a corecursive singleton only in some models. For
+example, if \ty{unit} is an uninterpreted type, \ty{b} and \ty{c} will be
+singleton types precisely in the models that interpret \ty{unit} as a
+singleton.
 
@@
 }{
-  \Val := \Val \{ v_{\ec{x}} \mapsto \mu v_{\ec{x}}. C( v_{\ec{t_1}}, \ldots, v_{\ec{t_n}} ) \}
+  \Val := \Val \{ v_{\ec{x}} \mapsto \mu\vthinspace v_{\ec{x}}. C( v_{\ec{t_1}}, \ldots, v_{\ec{t_n}} ) \}
 }
@@
 \rn{Ind} \ec{x}
-& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ), & \ec{z} \mapsto v_{\ec{z}}, \\
+& \{ & \ec{x} \mapsto \mu\vthinspace v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ), & \ec{z} \mapsto v_{\ec{z}}, \\
 &    & \ec{y} \mapsto v_{\ec{y}}, & \ec{w} \mapsto v_{\ec{w}} & \} \\
 \rn{Ind} \ec{y}
-& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ) ), & \ec{z} \mapsto v_{\ec{z}}, \\
-&    & \ec{y} \mapsto \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
+& \{ & \ec{x} \mapsto \mu\vthinspace v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu\vthinspace v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ) ), & \ec{z} \mapsto v_{\ec{z}}, \\
+&    & \ec{y} \mapsto \mu\vthinspace v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
 \rn{Ind} \ec{x}
-& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ) ), & \ec{z} \mapsto v_{\ec{z}}, \\
-&    & \ec{y} \mapsto \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ) ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
+& \{ & \ec{x} \mapsto \mu\vthinspace v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu\vthinspace v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ) ), & \ec{z} \mapsto v_{\ec{z}}, \\
+&    & \ec{y} \mapsto \mu\vthinspace v_{\ec{y}}. \const{C}( v_{\ec{z}}, \mu\vthinspace v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ) ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
 \rn{Base} \ec{z}
-& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu v_{\ec{y}}. \const{C}( \const{0}, v_{\ec{x}} ) ), & \ec{z} \mapsto \const{0}, \\
-&    & \ec{y} \mapsto \mu v_{\ec{y}}. \const{C}( \const{0}, \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ) ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
+& \{ & \ec{x} \mapsto \mu\vthinspace v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu\vthinspace v_{\ec{y}}. \const{C}( \const{0}, v_{\ec{x}} ) ), & \ec{z} \mapsto \const{0}, \\
+&    & \ec{y} \mapsto \mu\vthinspace v_{\ec{y}}. \const{C}( \const{0}, \mu\vthinspace v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ) ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
 \rn{Base} \ec{w}
-& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( \const{1}, \mu v_{\ec{y}}. \const{C}( \const{0}, v_{\ec{x}} ) ), & \ec{z} \mapsto \const{0}, \\
-&    & \ec{y} \mapsto \mu v_{\ec{y}}. \const{C}( \const{0}, \mu v_{\ec{x}}. \const{C}( \const{1}, v_{\ec{y}} ) ), & \ec{w} \mapsto \const{1} & \} \\
+& \{ & \ec{x} \mapsto \mu\vthinspace v_{\ec{x}}. \const{C}( \const{1}, \mu\vthinspace v_{\ec{y}}. \const{C}( \const{0}, v_{\ec{x}} ) ), & \ec{z} \mapsto \const{0}, \\
+&    & \ec{y} \mapsto \mu\vthinspace v_{\ec{y}}. \const{C}( \const{0}, \mu\vthinspace v_{\ec{x}}. \const{C}( \const{1}, v_{\ec{y}} ) ), & \ec{w} \mapsto \const{1} & \} \\
 \end{array}\]

### 2014-12-11 tex/codatadp.tex
more

@@
 
-\section{Theory of (Co)datatypes}
-\label{sec:theory-of-co-datatypes}
+\section{The Theory of (Co)datatypes}
+\label{sec:the-theory-of-co-datatypes}
 
@@
 For the decision procedure, it will be important to detect those finite
-codatatypes.
+codatatypes. Sometimes, a type might be a corecursive singleton only in some
+models. For example, if \ty{unit} is an uninterpreted type, \ty{b} and \ty{c}
+will be singleton types precisely in the models that interpret \ty{unit} as a
+singleton.
 
@@
 \begin{lemma}
-If the domain interpreting a corecursive codatatype is finite, it must be $1$.
+If the domain interpreting a corecursive codatatype is finite, it must be a
+singleton. Furthermore, such a type has necessarly a single constructor and
+all its arguments have types that are interpreted by singletons.
 \end{lemma}
@@
 \begin{proof}
-TODO.
+  * if two (either directly or indirectly) corecursive constructors
+    $\const{C}$ and $\const{D}$, it is
+    possible to encode infinitely many alternation patterns,
+    e.g. $\const{C}(\const{D}(\const{C}(\const{C}(\ldots))))$;
+    hence no finite models
+
+  * if one corecursive and one noncorecursive, then possibility to create
+    objects of arbitrary depths, e.g. $\const{C}(\ldots(\const{C}(\const{D}))\ldots)$,
+    and hence no finite models
+
+  * leaves us with the case of a single corecursive constructor
+
+  * if any of the noncorecursive arguments has cardinality greater than one,
+    then it is possible to encode alternation patterns using it,
+    e.g. $\const{C}(0,\: \const{C}(1,\: \const{C}(0,\: \const{C}(0,\: \ldots))))$.
+
+  * otherwise, by the coinduction principle, there is only one value that can
+    exist
+
+%  * assume for simplicity no indirect recursion, but this does not radically
+%    change the argument
+%  * assume there are at least two values built with C.
+%    they must be different at some point, e.g.
+%    C(0, C(1, C(0, ...))) and C(0, C(1, C(1, ...)))
+%    can use that to create infinitely alternation patterns
+%  * leaves us with the case of a single constructor
+% * since we are looking at a specific model, we can assume all ordinary
+%   types are finite datatypes with nullary constructors corresponding to their
+%   elements
+% * if two
+% * either there is only one ``path'' from ctr to itself
+%   or at least two;
 \qed
 \end{proof}
-
-An immediate consequence by the distinctness axiom is that these types have
-necessarly one constructor and that all arguments have types of cardinality 1.
-%
-Sometimes, a type might be a corecursive singleton only in some models. For
-example, if \ty{unit} is an uninterpreted type, \ty{b} and \ty{c} will be
-singleton types precisely in the models that interpret \ty{unit} as a
-singleton.
 
@@
 
-We introduced a decision procedure for datatypes and codatatypes and implemented
-it in the SMT solver CVC4. Our main theoretical contribution has been the
-addition of codatatypes. These are not very difficult
-per se, but there are tricky corner cases to take care of if we care about
-completeness (which we do). On the practical side, we obtained interesting
-results on benchmarks generated from Isabelle theories, including some rather
+We introduced a decision procedure for the ground theory of datatypes and
+codatatypes and implemented it in the SMT solver CVC4. Our main theoretical
+contribution has been the support for codatatypes. These are not very
+difficult per se, but there are tricky corner cases to take care of if we care
+about completeness (which we do).
+%% alliteration
+On the practical side, we obtained
+interesting results on benchmarks generated from Isabelle theories, including
+some rather
 %mind-boggling
@@
 alternative to the counterexample generator Nitpick
-\cite{blanchette-nipkow-2010}. We also see some opportunities to enrich CVC4
-with recursive and corecursive functions. Finally, it might be possible to go
-further in terms of supporting nested and mixed (co)recursion and quantified
-formulas over (co)datatypes in SMT solvers.
+\cite{blanchette-nipkow-2010}. We also see some opportunities to enrich
+SMT solvers with recursive and corecursive functions. Finally, it might be
+possible to go further in terms of supporting nested and mixed (co)recursion
+and quantified formulas over (co)datatypes in solvers.
 
@@
 We owe a great debt to Clark Barrett and Cesare Tinelli for datatype case blah
-blah blah.
+blah blah. \textbf{TODO FIXME}
 %
@@
 Cesare Tinelli, and Christoph Weidenbach, have either encouraged the research on
-codatatype or at least tolerated it, both of which we are thankful for.
-%
-Andrei Popescu helped clarify our thoughts about axiomatizing codatatypes.
-Andreas Lochbihler and Peter Gammie shared their private codatatype-based
+codatatype or at least endured
+%tolerated?
+it, both of which we are thankful for.
+%
+Andreas Lochbihler and Peter Gammie shared their private
 theories with us so we could include them in the benchmarks.
+Andrei Popescu helped clarify our thoughts regarding the axiomatization of
+codatatypes. Dmitriy Traytel took part in discussions about special
+cases.
 %

### 2014-12-11 tex/codatadp.tex
my proof

@@
 \begin{proof}
-  * if two (either directly or indirectly) corecursive constructors
-    $\const{C}$ and $\const{D}$, it is
-    possible to encode infinitely many alternation patterns,
-    e.g. $\const{C}(\const{D}(\const{C}(\const{C}(\ldots))))$;
-    hence no finite models
-
-  * if one corecursive and one noncorecursive, then possibility to create
-    objects of arbitrary depths, e.g. $\const{C}(\ldots(\const{C}(\const{D}))\ldots)$,
-    and hence no finite models
-
-  * leaves us with the case of a single corecursive constructor
-
-  * if any of the noncorecursive arguments has cardinality greater than one,
-    then it is possible to encode alternation patterns using it,
-    e.g. $\const{C}(0,\: \const{C}(1,\: \const{C}(0,\: \const{C}(0,\: \ldots))))$.
-
-  * otherwise, by the coinduction principle, there is only one value that can
-    exist
+By definition, the type is equipped with at least one (directly or indirectly)
+corecursive constructor $\const{C}$. If it additional has second
+corecursive constructor $\const{D}$, it is possible to encode infinitely many
+alternation patterns---e.g.,
+$\const{C}(\const{D}(\const{C}(\const{C}(\ldots))))$---all of which correspond
+to distinct values. If the type has a noncorecursive constructor
+$\const{E}$, it is possible to create terms of arbitrary depths---e.g.,
+$\const{C}(\ldots(\const{C}(\const{E}))\ldots)$. In both cases, there can be
+no finite models.
+
+Therefore, $\const{C}$ must be the only constructor.
+If any of its noncorecursive arguments has a cardinality greater than 1,
+it is possible to encode alternation patterns using it---e.g.,
+$\const{C}(0,\: \const{C}(1,\: \const{C}(0,\: \const{C}(0,\: \ldots))))$---which
+again excludes finite models. Otherwise, the coinduction principle ensures
+that there can be only one value.
+\qed
+\end{proof}
 
@@
 %   or at least two;
-\qed
-\end{proof}
 

### 2014-12-11 tex/codatadp.tex
Working on proof, model completion is more complicated than previously thought.

--- 2014-12-11 tex/codatadp.tex
@@
 \newcommand\Data{\Types_{\mathrm{dt}}}
-\newcommand\Codata{\Types_{\mathrm{codt}}}
+\newcommand\Codata{\Types_{\mathrm{cdt}}}
 \newcommand\Nondata{\Types_{\mathrm{ord}}}
@@
 
-Freely generated inductive datatypes are ubiquitous in functional programs and
+Freely generated datatypes are ubiquitous in functional programs and
 logical specifications. They are especially
@@
 To represent infinite values, % such as $\const{S}(\const{S}(\const{S}(\ldots))$,
-a natural choice is to turn to \emph{codatatypes} (or coinductive datatypes), the
-non-well-founded dual of (inductive) \relax{datatypes}.
+a natural choice is to turn to codatatypes (or coinductive datatypes), the
+non-well-founded dual of (inductive) datatypes.
 %
 Although they are sometimes perceived as esoteric, codatatypes have an important
-role to play in computer science. The verified C compiler CompCert
+role to play in computer science. For example, the verified C compiler CompCert
 \cite{leroy-2009}, the verified Java compiler Jinja\-Threads
@@
 with the more conventional, inductive definition
-\[\keyw{datatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat}) \,\mid\, \const{Infty}\]
-%
-the codatatype avoids one case by unifying the finite and infinite nonzero cases.
+\[\keyw{datatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat}) \,\mid\, \const{Inf}\]
+%
+the codatatype eliminates one case by unifying the finite and infinite cases.
 Moreover, equations such as $\const{S}(\infty) \teq \infty$ hold without additional
@@
 $\const{S}(\const{S}(\const{S}(\ldots)))$, which uniquely identifies the
-value $\infty$.
-
-\nopagebreak
+codatatype value $\infty$.
 
@@
 In this \thewordpaper, we introduce a unified decision procedure for ground
-problems involving datatypes and codatatypes in combination and discuss extensions 
-of the procedure for handling universal (non-ground) problems.
+problems involving both datatypes and codatatypes, and discuss extensions 
+of the procedure when handling universal (non-ground) problems.
 The procedure is
@@
 %
-\[\keyw{codatatype}~\;\ty{stream}_{\,\tau} \,=\, \const{SCons}(\tau,\: \ty{stream}_{\,\tau})\]
+\[\keyw{codatatype}~\;\ty{stream}_{\,\tau} \,=\, \const{SCons}(\tau,\, \ty{stream}_{\,\tau})\]
 %
@@
 values internally. For example,
-the term $\const{SCons}(1,\: \mu\vthinspace s.\; \const{SCons}(0,\: \const{SCons}(9,\: s)))$
+the term $\const{SCons}(1,\, \mu s.\; \const{SCons}(0,\, \const{SCons}(9,\, s)))$
 denotes the lasso-shaped sequence $1, 0, 9, 0, 9, 0, 9, \ldots$\afterLdots.
@@
 \begin{paper}
-Detailed proofs of soundness and completeness are included in the technical report
+The proofs of soundness and completeness are included in the technical report
 associated with this paper \cite{our-report}.
@@
 %\[
-%      \keyw{codatatype}\; \,\ty{llist} \,=\, \const{LNil} \,\mid\, \const{LCons}(\ty{int},\: \ty{llist})
+%      \keyw{codatatype}\; \,\ty{llist} \,=\, \const{LNil} \,\mid\, \const{LCons}(\ty{int},\, \ty{llist})
 %\]
@@
 Although it was written several years ago, the account of related work in
-Barrett et al.\ \cite{barrett-et-al-2007} is still a good reference.
+Barrett et al.\ \cite{barrett-et-al-2007} is still largely up to date.
 Since then, datatypes have been added to Z3 in unpublished work by Leonardo de
@@
 values $\{\bot, \top\}$. %, and may contain other interpreted types (e.g., \ty{int}, \ty{real}).
-The only predicate is equality ($\teq$) and belongs to the logical symbols.
+The only predicate is equality ($\teq$) and is considered a logical symbol.
 Other predicates can be represented as functions that return $\ty{bool}$.
-The metavariables $\delta,\:\varepsilon,\:\tau$ range over types, with $\delta$ reserved
+The metavariables $\delta, \tau$ range over types, with $\delta$ reserved
 for datatypes and codatatypes.
@@
 which collapses to $\const{f} : \tau$ if $n = 0$.
-For a term $t$, the notation $t : \tau$ indicates that it has type $\tau$.
+For a term $t$, the notation $t : \tau$ indicates that $t$ has type $\tau$.
 Functions are invoked in the standard way, with $\const{f}(t_1,\ldots,t_n)$
 applying the $n$-ary function symbol
-\const{f} to $n$ well-typed arguments $t_1 :\nobreak \tau_1$, \ldots, $t_n :
+\const{f} to $n$ well-typed arguments $t_1 : \tau_1$, \ldots, $t_n :
 \tau_n$. Nullary function symbols are called constants and can appear without
@@
 \mathrel{\uplus} \Sel$, where $\Ctr$ are the \emph{constructors} and $\Sel$ are the
-\emph{selectors}; there is no need to consider other function symbols because they can
+\emph{selectors}; there is no need to consider other function symbols, because they can
 be abstracted away as variables when combining theories.
@@
 
-\paragraph{Specifications.}
 A (co)datatype specification consists of $l$~mutually recursive types which are
@@
 arguments and returns a $\delta$ value. The argument types must be either
-ordinary, among the already known (co)datatypes, or among the (co)datatypes
-being introduced.
-%
-To every argument corresponds a selector. The names for the (co)data\-types, the
+ordinary, among the already know (co)datatypes, or among the (co)datatypes being
+introduced.
+%
+To every argument corresponds a selector. The introduced names for the (co)data\-types, the
 constructors, and the selectors must be distinct and different from already
@@
 (\keyw{co})\keyw{datatype}\;\,
-  \delta_1 & {}=\, \smash{\const{C}_{11}(\bigl[\const{s}_{11}^1{:}\bigr]\; \tau_{11}^1, \,\ldots,\, \bigl[\const{s}_{11}^{n_{11}}{:}\bigr]\; \tau_{11}^{n_{11}})} & \\
+  \delta_1 & {}=\, \const{C}_{1}\rlap{${}_{1}$}\phantom{{}_{m}}(\bigl[\const{s}_{11}^1{:}\bigr]\; \tau_{11}^1, \,\ldots,\, \bigl[\const{s}_{11}^{n_{11}}{:}\bigr]\; \tau_{11}^{n_{11}}) & \\
            & {}\;\phantom{=\,}\;{} \smash{\,\,\,\vdots} \\[-\jot] % {}\;\phantom{=}\llap{\ensuremath{\mid}}\; \cdots \\
-           & {}\;\phantom{=\,}\llap{\ensuremath{\,\mid\,}}\; \smash{\const{C}_{1m}(\bigl[\const{s}_{1m}^1{:}\bigr]\; \tau_{1m}^1, \,\ldots,\, \big[\const{s}_{1m}^{n_{1m}}{:}\big]\; \tau_{1m}^{n_{1m}})} \\[-1\jot]
+           & {}\;\phantom{=\,}\llap{\ensuremath{\,\mid\,}}\; \const{C}_{1m}(\bigl[\const{s}_{1m}^1{:}\bigr]\; \tau_{1m}^1, \,\ldots,\, \big[\const{s}_{1m}^{n_{1m}}{:}\big]\; \tau_{1m}^{n_{1m}}) \\[-1\jot]
    \smash{\vdots\,\,\,} \\[-1\jot]
@@
 $x = \const{C}_{i\negvthinspace j}\bigl(\const{s}_{i\negvthinspace j}^1(x), \ldots, \const{s}_{i\negvthinspace j}^{\,n_{\smash{{i\negvthinspace j}}}}(x)\bigr)$.
-This will simplify the presentation. % of the decision procedure.
+This will simplify the presentation of the decision procedure.
 
@@
 \[\begin{aligned}[t]
-      \keyw{datatype}\;\, \ty{tree}_{\,\tau} & \,=\, \const{Node}(\tau,\: \ty{forest}_{\,\tau}) \\[-\jot]
-      \keyw{and}\;\, \ty{forest}_{\,\tau} & \,=\, \const{FNil} \,\mid\, \const{FCons}(\ty{tree}_{\,\tau},\:\, \ty{forest}_{\,\tau}) \\%[-.5\jot]
-      \keyw{codatatype}\; \,\ty{llist}_{\,\tau} & \,=\, \const{LNil} \,\mid\, \const{LCons}(\const{lhead}{:}\; \tau,\; \const{ltail}{:}\;\ty{llist}_{\,\tau})
+      \keyw{datatype}\;\, \ty{tree}_{\,\tau} & \,=\, \const{Node}(\tau,\, \ty{forest}_{\,\tau}) \\[-\jot]
+      \keyw{and}\;\, \ty{forest}_{\,\tau} & \,=\, \const{FNil} \,\mid\, \const{FCons}(\ty{tree}_{\,\tau},\, \ty{forest}_{\,\tau}) \\%[-.5\jot]
+      \keyw{codatatype}\; \,\ty{llist}_{\,\tau} & \,=\, \const{LNil} \,\mid\, \const{LCons}(\tau,\, \ty{llist}_{\,\tau})
 \end{aligned}
@@
 that datatype. This rules out ill-founded specifications such as
-\[\keyw{datatype}\;\, \ty{fstream} \,=\, \const{FSCons}(\ty{int},\:\, \ty{fstream})\]
+\[\keyw{datatype}\;\, \ty{fstream} \,=\, \const{FSCons}(\ty{int},\,\ty{fstream})\]
 For codatatypes, no such check is necessary because there is always a term,
@@
 %
-%\[\keyw{codatatype}\;\, \ty{stream} \,=\, \const{SCons}(\ty{int},\:\ty{stream})\]
+%\[\keyw{codatatype}\;\, \ty{stream} \,=\, \const{SCons}(\ty{int},\,\ty{stream})\]
 %
@@
 
-\paragraph{Properties.}
 Datatypes and codatatypes share basic properties pertaining to constructors and
@@
 \text{Selection:}\quad
-  & s_{i\negvthinspace j}^{\,k}(\const{C}_{i\negvthinspace j}(x_1,\ldots,x_{n_{i\negvthinspace j}})) = x_k
+  & s_{i\negvthinspace j}^k(\const{C}_{i\negvthinspace j}(x_1,\ldots,x_{n_{i\negvthinspace j}})) = x_k
 \end{aligned}
@@
 exists, whereas distinctness and injectivity ensure that ``no confusion''
-arises. The result of selectors applied to the wrong constructor is left
-completely unspecified.
-
-%  * how to deal with "wrong" selectors, e.g.
-%        hd(nil1) = hd(nil2)?
-%      * leave them unspecified; hence hd(nil1) = hd(nil2) in some models,
-%        and not in other models
+arises.
 
@@
 for a datatype of natural numbers constructed from $\const{Z}$ and $\const{S}$,
-this rules out nonstandard models, which could contain cyclic values---e.g.,
-an $n$ such that $n \teq \const{S}(n)$---or even infinite acyclic values
+this rules out nonstandard models, which could contain cyclic values (e.g.,
+an $n$ such that $n \teq \const{S}(n)$) or even infinite acyclic values
 $\const{S}(\const{S}(\ldots))$.
@@
 this is type correct or $x \teq y$ otherwise.
+
+FIXME: Find out what ensures that codatatypes have infinite values.
+
 This axiom schema captures the idea that two values that yield the same
@@
 selectors and discriminators.
-
-In addition, codatatypes are guaranteed to contain all values corresponding to
-infinite ground constructor terms. This cannot be capture by a first-order
-axiomatization, since there may be uncountably many of them. For example,
-$\ty{stream}_{\,\ty{int}}$ is isomorphic to the uncountable function space
-$\ty{nat} \to \ty{int}$.
 
@@
 
-%      * again: when looking at the universal ground (?) theory, enough to
-%        consider bisimilarity/bisimulation
-
-% specifies the
-
-\paragraph{Semantics.}
+      * again: when looking at the universal ground (?) theory, enough to
+        consider bisimilarity/bisimulation
+
+specifies the
+
+  * how to deal with "wrong" selectors, e.g.
+        hd(nil1) = hd(nil2)?
+      * leave them unspecified; hence hd(nil1) = hd(nil2) in some models,
+        and not in other models
+
 Datatypes can be defined semantically as the initial model of an equational
@@
 
-A related semantic view of datatypes is as an initial algebra. Codatatypes are
-then defined dually as a final coalgebra. The datatypes are generated by their
+A related semantic view of datatypes is as initial algebra. Codatatypes are then
+defined dually as a final coalgebra. The datatypes are generated by their
 constructors, whereas the codatatypes are viewed through their selectors
 \cite{xxx}. By uniformly focusing on the constructors, the axiomatic approach
-emphasizes the commonality between datatypes and codatatypes, while
-sacrificing a theoretically fruitful notion of duality.
-
-\paragraph{Special Cases.}
-A type $\delta$ depends on another type $\varepsilon$ if $\varepsilon$ is the
-type of an argument to one of $\delta$'s constructors. Semantically, a set of
-types is mutually (co)recursive if and only if the associated dependency graph
-is strongly connected. Types can be declared together as mutually
-(co)recursive even if they are not actually (co)recursive; however, the
-semantic notion is more fruitful and is the one that interests us.
-
-Furthermore, nothing forbids specifications with no (co)recursion, such as
-\[\begin{aligned}[t]
-      \keyw{datatype}\;\, \ty{enat} & \,=\, \const{Nat}(\ty{nat}) \,\mid\, \const{Infty} \\[-.5\jot]
-      \keyw{codatatype}\; \,\ty{complex} & \,=\, \const{Complex}(\const{re}{:}\; \ty{real},\; \const{im}{:}\;\ty{real})
-\end{aligned}
-\]
-Semantically, it makes no difference whether the types are introduced as
-datatypes or as codatatypes.
-%Without loss of generality, we consider that these types are datatypes
-%instead of codatatypes.
-
-Finally, some codatatypes are so degenerate that they have infinite values
-even though they are finite. The simplest example is
-\[\keyw{datatype}\;\, \ty{a} \,=\, \const{A}(\ty{a})\]
-This is a finite codatatype, of cardinality one, whose single value is
-infinite: $\mu\vthinspace a.\; A(a)$. Other specimens are
-\[\begin{aligned}[t]
-      \keyw{codatatype}\;\, \ty{b} & \,=\, \const{B}(\ty{b},\: \ty{c},\: \ty{b},\: \ty{unit}) \\[-\jot]
-             \keyw{and}\;\, \ty{c} & \,=\, \const{C}(\ty{unit},\: \ty{c},\: \ty{b},\: \ty{c})
-\end{aligned}
-\]
-where \ty{unit} is a datatype with the single constructor $\const{Unity} :
-\ty{unit}$. We call such types \emph{corecursive singletons}.
-For the decision procedure, it will be important to detect those finite
-codatatypes.
-
-% In other words, all finite corecursive codatatypes are corecursive singletons.
-
-\begin{lemma}
-If the domain interpreting a corecursive codatatype is finite, it must be $1$.
-\end{lemma}
-
-\begin{proof}
-TODO.
-\qed
-\end{proof}
+emphasizes the commonality between datatypes and codatatypes, while sacrificing
+a theoretically fruitful notion of duality.
+
+    semantically, $k$ types are mutually (co)recursive if their dependency graph
+    (x -> y if x is the type of an argument to one of y's constructors) is
+    strongly connected
+
+    (co)datatypes can be declared together as mutually (co)recursive even if
+    they are not actually (co)recursive
+
+    the semantic notion is more fruitful and the one we will focus on in the
+    sequel
+
+    Also, nothing forbids specifications with no (co)recursion, e.g.
+
+      datatype enat = Nat(nat) | Infty
+      codatatype complex = Complex(re: real, im: real)
+
+    In this case, it makes no difference whether the type is introduced as a
+    datatype or as a codatatype. Without loss of generality, we consider that
+    these types are datatypes instead of codatatypes.
+
+Degenerate Codatatypes
+
+    Consider
+
+      codatatype a = A(a)
+
+    This is a finite codatatype, with a single value, which is infinite:
+    $\mu a. A(a)$
+
+    We need to detect those finite codatatypes. They are easy to
+    characterize semantically.
+
+    Lemma D.1. The only possible finite cardinality for the domain interpreting
+    a (corecursive) codatatype is 1.
+
+    Proof.
+      Assume you have at least 2 different values of a codatatype.
+      They differ at one point.
+      * 
+    Qed
 
     An immediate consequence by the distinctness axiom is that these types have
-necessarly one constructor and that all arguments have types of cardinality 1.
-%
-Sometimes, a type might be a corecursive singleton only in some models. For
-example, if \ty{unit} is an uninterpreted type, \ty{b} and \ty{c} will be
-singleton types precisely in the models that interpret \ty{unit} as a
-singleton.
+    necessarly one constructor, and that all arguments are of types of
+    cardinality 1.
+
+    Easy case: $k$ mutually recursive codatatypes, each equipped with one
+    constructor, all of which take each other in arguments in various
+    combinations, with potential other types that are known to have cardinality
+    1. e.g.
+
+      codatatype b = B b c b
+             and c = C c b c
+
+    Medium case: like above, but some of the arguments are other types not
+    participating in the corecursion that are known to have cardinality 1, e.g.
+
+      codatatype d = D unit d d unit
+
+    where "unit" is a datatype with the single constructor "Unity : unit".
+
+    Though case: like above, but with any other types where it's not clear
+    whether it could have cardinality one or not.
+
+      type e
+      codatatype f = F e f f unit
 
@@
 
-We present the calculus in three sections, shown in Figures~\ref{fig:cc-rules},~\ref{fig:ab-rules}, and~\ref{fig:split-rule}.
+We present the calculus in three parts.
+In the first part we compute the bidirectional closure of $\Ec$,
+in the second part we make inferences based on cyclicity and bisimilarity,
+and in the third part (when necessary) we branch on constructor types for various terms in $\tEc$.
 Following the conventions from~\cite{}, 
@@
   \quad
-  \ttpath{\Ec}{\ec{t}}{\ec{t}} \neq \emptyset
+  %\ttpath{\Ec}{\ec{t}}{\ec{t}} \neq \emptyset
+  \Val \ec{t_1} = \mu x. t_2
+  \quad
+  x \in FV( t_2 )
 }{
@@
 
+%For presentation of the rules in this step, we rely on a representation of terms in the $\mu$-notation.
+The rules of this step contain premises that involve terms containing $\mu$-bindings.
+Recall that these bindings can be used for representing (possibly cyclic) values,
+for instance, $\mu x. \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
+Formally, we recursively define a $\mu$-term as a term-like structure either of the form
+$\mu x. \const{C}( \vec t )$ for some constructor $\const{C} \in \Ctr$ and $\mu$-terms $\vec t$, 
+or a variable $x$.
+For consistency, we will also consider $\mu x. \const{c}$ to be a $\mu$-term
+in the case where $c$ is a constant of type $\tau \in \Nondata$.
+We consider only $\mu$-terms that are \emph{well-typed}, which we define recursively:
+if the type of $\const{C}$ is $\vec \tau \rightarrow \upsilon$
+and $\vec v$ are well-typed $\mu$-terms of type $\vec \tau$ under the assumption that $x$ has type $\upsilon$,
+then $\mu x. \const{C}( \vec v )$ is a well-typed term of type $\upsilon$.
+We say a $\mu$-term is \emph{closed} if and only if it contains no free variables.
+A $\mu$-term is \emph{cyclic} if and only if it contains a bound variable.
+We write $t \vsim s$ if $\mu$-terms $t$ and $s$ are syntactically equivalent for some renaming of $\mu$-bound variables,
+in which case we will commonly say $t$ and $s$ are \emph{bisimilar}.
+%, or informally are \emph{observationally equivalent}.
+%For convienience, we will use $\mu$-terms to refer to (classes of) values for both co-datatype and datatype terms,
+%where in the latter case, a $\mu$-term with a bound variable denotes an infeasible term.
+%where the latter case adds the restriction on $\mu$-terms $t$ that no subterms of $t$ contain are bound variables.
+
+Intuitively, a $\mu$-term $t$ of type $\tau$ describes a class of values of $\tau$, 
+either in the case when $\tau$ is a co-datatype or datatype type,
+where in the latter case, such a $t$ containing a bound variable describes an infeasible class of values.
+At the beginning of this step, 
+we compute a mapping $\Val$ from equivalence classes to $\mu$-terms 
+that describe the class of values that terms in that equivalence class can take in the models of $\Ec$. 
+For each equivalence class $\ec{x}$ of $\Ec$, we associate a fresh variable $v_{\ec{x}}$ of the same type as $x$ not occurring in $\tEc$,
+and initially say that $\Val$ contains $\ec{x} \mapsto v_{\ec{x}}$ for each equivalence class $\ec{x}$ of $\Ec$ 
+(in other words, there are initially no constraints on the values for any equivalence class).
+Then, $\Val$ is updated by applying the following rule exhaustively:
+
+\(
+\inferrule{
+  v_{\ec{x}} \in FV( \Val )
+  \quad
+  C( t_1, \ldots, t_n ) \in \ec{x}
+  \quad
+  C \in \Ctr
+}{
+  \Val := \Val \{ v_{\ec{x}} \mapsto \mu v_{\ec{x}}. C( v_{\ec{t_1}}, \ldots, v_{\ec{t_n}} ) \}
+}
+\)
+
+Here, we write $FV( \Val )$ to denote the \emph{free variables} occurring in $\Val$, which
+excludes variables that occur beneath $\mu$-binders.
+It is straightforward to see that the height of terms produced as a result of this rule
+is bounded by the number of equivalence classes of $\Ec$,
+and thus our construction of $\Val$ is terminating.
+
+\begin{example}
+Say that $\Ec$ contains (distinct) equivalence classes $\ec{x}$, $\ec{y}$, $\ec{z}$ and $\ec{w}$ where $\const{C}(z,x) \in \ec{y}$ and $\const{C}(w,y) \in \ec{x}$.
+A possible sequence of applications of the above rule is shown below.
+In the first column, we show the equivalence class that was considered,
+and in the second column, we show the content of $\Val$.
+\[\begin{array}{|l|@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l|}
+\hline
+\text{Eqc} & \multicolumn{3}{c}{\Val} & \ 
+\\
+\hline
+& \{ & \ec{x} \mapsto v_{\ec{x}}, & \ec{z} \mapsto v_{\ec{z}}, & \\
+&    & \ec{y} \mapsto v_{\ec{y}}, & \ec{w} \mapsto v_{\ec{w}} & \} \\
+\ec{x}
+& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ), & \ec{z} \mapsto v_{\ec{z}}, & \\
+&    & \ec{y} \mapsto v_{\ec{y}}, & \ec{w} \mapsto v_{\ec{w}} & \} \\
+\ec{y}
+& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ) ), & \ec{z} \mapsto v_{\ec{z}}, & \\
+&    & \ec{y} \mapsto \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
+\ec{x}
+& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ) ), & \ec{z} \mapsto v_{\ec{z}}, & \\
+&    & \ec{y} \mapsto \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ) ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
+\hline
+\end{array}\]
+Informally, this indicates that the values for $x$ (resp. $y$) in models of $\Ec$
+are of the form $\const{C}( v_{\ec{w}}, \const{C}( v_{\ec{z}}, \ldots ))$ (resp. $\const{C}( v_{\ec{z}}, \const{C}( v_{\ec{w}}, \ldots ))$) 
+for some value of $v_{\ec{w}}$ and $v_{\ec{z}}$.
+$\Box$
+\end{example}
+
+Given $\Val$, the rules for cyclicity and bisimilarity (shown in Figure~\ref{fig:ab-rules}) can be stated as follows.
+For cyclicity, if there is an equivalence class $\ec{t_1}$ of datatype type whose class of values $\Val \ec{t_1} = \mu x. t_2$ are cyclic,
+as indicated by the assumption $x \in FV( t_2 )$,
+then we can conclude that $\Ec$ is unsatisfiable.
+For bisimilarity, if there are two equivalence classes $\ec{t_1}$ and $\ec{t_2}$ 
+of co-datatype type whose classes of values $\Val \ec{t_1}$ and $\Val \ec{t_2}$ are equivalent modulo renaming of $\mu$-bound variables,
+then we can conclude $t_1 \teq t_2$.
+We demonstrate the bisimilarity rule concretely with the following example.
+
+\begin{example}
+Let $\const{C}$ be a constructor of type $\tau_1 \times \tau_2 \rightarrow \tau_2$,
+and let $\Ec$ be the set $\{ x \teq \const{C}(y,\const{C}(y,x)) \}$.
+After applying the rules in Figure~\ref{fig:cc-rules} to completion, the equivalence classes of $\Ec$ are
+$\{ x, \const{C}(y,\const{C}(y,x)) \}$, $\{ \const{C}(y,x) \}$, and $\{ y \}$.
+We then construct $\Val$ using the aforementioned steps, after which we obtain:
+\[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}r@{\hspace{.5em}}l@{\hspace{.5em}}r@{\hspace{.5em}}l}
+\Val \ec{x} & = & 
+\mu v_{\ec{x}}. & \const{C}( v_{\ec{y}}, & \mu v_{\ec{\const{C}(y,x)}}. & \const{C}( v_{\ec{y}}, v_{\ec{x}} ) ) \\
+\Val \ec{\const{C}(y,x)} & = & 
+\mu v_{\ec{\const{C}(y,x)}}. & \const{C}( v_{\ec{y}}, & \mu v_{\ec{x}}. & \const{C}( v_{\ec{y}}, v_{\ec{\const{C}(y,x)}} ) )
+\end{array}\]
+Notice that $\Val \ec{x}$ and $\Val \ec{\const{C}(y,x)}$ are syntactically equivalent for the renaming
+$\{ v_{\ec{x}} \mapsto v_{\ec{\const{C}(y,x)}}, v_{\ec{\const{C}(y,x)}} \mapsto v_{\ec{x}} \}$,
+and thus our calculus may infer $x \teq \const{C}(y,x)$.
+Intuitively, this equality holds, since the values of $x$ and $\const{C}( y, x )$ are observationally equivalent for any value of $v_{\ec{y}}$.
+\qed
+\end{example}
+
+\begin{comment}
 To explain the rules in Figure~\ref{fig:ab-rules}, we introduce the following notions.
@@
 in which case $t_1$ must be equal to $t_2$.
-We demonstrate this rule concretely with the following example.
-\rem{ensure that \const{LCons} understood}
-
-\begin{example}
-Let $\Ec$ be the set $\{ x \teq \const{LCons}(y,\const{LCons}(y,x)) \}$.
-After applying the rules in Figure~\ref{fig:cc-rules} to completion, the equivalence classes of $\Ec$ are
-$\{ x, \const{LCons}(y,\const{LCons}(y,x)) \}$, $\{ \const{LCons}(y,x) \}$, and $\{ y \}$.
-The set $\tpath{\Ec}{\ec{x}}$ is:
-\[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
-\{ & \ec{x} \rightarrow^{\const{LCons}}_1 & \ec{y}, & & \\ 
-   & \ec{x} \rightarrow^{\const{LCons}}_2 & \ec{\const{LCons}(y,x)} \rightarrow^{\const{LCons}}_1 & \ec{y},& \\
-   & \ec{x} \rightarrow^{\const{LCons}}_2 & \ec{\const{LCons}(y,x)} \rightarrow^{\const{LCons}}_2 & \ec{x} & \}.
-\end{array}\]
-The set $\tpath{\Ec}{\ec{\const{LCons}(y,x)}}$ is:
-\[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
-\{ & \ec{\const{LCons}(y,x)} \rightarrow^{\const{LCons}}_1 & \ec{y}, & & \\ 
-   & \ec{\const{LCons}(y,x)} \rightarrow^{\const{LCons}}_2 & \ec{x} \rightarrow^{\const{LCons}}_1 & \ec{y},& \\
-   & \ec{\const{LCons}(y,x)} \rightarrow^{\const{LCons}}_2 & \ec{x} \rightarrow^{\const{LCons}}_2 & \ec{\const{LCons}(y,x)} & \}.
-\end{array}\]
-Thus, we may infer $x \teq \const{LCons}(y,x)$ by the rule \rn{Bisimilar}
-for the mapping $\sigma :=$ $\{ \ec{x} \mapsto \ec{\const{LCons}(y,x)}, \ec{\const{LCons}(y,x)} \mapsto \ec{x} \}$.
-\qed
-\end{example}
+We demonstrate an application of this rule concretely with the following example.
+\end{comment}
 
@@
 
-\rem{Require : all co-datatypes either have infinite cardinality, or have cardinality 1.}
-
-Finally, if neither a rule from Figure~\ref{fig:cc-rules} or Figure~\ref{fig:ab-rules} applies to $\Ec$, 
+\paragraph{Step 3: Branching}
+If neither a rule from Figure~\ref{fig:cc-rules} or Figure~\ref{fig:ab-rules} applies to $\Ec$, 
 the rules from Figure~\ref{fig:split-rule} are applied.
@@
 then $\Ec$ is satisfiable $\thD$.
-
-%We first show that for any saturated node $\Fc$, 
-%additional equalities $\Gc$ can be added to $\Fc$ such that each codatatype/datatype equivalence class of $\Fc \cup \Gc$ has a constructor term.
-%For this purpose, we extend the calculus in this section.
-%For a subset of equivalence classes $S$ of $\Ec$, write $\dpath{\Ec}{C,n}{S}$ to denote the set
-%such that $e' \in \dpath{\Ec}{C,n}{S}$ if and only if there exists an $e \in S$ such that $e \rightarrow^{C}_n e'$ is induced by $\Ec$.
-
-\begin{comment}
-\begin{figure}[t]
-\centering
-\begin{tabular}{c}
-\rn{O-Finish}
-\(
-\inferrule{
-  \Ac := x
-  \quad
-  x : \tau
-  \quad
-  \tau \in \Nondata
-  \quad
-  c \not\in \tEc
-}{
-  \Ac := \emptyset \quad \Ec := \Ec \cup \{ x \teq c \}
-}
-\)
-\\[3.7ex]
-\rn{C-Finish}
-\(
-\inferrule{
-  \Ac := x
-  \quad
-  x : \tau
-  \quad
-  \tau \in \Codata
-  \quad
-  C_i \in \Ctr^\tau
-  \quad 
-  C_i \not\in \{ C_j \mid d \in \Dc, C_j( \vec t ) \in d \}
-}{
-  \Ac := \emptyset \quad \Ec := \Ec \cup \{ x \teq C_i( \vec y ) \} \quad \Rc := \Rc \cup \vec y
-}
-\)
-\\[3.7ex]
-\rn{C-Rec}
-\(
-\inferrule{
-  \Ac := x
-  \quad
-  x : \tau
-  \quad
-  \tau \in \Codata
-  \quad
-  C_i \in \Ctr^\tau
-  \quad
-  C_i \text{ infinite arg } j
-}{
-  \Ac := y_j \quad 
-  \Ec := \Ec \cup \{ x \teq C_i( \vec y ) \} \quad 
-  \Rc := \Rc \cup ( \vec y \setminus y_j ) \quad 
-  \Dc := \dpath{\Ec}{C_i,j}{\Dc}
-}
-\)
-\end{tabular}
-\caption{Rules for completion.
-}
-\label{fig:completion-rules}
-\end{figure}
-\end{comment}
-
-%and second show how to construct a valuation for the variables of $\Fc \cup \Gc$.
-
-%\begin{lemma}
-%\end{lemma}
-
-Given a set of disjoint paths $P$ where each $p \in P$
-we write $\interp{P}$ to denote the value
+First, we require additional terminology concerning $\mu$-terms.
+If $u$ is a subterm of $\mu$-term $v$, then the \emph{interpretation of $u$ in $v$} is the $\mu$-term $v\interp{u}^\emptyset$ as returned by the following recursive procedure,
+where $x \mapsto \mu x. C( \vec t ) \in v$ indicates that $\mu x. C( \vec t )$ is the subterm of $v$ that binds this occurrence of variable $x$.
+\[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
+v\interp{x}^M & = & v\interp{\mu x. C( \vec t )}^M & x \not\in M, x \mapsto \mu x. C( \vec t ) \in v  \\
+v\interp{x}^M & = & x & \text{otherwise} \\
+v\interp{\mu x. C( \vec t )}^M & = & \mu x. C( v\interp{ \vec t }^{M \cup \{ x \}} ) & x \not\in M \\
+v\interp{\mu x. C( \vec t )}^M & = & x & \text{otherwise}
+\end{array}\]
+We will write $v\interp{u}$ as shorthand for $v\interp{u}^\emptyset$.
+Given a $\mu$-term $v$ of the form $\mu x. C( \vec t )$, 
+we say $u$ is a \emph{bisimilar subterm} of $v$ if and only if $v\interp{u} \vsim v$.
+We say $v$ is \emph{normal} if and only if all of its bisimilar subterms are $x$,
+and all of its strict subterms are also normal.
+For instance, the $\mu$-term $v$ of the form $\mu x. \const{S}( \mu y. \const{S}( x ) )$ is not normal, 
+since 
+$v\interp{\mu y. \const{S}( x )}^\emptyset =$ 
+$\mu y. \const{S}( v\interp{x}^{\{y\}} ) =$ 
+$\mu y. \const{S}( v\interp{\mu x. \const{S}( \mu y. \const{S}( x ) ) }^{\{y\}} ) =$ 
+$\mu y. \const{S}( \mu x. \const{S}( v\interp{\mu y. \const{S}( x )}^{\{x,y\}} ) ) =$ 
+$\mu y. \const{S}( \mu x. \const{S}( y ) )$,
+which is syntactically equivalent to $v$ for the renaming $\{ x \mapsto y, y \mapsto x \}$.
+For any $\mu$-value $v$ of the form $\mu x. C( \vec w )$, we may recursively construct a normal value $\nf{v}$
+(which we will call the \emph{normal form} of $v$)
+that is a renaming of $v$ by replacing all of the bisimilar subvalues of $v$ with $x$,
+and by replacing all of the non-normal subterms of $v$ with their corresponding normal forms.
+For instance, $\nf{(\mu x. \const{S}( \mu y. \const{S}( x ) ))} = \mu x. \const{S}( x )$.
 
@@
 
-We construct a model $\M$ of $\thD$ that interprets constructors and correctly-applied selectors in the obvious way,
-interprets wrongly-applied selectors based on equality information in $\Fc$,
-interprets the variables and $x$ of $\Fc$ based on a mapping $\Val$ from the equivalence classes of $\Fc$ to values.
-Initially, $\Val$ contains $\ec{x} \mapsto v_{\ec{x}}$ for each equivalence class $\ec{x}$ of $\Fc$,
-where $v_{\ec{x}}$ is a fresh variable of the same type as $x$ not occurring in $\Fc$.
-Then, $\Val$ is updated according to the following rules:
-
-\begin{comment}
-\rn{Ord}
+We construct a model $\M$ of $\thD$ that interprets each type $\tau$
+as a maximal set of closed normal $\mu$-terms of that type that are 
+unique up to renaming of $\mu$-bound variables,
+and acyclic when $\tau \not\in \Codata$.
+The interpretation of constructor terms
+$\M( C( \vec t ) )$ for $C \in \Ctr$ is the value in $\M$ that is a renaming of $\nf{(\mu x. C( \M( \vec t ) ))}$ where $x \not\in FV( \vec t )$.
+The interpretation of (correctly-applied) selector terms
+$\M( s^j_i( t ) )$ for $s^j_i \in \Sel$ is the value in $\M$ that is a renaming of $\M( t )\interp{u_j}$
+where $\M( t )$ is of the form $\mu x. C_i( \vec u )$ for some $\vec u$.
+For instance, if $\M( t ) = \mu x. \const{C}_1( x )$ (call this term $v$), 
+then $\M( s_1^1( t ) )$ is a renaming of
+$v\interp{x} =$ 
+$v\interp{\mu x. \const{C}_1( x )}^\emptyset =$
+$\mu x. \const{C}_1( v\interp{x}^{\{x\}} ) =$
+$\mu x. \const{C}_1( x )$.
+%To interpret a correctly-applied selector in $\M$, if $v = \mu x. C_i( \vec t )$, then $\M( s^j_i( v ) )$ is the value in $\M( \tau )$ that is a renaming of $v\interp{t_j}$,
+%where $t_j$ has type $\tau$.
+%We assume $\M$ interprets constructors and correctly-applied selectors in the obvious way
+%\footnote{
+%The interpretation of $\M( C( \vec t ) )$ is the value in $\M( \tau )$ that is a renaming of $\nf{\mu x. C( \M( \vec t ) ) }$.
+%To interpret a correctly-applied selector, if $v = \mu x. C_i( \vec t )$, then $\M( s^j_i( v ) )$ is the value in $\M( \tau )$ that is a renaming of $v\interp{t_j}$,
+%where $t_j$ has type $\tau$.
+%},
+Additionally, $\M$ will interpret wrongly-applied selectors based on equality information in $\Fc$,
+and will interpret the variables of $\Fc$ based on the valuation $\Val$ which was constructed in step two of our calculus.
+
+Recall that $\Val$ is a mapping from the equivalence classes of $\Fc$ to $\mu$-terms.
+We first \emph{complete} $\Val$ so that it maps equivalence classes only to closed $\mu$-terms, that is, terms containing no free variables.
+This is done by exhaustively applying the following rule:
+
+\rn{Complete}
 \(
 \inferrule{
-  \ec{x} \not\in \vdom(\Val)
-  \quad
-  x : \tau
-  \quad
-  \tau \in \Nondata
-  \quad
-  c \not\in \vrange(\Val)
-}{
-  \Val := \Val \cup \ec{x} \mapsto c
-}
-\)
-
-\rn{D-Ind}
-\(
-\inferrule{
-  \ec{x} \not\in \vdom(\Val)
-  \quad
-  x : \tau
-  \quad
-  \tau \in \Data
-  \quad
-  C( \vec t ) \in \ec{x}
-  \quad
-  \ec{\vec t} \in \vdom(\Val)
-}{
-  \Val := \Val \cup \ec{x} \mapsto C( \Val \ec{\vec t}  )
-}
-\)
-
-\rn{D-Base}
-\(
-\inferrule{
-  \ec{x} \not\in \vdom(\Val)
-  \quad
-  x : \tau
-  \quad
-  \tau \in \Data
-  \quad
-  \ec{x} \text{ ctr-free}
-  \quad
-  c \not\in \vrange(\Val)
-}{
-  \Val := \Val \cup \ec{x} \mapsto c
-}
-\)
-\end{comment}
-
-\rn{Ind}
-\(
-\inferrule{
-  %x : \tau
-  %\quad
-  %\tau \in \Codata
   v_{\ec{x}} \in FV( \Val ) 
   \quad
-  C( t_1, \ldots, t_n ) \in \ec{x}
+  c \in \M
   \quad
-  C \in \Ctr
-}{
-  \Val := \Val \{ v_{\ec{x}} \mapsto \mu\vthinspace v_{\ec{x}}. C( v_{\ec{t_1}}, \ldots, v_{\ec{t_n}} ) \}
-}
-\)
-
-\rn{Base}
-\(
-\inferrule{
-  %x : \tau
-  %\quad
-  %\tau \in \Codata
-  %\quad
-  v_{\ec{x}} \in FV( \Val )
-  %\quad
-  %\ec{x} \text{ ctr-free}
-  \quad
-  c \not\sim c' \text{ for all } c' \in \vrange(\Val)
+  c \not\sim t \sigma \text{ for all } \sigma, t \in \vrange(\Val)
 }{
@@
 
-\rem{define $\sim$}
-We assume that first \rn{Ind} is applied exhaustively, after which \rn{Base} is applied exhaustively.
-%the rule \rn{D-Ind} is applied to completion before applying \rn{D-Base}.
-In the rule \rn{Ind}, we write $FV( \Val )$ to denote the free variables occurring in $\Val$.
-Notice that this excludes variables beneath $\mu$-binders.
-
-As an example,
-say that $\Fc$ contains equivalence classes $\ec{x}$ and $\ec{y}$ where $\const{C}(z,x) \in \ec{y}$ and $\const{C}(w,y) \in \ec{x}$, where $z$ and $w$ are variables of integer type.
-A possible application of the above rules, showing the state of $\Val$ after each application, is as follows:
-\[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
-\hline
-\text{Rule} & \multicolumn{4}{c}{\Val} \ 
-\\
-\hline
-& \{ & \ec{x} \mapsto v_{\ec{x}}, & \ec{z} \mapsto v_{\ec{z}}, \\
-&    & \ec{y} \mapsto v_{\ec{y}}, & \ec{w} \mapsto v_{\ec{w}} & \} \\
-\rn{Ind} \ec{x}
-& \{ & \ec{x} \mapsto \mu\vthinspace v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ), & \ec{z} \mapsto v_{\ec{z}}, \\
-&    & \ec{y} \mapsto v_{\ec{y}}, & \ec{w} \mapsto v_{\ec{w}} & \} \\
-\rn{Ind} \ec{y}
-& \{ & \ec{x} \mapsto \mu\vthinspace v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu\vthinspace v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ) ), & \ec{z} \mapsto v_{\ec{z}}, \\
-&    & \ec{y} \mapsto \mu\vthinspace v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
-\rn{Ind} \ec{x}
-& \{ & \ec{x} \mapsto \mu\vthinspace v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu\vthinspace v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ) ), & \ec{z} \mapsto v_{\ec{z}}, \\
-&    & \ec{y} \mapsto \mu\vthinspace v_{\ec{y}}. \const{C}( v_{\ec{z}}, \mu\vthinspace v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ) ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
-\rn{Base} \ec{z}
-& \{ & \ec{x} \mapsto \mu\vthinspace v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu\vthinspace v_{\ec{y}}. \const{C}( \const{0}, v_{\ec{x}} ) ), & \ec{z} \mapsto \const{0}, \\
-&    & \ec{y} \mapsto \mu\vthinspace v_{\ec{y}}. \const{C}( \const{0}, \mu\vthinspace v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ) ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
-\rn{Base} \ec{w}
-& \{ & \ec{x} \mapsto \mu\vthinspace v_{\ec{x}}. \const{C}( \const{1}, \mu\vthinspace v_{\ec{y}}. \const{C}( \const{0}, v_{\ec{x}} ) ), & \ec{z} \mapsto \const{0}, \\
-&    & \ec{y} \mapsto \mu\vthinspace v_{\ec{y}}. \const{C}( \const{0}, \mu\vthinspace v_{\ec{x}}. \const{C}( \const{1}, v_{\ec{y}} ) ), & \ec{w} \mapsto \const{1} & \} \\
-\end{array}\]
-
-It is clear that after applying these rules to completion, $\Val$ is a mapping from equivalence classes to ground values.
-
-\begin{eqnarray} \label{eqn:m-bij-dt}
-\Val \ec{x} = \Val \ec{y} \text{ if and only if } \ec{x} = \ec{y}
+%We write $\mathcal{T}(\Val)$ to denote the smallest set containing $v\interp{u}$ for each subterm $u$ of $v$ in the range of $\Val$.
+We write $\vrange(\Val)$ to denote the set of $\mu$-terms mapped to by $\Val$.
+Since this rule removes one variable from the set $FV( \Val )$ and does not add any additional variables to $FV( \Val )$, 
+it can only be applied a finite number of times.
+
+Let $\Val^\ast$ be the result of applying the above rule to completion on the mapping $\Val$ we constructed during step 2 of our calculus.
+We first argue that $FV( \Val^\ast ) = \emptyset$.
+Since \rn{Split} does not apply to $\Fc$,
+all equivalence classes of $\Fc$ finite datatype type contain constructor terms, 
+and thus $FV( \Val )$ contains no variables of finite datatype type.
+For each equivalence classes of $\Fc$ with finite co-datatype $\tau$, 
+we know $\tau$ must have cardinality one, due to \rem{reference}.
+Since \rn{Singleton} does not apply to $\Fc$, 
+there is only one equivalence class of type $\tau$, call it $\ec{x}$.
+%It cannot be the case that $v_{\ec{x}} \in FV(\Val^\ast \ec{x})$, then we may apply \rn{Complete} for the (single) value $c$ of type $\tau$, 
+%for which the second premise holds since $\ec{x}$ is the only equivalence class of type $\tau$.
+If it were the case that $v_{\ec{x}} \in FV(\Val^\ast )$,
+then \rn{Complete} would apply for the (single) value $c$ of type $\tau$, 
+for which the second premise holds since $\ec{x}$ is the only equivalence class of type $\tau$.
+Since all $\tau \in \Nondata$ are infinite, all variables in $FV( \Val^\ast )$ have infinite type.
+For any infinite type $\tau$,
+there exists a closed normal $\mu$-term $c_1$ that is not bisimilar to any instance of a term in $t \in \mathcal{T}(\Val^\ast)$.
+To argue this is the case, notice that either $\tau$ is a recursive type, or contains an infinite subfield of type $\upsilon$.
+In the former case, we choose a value $c_1$ whose depth is greater than any term in $\vrange(\Val)$.
+In the latter case, we choose a value $c_1$ containing a subterm of type $\upsilon$ not yet occurring in $\Val$.
+%since $\mathcal{T}(\Val^\ast)$ is a finite set. \rem{elaborate} % (noting that a closed $\mu$-term is only bisimilar to other closed $\mu$-terms).
+Thus, $FV( \Val^\ast )$ contains no variables of infinite type,
+and consequently $\Val^\ast$ is a mapping from the equivalence classes of $\Fc$ to closed $\mu$-terms.
+We now argue the following invariant is maintained by applications of \rn{Complete}:% for all $\ec{x}, \ec{y}$:
+\begin{eqnarray}
+%\label{eqn:m-normal} 
+%\Val \text{ maps to normal $\mu$-terms only } \\
+%\text{If } \Val \ec{x} \text{ is not closed and has subterm $u$, then } \Val \ec{x} \interp{ u } = \Val \ec{y} \text{ for some } \ec{y} \\
+%\Val \ec{x} = \Val \ec{y}\interp{u} \text{ for some $\ec{x}$, for all non-closed subterms } u \text{ of } \Val \ec{y},  \\
+%\Val \ec{x} = v\interp{u} \text{ for some } \ec{x}, \text{ for all non-closed } u \in \mathcal{T}( \Val ) \\
+%\Val \ec{x} \interp{\mu v_{\ec{y}}. t} = \Val \ec{y} \\
+\label{eqn:m-bij} 
+\Val \ec{x} \sim \Val \ec{y} \text{ if and only if } \ec{x} = \ec{y}
 \end{eqnarray}
-
-\begin{comment}
-We assume that:
-\begin{enumerate}
-\item Each equivalence class of $\Fc$ contains at least one variable,
-\item Each equivalence class $\ec{x}$ of $\Fc$ where $x : \tau$, $\tau \in \Nondata$ contains a constant,
-\item Each equivalence class $\ec{x}$ of $\Fc$ where $x : \tau$, $\tau \in \Data \cup \Codata$ contains a constructor term.
-\end{enumerate}
-The first statement comes with no loss of generality since new equalities of the form $y \teq t$ for fresh variable $y$ can be added to $\Fc$ without affecting its satisfiability.
-The second statement can be assumed for $x : \tau$, since $\tau$ is infinite.
-Hence, we can assign constants for each such equivalence class.
-Moreover, since $\Fc$ contains $\thD$-constraints only, these constants can be chosen arbitrarily.
-The third statement can be assumed by \rem{TODO}.
-
-For each equivalence class $e : \tau$, where $\tau \in \Nondata$ containing constant $c$,
-$\Val(x) = c$ for each $x \in e$.
-For each $e : \tau$, where $\tau \in \Data$ containing the constructor term $C( \vec u )$ where the equivalence classes $\ec{\vec u}$ contains the variables $\vec y$,
-$\Val(x) = C( \Val( \vec y ) )$ for each $x \in e$.
-Since \rn{Cyclic} does not apply, this definition is well-defined. 
-For each $e : \tau$, where $\tau \in \Codata$, 
-$\Val( x ) = \interp{ \tpath{\Fc}{e} }$ for each $x \in e$.
-Since neither \rn{Cong} or \rn{Bisimilar} applies to $\Fc$, we have that for all variables $x$ and $y$:
-\begin{eqnarray} \label{eqn:m-bij-dt}
-\Val(x) = \Val(y) \text{ if and only if } \ec{x} = \ec{y}
-\end{eqnarray}
-\end{comment}
-
-Now, we let $\M( x ) = \Val \ec{x}$ for each variable $x$ in $\Fc$.
+Initially, since \rn{Bisimilar} does not apply to $\Fc$,
+we have that $\Val \ec{x} \not\vsim \Val \ec{y}$ for all distinct $\ec{x}$ and $\ec{y}$,
+and thus~\ref{eqn:m-bij} holds.
+%Moreover, since an equivalence class $\ec{u}$ exists for all subterms $u$ of all terms $t \in \mathcal{T}(\Fc)$,
+%we have that $t\interp{u} = \Val \ec{u}$ for all such terms $u$ and $t$. \rem{elaborate?}
+Now, we assume $\Val$ satisfies~\ref{eqn:m-bij},
+and show that $\Val \sigma$ does as well, where $\sigma := \{ v_{\ec{x}} \mapsto c_1 \}$ is a substitution applied by \rn{Complete}.
+By contradiction, assume that $\Val \sigma \ec{y} \sim \Val \sigma \ec{z}$ for some distinct $\ec{y}$ and $\ec{z}$.
+Then it must have been the case that $\Val \ec{y}$ and $\Val \ec{z}$ contain a position $p$
+where, w.l.o.g., say $\Val \ec{y} \mid_p$ is $v_{\ec{x}}$ and $\Val \ec{z} \mid_p$ is a closed $\mu$-term $c_2$ where $c_1 \sim c_2$.
+%Since $\Val$ satisfies invariant~\ref{eqn:m-bij},
+%we have that $\Val \ec{w} \sim c_1$ for some $\ec{w}$, 
+However, this contradicts the premise of \rn{Complete} that $c_1$ is not bisimilar to any $c_2 \in \mathcal{T}( \Val )$.
+Thus, by induction on the number of applications of \rn{Complete}, statement~\ref{eqn:m-bij} holds for $\Val^\ast$.
+
+%we have that $\Val$ satisfies both~\ref{eqn:m-normal} and~\ref{eqn:m-bij}.
+%It satisfies the former since if a $\mu$-term $\mu v_{\ec{x}}. t_1$ had a bisimilar subterm $\mu v_{\ec{y}}. t_2$,
+%then the rule \rn{Bisimilar} would have applied to $x$ and $y$. \rem{elaborate?}
+%For the inductive case, we assume $\Val$ satisfies~\ref{eqn:m-normal} and~\ref{eqn:m-bij},
+%and show that $\Val \sigma$ does as well, where $\sigma := \{ v_{\ec{x}} \mapsto c_1 \}$ is a substitution applied by \rn{Complete}.
+%By contradiction, assume that $\Val \sigma \ec{y}$ is a non-normal $\mu$-term for some $\ec{y}$.
+%Since $c_1$ is normal, we have $\ec{x} \neq \ec{y}$, 
+%and thus $\Val \ec{y}$ must have been of the form $\mu v_{\ec{y}}. t$ where $\ec{x} \in FV( t )$.
+%Let $u$ be the (strict) subterm of $\Val \ec{y}$ such that $(\Val \ec{y} \sigma ) \interp{u \sigma } \sim (\Val \ec{y} \sigma)$. 
+%Since $\Val$ maps only to normal $\mu$-terms, we have $x \in FV( u )$.
+
+Now, we let $\M( x )$ be the value in $\M$ that is a renaming of$\Val^\ast \ec{x}$ for each variable $x$ in $\mathcal{T}(\Fc)$,
+which are guarenteed to be distinct values due to statement~\ref{eqn:m-bij} and our construction of $\M$.
 We first show $\M$ satisfies all equalities in $t_1 \teq t_2 \in \Fc$, where $t_1$ and $t_2$ have type $\tau$.
@@
 
-If $\tau \in \Data$,
-we first show that $\M( t ) = \Val \ec{x}$ for all non-selector terms $t$ that occur in datatype equivalence classes $\ec{x}$.
-We show this by induction on the size of $\Val \ec{x}$.
-In the case where $t$ is a variable, then $\M( t ) = \Val \ec{ t } = \Val \ec{ x }$.
+If $\tau \in \Data \cup \Codata$,
+we first show that $\M( t ) = \Val^\ast \ec{x}$ for all non-selector terms $t$ that occur in equivalence classes $\ec{x}$.
+We show this by induction on the size of $\Val^\ast \ec{x}$.
+In the case where $t$ is a variable, then $\M( t ) = \Val^\ast \ec{ t } = \Val^\ast \ec{ x }$.
 In the case where $t$ is a constructor term $C( \vec u )$, 
-then $\Val \ec{x}$ must have been assigned by \rn{Ind} for some constructor term $C'( \vec u' ) \in \ec{x}$.
+then $\Val^\ast \ec{x}$ must have been assigned by \rn{Ind} for some constructor term $C'( \vec u' ) \in \ec{x}$.
 Since \rn{Unify$_2$} does not apply to $\Fc$, we have that $C = C'$.
 Since \rn{Unify$_1$} does not apply to $\Fc$, we have that $\ec{\vec u} = \ec{\vec u'}$.
-By the induction hypothesis, we have $\M( \vec u ) = \M( \vec u' )$, and thus $\M( t ) = C( \M( \vec u ) ) = C'( \M( \vec u ' ) ) = \Val \ec{ x }$.
-
-We now show that $\M( t ) = \Val \ec{x}$ for all selector terms $s^j_k( y )$ that occur in datatype equivalence classes $\ec{x}$.
+By the induction hypothesis, we have $\M( \vec u ) = \M( \vec u' )$, and thus $\M( t ) = C( \M( \vec u ) ) = C'( \M( \vec u ' ) ) = \Val^\ast \ec{ x }$.
+
+We now show that $\M( t ) = \Val^\ast \ec{x}$ for all selector terms $s^j_k( y )$ that occur in datatype equivalence classes $\ec{x}$.
 Since \rn{Split} does not apply, 
@@
 By the previous paragraph, we have $\M( y ) = \M( C_i( s^1_i( y ), \ldots, s^{a_i}_i( y ) ) )$,
-which is $C_i( \Val \ec{ x_1 }, \ldots, \Val \ec{ x_{a_i} } )$.
-If $i = k$, then $\M( t ) = \M( s^j_k( y ) ) = s^j_k( \M( y ) ) = \Val \ec{x_j}$, which is $\Val \ec{ x }$ since $\ec{x_j} = \ec{s^j_i( y )} = \ec{x}$.
+which is $C_i( \Val^\ast \ec{ x_1 }, \ldots, \Val^\ast \ec{ x_{a_i} } )$.
+If $i = k$, then $\M( t ) = \M( s^j_k( y ) ) = s^j_k( \M( y ) ) = \Val^\ast \ec{x_j}$, which is $\Val^\ast \ec{ x }$ since $\ec{x_j} = \ec{s^j_i( y )} = \ec{x}$.
 If $i \neq k$, then $\thD$ does not constraint the value of $\M( t )$,
 and since $s^j_k( \M( y ) )$ is otherwise unconstrained in our model construction since \rn{Cong} does not apply, 
-we may interpret $s^j_k( \M( y ) )$ as $\Val \ec{ x }$. 
-Thus, for all datatype equivalence classes $\ec{x}$, we have $\M( t ) = \Val \ec{x}$ for each term $t \in \ec{x}$, which implies $\M \models t_1 \teq t_2$.
-
-If $\tau \in \Codata$, \rem{TODO}
+we may interpret $s^j_k( \M( y ) )$ as $\Val^\ast \ec{ x }$. 
+Thus, for all datatype equivalence classes $\ec{x}$, we have $\M( t ) = \Val^\ast \ec{x}$ for each term $t \in \ec{x}$, which implies $\M \models t_1 \teq t_2$.
 

### 2014-12-11 tex/codatadp.tex
more tuning sect 2

--- 2014-12-11 tex/codatadp.tex
@@
 %\cite{wand-weidenbach-201x}.
-Closely related is the support for structural
+Closely related is the automatic structural
 induction in CVC4 by Reynolds and Kuncak \cite{reynolds-kuncak-2015} and in
@@
 The only predicate is equality ($\teq$) and belongs to the logical symbols.
-Other predicates can be represented as functions that return $\ty{bool}$.
-The metavariables $\delta,\:\varepsilon,\:\tau$ range over types, with $\delta$ reserved
-for datatypes and codatatypes.
+Other predicates can be represented as functions that return $\ty{bool}$,
+with $\const{p}(\ldots)$ abbreviating $\const{p}(\ldots) \teq \const{true}$.
+The metavariables $\delta,\:\varepsilon$ range over (co)datatypes,
+whereas $\tau,\:\upsilon$ range over arbitrary types.
 
 Function symbols are written in a sans-serif font (e.g., $\const{f}$, $\const{g}$) to
-distinguish them from variables (e.g., $x$, $y$). Names starting with an uppercase
+distinguish them from variables (e.g., $x$, $y$). Symbols starting with an uppercase
 letter (e.g.,
@@
 %
-Finally, the notation $\bar x$ abbreviates a list or tuple $x_1,\ldots,x_n$.
-% and $\bigwedge_i \varphi_i$ abbeviates a conjunction $\varphi_1 \mathrel\land \cdots \mathrel\land \varphi_n$.
+Finally, the notation $\bar x$ abbreviates a list or tuple $x_1,\ldots,x_n$
+and $\bigwedge_{\,i} \varphi_i$ abbeviates a conjunction
+$\varphi_1 \mathrel\land \cdots \mathrel\land \varphi_n$.
 
@@
 To every argument corresponds a selector. The names for the (co)data\-types, the
-constructors, and the selectors must be distinct and different from already
-introduced ones%
+constructors, and the selectors must be distinct and different from
+existing names%
 \dotReportFootnote{It can be convenient to specify the same selector
@@
 it is not necessary to extend the signature:
-$\const{d}_{i\negvthinspace j}(x)$ can be seen as an abbreviation for
+The discriminator term $\const{d}_{i\negvthinspace j}(x)$ can be seen as an abbreviation for
 $x = \const{C}_{i\negvthinspace j}\bigl(\const{s}_{i\negvthinspace j}^1(x), \ldots, \const{s}_{i\negvthinspace j}^{\,n_{\smash{{i\negvthinspace j}}}}(x)\bigr)$.
-This will simplify the presentation. % of the decision procedure.
+This will simplify the presentation later. % of the decision procedure.
 
@@
 admissible only if a ground constructor term can be exhibited that belongs to
-that datatype. This rules out ill-founded specifications such as
+the datatype. This rules out non-well-founded specifications such as
 \[\keyw{datatype}\;\, \ty{fstream} \,=\, \const{FSCons}(\ty{int},\:\, \ty{fstream})\]
-For codatatypes, no such check is necessary because there is always a term,
+For codatatypes, no admissibility check is necessary because there is always a term,
 finite or infinite, that witnesses nonemptiness.
@@
 \paragraph{Properties.}
-Datatypes and codatatypes share basic properties pertaining to constructors and
-selectors. All properties below are implicitly universally quantified and range
-over all $i$, $j$, $\jPrime$, and $k$ within bounds:
+Datatypes and codatatypes share many basic properties pertaining to
+constructors and selectors. All properties below are implicitly universally
+quantified and range over all $i$, $j$, $\jPrime$, and $k$ within bounds:
 % and to all possible splits of the
@@
 \text{Induction:}\quad
-\AXC{\strut$\bigwedge_{i,j} \forall x_1 \ldots \vthinspace x_{n_{i\negvthinspace j}}.\; \bigl(\bigwedge_{\,k\,} \mathit{IH}_{i\negvthinspace j}^k[x_k]\bigr) \longrightarrow P_i[C_{i\negvthinspace j}(x_1,\ldots,x_{n_{ij}})]$}
-\UIC{\strut$\bigwedge_{i} P_i[v_i]$}
+\AXC{\strut$\bigwedge_{\,i,j} \forall x_1 \ldots \vthinspace x_{n_{i\negvthinspace j}}.\; \bigl(\bigwedge_{\,k\,} \mathit{IH}_{i\negvthinspace j}^k[x_k]\bigr) \longrightarrow P_i[C_{i\negvthinspace j}(x_1,\ldots,x_{n_{ij}})]$}
+\UIC{\strut$\bigwedge_{\,i} P_i[v_i]$}
 \DP
@@
 where the induction hypothesis $\mathit{IH}_{i\negvthinspace j}^k(x)$
-denotes either $P_{\iPrime}(x)$ if there exists some $i$ such that this is type
-correct or else $\top$.
-%
-The axiom schema ensures that we get a smallest model. For example,
+denotes either $P_{\iPrime}(x)$ if there exists some $\iPrime$ such that
+the formula is type-correct or else $\top$.
+%
+The axiom schema ensures that the interpretation of datatypes
+is a smallest model. For example,
 for a datatype of natural numbers constructed from $\const{Z}$ and $\const{S}$,
-this rules out nonstandard models, which could contain cyclic values---e.g.,
+induction rules out nonstandard models, which could contain cyclic values---e.g.,
 an $n$ such that $n \teq \const{S}(n)$---or even infinite acyclic values
@@
 
-For codatatype, the dual notion is that of a coinduction principle that
-depends on witnesses $R$ that are required to be bisimulations:
+For codatatypes, the dual notion is called coinduction. It depends on
+witnesses $R$ that are required to be bisimulations:
 %
@@
 \text{Coinduction:}\quad
-\AXC{\strut$\begin{gathered}\textstyle \bigwedge_i R_i[v_i, w_i] \\[-\jot]\textstyle
-\bigwedge_i \forall v\; w.\;\, R_i[v, w] \longrightarrow
-  \bigwedge_j \const{d}_j(v) \teq \const{d}_j(w)
+\AXC{\strut$\begin{gathered}\textstyle \bigwedge_{\,i} R_i[v_i, w_i] \\[-\jot]\textstyle
+\bigwedge_{\,i} \forall v\; w.\;\, R_i[v, w] \longrightarrow
+  \bigwedge_{\,j} \const{d}_j(v) \teq \const{d}_j(w)
   \mathrel\land
-  \const{d}_j(v) \longrightarrow \bigwedge_k \const{s}_{i\negvthinspace j}^{\,k}(v) \sim \const{s}_{i\negvthinspace j}^{\,k}(w)
+  \const{d}_j(v) \longrightarrow \bigwedge_{\,k} \const{s}_{i\negvthinspace j}^{\,k}(v) \sim \const{s}_{i\negvthinspace j}^{\,k}(w)
 \end{gathered}$}
-\UIC{\strut$\bigwedge_i v_i \teq w_i$}
+\UIC{\strut$\bigwedge_{\,i} v_i \teq w_i$}
 \DP
@@
 \]
-where $x \sim y$ denotes either $R_i[x, y]$ if there exists some $i$ such that
-this is type correct or $x \teq y$ otherwise.
+where $x \sim y$ denotes either $R_{\iPrime}[x, y]$ if there exists some
+$\iPrime$ such that the formula is type-correct or $x \teq y$ otherwise.
 This axiom schema captures the idea that two values that yield the same
@@
 In addition, codatatypes are guaranteed to contain all values corresponding to
-infinite ground constructor terms. This cannot be capture by a first-order
-axiomatization, since there may be uncountably many of them. For example,
-$\ty{stream}_{\,\ty{int}}$ is isomorphic to the uncountable function space
-$\ty{nat} \to \ty{int}$.
+infinite ground constructor terms. In general, this cannot be captured by a
+first-order axiomatization, since there may be uncountably many of them. For
+example, $\ty{stream}_{\,\ty{int}}$ is isomorphic to the uncountable function
+space $\ty{nat} \to \ty{int}$.
 
@@
 \paragraph{Semantics.}
-Datatypes can be defined semantically as the initial model of an equational
-specification of the selector--constructor equations. This is the view followed
-by Barrett et al.\ \cite{barrett-et-al-2010}. A disadvantage with this approach
-is that it does not naturally account for selectors applied to wrong
-constructors. Barrett et al.\ address this by parameterizing the construction by
-default values, but these give rise to spurious equalities between unrelated
-terms---for example, $\const{s}_{11}(\const{C}_2) \teq \const{s}_{11}(\const{C}_3)$.
-This could be corrected, but the added
-complexity seems to suggest that selectors are better specified axiomatically.
-
-A related semantic view of datatypes is as an initial algebra. Codatatypes are
-then defined dually as a final coalgebra. The datatypes are generated by their
-constructors, whereas the codatatypes are viewed through their selectors
-\cite{xxx}. By uniformly focusing on the constructors, the axiomatic approach
-emphasizes the commonality between datatypes and codatatypes, while
-sacrificing a theoretically fruitful notion of duality.
+Datatypes can be defined semantically as the initial model of the
+selector--constructor equations. This is the view followed by Barrett et al.\
+\cite{barrett-et-al-2010}. A drawback of this approach is that it does
+not naturally account for selectors applied to wrong constructors. Barrett et
+al.\ address this by parameterizing the construction by default values, but
+these give rise to spurious equalities between unrelated terms---e.g.,
+$\const{s}_{11}(\const{C}_2) \teq \const{s}_{11}(\const{C}_3)$. This could be
+corrected, but the added complexity seems to suggest that selectors are better
+characterized axiomatically.
+
+A related semantic view of datatypes is as initial algebras \cite{xxx}.
+Codatatypes are then defined dually as final coalgebras \cite{xxx}. The
+datatypes are generated by their constructors, whereas the codatatypes are
+viewed through their selectors. By uniformly focusing on the constructors, the
+axiomatic approach emphasizes the commonality between datatypes and
+codatatypes, while sacrificing a theoretically fruitful notion of duality.
 
@@
 (co)recursive even if they are not actually (co)recursive; however, the
-semantic notion is more fruitful and is the one that interests us.
-
-Furthermore, nothing forbids specifications with no (co)recursion, such as
+semantic notion is more precise and is the one that interests us.
+
+It is worth observing that nothing forbids non(co)recursive specifications
+such~as
 \[\begin{aligned}[t]
-      \keyw{datatype}\;\, \ty{enat} & \,=\, \const{Nat}(\ty{nat}) \,\mid\, \const{Infty} \\[-.5\jot]
+      \keyw{datatype}\;\, \ty{ext\_nat} & \,=\, \const{Nat}(\ty{nat}) \,\mid\, \const{Infty} \\[-.5\jot]
       \keyw{codatatype}\; \,\ty{complex} & \,=\, \const{Complex}(\const{re}{:}\; \ty{real},\; \const{im}{:}\;\ty{real})
@@
 \]
-Semantically, it makes no difference whether the types are introduced as
+At the semantic level, it makes no difference whether the types are introduced as
 datatypes or as codatatypes.
@@
 
-Finally, some codatatypes are so degenerate that they have infinite values
+Finally, some codatatypes are so degenerate as to have infinite values
 even though they are finite. The simplest example is
@@
       \keyw{codatatype}\;\, \ty{b} & \,=\, \const{B}(\ty{b},\: \ty{c},\: \ty{b},\: \ty{unit}) \\[-\jot]
-             \keyw{and}\;\, \ty{c} & \,=\, \const{C}(\ty{unit},\: \ty{c},\: \ty{b},\: \ty{c})
+             \keyw{and}\;\, \ty{c} & \,=\, \const{C}(\ty{a},\: \ty{unit},\: \ty{b},\: \ty{c})
 \end{aligned}
@@
 where \ty{unit} is a datatype with the single constructor $\const{Unity} :
-\ty{unit}$. We call such types \emph{corecursive singletons}.
-For the decision procedure, it will be important to detect those finite
-codatatypes. Sometimes, a type might be a corecursive singleton only in some
-models. For example, if \ty{unit} is an uninterpreted type, \ty{b} and \ty{c}
-will be singleton types precisely in the models that interpret \ty{unit} as a
-singleton.
+\ty{unit}$. We call such types \emph{corecursive singletons}. For the decision
+procedure, it will be important to detect these types. A type may also be a
+corecursive singleton only in some models. For example, if \ty{unit} is an
+uninterpreted type, \ty{b} and \ty{c} will be singleton types precisely in the
+models that interpret \ty{unit} as a singleton. Fortunately, given some
+cardinalities for the ordinary types, it is easy to characterize this
+degenerate case:
 
@@
 If the domain interpreting a corecursive codatatype is finite, it must be a
-singleton. Furthermore, such a type has necessarly a single constructor and
-all its arguments have types that are interpreted by singletons.
+singleton. Furthermore, such a codatatype has necessarly a single constructor,
+and all its arguments have types that are interpreted by singletons.
 \end{lemma}
@@
 $\const{C}(\const{D}(\const{C}(\const{C}(\ldots))))$---all of which correspond
-to distinct values. If the type has a noncorecursive constructor
-$\const{E}$, it is possible to create terms of arbitrary depths---e.g.,
-$\const{C}(\ldots(\const{C}(\const{E}))\ldots)$. In both cases, there can be
-no finite models.
+to distinct values (by distinctness and injectivity). If the type has a
+noncorecursive constructor $\const{E}$, it is possible to create terms of
+arbitrary depths---e.g., $\const{C}(\ldots(\const{C}(\const{E}))\ldots)$. In
+both cases, there can be no finite models.
 
@@
 again excludes finite models. Otherwise, the coinduction principle ensures
-that there can be only one value.
+that the type has only one value.
 \qed

### 2014-12-11 tex/codatadp.tex
more

@@
 even though they are finite. The simplest example is
-\[\keyw{datatype}\;\, \ty{a} \,=\, \const{A}(\ty{a})\]
+\[\keyw{codatatype}\;\, \ty{a} \,=\, \const{A}(\ty{a})\]
 This is a finite codatatype, of cardinality one, whose single value is
-infinite: $\mu\vthinspace a.\; A(a)$. Other specimens are
+infinite: \,$\mu\vthinspace a.\; A(a)$. Other specimens are
 \[\begin{aligned}[t]

### 2014-12-11 tex/codatadp.tex
More work, will revise.

--- 2014-12-11 tex/codatadp.tex
@@
   \quad
-  c \not\sim t \sigma \text{ for all } \sigma, t \in \vrange(\Val)
+  c \not\sim (\mu y. t) \sigma \text{ for all } \sigma, (\mu y. t) \in \mathcal{T}(\Val)
 }{
@@
 
-%We write $\mathcal{T}(\Val)$ to denote the smallest set containing $v\interp{u}$ for each subterm $u$ of $v$ in the range of $\Val$.
-We write $\vrange(\Val)$ to denote the set of $\mu$-terms mapped to by $\Val$.
+We write $\mathcal{T}(\Val)$ to denote the smallest set containing $\nf{(v\interp{u})}$ for each non-variable subterm $u$ of $v$ in the range of $\Val$.
+%We write $\vrange(\Val)$ to denote the set of $\mu$-terms mapped to by $\Val$.
 Since this rule removes one variable from the set $FV( \Val )$ and does not add any additional variables to $FV( \Val )$, 
@@
 then \rn{Complete} would apply for the (single) value $c$ of type $\tau$, 
-for which the second premise holds since $\ec{x}$ is the only equivalence class of type $\tau$.
+for which the third premise holds since $\ec{x}$ is the only equivalence class of type $\tau$.
 Since all $\tau \in \Nondata$ are infinite, all variables in $FV( \Val^\ast )$ have infinite type.
@@
 there exists a closed normal $\mu$-term $c_1$ that is not bisimilar to any instance of a term in $t \in \mathcal{T}(\Val^\ast)$.
-To argue this is the case, notice that either $\tau$ is a recursive type, or contains an infinite subfield of type $\upsilon$.
-In the former case, we choose a value $c_1$ whose depth is greater than any term in $\vrange(\Val)$.
-In the latter case, we choose a value $c_1$ containing a subterm of type $\upsilon$ not yet occurring in $\Val$.
+To argue this is the case, notice that either $\tau \in \Nondata$,
+$\tau$ is a recursive type, or contains an infinite subfield of type $\upsilon$.
+In the first case, we may choose an arbitrary constant for $c_1$ not yet occurring in $\vrange(\Val)$.
+In the second case, we choose a value $c_1$ whose depth is greater than any term in $\vrange(\Val)$.
+In the third case, we choose a value $c_1$ containing a subterm of type $\upsilon$ that is not bisimilar to any instance of a term $\mathcal{T}(\Val^\ast)$, 
+which is guarenteed to exist by applying this argument inductively.
 %since $\mathcal{T}(\Val^\ast)$ is a finite set. \rem{elaborate} % (noting that a closed $\mu$-term is only bisimilar to other closed $\mu$-terms).
@@
 and consequently $\Val^\ast$ is a mapping from the equivalence classes of $\Fc$ to closed $\mu$-terms.
-We now argue the following invariant is maintained by applications of \rn{Complete}:% for all $\ec{x}, \ec{y}$:
+We further argue that $\Val^\ast$ has the following properties:
+%the following invariants are maintained by applications of \rn{Complete}:% for all $\ec{x}, \ec{y}$:
 \begin{eqnarray}
-%\label{eqn:m-normal} 
-%\Val \text{ maps to normal $\mu$-terms only } \\
+\label{eqn:m-normal} 
+\Val^\ast \text{ maps to normal $\mu$-terms only } \\
 %\text{If } \Val \ec{x} \text{ is not closed and has subterm $u$, then } \Val \ec{x} \interp{ u } = \Val \ec{y} \text{ for some } \ec{y} \\
@@
 \label{eqn:m-bij} 
-\Val \ec{x} \sim \Val \ec{y} \text{ if and only if } \ec{x} = \ec{y}
+\Val^\ast \ec{x} \sim \Val^\ast \ec{y} \text{ if and only if } \ec{x} = \ec{y}
 \end{eqnarray}
-Initially, since \rn{Bisimilar} does not apply to $\Fc$,
+We show this by induction on the number of applications of \rn{Complete}.
+We additionally show that applications of \rn{Complete} maintain the invariant 
+for all subterms $\mu v_{\ec{x}}. t$ occurring in the range of $\Val$ for some $\ec{y}$:
+%\begin{eqnarray}
+%\label{eqn:m-inv}
+%\Val \ec{y} \interp{ \mu v_{\ec{x}}. t } = \Val \ec{x}
+%\end{eqnarray}
+For the base case, since \rn{Bisimilar} does not apply to $\Fc$,
 we have that $\Val \ec{x} \not\vsim \Val \ec{y}$ for all distinct $\ec{x}$ and $\ec{y}$,
-and thus~\ref{eqn:m-bij} holds.
+and thus~\ref{eqn:m-bij} holds for $\Val$.
+\rem{todo 1}
+%Due to our construction of $\Val$ in step 2 of our calculus, 
+%statement~\ref{eqn:m-inv} holds, which together with~\ref{eqn:m-bij} implies~\ref{eqn:m-normal}.
 %Moreover, since an equivalence class $\ec{u}$ exists for all subterms $u$ of all terms $t \in \mathcal{T}(\Fc)$,
 %we have that $t\interp{u} = \Val \ec{u}$ for all such terms $u$ and $t$. \rem{elaborate?}
-Now, we assume $\Val$ satisfies~\ref{eqn:m-bij},
-and show that $\Val \sigma$ does as well, where $\sigma := \{ v_{\ec{x}} \mapsto c_1 \}$ is a substitution applied by \rn{Complete}.
-By contradiction, assume that $\Val \sigma \ec{y} \sim \Val \sigma \ec{z}$ for some distinct $\ec{y}$ and $\ec{z}$.
-Then it must have been the case that $\Val \ec{y}$ and $\Val \ec{z}$ contain a position $p$
-where, w.l.o.g., say $\Val \ec{y} \mid_p$ is $v_{\ec{x}}$ and $\Val \ec{z} \mid_p$ is a closed $\mu$-term $c_2$ where $c_1 \sim c_2$.
-%Since $\Val$ satisfies invariant~\ref{eqn:m-bij},
-%we have that $\Val \ec{w} \sim c_1$ for some $\ec{w}$, 
-However, this contradicts the premise of \rn{Complete} that $c_1$ is not bisimilar to any $c_2 \in \mathcal{T}( \Val )$.
-Thus, by induction on the number of applications of \rn{Complete}, statement~\ref{eqn:m-bij} holds for $\Val^\ast$.
+For the inductive case, we assume an arbitrary $\Val_i$ satisfies~\ref{eqn:m-normal} and~\ref{eqn:m-bij},
+and show that $\Val_i \sigma_x$ does as well, where $\sigma_x := \{ v_{\ec{x}} \mapsto c_1 \}$ is a substitution applied by \rn{Complete}.
+By contradiction, assume that $\Val_i \sigma_x \ec{y} \sim \Val_i \sigma_x \ec{z}$ for some distinct $\ec{y}$ and $\ec{z}$.
+Then it must have been the case that $\Val_i \ec{y}$ and $\Val_i \ec{z}$ have a position $p$
+where, w.l.o.g., say $\Val_i \ec{y} \mid_p$ is $v_{\ec{x}}$ and $\Val_i \ec{z} \mid_p$ is a term of the form $\mu v_{\ec{w}}. t$ 
+and $c_1 \sim (\mu v_{\ec{w}}. t) \sigma_x$.
+%Since $\Val_i$ satisfies invariant~\ref{eqn:m-bij},
+%we have that $\Val_i \ec{w} \sim c_1$ for some $\ec{w}$, 
+However, this contradicts the premise of \rn{Complete} that $c_1$ is not bisimilar to any $t \sigma \in \mathcal{T}( \Val_i )$ for any $\sigma$.
+%Thus, by induction on the number of applications of \rn{Complete}, statements~\ref{eqn:m-normal} and~\ref{eqn:m-bij} hold for $\Val^\ast$.
+Thus, statements~\ref{eqn:m-normal} and~\ref{eqn:m-bij} hold for $\Val^\ast$.
 
@@
 
-Now, we let $\M( x )$ be the value in $\M$ that is a renaming of$\Val^\ast \ec{x}$ for each variable $x$ in $\mathcal{T}(\Fc)$,
-which are guarenteed to be distinct values due to statement~\ref{eqn:m-bij} and our construction of $\M$.
+Now, let $\M( x )$ be the value in $\M$ that is a renaming of$\Val^\ast \ec{x}$ for each variable $x$ in $\mathcal{T}(\Fc)$,
+each of which are guarenteed to exist in $\M$ due to statement~\ref{eqn:m-normal} and our construction of $\M$,
+and moreover are guarenteed to be distinct in $\M$ due to statement~\ref{eqn:m-bij}.
 We first show $\M$ satisfies all equalities in $t_1 \teq t_2 \in \Fc$, where $t_1$ and $t_2$ have type $\tau$.
@@
 
+\rem{revise for co-datatype}
 If $\tau \in \Data \cup \Codata$,

### 2014-12-11 tex/codatadp.tex
dumped email

--- 2014-12-11 tex/codatadp.tex
@@
 Finally, the notation $\bar x$ abbreviates a list or tuple $x_1,\ldots,x_n$
-and $\bigwedge_{\,i} \varphi_i$ abbeviates a conjunction
+and $\bigwedge_{\,i}\, \varphi_i$ abbeviates a conjunction
 $\varphi_1 \mathrel\land \cdots \mathrel\land \varphi_n$.
@@
 \mathrel{\uplus} \Sel$, where $\Ctr$ are the \emph{constructors} and $\Sel$ are the
-\emph{selectors}; there is no need to consider other function symbols because they can
-be abstracted away as variables when combining theories.
-%
+\emph{selectors}; there is no need to consider further function symbols
+because they can be abstracted away as variables when combining theories.
+
+\paragraph{Specifications.}
 In an SMT problem, the signature is normally given by specifying the
@@
 finally any other function symbols.
-
-\paragraph{Specifications.}
+%
 A (co)datatype specification consists of $l$~mutually recursive types which are
@@
 $x = \const{C}_{i\negvthinspace j}\bigl(\const{s}_{i\negvthinspace j}^1(x), \ldots, \const{s}_{i\negvthinspace j}^{\,n_{\smash{{i\negvthinspace j}}}}(x)\bigr)$.
-This will simplify the presentation later. % of the decision procedure.
+%This will simplify the presentation. % of the decision procedure.
 
@@
       \keyw{and}\;\, \ty{forest}_{\,\tau} & \,=\, \const{FNil} \,\mid\, \const{FCons}(\ty{tree}_{\,\tau},\:\, \ty{forest}_{\,\tau}) \\%[-.5\jot]
-      \keyw{codatatype}\; \,\ty{llist}_{\,\tau} & \,=\, \const{LNil} \,\mid\, \const{LCons}(\const{lhead}{:}\; \tau,\; \const{ltail}{:}\;\ty{llist}_{\,\tau})
+      \keyw{codatatype}\; \,\ty{llist}_{\,\tau} & \,=\, \const{LNil} \,\mid\, \const{LCons}(%\const{lhead}{:}\;
+      \tau,\: %\const{ltail}{:}\;
+      \ty{llist}_{\,\tau})
 \end{aligned}
@@
 %
-Expressed in the algebraic jargon, exhaustiveness ensures that ``no junk''
-exists, whereas distinctness and injectivity ensure that ``no confusion''
-arises. The result of selectors applied to the wrong constructor is left
-completely unspecified.
+Expressed in the algebraic jargon, exhaustiveness helps ensure that ``no
+junk'' exists, whereas distinctness and injectivity guarantee that ``no
+confusion'' can arise. The result of selectors applied to the wrong
+constructor is left completely unspecified.
 
@@
 \text{Induction:}\quad
-\AXC{\strut$\bigwedge_{\,i,j} \forall x_1 \ldots \vthinspace x_{n_{i\negvthinspace j}}.\; \bigl(\bigwedge_{\,k\,} \mathit{IH}_{i\negvthinspace j}^k[x_k]\bigr) \longrightarrow P_i[C_{i\negvthinspace j}(x_1,\ldots,x_{n_{ij}})]$}
-\UIC{\strut$\bigwedge_{\,i} P_i[v_i]$}
+\AXC{\strut$\bigwedge_{\,i,j}\, \forall x_1 \ldots \vthinspace x_{n_{i\negvthinspace j}}.\; \bigl(\bigwedge_{\,k}\, \mathit{IH}_{i\negvthinspace j}^k[x_k]\bigr) \longrightarrow P_i[C_{i\negvthinspace j}(x_1,\ldots,x_{n_{ij}})]$}
+\UIC{\strut$\bigwedge_{\,i}\, P_i[v_i]$}
 \DP
@@
 The axiom schema ensures that the interpretation of datatypes
-is a smallest model. For example,
+is a standard model. For example,
 for a datatype of natural numbers constructed from $\const{Z}$ and $\const{S}$,
@@
 For codatatypes, the dual notion is called coinduction. It depends on
-witnesses $R$ that are required to be bisimulations:
+witnesses $R_i$ that are required to be bisimulations:
 %
@@
 \begin{aligned}[t]
-\text{Coinduction:}\quad
-\AXC{\strut$\begin{gathered}\textstyle \bigwedge_{\,i} R_i[v_i, w_i] \\[-\jot]\textstyle
-\bigwedge_{\,i} \forall v\; w.\;\, R_i[v, w] \longrightarrow
-  \bigwedge_{\,j} \const{d}_j(v) \teq \const{d}_j(w)
+\text{Coinduction:}\kern.8em %%% TYPESETTING: should be \quad
+\AXC{\strut$\begin{gathered}\textstyle \bigwedge_{\,i}\, R_i[v_i, w_i] \\[-\jot]\textstyle
+\bigwedge_{\,i}\, \forall v\; w.\;\, R_i[v, w] \longrightarrow
+  \bigwedge_{\,j}\, \const{d}_j(v) \teq \const{d}_j(w)
   \mathrel\land
-  \const{d}_j(v) \longrightarrow \bigwedge_{\,k} \const{s}_{i\negvthinspace j}^{\,k}(v) \sim \const{s}_{i\negvthinspace j}^{\,k}(w)
+  \const{d}_j(v) \longrightarrow \bigwedge_{\,k}\, \const{s}_{i\negvthinspace j}^{\,k}(v) \sim \const{s}_{i\negvthinspace j}^{\,k}(w)
 \end{gathered}$}
-\UIC{\strut$\bigwedge_{\,i} v_i \teq w_i$}
+\UIC{\strut$\bigwedge_{\,i}\, v_i \teq w_i$}
 \DP
@@
 al.\ address this by parameterizing the construction by default values, but
-these give rise to spurious equalities between unrelated terms---e.g.,
-$\const{s}_{11}(\const{C}_2) \teq \const{s}_{11}(\const{C}_3)$. This could be
-corrected, but the added complexity seems to suggest that selectors are better
-characterized axiomatically.
+this gives rise to spurious equalities between unrelated terms---e.g.,
+$\const{s}_{11}(\const{C}_2) \teq \const{s}_{11}(\const{C}_3)$. This flaw
+could be corrected, but the added complexity seems to suggest that selectors
+are better characterized axiomatically.
 
@@
 where \ty{unit} is a datatype with the single constructor $\const{Unity} :
-\ty{unit}$. We call such types \emph{corecursive singletons}. For the decision
-procedure, it will be important to detect these types. A type may also be a
-corecursive singleton only in some models. For example, if \ty{unit} is an
-uninterpreted type, \ty{b} and \ty{c} will be singleton types precisely in the
-models that interpret \ty{unit} as a singleton. Fortunately, given some
-cardinalities for the ordinary types, it is easy to characterize this
-degenerate case:
+\ty{unit}$, as well as $\ty{stream}_{\,\ty{unit}\,}$. We call such types
+\emph{corecursive singletons}. For the decision procedure, it will be
+important to detect these types, even if they occur rarely in practice. A type
+may also be a corecursive singleton only in some models. For example, if
+\ty{unit} is an uninterpreted type, \ty{b} and \ty{c} will be singleton types
+precisely in the models that interpret \ty{unit} as a singleton. Fortunately,
+given some cardinalities for the ordinary types, it is easy to characterize
+this degenerate case:
 
@@
 \begin{lemma}
-If the domain interpreting a corecursive codatatype is finite, it must be a
-singleton. Furthermore, such a codatatype has necessarly a single constructor,
-and all its arguments have types that are interpreted by singletons.
+Let $\delta$ be a corecursive codatatype. If the domain interpreting $\delta$
+is finite, it must be a singleton. Furthermore, $\delta$ has necessarily a
+single constructor, whose arguments have types that are interpreted by
+singletons.
 \end{lemma}
@@
 
+  * to evaluate procedure, we generated problems from Isabelle theories using
+    Sledgehammer
+
+  * chose a selection of theory files:
+    * pretty much every theory file from Isabelle distribution and AFP that
+      defines a codatatype
+    * also included two private theories by Andreas Lochbihler and Peter Gammie
+    * (codatatypes being a recent addition to Isabelle \cite{itp},
+       they are not yet widely used)
+    * and included a number of theories on datatypes
+    * aimed at a good mixture between linear structures (lists, streams) and
+      tree structures
+    * the theory files were chosen before running the experiments
+
+  * for each goal, select 256 lemmas as axioms, before monomorphization
+    (after monomorphization, typically more formulas due to duplicates),
+    and encode with the goal
+
+  * then ran CVC4 for XXX seconds on the StarExec cluster \cite{xxx}
+
+{\footnotesize
+\begin{verbatim}
+For the evaluation of Section 5, I'm playing with various ideas. I certainly
+would like to keep it short and not waste too much of our time on it. We
+discussed earlier having (1) hand-crafted problems and (2) actual benchmarks
+(like those you already have). I suggest we drop (1); I never did hand-crafted
+before and this always looks suspicious anyway. For (2), I see three or four
+possible benchmark sets:
+
+1. without (co)datatype theories
+2. with datatypes
+(3. with codatatypes)
+4. with both
+
+The benchmark problems would be identical except for the declarations at the
+top ("declare-sort" and "declare-fun" vs. "declare-(co)datatype").
+Irrespective of whether interpreted (co)datatypes are used, Sledgehammer will
+typically include some (co)datatype axioms (i.e. tautologies if (co)datatypes
+are enabled). Hence the tests are a good indication of whether native
+(co)datatypes are useful to Sledgehammer. I assume the results will follow
+common sense, i.e. 4 > {3, 2} > 1
+
+As for the actual Isabelle theory files to use, for codatatypes there is not
+much question: We just use all I could get my hands on. But perhaps we should
+have a second benchmark set, ideally something a bit standard like, say, the
+Judgment Day (JD) suite. Or I can make some arbitrary a priori (before
+evaluation) decisions, but that always looks a bit fishy unless it can be
+justified easily. The trouble is, we have about 1000 datatypes lying around in
+Isabelle and the Archive of Formal Proofs (AFP). These would have no
+codatatypes in them, so we could only do 1 and 2 of the evaluation. We could
+at least try the "list" library, which is extremely developed and would have
+1600 goals already, and perhaps a few really basic types ("option", "string"),
+or perhaps red-black trees (about 450 goals).
+
+So we can show the success rates, in a table like this:
+
+                 Codatatype-oriented suite      Datatype (or JD) suite
+                      (2345 goals)                   (1240 goals)
+
+   w/o (co)data        1255                            699
+   w data              1268                            712
+   w codata            1270
+   w both              1273
+
+I can write the usual warning that an increase of a couple of percentage
+points in success rate in a general tool like Sledgehammer and general
+benchmark suites means a lot etc. Putting the datatype evaluations (i.e. 2 and
+4) in there offers the advantage that it will give some measure of how much to
+expect from the codatatypes. I.e. if the success rate doesn't go up by 50% for
+datatypes, nobody will expect this to happen for codatatypes. A priori, I
+would expect a similar gain, but I'm looking forward to see the experimental
+data.
+
+In addition to success rate (or number of solved goals), we should have other
+statistics, like how often do we get proofs that contain the cyclicity and
+bisimilarity rules. And finally, we look in more detail at one or two such
+proofs, by presenting it and explaining it (and, before that, understanding
+it).
+\end{verbatim}}
+
 \section{Conclusion}
@@
 This work is part of a larger program that aims at enriching automatic
-provers with more high-level features and reducing the gap between automatic and
+provers with high-level features and at reducing the gap between automatic and
 interactive theorem proving. As future work, we want to try out CVC4 and its

### 2014-12-11 tex/codatadp.tex
experimental stuff

@@
 
+  * how many problems
+  * how many problems have some codatatypes, how many have datatypes (including
+    XXX that have both)
+
   * for each goal, select 256 lemmas as axioms, before monomorphization
@@
 
+  * wanted to evaluate both the impact of datatypes and codatatypes
+  * datatypes are known to be very useful on SMT-LIB, e.g. blah
+  * but we wanted to see how much they make a difference in Isabelle
+
+  * ran all problems four times, to try all four combinations of datatypes and
+    codatatypes enabled or disabled
+
+  * lemmas/axioms often include basic properties of constructor and selectors,
+    which are tautologies and ignored if (co)datatypes are enabled
+  * allows a fair comparison: how much do (co)datatype support brings as
+    opposed to just letting Sledgehammer do what it would normally do
+
+  * experience with such features is not extremely high -- e.g. arithmetic,
+    a most useful theory on lots of benchmarks, increases Sledgehammer's
+    success rate by 2 percentage points, or 4\%, with Z3 in earlier work \cite{xxx}
+
+  * so when analyzing statistics of this kind, with highly varied problems,
+    we must not overestimate the importance of a single trick
+  * on the other hand: every percentage point counts! cite Tom Hales
+
 {\footnotesize
 \begin{verbatim}
-For the evaluation of Section 5, I'm playing with various ideas. I certainly
-would like to keep it short and not waste too much of our time on it. We
-discussed earlier having (1) hand-crafted problems and (2) actual benchmarks
-(like those you already have). I suggest we drop (1); I never did hand-crafted
-before and this always looks suspicious anyway. For (2), I see three or four
-possible benchmark sets:
-
-1. without (co)datatype theories
-2. with datatypes
-(3. with codatatypes)
-4. with both
-
-The benchmark problems would be identical except for the declarations at the
-top ("declare-sort" and "declare-fun" vs. "declare-(co)datatype").
-Irrespective of whether interpreted (co)datatypes are used, Sledgehammer will
-typically include some (co)datatype axioms (i.e. tautologies if (co)datatypes
-are enabled). Hence the tests are a good indication of whether native
-(co)datatypes are useful to Sledgehammer. I assume the results will follow
-common sense, i.e. 4 > {3, 2} > 1
-
-As for the actual Isabelle theory files to use, for codatatypes there is not
-much question: We just use all I could get my hands on. But perhaps we should
-have a second benchmark set, ideally something a bit standard like, say, the
-Judgment Day (JD) suite. Or I can make some arbitrary a priori (before
-evaluation) decisions, but that always looks a bit fishy unless it can be
-justified easily. The trouble is, we have about 1000 datatypes lying around in
-Isabelle and the Archive of Formal Proofs (AFP). These would have no
-codatatypes in them, so we could only do 1 and 2 of the evaluation. We could
-at least try the "list" library, which is extremely developed and would have
-1600 goals already, and perhaps a few really basic types ("option", "string"),
-or perhaps red-black trees (about 450 goals).
-
 So we can show the success rates, in a table like this:
 
-                 Codatatype-oriented suite      Datatype (or JD) suite
-                      (2345 goals)                   (1240 goals)
-
-   w/o (co)data        1255                            699
-   w data              1268                            712
-   w codata            1270
+                 Datatype suite     Mixed suite        Codatatype suite
+                (2345 problems)    (XXX problems)       (1240 problems)
+
+   w/o (co)data        800           1255                699
+   w data              815           1268                
+   w codata                          1270                712
    w both                            1273
-
-I can write the usual warning that an increase of a couple of percentage
-points in success rate in a general tool like Sledgehammer and general
-benchmark suites means a lot etc. Putting the datatype evaluations (i.e. 2 and
-4) in there offers the advantage that it will give some measure of how much to
-expect from the codatatypes. I.e. if the success rate doesn't go up by 50% for
-datatypes, nobody will expect this to happen for codatatypes. A priori, I
-would expect a similar gain, but I'm looking forward to see the experimental
-data.
-
-In addition to success rate (or number of solved goals), we should have other
-statistics, like how often do we get proofs that contain the cyclicity and
-bisimilarity rules. And finally, we look in more detail at one or two such
-proofs, by presenting it and explaining it (and, before that, understanding
-it).
 \end{verbatim}}
+
+* statistics, like how often do we get proofs that contain the cyclicity and
+  bisimilarity rules
+
+* and finally, we look in more detail at one or two such proofs, by presenting
+  it and explaining it (and, before that, understanding it)
 

### 2014-12-11 tex/codatadp.tex
more

@@
 \[\begin{aligned}[t]
-      \keyw{datatype}\;\, \ty{tree}_{\,\tau} & \,=\, \const{Node}(\tau,\: \ty{forest}_{\,\tau}) \\[-\jot]
-      \keyw{and}\;\, \ty{forest}_{\,\tau} & \,=\, \const{FNil} \,\mid\, \const{FCons}(\ty{tree}_{\,\tau},\:\, \ty{forest}_{\,\tau}) \\%[-.5\jot]
       \keyw{codatatype}\; \,\ty{llist}_{\,\tau} & \,=\, \const{LNil} \,\mid\, \const{LCons}(%\const{lhead}{:}\;
       \tau,\: %\const{ltail}{:}\;
-      \ty{llist}_{\,\tau})
+      \ty{llist}_{\,\tau}) \\%[-.5\jot]
+      \keyw{datatype}\;\, \ty{tree}_{\,\tau} & \,=\, \const{Node}(\tau,\:\, \ty{forest}_{\,\tau}) \\[-\jot]
+      \keyw{and}\;\,\, \ty{forest}_{\,\tau} & \,=\, \const{FNil} \,\mid\, \const{FCons}(\ty{tree}_{\,\tau},\:\, \ty{forest}_{\,\tau})
 \end{aligned}
@@
 the datatype. This rules out non-well-founded specifications such as
-\[\keyw{datatype}\;\, \ty{fstream} \,=\, \const{FSCons}(\ty{int},\:\, \ty{fstream})\]
+\[\keyw{datatype}\;\, \ty{fstream}_{\,\tau} \,=\, \const{FSCons}(\tau,\:\, \ty{fstream}_{\,\tau})\]
 For codatatypes, no admissibility check is necessary because there is always a term,
@@
 
-\begin{lemma}
-Let $\delta$ be a corecursive codatatype. If the domain interpreting $\delta$
-is finite, it must be a singleton. Furthermore, $\delta$ has necessarily a
+\begin{lemma}[Corecursive Singletons]%
+\label{lem:corecursive-singletons}%
+Let $\delta$ be a corecursive codatatype. The domain interpreting $\delta$ is
+either infinite or a singleton. In the latter case, $\delta$ has necessarily a
 single constructor, whose arguments have types that are interpreted by

### 2014-12-11 tex/codatadp.tex

@@
 
-We construct a model $\M$ of $\thD$ that interprets constructors and correctly-applied selectors in the obvious way,
-interprets wrongly-applied selectors based on equality information in $\Fc$,
-interprets the variables and $x$ of $\Fc$ based on a mapping $\Val$ from the equivalence classes of $\Fc$ to values.
-Initially, $\Val$ contains $\ec{x} \mapsto v_{\ec{x}}$ for each equivalence class $\ec{x}$ of $\Fc$,
-where $v_{\ec{x}}$ is a fresh variable of the same type as $x$ not occurring in $\Fc$.
-Then, $\Val$ is updated according to the following rules:
-
-\begin{comment}
-\rn{Ord}
+We construct a model $\M$ of $\thD$ that interprets each type $\tau$
+as a maximal set of closed normal $\mu$-terms of that type that are 
+unique up to renaming of $\mu$-bound variables,
+and acyclic when $\tau \not\in \Codata$.
+The interpretation of constructor terms
+$\M( C( \vec t ) )$ for $C \in \Ctr$ is the value in $\M$ that is a renaming of $\nf{(\mu x. C( \M( \vec t ) ))}$ where $x \not\in FV( \vec t )$.
+The interpretation of (correctly-applied) selector terms
+$\M( s^j_i( t ) )$ for $s^j_i \in \Sel$ is the value in $\M$ that is a renaming of $\M( t )\interp{u_j}$
+where $\M( t )$ is of the form $\mu x. C_i( \vec u )$ for some $\vec u$.
+For instance, if $\M( t ) = \mu x. \const{C}_1( x )$ (call this term $v$), 
+then $\M( s_1^1( t ) )$ is a renaming of
+$v\interp{x} =$ 
+$v\interp{\mu x. \const{C}_1( x )}^\emptyset =$
+$\mu x. \const{C}_1( v\interp{x}^{\{x\}} ) =$
+$\mu x. \const{C}_1( x )$.
+%To interpret a correctly-applied selector in $\M$, if $v = \mu x. C_i( \vec t )$, then $\M( s^j_i( v ) )$ is the value in $\M( \tau )$ that is a renaming of $v\interp{t_j}$,
+%where $t_j$ has type $\tau$.
+%We assume $\M$ interprets constructors and correctly-applied selectors in the obvious way
+%\footnote{
+%The interpretation of $\M( C( \vec t ) )$ is the value in $\M( \tau )$ that is a renaming of $\nf{\mu x. C( \M( \vec t ) ) }$.
+%To interpret a correctly-applied selector, if $v = \mu x. C_i( \vec t )$, then $\M( s^j_i( v ) )$ is the value in $\M( \tau )$ that is a renaming of $v\interp{t_j}$,
+%where $t_j$ has type $\tau$.
+%},
+Additionally, $\M$ will interpret wrongly-applied selectors based on equality information in $\Fc$,
+and will interpret the variables of $\Fc$ based on the valuation $\Val$ which was constructed in step two of our calculus.
+
+Recall that $\Val$ is a mapping from the equivalence classes of $\Fc$ to $\mu$-terms.
+We first \emph{complete} $\Val$ so that it maps equivalence classes only to closed $\mu$-terms, that is, terms containing no free variables.
+This is done by exhaustively applying the following rule:
+
+\rn{Complete}
 \(
 \inferrule{
-  \ec{x} \not\in \vdom(\Val)
-  \quad
-  x : \tau
-  \quad
-  \tau \in \Nondata
-  \quad
-  c \not\in \vrange(\Val)
-}{
-  \Val := \Val \cup \ec{x} \mapsto c
-}
-\)
-
-\rn{D-Ind}
-\(
-\inferrule{
-  \ec{x} \not\in \vdom(\Val)
-  \quad
-  x : \tau
-  \quad
-  \tau \in \Data
-  \quad
-  C( \vec t ) \in \ec{x}
-  \quad
-  \ec{\vec t} \in \vdom(\Val)
-}{
-  \Val := \Val \cup \ec{x} \mapsto C( \Val \ec{\vec t}  )
-}
-\)
-
-\rn{D-Base}
-\(
-\inferrule{
-  \ec{x} \not\in \vdom(\Val)
-  \quad
-  x : \tau
-  \quad
-  \tau \in \Data
-  \quad
-  \ec{x} \text{ ctr-free}
-  \quad
-  c \not\in \vrange(\Val)
-}{
-  \Val := \Val \cup \ec{x} \mapsto c
-}
-\)
-\end{comment}
-
-\rn{Ind}
-\(
-\inferrule{
-  %x : \tau
-  %\quad
-  %\tau \in \Codata
   v_{\ec{x}} \in FV( \Val ) 
   \quad
-  C( t_1, \ldots, t_n ) \in \ec{x}
+  c \in \M
   \quad
-  C \in \Ctr
-}{
-  \Val := \Val \{ v_{\ec{x}} \mapsto \mu\vthinspace v_{\ec{x}}. C( v_{\ec{t_1}}, \ldots, v_{\ec{t_n}} ) \}
-}
-\)
-
-\rn{Base}
-\(
-\inferrule{
-  %x : \tau
-  %\quad
-  %\tau \in \Codata
-  %\quad
-  v_{\ec{x}} \in FV( \Val )
-  %\quad
-  %\ec{x} \text{ ctr-free}
-  \quad
-  c \not\sim c' \text{ for all } c' \in \vrange(\Val)
+  c \not\sim (\mu y. t) \sigma \text{ for all } \sigma, (\mu y. t) \in \mathcal{T}(\Val)
 }{
@@
 
-\rem{define $\sim$}
-We assume that first \rn{Ind} is applied exhaustively, after which \rn{Base} is applied exhaustively.
-%the rule \rn{D-Ind} is applied to completion before applying \rn{D-Base}.
-In the rule \rn{Ind}, we write $FV( \Val )$ to denote the free variables occurring in $\Val$.
-Notice that this excludes variables beneath $\mu$-binders.
-
-As an example,
-say that $\Fc$ contains equivalence classes $\ec{x}$ and $\ec{y}$ where $\const{C}(z,x) \in \ec{y}$ and $\const{C}(w,y) \in \ec{x}$, where $z$ and $w$ are variables of integer type.
-A possible application of the above rules, showing the state of $\Val$ after each application, is as follows:
-\[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
-\hline
-\text{Rule} & \multicolumn{4}{c}{\Val} \ 
-\\
-\hline
-& \{ & \ec{x} \mapsto v_{\ec{x}}, & \ec{z} \mapsto v_{\ec{z}}, \\
-&    & \ec{y} \mapsto v_{\ec{y}}, & \ec{w} \mapsto v_{\ec{w}} & \} \\
-\rn{Ind} \ec{x}
-& \{ & \ec{x} \mapsto \mu\vthinspace v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ), & \ec{z} \mapsto v_{\ec{z}}, \\
-&    & \ec{y} \mapsto v_{\ec{y}}, & \ec{w} \mapsto v_{\ec{w}} & \} \\
-\rn{Ind} \ec{y}
-& \{ & \ec{x} \mapsto \mu\vthinspace v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu\vthinspace v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ) ), & \ec{z} \mapsto v_{\ec{z}}, \\
-&    & \ec{y} \mapsto \mu\vthinspace v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
-\rn{Ind} \ec{x}
-& \{ & \ec{x} \mapsto \mu\vthinspace v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu\vthinspace v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ) ), & \ec{z} \mapsto v_{\ec{z}}, \\
-&    & \ec{y} \mapsto \mu\vthinspace v_{\ec{y}}. \const{C}( v_{\ec{z}}, \mu\vthinspace v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ) ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
-\rn{Base} \ec{z}
-& \{ & \ec{x} \mapsto \mu\vthinspace v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu\vthinspace v_{\ec{y}}. \const{C}( \const{0}, v_{\ec{x}} ) ), & \ec{z} \mapsto \const{0}, \\
-&    & \ec{y} \mapsto \mu\vthinspace v_{\ec{y}}. \const{C}( \const{0}, \mu\vthinspace v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ) ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
-\rn{Base} \ec{w}
-& \{ & \ec{x} \mapsto \mu\vthinspace v_{\ec{x}}. \const{C}( \const{1}, \mu\vthinspace v_{\ec{y}}. \const{C}( \const{0}, v_{\ec{x}} ) ), & \ec{z} \mapsto \const{0}, \\
-&    & \ec{y} \mapsto \mu\vthinspace v_{\ec{y}}. \const{C}( \const{0}, \mu\vthinspace v_{\ec{x}}. \const{C}( \const{1}, v_{\ec{y}} ) ), & \ec{w} \mapsto \const{1} & \} \\
-\end{array}\]
-
-It is clear that after applying these rules to completion, $\Val$ is a mapping from equivalence classes to ground values.
-
-\begin{eqnarray} \label{eqn:m-bij-dt}
-\Val \ec{x} = \Val \ec{y} \text{ if and only if } \ec{x} = \ec{y}
+We write $\mathcal{T}(\Val)$ to denote the smallest set containing $\nf{(v\interp{u})}$ for each non-variable subterm $u$ of $v$ in the range of $\Val$.
+%We write $\vrange(\Val)$ to denote the set of $\mu$-terms mapped to by $\Val$.
+Since this rule removes one variable from the set $FV( \Val )$ and does not add any additional variables to $FV( \Val )$, 
+it can only be applied a finite number of times.
+
+Let $\Val^\ast$ be the result of applying the above rule to completion on the mapping $\Val$ we constructed during step 2 of our calculus.
+We first argue that $FV( \Val^\ast ) = \emptyset$.
+Since \rn{Split} does not apply to $\Fc$,
+all equivalence classes of $\Fc$ finite datatype type contain constructor terms, 
+and thus $FV( \Val )$ contains no variables of finite datatype type.
+For each equivalence classes of $\Fc$ with finite co-datatype $\tau$, 
+we know $\tau$ must have cardinality one, due to \rem{reference}.
+Since \rn{Singleton} does not apply to $\Fc$, 
+there is only one equivalence class of type $\tau$, call it $\ec{x}$.
+%It cannot be the case that $v_{\ec{x}} \in FV(\Val^\ast \ec{x})$, then we may apply \rn{Complete} for the (single) value $c$ of type $\tau$, 
+%for which the second premise holds since $\ec{x}$ is the only equivalence class of type $\tau$.
+If it were the case that $v_{\ec{x}} \in FV(\Val^\ast )$,
+then \rn{Complete} would apply for the (single) value $c$ of type $\tau$, 
+for which the third premise holds since $\ec{x}$ is the only equivalence class of type $\tau$.
+Since all $\tau \in \Nondata$ are infinite, all variables in $FV( \Val^\ast )$ have infinite type.
+For any infinite type $\tau$,
+there exists a closed normal $\mu$-term $c_1$ that is not bisimilar to any instance of a term in $t \in \mathcal{T}(\Val^\ast)$.
+To argue this is the case, notice that either $\tau \in \Nondata$,
+$\tau$ is a recursive type, or contains an infinite subfield of type $\upsilon$.
+In the first case, we may choose an arbitrary constant for $c_1$ not yet occurring in $\vrange(\Val)$.
+In the second case, we choose a value $c_1$ whose depth is greater than any term in $\vrange(\Val)$.
+In the third case, we choose a value $c_1$ containing a subterm of type $\upsilon$ that is not bisimilar to any instance of a term $\mathcal{T}(\Val^\ast)$, 
+which is guarenteed to exist by applying this argument inductively.
+%since $\mathcal{T}(\Val^\ast)$ is a finite set. \rem{elaborate} % (noting that a closed $\mu$-term is only bisimilar to other closed $\mu$-terms).
+Thus, $FV( \Val^\ast )$ contains no variables of infinite type,
+and consequently $\Val^\ast$ is a mapping from the equivalence classes of $\Fc$ to closed $\mu$-terms.
+We further argue that $\Val^\ast$ has the following properties:
+%the following invariants are maintained by applications of \rn{Complete}:% for all $\ec{x}, \ec{y}$:
+\begin{eqnarray}
+\label{eqn:m-normal} 
+\Val^\ast \text{ maps to normal $\mu$-terms only } \\
+%\text{If } \Val \ec{x} \text{ is not closed and has subterm $u$, then } \Val \ec{x} \interp{ u } = \Val \ec{y} \text{ for some } \ec{y} \\
+%\Val \ec{x} = \Val \ec{y}\interp{u} \text{ for some $\ec{x}$, for all non-closed subterms } u \text{ of } \Val \ec{y},  \\
+%\Val \ec{x} = v\interp{u} \text{ for some } \ec{x}, \text{ for all non-closed } u \in \mathcal{T}( \Val ) \\
+%\Val \ec{x} \interp{\mu v_{\ec{y}}. t} = \Val \ec{y} \\
+\label{eqn:m-bij} 
+\Val^\ast \ec{x} \sim \Val^\ast \ec{y} \text{ if and only if } \ec{x} = \ec{y}
 \end{eqnarray}
-
-\begin{comment}
-We assume that:
-\begin{enumerate}
-\item Each equivalence class of $\Fc$ contains at least one variable,
-\item Each equivalence class $\ec{x}$ of $\Fc$ where $x : \tau$, $\tau \in \Nondata$ contains a constant,
-\item Each equivalence class $\ec{x}$ of $\Fc$ where $x : \tau$, $\tau \in \Data \cup \Codata$ contains a constructor term.
-\end{enumerate}
-The first statement comes with no loss of generality since new equalities of the form $y \teq t$ for fresh variable $y$ can be added to $\Fc$ without affecting its satisfiability.
-The second statement can be assumed for $x : \tau$, since $\tau$ is infinite.
-Hence, we can assign constants for each such equivalence class.
-Moreover, since $\Fc$ contains $\thD$-constraints only, these constants can be chosen arbitrarily.
-The third statement can be assumed by \rem{TODO}.
-
-For each equivalence class $e : \tau$, where $\tau \in \Nondata$ containing constant $c$,
-$\Val(x) = c$ for each $x \in e$.
-For each $e : \tau$, where $\tau \in \Data$ containing the constructor term $C( \vec u )$ where the equivalence classes $\ec{\vec u}$ contains the variables $\vec y$,
-$\Val(x) = C( \Val( \vec y ) )$ for each $x \in e$.
-Since \rn{Cyclic} does not apply, this definition is well-defined. 
-For each $e : \tau$, where $\tau \in \Codata$, 
-$\Val( x ) = \interp{ \tpath{\Fc}{e} }$ for each $x \in e$.
-Since neither \rn{Cong} or \rn{Bisimilar} applies to $\Fc$, we have that for all variables $x$ and $y$:
-\begin{eqnarray} \label{eqn:m-bij-dt}
-\Val(x) = \Val(y) \text{ if and only if } \ec{x} = \ec{y}
-\end{eqnarray}
-\end{comment}
-
-Now, we let $\M( x ) = \Val \ec{x}$ for each variable $x$ in $\Fc$.
+We show this by induction on the number of applications of \rn{Complete}.
+We additionally show that applications of \rn{Complete} maintain the invariant 
+for all subterms $\mu v_{\ec{x}}. t$ occurring in the range of $\Val$ for some $\ec{y}$:
+%\begin{eqnarray}
+%\label{eqn:m-inv}
+%\Val \ec{y} \interp{ \mu v_{\ec{x}}. t } = \Val \ec{x}
+%\end{eqnarray}
+For the base case, since \rn{Bisimilar} does not apply to $\Fc$,
+we have that $\Val \ec{x} \not\vsim \Val \ec{y}$ for all distinct $\ec{x}$ and $\ec{y}$,
+and thus~\ref{eqn:m-bij} holds for $\Val$.
+\rem{todo 1}
+%Due to our construction of $\Val$ in step 2 of our calculus, 
+%statement~\ref{eqn:m-inv} holds, which together with~\ref{eqn:m-bij} implies~\ref{eqn:m-normal}.
+%Moreover, since an equivalence class $\ec{u}$ exists for all subterms $u$ of all terms $t \in \mathcal{T}(\Fc)$,
+%we have that $t\interp{u} = \Val \ec{u}$ for all such terms $u$ and $t$. \rem{elaborate?}
+For the inductive case, we assume an arbitrary $\Val_i$ satisfies~\ref{eqn:m-normal} and~\ref{eqn:m-bij},
+and show that $\Val_i \sigma_x$ does as well, where $\sigma_x := \{ v_{\ec{x}} \mapsto c_1 \}$ is a substitution applied by \rn{Complete}.
+By contradiction, assume that $\Val_i \sigma_x \ec{y} \sim \Val_i \sigma_x \ec{z}$ for some distinct $\ec{y}$ and $\ec{z}$.
+Then it must have been the case that $\Val_i \ec{y}$ and $\Val_i \ec{z}$ have a position $p$
+where, w.l.o.g., say $\Val_i \ec{y} \mid_p$ is $v_{\ec{x}}$ and $\Val_i \ec{z} \mid_p$ is a term of the form $\mu v_{\ec{w}}. t$ 
+and $c_1 \sim (\mu v_{\ec{w}}. t) \sigma_x$.
+%Since $\Val_i$ satisfies invariant~\ref{eqn:m-bij},
+%we have that $\Val_i \ec{w} \sim c_1$ for some $\ec{w}$, 
+However, this contradicts the premise of \rn{Complete} that $c_1$ is not bisimilar to any $t \sigma \in \mathcal{T}( \Val_i )$ for any $\sigma$.
+%Thus, by induction on the number of applications of \rn{Complete}, statements~\ref{eqn:m-normal} and~\ref{eqn:m-bij} hold for $\Val^\ast$.
+Thus, statements~\ref{eqn:m-normal} and~\ref{eqn:m-bij} hold for $\Val^\ast$.
+
+%we have that $\Val$ satisfies both~\ref{eqn:m-normal} and~\ref{eqn:m-bij}.
+%It satisfies the former since if a $\mu$-term $\mu v_{\ec{x}}. t_1$ had a bisimilar subterm $\mu v_{\ec{y}}. t_2$,
+%then the rule \rn{Bisimilar} would have applied to $x$ and $y$. \rem{elaborate?}
+%For the inductive case, we assume $\Val$ satisfies~\ref{eqn:m-normal} and~\ref{eqn:m-bij},
+%and show that $\Val \sigma$ does as well, where $\sigma := \{ v_{\ec{x}} \mapsto c_1 \}$ is a substitution applied by \rn{Complete}.
+%By contradiction, assume that $\Val \sigma \ec{y}$ is a non-normal $\mu$-term for some $\ec{y}$.
+%Since $c_1$ is normal, we have $\ec{x} \neq \ec{y}$, 
+%and thus $\Val \ec{y}$ must have been of the form $\mu v_{\ec{y}}. t$ where $\ec{x} \in FV( t )$.
+%Let $u$ be the (strict) subterm of $\Val \ec{y}$ such that $(\Val \ec{y} \sigma ) \interp{u \sigma } \sim (\Val \ec{y} \sigma)$. 
+%Since $\Val$ maps only to normal $\mu$-terms, we have $x \in FV( u )$.
+
+Now, let $\M( x )$ be the value in $\M$ that is a renaming of$\Val^\ast \ec{x}$ for each variable $x$ in $\mathcal{T}(\Fc)$,
+each of which are guarenteed to exist in $\M$ due to statement~\ref{eqn:m-normal} and our construction of $\M$,
+and moreover are guarenteed to be distinct in $\M$ due to statement~\ref{eqn:m-bij}.
 We first show $\M$ satisfies all equalities in $t_1 \teq t_2 \in \Fc$, where $t_1$ and $t_2$ have type $\tau$.
@@
 
-If $\tau \in \Data$,
-we first show that $\M( t ) = \Val \ec{x}$ for all non-selector terms $t$ that occur in datatype equivalence classes $\ec{x}$.
-We show this by induction on the size of $\Val \ec{x}$.
-In the case where $t$ is a variable, then $\M( t ) = \Val \ec{ t } = \Val \ec{ x }$.
+\rem{revise for co-datatype}
+If $\tau \in \Data \cup \Codata$,
+we first show that $\M( t ) = \Val^\ast \ec{x}$ for all non-selector terms $t$ that occur in equivalence classes $\ec{x}$.
+We show this by induction on the size of $\Val^\ast \ec{x}$.
+In the case where $t$ is a variable, then $\M( t ) = \Val^\ast \ec{ t } = \Val^\ast \ec{ x }$.
 In the case where $t$ is a constructor term $C( \vec u )$, 
-then $\Val \ec{x}$ must have been assigned by \rn{Ind} for some constructor term $C'( \vec u' ) \in \ec{x}$.
+then $\Val^\ast \ec{x}$ must have been assigned by \rn{Ind} for some constructor term $C'( \vec u' ) \in \ec{x}$.
 Since \rn{Unify$_2$} does not apply to $\Fc$, we have that $C = C'$.
 Since \rn{Unify$_1$} does not apply to $\Fc$, we have that $\ec{\vec u} = \ec{\vec u'}$.
-By the induction hypothesis, we have $\M( \vec u ) = \M( \vec u' )$, and thus $\M( t ) = C( \M( \vec u ) ) = C'( \M( \vec u ' ) ) = \Val \ec{ x }$.
-
-We now show that $\M( t ) = \Val \ec{x}$ for all selector terms $s^j_k( y )$ that occur in datatype equivalence classes $\ec{x}$.
+By the induction hypothesis, we have $\M( \vec u ) = \M( \vec u' )$, and thus $\M( t ) = C( \M( \vec u ) ) = C'( \M( \vec u ' ) ) = \Val^\ast \ec{ x }$.
+
+We now show that $\M( t ) = \Val^\ast \ec{x}$ for all selector terms $s^j_k( y )$ that occur in datatype equivalence classes $\ec{x}$.
 Since \rn{Split} does not apply, 

### 2014-12-12 tex/codatadp.tex
Minor updates.

@@
 
-\section{The Calculus}
+\section{A Ground Decision Procedure for (Co)-Datatypes}
 \label{sec:the-calculus}
@@
   * SMT background
-
-\subsection{The Ground Case}
 
@@
 A rule may have multiple conclusions separated by $\parallel$, which denotes a non-deterministic branching.
-An application of a rule is \emph{redundant} if it is non-terminal and at least one branch in its conclusion does not add a new equality to $\Ec$.
 
@@
 %For presentation of the rules in this step, we rely on a representation of terms in the $\mu$-notation.
-The rules of this step contain premises that involve terms containing $\mu$-bindings.
-Recall that these bindings can be used for representing (possibly cyclic) values,
+The rules of this step contain premises that involve a mapping $\Val$ from equivalence classes to terms containing $\mu$-bindings.
+Recall that $\mu$-bindings can be used for representing (possibly cyclic) terms and values,
 for instance, $\mu x. \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
@@
 For consistency, we will also consider $\mu x. \const{c}$ to be a $\mu$-term
-in the case where $c$ is a constant of type $\tau \in \Nondata$.
+in the case where $\const{c}$ is a constant of type $\tau \in \Nondata$.
 We consider only $\mu$-terms that are \emph{well-typed}, which we define recursively:
@@
 Intuitively, a $\mu$-term $t$ of type $\tau$ describes a class of values of $\tau$, 
-either in the case when $\tau$ is a co-datatype or datatype type,
-where in the latter case, such a $t$ containing a bound variable describes an infeasible class of values.
+%either in the case when $\tau$ is a co-datatype or datatype type,
+where a cyclic term $t$ describes an infeasible class of values in the case that $\tau \not\in \Codata$.
 At the beginning of this step, 
@@
 that describe the class of values that terms in that equivalence class can take in the models of $\Ec$. 
-For each equivalence class $\ec{x}$ of $\Ec$, we associate a fresh variable $v_{\ec{x}}$ of the same type as $x$ not occurring in $\tEc$,
-and initially say that $\Val$ contains $\ec{x} \mapsto v_{\ec{x}}$ for each equivalence class $\ec{x}$ of $\Ec$ 
-(in other words, there are initially no constraints on the values for any equivalence class).
+For each equivalence class $\ec{x}$, we associate a fresh variable $v_{\ec{x}}$ of the same type as $x$ not occurring in $\tEc$,
+and initially say that $\Val$ contains $\ec{x} \mapsto v_{\ec{x}}$,
+in other words, there are initially no constraints on the values for any equivalence class $\ec{x}$.
 Then, $\Val$ is updated by applying the following rule exhaustively:
@@
 and thus our calculus may infer $x \teq \const{C}(y,x)$.
-Intuitively, this equality holds, since the values of $x$ and $\const{C}( y, x )$ are observationally equivalent for any value of $v_{\ec{y}}$.
+Intuitively, this equality holds, since the values of $x$ and $\const{C}( y, x )$ 
+are equivalent for any value of $v_{\ec{y}}$.
 \qed
@@
 
+\
+
+An application of a rule is \emph{redundant} if it is non-terminal and at least one branch in its conclusion does not add a new equality to $\Ec$.
 A \emph{derivation tree} is a tree whose nodes are sets of equalities, where non-root nodes are obtained by 
@@
 and let $S^\ast_t$ be the fixed point of this sequence.
-We know this is a finite set for each $t$ since the values of $t$'s type are of finite size.
+We know this is a finite set for each $t$ since the values of the type of $t$ are of finite size.
 Let $S^\ast$ be the union of $S^\ast_t$ for all $t \in T$,
@@
 
-\subsection{Extension to Quantified Formulas}
+\section{Extension to Quantified Formulas}
 
@@
 
-\section{The Theory Solver}
+\section{Implementation}
 \label{sec:the-theory-solver}

### 2014-12-12 tex/codatadp.tex
Repair definition of normal mu-term.

@@
 First, we require additional terminology concerning $\mu$-terms.
-If $u$ is a subterm of $\mu$-term $v$, then the \emph{interpretation of $u$ in $v$} is the $\mu$-term $v\interp{u}^\emptyset$ as returned by the following recursive procedure,
-where $x \mapsto \mu x. C( \vec t ) \in v$ indicates that $\mu x. C( \vec t )$ is the subterm of $v$ that binds this occurrence of variable $x$.
+For a $\mu$-term $t_1$ with subterm $t_2$, 
+the \emph{interpretation of $t_2$ in $t_1$} is the $\mu$-term $t_1\interp{t_2}^\emptyset$ as returned by the following recursive procedure,
+where $x \mapsto \mu x. C( \vec u ) \in t_1$ indicates that $\mu x. C( \vec u )$ is the subterm of $t_1$ that binds this occurrence of variable $x$.
 \[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
-v\interp{x}^M & = & v\interp{\mu x. C( \vec t )}^M & x \not\in M, x \mapsto \mu x. C( \vec t ) \in v  \\
-v\interp{x}^M & = & x & \text{otherwise} \\
-v\interp{\mu x. C( \vec t )}^M & = & \mu x. C( v\interp{ \vec t }^{M \cup \{ x \}} ) & x \not\in M \\
-v\interp{\mu x. C( \vec t )}^M & = & x & \text{otherwise}
+t_1\interp{x}^M & = & \mu x. C( t_1\interp{ \vec u}^{M \cup \{ x \}}) & x \not\in M, x \mapsto \mu x. C( \vec u ) \in t_1  \\
+t_1\interp{x}^M & = & x & \text{otherwise} \\
+t_1\interp{\mu x. C( \vec u )}^M & = & \mu x. C( t_1\interp{ \vec u }^{M \cup \{ x \}} ) & x \not\in M \\
+t_1\interp{\mu x. C( \vec u )}^M & = & x & \text{otherwise}
 \end{array}\]
-We will write $v\interp{u}$ as shorthand for $v\interp{u}^\emptyset$.
-Given a $\mu$-term $v$ of the form $\mu x. C( \vec t )$, 
-we say $u$ is a \emph{bisimilar subterm} of $v$ if and only if $v\interp{u} \vsim v$.
-We say $v$ is \emph{normal} if and only if all of its bisimilar subterms are $x$,
+We will often write $t_1\interp{t_2}$ as shorthand for $t_1\interp{t_2}^\emptyset$.
+Given a $\mu$-term $t_1$ of the form $\mu x. C( u_1, \ldots, u_n )$ with subterm $t_2$,
+we say $t_2$ is a \emph{bisimilar subterm} of $t_1$ if and only if it is of the form
+$\mu y. C( v_1, \ldots, v_n )$ and $t_1\interp{ u_i } \sim t_1\interp{ v_i }$ for $i = 1,\ldots, n$.
+We say $t_1$ is \emph{normal} if and only if it does not contain a strict bisimilar subterm
 and all of its strict subterms are also normal.
-For instance, the $\mu$-term $v$ of the form $\mu x. \const{S}( \mu y. \const{S}( x ) )$ is not normal, 
-since 
-$v\interp{\mu y. \const{S}( x )}^\emptyset =$ 
-$\mu y. \const{S}( v\interp{x}^{\{y\}} ) =$ 
-$\mu y. \const{S}( v\interp{\mu x. \const{S}( \mu y. \const{S}( x ) ) }^{\{y\}} ) =$ 
-$\mu y. \const{S}( \mu x. \const{S}( v\interp{\mu y. \const{S}( x )}^{\{x,y\}} ) ) =$ 
-$\mu y. \const{S}( \mu x. \const{S}( y ) )$,
-which is syntactically equivalent to $v$ for the renaming $\{ x \mapsto y, y \mapsto x \}$.
+For instance, the $\mu$-term $t_1$ of the form $\mu x. \const{S}( \mu y. \const{S}( y ) )$ is not normal, 
+since $\mu y. \const{S}( y )$ is a bisimilar subterm of $t_1$.
+In particular, note that their arguments have the same interpretation in $t_1$, 
+that is,
+$t_1\interp{\mu y. \const{S}( y )}^\emptyset =$ 
+$\mu y. \const{S}( t_1\interp{y}^{\{y\}} ) =$ 
+$\mu y. \const{S}( y )$ 
+is syntactically equivalent to
+$t_1\interp{y}^\emptyset =$ 
+$\mu y. \const{S}( t_1\interp{y}^{\{y\}} ) =$ 
+$\mu y. \const{S}( y )$.
+The $\mu$-term $t_1$ of the form $\mu x. \const{S}( \mu y. \const{S}( x ) )$ is also not normal, 
+since $\mu y. \const{S}( x )$ is a bisimilar subterm of $t_1$,
+noting that
+$t_1\interp{\mu y. \const{S}( x )}^\emptyset =$ 
+$\mu y. \const{S}( t_1\interp{x}^{\{y\}} ) =$ 
+$\mu y. \const{S}( t_1\interp{\mu x. \const{S}( \mu y. \const{S}( x ) ) }^{\{y\}} ) =$ 
+$\mu y. \const{S}( \mu x. \const{S}( t_1\interp{\mu y. \const{S}( x )}^{\{x,y\}} ) ) =$ 
+$\mu y. \const{S}( \mu x. \const{S}( y ) )$
+is syntactically equivalent to $t_1\interp{x}^\emptyset = t_1$ for the renaming $\{ x \mapsto y, y \mapsto x \}$.
+
 For any $\mu$-value $v$ of the form $\mu x. C( \vec w )$, we may recursively construct a normal value $\nf{v}$

### 2014-12-12 tex/codatadp.tex
minor

--- 2014-12-11 tex/codatadp.tex
@@
 
-\excludeversion{report}
-\includeversion{paper}
-
 \def\thewordpaper{paper}
 \newcommand\dotReportFootnote[1]{.}
+
+\newcommand\afterDot{\;} %%% Too little space by default after "Lemma Foo."
 
@@
 \label{lem:corecursive-singletons}%
+\afterDot
 Let $\delta$ be a corecursive codatatype. The domain interpreting $\delta$ is
-either infinite or a singleton. In the latter case, $\delta$ has necessarily a
+either infinite or a singleton. In the latter case, $\delta$ necessarily has a
 single constructor, whose arguments have types that are interpreted by
@@
 
-\begin{lemma}[Termination] \label{lem:t}
+\begin{lemma}[Termination]%
+\label{lem:t}%
+\afterDot
 All derivation trees are finite.
@@
 
-\begin{lemma}[Refutation Soundness] \label{lem:rs}
+\begin{lemma}[Refutation Soundness]%
+\label{lem:rs}%
+\afterDot
 If there exists a closed derivation tree with root node $\Ec$, then $\Ec$ is unsatisfiable in $\thD$.
@@
 
-\begin{lemma}[Solution Soundness] \label{lem:ss}
+\begin{lemma}[Solution Soundness]%
+\label{lem:ss}%
+\afterDot
 If there exists a derivation tree with root node $\Ec$ containing a saturated node, then $\Ec$ is satisfiable in $\thD$.

### 2014-12-14 tex/codatadp.tex
Reorganize

@@
 \newcommand{\vsim}{\sim}
+\newcommand{\vsimv}[1]{\vsim_{#1}}
+\newcommand{\tpos}[2]{#1\!\mid_{#2}}
 
@@
 
-  * Rewriter
-  * Actual calculus
-
-  * Correctness
-
 \rem{Discuss signature $\Sig$, theory $\thD$}
@@
 A $\mu$-term is \emph{cyclic} if and only if it contains a bound variable.
-We write $t \vsim s$ if $\mu$-terms $t$ and $s$ are syntactically equivalent for some renaming of $\mu$-bound variables,
+We write $t \vsim s$ if $\mu$-terms $t$ and $s$ are syntactically equivalent for some capture-avoiding renaming of $\mu$-bound variables,
 in which case we will commonly say $t$ and $s$ are \emph{bisimilar}.
+For instance, 
+$\mu x. \const{C}( x ) \vsim \mu y. \const{C}( y )$ and
+$\mu x, \const{D}( y, x ) \vsim \mu z. \const{D}( y, z )$, but
+$\mu x. \const{D}( y, x ) \not\vsim \mu x. \const{C}( x )$,
+$\mu x. \const{D}( y, x ) \not\vsim \mu x. \const{D}( z, x )$ and
+$\mu x. \const{D}( y, x ) \not\vsim \mu y. \const{D}( y, x )$.
+Also, note that
+two $\mu$-terms may represent the same value while being not bisimilar, as in
+$\mu x. \const{C}( \mu y. \const{C}( x ) ) \not\vsim \mu x. \const{C}( x )$.
 %, or informally are \emph{observationally equivalent}.
@@
 %either in the case when $\tau$ is a co-datatype or datatype type,
-where a cyclic term $t$ describes an infeasible class of values in the case that $\tau \not\in \Codata$.
+where a cyclic $t$ describes an infeasible class of values when $\tau \not\in \Codata$.
 At the beginning of this step, 
@@
 in other words, there are initially no constraints on the values for any equivalence class $\ec{x}$.
-Then, $\Val$ is updated by applying the following rule exhaustively:
+Then, $\Val$ is modified by applying the following update exhaustively:
 
@@
   \quad
-  C( t_1, \ldots, t_n ) \in \ec{x}
+  \const{C}( t_1, \ldots, t_n ) \in \ec{x}
   \quad
-  C \in \Ctr
+  \const{C} \in \Ctr
 }{
-  \Val := \Val \{ v_{\ec{x}} \mapsto \mu v_{\ec{x}}. C( v_{\ec{t_1}}, \ldots, v_{\ec{t_n}} ) \}
+  \Val := \Val \{ v_{\ec{x}} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{t_1}}, \ldots, v_{\ec{t_n}} ) \}
 }
@@
 \begin{example}
-Say that $\Ec$ contains (distinct) equivalence classes $\ec{x}$, $\ec{y}$, $\ec{z}$ and $\ec{w}$ where $\const{C}(z,x) \in \ec{y}$ and $\const{C}(w,y) \in \ec{x}$.
-A possible sequence of applications of the above rule is shown below.
+Say that $\Ec$ contains (distinct) equivalence classes $\ec{x}$, $\ec{y}$, $\ec{z}$ and $\ec{w}$ 
+where $\const{C}(z,x) \in \ec{y}$ and $\const{C}(w,y) \in \ec{x}$ for some $\const{C} \in \Ctr$.
+A possible sequence of updates applied to $\Val$ is shown below.
 In the first column, we show the equivalence class that was considered,
@@
 \end{array}\]
-Informally, this indicates that the values for $x$ (resp. $y$) in models of $\Ec$
+This indicates that the values for $x$ (resp. $y$) in models of $\Ec$
 are of the form $\const{C}( v_{\ec{w}}, \const{C}( v_{\ec{z}}, \ldots ))$ (resp. $\const{C}( v_{\ec{z}}, \const{C}( v_{\ec{w}}, \ldots ))$) 
@@
 For bisimilarity, if there are two equivalence classes $\ec{t_1}$ and $\ec{t_2}$ 
-of co-datatype type whose classes of values $\Val \ec{t_1}$ and $\Val \ec{t_2}$ are equivalent modulo renaming of $\mu$-bound variables,
+of co-datatype type whose classes of values $\Val \ec{t_1}$ and $\Val \ec{t_2}$ are equivalent for a capture-avoiding renaming of $\mu$-bound variables,
 then we can conclude $t_1 \teq t_2$.
@@
 
-\begin{comment}
-To explain the rules in Figure~\ref{fig:ab-rules}, we introduce the following notions.
-First, an edge $\ec{t_1} \rightarrow^{C}_n \ec{t_2}$ is a transition between two equivalence classes $\ec{t_1}$ and $\ec{t_2}$ 
-(which we will call its \emph{source} and \emph{destination} respectively),
-labeled by a constructor $C$ and integer $n$.
-We say that edge $\ec{t_1} \rightarrow^{C}_n \ec{t_2}$ is \emph{induced by $\Ec$} if and only if $C( \vec u ) \in \ec{t_1}$ for some $\vec u$ and $\ec{t_2} = \ec{u_n}$.
-In other words, $t_1$ is equivalent to a $C$-application whose $n^{th}$ argument is equivalent to $t_2$.
-A \emph{path} $p$ is a non-empty ordered list of edges $e_1, \ldots e_n$ such that the destination of $e_i$ is the source of $e_{i+1}$ for $i = 1, \ldots (n-1)$,
-and the sources of $e_1, \ldots, e_n$ are distinct.
-We say that path $p$ is \emph{induced by $\Ec$} if and only if each of its edges are induced by $\Ec$.
-Notice that since $\Ec$ contains a finite set of equivalence classes, all paths induced by $\Ec$ have finite length.
-We call the \emph{source} of a path is the source of its first edge, and the \emph{destination} of a path is the destination of its last edge.
-We say a path $p$ is \emph{maximal} if its destination $\ec{t}$ is either
-the source of an edge in $p$, 
-or does not contain an application of a non-nullary constructor (including the case when $t$ is not of datatype/codatatype type).
-The set $\ttpath{\Ec}{\ec{t_1}}{\ec{t_2}}$ is the smallest set containing all maximal paths with source $\ec{t_1}$ and destination $\ec{t_2}$ that are induced by $\Ec$.
-The set $\tpath{\Ec}{\ec{t_1}}$ is the union of $\ttpath{\Ec}{\ec{t_1}}{\ec{t_2}}$ for all $t_2 \in \tEc$.
-We will write, for instance, $\ec{t_1} \rightarrow^{C_1}_{n_1} \ec{t_2} \rightarrow^{C_2}_{n_2} \ec{t_3}$ 
-as shorthand for the path containing an edge from $\ec{t_1}$ to $\ec{t_2}$, followed by one from $\ec{t_2}$ to $\ec{t_3}$.
-%A set of paths is \emph{disjoint} if it does not contain two paths $p_1$ and $p_2$ such that $p_1 = p_2 \sigma$ for some $\sigma$.
-
-The rule \rn{Cyclic} recognizes when a path exists from an equivalence class $\ec{t}$ to itself, as indicated
-by the assumption $\ttpath{\Ec}{\ec{t}}{\ec{t}} \neq \emptyset$.
-In this case, $\Ec$ entails that $t$ contains itself as a (strict) subterm,
-and thus if $t$ is a datatype, $\Ec$ must be unsatisfiable.
-The rule \rn{Bisimilar} recognizes when two equivalence classes $\ec{t_1}$ and $\ec{t_2}$ having co-datatype type are observationally equivalent.
-In this rule, $\sigma$ is a complete mapping over the equivalence classes that occur as sources in the edges of paths in $\tpath{\Ec}{\ec{t_2}}$.
-The rule applies if $\tpath{\Ec}{\ec{t_1}}$ is comprised of the paths $p \sigma$ for each (and only) the paths $p \in \tpath{\Ec}{\ec{t_2}}$,
-in which case $t_1$ must be equal to $t_2$.
-We demonstrate an application of this rule concretely with the following example.
-\end{comment}
-
 \begin{figure}[t]
@@
 \end{tabular}
-\caption{Rules for splitting.  
+\caption{Rules for branching.  
 %All pairs of terms whose type has cardinality 1 are entailed to be equal (the degenerate case).
@@
 
-\paragraph{Step 3: Branching}
+\paragraph{Step 3: Branching.}
 If neither a rule from Figure~\ref{fig:cc-rules} or Figure~\ref{fig:ab-rules} applies to $\Ec$, 
@@
 An application of a rule is \emph{redundant} if it is non-terminal and at least one branch in its conclusion does not add a new equality to $\Ec$.
-A \emph{derivation tree} is a tree whose nodes are sets of equalities, where non-root nodes are obtained by 
+A \emph{derivation tree} is a tree whose nodes are sets of equalities, where non-root nodes vtained by 
 a non-redundant application of a derivation rule to its parent node.
@@
 For \rn{Unify$_1$}, this is a consequence of injectivity.
-For \rn{Bisimilar}, our construction of $\Val$ indicates that the value of $t_1$ and $t_2$ are observationally equivalent in all models of $\Ec$.
+For \rn{Bisimilar}, our construction of $\Val$ indicates that the value of $t_1$ and $t_2$ are equivalent in all models of $\Ec$.
 For \rn{Singleton}, clearly $t_1$ and $t_2$ must have the same value since the cardinality of their type is one.
@@
 then $\Ec$ is satisfiable $\thD$.
-First, we require additional terminology concerning $\mu$-terms.
+We demonstrate a particular model $\M$ of $\thD$ that satisfies $\Ec$.
+Before doing so, we require additional terminology concerning $\mu$-terms.
+
 For a $\mu$-term $t_1$ with subterm $t_2$, 
@@
 We will often write $t_1\interp{t_2}$ as shorthand for $t_1\interp{t_2}^\emptyset$.
+We will write $\mathcal{T}( t_1 )$ to denote the set consisting of $t_1\interp{t_2}$ for all subterms $t_2$ of $t_1$,
+and write $\mathcal{T}( S )$ to denote the union of $\mathcal{T}( t )$ for each $t \in S$ where $S$ is a set of terms.
+
 Given a $\mu$-term $t_1$ of the form $\mu x. C( u_1, \ldots, u_n )$ with subterm $t_2$,
@@
 (which we will call the \emph{normal form} of $v$)
-that is a renaming of $v$ by replacing all of the bisimilar subvalues of $v$ with $x$,
+that is a renaming of $v$ by replacing all of the bisimilar subterms of $v$ with $x$,
 and by replacing all of the non-normal subterms of $v$ with their corresponding normal forms.
 For instance, $\nf{(\mu x. \const{S}( \mu y. \const{S}( x ) ))} = \mu x. \const{S}( x )$.
+
+\begin{definition}[Normal Model] \label{def:norm-model}
+A model $\M$ is \emph{normal} if and only if:
+\begin{enumerate}
+\item 
+for all types $\tau$, 
+$\M( \tau )$ is a maximal set of closed normal $\mu$-terms of that type that are 
+unique up to renaming of $\mu$-bound variables,
+and acyclic when $\tau \not\in \Codata$,
+\item 
+for all constructor terms $\const{C}( \vec t )$ of type $\tau$,
+$\M( \const{C}( \vec t ) )$ is the value 
+in $\M( \tau )$ that is a renaming of 
+$\nf{(\mu x. \const{C}( \M( \vec t ) ))}$ where $x \not\in FV( \vec t )$, and
+\item 
+for all selector terms $s^j_i( t )$ of type $\tau$ where $\M( t )$ is of the form $\mu x. \const{C}_i( \vec u )$.% for some $\vec u$,
+$\M( s^j_i( t ) )$ is the value
+in $\M( \tau )$ that is a renaming of 
+$\M( t )\interp{u_j}$.
+\end{enumerate}
+\end{definition}
+
+In the following, we consider only normal models.
+Thus, when constructing any model $\M$,
+it remains only to define how wrong-applied selector terms and the free variables are interpreted.
+For the latter, this will be based on the mapping $\Val$ constructed in step two of our calculus.
+We first state the following property of $\Val$, which amounts to saying that interpretation of an equivalence class is
+not sensisitive to the context in which it is interpretated.
+
+\begin{lemma}
+If $\Val$ is a mapping construced in step two of our calculus,
+then $t_1 \interp{ \mu v_{\ec{x}}. t_2 } = \Val \ec{x}$
+for all $t_1\interp{\mu v_{\ec{x}}. t_2} \in \mathcal{T}(\Val)$.
+\end{lemma}
+\begin{proof}
+\rem{TODO}
+\qed
+\end{proof}
+
+We say $t \vsimv{x} s$ if and only if $\mu$-terms $t$ and $s$ are syntactically equivalent
+for some renaming that avoids capturing any variable besides $x$.
+For instance, for $t:= \mu x. \const{C}( x, y )$ and $u := \mu y. \const{C}( x, y )$,
+we have $t \not\vsim u$, but it is the case that both
+$t \vsimv{x} u$ and $t \vsimv{y} u$.
+
+The \emph{completion} of $\Val$, which we will write as $\Val^\ast$, 
+is the result of exhaustively applying the following update to $\Val$:
+
+\(
+\inferrule{
+  x : \tau \in FV( \Val ) 
+  \quad
+  c \in \M( \tau )
+  \quad
+  c \not\sim_x \mu y. t \text{ for all } \mu y. t \in \mathcal{T}(\varna\Val)
+}{
+  \Val := \Val \{ x \mapsto c \}
+}
+\)
+
+\rem{revise from here on}
+
+\begin{lemma} \label{lem:normal-subterm-bisim}
+If $t_1 \vsim t_2$, then $t_1\interp{\tpos{t_1}{p}} \vsim t_2\interp{\tpos{t_2}{p}}$.
+\end{lemma}
+\begin{proof}
+\end{proof}
+
+\begin{lemma} \label{lem:normal-subs}
+If $t_1$ and $t_2$ are normal $\mu$-terms where $t_1 \not\vsim t_2$,
+and if $u_1$ is a normal $\mu$-term such that $u_1 \not\vsim u_2 \sigma$ for all $\sigma$, $u_2 \in \mathcal{T}( t_1 ) \cup \mathcal{T}( t_2 )$,
+then $\nf{(t_1 \sigma_x)} \not\vsim \nf{(t_2 \sigma_x)}$ for $\sigma_x := \{ x \mapsto u_1 \}$.
+\end{lemma}
+\begin{proof}
+By contradiction assume that 
+$t_1 \not\vsim t_2$ and 
+$\nf{(t_1 \sigma)} \vsim \nf{(t_2 \sigma)}$
+for some $\sigma := \{ x \mapsto u_1 \}$
+where $u_1 \not\vsim u_2 \sigma$ for all $\sigma$, $u_2 \in \mathcal{T}( t_1 ) \cup \mathcal{T}( t_2 )$ 
+and $t_1, t_2, u_1$ are normal.
+Then it must be the case that there is a position where, say $t_1$, 
+is such that $\tpos{t_1}{p} = x$ and $\tpos{t_2}{p} = \mu y. u$,
+and by Lemma~\ref{lem:normal-subterm-bisim}, where $\tpos{\nf{(t_1 \sigma)}}{p} \vsim \tpos{\nf{(t_2 \sigma)}}{p}$.
+
+%Since $\nf{(t_2 \interp{\mu y. u})} \in \mathcal{T}( t_2 )$.
+\end{proof}
 
@@
 and acyclic when $\tau \not\in \Codata$.
+
 The interpretation of constructor terms
@@
 This is done by exhaustively applying the following rule:
-
-\rn{Complete}
-\(
-\inferrule{
-  v_{\ec{x}} \in FV( \Val ) 
-  \quad
-  c \in \M
-  \quad
-  c \not\sim (\mu y. t) \sigma \text{ for all } \sigma, (\mu y. t) \in \mathcal{T}(\Val)
-}{
-  \Val := \Val \{ v_{\ec{x}} \mapsto c \}
-}
-\)
 

### 2014-12-14 tex/codatadp.tex
Fixes
minor

--- 2014-12-12 tex/codatadp.tex
@@
 infinit\-e constructor terms are also legitimate values for codatatypes
-(Section~\ref{sec:theory-of-co-datatypes}). Intuitively, the
+(Section~\ref{sec:the-theory-of-co-datatypes}). Intuitively, the
 values of a codatatype consist of all well-typed finite and infinite ground constructor

### 2014-12-14 tex/codatadp.tex

--- 2014-12-14 tex/codatadp.tex
@@
   \quad
-  c \not\sim_x \mu y. t \text{ for all } \mu y. t \in \mathcal{T}(\varna\Val)
+  c \not\sim_x t \text{ for all } t \in \mathcal{T}(\vrange(\Val))
 }{

### 2014-12-14 tex/codatadp.tex
edits here and there in Sects. 1 and 2

@@
 datatypes and codatatypes. The dual of the cyclicity rule for datatypes is a
-bisimilarity rule that identifies equal values, also in the presence of cycles.
-The procedure is complete for universal problems and is compatible with the
-Nelson--Oppen method. It has been implemented in the latest version of CVC4, a
-state-of-the-art SMT solver. An evaluation based on problems generated from
-Isabelle theories demonstrates the potential of the procedure.
+bisimilarity rule that identifies equal values, also in the presence of cyclic
+($\omega$-regular) data. The procedure is complete for ground problems and is
+composable via the Nelson--Oppen method. It has been implemented in the latest
+version of CVC4, a state-of-the-art SMT solver. An evaluation based on
+problems generated from Isabelle theories demonstrates the potential of the
+procedure.
 \end{abstract}
@@
 useful to represent finite data structures in computer science applications but
-also arise in mathematical formalizations.
+also arise when formalizing mathematics.
 They can be implemented efficiently and enjoy
@@
 To represent infinite values, % such as $\const{S}(\const{S}(\const{S}(\ldots))$,
-a natural choice is to turn to \emph{codatatypes} (or coinductive datatypes), the
-non-well-founded dual of (inductive) \relax{datatypes}.
+a natural choice is to turn to coinductive datatypes, or \emph{codatatypes},
+the non-well-founded dual of inductive \emph{datatypes}.
 %
@@
 processes.
-% as do the CoALP dialect of Prolog and the Dafny specification language.
 
@@
 values of a codatatype consist of all well-typed finite and infinite ground constructor
-terms, and only those. For example, the specification
+terms, and only those. For example, the coinductive specification
 %
@@
 
-Datatypes and codatatypes are an integral part of most popular proof assistants,
+Datatypes and codatatypes are an integral part of modern proof assistants,
 including Agda, Coq, Isabelle, Matita, and PVS. In recent years, some
@@
 In this \thewordpaper, we introduce a unified decision procedure for ground
-problems involving datatypes and codatatypes in combination and discuss extensions 
-of the procedure for handling universal (non-ground) problems.
-The procedure is
-described abstractly as a calculus (Section~\ref{sec:the-calculus}) and is
+problems involving datatypes and codatatypes in combination
+(Section~\ref{sec:a-ground-decision-procedure-for-co-datatypes})
+and discuss extensions  of the procedure for handling quantifiers
+(Section~\ref{sec:extension-to-quantified-formulas}).
+The procedure is described abstractly as a calculus and is
 designed to be composable via the Nelson--Oppen method \cite{nelson-oppen-1979}.
@@
 values internally. For example,
-the term $\const{SCons}(1,\: \mu\vthinspace s.\; \const{SCons}(0,\: \const{SCons}(9,\: s)))$
+the term $\const{SCons}(1,\: \mu s.\; \const{SCons}(0,\: \const{SCons}(9,\: s)))$
 denotes the lasso-shaped sequence $1, 0, 9, 0, 9, 0, 9, \ldots$\afterLdots.
@@
 (Section~\ref{sec:the-theory-solver}).
+\textbf{TODO: CHECK WITH IMPLEMENTATION, but round to nearest 50 or 1000.}
 It consists of about 2000 lines of \cpp{} code, among which 1600 are
@@
 %      * but theory solver cooperates
-%    * still, universal conjectures fall into the ground fragment via negation
-%      and skolemization (or, alternatively, herbrandization)
 
@@
 Although it was written several years ago, the account of related work in
-Barrett et al.\ \cite{barrett-et-al-2007} is still a good reference.
-Since then, datatypes have been added to Z3 in unpublished work by Leonardo de
-Moura and to a SPASS-like prototype prover called Pirate by Daniel Wand.
-%\cite{wand-weidenbach-201x}.
+Barrett et al.\ \cite{barrett-et-al-2007} is still a good starting point.
+Since then, datatypes have been added to the SMT solver
+Z3 in unpublished work by Leonardo de Moura and to a SPASS-like prototype
+superposition prover called Pirate by Daniel Wand.
 Closely related is the automatic structural
 induction in CVC4 by Reynolds and Kuncak \cite{reynolds-kuncak-2015} and in
-Pirate by Wand and Weidenbach \cite{wand-weidenbach-201x}, which depends on a
-notion of datatype.
+Pirate by Wand and Weidenbach \cite{wand-weidenbach-201x};
+both methods naturally depend on a notion of datatype.
 
@@
 is associated a list of argument types $\tau_1,\ldots,\tau_n$ (with $n \ge 0$)
-and a return type $\tau$. This connection can be expressed compactly as
-$\const{f} : \tau_1 \times \cdots \times \tau_n \to \tau$,
+and a return type $\tau$. This connection can be expressed %compactly
+as $\const{f} : \tau_1 \times \cdots \times \tau_n \to \tau$,
 which collapses to $\const{f} : \tau$ if $n = 0$.
@@
 $\Types = \Data \mathrel{\uplus} \Codata \mathrel{\uplus} \Nondata$, where $\Data$ are the
-\emph{datatypes}, $\Codata$ are the \emph{codatatypes}, and $\Nondata$ are the remaining
+\emph{datatypes}, $\Codata$ are the \emph{codatatypes}, and $\Nondata$ are the %remaining
 \emph{ordinary types}. The functions are partitioned into $\Funcs = \Ctr
@@
 
-Because the types are inhabited (nonempty), a datatype specification is
+Because all types must be inhabited (nonempty), a datatype specification is
 admissible only if a ground constructor term can be exhibited that belongs to
@@
 for a datatype of natural numbers constructed from $\const{Z}$ and $\const{S}$,
-induction rules out nonstandard models, which could contain cyclic values---e.g.,
+induction prohibits nonstandard models, which could contain cyclic values---e.g.,
 an $n$ such that $n \teq \const{S}(n)$---or even infinite acyclic values
@@
 \[\begin{aligned}[t]
-      \keyw{datatype}\;\, \ty{ext\_nat} & \,=\, \const{Nat}(\ty{nat}) \,\mid\, \const{Infty} \\[-.5\jot]
+      \keyw{datatype}\;\, \ty{enat} & \,=\, \const{Nat}(\ty{nat}) \,\mid\, \const{Infty} \\[-.5\jot]
       \keyw{codatatype}\; \,\ty{complex} & \,=\, \const{Complex}(\const{re}{:}\; \ty{real},\; \const{im}{:}\;\ty{real})
@@
 \]
-At the semantic level, it makes no difference whether the types are introduced as
-datatypes or as codatatypes.
+At the semantic level, it makes no difference whether such types are
+introduced as datatypes or as codatatypes.
 %Without loss of generality, we consider that these types are datatypes
@@
 This is a finite codatatype, of cardinality one, whose single value is
-infinite: \,$\mu\vthinspace a.\; A(a)$. Other specimens are
+infinite: \,$\mu a.\; A(a)$. Other specimens are
 \[\begin{aligned}[t]
@@
 \ty{unit} is an uninterpreted type, \ty{b} and \ty{c} will be singleton types
-precisely in the models that interpret \ty{unit} as a singleton. Fortunately,
-given some cardinalities for the ordinary types, it is easy to characterize
+precisely for the models that interpret \ty{unit} as a singleton. Fortunately,
+given cardinalities for the ordinary types, it is easy to characterize
 this degenerate case:
@@
 either infinite or a singleton. In the latter case, $\delta$ necessarily has a
-single constructor, whose arguments have types that are interpreted by
+single constructor, whose arguments have types that are interpreted as
 singletons.
@@
 again excludes finite models. Otherwise, the coinduction principle ensures
-that the type has only one value.
+that the type has at most one value.
 \qed
@@
 
-\section{A Ground Decision Procedure for (Co)-Datatypes}
-\label{sec:the-calculus}
+\section{A Ground Decision Procedure for (Co)datatypes}
+\label{sec:a-ground-decision-procedure-for-co-datatypes}
 
@@
 \section{Extension to Quantified Formulas}
+\label{sec:extension-to-quantified-formulas}
+
+  * universal conjecture is no problem:
+    * falls into the ground fragment via negation
+      and skolemization
+  * nor is existential axiom a problem
+  * what's interesting: universal axioms and existential conjectures
 

### 2014-12-14 tex/codatadp.tex
more data

@@
 bisimilarity rule that identifies equal values, also in the presence of cyclic
-($\omega$-regular) data. The procedure is complete for ground problems and is
-composable via the Nelson--Oppen method. It has been implemented in the latest
-version of CVC4, a state-of-the-art SMT solver. An evaluation based on
-problems generated from Isabelle theories demonstrates the potential of the
-procedure.
+($\omega$-regular) data. The procedure is complete for ground and universal
+problems and is composable via the Nelson--Oppen method. It has been
+implemented in the latest version of CVC4, a state-of-the-art SMT solver. An
+evaluation based on problems generated from Isabelle theories demonstrates the
+potential of the procedure.
 \end{abstract}
@@
 
-                 Datatype suite     Mixed suite        Codatatype suite
+                 Codatatype suite     Mixed suite        Datatype suite
                 (2345 problems)    (XXX problems)       (1240 problems)
 
-   w/o (co)data        800           1255                699
-   w data              815           1268                
-   w codata                          1270                712
-   w both                            1273
+   w/o (co)data        800           1122                699
+   w codata            815 (2)       1139 (3)
+   w data                            1153 (6)            712 (4)
+   w both                            1170 (3+6)
 \end{verbatim}}

### 2014-12-14 tex/codatadp.tex
prepared table

@@
 
+\newcommand\landg{L\,\&\,G}
+\newcommand\HD[1]{\hbox to2.5em{\hfill#1\hfill}}
+
+\begin{table}[t!]
+\begin{center}\begin{tabular}{@{}l@{\kern1em}c@{\kern.5em}c@{\kern.5em}c@{\kern1em}c@{}}
+  & \HD{Distro} & \HD{\emph{AFP}} & \HD{\landg} & \HD{All}
+\\[1pt] %%% TYPESETTING HACK
+\midrule
+%$\enatT$
+\\[-9pt] %%% TYPESETTING HACK
+No decision procedure
+  & 800 & 922 \\
+Codatatypes without bisimilarity
+  &  & 936 \\
+Codatatypes with bisimilarity
+  &  & 939 \\
+Datatypes without cyclicity
+  & 815 & 947 \\
+Datatypes with cyclicity
+  & 815 & 953 \\
+Full decision procedure
+  &  & 970 \\[\jot]
+Total number of goals
+  & 1011\phantom{0} & 1121\phantom{0} & 899 & 3001\phantom{0}
+\end{tabular}\end{center}
+\caption{Number of solved goals for three benchmark suites}
+\end{table}
+
+Do we lose any problems by enabling it?
+
 {\footnotesize
@@
 
-   w/o (co)data        800           1122                699
-   w codata            815 (2)       1139 (3)
-   w data                            1153 (6)            712 (4)
-   w both                            1170 (3+6)
+   w/o (co)data        800           922                699
+   w codata            815 (2)       939 (3)
+   w data                            953 (6)            712 (4)
+   w both                            970 (3+6)
 \end{verbatim}}

### 2014-12-14 tex/codatadp.tex
more

--- 2014-12-14 tex/codatadp.tex
@@
 
-  * to evaluate procedure, we generated problems from Isabelle theories using
-    Sledgehammer
-
-  * chose a selection of theory files:
-    * pretty much every theory file from Isabelle distribution and AFP that
-      defines a codatatype
-    * also included two private theories by Andreas Lochbihler and Peter Gammie
-    * (codatatypes being a recent addition to Isabelle \cite{itp},
-       they are not yet widely used)
-    * and included a number of theories on datatypes
-    * aimed at a good mixture between linear structures (lists, streams) and
-      tree structures
-    * the theory files were chosen before running the experiments
-
-  * how many problems
-  * how many problems have some codatatypes, how many have datatypes (including
-    XXX that have both)
+To evaluate the decision procedure, we generated problems from existing
+Isabelle formalizations using Sledgehammer \cite{paulson-blanchette-2010}.
+Codatatypes being a recent addition to Isabelle
+\cite{blanchette-et-al-2014-impl}, benchmarks are somewhat hard to come by. We
+included the theory files from the Isabelle distribution and the \emph{Archive
+of Formal Proofs} (AFP) \cite{klein-et-al-afp} that define codatatypes falling
+within the supported fragment. We also included two unpublished theories by
+Andreas Lochbihler and Peter Gammie. We complemented the benchark suite with
+some theories about various list and tree data structures, to also test the
+datatypes. The theory files were selected before running the experiments.
 
@@
 
-{\footnotesize
-\begin{verbatim}
-So we can show the success rates, in a table like this:
-
-                 Codatatype suite     Mixed suite        Datatype suite
-                (2345 problems)    (XXX problems)       (1240 problems)
-
-   w/o (co)data        800           1122                699
-   w codata            815 (2)       1139 (3)
-   w data                            1153 (6)            712 (4)
-   w both                            1170 (3+6)
-\end{verbatim}}
+\newcommand\landg{L\,\&\,G}
+\newcommand\HD[1]{\hbox to2.5em{\hfill#1\hfill}}
+
+\begin{table}[t!]
+\begin{center}\begin{tabular}{l@{\kern1.5em}c@{\kern.5em}c@{\kern.5em}c@{\kern1.5em}c}
+  & \HD{Distro} & \HD{AFP} & \HD{\landg} & All
+\\[1pt] %%% TYPESETTING HACK
+\midrule
+%$\enatT$
+\\[-9pt] %%% TYPESETTING HACK
+No decision procedure
+  & 800 & 922 & 450 & 2172 \\
+Codatatypes without bisimilarity
+  & 810 & 936 & 470 & 2216 \\
+Codatatypes with bisimilarity
+  & 811 & 939 & 472 & 2222 \\
+Datatypes without cyclicity
+  & 815 & 947 & 469 & 2231 \\
+Datatypes with cyclicity
+  & 818 & 953 & 470 & 2241 \\
+Full decision procedure
+  & 840 & 970 & 490 & 2300 \\[\jot]
+Total number of goals
+  & 1211\phantom{0} & 1121\phantom{0} & 899 & 3001
+\end{tabular}\end{center}
+\caption{Number of solved goals for three benchmark suites}
+\end{table}
+
+Do we lose any problems by enabling it?
 

### 2014-12-14 tex/codatadp.tex
more eval

@@
 within the supported fragment. We also included two unpublished theories by
-Andreas Lochbihler and Peter Gammie. We complemented the benchark suite with
-some theories about various list and tree data structures, to also test the
-datatypes. The theory files were selected before running the experiments.
-
-  * for each goal, select 256 lemmas as axioms, before monomorphization
-    (after monomorphization, typically more formulas due to duplicates),
-    and encode with the goal
-
-  * then ran CVC4 for XXX seconds on the StarExec cluster \cite{xxx}
-
-  * wanted to evaluate both the impact of datatypes and codatatypes
-  * datatypes are known to be very useful on SMT-LIB, e.g. blah
-  * but we wanted to see how much they make a difference in Isabelle
-
-  * ran all problems four times, to try all four combinations of datatypes and
-    codatatypes enabled or disabled
+Andreas Lochbihler and Peter Gammie. We complemented the benchmarks with some
+theories about various list and tree data structures, to also test the
+datatypes. The theory files were selected before running the experiments. Our
+experimental data is publicly available \cite{our-eval-data}.
+
+For each goal in each theory file, Sledgehammer was invoked to select
+256~lemmas as axioms, which were then monomorphized and translated to SMT-LIB
+along with the goal \cite{boehme-2012-phd}. The resulting problem was given to
+CVC4, running for up to \textbf{TODO: XXX} seconds on the StarExec cluster
+\cite{xxx}.
+
+  * ran all problems several times, to try all four combinations of datatypes
+    and codatatypes enabled or disabled -- also, for datatypes, we tried with
+    and without the cyclicity rule; and for codatatypes, with and without the
+    bisimilarity rule, so find out whether these somewhat expensive rules are
+    useful in practice
 
@@
 \end{tabular}\end{center}
-\caption{Number of solved goals for three benchmark suites}
+\caption{Number of solved goals for the three benchmark suites}
 \end{table}
+
+* compare with BBP 2011!! --- potentially makes a more significant difference
+  than arithmetic (although the benchmark suites are not directly comparable)
 

### 2014-12-14 tex/codatadp.tex
more

@@
 
-To evaluate the decision procedure, we generated problems from existing
-Isabelle formalizations using Sledgehammer \cite{paulson-blanchette-2010}.
-Codatatypes being a recent addition to Isabelle
-\cite{blanchette-et-al-2014-impl}, benchmarks are somewhat hard to come by. We
-included the theory files from the Isabelle distribution and the \emph{Archive
-of Formal Proofs} (AFP) \cite{klein-et-al-afp} that define codatatypes falling
-within the supported fragment. We also included two unpublished theories by
-Andreas Lochbihler and Peter Gammie. We complemented the benchmarks with some
-theories about various list and tree data structures, to also test the
-datatypes. The theory files were selected before running the experiments. Our
-experimental data is publicly available \cite{our-eval-data}.
-
-For each goal in each theory file, Sledgehammer was invoked to select
-256~lemmas as axioms, which were then monomorphized and translated to SMT-LIB
-along with the goal \cite{boehme-2012-phd}. The resulting problem was given to
-CVC4, running for up to \textbf{TODO: XXX} seconds on the StarExec cluster
-\cite{xxx}.
-
-  * ran all problems several times, to try all four combinations of datatypes
-    and codatatypes enabled or disabled -- also, for datatypes, we tried with
-    and without the cyclicity rule; and for codatatypes, with and without the
-    bisimilarity rule, so find out whether these somewhat expensive rules are
-    useful in practice
-
-  * lemmas/axioms often include basic properties of constructor and selectors,
-    which are tautologies and ignored if (co)datatypes are enabled
-  * allows a fair comparison: how much do (co)datatype support brings as
-    opposed to just letting Sledgehammer do what it would normally do
-
-  * experience with such features is not extremely high -- e.g. arithmetic,
-    a most useful theory on lots of benchmarks, increases Sledgehammer's
-    success rate by 2 percentage points, or 4\%, with Z3 in earlier work \cite{xxx}
-
-  * so when analyzing statistics of this kind, with highly varied problems,
-    we must not overestimate the importance of a single trick
-  * on the other hand: every percentage point counts! cite Tom Hales
-
 \newcommand\landg{L\,\&\,G}
@@
 
-\begin{table}[t!]
+\begin{table}[tbh!]
 \begin{center}\begin{tabular}{l@{\kern1.5em}c@{\kern.5em}c@{\kern.5em}c@{\kern1.5em}c}
@@
 \caption{Number of solved goals for the three benchmark suites}
+\label{tab:bench}
 \end{table}
 
-* compare with BBP 2011!! --- potentially makes a more significant difference
-  than arithmetic (although the benchmark suites are not directly comparable)
-
-Do we lose any problems by enabling it?
-
-* statistics, like how often do we get proofs that contain the cyclicity and
-  bisimilarity rules
-
-* and finally, we look in more detail at one or two such proofs, by presenting
-  it and explaining it (and, before that, understanding it)
+To evaluate the decision procedure, we generated problems from existing
+Isabelle formalizations using Sledgehammer \cite{paulson-blanchette-2010}.
+Codatatypes being a recent addition to Isabelle
+\cite{blanchette-et-al-2014-impl}, benchmarks are somewhat hard to come by. We
+included the theory files from the Isabelle distribution (the Distro) and the \emph{Archive
+of Formal Proofs} (AFP) \cite{klein-et-al-afp} that define codatatypes falling
+within the supported fragment. We also included two unpublished theories by
+Andreas Lochbihler and Peter Gammie (\landg). We complemented the benchmarks
+with some theories about various list and tree data structures, to also test
+the datatypes.
+The theory files were selected before running the experiments.
+Our experimental data is publicly available \cite{our-eval-data}.
+
+For each goal in each theory file, Sledgehammer was invoked to select
+256~lemmas as axioms, which were then monomorphized and translated to SMT-LIB
+along with the goal \cite{boehme-2012-phd}. The resulting problem was given to
+CVC4, running for up to \textbf{TODO: XXX} seconds on the StarExec cluster
+\cite{xxx}. Due to the lack of machinery for reconstructing inferences about
+(co)datatypes in Isabelle, CVC4 is trusted as an oracle in these experiments.
+
+CVC4 was run on each problem several times, with the support for datatypes and
+codatatypes either enabled or disabled. The contributions of the cyclicity and
+bisimilarity rules were also measured, so find out whether these somewhat
+expensive rules are useful in practice. Even when the decision procedure is
+disabled, some of the generated axioms typically include basic properties of
+constructors and selectors, which the decision procedure would recognize as
+tautologies. This allows us to answer the question, \relax{What are the
+benefits of activating the decision procedure as opposed to letting
+Sledgehammer do what it would normally do?}
+
+\newcommand\BAD[1]{\textcolor{red}{\textbf{#1}}}
+
+%  * experience with such features is not extremely high -- e.g. arithmetic,
+%    a most useful theory on lots of benchmarks, increases Sledgehammer's
+%    success rate by 2 percentage points, or 4\%, with Z3 in earlier work \cite{xxx}
+
+%  * so when analyzing statistics of this kind, with highly varied problems,
+%    we must not overestimate the importance of a single trick
+%  * on the other hand: every percentage point counts! cite Tom Hales
+
+The results are summarized in Table~\ref{tab:bench}. The decision procedure
+makes a difference across all theories, and an overall difference of
+\BAD{4.7\%}. Moreover, every aspect of the procedure, including the more
+expensive rules, make a contribution. What is not visible in the table, but
+readily available when looking at the raw data \cite{our-eval-data}, is that
+the stronger decision procedures subsume the weaker ones. In the context of
+Sledgehammer, the power of interpreted (co)datatype is comparable with that of
+arithmetic \cite{blanchette-et-al-2013-smt}.
+
+%* and finally, we look in more detail at one or two such proofs, by presenting
+%  it and explaining it (and, before that, understanding it)
 
@@
 
-This work is part of a larger program that aims at enriching automatic
-provers with high-level features and at reducing the gap between automatic and
-interactive theorem proving. As future work, we want to try out CVC4 and its
-(co)datatypes for higher-order model finding in Isabelle, as an
-alternative to the counterexample generator Nitpick
+This work is part of a larger program that aims at enriching automatic provers
+with high-level features and at reducing the gap between automatic and
+interactive theorem proving. As future work, we want to implement proof
+reconstruction for (co)datatype inferences in Isabelle. We also want to try
+out CVC4 and its (co)datatypes for higher-order model finding in Isabelle, as
+an alternative to the counterexample generator Nitpick
 \cite{blanchette-nipkow-2010}. We also see some opportunities to enrich

### 2014-12-14 tex/codatadp.tex
more

@@
 interactive theorem proving. As future work, we want to implement proof
-reconstruction for (co)datatype inferences in Isabelle. We also want to try
-out CVC4 and its (co)datatypes for higher-order model finding in Isabelle, as
-an alternative to the counterexample generator Nitpick
-\cite{blanchette-nipkow-2010}. We also see some opportunities to enrich
-SMT solvers with recursive and corecursive functions. Finally, it might be
-possible to go further in terms of supporting nested and mixed (co)recursion
-and quantified formulas over (co)datatypes in solvers.
+reconstruction for CVC4's (co)datatype inferences in Isabelle. We also want to
+try out CVC4 for higher-order model finding in Isabelle, as an alternative to
+the counterexample generator Nitpick \cite{blanchette-nipkow-2010}. We also
+see some opportunities to enrich SMT solvers with recursive and corecursive
+functions. Finally, it might be possible to go further in terms of supporting
+nested and mixed (co)recursion and quantified formulas over (co)datatypes in
+solvers.
 

### 2014-12-14 tex/codatadp.tex
More reorg, will rework text.

@@
 \newcommand{\tpos}[2]{#1\!\mid_{#2}}
+\newcommand{\muvar}{\mathcal{Var}}
 
@@
 
-It remains to show the inverse of the previous lemma, that when a derivation tree with root node $\Ec$ contains a saturated node $\Fc$,
+It remains to show the inverse of the previous lemma, that when a derivation tree with root node $\Ec$ contains a saturated node,
 then $\Ec$ is satisfiable $\thD$.
-We demonstrate a particular model $\M$ of $\thD$ that satisfies $\Ec$.
-Before doing so, we require additional terminology concerning $\mu$-terms.
+To do so, we demonstrate a particular model $\M$ of $\thD$ that satisfies $\Ec$,
+but first we require additional terminology concerning $\mu$-terms.
 
@@
 the \emph{interpretation of $t_2$ in $t_1$} is the $\mu$-term $t_1\interp{t_2}^\emptyset$ as returned by the following recursive procedure,
-where $x \mapsto \mu x. C( \vec u ) \in t_1$ indicates that $\mu x. C( \vec u )$ is the subterm of $t_1$ that binds this occurrence of variable $x$.
+where $x \mapsto \mu x. \const{C}( \vec u ) \in t_1$ indicates that $\mu x. \const{C}( \vec u )$ is the subterm of $t_1$ that binds this occurrence of variable $x$.
 \[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
-t_1\interp{x}^M & = & \mu x. C( t_1\interp{ \vec u}^{M \cup \{ x \}}) & x \not\in M, x \mapsto \mu x. C( \vec u ) \in t_1  \\
+t_1\interp{x}^M & = & \mu x. \const{C}( t_1\interp{ \vec u}^{M \cup \{ x \}}) & x \not\in M, x \mapsto \mu x. \const{C}( \vec u ) \in t_1  \\
 t_1\interp{x}^M & = & x & \text{otherwise} \\
-t_1\interp{\mu x. C( \vec u )}^M & = & \mu x. C( t_1\interp{ \vec u }^{M \cup \{ x \}} ) & x \not\in M \\
-t_1\interp{\mu x. C( \vec u )}^M & = & x & \text{otherwise}
+t_1\interp{\mu x. \const{C}( \vec u )}^M & = & \mu x. \const{C}( t_1\interp{ \vec u }^{M \cup \{ x \}} ) & x \not\in M \\
+t_1\interp{\mu x. \const{C}( \vec u )}^M & = & x & \text{otherwise}
 \end{array}\]
@@
 
-Given a $\mu$-term $t_1$ of the form $\mu x. C( u_1, \ldots, u_n )$ with subterm $t_2$,
+Given a $\mu$-term $t_1$ of the form $\mu x. \const{C}( u_1, \ldots, u_n )$ with subterm $t_2$,
 we say $t_2$ is a \emph{bisimilar subterm} of $t_1$ if and only if it is of the form
-$\mu y. C( v_1, \ldots, v_n )$ and $t_1\interp{ u_i } \sim t_1\interp{ v_i }$ for $i = 1,\ldots, n$.
+$\mu y. \const{C}( v_1, \ldots, v_n )$ and $t_1\interp{ u_i } \sim t_1\interp{ v_i }$ for $i = 1,\ldots, n$.
 We say $t_1$ is \emph{normal} if and only if it does not contain a strict bisimilar subterm
 and all of its strict subterms are also normal.
-For instance, the $\mu$-term $t_1$ of the form $\mu x. \const{S}( \mu y. \const{S}( y ) )$ is not normal, 
-since $\mu y. \const{S}( y )$ is a bisimilar subterm of $t_1$.
+For instance, the $\mu$-term $t_1$ of the form $\mu x. \const{C}( \mu y. \const{C}( y ) )$ is not normal, 
+since $\mu y. \const{C}( y )$ is a bisimilar subterm of $t_1$.
 In particular, note that their arguments have the same interpretation in $t_1$, 
 that is,
-$t_1\interp{\mu y. \const{S}( y )}^\emptyset =$ 
-$\mu y. \const{S}( t_1\interp{y}^{\{y\}} ) =$ 
-$\mu y. \const{S}( y )$ 
+$t_1\interp{\mu y. \const{C}( y )}^\emptyset =$ 
+$\mu y. \const{C}( t_1\interp{y}^{\{y\}} ) =$ 
+$\mu y. \const{C}( y )$ 
 is syntactically equivalent to
 $t_1\interp{y}^\emptyset =$ 
-$\mu y. \const{S}( t_1\interp{y}^{\{y\}} ) =$ 
-$\mu y. \const{S}( y )$.
-The $\mu$-term $t_1$ of the form $\mu x. \const{S}( \mu y. \const{S}( x ) )$ is also not normal, 
-since $\mu y. \const{S}( x )$ is a bisimilar subterm of $t_1$,
+$\mu y. \const{C}( t_1\interp{y}^{\{y\}} ) =$ 
+$\mu y. \const{C}( y )$.
+The $\mu$-term $t_1$ of the form $\mu x. \const{C}( \mu y. \const{C}( x ) )$ is also not normal, 
+since $\mu y. \const{C}( x )$ is a bisimilar subterm of $t_1$,
 noting that
-$t_1\interp{\mu y. \const{S}( x )}^\emptyset =$ 
-$\mu y. \const{S}( t_1\interp{x}^{\{y\}} ) =$ 
-$\mu y. \const{S}( t_1\interp{\mu x. \const{S}( \mu y. \const{S}( x ) ) }^{\{y\}} ) =$ 
-$\mu y. \const{S}( \mu x. \const{S}( t_1\interp{\mu y. \const{S}( x )}^{\{x,y\}} ) ) =$ 
-$\mu y. \const{S}( \mu x. \const{S}( y ) )$
+$t_1\interp{\mu y. \const{C}( x )}^\emptyset =$ 
+$\mu y. \const{C}( t_1\interp{x}^{\{y\}} ) =$ 
+$\mu y. \const{C}( t_1\interp{\mu x. \const{C}( \mu y. \const{C}( x ) ) }^{\{y\}} ) =$ 
+$\mu y. \const{C}( \mu x. \const{C}( t_1\interp{\mu y. \const{C}( x )}^{\{x,y\}} ) ) =$ 
+$\mu y. \const{C}( \mu x. \const{C}( y ) )$
 is syntactically equivalent to $t_1\interp{x}^\emptyset = t_1$ for the renaming $\{ x \mapsto y, y \mapsto x \}$.
@@
 and by replacing all of the non-normal subterms of $v$ with their corresponding normal forms.
-For instance, $\nf{(\mu x. \const{S}( \mu y. \const{S}( x ) ))} = \mu x. \const{S}( x )$.
+For instance, $\nf{(\mu x. \const{C}( \mu y. \const{C}( x ) ))} = \mu x. \const{C}( x )$.
+
+We say that $t \vsimv{x} s$ if and only if $\mu$-terms $t$ and $s$ are syntactically equivalent
+for some renaming that avoids capturing variables besides $x$.
+For instance, 
+$\mu x. \const{D}( x ) \vsimv{y} \mu x. \const{D}( y )$,
+$\mu x. \const{C}( x, x ) \vsimv{x} \mu y. \const{C}( x, y )$ and
+$\mu x. \const{C}( z, x ) \vsimv{z} \mu y. \const{C}( z, y )$,
+but
+$\mu x. \const{D}( x ) \not\vsimv{x} \mu x. \const{D}( y )$ and
+$\mu x. \const{C}( x, x ) \not\vsimv{y} \mu y. \const{C}( x, y )$.
+
+%\begin{lemma}
+%If $t_1$ and $u$ are normal, and if $u \not\vsimv{x} t_1\interp{t_2}$ for all $t_2 \in \mathcal{T}( t_1 )$,
+%then $t_1 \{ x \mapsto u \}$ is normal.
+%\end{lemma}
+%\begin{proof}
+%\rem{TODO}
+%\end{proof}
 
@@
 in $\M( \tau )$ that is a renaming of 
-$\nf{(\mu x. \const{C}( \M( \vec t ) ))}$ where $x \not\in FV( \vec t )$, and
+$\nf{(\mu x. \const{C}( \M( \vec t ) ))}$ where $x$ is fresh, and
 \item 
-for all selector terms $s^j_i( t )$ of type $\tau$ where $\M( t )$ is of the form $\mu x. \const{C}_i( \vec u )$.% for some $\vec u$,
+for all selector terms $s^j_i( t )$ of type $\tau$,% for some $\vec u$,
 $\M( s^j_i( t ) )$ is the value
 in $\M( \tau )$ that is a renaming of 
-$\M( t )\interp{u_j}$.
+$\M( t )\interp{u_j}$
+when $\M( t )$ is of the form $\mu x. \const{C}_i( \vec u )$.
 \end{enumerate}
@@
 In the following, we consider only normal models.
-Thus, when constructing any model $\M$,
-it remains only to define how wrong-applied selector terms and the free variables are interpreted.
+Thus, when constructing a model $\M$ for $\Ec$,
+it remains only to define how wrong-applied selector terms and variables in $\Ec$ are interpreted.
 For the latter, this will be based on the mapping $\Val$ constructed in step two of our calculus.
-We first state the following property of $\Val$, which amounts to saying that interpretation of an equivalence class is
-not sensisitive to the context in which it is interpretated.
-
-\begin{lemma}
-If $\Val$ is a mapping construced in step two of our calculus,
-then $t_1 \interp{ \mu v_{\ec{x}}. t_2 } = \Val \ec{x}$
-for all $t_1\interp{\mu v_{\ec{x}}. t_2} \in \mathcal{T}(\Val)$.
-\end{lemma}
-\begin{proof}
-\rem{TODO}
-\qed
-\end{proof}
-
-We say $t \vsimv{x} s$ if and only if $\mu$-terms $t$ and $s$ are syntactically equivalent
-for some renaming that avoids capturing any variable besides $x$.
-For instance, for $t:= \mu x. \const{C}( x, y )$ and $u := \mu y. \const{C}( x, y )$,
-we have $t \not\vsim u$, but it is the case that both
-$t \vsimv{x} u$ and $t \vsimv{y} u$.
-
-The \emph{completion} of $\Val$, which we will write as $\Val^\ast$, 
+Given a normal model $\M$, the \emph{completion} of $\Val$, which we will write as $\Val^\ast$, 
 is the result of exhaustively applying the following update to $\Val$:
@@
   \quad
-  c \in \M( \tau )
+  \mu x. t \sim v, v \in \M( \tau )
   \quad
-  c \not\sim_x t \text{ for all } t \in \mathcal{T}(\vrange(\Val))
+  (\mu x. t) \not\sim_x u \text{ for all } u \in \mathcal{T}(\vrange(\Val))
 }{
-  \Val := \Val \{ x \mapsto c \}
+  \Val := \Val \{ x \mapsto \mu x. t \}
 }
@@
 
-\rem{revise from here on}
-
-\begin{lemma} \label{lem:normal-subterm-bisim}
-If $t_1 \vsim t_2$, then $t_1\interp{\tpos{t_1}{p}} \vsim t_2\interp{\tpos{t_2}{p}}$.
+Since this rule removes one variable from the set $FV( \Val )$ and does not add any additional variables to $FV( \Val )$, 
+it can only be applied a finite number of times.
+We first state the following properties of $\Val$, 
+where we write $\muvar( t )$ to denote $t$ if $t$ is a variable, or $x$ if $t$ is of the form $\mu x. u$.
+%which amounts to saying that 
+%equivalence class are interpreted as values,
+%the interpretation of an equivalence class is not sensitive to the context in which it is interpreted,
+%and that each equivalence classes has a distinct interpretation.
+
+\begin{lemma} 
+\label{lem:model-completion}
+The following are true of $\Val^\ast$:
+\begin{enumerate}
+\item $\nf{(\Val^\ast \ec{x})} \in \M( \tau )$ for all $\ec{x} : \tau$ of $\Ec$,
+\item
+$t_1 \interp{ t_2 } = \nf{(\Val^\ast \ec{x})}$
+for all $t_1\interp{t_2} \in \mathcal{T}(\Val)$ where $\muvar(t_2)=v_{\ec{x}}$,
+\item 
+$\nf{(\Val^\ast \ec{x})} \vsim \nf{(\Val^\ast \ec{y})}$ if and only if $\ec{x} = \ec{y}$.
+\end{enumerate}
 \end{lemma}
 \begin{proof}
-\end{proof}
-
-\begin{lemma} \label{lem:normal-subs}
-If $t_1$ and $t_2$ are normal $\mu$-terms where $t_1 \not\vsim t_2$,
-and if $u_1$ is a normal $\mu$-term such that $u_1 \not\vsim u_2 \sigma$ for all $\sigma$, $u_2 \in \mathcal{T}( t_1 ) \cup \mathcal{T}( t_2 )$,
-then $\nf{(t_1 \sigma_x)} \not\vsim \nf{(t_2 \sigma_x)}$ for $\sigma_x := \{ x \mapsto u_1 \}$.
-\end{lemma}
-\begin{proof}
-By contradiction assume that 
-$t_1 \not\vsim t_2$ and 
-$\nf{(t_1 \sigma)} \vsim \nf{(t_2 \sigma)}$
-for some $\sigma := \{ x \mapsto u_1 \}$
-where $u_1 \not\vsim u_2 \sigma$ for all $\sigma$, $u_2 \in \mathcal{T}( t_1 ) \cup \mathcal{T}( t_2 )$ 
-and $t_1, t_2, u_1$ are normal.
-Then it must be the case that there is a position where, say $t_1$, 
-is such that $\tpos{t_1}{p} = x$ and $\tpos{t_2}{p} = \mu y. u$,
-and by Lemma~\ref{lem:normal-subterm-bisim}, where $\tpos{\nf{(t_1 \sigma)}}{p} \vsim \tpos{\nf{(t_2 \sigma)}}{p}$.
-
-%Since $\nf{(t_2 \interp{\mu y. u})} \in \mathcal{T}( t_2 )$.
+To show 1, we must argue that $FV( \Val^\ast ) = \emptyset$, and that $\nf{(\Val^\ast \ec{x})}$ is not cyclic for
+$\ec{x} : \tau$ when $\tau \not\in \Codata$.
+\rem{TODO}
+\qed
 \end{proof}
@@
 
-We construct a model $\M$ of $\thD$ that interprets each type $\tau$
-as a maximal set of closed normal $\mu$-terms of that type that are 
-unique up to renaming of $\mu$-bound variables,
-and acyclic when $\tau \not\in \Codata$.
-
-The interpretation of constructor terms
-$\M( C( \vec t ) )$ for $C \in \Ctr$ is the value in $\M$ that is a renaming of $\nf{(\mu x. C( \M( \vec t ) ))}$ where $x \not\in FV( \vec t )$.
-The interpretation of (correctly-applied) selector terms
-$\M( s^j_i( t ) )$ for $s^j_i \in \Sel$ is the value in $\M$ that is a renaming of $\M( t )\interp{u_j}$
-where $\M( t )$ is of the form $\mu x. C_i( \vec u )$ for some $\vec u$.
-For instance, if $\M( t ) = \mu x. \const{C}_1( x )$ (call this term $v$), 
-then $\M( s_1^1( t ) )$ is a renaming of
-$v\interp{x} =$ 
-$v\interp{\mu x. \const{C}_1( x )}^\emptyset =$
-$\mu x. \const{C}_1( v\interp{x}^{\{x\}} ) =$
-$\mu x. \const{C}_1( x )$.
-%To interpret a correctly-applied selector in $\M$, if $v = \mu x. C_i( \vec t )$, then $\M( s^j_i( v ) )$ is the value in $\M( \tau )$ that is a renaming of $v\interp{t_j}$,
-%where $t_j$ has type $\tau$.
-%We assume $\M$ interprets constructors and correctly-applied selectors in the obvious way
-%\footnote{
-%The interpretation of $\M( C( \vec t ) )$ is the value in $\M( \tau )$ that is a renaming of $\nf{\mu x. C( \M( \vec t ) ) }$.
-%To interpret a correctly-applied selector, if $v = \mu x. C_i( \vec t )$, then $\M( s^j_i( v ) )$ is the value in $\M( \tau )$ that is a renaming of $v\interp{t_j}$,
-%where $t_j$ has type $\tau$.
-%},
-Additionally, $\M$ will interpret wrongly-applied selectors based on equality information in $\Fc$,
-and will interpret the variables of $\Fc$ based on the valuation $\Val$ which was constructed in step two of our calculus.
-
-Recall that $\Val$ is a mapping from the equivalence classes of $\Fc$ to $\mu$-terms.
-We first \emph{complete} $\Val$ so that it maps equivalence classes only to closed $\mu$-terms, that is, terms containing no free variables.
-This is done by exhaustively applying the following rule:
-
-We write $\mathcal{T}(\Val)$ to denote the smallest set containing $\nf{(v\interp{u})}$ for each non-variable subterm $u$ of $v$ in the range of $\Val$.
-%We write $\vrange(\Val)$ to denote the set of $\mu$-terms mapped to by $\Val$.
-Since this rule removes one variable from the set $FV( \Val )$ and does not add any additional variables to $FV( \Val )$, 
-it can only be applied a finite number of times.
-
-Let $\Val^\ast$ be the result of applying the above rule to completion on the mapping $\Val$ we constructed during step 2 of our calculus.
-We first argue that $FV( \Val^\ast ) = \emptyset$.
-Since \rn{Split} does not apply to $\Fc$,
-all equivalence classes of $\Fc$ finite datatype type contain constructor terms, 
-and thus $FV( \Val )$ contains no variables of finite datatype type.
-For each equivalence classes of $\Fc$ with finite co-datatype $\tau$, 
-we know $\tau$ must have cardinality one, due to \rem{reference}.
-Since \rn{Singleton} does not apply to $\Fc$, 
-there is only one equivalence class of type $\tau$, call it $\ec{x}$.
-%It cannot be the case that $v_{\ec{x}} \in FV(\Val^\ast \ec{x})$, then we may apply \rn{Complete} for the (single) value $c$ of type $\tau$, 
-%for which the second premise holds since $\ec{x}$ is the only equivalence class of type $\tau$.
-If it were the case that $v_{\ec{x}} \in FV(\Val^\ast )$,
-then \rn{Complete} would apply for the (single) value $c$ of type $\tau$, 
-for which the third premise holds since $\ec{x}$ is the only equivalence class of type $\tau$.
-Since all $\tau \in \Nondata$ are infinite, all variables in $FV( \Val^\ast )$ have infinite type.
-For any infinite type $\tau$,
-there exists a closed normal $\mu$-term $c_1$ that is not bisimilar to any instance of a term in $t \in \mathcal{T}(\Val^\ast)$.
-To argue this is the case, notice that either $\tau \in \Nondata$,
-$\tau$ is a recursive type, or contains an infinite subfield of type $\upsilon$.
-In the first case, we may choose an arbitrary constant for $c_1$ not yet occurring in $\vrange(\Val)$.
-In the second case, we choose a value $c_1$ whose depth is greater than any term in $\vrange(\Val)$.
-In the third case, we choose a value $c_1$ containing a subterm of type $\upsilon$ that is not bisimilar to any instance of a term $\mathcal{T}(\Val^\ast)$, 
-which is guarenteed to exist by applying this argument inductively.
-%since $\mathcal{T}(\Val^\ast)$ is a finite set. \rem{elaborate} % (noting that a closed $\mu$-term is only bisimilar to other closed $\mu$-terms).
-Thus, $FV( \Val^\ast )$ contains no variables of infinite type,
-and consequently $\Val^\ast$ is a mapping from the equivalence classes of $\Fc$ to closed $\mu$-terms.
-We further argue that $\Val^\ast$ has the following properties:
-%the following invariants are maintained by applications of \rn{Complete}:% for all $\ec{x}, \ec{y}$:
+We consider a normal model $\M$ 
+that interprets wrongly-applied selectors based on equality information in $\Fc$,
+and interprets the variables of $\Fc$ based on the completion $\Val^\ast$ of the mapping $\Val$ from step two of our calculus.
+For the latter, let $\M( x )$ be the value in $\M$ that is a renaming of $\nf{(\Val^\ast \ec{x})}$ for each variable $x$ in $\mathcal{T}(\Fc)$.
+Notice that due to Lemma~\ref{lem:model-completion}.3, for variables $x, y \in FV( \Fc )$,
 \begin{eqnarray}
-\label{eqn:m-normal} 
-\Val^\ast \text{ maps to normal $\mu$-terms only } \\
-%\text{If } \Val \ec{x} \text{ is not closed and has subterm $u$, then } \Val \ec{x} \interp{ u } = \Val \ec{y} \text{ for some } \ec{y} \\
-%\Val \ec{x} = \Val \ec{y}\interp{u} \text{ for some $\ec{x}$, for all non-closed subterms } u \text{ of } \Val \ec{y},  \\
-%\Val \ec{x} = v\interp{u} \text{ for some } \ec{x}, \text{ for all non-closed } u \in \mathcal{T}( \Val ) \\
-%\Val \ec{x} \interp{\mu v_{\ec{y}}. t} = \Val \ec{y} \\
 \label{eqn:m-bij} 
-\Val^\ast \ec{x} \sim \Val^\ast \ec{y} \text{ if and only if } \ec{x} = \ec{y}
+\M( x ) = \M( y ) \text{ if and only if } \ec{x} = \ec{y}
 \end{eqnarray}
-We show this by induction on the number of applications of \rn{Complete}.
-We additionally show that applications of \rn{Complete} maintain the invariant 
-for all subterms $\mu v_{\ec{x}}. t$ occurring in the range of $\Val$ for some $\ec{y}$:
-%\begin{eqnarray}
-%\label{eqn:m-inv}
-%\Val \ec{y} \interp{ \mu v_{\ec{x}}. t } = \Val \ec{x}
-%\end{eqnarray}
-For the base case, since \rn{Bisimilar} does not apply to $\Fc$,
-we have that $\Val \ec{x} \not\vsim \Val \ec{y}$ for all distinct $\ec{x}$ and $\ec{y}$,
-and thus~\ref{eqn:m-bij} holds for $\Val$.
-\rem{todo 1}
-%Due to our construction of $\Val$ in step 2 of our calculus, 
-%statement~\ref{eqn:m-inv} holds, which together with~\ref{eqn:m-bij} implies~\ref{eqn:m-normal}.
-%Moreover, since an equivalence class $\ec{u}$ exists for all subterms $u$ of all terms $t \in \mathcal{T}(\Fc)$,
-%we have that $t\interp{u} = \Val \ec{u}$ for all such terms $u$ and $t$. \rem{elaborate?}
-For the inductive case, we assume an arbitrary $\Val_i$ satisfies~\ref{eqn:m-normal} and~\ref{eqn:m-bij},
-and show that $\Val_i \sigma_x$ does as well, where $\sigma_x := \{ v_{\ec{x}} \mapsto c_1 \}$ is a substitution applied by \rn{Complete}.
-By contradiction, assume that $\Val_i \sigma_x \ec{y} \sim \Val_i \sigma_x \ec{z}$ for some distinct $\ec{y}$ and $\ec{z}$.
-Then it must have been the case that $\Val_i \ec{y}$ and $\Val_i \ec{z}$ have a position $p$
-where, w.l.o.g., say $\Val_i \ec{y} \mid_p$ is $v_{\ec{x}}$ and $\Val_i \ec{z} \mid_p$ is a term of the form $\mu v_{\ec{w}}. t$ 
-and $c_1 \sim (\mu v_{\ec{w}}. t) \sigma_x$.
-%Since $\Val_i$ satisfies invariant~\ref{eqn:m-bij},
-%we have that $\Val_i \ec{w} \sim c_1$ for some $\ec{w}$, 
-However, this contradicts the premise of \rn{Complete} that $c_1$ is not bisimilar to any $t \sigma \in \mathcal{T}( \Val_i )$ for any $\sigma$.
-%Thus, by induction on the number of applications of \rn{Complete}, statements~\ref{eqn:m-normal} and~\ref{eqn:m-bij} hold for $\Val^\ast$.
-Thus, statements~\ref{eqn:m-normal} and~\ref{eqn:m-bij} hold for $\Val^\ast$.
-
-%we have that $\Val$ satisfies both~\ref{eqn:m-normal} and~\ref{eqn:m-bij}.
-%It satisfies the former since if a $\mu$-term $\mu v_{\ec{x}}. t_1$ had a bisimilar subterm $\mu v_{\ec{y}}. t_2$,
-%then the rule \rn{Bisimilar} would have applied to $x$ and $y$. \rem{elaborate?}
-%For the inductive case, we assume $\Val$ satisfies~\ref{eqn:m-normal} and~\ref{eqn:m-bij},
-%and show that $\Val \sigma$ does as well, where $\sigma := \{ v_{\ec{x}} \mapsto c_1 \}$ is a substitution applied by \rn{Complete}.
-%By contradiction, assume that $\Val \sigma \ec{y}$ is a non-normal $\mu$-term for some $\ec{y}$.
-%Since $c_1$ is normal, we have $\ec{x} \neq \ec{y}$, 
-%and thus $\Val \ec{y}$ must have been of the form $\mu v_{\ec{y}}. t$ where $\ec{x} \in FV( t )$.
-%Let $u$ be the (strict) subterm of $\Val \ec{y}$ such that $(\Val \ec{y} \sigma ) \interp{u \sigma } \sim (\Val \ec{y} \sigma)$. 
-%Since $\Val$ maps only to normal $\mu$-terms, we have $x \in FV( u )$.
-
-Now, let $\M( x )$ be the value in $\M$ that is a renaming of$\Val^\ast \ec{x}$ for each variable $x$ in $\mathcal{T}(\Fc)$,
-each of which are guarenteed to exist in $\M$ due to statement~\ref{eqn:m-normal} and our construction of $\M$,
-and moreover are guarenteed to be distinct in $\M$ due to statement~\ref{eqn:m-bij}.
+
 We first show $\M$ satisfies all equalities in $t_1 \teq t_2 \in \Fc$, where $t_1$ and $t_2$ have type $\tau$.
-
-If $\tau \in \Nondata$, $t_1$ and $t_2$ must be variables, and thus $\M( t_1 ) = \M( t_2 )$ by construction.
-
-\rem{revise for co-datatype}
-If $\tau \in \Data \cup \Codata$,
-we first show that $\M( t ) = \Val^\ast \ec{x}$ for all non-selector terms $t$ that occur in equivalence classes $\ec{x}$.
-We show this by induction on the size of $\Val^\ast \ec{x}$.
-In the case where $t$ is a variable, then $\M( t ) = \Val^\ast \ec{ t } = \Val^\ast \ec{ x }$.
-In the case where $t$ is a constructor term $C( \vec u )$, 
-then $\Val^\ast \ec{x}$ must have been assigned by \rn{Ind} for some constructor term $C'( \vec u' ) \in \ec{x}$.
-Since \rn{Unify$_2$} does not apply to $\Fc$, we have that $C = C'$.
-Since \rn{Unify$_1$} does not apply to $\Fc$, we have that $\ec{\vec u} = \ec{\vec u'}$.
-By the induction hypothesis, we have $\M( \vec u ) = \M( \vec u' )$, and thus $\M( t ) = C( \M( \vec u ) ) = C'( \M( \vec u ' ) ) = \Val^\ast \ec{ x }$.
-
-We now show that $\M( t ) = \Val^\ast \ec{x}$ for all selector terms $s^j_k( y )$ that occur in datatype equivalence classes $\ec{x}$.
-Since \rn{Split} does not apply, 
-$\ec{y}$ must contain a term of the form $C_i( s^1_i( y ), \ldots, s^{a_i}_i( y ) )$ for some $i$.
-Let $x_1, \ldots, x_n$ be variables in equivalence classes $\ec{s^1_i( y )}, \ldots, \ec{s^{a_i}_i( y )}$.
-By the previous paragraph, we have $\M( y ) = \M( C_i( s^1_i( y ), \ldots, s^{a_i}_i( y ) ) )$,
-which is $C_i( \Val^\ast \ec{ x_1 }, \ldots, \Val^\ast \ec{ x_{a_i} } )$.
-If $i = k$, then $\M( t ) = \M( s^j_k( y ) ) = s^j_k( \M( y ) ) = \Val^\ast \ec{x_j}$, which is $\Val^\ast \ec{ x }$ since $\ec{x_j} = \ec{s^j_i( y )} = \ec{x}$.
-If $i \neq k$, then $\thD$ does not constraint the value of $\M( t )$,
-and since $s^j_k( \M( y ) )$ is otherwise unconstrained in our model construction since \rn{Cong} does not apply, 
-we may interpret $s^j_k( \M( y ) )$ as $\Val^\ast \ec{ x }$. 
-Thus, for all datatype equivalence classes $\ec{x}$, we have $\M( t ) = \Val^\ast \ec{x}$ for each term $t \in \ec{x}$, which implies $\M \models t_1 \teq t_2$.
+We show that $\M( t ) = \Val^\ast \ec{t}$ for all terms $t \in \mathcal{T}( \Fc )$, which we show by induction on the size of $t$.
+
+If $t$ is a variable, then $\M( t ) = \Val^\ast \ec{ t }$ by construction.
+If $t$ is a constructor term $\const{C}( u_1, \ldots, u_n )$, 
+then since \rn{Unify$_2$} does not apply to $\Fc$,
+then $\Val^\ast \ec{t}$ must be a term of the form $\mu v_{\ec{t}}. \const{C}( w_1, \ldots, w_n )$,
+where say that $\muvar( w_i ) = v_{\ec{y_i}}$ for $i = 1, \ldots, n$.
+Since \rn{Unify$_1$} does not apply to $\Fc$, it must be that $\ec{u_i} = \ec{y_i}$ for $i = 1, \ldots, n $.
+%Due to Lemma~\ref{lem:model-completion}.2 and by induction
+\rem{TODO}
+
+%If $\tau \in \Nondata$, $t_1$ and $t_2$ must be variables, and thus $\M( t_1 ) = \M( t_2 )$ by construction.
+%If $\tau \in \Data \cup \Codata$, we show $\M \models t_1 \teq t_2$ by showing 
+%We show by induction on the term size of $t$ that $\M( t ) = \Val^\ast \ec{t}$.
+%If $t$ is a variable, then $\M( t ) = \Val^\ast \ec{ t } = \Val^\ast \ec{ x }$.
+%In the case where 
+%where say $\muvar( v_i ) = v_{\ec{y_i}}$ for $i = 1, \ldots, n$.
+%We have $\M( \const{C}( \vec u ) ) = \nf{(\mu y. \const{C}( \M( \vec u ) ))}$.
 

### 2014-12-15 tex/codatadp.tex
more

@@
 
-\newcommand\landg{L\,\&\,G}
+\newcommand\gandl{G\&L}
 \newcommand\HD[1]{\hbox to2.5em{\hfill#1\hfill}}
@@
 \begin{center}\begin{tabular}{l@{\kern1.5em}c@{\kern.5em}c@{\kern.5em}c@{\kern1.5em}c}
-  & \HD{Distro} & \HD{AFP} & \HD{\landg} & All
+  & \HD{Distro} & \HD{AFP} & \HD{\gandl} & All
 \\[1pt] %%% TYPESETTING HACK
@@
 within the supported fragment. We also included two unpublished theories by
-Andreas Lochbihler and Peter Gammie (\landg). We complemented the benchmarks
+Peter Gammie and Andreas Lochbihler (\gandl). We complemented the benchmarks
 with some theories about various list and tree data structures, to also test
@@
 %
-Andreas Lochbihler and Peter Gammie shared their private
+Peter Gammie and Andreas Lochbihler shared their private
 theories with us so we could include them in the benchmarks.

### 2014-12-15 tools/undata.c
more

@@ -256,17 +256,19 @@ int generate_one_file(const char *in_name, int keep_data, int keep_codata)
 
     strcpy(out_name, in_name);
 
+    char *dir = strstr(out_name, "both/");
+
     if (keep_data) {
         if (keep_codata) {
-            strcat(out_name, ".all");
+            memcpy(dir, "both", 4);
         } else {
-            strcat(out_name, ".data");
+            memcpy(dir, "data", 4);
         }
     } else {
         if (keep_codata) {
-            strcat(out_name, ".codata");
+            memcpy(dir, "coda", 4);
         } else {
-            strcat(out_name, ".nix");
+            memcpy(dir, "nada", 4);
         }
     }
 

### 2014-12-15 tex/codatadp.tex
more

@@
 No decision procedure
-  & 800 & 922 & 450 & 2172 \\
+  & 401 & 922 & \phantom{0}39 & 1122 \\
 Codatatypes without bisimilarity
-  & 810 & 936 & 470 & 2216 \\
+  & 410 & 936 & \phantom{0}45 & 1136 \\
 Codatatypes with bisimilarity
-  & 811 & 939 & 472 & 2222 \\
+  & 411 & 939 & \phantom{0}48 & 1139 \\
 Datatypes without cyclicity
-  & 815 & 947 & 469 & 2231 \\
+  & 415 & 947 & \phantom{0}43 & 1147 \\
 Datatypes with cyclicity
-  & 818 & 953 & 470 & 2241 \\
+  & 418 & 953 & \phantom{0}44 & 1153 \\
 Full decision procedure
-  & 840 & 970 & 490 & 2300 \\[\jot]
+  & 440 & 970 & \phantom{0}49 & 1170 \\[\jot]
 Total number of goals
-  & 1211\phantom{0} & 1121\phantom{0} & 899 & 3001
+  & 879 & 2974\phantom{0} & 117 & 3970
 \end{tabular}\end{center}

### 2014-12-15 tex/codatadp.tex
more

@@
 within the supported fragment. We also included two unpublished theories by
-Peter Gammie and Andreas Lochbihler (\gandl). We complemented the benchmarks
-with some theories about various list and tree data structures, to also test
-the datatypes.
-The theory files were selected before running the experiments.
-Our experimental data is publicly available \cite{our-eval-data}.
+Peter Gammie and Andreas Lochbihler (\gandl). To also exercise the the support
+for datatypes, the benchmarks are complemented by theories about various list
+and tree data structures. The theory files were selected before running the
+experiments. The experimental data is publicly available \cite{our-eval-data}.
 
@@
 CVC4, running for up to \textbf{TODO: XXX} seconds on the StarExec cluster
-\cite{xxx}. Due to the lack of machinery for reconstructing inferences about
-(co)datatypes in Isabelle, CVC4 is trusted as an oracle in these experiments.
+\cite{xxx}. Problems involving no (co)datatypes were filtered out.
+Due to the lack of machinery for reconstructing inferences about (co)datatypes
+in Isabelle, CVC4 is trusted as an oracle in these experiments.
 

### 2014-12-15 tex/codatadp.tex
more
Minor.

@@
 
-For any $\mu$-value $v$ of the form $\mu x. C( \vec w )$, we may recursively construct a normal value $\nf{v}$
-(which we will call the \emph{normal form} of $v$)
-that is a renaming of $v$ by replacing all of the bisimilar subterms of $v$ with $x$,
-and by replacing all of the non-normal subterms of $v$ with their corresponding normal forms.
+For any $\mu$-term $t$ of the form $\mu x. C( \vec u )$, we may recursively construct a normal $\mu$-term $\nf{t}$
+(which we will call the \emph{normal form} of $t$)
+that is a renaming of $v$ by replacing all of the bisimilar subterms of $t$ with $x$,
+and by replacing all of the non-normal subterms of $t$ with their corresponding normal forms.
 For instance, $\nf{(\mu x. \const{C}( \mu y. \const{C}( x ) ))} = \mu x. \const{C}( x )$.
@@
 We say that $t \vsimv{x} s$ if and only if $\mu$-terms $t$ and $s$ are syntactically equivalent
-for some renaming that avoids capturing variables besides $x$.
+for some renaming that avoids capturing any variable other than $x$.
 For instance, 
@@
 For the latter, this will be based on the mapping $\Val$ constructed in step two of our calculus.
-Given a normal model $\M$, the \emph{completion} of $\Val$, which we will write as $\Val^\ast$, 
-is the result of exhaustively applying the following update to $\Val$:
+Given a normal model $\M$, we compute the \emph{completion} of $\Val$, which we will write as $\Val^\ast$,
+by exhaustively applying the following update to $\Val$:
 
@@
 it can only be applied a finite number of times.
-We first state the following properties of $\Val$, 
+We then replace all values in the range of $\Val^\ast$ by their normal forms.
+We first state the following properties of $\Val^\ast$, 
 where we write $\muvar( t )$ to denote $t$ if $t$ is a variable, or $x$ if $t$ is of the form $\mu x. u$.
@@
 \begin{enumerate}
-\item $\nf{(\Val^\ast \ec{x})} \in \M( \tau )$ for all $\ec{x} : \tau$ of $\Ec$,
+\item $\Val^\ast \ec{x} \in \M( \tau )$ for all $\ec{x} : \tau$ of $\Ec$,
 \item
-$t_1 \interp{ t_2 } = \nf{(\Val^\ast \ec{x})}$
+$t_1 \interp{ t_2 } = \Val^\ast \ec{x}$
 for all $t_1\interp{t_2} \in \mathcal{T}(\Val)$ where $\muvar(t_2)=v_{\ec{x}}$,
 \item 
-$\nf{(\Val^\ast \ec{x})} \vsim \nf{(\Val^\ast \ec{y})}$ if and only if $\ec{x} = \ec{y}$.
+$\Val^\ast \ec{x} \vsim \Val^\ast \ec{y}$ if and only if $\ec{x} = \ec{y}$.
 \end{enumerate}
@@
 \begin{proof}
-To show 1, we must argue that $FV( \Val^\ast ) = \emptyset$, and that $\nf{(\Val^\ast \ec{x})}$ is not cyclic for
+To show 1, we must argue that $FV( \Val^\ast ) = \emptyset$, and that $\Val^\ast \ec{x}$ is not cyclic for
 $\ec{x} : \tau$ when $\tau \not\in \Codata$.
@@
 and interprets the variables of $\Fc$ based on the completion $\Val^\ast$ of the mapping $\Val$ from step two of our calculus.
-For the latter, let $\M( x )$ be the value in $\M$ that is a renaming of $\nf{(\Val^\ast \ec{x})}$ for each variable $x$ in $\mathcal{T}(\Fc)$.
+For the latter, let $\M( x )$ be the value in $\M$ that is a renaming of $\Val^\ast \ec{x}$ for each variable $x$ in $\mathcal{T}(\Fc)$.
 Notice that due to Lemma~\ref{lem:model-completion}.3, for variables $x, y \in FV( \Fc )$,
@@
 We first show $\M$ satisfies all equalities in $t_1 \teq t_2 \in \Fc$, where $t_1$ and $t_2$ have type $\tau$.
-We show that $\M( t ) = \Val^\ast \ec{t}$ for all terms $t \in \mathcal{T}( \Fc )$, which we show by induction on the size of $t$.
-
+We show,
+by induction on the size of $t$,
+that $\M( t ) \sim \Val^\ast \ec{t}$ for all terms $t \in \mathcal{T}( \Fc )$.
 If $t$ is a variable, then $\M( t ) = \Val^\ast \ec{ t }$ by construction.
 If $t$ is a constructor term $\const{C}( u_1, \ldots, u_n )$, 
-then since \rn{Unify$_2$} does not apply to $\Fc$,
-then $\Val^\ast \ec{t}$ must be a term of the form $\mu v_{\ec{t}}. \const{C}( w_1, \ldots, w_n )$,
-where say that $\muvar( w_i ) = v_{\ec{y_i}}$ for $i = 1, \ldots, n$.
-Since \rn{Unify$_1$} does not apply to $\Fc$, it must be that $\ec{u_i} = \ec{y_i}$ for $i = 1, \ldots, n $.
-%Due to Lemma~\ref{lem:model-completion}.2 and by induction
+then $\M( t )$ is a renaming of $\nf{(\mu x. \const{C}( \M( u_1 ), \ldots, \M( u_n ) ) )}$ for fresh $x$,
+which by the induction hypothesis is bisimilar to $\nf{(\mu x. \const{C}( \Val^\ast \ec{ u_1 }, \ldots, \Val^\ast \ec{ u_n } ) )}$, call this term $t_u$.
+Now, since \rn{Unify$_2$} does not apply to $\Fc$,
+then $\Val^\ast \ec{t}$ must be a term $t_w$ of the form $\mu v_{\ec{t}}. \const{C}( w_1, \ldots, w_n )$,
+that is, an application of the same construction $\const{C}$.
+For each $i = 1, \ldots, n$,
+say that $\muvar( w_i ) = v_{\ec{y_i}}$.
+Due to Lemma~\ref{lem:model-completion}.2, $t_w\interp{ w_i } = \Val^\ast \ec{ y_i }$.
+Due to our construction of $\Val^\ast$ and since \rn{Unify$_1$} does not apply to $\Fc$, we have $\ec{u_i} = \ec{y_i}$.
+
+If $t$ is a selector term $s^j_k( u )$,
+by the induction hypothesis, $\M( u ) \vsim \Val^\ast \ec{ u }$. 
+since \rn{Split} does not apply, 
+$\ec{u}$ must contain a term of the form $C_i( s^1_i( u ), \ldots, s^{a_i}_i( u ) )$ for some $i$.
+
+%we show that $w_i$ is bisimilar to the $i^{th}$ argument of
+%$\nf{(\mu x. \const{C}( \Val^\ast \ec{ u_1 }, \ldots, \Val^\ast \ec{ u_n } ) )}$.
+%In the case that $\mu
+%Since $t_c$ is normal
+
 \rem{TODO}
@@
 constructors and selectors, which the decision procedure would recognize as
-tautologies. This allows us to answer the question, \relax{What are the
-benefits of activating the decision procedure as opposed to letting
-Sledgehammer do what it would normally do?}
+tautologies.
+%This allows us to answer the question, \relax{What are the
+%benefits of activating the decision procedure as opposed to letting
+%Sledgehammer do what it would normally do?}
 

### 2014-12-15 tex/codatadp.tex
Finished draft of solution soundness, only lemmas remain.

@@
 excludes variables that occur beneath $\mu$-binders.
-It is straightforward to see that the height of terms produced as a result of this rule
+It is straightforward to see that the height of terms produced as a result of this update
 is bounded by the number of equivalence classes of $\Ec$,
@@
 the rules from Figure~\ref{fig:split-rule} are applied.
-The rule \rn{Singleton} handles the degenerate case when there are two terms from $\tEc$ having a singleton codatatype type,
+The rule \rn{Singleton} handles the case when there are two terms from $\tEc$ having a co-recursive singleton type,
 in which case they must be equivalent.
@@
 
+We now define a class of models that we consider.
+\rem{Clarify that non-codatatype terms are interpreted as $\mu$-terms, although dropping the $\mu$'s gives the standard representation?}
+
 \begin{definition}[Normal Model] \label{def:norm-model}
@@
 \begin{lemma}
+\label{lem:mu-norm-arg}
+If $\vec u$ are normal, then if $t = \nf{(\mu x. \const{C}( \vec u ) )} = \mu x. \const{C}( \vec w )$,
+then $\vec u \vsim t\interp{\vec w}$.
+\end{lemma}
+
+\begin{lemma}
+\label{lem:mu-cong}
+If $t_u = \mu x. \const{C}( \vec u )$ and $t_w = \mu y. \const{C}( \vec w )$ are normal,
+then $t_u\interp{ \vec u } \vsim t_w\interp{ \vec w }$ if and only if $t_u \vsim t_w$.
+\end{lemma}
+
+\begin{lemma} 
 \label{lem:model-completion}
@@
 \begin{proof}
-Let $\Fc_0$ be a saturated node in a derivation tree with root node $\Ec$.
-We will construct a model $\M$ for a set of equalities $\Fc$ that is equivalent to $\Fc_0$, and where 
-all equivalence classes of $\Fc$ contain at least one variable, and all selectors in $\Fc$ are applied to variables only.
-The former comes with no loss of generality since new equalities of the form $y \teq t$ for fresh variable $y$ can be added to $\Fc$ without affecting its satisfiability,
-The latter also comes with no loss of generality since nested applications of selectors in $\Fc_0$ can be replaced by fresh variables while adding additional equalities to $\Fc$.
-
+Let $\Fc$ be a saturated node in a derivation tree with root node $\Ec$.
+%We will construct a model $\M$ for a set of equalities $\Fc$ that is equivalent to $\Fc_0$, and where 
+%all equivalence classes of $\Fc$ contain at least one variable, and all selectors in $\Fc$ are applied to variables only.
+%The former comes with no loss of generality since new equalities of the form $y \teq t$ for fresh variable $y$ can be added to $\Fc$ without affecting its satisfiability,
+%The latter also comes with no loss of generality since nested applications of selectors in $\Fc_0$ can be replaced by fresh variables while adding additional equalities to $\Fc$.
 We consider a normal model $\M$ 
@@
 and interprets the variables of $\Fc$ based on the completion $\Val^\ast$ of the mapping $\Val$ from step two of our calculus.
-For the latter, let $\M( x )$ be the value in $\M$ that is a renaming of $\Val^\ast \ec{x}$ for each variable $x$ in $\mathcal{T}(\Fc)$.
-Notice that due to Lemma~\ref{lem:model-completion}.3, for variables $x, y \in FV( \Fc )$,
-\begin{eqnarray}
-\label{eqn:m-bij} 
-\M( x ) = \M( y ) \text{ if and only if } \ec{x} = \ec{y}
-\end{eqnarray}
+For the latter, let $\M( x )$ be the value in $\M$ that is bisimilar with $\Val^\ast \ec{x}$ for each variable $x$ in $\mathcal{T}(\Fc)$.
 
@@
 We show,
-by induction on the size of $t$,
-that $\M( t ) \sim \Val^\ast \ec{t}$ for all terms $t \in \mathcal{T}( \Fc )$.
-If $t$ is a variable, then $\M( t ) = \Val^\ast \ec{ t }$ by construction.
-If $t$ is a constructor term $\const{C}( u_1, \ldots, u_n )$, 
-then $\M( t )$ is a renaming of $\nf{(\mu x. \const{C}( \M( u_1 ), \ldots, \M( u_n ) ) )}$ for fresh $x$,
-which by the induction hypothesis is bisimilar to $\nf{(\mu x. \const{C}( \Val^\ast \ec{ u_1 }, \ldots, \Val^\ast \ec{ u_n } ) )}$, call this term $t_u$.
-Now, since \rn{Unify$_2$} does not apply to $\Fc$,
+by structural induction on $t$,
+that $\M( t ) \sim \Val^\ast \ec{t}$ for all terms $t \in \mathcal{T}( \Fc )$,
+which implies $\M \models t_1 \teq t_2$.
+
+If $t$ is a variable, then $\M( t ) \vsim \Val^\ast \ec{ t }$ by construction.
+
+If $t$ is a constructor term of the form $\const{C}( u_1, \ldots, u_n )$, 
+then $\M( t )$ is bisimilar with $\nf{(\mu x. \const{C}( \M( u_1 ), \ldots, \M( u_n ) ) )}$ for fresh $x$,
+which by the induction hypothesis is bisimilar with $\nf{(\mu x. \const{C}( \Val^\ast \ec{ u_1 }, \ldots, \Val^\ast \ec{ u_n } ) )}$, call this term $t_u$.
+Since \rn{Unify$_2$} does not apply to $\Fc$,
 then $\Val^\ast \ec{t}$ must be a term $t_w$ of the form $\mu v_{\ec{t}}. \const{C}( w_1, \ldots, w_n )$,
-that is, an application of the same construction $\const{C}$.
-For each $i = 1, \ldots, n$,
-say that $\muvar( w_i ) = v_{\ec{y_i}}$.
-Due to Lemma~\ref{lem:model-completion}.2, $t_w\interp{ w_i } = \Val^\ast \ec{ y_i }$.
-Due to our construction of $\Val^\ast$ and since \rn{Unify$_1$} does not apply to $\Fc$, we have $\ec{u_i} = \ec{y_i}$.
-
-If $t$ is a selector term $s^j_k( u )$,
-by the induction hypothesis, $\M( u ) \vsim \Val^\ast \ec{ u }$. 
+that is, an application of the same constructor.
+For each $i = 1, \ldots, n$, let $u_i'$ be the $i^{th}$ argument of $t_u$.
+By Lemma~\ref{lem:mu-norm-arg}, $t_u\interp{u_i'} \vsim \Val^\ast \ec{u_i}$.
+Say that $\muvar( w_i ) = v_{\ec{y_i}}$.
+Due to our construction of $\Val^\ast$ and since \rn{Unify$_1$} does not apply to $\Fc$, we have $\ec{u_i} = \ec{y_i}$,
+and thus $t_u\interp{u_i'} \vsim \Val^\ast \ec{y_i}$.
+By Lemma~\ref{lem:model-completion}.2, $\Val^\ast \ec{ y_i } = t_w\interp{ w_i }$,
+and thus $t_u\interp{u_i'} \vsim t_w\interp{ w_i }$ for each $i = 1, \ldots, n$.
+By Lemma~\ref{lem:mu-cong}, we have $t_u \vsim t_w$.
+Since $\vsim$ is a transitive relation and since $\M( t ) \vsim t_u \vsim t_w = \Val^\ast \ec{t}$,
+we have $\M( t ) \vsim \Val^\ast \ec{t}$.
+
+If $t$ is a selector term of the form $s^j_k( u )$,
 since \rn{Split} does not apply, 
-$\ec{u}$ must contain a term of the form $C_i( s^1_i( u ), \ldots, s^{a_i}_i( u ) )$ for some $i$.
-
-%we show that $w_i$ is bisimilar to the $i^{th}$ argument of
-%$\nf{(\mu x. \const{C}( \Val^\ast \ec{ u_1 }, \ldots, \Val^\ast \ec{ u_n } ) )}$.
-%In the case that $\mu
-%Since $t_c$ is normal
-
-\rem{TODO}
-
-%If $\tau \in \Nondata$, $t_1$ and $t_2$ must be variables, and thus $\M( t_1 ) = \M( t_2 )$ by construction.
-%If $\tau \in \Data \cup \Codata$, we show $\M \models t_1 \teq t_2$ by showing 
-%We show by induction on the term size of $t$ that $\M( t ) = \Val^\ast \ec{t}$.
-%If $t$ is a variable, then $\M( t ) = \Val^\ast \ec{ t } = \Val^\ast \ec{ x }$.
-%In the case where 
-%where say $\muvar( v_i ) = v_{\ec{y_i}}$ for $i = 1, \ldots, n$.
-%We have $\M( \const{C}( \vec u ) ) = \nf{(\mu y. \const{C}( \M( \vec u ) ))}$.
+$\ec{u}$ must contain a term of the form $C_i( s^1_i( u ), \ldots, s^{n}_i( u ) )$ for some $i$.
+By our construction and since \rn{Unify$_1$} and \rn{Unify$_2$} do not apply to $\Fc$, 
+$\Val^\ast \ec{ u }$ must be of the form $t_u = \mu v_{\ec{u}}. C_i( u_1, \ldots, u_n )$,
+where $\muvar( u_1 ) = v_{\ec{s^1_i(u)}}$, $\ldots$, $\muvar(u_n) = v_{\ec{s^n_i( u )}}$.
+Moreover, $\M( u )$ is of the form $t_w = \mu x. C_i( w_1, \ldots, w_n )$,
+which by the induction hypothesis is bisimilar with $\Val^\ast \ec{u}$.
+If $k = i$, then $\M( t )$ is bisimilar with $t_w\interp{w_j}$,
+which by Lemma~\ref{lem:mu-cong} is bisimilar with $t_u\interp{u_j}$,
+which by Lemma~\ref{lem:model-completion}.2 is equal to $\Val^\ast \ec{ s^j_k( u )}$.
+Thus, $\M( t ) \vsim \Val^\ast \ec{ s^j_k( u )} = \Val^\ast \ec{t}$.
+If $k \neq i$, then we interpret $s^j_k( \M( u ) )$ as the value in $\M( \tau )$ that is bisimilar with $\Val^\ast \ec{ t }$.
+Since \rn{Cong} does not apply and due to Lemma~\ref{lem:model-completion}.3, 
+pairs of wrongly-applied selectors are assigned the same value if and only if they reside in the same equivalence class,
+and thus this definition is well-defined.
 
@@
 Say $t_1 \tneq t_2 \in \Fc$.
-Since \rn{Conflict} does not apply, we know $t_1 \teq t_2 \not\in \Fc$ and thus $\ec{t_1}$ and $\ec{t_2}$ are distinct.
-By the previous paragraph and statement~\ref{eqn:m-bij}, $\M( t_1 ) \neq \M( t_2 )$, and thus $\M \models t_1 \tneq t_2$.
+Since \rn{Conflict} does not apply, $t_1 \teq t_2 \not\in \Fc$ and thus $\ec{t_1}$ and $\ec{t_2}$ are distinct.
+Since $\M( t_1 ) \vsim \Val^\ast \ec{t_1}$, $\M( t_2 ) \vsim \Val^\ast \ec{t_2}$
+and Lemma~\ref{lem:model-completion}.3, $\M( t_1 ) \neq \M( t_2 )$, and thus $\M \models t_1 \tneq t_2$.
 

### 2014-12-15 tex/codatadp.tex
more

@@
 \\[-9pt] %%% TYPESETTING HACK
-No decision procedure
+No (co)datatypes
   & 401 & 922 & \phantom{0}39 & 1122 \\
@@
   & 418 & 953 & \phantom{0}44 & 1153 \\
-Full decision procedure
+Full (co)datatypes
   & 440 & 970 & \phantom{0}49 & 1170 \\[\jot]

### 2014-12-15 tex/codatadp.tex
larger font
minor edits

@@
 bisimilarity rule that identifies equal values, also in the presence of cyclic
-($\omega$-regular) data. The procedure is complete for ground and universal
+($\omega$-regular) data. The procedure decides ground and universal
 problems and is composable via the Nelson--Oppen method. It has been
@@
 
-\section{A Ground Decision Procedure for (Co)datatypes}
+\section{The Decision Procedure} % for (Co)datatypes}
 \label{sec:a-ground-decision-procedure-for-co-datatypes}
@@
 %We will commonly denote tuples of terms $( t_1, \ldots, t_n )$ in bold font, as $\vec t$.
-We assume that all $\Sig$-terms in $\Ec$ are \emph{normalized}, meaning that all subterms of the form $s^i_j( \const{C}_j( \vec t ) )$
-are replaced by $t_i$, and moreover assume that all additional constraints added to $\Ec$ are normalized in this manner.
-
-We present the calculus in three parts.
-In the first part we compute the bidirectional closure of $\Ec$,
-in the second part we make inferences based on cyclicity and bisimilarity,
-and in the third part (when necessary) we branch on constructor types for various terms in $\tEc$.
+We assume that all $\Sig$-terms in $\Ec$ are \emph{normalized}, meaning that all subterms of the form $\const{s}^{\,k}_j( \const{C}_j(t_1,\ldots,t_n) )$
+are simplified to $t_i$, and moreover assume that all %additional
+constraints added to $\Ec$ are normalized in this manner.
+
+We present the calculus in three steps.
+In the first step, we compute the bidirectional closure of $\Ec$;
+in the second step, we make inferences based on cyclicity and bisimilarity;
+and in the third step, %(when necessary),
+we branch on constructor types for various terms in $\tEc$.
 Following the conventions from~\cite{}, 
@@
 where a rule can be applied to $\Ec$ if it meets all of the specified preconditions for $\Ec$.
-The conclusion of a rule either describes equalities to be added to $\Ec$ (in which case we will call an application of it to be \emph{non-terminal}),
+The conclusion of a rule either describes equalities to be added to $\Ec$ (in which case we will call an application of it to be \emph{nonterminal}),
 or is $\bot$ (in which case we will call an application of it to be \emph{terminal}).
-A rule may have multiple conclusions separated by $\parallel$, which denotes a non-deterministic branching.
+A rule may have multiple conclusions separated by $\parallel$, which denotes nondeterministic branching.
 
@@
 \end{tabular}
-\caption{Rules for bidirectional closure.
+\caption{Rules for bidirectional closure%.
 }
@@
 
-\paragraph{Step 1: Compute Birectional Closure}
+\paragraph{Step 1: Computing the Bidirectional Closure.}
 Figure~\ref{fig:cc-rules} gives the basic rules of our calculus.
-Together with the rules \rn{Refl}, \rn{Symm}, \rn{Trans}, the rule \rn{Cong} computes the (upwards) congruence closure,
-while the rules \rn{Unify$_1$} and \rn{Unify$_2$} together compute (downwards) unification.
+Together with the rules \rn{Refl}, \rn{Symm}, \rn{Trans}, the rule \rn{Cong} computes the (upward) congruence closure,
+while the rules \rn{Unify$_1$} and \rn{Unify$_2$} together compute (downward) unification.
 In the unification rules, additional equalities are inferred based on the injectivity of constructors by \rn{Unify$_1$},
@@
 \end{tabular}
-\caption{Rules for cyclicity and bisimilarity.
+\caption{Rules for cyclicity and bisimilarity%.
 }
@@
 
-\paragraph{Step 2: Apply cyclicity and bisimilarity.}
+\paragraph{Step 2: Applying Cyclicity and Bisimilarity.}
 In this step,
 we assume that $\Ec$ is saturated with respect to the rules in Figure~\ref{fig:cc-rules}, that is,
-no non-redudant application of one of its rules can be applied to $\Ec$.
+no nonredudant application of one of its rules can be applied to $\Ec$.
 When this is the case, it is easy to see that $\Ec$ induces an equivalence relation over $\tEc$ such that two terms $t_1$ and $t_2$ are equivalent if and only if $t_1 \teq t_2 \in \Ec$.
@@
 \end{tabular}
-\caption{Rules for branching.  
+\caption{Rules for branching%.  
 %All pairs of terms whose type has cardinality 1 are entailed to be equal (the degenerate case).
@@
 
-An application of a rule is \emph{redundant} if it is non-terminal and at least one branch in its conclusion does not add a new equality to $\Ec$.
+An application of a rule is \emph{redundant} if it is nonterminal and at least one branch in its conclusion does not add a new equality to $\Ec$.
 A \emph{derivation tree} is a tree whose nodes are sets of equalities, where non-root nodes vtained by 
-a non-redundant application of a derivation rule to its parent node.
+a nonredundant application of a derivation rule to its parent node.
 A derivation tree is \emph{closed} if all of its leaf nodes are $\bot$.
-A node is \emph{saturated} if no non-redundant application of a rule can be applied to it.
+A node is \emph{saturated} if no nonredundant application of a rule can be applied to it.
 If there exists a closed derivation tree with root node $\Ec$, then we conclude that $\Ec$ is $\thD$-unsatisfiable.
@@
 that is a renaming of $v$ by replacing all of the bisimilar subterms of $t$ with $x$,
-and by replacing all of the non-normal subterms of $t$ with their corresponding normal forms.
+%and by replacing all of the nonnormal subterms of $t$ with their corresponding normal forms.
+and by normalizing $t$'s subterms.
 For instance, $\nf{(\mu x. \const{C}( \mu y. \const{C}( x ) ))} = \mu x. \const{C}( x )$.
@@
 
-\begin{table}[tbh!]
+\begin{table*}[tbh!]
+\normalsize
 \begin{center}\begin{tabular}{l@{\kern1.5em}c@{\kern.5em}c@{\kern.5em}c@{\kern1.5em}c}
@@
 \label{tab:bench}
-\end{table}
+\end{table*}
 

### 2014-12-15 tex/codatadp.tex
played with headings
More reorg, prove model construction produces values.

--- 2014-12-15 tex/codatadp.tex
@@
 We will often write $t_1\interp{t_2}$ as shorthand for $t_1\interp{t_2}^\emptyset$.
-We will write $\mathcal{T}( t_1 )$ to denote the set consisting of $t_1\interp{t_2}$ for all subterms $t_2$ of $t_1$,
-and write $\mathcal{T}( S )$ to denote the union of $\mathcal{T}( t )$ for each $t \in S$ where $S$ is a set of terms.
 
@@
 
-We say that $t \vsimv{x} s$ if and only if $\mu$-terms $t$ and $s$ are syntactically equivalent
-for some renaming that avoids capturing any variable other than $x$.
-For instance, 
-$\mu x. \const{D}( x ) \vsimv{y} \mu x. \const{D}( y )$,
-$\mu x. \const{C}( x, x ) \vsimv{x} \mu y. \const{C}( x, y )$ and
-$\mu x. \const{C}( z, x ) \vsimv{z} \mu y. \const{C}( z, y )$,
-but
-$\mu x. \const{D}( x ) \not\vsimv{x} \mu x. \const{D}( y )$ and
-$\mu x. \const{C}( x, x ) \not\vsimv{y} \mu y. \const{C}( x, y )$.
-
-%\begin{lemma}
-%If $t_1$ and $u$ are normal, and if $u \not\vsimv{x} t_1\interp{t_2}$ for all $t_2 \in \mathcal{T}( t_1 )$,
-%then $t_1 \{ x \mapsto u \}$ is normal.
-%\end{lemma}
-%\begin{proof}
-%\rem{TODO}
-%\end{proof}
+\begin{lemma}
+\label{lem:mu-norm-arg}
+If $\vec u$ are normal, then if $t = \nf{(\mu x. \const{C}( \vec u ) )} = \mu x. \const{C}( \vec w )$,
+then $\vec u \vsim t\interp{\vec w}$.
+\end{lemma}
+\begin{proof}
+%Since $u_i$ is normal, $w_i$ is the result of replacing all subterms $v$ of $u_i$ such that $t\interp{v} \sim t$ with $x$.
+\rem{TODO}
+\end{proof}
+
+\begin{lemma}
+\label{lem:mu-cong}
+If $t_u = \mu x. \const{C}( \vec u )$ and $t_w = \mu y. \const{C}( \vec w )$ are normal,
+then $t_u\interp{ \vec u } \vsim t_w\interp{ \vec w }$ if and only if $t_u \vsim t_w$.
+\end{lemma}
+\begin{proof}
+\rem{TODO}
+\end{proof}
 
@@
 For the latter, this will be based on the mapping $\Val$ constructed in step two of our calculus.
-Given a normal model $\M$, we compute the \emph{completion} of $\Val$, which we will write as $\Val^\ast$,
-by exhaustively applying the following update to $\Val$:
+We compute the \emph{completion} of $\Val$ for normal model $\M$, which we will write as $\Val^\ast$,
+which we construct by assigning values from $\M$ to unassigned variables in the domain of $\Val$.
+
+First, we need the following definition.
+We say that $t \vsimv{x} s$ if and only if $\mu$-terms $t$ and $s$ are syntactically equivalent
+for some renaming that avoids capturing any variable other than $x$.
+For instance, 
+$\mu x. \const{D}( x ) \vsimv{y} \mu x. \const{D}( y )$,
+$\mu x. \const{C}( x, x ) \vsimv{x} \mu y. \const{C}( x, y )$ and
+$\mu x. \const{C}( z, x ) \vsimv{z} \mu y. \const{C}( z, y )$,
+but
+$\mu x. \const{D}( x ) \not\vsimv{x} \mu x. \const{D}( y )$ and
+$\mu x. \const{C}( x, x ) \not\vsimv{y} \mu y. \const{C}( x, y )$.
+For a variable $x$ of type $\tau$ and a fixed normal model $\M$,
+we write $\mathcal{T}^\M_x( \Val )$ to denote the set consisting of all values $v$ from $\M( \tau )$
+such that $v \vsim_x t_1\interp{t_2}$ for some term $t_1$ with subterm $t_2$ occurring in the range of $\Val$.
+Our completion $\Val^\ast$ is constructed by exhaustively applying the following update to $\Val$:
 
@@
   \quad
-  \mu x. t \sim v, v \in \M( \tau )
+  (\mu x. t) \sim v
   \quad
-  (\mu x. t) \not\sim_x u \text{ for all } u \in \mathcal{T}(\vrange(\Val))
+  v \in \M( \tau )
+  \quad
+  v \not\in \mathcal{T}^\M_x(\Val)
 }{
@@
 
-Since this rule removes one variable from the set $FV( \Val )$ and does not add any additional variables to $FV( \Val )$, 
+In this update, we choose an unassigned variable in $\Val$
+and assign it a value that is not bisimilar to one occurring in $\mathcal{T}^\M_x(\Val)$.
+Since this update removes one variable from the set $FV( \Val )$ and does not add any additional variables to $FV( \Val )$, 
 it can only be applied a finite number of times.
-We then replace all values in the range of $\Val^\ast$ by their normal forms.
+We then replace $\Val^\ast \ec{x}$ with $\nf{(\Val^\ast \ec{x})}$ for all $\ec{x}$.
 We first state the following properties of $\Val^\ast$, 
 where we write $\muvar( t )$ to denote $t$ if $t$ is a variable, or $x$ if $t$ is of the form $\mu x. u$.
-%which amounts to saying that 
-%equivalence class are interpreted as values,
-%the interpretation of an equivalence class is not sensitive to the context in which it is interpreted,
-%and that each equivalence classes has a distinct interpretation.
-
-\begin{lemma}
-\label{lem:mu-norm-arg}
-If $\vec u$ are normal, then if $t = \nf{(\mu x. \const{C}( \vec u ) )} = \mu x. \const{C}( \vec w )$,
-then $\vec u \vsim t\interp{\vec w}$.
-\end{lemma}
-
-\begin{lemma}
-\label{lem:mu-cong}
-If $t_u = \mu x. \const{C}( \vec u )$ and $t_w = \mu y. \const{C}( \vec w )$ are normal,
-then $t_u\interp{ \vec u } \vsim t_w\interp{ \vec w }$ if and only if $t_u \vsim t_w$.
-\end{lemma}
 
@@
 \label{lem:model-completion}
-The following are true of $\Val^\ast$:
+If $\Val$ is constructed for a saturated $\Ec$,
+and $\Val^\ast$ is a completion of $\Val$ for normal model $\M$, then:
 \begin{enumerate}
-\item $\Val^\ast \ec{x} \in \M( \tau )$ for all $\ec{x} : \tau$ of $\Ec$,
+\item $\Val^\ast \ec{x}$ is bisimilar with a value in $\M( \tau )$, for all $\ec{x} : \tau$ of $\Ec$,
 \item
 $t_1 \interp{ t_2 } = \Val^\ast \ec{x}$
-for all $t_1\interp{t_2} \in \mathcal{T}(\Val)$ where $\muvar(t_2)=v_{\ec{x}}$,
+for all $t_1$ with subterm $t_2$ in the range of $\Val^\ast$ where $\muvar(t_2)=v_{\ec{x}}$,
 \item 
@@
 \begin{proof}
-To show 1, we must argue that $FV( \Val^\ast ) = \emptyset$, and that $\Val^\ast \ec{x}$ is not cyclic for
-$\ec{x} : \tau$ when $\tau \not\in \Codata$.
+To show 1, we first show that $\Val^\ast$ contains no free variables.
+Assume by contradiction that $\Val^\ast$ contained a free variable $v_{\ec{y}}$ for some $\ec{y}$ of type $\tau$.
+Then it must be the case that $\ec{y}$ does not contain a constructor term,
+or else $v_{\ec{y}}$ would not occur as a free variable in $\Val$.
+Consider the case when $\tau$ is finite.
+By assumption, $\tau \not\in \Nondata$.
+Since \rn{Split} does not apply to $\Ec$, we have $\tau \not\in \Data$.
+If $\tau \in \Codata$, then by Lemma~\ref{lem:corecursive-singletons},
+the cardinality of $\tau$ must be one (call the value in its domain $c$).
+Since \rn{Singleton} does not apply to $\Ec$,
+there is only one equivalence class of type $\tau$ in $\Ec$,
+and thus there are no terms in $\mathcal{T}(\Val^\ast)$ of type $\tau$.
+This is a contradiction, since our model completion may assign $c$ to $v_{\ec{x}}$.
+Now, consider the case when $\tau$ is infinite.
+This is also a contradiction, 
+since there are only a finite number of closed terms in $\mathcal{T}^\M_x(\Val^\ast)$,
+and thus our model completion can assign a value not occurring in $\mathcal{T}^\M_x(\Val^\ast)$ to $v_{\ec{y}}$.
+By construction, $\Val^\ast \ec{x}$ is normal.
+Since \rn{Cyclic} does not apply, $\Val \ec{x}$ is acyclic when $\tau \in \Data$.
+Moreover, our construction of $\Val^\ast$ applies substitutions of the form 
+$\{ y \mapsto t \}$, where $t$ is acyclic when the type of $y$ is not a codatatype.
+Thus, $\Val^\ast \ec{x}$ is acyclic when $\tau \not\in \Codata$.
+Therefore, by definition, $\Val^\ast \ec{x}$ is bisimilar with a value in $\M( \tau )$.
+
 \rem{TODO}
@@
 $\ec{u}$ must contain a term of the form $C_i( s^1_i( u ), \ldots, s^{n}_i( u ) )$ for some $i$.
-By our construction and since \rn{Unify$_1$} and \rn{Unify$_2$} do not apply to $\Fc$, 
+By our construction of $\Val^\ast$ and since \rn{Unify$_1$} and \rn{Unify$_2$} do not apply to $\Fc$, 
 $\Val^\ast \ec{ u }$ must be of the form $t_u = \mu v_{\ec{u}}. C_i( u_1, \ldots, u_n )$,
-where $\muvar( u_1 ) = v_{\ec{s^1_i(u)}}$, $\ldots$, $\muvar(u_n) = v_{\ec{s^n_i( u )}}$.
-Moreover, $\M( u )$ is of the form $t_w = \mu x. C_i( w_1, \ldots, w_n )$,
+where $\muvar( u_1 ) = v_{\ec{s^1_i(u)}}$, $\ldots$, $\muvar(u_n) = v_{\ec{s^n_i( u )}}$,
+and $\M( u )$ is of the form $t_w = \mu x. C_i( w_1, \ldots, w_n )$,
 which by the induction hypothesis is bisimilar with $\Val^\ast \ec{u}$.
 If $k = i$, then $\M( t )$ is bisimilar with $t_w\interp{w_j}$,
-which by Lemma~\ref{lem:mu-cong} is bisimilar with $t_u\interp{u_j}$,
+which since $t_w \vsim t_u$ by Lemma~\ref{lem:mu-cong} is bisimilar with $t_u\interp{u_j}$,
 which by Lemma~\ref{lem:model-completion}.2 is equal to $\Val^\ast \ec{ s^j_k( u )}$.
@@
 
-By Lemmas~\ref{lem:t},~\ref{lem:rs}, and~\ref{lem:ss}, the calculus described in this section is sound and complete for $\thD$.
+By Lemmas~\ref{lem:t},~\ref{lem:rs}, and~\ref{lem:ss}, the calculus is sound and complete for $\thD$.
 

### 2014-12-15 tex/codatadp.tex
more

--- 2014-12-15 tex/codatadp.tex
@@
 
-\section{The Decision Procedure} % for (Co)datatypes}
+\section{The Ground Decision Procedure} % for (Co)datatypes}
 \label{sec:a-ground-decision-procedure-for-co-datatypes}
@@
 \begin{tabular}{c}
-\rn{Refl}
 \(
@@
 }{
-  \Ec := \Ec, t \teq t
+  \Ec := \Ec,\: t \teq t
 }
 \)
+\rn{Refl}
 \qquad
-\rn{Symm}
 \(
@@
 }{
- \Ec := \Ec, t_2 \teq t_1
+ \Ec := \Ec,\: t_2 \teq t_1
 }
 \)
+\rn{Symm}
 \qquad
-\rn{Conflict}
 \(
 \inferrule{
-  t_1 \teq t_2, t_1 \tneq t_2 \in \Ec
+  t_1 \teq t_2,\; t_1 \tneq t_2 \in \Ec
 }{
@@
 \)
-\\[3.7ex]
-\rn{Trans}
+\rn{Conflict}
+\\[5\jot]
 \(
 \inferrule{
-  t_1 \teq t_2, t_2 \teq t_3 \in \Ec
+  t_1 \teq t_2,\; t_2 \teq t_3 \in \Ec
 }{
-  \Ec := \Ec, t_1 \teq t_3
+  \Ec := \Ec,\: t_1 \teq t_3
 }
 \)
+\rn{Trans}
 \qquad
-\rn{Cong} 
 \(
 \inferrule{
-  \vec t \teq \vec u \in \Ec \quad f( \vec t ), f( \vec u ) \in \tEc
+  \vec t \teq \vec u \in \Ec \quad \const f( \vec t \,),\, \const f( \vec u ) \in \tEc
 }{
-  \Ec := \Ec, f( \vec t ) \teq f( \vec u )
+  \Ec := \Ec,\: \const f( \vec t \,) \teq \const f( \vec u )
 }
 \)
-\\[3.7ex]
-\rn{Unify$_1$} 
+\rn{Cong} 
+\\[5\jot]
 \(
 \inferrule{
-  \const{C}_1( \vec t ) \teq \const{C}_1( \vec u ) \in \Ec
+  \const{C}_1( \vec t \,) \teq \const{C}_1( \vec u ) \in \Ec
 }{
-  \Ec := \Ec, \vec t \teq \vec u
+  \Ec := \Ec,\: \vec t \teq \vec u
 }
 \)
+\rn{Unify$_1$} 
 \qquad
-\rn{Unify$_2$} 
 \(
 \inferrule{
-  \const{C}_1( \vec t ) \teq \const{C}_2( \vec u ) \in \Ec
+  \const{C}_1( \vec t \,) \teq \const{C}_2( \vec u ) \in \Ec
 }{
@@
 \)
+\rn{Unify$_2$} 
 \end{tabular}
@@
 \begin{tabular}{c}
-\rn{Cyclic}
 \(
 \inferrule{
+  \tau \in \Data
+  \quad
   t_1 : \tau
   \quad
-  \tau \in \Data
-  \quad
   %\ttpath{\Ec}{\ec{t}}{\ec{t}} \neq \emptyset
-  \Val \ec{t_1} = \mu x. t_2
+  \Val \ec{t_1} = (\vthinspace\mu x.\; t_2)
   \quad
@@
 \)
-\\[3.7ex]
-\rn{Bisimilar}
+\rn{Cyclic}
+\\[5\jot]
 \(
 \inferrule{
+ \tau \in \Codata
+ \quad
  t_1, t_2 : \tau
- \quad
- \tau \in \Codata
  \quad
@@
 }{
- \Ec := \Ec, t_1 \teq t_2
+ \Ec := \Ec,\: t_1 \teq t_2
 }
 \)
+\rn{Bisimilar}
 \end{tabular}
@@
 Formally, we recursively define a $\mu$-term as a term-like structure either of the form
-$\mu x. \const{C}( \vec t )$ for some constructor $\const{C} \in \Ctr$ and $\mu$-terms $\vec t$, 
+$\mu x. \const{C}( \vec t\, )$ for some constructor $\const{C} \in \Ctr$ and $\mu$-terms $\vec t$, 
 or a variable $x$.
@@
 \begin{tabular}{c}
-\rn{Singleton}
 \(
@@
 \)
-\\[3.7ex]
-\rn{Split} 
+\rn{Singleton}
+\\[5\jot]
 \(
@@
 \)
+\rn{Split} 
 \end{tabular}
@@
 \item 
-for all constructor terms $\const{C}( \vec t )$ of type $\tau$,
-$\M( \const{C}( \vec t ) )$ is the value 
+for all constructor terms $\const{C}( \vec t \,)$ of type $\tau$,
+$\M( \const{C}( \vec t \,) )$ is the value 
 in $\M( \tau )$ that is a renaming of 
-$\nf{(\mu x. \const{C}( \M( \vec t ) ))}$ where $x$ is fresh, and
+$\nf{(\mu x. \const{C}( \M( \vec t \,) ))}$ where $x$ is fresh, and
 \item 
@@
 
-\section{Implementation}
+\section{Implementation as a Theory Solver in CVC4}
 \label{sec:the-theory-solver}
@@
 
-\section{Experimental Results}
+\section{Evaluation on Isabelle Problems}
 \label{sec:experimental-results}

### 2014-12-15 tex/codatadp.tex
tiny spaces
renamed rules

@@
 \inferrule{
-  \const{C}_1( \vec t \,) \teq \const{C}_1( \vec u ) \in \Ec
+  \const{C}( \vec t \,) \teq \const{C}( \vec u ) \in \Ec
 }{
@@
 \)
-\rn{Unify$_1$} 
+\rn{Decomp} 
 \qquad
@@
 \inferrule{
-  \const{C}_1( \vec t \,) \teq \const{C}_2( \vec u ) \in \Ec
+  \const{C}( \vec t \,) \teq \const{D}( \vec u ) \in \Ec
+  \quad
+  \const{C} \not= \const{D}
 }{
@@
 \)
-\rn{Unify$_2$} 
+\rn{Clash}
 \end{tabular}
-\caption{Rules for bidirectional closure%.
+\caption{\,Rules for bidirectional closure%.
 }
@@
 Together with the rules \rn{Refl}, \rn{Symm}, \rn{Trans}, the rule \rn{Cong} computes the (upward) congruence closure,
-while the rules \rn{Unify$_1$} and \rn{Unify$_2$} together compute (downward) unification.
-In the unification rules, additional equalities are inferred based on the injectivity of constructors by \rn{Unify$_1$},
-and failures to unify equated terms are recognized by \rn{Unify$_2$}.
+while the rules \rn{Decomp} and \rn{Clash} together compute (downward) unification.
+In the unification rules, additional equalities are inferred based on the injectivity of constructors by \rn{Decomp},
+and failures to unify equated terms are recognized by \rn{Clash}.
 The rule \rn{Conflict} recognizes when an equality and its negation both occur in $\Ec$, in which case $\Ec$ has no model.
@@
 \end{tabular}
-\caption{Rules for cyclicity and bisimilarity%.
+\caption{\,Rules for cyclicity and bisimilarity%.
 }
@@
 \end{tabular}
-\caption{Rules for branching%.  
+\caption{\,Rules for branching%.  
 %All pairs of terms whose type has cardinality 1 are entailed to be equal (the degenerate case).
@@
 We prove this by structural induction on the (closed) derivation tree with root node $\Ec$.
-First, if the derivation tree is an application of \rn{Conflict}, \rn{Unify$_2$}, or \rn{Cyclic},
+First, if the derivation tree is an application of \rn{Conflict}, \rn{Clash}, or \rn{Cyclic},
 then $\Ec$ is unsatisfiable in $\thD$.
 For \rn{Conflict}, this is a consequence of equality reasoning.
-For \rn{Unify$_2$}, this is a consequence of distinctness.
+For \rn{Clash}, this is a consequence of distinctness.
 For \rn{Cyclic}, our construction of $\Val$ indicates that the class of values that $t_1$ can take in models of $\Ec$ is infeasible,
@@
 Otherwise, the child of $\Ec$ is a closed derivation tree 
-whose root node is $\Ec \cup t_1 \teq t_2$ obtained by applying one of the rules \rn{Refl}, \rn{Symm}, \rn{Trans}, \rn{Cong}, \rn{Unify$_1$}, \rn{Bisimilar}, or \rn{Singleton}.
+whose root node is $\Ec \cup t_1 \teq t_2$ obtained by applying one of the rules \rn{Refl}, \rn{Symm}, \rn{Trans}, \rn{Cong}, \rn{Decomp}, \rn{Bisimilar}, or \rn{Singleton}.
 In all cases, we have that $\Ec \models_{\thD} t_1 \teq t_2$.
 For \rn{Refl}, \rn{Symm}, \rn{Trans}, \rn{Cong}, this is a consequence of equality reasoning.
-For \rn{Unify$_1$}, this is a consequence of injectivity.
+For \rn{Decomp}, this is a consequence of injectivity.
 For \rn{Bisimilar}, our construction of $\Val$ indicates that the value of $t_1$ and $t_2$ are equivalent in all models of $\Ec$.
@@
 which by the induction hypothesis is bisimilar with $\nf{(\mu x. \const{C}( \Val^\ast \ec{ u_1 }, \ldots, \Val^\ast \ec{ u_n } ) )}$, call this term $t_u$.
-Since \rn{Unify$_2$} does not apply to $\Fc$,
+Since \rn{Clash} does not apply to $\Fc$,
 then $\Val^\ast \ec{t}$ must be a term $t_w$ of the form $\mu v_{\ec{t}}. \const{C}( w_1, \ldots, w_n )$,
@@
 Say that $\muvar( w_i ) = v_{\ec{y_i}}$.
-Due to our construction of $\Val^\ast$ and since \rn{Unify$_1$} does not apply to $\Fc$, we have $\ec{u_i} = \ec{y_i}$,
+Due to our construction of $\Val^\ast$ and since \rn{Decomp} does not apply to $\Fc$, we have $\ec{u_i} = \ec{y_i}$,
 and thus $t_u\interp{u_i'} \vsim \Val^\ast \ec{y_i}$.
@@
 $\ec{u}$ must contain a term of the form $C_i( s^1_i( u ), \ldots, s^{n}_i( u ) )$ for some $i$.
-By our construction and since \rn{Unify$_1$} and \rn{Unify$_2$} do not apply to $\Fc$, 
+By our construction and since \rn{Decomp} and \rn{Clash} do not apply to $\Fc$, 
 $\Val^\ast \ec{ u }$ must be of the form $t_u = \mu v_{\ec{u}}. C_i( u_1, \ldots, u_n )$,
@@
 \end{tabular}\end{center}
-\caption{Number of solved goals for the three benchmark suites}
+\caption{\,Number of solved goals for the three benchmark suites}
 \label{tab:bench}

### 2014-12-15 tex/codatadp.tex
merge?

@@
 \)
-\rn{Symm}
+\rn{Sym}
 \qquad
@@
 Figure~\ref{fig:cc-rules} gives the basic rules of our calculus.
-Together with the rules \rn{Refl}, \rn{Symm}, \rn{Trans}, the rule \rn{Cong} computes the (upward) congruence closure,
+Together with the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, the rule \rn{Cong} computes the (upward) congruence closure,
 while the rules \rn{Decomp} and \rn{Clash} together compute (downward) unification.
@@
 Otherwise, the child of $\Ec$ is a closed derivation tree 
-whose root node is $\Ec \cup t_1 \teq t_2$ obtained by applying one of the rules \rn{Refl}, \rn{Symm}, \rn{Trans}, \rn{Cong}, \rn{Decomp}, \rn{Bisimilar}, or \rn{Singleton}.
+whose root node is $\Ec \cup t_1 \teq t_2$ obtained by applying one of the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, \rn{Decomp}, \rn{Bisimilar}, or \rn{Singleton}.
 In all cases, we have that $\Ec \models_{\thD} t_1 \teq t_2$.
-For \rn{Refl}, \rn{Symm}, \rn{Trans}, \rn{Cong}, this is a consequence of equality reasoning.
+For \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, this is a consequence of equality reasoning.
 For \rn{Decomp}, this is a consequence of injectivity.
@@
 $\ec{u}$ must contain a term of the form $C_i( s^1_i( u ), \ldots, s^{n}_i( u ) )$ for some $i$.
-By our construction and since \rn{Decomp} and \rn{Clash} do not apply to $\Fc$, 
+By our construction of $\Val^\ast$ and since \rn{Decomp} and \rn{Clash} do not apply to $\Fc$, 
 $\Val^\ast \ec{ u }$ must be of the form $t_u = \mu v_{\ec{u}}. C_i( u_1, \ldots, u_n )$,
-where $\muvar( u_1 ) = v_{\ec{s^1_i(u)}}$, $\ldots$, $\muvar(u_n) = v_{\ec{s^n_i( u )}}$.
-Moreover, $\M( u )$ is of the form $t_w = \mu x. C_i( w_1, \ldots, w_n )$,
+where $\muvar( u_1 ) = v_{\ec{s^1_i(u)}}$, $\ldots$, $\muvar(u_n) = v_{\ec{s^n_i( u )}}$,
+and $\M( u )$ is of the form $t_w = \mu x. C_i( w_1, \ldots, w_n )$,
 which by the induction hypothesis is bisimilar with $\Val^\ast \ec{u}$.
 If $k = i$, then $\M( t )$ is bisimilar with $t_w\interp{w_j}$,
-which by Lemma~\ref{lem:mu-cong} is bisimilar with $t_u\interp{u_j}$,
+which since $t_w \vsim t_u$ by Lemma~\ref{lem:mu-cong} is bisimilar with $t_u\interp{u_j}$,
 which by Lemma~\ref{lem:model-completion}.2 is equal to $\Val^\ast \ec{ s^j_k( u )}$.

### 2014-12-15 tex/codatadp.tex
allow "first-order" parenthesis rules for \mu
minor

@@
 \paragraph{Step 1: Computing the Bidirectional Closure.}
-Figure~\ref{fig:cc-rules} gives the basic rules of our calculus.
-Together with the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, the rule \rn{Cong} computes the (upward) congruence closure,
-while the rules \rn{Decomp} and \rn{Clash} together compute (downward) unification.
+Figure~\ref{fig:cc-rules} gives the basic rules of the calculus.
+Together with \rn{Refl}, \rn{Sym}, and \rn{Trans}, the \rn{Cong} rule computes the (upward) congruence closure,
+whereas the \rn{Decomp} and \rn{Clash} rules together compute (downward) unification.
 In the unification rules, additional equalities are inferred based on the injectivity of constructors by \rn{Decomp},
@@
   %\ttpath{\Ec}{\ec{t}}{\ec{t}} \neq \emptyset
-  \Val \ec{t_1} = (\vthinspace\mu x.\; t_2)
+  \Val \ec{t_1} = \mu x.\; t_2
   \quad
@@
 \rn{Cyclic}
-\\[5\jot]
+\\[4\jot]
 \(
@@
 \paragraph{Step 2: Applying Cyclicity and Bisimilarity.}
-In this step,
-we assume that $\Ec$ is saturated with respect to the rules in Figure~\ref{fig:cc-rules}, that is,
-no nonredudant application of one of its rules can be applied to $\Ec$.
+In the second step,
+we assume that $\Ec$ is saturated with respect to the rules in Figure~\ref{fig:cc-rules}---that is,
+no rule can be applied to extend $\Ec$ further.
 When this is the case, it is easy to see that $\Ec$ induces an equivalence relation over $\tEc$ such that two terms $t_1$ and $t_2$ are equivalent if and only if $t_1 \teq t_2 \in \Ec$.
 Thus, we will in the following consider $\Ec$ as a set of equivalence classes of terms. 
-For term $t \in \tEc$, we will write $\ec{t}$ to denote the equivalence class in $\Ec$ that containing $t$.
+For term $t \in \tEc$, we will write $\ec{t}$ to denote the equivalence class in $\Ec$ that contains $t$.
 

### 2014-12-15 tex/codatadp.tex
Refactor lemma.

@@
 (which we will call the \emph{normal form} of $t$)
-that is a renaming of $v$ by replacing all of the bisimilar subterms of $t$ with $x$,
+that is a renaming of $t$ by replacing all of the bisimilar subterms of $t$ with $x$,
 %and by replacing all of the nonnormal subterms of $t$ with their corresponding normal forms.
@@
 Thus, when constructing a model $\M$ for $\Ec$,
-it remains only to define how wrong-applied selector terms and variables in $\Ec$ are interpreted.
+it remains only to define how $\M$ interprets wrong-applied selector terms and variables.
 For the latter, this will be based on the mapping $\Val$ constructed in step two of our calculus.
@@
 \item
-$t_1 \interp{ t_2 } = \Val^\ast \ec{x}$
-for all $t_1$ with subterm $t_2$ in the range of $\Val^\ast$ where $\muvar(t_2)=v_{\ec{x}}$,
+If $\const{C}( t_1, \ldots, t_n ) \in \ec{x}$,
+then $\Val^\ast \ec{x} = \mu v_{\ec{x}}. \const{C}( w_1, \ldots, w_n )$
+and $\Val^\ast \ec{x} \interp{ w_i } = \Val^\ast \ec{t_i}$ for $i = 1, \ldots, n$.
+%and $\muvar( w_i ) = v_{\ec{t_i}}$ for $i = 1, \ldots, n$,
+%\item
+%$t_1 \interp{ t_2 } = \Val^\ast \ec{x}$
+%for all $t_1$ with subterm $t_2$ in the range of $\Val^\ast$ where $\muvar(t_2)=v_{\ec{x}}$,
 \item 

### 2014-12-15 tex/codatadp.tex
Refactor proof for stronger lemma.

@@
 which by the induction hypothesis is bisimilar with $\nf{(\mu x. \const{C}( \Val^\ast \ec{ u_1 }, \ldots, \Val^\ast \ec{ u_n } ) )}$, call this term $t_u$.
-Since \rn{Clash} does not apply to $\Fc$,
-then $\Val^\ast \ec{t}$ must be a term $t_w$ of the form $\mu v_{\ec{t}}. \const{C}( w_1, \ldots, w_n )$,
-that is, an application of the same constructor.
+By Lemma~\ref{lem:model-completion}.2, $\Val^\ast \ec{t}$ is a term
+%Since \rn{Clash} does not apply to $\Fc$, then $\Val^\ast \ec{t}$ must be a term
+$t_w$ of the form $\mu v_{\ec{t}}. \const{C}( w_1, \ldots, w_n )$
+where $t_w\interp{w_i} = \Val^\ast \ec{u_i}$ for $i = 1, \ldots, n$.
+%that is, an application of the same constructor.
 For each $i = 1, \ldots, n$, let $u_i'$ be the $i^{th}$ argument of $t_u$.
 By Lemma~\ref{lem:mu-norm-arg}, $t_u\interp{u_i'} \vsim \Val^\ast \ec{u_i}$.
-Say that $\muvar( w_i ) = v_{\ec{y_i}}$.
-Due to our construction of $\Val^\ast$ and since \rn{Decomp} does not apply to $\Fc$, we have $\ec{u_i} = \ec{y_i}$,
-and thus $t_u\interp{u_i'} \vsim \Val^\ast \ec{y_i}$.
-By Lemma~\ref{lem:model-completion}.2, $\Val^\ast \ec{ y_i } = t_w\interp{ w_i }$,
-and thus $t_u\interp{u_i'} \vsim t_w\interp{ w_i }$ for each $i = 1, \ldots, n$.
+%Say that $\muvar( w_i ) = v_{\ec{y_i}}$.
+%Due to our construction of $\Val^\ast$ and since \rn{Decomp} does not apply to $\Fc$, we have $\ec{u_i} = \ec{y_i}$,
+%and thus $t_u\interp{u_i'} \vsim \Val^\ast \ec{y_i}$.
+%By Lemma~\ref{lem:model-completion}.2, $\Val^\ast \ec{ y_i } = t_w\interp{ w_i }$,
+and thus $t_u\interp{u_i'} \vsim t_w\interp{ w_i }$. % for each $i = 1, \ldots, n$.
 By Lemma~\ref{lem:mu-cong}, we have $t_u \vsim t_w$.
@@
 $\ec{u}$ must contain a term of the form $C_i( s^1_i( u ), \ldots, s^{n}_i( u ) )$ for some $i$.
-By our construction of $\Val^\ast$ and since \rn{Decomp} and \rn{Clash} do not apply to $\Fc$, 
-$\Val^\ast \ec{ u }$ must be of the form $t_u = \mu v_{\ec{u}}. C_i( u_1, \ldots, u_n )$,
-where $\muvar( u_1 ) = v_{\ec{s^1_i(u)}}$, $\ldots$, $\muvar(u_n) = v_{\ec{s^n_i( u )}}$,
-and $\M( u )$ is of the form $t_w = \mu x. C_i( w_1, \ldots, w_n )$,
-which by the induction hypothesis is bisimilar with $\Val^\ast \ec{u}$.
-If $k = i$, then $\M( t )$ is bisimilar with $t_w\interp{w_j}$,
-which since $t_w \vsim t_u$ by Lemma~\ref{lem:mu-cong} is bisimilar with $t_u\interp{u_j}$,
-which by Lemma~\ref{lem:model-completion}.2 is equal to $\Val^\ast \ec{ s^j_k( u )}$.
-Thus, $\M( t ) \vsim \Val^\ast \ec{ s^j_k( u )} = \Val^\ast \ec{t}$.
+By Lemma~\ref{lem:model-completion}.2, $\Val^\ast \ec{u}$ is a term $t_w$ of the form
+$\mu v_{\ec{u}}. C_i( w_1, \ldots, w_n )$,
+where $t_w \interp{ w_j } = \Val^\ast \ec{ s^j_k( u )}$. 
+%By our construction of $\Val^\ast$ and since \rn{Decomp} and \rn{Clash} do not apply to $\Fc$, 
+%$\Val^\ast \ec{ u }$ must be of the form $t_u = \mu v_{\ec{u}}. C_i( u_1, \ldots, u_n )$,
+%where $\muvar( u_1 ) = v_{\ec{s^1_i(u)}}$, $\ldots$, $\muvar(u_n) = v_{\ec{s^n_i( u )}}$,
+%and $\M( u )$ is of the form $t_w = \mu x. C_i( w_1, \ldots, w_n )$,
+%which by the induction hypothesis is bisimilar with $\Val^\ast \ec{u}$.
+If $k = i$, then $\M( t )$ is bisimilar with $t_w\interp{w_j} = \Val^\ast \ec{ s^j_k( u )}$,
+which is $\Val^\ast \ec{t}$.
+%which since $t_w \vsim t_u$ by Lemma~\ref{lem:mu-cong} is bisimilar with $t_u\interp{u_j}$,
+%which by Lemma~\ref{lem:model-completion}.2 is equal to $\Val^\ast \ec{ s^j_k( u )}$.
+%Thus, $\M( t ) \vsim \Val^\ast \ec{ s^j_k( u )} 
 If $k \neq i$, then we interpret $s^j_k( \M( u ) )$ as the value in $\M( \tau )$ that is bisimilar with $\Val^\ast \ec{ t }$.

### 2014-12-15 tex/codatadp.tex
dropped some needless 'will's

--- 2014-12-15 tex/codatadp.tex
@@
 \emph{corecursive singletons}. For the decision procedure, it will be
-important to detect these types, even if they occur rarely in practice. A type
+important to detect these types, even if they rarely arise in practice. A type
 may also be a corecursive singleton only in some models. For example, if
@@
 Our calculus consists of derivation rules that operate on a set $\Ec$ of equalities and disequalities between $\Sig$-terms.
-We will write $\tEc$ to denote the set of $\Sig$-terms occurring in $\Ec$.
+We write $\tEc$ to denote the set of $\Sig$-terms occurring in $\Ec$.
 %We will commonly denote tuples of terms $( t_1, \ldots, t_n )$ in bold font, as $\vec t$.
@@
 where a rule can be applied to $\Ec$ if it meets all of the specified preconditions for $\Ec$.
-The conclusion of a rule either describes equalities to be added to $\Ec$ (in which case we will call an application of it to be \emph{nonterminal}),
-or is $\bot$ (in which case we will call an application of it to be \emph{terminal}).
+The conclusion of a rule either describes equalities to be added to $\Ec$ (in which case we call an application of it \emph{nonterminal}),
+or is $\bot$ (in which case we call an application of it \emph{terminal}).
 A rule may have multiple conclusions separated by $\parallel$, which denotes nondeterministic branching.
@@
 When this is the case, it is easy to see that $\Ec$ induces an equivalence relation over $\tEc$ such that two terms $t_1$ and $t_2$ are equivalent if and only if $t_1 \teq t_2 \in \Ec$.
-Thus, we will in the following consider $\Ec$ as a set of equivalence classes of terms. 
-For term $t \in \tEc$, we will write $\ec{t}$ to denote the equivalence class in $\Ec$ that contains $t$.
+Thus, we can consider $\Ec$ as a set of equivalence classes of terms. 
+For a term $t \in \tEc$, we write $\ec{t}$ to denote the equivalence class in $\Ec$ that contains $t$.
 
@@
 The rules of this step contain premises that involve a mapping $\Val$ from equivalence classes to terms containing $\mu$-bindings.
-Recall that $\mu$-bindings can be used for representing (possibly cyclic) terms and values,
-for instance, $\mu x. \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
-Formally, we recursively define a $\mu$-term as a term-like structure either of the form
-$\mu x. \const{C}( \vec t\, )$ for some constructor $\const{C} \in \Ctr$ and $\mu$-terms $\vec t$, 
-or a variable $x$.
-For consistency, we will also consider $\mu x. \const{c}$ to be a $\mu$-term
-in the case where $\const{c}$ is a constant of type $\tau \in \Nondata$.
-We consider only $\mu$-terms that are \emph{well-typed}, which we define recursively:
-if the type of $\const{C}$ is $\vec \tau \rightarrow \upsilon$
-and $\vec v$ are well-typed $\mu$-terms of type $\vec \tau$ under the assumption that $x$ has type $\upsilon$,
-then $\mu x. \const{C}( \vec v )$ is a well-typed term of type $\upsilon$.
-We say a $\mu$-term is \emph{closed} if and only if it contains no free variables.
-A $\mu$-term is \emph{cyclic} if and only if it contains a bound variable.
+Recall that $\mu$-bindings can be used for representing possibly cyclic terms and values.
+%(Section~\ref{sec:introduction}).
+% for example, $\mu x. \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
+Formally, we recursively define a $\mu$-term as a term-like structure as either a variable $x$
+or a node of the form
+$\mu x.\; \const{C}( \vec t\, )$ for some constructor $\const{C} \in \Ctr$ and $\mu$-terms $\vec t$.
+For uniformity, we also consider $\mu x.\; \const{c}$ to be a $\mu$-term
+if $\const{c}$ is a constant of type $\tau \in \Nondata$.
+We consider only $\mu$-terms that are well-typed---i.e., whose function symbols
+are applied to arguments of the expected type.
+%%% I think this should be clear enough from the "Conventions" paragraph at the
+%%% end of the introduction.
+%%
+% which we define recursively:
+%if $\const{C} : \vec \tau \rightarrow \upsilon$
+%and $\vec v$ are well-typed $\mu$-terms of type $\vec \tau$ under the assumption that $x$ has type $\upsilon$,
+%then $\mu x. \const{C}( \vec v )$ is a well-typed term of type $\upsilon$.
+We say a $\mu$-term is \emph{closed} if %and only if --- by convention, "and only if" is not necessary for *definitions* (but add it back if you disagree)
+it contains no free variables.
+A $\mu$-term is \emph{cyclic} if %and only if --- ditto
+it contains an occurrence of a bound variable.
 We write $t \vsim s$ if $\mu$-terms $t$ and $s$ are syntactically equivalent for some capture-avoiding renaming of $\mu$-bound variables,
-in which case we will commonly say $t$ and $s$ are \emph{bisimilar}.
+in which case $t$ and $s$ are $\alpha$-\emph{equivalent}.
 For instance, 
@@
 %, or informally are \emph{observationally equivalent}.
-%For convienience, we will use $\mu$-terms to refer to (classes of) values for both co-datatype and datatype terms,
+%For convienience, we use $\mu$-terms to refer to (classes of) values for both co-datatype and datatype terms,
 %where in the latter case, a $\mu$-term with a bound variable denotes an infeasible term.
@@
 \end{array}\]
-We will often write $t_1\interp{t_2}$ as shorthand for $t_1\interp{t_2}^\emptyset$.
+We often write $t_1\interp{t_2}$ as shorthand for $t_1\interp{t_2}^\emptyset$.
 
@@
 For any $\mu$-term $t$ of the form $\mu x. C( \vec u )$, we may recursively construct a normal $\mu$-term $\nf{t}$
-(which we will call the \emph{normal form} of $t$)
+(which we call the \emph{normal form} of $t$)
 that is a renaming of $v$ by replacing all of the bisimilar subterms of $t$ with $x$,
@@
 For the latter, this will be based on the mapping $\Val$ constructed in step two of our calculus.
-We compute the \emph{completion} of $\Val$ for normal model $\M$, which we will write as $\Val^\ast$,
+We compute the \emph{completion} of $\Val$ for normal model $\M$, which we write as $\Val^\ast$,
 which we construct by assigning values from $\M$ to unassigned variables in the domain of $\Val$.

### 2014-12-15 tex/codatadp.tex
merge

@@
 \newcommand{\vsim}{\sim}
+\newcommand{\aequiv}{\mathrel{=_\alpha}}
 \newcommand{\vsimv}[1]{\vsim_{#1}}
@@
 The metavariables $\delta,\:\varepsilon$ range over (co)datatypes,
-whereas $\tau,\:\upsilon$ range over arbitrary types.
+whereas $\tau,\:\upsilon$ range over arbitrary types. The symbol $=$ is reserved
+for syntactic equality.
 
@@
 it contains an occurrence of a bound variable.
-We write $t \vsim s$ if $\mu$-terms $t$ and $s$ are syntactically equivalent for some capture-avoiding renaming of $\mu$-bound variables,
+We write $t \aequiv s$ if $\mu$-terms $t$ and $s$ are syntactically equivalent for some capture-avoiding renaming of $\mu$-bound variables,
 in which case $t$ and $s$ are $\alpha$-\emph{equivalent}.
@@
 that is a renaming of $v$ by replacing all of the bisimilar subterms of $t$ with $x$,
+(which we call the \emph{normal form} of $t$)
+that is a renaming of $t$ by replacing all of the bisimilar subterms of $t$ with $x$,
 %and by replacing all of the nonnormal subterms of $t$ with their corresponding normal forms.

### 2014-12-16 tex/codatadp.tex
more

@@
 to enumerate them.} This is necessary both for finite model finding (modulo theories)
-and for theory combinations. The $\mu$-binder notation is used to represent cyclic
-values internally. For example,
-the term $\const{SCons}(1,\: \mu s.\; \const{SCons}(0,\: \const{SCons}(9,\: s)))$
+and for theory combinations. The $\mu$-binder notation associates a name with
+a (sub)term; it is used to represent cyclic values in the generated models and
+in the metatheory. For example,
+the $\mu$-term $\const{SCons}(1,\: \mu s.\; \const{SCons}(0,\: \const{SCons}(9,\: s)))$
 denotes the lasso-shaped sequence $1, 0, 9, 0, 9, 0, 9, \ldots$\afterLdots.
@@
 we assume that $\Ec$ is saturated with respect to the rules in Figure~\ref{fig:cc-rules}---that is,
-no rule can be applied to extend $\Ec$ further.
+no rule can be applied to extend~$\Ec$ further.
 When this is the case, it is easy to see that $\Ec$ induces an equivalence relation over $\tEc$ such that two terms $t_1$ and $t_2$ are equivalent if and only if $t_1 \teq t_2 \in \Ec$.
@@
 % for example, $\mu x. \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
-Formally, we recursively define a $\mu$-term as a term-like structure as either a variable $x$
+Formally, $\mu$-terms are defined recursively as being either a variable $x$
 or a node of the form
-$\mu x.\; \const{C}( \vec t\, )$ for some constructor $\const{C} \in \Ctr$ and $\mu$-terms $\vec t$.
-For uniformity, we also consider $\mu x.\; \const{c}$ to be a $\mu$-term
-if $\const{c}$ is a constant of type $\tau \in \Nondata$.
-We consider only $\mu$-terms that are well-typed---i.e., whose function symbols
-are applied to arguments of the expected type.
+$\mu x.\; \const{C}( \vec t\, )$ for some constructor $\const{C} \in \Ctr$ and
+$\mu$-terms $\vec t$ with the expected types.
+The variable $x$ need not occur under the binder's body, in which case the $\mu$-prefix
+is optional.
+For uniformity, we also consider $%\mu x.\;
+\const{c}$ to be a $\mu$-term if $\const{c}$ is a constant of type $\tau \in \Nondata$.
 %%% I think this should be clear enough from the "Conventions" paragraph at the
@@
 %then $\mu x. \const{C}( \vec v )$ is a well-typed term of type $\upsilon$.
-We say a $\mu$-term is \emph{closed} if %and only if --- by convention, "and only if" is not necessary for *definitions* (but add it back if you disagree)
-it contains no free variables.
-A $\mu$-term is \emph{cyclic} if %and only if --- ditto
+
+A $\mu$-term is \emph{closed} if %and only if --- by convention, "and only if" is not necessary for *definitions* (but add it back if you disagree)
+it contains no free variables. It is \emph{cyclic} if %and only if --- ditto
 it contains an occurrence of a bound variable.
-We write $t \aequiv s$ if $\mu$-terms $t$ and $s$ are syntactically equivalent for some capture-avoiding renaming of $\mu$-bound variables,
-in which case $t$ and $s$ are $\alpha$-\emph{equivalent}.
-For instance, 
-$\mu x. \const{C}( x ) \vsim \mu y. \const{C}( y )$ and
-$\mu x, \const{D}( y, x ) \vsim \mu z. \const{D}( y, z )$, but
-$\mu x. \const{D}( y, x ) \not\vsim \mu x. \const{C}( x )$,
-$\mu x. \const{D}( y, x ) \not\vsim \mu x. \const{D}( z, x )$ and
-$\mu x. \const{D}( y, x ) \not\vsim \mu y. \const{D}( y, x )$.
-Also, note that
-two $\mu$-terms may represent the same value while being not bisimilar, as in
-$\mu x. \const{C}( \mu y. \const{C}( x ) ) \not\vsim \mu x. \const{C}( x )$.
+The notation $t \aequiv u$, expressing $\alpha$-\emph{equivalence},
+indicates that the $\mu$-terms $t$ and $u$
+are syntactically equivalent for some capture-avoiding renaming of $\mu$-bound variables.
+For example, 
+%$\mu x. \const{C}( x ) \aequiv \mu y. \const{C}( y )$ and
+$\mu x.\; \const{D}( y, x ) \aequiv \mu z.\; \const{D}( y, z )$,
+but
+$\mu x.\; \const{C}( x ) \not\aequiv \mu x.\; \const{D}( y, x ) \not\aequiv \mu x.\; \const{D}( z, x )
+\not\aequiv \mu y.\; \const{D}( y, x )$.
+Two $\mu$-terms may represent the same value while not being $\alpha$-equivalent, as in
+$\mu x. \const{C}( \const{C}( x ) ) \not\aequiv \mu x. \const{C}( x )$.
 %, or informally are \emph{observationally equivalent}.

### 2014-12-16 tex/codatadp.tex
more

@@
 Two $\mu$-terms may represent the same value while not being $\alpha$-equivalent, as in
-$\mu x. \const{C}( \const{C}( x ) ) \not\aequiv \mu x. \const{C}( x )$.
+$\mu x.\; \const{C}( \const{C}( x ) ) \not\aequiv \mu x.\; \const{C}( x )$.
 %, or informally are \emph{observationally equivalent}.
@@
 %either in the case when $\tau$ is a co-datatype or datatype type,
-where a cyclic $t$ describes an infeasible class of values when $\tau \not\in \Codata$.
+where a cyclic $t$ describes an infeasible class of values when
+%$\tau \not\in \Codata$.
+$\tau \in \Data$.
 At the beginning of this step, 

### 2014-12-16 tex/codatadp.tex
more

@@
 is optional.
-For uniformity, we also consider $%\mu x.\;
-\const{c}$ to be a $\mu$-term if $\const{c}$ is a constant of type $\tau \in \Nondata$.
+
+%%% But there are no constants of type $\tau \in \Nondata$ -- at most variables!
+%
+%For uniformity, we also consider $%\mu x.\;
+%\const{c}$ to be a $\mu$-term if $\const{c}$ is a constant of type $\tau \in \Nondata$.
+
 %%% I think this should be clear enough from the "Conventions" paragraph at the
@@
 Then, $\Val$ is modified by applying the following update exhaustively:
-
-\(
+\[
 \inferrule{
@@
 }
-\)
-
+\]
 Here, we write $FV( \Val )$ to denote the \emph{free variables} occurring in $\Val$, which
@@
 for some value of $v_{\ec{w}}$ and $v_{\ec{z}}$.
-$\Box$
+\qed
 \end{example}
@@
 \paragraph{Step 3: Branching.}
-If neither a rule from Figure~\ref{fig:cc-rules} or Figure~\ref{fig:ab-rules} applies to $\Ec$, 
+If neither a rule from Figure \ref{fig:cc-rules}~or~\ref{fig:ab-rules} applies to $\Ec$, 
 the rules from Figure~\ref{fig:split-rule} are applied.

### 2014-12-16 tex/codatadp.tex
clarified
a nicer schema (IMO)

@@
 The metavariables $\delta,\:\varepsilon$ range over (co)datatypes,
-whereas $\tau,\:\upsilon$ range over arbitrary types. The symbol $=$ is reserved
-for syntactic equality.
+whereas $\tau,\:\upsilon$ range over arbitrary types. When applied to terms,
+the symbol $=$ denotes syntactic equality.
 
@@
 \mathrel{\uplus} \Sel$, where $\Ctr$ are the \emph{constructors} and $\Sel$ are the
-\emph{selectors}; there is no need to consider further function symbols
+\emph{selectors}. There is no need to consider further function symbols
 because they can be abstracted away as variables when combining theories.
+Nonetheless, it is convenient to make an exception for natural number
+constants ($0$, $1$, \ldots) in examples.
 
@@
 
+\newcommand\elll{\kern.2ex l\kern.14ex}
+
 Each datatype $\delta$ is equipped with
@@
 \begin{aligned}[t]
-(\keyw{co})\keyw{datatype}\;\,
-  \delta_1 & {}=\, \smash{\const{C}_{11}(\bigl[\const{s}_{11}^1{:}\bigr]\; \tau_{11}^1, \,\ldots,\, \bigl[\const{s}_{11}^{n_{11}}{:}\bigr]\; \tau_{11}^{n_{11}})} & \\
-           & {}\;\phantom{=\,}\;{} \smash{\,\,\,\vdots} \\[-\jot] % {}\;\phantom{=}\llap{\ensuremath{\mid}}\; \cdots \\
-           & {}\;\phantom{=\,}\llap{\ensuremath{\,\mid\,}}\; \smash{\const{C}_{1m}(\bigl[\const{s}_{1m}^1{:}\bigr]\; \tau_{1m}^1, \,\ldots,\, \big[\const{s}_{1m}^{n_{1m}}{:}\big]\; \tau_{1m}^{n_{1m}})} \\[-1\jot]
+\!(\keyw{co})\keyw{datatype}\;\,
+  \delta_1 & {}= \smash{\const{C}_{11\!}(\bigl[\const{s}_{11\!}^1{:}\bigr]\vthinspace \tau_{11\!}^1, \ldots, \bigl[\const{s}_{11\!}^{n_{11\!}}{:}\bigr]\vthinspace \tau_{11\!}^{n_{11\!}})} \mid \cdots \mid \smash{\const{C}_{1m}(\bigl[\const{s}_{1m}^1{:}\bigr]\vthinspace \tau_{1m}^1, \ldots, \bigl[\const{s}_{1m}^{n_{1m}}{:}\bigr]\vthinspace \tau_{1m}^{n_{1m}})} \\
    \smash{\vdots\,\,\,} \\[-1\jot]
-  \keyw{and}\; \,\delta_l & =\, \ldots
+  \keyw{and}\; \,\delta_l & = \smash{\const{C}_{\elll 1\!}(\bigl[\const{s}_{\elll 1\!}^1{:}\bigr]\vthinspace \tau_{\elll 1\!}^1, \ldots, \bigl[\const{s}_{\elll 1\!}^{n_{\elll 1\!}}{:}\bigr]\vthinspace \tau_{\elll 1\!}^{n_{\elll 1\!}})} \mid \cdots \mid \smash{\const{C}_{\elll m}(\bigl[\const{s}_{\elll m}^1{:}\bigr]\vthinspace \tau_{\elll m}^1, \ldots, \bigl[\const{s}_{\elll m}^{n_{\elll m}}{:}\bigr]\vthinspace \tau_{\elll m}^{n_{\elll m}})}
 \end{aligned}

### 2014-12-16 tex/codatadp.tex
trying something else
nicer schema
shorten a bit for conference

--- 2014-12-16 tex/codatadp.tex
@@
 \begin{paper}
-Detailed proofs of soundness and completeness are included in the technical report
+Proofs of soundness and completeness are included in the technical report
 associated with this paper \cite{our-report}.
@@
 (Section~\ref{sec:the-theory-solver}).
-\textbf{TODO: CHECK WITH IMPLEMENTATION, but round to nearest 50 or 1000.}
+%\textbf{TODO: CHECK WITH IMPLEMENTATION, but round to nearest 50 or 1000.}
 It consists of about 2000 lines of \cpp{} code, among which 1600 are
@@
 The metavariables $\delta,\:\varepsilon$ range over (co)datatypes,
-whereas $\tau,\:\upsilon$ range over arbitrary types. The symbol $=$ is reserved
-for syntactic equality.
-
+whereas $\tau,\:\upsilon$ range over arbitrary types.
+\begin{report}When applied to terms, the symbol $=$ denotes syntactic equality.\end{report}
+
+\begin{report}
 Function symbols are written in a sans-serif font (e.g., $\const{f}$, $\const{g}$) to
-distinguish them from variables (e.g., $x$, $y$). Symbols starting with an uppercase
-letter (e.g.,
+distinguish them from variables (e.g., $x$, $y$).
+\end{report}
+Symbols starting with an uppercase letter (e.g.,
 $\const{S}$, $\const{SCons}$) are reserved for constructors. With each function symbol \const{f}
@@
 and a return type $\tau$. This connection can be expressed %compactly
-as $\const{f} : \tau_1 \times \cdots \times \tau_n \to \tau$,
-which collapses to $\const{f} : \tau$ if $n = 0$.
+as $\const{f} : \tau_1 \times \cdots \times \tau_n \to \tau$\begin{report},
+which collapses to $\const{f} : \tau$ if $n = 0$\end{report}.
 For a term $t$, the notation $t : \tau$ indicates that it has type $\tau$.
@@
 \emph{datatypes}, $\Codata$ are the \emph{codatatypes}, and $\Nondata$ are the %remaining
-\emph{ordinary types}. The functions are partitioned into $\Funcs = \Ctr
+\emph{ordinary types}. The function symbols are partitioned into $\Funcs = \Ctr
 \mathrel{\uplus} \Sel$, where $\Ctr$ are the \emph{constructors} and $\Sel$ are the
-\emph{selectors}; there is no need to consider further function symbols
+\emph{selectors}. There is no need to consider further function symbols
 because they can be abstracted away as variables when combining theories.
+Exceptionally, it is convenient to use natural number
+constants ($0$, $1$, \ldots)\ in examples.
 
@@
 
+\newcommand\elll{\kern.18ex l\kern.11ex}
+\newcommand\elllx{\kern.11ex l\kern.18ex}
+
 Each datatype $\delta$ is equipped with
@@
 \begin{aligned}[t]
-(\keyw{co})\keyw{datatype}\;\,
-  \delta_1 & {}=\, \smash{\const{C}_{11}(\bigl[\const{s}_{11}^1{:}\bigr]\; \tau_{11}^1, \,\ldots,\, \bigl[\const{s}_{11}^{n_{11}}{:}\bigr]\; \tau_{11}^{n_{11}})} & \\
-           & {}\;\phantom{=\,}\;{} \smash{\,\,\,\vdots} \\[-\jot] % {}\;\phantom{=}\llap{\ensuremath{\mid}}\; \cdots \\
-           & {}\;\phantom{=\,}\llap{\ensuremath{\,\mid\,}}\; \smash{\const{C}_{1m}(\bigl[\const{s}_{1m}^1{:}\bigr]\; \tau_{1m}^1, \,\ldots,\, \big[\const{s}_{1m}^{n_{1m}}{:}\big]\; \tau_{1m}^{n_{1m}})} \\[-1\jot]
+\!(\keyw{co})\keyw{datatype}\;\,
+  \delta_1 & {}= \smash{\const{C}_{11\!}(\bigl[\const{s}_{11\!}^1{:}\bigr]\vthinspace \tau_{11\!}^1, \ldots, \bigl[\const{s}_{11\!}^{n_{11\!}}{:}\bigr]\vthinspace \tau_{11\!}^{n_{11\!}})} \mid \cdots \mid \smash{\const{C}_{1m_1\!}(\ldots)} \\
    \smash{\vdots\,\,\,} \\[-1\jot]
-  \keyw{and}\; \,\delta_l & =\, \ldots
+  \keyw{and}\; \,\delta_{\elllx} & =\, \smash{\const{C}_{\elll 1\!}(\bigl[\const{s}_{\elll 1\!}^1{:}\bigr]\vthinspace \tau_{\elll 1\!}^1, \ldots, \bigl[\const{s}_{\elll 1\!}^{n_{\elll 1\!}}{:}\bigr]\vthinspace \tau_{\elll 1\!}^{n_{\elll 1\!}})} \mid \cdots \mid \smash{\const{C}_{\elll m_{\elllx}\!}(\ldots)}
 \end{aligned}
@@
 admissible only if a ground constructor term can be exhibited that belongs to
-the datatype. This rules out non-well-founded specifications such as
+the datatype. This rules out non-well-founded specifications\begin{paper}.\end{paper}
+\begin{report}%
+such as
 \[\keyw{datatype}\;\, \ty{fstream}_{\,\tau} \,=\, \const{FSCons}(\tau,\:\, \ty{fstream}_{\,\tau})\]
+\end{report}
 For codatatypes, no admissibility check is necessary because there is always a term,
@@
 
-\paragraph{Properties.}
+\paragraph{Characterization.}
+Datatypes can be defined semantically as the initial model of the
+selector--constructor equations. This is the view followed by Barrett et al.\
+\cite{barrett-et-al-2010}.
+\begin{report}
+A drawback of this approach is that it does
+not naturally account for selectors applied to wrong constructors. Barrett et
+al.\ address this by parameterizing the construction by default values, but
+this gives rise to spurious equalities between unrelated terms---e.g.,
+$\const{s}_{11}(\const{C}_2) \teq \const{s}_{11}(\const{C}_3)$. This flaw
+could be corrected, but the added complexity seems to suggest that selectors
+are better characterized axiomatically.
+
+\end{report}
+A related semantic view of datatypes is as initial algebras \cite{xxx}.
+Codatatypes are then defined dually as final coalgebras \cite{xxx}. The
+datatypes are generated by their constructors, whereas the codatatypes are
+viewed through their selectors. By uniformly focusing on the constructors, the
+axiomatic approach emphasizes the commonality between datatypes and
+codatatypes, while sacrificing a theoretically fruitful notion of duality.
+
 Datatypes and codatatypes share many basic properties pertaining to
@@
   & %\forall \bar x, \bar y.\;\,
-    \const{C}_{i\negvthinspace j}(\bar x) \tneq \const{C}_{i\negvthinspace \jPrime}(\bar y) \quad\text{if $j \not= \jPrime$}
-  \\
+    \smash{\const{C}_{i\negvthinspace j}(\bar x) \tneq \const{C}_{i\negvthinspace \jPrime}(\bar y) \quad\text{if $j \not= \jPrime$}}
+  \\[-.5\jot]
 \text{Injectivity:}\quad
   & %\forall x_1,\ldots,x_{n_{i\negvthinspace j}} y.\;\,
-    \const{C}_{i\negvthinspace j}(x_1,\ldots,x_{n_{i\negvthinspace j}}) \teq \const{C}_{i\negvthinspace j}(x_1,\ldots,x_{k-1},y,x_{k+1},\ldots,x_{n_{i\negvthinspace j}}) \longrightarrow x_k \teq y
-  \\
+    \smash{\const{C}_{i\negvthinspace j}(x_1,\ldots,x_{n_{i\negvthinspace j}}) \teq \const{C}_{i\negvthinspace j}(x_1,\ldots,x_{k-1},y,x_{k+1},\ldots,x_{n_{i\negvthinspace j}}) \longrightarrow x_k \teq y}
+  \\[-.5\jot]
 \text{Exhaustiveness:}\quad
-  & \is{C}_{i1}(x) \mathrel\lor \cdots \mathrel\lor \is{C}_{im_i}(x)
-  \\
+  & \smash{\is{C}_{i1}(x) \mathrel\lor \cdots \mathrel\lor \is{C}_{im_i}(x)}
+  \\[-.5\jot]
 \text{Selection:}\quad
-  & s_{i\negvthinspace j}^{\,k}(\const{C}_{i\negvthinspace j}(x_1,\ldots,x_{n_{i\negvthinspace j}})) = x_k
+  & \smash{\const{s}_{i\negvthinspace j}^{\,k}(\const{C}_{i\negvthinspace j}(x_1,\ldots,x_{n_{i\negvthinspace j}})) = x_k}
 \end{aligned}
@@
 %
+\begin{report}
 Expressed in the algebraic jargon, exhaustiveness helps ensure that ``no
 junk'' exists, whereas distinctness and injectivity guarantee that ``no
-confusion'' can arise. The result of selectors applied to the wrong
+confusion'' can arise.
+The result of selectors applied to the wrong
 constructor is left completely unspecified.
+\end{report}
 
@@
 
-Datatypes are additionally characterized by an induction axiom schema:
+Datatypes are additionally characterized by an induction axiom schema\begin{report}:
 %
@@
 denotes either $P_{\iPrime}(x)$ if there exists some $\iPrime$ such that
-the formula is type-correct or else $\top$.
+the formula is type-correct or else $\top$\end{report}.
 %
@@
 $\const{S}(\const{S}(\ldots))$.
-
-For codatatypes, the dual notion is called coinduction. It depends on
+%
+\begin{report}\par\end{report}
+%
+For codatatypes, the dual notion is called coinduction. \begin{report}It depends on
 witnesses $R_i$ that are required to be bisimulations:
@@
 $\iPrime$ such that the formula is type-correct or $x \teq y$ otherwise.
-This axiom schema captures the idea that two values that yield the same
+\end{report}%
+This axiom schema guarantees that two values that yield the same
 observations must be equal, where the observations are made by using the
 selectors and discriminators.
-
+%
+\begin{report}\par\end{report}
+  %
 In addition, codatatypes are guaranteed to contain all values corresponding to
 infinite ground constructor terms. In general, this cannot be captured by a
-first-order axiomatization, since there may be uncountably many of them. For
-example, $\ty{stream}_{\,\ty{int}}$ is isomorphic to the uncountable function
-space $\ty{nat} \to \ty{int}$.
+first-order axiomatization, since there may be uncountably many of them.
+\begin{report}
+For example, $\ty{stream}_{\,\ty{int}}$ is isomorphic to the uncountable
+function space $\ty{nat} \to \ty{int}$.
+\end{report}
 
@@
 % specifies the
-
-\paragraph{Semantics.}
-Datatypes can be defined semantically as the initial model of the
-selector--constructor equations. This is the view followed by Barrett et al.\
-\cite{barrett-et-al-2010}. A drawback of this approach is that it does
-not naturally account for selectors applied to wrong constructors. Barrett et
-al.\ address this by parameterizing the construction by default values, but
-this gives rise to spurious equalities between unrelated terms---e.g.,
-$\const{s}_{11}(\const{C}_2) \teq \const{s}_{11}(\const{C}_3)$. This flaw
-could be corrected, but the added complexity seems to suggest that selectors
-are better characterized axiomatically.
-
-A related semantic view of datatypes is as initial algebras \cite{xxx}.
-Codatatypes are then defined dually as final coalgebras \cite{xxx}. The
-datatypes are generated by their constructors, whereas the codatatypes are
-viewed through their selectors. By uniformly focusing on the constructors, the
-axiomatic approach emphasizes the commonality between datatypes and
-codatatypes, while sacrificing a theoretically fruitful notion of duality.
 
@@
 semantic notion is more precise and is the one that interests us.
-
-It is worth observing that nothing forbids non(co)recursive specifications
+%
+In addition, nothing forbids non(co)recursive specifications
 such~as
@@
 \]
+\begin{report}
 At the semantic level, it makes no difference whether such types are
 introduced as datatypes or as codatatypes.
+\end{report}
 %Without loss of generality, we consider that these types are datatypes
@@
 
-Finally, some codatatypes are so degenerate as to have infinite values
+Some codatatypes are so degenerate as to have infinite values
 even though they are finite. The simplest example is
-\[\keyw{codatatype}\;\, \ty{a} \,=\, \const{A}(\ty{a})\]
-This is a finite codatatype, of cardinality one, whose single value is
-infinite: \,$\mu a.\; A(a)$. Other specimens are
+\vthinspace$\keyw{codatatype}~\ty{a} = \const{A}(\ty{a})$, whose
+only value is $\mu a.\; A(a)$. Other specimens are
 \[\begin{aligned}[t]
-      \keyw{codatatype}\;\, \ty{b} & \,=\, \const{B}(\ty{b},\: \ty{c},\: \ty{b},\: \ty{unit}) \\[-\jot]
+      \keyw{codatatype}\;\, \ty{b} & \,=\, \const{B}(\ty{b},\: \ty{c},\: \ty{b},\: \ty{unit}) \\[-1\jot]
       \keyw{and}\;\, \ty{c} & \,=\, \const{C}(\ty{a},\: \ty{unit},\: \ty{b},\: \ty{c})
@@
 \emph{corecursive singletons}. For the decision procedure, it will be
-important to detect these types, even if they rarely arise in practice. A type
-may also be a corecursive singleton only in some models. For example, if
-\ty{unit} is an uninterpreted type, \ty{b} and \ty{c} will be singleton types
-precisely for the models that interpret \ty{unit} as a singleton. Fortunately,
-given cardinalities for the ordinary types, it is easy to characterize
-this degenerate case:
+important to detect these types. %, even if they rarely arise in practice.
+A type may also be a corecursive singleton only in some models. If the example
+above is altered to leave \ty{unit} uninterpreted, \ty{b} and \ty{c} will be
+singletons precisely when \ty{unit} is interpreted as a singleton.
+Fortunately, given cardinalities for the ordinary types, it is easy to
+characterize this degenerate case:
 
@@
 
+\begin{report}
 \begin{proof}
@@
 \end{proof}
+\end{report}
 

### 2014-12-17 tex/codatadp.tex
more

@@
 
-To represent infinite values, % such as $\const{S}(\const{S}(\const{S}(\ldots))$,
+To represent infinite objects, % such as $\const{S}(\const{S}(\const{S}(\ldots))$,
 a natural choice is to turn to coinductive datatypes, or \emph{codatatypes},
@@
 %
-Although they are sometimes perceived as esoteric, codatatypes have an important
+Despite their reputation for being esoteric, codatatypes have a
 role to play in computer science. The verified C compiler CompCert
@@
 infinity value $\infty = \const{S}(\const{S}(\const{S}(\ldots)))$. Compared
-with the more conventional, inductive definition
-\[\keyw{datatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat}) \,\mid\, \const{Infty}\]
+with the more conventional, inductive definition with a dedicated constructor
+for representing infinity,
+%\[\keyw{datatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat}) \,\mid\, \const{Infty}\]
 %
 the codatatype avoids one case by unifying the finite and infinite nonzero cases.
-Moreover, equations such as $\const{S}(\infty) \teq \infty$ hold without additional
-work, because both sides expand to the infinite term
+Moreover, equations such as $\const{S}(\infty) \teq \infty$ hold by default,
+because both sides expand to the infinite term
 $\const{S}(\const{S}(\const{S}(\ldots)))$, which uniquely identifies the
-value $\infty$.
+value~$\infty$.
 
@@
 Datatypes and codatatypes are an integral part of modern proof assistants,
-including Agda, Coq, Isabelle, Matita, and PVS. In recent years, some
-automatic theorem provers have been extended with dedicated reasoning facilities
-for datatypes. The SMT-LIB~2 \cite{barrett-et-al-2010} syntax, implemented by
-most SMT (satisfiability modulo theories) solvers, includes a theory of
-datatypes.
+including Agda, Coq, Isabelle, Matita, and PVS. In recent years, datatypes
+have made their appearance in a few automatic theorem provers. The SMT-LIB~2
+\cite{barrett-et-al-2010} syntax, implemented by most SMT (satisfiability
+modulo theories) solvers, includes a theory of datatypes.
+
+\pagebreak
 
@@
 of infinite sequences, or streams, over an element type $\tau$ is allowed; the
-corresponding datatype would be rejected as non-well-founded.
+corresponding datatype would be rejected as non-well-founded \cite{blanchette-et-al-2015-esop}.
 
@@
 $\const{y} \teq \const{S}(\const{x})$, it deduces $\const{x} \teq \const{y}$.
+The cyclicity and bisimilarity rules are needed to ensure completeness on
+ground problems and the absence of spurious models. They cannot be replaced by
+finite axiomatizations, so they naturally belong in a decision procedure.
+%    * in particular, acyclicity and bisimilarity are necessary for some proofs
+%%      (and cannot be axiomatized finitely)
+%    * and for model finding, without them we quickly get spurious models
+%    * explain how finite model finding works
+%(For the other (co)datatype properties---the injectivity, distinctness, and
+%exhaustiveness of constructors and the selector laws---it is
+%widely recognized that decision procedures can be more efficient than
+%axiomatizations.)
+%axiom.
 
@@
 the $\mu$-term $\const{SCons}(1,\: \mu s.\; \const{SCons}(0,\: \const{SCons}(9,\: s)))$
-denotes the lasso-shaped sequence $1, 0, 9, 0, 9, 0, 9, \ldots$\afterLdots.
+stands for the lasso-shaped sequence $1, 0, 9, 0, 9, 0, 9, \ldots$\afterLdots.
 
 %%% 1090909 is a prime number
-
-The cyclicity and bisimilarity rules are
-needed to ensure completeness on ground problems and the absence of spurious
-models in the context of finite model finding.
-They cannot be replaced by finite
-axiomatizations, so they naturally belong in a decision procedure. 
-%    * in particular, acyclicity and bisimilarity are necessary for some proofs
-%%      (and cannot be axiomatized finitely)
-%    * and for model finding, without them we quickly get spurious models
-%    * explain how finite model finding works
-For the other (co)datatype properties---the injectivity, distinctness, and
-exhaustiveness of constructors and the selector laws---it is
-widely recognized that decision procedures can be more efficient than
-axiomatizations.
-%axiom.
 
@@
 structure, and interpreted by nonempty domains. The set~$\Types$ must contain a
-distinguished type \ty{bool} interpreted as the set of truth
-values $\{\bot, \top\}$. %, and may contain other interpreted types (e.g., \ty{int}, \ty{real}).
-The only predicate is equality ($\teq$) and belongs to the logical symbols.
-Other predicates can be represented as functions that return $\ty{bool}$,
+distinguished type \ty{bool}\begin{report} interpreted as the set of truth
+values $\{\bot, \top\}$\end{report}. %, and may contain other interpreted types (e.g., \ty{int}, \ty{real}).
+The only predicate is equality ($\teq$)\begin{report} and belongs to the logical symbols\end{report}.
+Other predicates can be represented as functions to $\ty{bool}$,
 with $\const{p}(\ldots)$ abbreviating $\const{p}(\ldots) \teq \const{true}$.
+\begin{report}
 The metavariables $\delta,\:\varepsilon$ range over (co)datatypes,
 whereas $\tau,\:\upsilon$ range over arbitrary types.
-\begin{report}When applied to terms, the symbol $=$ denotes syntactic equality.\end{report}
+When applied to terms, the symbol $=$ denotes syntactic equality.\end{report}
 
@@
 
+\begin{report}%
 Because all types must be inhabited (nonempty), a datatype specification is
-admissible only if a ground constructor term can be exhibited that belongs to
-the datatype. This rules out non-well-founded specifications\begin{paper}.\end{paper}
-\begin{report}%
-such as
+admissible only if a ground constructor term can be exhibited.
+This rules out non-well-founded specifications such as
 \[\keyw{datatype}\;\, \ty{fstream}_{\,\tau} \,=\, \const{FSCons}(\tau,\:\, \ty{fstream}_{\,\tau})\]
+For codatatypes, no admissibility check is necessary because there is always a term,
+finite or infinite, that witnesses nonemptiness \cite{blanchette-et-al-2015-esop}.
 \end{report}
-For codatatypes, no admissibility check is necessary because there is always a term,
-finite or infinite, that witnesses nonemptiness.
 %Given the specification
@@
 such~as
+\begin{paper}
+\[\keyw{datatype}\;\, \ty{option}_{\,\tau} \,=\, \const{None} \,\mid\, \const{Some}(\tau)\]%
+\end{paper}%
+\begin{report}%
 \[\begin{aligned}[t]
-      \keyw{datatype}\;\, \ty{enat} & \,=\, \const{Nat}(\ty{nat}) \,\mid\, \const{Infty} \\[-.5\jot]
+      \keyw{datatype}\;\, \ty{option}_{\,\tau} & \,=\, \const{None} \mid \const{Some}(\tau) \\[-.5\jot]
       \keyw{codatatype}\; \,\ty{complex} & \,=\, \const{Complex}(\const{re}{:}\; \ty{real},\; \const{im}{:}\;\ty{real})
@@
 \]
-\begin{report}
 At the semantic level, it makes no difference whether such types are
 introduced as datatypes or as codatatypes.
-\end{report}
+\end{report}%
 %Without loss of generality, we consider that these types are datatypes

### 2014-12-17 tex/codatadp.tex
more

@@
 datatypes are generated by their constructors, whereas the codatatypes are
-viewed through their selectors. By uniformly focusing on the constructors, the
-axiomatic approach emphasizes the commonality between datatypes and
-codatatypes, while sacrificing a theoretically fruitful notion of duality.
+viewed through their selectors.
+%By uniformly focusing on the constructors, the
+%axiomatic approach emphasizes the commonality between datatypes and
+%codatatypes, while sacrificing a theoretically fruitful notion of duality.
 
@@
 such~as
-\begin{paper}
-\[\keyw{datatype}\;\, \ty{option}_{\,\tau} \,=\, \const{None} \,\mid\, \const{Some}(\tau)\]%
+\begin{paper}%
+\vthinspace$\keyw{datatype}~\vthinspace\ty{option}_{\,\tau} = \const{None} \mid \const{Some}(\tau)$.%
 \end{paper}%
@@
 only value is $\mu a.\; A(a)$. Other specimens are
+\begin{paper}%
+\vthinspace$\keyw{codatatype}~\ty{b} = \const{B}(\ty{b},\: \ty{c},\: \ty{b},\: \ty{unit})
+~\keyw{and}~ \ty{c} = \const{C}(\ty{a},\: \ty{unit},\: \ty{b},\: \ty{c})$,
+\end{paper}%
+\begin{report}%
 \[\begin{aligned}[t]
@@
 \]
+\end{report}
 where \ty{unit} is a datatype with the single constructor $\const{Unity} :
@@
 \begin{figure}[t]
+\normalsize
 \centering

### 2014-12-17 tex/codatadp.tex
more

@@
 No (co)datatypes
-  & 401 & 922 & \phantom{0}39 & 1122 \\
+  & 221 & 760 & \phantom{0}36 & 1017 \\
+Datatypes without cyclicity
+  & 227 & 765 & \phantom{0}36 & 1028 \\
+Datatypes with cyclicity
+  & 227 & 771 & \phantom{0}36 & 1034 \\
 Codatatypes without bisimilarity
-  & 410 & 936 & \phantom{0}45 & 1136 \\
+  & 217 & 787 & \phantom{0}34 & 1038 \\
 Codatatypes with bisimilarity
-  & 411 & 939 & \phantom{0}48 & 1139 \\
-Datatypes without cyclicity
-  & 415 & 947 & \phantom{0}43 & 1147 \\
-Datatypes with cyclicity
-  & 418 & 953 & \phantom{0}44 & 1153 \\
+  & 218 & 787 & \phantom{0}36 & 1041 \\
 Full (co)datatypes
-  & 440 & 970 & \phantom{0}49 & 1170 \\[\jot]
+  & 224 & 798 & \phantom{0}36 & 1058 \\[\jot]
 Total number of goals
@@
 \BAD{4.7\%}. Moreover, every aspect of the procedure, including the more
-expensive rules, make a contribution. What is not visible in the table, but
+expensive rules, make a contribution. \textbf{NOT TRUE}: What is not visible in the table, but
 readily available when looking at the raw data \cite{our-eval-data}, is that

### 2014-12-17 tex/codatadp.tex
Most of Section 5

@@
 
-  * Strategies
+This section describes how the calculus in Section~\ref{sec:the-calculus} is implemented within the SMT solver CVC4.
+
+\paragraph{Optimizations}
+We first discuss several optimizations that are not reflected in the presentation of the calculus, which closely follow the approach outlined by Barrett et al in~\cite{}.
+We only briefly mention these optimizations, since each applies to datatypes as well as co-datatype, and most have been presented thoroughly in~\cite{}.
+
+Discriminators are treated as predicate symbols, so a discriminator $\const{d}_i( t )$ is such that $\const{d}_i$ 
+is a predicate symbol, instead of reducing this constraint to $t \teq \const{C}_i( s^1_i( t ), \ldots, s^n_i( t ) )$.
+We have found that this leads to better performance since this reduction introduces terms more eagerly to $\tEc$.
+Handling discriminators requires extending the decision procedure with several additional rules (see~\cite{} for more details), which apply uniformly to both datatypes and co-datatypes.
+Also, selectors are collapsed eagerly:
+if $s^j_i( t )$ is a term in $\tEc$ and $t$ contains the constructor 
+$\const{C}_i( \vec u )$, then we immediately infer $s^j_i( t ) \teq u_j$, whereas the calculus we presented would apply \rn{Split} and \rn{Decomp} before inferring this equality.  
+For the sake of reducing the number of unique constraints considered by the calculus, we compute a normal form for constraints as a pre-processing step.
+In particular, we rewrite $t \teq s$ to $s \teq t$ if $s$ is less than $t$ based on a term ordering,
+replace $\const{C}_i( \vec t ) \teq \const{C}_j( \vec u )$ with $\bot$ when $i \neq j$,
+replace all selector terms of the form $\const{s}^{\,i}_j( \const{C}_j(t_1,\ldots,t_n) )$ by $t_i$,
+and replace all occurrences of discriminators $\const{d}_i( \const{C}_k( \vec t ) )$ by $\top$ if $i=k$ and $\bot$ if $i \neq k$.
+
+\paragraph{Integration into DPLL(T)}
+The calculus is implemented as \emph{theory solver} of CVC4,
+that is, a specialized solver for determining the satisfiability of conjunctions of constraints for its theory.
+%In DPLL(T) \rem{CDCL{T}?},
+%a SAT solver
+Given a theory $T = T_1 \cup \ldots T_n$ and a input set of clauses $F$ in CNF,
+the DPLL(T) \rem{CDCL{T}?} procedure
+(incrementally) builds partial assignments from the atomic formulas of $F$ to truth values such that no clause in $F$ is falsified.
+We may interpret this partial assignment as a set of literals where $a \in M$ for all atoms $a$ that it assigned the value $\top$, and $\neg a \in M$ for all atoms that it assigns the value $\bot$.
+Then, at a high level, by the Nelson-Oppen combination framework,
+each $T_i$-solver for $i = 1, \ldots, n$ takes as input a combination of (1) the purified form of $T_i$-constraints occurring in $M$ where terms containing symbols not belonging to $T_i$ are replaced by fresh variables of the same type,
+(2) additional equalities and disequalities between variables of types not belonging to $T_i$.
+Let us call this set $M_i$.
+Each $T_i$ solver either 
+reports that a subset $C$ of $M_i$ is unsatisfiable in which case $\neg C$ is added as a additional clause to $F$,
+adds an additional clause $C$ to $F$,
+or does nothing.
+When $M$ is a complete assignment for $F$, then a theory solver can only choose to do nothing when $M_i$ is indeed $T_i$-satisfiable.
+
+The rules of the calculus from Section~\ref{sec:the-calculus} are implemented as follows.
+First, for each equality, we associate an \emph{explanation} for that equality.
+\rem{TODO}
+
+Step 1 of the calculus (along with the optimization for collapsing selectors) and the \rn{Singleton} rule are performed eagerly, that is for partial satisfying assignments $M$, while step 2 and the \rn{Split} rule from step 3 are performed only for complete satisfying assignments $M$, and in that order.
+
   * Implementation
-    * mention that discriminators are their own function symbols
     * what is an SMT theory solver

### 2014-12-17 tex/codatadp.tex
Proof of lemma part 2.

@@
 such that $v \vsim_x t_1\interp{t_2}$ for some term $t_1$ with subterm $t_2$ occurring in the range of $\Val$.
-Our completion $\Val^\ast$ is constructed by exhaustively applying the following update to $\Val$:
+We first construct a mapping $\Val'$ by exhaustively applying the following update to $\Val$:
 
@@
 it can only be applied a finite number of times.
-We then replace $\Val^\ast \ec{x}$ with $\nf{(\Val^\ast \ec{x})}$ for all $\ec{x}$.
+We then let $\Val^\ast \ec{x}$ be $\nf{(\Val' \ec{x})}$ for all $\ec{x}$.
 We first state the following properties of $\Val^\ast$, 
@@
 \begin{enumerate}
-\item $\Val^\ast \ec{x}$ is bisimilar with a value in $\M( \tau )$, for all $\ec{x} : \tau$ of $\Ec$,
+\item $\Val^\ast \ec{x}$ is bisimilar with a value in $\M( \tau )$, for all $\ec{x}$ of type $\tau$ of $\Ec$,
 \item
@@
 If $\tau \in \Codata$, then by Lemma~\ref{lem:corecursive-singletons},
-the cardinality of $\tau$ must be one (call the value in its domain $c$).
+the cardinality of $\tau$ must be one.
 Since \rn{Singleton} does not apply to $\Ec$,
@@
 and thus there are no terms in $\mathcal{T}(\Val^\ast)$ of type $\tau$.
-This is a contradiction, since our model completion may assign $c$ to $v_{\ec{x}}$.
+This is a contradiction, since our model completion may assign the value in the domain of $\tau$ to $v_{\ec{x}}$.
 Now, consider the case when $\tau$ is infinite.
@@
 Therefore, by definition, $\Val^\ast \ec{x}$ is bisimilar with a value in $\M( \tau )$.
+
+To show 2, 
+first note that since \rn{Clash} does not apply to $\Ec$,
+for all equivalence classes $\ec{y}$,
+each pair of constructor terms $\const{C}_i( \vec t )$ and $\const{C}_j( \vec u )$ in $\ec{y}$
+are such that $i = j$ since \rn{Clash} does not apply
+are such that $\ec{ \vec t } = \ec{ \vec u }$ since \rn{Decomp} does not apply.
+Thus, 
+if $\const{C}_i( t_1, \ldots, t_n ) \in \ec{y}$,
+$\Val$ was constructed by applying a sequence of substitutions
+that replace $v_{\ec{y}}$ by $\mu v_{\ec{y}}. \const{C}_i( v_{\ec{t_1}}, \ldots, v_{\ec{t_n}} )$.
+Now, say that $\const{C}( t_1, \ldots, t_n )$ occurs in some equivalence class $\ec{x}$.
+Then $\Val \ec{x}$ must be of the form $\mu v_{\ec{x}}. \const{C}_i( w_1, \ldots, w_n )$,
+where for each $i=1,\ldots,n$, $w_i$ was obtained by applying a sequence of substitutions
+of the form mentioned above to $v_{\ec{t_i}}$.
+%Each $w_i$ is either $v_{\ec{x}}$,
+%a variable $v_{\ec{y}}$ where $\ec{y}$ does not contain a constructor,
+%or a term of the form $\mu v_{\ec{y}}. \const{C}_j( \vec u )$.
+While interpreting $w_i$ in $\Val \ec{x}$,
+all free occurrences of $v_{\ec{x}}$ in $\Val \ec{x} \interp{w_i}$ are replaced by $\Val \ec{x}$. \rem{better formalize.}
+Thus, since $\Val \ec{x} \interp{w_i}$ and $\Val \ec{t_i}$ are obtained by applying the same sequence of substitutions to $v_{\ec{t_i}}$,
+we have $\Val \ec{w_i} = \Val \ec{x} \interp{ w_i }$.
+Since $\Val'$ is obtained by applying the same substitution to all terms in the range of $\Val$,
+we have that $\Val' \ec{x} \interp{w_i} = \Val' \ec{t_i}$,
+and hence $\Val^\ast \ec{x} \interp{w_i} = \Val^\ast \ec{t_i}$.
 

### 2014-12-18 tex/codatadp.tex
compile
Add model completion example, more work on proof.

@@
 We then let $\Val^\ast \ec{x}$ be $\nf{(\Val' \ec{x})}$ for all $\ec{x}$.
-We first state the following properties of $\Val^\ast$, 
-where we write $\muvar( t )$ to denote $t$ if $t$ is a variable, or $x$ if $t$ is of the form $\mu x. u$.
+
+\begin{example}
+Let $\tau$ be a type in $\Codata$ with three constructors $\const{A}$, $\const{B}$ and $\const{C}$ each having type $\tau \rightarrow \tau$.
+Let $\Ec$ be the set consisting of equalities
+$\{$
+$y_1 \teq \const{A}( y_2 )$,
+$y_2 \teq \const{B}( x )$,
+$z_1 \teq \const{C}( z_2 )$,
+$z_2 \teq \const{A}( z_1 )$,
+$w_1 \teq \const{A}( x )$,
+$w_2 \teq \const{C}( w_3 )$,
+$x \not\teq y_2$,
+$\}$.
+After running the calculus to saturation on $\Ec$, the mapping $\Val$ contains:
+\[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{2.0em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
+\Val \ec{y_1} & = & \mu v_{\ec{y_1}}. & \const{A}( & \mu v_{\ec{y_2}}. & \const{B}( v_{\ec{x}} ) ) &
+\Val \ec{w_1} & = & \mu v_{\ec{w_1}}. & \const{A}( v_{\ec{x}} )      & \\ 
+\Val \ec{y_2} & = & \mu v_{\ec{y_2}}. & \multicolumn{2}{l}{\const{B}( v_{\ec{x}} )}      & &
+\Val \ec{w_2} & = & \mu v_{\ec{w_2}}. & \const{C}( v_{\ec{w_3}} ) \\
+\Val \ec{z_1} & = & \mu v_{\ec{z_1}}. & \const{C}( & \mu v_{\ec{z_2}}. & \const{A}( v_{\ec{z_1}} ) ) &
+\Val \ec{w_3} & = & v_{\ec{w_3}} \\
+\Val \ec{z_2} & = & \mu v_{\ec{z_2}}. & \const{A}( & \mu v_{\ec{z_1}}. & \const{C}( v_{\ec{z_2}} ) ) & 
+\Val \ec{x} & = & v_{\ec{x}} \\
+\end{array}\]
+
+To construct a completion of $\Val$, we must choose values for the free variables of $\Val$, $v_{\ec{x}}$ and $v_{\ec{w_3}}$.
+The content of set $\mathcal{T}^\M_{v_{\ec{x}}}( \Val )$ is $\alpha$-equivalent to
+\[\begin{array}{c}
+\{
+\mu x. \const{A}( \mu y. \const{B}( x )),
+\mu x. \const{A}( \mu y. \const{B}( y )),
+\mu x. \const{B}( x ),
+\mu x. \const{C}( \mu y. \const{A}( x )),
+\mu x. \const{A}( \mu y. \const{C}( x )),
+\mu x. \const{A}( x )
+\}
+\end{array}\]
+Say that $\M$ interprets variables in $\Ec$ based on the mapping $\Val$,
+for instance, $x$ is interpretted as $\Val \ec{x}$, $y_1$ as $\Val \ec{y_1}$ and so on.
+Assigning a value for $v_{\ec{x}}$ that is $\alpha$-equivalent to a value in $\mathcal{T}^\M_{v_{\ec{x}}}( \Val )$
+may cause values in the range of $\Val$ to become $\alpha$-equivalent, 
+which in turn may cause $\M$ to be inconsistent.
+For instance, say we assign $\mu v_{\ec{x}}. \const{B}( v_{\ec{x}} )$ for $v_{\ec{x}}$.
+After this substitution, $\Val \ec{y_2}$ is $\mu v_{\ec{y_2}}. \const{B}( \mu v_{\ec{x}}. \const{B}( v_{\ec{x}} ) )$, 
+which has normal form $\mu v_{\ec{y_2}}. \const{B}( v_{\ec{y_2}} )$,
+which is $\alpha$-equivalent to $\mu v_{\ec{x}}. \const{B}( v_{\ec{x}} )$.
+However, this contradicts the disequality $x \not\teq y_2$ in $\Ec$.
+
+Inversely, if the value assigned to $v_{\ec{x}}$ is not $\alpha$-equivalent to any term in $\mathcal{T}^\M_{v_{\ec{x}}}( \Val )$,
+then the values in the range of $\Val$ remain distinct.
+For instance, we may assign values such as 
+$\mu v_{\ec{x}}. \const{C}( v_{\ec{x}} )$,
+$\mu v_{\ec{x}}. \const{C}( \mu x_1. \const{A}( x_1) )$, and
+$\mu v_{\ec{x}}. \const{B}( \mu x_1. \const{A}( v_{\ec{x}} ) )$
+to $v_{\ec{x}}$.
+Notice that legal substitutions for $v_{\ec{x}}$ may cause the range of $\Val$ to contain non-normal terms,
+as in after applying the third substitution mentioned above,
+$\Val \ec{w_1}$ is $\mu v_{\ec{w_1}}. \const{A}( \mu v_{\ec{x}}. \const{B}( \mu x_1. \const{A}( v_{\ec{x}} ) ) )$.
+$\Box$
+\end{example}
+
+We first state the following properties of $\Val^\ast$,.
+%where we write $\muvar( t )$ to denote $t$ if $t$ is a variable, or $x$ if $t$ is of the form $\mu x. u$.
 
@@
 there is only one equivalence class of type $\tau$ in $\Ec$,
-and thus there are no terms in $\mathcal{T}(\Val^\ast)$ of type $\tau$.
+and thus there are no terms in $\mathcal{T}^\M_x(\Val^\ast)$ of type $\tau$.
 This is a contradiction, since our model completion may assign the value in the domain of $\tau$ to $v_{\ec{x}}$.
@@
 each pair of constructor terms $\const{C}_i( \vec t )$ and $\const{C}_j( \vec u )$ in $\ec{y}$
-are such that $i = j$ since \rn{Clash} does not apply
-are such that $\ec{ \vec t } = \ec{ \vec u }$ since \rn{Decomp} does not apply.
+are such that $i = j$ since \rn{Clash} does not apply to $\Ec$,
+and are such that $\ec{ \vec t } = \ec{ \vec u }$ since \rn{Decomp} does not apply to $\Ec$.
 Thus, 
@@
 
+To show 3,
+first note that since \rn{Bisimilar} does not apply,
+$\Val \ec{x} \sim \Val \ec{y}$ if and only if $\ec{x} = \ec{y}$.
+Since \rem{TODO}
+we have $\nf{(\Val \ec{x})} \sim \nf{(\Val \ec{y})}$ if and only if $\ec{x} = \ec{y}$.
+We show that for any substitution $\sigma$ applied to $\Val$ while constructing its completion is such that
+$\nf{(\Val \ec{x} \sigma)} \sim \nf{(\Val \ec{y} \sigma)}$ if and only if $\ec{x} = \ec{y}$.
 \rem{TODO}

### 2014-12-18 tex/codatadp.tex
updated data

--- 2014-12-17 tex/codatadp.tex
@@
 \def\thewordpaper{report}
-\newcommand\dotReportFootnote[1]{.\footnote{#1}}
+\renewcommand\dotReportFootnote[1]{.\footnote{#1}}
 \end{report}
@@
 No (co)datatypes
-  & 221 & 760 & \phantom{0}36 & 1017 \\
+  & 221 & 760 & \phantom{0}57 & 1038 \\
 Datatypes without cyclicity
-  & 227 & 765 & \phantom{0}36 & 1028 \\
+  & 227 & 765 & \phantom{0}57 & 1049 \\
 Datatypes with cyclicity
-  & 227 & 771 & \phantom{0}36 & 1034 \\
+  & 227 & 771 & \phantom{0}57 & 1055 \\
 Codatatypes without bisimilarity
-  & 217 & 787 & \phantom{0}34 & 1038 \\
+  & 217 & 787 & \phantom{0}55 & 1059 \\
 Codatatypes with bisimilarity
-  & 218 & 787 & \phantom{0}36 & 1041 \\
+  & 218 & 787 & \phantom{0}58 & 1063 \\
 Full (co)datatypes
-  & 224 & 798 & \phantom{0}36 & 1058 \\[\jot]
+  & 224 & 798 & \phantom{0}58 & 1080 \\[\jot]
 Total number of goals
-  & 879 & 2974\phantom{0} & 117 & 3970
+  & 879 & 2974\phantom{0} & 317 & 4170
 \end{tabular}\end{center}

### 2014-12-18 tex/codatadp.tex
more

@@
 The content of set $\mathcal{T}^\M_{v_{\ec{x}}}( \Val )$ is $\alpha$-equivalent to
-\[\begin{array}{c}
+\[%\begin{array}{c}
 \{
@@
 \}
-\end{array}\]
+%\end{array}
+\]
 Say that $\M$ interprets variables in $\Ec$ based on the mapping $\Val$,
@@
 the stronger decision procedures subsume the weaker ones. In the context of
-Sledgehammer, the power of interpreted (co)datatype is comparable with that of
-arithmetic \cite{blanchette-et-al-2013-smt}.
+Sledgehammer, the power of interpreted (co)datatypes is roughly
+comparable to that of arithmetic \cite{blanchette-et-al-2013-smt}.
 
@@
 %  it and explaining it (and, before that, understanding it)
+
+\begin{verbatim}
+primcorec FIXME_x :: "nat tree" where
+  "FIXME_x = Node 0 num FIXME_x"
+
+lemma FIXME_x_unique: "x = Node 0 num x \<Longrightarrow> x = FIXME_x"
+proof (coinduction arbitrary: x rule: tree.coinduct_strong)
+  case (Eq_tree x)
+  show ?case
+    by (subst (1 2 3 4) Eq_tree) (simp add: eqTrueI[OF Eq_tree])
+qed
+\end{verbatim}
 

### 2014-12-18 tex/codatadp.tex
more

@@
 
-\begin{verbatim}
-primcorec FIXME_x :: "nat tree" where
-  "FIXME_x = Node 0 num FIXME_x"
-
-lemma FIXME_x_unique: "x = Node 0 num x \<Longrightarrow> x = FIXME_x"
-proof (coinduction arbitrary: x rule: tree.coinduct_strong)
-  case (Eq_tree x)
-  show ?case
-    by (subst (1 2 3 4) Eq_tree) (simp add: eqTrueI[OF Eq_tree])
-qed
-\end{verbatim}
+Among the four bisimilarity proofs, three corresponded to one-line
+proofs in Isabelle, of the form \keyw{by}~\textit{coinduction}~\textit{auto}
+\cite{blanchette-et-al-2014-impl}. The fourth proof was somewhat more elaborate:
+%
+\begin{quote}
+\keyw{lemma} \,\textit{X0\_unique}: \,$x = \const{Node}\;0\;\const{num}\; x \Longrightarrow x = \const{X0}$ \\
+\keyw{proof} \,(\textit{coinduction arbitrary}: $x$ \textit{rule}: \textit{tree.coinduct\_strong}) \\
+\noindent\hbox{}\quad  \keyw{case} (\textit{Eq\_tree} $x$) \keyw{show} \textit{?case} \\
+\noindent\hbox{}\qquad  \keyw{by} (\textit{subst} (1 2 3 4) \textit{Eq\_tree}) (\textit{simp add}: \textit{eqTrueI}[OF \textit{Eq\_tree}]) \\
+\keyw{qed}
+\end{quote}
+%
+\noindent
+where \const{X0} is defined as $\const{X0} =
+\const{Node}\;0\;\const{num}\;\const{X0}$.
 

### 2014-12-18 tex/codatadp.tex
Minor

@@
 To construct a completion of $\Val$, we must choose values for the free variables of $\Val$, $v_{\ec{x}}$ and $v_{\ec{w_3}}$.
-The content of set $\mathcal{T}^\M_{v_{\ec{x}}}( \Val )$ is $\alpha$-equivalent to
+The content of the set $\mathcal{T}^\M_{v_{\ec{x}}}( \Val )$ is $\alpha$-equivalent to:
 \[%\begin{array}{c}
@@
 \]
-Say that $\M$ interprets variables in $\Ec$ based on the mapping $\Val$,
+Now, consider a model $\M$ that interprets variables in $\Ec$ based on the mapping $\Val$,
 for instance, $x$ is interpretted as $\Val \ec{x}$, $y_1$ as $\Val \ec{y_1}$ and so on.
@@
 However, this contradicts the disequality $x \not\teq y_2$ in $\Ec$.
-
 Inversely, if the value assigned to $v_{\ec{x}}$ is not $\alpha$-equivalent to any term in $\mathcal{T}^\M_{v_{\ec{x}}}( \Val )$,
-then the values in the range of $\Val$ remain distinct.
+then the values in the range of $\Val$ remain $\alpha$-disequivalent.
 For instance, we may assign values such as 

### 2014-12-18 tex/codatadp.tex
Assumptions for section 3

@@
 
-  * SMT background
-
-\rem{Discuss signature $\Sig$, theory $\thD$}
-\rem{DPLL(T) : we can restrict ourselves to conjunctions.}
-This section presents a calculus for determining the satisfiability of conjunctions of $\thD$-constraints.
-
-Our calculus consists of derivation rules that operate on a set $\Ec$ of equalities and disequalities between $\Sig$-terms.
+This section presents a decision procedure for \rem{introduce ``$\thD$" in previous section?}
+determining the $\thD$-satisfiability of sets of constraints built from a fixed signature $\Sig$.
+We will make the following assumptions about $\Sig$.
+First, all codatatypes in $\Codata$ are corecursive.
+This is not restrictive, since checking the $\thD$-satisfiability for constraints involving a non-corecursive co-datatype $\tau$ can be trivially reduced to
+checking the satisfiability of the same constraint involving a (non-recursive) datatype of the same form as $\tau$. \rem{revise?}
+Second, all types $\tau \in \Nondata$ have infinite cardinality.
+This also is not restrictive, since each interpreted type $\tau \in \Nondata$ having finite cardinality $n$
+can be replaced by a enumeration datatype with $n$ constructors
+\footnote{Doing so simplifies the presentation of the calculus, but is not done in practice. \rem{More?}}.
+Additionally, since $\Ec$ is ground, we know that $\Ec$ cannot entail an upper bound on the cardinality of any uninterpreted type $\tau \in \Nondata$, 
+and thus without loss of completeness we may treat all other types as infinite.
+
+Our calculus consists of derivation rules that operate on a set $\Ec$ of equalities and disequalities between $\Sig$-terms,
+where $\Ec$ is interpreted as a conjunction.
 We write $\tEc$ to denote the set of $\Sig$-terms occurring in $\Ec$.
@@
 Since \rn{Split} does not apply to $\Ec$, we have $\tau \not\in \Data$.
-If $\tau \in \Codata$, then by Lemma~\ref{lem:corecursive-singletons},
+If $\tau \in \Codata$, then $\tau$ is corecursive by assumption, and by Lemma~\ref{lem:corecursive-singletons},
 the cardinality of $\tau$ must be one.
@@
 We first discuss several optimizations that are not reflected in the presentation of the calculus, which closely follow the approach outlined by Barrett et al in~\cite{}.
-We only briefly mention these optimizations, since each applies to datatypes as well as co-datatype, and most have been presented thoroughly in~\cite{}.
+We only briefly mention these optimizations, since each applies to datatypes as well as codatatypes, and most have been presented thoroughly in~\cite{}.
 
@@
 We have found that this leads to better performance since this reduction introduces terms more eagerly to $\tEc$.
-Handling discriminators requires extending the decision procedure with several additional rules (see~\cite{} for more details), which apply uniformly to both datatypes and co-datatypes.
+Handling discriminators requires extending the decision procedure with several additional rules (see~\cite{} for more details), which apply uniformly to both datatypes and codatatypes.
 Also, selectors are collapsed eagerly:

### 2014-12-18 tex/codatadp.tex
More on implementation.

@@
 For the sake of reducing the number of unique constraints considered by the calculus, we compute a normal form for constraints as a pre-processing step.
-In particular, we rewrite $t \teq s$ to $s \teq t$ if $s$ is less than $t$ based on a term ordering,
+In particular, we 
+replace $t \teq s$ by $s \teq t$ if $s$ is less than $t$ based on a term ordering,
 replace $\const{C}_i( \vec t ) \teq \const{C}_j( \vec u )$ with $\bot$ when $i \neq j$,
@@
 Each $T_i$ solver either 
-reports that a subset $C$ of $M_i$ is unsatisfiable in which case $\neg C$ is added as a additional clause to $F$,
+reports that a subset $C$ of $M_i$ is $T_i$-unsatisfiable in which case $\neg C$ is added as a additional clause to $F$,
 adds an additional clause $C$ to $F$,
@@
 
-The rules of the calculus from Section~\ref{sec:the-calculus} are implemented as follows.
-First, for each equality, we associate an \emph{explanation} for that equality.
-\rem{TODO}
-
-Step 1 of the calculus (along with the optimization for collapsing selectors) and the \rn{Singleton} rule are performed eagerly, that is for partial satisfying assignments $M$, while step 2 and the \rn{Split} rule from step 3 are performed only for complete satisfying assignments $M$, and in that order.
-
-  * Implementation
-    * what is an SMT theory solver
-    * sharing is caring (and other "refinements" on calculus)
+Assume $\Ec$ is initially the set $M_i$ described above.
+For each equality added to $\Ec$, we associate a set of equalities from $M_i$ that together $\thD$-entail $t \teq s$,
+which we call its \emph{explanation}.
+Similarly, each $\Val \ec{x}$ is also associated an explanation, that is, 
+a set of equalities from $M_i$ that $\thD$-entail that the values of $\ec{x}$ in models of $\Ec$ are of the form $\Val \ec{x}$.
+For instance, if $x \teq \const{C}( x ) \in M_i$, then $x \teq \const{C}( x )$ is a (possible) explanation for $\Val \ec{x} = \mu v_{\ec{x}}. \const{C}( v_{\ec{x}} )$.
+Thus, the rules of the calculus are implemented as follows.
+For all rules with conclusion $\bot$,
+we report the union of the explanations for all premises is $\thD$-unsatisfiable.
+For the \rn{Split} rule, we add a clause of the form
+$(\const{C}_1( s^1_1( t ), \ldots, s^{a_1}_1( t ) ) \vee \ldots \vee t \teq \const{C}_n( s^1_n( t ), \ldots, s^{a_n}_n( t ) ))$
+to $F$.
+Notice that decisions on which branch to take are thus performed externally by the SAT solver, and not by the theory solver.
+All other rules add equalities to the internal state of the theory solver.
+
+Step 1 of the calculus (along with the optimization for collapsing selectors) and the \rn{Singleton} rule are performed eagerly, 
+that is for partial satisfying assignments $M$, while step 2 and the \rn{Split} rule from step 3 are performed only for complete satisfying assignments $M$, and in that order.
+
+    * sharing is caring
     * theory combination more complicated (and Theorem Completeness too weak)
-      * need something like the string paper
-    * cooperation with Nelson-Oppen / integration into SMT solver
-    * "mu" stuff
-    * no "internal search"
-  * Enumerator?
-    * for model generation
 

### 2014-12-19 tex/codatadp.tex
minor

--- 2014-12-18 tex/codatadp.tex
@@
 To construct a completion of $\Val$, we must choose values for the free variables of $\Val$, $v_{\ec{x}}$ and $v_{\ec{w_3}}$.
-The content of the set $\mathcal{T}^\M_{v_{\ec{x}}}( \Val )$ is $\alpha$-equivalent to:
+The content of set $\mathcal{T}^\M_{v_{\ec{x}}}( \Val )$ is $\alpha$-equivalent to
 \[%\begin{array}{c}
@@
 \]
-Now, consider a model $\M$ that interprets variables in $\Ec$ based on the mapping $\Val$,
+Say that $\M$ interprets variables in $\Ec$ based on the mapping $\Val$,
 for instance, $x$ is interpretted as $\Val \ec{x}$, $y_1$ as $\Val \ec{y_1}$ and so on.
@@
 However, this contradicts the disequality $x \not\teq y_2$ in $\Ec$.
+
 Inversely, if the value assigned to $v_{\ec{x}}$ is not $\alpha$-equivalent to any term in $\mathcal{T}^\M_{v_{\ec{x}}}( \Val )$,
-then the values in the range of $\Val$ remain $\alpha$-disequivalent.
+then the values in the range of $\Val$ remain distinct.
 For instance, we may assign values such as 
@@
 within the supported fragment. We also included two unpublished theories by
-Peter Gammie and Andreas Lochbihler (\gandl). To also exercise the the support
+Peter Gammie and Andreas Lochbihler (\gandl), about Bird and Stern--Brocot trees.
+To also exercise the the support
 for datatypes, the benchmarks are complemented by theories about various list

### 2014-12-19 tex/codatadp.tex
more
moved stuff around + ideas

--- 2014-12-18 tex/codatadp.tex
@@
 
-\paragraph{Specifications.}
+%\paragraph{Specifications.}
 In an SMT problem, the signature is normally given by specifying the
@@
 \end{report}
+
+A type $\delta$ depends on another type $\varepsilon$ if $\varepsilon$ is the
+type of an argument to one of $\delta$'s constructors. Semantically, a set of
+types is mutually (co)recursive if and only if the associated dependency graph
+is strongly connected. Types can be declared together as mutually
+(co)recursive even if they are not actually (co)recursive, but the
+semantic notion is more precise and is the one that interests us.
+%
+In addition, nothing forbids non(co)recursive specifications
+such~as
+\begin{paper}%
+\vthinspace$\keyw{datatype}~\vthinspace\ty{option}_{\,\tau} = \const{None} \mid \const{Some}(\tau)$.%
+\end{paper}%
+\begin{report}%
+\[\begin{aligned}[t]
+      \keyw{datatype}\;\, \ty{option}_{\,\tau} & \,=\, \const{None} \mid \const{Some}(\tau) \\[-.5\jot]
+      \keyw{codatatype}\; \,\ty{complex} & \,=\, \const{Complex}(\const{re}{:}\; \ty{real},\; \const{im}{:}\;\ty{real})
+\end{aligned}
+\]
+At the semantic level, it makes no difference whether such types are
+introduced as datatypes or as codatatypes.
+\end{report}%
+%Without loss of generality, we consider that these types are datatypes
+%instead of codatatypes.
+
 %Given the specification
@@
 
-\paragraph{Characterization.}
+%\paragraph{Characterization.}
 Datatypes can be defined semantically as the initial model of the
@@
 
-% TODO: where does this go?
-%      * when looking at the universal ground (?) theory:
+TODO: Andy, introduce theory here. There were some ideas below in the comments
+you might want to mention either here or in Section 3. I'm not sure what's the
+best way to justify it though, but for the conference at least we can just
+claim it. (we won't have the place for a proof anyway).
+
+%    * when looking at the ground theory:
 %        enough to consider acyclicity -- no way to specify infinite objects
@@
 %    * codatatypes:
-%      * coinduction axiom (schema)
-%        * consequence: infinite objects are allowed, but
-%          two objects yielding the same observations must be equal
-
-%      * again: when looking at the universal ground (?) theory, enough to
-%        consider bisimilarity/bisimulation
-
-% specifies the
-
-\paragraph{Special Cases.}
-A type $\delta$ depends on another type $\varepsilon$ if $\varepsilon$ is the
-type of an argument to one of $\delta$'s constructors. Semantically, a set of
-types is mutually (co)recursive if and only if the associated dependency graph
-is strongly connected. Types can be declared together as mutually
-(co)recursive even if they are not actually (co)recursive; however, the
-semantic notion is more precise and is the one that interests us.
-%
-In addition, nothing forbids non(co)recursive specifications
-such~as
-\begin{paper}%
-\vthinspace$\keyw{datatype}~\vthinspace\ty{option}_{\,\tau} = \const{None} \mid \const{Some}(\tau)$.%
-\end{paper}%
-\begin{report}%
-\[\begin{aligned}[t]
-      \keyw{datatype}\;\, \ty{option}_{\,\tau} & \,=\, \const{None} \mid \const{Some}(\tau) \\[-.5\jot]
-      \keyw{codatatype}\; \,\ty{complex} & \,=\, \const{Complex}(\const{re}{:}\; \ty{real},\; \const{im}{:}\;\ty{real})
-\end{aligned}
-\]
-At the semantic level, it makes no difference whether such types are
-introduced as datatypes or as codatatypes.
-\end{report}%
-%Without loss of generality, we consider that these types are datatypes
-%instead of codatatypes.
-
+%      * when looking at the ground theory, enough to
+%        consider bisimilarity -- no way to express that certain infinite
+%        acyclic objects do not exist
+
+%\paragraph{Special Cases.}
 Some codatatypes are so degenerate as to have infinite values
@@
 blah blah. \textbf{TODO FIXME}
+Morgan Deters?
 %

### 2014-12-19 tex/codatadp.tex
more

@@
 you might want to mention either here or in Section 3. I'm not sure what's the
-best way to justify it though, but for the conference at least we can just
-claim it. (we won't have the place for a proof anyway).
-
-%    * when looking at the ground theory:
+best way to justify it though, but for the conference at least we can just claim it. (we
+won't have the place for a proof anyway).
+
+When looking at the ground theory, induction can be replaced by the acyclicity
+axiom schema, which states that any ground constructor term cannot be equal to
+a variable it contains \cite{xxx}. Dually, coinduction can be replaced by the
+bisimilarity axiom schema, which asserts that two values having the same
+corecursive specification---e.g., $x \teq \const{C}(x)$ and $y \teq
+\const{C}(y)$---must be identical.
+
+\begin{report}
+PROVE?
+\end{report}
+
+%    * :
 %        enough to consider acyclicity -- no way to specify infinite objects

### 2014-12-19 tex/codatadp.tex
more

--- 2014-12-19 tex/codatadp.tex
@@
 you might want to mention either here or in Section 3. I'm not sure what's the
-best way to justify it though, but for the conference at least we can just
-claim it. (we won't have the place for a proof anyway).
-
-%    * when looking at the ground theory:
+best way to justify it though, but for the conference at least we can just claim it. (we
+won't have the place for a proof anyway).
+
+When looking at the ground theory, induction can be replaced by the acyclicity
+axiom schema, which states that any ground constructor term cannot be equal to
+a variable it contains \cite{xxx}. Dually, coinduction can be replaced by the
+bisimilarity axiom schema, which asserts that (co)recursive definitions giving
+rise to the same infinite expansion---e.g., $x \teq \const{C}(x)$ and $y \teq
+\const{C}(\const{C}(y))$---must be identical.
+
+\begin{report}
+TODO: State more precisely and proof?
+\end{report}
+
+%    * :
 %        enough to consider acyclicity -- no way to specify infinite objects

### 2014-12-19 tex/codatadp.tex
more

@@
 % change?
-\newcommand{\thO}{T_{o}}
-\newcommand{\thD}{T_{d}}
+\newcommand{\thO}{T_{\mathrm{o}}}
+\newcommand{\thD}{T_{\mathrm{cd}}}
 
@@
 (Section~\ref{sec:the-theory-of-co-datatypes}). Intuitively, the
-values of a codatatype consist of all well-typed finite and infinite ground constructor
+values of a codatatype consist of all well-typed finite and infinite ground
+constructor
 terms, and only those. For example, the coinductive specification
@@
 
-\section{The Theory of (Co)datatypes}
+\section{%The Theory of
+(Co)datatypes}
 \label{sec:the-theory-of-co-datatypes}
@@
 %\paragraph{Characterization.}
-Datatypes can be defined semantically as the initial model of the
-selector--constructor equations. This is the view followed by Barrett et al.\
-\cite{barrett-et-al-2010}.
+One way to define datatypes semantically is as the initial model of the
+selector--constructor equations \cite{barrett-et-al-2010}.
 \begin{report}
@@
 In addition, codatatypes are guaranteed to contain all values corresponding to
-infinite ground constructor terms. In general, this cannot be captured by a
+infinite ground constructor terms.
+\begin{report}
+In general, this cannot be captured by a
 first-order axiomatization, since there may be uncountably many of them.
-\begin{report}
 For example, $\ty{stream}_{\,\ty{int}}$ is isomorphic to the uncountable
@@
 
-This section presents a decision procedure for \rem{introduce ``$\thD$" in previous section?}
-determining the $\thD$-satisfiability of sets of constraints built from a fixed signature $\Sig$.
-We will make the following assumptions about $\Sig$.
-First, all codatatypes in $\Codata$ are corecursive.
-This is not restrictive, since checking the $\thD$-satisfiability for constraints involving a non-corecursive co-datatype $\tau$ can be trivially reduced to
-checking the satisfiability of the same constraint involving a (non-recursive) datatype of the same form as $\tau$. \rem{revise?}
-Second, all types $\tau \in \Nondata$ have infinite cardinality.
-This also is not restrictive, since each interpreted type $\tau \in \Nondata$ having finite cardinality $n$
-can be replaced by a enumeration datatype with $n$ constructors
-\footnote{Doing so simplifies the presentation of the calculus, but is not done in practice. \rem{More?}}.
-Additionally, since $\Ec$ is ground, we know that $\Ec$ cannot entail an upper bound on the cardinality of any uninterpreted type $\tau \in \Nondata$, 
-and thus without loss of completeness we may treat all other types as infinite.
-
-Our calculus consists of derivation rules that operate on a set $\Ec$ of equalities and disequalities between $\Sig$-terms,
-where $\Ec$ is interpreted as a conjunction.
+The decision procedure for the ground theory of (co)datatypes $\thD$ determines the
+$\thD$-satisfiability of sets of ground constraints over a fixed
+signature~$\Sig$. It is presented as a calculus, whose rules can be applied in
+any order until saturation.
+
+The calculus consists of derivation rules that operate on a set $\Ec$ of
+equalities and disequalities between $\Sig$-terms, where $\Ec$ is interpreted
+as a conjunction.
 We write $\tEc$ to denote the set of $\Sig$-terms occurring in $\Ec$.
@@
 constraints added to $\Ec$ are normalized in this manner.
+
+To simplify the presentation, we make the following assumptions about $\Sig$.
+First, all codatatypes in $\Codata$ are corecursive. This is not restrictive
+because noncorecursive codatatypes can be considered as nonrecursive
+datatypes.
+Second, all types $\tau \in \Nondata$ have infinite cardinality.
+This is not overly restrictive, because each interpreted type $\tau \in \Nondata$
+having finite cardinality~$n$
+can be replaced by an enumeration datatype with $n$~nullary constructors.
+Moreover, since the constraints are ground, they cannot entail an upper bound
+on the cardinality of any uninterpreted type $\tau \in \Nondata$; thus,
+we may without loss of completeness treat all other types as infinite.
 
@@
 %, or informally are \emph{observationally equivalent}.
-%For convienience, we use $\mu$-terms to refer to (classes of) values for both co-datatype and datatype terms,
+%For convienience, we use $\mu$-terms to refer to (classes of) values for both codatatype and datatype terms,
 %where in the latter case, a $\mu$-term with a bound variable denotes an infeasible term.
@@
 Intuitively, a $\mu$-term $t$ of type $\tau$ describes a class of values of $\tau$, 
-%either in the case when $\tau$ is a co-datatype or datatype type,
+%either in the case when $\tau$ is a codatatype or datatype type,
 where a cyclic $t$ describes an infeasible class of values when
@@
 For bisimilarity, if there are two equivalence classes $\ec{t_1}$ and $\ec{t_2}$ 
-of co-datatype type whose classes of values $\Val \ec{t_1}$ and $\Val \ec{t_2}$ are equivalent for a capture-avoiding renaming of $\mu$-bound variables,
+of codatatype type whose classes of values $\Val \ec{t_1}$ and $\Val \ec{t_2}$ are equivalent for a capture-avoiding renaming of $\mu$-bound variables,
 then we can conclude $t_1 \teq t_2$.

### 2014-12-19 tex/codatadp.tex
more

@@
 We assume that all $\Sig$-terms in $\Ec$ are \emph{normalized}, meaning that all subterms of the form $\const{s}^{\,k}_j( \const{C}_j(t_1,\ldots,t_n) )$
-are simplified to $t_i$, and moreover assume that all %additional
-constraints added to $\Ec$ are normalized in this manner.
+are simplified to $t_k$. All constraints added to $\Ec$ are also implicitly
+normalized in this manner.
 
@@
 
-We present the calculus in three steps.
-In the first step, we compute the bidirectional closure of $\Ec$;
-in the second step, we make inferences based on cyclicity and bisimilarity;
-and in the third step, %(when necessary),
-we branch on constructor types for various terms in $\tEc$.
+The calculus consists of three sets of rules corresponding to three
+successive phases.
+The first phase computes the bidirectional closure of $\Ec$.
+The second phase makes inferences based on cyclicity (for datatypes)
+and bisimilarity (for codatatypes).
+The third phase performs case splitting on constructors
+for various terms in $\tEc$.
 Following the conventions from~\cite{}, 

### 2014-12-19 tex/codatadp.tex
the rule needs to be called "Acyclic", otherwise there's a real dissonance with bisimilar

@@
 processes. We present a decision procedure that combines reasoning about
-datatypes and codatatypes. The dual of the cyclicity rule for datatypes is a
+datatypes and codatatypes. The dual of the acyclicity rule for datatypes is a
 bisimilarity rule that identifies equal values, also in the presence of cyclic
@@
 
-Second, \emph{a bisimilarity rule takes the place of the cyclicity rule of datatypes.}
+Second, \emph{a bisimilarity rule takes the place of the acyclicity rule of datatypes.}
 Cyclic constraints such as
@@
 $\const{y} \teq \const{S}(\const{x})$, it deduces $\const{x} \teq \const{y}$.
-The cyclicity and bisimilarity rules are needed to ensure completeness on
+The acyclicity and bisimilarity rules are needed to ensure completeness on
 ground problems and the absence of spurious models. They cannot be replaced by
@@
 The first phase computes the bidirectional closure of $\Ec$.
-The second phase makes inferences based on cyclicity (for datatypes)
+The second phase makes inferences based on acyclicity (for datatypes)
 and bisimilarity (for codatatypes).
@@
 \)
-\rn{Cyclic}
+\rn{Acyclic}
 \\[4\jot]
@@
 \end{tabular}
-\caption{\,Rules for cyclicity and bisimilarity%.
+\caption{\,Rules for acyclicity and bisimilarity%.
 }
@@
 
-\paragraph{Step 2: Applying Cyclicity and Bisimilarity.}
+\paragraph{Step 2: Applying Acyclicity and Bisimilarity.}
 In the second step,
@@
 
-Given $\Val$, the rules for cyclicity and bisimilarity (shown in Figure~\ref{fig:ab-rules}) can be stated as follows.
-For cyclicity, if there is an equivalence class $\ec{t_1}$ of datatype type whose class of values $\Val \ec{t_1} = \mu x. t_2$ are cyclic,
+Given $\Val$, the rules for acyclicity and bisimilarity (Figure~\ref{fig:ab-rules}) can be stated as follows.
+For acyclicity, if there is an equivalence class $\ec{t_1}$ of datatype type whose class of values $\Val \ec{t_1} = \mu x. t_2$ are cyclic,
 as indicated by the assumption $x \in FV( t_2 )$,
@@
 We prove this by structural induction on the (closed) derivation tree with root node $\Ec$.
-First, if the derivation tree is an application of \rn{Conflict}, \rn{Clash}, or \rn{Cyclic},
+First, if the derivation tree is an application of \rn{Conflict}, \rn{Clash}, or \rn{Acyclic},
 then $\Ec$ is unsatisfiable in $\thD$.
@@
 For \rn{Clash}, this is a consequence of distinctness.
-For \rn{Cyclic}, our construction of $\Val$ indicates that the class of values that $t_1$ can take in models of $\Ec$ is infeasible,
+For \rn{Acyclic}, our construction of $\Val$ indicates that the class of values that $t_1$ can take in models of $\Ec$ is infeasible,
 and thus $\Ec$ is unsatisfiable.
@@
 By construction, $\Val^\ast \ec{x}$ is normal.
-Since \rn{Cyclic} does not apply, $\Val \ec{x}$ is acyclic when $\tau \in \Data$.
+Since \rn{Acyclic} does not apply, $\Val \ec{x}$ is acyclic when $\tau \in \Data$.
 Moreover, our construction of $\Val^\ast$ applies substitutions of the form 
@@
   & 221 & 760 & \phantom{0}57 & 1038 \\
-Datatypes without cyclicity
+Datatypes without acyclicity
   & 227 & 765 & \phantom{0}57 & 1049 \\
-Datatypes with cyclicity
+Datatypes with acyclicity
   & 227 & 771 & \phantom{0}57 & 1055 \\
@@
 CVC4 was run on each problem several times, with the support for datatypes and
-codatatypes either enabled or disabled. The contributions of the cyclicity and
+codatatypes either enabled or disabled. The contributions of the acyclicity and
 bisimilarity rules were also measured, so find out whether these somewhat

### 2014-12-19 tex/codatadp.tex
step -> phase

@@
 
-\paragraph{Step 1: Computing the Bidirectional Closure.}
+\paragraph{Phase 1: Computing the Bidirectional Closure.}
 Figure~\ref{fig:cc-rules} gives the basic rules of the calculus.
@@
 
-\paragraph{Step 2: Applying Acyclicity and Bisimilarity.}
-In the second step,
+\paragraph{Phase 2: Applying Acyclicity and Bisimilarity.}
+In the second phase,
 we assume that $\Ec$ is saturated with respect to the rules in Figure~\ref{fig:cc-rules}---that is,
@@
 
-%For presentation of the rules in this step, we rely on a representation of terms in the $\mu$-notation.
-The rules of this step contain premises that involve a mapping $\Val$ from equivalence classes to terms containing $\mu$-bindings.
+%For presentation of the rules in this phase, we rely on a representation of terms in the $\mu$-notation.
+The rules of this phase contain premises that involve a mapping $\Val$ from equivalence classes to terms containing $\mu$-bindings.
 Recall that $\mu$-bindings can be used for representing possibly cyclic terms and values.
@@
 $\tau \in \Data$.
-At the beginning of this step, 
+At the beginning of this phase, 
 we compute a mapping $\Val$ from equivalence classes to $\mu$-terms 
@@
 $\{ x, \const{C}(y,\const{C}(y,x)) \}$, $\{ \const{C}(y,x) \}$, and $\{ y \}$.
-We then construct $\Val$ using the aforementioned steps, after which we obtain:
+We then construct $\Val$ using the aforementioned phases, after which we obtain:
 \[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}r@{\hspace{.5em}}l@{\hspace{.5em}}r@{\hspace{.5em}}l}
@@
 
-\paragraph{Step 3: Branching.}
+\paragraph{Phase 3: Branching.}
 If neither a rule from Figure \ref{fig:cc-rules}~or~\ref{fig:ab-rules} applies to $\Ec$, 
@@
 it remains only to define how $\M$ interprets wrong-applied selector terms and variables.
-For the latter, this will be based on the mapping $\Val$ constructed in step two of our calculus.
+For the latter, this will be based on the mapping $\Val$ constructed in the second phase of our calculus.
 We compute the \emph{completion} of $\Val$ for normal model $\M$, which we write as $\Val^\ast$,
@@
 that interprets wrongly-applied selectors based on equality information in $\Fc$,
-and interprets the variables of $\Fc$ based on the completion $\Val^\ast$ of the mapping $\Val$ from step two of our calculus.
+and interprets the variables of $\Fc$ based on the completion $\Val^\ast$ of the mapping $\Val$ from the second phase of our calculus.
 For the latter, let $\M( x )$ be the value in $\M$ that is bisimilar with $\Val^\ast \ec{x}$ for each variable $x$ in $\mathcal{T}(\Fc)$.
@@
 $\const{C}_i( \vec u )$, then we immediately infer $s^j_i( t ) \teq u_j$, whereas the calculus we presented would apply \rn{Split} and \rn{Decomp} before inferring this equality.  
-For the sake of reducing the number of unique constraints considered by the calculus, we compute a normal form for constraints as a pre-processing step.
+For the sake of reducing the number of unique constraints considered by the calculus, we compute a normal form for constraints as a preprocessing step.
 In particular, we 
@@
 
-Step 1 of the calculus (along with the optimization for collapsing selectors) and the \rn{Singleton} rule are performed eagerly, 
-that is for partial satisfying assignments $M$, while step 2 and the \rn{Split} rule from step 3 are performed only for complete satisfying assignments $M$, and in that order.
+Phase 1 (along with the optimization for collapsing selectors) and the \rn{Singleton} rule are performed eagerly, 
+that is for partial satisfying assignments $M$, while phase~2 and the \rn{Split} rule from phase~3 are performed only for complete satisfying assignments $M$, and in that order.
 

### 2014-12-19 tex/codatadp.tex
more edits

@@
 
-The calculus consists of derivation rules that operate on a set $\Ec$ of
-equalities and disequalities between $\Sig$-terms, where $\Ec$ is interpreted
-as a conjunction.
+The derivation rules operate on a set $\Ec$ of equalities and disequalities
+between $\Sig$-terms, where $\Ec$ is interpreted as a conjunction.
 We write $\tEc$ to denote the set of $\Sig$-terms occurring in $\Ec$.
@@
 for various terms in $\tEc$.
-Following the conventions from~\cite{}, 
-the derivation rules of our calculus are given in \emph{guarded assignment form},
-where a rule can be applied to $\Ec$ if it meets all of the specified preconditions for $\Ec$.
-The conclusion of a rule either describes equalities to be added to $\Ec$ (in which case we call an application of it \emph{nonterminal}),
-or is $\bot$ (in which case we call an application of it \emph{terminal}).
-A rule may have multiple conclusions separated by $\parallel$, which denotes nondeterministic branching.
+%Following the conventions from~\cite{}, 
+%The derivation rules are given in \relax{guarded assignment form},
+%where
+A rule can be applied to $\Ec$ if all the specified preconditions are met.
+The conclusion of a rule either describes equalities to be added to $\Ec$
+or is $\bot$.
+% (in which case we call an application of it \emph{nonterminal})
+% (in which case we call an application of it \emph{terminal}).
+A rule may have multiple conclusions separated by $\parallel$,
+denoting nondeterministic branching.
 
@@
 \inferrule{
- t_1 \teq t_2 \in \Ec
+ t \teq u \in \Ec
 }{
- \Ec := \Ec,\: t_2 \teq t_1
+ \Ec := \Ec,\: u \teq t
 }
@@
 \inferrule{
-  t_1 \teq t_2,\; t_1 \tneq t_2 \in \Ec
+  t \teq u,\; t \tneq u \in \Ec
 }{
@@
 \inferrule{
-  t_1 \teq t_2,\; t_2 \teq t_3 \in \Ec
+  s \teq t,\; t \teq u \in \Ec
 }{
-  \Ec := \Ec,\: t_1 \teq t_3
+  \Ec := \Ec,\: s \teq u
 }
@@
 \end{tabular}
-\caption{\,Rules for bidirectional closure%.
+\caption{\,Derivation rules for bidirectional closure%.
 }
@@
 Figure~\ref{fig:cc-rules} gives the basic rules of the calculus.
-Together with \rn{Refl}, \rn{Sym}, and \rn{Trans}, the \rn{Cong} rule computes the (upward) congruence closure,
+In conjunction with \rn{Refl}, \rn{Sym}, and \rn{Trans}, the \rn{Cong} rule computes the (upward) congruence closure,
 whereas the \rn{Decomp} and \rn{Clash} rules together compute (downward) unification.
-In the unification rules, additional equalities are inferred based on the injectivity of constructors by \rn{Decomp},
+For unification, additional equalities are inferred based on the injectivity of constructors by \rn{Decomp},
 and failures to unify equated terms are recognized by \rn{Clash}.
@@
   \quad
-  t_1 : \tau
+  t : \tau
   \quad
   %\ttpath{\Ec}{\ec{t}}{\ec{t}} \neq \emptyset
-  \Val \ec{t_1} = \mu x.\; t_2
+  \Val \ec{t} = \mu x.\; u
   \quad
-  x \in FV( t_2 )
+  x \in FV( u )
 }{
@@
  \quad
- t_1, t_2 : \tau
+ t, u : \tau
  \quad
- %\tpath{\Ec}{\ec{t_1}} = \tpath{\Ec}{\ec{t_2}}\sigma \neq \emptyset
- \Val \ec{t_1} \vsim \Val \ec{t_2}
+ %\tpath{\Ec}{\ec{t}} = \tpath{\Ec}{\ec{u}}\sigma \neq \emptyset
+ \Val \ec{t} \vsim \Val \ec{u}
 }{
- \Ec := \Ec,\: t_1 \teq t_2
+ \Ec := \Ec,\: t \teq u
 }
@@
 \end{tabular}
-\caption{\,Rules for acyclicity and bisimilarity%.
+\caption{\,Directional rules for acyclicity and bisimilarity%.
 }
@@
 no rule can be applied to extend~$\Ec$ further.
-When this is the case, it is easy to see that $\Ec$ induces an equivalence relation over $\tEc$ such that two terms $t_1$ and $t_2$ are equivalent if and only if $t_1 \teq t_2 \in \Ec$.
+When this is the case, it is easy to see that $\Ec$ induces an equivalence relation over $\tEc$ such that two terms $t$ and $u$ are equivalent if and only if $t \teq u \in \Ec$.
 Thus, we can consider $\Ec$ as a set of equivalence classes of terms. 
@@
 Given $\Val$, the rules for acyclicity and bisimilarity (Figure~\ref{fig:ab-rules}) can be stated as follows.
-For acyclicity, if there is an equivalence class $\ec{t_1}$ of datatype type whose class of values $\Val \ec{t_1} = \mu x. t_2$ are cyclic,
-as indicated by the assumption $x \in FV( t_2 )$,
+For acyclicity, if there is an equivalence class $\ec{t}$ of datatype type whose class of values $\Val \ec{t} = \mu x.\; u$ are cyclic,
+as indicated by the assumption $x \in FV( u )$,
 then we can conclude that $\Ec$ is unsatisfiable.
-For bisimilarity, if there are two equivalence classes $\ec{t_1}$ and $\ec{t_2}$ 
-of codatatype type whose classes of values $\Val \ec{t_1}$ and $\Val \ec{t_2}$ are equivalent for a capture-avoiding renaming of $\mu$-bound variables,
-then we can conclude $t_1 \teq t_2$.
+For bisimilarity, if there are two equivalence classes $\ec{t}$ and $\ec{u}$ 
+of codatatype type whose classes of values $\Val \ec{t}$ and $\Val \ec{u}$ are equivalent for a capture-avoiding renaming of $\mu$-bound variables,
+then we can conclude $t \teq u$.
 We demonstrate the bisimilarity rule concretely with the following example.
@@
 \end{tabular}
-\caption{\,Rules for branching%.  
+\caption{\,Derivation rules for branching%.  
 %All pairs of terms whose type has cardinality 1 are entailed to be equal (the degenerate case).
@@
 
-\
-
-An application of a rule is \emph{redundant} if it is nonterminal and at least one branch in its conclusion does not add a new equality to $\Ec$.
+An application of a rule is \emph{redundant} if at least one branch in the
+conclusion does not add a new equality to $\Ec$ and is not $\bot$.
 A \emph{derivation tree} is a tree whose nodes are sets of equalities, where non-root nodes vtained by 

### 2014-12-19 tex/codatadp.tex
Minor simplification

--- 2014-12-19 tex/codatadp.tex
@@
 % change?
-\newcommand{\thO}{T_{\mathrm{o}}}
-\newcommand{\thD}{T_{\mathrm{cd}}}
+\newcommand{\thO}{T_{o}}
+\newcommand{\thD}{T_{d}}
 
@@
 processes. We present a decision procedure that combines reasoning about
-datatypes and codatatypes. The dual of the acyclicity rule for datatypes is a
+datatypes and codatatypes. The dual of the cyclicity rule for datatypes is a
 bisimilarity rule that identifies equal values, also in the presence of cyclic
@@
 (Section~\ref{sec:the-theory-of-co-datatypes}). Intuitively, the
-values of a codatatype consist of all well-typed finite and infinite ground
-constructor
+values of a codatatype consist of all well-typed finite and infinite ground constructor
 terms, and only those. For example, the coinductive specification
@@
 
-Second, \emph{a bisimilarity rule takes the place of the acyclicity rule of datatypes.}
+Second, \emph{a bisimilarity rule takes the place of the cyclicity rule of datatypes.}
 Cyclic constraints such as
@@
 $\const{y} \teq \const{S}(\const{x})$, it deduces $\const{x} \teq \const{y}$.
-The acyclicity and bisimilarity rules are needed to ensure completeness on
+The cyclicity and bisimilarity rules are needed to ensure completeness on
 ground problems and the absence of spurious models. They cannot be replaced by
@@
 
-\section{%The Theory of
-(Co)datatypes}
+\section{The Theory of (Co)datatypes}
 \label{sec:the-theory-of-co-datatypes}
@@
 %\paragraph{Characterization.}
-One way to define datatypes semantically is as the initial model of the
-selector--constructor equations \cite{barrett-et-al-2010}.
+Datatypes can be defined semantically as the initial model of the
+selector--constructor equations. This is the view followed by Barrett et al.\
+\cite{barrett-et-al-2010}.
 \begin{report}
@@
 In addition, codatatypes are guaranteed to contain all values corresponding to
-infinite ground constructor terms.
+infinite ground constructor terms. In general, this cannot be captured by a
+first-order axiomatization, since there may be uncountably many of them.
 \begin{report}
-In general, this cannot be captured by a
-first-order axiomatization, since there may be uncountably many of them.
 For example, $\ty{stream}_{\,\ty{int}}$ is isomorphic to the uncountable
@@
 
-The decision procedure for the ground theory of (co)datatypes $\thD$ determines the
-$\thD$-satisfiability of sets of ground constraints over a fixed
-signature~$\Sig$. It is presented as a calculus, whose rules can be applied in
-any order until saturation.
-
-The calculus consists of derivation rules that operate on a set $\Ec$ of
-equalities and disequalities between $\Sig$-terms, where $\Ec$ is interpreted
-as a conjunction.
+This section presents a decision procedure for \rem{introduce ``$\thD$" in previous section?}
+determining the $\thD$-satisfiability of sets of constraints built from a fixed signature $\Sig$.
+We will make the following assumptions about $\Sig$.
+First, all codatatypes in $\Codata$ are corecursive.
+This is not restrictive, since checking the $\thD$-satisfiability for constraints involving a non-corecursive co-datatype $\tau$ can be trivially reduced to
+checking the satisfiability of the same constraint involving a (non-recursive) datatype of the same form as $\tau$. \rem{revise?}
+Second, all types $\tau \in \Nondata$ have infinite cardinality.
+This also is not restrictive, since each interpreted type $\tau \in \Nondata$ having finite cardinality $n$
+can be replaced by a enumeration datatype with $n$ constructors
+\footnote{Doing so simplifies the presentation of the calculus, but is not done in practice. \rem{More?}}.
+Additionally, since $\Ec$ is ground, we know that $\Ec$ cannot entail an upper bound on the cardinality of any uninterpreted type $\tau \in \Nondata$, 
+and thus without loss of completeness we may treat all other types as infinite.
+
+Our calculus consists of derivation rules that operate on a set $\Ec$ of equalities and disequalities between $\Sig$-terms,
+where $\Ec$ is interpreted as a conjunction.
 We write $\tEc$ to denote the set of $\Sig$-terms occurring in $\Ec$.
@@
 We assume that all $\Sig$-terms in $\Ec$ are \emph{normalized}, meaning that all subterms of the form $\const{s}^{\,k}_j( \const{C}_j(t_1,\ldots,t_n) )$
-are simplified to $t_k$. All constraints added to $\Ec$ are also implicitly
-normalized in this manner.
-
-To simplify the presentation, we make the following assumptions about $\Sig$.
-First, all codatatypes in $\Codata$ are corecursive. This is not restrictive
-because noncorecursive codatatypes can be considered as nonrecursive
-datatypes.
-Second, all types $\tau \in \Nondata$ have infinite cardinality.
-This is not overly restrictive, because each interpreted type $\tau \in \Nondata$
-having finite cardinality~$n$
-can be replaced by an enumeration datatype with $n$~nullary constructors.
-Moreover, since the constraints are ground, they cannot entail an upper bound
-on the cardinality of any uninterpreted type $\tau \in \Nondata$; thus,
-we may without loss of completeness treat all other types as infinite.
-
-The calculus consists of three sets of rules corresponding to three
-successive phases.
-The first phase computes the bidirectional closure of $\Ec$.
-The second phase makes inferences based on acyclicity (for datatypes)
-and bisimilarity (for codatatypes).
-The third phase performs case splitting on constructors
-for various terms in $\tEc$.
+are simplified to $t_i$, and moreover assume that all %additional
+constraints added to $\Ec$ are normalized in this manner.
+
+We present the calculus in three steps.
+In the first step, we compute the bidirectional closure of $\Ec$;
+in the second step, we make inferences based on cyclicity and bisimilarity;
+and in the third step, %(when necessary),
+we branch on constructor types for various terms in $\tEc$.
 Following the conventions from~\cite{}, 
@@
 \)
-\rn{Acyclic}
+\rn{Cyclic}
 \\[4\jot]
@@
 \end{tabular}
-\caption{\,Rules for acyclicity and bisimilarity%.
+\caption{\,Rules for cyclicity and bisimilarity%.
 }
@@
 
-\paragraph{Step 2: Applying Acyclicity and Bisimilarity.}
+\paragraph{Step 2: Applying Cyclicity and Bisimilarity.}
 In the second step,
@@
 %, or informally are \emph{observationally equivalent}.
-%For convienience, we use $\mu$-terms to refer to (classes of) values for both codatatype and datatype terms,
+%For convienience, we use $\mu$-terms to refer to (classes of) values for both co-datatype and datatype terms,
 %where in the latter case, a $\mu$-term with a bound variable denotes an infeasible term.
@@
 Intuitively, a $\mu$-term $t$ of type $\tau$ describes a class of values of $\tau$, 
-%either in the case when $\tau$ is a codatatype or datatype type,
+%either in the case when $\tau$ is a co-datatype or datatype type,
 where a cyclic $t$ describes an infeasible class of values when
@@
 
-Given $\Val$, the rules for acyclicity and bisimilarity (Figure~\ref{fig:ab-rules}) can be stated as follows.
-For acyclicity, if there is an equivalence class $\ec{t_1}$ of datatype type whose class of values $\Val \ec{t_1} = \mu x. t_2$ are cyclic,
+Given $\Val$, the rules for cyclicity and bisimilarity (shown in Figure~\ref{fig:ab-rules}) can be stated as follows.
+For cyclicity, if there is an equivalence class $\ec{t_1}$ of datatype type whose class of values $\Val \ec{t_1} = \mu x. t_2$ are cyclic,
 as indicated by the assumption $x \in FV( t_2 )$,
@@
 For bisimilarity, if there are two equivalence classes $\ec{t_1}$ and $\ec{t_2}$ 
-of codatatype type whose classes of values $\Val \ec{t_1}$ and $\Val \ec{t_2}$ are equivalent for a capture-avoiding renaming of $\mu$-bound variables,
+of co-datatype type whose classes of values $\Val \ec{t_1}$ and $\Val \ec{t_2}$ are equivalent for a capture-avoiding renaming of $\mu$-bound variables,
 then we can conclude $t_1 \teq t_2$.
@@
 We prove this by structural induction on the (closed) derivation tree with root node $\Ec$.
-First, if the derivation tree is an application of \rn{Conflict}, \rn{Clash}, or \rn{Acyclic},
+First, if the derivation tree is an application of \rn{Conflict}, \rn{Clash}, or \rn{Cyclic},
 then $\Ec$ is unsatisfiable in $\thD$.
@@
 For \rn{Clash}, this is a consequence of distinctness.
-For \rn{Acyclic}, our construction of $\Val$ indicates that the class of values that $t_1$ can take in models of $\Ec$ is infeasible,
+For \rn{Cyclic}, our construction of $\Val$ indicates that the class of values that $t_1$ can take in models of $\Ec$ is infeasible,
 and thus $\Ec$ is unsatisfiable.
@@
 \{
-\mu x. \const{A}( \mu y. \const{B}( x )),
-\mu x. \const{A}( \mu y. \const{B}( y )),
+\mu x. \const{A}( \const{B}( x )),
+\const{A}( \mu y. \const{B}( y )),
 \mu x. \const{B}( x ),
-\mu x. \const{C}( \mu y. \const{A}( x )),
-\mu x. \const{A}( \mu y. \const{C}( x )),
+\mu x. \const{C}( \const{A}( x )),
+\mu x. \const{A}( \const{C}( x )),
 \mu x. \const{A}( x )
@@
 $\mu v_{\ec{x}}. \const{C}( \mu x_1. \const{A}( x_1) )$, and
-$\mu v_{\ec{x}}. \const{B}( \mu x_1. \const{A}( v_{\ec{x}} ) )$
+$\mu v_{\ec{x}}. \const{B}( \const{A}( v_{\ec{x}} ) )$
 to $v_{\ec{x}}$.
@@
 as in after applying the third substitution mentioned above,
-$\Val \ec{w_1}$ is $\mu v_{\ec{w_1}}. \const{A}( \mu v_{\ec{x}}. \const{B}( \mu x_1. \const{A}( v_{\ec{x}} ) ) )$.
+$\Val \ec{w_1}$ is $\mu v_{\ec{w_1}}. \const{A}( \mu v_{\ec{x}}. \const{B}( \const{A}( v_{\ec{x}} ) ) )$.
 $\Box$
@@
 By construction, $\Val^\ast \ec{x}$ is normal.
-Since \rn{Acyclic} does not apply, $\Val \ec{x}$ is acyclic when $\tau \in \Data$.
+Since \rn{Cyclic} does not apply, $\Val \ec{x}$ is acyclic when $\tau \in \Data$.
 Moreover, our construction of $\Val^\ast$ applies substitutions of the form 
@@
   & 221 & 760 & \phantom{0}57 & 1038 \\
-Datatypes without acyclicity
+Datatypes without cyclicity
   & 227 & 765 & \phantom{0}57 & 1049 \\
-Datatypes with acyclicity
+Datatypes with cyclicity
   & 227 & 771 & \phantom{0}57 & 1055 \\
@@
 CVC4 was run on each problem several times, with the support for datatypes and
-codatatypes either enabled or disabled. The contributions of the acyclicity and
+codatatypes either enabled or disabled. The contributions of the cyclicity and
 bisimilarity rules were also measured, so find out whether these somewhat

### 2014-12-19 tex/codatadp.tex
more

--- 2014-12-19 tex/codatadp.tex
@@
 \newcommand\dotReportFootnote[1]{.}
+
+\newcommand\FV{\mathrm{FV}}
 
@@
 
-The calculus consists of three sets of rules corresponding to three
-successive phases.
-The first phase computes the bidirectional closure of $\Ec$.
-The second phase makes inferences based on acyclicity (for datatypes)
-and bisimilarity (for codatatypes).
-The third phase performs case splitting on constructors
-for various terms in $\tEc$.
+The calculus consists of three sets of rules (Figures \ref{fig:cc-rules} to
+\ref{fig:split-rule}) corresponding to three successive phases. The first
+phase computes the bidirectional closure of $\Ec$. The second phase makes
+inferences based on acyclicity (for datatypes) and bisimilarity (for
+codatatypes). The third phase performs case splitting on constructors for
+various terms in $\tEc$.
 %Following the conventions from~\cite{}, 
@@
 A rule may have multiple conclusions separated by $\parallel$,
-denoting nondeterministic branching.
+denoting nondeterministic branching. The rules associated with a phase are
+applied until saturation (i.e., until no rule can be applied to extend~$\Ec$
+further) before the next phase can begin.
 
@@
 \)
-\rn{Decomp} 
+\rn{Inject} 
 \qquad
@@
 \paragraph{Phase 1: Computing the Bidirectional Closure.}
-Figure~\ref{fig:cc-rules} gives the basic rules of the calculus.
 In conjunction with \rn{Refl}, \rn{Sym}, and \rn{Trans}, the \rn{Cong} rule computes the (upward) congruence closure,
-whereas the \rn{Decomp} and \rn{Clash} rules together compute (downward) unification.
-For unification, additional equalities are inferred based on the injectivity of constructors by \rn{Decomp},
+whereas the \rn{Inject} and \rn{Clash} rules together compute (downward) unification.
+For unification, additional equalities are inferred based on the injectivity of constructors by \rn{Inject},
 and failures to unify equated terms are recognized by \rn{Clash}.
 The rule \rn{Conflict} recognizes when an equality and its negation both occur in $\Ec$, in which case $\Ec$ has no model.
+
+At the end of this phase, $\Ec$ induces an equivalence
+relation over $\tEc$ such that two terms $t$ and $u$ are equivalent if and
+only if $t \teq u \in \Ec$.
+Thus, we can regard $\Ec$ as a set of
+equivalence classes of terms. For a term $t \in \tEc$, we write $\ec{t}$ to
+denote the equivalence class in $\Ec$ that contains $t$.
 
@@
   \quad
-  x \in FV( u )
+  x \in \FV( u )
 }{
@@
  %\tpath{\Ec}{\ec{t}} = \tpath{\Ec}{\ec{u}}\sigma \neq \emptyset
- \Val \ec{t} \vsim \Val \ec{u}
+ \Val \ec{t} \aequiv \Val \ec{u}
 }{
@@
 \end{tabular}
-\caption{\,Directional rules for acyclicity and bisimilarity%.
+\caption{\,Derivation rules for acyclicity and bisimilarity%.
 }
@@
 \paragraph{Phase 2: Applying Acyclicity and Bisimilarity.}
-In the second phase,
-we assume that $\Ec$ is saturated with respect to the rules in Figure~\ref{fig:cc-rules}---that is,
-no rule can be applied to extend~$\Ec$ further.
-When this is the case, it is easy to see that $\Ec$ induces an equivalence relation over $\tEc$ such that two terms $t$ and $u$ are equivalent if and only if $t \teq u \in \Ec$.
-Thus, we can consider $\Ec$ as a set of equivalence classes of terms. 
-For a term $t \in \tEc$, we write $\ec{t}$ to denote the equivalence class in $\Ec$ that contains $t$.
-
 %For presentation of the rules in this phase, we rely on a representation of terms in the $\mu$-notation.
-The rules of this phase contain premises that involve a mapping $\Val$ from equivalence classes to terms containing $\mu$-bindings.
+The rules of the second phase contain premises that involve a mapping $\Val$
+that associates a potentially cyclic $\mu$-term as representative with each equivalence class.
 Recall that $\mu$-bindings can be used for representing possibly cyclic terms and values.
@@
 $\mu$-terms $\vec t$ with the expected types.
-The variable $x$ need not occur under the binder's body, in which case the $\mu$-prefix
-is optional.
+The variable $x$ need not occur under the binder's body, in which case the
+binder can be omitted.
 
@@
 \not\aequiv \mu y.\; \const{D}( y, x )$.
-Two $\mu$-terms may represent the same value while not being $\alpha$-equivalent, as in
-$\mu x.\; \const{C}( \const{C}( x ) ) \not\aequiv \mu x.\; \const{C}( x )$.
+Two $\mu$-terms can represent the same value while not being $\alpha$-equivalent---e.g.,
+$\mu x.\; \const{C}( x ) \not\aequiv \mu x.\; \const{C}( \const{C}( x ) )$.
 %, or informally are \emph{observationally equivalent}.
@@
 \inferrule{
-  v_{\ec{x}} \in FV( \Val )
+  v_{\ec{x}} \in \FV( \Val )
   \quad
@@
 \]
-Here, we write $FV( \Val )$ to denote the \emph{free variables} occurring in $\Val$, which
-excludes variables that occur beneath $\mu$-binders.
+We write $\FV( \Val )$ to denote the \relax{free variables} occurring in $\Val$, which
+exclude variables that occur beneath $\mu$-binders.
 It is straightforward to see that the height of terms produced as a result of this update
@@
 For acyclicity, if there is an equivalence class $\ec{t}$ of datatype type whose class of values $\Val \ec{t} = \mu x.\; u$ are cyclic,
-as indicated by the assumption $x \in FV( u )$,
+as indicated by the assumption $x \in \FV( u )$,
 then we can conclude that $\Ec$ is unsatisfiable.
@@
 Otherwise, the child of $\Ec$ is a closed derivation tree 
-whose root node is $\Ec \cup t_1 \teq t_2$ obtained by applying one of the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, \rn{Decomp}, \rn{Bisimilar}, or \rn{Singleton}.
+whose root node is $\Ec \cup t_1 \teq t_2$ obtained by applying one of the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, \rn{Inject}, \rn{Bisimilar}, or \rn{Singleton}.
 In all cases, we have that $\Ec \models_{\thD} t_1 \teq t_2$.
 For \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, this is a consequence of equality reasoning.
-For \rn{Decomp}, this is a consequence of injectivity.
+For \rn{Inject}, this is a consequence of injectivity.
 For \rn{Bisimilar}, our construction of $\Val$ indicates that the value of $t_1$ and $t_2$ are equivalent in all models of $\Ec$.
@@
 \inferrule{
-  x : \tau \in FV( \Val ) 
+  x : \tau \in \FV( \Val ) 
   \quad
@@
 and assign it a value that is not bisimilar to one occurring in $\mathcal{T}^\M_x(\Val)$.
-Since this update removes one variable from the set $FV( \Val )$ and does not add any additional variables to $FV( \Val )$, 
+Since this update removes one variable from the set $\FV( \Val )$ and does not add any additional variables to $\FV( \Val )$, 
 it can only be applied a finite number of times.
@@
 are such that $i = j$ since \rn{Clash} does not apply to $\Ec$,
-and are such that $\ec{ \vec t } = \ec{ \vec u }$ since \rn{Decomp} does not apply to $\Ec$.
+and are such that $\ec{ \vec t } = \ec{ \vec u }$ since \rn{Inject} does not apply to $\Ec$.
 Thus, 
@@
 %Say that $\muvar( w_i ) = v_{\ec{y_i}}$.
-%Due to our construction of $\Val^\ast$ and since \rn{Decomp} does not apply to $\Fc$, we have $\ec{u_i} = \ec{y_i}$,
+%Due to our construction of $\Val^\ast$ and since \rn{Inject} does not apply to $\Fc$, we have $\ec{u_i} = \ec{y_i}$,
 %and thus $t_u\interp{u_i'} \vsim \Val^\ast \ec{y_i}$.
@@
 where $t_w \interp{ w_j } = \Val^\ast \ec{ s^j_k( u )}$. 
-%By our construction of $\Val^\ast$ and since \rn{Decomp} and \rn{Clash} do not apply to $\Fc$, 
+%By our construction of $\Val^\ast$ and since \rn{Inject} and \rn{Clash} do not apply to $\Fc$, 
 %$\Val^\ast \ec{ u }$ must be of the form $t_u = \mu v_{\ec{u}}. C_i( u_1, \ldots, u_n )$,
@@
 if $s^j_i( t )$ is a term in $\tEc$ and $t$ contains the constructor 
-$\const{C}_i( \vec u )$, then we immediately infer $s^j_i( t ) \teq u_j$, whereas the calculus we presented would apply \rn{Split} and \rn{Decomp} before inferring this equality.  
+$\const{C}_i( \vec u )$, then we immediately infer $s^j_i( t ) \teq u_j$, whereas the calculus we presented would apply \rn{Split} and \rn{Inject} before inferring this equality.  
 For the sake of reducing the number of unique constraints considered by the calculus, we compute a normal form for constraints as a preprocessing step.

### 2014-12-19 tex/codatadp.tex
moved around

@@
 The rules of the second phase contain premises that involve a mapping $\Val$
-that associates a potentially cyclic $\mu$-term as representative with each equivalence class.
-Recall that $\mu$-bindings can be used for representing possibly cyclic terms and values.
+that associates a $\mu$-term as representative with each equivalence class.
+%Recall that $\mu$-bindings can be used for representing possibly cyclic terms and values.
 %(Section~\ref{sec:introduction}).
 % for example, $\mu x. \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
+Intuitively, a $\mu$-term $t$ of type $\tau$ describes a class of values of $\tau$,
+%either in the case when $\tau$ is a codatatype or datatype type,
+where a cyclic $t$ describes an infeasible class of values when
+%$\tau \not\in \Codata$.
+$\tau \in \Data$.
 Formally, $\mu$-terms are defined recursively as being either a variable $x$
@@
 
-Intuitively, a $\mu$-term $t$ of type $\tau$ describes a class of values of $\tau$,
-%either in the case when $\tau$ is a codatatype or datatype type,
-where a cyclic $t$ describes an infeasible class of values when
-%$\tau \not\in \Codata$.
-$\tau \in \Data$.
 At the beginning of this phase, 

### 2014-12-19 tex/codatadp.tex
introduces \Var notation and made it consistent with \Val (although I'm open-minded here)

@@
 \newcommand{\Val}{\mathcal{A}}
+\newcommand{\Var}{\mathcal{V}}
 \newcommand{\nf}[1]{{#1}\!\!\downarrow}
@@
 % for example, $\mu x. \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
-Intuitively, a $\mu$-term $t$ of type $\tau$ describes a class of values of $\tau$,
+A $\mu$-term $t$ of type $\tau$ describes a class of values of
+$\tau$ that terms in that equivalence class can take in the models of $\Ec$;
 %either in the case when $\tau$ is a codatatype or datatype type,
-where a cyclic $t$ describes an infeasible class of values when
+when $\tau \in \Data$, a cyclic $t$ describes an infeasible class of values.
 %$\tau \not\in \Codata$.
-$\tau \in \Data$.
 Formally, $\mu$-terms are defined recursively as being either a variable $x$
@@
 $\mu$-terms $\vec t$ with the expected types.
-The variable $x$ need not occur under the binder's body, in which case the
+The variable $x$ need not occur freely under the binder's body, in which case the
 binder can be omitted.
+In the \rn{Acyclic} rule,
+$\FV( u )$ denotes the variables occurring freely in $u$.
 
@@
 
-At the beginning of this phase, 
-we compute a mapping $\Val$ from equivalence classes to $\mu$-terms 
-that describe the class of values that terms in that equivalence class can take in the models of $\Ec$. 
-For each equivalence class $\ec{x}$, we associate a fresh variable $v_{\ec{x}}$ of the same type as $x$ not occurring in $\tEc$,
-and initially say that $\Val$ contains $\ec{x} \mapsto v_{\ec{x}}$,
-in other words, there are initially no constraints on the values for any equivalence class $\ec{x}$.
-Then, $\Val$ is modified by applying the following update exhaustively:
+The mapping $\Val$ is computed as follows.
+%
+For each equivalence class $\ec{x}$, we associate a fresh variable $\Var{\ec{x}}$ of the same type as $x$
+%not occurring in $\tEc$,
+and let $\Val\ec{x} = \Var{\ec{x}}$.
+Thus, there are initially no constraints on the values for any equivalence class $\ec{x}$.
+The mapping $\Val$ is then modified by applying the following rule exhaustively:
 \[
+\hbox{\(
 \inferrule{
-  v_{\ec{x}} \in \FV( \Val )
+  \Var{\ec{x}} \in \FV( \Val )
   \quad
@@
 }{
-  \Val := \Val \{ v_{\ec{x}} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{t_1}}, \ldots, v_{\ec{t_n}} ) \}
+  \Val := \Val \{ \Var{\ec{x}} \mapsto \mu \Var{\ec{x}}.\; \const{C}( \Var{\ec{t_1}}, \ldots, \Var{\ec{t_n}} ) \}
 }
+\)
+\rn{Update}}
 \]
-We write $\FV( \Val )$ to denote the \relax{free variables} occurring in $\Val$, which
-exclude variables that occur beneath $\mu$-binders.
+In the rule, $\FV( \Val )$ denotes the union of the free variables occurring in $\Val$'s range.
 It is straightforward to see that the height of terms produced as a result of this update
@@
 \hline
-& \{ & \ec{x} \mapsto v_{\ec{x}}, & \ec{z} \mapsto v_{\ec{z}}, & \\
-&    & \ec{y} \mapsto v_{\ec{y}}, & \ec{w} \mapsto v_{\ec{w}} & \} \\
+& \{ & \ec{x} \mapsto \Var{\ec{x}}, & \ec{z} \mapsto \Var{\ec{z}}, & \\
+&    & \ec{y} \mapsto \Var{\ec{y}}, & \ec{w} \mapsto \Var{\ec{w}} & \} \\
 \ec{x}
-& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ), & \ec{z} \mapsto v_{\ec{z}}, & \\
-&    & \ec{y} \mapsto v_{\ec{y}}, & \ec{w} \mapsto v_{\ec{w}} & \} \\
+& \{ & \ec{x} \mapsto \mu \Var{\ec{x}}. \const{C}( \Var{\ec{w}}, \Var{\ec{y}} ), & \ec{z} \mapsto \Var{\ec{z}}, & \\
+&    & \ec{y} \mapsto \Var{\ec{y}}, & \ec{w} \mapsto \Var{\ec{w}} & \} \\
 \ec{y}
-& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ) ), & \ec{z} \mapsto v_{\ec{z}}, & \\
-&    & \ec{y} \mapsto \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
+& \{ & \ec{x} \mapsto \mu \Var{\ec{x}}. \const{C}( \Var{\ec{w}}, \mu \Var{\ec{y}}. \const{C}( \Var{\ec{z}}, \Var{\ec{x}} ) ), & \ec{z} \mapsto \Var{\ec{z}}, & \\
+&    & \ec{y} \mapsto \mu \Var{\ec{y}}. \const{C}( \Var{\ec{z}}, \Var{\ec{x}} ), & \ec{w} \mapsto \Var{\ec{w}} & \} \\
 \ec{x}
-& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ) ), & \ec{z} \mapsto v_{\ec{z}}, & \\
-&    & \ec{y} \mapsto \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ) ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
+& \{ & \ec{x} \mapsto \mu \Var{\ec{x}}. \const{C}( \Var{\ec{w}}, \mu \Var{\ec{y}}. \const{C}( \Var{\ec{z}}, \Var{\ec{x}} ) ), & \ec{z} \mapsto \Var{\ec{z}}, & \\
+&    & \ec{y} \mapsto \mu \Var{\ec{y}}. \const{C}( \Var{\ec{z}}, \mu \Var{\ec{x}}. \const{C}( \Var{\ec{w}}, \Var{\ec{y}} ) ), & \ec{w} \mapsto \Var{\ec{w}} & \} \\
 \hline
@@
 This indicates that the values for $x$ (resp. $y$) in models of $\Ec$
-are of the form $\const{C}( v_{\ec{w}}, \const{C}( v_{\ec{z}}, \ldots ))$ (resp. $\const{C}( v_{\ec{z}}, \const{C}( v_{\ec{w}}, \ldots ))$) 
-for some value of $v_{\ec{w}}$ and $v_{\ec{z}}$.
+are of the form $\const{C}( \Var{\ec{w}}, \const{C}( \Var{\ec{z}}, \ldots ))$ (resp. $\const{C}( \Var{\ec{z}}, \const{C}( \Var{\ec{w}}, \ldots ))$) 
+for some value of $\Var{\ec{w}}$ and $\Var{\ec{z}}$.
 \qed
@@
 \Val \ec{x} & = & 
-\mu v_{\ec{x}}. & \const{C}( v_{\ec{y}}, & \mu v_{\ec{\const{C}(y,x)}}. & \const{C}( v_{\ec{y}}, v_{\ec{x}} ) ) \\
+\mu \Var{\ec{x}}. & \const{C}( \Var{\ec{y}}, & \mu \Var{\ec{\const{C}(y,x)}}. & \const{C}( \Var{\ec{y}}, \Var{\ec{x}} ) ) \\
 \Val \ec{\const{C}(y,x)} & = & 
-\mu v_{\ec{\const{C}(y,x)}}. & \const{C}( v_{\ec{y}}, & \mu v_{\ec{x}}. & \const{C}( v_{\ec{y}}, v_{\ec{\const{C}(y,x)}} ) )
+\mu \Var{\ec{\const{C}(y,x)}}. & \const{C}( \Var{\ec{y}}, & \mu \Var{\ec{x}}. & \const{C}( \Var{\ec{y}}, \Var{\ec{\const{C}(y,x)}} ) )
 \end{array}\]
 Notice that $\Val \ec{x}$ and $\Val \ec{\const{C}(y,x)}$ are syntactically equivalent for the renaming
-$\{ v_{\ec{x}} \mapsto v_{\ec{\const{C}(y,x)}}, v_{\ec{\const{C}(y,x)}} \mapsto v_{\ec{x}} \}$,
+$\{ \Var{\ec{x}} \mapsto \Var{\ec{\const{C}(y,x)}}, \Var{\ec{\const{C}(y,x)}} \mapsto \Var{\ec{x}} \}$,
 and thus our calculus may infer $x \teq \const{C}(y,x)$.
 Intuitively, this equality holds, since the values of $x$ and $\const{C}( y, x )$ 
-are equivalent for any value of $v_{\ec{y}}$.
+are equivalent for any value of $\Var{\ec{y}}$.
 \qed
@@
 \[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{2.0em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
-\Val \ec{y_1} & = & \mu v_{\ec{y_1}}. & \const{A}( & \mu v_{\ec{y_2}}. & \const{B}( v_{\ec{x}} ) ) &
-\Val \ec{w_1} & = & \mu v_{\ec{w_1}}. & \const{A}( v_{\ec{x}} )      & \\ 
-\Val \ec{y_2} & = & \mu v_{\ec{y_2}}. & \multicolumn{2}{l}{\const{B}( v_{\ec{x}} )}      & &
-\Val \ec{w_2} & = & \mu v_{\ec{w_2}}. & \const{C}( v_{\ec{w_3}} ) \\
-\Val \ec{z_1} & = & \mu v_{\ec{z_1}}. & \const{C}( & \mu v_{\ec{z_2}}. & \const{A}( v_{\ec{z_1}} ) ) &
-\Val \ec{w_3} & = & v_{\ec{w_3}} \\
-\Val \ec{z_2} & = & \mu v_{\ec{z_2}}. & \const{A}( & \mu v_{\ec{z_1}}. & \const{C}( v_{\ec{z_2}} ) ) & 
-\Val \ec{x} & = & v_{\ec{x}} \\
+\Val \ec{y_1} & = & \mu \Var{\ec{y_1}}. & \const{A}( & \mu \Var{\ec{y_2}}. & \const{B}( \Var{\ec{x}} ) ) &
+\Val \ec{w_1} & = & \mu \Var{\ec{w_1}}. & \const{A}( \Var{\ec{x}} )      & \\ 
+\Val \ec{y_2} & = & \mu \Var{\ec{y_2}}. & \multicolumn{2}{l}{\const{B}( \Var{\ec{x}} )}      & &
+\Val \ec{w_2} & = & \mu \Var{\ec{w_2}}. & \const{C}( \Var{\ec{w_3}} ) \\
+\Val \ec{z_1} & = & \mu \Var{\ec{z_1}}. & \const{C}( & \mu \Var{\ec{z_2}}. & \const{A}( \Var{\ec{z_1}} ) ) &
+\Val \ec{w_3} & = & \Var{\ec{w_3}} \\
+\Val \ec{z_2} & = & \mu \Var{\ec{z_2}}. & \const{A}( & \mu \Var{\ec{z_1}}. & \const{C}( \Var{\ec{z_2}} ) ) & 
+\Val \ec{x} & = & \Var{\ec{x}} \\
 \end{array}\]
 
-To construct a completion of $\Val$, we must choose values for the free variables of $\Val$, $v_{\ec{x}}$ and $v_{\ec{w_3}}$.
-The content of the set $\mathcal{T}^\M_{v_{\ec{x}}}( \Val )$ is $\alpha$-equivalent to:
+To construct a completion of $\Val$, we must choose values for the free variables of $\Val$, $\Var{\ec{x}}$ and $\Var{\ec{w_3}}$.
+The content of the set $\mathcal{T}^\M_{\Var{\ec{x}}}( \Val )$ is $\alpha$-equivalent to:
 \[%\begin{array}{c}
@@
 for instance, $x$ is interpretted as $\Val \ec{x}$, $y_1$ as $\Val \ec{y_1}$ and so on.
-Assigning a value for $v_{\ec{x}}$ that is $\alpha$-equivalent to a value in $\mathcal{T}^\M_{v_{\ec{x}}}( \Val )$
+Assigning a value for $\Var{\ec{x}}$ that is $\alpha$-equivalent to a value in $\mathcal{T}^\M_{\Var{\ec{x}}}( \Val )$
 may cause values in the range of $\Val$ to become $\alpha$-equivalent, 
 which in turn may cause $\M$ to be inconsistent.
-For instance, say we assign $\mu v_{\ec{x}}. \const{B}( v_{\ec{x}} )$ for $v_{\ec{x}}$.
-After this substitution, $\Val \ec{y_2}$ is $\mu v_{\ec{y_2}}. \const{B}( \mu v_{\ec{x}}. \const{B}( v_{\ec{x}} ) )$, 
-which has normal form $\mu v_{\ec{y_2}}. \const{B}( v_{\ec{y_2}} )$,
-which is $\alpha$-equivalent to $\mu v_{\ec{x}}. \const{B}( v_{\ec{x}} )$.
+For instance, say we assign $\mu \Var{\ec{x}}. \const{B}( \Var{\ec{x}} )$ for $\Var{\ec{x}}$.
+After this substitution, $\Val \ec{y_2}$ is $\mu \Var{\ec{y_2}}. \const{B}( \mu \Var{\ec{x}}. \const{B}( \Var{\ec{x}} ) )$, 
+which has normal form $\mu \Var{\ec{y_2}}. \const{B}( \Var{\ec{y_2}} )$,
+which is $\alpha$-equivalent to $\mu \Var{\ec{x}}. \const{B}( \Var{\ec{x}} )$.
 However, this contradicts the disequality $x \not\teq y_2$ in $\Ec$.
-Inversely, if the value assigned to $v_{\ec{x}}$ is not $\alpha$-equivalent to any term in $\mathcal{T}^\M_{v_{\ec{x}}}( \Val )$,
+Inversely, if the value assigned to $\Var{\ec{x}}$ is not $\alpha$-equivalent to any term in $\mathcal{T}^\M_{\Var{\ec{x}}}( \Val )$,
 then the values in the range of $\Val$ remain $\alpha$-disequivalent.
 For instance, we may assign values such as 
-$\mu v_{\ec{x}}. \const{C}( v_{\ec{x}} )$,
-$\mu v_{\ec{x}}. \const{C}( \mu x_1. \const{A}( x_1) )$, and
-$\mu v_{\ec{x}}. \const{B}( \mu x_1. \const{A}( v_{\ec{x}} ) )$
-to $v_{\ec{x}}$.
-Notice that legal substitutions for $v_{\ec{x}}$ may cause the range of $\Val$ to contain non-normal terms,
+$\mu \Var{\ec{x}}. \const{C}( \Var{\ec{x}} )$,
+$\mu \Var{\ec{x}}. \const{C}( \mu x_1. \const{A}( x_1) )$, and
+$\mu \Var{\ec{x}}. \const{B}( \mu x_1. \const{A}( \Var{\ec{x}} ) )$
+to $\Var{\ec{x}}$.
+Notice that legal substitutions for $\Var{\ec{x}}$ may cause the range of $\Val$ to contain non-normal terms,
 as in after applying the third substitution mentioned above,
-$\Val \ec{w_1}$ is $\mu v_{\ec{w_1}}. \const{A}( \mu v_{\ec{x}}. \const{B}( \mu x_1. \const{A}( v_{\ec{x}} ) ) )$.
+$\Val \ec{w_1}$ is $\mu \Var{\ec{w_1}}. \const{A}( \mu \Var{\ec{x}}. \const{B}( \mu x_1. \const{A}( \Var{\ec{x}} ) ) )$.
 $\Box$
@@
 If $\const{C}( t_1, \ldots, t_n ) \in \ec{x}$,
-then $\Val^\ast \ec{x} = \mu v_{\ec{x}}. \const{C}( w_1, \ldots, w_n )$
+then $\Val^\ast \ec{x} = \mu \Var{\ec{x}}. \const{C}( w_1, \ldots, w_n )$
 and $\Val^\ast \ec{x} \interp{ w_i } = \Val^\ast \ec{t_i}$ for $i = 1, \ldots, n$.
-%and $\muvar( w_i ) = v_{\ec{t_i}}$ for $i = 1, \ldots, n$,
+%and $\muvar( w_i ) = \Var{\ec{t_i}}$ for $i = 1, \ldots, n$,
 %\item
 %$t_1 \interp{ t_2 } = \Val^\ast \ec{x}$
-%for all $t_1$ with subterm $t_2$ in the range of $\Val^\ast$ where $\muvar(t_2)=v_{\ec{x}}$,
+%for all $t_1$ with subterm $t_2$ in the range of $\Val^\ast$ where $\muvar(t_2)=\Var{\ec{x}}$,
 \item 
@@
 To show 1, we first show that $\Val^\ast$ contains no free variables.
-Assume by contradiction that $\Val^\ast$ contained a free variable $v_{\ec{y}}$ for some $\ec{y}$ of type $\tau$.
+Assume by contradiction that $\Val^\ast$ contained a free variable $\Var{\ec{y}}$ for some $\ec{y}$ of type $\tau$.
 Then it must be the case that $\ec{y}$ does not contain a constructor term,
-or else $v_{\ec{y}}$ would not occur as a free variable in $\Val$.
+or else $\Var{\ec{y}}$ would not occur as a free variable in $\Val$.
 Consider the case when $\tau$ is finite.
@@
 and thus there are no terms in $\mathcal{T}^\M_x(\Val^\ast)$ of type $\tau$.
-This is a contradiction, since our model completion may assign the value in the domain of $\tau$ to $v_{\ec{x}}$.
+This is a contradiction, since our model completion may assign the value in the domain of $\tau$ to $\Var{\ec{x}}$.
 Now, consider the case when $\tau$ is infinite.
@@
 since there are only a finite number of closed terms in $\mathcal{T}^\M_x(\Val^\ast)$,
-and thus our model completion can assign a value not occurring in $\mathcal{T}^\M_x(\Val^\ast)$ to $v_{\ec{y}}$.
+and thus our model completion can assign a value not occurring in $\mathcal{T}^\M_x(\Val^\ast)$ to $\Var{\ec{y}}$.
 By construction, $\Val^\ast \ec{x}$ is normal.
@@
 $\Val$ was constructed by applying a sequence of substitutions
-that replace $v_{\ec{y}}$ by $\mu v_{\ec{y}}. \const{C}_i( v_{\ec{t_1}}, \ldots, v_{\ec{t_n}} )$.
+that replace $\Var{\ec{y}}$ by $\mu \Var{\ec{y}}. \const{C}_i( \Var{\ec{t_1}}, \ldots, \Var{\ec{t_n}} )$.
 Now, say that $\const{C}( t_1, \ldots, t_n )$ occurs in some equivalence class $\ec{x}$.
-Then $\Val \ec{x}$ must be of the form $\mu v_{\ec{x}}. \const{C}_i( w_1, \ldots, w_n )$,
+Then $\Val \ec{x}$ must be of the form $\mu \Var{\ec{x}}. \const{C}_i( w_1, \ldots, w_n )$,
 where for each $i=1,\ldots,n$, $w_i$ was obtained by applying a sequence of substitutions
-of the form mentioned above to $v_{\ec{t_i}}$.
-%Each $w_i$ is either $v_{\ec{x}}$,
-%a variable $v_{\ec{y}}$ where $\ec{y}$ does not contain a constructor,
-%or a term of the form $\mu v_{\ec{y}}. \const{C}_j( \vec u )$.
+of the form mentioned above to $\Var{\ec{t_i}}$.
+%Each $w_i$ is either $\Var{\ec{x}}$,
+%a variable $\Var{\ec{y}}$ where $\ec{y}$ does not contain a constructor,
+%or a term of the form $\mu \Var{\ec{y}}. \const{C}_j( \vec u )$.
 While interpreting $w_i$ in $\Val \ec{x}$,
-all free occurrences of $v_{\ec{x}}$ in $\Val \ec{x} \interp{w_i}$ are replaced by $\Val \ec{x}$. \rem{better formalize.}
-Thus, since $\Val \ec{x} \interp{w_i}$ and $\Val \ec{t_i}$ are obtained by applying the same sequence of substitutions to $v_{\ec{t_i}}$,
+all free occurrences of $\Var{\ec{x}}$ in $\Val \ec{x} \interp{w_i}$ are replaced by $\Val \ec{x}$. \rem{better formalize.}
+Thus, since $\Val \ec{x} \interp{w_i}$ and $\Val \ec{t_i}$ are obtained by applying the same sequence of substitutions to $\Var{\ec{t_i}}$,
 we have $\Val \ec{w_i} = \Val \ec{x} \interp{ w_i }$.
@@
 %Since \rn{Clash} does not apply to $\Fc$, then $\Val^\ast \ec{t}$ must be a term
-$t_w$ of the form $\mu v_{\ec{t}}. \const{C}( w_1, \ldots, w_n )$
+$t_w$ of the form $\mu \Var{\ec{t}}. \const{C}( w_1, \ldots, w_n )$
 where $t_w\interp{w_i} = \Val^\ast \ec{u_i}$ for $i = 1, \ldots, n$.
@@
 By Lemma~\ref{lem:mu-norm-arg}, $t_u\interp{u_i'} \vsim \Val^\ast \ec{u_i}$.
-%Say that $\muvar( w_i ) = v_{\ec{y_i}}$.
+%Say that $\muvar( w_i ) = \Var{\ec{y_i}}$.
 %Due to our construction of $\Val^\ast$ and since \rn{Inject} does not apply to $\Fc$, we have $\ec{u_i} = \ec{y_i}$,
@@
 By Lemma~\ref{lem:model-completion}.2, $\Val^\ast \ec{u}$ is a term $t_w$ of the form
-$\mu v_{\ec{u}}. C_i( w_1, \ldots, w_n )$,
+$\mu \Var{\ec{u}}. C_i( w_1, \ldots, w_n )$,
 where $t_w \interp{ w_j } = \Val^\ast \ec{ s^j_k( u )}$. 
 %By our construction of $\Val^\ast$ and since \rn{Inject} and \rn{Clash} do not apply to $\Fc$, 
-%$\Val^\ast \ec{ u }$ must be of the form $t_u = \mu v_{\ec{u}}. C_i( u_1, \ldots, u_n )$,
-%where $\muvar( u_1 ) = v_{\ec{s^1_i(u)}}$, $\ldots$, $\muvar(u_n) = v_{\ec{s^n_i( u )}}$,
+%$\Val^\ast \ec{ u }$ must be of the form $t_u = \mu \Var{\ec{u}}. C_i( u_1, \ldots, u_n )$,
+%where $\muvar( u_1 ) = \Var{\ec{s^1_i(u)}}$, $\ldots$, $\muvar(u_n) = \Var{\ec{s^n_i( u )}}$,
 %and $\M( u )$ is of the form $t_w = \mu x. C_i( w_1, \ldots, w_n )$,
@@
 a set of equalities from $M_i$ that $\thD$-entail that the values of $\ec{x}$ in models of $\Ec$ are of the form $\Val \ec{x}$.
-For instance, if $x \teq \const{C}( x ) \in M_i$, then $x \teq \const{C}( x )$ is a (possible) explanation for $\Val \ec{x} = \mu v_{\ec{x}}. \const{C}( v_{\ec{x}} )$.
+For instance, if $x \teq \const{C}( x ) \in M_i$, then $x \teq \const{C}( x )$ is a (possible) explanation for $\Val \ec{x} = \mu \Var{\ec{x}}. \const{C}( \Var{\ec{x}} )$.
 Thus, the rules of the calculus are implemented as follows.

### 2014-12-19 tex/codatadp.tex
more

@@
 \)
-\rn{Update}}
+\rn{Refine}}
 \]
-In the rule, $\FV( \Val )$ denotes the union of the free variables occurring in $\Val$'s range.
+$\FV( \Val )$ denotes the union of the free variables occurring in $\Val$'s range.
 It is straightforward to see that the height of terms produced as a result of this update
 is bounded by the number of equivalence classes of $\Ec$,
-and thus our construction of $\Val$ is terminating.
+and thus the construction of $\Val$ will terminate.
 

### 2014-12-19 tex/codatadp.tex
Clean up

--- 2014-12-19 tex/codatadp.tex
@@
 \newcommand\dotReportFootnote[1]{.}
-
-\newcommand\FV{\mathrm{FV}}
 
@@
 \newcommand{\nf}[1]{{#1}\!\!\downarrow}
-\newcommand{\vsim}{\sim}
 \newcommand{\aequiv}{\mathrel{=_\alpha}}
-\newcommand{\vsimv}[1]{\vsim_{#1}}
+\newcommand{\vsim}{\aequiv}
+\newcommand{\vsimv}[1]{\mathrel{=^{#1}_\alpha}}
 \newcommand{\tpos}[2]{#1\!\mid_{#2}}
@@
 
-The decision procedure for the ground theory of (co)datatypes $\thD$ determines the
-$\thD$-satisfiability of sets of ground constraints over a fixed
-signature~$\Sig$. It is presented as a calculus, whose rules can be applied in
-any order until saturation.
-
-The derivation rules operate on a set $\Ec$ of equalities and disequalities
-between $\Sig$-terms, where $\Ec$ is interpreted as a conjunction.
+This section presents a decision procedure for \rem{introduce ``$\thD$" in previous section?}
+determining the $\thD$-satisfiability of sets of constraints built from a fixed signature $\Sig$.
+We will make the following assumptions about $\Sig$.
+First, all codatatypes in $\Codata$ are corecursive.
+This is not restrictive, since checking the $\thD$-satisfiability for constraints involving a non-corecursive co-datatype $\tau$ can be trivially reduced to
+checking the satisfiability of the same constraint involving a (non-recursive) datatype of the same form as $\tau$. \rem{revise?}
+Second, all types $\tau \in \Nondata$ have infinite cardinality.
+This also is not restrictive, since each interpreted type $\tau \in \Nondata$ having finite cardinality $n$
+can be replaced by a enumeration datatype with $n$ constructors
+\footnote{Doing so simplifies the presentation of the calculus, but is not done in practice. \rem{More?}}.
+Additionally, since $\Ec$ is ground, we know that $\Ec$ cannot entail an upper bound on the cardinality of any uninterpreted type $\tau \in \Nondata$, 
+and thus without loss of completeness we may treat all other types as infinite.
+
+Our calculus consists of derivation rules that operate on a set $\Ec$ of equalities and disequalities between $\Sig$-terms,
+where $\Ec$ is interpreted as a conjunction.
 We write $\tEc$ to denote the set of $\Sig$-terms occurring in $\Ec$.
-%We will commonly denote tuples of terms $( t_1, \ldots, t_n )$ in bold font, as $\vec t$.
-We assume that all $\Sig$-terms in $\Ec$ are \emph{normalized}, meaning that all subterms of the form $\const{s}^{\,k}_j( \const{C}_j(t_1,\ldots,t_n) )$
-are simplified to $t_k$. All constraints added to $\Ec$ are also implicitly
-normalized in this manner.
-
-To simplify the presentation, we make the following assumptions about $\Sig$.
-First, all codatatypes in $\Codata$ are corecursive. This is not restrictive
-because noncorecursive codatatypes can be considered as nonrecursive
-datatypes.
-Second, all types $\tau \in \Nondata$ have infinite cardinality.
-This is not overly restrictive, because each interpreted type $\tau \in \Nondata$
-having finite cardinality~$n$
-can be replaced by an enumeration datatype with $n$~nullary constructors.
-Moreover, since the constraints are ground, they cannot entail an upper bound
-on the cardinality of any uninterpreted type $\tau \in \Nondata$; thus,
-we may without loss of completeness treat all other types as infinite.
-
-The calculus consists of three sets of rules (Figures \ref{fig:cc-rules} to
-\ref{fig:split-rule}) corresponding to three successive phases. The first
-phase computes the bidirectional closure of $\Ec$. The second phase makes
-inferences based on acyclicity (for datatypes) and bisimilarity (for
-codatatypes). The third phase performs case splitting on constructors for
-various terms in $\tEc$.
-%Following the conventions from~\cite{}, 
-%The derivation rules are given in \relax{guarded assignment form},
-%where
-A rule can be applied to $\Ec$ if all the specified preconditions are met.
-The conclusion of a rule either describes equalities to be added to $\Ec$
-or is $\bot$.
-% (in which case we call an application of it \emph{nonterminal})
-% (in which case we call an application of it \emph{terminal}).
-A rule may have multiple conclusions separated by $\parallel$,
-denoting nondeterministic branching. The rules associated with a phase are
-applied until saturation (i.e., until no rule can be applied to extend~$\Ec$
-further) before the next phase can begin.
+%We assume that all $\Sig$-terms in $\Ec$ are \emph{normalized}, meaning that all subterms of the form $\const{s}^{\,k}_j( \const{C}_j(t_1,\ldots,t_n) )$
+%are simplified to $t_i$, and moreover assume that all %additional
+%constraints added to $\Ec$ are normalized in this manner.
+We present the calculus in three steps.
+In the first step, we compute the bidirectional closure of $\Ec$;
+in the second step, we make inferences based on cyclicity and bisimilarity;
+and in the third step, %(when necessary),
+we branch on constructor types for various terms in $\tEc$.
+Following the conventions from~\cite{}, 
+the derivation rules of our calculus are given in \emph{guarded assignment form},
+where a rule can be applied to $\Ec$ if it meets all of the specified preconditions for $\Ec$.
+The conclusion of a rule either describes equalities to be added to $\Ec$ (in which case we call an application of it \emph{nonterminal}),
+or is $\bot$ (in which case we call an application of it \emph{terminal}).
+A rule may have multiple conclusions separated by $\parallel$, which denotes nondeterministic branching.
 
@@
 \inferrule{
- t \teq u \in \Ec
+ t_1 \teq t_2 \in \Ec
 }{
- \Ec := \Ec,\: u \teq t
+ \Ec := \Ec,\: t_2 \teq t_1
 }
@@
 \inferrule{
-  t \teq u,\; t \tneq u \in \Ec
+  t_1 \teq t_2,\; t_1 \tneq t_2 \in \Ec
 }{
@@
 \inferrule{
-  s \teq t,\; t \teq u \in \Ec
+  t_1 \teq t_2,\; t_2 \teq t_3 \in \Ec
 }{
-  \Ec := \Ec,\: s \teq u
+  \Ec := \Ec,\: t_1 \teq t_3
 }
@@
 \)
-\rn{Inject} 
+\rn{Decomp} 
 \qquad
@@
 \end{tabular}
-\caption{\,Derivation rules for bidirectional closure%.
+\caption{\,Rules for bidirectional closure%.
 }
@@
 
-\paragraph{Phase 1: Computing the Bidirectional Closure.}
-In conjunction with \rn{Refl}, \rn{Sym}, and \rn{Trans}, the \rn{Cong} rule computes the (upward) congruence closure,
-whereas the \rn{Inject} and \rn{Clash} rules together compute (downward) unification.
-For unification, additional equalities are inferred based on the injectivity of constructors by \rn{Inject},
+\paragraph{Step 1: Computing the Bidirectional Closure.}
+Figure~\ref{fig:cc-rules} gives the basic rules of the calculus.
+Together with \rn{Refl}, \rn{Sym}, and \rn{Trans}, the \rn{Cong} rule computes the (upward) congruence closure,
+whereas the \rn{Decomp} and \rn{Clash} rules together compute (downward) unification.
+In the unification rules, additional equalities are inferred based on the injectivity of constructors by \rn{Decomp},
 and failures to unify equated terms are recognized by \rn{Clash}.
 The rule \rn{Conflict} recognizes when an equality and its negation both occur in $\Ec$, in which case $\Ec$ has no model.
-
-At the end of this phase, $\Ec$ induces an equivalence
-relation over $\tEc$ such that two terms $t$ and $u$ are equivalent if and
-only if $t \teq u \in \Ec$.
-Thus, we can regard $\Ec$ as a set of
-equivalence classes of terms. For a term $t \in \tEc$, we write $\ec{t}$ to
-denote the equivalence class in $\Ec$ that contains $t$.
 
@@
   \quad
-  t : \tau
+  t_1 : \tau
   \quad
   %\ttpath{\Ec}{\ec{t}}{\ec{t}} \neq \emptyset
-  \Val \ec{t} = \mu x.\; u
+  \Val \ec{t_1} = \mu x.\; t_2
   \quad
-  x \in \FV( u )
+  x \in FV( t_2 )
 }{
@@
  \quad
- t, u : \tau
+ t_1, t_2 : \tau
  \quad
- %\tpath{\Ec}{\ec{t}} = \tpath{\Ec}{\ec{u}}\sigma \neq \emptyset
- \Val \ec{t} \aequiv \Val \ec{u}
+ %\tpath{\Ec}{\ec{t_1}} = \tpath{\Ec}{\ec{t_2}}\sigma \neq \emptyset
+ \Val \ec{t_1} \vsim \Val \ec{t_2}
 }{
- \Ec := \Ec,\: t \teq u
+ \Ec := \Ec,\: t_1 \teq t_2
 }
@@
 \end{tabular}
-\caption{\,Derivation rules for acyclicity and bisimilarity%.
+\caption{\,Rules for cyclicity and bisimilarity%.
 }
@@
 
-\paragraph{Phase 2: Applying Acyclicity and Bisimilarity.}
-%For presentation of the rules in this phase, we rely on a representation of terms in the $\mu$-notation.
-The rules of the second phase contain premises that involve a mapping $\Val$
-that associates a $\mu$-term as representative with each equivalence class.
-%Recall that $\mu$-bindings can be used for representing possibly cyclic terms and values.
+\paragraph{Step 2: Applying Cyclicity and Bisimilarity.}
+In the second step,
+we assume that $\Ec$ is saturated with respect to the rules in Figure~\ref{fig:cc-rules}---that is,
+no rule can be applied to extend~$\Ec$ further.
+When this is the case, it is easy to see that $\Ec$ induces an equivalence relation over $\tEc$ such that two terms $t_1$ and $t_2$ are equivalent if and only if $t_1 \teq t_2 \in \Ec$.
+Thus, we can consider $\Ec$ as a set of equivalence classes of terms. 
+For a term $t \in \tEc$, we write $\ec{t}$ to denote the equivalence class in $\Ec$ that contains $t$.
+
+%For presentation of the rules in this step, we rely on a representation of terms in the $\mu$-notation.
+The rules of this step contain premises that involve a mapping $\Val$ from equivalence classes to terms containing $\mu$-bindings.
+Recall that $\mu$-bindings can be used for representing possibly cyclic terms and values.
 %(Section~\ref{sec:introduction}).
 % for example, $\mu x. \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
-Intuitively, a $\mu$-term $t$ of type $\tau$ describes a class of values of $\tau$,
-%either in the case when $\tau$ is a codatatype or datatype type,
-where a cyclic $t$ describes an infeasible class of values when
-%$\tau \not\in \Codata$.
-$\tau \in \Data$.
 Formally, $\mu$-terms are defined recursively as being either a variable $x$
@@
 $\mu$-terms $\vec t$ with the expected types.
-The variable $x$ need not occur under the binder's body, in which case the
-binder can be omitted.
+The variable $x$ need not occur under the binder's body, in which case the $\mu$-prefix
+is optional.
 
@@
 \not\aequiv \mu y.\; \const{D}( y, x )$.
-Two $\mu$-terms can represent the same value while not being $\alpha$-equivalent---e.g.,
-$\mu x.\; \const{C}( x ) \not\aequiv \mu x.\; \const{C}( \const{C}( x ) )$.
+Two $\mu$-terms may represent the same value while not being $\alpha$-equivalent, as in
+$\mu x.\; \const{C}( \const{C}( x ) ) \not\aequiv \mu x.\; \const{C}( x )$.
 %, or informally are \emph{observationally equivalent}.
-%For convienience, we use $\mu$-terms to refer to (classes of) values for both codatatype and datatype terms,
+%For convienience, we use $\mu$-terms to refer to (classes of) values for both co-datatype and datatype terms,
 %where in the latter case, a $\mu$-term with a bound variable denotes an infeasible term.
@@
 
-At the beginning of this phase, 
+Intuitively, a $\mu$-term $t$ of type $\tau$ describes a class of values of $\tau$, 
+%either in the case when $\tau$ is a co-datatype or datatype type,
+where a cyclic $t$ describes an infeasible class of values when
+%$\tau \not\in \Codata$.
+$\tau \in \Data$.
+At the beginning of this step, 
 we compute a mapping $\Val$ from equivalence classes to $\mu$-terms 
@@
 \inferrule{
-  v_{\ec{x}} \in \FV( \Val )
+  v_{\ec{x}} \in FV( \Val )
   \quad
@@
 \]
-We write $\FV( \Val )$ to denote the \relax{free variables} occurring in $\Val$, which
-exclude variables that occur beneath $\mu$-binders.
+Here, we write $FV( \Val )$ to denote the \emph{free variables} occurring in $\Val$, which
+excludes variables that occur beneath $\mu$-binders.
 It is straightforward to see that the height of terms produced as a result of this update
@@
 
-Given $\Val$, the rules for acyclicity and bisimilarity (Figure~\ref{fig:ab-rules}) can be stated as follows.
-For acyclicity, if there is an equivalence class $\ec{t}$ of datatype type whose class of values $\Val \ec{t} = \mu x.\; u$ are cyclic,
-as indicated by the assumption $x \in \FV( u )$,
+Given $\Val$, the rules for cyclicity and bisimilarity (shown in Figure~\ref{fig:ab-rules}) can be stated as follows.
+For cyclicity, if there is an equivalence class $\ec{t_1}$ of datatype type whose class of values $\Val \ec{t_1} = \mu x. t_2$ are cyclic,
+as indicated by the assumption $x \in FV( t_2 )$,
 then we can conclude that $\Ec$ is unsatisfiable.
-For bisimilarity, if there are two equivalence classes $\ec{t}$ and $\ec{u}$ 
-of codatatype type whose classes of values $\Val \ec{t}$ and $\Val \ec{u}$ are equivalent for a capture-avoiding renaming of $\mu$-bound variables,
-then we can conclude $t \teq u$.
+For bisimilarity, if there are two equivalence classes $\ec{t_1}$ and $\ec{t_2}$ 
+of co-datatype type whose classes of values $\Val \ec{t_1}$ and $\Val \ec{t_2}$ are equivalent for a capture-avoiding renaming of $\mu$-bound variables,
+then we can conclude $t_1 \teq t_2$.
 We demonstrate the bisimilarity rule concretely with the following example.
@@
 $\{ x, \const{C}(y,\const{C}(y,x)) \}$, $\{ \const{C}(y,x) \}$, and $\{ y \}$.
-We then construct $\Val$ using the aforementioned phases, after which we obtain:
+We then construct $\Val$ using the aforementioned steps, after which we obtain:
 \[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}r@{\hspace{.5em}}l@{\hspace{.5em}}r@{\hspace{.5em}}l}
@@
 \end{tabular}
-\caption{\,Derivation rules for branching%.  
+\caption{\,Rules for branching%.  
 %All pairs of terms whose type has cardinality 1 are entailed to be equal (the degenerate case).
@@
 
-\paragraph{Phase 3: Branching.}
+\paragraph{Step 3: Branching.}
 If neither a rule from Figure \ref{fig:cc-rules}~or~\ref{fig:ab-rules} applies to $\Ec$, 
@@
 
-An application of a rule is \emph{redundant} if at least one branch in the
-conclusion does not add a new equality to $\Ec$ and is not $\bot$.
+\
+
+An application of a rule is \emph{redundant} if it is nonterminal and at least one branch in its conclusion does not add a new equality to $\Ec$.
 A \emph{derivation tree} is a tree whose nodes are sets of equalities, where non-root nodes vtained by 
@@
 Otherwise, the child of $\Ec$ is a closed derivation tree 
-whose root node is $\Ec \cup t_1 \teq t_2$ obtained by applying one of the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, \rn{Inject}, \rn{Bisimilar}, or \rn{Singleton}.
+whose root node is $\Ec \cup t_1 \teq t_2$ obtained by applying one of the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, \rn{Decomp}, \rn{Bisimilar}, or \rn{Singleton}.
 In all cases, we have that $\Ec \models_{\thD} t_1 \teq t_2$.
 For \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, this is a consequence of equality reasoning.
-For \rn{Inject}, this is a consequence of injectivity.
+For \rn{Decomp}, this is a consequence of injectivity.
 For \rn{Bisimilar}, our construction of $\Val$ indicates that the value of $t_1$ and $t_2$ are equivalent in all models of $\Ec$.
@@
 
-We now define a class of models that we consider.
+We now define a class of models that we consider for a saturated set of equalities $\Ec$.
+In the following, we write $\M( \tau )$ to denote the interpretation type $\tau$ in $\M$,
+that is, a set of domain elements for that type.
+We write $\M( t )$ for a term $t$ to denote the interpretation of $t$ in $\M$,
+that is, an element of $\M( \tau )$ where $\tau$ is the type of $t$.
 \rem{Clarify that non-codatatype terms are interpreted as $\mu$-terms, although dropping the $\mu$'s gives the standard representation?}
@@
 $\M( \const{C}( \vec t \,) )$ is the value 
-in $\M( \tau )$ that is a renaming of 
+in $\M( \tau )$ that is $\alpha$-equivalent to
 $\nf{(\mu x. \const{C}( \M( \vec t \,) ))}$ where $x$ is fresh, and
@@
 $\M( s^j_i( t ) )$ is the value
-in $\M( \tau )$ that is a renaming of 
+in $\M( \tau )$ that is $\alpha$-equivalent to
 $\M( t )\interp{u_j}$
@@
 it remains only to define how $\M$ interprets wrong-applied selector terms and variables.
-For the latter, this will be based on the mapping $\Val$ constructed in the second phase of our calculus.
+For the latter, this will be based on the mapping $\Val$ constructed in step two of our calculus.
 We compute the \emph{completion} of $\Val$ for normal model $\M$, which we write as $\Val^\ast$,
@@
 we write $\mathcal{T}^\M_x( \Val )$ to denote the set consisting of all values $v$ from $\M( \tau )$
-such that $v \vsim_x t_1\interp{t_2}$ for some term $t_1$ with subterm $t_2$ occurring in the range of $\Val$.
+such that $v \vsimv{x} t_1\interp{t_2}$ for some term $t_1$ with subterm $t_2$ occurring in the range of $\Val$.
 We first construct a mapping $\Val'$ by exhaustively applying the following update to $\Val$:
@@
 \inferrule{
-  x : \tau \in \FV( \Val ) 
+  x : \tau \in FV( \Val ) 
   \quad
-  (\mu x. t) \sim v
+  (\mu x. t) \aequiv v
   \quad
@@
 In this update, we choose an unassigned variable in $\Val$
-and assign it a value that is not bisimilar to one occurring in $\mathcal{T}^\M_x(\Val)$.
-Since this update removes one variable from the set $\FV( \Val )$ and does not add any additional variables to $\FV( \Val )$, 
+and assign it a value that is not $\alpha$-equivalent to one occurring in $\mathcal{T}^\M_x(\Val)$.
+Since this update removes one variable from the set $FV( \Val )$ and does not add any additional variables to $FV( \Val )$, 
 it can only be applied a finite number of times.
@@
 $\mu v_{\ec{x}}. \const{C}( v_{\ec{x}} )$,
-$\mu v_{\ec{x}}. \const{C}( \mu x_1. \const{A}( x_1) )$, and
-$\mu v_{\ec{x}}. \const{B}( \mu x_1. \const{A}( v_{\ec{x}} ) )$
+$\mu v_{\ec{x}}. \const{C}( \mu x_1. \const{A}( x_1) )$, or
+$\mu v_{\ec{x}}. \const{B}( \const{A}( v_{\ec{x}} ) )$
 to $v_{\ec{x}}$.
@@
 
-We first state the following properties of $\Val^\ast$,.
+We first state the following properties of $\Val^\ast$.
 %where we write $\muvar( t )$ to denote $t$ if $t$ is a variable, or $x$ if $t$ is of the form $\mu x. u$.
@@
 \begin{enumerate}
-\item $\Val^\ast \ec{x}$ is bisimilar with a value in $\M( \tau )$, for all $\ec{x}$ of type $\tau$ of $\Ec$,
+\item $\Val^\ast \ec{x}$ is $\alpha$-equivalent to a value in $\M( \tau )$, for all $\ec{x}$ of type $\tau$ of $\Ec$,
 \item
@@
 Thus, $\Val^\ast \ec{x}$ is acyclic when $\tau \not\in \Codata$.
-Therefore, by definition, $\Val^\ast \ec{x}$ is bisimilar with a value in $\M( \tau )$.
+Therefore, by definition, $\Val^\ast \ec{x}$ is $\alpha$-equivalent to a value in $\M( \tau )$.
 
@@
 are such that $i = j$ since \rn{Clash} does not apply to $\Ec$,
-and are such that $\ec{ \vec t } = \ec{ \vec u }$ since \rn{Inject} does not apply to $\Ec$.
+and are such that $\ec{ \vec t } = \ec{ \vec u }$ since \rn{Decomp} does not apply to $\Ec$.
 Thus, 
@@
 that interprets wrongly-applied selectors based on equality information in $\Fc$,
-and interprets the variables of $\Fc$ based on the completion $\Val^\ast$ of the mapping $\Val$ from the second phase of our calculus.
-For the latter, let $\M( x )$ be the value in $\M$ that is bisimilar with $\Val^\ast \ec{x}$ for each variable $x$ in $\mathcal{T}(\Fc)$.
+and interprets the variables of $\Fc$ based on the completion $\Val^\ast$ of the mapping $\Val$ from step two of our calculus.
+For the latter, let $\M( x )$ be the value in $\M$ that is $\alpha$-equivalent with $\Val^\ast \ec{x}$ for each variable $x$ in $\mathcal{T}(\Fc)$.
 
@@
 by structural induction on $t$,
-that $\M( t ) \sim \Val^\ast \ec{t}$ for all terms $t \in \mathcal{T}( \Fc )$,
-which implies $\M \models t_1 \teq t_2$.
+that $\M( t ) \aequiv \Val^\ast \ec{t}$ for all terms $t \in \mathcal{T}( \Fc )$,
+which since $\M$ is normal implies $\M \models t_1 \teq t_2$.
 
@@
 If $t$ is a constructor term of the form $\const{C}( u_1, \ldots, u_n )$, 
-then $\M( t )$ is bisimilar with $\nf{(\mu x. \const{C}( \M( u_1 ), \ldots, \M( u_n ) ) )}$ for fresh $x$,
-which by the induction hypothesis is bisimilar with $\nf{(\mu x. \const{C}( \Val^\ast \ec{ u_1 }, \ldots, \Val^\ast \ec{ u_n } ) )}$, call this term $t_u$.
+then $\M( t )$ is $\alpha$-equivalent with $\nf{(\mu x. \const{C}( \M( u_1 ), \ldots, \M( u_n ) ) )}$ for fresh $x$,
+which by the induction hypothesis is $\alpha$-equivalent with $\nf{(\mu x. \const{C}( \Val^\ast \ec{ u_1 }, \ldots, \Val^\ast \ec{ u_n } ) )}$, call this term $t_u$.
 By Lemma~\ref{lem:model-completion}.2, $\Val^\ast \ec{t}$ is a term
@@
 %Say that $\muvar( w_i ) = v_{\ec{y_i}}$.
-%Due to our construction of $\Val^\ast$ and since \rn{Inject} does not apply to $\Fc$, we have $\ec{u_i} = \ec{y_i}$,
+%Due to our construction of $\Val^\ast$ and since \rn{Decomp} does not apply to $\Fc$, we have $\ec{u_i} = \ec{y_i}$,
 %and thus $t_u\interp{u_i'} \vsim \Val^\ast \ec{y_i}$.
@@
 where $t_w \interp{ w_j } = \Val^\ast \ec{ s^j_k( u )}$. 
-%By our construction of $\Val^\ast$ and since \rn{Inject} and \rn{Clash} do not apply to $\Fc$, 
+%By our construction of $\Val^\ast$ and since \rn{Decomp} and \rn{Clash} do not apply to $\Fc$, 
 %$\Val^\ast \ec{ u }$ must be of the form $t_u = \mu v_{\ec{u}}. C_i( u_1, \ldots, u_n )$,
@@
 %and $\M( u )$ is of the form $t_w = \mu x. C_i( w_1, \ldots, w_n )$,
-%which by the induction hypothesis is bisimilar with $\Val^\ast \ec{u}$.
-If $k = i$, then $\M( t )$ is bisimilar with $t_w\interp{w_j} = \Val^\ast \ec{ s^j_k( u )}$,
+%which by the induction hypothesis is $\alpha$-equivalent with $\Val^\ast \ec{u}$.
+If $k = i$, then $\M( t )$ is $\alpha$-equivalent with $t_w\interp{w_j} = \Val^\ast \ec{ s^j_k( u )}$,
 which is $\Val^\ast \ec{t}$.
@@
 %Thus, $\M( t ) \vsim \Val^\ast \ec{ s^j_k( u )} 
-If $k \neq i$, then we interpret $s^j_k( \M( u ) )$ as the value in $\M( \tau )$ that is bisimilar with $\Val^\ast \ec{ t }$.
+If $k \neq i$, then we interpret $s^j_k( \M( u ) )$ as the value in $\M( \tau )$ that is $\alpha$-equivalent with $\Val^\ast \ec{ t }$.
 Since \rn{Cong} does not apply and due to Lemma~\ref{lem:model-completion}.3, 
@@
 
-This section describes how the calculus in Section~\ref{sec:the-calculus} is implemented within the SMT solver CVC4.
+This section describes how the calculus in Section~\ref{sec:a-ground-decision-procedure-for-co-datatypes} is implemented within the SMT solver CVC4.
 
@@
 if $s^j_i( t )$ is a term in $\tEc$ and $t$ contains the constructor 
-$\const{C}_i( \vec u )$, then we immediately infer $s^j_i( t ) \teq u_j$, whereas the calculus we presented would apply \rn{Split} and \rn{Inject} before inferring this equality.  
-For the sake of reducing the number of unique constraints considered by the calculus, we compute a normal form for constraints as a preprocessing step.
+$\const{C}_i( \vec u )$, then we immediately infer $s^j_i( t ) \teq u_j$, whereas the calculus we presented would apply \rn{Split} and \rn{Decomp} before inferring this equality.  
+For the sake of reducing the number of unique constraints considered by the calculus, we compute a normal form for constraints as a pre-processing step.
 In particular, we 
@@
 that is, a specialized solver for determining the satisfiability of conjunctions of constraints for its theory.
-%In DPLL(T) \rem{CDCL{T}?},
-%a SAT solver
 Given a theory $T = T_1 \cup \ldots T_n$ and a input set of clauses $F$ in CNF,
-the DPLL(T) \rem{CDCL{T}?} procedure
+the DPLL(T) \rem{CDCL(T)?} procedure
 (incrementally) builds partial assignments from the atomic formulas of $F$ to truth values such that no clause in $F$ is falsified.
@@
 All other rules add equalities to the internal state of the theory solver.
-
-Phase 1 (along with the optimization for collapsing selectors) and the \rn{Singleton} rule are performed eagerly, 
-that is for partial satisfying assignments $M$, while phase~2 and the \rn{Split} rule from phase~3 are performed only for complete satisfying assignments $M$, and in that order.
+The rules in step 1 of the calculus (along with the optimization for collapsing selectors) and the \rn{Singleton} rule are performed eagerly, 
+that is for partial satisfying assignments $M$, while the rules in step 2 and the \rn{Split} rule are performed only for complete satisfying assignments $M$, 
+and in that order.
 

### 2014-12-19 tex/codatadp.tex
more

--- 2014-12-19 tex/codatadp.tex
@@
 \usepackage[all]{xy}
+
+\newcommand\MIDRULE{
+\\[-1pt] %%% TYPESETTING HACK
+\midrule
+%$\enatT$
+\\[-11pt] %%% TYPESETTING HACK
+}
 
@@
 indicates that the $\mu$-terms $t$ and $u$
-are syntactically equivalent for some capture-avoiding renaming of $\mu$-bound variables.
-For example, 
+are syntactically equivalent for some capture-avoiding renaming of $\mu$-bound variables---e.g.,
 %$\mu x. \const{C}( x ) \aequiv \mu y. \const{C}( y )$ and
-$\mu x.\; \const{D}( y, x ) \aequiv \mu z.\; \const{D}( y, z )$,
+$\mu x.\; \const{D}( y, x ) \aequiv \mu z.\; \const{D}( y, z )$\begin{report},
 but
 $\mu x.\; \const{C}( x ) \not\aequiv \mu x.\; \const{D}( y, x ) \not\aequiv \mu x.\; \const{D}( z, x )
-\not\aequiv \mu y.\; \const{D}( y, x )$.
+\not\aequiv \mu y.\; \const{D}( y, x )$\end{report}.
 Two $\mu$-terms can represent the same value while not being $\alpha$-equivalent---e.g.,
@@
 \)
-\rn{Refine}}
+\rn{Update}}
 \]
@@
 \begin{example}
-Say that $\Ec$ contains (distinct) equivalence classes $\ec{x}$, $\ec{y}$, $\ec{z}$ and $\ec{w}$ 
+Suppose that $\Ec$ contains distinct equivalence classes $\ec{w}$, $\ec{x}$, $\ec{y}$, and $\ec{z}$
 where $\const{C}(z,x) \in \ec{y}$ and $\const{C}(w,y) \in \ec{x}$ for some $\const{C} \in \Ctr$.
-A possible sequence of updates applied to $\Val$ is shown below.
-In the first column, we show the equivalence class that was considered,
-and in the second column, we show the content of $\Val$.
-\[\begin{array}{|l|@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l|}
-\hline
-\text{Eqc} & \multicolumn{3}{c}{\Val} & \ 
-\\
-\hline
+A possible sequence of \rn{Update} rule applications is presented below.
+\[\begin{array}{l@{\hspace{1em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
+%\hline
+%\text{Step} & \multicolumn{3}{l}{\text{Mapping }\Val}
+%\MIDRULE
+\text{1. Initial:}
 & \{ & \ec{x} \mapsto \Var{\ec{x}}, & \ec{z} \mapsto \Var{\ec{z}}, & \\
-&    & \ec{y} \mapsto \Var{\ec{y}}, & \ec{w} \mapsto \Var{\ec{w}} & \} \\
-\ec{x}
+&    & \ec{y} \mapsto \Var{\ec{y}}, & \ec{w} \mapsto \Var{\ec{w}} & \} \\[\jot]
+\text{2. Update w.r.t.\ }\ec{x}\text{:}
 & \{ & \ec{x} \mapsto \mu \Var{\ec{x}}. \const{C}( \Var{\ec{w}}, \Var{\ec{y}} ), & \ec{z} \mapsto \Var{\ec{z}}, & \\
-&    & \ec{y} \mapsto \Var{\ec{y}}, & \ec{w} \mapsto \Var{\ec{w}} & \} \\
-\ec{y}
+&    & \ec{y} \mapsto \Var{\ec{y}}, & \ec{w} \mapsto \Var{\ec{w}} & \} \\[\jot]
+\text{3. Update w.r.t.\ }\ec{y}\text{:}
 & \{ & \ec{x} \mapsto \mu \Var{\ec{x}}. \const{C}( \Var{\ec{w}}, \mu \Var{\ec{y}}. \const{C}( \Var{\ec{z}}, \Var{\ec{x}} ) ), & \ec{z} \mapsto \Var{\ec{z}}, & \\
-&    & \ec{y} \mapsto \mu \Var{\ec{y}}. \const{C}( \Var{\ec{z}}, \Var{\ec{x}} ), & \ec{w} \mapsto \Var{\ec{w}} & \} \\
-\ec{x}
+&    & \ec{y} \mapsto \mu \Var{\ec{y}}. \const{C}( \Var{\ec{z}}, \Var{\ec{x}} ), & \ec{w} \mapsto \Var{\ec{w}} & \} \\[\jot]
+\text{4. Update w.r.t.\ }\ec{x}\text{:}
 & \{ & \ec{x} \mapsto \mu \Var{\ec{x}}. \const{C}( \Var{\ec{w}}, \mu \Var{\ec{y}}. \const{C}( \Var{\ec{z}}, \Var{\ec{x}} ) ), & \ec{z} \mapsto \Var{\ec{z}}, & \\
-&    & \ec{y} \mapsto \mu \Var{\ec{y}}. \const{C}( \Var{\ec{z}}, \mu \Var{\ec{x}}. \const{C}( \Var{\ec{w}}, \Var{\ec{y}} ) ), & \ec{w} \mapsto \Var{\ec{w}} & \} \\
-\hline
+&    & \ec{y} \mapsto \mu \Var{\ec{y}}. \const{C}( \Var{\ec{z}}, \mu \Var{\ec{x}}. \const{C}( \Var{\ec{w}}, \Var{\ec{y}} ) ), & \ec{w} \mapsto \Var{\ec{w}} & \}
+%\hline
 \end{array}\]
+%The first column shows the equivalence class that was considered;
+%the second column shows the content of $\Val$.
 This indicates that the values for $x$ (resp. $y$) in models of $\Ec$
@@
   & \HD{Distro} & \HD{AFP} & \HD{\gandl} & All
-\\[1pt] %%% TYPESETTING HACK
-\midrule
-%$\enatT$
-\\[-9pt] %%% TYPESETTING HACK
+\MIDRULE
 No (co)datatypes

### 2014-12-19 tex/codatadp.tex
more

@@
 \newcommand{\dpath}[3]{\delta^{#1}_{#2}( #3 )}
-\newcommand{\ec}[1]{ [ #1 ] }
+\newcommand{\ec}[1]{[#1]}
 \newcommand{\M}{\mathcal{M}}
@@
 \newcommand{\Var}{\mathcal{V}}
+%\newcommand{\Varec}[1]{\Var{\ec{#1}}}
+\newcommand{\Varec}[1]{\widetilde{#1}}
 \newcommand{\nf}[1]{{#1}\!\!\downarrow}
@@
 %
-For each equivalence class $\ec{x}$, we associate a fresh variable $\Var{\ec{x}}$ of the same type as $x$
+For each equivalence class $\ec{x}$, we associate a fresh variable ${\Varec{x}}$ of the same type as $x$
 %not occurring in $\tEc$,
-and let $\Val\ec{x} = \Var{\ec{x}}$.
+and let $\Val\ec{x} = {\Varec{x}}$.
 Thus, there are initially no constraints on the values for any equivalence class $\ec{x}$.
@@
 \inferrule{
-  \Var{\ec{x}} \in \FV( \Val )
+  {\Varec{x}} \in \FV( \Val )
   \quad
@@
 }{
-  \Val := \Val \{ \Var{\ec{x}} \mapsto \mu \Var{\ec{x}}.\; \const{C}( \Var{\ec{t_1}}, \ldots, \Var{\ec{t_n}} ) \}
+  \Val := \Val \{ {\Varec{x}} \mapsto \mu {\Varec{x}}.\; \const{C}( {\Varec{t_1}}, \ldots, {\Varec{t_n}} ) \}
 }
@@
 where $\const{C}(z,x) \in \ec{y}$ and $\const{C}(w,y) \in \ec{x}$ for some $\const{C} \in \Ctr$.
-A possible sequence of \rn{Update} rule applications is presented below.
-\[\begin{array}{l@{\hspace{1em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
-%\hline
-%\text{Step} & \multicolumn{3}{l}{\text{Mapping }\Val}
-%\MIDRULE
-\text{1. Initial:}
-& \{ & \ec{x} \mapsto \Var{\ec{x}}, & \ec{z} \mapsto \Var{\ec{z}}, & \\
-&    & \ec{y} \mapsto \Var{\ec{y}}, & \ec{w} \mapsto \Var{\ec{w}} & \} \\[\jot]
-\text{2. Update w.r.t.\ }\ec{x}\text{:}
-& \{ & \ec{x} \mapsto \mu \Var{\ec{x}}. \const{C}( \Var{\ec{w}}, \Var{\ec{y}} ), & \ec{z} \mapsto \Var{\ec{z}}, & \\
-&    & \ec{y} \mapsto \Var{\ec{y}}, & \ec{w} \mapsto \Var{\ec{w}} & \} \\[\jot]
-\text{3. Update w.r.t.\ }\ec{y}\text{:}
-& \{ & \ec{x} \mapsto \mu \Var{\ec{x}}. \const{C}( \Var{\ec{w}}, \mu \Var{\ec{y}}. \const{C}( \Var{\ec{z}}, \Var{\ec{x}} ) ), & \ec{z} \mapsto \Var{\ec{z}}, & \\
-&    & \ec{y} \mapsto \mu \Var{\ec{y}}. \const{C}( \Var{\ec{z}}, \Var{\ec{x}} ), & \ec{w} \mapsto \Var{\ec{w}} & \} \\[\jot]
-\text{4. Update w.r.t.\ }\ec{x}\text{:}
-& \{ & \ec{x} \mapsto \mu \Var{\ec{x}}. \const{C}( \Var{\ec{w}}, \mu \Var{\ec{y}}. \const{C}( \Var{\ec{z}}, \Var{\ec{x}} ) ), & \ec{z} \mapsto \Var{\ec{z}}, & \\
-&    & \ec{y} \mapsto \mu \Var{\ec{y}}. \const{C}( \Var{\ec{z}}, \mu \Var{\ec{x}}. \const{C}( \Var{\ec{w}}, \Var{\ec{y}} ) ), & \ec{w} \mapsto \Var{\ec{w}} & \}
-%\hline
-\end{array}\]
-%The first column shows the equivalence class that was considered;
-%the second column shows the content of $\Val$.
-This indicates that the values for $x$ (resp. $y$) in models of $\Ec$
-are of the form $\const{C}( \Var{\ec{w}}, \const{C}( \Var{\ec{z}}, \ldots ))$ (resp. $\const{C}( \Var{\ec{z}}, \const{C}( \Var{\ec{w}}, \ldots ))$) 
-for some value of $\Var{\ec{w}}$ and $\Var{\ec{z}}$.
-\qed
+A possible sequence of \rn{Update} rule applications is presented below, omitting
+trivial entries such as $\ec{w} \mapsto {\Varec{w}}$.
+%
+\begin{enumerate}
+\item \noindent\rlap{Initial state:}\phantom{Update w.r.t.\ $\ec{x}$:\enskip}$\{  \}$
+\item \noindent\rlap{Update w.r.t.\ $\ec{x}$:}\phantom{Update w.r.t.\ $\ec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \mu {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) \}$
+\item \noindent\rlap{Update w.r.t.\ $\ec{y}$:}\phantom{Update w.r.t.\ $\ec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \mu {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \mu {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\;
+  \ec{y} \mapsto \mu {\Varec{y}}. \const{C}( {\Varec{z}},\, {\Varec{x}} ) \vthinspace\}$
+\item \noindent\rlap{Update w.r.t.\ $\ec{x}$:}\phantom{Update w.r.t.\ $\ec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \mu {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \mu {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\;
+  \ec{y} \mapsto \mu {\Varec{y}}.\; \const{C}( {\Varec{z}},\; \mu {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) ) \vthinspace\}$
+\end{enumerate}
+%
+This indicates that the values for $x$ and $y$ in models of $\Ec$
+are of the form $\const{C}( {\Varec{w}}, \const{C}( {\Varec{z}}, \ldots ))$
+and $\const{C}( {\Varec{z}}, \const{C}( {\Varec{w}}, \ldots ))$,
+respectively, for some value of ${\Varec{w}}$ and ${\Varec{z}}$.
 \end{example}
@@
 \Val \ec{x} & = & 
-\mu \Var{\ec{x}}. & \const{C}( \Var{\ec{y}}, & \mu \Var{\ec{\const{C}(y,x)}}. & \const{C}( \Var{\ec{y}}, \Var{\ec{x}} ) ) \\
+\mu {\Varec{x}}. & \const{C}( {\Varec{y}}, & \mu {\Varec{\const{C}(y,x)}}. & \const{C}( {\Varec{y}}, {\Varec{x}} ) ) \\
 \Val \ec{\const{C}(y,x)} & = & 
-\mu \Var{\ec{\const{C}(y,x)}}. & \const{C}( \Var{\ec{y}}, & \mu \Var{\ec{x}}. & \const{C}( \Var{\ec{y}}, \Var{\ec{\const{C}(y,x)}} ) )
+\mu {\Varec{\const{C}(y,x)}}. & \const{C}( {\Varec{y}}, & \mu {\Varec{x}}. & \const{C}( {\Varec{y}}, {\Varec{\const{C}(y,x)}} ) )
 \end{array}\]
 Notice that $\Val \ec{x}$ and $\Val \ec{\const{C}(y,x)}$ are syntactically equivalent for the renaming
-$\{ \Var{\ec{x}} \mapsto \Var{\ec{\const{C}(y,x)}}, \Var{\ec{\const{C}(y,x)}} \mapsto \Var{\ec{x}} \}$,
+$\{ {\Varec{x}} \mapsto {\Varec{\const{C}(y,x)}}, {\Varec{\const{C}(y,x)}} \mapsto {\Varec{x}} \}$,
 and thus our calculus may infer $x \teq \const{C}(y,x)$.
 Intuitively, this equality holds, since the values of $x$ and $\const{C}( y, x )$ 
-are equivalent for any value of $\Var{\ec{y}}$.
+are equivalent for any value of ${\Varec{y}}$.
 \qed
@@
 \inferrule{
+  \tau \in \Codata
+  \quad
+  \left|\tau\right| = 1
+  \quad
   t, u : \tau \in \tEc
-  \quad
-  \tau \in \Codata,
-  \mid \tau \mid = 1
 }{
-  \Ec := \Ec, t \teq u
+  \Ec := \Ec,\: t \teq u
 }
@@
 }{
-  \Ec := \Ec, t \teq \const{C}_1( s^1_1( t ), \ldots, s^{a_1}_1( t ) ) \ror \ldots \ror \Ec := \Ec, t \teq \const{C}_n( s^1_n( t ), \ldots, s^{a_n}_n( t ) ) 
+  \Ec := \Ec,\: t \teq \const{C}_1( s^1_1( t ), \ldots, s^{a_1}_1( t ) ) \ror \cdots \ror \Ec := \Ec, t \teq \const{C}_n( s^1_n( t ), \ldots, s^{a_n}_n( t ) ) 
 }
@@
 the rules from Figure~\ref{fig:split-rule} are applied.
-The rule \rn{Singleton} handles the case when there are two terms from $\tEc$ having a co-recursive singleton type,
+The \rn{Singleton} rule handles the case when there are two terms from $\tEc$ having a co-recursive singleton type,
 in which case they must be equivalent.
@@
 \[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{2.0em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
-\Val \ec{y_1} & = & \mu \Var{\ec{y_1}}. & \const{A}( & \mu \Var{\ec{y_2}}. & \const{B}( \Var{\ec{x}} ) ) &
-\Val \ec{w_1} & = & \mu \Var{\ec{w_1}}. & \const{A}( \Var{\ec{x}} )      & \\ 
-\Val \ec{y_2} & = & \mu \Var{\ec{y_2}}. & \multicolumn{2}{l}{\const{B}( \Var{\ec{x}} )}      & &
-\Val \ec{w_2} & = & \mu \Var{\ec{w_2}}. & \const{C}( \Var{\ec{w_3}} ) \\
-\Val \ec{z_1} & = & \mu \Var{\ec{z_1}}. & \const{C}( & \mu \Var{\ec{z_2}}. & \const{A}( \Var{\ec{z_1}} ) ) &
-\Val \ec{w_3} & = & \Var{\ec{w_3}} \\
-\Val \ec{z_2} & = & \mu \Var{\ec{z_2}}. & \const{A}( & \mu \Var{\ec{z_1}}. & \const{C}( \Var{\ec{z_2}} ) ) & 
-\Val \ec{x} & = & \Var{\ec{x}} \\
+\Val \ec{y_1} & = & \mu {\Varec{y_1}}. & \const{A}( & \mu {\Varec{y_2}}. & \const{B}( {\Varec{x}} ) ) &
+\Val \ec{w_1} & = & \mu {\Varec{w_1}}. & \const{A}( {\Varec{x}} )      & \\ 
+\Val \ec{y_2} & = & \mu {\Varec{y_2}}. & \multicolumn{2}{l}{\const{B}( {\Varec{x}} )}      & &
+\Val \ec{w_2} & = & \mu {\Varec{w_2}}. & \const{C}( {\Varec{w_3}} ) \\
+\Val \ec{z_1} & = & \mu {\Varec{z_1}}. & \const{C}( & \mu {\Varec{z_2}}. & \const{A}( {\Varec{z_1}} ) ) &
+\Val \ec{w_3} & = & {\Varec{w_3}} \\
+\Val \ec{z_2} & = & \mu {\Varec{z_2}}. & \const{A}( & \mu {\Varec{z_1}}. & \const{C}( {\Varec{z_2}} ) ) & 
+\Val \ec{x} & = & {\Varec{x}} \\
 \end{array}\]
 
-To construct a completion of $\Val$, we must choose values for the free variables of $\Val$, $\Var{\ec{x}}$ and $\Var{\ec{w_3}}$.
-The content of the set $\mathcal{T}^\M_{\Var{\ec{x}}}( \Val )$ is $\alpha$-equivalent to:
+To construct a completion of $\Val$, we must choose values for the free variables of $\Val$, ${\Varec{x}}$ and ${\Varec{w_3}}$.
+The content of the set $\mathcal{T}^\M_{{\Varec{x}}}( \Val )$ is $\alpha$-equivalent to:
 \[%\begin{array}{c}
@@
 for instance, $x$ is interpretted as $\Val \ec{x}$, $y_1$ as $\Val \ec{y_1}$ and so on.
-Assigning a value for $\Var{\ec{x}}$ that is $\alpha$-equivalent to a value in $\mathcal{T}^\M_{\Var{\ec{x}}}( \Val )$
+Assigning a value for ${\Varec{x}}$ that is $\alpha$-equivalent to a value in $\mathcal{T}^\M_{{\Varec{x}}}( \Val )$
 may cause values in the range of $\Val$ to become $\alpha$-equivalent, 
 which in turn may cause $\M$ to be inconsistent.
-For instance, say we assign $\mu \Var{\ec{x}}. \const{B}( \Var{\ec{x}} )$ for $\Var{\ec{x}}$.
-After this substitution, $\Val \ec{y_2}$ is $\mu \Var{\ec{y_2}}. \const{B}( \mu \Var{\ec{x}}. \const{B}( \Var{\ec{x}} ) )$, 
-which has normal form $\mu \Var{\ec{y_2}}. \const{B}( \Var{\ec{y_2}} )$,
-which is $\alpha$-equivalent to $\mu \Var{\ec{x}}. \const{B}( \Var{\ec{x}} )$.
+For instance, say we assign $\mu {\Varec{x}}. \const{B}( {\Varec{x}} )$ for ${\Varec{x}}$.
+After this substitution, $\Val \ec{y_2}$ is $\mu {\Varec{y_2}}. \const{B}( \mu {\Varec{x}}. \const{B}( {\Varec{x}} ) )$, 
+which has normal form $\mu {\Varec{y_2}}. \const{B}( {\Varec{y_2}} )$,
+which is $\alpha$-equivalent to $\mu {\Varec{x}}. \const{B}( {\Varec{x}} )$.
 However, this contradicts the disequality $x \not\teq y_2$ in $\Ec$.
-Inversely, if the value assigned to $\Var{\ec{x}}$ is not $\alpha$-equivalent to any term in $\mathcal{T}^\M_{\Var{\ec{x}}}( \Val )$,
+Inversely, if the value assigned to ${\Varec{x}}$ is not $\alpha$-equivalent to any term in $\mathcal{T}^\M_{{\Varec{x}}}( \Val )$,
 then the values in the range of $\Val$ remain $\alpha$-disequivalent.
 For instance, we may assign values such as 
-$\mu \Var{\ec{x}}. \const{C}( \Var{\ec{x}} )$,
-$\mu \Var{\ec{x}}. \const{C}( \mu x_1. \const{A}( x_1) )$, and
-$\mu \Var{\ec{x}}. \const{B}( \mu x_1. \const{A}( \Var{\ec{x}} ) )$
-to $\Var{\ec{x}}$.
-Notice that legal substitutions for $\Var{\ec{x}}$ may cause the range of $\Val$ to contain non-normal terms,
+$\mu {\Varec{x}}. \const{C}( {\Varec{x}} )$,
+$\mu {\Varec{x}}. \const{C}( \mu x_1. \const{A}( x_1) )$, and
+$\mu {\Varec{x}}. \const{B}( \mu x_1. \const{A}( {\Varec{x}} ) )$
+to ${\Varec{x}}$.
+Notice that legal substitutions for ${\Varec{x}}$ may cause the range of $\Val$ to contain non-normal terms,
 as in after applying the third substitution mentioned above,
-$\Val \ec{w_1}$ is $\mu \Var{\ec{w_1}}. \const{A}( \mu \Var{\ec{x}}. \const{B}( \mu x_1. \const{A}( \Var{\ec{x}} ) ) )$.
+$\Val \ec{w_1}$ is $\mu {\Varec{w_1}}. \const{A}( \mu {\Varec{x}}. \const{B}( \mu x_1. \const{A}( {\Varec{x}} ) ) )$.
 $\Box$
@@
 If $\const{C}( t_1, \ldots, t_n ) \in \ec{x}$,
-then $\Val^\ast \ec{x} = \mu \Var{\ec{x}}. \const{C}( w_1, \ldots, w_n )$
+then $\Val^\ast \ec{x} = \mu {\Varec{x}}. \const{C}( w_1, \ldots, w_n )$
 and $\Val^\ast \ec{x} \interp{ w_i } = \Val^\ast \ec{t_i}$ for $i = 1, \ldots, n$.
-%and $\muvar( w_i ) = \Var{\ec{t_i}}$ for $i = 1, \ldots, n$,
+%and $\muvar( w_i ) = {\Varec{t_i}}$ for $i = 1, \ldots, n$,
 %\item
 %$t_1 \interp{ t_2 } = \Val^\ast \ec{x}$
-%for all $t_1$ with subterm $t_2$ in the range of $\Val^\ast$ where $\muvar(t_2)=\Var{\ec{x}}$,
+%for all $t_1$ with subterm $t_2$ in the range of $\Val^\ast$ where $\muvar(t_2)={\Varec{x}}$,
 \item 
@@
 To show 1, we first show that $\Val^\ast$ contains no free variables.
-Assume by contradiction that $\Val^\ast$ contained a free variable $\Var{\ec{y}}$ for some $\ec{y}$ of type $\tau$.
+Assume by contradiction that $\Val^\ast$ contained a free variable ${\Varec{y}}$ for some $\ec{y}$ of type $\tau$.
 Then it must be the case that $\ec{y}$ does not contain a constructor term,
-or else $\Var{\ec{y}}$ would not occur as a free variable in $\Val$.
+or else ${\Varec{y}}$ would not occur as a free variable in $\Val$.
 Consider the case when $\tau$ is finite.
@@
 and thus there are no terms in $\mathcal{T}^\M_x(\Val^\ast)$ of type $\tau$.
-This is a contradiction, since our model completion may assign the value in the domain of $\tau$ to $\Var{\ec{x}}$.
+This is a contradiction, since our model completion may assign the value in the domain of $\tau$ to ${\Varec{x}}$.
 Now, consider the case when $\tau$ is infinite.
@@
 since there are only a finite number of closed terms in $\mathcal{T}^\M_x(\Val^\ast)$,
-and thus our model completion can assign a value not occurring in $\mathcal{T}^\M_x(\Val^\ast)$ to $\Var{\ec{y}}$.
+and thus our model completion can assign a value not occurring in $\mathcal{T}^\M_x(\Val^\ast)$ to ${\Varec{y}}$.
 By construction, $\Val^\ast \ec{x}$ is normal.
@@
 $\Val$ was constructed by applying a sequence of substitutions
-that replace $\Var{\ec{y}}$ by $\mu \Var{\ec{y}}. \const{C}_i( \Var{\ec{t_1}}, \ldots, \Var{\ec{t_n}} )$.
+that replace ${\Varec{y}}$ by $\mu {\Varec{y}}. \const{C}_i( {\Varec{t_1}}, \ldots, {\Varec{t_n}} )$.
 Now, say that $\const{C}( t_1, \ldots, t_n )$ occurs in some equivalence class $\ec{x}$.
-Then $\Val \ec{x}$ must be of the form $\mu \Var{\ec{x}}. \const{C}_i( w_1, \ldots, w_n )$,
+Then $\Val \ec{x}$ must be of the form $\mu {\Varec{x}}. \const{C}_i( w_1, \ldots, w_n )$,
 where for each $i=1,\ldots,n$, $w_i$ was obtained by applying a sequence of substitutions
-of the form mentioned above to $\Var{\ec{t_i}}$.
-%Each $w_i$ is either $\Var{\ec{x}}$,
-%a variable $\Var{\ec{y}}$ where $\ec{y}$ does not contain a constructor,
-%or a term of the form $\mu \Var{\ec{y}}. \const{C}_j( \vec u )$.
+of the form mentioned above to ${\Varec{t_i}}$.
+%Each $w_i$ is either ${\Varec{x}}$,
+%a variable ${\Varec{y}}$ where $\ec{y}$ does not contain a constructor,
+%or a term of the form $\mu {\Varec{y}}. \const{C}_j( \vec u )$.
 While interpreting $w_i$ in $\Val \ec{x}$,
-all free occurrences of $\Var{\ec{x}}$ in $\Val \ec{x} \interp{w_i}$ are replaced by $\Val \ec{x}$. \rem{better formalize.}
-Thus, since $\Val \ec{x} \interp{w_i}$ and $\Val \ec{t_i}$ are obtained by applying the same sequence of substitutions to $\Var{\ec{t_i}}$,
+all free occurrences of ${\Varec{x}}$ in $\Val \ec{x} \interp{w_i}$ are replaced by $\Val \ec{x}$. \rem{better formalize.}
+Thus, since $\Val \ec{x} \interp{w_i}$ and $\Val \ec{t_i}$ are obtained by applying the same sequence of substitutions to ${\Varec{t_i}}$,
 we have $\Val \ec{w_i} = \Val \ec{x} \interp{ w_i }$.
@@
 %Since \rn{Clash} does not apply to $\Fc$, then $\Val^\ast \ec{t}$ must be a term
-$t_w$ of the form $\mu \Var{\ec{t}}. \const{C}( w_1, \ldots, w_n )$
+$t_w$ of the form $\mu {\Varec{t}}. \const{C}( w_1, \ldots, w_n )$
 where $t_w\interp{w_i} = \Val^\ast \ec{u_i}$ for $i = 1, \ldots, n$.
@@
 By Lemma~\ref{lem:mu-norm-arg}, $t_u\interp{u_i'} \vsim \Val^\ast \ec{u_i}$.
-%Say that $\muvar( w_i ) = \Var{\ec{y_i}}$.
+%Say that $\muvar( w_i ) = {\Varec{y_i}}$.
 %Due to our construction of $\Val^\ast$ and since \rn{Inject} does not apply to $\Fc$, we have $\ec{u_i} = \ec{y_i}$,
@@
 By Lemma~\ref{lem:model-completion}.2, $\Val^\ast \ec{u}$ is a term $t_w$ of the form
-$\mu \Var{\ec{u}}. C_i( w_1, \ldots, w_n )$,
+$\mu {\Varec{u}}. C_i( w_1, \ldots, w_n )$,
 where $t_w \interp{ w_j } = \Val^\ast \ec{ s^j_k( u )}$. 
 %By our construction of $\Val^\ast$ and since \rn{Inject} and \rn{Clash} do not apply to $\Fc$, 
-%$\Val^\ast \ec{ u }$ must be of the form $t_u = \mu \Var{\ec{u}}. C_i( u_1, \ldots, u_n )$,
-%where $\muvar( u_1 ) = \Var{\ec{s^1_i(u)}}$, $\ldots$, $\muvar(u_n) = \Var{\ec{s^n_i( u )}}$,
+%$\Val^\ast \ec{ u }$ must be of the form $t_u = \mu {\Varec{u}}. C_i( u_1, \ldots, u_n )$,
+%where $\muvar( u_1 ) = {\Varec{s^1_i(u)}}$, $\ldots$, $\muvar(u_n) = {\Varec{s^n_i( u )}}$,
 %and $\M( u )$ is of the form $t_w = \mu x. C_i( w_1, \ldots, w_n )$,
@@
 a set of equalities from $M_i$ that $\thD$-entail that the values of $\ec{x}$ in models of $\Ec$ are of the form $\Val \ec{x}$.
-For instance, if $x \teq \const{C}( x ) \in M_i$, then $x \teq \const{C}( x )$ is a (possible) explanation for $\Val \ec{x} = \mu \Var{\ec{x}}. \const{C}( \Var{\ec{x}} )$.
+For instance, if $x \teq \const{C}( x ) \in M_i$, then $x \teq \const{C}( x )$ is a (possible) explanation for $\Val \ec{x} = \mu {\Varec{x}}. \const{C}( {\Varec{x}} )$.
 Thus, the rules of the calculus are implemented as follows.

### 2014-12-19 tex/codatadp.tex
theory def, cleanup.

--- 2014-12-19 tex/codatadp.tex
@@
 
-TODO: Andy, introduce theory here. There were some ideas below in the comments
-you might want to mention either here or in Section 3. I'm not sure what's the
-best way to justify it though, but for the conference at least we can just claim it. (we
-won't have the place for a proof anyway).
+%TODO: Andy, introduce theory here. There were some ideas below in the comments
+%you might want to mention either here or in Section 3. I'm not sure what's the
+%best way to justify it though, but for the conference at least we can just claim it. (we
+%won't have the place for a proof anyway).
+
+Given a signature $\Sig$, we write $\thD$ to refer to the \emph{theory} of (co-)datatypes,
+which in addition to $\Sig$ defines a class of $\Sig$-interpretations $\mathbf{I}$,
+namely the ones that satisfy the axioms mentioned in this section.
+In particular, this means that each interpretation in $\mathbf{I}$ has a fixed interpretation
+for constructor terms and correctly-applied selector terms, but may differ on its interpretation for 
+wrongly-applied selector terms and variables.
+A formula $\varphi$ is $\thD$-satisfiable if there is an interpretation in $\mathbf{I}$ that satisfies $\varphi$,
+and $\thD$-unsatisfiable otherwise.
+\rem{more?}
 
@@
 \afterDot
-If there exists a closed derivation tree with root node $\Ec$, then $\Ec$ is unsatisfiable in $\thD$.
+If there exists a closed derivation tree with root node $\Ec$, then $\Ec$ is $\thD$-unsatisfiable.
 \end{lemma}
@@
 First, if the derivation tree is an application of \rn{Conflict}, \rn{Clash}, or \rn{Cyclic},
-then $\Ec$ is unsatisfiable in $\thD$.
+then $\Ec$ is $\thD$-unsatisfiable.
 For \rn{Conflict}, this is a consequence of equality reasoning.
@@
 Since by exhaustiveness, all models of $\thD$ entail (exactly) one equality $t \teq C_i( s^1_i( t ), \ldots, s^{a_i}_i( t ) )$,
-we know $\Ec$ is unsatisfiable in $\thD$.
+we know $\Ec$ is $\thD$-unsatisfiable.
 Otherwise, the child of $\Ec$ is a closed derivation tree 
@@
 For \rn{Singleton}, clearly $t_1$ and $t_2$ must have the same value since the cardinality of their type is one.
-By the induction hypothesis, we have $\Ec \cup t_1 \teq t_2$ is unsatisfiable in $\thD$, 
-and thus $\Ec$ is unsatisfiable in $\thD$.
+By the induction hypothesis, we have $\Ec \cup t_1 \teq t_2$ is $\thD$-unsatisfiable, 
+and thus $\Ec$ is $\thD$-unsatisfiable.
 \qed
@@
 such that $v \vsimv{x} t_1\interp{t_2}$ for some term $t_1$ with subterm $t_2$ occurring in the range of $\Val$.
-We first construct a mapping $\Val'$ by exhaustively applying the following update to $\Val$:
+We construct $\Val^\ast$ by exhaustively applying the following update to $\Val$:
 
@@
 }{
-  \Val := \Val \{ x \mapsto \mu x. t \}
+  \Val := \nf{(\Val \{ x \mapsto \mu x. t \})}
 }
@@
 it can only be applied a finite number of times.
-We then let $\Val^\ast \ec{x}$ be $\nf{(\Val' \ec{x})}$ for all $\ec{x}$.
+After assigning this value, we normalize all terms in the range of $\Val$.
+%We then let $\Val^\ast \ec{x}$ be $\nf{(\Val' \ec{x})}$ for all $\ec{x}$.
 
@@
 Notice that legal substitutions for $v_{\ec{x}}$ may cause the range of $\Val$ to contain non-normal terms,
-as in after applying the third substitution mentioned above,
-$\Val \ec{w_1}$ is $\mu v_{\ec{w_1}}. \const{A}( \mu v_{\ec{x}}. \const{B}( \const{A}( v_{\ec{x}} ) ) )$.
+as in after applying the substitution for the third term mentioned above,
+$\Val \ec{w_1}$ is $\mu v_{\ec{w_1}}. \const{A}( \mu v_{\ec{x}}. \const{B}( \const{A}( v_{\ec{x}} ) ) )$,
+which has normal form
+$\mu v_{\ec{w_1}}. \const{A}( \mu v_{\ec{x}}. \const{B}( v_{\ec{w_1}} ) )$.
 $\Box$
@@
 \begin{enumerate}
-\item $\Val^\ast \ec{x}$ is $\alpha$-equivalent to a value in $\M( \tau )$, for all $\ec{x}$ of type $\tau$ of $\Ec$,
+\item $\Val^\ast \ec{x}$ is $\alpha$-equivalent to a value in $\M( \tau )$, where $\ec{x}$ has type $\tau$,
 \item
@@
 \afterDot
-If there exists a derivation tree with root node $\Ec$ containing a saturated node, then $\Ec$ is satisfiable in $\thD$.
+If there exists a derivation tree with root node $\Ec$ containing a saturated node, then $\Ec$ is $\thD$-satisfiable.
 \end{lemma}
@@
 that $\M( t ) \aequiv \Val^\ast \ec{t}$ for all terms $t \in \mathcal{T}( \Fc )$,
-which since $\M$ is normal implies $\M \models t_1 \teq t_2$.
+which since $\M$ is normal, implies $\M \models t_1 \teq t_2$.
 
@@
 Since \rn{Conflict} does not apply, $t_1 \teq t_2 \not\in \Fc$ and thus $\ec{t_1}$ and $\ec{t_2}$ are distinct.
-Since $\M( t_1 ) \vsim \Val^\ast \ec{t_1}$, $\M( t_2 ) \vsim \Val^\ast \ec{t_2}$
-and Lemma~\ref{lem:model-completion}.3, $\M( t_1 ) \neq \M( t_2 )$, and thus $\M \models t_1 \tneq t_2$.
+Since $\M( t_1 ) \vsim \Val^\ast \ec{t_1}$ and $\M( t_2 ) \vsim \Val^\ast \ec{t_2}$,
+by Lemma~\ref{lem:model-completion}.3, $\M( t_1 ) \neq \M( t_2 )$, and thus $\M \models t_1 \tneq t_2$.
 
@@
 Each $T_i$ solver either 
-reports that a subset $C$ of $M_i$ is $T_i$-unsatisfiable in which case $\neg C$ is added as a additional clause to $F$,
+reports that a subset $C$ of $M_i$ is $T_i$-unsatisfiable, in which case $\neg C$ is added as a additional clause to $F$,
 adds an additional clause $C$ to $F$,
@@
 Assume $\Ec$ is initially the set $M_i$ described above.
-For each equality added to $\Ec$, we associate a set of equalities from $M_i$ that together $\thD$-entail $t \teq s$,
+For each equality added to $\Ec$, we associate a set of equalities from $M_i$ that together entail $t \teq s$,
 which we call its \emph{explanation}.
 Similarly, each $\Val \ec{x}$ is also associated an explanation, that is, 
-a set of equalities from $M_i$ that $\thD$-entail that the values of $\ec{x}$ in models of $\Ec$ are of the form $\Val \ec{x}$.
+a set of equalities from $M_i$ that entail that the values of $\ec{x}$ in models of $\Ec$ are of the form $\Val \ec{x}$.
 For instance, if $x \teq \const{C}( x ) \in M_i$, then $x \teq \const{C}( x )$ is a (possible) explanation for $\Val \ec{x} = \mu v_{\ec{x}}. \const{C}( v_{\ec{x}} )$.

### 2014-12-19 tex/codatadp.tex
more edits

--- 2014-12-19 tex/codatadp.tex
@@
 \global\def\figurename{Figure}
+\global\def\figuresname{Figures}
 
@@
 
-The calculus consists of three sets of rules (Figures \ref{fig:cc-rules} to
+The calculus consists of three sets of rules (\figuresname~\ref{fig:cc-rules} to
 \ref{fig:split-rule}) corresponding to three successive phases. The first
@@
 \inferrule{
-  \tau \in \Data
+  \delta \in \Data
   \quad
-  t : \tau
+  t : \delta
   \quad
@@
 \inferrule{
- \tau \in \Codata
+ \delta \in \Codata
  \quad
- t, u : \tau
+ t, u : \delta
  \quad
@@
 where $\const{C}(z,x) \in \ec{y}$ and $\const{C}(w,y) \in \ec{x}$ for some $\const{C} \in \Ctr$.
-A possible sequence of \rn{Update} rule applications is presented below, omitting
+A possible sequence of \rn{Update} rule applications is given below, omitting
 trivial entries such as $\ec{w} \mapsto {\Varec{w}}$.
@@
 \item \noindent\rlap{Update w.r.t.\ $\ec{y}$:}\phantom{Update w.r.t.\ $\ec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \mu {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \mu {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\;
-  \ec{y} \mapsto \mu {\Varec{y}}. \const{C}( {\Varec{z}},\, {\Varec{x}} ) \vthinspace\}$
+  \ec{y} \mapsto \mu {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) \vthinspace\}$
 \item \noindent\rlap{Update w.r.t.\ $\ec{x}$:}\phantom{Update w.r.t.\ $\ec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \mu {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \mu {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\;
@@
 This indicates that the values for $x$ and $y$ in models of $\Ec$
-are of the form $\const{C}( {\Varec{w}}, \const{C}( {\Varec{z}}, \ldots ))$
-and $\const{C}( {\Varec{z}}, \const{C}( {\Varec{w}}, \ldots ))$,
-respectively, for some value of ${\Varec{w}}$ and ${\Varec{z}}$.
+are of the forms $\const{C}( {\Varec{w}}, \const{C}( {\Varec{z}}, \const{C}( {\Varec{w}},\allowbreak \const{C}( {\Varec{z}},\ldots ))))$
+and $\const{C}( {\Varec{z}}, \const{C}( {\Varec{w}}, \const{C}( {\Varec{z}}, \const{C}( {\Varec{w}}, \ldots ))))$,
+respectively. %, for some values of ${\Varec{w}}$ and ${\Varec{z}}$.
 \end{example}
 
-Given $\Val$, the rules for acyclicity and bisimilarity (Figure~\ref{fig:ab-rules}) can be stated as follows.
+Given $\Val$, the rules for acyclicity and bisimilarity can be stated as follows.
 For acyclicity, if there is an equivalence class $\ec{t}$ of datatype type whose class of values $\Val \ec{t} = \mu x.\; u$ are cyclic,
@@
 and let $\Ec$ be the set $\{ x \teq \const{C}(y,\const{C}(y,x)) \}$.
-After applying the rules in Figure~\ref{fig:cc-rules} to completion, the equivalence classes of $\Ec$ are
+After phase~1, the equivalence classes of $\Ec$ are
 $\{ x, \const{C}(y,\const{C}(y,x)) \}$, $\{ \const{C}(y,x) \}$, and $\{ y \}$.
@@
 \paragraph{Phase 3: Branching.}
-If neither a rule from Figure \ref{fig:cc-rules}~or~\ref{fig:ab-rules} applies to $\Ec$, 
-the rules from Figure~\ref{fig:split-rule} are applied.
-The \rn{Singleton} rule handles the case when there are two terms from $\tEc$ having a co-recursive singleton type,
-in which case they must be equivalent.
+In the third phase,
+the \rn{Singleton} rule handles the case when there are two terms from $\tEc$
+having a co-recursive singleton type, in which case they must be equivalent.
 Otherwise, in the case there is a selector term applied to $t$, or if $t$ has a finite datatype type, 

### 2014-12-19 tex/codatadp.tex

@@
 %We will commonly denote tuples of terms $( t_1, \ldots, t_n )$ in bold font, as $\vec t$.
-We assume that all $\Sig$-terms in $\Ec$ are \emph{normalized}, meaning that all subterms of the form $\const{s}^{\,k}_j( \const{C}_j(t_1,\ldots,t_n) )$
-are simplified to $t_k$. All constraints added to $\Ec$ are also implicitly
-normalized in this manner.
+%We assume that all $\Sig$-terms in $\Ec$ are \emph{normalized}, meaning that all subterms of the form $\const{s}^{\,k}_j( \const{C}_j(t_1,\ldots,t_n) )$
+%are simplified to $t_k$. All constraints added to $\Ec$ are also implicitly
+%normalized in this manner.
 
@@
 further) before the next phase can begin.
+
+%We assume that all $\Sig$-terms in $\Ec$ are \emph{normalized}, meaning that all subterms of the form $\const{s}^{\,k}_j( \const{C}_j(t_1,\ldots,t_n) )$
+%are simplified to $t_i$, and moreover assume that all %additional
+%constraints added to $\Ec$ are normalized in this manner.
+%We present the calculus in three steps.
+%In the first step, we compute the bidirectional closure of $\Ec$;
+%in the second step, we make inferences based on cyclicity and bisimilarity;
+%and in the third step, %(when necessary),
+%we branch on constructor types for various terms in $\tEc$.
+%Following the conventions from~\cite{}, 
+%the derivation rules of our calculus are given in \emph{guarded assignment form},
+%where a rule can be applied to $\Ec$ if it meets all of the specified preconditions for $\Ec$.
+%The conclusion of a rule either describes equalities to be added to $\Ec$ (in which case we call an application of it \emph{nonterminal}),
+%or is $\bot$ (in which case we call an application of it \emph{terminal}).
+%A rule may have multiple conclusions separated by $\parallel$, which denotes nondeterministic branching.
 
@@
 respectively. %, for some values of ${\Varec{w}}$ and ${\Varec{z}}$.
+$\Box$
 \end{example}
@@
 are equivalent for any value of ${\Varec{y}}$.
-\qed
+$\Box$
 \end{example}
@@
 First, if the derivation tree is an application of \rn{Conflict}, \rn{Clash}, or \rn{Acyclic},
-then $\Ec$ is unsatisfiable in $\thD$.
+then $\Ec$ is $\thD$-unsatisfiable.
 For \rn{Conflict}, this is a consequence of equality reasoning.
@@
 we write $\mathcal{T}^\M_x( \Val )$ to denote the set consisting of all values $v$ from $\M( \tau )$
-such that $v \vsim_x t_1\interp{t_2}$ for some term $t_1$ with subterm $t_2$ occurring in the range of $\Val$.
-We first construct a mapping $\Val'$ by exhaustively applying the following update to $\Val$:
+such that $v \vsimv{x} t_1\interp{t_2}$ for some term $t_1$ with subterm $t_2$ occurring in the range of $\Val$.
+We construct $\Val^\ast$ by exhaustively applying the following update to $\Val$:
 
@@
   \quad
-  (\mu x. t) \sim v
+  (\mu x. t) \aequiv v
   \quad
@@
 In this update, we choose an unassigned variable in $\Val$
-and assign it a value that is not bisimilar to one occurring in $\mathcal{T}^\M_x(\Val)$.
-Since this update removes one variable from the set $\FV( \Val )$ and does not add any additional variables to $\FV( \Val )$, 
+and assign it a value that is not $\alpha$-equivalent to one occurring in $\mathcal{T}^\M_x(\Val)$.
+Since this update removes one variable from the set $FV( \Val )$ and does not add any additional variables to $\FV( \Val )$, 
 it can only be applied a finite number of times.
-We then let $\Val^\ast \ec{x}$ be $\nf{(\Val' \ec{x})}$ for all $\ec{x}$.
+After assigning this value, we normalize all terms in the range of $\Val$.
+%We then let $\Val^\ast \ec{x}$ be $\nf{(\Val' \ec{x})}$ for all $\ec{x}$.
 
@@
 $\mu {\Varec{x}}. \const{C}( {\Varec{x}} )$,
-$\mu {\Varec{x}}. \const{C}( \mu x_1. \const{A}( x_1) )$, and
-$\mu {\Varec{x}}. \const{B}( \mu x_1. \const{A}( {\Varec{x}} ) )$
+$\mu {\Varec{x}}. \const{C}( \mu x_1. \const{A}( x_1) )$, or
+$\mu {\Varec{x}}. \const{B}( \const{A}( {\Varec{x}} ) )$
 to ${\Varec{x}}$.
 Notice that legal substitutions for ${\Varec{x}}$ may cause the range of $\Val$ to contain non-normal terms,
-as in after applying the third substitution mentioned above,
-$\Val \ec{w_1}$ is $\mu {\Varec{w_1}}. \const{A}( \mu {\Varec{x}}. \const{B}( \mu x_1. \const{A}( {\Varec{x}} ) ) )$.
+as in after applying the substitution for the third term mentioned above,
+$\Val \ec{w_1}$ is $\mu {\Varec{w_1}}. \const{A}( \mu {\Varec{x}}. \const{B}( \const{A}( {\Varec{x}} ) ) )$,
+which has normal form
+$\mu {\Varec{w_1}}. \const{A}( \mu {\Varec{x}}. \const{B}( {\Varec{w_1}} ) )$.
 $\Box$
@@
 \begin{enumerate}
-\item $\Val^\ast \ec{x}$ is bisimilar with a value in $\M( \tau )$, for all $\ec{x}$ of type $\tau$ of $\Ec$,
+\item $\Val^\ast \ec{x}$ is $\alpha$-equivalent to a value in $\M( \tau )$, where $\ec{x}$ has type $\tau$,
 \item
@@
 that interprets wrongly-applied selectors based on equality information in $\Fc$,
-and interprets the variables of $\Fc$ based on the completion $\Val^\ast$ of the mapping $\Val$ from the second phase of our calculus.
-For the latter, let $\M( x )$ be the value in $\M$ that is bisimilar with $\Val^\ast \ec{x}$ for each variable $x$ in $\mathcal{T}(\Fc)$.
+and interprets the variables of $\Fc$ based on the completion $\Val^\ast$ of the mapping $\Val$ from phase 2.
+For the latter, let $\M( x )$ be the value in $\M$ that is $\alpha$-equivalent with $\Val^\ast \ec{x}$ for each variable $x$ in $\mathcal{T}(\Fc)$.
 
@@
 by structural induction on $t$,
-that $\M( t ) \sim \Val^\ast \ec{t}$ for all terms $t \in \mathcal{T}( \Fc )$,
-which implies $\M \models t_1 \teq t_2$.
+that $\M( t ) \aequiv \Val^\ast \ec{t}$ for all terms $t \in \mathcal{T}( \Fc )$,
+which since $\M$ is normal, implies $\M \models t_1 \teq t_2$.
 
@@
 Assume $\Ec$ is initially the set $M_i$ described above.
-For each equality added to $\Ec$, we associate a set of equalities from $M_i$ that together $\thD$-entail $t \teq s$,
+For each equality added to $\Ec$, we associate a set of equalities from $M_i$ that together entail $t \teq s$,
 which we call its \emph{explanation}.
 Similarly, each $\Val \ec{x}$ is also associated an explanation, that is, 
-a set of equalities from $M_i$ that $\thD$-entail that the values of $\ec{x}$ in models of $\Ec$ are of the form $\Val \ec{x}$.
+a set of equalities from $M_i$ that entail that the values of $\ec{x}$ in models of $\Ec$ are of the form $\Val \ec{x}$.
 For instance, if $x \teq \const{C}( x ) \in M_i$, then $x \teq \const{C}( x )$ is a (possible) explanation for $\Val \ec{x} = \mu {\Varec{x}}. \const{C}( {\Varec{x}} )$.
@@
 All other rules add equalities to the internal state of the theory solver.
-
-Phase 1 (along with the optimization for collapsing selectors) and the \rn{Singleton} rule are performed eagerly, 
-that is for partial satisfying assignments $M$, while phase~2 and the \rn{Split} rule from phase~3 are performed only for complete satisfying assignments $M$, and in that order.
+The rules in phase 1 (along with the optimization for collapsing selectors) and the \rn{Singleton} rule are performed eagerly, 
+that is for partial satisfying assignments $M$, while the rules in phase 2 and the \rn{Split} rule are performed only for complete satisfying assignments $M$, 
+and in that order.
 

### 2014-12-19 tex/codatadp.tex
minor

@@
 it remains only to define how $\M$ interprets wrong-applied selector terms and variables.
-For the latter, this will be based on the mapping $\Val$ constructed in the second phase of our calculus.
+For the latter, this will be based on the mapping $\Val$ constructed in phase~2 of the calculus.
 We compute the \emph{completion} of $\Val$ for normal model $\M$, which we write as $\Val^\ast$,
@@
 that interprets wrongly-applied selectors based on equality information in $\Fc$,
-and interprets the variables of $\Fc$ based on the completion $\Val^\ast$ of the mapping $\Val$ from phase 2.
+and interprets the variables of $\Fc$ based on the completion $\Val^\ast$ of the mapping $\Val$ from phase~2.
 For the latter, let $\M( x )$ be the value in $\M$ that is $\alpha$-equivalent with $\Val^\ast \ec{x}$ for each variable $x$ in $\mathcal{T}(\Fc)$.
@@
 All other rules add equalities to the internal state of the theory solver.
-The rules in phase 1 (along with the optimization for collapsing selectors) and the \rn{Singleton} rule are performed eagerly, 
-that is for partial satisfying assignments $M$, while the rules in phase 2 and the \rn{Split} rule are performed only for complete satisfying assignments $M$, 
+The rules in phase~1 (along with the optimization for collapsing selectors) and the \rn{Singleton} rule are performed eagerly, 
+that is for partial satisfying assignments $M$, while the rules in phase~2 and the \rn{Split} rule are performed only for complete satisfying assignments $M$, 
 and in that order.

### 2014-12-19 tex/codatadp.tex
whitespace
more

@@
 datatypes and codatatypes. The dual of the acyclicity rule for datatypes is a
-bisimilarity rule that identifies equal values, also in the presence of cyclic
+uniqueness rule that identifies bisimilar values, also in the presence of cyclic
 ($\omega$-regular) data. The procedure decides ground and universal
@@
 
-Second, \emph{a bisimilarity rule takes the place of the acyclicity rule of datatypes.}
+Second, \emph{a uniqueness rule takes the place of the acyclicity rule of datatypes.}
 Cyclic constraints such as
@@
 are unsatisfiable for datatypes but satisfiable for codatatypes.
-The bisimilarity principle states that two values having the same (possibly
+The uniqueness principle states that two values having the same (possibly
 infinite) expansion must be equal; from $\const{x} \teq \const{S}(\const{y})$ and
 $\const{y} \teq \const{S}(\const{x})$, it deduces $\const{x} \teq \const{y}$.
-The acyclicity and bisimilarity rules are needed to ensure completeness on
+The acyclicity and uniqueness rules are needed to ensure completeness on
 ground problems and the absence of spurious models. They cannot be replaced by
 finite axiomatizations, so they naturally belong in a decision procedure.
-%    * in particular, acyclicity and bisimilarity are necessary for some proofs
+%    * in particular, acyclicity and uniqueness are necessary for some proofs
 %%      (and cannot be axiomatized finitely)
@@
 a variable it contains \cite{xxx}. Dually, coinduction can be replaced by the
-bisimilarity axiom schema, which asserts that (co)recursive definitions giving
+uniqueness axiom schema, which asserts that (co)recursive definitions giving
 rise to the same infinite expansion---e.g., $x \teq \const{C}(x)$ and $y \teq
-\const{C}(\const{C}(y))$---must be identical.
+\const{C}(\const{C}(y))$---must be equal.
 
@@
 %      * when looking at the ground theory, enough to
-%        consider bisimilarity -- no way to express that certain infinite
+%        consider uniqueness -- no way to express that certain infinite
 %        acyclic objects do not exist
@@
 phase computes the bidirectional closure of $\Ec$. The second phase makes
-inferences based on acyclicity (for datatypes) and bisimilarity (for
+inferences based on acyclicity (for datatypes) and uniqueness (for
 codatatypes). The third phase performs case splitting on constructors for
@@
 %In the first step, we compute the bidirectional closure of $\Ec$;
-%in the second step, we make inferences based on cyclicity and bisimilarity;
+%in the second step, we make inferences based on cyclicity and uniqueness;
 %and in the third step, %(when necessary),
@@
 \)
-\rn{Bisimilar}
+\rn{Unique}
 \end{tabular}
-\caption{\,Derivation rules for acyclicity and bisimilarity%.
+\caption{\,Derivation rules for acyclicity and uniqueness%.
 }
@@
 
-\paragraph{Phase 2: Applying Acyclicity and Bisimilarity.}
+\paragraph{Phase 2: Applying Acyclicity and Uniqueness.}
 %For presentation of the rules in this phase, we rely on a representation of terms in the $\mu$-notation.
@@
 
-Given $\Val$, the rules for acyclicity and bisimilarity can be stated as follows.
+Given $\Val$, the rules for acyclicity and uniqueness can be stated as follows.
 For acyclicity, if there is an equivalence class $\ec{t}$ of datatype type whose class of values $\Val \ec{t} = \mu x.\; u$ are cyclic,
@@
 then we can conclude that $\Ec$ is unsatisfiable.
-For bisimilarity, if there are two equivalence classes $\ec{t}$ and $\ec{u}$ 
+For uniqueness, if there are two equivalence classes $\ec{t}$ and $\ec{u}$
 of codatatype type whose classes of values $\Val \ec{t}$ and $\Val \ec{u}$ are equivalent for a capture-avoiding renaming of $\mu$-bound variables,
 then we can conclude $t \teq u$.
-We demonstrate the bisimilarity rule concretely with the following example.
+We demonstrate the uniqueness rule concretely with the following example.
 
@@
 Otherwise, the child of $\Ec$ is a closed derivation tree
-whose root node is $\Ec \cup t_1 \teq t_2$ obtained by applying one of the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, \rn{Inject}, \rn{Bisimilar}, or \rn{Singleton}.
+whose root node is $\Ec \cup t_1 \teq t_2$ obtained by applying one of the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, \rn{Inject}, \rn{Unique}, or \rn{Singleton}.
 In all cases, we have that $\Ec \models_{\thD} t_1 \teq t_2$.
@@
 For \rn{Inject}, this is a consequence of injectivity.
-For \rn{Bisimilar}, our construction of $\Val$ indicates that the value of $t_1$ and $t_2$ are equivalent in all models of $\Ec$.
+For \rn{Unique}, our construction of $\Val$ indicates that the value of $t_1$ and $t_2$ are equivalent in all models of $\Ec$.
 For \rn{Singleton}, clearly $t_1$ and $t_2$ must have the same value since the cardinality of their type is one.
@@
 To show 3,
-first note that since \rn{Bisimilar} does not apply,
+first note that since \rn{Unique} does not apply,
 $\Val \ec{x} \sim \Val \ec{y}$ if and only if $\ec{x} = \ec{y}$.
@@
   & 227 & 771 & \phantom{0}57 & 1055 \\
-Codatatypes without bisimilarity
+Codatatypes without uniqueness
   & 217 & 787 & \phantom{0}55 & 1059 \\
-Codatatypes with bisimilarity
+Codatatypes with uniqueness
   & 218 & 787 & \phantom{0}58 & 1063 \\
@@
 codatatypes either enabled or disabled. The contributions of the acyclicity and
-bisimilarity rules were also measured, so find out whether these somewhat
+uniqueness rules were also measured, so find out whether these somewhat
 expensive rules are useful in practice. Even when the decision procedure is
@@
 
-Among the four bisimilarity proofs, three corresponded to one-line
+Among the four uniqueness proofs, three corresponded to one-line
 proofs in Isabelle, of the form \keyw{by}~\textit{coinduction}~\textit{auto}
@@
 %% alliteration
-On the practical side, we obtained
-interesting results on benchmarks generated from Isabelle theories, including
-some rather
-%mind-boggling
-puzzling bisimulation proofs.
+On the practical side, we obtained respectable results on benchmarks generated
+from Isabelle theories.
 

### 2014-12-19 tex/codatadp.tex
rephrased

@@
 datatypes and codatatypes. The dual of the acyclicity rule for datatypes is a
-uniqueness rule that identifies bisimilar values, also in the presence of cyclic
+uniqueness rule that identifies observationally equal values, also in the presence of cyclic
 ($\omega$-regular) data. The procedure decides ground and universal
@@
 
-Second, \emph{a uniqueness rule takes the place of the acyclicity rule of datatypes.}
+Second, \emph{a uniqueness rule replaces the acyclicity rule of datatypes.}
 Cyclic constraints such as
@@
 are unsatisfiable for datatypes but satisfiable for codatatypes.
-The uniqueness principle states that two values having the same (possibly
-infinite) expansion must be equal; from $\const{x} \teq \const{S}(\const{y})$ and
+However, the uniqueness principle states that two bisimilar values---i.e., two values
+having the same possibly infinite expansion---must be equal; from $\const{x}
+\teq \const{S}(\const{y})$ and
 $\const{y} \teq \const{S}(\const{x})$, it deduces $\const{x} \teq \const{y}$.

### 2014-12-19 tex/codatadp.tex
More work.

@@
 
+\newcommand\xend{{\hfill\footnotesize$\blacksquare$}}
+
 %\renewcommand{\vec}[1]{\mathbf{#1}}
@@
 \newcommand{\is}[1]{\const{is#1}}
-\newcommand{\ror}{\quad \parallel \quad}
+\newcommand{\ror}{\kern.45em \parallel \kern.45em}
 \newcommand{\tpath}[2]{\mathcal{P}_{ #2 \rightarrow \_ }( #1 )}
@@
 %
-Finally, the notation $\bar x$ abbreviates a list or tuple $x_1,\ldots,x_n$
+Finally, $\bar x$ abbreviates a list or tuple $x_1,\ldots,x_n$\begin{report}
 and $\bigwedge_{\,i}\, \varphi_i$ abbeviates a conjunction
-$\varphi_1 \mathrel\land \cdots \mathrel\land \varphi_n$.
+$\varphi_1 \mathrel\land \cdots \mathrel\land \varphi_n$\end{report}.
 
@@
 $\thD$-satisfiability of sets of ground constraints over a fixed
-signature~$\Sig$. It is presented as a calculus, whose rules can be applied in
+signature~$\Sig$. It is formulated as a calculus, whose rules can be applied in
 any order until saturation.
-
-The derivation rules operate on a set $\Ec$ of equalities and disequalities
+%
+The rules operate on a set $\Ec$ of equalities and disequalities
 between $\Sig$-terms, where $\Ec$ is interpreted as a conjunction.
@@
 on the cardinality of any uninterpreted type $\tau \in \Nondata$; thus,
-we may without loss of completeness treat all other types as infinite.
-
-The calculus consists of three sets of rules (\figuresname~\ref{fig:cc-rules} to
-\ref{fig:split-rule}) corresponding to three successive phases. The first
+we may without loss of generality or completeness treat all types in $\Nondata$ as infinite.
+
+The calculus consists of three sets of rules, given in \figuresname~\ref{fig:cc-rules} to
+\ref{fig:split-rule}, corresponding to three successive phases. The first
 phase computes the bidirectional closure of $\Ec$. The second phase makes
@@
 respectively. %, for some values of ${\Varec{w}}$ and ${\Varec{z}}$.
-$\Box$
+\xend
 \end{example}
@@
 are equivalent for any value of ${\Varec{y}}$.
-$\Box$
+\xend
 \end{example}
@@
   \quad
-  ( s( t ) \in \tEc, s \in \Sel^\tau )
+  ( s( t ) \in \tEc \text{ and } s \in \Sel^\tau )
   \text{ or }
-  ( \tau \in \Data, \tau \text{ finite } )
+  ( \tau \in \Data \text{ and } \tau \text{ finite} )
 }{
-  \Ec := \Ec,\: t \teq \const{C}_1( s^1_1( t ), \ldots, s^{a_1}_1( t ) ) \ror \cdots \ror \Ec := \Ec, t \teq \const{C}_n( s^1_n( t ), \ldots, s^{a_n}_n( t ) )
+  \Ec := \Ec,\: t \teq \const{C}_1(\const s^1_1( t ), \ldots,\const s^{n_1}_1( t ) ) \ror \cdots \ror \Ec := \Ec,\: t \teq \const{C}_m(\const s^1_m( t ), \ldots,\const s^{n_m}_m( t ) )
 }
@@
 we say $t_2$ is a \emph{bisimilar subterm} of $t_1$ if and only if it is of the form
-$\mu y. \const{C}( v_1, \ldots, v_n )$ and $t_1\interp{ u_i } \sim t_1\interp{ v_i }$ for $i = 1,\ldots, n$.
+$\mu y. \const{C}( v_1, \ldots, v_n )$ and $t_1\interp{ u_i } \aequiv t_1\interp{ v_i }$ for $i = 1,\ldots, n$.
 We say $t_1$ is \emph{normal} if and only if it does not contain a strict bisimilar subterm
@@
 \begin{proof}
-%Since $u_i$ is normal, $w_i$ is the result of replacing all subterms $v$ of $u_i$ such that $t\interp{v} \sim t$ with $x$.
+%Since $u_i$ is normal, $w_i$ is the result of replacing all subterms $v$ of $u_i$ such that $t\interp{v} \aequiv t$ with $x$.
 \rem{TODO}
@@
 Let $\tau$ be a type in $\Codata$ with three constructors $\const{A}$, $\const{B}$ and $\const{C}$ each having type $\tau \rightarrow \tau$.
-Let $\Ec$ be the set consisting of equalities
+Let $\Ec$ be the set
 $\{$
@@
 \]
-Now, consider a model $\M$ that interprets variables in $\Ec$ based on the mapping $\Val$,
+Consider a model $\M$ that interprets variables in $\Ec$ based on the mapping $\Val$,
 for instance, $x$ is interpretted as $\Val \ec{x}$, $y_1$ as $\Val \ec{y_1}$ and so on.
@@
 to ${\Varec{x}}$.
-Notice that legal substitutions for ${\Varec{x}}$ may cause the range of $\Val$ to contain non-normal terms,
-as in after applying the substitution for the third term mentioned above,
+Legal substitutions for $v_{\ec{x}}$ may cause the range of $\Val$ to contain non-normal terms.
+After applying the substitution for the third term mentioned above,
 $\Val \ec{w_1}$ is $\mu {\Varec{w_1}}. \const{A}( \mu {\Varec{x}}. \const{B}( \const{A}( {\Varec{x}} ) ) )$,
@@
 $\mu {\Varec{w_1}}. \const{A}( \mu {\Varec{x}}. \const{B}( {\Varec{w_1}} ) )$.
-$\Box$
+\xend
 \end{example}
 
-We first state the following properties of $\Val^\ast$.
+%We first state the following properties of $\Val^\ast$.
 %where we write $\muvar( t )$ to denote $t$ if $t$ is a variable, or $x$ if $t$ is of the form $\mu x. u$.
@@
 \item
-If $\const{C}( t_1, \ldots, t_n ) \in \ec{x}$,
-then $\Val^\ast \ec{x} = \mu {\Varec{x}}. \const{C}( w_1, \ldots, w_n )$
-and $\Val^\ast \ec{x} \interp{ w_i } = \Val^\ast \ec{t_i}$ for $i = 1, \ldots, n$.
-%and $\muvar( w_i ) = {\Varec{t_i}}$ for $i = 1, \ldots, n$,
+All subterms $\mu {\Varec{x}}. t$ of $\Val^\ast \ec{y}$ are such that
+$\Val^\ast \ec{y} \interp{ \mu {\Varec{x}}. t } = \Val^\ast \ec{x}$. 
+%If $\const{C}( t_1, \ldots, t_n ) \in \ec{x}$,
+%then all subterms $\mu v_{\ec{x}}. u$ occurring in $\Val^\ast \ec{y}$ 
+%are of the form $\mu v_{\ec{x}}. \const{C}( w_1, \ldots, w_n )$,
+%where 
+%and $\Val^\ast \ec{y} \interp{ w_i } = \Val^\ast \ec{t_i}$ for $i = 1, \ldots, n$.
+%If $\Val^\ast \ec{x}$ contains a subterm of the form $\mu v_{\ec{x}}. \const{C}( w_1, \ldots, w_n )$,
+%and $\muvar( w_i ) = v_{\ec{t_i}}$ for $i = 1, \ldots, n$,
 %\item
@@
 first note that since \rn{Clash} does not apply to $\Ec$,
-for all equivalence classes $\ec{y}$,
-each pair of constructor terms $\const{C}_i( \vec t )$ and $\const{C}_j( \vec u )$ in $\ec{y}$
+for all equivalence classes $\ec{z}$,
+each pair of constructor terms $\const{C}_i( \vec t )$ and $\const{C}_j( \vec u )$ in $\ec{z}$
 are such that $i = j$ since \rn{Clash} does not apply to $\Ec$,
-and are such that $\ec{ \vec t } = \ec{ \vec u }$ since \rn{Inject} does not apply to $\Ec$.
+and are such that $\ec{ \vec t } = \ec{ \vec u }$ since \rn{Decomp} does not apply to $\Ec$.
 Thus, 
-if $\const{C}_i( t_1, \ldots, t_n ) \in \ec{y}$,
+if $\const{C}_i( t_1, \ldots, t_n ) \in \ec{z}$,
 $\Val$ was constructed by applying a sequence of substitutions
-that replace ${\Varec{y}}$ by $\mu {\Varec{y}}. \const{C}_i( {\Varec{t_1}}, \ldots, {\Varec{t_n}} )$.
-Now, say that $\const{C}( t_1, \ldots, t_n )$ occurs in some equivalence class $\ec{x}$.
-Then $\Val \ec{x}$ must be of the form $\mu {\Varec{x}}. \const{C}_i( w_1, \ldots, w_n )$,
-where for each $i=1,\ldots,n$, $w_i$ was obtained by applying a sequence of substitutions
-of the form mentioned above to ${\Varec{t_i}}$.
-%Each $w_i$ is either ${\Varec{x}}$,
-%a variable ${\Varec{y}}$ where $\ec{y}$ does not contain a constructor,
-%or a term of the form $\mu {\Varec{y}}. \const{C}_j( \vec u )$.
-While interpreting $w_i$ in $\Val \ec{x}$,
-all free occurrences of ${\Varec{x}}$ in $\Val \ec{x} \interp{w_i}$ are replaced by $\Val \ec{x}$. \rem{better formalize.}
-Thus, since $\Val \ec{x} \interp{w_i}$ and $\Val \ec{t_i}$ are obtained by applying the same sequence of substitutions to ${\Varec{t_i}}$,
-we have $\Val \ec{w_i} = \Val \ec{x} \interp{ w_i }$.
-Since $\Val'$ is obtained by applying the same substitution to all terms in the range of $\Val$,
-we have that $\Val' \ec{x} \interp{w_i} = \Val' \ec{t_i}$,
-and hence $\Val^\ast \ec{x} \interp{w_i} = \Val^\ast \ec{t_i}$.
+where ${\Varec{z}}$ was replaced only by $\mu {\Varec{z}}. \const{C}_i( {\Varec{t_1}}, \ldots, {\Varec{t_n}} )$.
+Assume $\Val \ec{y}$ has a subterm $\mu {\Varec{x}}. \const{C}( \vec t )$.
+We first show that our construction of $\Val$ is such that $\Val \ec{x} = \Val \ec{y}\interp{ \mu {\Varec{x}}. t }$.
+Notice that both $\Val \ec{x}$ and the subterm $\mu {\Varec{x}}. t$ of $\Val \ec{y}$ were
+constructed by applying a sequence of substitutions of the form mentioned above to ${\Varec{x}}$.
+We show that $\Val \ec{x} \interp{ \Val \ec{x} } = \Val \ec{y} \interp{ \mu {\Varec{x}}.t }$.
+\rem{TODO}
+%Thus, the set of subterms which $\Val \ec{x}$ and $\mu v_{\ec{x}}. t$ differ 
+%of the form $\mu {\Varec{z}}. u$ and $v_{\ec{z}}$ respectively, where $v_{\ec{z}}$ is a bound variable in $\Val \ec{y}$.
+%In the latter case, $\Val \ec{y} \interp{ {\Varec{z}} } = \mu {\Varec{z}}. \const{C}( \Val \ec{y} \interp{\vec u} )$,
+%where $\mu v_{\ec{z}}. u$ and $\mu v_{\ec{z}}. u'$ were constructed by applying a sequence of substitutions of the form above to $v_{\ec{z}}$.
+Since $\Val^\ast$ is obtained by applying the same substitutions and normalizations to all terms in the range of $\Val$,
+we have $\Val^\ast \ec{y} \interp{ \mu {\Varec{x}}. t} = \Val^\ast \ec{x}$.
 
@@
 first note that since \rn{Unique} does not apply,
-$\Val \ec{x} \sim \Val \ec{y}$ if and only if $\ec{x} = \ec{y}$.
+$\Val \ec{x} \aequiv \Val \ec{y}$ if and only if $\ec{x} = \ec{y}$.
 Since \rem{TODO}
-we have $\nf{(\Val \ec{x})} \sim \nf{(\Val \ec{y})}$ if and only if $\ec{x} = \ec{y}$.
+we have $\nf{(\Val \ec{x})} \aequiv \nf{(\Val \ec{y})}$ if and only if $\ec{x} = \ec{y}$.
 We show that for any substitution $\sigma$ applied to $\Val$ while constructing its completion is such that
-$\nf{(\Val \ec{x} \sigma)} \sim \nf{(\Val \ec{y} \sigma)}$ if and only if $\ec{x} = \ec{y}$.
+$\nf{(\Val \ec{x} \sigma)} \aequiv \nf{(\Val \ec{y} \sigma)}$ if and only if $\ec{x} = \ec{y}$.
 \rem{TODO}
+Let $P$ be the set of positions for which $\Val \ec{x}$ and $\Val \ec{y}$ are not $\alpha$-equivalent,
+which is non-empty by assumption.
+
 \qed
@@
 which by the induction hypothesis is $\alpha$-equivalent with $\nf{(\mu x. \const{C}( \Val^\ast \ec{ u_1 }, \ldots, \Val^\ast \ec{ u_n } ) )}$, call this term $t_u$.
-By Lemma~\ref{lem:model-completion}.2, $\Val^\ast \ec{t}$ is a term
-%Since \rn{Clash} does not apply to $\Fc$, then $\Val^\ast \ec{t}$ must be a term
+By Lemma~\ref{lem:model-completion}.2 and due to our construction of $\Val^\ast$, we have $\Val^\ast \ec{t}$ is a term
 $t_w$ of the form $\mu {\Varec{t}}. \const{C}( w_1, \ldots, w_n )$
 where $t_w\interp{w_i} = \Val^\ast \ec{u_i}$ for $i = 1, \ldots, n$.
-%that is, an application of the same constructor.
 For each $i = 1, \ldots, n$, let $u_i'$ be the $i^{th}$ argument of $t_u$.
-By Lemma~\ref{lem:mu-norm-arg}, $t_u\interp{u_i'} \vsim \Val^\ast \ec{u_i}$.
-%Say that $\muvar( w_i ) = {\Varec{y_i}}$.
-%Due to our construction of $\Val^\ast$ and since \rn{Inject} does not apply to $\Fc$, we have $\ec{u_i} = \ec{y_i}$,
-%and thus $t_u\interp{u_i'} \vsim \Val^\ast \ec{y_i}$.
-%By Lemma~\ref{lem:model-completion}.2, $\Val^\ast \ec{ y_i } = t_w\interp{ w_i }$,
+By Lemma~\ref{lem:mu-norm-arg}, $t_u\interp{u_i'} \vsim \Val^\ast \ec{u_i}$,
 and thus $t_u\interp{u_i'} \vsim t_w\interp{ w_i }$. % for each $i = 1, \ldots, n$.
 By Lemma~\ref{lem:mu-cong}, we have $t_u \vsim t_w$.
-Since $\vsim$ is a transitive relation and since $\M( t ) \vsim t_u \vsim t_w = \Val^\ast \ec{t}$,
+Since $\M( t ) \vsim t_u \vsim t_w = \Val^\ast \ec{t}$,
 we have $\M( t ) \vsim \Val^\ast \ec{t}$.
@@
 $\ec{u}$ must contain a term of the form $C_i( s^1_i( u ), \ldots, s^{n}_i( u ) )$ for some $i$.
-By Lemma~\ref{lem:model-completion}.2, $\Val^\ast \ec{u}$ is a term $t_w$ of the form
+By Lemma~\ref{lem:model-completion}.2 and due to our construction of $\Val^\ast$, we have $\Val^\ast \ec{u}$ is a term $t_w$ of the form
 $\mu {\Varec{u}}. C_i( w_1, \ldots, w_n )$,
 where $t_w \interp{ w_j } = \Val^\ast \ec{ s^j_k( u )}$. 
-%By our construction of $\Val^\ast$ and since \rn{Inject} and \rn{Clash} do not apply to $\Fc$,
-%$\Val^\ast \ec{ u }$ must be of the form $t_u = \mu {\Varec{u}}. C_i( u_1, \ldots, u_n )$,
-%where $\muvar( u_1 ) = {\Varec{s^1_i(u)}}$, $\ldots$, $\muvar(u_n) = {\Varec{s^n_i( u )}}$,
-%and $\M( u )$ is of the form $t_w = \mu x. C_i( w_1, \ldots, w_n )$,
-%which by the induction hypothesis is $\alpha$-equivalent with $\Val^\ast \ec{u}$.
 If $k = i$, then $\M( t )$ is $\alpha$-equivalent with $t_w\interp{w_j} = \Val^\ast \ec{ s^j_k( u )}$,
 which is $\Val^\ast \ec{t}$.
-%which since $t_w \vsim t_u$ by Lemma~\ref{lem:mu-cong} is bisimilar with $t_u\interp{u_j}$,
-%which by Lemma~\ref{lem:model-completion}.2 is equal to $\Val^\ast \ec{ s^j_k( u )}$.
-%Thus, $\M( t ) \vsim \Val^\ast \ec{ s^j_k( u )}
 If $k \neq i$, then we interpret $s^j_k( \M( u ) )$ as the value in $\M( \tau )$ that is $\alpha$-equivalent with $\Val^\ast \ec{ t }$.

### 2014-12-19 tex/codatadp.tex
more
tuned example

@@
 
+%%% @ANDY: I don't like using \qed or anything too similar (e.g. \Box) for
+%%% anything else than "quid erat demonstrandum". I'm open to other symbols,
+%%% e.g. a filled black square. But if the example are short enough, which
+%%% they currently are, we could live with nothing.
+%%%
+%\newcommand\xend{{\hfill\footnotesize$\blacksquare$}}
+%\newcommand\xend{{\qed}}
+\newcommand\xend{{}}
+
 %\renewcommand{\vec}[1]{\mathbf{#1}}
@@
 \newcommand{\Varec}[1]{\widetilde{#1}}
+\newcommand{\VAREC}[1]{\widetilde{\vphantom{\scriptstyle x^i}\smash{#1}}}
 \newcommand{\nf}[1]{{#1}\!\!\downarrow}
@@
 Thus, there are initially no constraints on the values for any equivalence class $\ec{x}$.
-The mapping $\Val$ is then modified by applying the following rule exhaustively:
+The mapping $\Val$ is then modified by applying the following unfolding rule exhaustively:
 \[
@@
 \)
-\rn{Update}}
+\rn{Unfold}}
 \]
 $\FV( \Val )$ denotes the union of the free variables occurring in $\Val$'s range.
-It is straightforward to see that the height of terms produced as a result of this update
+It is easy to see that the height of terms produced as a result of the unfolding
 is bounded by the number of equivalence classes of $\Ec$,
@@
 where $\const{C}(z,x) \in \ec{y}$ and $\const{C}(w,y) \in \ec{x}$ for some $\const{C} \in \Ctr$.
-A possible sequence of \rn{Update} rule applications is given below, omitting
+A possible sequence of \rn{Unfold} rule applications is given below, omitting
 trivial entries such as $\ec{w} \mapsto {\Varec{w}}$.
@@
 \begin{enumerate}
-\item \noindent\rlap{Initial state:}\phantom{Update w.r.t.\ $\ec{x}$:\enskip}$\{  \}$
-\item \noindent\rlap{Update w.r.t.\ $\ec{x}$:}\phantom{Update w.r.t.\ $\ec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \mu {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) \}$
-\item \noindent\rlap{Update w.r.t.\ $\ec{y}$:}\phantom{Update w.r.t.\ $\ec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \mu {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \mu {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\;
+\item \noindent\rlap{Initial state:}\phantom{Unfold $\Varec{x}$:\enskip}$\{  \}$
+\item \noindent\rlap{Unfold $\Varec{x}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \mu {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) \}$
+\item \noindent\rlap{Unfold $\Varec{y}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \mu {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \mu {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\;
   \ec{y} \mapsto \mu {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) \vthinspace\}$
-\item \noindent\rlap{Update w.r.t.\ $\ec{x}$:}\phantom{Update w.r.t.\ $\ec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \mu {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \mu {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\;
+\item \noindent\rlap{Unfold $\Varec{x}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \mu {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \mu {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\;
   \ec{y} \mapsto \mu {\Varec{y}}.\; \const{C}( {\Varec{z}},\; \mu {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) ) \vthinspace\}$
@@
 
+%% @ANDY: Minor terminology point:
+%% A function symbol can have a "type signature" (but we're trying to avoid that terminology);
+%% unless it's a constant, it doesn't have a (first-order) type. So I'm rephrasing.
+
 \begin{example}
-Let $\const{C}$ be a constructor of type $\tau_1 \times \tau_2 \rightarrow \tau_2$,
-and let $\Ec$ be the set $\{ x \teq \const{C}(y,\const{C}(y,x)) \}$.
+Let $\const{C} : $ $\tau_1 \times \tau_2 \rightarrow \tau_2$ be a constructor,
+and let $\Ec = \{ x \teq \const{C}(y,\const{C}(y,x)) \}$.
 After phase~1, the equivalence classes of $\Ec$ are
 $\{ x, \const{C}(y,\const{C}(y,x)) \}$, $\{ \const{C}(y,x) \}$, and $\{ y \}$.
-We then construct $\Val$ using the aforementioned phases, after which we obtain:
-\[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}r@{\hspace{.5em}}l@{\hspace{.5em}}r@{\hspace{.5em}}l}
-\Val \ec{x} & = &
-\mu {\Varec{x}}. & \const{C}( {\Varec{y}}, & \mu {\Varec{\const{C}(y,x)}}. & \const{C}( {\Varec{y}}, {\Varec{x}} ) ) \\
+Constructing $\Val$ as described above yields
+\[\begin{array}{rcrlrl}
+\Val \ec{x} & \,=\, &
+\mu {\Varec{x}}. & \const{C}( {\Varec{y}},\, & \mu {\VAREC{\const{C}(y, x)}}. & \const{C}( {\Varec{y}},\, {\Varec{x}} ) ) \\
 \Val \ec{\const{C}(y,x)} & = &
-\mu {\Varec{\const{C}(y,x)}}. & \const{C}( {\Varec{y}}, & \mu {\Varec{x}}. & \const{C}( {\Varec{y}}, {\Varec{\const{C}(y,x)}} ) )
+\mu {\VAREC{\const{C}(y,x)}}. & \const{C}( {\Varec{y}},\, & \mu {\Varec{x}}. & \const{C}( {\Varec{y}},\, {\VAREC{\const{C}(y,x)}} ) )
 \end{array}\]
-Notice that $\Val \ec{x}$ and $\Val \ec{\const{C}(y,x)}$ are syntactically equivalent for the renaming
-$\{ {\Varec{x}} \mapsto {\Varec{\const{C}(y,x)}}, {\Varec{\const{C}(y,x)}} \mapsto {\Varec{x}} \}$,
-and thus our calculus may infer $x \teq \const{C}(y,x)$.
-Intuitively, this equality holds, since the values of $x$ and $\const{C}( y, x )$
-are equivalent for any value of ${\Varec{y}}$.
-$\Box$
+Since $\Val \ec{x}$ and $\Val \ec{\const{C}(y,x)}$ are $\alpha$-equivalent,
+%for the renaming
+%$\{ {\Varec{x}} \mapsto {\VAREC{\const{C}(y,x)}}, {\VAREC{\const{C}(y,x)}} \mapsto {\Varec{x}} \}$,
+the calculus can infer $x \teq \const{C}(y,x)$.
+%Intuitively, this equality holds, since the values of $x$ and $\const{C}( y, x )$
+%are equivalent for any value of ${\Varec{y}}$.
+\xend
 \end{example}
@@
   \quad
-  ( s( t ) \in \tEc, s \in \Sel^\tau )
+  ( \const s( t ) \in \tEc \text{ and } \const s \in \Sel^\tau )
   \text{ or }
-  ( \tau \in \Data, \tau \text{ finite } )
+  ( \tau \in \Data \text{ and } \tau \text{ finite} )
 }{
-  \Ec := \Ec,\: t \teq \const{C}_1( s^1_1( t ), \ldots, s^{a_1}_1( t ) ) \ror \cdots \ror \Ec := \Ec, t \teq \const{C}_n( s^1_n( t ), \ldots, s^{a_n}_n( t ) )
+  \Ec := \Ec,\: t \teq \const{C}_1(\const s^1_1( t ), \ldots,\const s^{n_1}_1( t ) ) \ror \cdots \ror \Ec := \Ec,\: t \teq \const{C}_m(\const s^1_m( t ), \ldots,\const s^{n_m}_m( t ) )
 }
@@
 
-\begin{definition}[Normal Model] \label{def:norm-model}
+\begin{definition}[Normal Model]
+\afterDot%
+\label{def:norm-model}%
+\rm
 A model $\M$ is \emph{normal} if and only if:
@@
 $\mu {\Varec{w_1}}. \const{A}( \mu {\Varec{x}}. \const{B}( {\Varec{w_1}} ) )$.
-$\Box$
+\xend
 \end{example}

### 2014-12-19 tex/codatadp.tex

--- 2014-12-19 tex/codatadp.tex
@@
 
-\newcommand\xend{{\hfill\footnotesize$\blacksquare$}}
+%%% @ANDY: I don't like using \qed or anything too similar (e.g. \Box) for
+%%% anything else than "quid erat demonstrandum". I'm open to other symbols,
+%%% e.g. a filled black square. But if the example are short enough, which
+%%% they currently are, we could live with nothing.
+%%%
+%\newcommand\xend{{\hfill\footnotesize$\blacksquare$}}
+%\newcommand\xend{{\qed}}
+\newcommand\xend{{}}
 
@@
 
+%% @ANDY: Minor terminology point:
+%% A function symbol can have a "type signature" (but we're trying to avoid that terminology);
+%% unless it's a constant, it doesn't have a (first-order) type. So I'm rephrasing.
+
 \begin{example}
-Let $\const{C}$ be a constructor of type $\tau_1 \times \tau_2 \rightarrow \tau_2$,
-and let $\Ec$ be the set $\{ x \teq \const{C}(y,\const{C}(y,x)) \}$.
+Let $\const{C} : $ $\tau_1 \times \tau_2 \rightarrow \tau_2$ be a constructor,
+and let $\Ec = \{ x \teq \const{C}(y,\const{C}(y,x)) \}$.
 After phase~1, the equivalence classes of $\Ec$ are
 $\{ x, \const{C}(y,\const{C}(y,x)) \}$, $\{ \const{C}(y,x) \}$, and $\{ y \}$.
-We then construct $\Val$ using the aforementioned phases, after which we obtain:
-\[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}r@{\hspace{.5em}}l@{\hspace{.5em}}r@{\hspace{.5em}}l}
-\Val \ec{x} & = &
-\mu {\Varec{x}}. & \const{C}( {\Varec{y}}, & \mu {\Varec{\const{C}(y,x)}}. & \const{C}( {\Varec{y}}, {\Varec{x}} ) ) \\
+Constructing $\Val$ as described above yields
+\[\begin{array}{rcrlrl}
+\Val \ec{x} & \,=\, &
+\mu {\Varec{x}}. & \const{C}( {\Varec{y}},\, & \mu {\VAREC{\const{C}(y, x)}}. & \const{C}( {\Varec{y}},\, {\Varec{x}} ) ) \\
 \Val \ec{\const{C}(y,x)} & = &
-\mu {\Varec{\const{C}(y,x)}}. & \const{C}( {\Varec{y}}, & \mu {\Varec{x}}. & \const{C}( {\Varec{y}}, {\Varec{\const{C}(y,x)}} ) )
+\mu {\VAREC{\const{C}(y,x)}}. & \const{C}( {\Varec{y}},\, & \mu {\Varec{x}}. & \const{C}( {\Varec{y}},\, {\VAREC{\const{C}(y,x)}} ) )
 \end{array}\]
-Notice that $\Val \ec{x}$ and $\Val \ec{\const{C}(y,x)}$ are syntactically equivalent for the renaming
-$\{ {\Varec{x}} \mapsto {\Varec{\const{C}(y,x)}}, {\Varec{\const{C}(y,x)}} \mapsto {\Varec{x}} \}$,
-and thus our calculus may infer $x \teq \const{C}(y,x)$.
-Intuitively, this equality holds, since the values of $x$ and $\const{C}( y, x )$
-are equivalent for any value of ${\Varec{y}}$.
+Since $\Val \ec{x}$ and $\Val \ec{\const{C}(y,x)}$ are $\alpha$-equivalent,
+%for the renaming
+%$\{ {\Varec{x}} \mapsto {\VAREC{\const{C}(y,x)}}, {\VAREC{\const{C}(y,x)}} \mapsto {\Varec{x}} \}$,
+the calculus can infer $x \teq \const{C}(y,x)$.
+%Intuitively, this equality holds, since the values of $x$ and $\const{C}( y, x )$
+%are equivalent for any value of ${\Varec{y}}$.
 \xend
@@
   \quad
-  ( s( t ) \in \tEc \text{ and } s \in \Sel^\tau )
+  ( \const s( t ) \in \tEc \text{ and } \const s \in \Sel^\tau )
   \text{ or }

### 2014-12-19 tex/codatadp.tex
more

@@
 
-Given a signature $\Sig$, we write $\thD$ to refer to the \emph{theory} of (co-)datatypes,
+Given a signature $\Sig$, we write $\thD$ to refer to the \emph{theory} of (co)datatypes,
 which in addition to $\Sig$ defines a class of $\Sig$-interpretations $\mathbf{I}$,
@@
 In particular, this means that each interpretation in $\mathbf{I}$ has a fixed interpretation
-for constructor terms and correctly-applied selector terms, but may differ on its interpretation for
-wrongly-applied selector terms and variables.
-A formula $\varphi$ is $\thD$-satisfiable if there is an interpretation in $\mathbf{I}$ that satisfies $\varphi$,
-and $\thD$-unsatisfiable otherwise.
+for constructor terms and correctly applied selector terms, but may differ on its interpretation for
+wrongly applied selector terms and variables.
+A formula $\varphi$ is $\thD$-satisfiable if there exists an interpretation in $\mathbf{I}$ that satisfies $\varphi$;
+otherwise, it is $\thD$-unsatisfiable.
 \rem{more?}
@@
 \begin{enumerate}
-\item \noindent\rlap{Initial state:}\phantom{Unfold $\Varec{x}$:\enskip}$\{  \}$
+\item \noindent\rlap{Initially:}\phantom{Unfold $\Varec{x}$:\enskip}$\{  \}$
 \item \noindent\rlap{Unfold $\Varec{x}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \mu {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) \}$
@@
 \begin{example}
-Let $\const{C} : $ $\tau_1 \times \tau_2 \rightarrow \tau_2$ be a constructor,
+Let $\const{C} : $ $\tau \times \delta \rightarrow \delta$ be a constructor,
 and let $\Ec = \{ x \teq \const{C}(y,\const{C}(y,x)) \}$.
@@
 the \rn{Singleton} rule handles the case when there are two terms from $\tEc$
-having a co-recursive singleton type, in which case they must be equivalent.
+having a corecursive singleton type, in which case they must be equivalent.
 Otherwise, in the case there is a selector term applied to $t$, or if $t$ has a finite datatype type,

### 2014-12-19 tex/codatadp.tex
more

@@
 \item \noindent\rlap{Initially:}\phantom{Unfold $\Varec{x}$:\enskip}$\{  \}$
-\item \noindent\rlap{Unfold $\Varec{x}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \mu {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) \}$
+\item \noindent\rlap{Unfold $\Varec{x}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \mu {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) \vthinspace\}$
 \item \noindent\rlap{Unfold $\Varec{y}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \mu {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \mu {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\;
@@
 After phase~1, the equivalence classes of $\Ec$ are
-$\{ x, \const{C}(y,\const{C}(y,x)) \}$, $\{ \const{C}(y,x) \}$, and $\{ y \}$.
+$\{ x,\, \const{C}(y,\const{C}(y,x)) \}$, $\{ \const{C}(y,x) \}$, and $\{ y \}$.
 Constructing $\Val$ as described above yields
@@
 \inferrule{
+  t : \tau \in \tEc
+  \quad
+  ( \const s( t ) \in \tEc \text{ and } \const s \in \Sel^\tau )
+  \text{ or }
+  ( \tau \in \Data \text{ and } \tau \text{ finite} )
+}{
+  \Ec := \Ec,\: t \teq \const{C}_1(\const s^1_1( t ), \ldots,\const s^{n_1}_1( t ) ) \ror \cdots \ror \Ec := \Ec,\: t \teq \const{C}_m(\const s^1_m( t ), \ldots,\const s^{n_m}_m( t ) )
+}
+\)
+\rn{Split}
+\\[5\jot]
+\(
+\inferrule{
   \tau \in \Codata
@@
 \rn{Singleton}
-\\[5\jot]
-\(
-\inferrule{
-  t : \tau \in \tEc
-  \quad
-  ( \const s( t ) \in \tEc \text{ and } \const s \in \Sel^\tau )
-  \text{ or }
-  ( \tau \in \Data \text{ and } \tau \text{ finite} )
-}{
-  \Ec := \Ec,\: t \teq \const{C}_1(\const s^1_1( t ), \ldots,\const s^{n_1}_1( t ) ) \ror \cdots \ror \Ec := \Ec,\: t \teq \const{C}_m(\const s^1_m( t ), \ldots,\const s^{n_m}_m( t ) )
-}
-\)
-\rn{Split}
 \end{tabular}
@@
 \paragraph{Phase 3: Branching.}
-In the third phase,
-the \rn{Singleton} rule handles the case when there are two terms from $\tEc$
-having a corecursive singleton type, in which case they must be equivalent.
 Otherwise, in the case there is a selector term applied to $t$, or if $t$ has a finite datatype type,
-then the equivalence class of $t$ must contain an application of one of the constructors of $\tau$ (where $\Ctr^\tau = \{ C_1, \ldots C_n \}$),
-which is enforced by the rule \rn{Split}.
+the equivalence class of $t$ must contain an application of one of the constructors of $\tau$ (where $\Ctr^\tau = \{ C_1, \ldots C_n \}$).
+This is enforced in the third phase by the \rn{Split} rule.
+The \rn{Singleton} rule handles the degenerate case where two terms from $\tEc$
+have a corecursive singleton type, in which case they must be equivalent.
 
@@
 conclusion does not add a new equality to $\Ec$ and is not $\bot$.
-A \emph{derivation tree} is a tree whose nodes are sets of equalities, where non-root nodes vtained by
+A \emph{derivation tree} is a tree whose nodes are sets of equalities, where non-root nodes are obtained by
 a nonredundant application of a derivation rule to its parent node.

### 2014-12-19 tex/codatadp.tex
more

@@
 
+\paragraph{Correctness.}
 An application of a rule is \emph{redundant} if at least one branch in the
@@
 If there exists a derivation tree with root node $\Ec$ that contains a saturated node, then we conclude that $\Ec$ is $\thD$-satisfiable.
-
-\paragraph{Correctness.}
 
@@
 \afterDot
-If there exists a closed derivation tree with root node $\Ec$, then $\Ec$ is $\thD$-unsatisfiable.
+If there exists a closed derivation tree with root node\/ $\Ec$, then\/ $\Ec$ is $\thD$-unsatisfiable.
 \end{lemma}
@@
 
-It remains to show the inverse of the previous lemma, that when a derivation tree with root node $\Ec$ contains a saturated node,
-then $\Ec$ is satisfiable $\thD$.
-To do so, we demonstrate a particular model $\M$ of $\thD$ that satisfies $\Ec$,
-but first we require additional terminology concerning $\mu$-terms.
+It remains to show the converse of the previous lemma: When a derivation tree
+with root node $\Ec$ contains a saturated node, then $\Ec$ is
+$\thD$-satisfiable.
+To do so, we exhibit a particular model $\M$ of $\thD$ that satisfies $\Ec$.
+First, we require additional terminology concerning $\mu$-terms.
 

### 2014-12-19 tex/codatadp.tex
more

@@
 
-\begin{lemma}[Termination]%
-\label{lem:t}%
+\begin{theorem}[Termination]%
+\label{thm:t}%
 \afterDot
 All derivation trees are finite.
-\end{lemma}
+\end{theorem}
 \begin{proof}
@@
 
-\begin{lemma}[Refutation Soundness]%
-\label{lem:rs}%
+\begin{theorem}[Refutation Soundness]%
+\label{thm:rs}%
 \afterDot
 If there exists a closed derivation tree with root node\/ $\Ec$, then\/ $\Ec$ is $\thD$-unsatisfiable.
-\end{lemma}
+\end{theorem}
 \begin{proof}
@@
 
-It remains to show the converse of the previous lemma: When a derivation tree
+It remains to show the converse of the previous theorem: When a derivation tree
 with root node $\Ec$ contains a saturated node, then $\Ec$ is
@@
 where $x \mapsto \mu x. \const{C}( \vec u ) \in t_1$ indicates that $\mu x. \const{C}( \vec u )$ is the subterm of $t_1$ that binds this occurrence of variable $x$.
-\[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
-t_1\interp{x}^M & = & \mu x. \const{C}( t_1\interp{ \vec u}^{M \cup \{ x \}}) & x \not\in M, x \mapsto \mu x. \const{C}( \vec u ) \in t_1  \\
-t_1\interp{x}^M & = & x & \text{otherwise} \\
-t_1\interp{\mu x. \const{C}( \vec u )}^M & = & \mu x. \const{C}( t_1\interp{ \vec u }^{M \cup \{ x \}} ) & x \not\in M \\
-t_1\interp{\mu x. \const{C}( \vec u )}^M & = & x & \text{otherwise}
+\[\begin{array}{r@{}c@{}l}
+t_1\interp{x}^M & {} \,=\, {} &
+\textstyle\begin{cases}
+    \mu x.\; \const{C}( t_1\interp{ \vec u}^{M \cup \{ x \}}) & \text{if }x \not\in M, x \mapsto \mu x. \const{C}( \vec u ) \in t_1  \\[-\jot]
+    x & \text{otherwise}
+\end{cases} \\
+t_1\interp{\mu x. \const{C}( \vec u )}^M & = &
+\textstyle\begin{cases}
+    \mu x. \const{C}( t_1\interp{ \vec u }^{M \cup \{ x \}} ) & \text{if }x \not\in M \\[-\jot]
+    x & \text{otherwise}
+\end{cases}
 \end{array}\]
@@
 
-\begin{lemma}[Solution Soundness]%
-\label{lem:ss}%
+\begin{theorem}[Solution Soundness]%
+\label{thm:ss}%
 \afterDot
 If there exists a derivation tree with root node $\Ec$ containing a saturated node, then $\Ec$ is $\thD$-satisfiable.
-\end{lemma}
+\end{theorem}
 \begin{proof}
@@
 
-By Lemmas~\ref{lem:t},~\ref{lem:rs}, and~\ref{lem:ss}, the calculus is sound and complete for $\thD$.
+By Theorems~\ref{thm:t},~\ref{thm:rs}, and~\ref{thm:ss}, the calculus is sound and complete for $\thD$.
 

### 2014-12-19 tex/codatadp.tex
more

@@
 \usepackage[all]{xy}
+
+\makeatletter
+\def\spnXwtheorem#1#2#3#4{\@spynthm{#1}{#2}{#3}{#4}%
+                         \@addtoreset{#1}{chapter}}%
+\makeatother
+
+%%% not part of shared counter
+\spnXwtheorem{examplex}{Example}{\itshape}{\rmfamily}
 
@@
 
-\begin{lemma}[Corecursive Singletons]%
+\begin{lemma}%[Corecursive Singletons]%
 \label{lem:corecursive-singletons}%
@@
 
-\begin{example}
+\begin{examplex}
 Suppose that $\Ec$ contains distinct equivalence classes $\ec{w}$, $\ec{x}$, $\ec{y}$, and $\ec{z}$
@@
 \xend
-\end{example}
+\end{examplex}
 
@@
 
-\begin{example}
+\begin{examplex}
 Let $\const{C} : $ $\tau \times \delta \rightarrow \delta$ be a constructor,
@@
 \Val \ec{x} & \,=\, &
-\mu {\Varec{x}}. & \const{C}( {\Varec{y}},\, & \mu {\VAREC{\const{C}(y, x)}}. & \const{C}( {\Varec{y}},\, {\Varec{x}} ) ) \\
+\mu {\Varec{x}}. & \const{C}( {\Varec{y}},\, & \smash{\mu {\vthinspace\VAREC{\const{C}(y, x)}}}. & \const{C}( {\Varec{y}},\, {\Varec{x}} ) ) \\
 \Val \ec{\const{C}(y,x)} & = &
-\mu {\VAREC{\const{C}(y,x)}}. & \const{C}( {\Varec{y}},\, & \mu {\Varec{x}}. & \const{C}( {\Varec{y}},\, {\VAREC{\const{C}(y,x)}} ) )
+\mu {\vthinspace\VAREC{\const{C}(y,x)}}. & \const{C}( {\Varec{y}},\, & \mu {\Varec{x}}. & \const{C}( {\Varec{y}},\, {\VAREC{\const{C}(y,x)}} ) )
 \end{array}\]
@@
 \inferrule{
-  t : \tau \in \tEc
+  t : \delta \in \tEc
   \quad
-  ( \const s( t ) \in \tEc \text{ and } \const s \in \Sel^\tau )
+  ( \const s( t ) \in \tEc \text{ and } \const s \in \Sel^\delta )
   \text{ or }
-  ( \tau \in \Data \text{ and } \tau \text{ finite} )
+  ( \delta \in \Data \text{ and } \delta \text{ finite} )
 }{
@@
 \inferrule{
-  \tau \in \Codata
+  \delta \in \Codata
   \quad
-  \left|\tau\right| = 1
+  \left|\delta\right| = 1
   \quad
-  t, u : \tau \in \tEc
+  t, u : \delta \in \tEc
 }{
@@
 \paragraph{Phase 3: Branching.}
-Otherwise, in the case there is a selector term applied to $t$, or if $t$ has a finite datatype type,
-the equivalence class of $t$ must contain an application of one of the constructors of $\tau$ (where $\Ctr^\tau = \{ C_1, \ldots C_n \}$).
+If a selector is applied to a term $t$, or if $t$'s type is a finite datatype,
+the equivalence class of $t$ must contain an application of one of the constructors of $\delta$ (where $\Ctr^\delta = \{ \const{C}_1, \ldots \const{C}_m \}$).
 This is enforced in the third phase by the \rn{Split} rule.
@@
 
-\begin{example}
+\begin{examplex}
 Let $\tau$ be a type in $\Codata$ with three constructors $\const{A}$, $\const{B}$ and $\const{C}$ each having type $\tau \rightarrow \tau$.

### 2014-12-19 tex/codatadp.tex
got the phases a bit better now

@@
 The calculus consists of three sets of rules, given in \figuresname~\ref{fig:cc-rules} to
-\ref{fig:split-rule}, corresponding to three successive phases. The first
+\ref{fig:split-rule}, corresponding to three phases. The first
 phase computes the bidirectional closure of $\Ec$. The second phase makes
@@
 A rule may have multiple conclusions separated by $\parallel$,
-denoting nondeterministic branching. The rules associated with a phase are
-applied until saturation (i.e., until no rule can be applied to extend~$\Ec$
-further) before the next phase can begin.
+denoting nondeterministic branching. The rules associated with a phase have
+priority over those of the previous phases. The rules are applied until
+saturation (i.e., until no rule can be applied to extend~$\Ec$
+further).
 
@@
 If a selector is applied to a term $t$, or if $t$'s type is a finite datatype,
-the equivalence class of $t$ must contain an application of one of the constructors of $\delta$ (where $\Ctr^\delta = \{ \const{C}_1, \ldots \const{C}_m \}$).
+the equivalence class of $t$ must contain an application of a $\delta$ constructor
+(where $\Ctr^\delta = \{ \const{C}_1, \ldots, \const{C}_m \}$).
 This is enforced in the third phase by the \rn{Split} rule.
-The \rn{Singleton} rule handles the degenerate case where two terms from $\tEc$
-have a corecursive singleton type, in which case they must be equivalent.
+A companion rule, \rn{Singleton}, handles the degenerate case where two terms
+from $\tEc$ have a corecursive singleton type, in which case they must be
+equivalent.
 

### 2014-12-19 tex/codatadp.tex
more

@@
 \makeatother
+
+\newcommand\MU{\mu\vvthinspace}
 
@@
 
+\newcommand\vvthinspace{\kern+0.041667em}
 \newcommand\vthinspace{\kern+0.083333em}
@@
 in the metatheory. For example,
-the $\mu$-term $\const{SCons}(1,\: \mu s.\; \const{SCons}(0,\: \const{SCons}(9,\: s)))$
+the $\mu$-term $\const{SCons}(1,\: \MU s.\; \const{SCons}(0,\: \const{SCons}(9,\: s)))$
 stands for the lasso-shaped sequence $1, 0, 9, 0, 9, 0, 9, \ldots$\afterLdots.
@@
 %    * mutual recursion, but no polymorphism, nested recursion
-%  * integrated with Nelson-Oppen
+%  * integrated with Nelson--Oppen
 
@@
 \vthinspace$\keyw{codatatype}~\ty{a} = \const{A}(\ty{a})$, whose
-only value is $\mu a.\; A(a)$. Other specimens are
+only value is $\MU a.\; A(a)$. Other specimens are
 \begin{paper}%
@@
   %\ttpath{\Ec}{\ec{t}}{\ec{t}} \neq \emptyset
-  \Val \ec{t} = \mu x.\; u
+  \Val \ec{t} = \MU x.\; u
   \quad
@@
 %(Section~\ref{sec:introduction}).
-% for example, $\mu x. \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
+% for example, $\MU x. \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
 A $\mu$-term $t$ of type $\tau$ describes a class of values of
@@
 or a node of the form
-$\mu x.\; \const{C}( \vec t\, )$ for some constructor $\const{C} \in \Ctr$ and
+$\MU x.\; \const{C}( \vec t\, )$ for some constructor $\const{C} \in \Ctr$ and
 $\mu$-terms $\vec t$ with the expected types.
@@
 %
-%For uniformity, we also consider $%\mu x.\;
+%For uniformity, we also consider $%\MU x.\;
 %\const{c}$ to be a $\mu$-term if $\const{c}$ is a constant of type $\tau \in \Nondata$.
@@
 %and $\vec v$ are well-typed $\mu$-terms of type $\vec \tau$ under the assumption that $x$ has type $\upsilon$,
-%then $\mu x. \const{C}( \vec v )$ is a well-typed term of type $\upsilon$.
+%then $\MU x. \const{C}( \vec v )$ is a well-typed term of type $\upsilon$.
 
@@
 are syntactically equivalent for some capture-avoiding renaming of $\mu$-bound variables---e.g.,
-%$\mu x. \const{C}( x ) \aequiv \mu y. \const{C}( y )$ and
-$\mu x.\; \const{D}( y, x ) \aequiv \mu z.\; \const{D}( y, z )$\begin{report},
+%$\MU x. \const{C}( x ) \aequiv \MU y. \const{C}( y )$ and
+$\MU x.\; \const{D}( y, x ) \aequiv \MU z.\; \const{D}( y, z )$\begin{report},
 but
-$\mu x.\; \const{C}( x ) \not\aequiv \mu x.\; \const{D}( y, x ) \not\aequiv \mu x.\; \const{D}( z, x )
-\not\aequiv \mu y.\; \const{D}( y, x )$\end{report}.
+$\MU x.\; \const{C}( x ) \not\aequiv \MU x.\; \const{D}( y, x ) \not\aequiv \MU x.\; \const{D}( z, x )
+\not\aequiv \MU y.\; \const{D}( y, x )$\end{report}.
 Two $\mu$-terms can represent the same value while not being $\alpha$-equivalent---e.g.,
-$\mu x.\; \const{C}( x ) \not\aequiv \mu x.\; \const{C}( \const{C}( x ) )$.
+$\MU x.\; \const{C}( x ) \not\aequiv \MU x.\; \const{C}( \const{C}( x ) )$.
 %, or informally are \emph{observationally equivalent}.
@@
 }{
-  \Val := \Val \{ {\Varec{x}} \mapsto \mu {\Varec{x}}.\; \const{C}( {\Varec{t_1}}, \ldots, {\Varec{t_n}} ) \}
+  \Val := \Val \{ {\Varec{x}} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{t_1}}, \ldots, {\Varec{t_n}} ) \}
 }
@@
 \item \noindent\rlap{Initially:}\phantom{Unfold $\Varec{x}$:\enskip}$\{  \}$
-\item \noindent\rlap{Unfold $\Varec{x}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \mu {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) \vthinspace\}$
-\item \noindent\rlap{Unfold $\Varec{y}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \mu {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \mu {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\;
-  \ec{y} \mapsto \mu {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) \vthinspace\}$
-\item \noindent\rlap{Unfold $\Varec{x}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \mu {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \mu {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\;
-  \ec{y} \mapsto \mu {\Varec{y}}.\; \const{C}( {\Varec{z}},\; \mu {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) ) \vthinspace\}$
+\item \noindent\rlap{Unfold $\Varec{x}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) \vthinspace\}$
+\item \noindent\rlap{Unfold $\Varec{y}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\;
+  \ec{y} \mapsto \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) \vthinspace\}$
+\item \noindent\rlap{Unfold $\Varec{x}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\;
+  \ec{y} \mapsto \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\; \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) ) \vthinspace\}$
 \end{enumerate}
@@
 Given $\Val$, the rules for acyclicity and uniqueness can be stated as follows.
-For acyclicity, if there is an equivalence class $\ec{t}$ of datatype type whose class of values $\Val \ec{t} = \mu x.\; u$ are cyclic,
+For acyclicity, if there is an equivalence class $\ec{t}$ of datatype type whose class of values $\Val \ec{t} = \MU x.\; u$ are cyclic,
 as indicated by the assumption $x \in \FV( u )$,
@@
 \Val \ec{x} & \,=\, &
-\mu {\Varec{x}}. & \const{C}( {\Varec{y}},\, & \smash{\mu {\vthinspace\VAREC{\const{C}(y, x)}}}. & \const{C}( {\Varec{y}},\, {\Varec{x}} ) ) \\
+\MU {\Varec{x}}. & \const{C}( {\Varec{y}},\, & \smash{\MU {\VAREC{\const{C}(y, x)}}}. & \const{C}( {\Varec{y}},\, {\Varec{x}} ) ) \\
 \Val \ec{\const{C}(y,x)} & = &
-\mu {\vthinspace\VAREC{\const{C}(y,x)}}. & \const{C}( {\Varec{y}},\, & \mu {\Varec{x}}. & \const{C}( {\Varec{y}},\, {\VAREC{\const{C}(y,x)}} ) )
+\MU {\VAREC{\const{C}(y,x)}}. & \const{C}( {\Varec{y}},\, & \MU {\Varec{x}}. & \const{C}( {\Varec{y}},\, {\VAREC{\const{C}(y,x)}} ) )
 \end{array}\]
@@
 the \emph{interpretation of $t_2$ in $t_1$} is the $\mu$-term $t_1\interp{t_2}^\emptyset$ as returned by the following recursive procedure,
-where $x \mapsto \mu x. \const{C}( \vec u ) \in t_1$ indicates that $\mu x. \const{C}( \vec u )$ is the subterm of $t_1$ that binds this occurrence of variable $x$.
+where $x \mapsto \MU x. \const{C}( \vec u ) \in t_1$ indicates that $\MU x. \const{C}( \vec u )$ is the subterm of $t_1$ that binds this occurrence of variable $x$.
 \[\begin{array}{r@{}c@{}l}
@@
 \textstyle\begin{cases}
-    \mu x.\; \const{C}( t_1\interp{ \vec u}^{M \cup \{ x \}}) & \text{if }x \not\in M, x \mapsto \mu x. \const{C}( \vec u ) \in t_1  \\[-\jot]
+    \MU x.\; \const{C}( t_1\interp{ \vec u}^{M \cup \{ x \}}) & \text{if }x \not\in M, x \mapsto \MU x. \const{C}( \vec u ) \in t_1  \\[-\jot]
     x & \text{otherwise}
 \end{cases} \\
-t_1\interp{\mu x. \const{C}( \vec u )}^M & = &
+t_1\interp{\MU x. \const{C}( \vec u )}^M & = &
 \textstyle\begin{cases}
-    \mu x. \const{C}( t_1\interp{ \vec u }^{M \cup \{ x \}} ) & \text{if }x \not\in M \\[-\jot]
+    \MU x. \const{C}( t_1\interp{ \vec u }^{M \cup \{ x \}} ) & \text{if }x \not\in M \\[-\jot]
     x & \text{otherwise}
@@
 
-Given a $\mu$-term $t_1$ of the form $\mu x. \const{C}( u_1, \ldots, u_n )$ with subterm $t_2$,
+Given a $\mu$-term $t_1$ of the form $\MU x. \const{C}( u_1, \ldots, u_n )$ with subterm $t_2$,
 we say $t_2$ is a \emph{bisimilar subterm} of $t_1$ if and only if it is of the form
-$\mu y. \const{C}( v_1, \ldots, v_n )$ and $t_1\interp{ u_i } \aequiv t_1\interp{ v_i }$ for $i = 1,\ldots, n$.
-We say $t_1$ is \emph{normal} if and only if it does not contain a strict bisimilar subterm
+$\MU y. \const{C}( v_1, \ldots, v_n )$ and $t_1\interp{ u_i } \aequiv t_1\interp{ v_i }$ for $i = 1,\ldots, n$.
+The term $t_1$ is \emph{normal} if it does not contain a strict, bisimilar subterm
 and all of its strict subterms are also normal.
-For instance, the $\mu$-term $t_1$ of the form $\mu x. \const{C}( \mu y. \const{C}( y ) )$ is not normal,
-since $\mu y. \const{C}( y )$ is a bisimilar subterm of $t_1$.
+For example, the $\mu$-term $t_1$ of the form $\MU x. \const{C}( \MU y. \const{C}( y ) )$ is not normal,
+since $\MU y. \const{C}( y )$ is a bisimilar subterm of $t_1$.
 In particular, note that their arguments have the same interpretation in $t_1$,
 that is,
-$t_1\interp{\mu y. \const{C}( y )}^\emptyset =$
-$\mu y. \const{C}( t_1\interp{y}^{\{y\}} ) =$
-$\mu y. \const{C}( y )$
+$t_1\interp{\MU y. \const{C}( y )}^\emptyset =$
+$\MU y. \const{C}( t_1\interp{y}^{\{y\}} ) =$
+$\MU y. \const{C}( y )$
 is syntactically equivalent to
 $t_1\interp{y}^\emptyset =$
-$\mu y. \const{C}( t_1\interp{y}^{\{y\}} ) =$
-$\mu y. \const{C}( y )$.
-The $\mu$-term $t_1$ of the form $\mu x. \const{C}( \mu y. \const{C}( x ) )$ is also not normal,
-since $\mu y. \const{C}( x )$ is a bisimilar subterm of $t_1$,
+$\MU y. \const{C}( t_1\interp{y}^{\{y\}} ) =$
+$\MU y. \const{C}( y )$.
+The $\mu$-term $t_1$ of the form $\MU x. \const{C}( \MU y. \const{C}( x ) )$ is also not normal,
+since $\MU y. \const{C}( x )$ is a bisimilar subterm of $t_1$,
 noting that
-$t_1\interp{\mu y. \const{C}( x )}^\emptyset =$
-$\mu y. \const{C}( t_1\interp{x}^{\{y\}} ) =$
-$\mu y. \const{C}( t_1\interp{\mu x. \const{C}( \mu y. \const{C}( x ) ) }^{\{y\}} ) =$
-$\mu y. \const{C}( \mu x. \const{C}( t_1\interp{\mu y. \const{C}( x )}^{\{x,y\}} ) ) =$
-$\mu y. \const{C}( \mu x. \const{C}( y ) )$
+$t_1\interp{\MU y. \const{C}( x )}^\emptyset =$
+$\MU y. \const{C}( t_1\interp{x}^{\{y\}} ) =$
+$\MU y. \const{C}( t_1\interp{\MU x. \const{C}( \MU y. \const{C}( x ) ) }^{\{y\}} ) =$
+$\MU y. \const{C}( \MU x. \const{C}( t_1\interp{\MU y. \const{C}( x )}^{\{x,y\}} ) ) =$
+$\MU y. \const{C}( \MU x. \const{C}( y ) )$
 is syntactically equivalent to $t_1\interp{x}^\emptyset = t_1$ for the renaming $\{ x \mapsto y, y \mapsto x \}$.
 
-For any $\mu$-term $t$ of the form $\mu x. C( \vec u )$, we may recursively construct a normal $\mu$-term $\nf{t}$
+For any $\mu$-term $t$ of the form $\MU x. C( \vec u )$, we may recursively construct a normal $\mu$-term $\nf{t}$
 (which we call the \emph{normal form} of $t$)
@@
 and by normalizing $t$'s subterms.
-For instance, $\nf{(\mu x. \const{C}( \mu y. \const{C}( x ) ))} = \mu x. \const{C}( x )$.
+For example, $\nf{(\MU x. \const{C}( \MU y. \const{C}( x ) ))} = \MU x. \const{C}( x )$.
 
@@
 \label{lem:mu-norm-arg}
-If $\vec u$ are normal, then if $t = \nf{(\mu x. \const{C}( \vec u ) )} = \mu x. \const{C}( \vec w )$,
+If $\vec u$ are normal, then if $t = \nf{(\MU x. \const{C}( \vec u ) )} = \MU x. \const{C}( \vec w )$,
 then $\vec u \vsim t\interp{\vec w}$.
@@
 \label{lem:mu-cong}
-If $t_u = \mu x. \const{C}( \vec u )$ and $t_w = \mu y. \const{C}( \vec w )$ are normal,
+If $t_u = \MU x. \const{C}( \vec u )$ and $t_w = \MU y. \const{C}( \vec w )$ are normal,
 then $t_u\interp{ \vec u } \vsim t_w\interp{ \vec w }$ if and only if $t_u \vsim t_w$.
@@
 in $\M( \tau )$ that is $\alpha$-equivalent to
-$\nf{(\mu x. \const{C}( \M( \vec t \,) ))}$ where $x$ is fresh, and
+$\nf{(\MU x. \const{C}( \M( \vec t \,) ))}$ where $x$ is fresh, and
 \item
@@
 $\M( t )\interp{u_j}$
-when $\M( t )$ is of the form $\mu x. \const{C}_i( \vec u )$.
+when $\M( t )$ is of the form $\MU x. \const{C}_i( \vec u )$.
 \end{enumerate}
@@
 First, we need the following definition.
-We say that $t \vsimv{x} s$ if and only if $\mu$-terms $t$ and $s$ are syntactically equivalent
+We write $t \vsimv{x} s$ if $\mu$-terms $t$ and $s$ are syntactically equivalent
 for some renaming that avoids capturing any variable other than $x$.
-For instance,
-$\mu x. \const{D}( x ) \vsimv{y} \mu x. \const{D}( y )$,
-$\mu x. \const{C}( x, x ) \vsimv{x} \mu y. \const{C}( x, y )$ and
-$\mu x. \const{C}( z, x ) \vsimv{z} \mu y. \const{C}( z, y )$,
+For example,
+$\MU x. \const{D}( x ) \vsimv{y} \MU x. \const{D}( y )$,
+$\MU x. \const{C}( x, x ) \vsimv{x} \MU y. \const{C}( x, y )$ and
+$\MU x. \const{C}( z, x ) \vsimv{z} \MU y. \const{C}( z, y )$,
 but
-$\mu x. \const{D}( x ) \not\vsimv{x} \mu x. \const{D}( y )$ and
-$\mu x. \const{C}( x, x ) \not\vsimv{y} \mu y. \const{C}( x, y )$.
+$\MU x. \const{D}( x ) \not\vsimv{x} \MU x. \const{D}( y )$ and
+$\MU x. \const{C}( x, x ) \not\vsimv{y} \MU y. \const{C}( x, y )$.
 For a variable $x$ of type $\tau$ and a fixed normal model $\M$,
@@
   \quad
-  (\mu x. t) \aequiv v
+  (\MU x. t) \aequiv v
   \quad
@@
 }{
-  \Val := \nf{(\Val \{ x \mapsto \mu x. t \})}
+  \Val := \nf{(\Val \{ x \mapsto \MU x. t \})}
 }
@@
 \[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{2.0em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
-\Val \ec{y_1} & = & \mu {\Varec{y_1}}. & \const{A}( & \mu {\Varec{y_2}}. & \const{B}( {\Varec{x}} ) ) &
-\Val \ec{w_1} & = & \mu {\Varec{w_1}}. & \const{A}( {\Varec{x}} )      & \\
-\Val \ec{y_2} & = & \mu {\Varec{y_2}}. & \multicolumn{2}{l}{\const{B}( {\Varec{x}} )}      & &
-\Val \ec{w_2} & = & \mu {\Varec{w_2}}. & \const{C}( {\Varec{w_3}} ) \\
-\Val \ec{z_1} & = & \mu {\Varec{z_1}}. & \const{C}( & \mu {\Varec{z_2}}. & \const{A}( {\Varec{z_1}} ) ) &
+\Val \ec{y_1} & = & \MU {\Varec{y_1}}. & \const{A}( & \MU {\Varec{y_2}}. & \const{B}( {\Varec{x}} ) ) &
+\Val \ec{w_1} & = & \MU {\Varec{w_1}}. & \const{A}( {\Varec{x}} )      & \\
+\Val \ec{y_2} & = & \MU {\Varec{y_2}}. & \multicolumn{2}{l}{\const{B}( {\Varec{x}} )}      & &
+\Val \ec{w_2} & = & \MU {\Varec{w_2}}. & \const{C}( {\Varec{w_3}} ) \\
+\Val \ec{z_1} & = & \MU {\Varec{z_1}}. & \const{C}( & \MU {\Varec{z_2}}. & \const{A}( {\Varec{z_1}} ) ) &
 \Val \ec{w_3} & = & {\Varec{w_3}} \\
-\Val \ec{z_2} & = & \mu {\Varec{z_2}}. & \const{A}( & \mu {\Varec{z_1}}. & \const{C}( {\Varec{z_2}} ) ) &
+\Val \ec{z_2} & = & \MU {\Varec{z_2}}. & \const{A}( & \MU {\Varec{z_1}}. & \const{C}( {\Varec{z_2}} ) ) &
 \Val \ec{x} & = & {\Varec{x}} \\
@@
 \{
-\mu x. \const{A}( \const{B}( x )),
-\const{A}( \mu y. \const{B}( y )),
-\mu x. \const{B}( x ),
-\mu x. \const{C}( \const{A}( x )),
-\mu x. \const{A}( \const{C}( x )),
-\mu x. \const{A}( x )
+\MU x. \const{A}( \const{B}( x )),
+\const{A}( \MU y. \const{B}( y )),
+\MU x. \const{B}( x ),
+\MU x. \const{C}( \const{A}( x )),
+\MU x. \const{A}( \const{C}( x )),
+\MU x. \const{A}( x )
 \}
@@
 Consider a model $\M$ that interprets variables in $\Ec$ based on the mapping $\Val$,
-for instance, $x$ is interpretted as $\Val \ec{x}$, $y_1$ as $\Val \ec{y_1}$ and so on.
+for example, $x$ is interpretted as $\Val \ec{x}$, $y_1$ as $\Val \ec{y_1}$ and so on.
 Assigning a value for ${\Varec{x}}$ that is $\alpha$-equivalent to a value in $\mathcal{T}^\M_{{\Varec{x}}}( \Val )$
@@
 which in turn may cause $\M$ to be inconsistent.
-For instance, say we assign $\mu {\Varec{x}}. \const{B}( {\Varec{x}} )$ for ${\Varec{x}}$.
-After this substitution, $\Val \ec{y_2}$ is $\mu {\Varec{y_2}}. \const{B}( \mu {\Varec{x}}. \const{B}( {\Varec{x}} ) )$,
-which has normal form $\mu {\Varec{y_2}}. \const{B}( {\Varec{y_2}} )$,
-which is $\alpha$-equivalent to $\mu {\Varec{x}}. \const{B}( {\Varec{x}} )$.
+For example, say we assign $\MU {\Varec{x}}. \const{B}( {\Varec{x}} )$ for ${\Varec{x}}$.
+After this substitution, $\Val \ec{y_2}$ is $\MU {\Varec{y_2}}. \const{B}( \MU {\Varec{x}}. \const{B}( {\Varec{x}} ) )$,
+which has normal form $\MU {\Varec{y_2}}. \const{B}( {\Varec{y_2}} )$,
+which is $\alpha$-equivalent to $\MU {\Varec{x}}. \const{B}( {\Varec{x}} )$.
 However, this contradicts the disequality $x \not\teq y_2$ in $\Ec$.
@@
 then the values in the range of $\Val$ remain $\alpha$-disequivalent.
-For instance, we may assign values such as 
-$\mu {\Varec{x}}. \const{C}( {\Varec{x}} )$,
-$\mu {\Varec{x}}. \const{C}( \mu x_1. \const{A}( x_1) )$, or
-$\mu {\Varec{x}}. \const{B}( \const{A}( {\Varec{x}} ) )$
+For example, we may assign values such as 
+$\MU {\Varec{x}}. \const{C}( {\Varec{x}} )$,
+$\MU {\Varec{x}}. \const{C}( \MU x_1. \const{A}( x_1) )$, or
+$\MU {\Varec{x}}. \const{B}( \const{A}( {\Varec{x}} ) )$
 to ${\Varec{x}}$.
@@
 After applying the substitution for the third term mentioned above,
-$\Val \ec{w_1}$ is $\mu {\Varec{w_1}}. \const{A}( \mu {\Varec{x}}. \const{B}( \const{A}( {\Varec{x}} ) ) )$,
+$\Val \ec{w_1}$ is $\MU {\Varec{w_1}}. \const{A}( \MU {\Varec{x}}. \const{B}( \const{A}( {\Varec{x}} ) ) )$,
 which has normal form
-$\mu {\Varec{w_1}}. \const{A}( \mu {\Varec{x}}. \const{B}( {\Varec{w_1}} ) )$.
+$\MU {\Varec{w_1}}. \const{A}( \MU {\Varec{x}}. \const{B}( {\Varec{w_1}} ) )$.
 \xend
@@
 %We first state the following properties of $\Val^\ast$.
-%where we write $\muvar( t )$ to denote $t$ if $t$ is a variable, or $x$ if $t$ is of the form $\mu x. u$.
+%where we write $\muvar( t )$ to denote $t$ if $t$ is a variable, or $x$ if $t$ is of the form $\MU x. u$.
 
@@
 \item
-All subterms $\mu {\Varec{x}}. t$ of $\Val^\ast \ec{y}$ are such that
-$\Val^\ast \ec{y} \interp{ \mu {\Varec{x}}. t } = \Val^\ast \ec{x}$. 
+All subterms $\MU {\Varec{x}}. t$ of $\Val^\ast \ec{y}$ are such that
+$\Val^\ast \ec{y} \interp{ \MU {\Varec{x}}. t } = \Val^\ast \ec{x}$. 
 %If $\const{C}( t_1, \ldots, t_n ) \in \ec{x}$,
-%then all subterms $\mu v_{\ec{x}}. u$ occurring in $\Val^\ast \ec{y}$ 
-%are of the form $\mu v_{\ec{x}}. \const{C}( w_1, \ldots, w_n )$,
+%then all subterms $\MU v_{\ec{x}}. u$ occurring in $\Val^\ast \ec{y}$ 
+%are of the form $\MU v_{\ec{x}}. \const{C}( w_1, \ldots, w_n )$,
 %where 
 %and $\Val^\ast \ec{y} \interp{ w_i } = \Val^\ast \ec{t_i}$ for $i = 1, \ldots, n$.
-%If $\Val^\ast \ec{x}$ contains a subterm of the form $\mu v_{\ec{x}}. \const{C}( w_1, \ldots, w_n )$,
+%If $\Val^\ast \ec{x}$ contains a subterm of the form $\MU v_{\ec{x}}. \const{C}( w_1, \ldots, w_n )$,
 %and $\muvar( w_i ) = v_{\ec{t_i}}$ for $i = 1, \ldots, n$,
@@
 $\Val$ was constructed by applying a sequence of substitutions
-where ${\Varec{z}}$ was replaced only by $\mu {\Varec{z}}. \const{C}_i( {\Varec{t_1}}, \ldots, {\Varec{t_n}} )$.
-Assume $\Val \ec{y}$ has a subterm $\mu {\Varec{x}}. \const{C}( \vec t )$.
-We first show that our construction of $\Val$ is such that $\Val \ec{x} = \Val \ec{y}\interp{ \mu {\Varec{x}}. t }$.
-Notice that both $\Val \ec{x}$ and the subterm $\mu {\Varec{x}}. t$ of $\Val \ec{y}$ were
+where ${\Varec{z}}$ was replaced only by $\MU {\Varec{z}}. \const{C}_i( {\Varec{t_1}}, \ldots, {\Varec{t_n}} )$.
+Assume $\Val \ec{y}$ has a subterm $\MU {\Varec{x}}. \const{C}( \vec t )$.
+We first show that our construction of $\Val$ is such that $\Val \ec{x} = \Val \ec{y}\interp{ \MU {\Varec{x}}. t }$.
+Notice that both $\Val \ec{x}$ and the subterm $\MU {\Varec{x}}. t$ of $\Val \ec{y}$ were
 constructed by applying a sequence of substitutions of the form mentioned above to ${\Varec{x}}$.
-We show that $\Val \ec{x} \interp{ \Val \ec{x} } = \Val \ec{y} \interp{ \mu {\Varec{x}}.t }$.
+We show that $\Val \ec{x} \interp{ \Val \ec{x} } = \Val \ec{y} \interp{ \MU {\Varec{x}}.t }$.
 \rem{TODO}
-%Thus, the set of subterms which $\Val \ec{x}$ and $\mu v_{\ec{x}}. t$ differ 
-%of the form $\mu {\Varec{z}}. u$ and $v_{\ec{z}}$ respectively, where $v_{\ec{z}}$ is a bound variable in $\Val \ec{y}$.
-%In the latter case, $\Val \ec{y} \interp{ {\Varec{z}} } = \mu {\Varec{z}}. \const{C}( \Val \ec{y} \interp{\vec u} )$,
-%where $\mu v_{\ec{z}}. u$ and $\mu v_{\ec{z}}. u'$ were constructed by applying a sequence of substitutions of the form above to $v_{\ec{z}}$.
+%Thus, the set of subterms which $\Val \ec{x}$ and $\MU v_{\ec{x}}. t$ differ 
+%of the form $\MU {\Varec{z}}. u$ and $v_{\ec{z}}$ respectively, where $v_{\ec{z}}$ is a bound variable in $\Val \ec{y}$.
+%In the latter case, $\Val \ec{y} \interp{ {\Varec{z}} } = \MU {\Varec{z}}. \const{C}( \Val \ec{y} \interp{\vec u} )$,
+%where $\MU v_{\ec{z}}. u$ and $\MU v_{\ec{z}}. u'$ were constructed by applying a sequence of substitutions of the form above to $v_{\ec{z}}$.
 Since $\Val^\ast$ is obtained by applying the same substitutions and normalizations to all terms in the range of $\Val$,
-we have $\Val^\ast \ec{y} \interp{ \mu {\Varec{x}}. t} = \Val^\ast \ec{x}$.
+we have $\Val^\ast \ec{y} \interp{ \MU {\Varec{x}}. t} = \Val^\ast \ec{x}$.
 
@@
 If $t$ is a constructor term of the form $\const{C}( u_1, \ldots, u_n )$,
-then $\M( t )$ is $\alpha$-equivalent with $\nf{(\mu x. \const{C}( \M( u_1 ), \ldots, \M( u_n ) ) )}$ for fresh $x$,
-which by the induction hypothesis is $\alpha$-equivalent with $\nf{(\mu x. \const{C}( \Val^\ast \ec{ u_1 }, \ldots, \Val^\ast \ec{ u_n } ) )}$, call this term $t_u$.
+then $\M( t )$ is $\alpha$-equivalent with $\nf{(\MU x. \const{C}( \M( u_1 ), \ldots, \M( u_n ) ) )}$ for fresh $x$,
+which by the induction hypothesis is $\alpha$-equivalent with $\nf{(\MU x. \const{C}( \Val^\ast \ec{ u_1 }, \ldots, \Val^\ast \ec{ u_n } ) )}$, call this term $t_u$.
 By Lemma~\ref{lem:model-completion}.2 and due to our construction of $\Val^\ast$, we have $\Val^\ast \ec{t}$ is a term
-$t_w$ of the form $\mu {\Varec{t}}. \const{C}( w_1, \ldots, w_n )$
+$t_w$ of the form $\MU {\Varec{t}}. \const{C}( w_1, \ldots, w_n )$
 where $t_w\interp{w_i} = \Val^\ast \ec{u_i}$ for $i = 1, \ldots, n$.
@@
 By Lemma~\ref{lem:model-completion}.2 and due to our construction of $\Val^\ast$, we have $\Val^\ast \ec{u}$ is a term $t_w$ of the form
-$\mu {\Varec{u}}. C_i( w_1, \ldots, w_n )$,
+$\MU {\Varec{u}}. C_i( w_1, \ldots, w_n )$,
 where $t_w \interp{ w_j } = \Val^\ast \ec{ s^j_k( u )}$. 
@@
 We may interpret this partial assignment as a set of literals where $a \in M$ for all atoms $a$ that it assigned the value $\top$, and $\neg a \in M$ for all atoms that it assigns the value $\bot$.
-Then, at a high level, by the Nelson-Oppen combination framework,
+Then, at a high level, by the Nelson--Oppen combination framework,
 each $T_i$-solver for $i = 1, \ldots, n$ takes as input a combination of (1) the purified form of $T_i$-constraints occurring in $M$ where terms containing symbols not belonging to $T_i$ are replaced by fresh variables of the same type,
@@
 a set of equalities from $M_i$ that entail that the values of $\ec{x}$ in models of $\Ec$ are of the form $\Val \ec{x}$.
-For instance, if $x \teq \const{C}( x ) \in M_i$, then $x \teq \const{C}( x )$ is a (possible) explanation for $\Val \ec{x} = \mu {\Varec{x}}. \const{C}( {\Varec{x}} )$.
+For example, if $x \teq \const{C}( x ) \in M_i$, then $x \teq \const{C}( x )$ is a (possible) explanation for $\Val \ec{x} = \MU {\Varec{x}}. \const{C}( {\Varec{x}} )$.
 Thus, the rules of the calculus are implemented as follows.
@@
 
-    * sharing is caring
+    * sharing is caring (let's cite it for a change)
     * theory combination more complicated (and Theorem Completeness too weak)

### 2014-12-19 tex/codatadp.tex
more

@@
 If a selector is applied to a term $t$, or if $t$'s type is a finite datatype,
-the equivalence class of $t$ must contain an application of a $\delta$ constructor
-(where $\Ctr^\delta = \{ \const{C}_1, \ldots, \const{C}_m \}$).
+the equivalence class of $t$ must contain an application of one of
+$\delta$'s constructors
+$\Ctr^\delta = \{ \const{C}_1, \ldots, \const{C}_m \}$.
 This is enforced in the third phase by the \rn{Split} rule.
@@
 A derivation tree is \emph{closed} if all of its leaf nodes are $\bot$.
-A node is \emph{saturated} if no nonredundant application of a rule can be applied to it.
-If there exists a closed derivation tree with root node $\Ec$, then we conclude that $\Ec$ is $\thD$-unsatisfiable.
-If there exists a derivation tree with root node $\Ec$ that contains a saturated node, then we conclude that $\Ec$ is $\thD$-satisfiable.
+A node is \emph{saturated} if no nonredundant instance of a rule can be applied to it.
+%
+If there exists a closed derivation tree with root node $\Ec$, we conclude that $\Ec$ is $\thD$-unsatisfiable.
+If there exists a derivation tree with root node $\Ec$ that contains a saturated node, we conclude that $\Ec$ is $\thD$-satisfiable.
 
@@
 let $S^0_t$ be the set $\{ t \}$,
-for $i \geq 0$ let $S^{i+1}_t =$ $S^i_t \cup \{ s( u ) \mid u : \upsilon \in S^i, \upsilon \in \Data, \upsilon \text{ finite}, s \in \Sel^{\upsilon}  \}$,
+for $i \geq 0$ let $S^{i+1}_t =$ $S^i_t \mathrel\cup \{ s( u ) \mid u : \upsilon \in S^i, \upsilon \in \Data, \upsilon \text{ finite}, s \in \Sel^{\upsilon}  \}$,
 and let $S^\ast_t$ be the fixed point of this sequence.
@@
 Let $S^\ast$ be the union of $S^\ast_t$ for all $t \in T$,
-and let $\tcEc$ be the set of subterms of $\Ec \cup \{ C_i( s^1_i( t ), \ldots, s^{a_i}_i( t ) ) \mid t : \tau \in S \cup S^\ast, C_i \in \Ctr^\tau \}$.
+and let $\tcEc$ be the set of subterms of $\Ec \mathrel\cup \{ C_i( s^1_i( t ), \ldots, s^{a_i}_i( t ) ) \mid t : \tau \in S \mathrel\cup S^\ast, C_i \in \Ctr^\tau \}$.
 In a derivation tree with root node $\Ec$,
@@
 $\mathcal{T}(\Fc) \subseteq \tcEc$, and moreover thus contains an equality between two terms from $\tcEc$ not occurring in its parent node.
-Thus, the depth of a branch in a derivation tree with root node $\Ec$ is at most $\mid \tcEc \mid^2$,
+Thus, the depth of a branch in a derivation tree with root node $\Ec$ is at most $\left| \tcEc \right|^2$,
 which is finite since $\tcEc$ is finite.
@@
 whose root node is the result of applying the rule \rn{Split} on term $t$ of type $\tau$,
-then by the induction hypothesis $\Ec \cup t \teq C_i( s^1_i( t ), \ldots, s^{a_i}_i( t ) )$ is unsatisfiable
+then by the induction hypothesis $\Ec \mathrel\cup t \teq C_i( s^1_i( t ), \ldots, s^{a_i}_i( t ) )$ is unsatisfiable
 for each $C_i \in \Ctr^\tau$.
@@
 Otherwise, the child of $\Ec$ is a closed derivation tree
-whose root node is $\Ec \cup t_1 \teq t_2$ obtained by applying one of the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, \rn{Inject}, \rn{Unique}, or \rn{Singleton}.
+whose root node is $\Ec \mathrel\cup t_1 \teq t_2$ obtained by applying one of the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, \rn{Inject}, \rn{Unique}, or \rn{Singleton}.
 In all cases, we have that $\Ec \models_{\thD} t_1 \teq t_2$.
@@
 For \rn{Singleton}, clearly $t_1$ and $t_2$ must have the same value since the cardinality of their type is one.
-By the induction hypothesis, we have $\Ec \cup t_1 \teq t_2$ is $\thD$-unsatisfiable,
+By the induction hypothesis, we have $\Ec \mathrel\cup t_1 \teq t_2$ is $\thD$-unsatisfiable,
 and thus $\Ec$ is $\thD$-unsatisfiable.
@@
 \textstyle\begin{cases}
-    \MU x.\; \const{C}( t_1\interp{ \vec u}^{M \cup \{ x \}}) & \text{if }x \not\in M, x \mapsto \MU x. \const{C}( \vec u ) \in t_1  \\[-\jot]
+    \MU x.\; \const{C}( t_1\interp{ \vec u}^{M \mathrel\cup \{ x \}}) & \text{if }x \not\in M, x \mapsto \MU x. \const{C}( \vec u ) \in t_1  \\[-\jot]
     x & \text{otherwise}
@@
 \textstyle\begin{cases}
-    \MU x. \const{C}( t_1\interp{ \vec u }^{M \cup \{ x \}} ) & \text{if }x \not\in M \\[-\jot]
+    \MU x. \const{C}( t_1\interp{ \vec u }^{M \mathrel\cup \{ x \}} ) & \text{if }x \not\in M \\[-\jot]
     x & \text{otherwise}
@@
 that is, a specialized solver for determining the satisfiability of conjunctions of constraints for its theory.
-Given a theory $T = T_1 \cup \ldots T_n$ and a input set of clauses $F$ in CNF,
+Given a theory $T = T_1 \mathrel\cup \ldots T_n$ and a input set of clauses $F$ in CNF,
 the DPLL(T) \rem{CDCL(T)?} procedure

### 2014-12-19 tex/codatadp.tex
more

@@
 \text{Induction:}\quad
-\AXC{\strut$\bigwedge_{\,i,j}\, \forall x_1 \ldots \vthinspace x_{n_{i\negvthinspace j}}.\; \bigl(\bigwedge_{\,k}\, \mathit{IH}_{i\negvthinspace j}^k[x_k]\bigr) \longrightarrow P_i[C_{i\negvthinspace j}(x_1,\ldots,x_{n_{ij}})]$}
+\AXC{\strut$\bigwedge_{\,i,j}\, \forall x_1 \ldots \vthinspace x_{n_{i\negvthinspace j}}.\; \bigl(\bigwedge_{\,k}\, \mathit{IH}_{i\negvthinspace j}^k[x_k]\bigr) \longrightarrow P_i[\const C_{i\negvthinspace j}(x_1,\ldots,x_{n_{ij}})]$}
 \UIC{\strut$\bigwedge_{\,i}\, P_i[v_i]$}
@@
 Let $S^\ast$ be the union of $S^\ast_t$ for all $t \in T$,
-and let $\tcEc$ be the set of subterms of $\Ec \mathrel\cup \{ C_i( s^1_i( t ), \ldots, s^{a_i}_i( t ) ) \mid t : \tau \in S \mathrel\cup S^\ast, C_i \in \Ctr^\tau \}$.
+and let $\tcEc$ be the set of subterms of $\Ec \mathrel\cup \{ \const C_i( \const s^1_i( t ), \ldots, \const s^{a_i}_i( t ) ) \mid t : \tau \in S \mathrel\cup S^\ast, \const C_i \in \Ctr^\tau \}$.
 In a derivation tree with root node $\Ec$,
@@
 whose root node is the result of applying the rule \rn{Split} on term $t$ of type $\tau$,
-then by the induction hypothesis $\Ec \mathrel\cup t \teq C_i( s^1_i( t ), \ldots, s^{a_i}_i( t ) )$ is unsatisfiable
-for each $C_i \in \Ctr^\tau$.
-Since by exhaustiveness, all models of $\thD$ entail (exactly) one equality $t \teq C_i( s^1_i( t ), \ldots, s^{a_i}_i( t ) )$,
+then by the induction hypothesis $\Ec \mathrel\cup t \teq \const C_i( \const s^1_i( t ), \ldots, \const s^{a_i}_i( t ) )$ is unsatisfiable
+for each $\const C_i \in \Ctr^\tau$.
+Since by exhaustiveness, all models of $\thD$ entail (exactly) one equality $t \teq \const C_i( \const s^1_i( t ), \ldots, \const s^{a_i}_i( t ) )$,
 we know $\Ec$ is $\thD$-unsatisfiable.
@@
 \item
-for all selector terms $s^j_i( t )$ of type $\tau$,% for some $\vec u$,
-$\M( s^j_i( t ) )$ is the value
+for all selector terms $\const s^j_i( t )$ of type $\tau$,% for some $\vec u$,
+$\M( \const s^j_i( t ) )$ is the value
 in $\M( \tau )$ that is $\alpha$-equivalent to
@@
 
-If $t$ is a selector term of the form $s^j_k( u )$,
+If $t$ is a selector term of the form $\const s^j_k( u )$,
 since \rn{Split} does not apply,
-$\ec{u}$ must contain a term of the form $C_i( s^1_i( u ), \ldots, s^{n}_i( u ) )$ for some $i$.
+$\ec{u}$ must contain a term of the form $\const C_i( \const s^1_i( u ), \ldots, \const s^{n}_i( u ) )$ for some $i$.
 By Lemma~\ref{lem:model-completion}.2 and due to our construction of $\Val^\ast$, we have $\Val^\ast \ec{u}$ is a term $t_w$ of the form
-$\MU {\Varec{u}}. C_i( w_1, \ldots, w_n )$,
-where $t_w \interp{ w_j } = \Val^\ast \ec{ s^j_k( u )}$. 
-If $k = i$, then $\M( t )$ is $\alpha$-equivalent with $t_w\interp{w_j} = \Val^\ast \ec{ s^j_k( u )}$,
+$\MU {\Varec{u}}.\; \const C_i( w_1, \ldots, w_n )$,
+where $t_w \interp{ w_j } = \Val^\ast \ec{ \const s^j_k( u )}$. 
+If $k = i$, then $\M( t )$ is $\alpha$-equivalent with $t_w\interp{w_j} = \Val^\ast \ec{ \const s^j_k( u )}$,
 which is $\Val^\ast \ec{t}$.
-If $k \neq i$, then we interpret $s^j_k( \M( u ) )$ as the value in $\M( \tau )$ that is $\alpha$-equivalent with $\Val^\ast \ec{ t }$.
+If $k \neq i$, then we interpret $\const s^j_k( \M( u ) )$ as the value in $\M( \tau )$ that is $\alpha$-equivalent with $\Val^\ast \ec{ t }$.
 Since \rn{Cong} does not apply and due to Lemma~\ref{lem:model-completion}.3,
@@
 Discriminators are treated as predicate symbols, so a discriminator $\const{d}_i( t )$ is such that $\const{d}_i$
-is a predicate symbol, instead of reducing this constraint to $t \teq \const{C}_i( s^1_i( t ), \ldots, s^n_i( t ) )$.
+is a predicate symbol, instead of reducing this constraint to $t \teq \const{C}_i( \const s^1_i( t ), \ldots, \const s^n_i( t ) )$.
 We have found that this leads to better performance since this reduction introduces terms more eagerly to $\tEc$.
@@
 Also, selectors are collapsed eagerly:
-if $s^j_i( t )$ is a term in $\tEc$ and $t$ contains the constructor
-$\const{C}_i( \vec u )$, then we immediately infer $s^j_i( t ) \teq u_j$, whereas the calculus we presented would apply \rn{Split} and \rn{Inject} before inferring this equality.
+if $\const s^j_i( t )$ is a term in $\tEc$ and $t$ contains the constructor
+$\const{C}_i( \vec u )$, then we immediately infer $\const s^j_i( t ) \teq u_j$, whereas the calculus we presented would apply \rn{Split} and \rn{Inject} before inferring this equality.
 For the sake of reducing the number of unique constraints considered by the calculus, we compute a normal form for constraints as a preprocessing step.
@@
 For the \rn{Split} rule, we add a clause of the form
-$(\const{C}_1( s^1_1( t ), \ldots, s^{a_1}_1( t ) ) \vee \ldots \vee t \teq \const{C}_n( s^1_n( t ), \ldots, s^{a_n}_n( t ) ))$
+$(\const{C}_1( \const s^1_1( t ), \ldots, \const s^{a_1}_1( t ) ) \vee \ldots \vee t \teq \const{C}_n( \const s^1_n( t ), \ldots, \const s^{a_n}_n( t ) ))$
 to $F$.

### 2014-12-19 tex/codatadp.tex
repair kerning

@@
 
+\newcommand\SSS{\mathit{S\vthinspace}}
+\newcommand\SSSS[1]{\mathit{S}^{\,#1}}
+
 % change?
@@
 %
-The rules operate on a set $\Ec$ of equalities and disequalities
+The rules operate on a finite set $\Ec$ of equalities and disequalities
 between $\Sig$-terms, where $\Ec$ is interpreted as a conjunction.
@@
 conclusion does not add a new equality to $\Ec$ and is not $\bot$.
-A \emph{derivation tree} is a tree whose nodes are sets of equalities, where non-root nodes are obtained by
+A \emph{derivation tree} is a tree whose nodes are finite sets of equalities, where non-root nodes are obtained by
 a nonredundant application of a derivation rule to its parent node.
@@
 Consider a derivation tree with root node $\Ec$.
-Let $S$ be the set of terms occurring as the argument of a selector in $\tEc$.
-Let $T$ be the set of terms in $\tEc$ of finite datatype type.
+Let $\SSS$ be the set of terms occurring as the argument of a selector in $\tEc$.
+Let $T$ be the set of terms in $\tEc$ whose types are finite datatype.
 For each term $t \in T$,
-let $S^0_t$ be the set $\{ t \}$,
-for $i \geq 0$ let $S^{i+1}_t =$ $S^i_t \mathrel\cup \{ s( u ) \mid u : \upsilon \in S^i, \upsilon \in \Data, \upsilon \text{ finite}, s \in \Sel^{\upsilon}  \}$,
-and let $S^\ast_t$ be the fixed point of this sequence.
+let $\SSSS 0_t$ be the set $\{ t \}$,
+for $i \geq 0$ let $\SSSS {i+1}_t =$ $\SSSS i_t \mathrel\cup \{ \const s( u ) \mid u : \upsilon \in \SSSS i, \upsilon \in \Data, \upsilon \text{ finite}, s \in \Sel^{\upsilon}  \}$,
+and let $\SSSS \ast_t$ be the fixed point of this sequence.
 We know this is a finite set for each $t$ since the values of the type of $t$ are of finite size.
-Let $S^\ast$ be the union of $S^\ast_t$ for all $t \in T$,
-and let $\tcEc$ be the set of subterms of $\Ec \mathrel\cup \{ \const C_i( \const s^1_i( t ), \ldots, \const s^{a_i}_i( t ) ) \mid t : \tau \in S \mathrel\cup S^\ast, \const C_i \in \Ctr^\tau \}$.
+Let $\SSSS \ast$ be the union of $\SSSS \ast_t$ for all $t \in T$,
+and let $\tcEc$ be the set of subterms of $\Ec \mathrel\cup \{ \const C_i( \const s^1_i( t ), \ldots, \const s^{a_i}_i( t ) ) \mid t : \tau \in \SSS \mathrel\cup \SSSS \ast, \const C_i \in \Ctr^\tau \}$.
 In a derivation tree with root node $\Ec$,

### 2014-12-19 tex/codatadp.tex
tuning

@@
 %(Section~\ref{sec:introduction}).
-% for example, $\MU x. \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
+% for example, $\MU x.\; \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
 A $\mu$-term $t$ of type $\tau$ describes a class of values of
@@
 %and $\vec v$ are well-typed $\mu$-terms of type $\vec \tau$ under the assumption that $x$ has type $\upsilon$,
-%then $\MU x. \const{C}( \vec v )$ is a well-typed term of type $\upsilon$.
+%then $\MU x.\; \const{C}( \vec v )$ is a well-typed term of type $\upsilon$.
 
@@
 are syntactically equivalent for some capture-avoiding renaming of $\mu$-bound variables---e.g.,
-%$\MU x. \const{C}( x ) \aequiv \MU y. \const{C}( y )$ and
+%$\MU x.\; \const{C}( x ) \aequiv \MU y.\; \const{C}( y )$ and
 $\MU x.\; \const{D}( y, x ) \aequiv \MU z.\; \const{D}( y, z )$\begin{report},
@@
 the \emph{interpretation of $t_2$ in $t_1$} is the $\mu$-term $t_1\interp{t_2}^\emptyset$ as returned by the following recursive procedure,
-where $x \mapsto \MU x. \const{C}( \vec u ) \in t_1$ indicates that $\MU x. \const{C}( \vec u )$ is the subterm of $t_1$ that binds this occurrence of variable $x$.
+where $x \mapsto \MU x.\; \const{C}( \vec u ) \in t_1$ indicates that $\MU x.\; \const{C}( \vec u )$ is the subterm of $t_1$ that binds this occurrence of variable $x$.
 \[\begin{array}{r@{}c@{}l}
@@
 \textstyle\begin{cases}
-    \MU x.\; \const{C}( t_1\interp{ \vec u}^{M \mathrel\cup \{ x \}}) & \text{if }x \not\in M, x \mapsto \MU x. \const{C}( \vec u ) \in t_1  \\[-\jot]
+    \MU x.\; \const{C}( t_1\interp{ \vec u}^{M \mathrel\cup \{ x \}}) & \text{if }x \not\in M, x \mapsto \MU x.\; \const{C}( \vec u ) \in t_1  \\[-\jot]
     x & \text{otherwise}
 \end{cases} \\
-t_1\interp{\MU x. \const{C}( \vec u )}^M & = &
+t_1\interp{\MU x.\; \const{C}( \vec u )}^M & = &
 \textstyle\begin{cases}
-    \MU x. \const{C}( t_1\interp{ \vec u }^{M \mathrel\cup \{ x \}} ) & \text{if }x \not\in M \\[-\jot]
+    \MU x.\; \const{C}( t_1\interp{ \vec u }^{M \mathrel\cup \{ x \}} ) & \text{if }x \not\in M \\[-\jot]
     x & \text{otherwise}
@@
 
-Given a $\mu$-term $t_1$ of the form $\MU x. \const{C}( u_1, \ldots, u_n )$ with subterm $t_2$,
+Given a $\mu$-term $t_1$ of the form $\MU x.\; \const{C}( u_1, \ldots, u_n )$ with subterm $t_2$,
 we say $t_2$ is a \emph{bisimilar subterm} of $t_1$ if and only if it is of the form
-$\MU y. \const{C}( v_1, \ldots, v_n )$ and $t_1\interp{ u_i } \aequiv t_1\interp{ v_i }$ for $i = 1,\ldots, n$.
+$\MU y.\; \const{C}( v_1, \ldots, v_n )$ and $t_1\interp{ u_i } \aequiv t_1\interp{ v_i }$ for $i = 1,\ldots, n$.
 The term $t_1$ is \emph{normal} if it does not contain a strict, bisimilar subterm
 and all of its strict subterms are also normal.
-For example, the $\mu$-term $t_1$ of the form $\MU x. \const{C}( \MU y. \const{C}( y ) )$ is not normal,
-since $\MU y. \const{C}( y )$ is a bisimilar subterm of $t_1$.
+For example, the $\mu$-term $t_1$ of the form $\MU x.\; \const{C}( \MU y.\; \const{C}( y ) )$ is not normal,
+since $\MU y.\; \const{C}( y )$ is a bisimilar subterm of $t_1$.
 In particular, note that their arguments have the same interpretation in $t_1$,
 that is,
-$t_1\interp{\MU y. \const{C}( y )}^\emptyset =$
-$\MU y. \const{C}( t_1\interp{y}^{\{y\}} ) =$
-$\MU y. \const{C}( y )$
+$t_1\interp{\MU y.\; \const{C}( y )}^\emptyset =$
+$\MU y.\; \const{C}( t_1\interp{y}^{\{y\}} ) =$
+$\MU y.\; \const{C}( y )$
 is syntactically equivalent to
 $t_1\interp{y}^\emptyset =$
-$\MU y. \const{C}( t_1\interp{y}^{\{y\}} ) =$
-$\MU y. \const{C}( y )$.
-The $\mu$-term $t_1$ of the form $\MU x. \const{C}( \MU y. \const{C}( x ) )$ is also not normal,
-since $\MU y. \const{C}( x )$ is a bisimilar subterm of $t_1$,
+$\MU y.\; \const{C}( t_1\interp{y}^{\{y\}} ) =$
+$\MU y.\; \const{C}( y )$.
+The $\mu$-term $t_1$ of the form $\MU x.\; \const{C}( \MU y.\; \const{C}( x ) )$ is also not normal,
+since $\MU y.\; \const{C}( x )$ is a bisimilar subterm of $t_1$,
 noting that
-$t_1\interp{\MU y. \const{C}( x )}^\emptyset =$
-$\MU y. \const{C}( t_1\interp{x}^{\{y\}} ) =$
-$\MU y. \const{C}( t_1\interp{\MU x. \const{C}( \MU y. \const{C}( x ) ) }^{\{y\}} ) =$
-$\MU y. \const{C}( \MU x. \const{C}( t_1\interp{\MU y. \const{C}( x )}^{\{x,y\}} ) ) =$
-$\MU y. \const{C}( \MU x. \const{C}( y ) )$
+$t_1\interp{\MU y.\; \const{C}( x )}^\emptyset =$
+$\MU y.\; \const{C}( t_1\interp{x}^{\{y\}} ) =$
+$\MU y.\; \const{C}( t_1\interp{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) ) }^{\{y\}} ) =$
+$\MU y.\; \const{C}( \MU x.\; \const{C}( t_1\interp{\MU y.\; \const{C}( x )}^{\{x,y\}} ) ) =$
+$\MU y.\; \const{C}( \MU x.\; \const{C}( y ) )$
 is syntactically equivalent to $t_1\interp{x}^\emptyset = t_1$ for the renaming $\{ x \mapsto y, y \mapsto x \}$.
 
-For any $\mu$-term $t$ of the form $\MU x. C( \vec u )$, we may recursively construct a normal $\mu$-term $\nf{t}$
+For any $\mu$-term $t$ of the form $\MU x.\; \const C( \vec u )$, we may recursively construct a normal $\mu$-term $\nf{t}$
 (which we call the \emph{normal form} of $t$)
@@
 and by normalizing $t$'s subterms.
-For example, $\nf{(\MU x. \const{C}( \MU y. \const{C}( x ) ))} = \MU x. \const{C}( x )$.
+For example, $\nf{(\MU x.\; \const{C}( \MU y.\; \const{C}( x ) ))} = \MU x.\; \const{C}( x )$.
 
@@
 \label{lem:mu-norm-arg}
-If $\vec u$ are normal, then if $t = \nf{(\MU x. \const{C}( \vec u ) )} = \MU x. \const{C}( \vec w )$,
+If $\vec u$ are normal, then if $t = \nf{(\MU x.\; \const{C}( \vec u ) )} = \MU x.\; \const{C}( \vec w )$,
 then $\vec u \vsim t\interp{\vec w}$.
@@
 \label{lem:mu-cong}
-If $t_u = \MU x. \const{C}( \vec u )$ and $t_w = \MU y. \const{C}( \vec w )$ are normal,
+If $t_u = \MU x.\; \const{C}( \vec u )$ and $t_w = \MU y.\; \const{C}( \vec w )$ are normal,
 then $t_u\interp{ \vec u } \vsim t_w\interp{ \vec w }$ if and only if $t_u \vsim t_w$.
@@
 in $\M( \tau )$ that is $\alpha$-equivalent to
-$\nf{(\MU x. \const{C}( \M( \vec t \,) ))}$ where $x$ is fresh, and
+$\nf{(\MU x.\; \const{C}( \M( \vec t \,) ))}$ where $x$ is fresh, and
 \item
@@
 $\M( t )\interp{u_j}$
-when $\M( t )$ is of the form $\MU x. \const{C}_i( \vec u )$.
+when $\M( t )$ is of the form $\MU x.\; \const{C}_i( \vec u )$.
 \end{enumerate}
@@
 For example,
-$\MU x. \const{D}( x ) \vsimv{y} \MU x. \const{D}( y )$,
-$\MU x. \const{C}( x, x ) \vsimv{x} \MU y. \const{C}( x, y )$ and
-$\MU x. \const{C}( z, x ) \vsimv{z} \MU y. \const{C}( z, y )$,
+$\MU x.\; \const{D}( x ) \vsimv{y} \MU x.\; \const{D}( y )$,
+$\MU x.\; \const{C}( x, x ) \vsimv{x} \MU y.\; \const{C}( x, y )$ and
+$\MU x.\; \const{C}( z, x ) \vsimv{z} \MU y.\; \const{C}( z, y )$,
 but
-$\MU x. \const{D}( x ) \not\vsimv{x} \MU x. \const{D}( y )$ and
-$\MU x. \const{C}( x, x ) \not\vsimv{y} \MU y. \const{C}( x, y )$.
+$\MU x.\; \const{D}( x ) \not\vsimv{x} \MU x.\; \const{D}( y )$ and
+$\MU x.\; \const{C}( x, x ) \not\vsimv{y} \MU y.\; \const{C}( x, y )$.
 For a variable $x$ of type $\tau$ and a fixed normal model $\M$,
@@
   \quad
-  (\MU x. t) \aequiv v
+  (\MU x.\; t) \aequiv v
   \quad
@@
 }{
-  \Val := \nf{(\Val \{ x \mapsto \MU x. t \})}
+  \Val := \nf{(\Val \{ x \mapsto \MU x.\; t \})}
 }
@@
 $\}$.
-After running the calculus to saturation on $\Ec$, the mapping $\Val$ contains:
-\[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{2.0em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
-\Val \ec{y_1} & = & \MU {\Varec{y_1}}. & \const{A}( & \MU {\Varec{y_2}}. & \const{B}( {\Varec{x}} ) ) &
-\Val \ec{w_1} & = & \MU {\Varec{w_1}}. & \const{A}( {\Varec{x}} )      & \\
+After running the calculus to saturation on $\Ec$, the mapping $\Val$ contains
+\[\begin{array}{r@{}c@{}rlll@{\qquad}r@{}c@{}lllll}
+\Val \ec{y_1} & {}\,=\,{} & \MU {\Varec{y_1}}. & \const{A}( & \MU {\Varec{y_2}}. & \const{B}( {\Varec{x}} ) ) &
+\Val \ec{w_1} & {}\,=\,{} & \MU {\Varec{w_1}}. & \const{A}( {\Varec{x}} )      & \\
 \Val \ec{y_2} & = & \MU {\Varec{y_2}}. & \multicolumn{2}{l}{\const{B}( {\Varec{x}} )}      & &
@@
 To construct a completion of $\Val$, we must choose values for the free variables of $\Val$, ${\Varec{x}}$ and ${\Varec{w_3}}$.
-The content of the set $\mathcal{T}^\M_{{\Varec{x}}}( \Val )$ is $\alpha$-equivalent to:
+The content of the set $\mathcal{T}^\M_{{\Varec{x}}}( \Val )$ is $\alpha$-equivalent to
 \[%\begin{array}{c}
 \{
-\MU x. \const{A}( \const{B}( x )),
-\const{A}( \MU y. \const{B}( y )),
-\MU x. \const{B}( x ),
-\MU x. \const{C}( \const{A}( x )),
-\MU x. \const{A}( \const{C}( x )),
-\MU x. \const{A}( x )
+\MU x.\; \const{A}( \const{B}( x )),
+\const{A}( \MU y.\; \const{B}( y )),
+\MU x.\; \const{B}( x ),
+\MU x.\; \const{C}( \const{A}( x )),
+\MU x.\; \const{A}( \const{C}( x )),
+\MU x.\; \const{A}( x )
 \}
@@
 which in turn may cause $\M$ to be inconsistent.
-For example, say we assign $\MU {\Varec{x}}. \const{B}( {\Varec{x}} )$ for ${\Varec{x}}$.
-After this substitution, $\Val \ec{y_2}$ is $\MU {\Varec{y_2}}. \const{B}( \MU {\Varec{x}}. \const{B}( {\Varec{x}} ) )$,
-which has normal form $\MU {\Varec{y_2}}. \const{B}( {\Varec{y_2}} )$,
-which is $\alpha$-equivalent to $\MU {\Varec{x}}. \const{B}( {\Varec{x}} )$.
+For example, say we assign $\MU {\Varec{x}}.\; \const{B}( {\Varec{x}} )$ for ${\Varec{x}}$.
+After this substitution, $\Val \ec{y_2}$ is $\MU {\Varec{y_2}}.\; \const{B}( \MU {\Varec{x}}.\; \const{B}( {\Varec{x}} ) )$,
+which has normal form $\MU {\Varec{y_2}}.\; \const{B}( {\Varec{y_2}} )$,
+which is $\alpha$-equivalent to $\MU {\Varec{x}}.\; \const{B}( {\Varec{x}} )$.
 However, this contradicts the disequality $x \not\teq y_2$ in $\Ec$.
@@
 For example, we may assign values such as 
-$\MU {\Varec{x}}. \const{C}( {\Varec{x}} )$,
-$\MU {\Varec{x}}. \const{C}( \MU x_1. \const{A}( x_1) )$, or
-$\MU {\Varec{x}}. \const{B}( \const{A}( {\Varec{x}} ) )$
+$\MU {\Varec{x}}.\; \const{C}( {\Varec{x}} )$,
+$\MU {\Varec{x}}.\; \const{C}( \MU x_1.\; \const{A}( x_1) )$, or
+$\MU {\Varec{x}}.\; \const{B}( \const{A}( {\Varec{x}} ) )$
 to ${\Varec{x}}$.
@@
 After applying the substitution for the third term mentioned above,
-$\Val \ec{w_1}$ is $\MU {\Varec{w_1}}. \const{A}( \MU {\Varec{x}}. \const{B}( \const{A}( {\Varec{x}} ) ) )$,
+$\Val \ec{w_1}$ is $\MU {\Varec{w_1}}.\; \const{A}( \MU {\Varec{x}}.\; \const{B}( \const{A}( {\Varec{x}} ) ) )$,
 which has normal form
-$\MU {\Varec{w_1}}. \const{A}( \MU {\Varec{x}}. \const{B}( {\Varec{w_1}} ) )$.
+$\MU {\Varec{w_1}}.\; \const{A}( \MU {\Varec{x}}.\; \const{B}( {\Varec{w_1}} ) )$.
 \xend
@@
 %We first state the following properties of $\Val^\ast$.
-%where we write $\muvar( t )$ to denote $t$ if $t$ is a variable, or $x$ if $t$ is of the form $\MU x. u$.
+%where we write $\muvar( t )$ to denote $t$ if $t$ is a variable, or $x$ if $t$ is of the form $\MU x.\; u$.
 
@@
 \item
-All subterms $\MU {\Varec{x}}. t$ of $\Val^\ast \ec{y}$ are such that
-$\Val^\ast \ec{y} \interp{ \MU {\Varec{x}}. t } = \Val^\ast \ec{x}$. 
+All subterms $\MU {\Varec{x}}.\; t$ of $\Val^\ast \ec{y}$ are such that
+$\Val^\ast \ec{y} \interp{ \MU {\Varec{x}}.\; t } = \Val^\ast \ec{x}$. 
 %If $\const{C}( t_1, \ldots, t_n ) \in \ec{x}$,
-%then all subterms $\MU v_{\ec{x}}. u$ occurring in $\Val^\ast \ec{y}$ 
-%are of the form $\MU v_{\ec{x}}. \const{C}( w_1, \ldots, w_n )$,
+%then all subterms $\MU v_{\ec{x}}.\; u$ occurring in $\Val^\ast \ec{y}$ 
+%are of the form $\MU v_{\ec{x}}.\; \const{C}( w_1, \ldots, w_n )$,
 %where 
 %and $\Val^\ast \ec{y} \interp{ w_i } = \Val^\ast \ec{t_i}$ for $i = 1, \ldots, n$.
-%If $\Val^\ast \ec{x}$ contains a subterm of the form $\MU v_{\ec{x}}. \const{C}( w_1, \ldots, w_n )$,
+%If $\Val^\ast \ec{x}$ contains a subterm of the form $\MU v_{\ec{x}}.\; \const{C}( w_1, \ldots, w_n )$,
 %and $\muvar( w_i ) = v_{\ec{t_i}}$ for $i = 1, \ldots, n$,
@@
 $\Val$ was constructed by applying a sequence of substitutions
-where ${\Varec{z}}$ was replaced only by $\MU {\Varec{z}}. \const{C}_i( {\Varec{t_1}}, \ldots, {\Varec{t_n}} )$.
-Assume $\Val \ec{y}$ has a subterm $\MU {\Varec{x}}. \const{C}( \vec t )$.
-We first show that our construction of $\Val$ is such that $\Val \ec{x} = \Val \ec{y}\interp{ \MU {\Varec{x}}. t }$.
-Notice that both $\Val \ec{x}$ and the subterm $\MU {\Varec{x}}. t$ of $\Val \ec{y}$ were
+where ${\Varec{z}}$ was replaced only by $\MU {\Varec{z}}.\; \const{C}_i( {\Varec{t_1}}, \ldots, {\Varec{t_n}} )$.
+Assume $\Val \ec{y}$ has a subterm $\MU {\Varec{x}}.\; \const{C}( \vec t )$.
+We first show that our construction of $\Val$ is such that $\Val \ec{x} = \Val \ec{y}\interp{ \MU {\Varec{x}}.\; t }$.
+Notice that both $\Val \ec{x}$ and the subterm $\MU {\Varec{x}}.\; t$ of $\Val \ec{y}$ were
 constructed by applying a sequence of substitutions of the form mentioned above to ${\Varec{x}}$.
@@
 \rem{TODO}
-%Thus, the set of subterms which $\Val \ec{x}$ and $\MU v_{\ec{x}}. t$ differ 
-%of the form $\MU {\Varec{z}}. u$ and $v_{\ec{z}}$ respectively, where $v_{\ec{z}}$ is a bound variable in $\Val \ec{y}$.
-%In the latter case, $\Val \ec{y} \interp{ {\Varec{z}} } = \MU {\Varec{z}}. \const{C}( \Val \ec{y} \interp{\vec u} )$,
-%where $\MU v_{\ec{z}}. u$ and $\MU v_{\ec{z}}. u'$ were constructed by applying a sequence of substitutions of the form above to $v_{\ec{z}}$.
+%Thus, the set of subterms which $\Val \ec{x}$ and $\MU v_{\ec{x}}.\; t$ differ 
+%of the form $\MU {\Varec{z}}.\; u$ and $v_{\ec{z}}$ respectively, where $v_{\ec{z}}$ is a bound variable in $\Val \ec{y}$.
+%In the latter case, $\Val \ec{y} \interp{ {\Varec{z}} } = \MU {\Varec{z}}.\; \const{C}( \Val \ec{y} \interp{\vec u} )$,
+%where $\MU v_{\ec{z}}.\; u$ and $\MU v_{\ec{z}}.\; u'$ were constructed by applying a sequence of substitutions of the form above to $v_{\ec{z}}$.
 Since $\Val^\ast$ is obtained by applying the same substitutions and normalizations to all terms in the range of $\Val$,
-we have $\Val^\ast \ec{y} \interp{ \MU {\Varec{x}}. t} = \Val^\ast \ec{x}$.
+we have $\Val^\ast \ec{y} \interp{ \MU {\Varec{x}}.\; t} = \Val^\ast \ec{x}$.
 
@@
 If $t$ is a constructor term of the form $\const{C}( u_1, \ldots, u_n )$,
-then $\M( t )$ is $\alpha$-equivalent with $\nf{(\MU x. \const{C}( \M( u_1 ), \ldots, \M( u_n ) ) )}$ for fresh $x$,
-which by the induction hypothesis is $\alpha$-equivalent with $\nf{(\MU x. \const{C}( \Val^\ast \ec{ u_1 }, \ldots, \Val^\ast \ec{ u_n } ) )}$, call this term $t_u$.
+then $\M( t )$ is $\alpha$-equivalent with $\nf{(\MU x.\; \const{C}( \M( u_1 ), \ldots, \M( u_n ) ) )}$ for fresh $x$,
+which by the induction hypothesis is $\alpha$-equivalent with $\nf{(\MU x.\; \const{C}( \Val^\ast \ec{ u_1 }, \ldots, \Val^\ast \ec{ u_n } ) )}$, call this term $t_u$.
 By Lemma~\ref{lem:model-completion}.2 and due to our construction of $\Val^\ast$, we have $\Val^\ast \ec{t}$ is a term
-$t_w$ of the form $\MU {\Varec{t}}. \const{C}( w_1, \ldots, w_n )$
+$t_w$ of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$
 where $t_w\interp{w_i} = \Val^\ast \ec{u_i}$ for $i = 1, \ldots, n$.
@@
 a set of equalities from $M_i$ that entail that the values of $\ec{x}$ in models of $\Ec$ are of the form $\Val \ec{x}$.
-For example, if $x \teq \const{C}( x ) \in M_i$, then $x \teq \const{C}( x )$ is a (possible) explanation for $\Val \ec{x} = \MU {\Varec{x}}. \const{C}( {\Varec{x}} )$.
+For example, if $x \teq \const{C}( x ) \in M_i$, then $x \teq \const{C}( x )$ is a (possible) explanation for $\Val \ec{x} = \MU {\Varec{x}}.\; \const{C}( {\Varec{x}} )$.
 Thus, the rules of the calculus are implemented as follows.

### 2014-12-19 tex/codatadp.tex
More work on model completion lemma.

--- 2014-12-19 tex/codatadp.tex
@@
 \usepackage[all]{xy}
-
-\makeatletter
-\def\spnXwtheorem#1#2#3#4{\@spynthm{#1}{#2}{#3}{#4}%
-                         \@addtoreset{#1}{chapter}}%
-\makeatother
-
-\newcommand\MU{\mu\vvthinspace}
-
-%%% not part of shared counter
-\spnXwtheorem{examplex}{Example}{\itshape}{\rmfamily}
 
@@
 
-\newcommand\vvthinspace{\kern+0.041667em}
 \newcommand\vthinspace{\kern+0.083333em}
@@
 in the metatheory. For example,
-the $\mu$-term $\const{SCons}(1,\: \MU s.\; \const{SCons}(0,\: \const{SCons}(9,\: s)))$
+the $\mu$-term $\const{SCons}(1,\: \mu s.\; \const{SCons}(0,\: \const{SCons}(9,\: s)))$
 stands for the lasso-shaped sequence $1, 0, 9, 0, 9, 0, 9, \ldots$\afterLdots.
@@
 %    * mutual recursion, but no polymorphism, nested recursion
-%  * integrated with Nelson--Oppen
+%  * integrated with Nelson-Oppen
 
@@
 \text{Induction:}\quad
-\AXC{\strut$\bigwedge_{\,i,j}\, \forall x_1 \ldots \vthinspace x_{n_{i\negvthinspace j}}.\; \bigl(\bigwedge_{\,k}\, \mathit{IH}_{i\negvthinspace j}^k[x_k]\bigr) \longrightarrow P_i[\const C_{i\negvthinspace j}(x_1,\ldots,x_{n_{ij}})]$}
+\AXC{\strut$\bigwedge_{\,i,j}\, \forall x_1 \ldots \vthinspace x_{n_{i\negvthinspace j}}.\; \bigl(\bigwedge_{\,k}\, \mathit{IH}_{i\negvthinspace j}^k[x_k]\bigr) \longrightarrow P_i[C_{i\negvthinspace j}(x_1,\ldots,x_{n_{ij}})]$}
 \UIC{\strut$\bigwedge_{\,i}\, P_i[v_i]$}
@@
 \vthinspace$\keyw{codatatype}~\ty{a} = \const{A}(\ty{a})$, whose
-only value is $\MU a.\; A(a)$. Other specimens are
+only value is $\mu a.\; A(a)$. Other specimens are
 \begin{paper}%
@@
 
-\begin{lemma}%[Corecursive Singletons]%
+\begin{lemma}[Corecursive Singletons]%
 \label{lem:corecursive-singletons}%
@@
 The calculus consists of three sets of rules, given in \figuresname~\ref{fig:cc-rules} to
-\ref{fig:split-rule}, corresponding to three phases. The first
+\ref{fig:split-rule}, corresponding to three successive phases. The first
 phase computes the bidirectional closure of $\Ec$. The second phase makes
@@
 A rule may have multiple conclusions separated by $\parallel$,
-denoting nondeterministic branching. The rules associated with a phase have
-priority over those of the previous phases. The rules are applied until
-saturation (i.e., until no rule can be applied to extend~$\Ec$
-further).
+denoting nondeterministic branching. The rules associated with a phase are
+applied until saturation (i.e., until no rule can be applied to extend~$\Ec$
+further) before the next phase can begin.
 
@@
   %\ttpath{\Ec}{\ec{t}}{\ec{t}} \neq \emptyset
-  \Val \ec{t} = \MU x.\; u
+  \Val \ec{t} = \mu x.\; u
   \quad
@@
 %(Section~\ref{sec:introduction}).
-% for example, $\MU x. \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
+% for example, $\mu x. \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
 A $\mu$-term $t$ of type $\tau$ describes a class of values of
@@
 or a node of the form
-$\MU x.\; \const{C}( \vec t\, )$ for some constructor $\const{C} \in \Ctr$ and
+$\mu x.\; \const{C}( \vec t\, )$ for some constructor $\const{C} \in \Ctr$ and
 $\mu$-terms $\vec t$ with the expected types.
@@
 %
-%For uniformity, we also consider $%\MU x.\;
+%For uniformity, we also consider $%\mu x.\;
 %\const{c}$ to be a $\mu$-term if $\const{c}$ is a constant of type $\tau \in \Nondata$.
@@
 %and $\vec v$ are well-typed $\mu$-terms of type $\vec \tau$ under the assumption that $x$ has type $\upsilon$,
-%then $\MU x. \const{C}( \vec v )$ is a well-typed term of type $\upsilon$.
+%then $\mu x. \const{C}( \vec v )$ is a well-typed term of type $\upsilon$.
 
@@
 are syntactically equivalent for some capture-avoiding renaming of $\mu$-bound variables---e.g.,
-%$\MU x. \const{C}( x ) \aequiv \MU y. \const{C}( y )$ and
-$\MU x.\; \const{D}( y, x ) \aequiv \MU z.\; \const{D}( y, z )$\begin{report},
+%$\mu x. \const{C}( x ) \aequiv \mu y. \const{C}( y )$ and
+$\mu x.\; \const{D}( y, x ) \aequiv \mu z.\; \const{D}( y, z )$\begin{report},
 but
-$\MU x.\; \const{C}( x ) \not\aequiv \MU x.\; \const{D}( y, x ) \not\aequiv \MU x.\; \const{D}( z, x )
-\not\aequiv \MU y.\; \const{D}( y, x )$\end{report}.
+$\mu x.\; \const{C}( x ) \not\aequiv \mu x.\; \const{D}( y, x ) \not\aequiv \mu x.\; \const{D}( z, x )
+\not\aequiv \mu y.\; \const{D}( y, x )$\end{report}.
 Two $\mu$-terms can represent the same value while not being $\alpha$-equivalent---e.g.,
-$\MU x.\; \const{C}( x ) \not\aequiv \MU x.\; \const{C}( \const{C}( x ) )$.
+$\mu x.\; \const{C}( x ) \not\aequiv \mu x.\; \const{C}( \const{C}( x ) )$.
 %, or informally are \emph{observationally equivalent}.
@@
 }{
-  \Val := \Val \{ {\Varec{x}} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{t_1}}, \ldots, {\Varec{t_n}} ) \}
+  \Val := \Val \{ {\Varec{x}} \mapsto \mu {\Varec{x}}.\; \const{C}( {\Varec{t_1}}, \ldots, {\Varec{t_n}} ) \}
 }
@@
 
-\begin{examplex}
+\begin{example}
 Suppose that $\Ec$ contains distinct equivalence classes $\ec{w}$, $\ec{x}$, $\ec{y}$, and $\ec{z}$
@@
 \item \noindent\rlap{Initially:}\phantom{Unfold $\Varec{x}$:\enskip}$\{  \}$
-\item \noindent\rlap{Unfold $\Varec{x}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) \vthinspace\}$
-\item \noindent\rlap{Unfold $\Varec{y}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\;
-  \ec{y} \mapsto \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) \vthinspace\}$
-\item \noindent\rlap{Unfold $\Varec{x}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\;
-  \ec{y} \mapsto \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\; \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) ) \vthinspace\}$
+\item \noindent\rlap{Unfold $\Varec{x}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \mu {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) \vthinspace\}$
+\item \noindent\rlap{Unfold $\Varec{y}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \mu {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \mu {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\;
+  \ec{y} \mapsto \mu {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) \vthinspace\}$
+\item \noindent\rlap{Unfold $\Varec{x}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \mu {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \mu {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\;
+  \ec{y} \mapsto \mu {\Varec{y}}.\; \const{C}( {\Varec{z}},\; \mu {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) ) \vthinspace\}$
 \end{enumerate}
@@
 \xend
-\end{examplex}
+\end{example}
 
 Given $\Val$, the rules for acyclicity and uniqueness can be stated as follows.
-For acyclicity, if there is an equivalence class $\ec{t}$ of datatype type whose class of values $\Val \ec{t} = \MU x.\; u$ are cyclic,
+For acyclicity, if there is an equivalence class $\ec{t}$ of datatype type whose class of values $\Val \ec{t} = \mu x.\; u$ are cyclic,
 as indicated by the assumption $x \in \FV( u )$,
@@
 
-\begin{examplex}
+\begin{example}
 Let $\const{C} : $ $\tau \times \delta \rightarrow \delta$ be a constructor,
@@
 \Val \ec{x} & \,=\, &
-\MU {\Varec{x}}. & \const{C}( {\Varec{y}},\, & \smash{\MU {\VAREC{\const{C}(y, x)}}}. & \const{C}( {\Varec{y}},\, {\Varec{x}} ) ) \\
+\mu {\Varec{x}}. & \const{C}( {\Varec{y}},\, & \mu {\VAREC{\const{C}(y, x)}}. & \const{C}( {\Varec{y}},\, {\Varec{x}} ) ) \\
 \Val \ec{\const{C}(y,x)} & = &
-\MU {\VAREC{\const{C}(y,x)}}. & \const{C}( {\Varec{y}},\, & \MU {\Varec{x}}. & \const{C}( {\Varec{y}},\, {\VAREC{\const{C}(y,x)}} ) )
+\mu {\VAREC{\const{C}(y,x)}}. & \const{C}( {\Varec{y}},\, & \mu {\Varec{x}}. & \const{C}( {\Varec{y}},\, {\VAREC{\const{C}(y,x)}} ) )
 \end{array}\]
@@
 \xend
-\end{examplex}
+\end{example}
 
@@
 \inferrule{
-  t : \delta \in \tEc
+  t : \tau \in \tEc
   \quad
-  ( \const s( t ) \in \tEc \text{ and } \const s \in \Sel^\delta )
+  ( \const s( t ) \in \tEc \text{ and } \const s \in \Sel^\tau )
   \text{ or }
-  ( \delta \in \Data \text{ and } \delta \text{ finite} )
+  ( \tau \in \Data \text{ and } \tau \text{ finite} )
 }{
@@
 \inferrule{
-  \delta \in \Codata
+  \tau \in \Codata
   \quad
-  \left|\delta\right| = 1
+  \left|\tau\right| = 1
   \quad
-  t, u : \delta \in \tEc
+  t, u : \tau \in \tEc
 }{
@@
 \paragraph{Phase 3: Branching.}
-If a selector is applied to a term $t$, or if $t$'s type is a finite datatype,
-the equivalence class of $t$ must contain an application of one of
-$\delta$'s constructors
-$\Ctr^\delta = \{ \const{C}_1, \ldots, \const{C}_m \}$.
+Otherwise, in the case there is a selector term applied to $t$, or if $t$ has a finite datatype type,
+the equivalence class of $t$ must contain an application of one of the constructors of $\tau$ (where $\Ctr^\tau = \{ C_1, \ldots C_n \}$).
 This is enforced in the third phase by the \rn{Split} rule.
-A companion rule, \rn{Singleton}, handles the degenerate case where two terms
-from $\tEc$ have a corecursive singleton type, in which case they must be
-equivalent.
+The \rn{Singleton} rule handles the degenerate case where two terms from $\tEc$
+have a corecursive singleton type, in which case they must be equivalent.
 
@@
 A derivation tree is \emph{closed} if all of its leaf nodes are $\bot$.
-A node is \emph{saturated} if no nonredundant instance of a rule can be applied to it.
-%
-If there exists a closed derivation tree with root node $\Ec$, we conclude that $\Ec$ is $\thD$-unsatisfiable.
-If there exists a derivation tree with root node $\Ec$ that contains a saturated node, we conclude that $\Ec$ is $\thD$-satisfiable.
+A node is \emph{saturated} if no nonredundant application of a rule can be applied to it.
+If there exists a closed derivation tree with root node $\Ec$, then we conclude that $\Ec$ is $\thD$-unsatisfiable.
+If there exists a derivation tree with root node $\Ec$ that contains a saturated node, then we conclude that $\Ec$ is $\thD$-satisfiable.
 
@@
 let $S^0_t$ be the set $\{ t \}$,
-for $i \geq 0$ let $S^{i+1}_t =$ $S^i_t \mathrel\cup \{ s( u ) \mid u : \upsilon \in S^i, \upsilon \in \Data, \upsilon \text{ finite}, s \in \Sel^{\upsilon}  \}$,
+for $i \geq 0$ let $S^{i+1}_t =$ $S^i_t \cup \{ s( u ) \mid u : \upsilon \in S^i, \upsilon \in \Data, \upsilon \text{ finite}, s \in \Sel^{\upsilon}  \}$,
 and let $S^\ast_t$ be the fixed point of this sequence.
@@
 Let $S^\ast$ be the union of $S^\ast_t$ for all $t \in T$,
-and let $\tcEc$ be the set of subterms of $\Ec \mathrel\cup \{ \const C_i( \const s^1_i( t ), \ldots, \const s^{a_i}_i( t ) ) \mid t : \tau \in S \mathrel\cup S^\ast, \const C_i \in \Ctr^\tau \}$.
+and let $\tcEc$ be the set of subterms of $\Ec \cup \{ C_i( s^1_i( t ), \ldots, s^{a_i}_i( t ) ) \mid t : \tau \in S \cup S^\ast, C_i \in \Ctr^\tau \}$.
 In a derivation tree with root node $\Ec$,
@@
 $\mathcal{T}(\Fc) \subseteq \tcEc$, and moreover thus contains an equality between two terms from $\tcEc$ not occurring in its parent node.
-Thus, the depth of a branch in a derivation tree with root node $\Ec$ is at most $\left| \tcEc \right|^2$,
+Thus, the depth of a branch in a derivation tree with root node $\Ec$ is at most $\mid \tcEc \mid^2$,
 which is finite since $\tcEc$ is finite.
@@
 whose root node is the result of applying the rule \rn{Split} on term $t$ of type $\tau$,
-then by the induction hypothesis $\Ec \mathrel\cup t \teq \const C_i( \const s^1_i( t ), \ldots, \const s^{a_i}_i( t ) )$ is unsatisfiable
-for each $\const C_i \in \Ctr^\tau$.
-Since by exhaustiveness, all models of $\thD$ entail (exactly) one equality $t \teq \const C_i( \const s^1_i( t ), \ldots, \const s^{a_i}_i( t ) )$,
+then by the induction hypothesis $\Ec \cup t \teq C_i( s^1_i( t ), \ldots, s^{a_i}_i( t ) )$ is unsatisfiable
+for each $C_i \in \Ctr^\tau$.
+Since by exhaustiveness, all models of $\thD$ entail (exactly) one equality $t \teq C_i( s^1_i( t ), \ldots, s^{a_i}_i( t ) )$,
 we know $\Ec$ is $\thD$-unsatisfiable.
 Otherwise, the child of $\Ec$ is a closed derivation tree
-whose root node is $\Ec \mathrel\cup t_1 \teq t_2$ obtained by applying one of the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, \rn{Inject}, \rn{Unique}, or \rn{Singleton}.
+whose root node is $\Ec \cup t_1 \teq t_2$ obtained by applying one of the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, \rn{Inject}, \rn{Unique}, or \rn{Singleton}.
 In all cases, we have that $\Ec \models_{\thD} t_1 \teq t_2$.
@@
 For \rn{Singleton}, clearly $t_1$ and $t_2$ must have the same value since the cardinality of their type is one.
-By the induction hypothesis, we have $\Ec \mathrel\cup t_1 \teq t_2$ is $\thD$-unsatisfiable,
+By the induction hypothesis, we have $\Ec \cup t_1 \teq t_2$ is $\thD$-unsatisfiable,
 and thus $\Ec$ is $\thD$-unsatisfiable.
@@
 the \emph{interpretation of $t_2$ in $t_1$} is the $\mu$-term $t_1\interp{t_2}^\emptyset$ as returned by the following recursive procedure,
-where $x \mapsto \MU x. \const{C}( \vec u ) \in t_1$ indicates that $\MU x. \const{C}( \vec u )$ is the subterm of $t_1$ that binds this occurrence of variable $x$.
+where $x \mapsto \mu x. \const{C}( \vec u ) \in t_1$ indicates that $\mu x. \const{C}( \vec u )$ is the subterm of $t_1$ that binds this occurrence of variable $x$.
 \[\begin{array}{r@{}c@{}l}
@@
 \textstyle\begin{cases}
-    \MU x.\; \const{C}( t_1\interp{ \vec u}^{M \mathrel\cup \{ x \}}) & \text{if }x \not\in M, x \mapsto \MU x. \const{C}( \vec u ) \in t_1  \\[-\jot]
+    \mu x.\; \const{C}( t_1\interp{ \vec u}^{M \cup \{ x \}}) & \text{if }x \not\in M, x \mapsto \mu x. \const{C}( \vec u ) \in t_1  \\[-\jot]
     x & \text{otherwise}
 \end{cases} \\
-t_1\interp{\MU x. \const{C}( \vec u )}^M & = &
+t_1\interp{\mu x. \const{C}( \vec u )}^M & = &
 \textstyle\begin{cases}
-    \MU x. \const{C}( t_1\interp{ \vec u }^{M \mathrel\cup \{ x \}} ) & \text{if }x \not\in M \\[-\jot]
+    \mu x. \const{C}( t_1\interp{ \vec u }^{M \cup \{ x \}} ) & \text{if }x \not\in M \\[-\jot]
     x & \text{otherwise}
@@
 
-Given a $\mu$-term $t_1$ of the form $\MU x. \const{C}( u_1, \ldots, u_n )$ with subterm $t_2$,
+Given a $\mu$-term $t_1$ of the form $\mu x. \const{C}( u_1, \ldots, u_n )$ with subterm $t_2$,
 we say $t_2$ is a \emph{bisimilar subterm} of $t_1$ if and only if it is of the form
-$\MU y. \const{C}( v_1, \ldots, v_n )$ and $t_1\interp{ u_i } \aequiv t_1\interp{ v_i }$ for $i = 1,\ldots, n$.
-The term $t_1$ is \emph{normal} if it does not contain a strict, bisimilar subterm
+$\mu y. \const{C}( v_1, \ldots, v_n )$ and $t_1\interp{ u_i } \aequiv t_1\interp{ v_i }$ for $i = 1,\ldots, n$.
+We say $t_1$ is \emph{normal} if and only if it does not contain a strict bisimilar subterm
 and all of its strict subterms are also normal.
-For example, the $\mu$-term $t_1$ of the form $\MU x. \const{C}( \MU y. \const{C}( y ) )$ is not normal,
-since $\MU y. \const{C}( y )$ is a bisimilar subterm of $t_1$.
+For instance, the $\mu$-term $t_1$ of the form $\mu x. \const{C}( \mu y. \const{C}( y ) )$ is not normal,
+since $\mu y. \const{C}( y )$ is a bisimilar subterm of $t_1$.
 In particular, note that their arguments have the same interpretation in $t_1$,
 that is,
-$t_1\interp{\MU y. \const{C}( y )}^\emptyset =$
-$\MU y. \const{C}( t_1\interp{y}^{\{y\}} ) =$
-$\MU y. \const{C}( y )$
+$t_1\interp{\mu y. \const{C}( y )}^\emptyset =$
+$\mu y. \const{C}( t_1\interp{y}^{\{y\}} ) =$
+$\mu y. \const{C}( y )$
 is syntactically equivalent to
 $t_1\interp{y}^\emptyset =$
-$\MU y. \const{C}( t_1\interp{y}^{\{y\}} ) =$
-$\MU y. \const{C}( y )$.
-The $\mu$-term $t_1$ of the form $\MU x. \const{C}( \MU y. \const{C}( x ) )$ is also not normal,
-since $\MU y. \const{C}( x )$ is a bisimilar subterm of $t_1$,
+$\mu y. \const{C}( t_1\interp{y}^{\{y\}} ) =$
+$\mu y. \const{C}( y )$.
+The $\mu$-term $t_1$ of the form $\mu x. \const{C}( \mu y. \const{C}( x ) )$ is also not normal,
+since $\mu y. \const{C}( x )$ is a bisimilar subterm of $t_1$,
 noting that
-$t_1\interp{\MU y. \const{C}( x )}^\emptyset =$
-$\MU y. \const{C}( t_1\interp{x}^{\{y\}} ) =$
-$\MU y. \const{C}( t_1\interp{\MU x. \const{C}( \MU y. \const{C}( x ) ) }^{\{y\}} ) =$
-$\MU y. \const{C}( \MU x. \const{C}( t_1\interp{\MU y. \const{C}( x )}^{\{x,y\}} ) ) =$
-$\MU y. \const{C}( \MU x. \const{C}( y ) )$
+$t_1\interp{\mu y. \const{C}( x )}^\emptyset =$
+$\mu y. \const{C}( t_1\interp{x}^{\{y\}} ) =$
+$\mu y. \const{C}( t_1\interp{\mu x. \const{C}( \mu y. \const{C}( x ) ) }^{\{y\}} ) =$
+$\mu y. \const{C}( \mu x. \const{C}( t_1\interp{\mu y. \const{C}( x )}^{\{x,y\}} ) ) =$
+$\mu y. \const{C}( \mu x. \const{C}( y ) )$
 is syntactically equivalent to $t_1\interp{x}^\emptyset = t_1$ for the renaming $\{ x \mapsto y, y \mapsto x \}$.
 
-For any $\mu$-term $t$ of the form $\MU x. C( \vec u )$, we may recursively construct a normal $\mu$-term $\nf{t}$
+For any $\mu$-term $t$ of the form $\mu x. C( \vec u )$, we may recursively construct a normal $\mu$-term $\nf{t}$
 (which we call the \emph{normal form} of $t$)
@@
 and by normalizing $t$'s subterms.
-For example, $\nf{(\MU x. \const{C}( \MU y. \const{C}( x ) ))} = \MU x. \const{C}( x )$.
+For instance, $\nf{(\mu x. \const{C}( \mu y. \const{C}( x ) ))} = \mu x. \const{C}( x )$.
 
@@
 \label{lem:mu-norm-arg}
-If $\vec u$ are normal, then if $t = \nf{(\MU x. \const{C}( \vec u ) )} = \MU x. \const{C}( \vec w )$,
+If $\vec u$ are normal, then if $t = \nf{(\mu x. \const{C}( \vec u ) )} = \mu x. \const{C}( \vec w )$,
 then $\vec u \vsim t\interp{\vec w}$.
@@
 \label{lem:mu-cong}
-If $t_u = \MU x. \const{C}( \vec u )$ and $t_w = \MU y. \const{C}( \vec w )$ are normal,
+If $t_u = \mu x. \const{C}( \vec u )$ and $t_w = \mu y. \const{C}( \vec w )$ are normal,
 then $t_u\interp{ \vec u } \vsim t_w\interp{ \vec w }$ if and only if $t_u \vsim t_w$.
@@
 in $\M( \tau )$ that is $\alpha$-equivalent to
-$\nf{(\MU x. \const{C}( \M( \vec t \,) ))}$ where $x$ is fresh, and
+$\nf{(\mu x. \const{C}( \M( \vec t \,) ))}$ where $x$ is fresh, and
 \item
-for all selector terms $\const s^j_i( t )$ of type $\tau$,% for some $\vec u$,
-$\M( \const s^j_i( t ) )$ is the value
+for all selector terms $s^j_i( t )$ of type $\tau$,% for some $\vec u$,
+$\M( s^j_i( t ) )$ is the value
 in $\M( \tau )$ that is $\alpha$-equivalent to
 $\M( t )\interp{u_j}$
-when $\M( t )$ is of the form $\MU x. \const{C}_i( \vec u )$.
+when $\M( t )$ is of the form $\mu x. \const{C}_i( \vec u )$.
 \end{enumerate}
@@
 First, we need the following definition.
-We write $t \vsimv{x} s$ if $\mu$-terms $t$ and $s$ are syntactically equivalent
+We say that $t \vsimv{x} s$ if and only if $\mu$-terms $t$ and $s$ are syntactically equivalent
 for some renaming that avoids capturing any variable other than $x$.
-For example,
-$\MU x. \const{D}( x ) \vsimv{y} \MU x. \const{D}( y )$,
-$\MU x. \const{C}( x, x ) \vsimv{x} \MU y. \const{C}( x, y )$ and
-$\MU x. \const{C}( z, x ) \vsimv{z} \MU y. \const{C}( z, y )$,
+For instance,
+$\mu x. \const{D}( x ) \vsimv{y} \mu x. \const{D}( y )$,
+$\mu x. \const{C}( x, x ) \vsimv{x} \mu y. \const{C}( x, y )$ and
+$\mu x. \const{C}( z, x ) \vsimv{z} \mu y. \const{C}( z, y )$,
 but
-$\MU x. \const{D}( x ) \not\vsimv{x} \MU x. \const{D}( y )$ and
-$\MU x. \const{C}( x, x ) \not\vsimv{y} \MU y. \const{C}( x, y )$.
+$\mu x. \const{D}( x ) \not\vsimv{x} \mu x. \const{D}( y )$ and
+$\mu x. \const{C}( x, x ) \not\vsimv{y} \mu y. \const{C}( x, y )$.
 For a variable $x$ of type $\tau$ and a fixed normal model $\M$,
@@
 such that $v \vsimv{x} t_1\interp{t_2}$ for some term $t_1$ with subterm $t_2$ occurring in the range of $\Val$.
-We construct $\Val^\ast$ by exhaustively applying the following update to $\Val$:
+We construct $\Val^\ast$ by exhaustively applying the following update to $\nf{(\Val)}$:
 
@@
 \inferrule{
-  x : \tau \in \FV( \Val )
+  {\Varec{x}} : \tau \in \FV( \Val )
   \quad
-  (\MU x. t) \aequiv v
+  (\mu {\Varec{x}}. t) \aequiv v
   \quad
@@
   \quad
-  v \not\in \mathcal{T}^\M_x(\Val)
+  v \not\in \mathcal{T}^\M_{\Varec{x}}(\Val)
 }{
-  \Val := \nf{(\Val \{ x \mapsto \MU x. t \})}
+  \Val := \nf{(\Val \{ {\Varec{x}} \mapsto \mu {\Varec{x}}. t \})}
 }
@@
 After assigning this value, we normalize all terms in the range of $\Val$.
-%We then let $\Val^\ast \ec{x}$ be $\nf{(\Val' \ec{x})}$ for all $\ec{x}$.
-
-\begin{examplex}
+
+\begin{example}
 Let $\tau$ be a type in $\Codata$ with three constructors $\const{A}$, $\const{B}$ and $\const{C}$ each having type $\tau \rightarrow \tau$.
@@
 \[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{2.0em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
-\Val \ec{y_1} & = & \MU {\Varec{y_1}}. & \const{A}( & \MU {\Varec{y_2}}. & \const{B}( {\Varec{x}} ) ) &
-\Val \ec{w_1} & = & \MU {\Varec{w_1}}. & \const{A}( {\Varec{x}} )      & \\
-\Val \ec{y_2} & = & \MU {\Varec{y_2}}. & \multicolumn{2}{l}{\const{B}( {\Varec{x}} )}      & &
-\Val \ec{w_2} & = & \MU {\Varec{w_2}}. & \const{C}( {\Varec{w_3}} ) \\
-\Val \ec{z_1} & = & \MU {\Varec{z_1}}. & \const{C}( & \MU {\Varec{z_2}}. & \const{A}( {\Varec{z_1}} ) ) &
+\Val \ec{y_1} & = & \mu {\Varec{y_1}}. & \const{A}( & \mu {\Varec{y_2}}. & \const{B}( {\Varec{x}} ) ) &
+\Val \ec{w_1} & = & \mu {\Varec{w_1}}. & \const{A}( {\Varec{x}} )      & \\
+\Val \ec{y_2} & = & \mu {\Varec{y_2}}. & \multicolumn{2}{l}{\const{B}( {\Varec{x}} )}      & &
+\Val \ec{w_2} & = & \mu {\Varec{w_2}}. & \const{C}( {\Varec{w_3}} ) \\
+\Val \ec{z_1} & = & \mu {\Varec{z_1}}. & \const{C}( & \mu {\Varec{z_2}}. & \const{A}( {\Varec{z_1}} ) ) &
 \Val \ec{w_3} & = & {\Varec{w_3}} \\
-\Val \ec{z_2} & = & \MU {\Varec{z_2}}. & \const{A}( & \MU {\Varec{z_1}}. & \const{C}( {\Varec{z_2}} ) ) &
+\Val \ec{z_2} & = & \mu {\Varec{z_2}}. & \const{A}( & \mu {\Varec{z_1}}. & \const{C}( {\Varec{z_2}} ) ) &
 \Val \ec{x} & = & {\Varec{x}} \\
@@
 \{
-\MU x. \const{A}( \const{B}( x )),
-\const{A}( \MU y. \const{B}( y )),
-\MU x. \const{B}( x ),
-\MU x. \const{C}( \const{A}( x )),
-\MU x. \const{A}( \const{C}( x )),
-\MU x. \const{A}( x )
+\mu x. \const{A}( \const{B}( x )),
+\const{A}( \mu y. \const{B}( y )),
+\mu x. \const{B}( x ),
+\mu x. \const{C}( \const{A}( x )),
+\mu x. \const{A}( \const{C}( x )),
+\mu x. \const{A}( x )
 \}
@@
 Consider a model $\M$ that interprets variables in $\Ec$ based on the mapping $\Val$,
-for example, $x$ is interpretted as $\Val \ec{x}$, $y_1$ as $\Val \ec{y_1}$ and so on.
+for instance, $x$ is interpretted as $\Val \ec{x}$, $y_1$ as $\Val \ec{y_1}$ and so on.
 Assigning a value for ${\Varec{x}}$ that is $\alpha$-equivalent to a value in $\mathcal{T}^\M_{{\Varec{x}}}( \Val )$
@@
 which in turn may cause $\M$ to be inconsistent.
-For example, say we assign $\MU {\Varec{x}}. \const{B}( {\Varec{x}} )$ for ${\Varec{x}}$.
-After this substitution, $\Val \ec{y_2}$ is $\MU {\Varec{y_2}}. \const{B}( \MU {\Varec{x}}. \const{B}( {\Varec{x}} ) )$,
-which has normal form $\MU {\Varec{y_2}}. \const{B}( {\Varec{y_2}} )$,
-which is $\alpha$-equivalent to $\MU {\Varec{x}}. \const{B}( {\Varec{x}} )$.
+For instance, say we assign $\mu {\Varec{x}}. \const{B}( {\Varec{x}} )$ for ${\Varec{x}}$.
+After this substitution, $\Val \ec{y_2}$ is $\mu {\Varec{y_2}}. \const{B}( \mu {\Varec{x}}. \const{B}( {\Varec{x}} ) )$,
+which has normal form $\mu {\Varec{y_2}}. \const{B}( {\Varec{y_2}} )$,
+which is $\alpha$-equivalent to $\mu {\Varec{x}}. \const{B}( {\Varec{x}} )$.
 However, this contradicts the disequality $x \not\teq y_2$ in $\Ec$.
@@
 then the values in the range of $\Val$ remain $\alpha$-disequivalent.
-For example, we may assign values such as 
-$\MU {\Varec{x}}. \const{C}( {\Varec{x}} )$,
-$\MU {\Varec{x}}. \const{C}( \MU x_1. \const{A}( x_1) )$, or
-$\MU {\Varec{x}}. \const{B}( \const{A}( {\Varec{x}} ) )$
+For instance, we may assign values such as 
+$\mu {\Varec{x}}. \const{C}( {\Varec{x}} )$,
+$\mu {\Varec{x}}. \const{C}( \mu x_1. \const{A}( x_1) )$, or
+$\mu {\Varec{x}}. \const{B}( \const{A}( {\Varec{x}} ) )$
 to ${\Varec{x}}$.
-Legal substitutions for $v_{\ec{x}}$ may cause the range of $\Val$ to contain non-normal terms.
+Legal substitutions for ${\Varec{x}}$ may cause the range of $\Val$ to contain non-normal terms.
 After applying the substitution for the third term mentioned above,
-$\Val \ec{w_1}$ is $\MU {\Varec{w_1}}. \const{A}( \MU {\Varec{x}}. \const{B}( \const{A}( {\Varec{x}} ) ) )$,
+$\Val \ec{w_1}$ is $\mu {\Varec{w_1}}. \const{A}( \mu {\Varec{x}}. \const{B}( \const{A}( {\Varec{x}} ) ) )$,
 which has normal form
-$\MU {\Varec{w_1}}. \const{A}( \MU {\Varec{x}}. \const{B}( {\Varec{w_1}} ) )$.
+$\mu {\Varec{w_1}}. \const{A}( \mu {\Varec{x}}. \const{B}( {\Varec{w_1}} ) )$.
 \xend
-\end{examplex}
-
-%We first state the following properties of $\Val^\ast$.
-%where we write $\muvar( t )$ to denote $t$ if $t$ is a variable, or $x$ if $t$ is of the form $\MU x. u$.
+\end{example}
+
+We first state the following properties of $\Val^\ast$,
+where we write $\muvar( t )$ to denote $t$ if $t$ is a variable, or ${\Varec{x}}$ if $t$ is of the form $\mu {\Varec{x}}. u$.
 
@@
 \item
-All subterms $\MU {\Varec{x}}. t$ of $\Val^\ast \ec{y}$ are such that
-$\Val^\ast \ec{y} \interp{ \MU {\Varec{x}}. t } = \Val^\ast \ec{x}$. 
-%If $\const{C}( t_1, \ldots, t_n ) \in \ec{x}$,
-%then all subterms $\MU v_{\ec{x}}. u$ occurring in $\Val^\ast \ec{y}$ 
-%are of the form $\MU v_{\ec{x}}. \const{C}( w_1, \ldots, w_n )$,
-%where 
-%and $\Val^\ast \ec{y} \interp{ w_i } = \Val^\ast \ec{t_i}$ for $i = 1, \ldots, n$.
-%If $\Val^\ast \ec{x}$ contains a subterm of the form $\MU v_{\ec{x}}. \const{C}( w_1, \ldots, w_n )$,
-%and $\muvar( w_i ) = v_{\ec{t_i}}$ for $i = 1, \ldots, n$,
-%\item
-%$t_1 \interp{ t_2 } = \Val^\ast \ec{x}$
-%for all $t_1$ with subterm $t_2$ in the range of $\Val^\ast$ where $\muvar(t_2)={\Varec{x}}$,
+All subterms $t$ of $\Val^\ast \ec{y}$ where $\muvar(t) = {\Varec{x}}$ are such that
+$\Val^\ast \ec{y} \interp{ t } = \Val^\ast \ec{x}$. 
 \item
@@
 
-To show 2,
-first note that since \rn{Clash} does not apply to $\Ec$,
+We first show that 2 and 3 hold initially for $\Val$.
+Since \rn{Clash} does not apply to $\Ec$,
 for all equivalence classes $\ec{z}$,
@@
 Thus, 
-if $\const{C}_i( t_1, \ldots, t_n ) \in \ec{z}$,
 $\Val$ was constructed by applying a sequence of substitutions
-where ${\Varec{z}}$ was replaced only by $\MU {\Varec{z}}. \const{C}_i( {\Varec{t_1}}, \ldots, {\Varec{t_n}} )$.
-Assume $\Val \ec{y}$ has a subterm $\MU {\Varec{x}}. \const{C}( \vec t )$.
-We first show that our construction of $\Val$ is such that $\Val \ec{x} = \Val \ec{y}\interp{ \MU {\Varec{x}}. t }$.
-Notice that both $\Val \ec{x}$ and the subterm $\MU {\Varec{x}}. t$ of $\Val \ec{y}$ were
+where all substitutions for each variable ${\Varec{z}}$ 
+were uniquely of the form $\{ {\Varec{z}} \mapsto \const{C}_i( {\Varec{t_1}}, \ldots, {\Varec{t_n}} ) \}$
+for $\ec{z}$ that contain a constructor $\const{C}_i( t_1, \ldots, t_n )$.
+Say $\Val \ec{y}$ has a subterm $t$ where $\muvar(t) = {\Varec{x}}$.
+Both $\Val \ec{x}$ and the subterm $t$ of $\Val \ec{y}$ were
 constructed by applying a sequence of substitutions of the form mentioned above to ${\Varec{x}}$.
-We show that $\Val \ec{x} \interp{ \Val \ec{x} } = \Val \ec{y} \interp{ \MU {\Varec{x}}.t }$.
-\rem{TODO}
-%Thus, the set of subterms which $\Val \ec{x}$ and $\MU v_{\ec{x}}. t$ differ 
-%of the form $\MU {\Varec{z}}. u$ and $v_{\ec{z}}$ respectively, where $v_{\ec{z}}$ is a bound variable in $\Val \ec{y}$.
-%In the latter case, $\Val \ec{y} \interp{ {\Varec{z}} } = \MU {\Varec{z}}. \const{C}( \Val \ec{y} \interp{\vec u} )$,
-%where $\MU v_{\ec{z}}. u$ and $\MU v_{\ec{z}}. u'$ were constructed by applying a sequence of substitutions of the form above to $v_{\ec{z}}$.
-Since $\Val^\ast$ is obtained by applying the same substitutions and normalizations to all terms in the range of $\Val$,
-we have $\Val^\ast \ec{y} \interp{ \MU {\Varec{x}}. t} = \Val^\ast \ec{x}$.
-
-To show 3,
-first note that since \rn{Unique} does not apply,
-$\Val \ec{x} \aequiv \Val \ec{y}$ if and only if $\ec{x} = \ec{y}$.
-Since \rem{TODO}
-we have $\nf{(\Val \ec{x})} \aequiv \nf{(\Val \ec{y})}$ if and only if $\ec{x} = \ec{y}$.
-We show that for any substitution $\sigma$ applied to $\Val$ while constructing its completion is such that
-$\nf{(\Val \ec{x} \sigma)} \aequiv \nf{(\Val \ec{y} \sigma)}$ if and only if $\ec{x} = \ec{y}$.
-\rem{TODO}
-Let $P$ be the set of positions for which $\Val \ec{x}$ and $\Val \ec{y}$ are not $\alpha$-equivalent,
-which is non-empty by assumption.
+Moreover, free variables ${\Varec{z}}$ in $\mu {\Varec{x}}.t$ that are bound in $\Val \ec{y}$ are interpreted in 
+the expansion of $\Val \ec{y} \interp{ \mu {\Varec{x}}.t }$ as a term
+constructed by a sequence of substitutions of the form mentioned above to ${\Varec{z}}$. \rem{more detail}
+Thus, we have
+$\Val \ec{y} \interp{ \mu {\Varec{x}}.t } = \Val \ec{x} \interp{ \Val \ec{x} } = \Val \ec{x}$, and thus 2 holds for $\Val$.
+Statement 3 holds for $\Val$ since \rn{Unique} does not apply.
+
+We now show that $\Val = \nf{(\Val)}$.
+Assume by contradition $\Val \ec{x} \neq \nf{(\Val \ec{x})}$ for some $\Val \ec{x}$ of minimal size.
+We have that $\Val \ec{x}$ is of the form $\mu {\Varec{x}}. \const{C}( t_1, \ldots, t_n )$.
+Due to our construction of $\Val$, 
+we know $\ec{x}$ contains a constructor $\const{C}( z_1, \ldots, z_n )$ and $\muvar( t_i ) = z_i$ for $i = 1, \ldots, n$.
+Since $\Val \ec{x}$ is a minimal, it
+contain a subterm of the form $\mu {\Varec{y}}. \const{C}( u_1, \ldots, u_n )$
+where $\Val \ec{x} \interp{ t_i } \aequiv \Val \ec{x} \interp{ u_i }$ for $i = 1, \ldots, n$.
+Due to our construction of $\Val$, $\ec{y}$ contains a constructor $\const{C}( w_1, \ldots, w_n )$ and $\muvar( u_i ) = w_i$ for $i = 1, \ldots, n$.
+Since \rn{Cong} does not apply to $\Ec$
+we have $\ec{w_j}$ and $\ec{z_j}$ are distinct for some $j$.
+By statement 2, $\Val \ec{x} \interp{ t_j } = \Val \ec{z_j}$
+and $\Val \ec{x} \interp{ u_j } = \Val \ec{w_j}$,
+which are not $\alpha$-equivalent by statement 3, contradicting the fact that $\mu {\Varec{y}}. \const{C}( u_1, \ldots, u_n )$ is a bisimilar subterm of $\Val \ec{x}$.
+Thus, $\Val = \nf{(\Val)}$, and statements 2 and 3 hold for $\nf{(\Val)}$.
+%For each $i = 1, \ldots, n$,
+%if ${\Varec{z_i}} = {\Varec{w_i}}$, then $w_i \teq z_i \in \Ec$.
+%If ${\Varec{z_i}} \neq {\Varec{w_i}}$, then by statement 2, $\Val \ec{x} \interp{ t_i } = \Val \ec{z_i}$
+%and $\Val \ec{x} \interp{ u_i } = \Val \ec{w_i}$.
+%By statement 3, we have that $\ec{z_i} = \ec{w_i}$, and thus $u_i \teq t_i \in \Ec$ for each $i = 1, \ldots, n$.
+%This is a contradiction since $\ec{x}$ and $\ec{y}$ are distinct, and \rn{Cong} does not apply to $\Ec$.
+%Thus, statements 2 and 3 hold for $\nf{(\Val)}$.
+
+We now show that if 2 and 3 hold for some $\Val_1$,
+then 2 and 3 also hold for $\nf{(\Val_1 \sigma)}$,
+where $\sigma$ is a substitution of the form $\{ {\Varec{x}} \mapsto \mu {\Varec{x}}. t \}$,
+where $\mu {\Varec{x}}. t$ is not $\alpha$-equivalent to a term in $\mathcal{T}^\M_{\Varec{x}}(\Val)$.
+
+%\rem{TODO}
+%Thus, the set of subterms which $\Val \ec{x}$ and $\mu v_{\ec{x}}. t$ differ 
+%of the form $\mu {\Varec{z}}. u$ and $v_{\ec{z}}$ respectively, where $v_{\ec{z}}$ is a bound variable in $\Val \ec{y}$.
+%In the latter case, $\Val \ec{y} \interp{ {\Varec{z}} } = \mu {\Varec{z}}. \const{C}( \Val \ec{y} \interp{\vec u} )$,
+%where $\mu v_{\ec{z}}. u$ and $\mu v_{\ec{z}}. u'$ were constructed by applying a sequence of substitutions of the form above to $v_{\ec{z}}$.
+%Since $\Val^\ast$ is obtained by applying the same substitutions and normalizations to all terms in the range of $\Val$,
+%we have $\Val^\ast \ec{y} \interp{ \mu {\Varec{x}}. t} = \Val^\ast \ec{x}$.
+
+%To show 3,
+%first note that since \rn{Unique} does not apply,
+%$\Val \ec{x} \aequiv \Val \ec{y}$ if and only if $\ec{x} = \ec{y}$.
+%Due to part 2, we have $\Val = \nf{(\Val)}$, and thus
+%$\nf{(\Val \ec{x})} \aequiv \nf{(\Val \ec{y})}$ if and only if $\ec{x} = \ec{y}$.
+%We show that for any substitution $\sigma$ applied to $\Val$ while constructing its completion is such that
+%$\nf{(\Val \ec{x} \sigma)} \aequiv \nf{(\Val \ec{y} \sigma)}$ if and only if $\ec{x} = \ec{y}$.
+%\rem{TODO}
+%Let $P$ be the set of positions for which $\Val \ec{x}$ and $\Val \ec{y}$ are not $\alpha$-equivalent,
+%which is non-empty by assumption.
 
@@
 If $t$ is a constructor term of the form $\const{C}( u_1, \ldots, u_n )$,
-then $\M( t )$ is $\alpha$-equivalent with $\nf{(\MU x. \const{C}( \M( u_1 ), \ldots, \M( u_n ) ) )}$ for fresh $x$,
-which by the induction hypothesis is $\alpha$-equivalent with $\nf{(\MU x. \const{C}( \Val^\ast \ec{ u_1 }, \ldots, \Val^\ast \ec{ u_n } ) )}$, call this term $t_u$.
+then $\M( t )$ is $\alpha$-equivalent with $\nf{(\mu x. \const{C}( \M( u_1 ), \ldots, \M( u_n ) ) )}$ for fresh $x$,
+which by the induction hypothesis is $\alpha$-equivalent with $\nf{(\mu x. \const{C}( \Val^\ast \ec{ u_1 }, \ldots, \Val^\ast \ec{ u_n } ) )}$, call this term $t_u$.
 By Lemma~\ref{lem:model-completion}.2 and due to our construction of $\Val^\ast$, we have $\Val^\ast \ec{t}$ is a term
-$t_w$ of the form $\MU {\Varec{t}}. \const{C}( w_1, \ldots, w_n )$
+$t_w$ of the form $\mu {\Varec{t}}. \const{C}( w_1, \ldots, w_n )$
 where $t_w\interp{w_i} = \Val^\ast \ec{u_i}$ for $i = 1, \ldots, n$.
@@
 
-If $t$ is a selector term of the form $\const s^j_k( u )$,
+If $t$ is a selector term of the form $s^j_k( u )$,
 since \rn{Split} does not apply,
-$\ec{u}$ must contain a term of the form $\const C_i( \const s^1_i( u ), \ldots, \const s^{n}_i( u ) )$ for some $i$.
+$\ec{u}$ must contain a term of the form $C_i( s^1_i( u ), \ldots, s^{n}_i( u ) )$ for some $i$.
 By Lemma~\ref{lem:model-completion}.2 and due to our construction of $\Val^\ast$, we have $\Val^\ast \ec{u}$ is a term $t_w$ of the form
-$\MU {\Varec{u}}.\; \const C_i( w_1, \ldots, w_n )$,
-where $t_w \interp{ w_j } = \Val^\ast \ec{ \const s^j_k( u )}$. 
-If $k = i$, then $\M( t )$ is $\alpha$-equivalent with $t_w\interp{w_j} = \Val^\ast \ec{ \const s^j_k( u )}$,
+$\mu {\Varec{u}}. C_i( w_1, \ldots, w_n )$,
+where $t_w \interp{ w_j } = \Val^\ast \ec{ s^j_k( u )}$. 
+If $k = i$, then $\M( t )$ is $\alpha$-equivalent with $t_w\interp{w_j} = \Val^\ast \ec{ s^j_k( u )}$,
 which is $\Val^\ast \ec{t}$.
-If $k \neq i$, then we interpret $\const s^j_k( \M( u ) )$ as the value in $\M( \tau )$ that is $\alpha$-equivalent with $\Val^\ast \ec{ t }$.
+If $k \neq i$, then we interpret $s^j_k( \M( u ) )$ as the value in $\M( \tau )$ that is $\alpha$-equivalent with $\Val^\ast \ec{ t }$.
 Since \rn{Cong} does not apply and due to Lemma~\ref{lem:model-completion}.3,
@@
 Discriminators are treated as predicate symbols, so a discriminator $\const{d}_i( t )$ is such that $\const{d}_i$
-is a predicate symbol, instead of reducing this constraint to $t \teq \const{C}_i( \const s^1_i( t ), \ldots, \const s^n_i( t ) )$.
+is a predicate symbol, instead of reducing this constraint to $t \teq \const{C}_i( s^1_i( t ), \ldots, s^n_i( t ) )$.
 We have found that this leads to better performance since this reduction introduces terms more eagerly to $\tEc$.
@@
 Also, selectors are collapsed eagerly:
-if $\const s^j_i( t )$ is a term in $\tEc$ and $t$ contains the constructor
-$\const{C}_i( \vec u )$, then we immediately infer $\const s^j_i( t ) \teq u_j$, whereas the calculus we presented would apply \rn{Split} and \rn{Inject} before inferring this equality.
+if $s^j_i( t )$ is a term in $\tEc$ and $t$ contains the constructor
+$\const{C}_i( \vec u )$, then we immediately infer $s^j_i( t ) \teq u_j$, whereas the calculus we presented would apply \rn{Split} and \rn{Inject} before inferring this equality.
 For the sake of reducing the number of unique constraints considered by the calculus, we compute a normal form for constraints as a preprocessing step.
@@
 that is, a specialized solver for determining the satisfiability of conjunctions of constraints for its theory.
-Given a theory $T = T_1 \mathrel\cup \ldots T_n$ and a input set of clauses $F$ in CNF,
+Given a theory $T = T_1 \cup \ldots T_n$ and a input set of clauses $F$ in CNF,
 the DPLL(T) \rem{CDCL(T)?} procedure
@@
 We may interpret this partial assignment as a set of literals where $a \in M$ for all atoms $a$ that it assigned the value $\top$, and $\neg a \in M$ for all atoms that it assigns the value $\bot$.
-Then, at a high level, by the Nelson--Oppen combination framework,
+Then, at a high level, by the Nelson-Oppen combination framework,
 each $T_i$-solver for $i = 1, \ldots, n$ takes as input a combination of (1) the purified form of $T_i$-constraints occurring in $M$ where terms containing symbols not belonging to $T_i$ are replaced by fresh variables of the same type,
@@
 a set of equalities from $M_i$ that entail that the values of $\ec{x}$ in models of $\Ec$ are of the form $\Val \ec{x}$.
-For example, if $x \teq \const{C}( x ) \in M_i$, then $x \teq \const{C}( x )$ is a (possible) explanation for $\Val \ec{x} = \MU {\Varec{x}}. \const{C}( {\Varec{x}} )$.
+For instance, if $x \teq \const{C}( x ) \in M_i$, then $x \teq \const{C}( x )$ is a (possible) explanation for $\Val \ec{x} = \mu {\Varec{x}}. \const{C}( {\Varec{x}} )$.
 Thus, the rules of the calculus are implemented as follows.
@@
 For the \rn{Split} rule, we add a clause of the form
-$(\const{C}_1( \const s^1_1( t ), \ldots, \const s^{a_1}_1( t ) ) \vee \ldots \vee t \teq \const{C}_n( \const s^1_n( t ), \ldots, \const s^{a_n}_n( t ) ))$
+$(\const{C}_1( s^1_1( t ), \ldots, s^{a_1}_1( t ) ) \vee \ldots \vee t \teq \const{C}_n( s^1_n( t ), \ldots, s^{a_n}_n( t ) ))$
 to $F$.
@@
 
-    * sharing is caring (let's cite it for a change)
+    * sharing is caring
     * theory combination more complicated (and Theorem Completeness too weak)

### 2014-12-19 tex/codatadp.tex
more:wq

--- 2014-12-19 tex/codatadp.tex
@@
 
-\newcommand\MU{\mu\vvthinspace}
+\newcommand\MU{\vvthinspace\mu\vvthinspace}
 
@@
 \renewcommand{\vec}[1]{\bar #1}
-\newcommand{\Ec}{\mathsf{E}}
-\newcommand{\Fc}{\mathsf{F}}
-\newcommand{\Gc}{\mathsf{G}}
-\newcommand{\Ac}{\mathsf{A}}
-\newcommand{\Dc}{\mathsf{D}}
-\newcommand{\Rc}{\mathsf{R}}
+%\newcommand{\Ec}{\mathsf{E}}
+%\newcommand{\Ec}{E}
+\newcommand{\Ec}{\mathcal{E}}
+%\newcommand{\Fc}{\mathsf{F}}
+%\newcommand{\Fc}{F}
+\newcommand{\Fc}{\mathcal{F}}
+%\newcommand{\Gc}{\mathsf{G}}
+%\newcommand{\Ac}{\mathsf{A}}
+%\newcommand{\Dc}{\mathsf{D}}
+%\newcommand{\Rc}{\mathsf{R}}
 \newcommand{\vrange}{\mathsf{range}}
@@
 or a node of the form
-$\MU x.\; \const{C}( \vec t\, )$ for some constructor $\const{C} \in \Ctr$ and
+$\MU x.\: \const{C}( \vec t\, )$ for some constructor $\const{C} \in \Ctr$ and
 $\mu$-terms $\vec t$ with the expected types.
@@
 from $\tEc$ have a corecursive singleton type, in which case they must be
-equivalent.
+equivalent. Both the finiteness assumption on the datatype in \rn{Split} and
+the cardinality-1 constraint on the codatatype in \rn{Singleton} are easy to
+check based on a recursive computation of the cardinalities of the
+constructors' argument types.
 
@@
 Let $\SSS$ be the set of terms occurring as the argument of a selector in $\tEc$.
-Let $T$ be the set of terms in $\tEc$ whose types are finite datatype.
+Let $T$ be the set of terms in $\tEc$ whose types are finite datatypes.
 For each term $t \in T$,
@@
 \begin{examplex}
-Let $\tau$ be a type in $\Codata$ with three constructors $\const{A}$, $\const{B}$ and $\const{C}$ each having type $\tau \rightarrow \tau$.
+Let $\tau$ be a type in $\Codata$ with three constructors $\const{C}$, $\const{D}$, and $\const{E}$ each having type $\tau \rightarrow \tau$.
 Let $\Ec$ be the set
 $\{$
-$y_1 \teq \const{A}( y_2 )$,
-$y_2 \teq \const{B}( x )$,
-$z_1 \teq \const{C}( z_2 )$,
-$z_2 \teq \const{A}( z_1 )$,
-$w_1 \teq \const{A}( x )$,
-$w_2 \teq \const{C}( w_3 )$,
+$y_1 \teq \const{C}( y_2 )$,
+$y_2 \teq \const{D}( x )$,
+$z_1 \teq \const{E}( z_2 )$,
+$z_2 \teq \const{C}( z_1 )$,
+$w_1 \teq \const{C}( x )$,
+$w_2 \teq \const{E}( w_3 )$,
 $x \not\teq y_2$,
@@
 \[\begin{array}{r@{}c@{}rlll@{\qquad}r@{}c@{}lllll}
-\Val \ec{y_1} & {}\,=\,{} & \MU {\Varec{y_1}}. & \const{A}( & \MU {\Varec{y_2}}. & \const{B}( {\Varec{x}} ) ) &
-\Val \ec{w_1} & {}\,=\,{} & \MU {\Varec{w_1}}. & \const{A}( {\Varec{x}} )      & \\
-\Val \ec{y_2} & = & \MU {\Varec{y_2}}. & \multicolumn{2}{l}{\const{B}( {\Varec{x}} )}      & &
-\Val \ec{w_2} & = & \MU {\Varec{w_2}}. & \const{C}( {\Varec{w_3}} ) \\
-\Val \ec{z_1} & = & \MU {\Varec{z_1}}. & \const{C}( & \MU {\Varec{z_2}}. & \const{A}( {\Varec{z_1}} ) ) &
+\Val \ec{y_1} & {}\,=\,{} & \MU {\Varec{y_1}}. & \const{C}( & \MU {\Varec{y_2}}. & \const{D}( {\Varec{x}} ) ) &
+\Val \ec{w_1} & {}\,=\,{} & \MU {\Varec{w_1}}. & \const{C}( {\Varec{x}} )      & \\
+\Val \ec{y_2} & = & \MU {\Varec{y_2}}. & \multicolumn{2}{l}{\const{D}( {\Varec{x}} )}      & &
+\Val \ec{w_2} & = & \MU {\Varec{w_2}}. & \const{E}( {\Varec{w_3}} ) \\
+\Val \ec{z_1} & = & \MU {\Varec{z_1}}. & \const{E}( & \MU {\Varec{z_2}}. & \const{C}( {\Varec{z_1}} ) ) &
 \Val \ec{w_3} & = & {\Varec{w_3}} \\
-\Val \ec{z_2} & = & \MU {\Varec{z_2}}. & \const{A}( & \MU {\Varec{z_1}}. & \const{C}( {\Varec{z_2}} ) ) &
+\Val \ec{z_2} & = & \MU {\Varec{z_2}}. & \const{C}( & \MU {\Varec{z_1}}. & \const{E}( {\Varec{z_2}} ) ) &
 \Val \ec{x} & = & {\Varec{x}} \\
@@
 \{
-\MU x.\; \const{A}( \const{B}( x )),
-\const{A}( \MU y.\; \const{B}( y )),
-\MU x.\; \const{B}( x ),
-\MU x.\; \const{C}( \const{A}( x )),
-\MU x.\; \const{A}( \const{C}( x )),
-\MU x.\; \const{A}( x )
+\MU x.\; \const{C}( \const{D}( x )),
+\const{C}( \MU y.\; \const{D}( y )),
+\MU x.\; \const{D}( x ),
+\MU x.\; \const{E}( \const{C}( x )),
+\MU x.\; \const{C}( \const{E}( x )),
+\MU x.\; \const{C}( x )
 \}
@@
 which in turn may cause $\M$ to be inconsistent.
-For example, say we assign $\MU {\Varec{x}}.\; \const{B}( {\Varec{x}} )$ for ${\Varec{x}}$.
-After this substitution, $\Val \ec{y_2}$ is $\MU {\Varec{y_2}}.\; \const{B}( \MU {\Varec{x}}.\; \const{B}( {\Varec{x}} ) )$,
-which has normal form $\MU {\Varec{y_2}}.\; \const{B}( {\Varec{y_2}} )$,
-which is $\alpha$-equivalent to $\MU {\Varec{x}}.\; \const{B}( {\Varec{x}} )$.
+For example, say we assign $\MU {\Varec{x}}.\; \const{D}( {\Varec{x}} )$ for ${\Varec{x}}$.
+After this substitution, $\Val \ec{y_2}$ is $\MU {\Varec{y_2}}.\; \const{D}( \MU {\Varec{x}}.\; \const{D}( {\Varec{x}} ) )$,
+which has normal form $\MU {\Varec{y_2}}.\; \const{D}( {\Varec{y_2}} )$,
+which is $\alpha$-equivalent to $\MU {\Varec{x}}.\; \const{D}( {\Varec{x}} )$.
 However, this contradicts the disequality $x \not\teq y_2$ in $\Ec$.
@@
 For example, we may assign values such as 
-$\MU {\Varec{x}}.\; \const{C}( {\Varec{x}} )$,
-$\MU {\Varec{x}}.\; \const{C}( \MU x_1.\; \const{A}( x_1) )$, or
-$\MU {\Varec{x}}.\; \const{B}( \const{A}( {\Varec{x}} ) )$
+$\MU {\Varec{x}}.\; \const{E}( {\Varec{x}} )$,
+$\MU {\Varec{x}}.\; \const{E}( \MU x_1.\; \const{C}( x_1) )$, or
+$\MU {\Varec{x}}.\; \const{D}( \const{C}( {\Varec{x}} ) )$
 to ${\Varec{x}}$.
@@
 After applying the substitution for the third term mentioned above,
-$\Val \ec{w_1}$ is $\MU {\Varec{w_1}}.\; \const{A}( \MU {\Varec{x}}.\; \const{B}( \const{A}( {\Varec{x}} ) ) )$,
+$\Val \ec{w_1}$ is $\MU {\Varec{w_1}}.\; \const{C}( \MU {\Varec{x}}.\; \const{D}( \const{C}( {\Varec{x}} ) ) )$,
 which has normal form
-$\MU {\Varec{w_1}}.\; \const{A}( \MU {\Varec{x}}.\; \const{B}( {\Varec{w_1}} ) )$.
+$\MU {\Varec{w_1}}.\; \const{C}( \MU {\Varec{x}}.\; \const{D}( {\Varec{w_1}} ) )$.
 \xend

### 2014-12-19 tex/codatadp.tex
more edits:wq

@@
 it can be shown by case analysis on the rules of the calculus that each non-root node $\Fc$ is such that
-$\mathcal{T}(\Fc) \subseteq \tcEc$, and moreover thus contains an equality between two terms from $\tcEc$ not occurring in its parent node.
+$\mathcal{T}(\Fc) \subseteq \tcEc$, and hence contains an equality between two terms from $\tcEc$ not occurring in its parent node.
 Thus, the depth of a branch in a derivation tree with root node $\Ec$ is at most $\left| \tcEc \right|^2$,

### 2014-12-19 tex/codatadp.tex
more

@@
 %\newcommand{\Ec}{\mathsf{E}}
-%\newcommand{\Ec}{E}
-\newcommand{\Ec}{\mathcal{E}}
+\newcommand{\Ec}{E}
+%\newcommand{\Ec}{\mathcal{E}}
 %\newcommand{\Fc}{\mathsf{F}}
-%\newcommand{\Fc}{F}
-\newcommand{\Fc}{\mathcal{F}}
+\newcommand{\Fc}{F}
+%\newcommand{\Fc}{\mathcal{F}}
 %\newcommand{\Gc}{\mathsf{G}}
@@
 \newcommand{\thO}{T_{\mathrm{o}}}
-\newcommand{\thD}{T_{\mathrm{cd}}}
+\newcommand{\thD}{T_{\!\mathrm{cd}}}
 
@@
 Let $\SSSS \ast$ be the union of $\SSSS \ast_t$ for all $t \in T$,
-and let $\tcEc$ be the set of subterms of $\Ec \mathrel\cup \{ \const C_i( \const s^1_i( t ), \ldots, \const s^{a_i}_i( t ) ) \mid t : \tau \in \SSS \mathrel\cup \SSSS \ast, \const C_i \in \Ctr^\tau \}$.
-In a derivation tree with root node $\Ec$,
+and let $\tcEc$ be the set of subterms of $\Ec \mathrel\cup \{ \const C_i( \const s^1_i( t ), \ldots, \const s^{a_i}_i( t ) ) \mid t : \delta \in \SSS \mathrel\cup \SSSS \ast,\, \const C_i \in \Ctr^\delta \}$.
+In a derivation tree with root node~$\Ec$,
 it can be shown by case analysis on the rules of the calculus that each non-root node $\Fc$ is such that
 $\mathcal{T}(\Fc) \subseteq \tcEc$, and hence contains an equality between two terms from $\tcEc$ not occurring in its parent node.
-Thus, the depth of a branch in a derivation tree with root node $\Ec$ is at most $\left| \tcEc \right|^2$,
+Thus, the depth of a branch in a derivation tree with root node $\Ec$ is at most $\left| \smash{\tcEc}\vphantom{X} \right|{\!}^2$,
 which is finite since $\tcEc$ is finite.
@@
 \begin{proof}
-We prove this by structural induction on the (closed) derivation tree with root node $\Ec$.
-First, if the derivation tree is an application of \rn{Conflict}, \rn{Clash}, or \rn{Acyclic},
+The proof is by structural induction on the derivation tree with root node $\Ec$.
+If the tree is an application of \rn{Conflict}, \rn{Clash}, or \rn{Acyclic},
 then $\Ec$ is $\thD$-unsatisfiable.
@@
 For \rn{Clash}, this is a consequence of distinctness.
-For \rn{Acyclic}, our construction of $\Val$ indicates that the class of values that $t_1$ can take in models of $\Ec$ is infeasible,
+For \rn{Acyclic}, the construction of $\Val$ indicates that the class of values that $t_1$ can take in models of $\Ec$ is infeasible,
 and thus $\Ec$ is unsatisfiable.
 If the children of $\Ec$ are closed derivation trees
-whose root node is the result of applying the rule \rn{Split} on term $t$ of type $\tau$,
-then by the induction hypothesis $\Ec \mathrel\cup t \teq \const C_i( \const s^1_i( t ), \ldots, \const s^{a_i}_i( t ) )$ is unsatisfiable
-for each $\const C_i \in \Ctr^\tau$.
+whose root node is the result of applying the rule \rn{Split} on term $t$ of type $\delta$,
+by the induction hypothesis $\Ec \mathrel\cup t \teq \const C_i( \const s^1_i( t ), \ldots, \const s^{a_i}_i( t ) )$ is unsatisfiable
+for each $\const C_i \in \Ctr^\delta$.
 Since by exhaustiveness, all models of $\thD$ entail (exactly) one equality $t \teq \const C_i( \const s^1_i( t ), \ldots, \const s^{a_i}_i( t ) )$,
@@
 \label{lem:mu-norm-arg}
-If $\vec u$ are normal, then if $t = \nf{(\MU x.\; \const{C}( \vec u ) )} = \MU x.\; \const{C}( \vec w )$,
+If $\vec u$ are normal and $t = \nf{(\MU x.\; \const{C}( \vec u ) )} = \MU x.\; \const{C}( \vec w )$,
 then $\vec u \vsim t\interp{\vec w}$.
@@
 Inversely, if the value assigned to ${\Varec{x}}$ is not $\alpha$-equivalent to any term in $\mathcal{T}^\M_{{\Varec{x}}}( \Val )$,
-then the values in the range of $\Val$ remain $\alpha$-disequivalent.
+the values in the range of $\Val$ remain $\alpha$-disequivalent.
 For example, we may assign values such as 
@@
 If $\Val$ is constructed for a saturated $\Ec$,
-and $\Val^\ast$ is a completion of $\Val$ for normal model $\M$, then:
+and $\Val^\ast$ is a completion of $\Val$ for normal model $\M$, then
 \begin{enumerate}
@@
 which is $\Val^\ast \ec{t}$.
-If $k \neq i$, then we interpret $\const s^j_k( \M( u ) )$ as the value in $\M( \tau )$ that is $\alpha$-equivalent with $\Val^\ast \ec{ t }$.
+If $k \neq i$, we interpret $\const s^j_k( \M( u ) )$ as the value in $\M( \tau )$ that is $\alpha$-equivalent with $\Val^\ast \ec{ t }$.
 Since \rn{Cong} does not apply and due to Lemma~\ref{lem:model-completion}.3,
@@
 if $\const s^j_i( t )$ is a term in $\tEc$ and $t$ contains the constructor
-$\const{C}_i( \vec u )$, then we immediately infer $\const s^j_i( t ) \teq u_j$, whereas the calculus we presented would apply \rn{Split} and \rn{Inject} before inferring this equality.
+$\const{C}_i( \vec u )$, we immediately infer $\const s^j_i( t ) \teq u_j$, whereas the calculus we presented would apply \rn{Split} and \rn{Inject} before inferring this equality.
 For the sake of reducing the number of unique constraints considered by the calculus, we compute a normal form for constraints as a preprocessing step.
@@
 or does nothing.
-When $M$ is a complete assignment for $F$, then a theory solver can only choose to do nothing when $M_i$ is indeed $T_i$-satisfiable.
+When $M$ is a complete assignment for $F$, a theory solver can only choose to do nothing when $M_i$ is indeed $T_i$-satisfiable.
 

### 2014-12-19 tex/codatadp.tex
more

@@
 Let $\SSSS \ast$ be the union of $\SSSS \ast_t$ for all $t \in T$,
-and let $\tcEc$ be the set of subterms of $\Ec \mathrel\cup \{ \const C_i( \const s^1_i( t ), \ldots, \const s^{a_i}_i( t ) ) \mid t : \delta \in \SSS \mathrel\cup \SSSS \ast,\, \const C_i \in \Ctr^\delta \}$.
+and let $\tcEc$ be the set of subterms of $\Ec \mathrel\cup \{ \const C_i( \const s^1_i( t ), \ldots, \const s^{n_i}_i( t ) ) \mid t : \delta \in \SSS \mathrel\cup \SSSS \ast,\, \const C_i \in \Ctr^\delta \}$.
 In a derivation tree with root node~$\Ec$,
@@
 whose root node is the result of applying the rule \rn{Split} on term $t$ of type $\delta$,
-by the induction hypothesis $\Ec \mathrel\cup t \teq \const C_i( \const s^1_i( t ), \ldots, \const s^{a_i}_i( t ) )$ is unsatisfiable
+by the induction hypothesis $\Ec \mathrel\cup t \teq \const C_i( \const s^1_i( t ), \ldots, \const s^{n_i}_i( t ) )$ is unsatisfiable
 for each $\const C_i \in \Ctr^\delta$.
-Since by exhaustiveness, all models of $\thD$ entail (exactly) one equality $t \teq \const C_i( \const s^1_i( t ), \ldots, \const s^{a_i}_i( t ) )$,
+Since by exhaustiveness, all models of $\thD$ entail (exactly) one equality $t \teq \const C_i( \const s^1_i( t ), \ldots, \const s^{n_i}_i( t ) )$,
 we know $\Ec$ is $\thD$-unsatisfiable.
@@
 For the \rn{Split} rule, we add a clause of the form
-$(\const{C}_1( \const s^1_1( t ), \ldots, \const s^{a_1}_1( t ) ) \vee \ldots \vee t \teq \const{C}_n( \const s^1_n( t ), \ldots, \const s^{a_n}_n( t ) ))$
+$(\const{C}_1( \const s^1_1( t ), \ldots, \const s^{n_1}_1( t ) ) \vee \ldots \vee t \teq \const{C}_m( \const s^1_m( t ), \ldots, \const s^{n_m}_m( t ) ))$
 to $F$.

### 2014-12-19 tex/codatadp.tex
more

@@
 \newcommand{\ttpath}[3]{\mathcal{P}_{ #2 \rightarrow #3 }( #1 )}
-\newcommand{\interp}[1]{[\![ #1 ]\!]}
+\newcommand{\interp}[1]{\llbracket #1 \rrbracket}
 \newcommand{\dpath}[3]{\delta^{#1}_{#2}( #3 )}
@@
 For \rn{Clash}, this is a consequence of distinctness.
-For \rn{Acyclic}, the construction of $\Val$ indicates that the class of values that $t_1$ can take in models of $\Ec$ is infeasible,
+For \rn{Acyclic}, the construction of $\Val$ indicates that the class of values that $t$ can take in models of $\Ec$ is infeasible,
 and thus $\Ec$ is unsatisfiable.
@@
 Otherwise, the child of $\Ec$ is a closed derivation tree
-whose root node is $\Ec \mathrel\cup t_1 \teq t_2$ obtained by applying one of the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, \rn{Inject}, \rn{Unique}, or \rn{Singleton}.
-In all cases, we have that $\Ec \models_{\thD} t_1 \teq t_2$.
+whose root node is $\Ec \mathrel\cup t \teq u$ obtained by applying one of the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, \rn{Inject}, \rn{Unique}, or \rn{Singleton}.
+In all cases, we have that $\Ec \models_{\thD} t \teq u$.
 For \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, this is a consequence of equality reasoning.
 For \rn{Inject}, this is a consequence of injectivity.
-For \rn{Unique}, our construction of $\Val$ indicates that the value of $t_1$ and $t_2$ are equivalent in all models of $\Ec$.
-For \rn{Singleton}, clearly $t_1$ and $t_2$ must have the same value since the cardinality of their type is one.
-By the induction hypothesis, we have $\Ec \mathrel\cup t_1 \teq t_2$ is $\thD$-unsatisfiable,
+For \rn{Unique}, our construction of $\Val$ indicates that the value of $t$ and $u$ are equivalent in all models of $\Ec$.
+For \rn{Singleton}, $t$~and~$u$ must have the same value since the cardinality of their type is $1$.
+By the induction hypothesis, we have $\Ec \mathrel\cup t \teq u$ is $\thD$-unsatisfiable,
 and thus $\Ec$ is $\thD$-unsatisfiable.
@@
 $\thD$-satisfiable.
-To do so, we exhibit a particular model $\M$ of $\thD$ that satisfies $\Ec$.
-First, we require additional terminology concerning $\mu$-terms.
-
-For a $\mu$-term $t_1$ with subterm $t_2$,
-the \emph{interpretation of $t_2$ in $t_1$} is the $\mu$-term $t_1\interp{t_2}^\emptyset$ as returned by the following recursive procedure,
-where $x \mapsto \MU x.\; \const{C}( \vec u ) \in t_1$ indicates that $\MU x.\; \const{C}( \vec u )$ is the subterm of $t_1$ that binds this occurrence of variable $x$.
+To do so, we exhibit a specific model $\M$ of $\thD$ that satisfies $\Ec$.
+
+First, we need additional terminology concerning $\mu$-terms.
+For a $\mu$-term $t$ with subterm $u$,
+the \emph{interpretation of\/ $u$ in\/ $t$} is the $\mu$-term $t\interp{u}^\emptyset$ as returned by the following recursive procedure,
+where $x \mapsto \MU x.\; \const{C}( \vec u ) \in t$ indicates that $\MU x.\; \const{C}( \vec u )$ is the subterm of $t$ that binds this occurrence of variable $x$:
 \[\begin{array}{r@{}c@{}l}
-t_1\interp{x}^M & {} \,=\, {} &
+t\interp{x}^M & {} \,=\, {} &
 \textstyle\begin{cases}
-    \MU x.\; \const{C}( t_1\interp{ \vec u}^{M \mathrel\cup \{ x \}}) & \text{if }x \not\in M, x \mapsto \MU x.\; \const{C}( \vec u ) \in t_1  \\[-\jot]
+    \MU x.\; \const{C}( t\interp{ \vec u}^{M \mathrel\cup \{ x \}}) & \text{if }x \not\in M, x \mapsto \MU x.\; \const{C}( \vec u ) \in t  \\[-\jot]
     x & \text{otherwise}
 \end{cases} \\
-t_1\interp{\MU x.\; \const{C}( \vec u )}^M & = &
+t\interp{\MU x.\; \const{C}( \vec u )}^M & = &
 \textstyle\begin{cases}
-    \MU x.\; \const{C}( t_1\interp{ \vec u }^{M \mathrel\cup \{ x \}} ) & \text{if }x \not\in M \\[-\jot]
+    \MU x.\; \const{C}( t\interp{ \vec u }^{M \mathrel\cup \{ x \}} ) & \text{if }x \not\in M \\[-\jot]
     x & \text{otherwise}
@@
 \end{array}\]
-We often write $t_1\interp{t_2}$ as shorthand for $t_1\interp{t_2}^\emptyset$.
-
-Given a $\mu$-term $t_1$ of the form $\MU x.\; \const{C}( u_1, \ldots, u_n )$ with subterm $t_2$,
-we say $t_2$ is a \emph{bisimilar subterm} of $t_1$ if and only if it is of the form
-$\MU y.\; \const{C}( v_1, \ldots, v_n )$ and $t_1\interp{ u_i } \aequiv t_1\interp{ v_i }$ for $i = 1,\ldots, n$.
-The term $t_1$ is \emph{normal} if it does not contain a strict, bisimilar subterm
+We write $t\interp{u}$ to abbreviate $t\interp{u}^\emptyset$.
+
+Given a $\mu$-term $t$ of the form $\MU x.\; \const{C}( t_1, \ldots, t_n )$ with subterm $u$,
+we say $u$ is a \emph{bisimilar subterm} of $t$ if it is of the form
+$\MU y.\; \const{C}( u_1, \ldots, u_n )$ and $t\interp{ t_i } \aequiv t\interp{ u_i }$ for $i = 1,\ldots, n$.
+The term $t$ is \emph{normal} if it does not contain a strict, bisimilar subterm
 and all of its strict subterms are also normal.
-For example, the $\mu$-term $t_1$ of the form $\MU x.\; \const{C}( \MU y.\; \const{C}( y ) )$ is not normal,
-since $\MU y.\; \const{C}( y )$ is a bisimilar subterm of $t_1$.
-In particular, note that their arguments have the same interpretation in $t_1$,
+For example, the $\mu$-term $t = \MU x.\; \const{C}( \MU y.\; \const{C}( y ) )$ is subnormal,
+since $\MU y.\; \const{C}( y )$ is a bisimilar subterm of $t$.
+In particular, note that their arguments have the same interpretation in $t$,
 that is,
-$t_1\interp{\MU y.\; \const{C}( y )}^\emptyset =$
-$\MU y.\; \const{C}( t_1\interp{y}^{\{y\}} ) =$
+$t\interp{\MU y.\; \const{C}( y )}^\emptyset =$
+$\MU y.\; \const{C}( t\interp{y}^{\{y\}} ) =$
 $\MU y.\; \const{C}( y )$
 is syntactically equivalent to
-$t_1\interp{y}^\emptyset =$
-$\MU y.\; \const{C}( t_1\interp{y}^{\{y\}} ) =$
+$t\interp{y}^\emptyset =$
+$\MU y.\; \const{C}( t\interp{y}^{\{y\}} ) =$
 $\MU y.\; \const{C}( y )$.
-The $\mu$-term $t_1$ of the form $\MU x.\; \const{C}( \MU y.\; \const{C}( x ) )$ is also not normal,
-since $\MU y.\; \const{C}( x )$ is a bisimilar subterm of $t_1$,
+The $\mu$-term $t$ of the form $\MU x.\; \const{C}( \MU y.\; \const{C}( x ) )$ is also not normal,
+since $\MU y.\; \const{C}( x )$ is a bisimilar subterm of $t$,
 noting that
-$t_1\interp{\MU y.\; \const{C}( x )}^\emptyset =$
-$\MU y.\; \const{C}( t_1\interp{x}^{\{y\}} ) =$
-$\MU y.\; \const{C}( t_1\interp{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) ) }^{\{y\}} ) =$
-$\MU y.\; \const{C}( \MU x.\; \const{C}( t_1\interp{\MU y.\; \const{C}( x )}^{\{x,y\}} ) ) =$
+$t\interp{\MU y.\; \const{C}( x )}^\emptyset =$
+$\MU y.\; \const{C}( t\interp{x}^{\{y\}} ) =$
+$\MU y.\; \const{C}( t\interp{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) ) }^{\{y\}} ) =$
+$\MU y.\; \const{C}( \MU x.\; \const{C}( t\interp{\MU y.\; \const{C}( x )}^{\{x,y\}} ) ) =$
 $\MU y.\; \const{C}( \MU x.\; \const{C}( y ) )$
-is syntactically equivalent to $t_1\interp{x}^\emptyset = t_1$ for the renaming $\{ x \mapsto y, y \mapsto x \}$.
+is syntactically equivalent to $t\interp{x}^\emptyset = t$ for the renaming $\{ x \mapsto y, y \mapsto x \}$.
 
@@
 to ${\Varec{x}}$.
-Legal substitutions for $v_{\ec{x}}$ may cause the range of $\Val$ to contain non-normal terms.
+Legal substitutions for $v_{\ec{x}}$ may cause the range of $\Val$ to contain subnormal terms.
 After applying the substitution for the third term mentioned above,

### 2014-12-19 tex/codatadp.tex
more

@@
 \newcommand{\VAREC}[1]{\widetilde{\vphantom{\scriptstyle x^i}\smash{#1}}}
-\newcommand{\nf}[1]{{#1}\!\!\downarrow}
+\newcommand{\nf}[1]{{{#1}{\downarrow}}}
 \newcommand{\aequiv}{\mathrel{=_\alpha}}
@@
 and all of its strict subterms are also normal.
-For example, the $\mu$-term $t = \MU x.\; \const{C}( \MU y.\; \const{C}( y ) )$ is subnormal,
+For example, the $\mu$-term $t = \MU x.\; \const{C}( \MU y.\; \const{C}( y ) )$ is abnormal,
 since $\MU y.\; \const{C}( y )$ is a bisimilar subterm of $t$.
@@
 $\MU y.\; \const{C}( y )$
-is syntactically equivalent to
+is $\alpha$-equivalent to
 $t\interp{y}^\emptyset =$
@@
 $\MU y.\; \const{C}( \MU x.\; \const{C}( y ) )$
-is syntactically equivalent to $t\interp{x}^\emptyset = t$ for the renaming $\{ x \mapsto y, y \mapsto x \}$.
-
-For any $\mu$-term $t$ of the form $\MU x.\; \const C( \vec u )$, we may recursively construct a normal $\mu$-term $\nf{t}$
-(which we call the \emph{normal form} of $t$)
-that is a renaming of $v$ by replacing all of the bisimilar subterms of $t$ with $x$,
-(which we call the \emph{normal form} of $t$)
-that is a renaming of $t$ by replacing all of the bisimilar subterms of $t$ with $x$,
+is $\alpha$-equivalent to $t\interp{x}^\emptyset = t$.
+% for the renaming $\{ x \mapsto y, y \mapsto x \}$.
+
+For any $\mu$-term $t$ of the form $\MU x.\; \const C( \vec u )$, we can
+recursively construct the \emph{normal form} $\nf{t}$ of $t$,
+that is a renaming of $v$ by replacing all of the bisimilar subterms of $t$ with $x$
 %and by replacing all of the nonnormal subterms of $t$ with their corresponding normal forms.
@@
 to ${\Varec{x}}$.
-Legal substitutions for $v_{\ec{x}}$ may cause the range of $\Val$ to contain subnormal terms.
+Legal substitutions for $v_{\ec{x}}$ may cause the range of $\Val$ to contain abnormal terms.
 After applying the substitution for the third term mentioned above,

### 2014-12-19 tex/codatadp.tex
use a bracketing notation for normal forms, so we don't need an ugly combination of parentheses and arrow (if it were not for the parentheses, I would have liked the arrow)

@@
 \newcommand{\VAREC}[1]{\widetilde{\vphantom{\scriptstyle x^i}\smash{#1}}}
-\newcommand{\nf}[1]{{{#1}{\downarrow}}}
+%\newcommand{\nf}[1]{{{#1}{\downarrow}}}
+\newcommand{\nf}[1]{\lfloor#1\rfloor}
 \newcommand{\aequiv}{\mathrel{=_\alpha}}
@@
 and by normalizing $t$'s subterms.
-For example, $\nf{(\MU x.\; \const{C}( \MU y.\; \const{C}( x ) ))} = \MU x.\; \const{C}( x )$.
+For example, $\nf{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) )} = \MU x.\; \const{C}( x )$.
 
@@
 \label{lem:mu-norm-arg}
-If $\vec u$ are normal and $t = \nf{(\MU x.\; \const{C}( \vec u ) )} = \MU x.\; \const{C}( \vec w )$,
+If $\vec u$ are normal and $t = \nf{\MU x.\; \const{C}( \vec u )} = \MU x.\; \const{C}( \vec w )$,
 then $\vec u \vsim t\interp{\vec w}$.
@@
 in $\M( \tau )$ that is $\alpha$-equivalent to
-$\nf{(\MU x.\; \const{C}( \M( \vec t \,) ))}$ where $x$ is fresh, and
+$\nf{\MU x.\; \const{C}( \M( \vec t \,) )}$ where $x$ is fresh, and
 \item
@@
 }{
-  \Val := \nf{(\Val \{ x \mapsto \MU x.\; t \})}
+  \Val := \nf{\Val \{ x \mapsto \MU x.\; t \}}
 }
@@
 After assigning this value, we normalize all terms in the range of $\Val$.
-%We then let $\Val^\ast \ec{x}$ be $\nf{(\Val' \ec{x})}$ for all $\ec{x}$.
+%We then let $\Val^\ast \ec{x}$ be $\nf{\Val' \ec{x}}$ for all $\ec{x}$.
 
@@
 Since \rem{TODO}
-we have $\nf{(\Val \ec{x})} \aequiv \nf{(\Val \ec{y})}$ if and only if $\ec{x} = \ec{y}$.
+we have $\nf{\Val \ec{x}} \aequiv \nf{\Val \ec{y}}$ if and only if $\ec{x} = \ec{y}$.
 We show that for any substitution $\sigma$ applied to $\Val$ while constructing its completion is such that
-$\nf{(\Val \ec{x} \sigma)} \aequiv \nf{(\Val \ec{y} \sigma)}$ if and only if $\ec{x} = \ec{y}$.
+$\nf{\Val \ec{x} \sigma} \aequiv \nf{\Val \ec{y} \sigma}$ if and only if $\ec{x} = \ec{y}$.
 \rem{TODO}
@@
 If $t$ is a constructor term of the form $\const{C}( u_1, \ldots, u_n )$,
-then $\M( t )$ is $\alpha$-equivalent with $\nf{(\MU x.\; \const{C}( \M( u_1 ), \ldots, \M( u_n ) ) )}$ for fresh $x$,
-which by the induction hypothesis is $\alpha$-equivalent with $\nf{(\MU x.\; \const{C}( \Val^\ast \ec{ u_1 }, \ldots, \Val^\ast \ec{ u_n } ) )}$, call this term $t_u$.
+then $\M( t )$ is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \M( u_1 ), \ldots, \M( u_n ) ) }$ for some fresh $x$,
+which by the induction hypothesis is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \Val^\ast \ec{ u_1 }, \ldots, \Val^\ast \ec{ u_n } )}$, call this term $t_u$.
 By Lemma~\ref{lem:model-completion}.2 and due to our construction of $\Val^\ast$, we have $\Val^\ast \ec{t}$ is a term

### 2014-12-19 tex/codatadp.tex
more

@@
 $\MU y.\; \const{C}( y )$.
-The $\mu$-term $t$ of the form $\MU x.\; \const{C}( \MU y.\; \const{C}( x ) )$ is also not normal,
+The $\mu$-term $t$ of the form $\MU x.\; \const{C}( \MU y.\; \const{C}( x ) )$ is also abnormal,
 since $\MU y.\; \const{C}( x )$ is a bisimilar subterm of $t$,
 noting that
-$t\interp{\MU y.\; \const{C}( x )}^\emptyset =$
-$\MU y.\; \const{C}( t\interp{x}^{\{y\}} ) =$
-$\MU y.\; \const{C}( t\interp{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) ) }^{\{y\}} ) =$
-$\MU y.\; \const{C}( \MU x.\; \const{C}( t\interp{\MU y.\; \const{C}( x )}^{\{x,y\}} ) ) =$
-$\MU y.\; \const{C}( \MU x.\; \const{C}( y ) )$
+$t\interp{\MU y.\; \const{C}( x )}^\emptyset =
+\MU y.\; \const{C}( t\interp{x}^{\{y\}} ) =
+\MU y.\; \const{C}( t\interp{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) ) }^{\{y\}} ) =
+\MU y.\; \const{C}( \MU x.\; \const{C}( t\interp{\MU y.\; \const{C}( x )}^{\{x,y\}} ) ) =
+\MU y.\allowbreak\; \const{C}( \MU x.\; \const{C}( y ) )$
 is $\alpha$-equivalent to $t\interp{x}^\emptyset = t$.
@@
 For any $\mu$-term $t$ of the form $\MU x.\; \const C( \vec u )$, we can
-recursively construct the \emph{normal form} $\nf{t}$ of $t$,
-that is a renaming of $v$ by replacing all of the bisimilar subterms of $t$ with $x$
-%and by replacing all of the nonnormal subterms of $t$ with their corresponding normal forms.
+recursively construct its \emph{normal form} $\nf{t}$,
+which is a renaming of $v$ by replacing all of the bisimilar subterms of $t$ with $x$
+%and by replacing all of the abnormal subterms of $t$ with their corresponding normal forms.
 and by normalizing $t$'s subterms.
@@
 \label{lem:mu-norm-arg}
-If $\vec u$ are normal and $t = \nf{\MU x.\; \const{C}( \vec u )} = \MU x.\; \const{C}( \vec w )$,
+If $\vec u$ are in normal form and $t = \nf{\MU x.\; \const{C}( \vec u )} = \MU x.\; \const{C}( \vec w )$,
 then $\vec u \vsim t\interp{\vec w}$.
@@
 \label{lem:mu-cong}
-If $t_u = \MU x.\; \const{C}( \vec u )$ and $t_w = \MU y.\; \const{C}( \vec w )$ are normal,
+If $t_u = \MU x.\; \const{C}( \vec u )$ and $t_w = \MU y.\; \const{C}( \vec w )$ are in normal form,
 then $t_u\interp{ \vec u } \vsim t_w\interp{ \vec w }$ if and only if $t_u \vsim t_w$.
@@
 
-We now define a class of models that we consider for a saturated set of equalities $\Ec$.
-In the following, we write $\M( \tau )$ to denote the interpretation type $\tau$ in $\M$,
+We now define a class of models that we consider for a saturated set of equalities~$\Ec$.
+Let $\M( \tau )$ denote the interpretation type $\tau$ in $\M$,
 that is, a set of domain elements for that type.
-We write $\M( t )$ for a term $t$ to denote the interpretation of $t$ in $\M$,
+For a term $t$, let $\M( t )$ denote the interpretation of $t$ in $\M$,
 that is, an element of $\M( \tau )$ where $\tau$ is the type of $t$.
-\rem{Clarify that non-codatatype terms are interpreted as $\mu$-terms, although dropping the $\mu$'s gives the standard representation?}
+\rem{Clarify that non-codatatype terms are interpreted as $\mu$-terms, although dropping the $\mu$'s gives the standard representation?
+(GOOD IDEA. A BIT OF REDUNDANCY CANNOT HURT.)}
 
@@
 \rm
-A model $\M$ is \emph{normal} if and only if:
+A model $\M$ is \emph{normal} if these conditions are met:
 \begin{enumerate}
 \item
-for all types $\tau$,
+For all types $\tau$,
 $\M( \tau )$ is a maximal set of closed normal $\mu$-terms of that type that are
 unique up to renaming of $\mu$-bound variables,
-and acyclic when $\tau \not\in \Codata$,
+and acyclic when $\tau \not\in \Codata$.
 \item
-for all constructor terms $\const{C}( \vec t \,)$ of type $\tau$,
+For all constructor terms $\const{C}( \vec t \,)$ of type $\tau$,
 $\M( \const{C}( \vec t \,) )$ is the value
 in $\M( \tau )$ that is $\alpha$-equivalent to
-$\nf{\MU x.\; \const{C}( \M( \vec t \,) )}$ where $x$ is fresh, and
+$\nf{\MU x.\; \const{C}( \M( \vec t \,) )}$ where $x$ is fresh.
 \item
-for all selector terms $\const s^j_i( t )$ of type $\tau$,% for some $\vec u$,
-$\M( \const s^j_i( t ) )$ is the value
+For all selector terms $\const s^k_i( t )$ of type $\tau$, % for some $\vec u$,
+$\M( \const s^k_i( t ) )$ is the value
 in $\M( \tau )$ that is $\alpha$-equivalent to
@@
 Thus, when constructing a model $\M$ for $\Ec$,
-it remains only to define how $\M$ interprets wrong-applied selector terms and variables.
+it remains only to define how $\M$ interprets wrongly applied selector terms and variables.
 For the latter, this will be based on the mapping $\Val$ constructed in phase~2 of the calculus.
-We compute the \emph{completion} of $\Val$ for normal model $\M$, which we write as $\Val^\ast$,
-which we construct by assigning values from $\M$ to unassigned variables in the domain of $\Val$.
-
-First, we need the following definition.
+We compute the \emph{completion} $\Val^\ast$ of $\Val$ for normal model $\M$
+by assigning values from $\M$ to unassigned variables in the domain of $\Val$.
+
+%First, we need the following definition.
 We write $t \vsimv{x} s$ if $\mu$-terms $t$ and $s$ are syntactically equivalent
@@
 such that $v \vsimv{x} t_1\interp{t_2}$ for some term $t_1$ with subterm $t_2$ occurring in the range of $\Val$.
-We construct $\Val^\ast$ by exhaustively applying the following update to $\Val$:
-
-\(
+
+We construct $\Val^\ast$ by exhaustively applying the following rule to $\Val$:
+\[
+\hbox{\(
 \inferrule{
@@
   \quad
-  (\MU x.\; t) \aequiv v
+  \MU x.\; t \aequiv v
   \quad
@@
 \)
-
+\rn{Assign}}
+\]
+%
 In this update, we choose an unassigned variable in $\Val$
@@
 We consider a normal model $\M$
-that interprets wrongly-applied selectors based on equality information in $\Fc$,
+that interprets wrongly applied selectors based on equality information in $\Fc$,
 and interprets the variables of $\Fc$ based on the completion $\Val^\ast$ of the mapping $\Val$ from phase~2.
@@
 Since \rn{Cong} does not apply and due to Lemma~\ref{lem:model-completion}.3,
-pairs of wrongly-applied selectors are assigned the same value if and only if they reside in the same equivalence class,
+pairs of wrongly applied selectors are assigned the same value if and only if they reside in the same equivalence class,
 and thus this definition is well-defined.

### 2014-12-19 tex/codatadp.tex
"add additional" etc.

@@
 and assign it a value that is not $\alpha$-equivalent to one occurring in $\mathcal{T}^\M_x(\Val)$.
-Since this update removes one variable from the set $FV( \Val )$ and does not add any additional variables to $\FV( \Val )$,
+Since this update removes one variable from the set $FV( \Val )$ and does not add any variables to $\FV( \Val )$,
 it can only be applied a finite number of times.
@@
 \begin{examplex}
-Let $\tau$ be a type in $\Codata$ with three constructors $\const{C}$, $\const{D}$, and $\const{E}$ each having type $\tau \rightarrow \tau$.
+Let $\delta$ be a codatatype with three constructors $\const{C}, \const{D}, \const{E} :\delta \rightarrow \delta$.
 Let $\Ec$ be the set
-$\{$
-$y_1 \teq \const{C}( y_2 )$,
-$y_2 \teq \const{D}( x )$,
-$z_1 \teq \const{E}( z_2 )$,
-$z_2 \teq \const{C}( z_1 )$,
-$w_1 \teq \const{C}( x )$,
-$w_2 \teq \const{E}( w_3 )$,
-$x \not\teq y_2$,
-$\}$.
+$\{y_1 \teq \const{C}( y_2 ),\;
+y_2 \teq \const{D}( x ),\;
+z_1 \teq \const{E}( z_2 ),\;
+z_2 \teq \const{C}( z_1 ),\;
+w_1 \teq \const{C}( x ),\;
+w_2 \teq \const{E}( w_3 ),\;
+x \not\teq y_2\}$.
 After running the calculus to saturation on $\Ec$, the mapping $\Val$ contains
@@
 %The former comes with no loss of generality since new equalities of the form $y \teq t$ for fresh variable $y$ can be added to $\Fc$ without affecting its satisfiability,
-%The latter also comes with no loss of generality since nested applications of selectors in $\Fc_0$ can be replaced by fresh variables while adding additional equalities to $\Fc$.
+%The latter also comes with no loss of generality since nested applications of selectors in $\Fc_0$ can be replaced by fresh variables while adding equalities to $\Fc$.
 We consider a normal model $\M$
@@
 Each $T_i$ solver either
-reports that a subset $C$ of $M_i$ is $T_i$-unsatisfiable, in which case $\neg C$ is added as a additional clause to $F$,
-adds an additional clause $C$ to $F$,
+reports that a subset $C$ of $M_i$ is $T_i$-unsatisfiable, in which case $\neg C$ is added to $F$,
+adds a clause $C$ to $F$,
 or does nothing.

### 2014-12-19 tex/codatadp.tex
more

@@
 In the following, we consider only normal models.
-Thus, when constructing a model $\M$ for $\Ec$,
+When constructing a model $\M$ for $\Ec$,
 it remains only to define how $\M$ interprets wrongly applied selector terms and variables.
@@
 %First, we need the following definition.
-We write $t \vsimv{x} s$ if $\mu$-terms $t$ and $s$ are syntactically equivalent
+We write $t \vsimv{x} u$ if $\mu$-terms $t$ and $u$ are syntactically equivalent
 for some renaming that avoids capturing any variable other than $x$.
@@
 $\MU x.\; \const{C}( x, x ) \not\vsimv{y} \MU y.\; \const{C}( x, y )$.
-For a variable $x$ of type $\tau$ and a fixed normal model $\M$,
+For a variable $x : \tau$ and a fixed normal model $\M$,
 we write $\mathcal{T}^\M_x( \Val )$ to denote the set consisting of all values $v$ from $\M( \tau )$
-such that $v \vsimv{x} t_1\interp{t_2}$ for some term $t_1$ with subterm $t_2$ occurring in the range of $\Val$.
+such that $v \vsimv{x} t\interp{u}$ for some term $t$ with subterm $u$ occurring in the range of $\Val$.
 
@@
 %
-In this update, we choose an unassigned variable in $\Val$
-and assign it a value that is not $\alpha$-equivalent to one occurring in $\mathcal{T}^\M_x(\Val)$.
-Since this update removes one variable from the set $FV( \Val )$ and does not add any variables to $\FV( \Val )$,
+The rule chooses an unassigned variable in $\Val$
+and assigns it a value that is not $\alpha$-equivalent to one occurring in $\mathcal{T}^\M_x(\Val)$.
+Since this update removes one variable from the set $\FV( \Val )$ and does not add any variables to $\FV( \Val )$,
 it can only be applied a finite number of times.

### 2014-12-19 tex/codatadp.tex
end of ex
tuned table

@@
 %%%
-%\newcommand\xend{{\hfill\footnotesize$\blacksquare$}}
+\newcommand\xend{{\hfill$\scriptstyle\blacksquare$}}
 %\newcommand\xend{{\qed}}
-\newcommand\xend{{}}
+%\newcommand\xend{{}}
 
@@
 After running the calculus to saturation on $\Ec$, the mapping $\Val$ contains
-\[\begin{array}{r@{}c@{}rlll@{\qquad}r@{}c@{}lllll}
+\[\begin{array}{r@{}c@{}r@{\;}l@{}ll@{\qquad}r@{}c@{}l@{\;}l}
 \Val \ec{y_1} & {}\,=\,{} & \MU {\Varec{y_1}}. & \const{C}( & \MU {\Varec{y_2}}. & \const{D}( {\Varec{x}} ) ) &
-\Val \ec{w_1} & {}\,=\,{} & \MU {\Varec{w_1}}. & \const{C}( {\Varec{x}} )      & \\
-\Val \ec{y_2} & = & \MU {\Varec{y_2}}. & \multicolumn{2}{l}{\const{D}( {\Varec{x}} )}      & &
+\Val \ec{w_1} & {}\,=\,{} & \MU {\Varec{w_1}}. & \const{C}( {\Varec{x}} )      \\
+\Val \ec{y_2} & = & \MU {\Varec{y_2}}. & \multicolumn{2}{@{}l@{}}{\const{D}( {\Varec{x}} )}      & &
 \Val \ec{w_2} & = & \MU {\Varec{w_2}}. & \const{E}( {\Varec{w_3}} ) \\
@@
 \Val \ec{z_2} & = & \MU {\Varec{z_2}}. & \const{C}( & \MU {\Varec{z_1}}. & \const{E}( {\Varec{z_2}} ) ) &
-\Val \ec{x} & = & {\Varec{x}} \\
+\Val \ec{x} & = & {\Varec{x}}
 \end{array}\]
-
+%
 To construct a completion of $\Val$, we must choose values for the free variables of $\Val$, ${\Varec{x}}$ and ${\Varec{w_3}}$.
@@
 \{
-\MU x.\; \const{C}( \const{D}( x )),
-\const{C}( \MU y.\; \const{D}( y )),
-\MU x.\; \const{D}( x ),
-\MU x.\; \const{E}( \const{C}( x )),
-\MU x.\; \const{C}( \const{E}( x )),
+\MU x.\; \const{C}( \const{D}( x )),\;
+\const{C}( \MU y.\; \const{D}( y )),\;
+\MU x.\; \const{D}( x ),\;
+\MU x.\; \const{E}( \const{C}( x )),\;
+\MU x.\; \const{C}( \const{E}( x )),\;
 \MU x.\; \const{C}( x )
@@
 Consider a model $\M$ that interprets variables in $\Ec$ based on the mapping $\Val$,
-for example, $x$ is interpretted as $\Val \ec{x}$, $y_1$ as $\Val \ec{y_1}$ and so on.
+for example, $x$ is interpreted as $\Val \ec{x}$, $y_1$ as $\Val \ec{y_1}$ and so on.
 Assigning a value for ${\Varec{x}}$ that is $\alpha$-equivalent to a value in $\mathcal{T}^\M_{{\Varec{x}}}( \Val )$

### 2014-12-19 tex/codatadp.tex
more

@@
 \begin{examplex}
-Let $\delta$ be a codatatype with three constructors $\const{C}, \const{D}, \const{E} :\delta \rightarrow \delta$.
+Let $\delta$ be a codatatype with the constructors $\const{C}, \const{D}, \const{E} :\delta \rightarrow \delta$.
 Let $\Ec$ be the set
-$\{y_1 \teq \const{C}( y_2 ),\;
+$\{
+w_1 \teq \const{C}( x ),\;
+w_2 \teq \const{E}( w_3 ),\;
+x \not\teq y_2,\;
+y_1 \teq \const{C}( y_2 ),\;
 y_2 \teq \const{D}( x ),\;
 z_1 \teq \const{E}( z_2 ),\;
-z_2 \teq \const{C}( z_1 ),\;
-w_1 \teq \const{C}( x ),\;
-w_2 \teq \const{E}( w_3 ),\;
-x \not\teq y_2\}$.
+z_2 \teq \const{C}( z_1 )
+\}$.
 After running the calculus to saturation on $\Ec$, the mapping $\Val$ contains
-\[\begin{array}{r@{}c@{}r@{\;}l@{}ll@{\qquad}r@{}c@{}l@{\;}l}
-\Val \ec{y_1} & {}\,=\,{} & \MU {\Varec{y_1}}. & \const{C}( & \MU {\Varec{y_2}}. & \const{D}( {\Varec{x}} ) ) &
-\Val \ec{w_1} & {}\,=\,{} & \MU {\Varec{w_1}}. & \const{C}( {\Varec{x}} )      \\
-\Val \ec{y_2} & = & \MU {\Varec{y_2}}. & \multicolumn{2}{@{}l@{}}{\const{D}( {\Varec{x}} )}      & &
-\Val \ec{w_2} & = & \MU {\Varec{w_2}}. & \const{E}( {\Varec{w_3}} ) \\
-\Val \ec{z_1} & = & \MU {\Varec{z_1}}. & \const{E}( & \MU {\Varec{z_2}}. & \const{C}( {\Varec{z_1}} ) ) &
-\Val \ec{w_3} & = & {\Varec{w_3}} \\
-\Val \ec{z_2} & = & \MU {\Varec{z_2}}. & \const{C}( & \MU {\Varec{z_1}}. & \const{E}( {\Varec{z_2}} ) ) &
-\Val \ec{x} & = & {\Varec{x}}
+\[\begin{array}{r@{}c@{}l@{\;}l@{\qquad}r@{}c@{}r@{\;}l@{}l@{\;}l}
+\Val \ec{w_1} & {}\,=\,{} & \MU {\Varec{w_1}}. & \const{C}( {\Varec{x}} ) &
+\Val \ec{y_1} & {}\,=\,{} & \MU {\Varec{y_1}}. & \const{C}( & \MU {\Varec{y_2}}. & \const{D}( {\Varec{x}} ) ) \\
+\Val \ec{w_2} & = & \MU {\Varec{w_2}}. & \const{E}( {\Varec{w_3}} ) &
+\Val \ec{y_2} & = & \MU {\Varec{y_2}}. & \multicolumn{2}{@{}l@{}}{\const{D}( {\Varec{x}} )} \\
+\Val \ec{w_3} & = & {\Varec{w_3}} &&
+\Val \ec{z_1} & = & \MU {\Varec{z_1}}. & \const{E}( & \MU {\Varec{z_2}}. & \const{C}( {\Varec{z_1}} ) ) \\
+\Val \ec{x} & = & {\Varec{x}} &&
+\Val \ec{z_2} & = & \MU {\Varec{z_2}}. & \const{C}( & \MU {\Varec{z_1}}. & \const{E}( {\Varec{z_2}} ) )
 \end{array}\]
 %
-To construct a completion of $\Val$, we must choose values for the free variables of $\Val$, ${\Varec{x}}$ and ${\Varec{w_3}}$.
-The content of the set $\mathcal{T}^\M_{{\Varec{x}}}( \Val )$ is $\alpha$-equivalent to
+To construct a completion of $\Val$, we must choose values for the free variables of $\Val$,
+namely ${\Varec{w_3}}$ and ${\Varec{x}}$.
+The content of the set \smash{$\mathcal{T}^\M_{{\Varec{x}}}( \Val )$} is $\alpha$-equivalent to
 \[%\begin{array}{c}
 \{
+\MU x.\; \const{C}( x ),\;
 \MU x.\; \const{C}( \const{D}( x )),\;
-\const{C}( \MU y.\; \const{D}( y )),\;
+\const{C}( \MU x.\; \const{D}( x )),\;
+\MU x.\; \const{C}( \const{E}( x )),\;
 \MU x.\; \const{D}( x ),\;
-\MU x.\; \const{E}( \const{C}( x )),\;
-\MU x.\; \const{C}( \const{E}( x )),\;
-\MU x.\; \const{C}( x )
+\MU x.\; \const{E}( \const{C}( x ))
 \}
@@
 \]
-Consider a model $\M$ that interprets variables in $\Ec$ based on the mapping $\Val$,
-for example, $x$ is interpreted as $\Val \ec{x}$, $y_1$ as $\Val \ec{y_1}$ and so on.
+Consider a model $\M$ that interprets variables in $\Ec$ based on the mapping $\Val$:
+$x$ is interpreted as $\Val \ec{x}$, $y_1$ as $\Val \ec{y_1}$, and so on.
 Assigning a value for ${\Varec{x}}$ that is $\alpha$-equivalent to a value in $\mathcal{T}^\M_{{\Varec{x}}}( \Val )$
@@
 which in turn may cause $\M$ to be inconsistent.
-For example, say we assign $\MU {\Varec{x}}.\; \const{D}( {\Varec{x}} )$ for ${\Varec{x}}$.
+For example, assign $\MU {\Varec{x}}.\; \const{D}( {\Varec{x}} )$ for ${\Varec{x}}$.
 After this substitution, $\Val \ec{y_2}$ is $\MU {\Varec{y_2}}.\; \const{D}( \MU {\Varec{x}}.\; \const{D}( {\Varec{x}} ) )$,
@@
 which is $\alpha$-equivalent to $\MU {\Varec{x}}.\; \const{D}( {\Varec{x}} )$.
-However, this contradicts the disequality $x \not\teq y_2$ in $\Ec$.
-Inversely, if the value assigned to ${\Varec{x}}$ is not $\alpha$-equivalent to any term in $\mathcal{T}^\M_{{\Varec{x}}}( \Val )$,
+However, this contradicts the disequality $x \not\teq y_2$ in~$\Ec$.
+On the other hand, if the value assigned to ${\Varec{x}}$ is not $\alpha$-equivalent to any term in $\mathcal{T}^\M_{{\Varec{x}}}( \Val )$,
 the values in the range of $\Val$ remain $\alpha$-disequivalent.
@@
 \label{lem:model-completion}
-If $\Val$ is constructed for a saturated $\Ec$,
+If $\Val$ is constructed for a saturated set $\Ec$
 and $\Val^\ast$ is a completion of $\Val$ for normal model $\M$, then
 \begin{enumerate}
-\item $\Val^\ast \ec{x}$ is $\alpha$-equivalent to a value in $\M( \tau )$, where $\ec{x}$ has type $\tau$,
+\item $\Val^\ast \ec{x}$ is $\alpha$-equivalent to a value in $\M( \tau )$, where $\ec{x}$ has type $\tau$.
 \item

### 2014-12-19 tex/codatadp.tex
more

@@
 If $\Val$ is constructed for a saturated set $\Ec$
-and $\Val^\ast$ is a completion of $\Val$ for normal model $\M$, then
+and $\Val^\ast$ is a completion of $\Val$ for normal model $\M$, the following holds:
 \begin{enumerate}
-\item $\Val^\ast \ec{x}$ is $\alpha$-equivalent to a value in $\M( \tau )$, where $\ec{x}$ has type $\tau$.
-\item
+\item[\rm 1.] $\Val^\ast \ec{x}$ is $\alpha$-equivalent to a value in $\M( \tau )$, where $\ec{x}$ has type $\tau$.
+\item[\rm 2.]
 All subterms $\MU {\Varec{x}}.\; t$ of $\Val^\ast \ec{y}$ are such that
@@
 %for all $t_1$ with subterm $t_2$ in the range of $\Val^\ast$ where $\muvar(t_2)={\Varec{x}}$,
-\item
+\item[\rm 3.]
 $\Val^\ast \ec{x} \vsim \Val^\ast \ec{y}$ if and only if $\ec{x} = \ec{y}$.
@@
 If $\tau \in \Codata$, then $\tau$ is corecursive by assumption, and by Lemma~\ref{lem:corecursive-singletons},
-the cardinality of $\tau$ must be one.
+the cardinality of $\tau$ must be $1$.
 Since \rn{Singleton} does not apply to $\Ec$,

### 2014-12-19 tex/codatadp.tex
mess with indices

@@
 Let $\SSSS \ast$ be the union of $\SSSS \ast_t$ for all $t \in T$,
-and let $\tcEc$ be the set of subterms of $\Ec \mathrel\cup \{ \const C_i( \const s^1_i( t ), \ldots, \const s^{n_i}_i( t ) ) \mid t : \delta \in \SSS \mathrel\cup \SSSS \ast,\, \const C_i \in \Ctr^\delta \}$.
+and let $\tcEc$ be the set of subterms of $\Ec \mathrel\cup \{ \const C_j( \const s^1_j( t ), \ldots, \const s^{n_j}_j( t ) ) \mid t : \delta \in \SSS \mathrel\cup \SSSS \ast,\, \const C_j \in \Ctr^\delta \}$.
 In a derivation tree with root node~$\Ec$,
@@
 whose root node is the result of applying the rule \rn{Split} on term $t$ of type $\delta$,
-by the induction hypothesis $\Ec \mathrel\cup t \teq \const C_i( \const s^1_i( t ), \ldots, \const s^{n_i}_i( t ) )$ is unsatisfiable
-for each $\const C_i \in \Ctr^\delta$.
-Since by exhaustiveness, all models of $\thD$ entail (exactly) one equality $t \teq \const C_i( \const s^1_i( t ), \ldots, \const s^{n_i}_i( t ) )$,
+by the induction hypothesis $\Ec \mathrel\cup t \teq \const C_j( \const s^1_j( t ), \ldots, \const s^{n_j}_j( t ) )$ is unsatisfiable
+for each $\const C_j \in \Ctr^\delta$.
+Since by exhaustiveness, all models of $\thD$ entail (exactly) one equality $t \teq \const C_j( \const s^1_j( t ), \ldots, \const s^{n_j}_j( t ) )$,
 we know $\Ec$ is $\thD$-unsatisfiable.
@@
 we say $u$ is a \emph{bisimilar subterm} of $t$ if it is of the form
-$\MU y.\; \const{C}( u_1, \ldots, u_n )$ and $t\interp{ t_i } \aequiv t\interp{ u_i }$ for $i = 1,\ldots, n$.
+$\MU y.\; \const{C}( u_1, \ldots, u_n )$ and $t\interp{ t_k } \aequiv t\interp{ u_k }$ for $k = 1,\ldots, n$.
 The term $t$ is \emph{normal} if it does not contain a strict, bisimilar subterm
@@
 \begin{proof}
-%Since $u_i$ is normal, $w_i$ is the result of replacing all subterms $v$ of $u_i$ such that $t\interp{v} \aequiv t$ with $x$.
+%Since $u_k$ is normal, $w_k$ is the result of replacing all subterms $v$ of $u_k$ such that $t\interp{v} \aequiv t$ with $x$.
 \rem{TODO}
@@
 \item
-For all selector terms $\const s^k_i( t )$ of type $\tau$, % for some $\vec u$,
-$\M( \const s^k_i( t ) )$ is the value
+For all selector terms $\const s^k_j( t )$ of type $\tau$, % for some $\vec u$,
+$\M( \const s^k_j( t ) )$ is the value
 in $\M( \tau )$ that is $\alpha$-equivalent to
 $\M( t )\interp{u_j}$
-when $\M( t )$ is of the form $\MU x.\; \const{C}_i( \vec u )$.
+when $\M( t )$ is of the form $\MU x.\; \const{C}_j( \vec u )$.
 \end{enumerate}
@@
 for all equivalence classes $\ec{z}$,
-each pair of constructor terms $\const{C}_i( \vec t )$ and $\const{C}_j( \vec u )$ in $\ec{z}$
-are such that $i = j$ since \rn{Clash} does not apply to $\Ec$,
+each pair of constructor terms $\const{C}_j( \vec t )$ and $\const{C}_{j'}( \vec u )$ in $\ec{z}$
+are such that $j = j'$ since \rn{Clash} does not apply to $\Ec$,
 and are such that $\ec{ \vec t } = \ec{ \vec u }$ since \rn{Decomp} does not apply to $\Ec$.
 Thus, 
-if $\const{C}_i( t_1, \ldots, t_n ) \in \ec{z}$,
+if $\const{C}_j( t_1, \ldots, t_n ) \in \ec{z}$,
 $\Val$ was constructed by applying a sequence of substitutions
-where ${\Varec{z}}$ was replaced only by $\MU {\Varec{z}}.\; \const{C}_i( {\Varec{t_1}}, \ldots, {\Varec{t_n}} )$.
+where ${\Varec{z}}$ was replaced only by $\MU {\Varec{z}}.\; \const{C}_j( {\Varec{t_1}}, \ldots, {\Varec{t_n}} )$.
 Assume $\Val \ec{y}$ has a subterm $\MU {\Varec{x}}.\; \const{C}( \vec t )$.
@@
 where $t_w\interp{w_i} = \Val^\ast \ec{u_i}$ for $i = 1, \ldots, n$.
-For each $i = 1, \ldots, n$, let $u_i'$ be the $i^{th}$ argument of $t_u$.
+For each $i = 1, \ldots, n$, let $u_i'$ be the $i$\vvthinspace th argument of $t_u$.
 By Lemma~\ref{lem:mu-norm-arg}, $t_u\interp{u_i'} \vsim \Val^\ast \ec{u_i}$,
@@
 since \rn{Split} does not apply,
-$\ec{u}$ must contain a term of the form $\const C_i( \const s^1_i( u ), \ldots, \const s^{n}_i( u ) )$ for some $i$.
+$\ec{u}$ must contain a term of the form $\const C_j( \const s^1_j( u ), \ldots, \const s^{n}_j( u ) )$ for some $j$.
 By Lemma~\ref{lem:model-completion}.2 and due to our construction of $\Val^\ast$, we have $\Val^\ast \ec{u}$ is a term $t_w$ of the form
-$\MU {\Varec{u}}.\; \const C_i( w_1, \ldots, w_n )$,
+$\MU {\Varec{u}}.\; \const C_j( w_1, \ldots, w_n )$,
 where $t_w \interp{ w_j } = \Val^\ast \ec{ \const s^j_k( u )}$. 
@@
 
-Discriminators are treated as predicate symbols, so a discriminator $\const{d}_i( t )$ is such that $\const{d}_i$
-is a predicate symbol, instead of reducing this constraint to $t \teq \const{C}_i( \const s^1_i( t ), \ldots, \const s^n_i( t ) )$.
+Discriminators are treated as predicate symbols, so a discriminator $\const{d}_j( t )$ is such that $\const{d}_j$
+is a predicate symbol, instead of reducing this constraint to $t \teq \const{C}_j( \const s^1_j( t ), \ldots, \const s^n_j( t ) )$.
 We have found that this leads to better performance since this reduction introduces terms more eagerly to $\tEc$.
@@
 Also, selectors are collapsed eagerly:
-if $\const s^j_i( t )$ is a term in $\tEc$ and $t$ contains the constructor
-$\const{C}_i( \vec u )$, we immediately infer $\const s^j_i( t ) \teq u_j$, whereas the calculus we presented would apply \rn{Split} and \rn{Inject} before inferring this equality.
+if $\const s^k_j( t )$ is a term in $\tEc$ and $t$ contains the constructor
+$\const{C}_j( \vec u )$, we immediately infer $\const s^k_j( t ) \teq u_k$, whereas the calculus we presented would apply \rn{Split} and \rn{Inject} before inferring this equality.
 For the sake of reducing the number of unique constraints considered by the calculus, we compute a normal form for constraints as a preprocessing step.
@@
 replace $t \teq s$ by $s \teq t$ if $s$ is less than $t$ based on a term ordering,
-replace $\const{C}_i( \vec t ) \teq \const{C}_j( \vec u )$ with $\bot$ when $i \neq j$,
-replace all selector terms of the form $\const{s}^{\,i}_j( \const{C}_j(t_1,\ldots,t_n) )$ by $t_i$,
-and replace all occurrences of discriminators $\const{d}_i( \const{C}_k( \vec t ) )$ by $\top$ if $i=k$ and $\bot$ if $i \neq k$.
+replace $\const{C}_j( \vec t ) \teq \const{C}_{j'}( \vec u )$ with $\bot$ when $j \neq j'$,
+replace all selector terms of the form $\const{s}^{\,k}_j( \const{C}_j(t_1,\ldots,t_n) )$ by $t_k$,
+and replace all occurrences of discriminators $\const{d}_j( \const{C}_{j'}( \vec t ) )$ by $\top$ if $j=j'$ and $\bot$ otherwise.
 
@@
 Then, at a high level, by the Nelson--Oppen combination framework,
-each $T_i$-solver for $i = 1, \ldots, n$ takes as input a combination of (1) the purified form of $T_i$-constraints occurring in $M$ where terms containing symbols not belonging to $T_i$ are replaced by fresh variables of the same type,
+each $T_i$-solver for $i = 1, \ldots, l$ takes as input a combination of (1) the purified form of $T_i$-constraints occurring in $M$ where terms containing symbols not belonging to $T_i$ are replaced by fresh variables of the same type,
 (2) additional equalities and disequalities between variables of types not belonging to $T_i$.

### 2014-12-19 tex/codatadp.tex
minor

@@
 
-\hyphenation{data-type data-types co-data-type co-data-types}
+\hyphenation{data-type data-types co-data-type co-data-types isa-belle sledge-hammer}
 
@@
 \cite{blanchette-et-al-2014-impl}, benchmarks are somewhat hard to come by. We
-included the theory files from the Isabelle distribution (the Distro) and the \emph{Archive
+included the theory files from the Isabelle distribution (Distro) and the \emph{Archive
 of Formal Proofs} (AFP) \cite{klein-et-al-afp} that define codatatypes falling

### 2014-12-19 tex/codatadp.tex

@@
 
-We construct $\Val^\ast$ by exhaustively applying the following rule to $\Val$:
+We construct $\Val^\ast$ by exhaustively applying the following rule to $\nf{\Val}$:
 \[
@@
 \inferrule{
-  x : \tau \in \FV( \Val )
+  {\Varec{x}} : \tau \in \FV( \Val )
   \quad
-  \MU x.\; t \aequiv v
+  \MU {\Varec{x}}.\; t \aequiv v
   \quad
@@
   \quad
-  v \not\in \mathcal{T}^\M_x(\Val)
+  v \not\in \mathcal{T}^\M_{\Varec{x}}(\Val)
 }{
-  \Val := \nf{\Val \{ x \mapsto \MU x.\; t \}}
+  \Val := \nf{\Val \{ {\Varec{x}} \mapsto \MU {\Varec{x}}.\; t \}}
 }
@@
 After assigning this value, we normalize all terms in the range of $\Val$.
-%We then let $\Val^\ast \ec{x}$ be $\nf{\Val' \ec{x}}$ for all $\ec{x}$.
 
@@
 to ${\Varec{x}}$.
-Legal substitutions for $v_{\ec{x}}$ may cause the range of $\Val$ to contain abnormal terms.
+Legal substitutions for ${\Varec{x}}$ may cause the range of $\Val$ to contain abnormal terms.
 After applying the substitution for the third term mentioned above,
@@
 
-%We first state the following properties of $\Val^\ast$.
-%where we write $\muvar( t )$ to denote $t$ if $t$ is a variable, or $x$ if $t$ is of the form $\MU x.\; u$.
+We first state the following properties of $\Val^\ast$,
+where we write $\muvar( t )$ to denote $t$ if $t$ is a variable, or ${\Varec{x}}$ if $t$ is of the form $\MU {\Varec{x}}.\; u$.
 
@@
 \item[\rm 2.]
-All subterms $\MU {\Varec{x}}.\; t$ of $\Val^\ast \ec{y}$ are such that
-$\Val^\ast \ec{y} \interp{ \MU {\Varec{x}}.\; t } = \Val^\ast \ec{x}$. 
+All subterms $t$ of $\Val^\ast \ec{y}$ with $\muvar(t) = {\Varec{x}}$ are such that
+$\Val^\ast \ec{y} \interp{ t } = \Val^\ast \ec{x}$. 
 %If $\const{C}( t_1, \ldots, t_n ) \in \ec{x}$,
@@
 Thus, 
-if $\const{C}_j( t_1, \ldots, t_n ) \in \ec{z}$,
 $\Val$ was constructed by applying a sequence of substitutions
-where ${\Varec{z}}$ was replaced only by $\MU {\Varec{z}}.\; \const{C}_j( {\Varec{t_1}}, \ldots, {\Varec{t_n}} )$.
-Assume $\Val \ec{y}$ has a subterm $\MU {\Varec{x}}.\; \const{C}( \vec t )$.
-We first show that our construction of $\Val$ is such that $\Val \ec{x} = \Val \ec{y}\interp{ \MU {\Varec{x}}.\; t }$.
-Notice that both $\Val \ec{x}$ and the subterm $\MU {\Varec{x}}.\; t$ of $\Val \ec{y}$ were
+where all substitutions for each variable ${\Varec{z}}$ 
+were uniquely of the form $\{ {\Varec{z}} \mapsto \const{C}_j( {\Varec{t_1}}, \ldots, {\Varec{t_n}} ) \}$
+for $\ec{z}$ that contain a constructor $\const{C}_j( t_1, \ldots, t_n )$.
+Say $\Val \ec{y}$ has a subterm $t$ where $\muvar(t) = {\Varec{x}}$.
+Both $\Val \ec{x}$ and the subterm $t$ of $\Val \ec{y}$ were
 constructed by applying a sequence of substitutions of the form mentioned above to ${\Varec{x}}$.
-We show that $\Val \ec{x} \interp{ \Val \ec{x} } = \Val \ec{y} \interp{ \MU {\Varec{x}}.t }$.
-\rem{TODO}
+Moreover, free variables ${\Varec{z}}$ in $\MU {\Varec{x}}.\; t$ that are bound in $\Val \ec{y}$ are interpreted in 
+the expansion of $\Val \ec{y} \interp{ \MU {\Varec{x}}.\; t }$ as a term
+constructed by a sequence of substitutions of the form mentioned above to ${\Varec{z}}$. \rem{more detail}
+Thus, we have
+$\Val \ec{y} \interp{ \MU {\Varec{x}}.\; t } = \Val \ec{x} \interp{ \Val \ec{x} } = \Val \ec{x}$, and thus 2 holds for $\Val$.
+Statement 3 holds for $\Val$ since \rn{Unique} does not apply.
+
+We now show that $\Val = \nf{(\Val)}$.
+Assume by contradition $\Val \ec{x} \neq \nf{(\Val \ec{x})}$ for some $\Val \ec{x}$ of minimal size.
+We have that $\Val \ec{x}$ is of the form $\MU {\Varec{x}}.\; \const{C}( t_1, \ldots, t_n )$.
+Due to our construction of $\Val$, 
+we know $\ec{x}$ contains a constructor $\const{C}( z_1, \ldots, z_n )$ and $\muvar( t_i ) = z_i$ for $i = 1, \ldots, n$.
+Since $\Val \ec{x}$ is a minimal, it
+contain a subterm of the form $\MU {\Varec{y}}.\; \const{C}( u_1, \ldots, u_n )$
+where $\Val \ec{x} \interp{ t_i } \aequiv \Val \ec{x} \interp{ u_i }$ for $i = 1, \ldots, n$.
+Due to our construction of $\Val$, $\ec{y}$ contains a constructor $\const{C}( w_1, \ldots, w_n )$ and $\muvar( u_i ) = w_i$ for $i = 1, \ldots, n$.
+Since \rn{Cong} does not apply to $\Ec$
+we have $\ec{w_j}$ and $\ec{z_j}$ are distinct for some $j$.
+By statement 2, $\Val \ec{x} \interp{ t_j } = \Val \ec{z_j}$
+and $\Val \ec{x} \interp{ u_j } = \Val \ec{w_j}$,
+which are not $\alpha$-equivalent by statement 3, contradicting the fact that $\MU {\Varec{y}}.\; \const{C}( u_1, \ldots, u_n )$ is a bisimilar subterm of $\Val \ec{x}$.
+Thus, $\Val = \nf{(\Val)}$, and statements 2 and 3 hold for $\nf{(\Val)}$.
+%For each $i = 1, \ldots, n$,
+%if ${\Varec{z_i}} = {\Varec{w_i}}$, then $w_i \teq z_i \in \Ec$.
+%If ${\Varec{z_i}} \neq {\Varec{w_i}}$, then by statement 2, $\Val \ec{x} \interp{ t_i } = \Val \ec{z_i}$
+%and $\Val \ec{x} \interp{ u_i } = \Val \ec{w_i}$.
+%By statement 3, we have that $\ec{z_i} = \ec{w_i}$, and thus $u_i \teq t_i \in \Ec$ for each $i = 1, \ldots, n$.
+%This is a contradiction since $\ec{x}$ and $\ec{y}$ are distinct, and \rn{Cong} does not apply to $\Ec$.
+%Thus, statements 2 and 3 hold for $\nf{(\Val)}$.
+
+We now show that if 2 and 3 hold for some $\Val_1$,
+then 2 and 3 also hold for $\nf{(\Val_1 \sigma)}$,
+where $\sigma$ is a substitution of the form $\{ {\Varec{x}} \mapsto \MU {\Varec{x}}.\; t \}$,
+where $\MU {\Varec{x}}.\; t$ is not $\alpha$-equivalent to a term in $\mathcal{T}^\M_{\Varec{x}}(\Val)$.
+
+%\rem{TODO}
 %Thus, the set of subterms which $\Val \ec{x}$ and $\MU v_{\ec{x}}.\; t$ differ 
@@
 %where $\MU v_{\ec{z}}.\; u$ and $\MU v_{\ec{z}}.\; u'$ were constructed by applying a sequence of substitutions of the form above to $v_{\ec{z}}$.
-Since $\Val^\ast$ is obtained by applying the same substitutions and normalizations to all terms in the range of $\Val$,
-we have $\Val^\ast \ec{y} \interp{ \MU {\Varec{x}}.\; t} = \Val^\ast \ec{x}$.
-
-To show 3,
-first note that since \rn{Unique} does not apply,
-$\Val \ec{x} \aequiv \Val \ec{y}$ if and only if $\ec{x} = \ec{y}$.
-Since \rem{TODO}
-we have $\nf{\Val \ec{x}} \aequiv \nf{\Val \ec{y}}$ if and only if $\ec{x} = \ec{y}$.
-We show that for any substitution $\sigma$ applied to $\Val$ while constructing its completion is such that
-$\nf{\Val \ec{x} \sigma} \aequiv \nf{\Val \ec{y} \sigma}$ if and only if $\ec{x} = \ec{y}$.
-\rem{TODO}
-Let $P$ be the set of positions for which $\Val \ec{x}$ and $\Val \ec{y}$ are not $\alpha$-equivalent,
-which is non-empty by assumption.
+%Since $\Val^\ast$ is obtained by applying the same substitutions and normalizations to all terms in the range of $\Val$,
+%we have $\Val^\ast \ec{y} \interp{ \MU {\Varec{x}}.\; t} = \Val^\ast \ec{x}$.
+
+%To show 3,
+%first note that since \rn{Unique} does not apply,
+%$\Val \ec{x} \aequiv \Val \ec{y}$ if and only if $\ec{x} = \ec{y}$.
+%Due to part 2, we have $\Val = \nf{(\Val)}$, and thus
+%$\nf{(\Val \ec{x})} \aequiv \nf{(\Val \ec{y})}$ if and only if $\ec{x} = \ec{y}$.
+%We show that for any substitution $\sigma$ applied to $\Val$ while constructing its completion is such that
+%$\nf{(\Val \ec{x} \sigma)} \aequiv \nf{(\Val \ec{y} \sigma)}$ if and only if $\ec{x} = \ec{y}$.
+%\rem{TODO}
+%Let $P$ be the set of positions for which $\Val \ec{x}$ and $\Val \ec{y}$ are not $\alpha$-equivalent,
+%which is non-empty by assumption.
 

### 2014-12-19 tex/codatadp.tex
omit names

@@
 $\thD$-satisfiability of sets of ground constraints over a fixed
-signature~$\Sig$. It is formulated as a calculus, whose rules can be applied in
-any order until saturation.
+signature~$\Sig$. It is formulated as a calculus, whose rules are applied
+until a contradiction is detected or until saturation.
 %
@@
 \)
-\rn{Unfold}}
+%\rn{Unfold}}
 \]
@@
 where $\const{C}(z,x) \in \ec{y}$ and $\const{C}(w,y) \in \ec{x}$ for some $\const{C} \in \Ctr$.
-A possible sequence of \rn{Unfold} rule applications is given below, omitting
+A possible sequence of unfoldings is given below, omitting
 trivial entries such as $\ec{w} \mapsto {\Varec{w}}$.
@@
 \)
-\rn{Assign}}
+%\rn{Assign}}
 \]

### 2014-12-20 tex/codatadp.tex
More work on Lemma

@@
 %\rn{Unfold}}
+}
 \]
@@
 If the children of $\Ec$ are closed derivation trees
-whose root node is the result of applying the rule \rn{Split} on term $t$ of type $\delta$,
+whose root nodes are the result of applying the rule \rn{Split} on term $t$ of type $\delta$,
 by the induction hypothesis $\Ec \mathrel\cup t \teq \const C_j( \const s^1_j( t ), \ldots, \const s^{n_j}_j( t ) )$ is unsatisfiable
@@
 recursively construct its \emph{normal form} $\nf{t}$,
-which is a renaming of $v$ by replacing all of the bisimilar subterms of $t$ with $x$
+by replacing all of the bisimilar subterms of $t$ with $x$
 %and by replacing all of the abnormal subterms of $t$ with their corresponding normal forms.
@@
 
-\begin{lemma}
-\label{lem:mu-cong}
-If $t_u = \MU x.\; \const{C}( \vec u )$ and $t_w = \MU y.\; \const{C}( \vec w )$ are in normal form,
-then $t_u\interp{ \vec u } \vsim t_w\interp{ \vec w }$ if and only if $t_u \vsim t_w$.
-\end{lemma}
-\begin{proof}
-\rem{TODO}
-\end{proof}
+%\begin{lemma}
+%\label{lem:mu-cong}
+%If $t_u = \MU x.\; \const{C}( \vec u )$ and $t_w = \MU y.\; \const{C}( \vec w )$ are in normal form,
+%then $t_u\interp{ \vec u } \vsim t_w\interp{ \vec w }$ if and only if $t_u \vsim t_w$.
+%\end{lemma}
+%\begin{proof}
+%\rem{TODO}
+%\end{proof}
 
@@
 %\rn{Assign}}
+}
 \]
@@
 We first show that 2 and 3 hold initially for $\Val$.
-Since \rn{Clash} does not apply to $\Ec$,
-for all equivalence classes $\ec{z}$,
+For all equivalence classes $\ec{z}$,
 each pair of constructor terms $\const{C}_j( \vec t )$ and $\const{C}_{j'}( \vec u )$ in $\ec{z}$
 are such that $j = j'$ since \rn{Clash} does not apply to $\Ec$,
-and are such that $\ec{ \vec t } = \ec{ \vec u }$ since \rn{Decomp} does not apply to $\Ec$.
+and are such that $\ec{ \vec t } = \ec{ \vec u }$ since \rn{Inject} does not apply to $\Ec$.
 Thus, 
 $\Val$ was constructed by applying a sequence of substitutions
-where all substitutions for each variable ${\Varec{z}}$ 
+where all substitutions for variables ${\Varec{z}}$ 
 were uniquely of the form $\{ {\Varec{z}} \mapsto \const{C}_j( {\Varec{t_1}}, \ldots, {\Varec{t_n}} ) \}$
-for $\ec{z}$ that contain a constructor $\const{C}_j( t_1, \ldots, t_n )$.
+when $\ec{z}$ contains a constructor $\const{C}_j( t_1, \ldots, t_n )$.
 Say $\Val \ec{y}$ has a subterm $t$ where $\muvar(t) = {\Varec{x}}$.
@@
 Thus, $\Val = \nf{(\Val)}$, and statements 2 and 3 hold for $\nf{(\Val)}$.
-%For each $i = 1, \ldots, n$,
-%if ${\Varec{z_i}} = {\Varec{w_i}}$, then $w_i \teq z_i \in \Ec$.
-%If ${\Varec{z_i}} \neq {\Varec{w_i}}$, then by statement 2, $\Val \ec{x} \interp{ t_i } = \Val \ec{z_i}$
-%and $\Val \ec{x} \interp{ u_i } = \Val \ec{w_i}$.
-%By statement 3, we have that $\ec{z_i} = \ec{w_i}$, and thus $u_i \teq t_i \in \Ec$ for each $i = 1, \ldots, n$.
-%This is a contradiction since $\ec{x}$ and $\ec{y}$ are distinct, and \rn{Cong} does not apply to $\Ec$.
-%Thus, statements 2 and 3 hold for $\nf{(\Val)}$.
-
-We now show that if 2 and 3 hold for some $\Val_1$,
+
+We now show that if statements 2 and 3 hold for some $\Val_1$,
 then 2 and 3 also hold for $\nf{(\Val_1 \sigma)}$,
-where $\sigma$ is a substitution of the form $\{ {\Varec{x}} \mapsto \MU {\Varec{x}}.\; t \}$,
-where $\MU {\Varec{x}}.\; t$ is not $\alpha$-equivalent to a term in $\mathcal{T}^\M_{\Varec{x}}(\Val)$.
-
-%\rem{TODO}
-%Thus, the set of subterms which $\Val \ec{x}$ and $\MU v_{\ec{x}}.\; t$ differ 
-%of the form $\MU {\Varec{z}}.\; u$ and $v_{\ec{z}}$ respectively, where $v_{\ec{z}}$ is a bound variable in $\Val \ec{y}$.
-%In the latter case, $\Val \ec{y} \interp{ {\Varec{z}} } = \MU {\Varec{z}}.\; \const{C}( \Val \ec{y} \interp{\vec u} )$,
-%where $\MU v_{\ec{z}}.\; u$ and $\MU v_{\ec{z}}.\; u'$ were constructed by applying a sequence of substitutions of the form above to $v_{\ec{z}}$.
-%Since $\Val^\ast$ is obtained by applying the same substitutions and normalizations to all terms in the range of $\Val$,
-%we have $\Val^\ast \ec{y} \interp{ \MU {\Varec{x}}.\; t} = \Val^\ast \ec{x}$.
-
-%To show 3,
-%first note that since \rn{Unique} does not apply,
-%$\Val \ec{x} \aequiv \Val \ec{y}$ if and only if $\ec{x} = \ec{y}$.
-%Due to part 2, we have $\Val = \nf{(\Val)}$, and thus
-%$\nf{(\Val \ec{x})} \aequiv \nf{(\Val \ec{y})}$ if and only if $\ec{x} = \ec{y}$.
-%We show that for any substitution $\sigma$ applied to $\Val$ while constructing its completion is such that
-%$\nf{(\Val \ec{x} \sigma)} \aequiv \nf{(\Val \ec{y} \sigma)}$ if and only if $\ec{x} = \ec{y}$.
-%\rem{TODO}
-%Let $P$ be the set of positions for which $\Val \ec{x}$ and $\Val \ec{y}$ are not $\alpha$-equivalent,
-%which is non-empty by assumption.
-
+where $\sigma$ is a substitution of the form $\{ {\Varec{x}} \mapsto \mu {\Varec{x}}. t \}$
+and $\mu {\Varec{x}}. t$ is not $\alpha$-equivalent to a term in $\mathcal{T}^\M_{\Varec{x}}(\Val_1)$.
+To show 2,
+by assumption of 2 on $\Val_1$, we have $\Val_1 \ec{y} \interp{ t } = \Val_1 \ec{x}$ for
+all subterms $t$ of $\Val_1 \ec{y}$ where $\muvar(t) = {\Varec{x}}$.
+Thus, $\Val_1 \sigma \ec{y} \interp{ t } = \Val_1 \sigma \ec{x}$ and
+$\nf{(\Val_1 \sigma)} \ec{y} \interp{ t } = \nf{(\Val_1 \sigma)} \ec{x}$.
+To show 3, 
+consider two distinct equivalence classes $\ec{x}$ and $\ec{y}$,
+and assume by contradiction that $\Val_1 \sigma \ec{x} \aequiv \Val_1 \sigma \ec{y}$.
+Due to statement 3 for $\Val_1$,
+$\ec{y}$ and $\ec{z}$ must have (minimal) subterms 
+where $t_1$ occurs in $\Val_1 \ec{y}$ the same position as $t_2$ occurs in $\Val_1 \ec{z}$,
+and $t_1 \not\aequiv t_2$.
+If $t_1$ and $t_2$ are distinct free variables,
+then $t_1 \sigma \not \aequiv t_2 \sigma$.
+If $t_1$ is of the form $\mu {\Varec{w_1}}. \const{C}( \vec t )$ 
+and $t_2$ is of the form $\mu {\Varec{w_2}}. \const{D}( \vec u )$,
+then $t_1 \sigma \not \aequiv t_2 \sigma$.
+Since $t_1$ and $t_2$ are minimal,
+$t_1$ must be of the form $\mu {\Varec{z}}. \const{C}( \vec t )$,
+and $t_2$ must be a free variable.
+If $t_2$ is not ${\Varec{x}}$,
+then $t_1 \sigma \not \aequiv t_2 \sigma$.
+Thus, $t_2$ is ${\Varec{x}}$.
+\rem{Must show positions do not collapse.}
 \qed
@@
 and thus $t_u\interp{u_i'} \vsim t_w\interp{ w_i }$. % for each $i = 1, \ldots, n$.
-By Lemma~\ref{lem:mu-cong}, we have $t_u \vsim t_w$.
-Since $\M( t ) \vsim t_u \vsim t_w = \Val^\ast \ec{t}$,
+%By Lemma~\ref{lem:mu-cong}, we have $t_u \vsim t_w$.
+Thus, since $\M( t ) \vsim t_u \vsim t_w = \Val^\ast \ec{t}$,
 we have $\M( t ) \vsim \Val^\ast \ec{t}$.

### 2014-12-20 tex/codatadp.tex
more

--- 2014-12-19 tex/codatadp.tex
@@
 The acyclicity and uniqueness rules are needed to ensure completeness on
-ground problems and the absence of spurious models. They cannot be replaced by
+ground problems. %and the absence of spurious models.
+They cannot be replaced by
 finite axiomatizations, so they naturally belong in a decision procedure.
@@
 \)
-%\rn{Unfold}}
+%\rn{Unfold}
+}
 \]
@@
 \)
-%\rn{Assign}}
+%\rn{Assign}
+}
 \]
@@
 No (co)datatypes
-  & 221 & 760 & \phantom{0}57 & 1038 \\
+  & 221 & 760 & \phantom{0}57 & 1038 \\[\jot]
+Datatype simplication only
+  &  \\
 Datatypes without acyclicity
-  & 227 & 765 & \phantom{0}57 & 1049 \\
+  & \bfseries 227 & 765 & \phantom{0}57 & 1049 \\
 Datatypes with acyclicity
-  & 227 & 771 & \phantom{0}57 & 1055 \\
+  & \bfseries 227 & 771 & \phantom{0}57 & 1055 \\[\jot]
+Codatatype simplication only
+  &  \\
 Codatatypes without uniqueness
@@
 Codatatypes with uniqueness
-  & 218 & 787 & \phantom{0}58 & 1063 \\
+  & 218 & 787 & \bfseries \phantom{0}58 & 1063 \\[\jot]
 Full (co)datatypes
-  & 224 & 798 & \phantom{0}58 & 1080 \\[\jot]
+  & 224 & \bfseries 798 & \bfseries \phantom{0}58 & \bfseries 1080 \\[\jot]
 Total number of goals
@@
 We introduced a decision procedure for the ground theory of datatypes and
-codatatypes and implemented it in the SMT solver CVC4. Our main theoretical
-contribution has been the support for codatatypes. These are not very
-difficult per se, but there are tricky corner cases to take care of if we care
-about completeness (which we do).
+codatatypes and implemented it in the SMT solver CVC4. Our main
+contribution has been the support for codatatypes. Both the metatheory and
+the implementation rely heavily on $\mu$-terms to represent cyclic values,
+a complication specific to codatatypes.
+% are not very
+% difficult per se, but there are tricky corner cases to take care of if we care
+% about completeness (which we do).
 %% alliteration
-On the practical side, we obtained respectable results on benchmarks generated
-from Isabelle theories.
+On the practical side, we obtained respectable results on Isabelle benchmarks.
 
@@
 try out CVC4 for higher-order model finding in Isabelle, as an alternative to
-the counterexample generator Nitpick \cite{blanchette-nipkow-2010}. We also
+the counterexample generator Nitpick \cite{blanchette-nipkow-2010}. We
 see some opportunities to enrich SMT solvers with recursive and corecursive
@@
 Andrei Popescu helped clarify our thoughts regarding the axiomatization of
-codatatypes. Dmitriy Traytel took part in discussions about special
-cases.
+codatatypes. Dmitriy Traytel took part in discussions about degenerate
+codatatypes.
 %

### 2014-12-20 tex/codatadp.tex

--- 2014-12-20 tex/codatadp.tex
@@
 \)
-%\rn{Unfold}}
+%\rn{Unfold}
 }
@@
 \)
-%\rn{Assign}}
+%\rn{Assign}
 }

### 2014-12-20 tex/codatadp.tex
Finish first draft of lemma.

@@
 \begin{proof}
-%Since $u_k$ is normal, $w_k$ is the result of replacing all subterms $v$ of $u_k$ such that $t\interp{v} \aequiv t$ with $x$.
+\rem{TODO}
+\end{proof}
+
+\begin{lemma}
+\label{lem:mu-norm-interp}
+If $t$ is normal and has subterm $u$, then $t \interp{u}$ is normal.
+\end{lemma}
+\begin{proof}
 \rem{TODO}
@@
 there is only one equivalence class of type $\tau$ in $\Ec$,
-and thus there are no terms in $\mathcal{T}^\M_x(\Val^\ast)$ of type $\tau$.
-This is a contradiction, since our model completion may assign the value in the domain of $\tau$ to ${\Varec{x}}$.
+and thus there are no terms in $\mathcal{T}^\M_{\Varec{y}}(\Val^\ast)$ of type $\tau$.
+This is a contradiction, since our model completion may assign the value in the domain of $\tau$ to ${\Varec{y}}$.
 Now, consider the case when $\tau$ is infinite.
 This is also a contradiction,
-since there are only a finite number of closed terms in $\mathcal{T}^\M_x(\Val^\ast)$,
-and thus our model completion can assign a value not occurring in $\mathcal{T}^\M_x(\Val^\ast)$ to ${\Varec{y}}$.
+since there are only a finite number of closed terms in $\mathcal{T}^\M_{\Varec{y}}(\Val^\ast)$,
+and thus our model completion can assign a value not occurring in $\mathcal{T}^\M_{\Varec{y}}(\Val^\ast)$ to ${\Varec{y}}$.
 By construction, $\Val^\ast \ec{x}$ is normal.
@@
 Moreover, our construction of $\Val^\ast$ applies substitutions of the form
-$\{ y \mapsto t \}$, where $t$ is acyclic when the type of $y$ is not a codatatype.
+$\{ {\Varec{y}} \mapsto t \}$, where $t$ is acyclic when the type of ${\Varec{y}}$ is not a codatatype.
 Thus, $\Val^\ast \ec{x}$ is acyclic when $\tau \not\in \Codata$.
@@
 constructed by applying a sequence of substitutions of the form mentioned above to ${\Varec{x}}$.
-Moreover, free variables ${\Varec{z}}$ in $\MU {\Varec{x}}.\; t$ that are bound in $\Val \ec{y}$ are interpreted in 
-the expansion of $\Val \ec{y} \interp{ \MU {\Varec{x}}.\; t }$ as a term
+Moreover, free variables ${\Varec{z}}$ in $t$ that are bound in $\Val \ec{y}$ are interpreted in 
+the expansion of $\Val \ec{y} \interp{ t }$ as a term
 constructed by a sequence of substitutions of the form mentioned above to ${\Varec{z}}$. \rem{more detail}
 Thus, we have
-$\Val \ec{y} \interp{ \MU {\Varec{x}}.\; t } = \Val \ec{x} \interp{ \Val \ec{x} } = \Val \ec{x}$, and thus 2 holds for $\Val$.
+$\Val \ec{y} \interp{ t } = \Val \ec{x} \interp{ \Val \ec{x} } = \Val \ec{x}$, and thus 2 holds for $\Val$.
 Statement 3 holds for $\Val$ since \rn{Unique} does not apply.
@@
 Since $\Val \ec{x}$ is a minimal, it
-contain a subterm of the form $\MU {\Varec{y}}.\; \const{C}( u_1, \ldots, u_n )$
+contains a subterm of the form $\MU {\Varec{y}}.\; \const{C}( u_1, \ldots, u_n )$
 where $\Val \ec{x} \interp{ t_i } \aequiv \Val \ec{x} \interp{ u_i }$ for $i = 1, \ldots, n$.
@@
 then 2 and 3 also hold for $\nf{(\Val_1 \sigma)}$,
-where $\sigma$ is a substitution of the form $\{ {\Varec{x}} \mapsto \mu {\Varec{x}}. t \}$
+where $\sigma$ is a substitution of the form $\{ {\Varec{x}} \mapsto \mu {\Varec{x}}. t \}$,
+${\Varec{x}} \in FV( \Val_1 )$,
 and $\mu {\Varec{x}}. t$ is not $\alpha$-equivalent to a term in $\mathcal{T}^\M_{\Varec{x}}(\Val_1)$.
 To show 2,
-by assumption of 2 on $\Val_1$, we have $\Val_1 \ec{y} \interp{ t } = \Val_1 \ec{x}$ for
-all subterms $t$ of $\Val_1 \ec{y}$ where $\muvar(t) = {\Varec{x}}$.
-Thus, $\Val_1 \sigma \ec{y} \interp{ t } = \Val_1 \sigma \ec{x}$ and
-$\nf{(\Val_1 \sigma)} \ec{y} \interp{ t } = \nf{(\Val_1 \sigma)} \ec{x}$.
+by assumption of 2 on $\Val_1$, we have $\Val_1 \ec{y} \interp{ u } = \Val_1 \ec{x}$ for
+all subterms $u$ of $\Val_1 \ec{y}$ where $\muvar(t) = {\Varec{x}}$.
+Thus, $\Val_1 \sigma \ec{y} \interp{ u } = \Val_1 \sigma \ec{x}$ and
+$\nf{(\Val_1 \sigma)} \ec{y} \interp{ u } = \nf{(\Val_1 \sigma)} \ec{x}$.
 To show 3, 
-consider two distinct equivalence classes $\ec{x}$ and $\ec{y}$,
-and assume by contradiction that $\Val_1 \sigma \ec{x} \aequiv \Val_1 \sigma \ec{y}$.
+consider two distinct equivalence classes $\ec{y}$ and $\ec{z}$,
+and assume by contradiction that $\nf{(\Val_1 \sigma)} \ec{y} \aequiv \nf{(\Val_1 \sigma)} \ec{z}$.
 Due to statement 3 for $\Val_1$,
 $\ec{y}$ and $\ec{z}$ must have (minimal) subterms 
-where $t_1$ occurs in $\Val_1 \ec{y}$ the same position as $t_2$ occurs in $\Val_1 \ec{z}$,
+where $t_1$ occurs in $\Val_1 \ec{y}$ the same position (call it $p$) as $t_2$ occurs in $\Val_1 \ec{z}$,
 and $t_1 \not\aequiv t_2$.
-If $t_1$ and $t_2$ are distinct free variables,
-then $t_1 \sigma \not \aequiv t_2 \sigma$.
-If $t_1$ is of the form $\mu {\Varec{w_1}}. \const{C}( \vec t )$ 
-and $t_2$ is of the form $\mu {\Varec{w_2}}. \const{D}( \vec u )$,
-then $t_1 \sigma \not \aequiv t_2 \sigma$.
+%We first show that $\nf{(\Val_1 \sigma)} \ec{y}$ (and for the same reasons, $\nf{(\Val_1 \sigma)} \ec{z}$)
+%also contains position $p$.
+%If this were not the case,
+%then without loss of generality we may assume
+%$\Val_1 \sigma \ec{y}$ contains a strict bisimilar subterm 
+%$\mu {\Varec{w}}. u$ that strictly contains $t_1 \sigma$.
+If $t_1$ (resp. $t_2$) is a free variable that is not ${\Varec{x}}$,
+then $\nf{(\Val_1 \sigma)} \ec{y}$ (resp. $\nf{(\Val_1 \sigma)} \ec{z}$)
+contains $t_1$ (resp. $t_2$) at position $p$,
+and $\nf{(\Val_1 \sigma)} \ec{z}$ (resp. $\nf{(\Val_1 \sigma)} \ec{y}$) does not.
+If $t_1$ is of the form $\mu w_1. \const{C}( \vec t )$ 
+and $t_2$ is of the form $\mu w_2. \const{D}( \vec u )$,
+then the expansion \rem{define} of $\nf{(\Val_1 \sigma)} \ec{y}$ and $\nf{(\Val_1 \sigma)} \ec{z}$
+are $\alpha$-disequivalent at position $p$.
 Since $t_1$ and $t_2$ are minimal,
-$t_1$ must be of the form $\mu {\Varec{z}}. \const{C}( \vec t )$,
-and $t_2$ must be a free variable.
-If $t_2$ is not ${\Varec{x}}$,
-then $t_1 \sigma \not \aequiv t_2 \sigma$.
-Thus, $t_2$ is ${\Varec{x}}$.
-\rem{Must show positions do not collapse.}
+say $t_1$ is of the form $\mu w. \const{C}( \vec t )$,
+and $t_2$ is ${\Varec{x}}$.
+Since $\sigma$ maps ${\Varec{x}}$ to a closed $\mu$-term $\mu {\Varec{x}}. t$,
+we have that $FV( t_1 ) \subseteq \{ {\Varec{x}} \}$,
+or else the expansion of $\nf{(\Val_1 \sigma)} \ec{y}$ and $\nf{(\Val_1 \sigma)} \ec{z}$
+are $\alpha$-disequivalent at position $p$ since they do not contain the same free variables.
+Since $\Val_1 \ec{y} \interp{t_1}$ is normal,
+there is a closed $\mu$-term $v \in \mathcal{T}^\M_{\Varec{x}}( \Val_1 )$
+such that $v \vsimv{\Varec{x}} \Val_1 \ec{y} \interp{t_1}$.
+Thus, by assumption on our selection of $\mu {\Varec{x}}. t$,
+we have $\mu {\Varec{x}}. t \not\vsimv{\Varec{x}} \Val_1 \ec{y} \interp{t_1}$,
+which implies that the expansion of $\nf{(\Val_1 \sigma)} \ec{y}$ and $\nf{(\Val_1 \sigma)} \ec{z}$
+are $\alpha$-disequivalent at position $p$.
+
+Thus, by induction on the number of applications of the above rule used to obtain $\Val^\ast$,
+we have that $\Val^\ast$ satisfies statements 2 and 3.
 \qed

### 2014-12-20 tex/codatadp.tex
Minor cleanup

@@
 \item[\rm 2.]
-All subterms $t$ of $\Val^\ast \ec{y}$ with $\muvar(t) = {\Varec{x}}$ are such that
-$\Val^\ast \ec{y} \interp{ t } = \Val^\ast \ec{x}$. 
-%If $\const{C}( t_1, \ldots, t_n ) \in \ec{x}$,
-%then all subterms $\MU v_{\ec{x}}.\; u$ occurring in $\Val^\ast \ec{y}$ 
-%are of the form $\MU v_{\ec{x}}.\; \const{C}( w_1, \ldots, w_n )$,
-%where 
-%and $\Val^\ast \ec{y} \interp{ w_i } = \Val^\ast \ec{t_i}$ for $i = 1, \ldots, n$.
-%If $\Val^\ast \ec{x}$ contains a subterm of the form $\MU v_{\ec{x}}.\; \const{C}( w_1, \ldots, w_n )$,
-%and $\muvar( w_i ) = v_{\ec{t_i}}$ for $i = 1, \ldots, n$,
-%\item
-%$t_1 \interp{ t_2 } = \Val^\ast \ec{x}$
-%for all $t_1$ with subterm $t_2$ in the range of $\Val^\ast$ where $\muvar(t_2)={\Varec{x}}$,
+$\Val^\ast \ec{x} = \Val^\ast \ec{y} \interp{ t }$ for
+all subterms $t$ of $\Val^\ast \ec{y}$ with $\muvar(t) = {\Varec{x}}$. 
 \item[\rm 3.]
@@
 which implies that the expansion of $\nf{(\Val_1 \sigma)} \ec{y}$ and $\nf{(\Val_1 \sigma)} \ec{z}$
-are $\alpha$-disequivalent at position $p$.
+are $\alpha$-disequivalent at position $p$. \rem{more detail}
 
@@
 By Lemma~\ref{lem:model-completion}.2 and due to our construction of $\Val^\ast$, we have $\Val^\ast \ec{t}$ is a term
-$t_w$ of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$
-where $t_w\interp{w_i} = \Val^\ast \ec{u_i}$ for $i = 1, \ldots, n$.
+of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$
+where $\Val^\ast \ec{t}\interp{w_i} = \Val^\ast \ec{u_i}$ for $i = 1, \ldots, n$.
 For each $i = 1, \ldots, n$, let $u_i'$ be the $i$\vvthinspace th argument of $t_u$.
 By Lemma~\ref{lem:mu-norm-arg}, $t_u\interp{u_i'} \vsim \Val^\ast \ec{u_i}$,
-and thus $t_u\interp{u_i'} \vsim t_w\interp{ w_i }$. % for each $i = 1, \ldots, n$.
+and thus $t_u\interp{u_i'} \vsim \Val^\ast \ec{t}\interp{ w_i }$. % for each $i = 1, \ldots, n$.
 %By Lemma~\ref{lem:mu-cong}, we have $t_u \vsim t_w$.
-Thus, since $\M( t ) \vsim t_u \vsim t_w = \Val^\ast \ec{t}$,
+Thus, $\M( t ) \vsim t_u \vsim \Val^\ast \ec{t}$, and
 we have $\M( t ) \vsim \Val^\ast \ec{t}$.
@@
 $\ec{u}$ must contain a term of the form $\const C_j( \const s^1_j( u ), \ldots, \const s^{n}_j( u ) )$ for some $j$.
-By Lemma~\ref{lem:model-completion}.2 and due to our construction of $\Val^\ast$, we have $\Val^\ast \ec{u}$ is a term $t_w$ of the form
-$\MU {\Varec{u}}.\; \const C_j( w_1, \ldots, w_n )$,
-where $t_w \interp{ w_j } = \Val^\ast \ec{ \const s^j_k( u )}$. 
-If $k = i$, then $\M( t )$ is $\alpha$-equivalent with $t_w\interp{w_j} = \Val^\ast \ec{ \const s^j_k( u )}$,
+By Lemma~\ref{lem:model-completion}.2 and due to our construction of $\Val^\ast$, we have $\Val^\ast \ec{u}$ is of the form
+$\MU {\Varec{u}}.\; \const C_l( w_1, \ldots, w_n )$,
+where $\Val^\ast \ec{u} \interp{ w_j } = \Val^\ast \ec{ \const s^j_k( u )}$. 
+If $k = l$, then $\M( t )$ is $\alpha$-equivalent with $\Val^\ast \ec{u} \interp{w_j} = \Val^\ast \ec{ \const s^j_k( u )}$,
 which is $\Val^\ast \ec{t}$.
-If $k \neq i$, we interpret $\const s^j_k( \M( u ) )$ as the value in $\M( \tau )$ that is $\alpha$-equivalent with $\Val^\ast \ec{ t }$.
+If $k \neq l$, we interpret $\const s^j_k( \M( u ) )$ as the value in $\M( \tau )$ that is $\alpha$-equivalent with $\Val^\ast \ec{ t }$.
 Since \rn{Cong} does not apply and due to Lemma~\ref{lem:model-completion}.3,

### 2014-12-21 tex/codatadp.tex
Minor fixes

@@
 equivalent. Both the finiteness assumption on the datatype in \rn{Split} and
-the cardinality-1 constraint on the codatatype in \rn{Singleton} are easy to
-check based on a recursive computation of the cardinalities of the
+the cardinality-1 constraint on the codatatype in \rn{Singleton} are 
+are computed statically
+based on a recursive computation of the cardinalities of the
 constructors' argument types.
@@
 
-If $t$ is a selector term of the form $\const s^j_k( u )$,
+If $t$ is a selector term of the form $\const s^k_j( u )$,
 since \rn{Split} does not apply,
-$\ec{u}$ must contain a term of the form $\const C_j( \const s^1_j( u ), \ldots, \const s^{n}_j( u ) )$ for some $j$.
+$\ec{u}$ must contain a term of the form $\const C_{j'}( \const s^1_{j'}( u ), \ldots, \const s^{n}_{j'}( u ) )$ for some $j'$.
 By Lemma~\ref{lem:model-completion}.2 and due to our construction of $\Val^\ast$, we have $\Val^\ast \ec{u}$ is of the form
-$\MU {\Varec{u}}.\; \const C_l( w_1, \ldots, w_n )$,
-where $\Val^\ast \ec{u} \interp{ w_j } = \Val^\ast \ec{ \const s^j_k( u )}$. 
-If $k = l$, then $\M( t )$ is $\alpha$-equivalent with $\Val^\ast \ec{u} \interp{w_j} = \Val^\ast \ec{ \const s^j_k( u )}$,
+$\MU {\Varec{u}}.\; \const C_{j'}( w_1, \ldots, w_n )$,
+where $\Val^\ast \ec{u} \interp{ w_i } = \Val^\ast \ec{ \const s^i_{j'}( u )}$ for $i = 1, \ldots, n$. 
+If $j = j'$, then $\M( t )$ is $\alpha$-equivalent with $\Val^\ast \ec{u} \interp{w_k}$, which is equal to $\Val^\ast \ec{ \const s^k_j( u )}$,
 which is $\Val^\ast \ec{t}$.
-If $k \neq l$, we interpret $\const s^j_k( \M( u ) )$ as the value in $\M( \tau )$ that is $\alpha$-equivalent with $\Val^\ast \ec{ t }$.
+If $j \neq j'$, we are free to interpret $\const s^k_j( \M( u ) )$ as the value in $\M( \tau )$ that is $\alpha$-equivalent with $\Val^\ast \ec{ t }$.
 Since \rn{Cong} does not apply and due to Lemma~\ref{lem:model-completion}.3,

### 2014-12-21 tex/codatadp.tex
Split report/paper versions.  Minor.

@@
 % for the renaming $\{ x \mapsto y, y \mapsto x \}$.
-
 For any $\mu$-term $t$ of the form $\MU x.\; \const C( \vec u )$, we can
@@
 
+\begin{report}
 \begin{lemma}
@@
 %\end{proof}
+\end{report}
 
@@
 that is, an element of $\M( \tau )$ where $\tau$ is the type of $t$.
-\rem{Clarify that non-codatatype terms are interpreted as $\mu$-terms, although dropping the $\mu$'s gives the standard representation?
-(GOOD IDEA. A BIT OF REDUNDANCY CANNOT HURT.)}
+In the following,
+\emph{all} types are interpretated as a set of $\mu$-terms.
+This is done to simplify the presentation.
+Since all values of types in $\Data$ and $\Nondata$ are acyclic,
+a typical representation for these values can be obtained by
+dropping each of its $\mu$-bindings. 
+For instance, the $\mu$-term $\mu x. \const{0}$ representing the integer constant zero
+is simply $\const{0}$ after dropping the $\mu$-binding of $x$.
 
@@
 in $\M( \tau )$ that is $\alpha$-equivalent to
-$\M( t )\interp{u_j}$
+$\M( t )\interp{u_k}$
 when $\M( t )$ is of the form $\MU x.\; \const{C}_j( \vec u )$.
@@
 where we write $\muvar( t )$ to denote $t$ if $t$ is a variable, or ${\Varec{x}}$ if $t$ is of the form $\MU {\Varec{x}}.\; u$.
+\begin{paper} 
+The proof of this lemma can be found in the extended version of this paper.
+\end{paper}
 
@@
 \end{lemma}
+\begin{report}
 \begin{proof}

### 2014-12-21 tex/codatadp.tex
Minor clarification

@@
 
-First, we need additional terminology concerning $\mu$-terms.
+First, we need to define the set of interpretations of the theory $\thD$,
+which requires additional terminology concerning $\mu$-terms.
 For a $\mu$-term $t$ with subterm $u$,
@@
 
-We now define a class of models that we consider for a saturated set of equalities~$\Ec$.
-Let $\M( \tau )$ denote the interpretation type $\tau$ in $\M$,
+We now define the class of interpretations for $\thD$.
+In the following, $\M( \tau )$ denotes the interpretation type $\tau$ in $\M$,
 that is, a set of domain elements for that type.
-For a term $t$, let $\M( t )$ denote the interpretation of $t$ in $\M$,
+For a term $t$, $\M( t )$ denotes the interpretation of $t$ in $\M$,
 that is, an element of $\M( \tau )$ where $\tau$ is the type of $t$.
@@
 which by the induction hypothesis is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \Val^\ast \ec{ u_1 }, \ldots, \Val^\ast \ec{ u_n } )}$, call this term $t_u$.
-By Lemma~\ref{lem:model-completion}.2 and due to our construction of $\Val^\ast$, we have $\Val^\ast \ec{t}$ is a term
+%By Lemma~\ref{lem:model-completion}.2 and due to our construction of $\Val^\ast$, we have $\Val^\ast \ec{t}$ is a term
+%of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$
+%where $\Val^\ast \ec{t}\interp{w_i} = \Val^\ast \ec{u_i}$ for $i = 1, \ldots, n$.
+Since \rn{Inject} and \rn{Clash} do not apply to $\Fc$, 
+by our construction of $\Val^\ast$ we have $\Val^\ast \ec{t}$ is a term
 of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$ 
-where $\Val^\ast \ec{t}\interp{w_i} = \Val^\ast \ec{u_i}$ for $i = 1, \ldots, n$.
+where for each $i = 1, \ldots, n$, 
+$\muvar(w_i) = {\Varec{u_i}}$ and thus by Lemma~\ref{lem:model-completion}.2, 
+$\Val^\ast \ec{t}\interp{w_i} = \Val^\ast \ec{u_i}$.
 For each $i = 1, \ldots, n$, let $u_i'$ be the $i$\vvthinspace th argument of $t_u$.
-By Lemma~\ref{lem:mu-norm-arg}, $t_u\interp{u_i'} \vsim \Val^\ast \ec{u_i}$,
-and thus $t_u\interp{u_i'} \vsim \Val^\ast \ec{t}\interp{ w_i }$. % for each $i = 1, \ldots, n$.
+By Lemma~\ref{lem:mu-norm-arg}, $t_u\interp{u_i'} \vsim \Val^\ast \ec{u_i}$.
+Thus $t_u\interp{u_i'} \vsim \Val^\ast \ec{t}\interp{ w_i }$. % for each $i = 1, \ldots, n$.
 %By Lemma~\ref{lem:mu-cong}, we have $t_u \vsim t_w$.
@@
 If $t$ is a selector term of the form $\const s^k_j( u )$,
-since \rn{Split} does not apply,
+since \rn{Split} does not apply to $\Fc$,
 $\ec{u}$ must contain a term of the form $\const C_{j'}( \const s^1_{j'}( u ), \ldots, \const s^{n}_{j'}( u ) )$ for some $j'$.
-By Lemma~\ref{lem:model-completion}.2 and due to our construction of $\Val^\ast$, we have $\Val^\ast \ec{u}$ is of the form
+%By Lemma~\ref{lem:model-completion}.2 and due to our construction of $\Val^\ast$, we have $\Val^\ast \ec{u}$ is of the form
+%$\MU {\Varec{u}}.\; \const C_{j'}( w_1, \ldots, w_n )$,
+%where $\Val^\ast \ec{u} \interp{ w_i } = \Val^\ast \ec{ \const s^i_{j'}( u )}$ for $i = 1, \ldots, n$. 
+Since \rn{Inject} and \rn{Clash} do not apply to $\Fc$, 
+by our construction of $\Val^\ast$ we have $\Val^\ast \ec{u}$ is of the form
 $\MU {\Varec{u}}.\; \const C_{j'}( w_1, \ldots, w_n )$,
-where $\Val^\ast \ec{u} \interp{ w_i } = \Val^\ast \ec{ \const s^i_{j'}( u )}$ for $i = 1, \ldots, n$. 
+where for each $i = 1, \ldots, n$,
+$\muvar(w_i) = {\Varec{\const s^i_{j'}( u )}}$ and thus by Lemma~\ref{lem:model-completion}.2, 
+$\Val^\ast \ec{u} \interp{ w_i } = \Val^\ast \ec{ \const s^i_{j'}( u )}$. 
 If $j = j'$, then $\M( t )$ is $\alpha$-equivalent with $\Val^\ast \ec{u} \interp{w_k}$, which is equal to $\Val^\ast \ec{ \const s^k_j( u )}$,

### 2014-12-22 tex/codatadp.tex
tweak intro

--- 2014-12-21 tex/codatadp.tex
@@
 using Peano notation but extended with an
-infinity value $\infty = \const{S}(\const{S}(\const{S}(\ldots)))$. Compared
+infinity value $\infty = \const{S}(\const{S}(\const{S}(\ldots)))$.
+The equation $\const{S}(\infty) \teq \infty$ holds as expected,
+because both sides expand to the infinite term
+$\const{S}(\const{S}(\const{S}(\ldots)))$, which uniquely identifies the
+value~$\infty$.
+Compared
 with the more conventional, inductive definition with a dedicated constructor
@@
 the codatatype avoids one case by unifying the finite and infinite nonzero cases.
-Moreover, equations such as $\const{S}(\infty) \teq \infty$ hold by default,
-because both sides expand to the infinite term
-$\const{S}(\const{S}(\const{S}(\ldots)))$, which uniquely identifies the
-value~$\infty$.
 

### 2014-12-22 tex/codatadp.tex
more

@@
 No (co)datatypes
-  & 221 & 760 & \phantom{0}57 & 1038 \\[\jot]
-Datatype simplication only
-  &  \\
+  & 221 & 760 & \phantom{0}55 & 1036 \\[\jot]
+%Datatype simplication only
+%  &  \\
 Datatypes without acyclicity
-  & \bfseries 227 & 765 & \phantom{0}57 & 1049 \\
+  & \bfseries 227 & 765 & \phantom{0}55 & 1047 \\
 Datatypes with acyclicity
-  & \bfseries 227 & 771 & \phantom{0}57 & 1055 \\[\jot]
-Codatatype simplication only
-  &  \\
+  & \bfseries 227 & 771 & \phantom{0}55 & 1053 \\[\jot]
+%Codatatype simplication only
+%  &  \\
 Codatatypes without uniqueness
-  & 217 & 787 & \phantom{0}55 & 1059 \\
+  & 215 & 789 & \phantom{0}56 & 1060 \\
 Codatatypes with uniqueness
-  & 218 & 787 & \bfseries \phantom{0}58 & 1063 \\[\jot]
+  & 216 & 789 & \bfseries \phantom{0}59 & 1064 \\[\jot]
 Full (co)datatypes
-  & 224 & \bfseries 798 & \bfseries \phantom{0}58 & \bfseries 1080 \\[\jot]
+  & 222 & \bfseries 800 & \bfseries \phantom{0}59 & \bfseries 1081 \\[\jot]
 Total number of goals

### 2014-12-22 tex/codatadp.tex
edits

@@
 signature~$\Sig$. It is formulated as a calculus, whose rules are applied
-until a contradiction is detected or until saturation.
+until saturation or contradiction.
 %
@@
 denoting nondeterministic branching. The rules associated with a phase have
-priority over those of the previous phases. The rules are applied until
-saturation (i.e., until no rule can be applied to extend~$\Ec$
-further).
+priority over those of the previous phases. The rules are applied
+until a contradiction ($\bot$) is derived or until no rule can be applied to
+extend~$\Ec$ further.
 
@@
 
-We first state the following properties of $\Val^\ast$,
-where we write $\muvar( t )$ to denote $t$ if $t$ is a variable, or ${\Varec{x}}$ if $t$ is of the form $\MU {\Varec{x}}.\; u$.
-\begin{paper} 
-The proof of this lemma can be found in the extended version of this paper.
-\end{paper}
+%We first state the following properties of $\Val^\ast$,
+In the following lemma about $\Val^\ast$,
+$\muvar( t )$ denotes $t$ if $t$ is a variable and ${\Varec{x}}$ if $t$ is of the form $\MU {\Varec{x}}.\; u$.
+
+%%% @ANDY: I think it's enough to point this out clearly in the introduction.
+%%% At least, that always worked fine for me. :)
+%
+%\begin{paper} 
+%The proof of this lemma can be found in the extended version of this paper.
+%\end{paper}
 

### 2014-12-22 tex/codatadp.tex
more

@@
 %
-\begin{report}
+\begin{report}%
 Expressed in the algebraic jargon, exhaustiveness helps ensure that ``no
@@
 which in addition to $\Sig$ defines a class of $\Sig$-interpretations $\mathbf{I}$,
-namely the ones that satisfy the axioms mentioned in this section.
-In particular, this means that each interpretation in $\mathbf{I}$ has a fixed interpretation
+namely the ones that satisfy the %axioms --- existence of all infinite codatatype values cannot be axiomatized
+properties mentioned in this section.
+In particular, this means that each interpretation in~$\mathbf{I}$ has a fixed interpretation
 for constructor terms and correctly applied selector terms, but may differ on its interpretation for
@@
 \]
-\end{report}
+\end{report}%
 where \ty{unit} is a datatype with the single constructor $\const{Unity} :
@@
 
-\begin{figure}[t]
+\begin{figure}[bth]
 \normalsize
@@
 and failures to unify equated terms are recognized by \rn{Clash}.
-The rule \rn{Conflict} recognizes when an equality and its negation both occur in $\Ec$, in which case $\Ec$ has no model.
+The \rn{Conflict} rule recognizes when an equality and its negation both occur in $\Ec$, in which case $\Ec$ has no model.
 
@@
 
-\begin{figure}[t]
+\begin{figure}[b!]
 \normalsize
@@
 
-\begin{figure}[t]
+\begin{figure}[tbh]
 \normalsize

### 2014-12-22 tex/codatadp.tex
more

@@
 denoting nondeterministic branching. The rules associated with a phase have
-priority over those of the previous phases. The rules are applied
+priority over those of the subsequent phases. The rules are applied
 until a contradiction ($\bot$) is derived or until no rule can be applied to
@@
 \inferrule{
-  t \teq u,\; t \tneq u \in \Ec
-}{
-  \bot
-}
-\)
-\rn{Conflict}
-\\[5\jot]
-\(
-\inferrule{
   s \teq t,\; t \teq u \in \Ec
@@
 \rn{Trans}
-\qquad
+\\[5\jot]
 \(
@@
 \rn{Cong}
+\qquad
+\(
+\inferrule{
+  t \teq u,\; t \tneq u \in \Ec
+}{
+  \bot
+}
+\)
+\rn{Conflict}
 \\[5\jot]
@@
 %For presentation of the rules in this phase, we rely on a representation of terms in the $\mu$-notation.
-The rules of the second phase contain premises that involve a mapping $\Val$
+In the second phase, the premises of the rules refer to a mapping $\Val$
 that associates a $\mu$-term as representative with each equivalence class.
@@
 A $\mu$-term $t$ of type $\tau$ describes a class of values of
-$\tau$ that terms in that equivalence class can take in the models of $\Ec$;
+$\tau$ that terms in that equivalence class can take in the models of $\Ec$.
 %either in the case when $\tau$ is a codatatype or datatype type,
-when $\tau \in \Data$, a cyclic $t$ describes an infeasible class of values.
+When $\tau$ is a datatype, % \in \Data$,
+a cyclic $t$ describes an infeasible class of values.
 %$\tau \not\in \Codata$.
+
 Formally, $\mu$-terms are defined recursively as being either a variable $x$
@@
 The variable $x$ need not occur freely under the binder's body, in which case the
-binder can be omitted.
+binder will sometimes be omitted in examples.
 In the \rn{Acyclic} rule,

### 2014-12-22 tex/codatadp.tex
more

@@
 %For presentation of the rules in this phase, we rely on a representation of terms in the $\mu$-notation.
-In the second phase, the premises of the rules refer to a mapping $\Val$
+The premises of the rules that make up the second phase refer to a mapping $\Val$
 that associates a $\mu$-term as representative with each equivalence class.
@@
 
-Formally, $\mu$-terms are defined recursively as being either a variable $x$
+Formally, $\mu$-\emph{terms} are defined recursively as being either a variable $x$
 or a node of the form
@@
 %not occurring in $\tEc$,
-and let $\Val\ec{x} = {\Varec{x}}$.
+and set $\Val\ec{x} = {\Varec{x}}$.
 Thus, there are initially no constraints on the values for any equivalence class $\ec{x}$.
-The mapping $\Val$ is then modified by applying the following unfolding rule exhaustively:
+The mapping $\Val$ is refined by applying the following unfolding rule exhaustively:\strut
 \[
@@
 \]
-$\FV( \Val )$ denotes the union of the free variables occurring in $\Val$'s range.
+$\FV( \Val )$ denotes the union of the free variables occurring in $\Val$'s range,
+and $\Val\{{\Varec{x}} \mapsto \ldots\}$ denotes the corresponding substitution.
 It is easy to see that the height of terms produced as a result of the unfolding
@@
 \begin{examplex}
-Suppose that $\Ec$ contains distinct equivalence classes $\ec{w}$, $\ec{x}$, $\ec{y}$, and $\ec{z}$
-where $\const{C}(z,x) \in \ec{y}$ and $\const{C}(w,y) \in \ec{x}$ for some $\const{C} \in \Ctr$.
+Suppose that $\Ec$ contains four distinct equivalence classes $\ec{w}$, $\ec{x}$, $\ec{y}$, and~$\ec{z}$
+such that $\const{C}(w,y) \in \ec{x}$ and $\const{C}(z,x) \in \ec{y}$ for some $\const{C} \in \Ctr$.
 A possible sequence of unfoldings is given below, omitting

### 2014-12-22 tex/codatadp.tex
more

@@
 %
-This indicates that the values for $x$ and $y$ in models of $\Ec$
-are of the forms $\const{C}( {\Varec{w}}, \const{C}( {\Varec{z}}, \const{C}( {\Varec{w}},\allowbreak \const{C}( {\Varec{z}},\ldots ))))$
+The result indicates that the values for $x$ and $y$ in models of $\Ec$
+are of the forms $\const{C}( {\Varec{w}},\allowbreak \const{C}( {\Varec{z}},\allowbreak \const{C}( {\Varec{w}},\allowbreak \const{C}( {\Varec{z}},\ldots ))))$
 and $\const{C}( {\Varec{z}}, \const{C}( {\Varec{w}}, \const{C}( {\Varec{z}}, \const{C}( {\Varec{w}}, \ldots ))))$,
@@
 
-Given $\Val$, the rules for acyclicity and uniqueness can be stated as follows.
-For acyclicity, if there is an equivalence class $\ec{t}$ of datatype type whose class of values $\Val \ec{t} = \MU x.\; u$ are cyclic,
+Given the mapping $\Val$, the \rn{Acyclic} and \rn{Unique} rules can be stated as follows.
+For acyclicity, if there exists an equivalence class $\ec{t}$
+where $t$'s type is a datatype and
+whose %class of  -- "whose" already refers to "equivalence class"
+values $\Val \ec{t} = \MU x.\; u$ are cyclic,
 as indicated by the assumption $x \in \FV( u )$,
-then we can conclude that $\Ec$ is unsatisfiable.
-For uniqueness, if there are two equivalence classes $\ec{t}$ and $\ec{u}$
-of codatatype type whose classes of values $\Val \ec{t}$ and $\Val \ec{u}$ are equivalent for a capture-avoiding renaming of $\mu$-bound variables,
-then we can conclude $t \teq u$.
-We demonstrate the uniqueness rule concretely with the following example.
+then $\Ec$ is unsatisfiable.
+For uniqueness, if there exist two equivalence classes $\ec{t}$ and $\ec{u}$
+where $t$ and $u$'s type is a codatatype and
+whose %classes of  -- "whose" already refers to "equivalence classes"
+values $\Val \ec{t}$ and $\Val \ec{u}$ are $\alpha$-equivalent,
+%for a capture-avoiding renaming of $\mu$-bound variables,
+then $t \teq u$. This rule may seem too restrictive, since
+$\MU x.\; \const{S}(x)$ and $\MU y.\; \const{S}(\const{S}(y))$ denote the same
+value despite being $\alpha$-disequivalent, but the construction also works in
+such cases, as illustrated by the following example.
 
@@
 and let $\Ec = \{ x \teq \const{C}(y,\const{C}(y,x)) \}$.
-After phase~1, the equivalence classes of $\Ec$ are
-$\{ x,\, \const{C}(y,\const{C}(y,x)) \}$, $\{ \const{C}(y,x) \}$, and $\{ y \}$.
-Constructing $\Val$ as described above yields
+After phase~1, the equivalence classes %of $\Ec$
+are
+$\{ y \}$, $\{ \const{C}(y,x) \}$, and $\{ x,\, \const{C}(y,\const{C}(y,x)) \}$.
+Constructing $\Val$ yields
 \[\begin{array}{rcrlrl}
@@
 \end{array}\]
-Since $\Val \ec{x}$ and $\Val \ec{\const{C}(y,x)}$ are $\alpha$-equivalent,
+Since $\Val \ec{x} \aequiv \Val \ec{\const{C}(y,x)}$,
 %for the renaming
 %$\{ {\Varec{x}} \mapsto {\VAREC{\const{C}(y,x)}}, {\VAREC{\const{C}(y,x)}} \mapsto {\Varec{x}} \}$,
-the calculus can infer $x \teq \const{C}(y,x)$.
+the \rn{Unique} rule applies to derive $x \teq \const{C}(y,x)$.
 %Intuitively, this equality holds, since the values of $x$ and $\const{C}( y, x )$

### 2014-12-22 tex/codatadp.tex
minor

@@
 \inferrule{
+  t : \delta
+  \quad
   \delta \in \Data
-  \quad
-  t : \delta
   \quad
@@
 \inferrule{
+ t, u : \delta
+ \quad
  \delta \in \Codata
- \quad
- t, u : \delta
  \quad
@@
   \quad
-  ( \const s( t ) \in \tEc \text{ and } \const s \in \Sel^\delta )
+  \bigl( \const s( t ) \in \tEc \text{ and } \const s \in \Sel^\delta \bigr)
   \text{ or }
-  ( \delta \in \Data \text{ and } \delta \text{ finite} )
+  \bigl( \delta \in \Data \text{ and } \delta \text{ finite} \bigr)
 }{
@@
 \inferrule{
+  t, u : \delta \in \tEc
+  \quad
   \delta \in \Codata
@@
   \left|\delta\right| = 1
-  \quad
-  t, u : \delta \in \tEc
 }{
@@
 A companion rule, \rn{Singleton}, handles the degenerate case where two terms
-from $\tEc$ have a corecursive singleton type, in which case they must be
+from $\tEc$ have a corecursive singleton type
+(Section~\ref{sec:the-theory-of-co-datatypes}), in which case they must be
 equivalent. Both the finiteness assumption on the datatype in \rn{Split} and

### 2014-12-22 tex/codatadp.tex
more

@@
   \text{ or }
-  \bigl( \delta \in \Data \text{ and } \delta \text{ finite} \bigr)
+  \bigl( \delta \in \Data \text{ and } \left|\delta\right| \text{ is finite} \bigr)
 }{
@@
 This is enforced in the third phase by the \rn{Split} rule.
-A companion rule, \rn{Singleton}, handles the degenerate case where two terms
-from $\tEc$ have a corecursive singleton type
+A companion rule, \rn{Singleton}, focuses on the degenerate case where two
+terms from $\tEc$ have a corecursive singleton type
 (Section~\ref{sec:the-theory-of-co-datatypes}), in which case they must be

### 2014-12-22 tex/codatadp.tex
shortened rule name

@@
 \)
-\rn{Singleton}
+\rn{Single}
 \end{tabular}
@@
 This is enforced in the third phase by the \rn{Split} rule.
-A companion rule, \rn{Singleton}, focuses on the degenerate case where two
+A companion rule, \rn{Single}, focuses on the degenerate case where two
 terms from $\tEc$ have a corecursive singleton type
@@
 equivalent. Both the finiteness assumption on the datatype in \rn{Split} and
-the cardinality-1 constraint on the codatatype in \rn{Singleton} are 
+the singleton constraint on the codatatype in \rn{Single} are 
 are computed statically
@@
 Otherwise, the child of $\Ec$ is a closed derivation tree
-whose root node is $\Ec \mathrel\cup t \teq u$ obtained by applying one of the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, \rn{Inject}, \rn{Unique}, or \rn{Singleton}.
+whose root node is $\Ec \mathrel\cup t \teq u$ obtained by applying one of the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, \rn{Inject}, \rn{Unique}, or \rn{Single}.
 In all cases, we have that $\Ec \models_{\thD} t \teq u$.
@@
 For \rn{Unique}, our construction of $\Val$ indicates that the value of $t$ and $u$ are equivalent in all models of $\Ec$.
-For \rn{Singleton}, $t$~and~$u$ must have the same value since the cardinality of their type is $1$.
+For \rn{Single}, $t$~and~$u$ must have the same value since the cardinality of their type is $1$.
 By the induction hypothesis, we have $\Ec \mathrel\cup t \teq u$ is $\thD$-unsatisfiable,
@@
 the cardinality of $\tau$ must be $1$.
-Since \rn{Singleton} does not apply to $\Ec$,
+Since \rn{Single} does not apply to $\Ec$,
 there is only one equivalence class of type $\tau$ in $\Ec$,
@@
 All other rules add equalities to the internal state of the theory solver.
-The rules in phase~1 (along with the optimization for collapsing selectors) and the \rn{Singleton} rule are performed eagerly,
+The rules in phase~1 (along with the optimization for collapsing selectors) and the \rn{Single} rule are performed eagerly,
 that is for partial satisfying assignments $M$, while the rules in phase~2 and the \rn{Split} rule are performed only for complete satisfying assignments $M$,

### 2014-12-22 tex/codatadp.tex
minor

@@
 equivalent. Both the finiteness assumption on the datatype in \rn{Split} and
-the singleton constraint on the codatatype in \rn{Single} are 
-are computed statically
+the singleton constraint on the codatatype in \rn{Single} can be 
+computed statically
 based on a recursive computation of the cardinalities of the
@@
 conclusion does not add a new equality to $\Ec$ and is not $\bot$.
+%%% @ANDY: Is the definition or redundant really OK? Or should it read "if none of the branches add a new eq. or is \bot"?
 A \emph{derivation tree} is a tree whose nodes are finite sets of equalities, where non-root nodes are obtained by

### 2014-12-22 tex/codatadp.tex
minor

@@
 
-\hyphenation{data-type data-types co-data-type co-data-types isa-belle sledge-hammer}
+\hyphenation{data-type data-types co-data-type co-data-types isa-belle sledge-hammer
+non-redun-dant}
 
@@
 conclusion does not add a new equality to $\Ec$ and is not $\bot$.
-%%% @ANDY: Is the definition or redundant really OK? Or should it read "if none of the branches add a new eq. or is \bot"?
-A \emph{derivation tree} is a tree whose nodes are finite sets of equalities, where non-root nodes are obtained by
+A \emph{derivation tree} is a tree whose nodes are finite sets of equalities and
+where each non-root node is obtained by
 a nonredundant application of a derivation rule to its parent node.
@@
 A node is \emph{saturated} if no nonredundant instance of a rule can be applied to it.
-%
+
 If there exists a closed derivation tree with root node $\Ec$, we conclude that $\Ec$ is $\thD$-unsatisfiable.
-If there exists a derivation tree with root node $\Ec$ that contains a saturated node, we conclude that $\Ec$ is $\thD$-satisfiable.
+And if there exists a derivation tree with root node $\Ec$ that contains a saturated node, we conclude that $\Ec$ is $\thD$-satisfiable.
 
@@
 dropping each of its $\mu$-bindings. 
-For instance, the $\mu$-term $\mu x. \const{0}$ representing the integer constant zero
+For example, the $\mu$-term $\MU x.\; \const{0}$ representing the integer constant zero
 is simply $\const{0}$ after dropping the $\mu$-binding of $x$.
@@
 then 2 and 3 also hold for $\nf{(\Val_1 \sigma)}$,
-where $\sigma$ is a substitution of the form $\{ {\Varec{x}} \mapsto \mu {\Varec{x}}. t \}$,
+where $\sigma$ is a substitution of the form $\{ {\Varec{x}} \mapsto \MU{\Varec{x}}.\; t \}$,
 ${\Varec{x}} \in FV( \Val_1 )$,
-and $\mu {\Varec{x}}. t$ is not $\alpha$-equivalent to a term in $\mathcal{T}^\M_{\Varec{x}}(\Val_1)$.
+and $\MU {\Varec{x}}.\; t$ is not $\alpha$-equivalent to a term in $\mathcal{T}^\M_{\Varec{x}}(\Val_1)$.
 To show 2,
@@
 %$\Val_1 \sigma \ec{y}$ contains a strict bisimilar subterm 
-%$\mu {\Varec{w}}. u$ that strictly contains $t_1 \sigma$.
+%$\MU {\Varec{w}}.\; u$ that strictly contains $t_1 \sigma$.
 If $t_1$ (resp. $t_2$) is a free variable that is not ${\Varec{x}}$,
@@
 and $\nf{(\Val_1 \sigma)} \ec{z}$ (resp. $\nf{(\Val_1 \sigma)} \ec{y}$) does not.
-If $t_1$ is of the form $\mu w_1. \const{C}( \vec t )$ 
-and $t_2$ is of the form $\mu w_2. \const{D}( \vec u )$,
+If $t_1$ is of the form $\MU w_1.\; \const{C}( \vec t )$ 
+and $t_2$ is of the form $\MU w_2.\; \const{D}( \vec u )$,
 then the expansion \rem{define} of $\nf{(\Val_1 \sigma)} \ec{y}$ and $\nf{(\Val_1 \sigma)} \ec{z}$
@@
 Since $t_1$ and $t_2$ are minimal,
-say $t_1$ is of the form $\mu w. \const{C}( \vec t )$,
+say $t_1$ is of the form $\MU w.\; \const{C}( \vec t )$,
 and $t_2$ is ${\Varec{x}}$.
-Since $\sigma$ maps ${\Varec{x}}$ to a closed $\mu$-term $\mu {\Varec{x}}. t$,
+Since $\sigma$ maps ${\Varec{x}}$ to a closed $\mu$-term $\MU {\Varec{x}}.\; t$,
 we have that $FV( t_1 ) \subseteq \{ {\Varec{x}} \}$,
@@
 such that $v \vsimv{\Varec{x}} \Val_1 \ec{y} \interp{t_1}$.
-Thus, by assumption on our selection of $\mu {\Varec{x}}. t$,
-we have $\mu {\Varec{x}}. t \not\vsimv{\Varec{x}} \Val_1 \ec{y} \interp{t_1}$,
+Thus, by assumption on our selection of $\MU {\Varec{x}}.\; t$,
+we have $\MU {\Varec{x}}.\; t \not\vsimv{\Varec{x}} \Val_1 \ec{y} \interp{t_1}$,
 which implies that the expansion of $\nf{(\Val_1 \sigma)} \ec{y}$ and $\nf{(\Val_1 \sigma)} \ec{z}$

### 2014-12-22 tex/codatadp.tex
more

@@
 
-If there exists a closed derivation tree with root node $\Ec$, we conclude that $\Ec$ is $\thD$-unsatisfiable.
-And if there exists a derivation tree with root node $\Ec$ that contains a saturated node, we conclude that $\Ec$ is $\thD$-satisfiable.
+Correctness means that if there exists a closed derivation tree with root node
+$\Ec$, we can conclude that $\Ec$ is $\thD$-unsatisfiable; and if there exists
+a derivation tree with root node $\Ec$ that contains a saturated node, we can
+conclude that $\Ec$ is $\thD$-satisfiable.
 

### 2014-12-22 tex/codatadp.tex
more

@@
 
+%\newcommand\AST{\ast}
+\newcommand\AST{\infty} %%% matter of taste -- revert if you don't like
+
 \newcommand\MU{\vvthinspace\mu\vvthinspace}
@@
 \newcommand{\tEc}{\mathcal{T}(\Ec)}
-\newcommand{\tcEc}{\mathcal{T}^\ast(\Ec)}
+\newcommand{\tcEc}{\mathcal{T}^\AST(\Ec)}
 \newcommand{\rn}[1]{\textsf{\small #1}}
@@
 and $\const{s}_{i\negvthinspace j}^{\,k} : \delta_i \to \tau_{i\negvthinspace j}^{\,k}$. Defaults are assumed for
-the selector names if they are omitted.
+the selector names if they are omitted. Given a type $\delta$, its
+constructors and selectors are denoted by $\Ctr^\delta$ and $\Sel^\delta$,
+respectively.
 
@@
 
+%%% @ANDY: I explicitly write "|\delta| is finite" instead of "\delta is finite"
+%%% to emphasize the dependency on the model. In principle, some types are
+%%% interpreted finitely in some models and infinitely in others. In Isabelle,
+%%% those wouldn't be considered "finite" types, but they wouldn't be "infinite"
+%%% either. (We have Haskell-style type classes for both of these concepts.)
+%%% Naturally, thanks to our assumptions at the beginning of Sect. 3, our types
+%%% are either finite or infinite.
+
 \begin{figure}[tbh]
@@
   \quad
+  \Ctr^\delta = \{ \const{C}_1, \ldots, \const{C}_m \}
+\\
   \bigl( \const s( t ) \in \tEc \text{ and } \const s \in \Sel^\delta \bigr)
@@
 the equivalence class of $t$ must contain an application of one of
-$\delta$'s constructors
-$\Ctr^\delta = \{ \const{C}_1, \ldots, \const{C}_m \}$.
+$\delta$'s constructors.
 This is enforced in the third phase by the \rn{Split} rule.
@@
 \paragraph{Correctness.}
-An application of a rule is \emph{redundant} if at least one branch in the
-conclusion does not add a new equality to $\Ec$ and is not $\bot$.
+An application of a rule is \emph{redundant} if
+at least one branch in the conclusion is an assignment to $\Ec$ that leaves it
+unchanged.
 A \emph{derivation tree} is a tree whose nodes are finite sets of equalities and
@@
 Consider a derivation tree with root node $\Ec$.
-Let $\SSS$ be the set of terms occurring as the argument of a selector in $\tEc$.
-Let $T$ be the set of terms in $\tEc$ whose types are finite datatypes.
-For each term $t \in T$,
-let $\SSSS 0_t$ be the set $\{ t \}$,
-for $i \geq 0$ let $\SSSS {i+1}_t =$ $\SSSS i_t \mathrel\cup \{ \const s( u ) \mid u : \upsilon \in \SSSS i, \upsilon \in \Data, \upsilon \text{ finite}, s \in \Sel^{\upsilon}  \}$,
-and let $\SSSS \ast_t$ be the fixed point of this sequence.
-We know this is a finite set for each $t$ since the values of the type of $t$ are of finite size.
-Let $\SSSS \ast$ be the union of $\SSSS \ast_t$ for all $t \in T$,
-and let $\tcEc$ be the set of subterms of $\Ec \mathrel\cup \{ \const C_j( \const s^1_j( t ), \ldots, \const s^{n_j}_j( t ) ) \mid t : \delta \in \SSS \mathrel\cup \SSSS \ast,\, \const C_j \in \Ctr^\delta \}$.
+Let $\SSS \subseteq \tEc$ be the set of terms occurring as arguments of selectors.
+Let $T \subseteq \tEc$ be the set of terms whose types are finite datatypes.
+For each term $t \in T$\negvthinspace,
+let
+\begin{align*}
+\smash{\SSSS 0_t} & = \{ t \}
+& \smash{\SSSS {i+1}_t} & = \smash{\SSSS i_t \mathrel\cup \{ \const s( u ) \mid u : \delta \in \SSSS i{,}~ \delta \in \Data{,}~ \left|\delta\right| \text{ is finite}{,}~ \const{s} \in \Sel^{\delta}}  \}
+\end{align*}
+%for $i \geq 0$, let ;
+and let $\SSSS \AST_t$ be the limit of this sequence.
+This is a finite set for each $t$ because the values of the type of $t$ are of finite size.
+Let $\SSSS \AST$ be the union of $\SSSS \AST_t$ for all $t \in T$,
+and let $\tcEc$ be the set of subterms of $\Ec \mathrel\cup \{ \const C_j( \const s^1_j( t ), \ldots, \const s^{n_j}_j( t ) ) \mid t : \delta \in \SSS \mathrel\cup \SSSS \AST,\, \const C_j \in \Ctr^\delta \}$.
 In a derivation tree with root node~$\Ec$,
@@
 For the latter, this will be based on the mapping $\Val$ constructed in phase~2 of the calculus.
-We compute the \emph{completion} $\Val^\ast$ of $\Val$ for normal model $\M$
+We compute the \emph{completion} $\Val^\AST$ of $\Val$ for normal model $\M$
 by assigning values from $\M$ to unassigned variables in the domain of $\Val$.
@@
 
-We construct $\Val^\ast$ by exhaustively applying the following rule to $\nf{\Val}$:
+We construct $\Val^\AST$ by exhaustively applying the following rule to $\nf{\Val}$:
 \[
@@
 
-%We first state the following properties of $\Val^\ast$,
-In the following lemma about $\Val^\ast$,
+%We first state the following properties of $\Val^\AST$,
+In the following lemma about $\Val^\AST$,
 $\muvar( t )$ denotes $t$ if $t$ is a variable and ${\Varec{x}}$ if $t$ is of the form $\MU {\Varec{x}}.\; u$.
@@
 %%% @ANDY: I think it's enough to point this out clearly in the introduction.
-%%% At least, that always worked fine for me. :)
+%%% At least, that always worked fine for my other papers. :)
 %
@@
 If $\Val$ is constructed for a saturated set $\Ec$
-and $\Val^\ast$ is a completion of $\Val$ for normal model $\M$, the following holds:
+and $\Val^\AST$ is a completion of $\Val$ for normal model $\M$, the following holds:
 \begin{enumerate}
-\item[\rm 1.] $\Val^\ast \ec{x}$ is $\alpha$-equivalent to a value in $\M( \tau )$, where $\ec{x}$ has type $\tau$.
+\item[\rm 1.] $\Val^\AST \ec{x}$ is $\alpha$-equivalent to a value in $\M( \tau )$, where $\ec{x}$ has type $\tau$.
 \item[\rm 2.]
-$\Val^\ast \ec{x} = \Val^\ast \ec{y} \interp{ t }$ for
-all subterms $t$ of $\Val^\ast \ec{y}$ with $\muvar(t) = {\Varec{x}}$. 
+$\Val^\AST \ec{x} = \Val^\AST \ec{y} \interp{ t }$ for
+all subterms $t$ of $\Val^\AST \ec{y}$ with $\muvar(t) = {\Varec{x}}$. 
 \item[\rm 3.]
-$\Val^\ast \ec{x} \vsim \Val^\ast \ec{y}$ if and only if $\ec{x} = \ec{y}$.
+$\Val^\AST \ec{x} \vsim \Val^\AST \ec{y}$ if and only if $\ec{x} = \ec{y}$.
 \end{enumerate}
@@
 \begin{proof}
-To show 1, we first show that $\Val^\ast$ contains no free variables.
-Assume by contradiction that $\Val^\ast$ contained a free variable ${\Varec{y}}$ for some $\ec{y}$ of type $\tau$.
+To show 1, we first show that $\Val^\AST$ contains no free variables.
+Assume by contradiction that $\Val^\AST$ contained a free variable ${\Varec{y}}$ for some $\ec{y}$ of type $\tau$.
 Then it must be the case that $\ec{y}$ does not contain a constructor term,
@@
 there is only one equivalence class of type $\tau$ in $\Ec$,
-and thus there are no terms in $\mathcal{T}^\M_{\Varec{y}}(\Val^\ast)$ of type $\tau$.
+and thus there are no terms in $\mathcal{T}^\M_{\Varec{y}}(\Val^\AST)$ of type $\tau$.
 This is a contradiction, since our model completion may assign the value in the domain of $\tau$ to ${\Varec{y}}$.
@@
 This is also a contradiction,
-since there are only a finite number of closed terms in $\mathcal{T}^\M_{\Varec{y}}(\Val^\ast)$,
-and thus our model completion can assign a value not occurring in $\mathcal{T}^\M_{\Varec{y}}(\Val^\ast)$ to ${\Varec{y}}$.
-By construction, $\Val^\ast \ec{x}$ is normal.
+since there are only a finite number of closed terms in $\mathcal{T}^\M_{\Varec{y}}(\Val^\AST)$,
+and thus our model completion can assign a value not occurring in $\mathcal{T}^\M_{\Varec{y}}(\Val^\AST)$ to ${\Varec{y}}$.
+By construction, $\Val^\AST \ec{x}$ is normal.
 Since \rn{Acyclic} does not apply, $\Val \ec{x}$ is acyclic when $\tau \in \Data$.
-Moreover, our construction of $\Val^\ast$ applies substitutions of the form
+Moreover, our construction of $\Val^\AST$ applies substitutions of the form
 $\{ {\Varec{y}} \mapsto t \}$, where $t$ is acyclic when the type of ${\Varec{y}}$ is not a codatatype.
-Thus, $\Val^\ast \ec{x}$ is acyclic when $\tau \not\in \Codata$.
-Therefore, by definition, $\Val^\ast \ec{x}$ is $\alpha$-equivalent to a value in $\M( \tau )$.
+Thus, $\Val^\AST \ec{x}$ is acyclic when $\tau \not\in \Codata$.
+Therefore, by definition, $\Val^\AST \ec{x}$ is $\alpha$-equivalent to a value in $\M( \tau )$.
 
@@
 
-Thus, by induction on the number of applications of the above rule used to obtain $\Val^\ast$,
-we have that $\Val^\ast$ satisfies statements 2 and 3.
+Thus, by induction on the number of applications of the above rule used to obtain $\Val^\AST$,
+we have that $\Val^\AST$ satisfies statements 2 and 3.
 \qed
@@
 that interprets wrongly applied selectors based on equality information in $\Fc$,
-and interprets the variables of $\Fc$ based on the completion $\Val^\ast$ of the mapping $\Val$ from phase~2.
-For the latter, let $\M( x )$ be the value in $\M$ that is $\alpha$-equivalent with $\Val^\ast \ec{x}$ for each variable $x$ in $\mathcal{T}(\Fc)$.
+and interprets the variables of $\Fc$ based on the completion $\Val^\AST$ of the mapping $\Val$ from phase~2.
+For the latter, let $\M( x )$ be the value in $\M$ that is $\alpha$-equivalent with $\Val^\AST \ec{x}$ for each variable $x$ in $\mathcal{T}(\Fc)$.
 
@@
 by structural induction on $t$,
-that $\M( t ) \aequiv \Val^\ast \ec{t}$ for all terms $t \in \mathcal{T}( \Fc )$,
+that $\M( t ) \aequiv \Val^\AST \ec{t}$ for all terms $t \in \mathcal{T}( \Fc )$,
 which since $\M$ is normal, implies $\M \models t_1 \teq t_2$.
 
-If $t$ is a variable, then $\M( t ) \vsim \Val^\ast \ec{ t }$ by construction.
+If $t$ is a variable, then $\M( t ) \vsim \Val^\AST \ec{ t }$ by construction.
 
@@
 then $\M( t )$ is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \M( u_1 ), \ldots, \M( u_n ) ) }$ for some fresh $x$,
-which by the induction hypothesis is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \Val^\ast \ec{ u_1 }, \ldots, \Val^\ast \ec{ u_n } )}$, call this term $t_u$.
-%By Lemma~\ref{lem:model-completion}.2 and due to our construction of $\Val^\ast$, we have $\Val^\ast \ec{t}$ is a term
+which by the induction hypothesis is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \Val^\AST \ec{ u_1 }, \ldots, \Val^\AST \ec{ u_n } )}$, call this term $t_u$.
+%By Lemma~\ref{lem:model-completion}.2 and due to our construction of $\Val^\AST$, we have $\Val^\AST \ec{t}$ is a term
 %of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$
-%where $\Val^\ast \ec{t}\interp{w_i} = \Val^\ast \ec{u_i}$ for $i = 1, \ldots, n$.
+%where $\Val^\AST \ec{t}\interp{w_i} = \Val^\AST \ec{u_i}$ for $i = 1, \ldots, n$.
 Since \rn{Inject} and \rn{Clash} do not apply to $\Fc$, 
-by our construction of $\Val^\ast$ we have $\Val^\ast \ec{t}$ is a term
+by our construction of $\Val^\AST$ we have $\Val^\AST \ec{t}$ is a term
 of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$ 
@@
 $\muvar(w_i) = {\Varec{u_i}}$ and thus by Lemma~\ref{lem:model-completion}.2, 
-$\Val^\ast \ec{t}\interp{w_i} = \Val^\ast \ec{u_i}$.
+$\Val^\AST \ec{t}\interp{w_i} = \Val^\AST \ec{u_i}$.
 For each $i = 1, \ldots, n$, let $u_i'$ be the $i$\vvthinspace th argument of $t_u$.
-By Lemma~\ref{lem:mu-norm-arg}, $t_u\interp{u_i'} \vsim \Val^\ast \ec{u_i}$.
-Thus $t_u\interp{u_i'} \vsim \Val^\ast \ec{t}\interp{ w_i }$. % for each $i = 1, \ldots, n$.
+By Lemma~\ref{lem:mu-norm-arg}, $t_u\interp{u_i'} \vsim \Val^\AST \ec{u_i}$.
+Thus $t_u\interp{u_i'} \vsim \Val^\AST \ec{t}\interp{ w_i }$. % for each $i = 1, \ldots, n$.
 %By Lemma~\ref{lem:mu-cong}, we have $t_u \vsim t_w$.
-Thus, $\M( t ) \vsim t_u \vsim \Val^\ast \ec{t}$, and
-we have $\M( t ) \vsim \Val^\ast \ec{t}$.
+Thus, $\M( t ) \vsim t_u \vsim \Val^\AST \ec{t}$, and
+we have $\M( t ) \vsim \Val^\AST \ec{t}$.
 
@@
 $\ec{u}$ must contain a term of the form $\const C_{j'}( \const s^1_{j'}( u ), \ldots, \const s^{n}_{j'}( u ) )$ for some $j'$.
-%By Lemma~\ref{lem:model-completion}.2 and due to our construction of $\Val^\ast$, we have $\Val^\ast \ec{u}$ is of the form
+%By Lemma~\ref{lem:model-completion}.2 and due to our construction of $\Val^\AST$, we have $\Val^\AST \ec{u}$ is of the form
 %$\MU {\Varec{u}}.\; \const C_{j'}( w_1, \ldots, w_n )$,
-%where $\Val^\ast \ec{u} \interp{ w_i } = \Val^\ast \ec{ \const s^i_{j'}( u )}$ for $i = 1, \ldots, n$. 
+%where $\Val^\AST \ec{u} \interp{ w_i } = \Val^\AST \ec{ \const s^i_{j'}( u )}$ for $i = 1, \ldots, n$. 
 Since \rn{Inject} and \rn{Clash} do not apply to $\Fc$, 
-by our construction of $\Val^\ast$ we have $\Val^\ast \ec{u}$ is of the form
+by our construction of $\Val^\AST$ we have $\Val^\AST \ec{u}$ is of the form
 $\MU {\Varec{u}}.\; \const C_{j'}( w_1, \ldots, w_n )$,
@@
 $\muvar(w_i) = {\Varec{\const s^i_{j'}( u )}}$ and thus by Lemma~\ref{lem:model-completion}.2, 
-$\Val^\ast \ec{u} \interp{ w_i } = \Val^\ast \ec{ \const s^i_{j'}( u )}$. 
-If $j = j'$, then $\M( t )$ is $\alpha$-equivalent with $\Val^\ast \ec{u} \interp{w_k}$, which is equal to $\Val^\ast \ec{ \const s^k_j( u )}$,
-which is $\Val^\ast \ec{t}$.
-If $j \neq j'$, we are free to interpret $\const s^k_j( \M( u ) )$ as the value in $\M( \tau )$ that is $\alpha$-equivalent with $\Val^\ast \ec{ t }$.
+$\Val^\AST \ec{u} \interp{ w_i } = \Val^\AST \ec{ \const s^i_{j'}( u )}$. 
+If $j = j'$, then $\M( t )$ is $\alpha$-equivalent with $\Val^\AST \ec{u} \interp{w_k}$, which is equal to $\Val^\AST \ec{ \const s^k_j( u )}$,
+which is $\Val^\AST \ec{t}$.
+If $j \neq j'$, we are free to interpret $\const s^k_j( \M( u ) )$ as the value in $\M( \tau )$ that is $\alpha$-equivalent with $\Val^\AST \ec{ t }$.
 Since \rn{Cong} does not apply and due to Lemma~\ref{lem:model-completion}.3,
@@
 Since \rn{Conflict} does not apply, $t_1 \teq t_2 \not\in \Fc$ and thus $\ec{t_1}$ and $\ec{t_2}$ are distinct.
-Since $\M( t_1 ) \vsim \Val^\ast \ec{t_1}$ and $\M( t_2 ) \vsim \Val^\ast \ec{t_2}$,
+Since $\M( t_1 ) \vsim \Val^\AST \ec{t_1}$ and $\M( t_2 ) \vsim \Val^\AST \ec{t_2}$,
 by Lemma~\ref{lem:model-completion}.3, $\M( t_1 ) \neq \M( t_2 )$, and thus $\M \models t_1 \tneq t_2$.

### 2014-12-22 tex/codatadp.tex
more

@@
 \bigwedge_{\,i}\, \forall v\; w.\;\, R_i[v, w] \longrightarrow
-  \bigwedge_{\,j}\, \const{d}_j(v) \teq \const{d}_j(w)
+  \bigwedge_{\,j}\, \const{d}_{\negvthinspace j}(v) \teq \const{d}_{\negvthinspace j}(w)
   \mathrel\land
-  \const{d}_j(v) \longrightarrow \bigwedge_{\,k}\, \const{s}_{i\negvthinspace j}^{\,k}(v) \sim \const{s}_{i\negvthinspace j}^{\,k}(w)
+  \const{d}_{\negvthinspace j}(v) \longrightarrow \bigwedge_{\,k}\, \const{s}_{i\negvthinspace j}^{\,k}(v) \sim \const{s}_{i\negvthinspace j}^{\,k}(w)
 \end{gathered}$}
@@
 %We will commonly denote tuples of terms $( t_1, \ldots, t_n )$ in bold font, as $\vec t$.
-%We assume that all $\Sig$-terms in $\Ec$ are \emph{normalized}, meaning that all subterms of the form $\const{s}^{\,k}_j( \const{C}_j(t_1,\ldots,t_n) )$
+%We assume that all $\Sig$-terms in $\Ec$ are \emph{normalized}, meaning that all subterms of the form $\const{s}^{\,k}_{\negvthinspace j}( \const{C}_{\negvthinspace j}(t_1,\ldots,t_n) )$
 %are simplified to $t_k$. All constraints added to $\Ec$ are also implicitly
@@
 
-%We assume that all $\Sig$-terms in $\Ec$ are \emph{normalized}, meaning that all subterms of the form $\const{s}^{\,k}_j( \const{C}_j(t_1,\ldots,t_n) )$
+%We assume that all $\Sig$-terms in $\Ec$ are \emph{normalized}, meaning that all subterms of the form $\const{s}^{\,k}_{\negvthinspace j}( \const{C}_{\negvthinspace j}(t_1,\ldots,t_n) )$
 %are simplified to $t_i$, and moreover assume that all %additional
@@
 This is a finite set for each $t$ because the values of the type of $t$ are of finite size.
-Let $\SSSS \AST$ be the union of $\SSSS \AST_t$ for all $t \in T$,
-and let $\tcEc$ be the set of subterms of $\Ec \mathrel\cup \{ \const C_j( \const s^1_j( t ), \ldots, \const s^{n_j}_j( t ) ) \mid t : \delta \in \SSS \mathrel\cup \SSSS \AST,\, \const C_j \in \Ctr^\delta \}$.
+Let $\SSSS \AST$ be the union of all sets $\SSSS \AST_t$ where $t \in T$,
+and let $\tcEc$ be the set of subterms of $\Ec \mathrel\cup \{ \const C_{\negvthinspace j}( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\smash{\,n_{\negvthinspace j}}}_{\negvthinspace j}( t ) ) \mid t : \delta \in \SSS \mathrel\cup \SSSS \AST,\, \const C_{\negvthinspace j} \in \Ctr^\delta \}$.
 In a derivation tree with root node~$\Ec$,
@@
 whose root nodes are the result of applying the rule \rn{Split} on term $t$ of type $\delta$,
-by the induction hypothesis $\Ec \mathrel\cup t \teq \const C_j( \const s^1_j( t ), \ldots, \const s^{n_j}_j( t ) )$ is unsatisfiable
-for each $\const C_j \in \Ctr^\delta$.
-Since by exhaustiveness, all models of $\thD$ entail (exactly) one equality $t \teq \const C_j( \const s^1_j( t ), \ldots, \const s^{n_j}_j( t ) )$,
+by the induction hypothesis $\Ec \mathrel\cup t \teq \const C_{\negvthinspace j}( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\,\smash{n_{\negvthinspace j}}}_{\negvthinspace j}( t ) )$ is unsatisfiable
+for each $\const C_{\negvthinspace j} \in \Ctr^\delta$.
+Since by exhaustiveness, all models of $\thD$ entail (exactly) one equality $t \teq \const C_{\negvthinspace j}( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\,\smash{n_{\negvthinspace j}}}_{\negvthinspace j}( t ) )$,
 we know $\Ec$ is $\thD$-unsatisfiable.
@@
 \item
-For all selector terms $\const s^k_j( t )$ of type $\tau$, % for some $\vec u$,
-$\M( \const s^k_j( t ) )$ is the value
+For all selector terms $\const s^k_{\negvthinspace j}( t )$ of type $\tau$, % for some $\vec u$,
+$\M( \const s^k_{\negvthinspace j}( t ) )$ is the value
 in $\M( \tau )$ that is $\alpha$-equivalent to
 $\M( t )\interp{u_k}$
-when $\M( t )$ is of the form $\MU x.\; \const{C}_j( \vec u )$.
+when $\M( t )$ is of the form $\MU x.\; \const{C}_{\negvthinspace j}( \vec u )$.
 \end{enumerate}
@@
 For all equivalence classes $\ec{z}$,
-each pair of constructor terms $\const{C}_j( \vec t )$ and $\const{C}_{j'}( \vec u )$ in $\ec{z}$
+each pair of constructor terms $\const{C}_{\negvthinspace j}( \vec t )$ and $\const{C}_{j'}( \vec u )$ in $\ec{z}$
 are such that $j = j'$ since \rn{Clash} does not apply to $\Ec$,
@@
 where all substitutions for variables ${\Varec{z}}$ 
-were uniquely of the form $\{ {\Varec{z}} \mapsto \const{C}_j( {\Varec{t_1}}, \ldots, {\Varec{t_n}} ) \}$
-when $\ec{z}$ contains a constructor $\const{C}_j( t_1, \ldots, t_n )$.
+were uniquely of the form $\{ {\Varec{z}} \mapsto \const{C}_{\negvthinspace j}( {\Varec{t_1}}, \ldots, {\Varec{t_n}} ) \}$
+when $\ec{z}$ contains a constructor $\const{C}_{\negvthinspace j}( t_1, \ldots, t_n )$.
 Say $\Val \ec{y}$ has a subterm $t$ where $\muvar(t) = {\Varec{x}}$.
@@
 Since \rn{Cong} does not apply to $\Ec$
-we have $\ec{w_j}$ and $\ec{z_j}$ are distinct for some $j$.
-By statement 2, $\Val \ec{x} \interp{ t_j } = \Val \ec{z_j}$
-and $\Val \ec{x} \interp{ u_j } = \Val \ec{w_j}$,
+we have $\ec{w_{\negvthinspace j}}$ and $\ec{z_{\negvthinspace j}}$ are distinct for some $j$.
+By statement 2, $\Val \ec{x} \interp{ t_{\negvthinspace j} } = \Val \ec{z_{\negvthinspace j}}$
+and $\Val \ec{x} \interp{ u_{\negvthinspace j} } = \Val \ec{w_{\negvthinspace j}}$,
 which are not $\alpha$-equivalent by statement 3, contradicting the fact that $\MU {\Varec{y}}.\; \const{C}( u_1, \ldots, u_n )$ is a bisimilar subterm of $\Val \ec{x}$.
@@
 
-If $t$ is a selector term of the form $\const s^k_j( u )$,
+If $t$ is a selector term of the form $\const s^k_{\negvthinspace j}( u )$,
 since \rn{Split} does not apply to $\Fc$,
@@
 $\Val^\AST \ec{u} \interp{ w_i } = \Val^\AST \ec{ \const s^i_{j'}( u )}$. 
-If $j = j'$, then $\M( t )$ is $\alpha$-equivalent with $\Val^\AST \ec{u} \interp{w_k}$, which is equal to $\Val^\AST \ec{ \const s^k_j( u )}$,
+If $j = j'$, then $\M( t )$ is $\alpha$-equivalent with $\Val^\AST \ec{u} \interp{w_k}$, which is equal to $\Val^\AST \ec{ \const s^k_{\negvthinspace j}( u )}$,
 which is $\Val^\AST \ec{t}$.
-If $j \neq j'$, we are free to interpret $\const s^k_j( \M( u ) )$ as the value in $\M( \tau )$ that is $\alpha$-equivalent with $\Val^\AST \ec{ t }$.
+If $j \neq j'$, we are free to interpret $\const s^k_{\negvthinspace j}( \M( u ) )$ as the value in $\M( \tau )$ that is $\alpha$-equivalent with $\Val^\AST \ec{ t }$.
 Since \rn{Cong} does not apply and due to Lemma~\ref{lem:model-completion}.3,
@@
 
-Discriminators are treated as predicate symbols, so a discriminator $\const{d}_j( t )$ is such that $\const{d}_j$
-is a predicate symbol, instead of reducing this constraint to $t \teq \const{C}_j( \const s^1_j( t ), \ldots, \const s^n_j( t ) )$.
+Discriminators are treated as predicate symbols, so a discriminator $\const{d}_{\negvthinspace j}( t )$ is such that $\const{d}_{\negvthinspace j}$
+is a predicate symbol, instead of reducing this constraint to $t \teq \const{C}_{\negvthinspace j}( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^n_{\negvthinspace j}( t ) )$.
 We have found that this leads to better performance since this reduction introduces terms more eagerly to $\tEc$.
@@
 Also, selectors are collapsed eagerly:
-if $\const s^k_j( t )$ is a term in $\tEc$ and $t$ contains the constructor
-$\const{C}_j( \vec u )$, we immediately infer $\const s^k_j( t ) \teq u_k$, whereas the calculus we presented would apply \rn{Split} and \rn{Inject} before inferring this equality.
+if $\const s^k_{\negvthinspace j}( t )$ is a term in $\tEc$ and $t$ contains the constructor
+$\const{C}_{\negvthinspace j}( \vec u )$, we immediately infer $\const s^k_{\negvthinspace j}( t ) \teq u_k$, whereas the calculus we presented would apply \rn{Split} and \rn{Inject} before inferring this equality.
 For the sake of reducing the number of unique constraints considered by the calculus, we compute a normal form for constraints as a preprocessing step.
@@
 replace $t \teq s$ by $s \teq t$ if $s$ is less than $t$ based on a term ordering,
-replace $\const{C}_j( \vec t ) \teq \const{C}_{j'}( \vec u )$ with $\bot$ when $j \neq j'$,
-replace all selector terms of the form $\const{s}^{\,k}_j( \const{C}_j(t_1,\ldots,t_n) )$ by $t_k$,
-and replace all occurrences of discriminators $\const{d}_j( \const{C}_{j'}( \vec t ) )$ by $\top$ if $j=j'$ and $\bot$ otherwise.
+replace $\const{C}_{\negvthinspace j}( \vec t ) \teq \const{C}_{j'}( \vec u )$ with $\bot$ when $j \neq j'$,
+replace all selector terms of the form $\const{s}^{\,k}_{\negvthinspace j}( \const{C}_{\negvthinspace j}(t_1,\ldots,t_n) )$ by $t_k$,
+and replace all occurrences of discriminators $\const{d}_{\negvthinspace j}( \const{C}_{j'}( \vec t ) )$ by $\top$ if $j=j'$ and $\bot$ otherwise.
 

### 2014-12-22 tex/codatadp.tex
more

@@
 }
+
+\renewcommand\models{\vDash} %%% matter of taste
 
@@
 Let $\SSSS \AST$ be the union of all sets $\SSSS \AST_t$ where $t \in T$,
-and let $\tcEc$ be the set of subterms of $\Ec \mathrel\cup \{ \const C_{\negvthinspace j}( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\smash{\,n_{\negvthinspace j}}}_{\negvthinspace j}( t ) ) \mid t : \delta \in \SSS \mathrel\cup \SSSS \AST,\, \const C_{\negvthinspace j} \in \Ctr^\delta \}$.
+and let $\tcEc$ be the set of subterms of $\Ec \mathrel\cup \{ \const C_{\negvthinspace j}( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\smash{\,n_{\negvthinspace j}}}_{\negvthinspace j}( t ) ) \mid t : \delta \in \SSS \mathrel\cup \SSSS \AST{,}~ \const C_{\negvthinspace j} \in \Ctr^\delta \}$.
 In a derivation tree with root node~$\Ec$,
@@
 For \rn{Clash}, this is a consequence of distinctness.
-For \rn{Acyclic}, the construction of $\Val$ indicates that the class of values that $t$ can take in models of $\Ec$ is infeasible,
+For \rn{Acyclic}, the construction of $\Val$ indicates that the class of values that term~$t$ can take in models of $\Ec$ is infeasible,
 and thus $\Ec$ is unsatisfiable.
 If the children of $\Ec$ are closed derivation trees
-whose root nodes are the result of applying the rule \rn{Split} on term $t$ of type $\delta$,
-by the induction hypothesis $\Ec \mathrel\cup t \teq \const C_{\negvthinspace j}( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\,\smash{n_{\negvthinspace j}}}_{\negvthinspace j}( t ) )$ is unsatisfiable
-for each $\const C_{\negvthinspace j} \in \Ctr^\delta$.
+whose root nodes are the result of applying the rule \rn{Split} on $t : \delta$,
+by the induction hypothesis $\Ec \mathrel\cup t \teq \smash{\const C_{\negvthinspace j}( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\,\smash{n_{\negvthinspace j}}}_{\negvthinspace j}( t ) )}$ is unsatisfiable
+for each $\const C_{\negvthinspace j} \in \smash{\Ctr^\delta}$.
 Since by exhaustiveness, all models of $\thD$ entail (exactly) one equality $t \teq \const C_{\negvthinspace j}( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\,\smash{n_{\negvthinspace j}}}_{\negvthinspace j}( t ) )$,
-we know $\Ec$ is $\thD$-unsatisfiable.
+this means $\Ec$ is $\thD$-unsatisfiable.
 Otherwise, the child of $\Ec$ is a closed derivation tree

### 2014-12-22 tex/codatadp.tex
more

@@
 \newcommand{\thO}{T_{\mathrm{o}}}
-\newcommand{\thD}{T_{\!\mathrm{cd}}}
+%\newcommand{\thD}{T_{\!\mathrm{cd}}}
+%%% @ANDY: the italicized "T" makes it look as though it were a variable;
+%%% and the subscript means we have two layers of subscripts in things like
+%%% \model_\thD (although we could drop the "T" and keep only the subscript
+%%% like you did in the string paper).
+%%% How about the following? I'm pretty sure \mathcal has been used for theories
+%%% before by other authors.
+\newcommand{\thD}{\mathcal{CoDt}}
 
@@
 }{
-  \Ec := \Ec,\: t \teq \const{C}_1(\const s^1_1( t ), \ldots,\const s^{n_1}_1( t ) ) \ror \cdots \ror \Ec := \Ec,\: t \teq \const{C}_m(\const s^1_m( t ), \ldots,\const s^{n_m}_m( t ) )
+  \Ec := \Ec,\: t \teq \const{C}_1\bigl(\const s^1_1( t ), \ldots,\const s^{n_1}_1( t ) \bigr) \ror \cdots \ror \Ec := \Ec,\: t \teq \const{C}_m\bigl(\const s^1_m( t ), \ldots,\const s^{n_m}_m( t ) \bigr)
 }
@@
 Let $\SSSS \AST$ be the union of all sets $\SSSS \AST_t$ where $t \in T$,
-and let $\tcEc$ be the set of subterms of $\Ec \mathrel\cup \{ \const C_{\negvthinspace j}( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\smash{\,n_{\negvthinspace j}}}_{\negvthinspace j}( t ) ) \mid t : \delta \in \SSS \mathrel\cup \SSSS \AST{,}~ \const C_{\negvthinspace j} \in \Ctr^\delta \}$.
+and let $\tcEc$ be the set of subterms of $\Ec \mathrel\cup \{ \const C_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\smash{\,n_{\negvthinspace j}}}_{\negvthinspace j}( t ) \bigr) \mid t : \delta \in \SSS \mathrel\cup \SSSS \AST{,}~ \const C_{\negvthinspace j} \in \Ctr^\delta \}$.
 In a derivation tree with root node~$\Ec$,
@@
 whose root nodes are the result of applying the rule \rn{Split} on $t : \delta$,
-by the induction hypothesis $\Ec \mathrel\cup t \teq \smash{\const C_{\negvthinspace j}( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\,\smash{n_{\negvthinspace j}}}_{\negvthinspace j}( t ) )}$ is unsatisfiable
+by the induction hypothesis $\Ec \mathrel\cup t \teq \smash{\const C_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\,\smash{n_{\negvthinspace j}}}_{\negvthinspace j}( t ) \bigr)}$ is unsatisfiable
 for each $\const C_{\negvthinspace j} \in \smash{\Ctr^\delta}$.
-Since by exhaustiveness, all models of $\thD$ entail (exactly) one equality $t \teq \const C_{\negvthinspace j}( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\,\smash{n_{\negvthinspace j}}}_{\negvthinspace j}( t ) )$,
+Since by exhaustiveness, all models of $\thD$ entail (exactly) one equality $t \teq \const C_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\,\smash{n_{\negvthinspace j}}}_{\negvthinspace j}( t ) \bigr)$,
 this means $\Ec$ is $\thD$-unsatisfiable.
@@
 For all selector terms $\const s^k_{\negvthinspace j}( t )$ of type $\tau$, % for some $\vec u$,
-$\M( \const s^k_{\negvthinspace j}( t ) )$ is the value
+$\M\bigl( \const s^k_{\negvthinspace j}( t ) \bigr)$ is the value
 in $\M( \tau )$ that is $\alpha$-equivalent to
@@
 since \rn{Split} does not apply to $\Fc$,
-$\ec{u}$ must contain a term of the form $\const C_{j'}( \const s^1_{j'}( u ), \ldots, \const s^{n}_{j'}( u ) )$ for some $j'$.
+$\ec{u}$ must contain a term of the form $\const C_{j'}\bigl( \const s^1_{j'}( u ), \ldots, \const s^{n}_{j'}( u ) \bigr)$ for some $j'$.
 %By Lemma~\ref{lem:model-completion}.2 and due to our construction of $\Val^\AST$, we have $\Val^\AST \ec{u}$ is of the form
@@
 Discriminators are treated as predicate symbols, so a discriminator $\const{d}_{\negvthinspace j}( t )$ is such that $\const{d}_{\negvthinspace j}$
-is a predicate symbol, instead of reducing this constraint to $t \teq \const{C}_{\negvthinspace j}( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^n_{\negvthinspace j}( t ) )$.
+is a predicate symbol, instead of reducing this constraint to $t \teq \const{C}_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^n_{\negvthinspace j}( t ) \bigr)$.
 We have found that this leads to better performance since this reduction introduces terms more eagerly to $\tEc$.
@@
 For the \rn{Split} rule, we add a clause of the form
-$(\const{C}_1( \const s^1_1( t ), \ldots, \const s^{n_1}_1( t ) ) \vee \ldots \vee t \teq \const{C}_m( \const s^1_m( t ), \ldots, \const s^{n_m}_m( t ) ))$
+$\const{C}_1\bigl( \const s^1_1( t ), \ldots, \const s^{n_1}_1( t ) \bigr) \mathrel\lor \cdots \mathrel\lor t \teq \const{C}_m\bigl( \const s^1_m( t ), \ldots, \const s^{n_m}_m( t ) \bigr)$
 to $F$.
-Notice that decisions on which branch to take are thus performed externally by the SAT solver, and not by the theory solver.
+Decisions on which branch to take are thus performed externally by the SAT solver, and not by the theory solver.
 All other rules add equalities to the internal state of the theory solver.

### 2014-12-22 tex/codatadp.tex
more

@@
 For all equivalence classes $\ec{z}$,
-each pair of constructor terms $\const{C}_{\negvthinspace j}( \vec t )$ and $\const{C}_{j'}( \vec u )$ in $\ec{z}$
-are such that $j = j'$ since \rn{Clash} does not apply to $\Ec$,
+each pair of constructor terms $\const{C}_{\negvthinspace j}( \vec t )$ and $\const{C}_{\vthinspace\jPrime}( \vec u )$ in $\ec{z}$
+are such that $j = \jPrime$ since \rn{Clash} does not apply to $\Ec$,
 and are such that $\ec{ \vec t } = \ec{ \vec u }$ since \rn{Inject} does not apply to $\Ec$.
@@
 $\Val^\AST \ec{t}\interp{w_i} = \Val^\AST \ec{u_i}$.
-For each $i = 1, \ldots, n$, let $u_i'$ be the $i$\vvthinspace th argument of $t_u$.
-By Lemma~\ref{lem:mu-norm-arg}, $t_u\interp{u_i'} \vsim \Val^\AST \ec{u_i}$.
-Thus $t_u\interp{u_i'} \vsim \Val^\AST \ec{t}\interp{ w_i }$. % for each $i = 1, \ldots, n$.
+For each $i = 1, \ldots, n$, let $u_\iPrime$ be the $i$\vvthinspace th argument of $t_u$.
+By Lemma~\ref{lem:mu-norm-arg}, $t_u\interp{u_\iPrime} \vsim \Val^\AST \ec{u_i}$.
+Thus $t_u\interp{u_\iPrime} \vsim \Val^\AST \ec{t}\interp{ w_i }$. % for each $i = 1, \ldots, n$.
 %By Lemma~\ref{lem:mu-cong}, we have $t_u \vsim t_w$.
@@
 since \rn{Split} does not apply to $\Fc$,
-$\ec{u}$ must contain a term of the form $\const C_{j'}\bigl( \const s^1_{j'}( u ), \ldots, \const s^{n}_{j'}( u ) \bigr)$ for some $j'$.
+$\ec{u}$ must contain a term of the form $\const C_{\vthinspace\jPrime}\bigl( \const s^1_{\vthinspace\jPrime}( u ), \ldots, \const s^{n}_{\vthinspace\jPrime}( u ) \bigr)$ for some $\jPrime$.
 %By Lemma~\ref{lem:model-completion}.2 and due to our construction of $\Val^\AST$, we have $\Val^\AST \ec{u}$ is of the form
-%$\MU {\Varec{u}}.\; \const C_{j'}( w_1, \ldots, w_n )$,
-%where $\Val^\AST \ec{u} \interp{ w_i } = \Val^\AST \ec{ \const s^i_{j'}( u )}$ for $i = 1, \ldots, n$. 
+%$\MU {\Varec{u}}.\; \const C_{\vthinspace\jPrime}( w_1, \ldots, w_n )$,
+%where $\Val^\AST \ec{u} \interp{ w_i } = \Val^\AST \ec{ \const s^i_{\vthinspace\jPrime}( u )}$ for $i = 1, \ldots, n$. 
 Since \rn{Inject} and \rn{Clash} do not apply to $\Fc$, 
 by our construction of $\Val^\AST$ we have $\Val^\AST \ec{u}$ is of the form
-$\MU {\Varec{u}}.\; \const C_{j'}( w_1, \ldots, w_n )$,
+$\MU {\Varec{u}}.\; \const C_{\vthinspace\jPrime}( w_1, \ldots, w_n )$,
 where for each $i = 1, \ldots, n$,
-$\muvar(w_i) = {\Varec{\const s^i_{j'}( u )}}$ and thus by Lemma~\ref{lem:model-completion}.2, 
-$\Val^\AST \ec{u} \interp{ w_i } = \Val^\AST \ec{ \const s^i_{j'}( u )}$. 
-If $j = j'$, then $\M( t )$ is $\alpha$-equivalent with $\Val^\AST \ec{u} \interp{w_k}$, which is equal to $\Val^\AST \ec{ \const s^k_{\negvthinspace j}( u )}$,
+$\muvar(w_i) = {\Varec{\const s^i_{\vthinspace\jPrime}( u )}}$ and thus by Lemma~\ref{lem:model-completion}.2, 
+$\Val^\AST \ec{u} \interp{ w_i } = \Val^\AST \ec{ \const s^i_{\vthinspace\jPrime}( u )}$. 
+If $j = \jPrime$, then $\M( t )$ is $\alpha$-equivalent with $\Val^\AST \ec{u} \interp{w_k}$, which is equal to $\Val^\AST \ec{ \const s^k_{\negvthinspace j}( u )}$,
 which is $\Val^\AST \ec{t}$.
-If $j \neq j'$, we are free to interpret $\const s^k_{\negvthinspace j}( \M( u ) )$ as the value in $\M( \tau )$ that is $\alpha$-equivalent with $\Val^\AST \ec{ t }$.
+If $j \neq \jPrime$, we are free to interpret $\const s^k_{\negvthinspace j}( \M( u ) )$ as the value in $\M( \tau )$ that is $\alpha$-equivalent with $\Val^\AST \ec{ t }$.
 Since \rn{Cong} does not apply and due to Lemma~\ref{lem:model-completion}.3,
@@
 
-\paragraph{Optimizations}
+\paragraph{Optimizations.}
 We first discuss several optimizations that are not reflected in the presentation of the calculus, which closely follow the approach outlined by Barrett et al in~\cite{}.
@@
 replace $t \teq s$ by $s \teq t$ if $s$ is less than $t$ based on a term ordering,
-replace $\const{C}_{\negvthinspace j}( \vec t ) \teq \const{C}_{j'}( \vec u )$ with $\bot$ when $j \neq j'$,
+replace $\const{C}_{\negvthinspace j}( \vec t ) \teq \const{C}_{\vthinspace\jPrime}( \vec u )$ with $\bot$ when $j \neq \jPrime$,
 replace all selector terms of the form $\const{s}^{\,k}_{\negvthinspace j}( \const{C}_{\negvthinspace j}(t_1,\ldots,t_n) )$ by $t_k$,
-and replace all occurrences of discriminators $\const{d}_{\negvthinspace j}( \const{C}_{j'}( \vec t ) )$ by $\top$ if $j=j'$ and $\bot$ otherwise.
-
-\paragraph{Integration into DPLL(T)}
+and replace all occurrences of discriminators $\const{d}_{\negvthinspace j}( \const{C}_{\vthinspace\jPrime}( \vec t ) )$ by $\top$ if $j = \jPrime$ and $\bot$ otherwise.
+
+\paragraph{Integration into DPLL(T).}
 The calculus is implemented as \emph{theory solver} of CVC4,
@@
 Given a theory $T = T_1 \mathrel\cup \ldots T_n$ and a input set of clauses $F$ in CNF,
-the DPLL(T) \rem{CDCL(T)?} procedure
+the DPLL($T$) \rem{CDCL($T$)?} procedure
 (incrementally) builds partial assignments from the atomic formulas of $F$ to truth values such that no clause in $F$ is falsified.

### 2014-12-22 tex/codatadp.tex
more

@@
 \makeatother
+
+%\newcommand\PART[1]{.#1}
+\newcommand\PART[1]{(#1)}
 
@@
 If $\Val$ is constructed for a saturated set $\Ec$
-and $\Val^\AST$ is a completion of $\Val$ for normal model $\M$, the following holds:
+and $\Val^\AST$ is a completion of $\Val$ for normal model $\M$, the following properties hold:
 \begin{enumerate}
-\item[\rm 1.] $\Val^\AST \ec{x}$ is $\alpha$-equivalent to a value in $\M( \tau )$, where $\ec{x}$ has type $\tau$.
-\item[\rm 2.]
+\item[\rm (1)] $\Val^\AST \ec{x}$ is $\alpha$-equivalent to a value in $\M( \tau )$, where $\ec{x}$ has type $\tau$.
+\item[\rm (2)]
 $\Val^\AST \ec{x} = \Val^\AST \ec{y} \interp{ t }$ for
 all subterms $t$ of $\Val^\AST \ec{y}$ with $\muvar(t) = {\Varec{x}}$. 
-\item[\rm 3.]
+\item[\rm (3)]
 $\Val^\AST \ec{x} \vsim \Val^\AST \ec{y}$ if and only if $\ec{x} = \ec{y}$.
@@
 For all equivalence classes $\ec{z}$,
-each pair of constructor terms $\const{C}_{\negvthinspace j}( \vec t )$ and $\const{C}_{\vthinspace\jPrime}( \vec u )$ in $\ec{z}$
+each pair of constructor terms $\const{C}_{\negvthinspace j}( \vec t )$ and $\const{C}_{\negvthinspace\jPrime}( \vec u )$ in $\ec{z}$
 are such that $j = \jPrime$ since \rn{Clash} does not apply to $\Ec$,
@@
 which by the induction hypothesis is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \Val^\AST \ec{ u_1 }, \ldots, \Val^\AST \ec{ u_n } )}$, call this term $t_u$.
-%By Lemma~\ref{lem:model-completion}.2 and due to our construction of $\Val^\AST$, we have $\Val^\AST \ec{t}$ is a term
+%By Lemma~\ref{lem:model-completion}\PART{2} and due to our construction of $\Val^\AST$, we have $\Val^\AST \ec{t}$ is a term
 %of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$
@@
 where for each $i = 1, \ldots, n$, 
-$\muvar(w_i) = {\Varec{u_i}}$ and thus by Lemma~\ref{lem:model-completion}.2, 
+$\muvar(w_i) = {\Varec{u_i}}$ and thus by Lemma~\ref{lem:model-completion}\PART{2}, 
 $\Val^\AST \ec{t}\interp{w_i} = \Val^\AST \ec{u_i}$.
@@
 since \rn{Split} does not apply to $\Fc$,
-$\ec{u}$ must contain a term of the form $\const C_{\vthinspace\jPrime}\bigl( \const s^1_{\vthinspace\jPrime}( u ), \ldots, \const s^{n}_{\vthinspace\jPrime}( u ) \bigr)$ for some $\jPrime$.
-%By Lemma~\ref{lem:model-completion}.2 and due to our construction of $\Val^\AST$, we have $\Val^\AST \ec{u}$ is of the form
-%$\MU {\Varec{u}}.\; \const C_{\vthinspace\jPrime}( w_1, \ldots, w_n )$,
-%where $\Val^\AST \ec{u} \interp{ w_i } = \Val^\AST \ec{ \const s^i_{\vthinspace\jPrime}( u )}$ for $i = 1, \ldots, n$. 
+$\ec{u}$ must contain a term of the form $\const C_{\negvthinspace\jPrime}\bigl( \const s^1_{\negvthinspace\jPrime}( u ), \ldots, \const s^{n}_{\negvthinspace\jPrime}( u ) \bigr)$ for some $\jPrime$.
+%By Lemma~\ref{lem:model-completion}\PART{2} and due to our construction of $\Val^\AST$, we have $\Val^\AST \ec{u}$ is of the form
+%$\MU {\Varec{u}}.\; \const C_{\negvthinspace\jPrime}( w_1, \ldots, w_n )$,
+%where $\Val^\AST \ec{u} \interp{ w_i } = \Val^\AST \ec{ \const s^i_{\negvthinspace\jPrime}( u )}$ for $i = 1, \ldots, n$. 
 Since \rn{Inject} and \rn{Clash} do not apply to $\Fc$, 
-by our construction of $\Val^\AST$ we have $\Val^\AST \ec{u}$ is of the form
-$\MU {\Varec{u}}.\; \const C_{\vthinspace\jPrime}( w_1, \ldots, w_n )$,
+by construction $\Val^\AST$ must be of the form
+$\MU {\Varec{u}}.\; \const C_{\negvthinspace\jPrime}( w_1, \ldots, w_n )$,
 where for each $i = 1, \ldots, n$,
-$\muvar(w_i) = {\Varec{\const s^i_{\vthinspace\jPrime}( u )}}$ and thus by Lemma~\ref{lem:model-completion}.2, 
-$\Val^\AST \ec{u} \interp{ w_i } = \Val^\AST \ec{ \const s^i_{\vthinspace\jPrime}( u )}$. 
-If $j = \jPrime$, then $\M( t )$ is $\alpha$-equivalent with $\Val^\AST \ec{u} \interp{w_k}$, which is equal to $\Val^\AST \ec{ \const s^k_{\negvthinspace j}( u )}$,
+$\muvar(w_i) = \smash{\Varec{\vphantom{I}\smash{\const s^i_{\negvthinspace\jPrime}( u )}}}$ and thus by Lemma~\ref{lem:model-completion}\PART{2}, 
+$\Val^\AST \ec{u} \interp{ w_i } = \Val^\AST \ec{ \smash{\const s^i_{\negvthinspace\jPrime}( u )}}$. 
+If $j = \jPrime$, then $\M( t )$ is $\alpha$-equivalent with $\Val^\AST \ec{u} \interp{w_k}$, which is equal to $\Val^\AST \ec{ \smash{ \const s^k_{\negvthinspace j}( u )}}$,
 which is $\Val^\AST \ec{t}$.
-If $j \neq \jPrime$, we are free to interpret $\const s^k_{\negvthinspace j}( \M( u ) )$ as the value in $\M( \tau )$ that is $\alpha$-equivalent with $\Val^\AST \ec{ t }$.
-Since \rn{Cong} does not apply and due to Lemma~\ref{lem:model-completion}.3,
+If $j \neq \jPrime$, we are free to interpret $\smash{\const s^k_{\negvthinspace j}( \M( u ) )}$ as the value in $\M( \tau )$ that is $\alpha$-equivalent with $\Val^\AST \ec{ t }$.
+Since \rn{Cong} does not apply and due to Lemma~\ref{lem:model-completion}\PART{3},
 pairs of wrongly applied selectors are assigned the same value if and only if they reside in the same equivalence class,
@@
 Since $\M( t_1 ) \vsim \Val^\AST \ec{t_1}$ and $\M( t_2 ) \vsim \Val^\AST \ec{t_2}$,
-by Lemma~\ref{lem:model-completion}.3, $\M( t_1 ) \neq \M( t_2 )$, and thus $\M \models t_1 \tneq t_2$.
+by Lemma~\ref{lem:model-completion}\PART{3}, $\M( t_1 ) \neq \M( t_2 )$, and thus $\M \models t_1 \tneq t_2$.
 
@@
 replace $t \teq s$ by $s \teq t$ if $s$ is less than $t$ based on a term ordering,
-replace $\const{C}_{\negvthinspace j}( \vec t ) \teq \const{C}_{\vthinspace\jPrime}( \vec u )$ with $\bot$ when $j \neq \jPrime$,
+replace $\const{C}_{\negvthinspace j}( \vec t ) \teq \const{C}_{\negvthinspace\jPrime}( \vec u )$ with $\bot$ when $j \neq \jPrime$,
 replace all selector terms of the form $\const{s}^{\,k}_{\negvthinspace j}( \const{C}_{\negvthinspace j}(t_1,\ldots,t_n) )$ by $t_k$,
-and replace all occurrences of discriminators $\const{d}_{\negvthinspace j}( \const{C}_{\vthinspace\jPrime}( \vec t ) )$ by $\top$ if $j = \jPrime$ and $\bot$ otherwise.
+and replace all occurrences of discriminators $\const{d}_{\negvthinspace j}( \const{C}_{\negvthinspace\jPrime}( \vec t ) )$ by $\top$ if $j = \jPrime$ and $\bot$ otherwise.
 

### 2014-12-22 tex/codatadp.tex
more

@@
 Given the mapping $\Val$, the \rn{Acyclic} and \rn{Unique} rules can be stated as follows.
-For acyclicity, if there exists an equivalence class $\ec{t}$
-where $t$'s type is a datatype and
+For acyclicity, if $\ec{t : \delta}$ is an equivalence class such that $\delta$ is a datatype and
 whose %class of  -- "whose" already refers to "equivalence class"
@@
 then $\Ec$ is unsatisfiable.
-For uniqueness, if there exist two equivalence classes $\ec{t}$ and $\ec{u}$
-where $t$ and $u$'s type is a codatatype and
+For uniqueness, if $\ec{t : \delta}$ and $\ec{u : \delta}$ are two equivalence classes such
+that $\delta$ is a codatatype and
 whose %classes of  -- "whose" already refers to "equivalence classes"
@@
 In the following lemma about $\Val^\AST$,
-$\muvar( t )$ denotes $t$ if $t$ is a variable and ${\Varec{x}}$ if $t$ is of the form $\MU {\Varec{x}}.\; u$.
+$\muvar( t ) = \begin{cases}
+  x & \text{if $t = \MU {x}.\; u$} \\[-\jot]
+  t & \text{otherwise ($t$ is a variable)}
+\end{cases}$
 

### 2014-12-22 tex/codatadp.tex
more

@@
 \makeatother
+
+\newcommand\typ[1]{: #1}
 
@@
 which collapses to $\const{f} : \tau$ if $n = 0$\end{report}.
-For a term $t$, the notation $t : \tau$ indicates that it has type $\tau$.
+For a term $t$, the notation $t\typ{\tau}$ indicates that it has type $\tau$.
 Functions are invoked in the standard way, with $\const{f}(t_1,\ldots,t_n)$
@@
 \inferrule{
-  t : \delta
+  t\typ{\delta}
   \quad
@@
 \inferrule{
- t, u : \delta
+ t, u\typ{\delta}
  \quad
@@
 Given the mapping $\Val$, the \rn{Acyclic} and \rn{Unique} rules can be stated as follows.
-For acyclicity, if $\ec{t : \delta}$ is an equivalence class such that $\delta$ is a datatype and
+For acyclicity, if $\ec{t\typ{\delta}}$ is an equivalence class such that $\delta$ is a datatype and
 whose %class of  -- "whose" already refers to "equivalence class"
@@
 then $\Ec$ is unsatisfiable.
-For uniqueness, if $\ec{t : \delta}$ and $\ec{u : \delta}$ are two equivalence classes such
+For uniqueness, if $\ec{t\typ{\delta}}$ and $\ec{u\typ{\delta}}$ are two equivalence classes such
 that $\delta$ is a codatatype and
@@
 \inferrule{
-  t : \delta \in \tEc
+  t\typ{\delta} \in \tEc
   \quad
@@
 \inferrule{
-  t, u : \delta \in \tEc
+  t, u\typ{\delta} \in \tEc
   \quad
@@
 \smash{\SSSS 0_t} & = \{ t \}
-& \smash{\SSSS {i+1}_t} & = \smash{\SSSS i_t \mathrel\cup \{ \const s( u ) \mid u : \delta \in \SSSS i{,}~ \delta \in \Data{,}~ \left|\delta\right| \text{ is finite}{,}~ \const{s} \in \Sel^{\delta}}  \}
+& \smash{\SSSS {i+1}_t} & = \smash{\SSSS i_t \mathrel\cup \{ \const s( u ) \mid u\typ{\delta} \in \SSSS i{,}~ \delta \in \Data{,}~ \left|\delta\right| \text{ is finite}{,}~ \const{s} \in \Sel^{\delta}}  \}
 \end{align*}
@@
 Let $\SSSS \AST$ be the union of all sets $\SSSS \AST_t$ where $t \in T$,
-and let $\tcEc$ be the set of subterms of $\Ec \mathrel\cup \{ \const C_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\smash{\,n_{\negvthinspace j}}}_{\negvthinspace j}( t ) \bigr) \mid t : \delta \in \SSS \mathrel\cup \SSSS \AST{,}~ \const C_{\negvthinspace j} \in \Ctr^\delta \}$.
+and let $\tcEc$ be the set of subterms of $\Ec \mathrel\cup \{ \const C_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\smash{\,n_{\negvthinspace j}}}_{\negvthinspace j}( t ) \bigr) \mid t\typ{\delta} \in \SSS \mathrel\cup \SSSS \AST{,}~ \const C_{\negvthinspace j} \in \Ctr^\delta \}$.
 In a derivation tree with root node~$\Ec$,
@@
 If the children of $\Ec$ are closed derivation trees
-whose root nodes are the result of applying the rule \rn{Split} on $t : \delta$,
+whose root nodes are the result of applying the rule \rn{Split} on $t\typ{\delta}$,
 by the induction hypothesis $\Ec \mathrel\cup t \teq \smash{\const C_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\,\smash{n_{\negvthinspace j}}}_{\negvthinspace j}( t ) \bigr)}$ is unsatisfiable
@@
 $\MU x.\; \const{C}( x, x ) \not\vsimv{y} \MU y.\; \const{C}( x, y )$.
-For a variable $x : \tau$ and a fixed normal model $\M$,
+For a variable $x\typ{\tau}$ and a fixed normal model $\M$,
 we write $\mathcal{T}^\M_x( \Val )$ to denote the set consisting of all values $v$ from $\M( \tau )$
@@
 \inferrule{
-  {\Varec{x}} : \tau \in \FV( \Val )
+  {\Varec{x}}\typ{\tau} \in \FV( \Val )
   \quad

### 2014-12-22 tex/codatadp.tex
trying more notations

@@
 
-\newcommand\typ[1]{: #1}
+%\newcommand\typ[1]{: #1}
+\newcommand\typ[1]{^{\vthinspace #1}}
 
@@
 %%% before by other authors.
-\newcommand{\thD}{\mathcal{CoDt}}
+%\newcommand{\thD}{\mathcal{CoDt}}
+%\newcommand{\thD}{\mathcal{TDC}}
+\newcommand{\thD}{\mathcal{DC}}
 
@@
 We now define the class of interpretations for $\thD$.
-In the following, $\M( \tau )$ denotes the interpretation type $\tau$ in $\M$,
+$\M( \tau )$ denotes the interpretation type $\tau$ in $\M$,
 that is, a set of domain elements for that type.
@@
 that is, an element of $\M( \tau )$ where $\tau$ is the type of $t$.
-In the following,
-\emph{all} types are interpretated as a set of $\mu$-terms.
+\emph{All} types are interpreted as a set of $\mu$-terms.
 This is done to simplify the presentation.
@@
 
+%%% TYPESETTING: Evil hack!!
 %We first state the following properties of $\Val^\AST$,
 In the following lemma about $\Val^\AST$,
-$\muvar( t ) = \begin{cases}
-  x & \text{if $t = \MU {x}.\; u$} \\[-\jot]
-  t & \text{otherwise ($t$ is a variable)}
-\end{cases}$
+$\muvar( t ) = \smash{\begin{cases}
+  x & \text{if $t$ is of the form $\MU {x}.\; u$} \\[-\jot]
+  t & \text{otherwise.}
+\end{cases}}$
 
@@
 If $\Val$ is constructed for a saturated set $\Ec$
-and $\Val^\AST$ is a completion of $\Val$ for normal model $\M$, the following properties hold:
+and $\Val^\AST$ is a completion of $\Val$ for a normal model $\M$, the following properties hold:
 \begin{enumerate}

### 2014-12-22 tex/codatadp.tex
made \interp binary

@@
 \newcommand{\ttpath}[3]{\mathcal{P}_{ #2 \rightarrow #3 }( #1 )}
-\newcommand{\interp}[1]{\llbracket #1 \rrbracket}
+\newcommand{\interp}[2]{#1\llbracket #2 \rrbracket}
 \newcommand{\dpath}[3]{\delta^{#1}_{#2}( #3 )}
@@
 For \rn{Single}, $t$~and~$u$ must have the same value since the cardinality of their type is $1$.
-By the induction hypothesis, we have $\Ec \mathrel\cup t \teq u$ is $\thD$-unsatisfiable,
+By the induction hypothesis, $\Ec \mathrel\cup t \teq u$ is $\thD$-unsatisfiable
 and thus $\Ec$ is $\thD$-unsatisfiable.
@@
 For a $\mu$-term $t$ with subterm $u$,
-the \emph{interpretation of\/ $u$ in\/ $t$} is the $\mu$-term $t\interp{u}^\emptyset$ as returned by the following recursive procedure,
-where $x \mapsto \MU x.\; \const{C}( \vec u ) \in t$ indicates that $\MU x.\; \const{C}( \vec u )$ is the subterm of $t$ that binds this occurrence of variable $x$:
+the \emph{interpretation of $u$ in $t$} is the $\mu$-term $\interp{t}{u}^\emptyset$ as returned by the following recursive procedure,
+where $x \leadsto \MU x.\; \const{C}( \vec u ) \in t$ indicates that $\MU x.\;
+\const{C}( \vec u )$ is the subterm of $t$ that binds this occurrence of
+variable $x$:
 \[\begin{array}{r@{}c@{}l}
-t\interp{x}^M & {} \,=\, {} &
+\interp{t}{x}^M & {} \,=\, {} &
 \textstyle\begin{cases}
-    \MU x.\; \const{C}( t\interp{ \vec u}^{M \mathrel\cup \{ x \}}) & \text{if }x \not\in M, x \mapsto \MU x.\; \const{C}( \vec u ) \in t  \\[-\jot]
+    \MU x.\; \const{C}( \interp{t}{ \vec u}^{M \mathrel\cup \{ x \}}) & \text{if }x \not\in M\text{~and $x$ is
+    bound by $\MU x.\;
+\const{C}( \vec u )$}  \\[-\jot]
     x & \text{otherwise}
 \end{cases} \\
-t\interp{\MU x.\; \const{C}( \vec u )}^M & = &
+\interp{t}{\MU x.\; \const{C}( \vec u )}^M & = &
 \textstyle\begin{cases}
-    \MU x.\; \const{C}( t\interp{ \vec u }^{M \mathrel\cup \{ x \}} ) & \text{if }x \not\in M \\[-\jot]
+    \MU x.\; \const{C}( \interp{t}{ \vec u }^{M \mathrel\cup \{ x \}} ) & \text{if }x \not\in M \\[-\jot]
     x & \text{otherwise}
@@
 \end{array}\]
-We write $t\interp{u}$ to abbreviate $t\interp{u}^\emptyset$.
+We write $\interp{t}{u}$ to abbreviate $\interp{t}{u}^\emptyset$.
 
@@
 we say $u$ is a \emph{bisimilar subterm} of $t$ if it is of the form
-$\MU y.\; \const{C}( u_1, \ldots, u_n )$ and $t\interp{ t_k } \aequiv t\interp{ u_k }$ for $k = 1,\ldots, n$.
+$\MU y.\; \const{C}( u_1, \ldots, u_n )$ and $\interp{t}{ t_k } \aequiv \interp{t}{ u_k }$ for $k = 1,\ldots, n$.
 The term $t$ is \emph{normal} if it does not contain a strict, bisimilar subterm
@@
 that is,
-$t\interp{\MU y.\; \const{C}( y )}^\emptyset =$
-$\MU y.\; \const{C}( t\interp{y}^{\{y\}} ) =$
+$\interp{t}{\MU y.\; \const{C}( y )}^\emptyset =$
+$\MU y.\; \const{C}( \interp{t}{y}^{\{y\}} ) =$
 $\MU y.\; \const{C}( y )$
 is $\alpha$-equivalent to
-$t\interp{y}^\emptyset =$
-$\MU y.\; \const{C}( t\interp{y}^{\{y\}} ) =$
+$\interp{t}{y}^\emptyset =$
+$\MU y.\; \const{C}( \interp{t}{y}^{\{y\}} ) =$
 $\MU y.\; \const{C}( y )$.
@@
 noting that
-$t\interp{\MU y.\; \const{C}( x )}^\emptyset =
-\MU y.\; \const{C}( t\interp{x}^{\{y\}} ) =
-\MU y.\; \const{C}( t\interp{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) ) }^{\{y\}} ) =
-\MU y.\; \const{C}( \MU x.\; \const{C}( t\interp{\MU y.\; \const{C}( x )}^{\{x,y\}} ) ) =
+$\interp{t}{\MU y.\; \const{C}( x )}^\emptyset =
+\MU y.\; \const{C}( \interp{t}{x}^{\{y\}} ) =
+\MU y.\; \const{C}( \interp{t}{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) ) }^{\{y\}} ) =
+\MU y.\; \const{C}( \MU x.\; \const{C}( \interp{t}{\MU y.\; \const{C}( x )}^{\{x,y\}} ) ) =
 \MU y.\allowbreak\; \const{C}( \MU x.\; \const{C}( y ) )$
-is $\alpha$-equivalent to $t\interp{x}^\emptyset = t$.
+is $\alpha$-equivalent to $\interp{t}{x}^\emptyset = t$.
 % for the renaming $\{ x \mapsto y, y \mapsto x \}$.
@@
 If $\vec u$ are in normal form and $t = \nf{\MU x.\; \const{C}( \vec u )} = \MU x.\; \const{C}( \vec w )$,
-then $\vec u \vsim t\interp{\vec w}$.
+then $\vec u \vsim \interp{t}{\vec w}$.
 \end{lemma}
@@
 \label{lem:mu-norm-interp}
-If $t$ is normal and has subterm $u$, then $t \interp{u}$ is normal.
+If $t$ is normal and has subterm $u$, then $\interp{t}{u}$ is normal.
 \end{lemma}
@@
 %If $t_u = \MU x.\; \const{C}( \vec u )$ and $t_w = \MU y.\; \const{C}( \vec w )$ are in normal form,
-%then $t_u\interp{ \vec u } \vsim t_w\interp{ \vec w }$ if and only if $t_u \vsim t_w$.
+%then $\interp{t_u}{ \vec u } \vsim \interp{t_w}{ \vec w }$ if and only if $t_u \vsim t_w$.
 %\end{lemma}
@@
 in $\M( \tau )$ that is $\alpha$-equivalent to
-$\M( t )\interp{u_k}$
+$\interp{\M( t )}{u_k}$
 when $\M( t )$ is of the form $\MU x.\; \const{C}_{\negvthinspace j}( \vec u )$.
@@
 we write $\mathcal{T}^\M_x( \Val )$ to denote the set consisting of all values $v$ from $\M( \tau )$
-such that $v \vsimv{x} t\interp{u}$ for some term $t$ with subterm $u$ occurring in the range of $\Val$.
+such that $v \vsimv{x} \interp{t}{u}$ for some term $t$ with subterm $u$ occurring in the range of $\Val$.
 
@@
 \item[\rm (2)]
-$\Val^\AST \ec{x} = \Val^\AST \ec{y} \interp{ t }$ for
+$\Val^\AST \ec{x} = \interp{\Val^\AST \ec{y}}{ t }$ for
 all subterms $t$ of $\Val^\AST \ec{y}$ with $\muvar(t) = {\Varec{x}}$. 
@@
 Moreover, free variables ${\Varec{z}}$ in $t$ that are bound in $\Val \ec{y}$ are interpreted in 
-the expansion of $\Val \ec{y} \interp{ t }$ as a term
+the expansion of $\interp{\Val \ec{y}}{ t }$ as a term
 constructed by a sequence of substitutions of the form mentioned above to ${\Varec{z}}$. \rem{more detail}
 Thus, we have
-$\Val \ec{y} \interp{ t } = \Val \ec{x} \interp{ \Val \ec{x} } = \Val \ec{x}$, and thus 2 holds for $\Val$.
+$\interp{\Val \ec{y}}{ t } = \interp{\Val \ec{x}}{ \Val \ec{x} } = \Val \ec{x}$, and thus 2 holds for $\Val$.
 Statement 3 holds for $\Val$ since \rn{Unique} does not apply.
@@
 contains a subterm of the form $\MU {\Varec{y}}.\; \const{C}( u_1, \ldots, u_n )$
-where $\Val \ec{x} \interp{ t_i } \aequiv \Val \ec{x} \interp{ u_i }$ for $i = 1, \ldots, n$.
+where $\interp{\Val \ec{x}}{ t_i } \aequiv \interp{\Val \ec{x} }{ u_i }$ for $i = 1, \ldots, n$.
 Due to our construction of $\Val$, $\ec{y}$ contains a constructor $\const{C}( w_1, \ldots, w_n )$ and $\muvar( u_i ) = w_i$ for $i = 1, \ldots, n$.
@@
 we have $\ec{w_{\negvthinspace j}}$ and $\ec{z_{\negvthinspace j}}$ are distinct for some $j$.
-By statement 2, $\Val \ec{x} \interp{ t_{\negvthinspace j} } = \Val \ec{z_{\negvthinspace j}}$
-and $\Val \ec{x} \interp{ u_{\negvthinspace j} } = \Val \ec{w_{\negvthinspace j}}$,
+By statement 2, $\interp{\Val \ec{x}}{ t_{\negvthinspace j} } = \Val \ec{z_{\negvthinspace j}}$
+and $\interp{\Val \ec{x}}{ u_{\negvthinspace j} } = \Val \ec{w_{\negvthinspace j}}$,
 which are not $\alpha$-equivalent by statement 3, contradicting the fact that $\MU {\Varec{y}}.\; \const{C}( u_1, \ldots, u_n )$ is a bisimilar subterm of $\Val \ec{x}$.
@@
 To show 2,
-by assumption of 2 on $\Val_1$, we have $\Val_1 \ec{y} \interp{ u } = \Val_1 \ec{x}$ for
+by assumption of 2 on $\Val_1$, we have $\interp{\Val_1 \ec{y}}{ u } = \Val_1 \ec{x}$ for
 all subterms $u$ of $\Val_1 \ec{y}$ where $\muvar(t) = {\Varec{x}}$.
-Thus, $\Val_1 \sigma \ec{y} \interp{ u } = \Val_1 \sigma \ec{x}$ and
-$\nf{(\Val_1 \sigma)} \ec{y} \interp{ u } = \nf{(\Val_1 \sigma)} \ec{x}$.
+Thus, $\interp{\Val_1 \sigma \ec{y}}{ u } = \Val_1 \sigma \ec{x}$ and
+$\interp{\nf{(\Val_1 \sigma)} \ec{y}}{ u } = \nf{(\Val_1 \sigma)} \ec{x}$.
 To show 3, 
@@
 are $\alpha$-disequivalent at position $p$ since they do not contain the same free variables.
-Since $\Val_1 \ec{y} \interp{t_1}$ is normal,
+Since $\interp{\Val_1 \ec{y}}{t_1}$ is normal,
 there is a closed $\mu$-term $v \in \mathcal{T}^\M_{\Varec{x}}( \Val_1 )$
-such that $v \vsimv{\Varec{x}} \Val_1 \ec{y} \interp{t_1}$.
+such that $v \vsimv{\Varec{x}} \interp{\Val_1 \ec{y}}{t_1}$.
 Thus, by assumption on our selection of $\MU {\Varec{x}}.\; t$,
-we have $\MU {\Varec{x}}.\; t \not\vsimv{\Varec{x}} \Val_1 \ec{y} \interp{t_1}$,
+we have $\MU {\Varec{x}}.\; t \not\vsimv{\Varec{x}} \interp{\Val_1 \ec{y}}{t_1}$,
 which implies that the expansion of $\nf{(\Val_1 \sigma)} \ec{y}$ and $\nf{(\Val_1 \sigma)} \ec{z}$
@@
 %of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$
-%where $\Val^\AST \ec{t}\interp{w_i} = \Val^\AST \ec{u_i}$ for $i = 1, \ldots, n$.
+%where $\interp{\Val^\AST \ec{t}}{w_i} = \Val^\AST \ec{u_i}$ for $i = 1, \ldots, n$.
 Since \rn{Inject} and \rn{Clash} do not apply to $\Fc$, 
@@
 $\muvar(w_i) = {\Varec{u_i}}$ and thus by Lemma~\ref{lem:model-completion}\PART{2}, 
-$\Val^\AST \ec{t}\interp{w_i} = \Val^\AST \ec{u_i}$.
+$\interp{\Val^\AST \ec{t}}{w_i} = \Val^\AST \ec{u_i}$.
 For each $i = 1, \ldots, n$, let $u_\iPrime$ be the $i$\vvthinspace th argument of $t_u$.
-By Lemma~\ref{lem:mu-norm-arg}, $t_u\interp{u_\iPrime} \vsim \Val^\AST \ec{u_i}$.
-Thus $t_u\interp{u_\iPrime} \vsim \Val^\AST \ec{t}\interp{ w_i }$. % for each $i = 1, \ldots, n$.
+By Lemma~\ref{lem:mu-norm-arg}, $\interp{t_u}{u_\iPrime} \vsim \Val^\AST \ec{u_i}$.
+Thus $\interp{t_u}{u_\iPrime} \vsim \interp{\Val^\AST \ec{t}}{ w_i }$. % for each $i = 1, \ldots, n$.
 %By Lemma~\ref{lem:mu-cong}, we have $t_u \vsim t_w$.
@@
 %$\MU {\Varec{u}}.\; \const C_{\negvthinspace\jPrime}( w_1, \ldots, w_n )$,
-%where $\Val^\AST \ec{u} \interp{ w_i } = \Val^\AST \ec{ \const s^i_{\negvthinspace\jPrime}( u )}$ for $i = 1, \ldots, n$. 
+%where $\interp{\Val^\AST \ec{u} }{ w_i } = \Val^\AST \ec{ \const s^i_{\negvthinspace\jPrime}( u )}$ for $i = 1, \ldots, n$. 
 Since \rn{Inject} and \rn{Clash} do not apply to $\Fc$, 
@@
 $\muvar(w_i) = \smash{\Varec{\vphantom{I}\smash{\const s^i_{\negvthinspace\jPrime}( u )}}}$ and thus by Lemma~\ref{lem:model-completion}\PART{2}, 
-$\Val^\AST \ec{u} \interp{ w_i } = \Val^\AST \ec{ \smash{\const s^i_{\negvthinspace\jPrime}( u )}}$. 
-If $j = \jPrime$, then $\M( t )$ is $\alpha$-equivalent with $\Val^\AST \ec{u} \interp{w_k}$, which is equal to $\Val^\AST \ec{ \smash{ \const s^k_{\negvthinspace j}( u )}}$,
+$\interp{\Val^\AST \ec{u}}{ w_i } = \Val^\AST \ec{ \smash{\const s^i_{\negvthinspace\jPrime}( u )}}$. 
+If $j = \jPrime$, then $\M( t )$ is $\alpha$-equivalent with $\interp{\Val^\AST \ec{u}}{w_k}$, which is equal to $\Val^\AST \ec{ \smash{ \const s^k_{\negvthinspace j}( u )}}$,
 which is $\Val^\AST \ec{t}$.

### 2014-12-22 tex/codatadp.tex
more

@@
 In the following lemma about $\Val^\AST$,
-$\muvar( t ) = \smash{\begin{cases}
+$\vphantom{\Bigl(_j}\muvar( t ) = \smash{\begin{cases}
   x & \text{if $t$ is of the form $\MU {x}.\; u$} \\[-\jot]

### 2014-12-22 tex/codatadp.tex
more

@@
 
-We now show that $\Val = \nf{(\Val)}$.
-Assume by contradition $\Val \ec{x} \neq \nf{(\Val \ec{x})}$ for some $\Val \ec{x}$ of minimal size.
+We now show that $\Val = \nf{\Val}$.
+Assume by contradition $\Val \ec{x} \neq \nf{\Val \ec{x}}$ for some $\Val \ec{x}$ of minimal size.
 We have that $\Val \ec{x}$ is of the form $\MU {\Varec{x}}.\; \const{C}( t_1, \ldots, t_n )$.
@@
 which are not $\alpha$-equivalent by statement 3, contradicting the fact that $\MU {\Varec{y}}.\; \const{C}( u_1, \ldots, u_n )$ is a bisimilar subterm of $\Val \ec{x}$.
-Thus, $\Val = \nf{(\Val)}$, and statements 2 and 3 hold for $\nf{(\Val)}$.
+Thus, $\Val = \nf{\Val}$, and statements 2 and 3 hold for $\nf{\Val}$.
 
 We now show that if statements 2 and 3 hold for some $\Val_1$,
-then 2 and 3 also hold for $\nf{(\Val_1 \sigma)}$,
+then 2 and 3 also hold for $\nf{\Val_1 \sigma}$,
 where $\sigma$ is a substitution of the form $\{ {\Varec{x}} \mapsto \MU{\Varec{x}}.\; t \}$,
@@
 Thus, $\interp{\Val_1 \sigma \ec{y}}{ u } = \Val_1 \sigma \ec{x}$ and
-$\interp{\nf{(\Val_1 \sigma)} \ec{y}}{ u } = \nf{(\Val_1 \sigma)} \ec{x}$.
+$\interp{\nf{\Val_1 \sigma} \ec{y}}{ u } = \nf{\Val_1 \sigma} \ec{x}$.
 To show 3, 
 consider two distinct equivalence classes $\ec{y}$ and $\ec{z}$,
-and assume by contradiction that $\nf{(\Val_1 \sigma)} \ec{y} \aequiv \nf{(\Val_1 \sigma)} \ec{z}$.
+and assume by contradiction that $\nf{\Val_1 \sigma} \ec{y} \aequiv \nf{\Val_1 \sigma} \ec{z}$.
 Due to statement 3 for $\Val_1$,
@@
 and $t_1 \not\aequiv t_2$.
-%We first show that $\nf{(\Val_1 \sigma)} \ec{y}$ (and for the same reasons, $\nf{(\Val_1 \sigma)} \ec{z}$)
+%We first show that $\nf{\Val_1 \sigma} \ec{y}$ (and for the same reasons, $\nf{\Val_1 \sigma} \ec{z}$)
 %also contains position $p$.
@@
 If $t_1$ (resp. $t_2$) is a free variable that is not ${\Varec{x}}$,
-then $\nf{(\Val_1 \sigma)} \ec{y}$ (resp. $\nf{(\Val_1 \sigma)} \ec{z}$)
+then $\nf{\Val_1 \sigma} \ec{y}$ (resp. $\nf{\Val_1 \sigma} \ec{z}$)
 contains $t_1$ (resp. $t_2$) at position $p$,
-and $\nf{(\Val_1 \sigma)} \ec{z}$ (resp. $\nf{(\Val_1 \sigma)} \ec{y}$) does not.
+and $\nf{\Val_1 \sigma} \ec{z}$ (resp. $\nf{\Val_1 \sigma} \ec{y}$) does not.
 If $t_1$ is of the form $\MU w_1.\; \const{C}( \vec t )$ 
 and $t_2$ is of the form $\MU w_2.\; \const{D}( \vec u )$,
-then the expansion \rem{define} of $\nf{(\Val_1 \sigma)} \ec{y}$ and $\nf{(\Val_1 \sigma)} \ec{z}$
+then the expansion \rem{define} of $\nf{\Val_1 \sigma} \ec{y}$ and $\nf{\Val_1 \sigma} \ec{z}$
 are $\alpha$-disequivalent at position $p$.
@@
 we have that $FV( t_1 ) \subseteq \{ {\Varec{x}} \}$,
-or else the expansion of $\nf{(\Val_1 \sigma)} \ec{y}$ and $\nf{(\Val_1 \sigma)} \ec{z}$
+or else the expansion of $\nf{\Val_1 \sigma} \ec{y}$ and $\nf{\Val_1 \sigma} \ec{z}$
 are $\alpha$-disequivalent at position $p$ since they do not contain the same free variables.
@@
 we have $\MU {\Varec{x}}.\; t \not\vsimv{\Varec{x}} \interp{\Val_1 \ec{y}}{t_1}$,
-which implies that the expansion of $\nf{(\Val_1 \sigma)} \ec{y}$ and $\nf{(\Val_1 \sigma)} \ec{z}$
+which implies that the expansion of $\nf{\Val_1 \sigma} \ec{y}$ and $\nf{\Val_1 \sigma} \ec{z}$
 are $\alpha$-disequivalent at position $p$. \rem{more detail}

### 2014-12-22 tex/codatadp.tex
more

@@
 \newcommand{\ttpath}[3]{\mathcal{P}_{ #2 \rightarrow #3 }( #1 )}
-\newcommand{\interp}[2]{#1\llbracket #2 \rrbracket}
+%\newcommand{\interp}[2]{#1\llbracket #2 \rrbracket}
+%%% @ANDY: I prefer a notation that puts more emphasis on #2 (which is, in essence,
+%%% the object that is returned) and less on #1
+%\newcommand{\interp}[2]{\llbracket #2 \rrbracket_{#1}}
+%%% And one that doesn't look so semantic
+\newcommand{\interp}[2]{\lceil #2 \rceil_{#1}}
 \newcommand{\dpath}[3]{\delta^{#1}_{#2}( #3 )}
@@
 
-First, we need to define the set of interpretations of the theory $\thD$,
+First, we define the set of interpretations of the theory $\thD$,
 which requires additional terminology concerning $\mu$-terms.
 For a $\mu$-term $t$ with subterm $u$,
-the \emph{interpretation of $u$ in $t$} is the $\mu$-term $\interp{t}{u}^\emptyset$ as returned by the following recursive procedure,
-where $x \leadsto \MU x.\; \const{C}( \vec u ) \in t$ indicates that $\MU x.\;
-\const{C}( \vec u )$ is the subterm of $t$ that binds this occurrence of
-variable $x$:
+%%% @ANDY: I'm taking the liberty of renaming "interpretation" to "expansion". Interpretation
+%%% sounds semantic to me (although it is sometimes used for syntactic transformations).
+the \emph{expansion of $u$ with respect to $t$} is the $\mu$-term $\interp{t}{u}^\emptyset$,
+abbreviated to
+$\interp{t}{u}$, as returned by the following recursive function:
+%where $x \leadsto \MU x.\; \const{C}( \vec u ) \in t$ indicates that $\MU x.\;
+%\const{C}( \vec u )$ is the subterm of $t$ that binds this occurrence of
+%variable $x$:
 \[\begin{array}{r@{}c@{}l}
-\interp{t}{x}^M & {} \,=\, {} &
+\interp{t}{x}^B & {} \,=\, {} &
 \textstyle\begin{cases}
-    \MU x.\; \const{C}( \interp{t}{ \vec u}^{M \mathrel\cup \{ x \}}) & \text{if }x \not\in M\text{~and $x$ is
-    bound by $\MU x.\;
-\const{C}( \vec u )$}  \\[-\jot]
+    \MU x.\; \const{C}( \interp{t}{ \vec u}^{\smash{B \mathrel\cup \{ x \}}}) & \text{if this $x \notin B$ is
+    bound by $\MU x.\; \const{C}( \vec u )$ in $t$}  \\[-\jot]
     x & \text{otherwise}
 \end{cases} \\
-\interp{t}{\MU x.\; \const{C}( \vec u )}^M & = &
+\interp{t}{\MU x.\; \const{C}( \vec u )}^B & = &
 \textstyle\begin{cases}
-    \MU x.\; \const{C}( \interp{t}{ \vec u }^{M \mathrel\cup \{ x \}} ) & \text{if }x \not\in M \\[-\jot]
+    \MU x.\; \const{C}( \interp{t}{ \vec u }^{\smash{B \mathrel\cup \{ x \}}} ) & \text{if }x \not\in B \\[-\jot]
     x & \text{otherwise}
@@
 \end{array}\]
-We write $\interp{t}{u}$ to abbreviate $\interp{t}{u}^\emptyset$.
-
-Given a $\mu$-term $t$ of the form $\MU x.\; \const{C}( t_1, \ldots, t_n )$ with subterm $u$,
-we say $u$ is a \emph{bisimilar subterm} of $t$ if it is of the form
-$\MU y.\; \const{C}( u_1, \ldots, u_n )$ and $\interp{t}{ t_k } \aequiv \interp{t}{ u_k }$ for $k = 1,\ldots, n$.
+
+Given a $\mu$-term $t$ of the form $\MU x.\; \const{C}( t_1, \ldots, t_n )$,
+a subterm $u$ is a \emph{bisimilar subterm} of~$t$ if it is of the form
+$\MU y.\; \const{C}( u_1, \ldots, u_n )$ and $\interp{t}{ t_k } \aequiv \interp{t}{ u_k }$ for $k \in \{1,\ldots,n\}$.
 The term $t$ is \emph{normal} if it does not contain a strict, bisimilar subterm
@@
 since $\MU y.\; \const{C}( y )$ is a bisimilar subterm of $t$.
-In particular, note that their arguments have the same interpretation in $t$,
+In particular, note that their arguments have the same expansion with respect to $t$,
 that is,
-$\interp{t}{\MU y.\; \const{C}( y )}^\emptyset =$
+$\interp{t}{\MU y.\; \const{C}( y )} =$
 $\MU y.\; \const{C}( \interp{t}{y}^{\{y\}} ) =$
@@
 is $\alpha$-equivalent to
-$\interp{t}{y}^\emptyset =$
+$\interp{t}{y} =$
 $\MU y.\; \const{C}( \interp{t}{y}^{\{y\}} ) =$
@@
 noting that
-$\interp{t}{\MU y.\; \const{C}( x )}^\emptyset =
+$\interp{t}{\MU y.\; \const{C}( x )} =
 \MU y.\; \const{C}( \interp{t}{x}^{\{y\}} ) =
@@
 \MU y.\allowbreak\; \const{C}( \MU x.\; \const{C}( y ) )$
-is $\alpha$-equivalent to $\interp{t}{x}^\emptyset = t$.
+is $\alpha$-equivalent to $\interp{t}{x} = t$.
 % for the renaming $\{ x \mapsto y, y \mapsto x \}$.
@@
 Due to our construction of $\Val$, 
-we know $\ec{x}$ contains a constructor $\const{C}( z_1, \ldots, z_n )$ and $\muvar( t_i ) = z_i$ for $i = 1, \ldots, n$.
+we know $\ec{x}$ contains a constructor $\const{C}( z_1, \ldots, z_n )$ and $\muvar( t_i ) = z_i$ for $i \in \{1,\ldots,n\}$.
 Since $\Val \ec{x}$ is a minimal, it
 contains a subterm of the form $\MU {\Varec{y}}.\; \const{C}( u_1, \ldots, u_n )$
-where $\interp{\Val \ec{x}}{ t_i } \aequiv \interp{\Val \ec{x} }{ u_i }$ for $i = 1, \ldots, n$.
-Due to our construction of $\Val$, $\ec{y}$ contains a constructor $\const{C}( w_1, \ldots, w_n )$ and $\muvar( u_i ) = w_i$ for $i = 1, \ldots, n$.
+where $\interp{\Val \ec{x}}{ t_i } \aequiv \interp{\Val \ec{x} }{ u_i }$ for $i \in \{1,\ldots,n\}$.
+Due to our construction of $\Val$, $\ec{y}$ contains a constructor $\const{C}( w_1, \ldots, w_n )$ and $\muvar( u_i ) = w_i$ for $i \in \{1,\ldots,n\}$.
 Since \rn{Cong} does not apply to $\Ec$
@@
 %of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$
-%where $\interp{\Val^\AST \ec{t}}{w_i} = \Val^\AST \ec{u_i}$ for $i = 1, \ldots, n$.
+%where $\interp{\Val^\AST \ec{t}}{w_i} = \Val^\AST \ec{u_i}$ for $i \in \{1,\ldots,n\}$.
 Since \rn{Inject} and \rn{Clash} do not apply to $\Fc$, 
@@
 of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$ 
-where for each $i = 1, \ldots, n$, 
+where for each $i \in \{1,\ldots,n\}$, 
 $\muvar(w_i) = {\Varec{u_i}}$ and thus by Lemma~\ref{lem:model-completion}\PART{2}, 
 $\interp{\Val^\AST \ec{t}}{w_i} = \Val^\AST \ec{u_i}$.
-For each $i = 1, \ldots, n$, let $u_\iPrime$ be the $i$\vvthinspace th argument of $t_u$.
+For each $i \in \{1,\ldots,n\}$, let $u_\iPrime$ be the $i$\vvthinspace th argument of $t_u$.
 By Lemma~\ref{lem:mu-norm-arg}, $\interp{t_u}{u_\iPrime} \vsim \Val^\AST \ec{u_i}$.
-Thus $\interp{t_u}{u_\iPrime} \vsim \interp{\Val^\AST \ec{t}}{ w_i }$. % for each $i = 1, \ldots, n$.
+Thus $\interp{t_u}{u_\iPrime} \vsim \interp{\Val^\AST \ec{t}}{ w_i }$. % for each $i \in \{1,\ldots,n\}$.
 %By Lemma~\ref{lem:mu-cong}, we have $t_u \vsim t_w$.
@@
 %$\MU {\Varec{u}}.\; \const C_{\negvthinspace\jPrime}( w_1, \ldots, w_n )$,
-%where $\interp{\Val^\AST \ec{u} }{ w_i } = \Val^\AST \ec{ \const s^i_{\negvthinspace\jPrime}( u )}$ for $i = 1, \ldots, n$. 
+%where $\interp{\Val^\AST \ec{u} }{ w_i } = \Val^\AST \ec{ \const s^i_{\negvthinspace\jPrime}( u )}$ for $i \in \{1,\ldots,n\}$. 
 Since \rn{Inject} and \rn{Clash} do not apply to $\Fc$, 
@@
 $\MU {\Varec{u}}.\; \const C_{\negvthinspace\jPrime}( w_1, \ldots, w_n )$,
-where for each $i = 1, \ldots, n$,
+where for each $i \in \{1,\ldots,n\}$,
 $\muvar(w_i) = \smash{\Varec{\vphantom{I}\smash{\const s^i_{\negvthinspace\jPrime}( u )}}}$ and thus by Lemma~\ref{lem:model-completion}\PART{2}, 

### 2014-12-22 tex/codatadp.tex
more

@@
 %%% And one that doesn't look so semantic
-\newcommand{\interp}[2]{\lceil #2 \rceil_{#1}}
+%\newcommand{\interp}[2]{\lceil #2 \rceil_{#1}}
+\newcommand{\interp}[2]{\left< \smash{#2} \right>_{#1}}
 \newcommand{\dpath}[3]{\delta^{#1}_{#2}( #3 )}
@@
 Symbols starting with an uppercase letter (e.g.,
-$\const{S}$, $\const{SCons}$) are reserved for constructors. With each function symbol \const{f}
+$\const{C}, \const{S}$, $\const{SCons}$) are reserved for constructors. With each function symbol \const{f}
 is associated a list of argument types $\tau_1,\ldots,\tau_n$ (with $n \ge 0$)
@@
 parentheses in terms.
-%
-Finally, $\bar x$ abbreviates a list or tuple $x_1,\ldots,x_n$\begin{report}
-and $\bigwedge_{\,i}\, \varphi_i$ abbeviates a conjunction
-$\varphi_1 \mathrel\land \cdots \mathrel\land \varphi_n$\end{report}.
+\begin{report}%
+The operator $\bigwedge_{\,i}\, \varphi_i$ abbeviates a conjunction
+$\varphi_1 \mathrel\land \cdots \mathrel\land \varphi_n$. \end{report}%
+Finally, $[n]$ denotes the set $\{1,\ldots, n\}$, and
+$\bar x$ abbreviates a list or tuple $x_1,\ldots,x_n$.
 
@@
 
-Given a $\mu$-term $t$ of the form $\MU x.\; \const{C}( t_1, \ldots, t_n )$,
-a subterm $u$ is a \emph{bisimilar subterm} of~$t$ if it is of the form
-$\MU y.\; \const{C}( u_1, \ldots, u_n )$ and $\interp{t}{ t_k } \aequiv \interp{t}{ u_k }$ for $k \in \{1,\ldots,n\}$.
+Given a $\mu$-term $t$ and a subterm $u$,
+$u$ is a \emph{bisimilar subterm} of~$t$ if
+$t$ is of the form $\MU x.\; \const{C}( t_1, \ldots, t_n )$,
+$u$ it is of the form $\MU y.\; \const{C}( u_1, \ldots, u_n )$,
+and $\interp{t}{ t_k } \aequiv \interp{t}{ u_k }$ for all $k \in [n]$.
 The term $t$ is \emph{normal} if it does not contain a strict, bisimilar subterm
@@
 Due to our construction of $\Val$, 
-we know $\ec{x}$ contains a constructor $\const{C}( z_1, \ldots, z_n )$ and $\muvar( t_i ) = z_i$ for $i \in \{1,\ldots,n\}$.
+we know $\ec{x}$ contains a constructor $\const{C}( z_1, \ldots, z_n )$ and $\muvar( t_i ) = z_i$ for $i \in [n]$.
 Since $\Val \ec{x}$ is a minimal, it
 contains a subterm of the form $\MU {\Varec{y}}.\; \const{C}( u_1, \ldots, u_n )$
-where $\interp{\Val \ec{x}}{ t_i } \aequiv \interp{\Val \ec{x} }{ u_i }$ for $i \in \{1,\ldots,n\}$.
-Due to our construction of $\Val$, $\ec{y}$ contains a constructor $\const{C}( w_1, \ldots, w_n )$ and $\muvar( u_i ) = w_i$ for $i \in \{1,\ldots,n\}$.
+where $\interp{\Val \ec{x}}{ t_i } \aequiv \interp{\Val \ec{x} }{ u_i }$ for $i \in [n]$.
+Due to our construction of $\Val$, $\ec{y}$ contains a constructor $\const{C}( w_1, \ldots, w_n )$ and $\muvar( u_i ) = w_i$ for $i \in [n]$.
 Since \rn{Cong} does not apply to $\Ec$
@@
 and interprets the variables of $\Fc$ based on the completion $\Val^\AST$ of the mapping $\Val$ from phase~2.
-For the latter, let $\M( x )$ be the value in $\M$ that is $\alpha$-equivalent with $\Val^\AST \ec{x}$ for each variable $x$ in $\mathcal{T}(\Fc)$.
+For the latter, let $\M( x )$ be the value in $\M$ that is $\alpha$-equivalent with $\Val^\AST \ec{x}$ for each variable $x \in \mathcal{T}(\Fc)$.
 
@@
 %of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$
-%where $\interp{\Val^\AST \ec{t}}{w_i} = \Val^\AST \ec{u_i}$ for $i \in \{1,\ldots,n\}$.
+%where $\interp{\Val^\AST \ec{t}}{w_i} = \Val^\AST \ec{u_i}$ for $i \in [n]$.
 Since \rn{Inject} and \rn{Clash} do not apply to $\Fc$, 
@@
 of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$ 
-where for each $i \in \{1,\ldots,n\}$, 
+where for each $i \in [n]$, 
 $\muvar(w_i) = {\Varec{u_i}}$ and thus by Lemma~\ref{lem:model-completion}\PART{2}, 
 $\interp{\Val^\AST \ec{t}}{w_i} = \Val^\AST \ec{u_i}$.
-For each $i \in \{1,\ldots,n\}$, let $u_\iPrime$ be the $i$\vvthinspace th argument of $t_u$.
+For each $i \in [n]$, let $u_\iPrime$ be the $i$\vvthinspace th argument of $t_u$.
 By Lemma~\ref{lem:mu-norm-arg}, $\interp{t_u}{u_\iPrime} \vsim \Val^\AST \ec{u_i}$.
-Thus $\interp{t_u}{u_\iPrime} \vsim \interp{\Val^\AST \ec{t}}{ w_i }$. % for each $i \in \{1,\ldots,n\}$.
+Thus $\interp{t_u}{u_\iPrime} \vsim \interp{\Val^\AST \ec{t}}{ w_i }$. % for each $i \in [n]$.
 %By Lemma~\ref{lem:mu-cong}, we have $t_u \vsim t_w$.
@@
 %$\MU {\Varec{u}}.\; \const C_{\negvthinspace\jPrime}( w_1, \ldots, w_n )$,
-%where $\interp{\Val^\AST \ec{u} }{ w_i } = \Val^\AST \ec{ \const s^i_{\negvthinspace\jPrime}( u )}$ for $i \in \{1,\ldots,n\}$. 
+%where $\interp{\Val^\AST \ec{u} }{ w_i } = \Val^\AST \ec{ \const s^i_{\negvthinspace\jPrime}( u )}$ for $i \in [n]$. 
 Since \rn{Inject} and \rn{Clash} do not apply to $\Fc$, 
@@
 $\MU {\Varec{u}}.\; \const C_{\negvthinspace\jPrime}( w_1, \ldots, w_n )$,
-where for each $i \in \{1,\ldots,n\}$,
+where for each $i \in [n]$,
 $\muvar(w_i) = \smash{\Varec{\vphantom{I}\smash{\const s^i_{\negvthinspace\jPrime}( u )}}}$ and thus by Lemma~\ref{lem:model-completion}\PART{2}, 

### 2014-12-22 tex/codatadp.tex
more
no confusion between syntactic concepts (bisimilar subterm) and a somewhat different semantic notion

@@
 \interp{t}{x}^B & {} \,=\, {} &
-\textstyle\begin{cases}
+\textstyle\smash{\begin{cases}
     \MU x.\; \const{C}( \interp{t}{ \vec u}^{\smash{B \mathrel\cup \{ x \}}}) & \text{if this $x \notin B$ is
@@
     x & \text{otherwise}
-\end{cases} \\
+\end{cases}} ~ \\[1.5ex] %% TYPESETTING: trick with \smash
 \interp{t}{\MU x.\; \const{C}( \vec u )}^B & = &
@@
 Given a $\mu$-term $t$ and a subterm $u$,
-$u$ is a \emph{bisimilar subterm} of~$t$ if
+$u$ is an \emph{obviously bisimilar subterm} of~$t$ if
 $t$ is of the form $\MU x.\; \const{C}( t_1, \ldots, t_n )$,
@@
 and $\interp{t}{ t_k } \aequiv \interp{t}{ u_k }$ for all $k \in [n]$.
-The term $t$ is \emph{normal} if it does not contain a strict, bisimilar subterm
+The term $t$ is \emph{normal} if it does not contain an obviously bisimilar strict subterm
 and all of its strict subterms are also normal.
 For example, the $\mu$-term $t = \MU x.\; \const{C}( \MU y.\; \const{C}( y ) )$ is abnormal,
-since $\MU y.\; \const{C}( y )$ is a bisimilar subterm of $t$.
+since $\MU y.\; \const{C}( y )$ is an obviously bisimilar subterm of $t$.
 In particular, note that their arguments have the same expansion with respect to $t$,
@@
 The $\mu$-term $t$ of the form $\MU x.\; \const{C}( \MU y.\; \const{C}( x ) )$ is also abnormal,
-since $\MU y.\; \const{C}( x )$ is a bisimilar subterm of $t$,
+since $\MU y.\; \const{C}( x )$ is an obviously bisimilar subterm of $t$,
 noting that
@@
 recursively construct its \emph{normal form} $\nf{t}$,
-by replacing all of the bisimilar subterms of $t$ with $x$
+by replacing all of the obviously bisimilar subterms of $t$ with $x$
 %and by replacing all of the abnormal subterms of $t$ with their corresponding normal forms.
@@
 and $\interp{\Val \ec{x}}{ u_{\negvthinspace j} } = \Val \ec{w_{\negvthinspace j}}$,
-which are not $\alpha$-equivalent by statement 3, contradicting the fact that $\MU {\Varec{y}}.\; \const{C}( u_1, \ldots, u_n )$ is a bisimilar subterm of $\Val \ec{x}$.
+which are not $\alpha$-equivalent by statement 3, contradicting the fact that $\MU {\Varec{y}}.\; \const{C}( u_1, \ldots, u_n )$ is an obviously bisimilar subterm of $\Val \ec{x}$.
 Thus, $\Val = \nf{\Val}$, and statements 2 and 3 hold for $\nf{\Val}$.
@@
 %then without loss of generality we may assume
-%$\Val_1 \sigma \ec{y}$ contains a strict bisimilar subterm 
+%$\Val_1 \sigma \ec{y}$ contains an obviously bisimilar strict subterm 
 %$\MU {\Varec{w}}.\; u$ that strictly contains $t_1 \sigma$.

### 2014-12-22 tex/codatadp.tex
make definition more robust (no precondition)

@@
 
-Given a $\mu$-term $t$ and a subterm $u$,
-$u$ is an \emph{obviously bisimilar subterm} of~$t$ if
+The $\mu$-term $u$
+is an \emph{obviously bisimilar subterm} of~$t$ if
+$u$ is a strict subterm of $t$,
 $t$ is of the form $\MU x.\; \const{C}( t_1, \ldots, t_n )$,

### 2014-12-22 tex/codatadp.tex
trying another terminology

--- 2014-12-22 tex/codatadp.tex
@@
 
-Given a $\mu$-term $t$ and a subterm $u$,
-$u$ is an \emph{obviously bisimilar subterm} of~$t$ if
+The $\mu$-term $u$
+is a \emph{self-similar subterm} of~$t$ if
+$u$ is a strict subterm of $t$,
 $t$ is of the form $\MU x.\; \const{C}( t_1, \ldots, t_n )$,
@@
 and $\interp{t}{ t_k } \aequiv \interp{t}{ u_k }$ for all $k \in [n]$.
-The term $t$ is \emph{normal} if it does not contain an obviously bisimilar strict subterm
+The term $t$ is \emph{normal} if it does not contain a self-similar subterm
 and all of its strict subterms are also normal.
 For example, the $\mu$-term $t = \MU x.\; \const{C}( \MU y.\; \const{C}( y ) )$ is abnormal,
-since $\MU y.\; \const{C}( y )$ is an obviously bisimilar subterm of $t$.
+since $\MU y.\; \const{C}( y )$ is a self-similar subterm of $t$.
 In particular, note that their arguments have the same expansion with respect to $t$,
@@
 The $\mu$-term $t$ of the form $\MU x.\; \const{C}( \MU y.\; \const{C}( x ) )$ is also abnormal,
-since $\MU y.\; \const{C}( x )$ is an obviously bisimilar subterm of $t$,
+since $\MU y.\; \const{C}( x )$ is a self-similar subterm of $t$,
 noting that
@@
 recursively construct its \emph{normal form} $\nf{t}$,
-by replacing all of the obviously bisimilar subterms of $t$ with $x$
+by replacing all of the self-similar subterms of $t$ with $x$
 %and by replacing all of the abnormal subterms of $t$ with their corresponding normal forms.
@@
 and $\interp{\Val \ec{x}}{ u_{\negvthinspace j} } = \Val \ec{w_{\negvthinspace j}}$,
-which are not $\alpha$-equivalent by statement 3, contradicting the fact that $\MU {\Varec{y}}.\; \const{C}( u_1, \ldots, u_n )$ is an obviously bisimilar subterm of $\Val \ec{x}$.
+which are not $\alpha$-equivalent by statement 3, contradicting the fact that $\MU {\Varec{y}}.\; \const{C}( u_1, \ldots, u_n )$ is a self-similar subterm of $\Val \ec{x}$.
 Thus, $\Val = \nf{\Val}$, and statements 2 and 3 hold for $\nf{\Val}$.
@@
 %then without loss of generality we may assume
-%$\Val_1 \sigma \ec{y}$ contains an obviously bisimilar strict subterm 
+%$\Val_1 \sigma \ec{y}$ contains a self-similar subterm 
 %$\MU {\Varec{w}}.\; u$ that strictly contains $t_1 \sigma$.

### 2014-12-22 tex/codatadp.tex
more

@@
 in $\M( \tau )$ that is $\alpha$-equivalent to
-$\nf{\MU x.\; \const{C}( \M( \vec t \,) )}$ where $x$ is fresh.
+$\nf{\MU x.\; \const{C}( \M( \vec t \,) )}$, where $x$ is fresh.
 \item
@@
 in $\M( \tau )$ that is $\alpha$-equivalent to
-$\interp{\M( t )}{u_k}$
-when $\M( t )$ is of the form $\MU x.\; \const{C}_{\negvthinspace j}( \vec u )$.
+$\interp{\M( t )}{u_k}$,
+assuming $\M( t )$ is of the form $\MU x.\; \const{C}_{\negvthinspace j}( \vec u )$.
 \end{enumerate}
@@
 we have $\ec{w_{\negvthinspace j}}$ and $\ec{z_{\negvthinspace j}}$ are distinct for some $j$.
-By statement 2, $\interp{\Val \ec{x}}{ t_{\negvthinspace j} } = \Val \ec{z_{\negvthinspace j}}$
+By (2), $\interp{\Val \ec{x}}{ t_{\negvthinspace j} } = \Val \ec{z_{\negvthinspace j}}$
 and $\interp{\Val \ec{x}}{ u_{\negvthinspace j} } = \Val \ec{w_{\negvthinspace j}}$,
-which are not $\alpha$-equivalent by statement 3, contradicting the fact that $\MU {\Varec{y}}.\; \const{C}( u_1, \ldots, u_n )$ is a self-similar subterm of $\Val \ec{x}$.
-Thus, $\Val = \nf{\Val}$, and statements 2 and 3 hold for $\nf{\Val}$.
-
-We now show that if statements 2 and 3 hold for some $\Val_1$,
-then 2 and 3 also hold for $\nf{\Val_1 \sigma}$,
+which are not $\alpha$-equivalent by (3), contradicting the fact that $\MU {\Varec{y}}.\; \const{C}( u_1, \ldots, u_n )$ is a self-similar subterm of $\Val \ec{x}$.
+Thus, $\Val = \nf{\Val}$, and (2) and (3) hold for $\nf{\Val}$.
+
+We now show that if (2) and (3) hold for some $\Val_1$,
+they also hold for $\nf{\Val_1 \sigma}$,
 where $\sigma$ is a substitution of the form $\{ {\Varec{x}} \mapsto \MU{\Varec{x}}.\; t \}$,
@@
 and assume by contradiction that $\nf{\Val_1 \sigma} \ec{y} \aequiv \nf{\Val_1 \sigma} \ec{z}$.
-Due to statement 3 for $\Val_1$,
+Due to (3) for $\Val_1$,
 $\ec{y}$ and $\ec{z}$ must have (minimal) subterms 
@@
 Thus, by induction on the number of applications of the above rule used to obtain $\Val^\AST$,
-we have that $\Val^\AST$ satisfies statements 2 and 3.
+we have that $\Val^\AST$ satisfies (2) and (3).
 \qed

### 2014-12-22 tex/codatadp.tex
more edits, including new Tab. 1 data

@@
 \newcommand{\Val}{\mathcal{A}}
+\newcommand{\ValC}{\mathcal{A}^\star}
 \newcommand{\Var}{\mathcal{V}}
@@
 and all of its strict subterms are also normal.
-For example, the $\mu$-term $t = \MU x.\; \const{C}( \MU y.\; \const{C}( y ) )$ is abnormal,
-since $\MU y.\; \const{C}( y )$ is a self-similar subterm of $t$.
+For example, $t = \MU x.\; \const{C}( \MU y.\; \const{C}( y ) )$ is abnormal,
+because $\MU y.\; \const{C}( y )$ is a self-similar subterm of $t$.
 In particular, note that their arguments have the same expansion with respect to $t$,
@@
 For the latter, this will be based on the mapping $\Val$ constructed in phase~2 of the calculus.
-We compute the \emph{completion} $\Val^\AST$ of $\Val$ for normal model $\M$
+We compute the \emph{completion} $\ValC$ of $\Val$ for normal model $\M$
 by assigning values from $\M$ to unassigned variables in the domain of $\Val$.
@@
 
-We construct $\Val^\AST$ by exhaustively applying the following rule to $\nf{\Val}$:
+We construct $\ValC$ by exhaustively applying the following rule to $\nf{\Val}$:
 \[
@@
 %%% TYPESETTING: Evil hack!!
-%We first state the following properties of $\Val^\AST$,
-In the following lemma about $\Val^\AST$,
+%We first state the following properties of $\ValC$,
+In the following lemma about $\ValC$,
 $\vphantom{\Bigl(_j}\muvar( t ) = \smash{\begin{cases}
@@
 If $\Val$ is constructed for a saturated set $\Ec$
-and $\Val^\AST$ is a completion of $\Val$ for a normal model $\M$, the following properties hold:
+and $\ValC$ is a completion of $\Val$ for a normal model $\M$, the following properties hold:
 \begin{enumerate}
-\item[\rm (1)] $\Val^\AST \ec{x}$ is $\alpha$-equivalent to a value in $\M( \tau )$, where $\ec{x}$ has type $\tau$.
+\item[\rm (1)] $\ValC \ec{x}$ is $\alpha$-equivalent to a value in $\M( \tau )$, where $\ec{x}$ has type $\tau$.
 \item[\rm (2)]
-$\Val^\AST \ec{x} = \interp{\Val^\AST \ec{y}}{ t }$ for
-all subterms $t$ of $\Val^\AST \ec{y}$ with $\muvar(t) = {\Varec{x}}$. 
+$\ValC \ec{x} = \interp{\ValC \ec{y}}{ t }$ for
+all subterms $t$ of $\ValC \ec{y}$ with $\muvar(t) = {\Varec{x}}$. 
 \item[\rm (3)]
-$\Val^\AST \ec{x} \vsim \Val^\AST \ec{y}$ if and only if $\ec{x} = \ec{y}$.
+$\ValC \ec{x} \vsim \ValC \ec{y}$ if and only if $\ec{x} = \ec{y}$.
 \end{enumerate}
@@
 \begin{proof}
-To show 1, we first show that $\Val^\AST$ contains no free variables.
-Assume by contradiction that $\Val^\AST$ contained a free variable ${\Varec{y}}$ for some $\ec{y}$ of type $\tau$.
+To show 1, we first show that $\ValC$ contains no free variables.
+Assume by contradiction that $\ValC$ contained a free variable ${\Varec{y}}$ for some $\ec{y}$ of type $\tau$.
 Then it must be the case that $\ec{y}$ does not contain a constructor term,
@@
 there is only one equivalence class of type $\tau$ in $\Ec$,
-and thus there are no terms in $\mathcal{T}^\M_{\Varec{y}}(\Val^\AST)$ of type $\tau$.
+and thus there are no terms in $\mathcal{T}^\M_{\Varec{y}}(\ValC)$ of type $\tau$.
 This is a contradiction, since our model completion may assign the value in the domain of $\tau$ to ${\Varec{y}}$.
@@
 This is also a contradiction,
-since there are only a finite number of closed terms in $\mathcal{T}^\M_{\Varec{y}}(\Val^\AST)$,
-and thus our model completion can assign a value not occurring in $\mathcal{T}^\M_{\Varec{y}}(\Val^\AST)$ to ${\Varec{y}}$.
-By construction, $\Val^\AST \ec{x}$ is normal.
+since there are only a finite number of closed terms in $\mathcal{T}^\M_{\Varec{y}}(\ValC)$,
+and thus our model completion can assign a value not occurring in $\mathcal{T}^\M_{\Varec{y}}(\ValC)$ to ${\Varec{y}}$.
+By construction, $\ValC \ec{x}$ is normal.
 Since \rn{Acyclic} does not apply, $\Val \ec{x}$ is acyclic when $\tau \in \Data$.
-Moreover, our construction of $\Val^\AST$ applies substitutions of the form
+Moreover, our construction of $\ValC$ applies substitutions of the form
 $\{ {\Varec{y}} \mapsto t \}$, where $t$ is acyclic when the type of ${\Varec{y}}$ is not a codatatype.
-Thus, $\Val^\AST \ec{x}$ is acyclic when $\tau \not\in \Codata$.
-Therefore, by definition, $\Val^\AST \ec{x}$ is $\alpha$-equivalent to a value in $\M( \tau )$.
+Thus, $\ValC \ec{x}$ is acyclic when $\tau \not\in \Codata$.
+Therefore, by definition, $\ValC \ec{x}$ is $\alpha$-equivalent to a value in $\M( \tau )$.
 
@@
 
-Thus, by induction on the number of applications of the above rule used to obtain $\Val^\AST$,
-we have that $\Val^\AST$ satisfies (2) and (3).
+Thus, by induction on the number of applications of the above rule used to obtain $\ValC$,
+we have that $\ValC$ satisfies (2) and (3).
 \qed
@@
 that interprets wrongly applied selectors based on equality information in $\Fc$,
-and interprets the variables of $\Fc$ based on the completion $\Val^\AST$ of the mapping $\Val$ from phase~2.
-For the latter, let $\M( x )$ be the value in $\M$ that is $\alpha$-equivalent with $\Val^\AST \ec{x}$ for each variable $x \in \mathcal{T}(\Fc)$.
+and interprets the variables of $\Fc$ based on the completion $\ValC$ of the mapping $\Val$ from phase~2.
+For the latter, let $\M( x )$ be the value in $\M$ that is $\alpha$-equivalent with $\ValC \ec{x}$ for each variable $x \in \mathcal{T}(\Fc)$.
 
@@
 by structural induction on $t$,
-that $\M( t ) \aequiv \Val^\AST \ec{t}$ for all terms $t \in \mathcal{T}( \Fc )$,
+that $\M( t ) \aequiv \ValC \ec{t}$ for all terms $t \in \mathcal{T}( \Fc )$,
 which since $\M$ is normal, implies $\M \models t_1 \teq t_2$.
 
-If $t$ is a variable, then $\M( t ) \vsim \Val^\AST \ec{ t }$ by construction.
+If $t$ is a variable, then $\M( t ) \vsim \ValC \ec{ t }$ by construction.
 
@@
 then $\M( t )$ is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \M( u_1 ), \ldots, \M( u_n ) ) }$ for some fresh $x$,
-which by the induction hypothesis is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \Val^\AST \ec{ u_1 }, \ldots, \Val^\AST \ec{ u_n } )}$, call this term $t_u$.
-%By Lemma~\ref{lem:model-completion}\PART{2} and due to our construction of $\Val^\AST$, we have $\Val^\AST \ec{t}$ is a term
+which by the induction hypothesis is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \ValC \ec{ u_1 }, \ldots, \ValC \ec{ u_n } )}$, call this term $t_u$.
+%By Lemma~\ref{lem:model-completion}\PART{2} and due to our construction of $\ValC$, we have $\ValC \ec{t}$ is a term
 %of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$
-%where $\interp{\Val^\AST \ec{t}}{w_i} = \Val^\AST \ec{u_i}$ for $i \in [n]$.
+%where $\interp{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$ for $i \in [n]$.
 Since \rn{Inject} and \rn{Clash} do not apply to $\Fc$, 
-by our construction of $\Val^\AST$ we have $\Val^\AST \ec{t}$ is a term
+by our construction of $\ValC$ we have $\ValC \ec{t}$ is a term
 of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$ 
@@
 $\muvar(w_i) = {\Varec{u_i}}$ and thus by Lemma~\ref{lem:model-completion}\PART{2}, 
-$\interp{\Val^\AST \ec{t}}{w_i} = \Val^\AST \ec{u_i}$.
+$\interp{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$.
 For each $i \in [n]$, let $u_\iPrime$ be the $i$\vvthinspace th argument of $t_u$.
-By Lemma~\ref{lem:mu-norm-arg}, $\interp{t_u}{u_\iPrime} \vsim \Val^\AST \ec{u_i}$.
-Thus $\interp{t_u}{u_\iPrime} \vsim \interp{\Val^\AST \ec{t}}{ w_i }$. % for each $i \in [n]$.
+By Lemma~\ref{lem:mu-norm-arg}, $\interp{t_u}{u_\iPrime} \vsim \ValC \ec{u_i}$.
+Thus $\interp{t_u}{u_\iPrime} \vsim \interp{\ValC \ec{t}}{ w_i }$. % for each $i \in [n]$.
 %By Lemma~\ref{lem:mu-cong}, we have $t_u \vsim t_w$.
-Thus, $\M( t ) \vsim t_u \vsim \Val^\AST \ec{t}$, and
-we have $\M( t ) \vsim \Val^\AST \ec{t}$.
+Thus, $\M( t ) \vsim t_u \vsim \ValC \ec{t}$, and
+we have $\M( t ) \vsim \ValC \ec{t}$.
 
@@
 $\ec{u}$ must contain a term of the form $\const C_{\negvthinspace\jPrime}\bigl( \const s^1_{\negvthinspace\jPrime}( u ), \ldots, \const s^{n}_{\negvthinspace\jPrime}( u ) \bigr)$ for some $\jPrime$.
-%By Lemma~\ref{lem:model-completion}\PART{2} and due to our construction of $\Val^\AST$, we have $\Val^\AST \ec{u}$ is of the form
+%By Lemma~\ref{lem:model-completion}\PART{2} and due to our construction of $\ValC$, we have $\ValC \ec{u}$ is of the form
 %$\MU {\Varec{u}}.\; \const C_{\negvthinspace\jPrime}( w_1, \ldots, w_n )$,
-%where $\interp{\Val^\AST \ec{u} }{ w_i } = \Val^\AST \ec{ \const s^i_{\negvthinspace\jPrime}( u )}$ for $i \in [n]$. 
+%where $\interp{\ValC \ec{u} }{ w_i } = \ValC \ec{ \const s^i_{\negvthinspace\jPrime}( u )}$ for $i \in [n]$. 
 Since \rn{Inject} and \rn{Clash} do not apply to $\Fc$, 
-by construction $\Val^\AST$ must be of the form
+by construction $\ValC$ must be of the form
 $\MU {\Varec{u}}.\; \const C_{\negvthinspace\jPrime}( w_1, \ldots, w_n )$,
@@
 $\muvar(w_i) = \smash{\Varec{\vphantom{I}\smash{\const s^i_{\negvthinspace\jPrime}( u )}}}$ and thus by Lemma~\ref{lem:model-completion}\PART{2}, 
-$\interp{\Val^\AST \ec{u}}{ w_i } = \Val^\AST \ec{ \smash{\const s^i_{\negvthinspace\jPrime}( u )}}$. 
-If $j = \jPrime$, then $\M( t )$ is $\alpha$-equivalent with $\interp{\Val^\AST \ec{u}}{w_k}$, which is equal to $\Val^\AST \ec{ \smash{ \const s^k_{\negvthinspace j}( u )}}$,
-which is $\Val^\AST \ec{t}$.
-If $j \neq \jPrime$, we are free to interpret $\smash{\const s^k_{\negvthinspace j}( \M( u ) )}$ as the value in $\M( \tau )$ that is $\alpha$-equivalent with $\Val^\AST \ec{ t }$.
+$\interp{\ValC \ec{u}}{ w_i } = \ValC \ec{ \smash{\const s^i_{\negvthinspace\jPrime}( u )}}$. 
+If $j = \jPrime$, then $\M( t )$ is $\alpha$-equivalent with $\interp{\ValC \ec{u}}{w_k}$, which is equal to $\ValC \ec{ \smash{ \const s^k_{\negvthinspace j}( u )}}$,
+which is $\ValC \ec{t}$.
+If $j \neq \jPrime$, we are free to interpret $\smash{\const s^k_{\negvthinspace j}( \M( u ) )}$ as the value in $\M( \tau )$ that is $\alpha$-equivalent with $\ValC \ec{ t }$.
 Since \rn{Cong} does not apply and due to Lemma~\ref{lem:model-completion}\PART{3},
@@
 Since \rn{Conflict} does not apply, $t_1 \teq t_2 \not\in \Fc$ and thus $\ec{t_1}$ and $\ec{t_2}$ are distinct.
-Since $\M( t_1 ) \vsim \Val^\AST \ec{t_1}$ and $\M( t_2 ) \vsim \Val^\AST \ec{t_2}$,
+Since $\M( t_1 ) \vsim \ValC \ec{t_1}$ and $\M( t_2 ) \vsim \ValC \ec{t_2}$,
 by Lemma~\ref{lem:model-completion}\PART{3}, $\M( t_1 ) \neq \M( t_2 )$, and thus $\M \models t_1 \tneq t_2$.
@@
 Codatatypes without uniqueness
-  & 215 & 789 & \phantom{0}56 & 1060 \\
+  & 222 & 788 & \phantom{0}56 & 1066 \\
 Codatatypes with uniqueness
-  & 216 & 789 & \bfseries \phantom{0}59 & 1064 \\[\jot]
+  & 223 & 788 & \bfseries \phantom{0}59 & 1070 \\[\jot]
 Full (co)datatypes
-  & 222 & \bfseries 800 & \bfseries \phantom{0}59 & \bfseries 1081 \\[\jot]
+  & 229 & \bfseries 800 & \bfseries \phantom{0}59 & \bfseries 1088 \\[\jot]
 Total number of goals

### 2014-12-22 tex/codatadp.tex
fixed typo

@@
 Codatatypes without uniqueness
-  & 222 & 788 & \phantom{0}56 & 1066 \\
+  & 222 & 789 & \phantom{0}56 & 1067 \\
 Codatatypes with uniqueness
-  & 223 & 788 & \bfseries \phantom{0}59 & 1070 \\[\jot]
+  & 223 & 789 & \bfseries \phantom{0}59 & 1071 \\[\jot]
 Full (co)datatypes

### 2014-12-22 tex/codatadp.tex
more

@@
 Datatypes without acyclicity
-  & \bfseries 227 & 765 & \phantom{0}55 & 1047 \\
+  & 227 & 765 & \phantom{0}55 & 1047 \\
 Datatypes with acyclicity
-  & \bfseries 227 & 771 & \phantom{0}55 & 1053 \\[\jot]
+  & 227 & 771 & \phantom{0}55 & 1053 \\[\jot]
 %Codatatype simplication only
@@
 Full (co)datatypes
-  & 229 & \bfseries 800 & \bfseries \phantom{0}59 & \bfseries 1088 \\[\jot]
+  & \bfseries 229 & \bfseries 800 & \bfseries \phantom{0}59 & \bfseries 1088 \\[\jot]
 Total number of goals

### 2014-12-22 tex/codatadp.tex
more

@@
 and all of its strict subterms are also normal.
-For example, $t = \MU x.\; \const{C}( \MU y.\; \const{C}( y ) )$ is abnormal,
+For example, $t = \MU x.\; \const{C}( \MU y.\; \const{C}( y ) )$ is abnormal
 because $\MU y.\; \const{C}( y )$ is a self-similar subterm of $t$.
@@
 $\interp{t}{\MU y.\; \const{C}( y )} =$
-$\MU y.\; \const{C}( \interp{t}{y}^{\{y\}} ) =$
+$\MU y.\; \const{C}( \interp{t}{y}^{\smash{\{y\}}} ) =$
 $\MU y.\; \const{C}( y )$
@@
 $\interp{t}{y} =$
-$\MU y.\; \const{C}( \interp{t}{y}^{\{y\}} ) =$
+$\MU y.\; \const{C}( \interp{t}{y}^{\smash{\{y\}}} ) =$
 $\MU y.\; \const{C}( y )$.
-The $\mu$-term $t$ of the form $\MU x.\; \const{C}( \MU y.\; \const{C}( x ) )$ is also abnormal,
+The term $t$ of the form $\MU x.\; \const{C}( \MU y.\; \const{C}( x ) )$ is also abnormal,
 since $\MU y.\; \const{C}( x )$ is a self-similar subterm of $t$,
 noting that
-$\interp{t}{\MU y.\; \const{C}( x )} =
-\MU y.\; \const{C}( \interp{t}{x}^{\{y\}} ) =
-\MU y.\; \const{C}( \interp{t}{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) ) }^{\{y\}} ) =
-\MU y.\; \const{C}( \MU x.\; \const{C}( \interp{t}{\MU y.\; \const{C}( x )}^{\{x,y\}} ) ) =
-\MU y.\allowbreak\; \const{C}( \MU x.\; \const{C}( y ) )$
+%
+\begin{align*}
+\interp{t}{\MU y.\; \const{C}( x )}
+& = \MU y.\; \const{C}( \interp{t}{x}^{\smash{\{y\}}} ) \\
+& = \MU y.\; \const{C}( \interp{t}{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) ) }^{\smash{\{y\}}} ) \\
+& = \MU y.\; \const{C}( \MU x.\; \const{C}( \interp{t}{\MU y.\; \const{C}( x )}^{\smash{\{x,y\}}} ) ) \\
+& = \MU y.\allowbreak\; \const{C}( \MU x.\allowbreak\; \const{C}( y ) )
+\end{align*}
 is $\alpha$-equivalent to $\interp{t}{x} = t$.
 % for the renaming $\{ x \mapsto y, y \mapsto x \}$.
+
 For any $\mu$-term $t$ of the form $\MU x.\; \const C( \vec u )$, we can

### 2014-12-22 tex/codatadp.tex
comma

@@
 For any $\mu$-term $t$ of the form $\MU x.\; \const C( \vec u )$, we can
-recursively construct its \emph{normal form} $\nf{t}$,
+recursively construct its \emph{normal form} $\nf{t}$
 by replacing all of the self-similar subterms of $t$ with $x$

### 2014-12-23 tex/codatadp.tex
more

@@
 and thus $\Ec$ is unsatisfiable.
-If the children of $\Ec$ are closed derivation trees
+If the child nodes of $\Ec$ are closed derivation trees
 whose root nodes are the result of applying the rule \rn{Split} on $t\typ{\delta}$,
@@
 this means $\Ec$ is $\thD$-unsatisfiable.
-Otherwise, the child of $\Ec$ is a closed derivation tree
+Otherwise, the child node of $\Ec$ is a closed derivation tree
 whose root node is $\Ec \mathrel\cup t \teq u$ obtained by applying one of the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, \rn{Inject}, \rn{Unique}, or \rn{Single}.
@@
 $\MU y.\; \const{C}( y )$.
-The term $t$ of the form $\MU x.\; \const{C}( \MU y.\; \const{C}( x ) )$ is also abnormal,
+The term $t = \MU x.\; \const{C}( \MU y.\; \const{C}( x ) )$ is also abnormal,
 since $\MU y.\; \const{C}( x )$ is a self-similar subterm of $t$,
@@
 %and by replacing all of the abnormal subterms of $t$ with their corresponding normal forms.
-and by normalizing $t$'s subterms.
+and by normalizing the subterms.
 For example, $\nf{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) )} = \MU x.\; \const{C}( x )$.
@@
 \label{lem:mu-norm-arg}
-If $\vec u$ are in normal form and $t = \nf{\MU x.\; \const{C}( \vec u )} = \MU x.\; \const{C}( \vec w )$,
-then $\vec u \vsim \interp{t}{\vec w}$.
+If the $\mu$-terms $\vec u$ are in normal form and $t = \nf{\MU x.\; \const{C}( \vec u )} = \MU x.\; \const{C}( \vec w )$,
+then $\vec u \vsim \interp{t}{\vec w}$\vthinspace.
 \end{lemma}
@@
 \label{lem:mu-norm-interp}
-If $t$ is normal and has subterm $u$, then $\interp{t}{u}$ is normal.
+If the $\mu$-term $t$ is normal and $u$ is a subterm of $t$, then $\interp{t}{u}$ is normal.
 \end{lemma}
@@
 that is, a set of domain elements for that type.
-For a term $t$, $\M( t )$ denotes the interpretation of $t$ in $\M$,
-that is, an element of $\M( \tau )$ where $\tau$ is the type of $t$.
-\emph{All} types are interpreted as a set of $\mu$-terms.
-This is done to simplify the presentation.
+For a term $t\typ{\tau}$, $\M( t ) \in \M( \tau )$ denotes the interpretation of $t$ in $\M$.
+To simplify the presentation, all types are interpreted as sets of $\mu$-terms.
 Since all values of types in $\Data$ and $\Nondata$ are acyclic,
@@
 dropping each of its $\mu$-bindings. 
-For example, the $\mu$-term $\MU x.\; \const{0}$ representing the integer constant zero
-is simply $\const{0}$ after dropping the $\mu$-binding of $x$.
+For example, the $\mu$-term $\MU x.\; \const{Z}$ representing the integer constant zero
+is simply $\const{Z}$ after dropping the $\mu$-binding of $x$.
 

### 2014-12-23 tex/codatadp.tex
more

@@
 For a term $t\typ{\tau}$, $\M( t ) \in \M( \tau )$ denotes the interpretation of $t$ in $\M$.
-To simplify the presentation, all types are interpreted as sets of $\mu$-terms.
-Since all values of types in $\Data$ and $\Nondata$ are acyclic,
-a typical representation for these values can be obtained by
-dropping each of its $\mu$-bindings. 
-For example, the $\mu$-term $\MU x.\; \const{Z}$ representing the integer constant zero
-is simply $\const{Z}$ after dropping the $\mu$-binding of $x$.
+To simplify the presentation, all types are interpreted as sets of $\mu$-terms; but
+only values of types in $\Codata$ may contain cycles.
+%Since all values of types in $\Data$ and $\Nondata$ are acyclic,
+%a typical representation for these values can be obtained by
+%dropping each of its $\mu$-bindings. 
+%For example, the $\mu$-term $\MU x.\; \const{Z}$ representing the integer constant zero
+%is simply $\const{Z}$ after dropping the $\mu$-binding of $x$.
 

### 2014-12-23 tex/codatadp.tex
more

@@
 $\M( \tau )$ is a maximal set of closed normal $\mu$-terms of that type that are
-unique up to renaming of $\mu$-bound variables,
-and acyclic when $\tau \not\in \Codata$.
+unique up to $\alpha$-equivalence and acyclic when $\tau \not\in \Codata$.
 \item

### 2014-12-23 tex/codatadp.tex
more

@@
 \item
-For all types $\tau$,
+%%% TODO: What does maximal set mean? Does it make sense for \Nondata?
+For each type $\tau$,
 $\M( \tau )$ is a maximal set of closed normal $\mu$-terms of that type that are
@@
 \item
-For all constructor terms $\const{C}( \vec t \,)$ of type $\tau$,
+For each constructor term $\const{C}( \vec t \,)$ of type $\tau$,
 $\M( \const{C}( \vec t \,) )$ is the value
@@
 $\nf{\MU x.\; \const{C}( \M( \vec t \,) )}$, where $x$ is fresh.
+%%% TODO: I'm a bit worried to see a semantic value \M under a syntactic
+%%% constructor. I know our models consist of \mu-terms, but is this intended?
 \item

### 2015-02-02 tex/codatadp.tex
compile on new mac
more edits
more

@@
 \usepackage{amssymb}
-\usepackage{arydshln}
+%\usepackage{arydshln}
 \usepackage[scaled=.82]{beramono}
@@
 \begin{abstract}
-Codatatypes naturally capture potentially infinite data structures and
-processes. We present a decision procedure that combines reasoning about
+%Codatatypes naturally capture potentially infinite data structures and
+%processes.
+We present a decision procedure that combines reasoning about
 datatypes and codatatypes. The dual of the acyclicity rule for datatypes is a
@@
 \end{report}
-A related semantic view of datatypes is as initial algebras \cite{xxx}.
-Codatatypes are then defined dually as final coalgebras \cite{xxx}. The
-datatypes are generated by their constructors, whereas the codatatypes are
-viewed through their selectors.
+A related semantic view of datatypes is as initial algebras. Codatatypes are
+then defined dually as final coalgebras \cite{rutten-2000}. The datatypes are
+generated by their constructors, whereas the codatatypes are viewed through
+their selectors.
 %By uniformly focusing on the constructors, the
@@
 axiom schema, which states that any ground constructor term cannot be equal to
-a variable it contains \cite{xxx}. Dually, coinduction can be replaced by the
-uniqueness axiom schema, which asserts that (co)recursive definitions giving
-rise to the same infinite expansion---e.g., $x \teq \const{C}(x)$ and $y \teq
-\const{C}(\const{C}(y))$---must be equal.
+a variable it contains \cite{barrett-et-al-2007}. Dually, coinduction can be
+replaced by the uniqueness axiom schema, which asserts that (co)recursive
+definitions giving rise to the same infinite expansion---e.g., $x \teq
+\const{C}(x)$ and $y \teq \const{C}(\const{C}(y))$---must be equal.
 
@@
 \paragraph{Optimizations.}
-We first discuss several optimizations that are not reflected in the presentation of the calculus, which closely follow the approach outlined by Barrett et al in~\cite{}.
-We only briefly mention these optimizations, since each applies to datatypes as well as codatatypes, and most have been presented thoroughly in~\cite{}.
+We first discuss several optimizations that are not reflected in the presentation of the calculus, which closely follow the approach outlined by Barrett et al.\ \cite{barrett-et-al-2007}.
+We only briefly mention these optimizations, since each applies to datatypes as well as codatatypes, and most have been presented thoroughly in Barrett et al.
 
@@
 We have found that this leads to better performance since this reduction introduces terms more eagerly to $\tEc$.
-Handling discriminators requires extending the decision procedure with several additional rules (see~\cite{} for more details), which apply uniformly to both datatypes and codatatypes.
+Handling discriminators requires extending the decision procedure with several
+additional rules (see~Barrett et al.\ \cite{barrett-et-al-2007} for more
+details), which apply uniformly to both datatypes and codatatypes.
 Also, selectors are collapsed eagerly:
@@
 CVC4, running for up to \textbf{TODO: XXX} seconds on the StarExec cluster
-\cite{xxx}. Problems involving no (co)datatypes were filtered out.
-Due to the lack of machinery for reconstructing inferences about (co)datatypes
-in Isabelle, CVC4 is trusted as an oracle in these experiments.
+\cite{stump-et-al-2014-starexec}. Problems involving no (co)datatypes were
+filtered out. Due to the lack of machinery for reconstructing inferences about
+(co)datatypes in Isabelle, CVC4 is trusted as an oracle in these experiments.
 

### 2015-02-02 tex/body.tex
paper-agnostic name, cf. 'conf', 'report'

--- 2015-02-02 tex/codatadp.tex
@@
 using Peano notation but extended with an
-infinity value $\infty = \const{S}(\const{S}(\const{S}(\ldots)))$.
+infinite value $\infty = \const{S}(\const{S}(\const{S}(\ldots)))$.
 The equation $\const{S}(\infty) \teq \infty$ holds as expected,
@@
 Compared
-with the more conventional, inductive definition with a dedicated constructor
-for representing infinity,
-%\[\keyw{datatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat}) \,\mid\, \const{Infty}\]
-%
-the codatatype avoids one case by unifying the finite and infinite nonzero cases.
+with the conventional inductive definition
+\[\keyw{datatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat}) \,\mid\, \const{Infty}\]
+with a dedicated constructor for representing infinity, the codatatype avoids
+one case by unifying the finite and infinite nonzero cases.
 

### 2015-02-02 tex/body.tex
more consistent naming

@@
 
-\begin{report}
+\begin{rep}
 \def\thewordpaper{report}
 \renewcommand\dotReportFootnote[1]{.\footnote{#1}}
-\end{report}
+\end{rep}
 
@@
 
-Datatypes and codatatypes are an integral part of modern proof assistants,
+Datatypes and codatatypes are an integral part of most modern proof assistants,
 including Agda, Coq, Isabelle, Matita, and PVS. In recent years, datatypes
@@
 
-\begin{paper}
+\begin{conf}
 Proofs of soundness and completeness are included in the technical report
 associated with this paper \cite{our-report}.
-\end{paper}%
+\end{conf}%
 The procedure is implemented in the SMT solver CVC4 as a combination
@@
 structure, and interpreted by nonempty domains. The set~$\Types$ must contain a
-distinguished type \ty{bool}\begin{report} interpreted as the set of truth
-values $\{\bot, \top\}$\end{report}. %, and may contain other interpreted types (e.g., \ty{int}, \ty{real}).
-The only predicate is equality ($\teq$)\begin{report} and belongs to the logical symbols\end{report}.
+distinguished type \ty{bool}\begin{rep} interpreted as the set of truth
+values $\{\bot, \top\}$\end{rep}. %, and may contain other interpreted types (e.g., \ty{int}, \ty{real}).
+The only predicate is equality ($\teq$)\begin{rep} and belongs to the logical symbols\end{rep}.
 Other predicates can be represented as functions to $\ty{bool}$,
 with $\const{p}(\ldots)$ abbreviating $\const{p}(\ldots) \teq \const{true}$.
-\begin{report}
+\begin{rep}
 The metavariables $\delta,\:\varepsilon$ range over (co)datatypes,
 whereas $\tau,\:\upsilon$ range over arbitrary types.
-When applied to terms, the symbol $=$ denotes syntactic equality.\end{report}
-
-\begin{report}
+When applied to terms, the symbol $=$ denotes syntactic equality.\end{rep}
+
+\begin{rep}
 Function symbols are written in a sans-serif font (e.g., $\const{f}$, $\const{g}$) to
 distinguish them from variables (e.g., $x$, $y$).
-\end{report}
+\end{rep}
 Symbols starting with an uppercase letter (e.g.,
@@
 and a return type $\tau$. This connection can be expressed %compactly
-as $\const{f} : \tau_1 \times \cdots \times \tau_n \to \tau$\begin{report},
-which collapses to $\const{f} : \tau$ if $n = 0$\end{report}.
+as $\const{f} : \tau_1 \times \cdots \times \tau_n \to \tau$\begin{rep},
+which collapses to $\const{f} : \tau$ if $n = 0$\end{rep}.
 For a term $t$, the notation $t\typ{\tau}$ indicates that it has type $\tau$.
@@
 parentheses in terms.
-\begin{report}%
+\begin{rep}%
 The operator $\bigwedge_{\,i}\, \varphi_i$ abbeviates a conjunction
-$\varphi_1 \mathrel\land \cdots \mathrel\land \varphi_n$. \end{report}%
+$\varphi_1 \mathrel\land \cdots \mathrel\land \varphi_n$. \end{rep}%
 Finally, $[n]$ denotes the set $\{1,\ldots, n\}$, and
@@
 
-\begin{report}%
+\begin{rep}%
 Because all types must be inhabited (nonempty), a datatype specification is
@@
 finite or infinite, that witnesses nonemptiness \cite{blanchette-et-al-2015-esop}.
-\end{report}
+\end{rep}
 
@@
 such~as
-\begin{paper}%
+\begin{conf}%
 \vthinspace$\keyw{datatype}~\vthinspace\ty{option}_{\,\tau} = \const{None} \mid \const{Some}(\tau)$.%
-\end{paper}%
-\begin{report}%
+\end{conf}%
+\begin{rep}%
 \[\begin{aligned}[t]
@@
 introduced as datatypes or as codatatypes.
-\end{report}%
+\end{rep}%
 %Without loss of generality, we consider that these types are datatypes
@@
 selector--constructor equations \cite{barrett-et-al-2010}.
-\begin{report}
+\begin{rep}
 A drawback of this approach is that it does
@@
 
-\end{report}
+\end{rep}
 A related semantic view of datatypes is as initial algebras. Codatatypes are
@@
 %
-\begin{report}%
+\begin{rep}%
 Expressed in the algebraic jargon, exhaustiveness helps ensure that ``no
@@
 constructor is left completely unspecified.
-\end{report}
+\end{rep}
 
@@
 
-Datatypes are additionally characterized by an induction axiom schema\begin{report}:
+Datatypes are additionally characterized by an induction axiom schema\begin{rep}:
 %
@@
 denotes either $P_{\iPrime}(x)$ if there exists some $\iPrime$ such that
-the formula is type-correct or else $\top$\end{report}.
+the formula is type-correct or else $\top$\end{rep}.
 %
@@
 %
-\begin{report}\par\end{report}
-%
-For codatatypes, the dual notion is called coinduction. \begin{report}It depends on
+\begin{rep}\par\end{rep}
+%
+For codatatypes, the dual notion is called coinduction. \begin{rep}It depends on
 witnesses $R_i$ that are required to be bisimulations:
@@
 $\iPrime$ such that the formula is type-correct or $x \teq y$ otherwise.
-\end{report}%
+\end{rep}%
 This axiom schema guarantees that two values that yield the same
@@
 %
-\begin{report}\par\end{report}
+\begin{rep}\par\end{rep}
   %
@@
 infinite ground constructor terms.
-\begin{report}
+\begin{rep}
 In general, this cannot be captured by a
@@
 function space $\ty{nat} \to \ty{int}$.
-\end{report}
+\end{rep}
 
@@
 
-\begin{report}
+\begin{rep}
 TODO: State more precisely and proof?
-\end{report}
+\end{rep}
 
@@
 only value is $\MU a.\; A(a)$. Other specimens are
-\begin{paper}%
+\begin{conf}%
 \vthinspace$\keyw{codatatype}~\ty{b} = \const{B}(\ty{b},\: \ty{c},\: \ty{b},\: \ty{unit})
 ~\keyw{and}~ \ty{c} = \const{C}(\ty{a},\: \ty{unit},\: \ty{b},\: \ty{c})$,
-\end{paper}%
-\begin{report}%
+\end{conf}%
+\begin{rep}%
 \[\begin{aligned}[t]
@@
 \]
-\end{report}%
+\end{rep}%
 where \ty{unit} is a datatype with the single constructor $\const{Unity} :
@@
 
-\begin{report}
+\begin{rep}
 \begin{proof}
@@
 \end{proof}
-\end{report}
+\end{rep}
 
@@
 %$\MU x.\; \const{C}( x ) \aequiv \MU y.\; \const{C}( y )$ and
-$\MU x.\; \const{D}( y, x ) \aequiv \MU z.\; \const{D}( y, z )$\begin{report},
+$\MU x.\; \const{D}( y, x ) \aequiv \MU z.\; \const{D}( y, z )$\begin{rep},
 but
 $\MU x.\; \const{C}( x ) \not\aequiv \MU x.\; \const{D}( y, x ) \not\aequiv \MU x.\; \const{D}( z, x )
-\not\aequiv \MU y.\; \const{D}( y, x )$\end{report}.
+\not\aequiv \MU y.\; \const{D}( y, x )$\end{rep}.
 Two $\mu$-terms can represent the same value while not being $\alpha$-equivalent---e.g.,
@@
 
-\begin{report}
+\begin{rep}
 \begin{lemma}
@@
 %\end{proof}
-\end{report}
+\end{rep}
 
@@
 %
-%\begin{paper} 
+%\begin{conf} 
 %The proof of this lemma can be found in the extended version of this paper.
-%\end{paper}
+%\end{conf}
 
@@
 \end{lemma}
-\begin{report}
+\begin{rep}
 \begin{proof}

### 2015-02-02 tex/body.tex
more edits to Sect. 2

@@
 including Agda, Coq, Isabelle, Matita, and PVS. In recent years, datatypes
-have made their appearance in a few automatic theorem provers. The SMT-LIB~2
+have emerged in a few automatic theorem provers as well. The SMT-LIB~2
 \cite{barrett-et-al-2010} syntax, implemented by most SMT (satisfiability
@@
 problems involving datatypes and codatatypes in combination
-(Section~\ref{sec:a-ground-decision-procedure-for-co-datatypes})
-and discuss extensions  of the procedure for handling quantifiers
+(Section~\ref{sec:a-ground-decision-procedure-for-co-datatypes}).
+We also discuss extensions of the procedure for handling quantifiers
 (Section~\ref{sec:extension-to-quantified-formulas}).
@@
 
-Datatypes and codatatypes share many of the same properties, so it makes sense
+Datatypes and codatatypes share many properties, so it makes sense
 to consider them together. There are, however, at least three important
@@
 having the same possibly infinite expansion---must be equal; from $\const{x}
-\teq \const{S}(\const{y})$ and
-$\const{y} \teq \const{S}(\const{x})$, it deduces $\const{x} \teq \const{y}$.
+\teq \const{S}(\const{x})$ and
+$\const{y} \teq \const{S}(\const{y})$, it deduces $\const{x} \teq \const{y}$.
 The acyclicity and uniqueness rules are needed to ensure completeness on
 ground problems. %and the absence of spurious models.
-They cannot be replaced by
-finite axiomatizations, so they naturally belong in a decision procedure.
+They cannot be expressed as finite axiomatizations, so they naturally belong in
+a decision procedure.
 %    * in particular, acyclicity and uniqueness are necessary for some proofs
@@
 and for theory combinations. The $\mu$-binder notation associates a name with
-a (sub)term; it is used to represent cyclic values in the generated models and
+a subterm; it is used to represent cyclic values in the generated models and
 in the metatheory. For example,
@@
 \begin{conf}
-Proofs of soundness and completeness are included in the technical report
+Detailed proofs of soundness and completeness are included in the technical report
 associated with this paper \cite{our-report}.
@@
 (Section~\ref{sec:the-theory-solver}).
-%\textbf{TODO: CHECK WITH IMPLEMENTATION, but round to nearest 50 or 1000.}
+\rem{CHECK WITH IMPLEMENTATION, but round to nearest 50 or 1000.}
 It consists of about 2000 lines of \cpp{} code, among which 1600 are
@@
 
-\paragraph{Related Work.}
-Although it was written several years ago, the account of related work in
-Barrett et al.\ \cite{barrett-et-al-2007} is still a good starting point.
-Since then, datatypes have been added to the SMT solver
-Z3 in unpublished work by Leonardo de Moura and to a SPASS-like prototype
+%\paragraph{Related Work.}
+Although it was written several years ago,
+Barrett et al.\ \cite{barrett-et-al-2007} still provides a good account of
+related work. Since then, datatypes have been added to the SMT solver
+Z3 in unpublished work by Leonardo de Moura and to a SPASS-like prototypical
 superposition prover called Pirate by Daniel Wand.
@@
 induction in CVC4 by Reynolds and Kuncak \cite{reynolds-kuncak-2015} and in
-Pirate by Wand and Weidenbach \cite{wand-weidenbach-201x};
+the SPASS-like prototype by Wand and Weidenbach \cite{wand-weidenbach-201x};
 both methods naturally depend on a notion of datatype.
@@
 %       * (deal more directly with finite sorts than Barrett et al., Section 6.1)
-%     * SPASS-Pirate
 %     * additional ones since then (e.g. strings?)
@@
 Symbols starting with an uppercase letter (e.g.,
-$\const{C}, \const{S}$, $\const{SCons}$) are reserved for constructors. With each function symbol \const{f}
+$\const{S}$, $\const{Infty}$) are reserved for constructors. With each function symbol \const{f}
 is associated a list of argument types $\tau_1,\ldots,\tau_n$ (with $n \ge 0$)
@@
 
-We fix a signature $\Sig = (\Types, \Funcs)$. The types are partioned into
+We fix a signature $\Sig = (\Types, \Funcs)$. The types are partitioned into
 $\Types = \Data \mathrel{\uplus} \Codata \mathrel{\uplus} \Nondata$, where $\Data$ are the
@@
 because they can be abstracted away as variables when combining theories.
-Exceptionally, it is convenient to use natural number
-constants ($0$, $1$, \ldots)\ in examples.
+Exceptionally, it is convenient to use numeric constants ($0$, $1$, \ldots)\ in
+examples.
 
 %\paragraph{Specifications.}
-In an SMT problem, the signature is normally given by specifying the
-uninterpreted types in any order, the (co)datatypes with their constructors
+In an SMT problem, the signature is normally given by specifying first the
+uninterpreted types in any order, then the (co)datatypes with their constructors
 and selectors in groups of mutually (co)recursive groups of (co)datatypes, and
@@
 %
-A (co)datatype specification consists of $l$~mutually recursive types which are
+Each (co)datatype specification consists of $l$~mutually recursive types which are
 either all datatypes or all codatatypes. Polymorphic types, nested
@@
 type of an argument to one of $\delta$'s constructors. Semantically, a set of
-types is mutually (co)recursive if and only if the associated dependency graph
-is strongly connected. Types can be declared together as mutually
-(co)recursive even if they are not actually (co)recursive, but the
-semantic notion is more precise and is the one that interests us.
+types is \emph{mutually} (\emph{co})\emph{recursive} if and only if the
+associated dependency graph is strongly connected. Types can be declared
+together in a mutual fashion even if they are not actually (co)recursive.
+The semantic notion is more precise and is the one that interests us.
 %
@@
 The axiom schema ensures that the interpretation of datatypes
-is a standard model. For example,
-for a datatype of natural numbers constructed from $\const{Z}$ and $\const{S}$,
+is standard.
+For the natural numbers constructed from $\const{Z}$ and $\const{S}$,
 induction prohibits nonstandard models, which could contain cyclic values---e.g.,
@@
 
-Given a signature $\Sig$, we write $\thD$ to refer to the \emph{theory} of (co)datatypes,
+Given a signature $\Sig$, we write $\thD$ to refer to the \emph{theory of}
+(\emph{co})\emph{datatypes},
 which in addition to $\Sig$ defines a class of $\Sig$-interpretations $\mathbf{I}$,
 namely the ones that satisfy the %axioms --- existence of all infinite codatatype values cannot be axiomatized
-properties mentioned in this section.
-In particular, this means that each interpretation in~$\mathbf{I}$ has a fixed interpretation
-for constructor terms and correctly applied selector terms, but may differ on its interpretation for
-wrongly applied selector terms and variables.
-A formula $\varphi$ is $\thD$-satisfiable if there exists an interpretation in $\mathbf{I}$ that satisfies $\varphi$;
-otherwise, it is $\thD$-unsatisfiable.
+properties mentioned in this section. In particular, each interpretation
+in~$\mathbf{I}$ has a fixed interpretation for constructor terms and correctly
+applied selector terms, but may differ on its interpretation for wrongly applied
+selector terms and variables. A formula $\varphi$ is $\thD$-\emph{satisfiable} if there
+exists an interpretation in $\mathbf{I}$ that satisfies $\varphi$.
+% otherwise, it is $\thD$-unsatisfiable.  (I trust the readers' intelligence when
+% it comes to English prefixes like non-, in-, un-, and a-.)
 \rem{more?}
@@
 Some codatatypes are so degenerate as to have infinite values
-even though they are finite. The simplest example is
+even though they are finite. A simple example is
 \vthinspace$\keyw{codatatype}~\ty{a} = \const{A}(\ty{a})$, whose
 only value is $\MU a.\; A(a)$. Other specimens are
+$\ty{stream}_{\,\ty{unit}\,}$
+and
 \begin{conf}%
@@
 where \ty{unit} is a datatype with the single constructor $\const{Unity} :
-\ty{unit}$, as well as $\ty{stream}_{\,\ty{unit}\,}$. We call such types
+\ty{unit}$. We call such types
 \emph{corecursive singletons}. For the decision procedure, it will be
@@
 Fortunately, given cardinalities for the ordinary types, it is easy to
-characterize this degenerate case:
+characterize this degenerate case.
 

### 2015-02-02 tex/body.tex
more

@@
 \DeclareFontShape{OT1}{pzc}{m}{it}{<-> s * [1.10] pzcmi7t}{}
-\DeclareMathAlphabet{\mathscr}{OT1}{pzc}{m}{it}
-
 \DeclareMathAlphabet{\mathcal}{OT1}{pzc}{m}{it}
@@
 %%% before by other authors.
-%\newcommand{\thD}{\mathcal{CoDt}}
-%\newcommand{\thD}{\mathcal{TDC}}
-\newcommand{\thD}{\mathcal{DC}}
+\newcommand{\thD}{\mathcal{CD}}
 
@@
 
-\newcommand\Ctr{\Funcs_{\mathrm{ctr}}}
-\newcommand\Sel{\Funcs_{\mathrm{sel}}}
+\newcommand\Ctr{\Funcs_{\smash{\mathrm{ctr}}}}
+\newcommand\Sel{\Funcs_{\smash{\mathrm{sel}}}}
 %\newcommand\Plainfuncs{\Funcs_{\mathrm{other}}}
@@
 
-\pagebreak
+\pagebreak[3]
 
@@
 To every argument corresponds a selector. The names for the (co)data\-types, the
-constructors, and the selectors must be distinct and different from
-existing names%
+constructors, and the selectors must be
+\begin{conf}fresh\end{conf}\begin{rep}distinct and different from
+existing names\end{rep}%
 \dotReportFootnote{It can be convenient to specify the same selector
@@
 with
-$\const{C}_{i\negvthinspace j} : \tau_{i\negvthinspace j}^1\times\cdots\times\tau_{i\negvthinspace j}^{\,k_{\smash{i\negvthinspace j}}} \to \delta_i$
-and $\const{s}_{i\negvthinspace j}^{\,k} : \delta_i \to \tau_{i\negvthinspace j}^{\,k}$. Defaults are assumed for
-the selector names if they are omitted. Given a type $\delta$, its
-constructors and selectors are denoted by $\Ctr^\delta$ and $\Sel^\delta$,
-respectively.
+$\smash{\const{C}_{i\negvthinspace j} : \tau_{i\negvthinspace j}^1\times\cdots\times\tau_{i\negvthinspace j}^{\,k_{\smash{i\negvthinspace j}}} \to \delta_i}$
+and $\smash{\const{s}_{i\negvthinspace j}^{\,k} : \delta_i \to \tau_{i\negvthinspace j}^{\,k}}$.
+Defaults are assumed for the selector names if they are omitted.
+The $\delta$ constructors and selectors are denoted by $\Ctr^\delta$ and
+$\Sel^\delta$.
 
@@
 
+\begin{rep}%
 Here are a few examples of legal specifications of (co)datatype families:
@@
 
-\begin{rep}%
 Because all types must be inhabited (nonempty), a datatype specification is
@@
 namely the ones that satisfy the %axioms --- existence of all infinite codatatype values cannot be axiomatized
-properties mentioned in this section. In particular, each interpretation
-in~$\mathbf{I}$ has a fixed interpretation for constructor terms and correctly
-applied selector terms, but may differ on its interpretation for wrongly applied
-selector terms and variables. A formula $\varphi$ is $\thD$-\emph{satisfiable} if there
+properties mentioned in this section. In particular, the interpretations
+in~$\mathbf{I}$ share the same fixed interpretation for constructor terms and correctly
+applied selector terms but may differ on wrongly applied
+selector terms. \rem{and variables?}
+%%% @ANDY: Variables are normally not part of the interpretation, but given by
+%%% an additional environment. Or is this done differently in the SMT community?
+A formula $\varphi$ is $\thD$-\emph{satisfiable} if there
 exists an interpretation in $\mathbf{I}$ that satisfies $\varphi$.
@@
 replaced by the uniqueness axiom schema, which asserts that (co)recursive
-definitions giving rise to the same infinite expansion---e.g., $x \teq
+equations giving rise to the same infinite expansion---e.g., $x \teq
 \const{C}(x)$ and $y \teq \const{C}(\const{C}(y))$---must be equal.
@@
 \emph{corecursive singletons}. For the decision procedure, it will be
-important to detect these types. %, even if they rarely arise in practice.
+crucial to detect these types. %, even if they rarely arise in practice.
 A type may also be a corecursive singleton only in some models. If the example

### 2015-02-02 tex/body.tex
more

@@
 The rules operate on a finite set $\Ec$ of equalities and disequalities
-between $\Sig$-terms, where $\Ec$ is interpreted as a conjunction.
+between $\Sig$-terms, where $\Ec$ is taken as a conjunction.
 We write $\tEc$ to denote the set of $\Sig$-terms occurring in $\Ec$.
@@
 First, all codatatypes in $\Codata$ are corecursive. This is not restrictive
-because noncorecursive codatatypes can be considered as nonrecursive
+because noncorecursive codatatypes can be seen instead as nonrecursive
 datatypes.
@@
 on the cardinality of any uninterpreted type $\tau \in \Nondata$; thus,
-we may without loss of generality or completeness treat all types in $\Nondata$ as infinite.
+we may without loss of generality %or completeness
+treat all types in $\Nondata$ as infinite.
 
@@
 The conclusion of a rule either describes equalities to be added to $\Ec$
-or is $\bot$.
+or is $\bot$ (contradiction).
 % (in which case we call an application of it \emph{nonterminal})
@@
 A rule may have multiple conclusions separated by $\parallel$,
-denoting nondeterministic branching. The rules associated with a phase have
-priority over those of the subsequent phases. The rules are applied
-until a contradiction ($\bot$) is derived or until no rule can be applied to
-extend~$\Ec$ further.
+denoting nondeterministic branching. The rules belonging to a phase have
+priority over those of the subsequent phases. The rules are applied until $\bot$
+is derived or until no rule can be applied to extend~$\Ec$ further.
 

### 2015-02-02 tex/body.tex
more

@@
 
-\begin{figure}[bth]
+\begin{figure}[b!]
 \normalsize
@@
 $\MU x.\: \const{C}( \vec t\, )$ for some constructor $\const{C} \in \Ctr$ and
-$\mu$-terms $\vec t$ with the expected types.
-The variable $x$ need not occur freely under the binder's body, in which case the
-binder will sometimes be omitted in examples.
+$\mu$-terms $\vec t$ of the expected types.
+The variable $x$ need not occur freely in the binder's body, in which case
+the binder can be omitted.
 In the \rn{Acyclic} rule,
-$\FV( u )$ denotes the variables occurring freely in $u$.
+$\FV( u )$ denotes the variables occurring freely in~$u$.
 
@@
 Two $\mu$-terms can represent the same value while not being $\alpha$-equivalent---e.g.,
-$\MU x.\; \const{C}( x ) \not\aequiv \MU x.\; \const{C}( \const{C}( x ) )$.
+$\MU x.\; \const{C}( x ) \not\aequiv \MU y.\; \const{C}( \const{C}( y ) )$.
 %, or informally are \emph{observationally equivalent}.

### 2015-02-02 tex/body.tex
more
clarify

@@
 \newcommand\FV{\mathrm{FV}}
+
+\newcommand\betweenantes{\kern1.5em}
 
@@
 \inferrule{
-  \vec t \teq \vec u \in \Ec \quad \const f( \vec t \,),\, \const f( \vec u ) \in \tEc
+  \vec t \teq \vec u \in \Ec
+  \betweenantes
+  \const f( \vec t \,),\, \const f( \vec u ) \in \tEc
 }{
@@
   \const{C}( \vec t \,) \teq \const{D}( \vec u ) \in \Ec
-  \quad
+  \betweenantes
   \const{C} \not= \const{D}
@@
   t\typ{\delta}
-  \quad
+  \betweenantes
   \delta \in \Data
-  \quad
+  \betweenantes
   %\ttpath{\Ec}{\ec{t}}{\ec{t}} \neq \emptyset
   \Val \ec{t} = \MU x.\; u
-  \quad
+  \betweenantes
   x \in \FV( u )
@@
  t, u\typ{\delta}
- \quad
+ \betweenantes
  \delta \in \Codata
- \quad
+ \betweenantes
  %\tpath{\Ec}{\ec{t}} = \tpath{\Ec}{\ec{u}}\sigma \neq \emptyset
@@
 %
-For each equivalence class $\ec{x}$, we associate a fresh variable ${\Varec{x}}$ of the same type as $x$
+With each equivalence class $\ec{u}$, we associate a fresh variable $\Varec{u}$
+of the same type as $u$
 %not occurring in $\tEc$,
-and set $\Val\ec{x} = {\Varec{x}}$.
-Thus, there are initially no constraints on the values for any equivalence class $\ec{x}$.
+and set $\Val\ec{u} = {\Varec{u}}$.
+Thus, there are initially no constraints on the values for any equivalence class $\ec{u}$.
 The mapping $\Val$ is refined by applying the following unfolding rule exhaustively:\strut
@@
 \inferrule{
-  {\Varec{x}} \in \FV( \Val )
-  \quad
-  \const{C}( t_1, \ldots, t_n ) \in \ec{x}
-  \quad
+  {\Varec{u}} \in \FV( \Val )
+  \betweenantes
+  \const{C}( t_1, \ldots, t_n ) \in \ec{u}
+  \betweenantes
   \const{C} \in \Ctr
 }{
-  \Val := \Val \{ {\Varec{x}} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{t_1}}, \ldots, {\Varec{t_n}} ) \}
+  \Val := \Val [{\Varec{u}} \mapsto \MU {\Varec{u}}.\; \const{C}( {\Varec{t_1}}, \ldots, {\Varec{t_n}} )]
 }
@@
 $\FV( \Val )$ denotes the union of the free variables occurring in $\Val$'s range,
-and $\Val\{{\Varec{x}} \mapsto \ldots\}$ denotes the corresponding substitution.
+and $\Val[x \mapsto t]$ denotes the substitution of $t$ for $x$ in $\Val$'s range.
 It is easy to see that the height of terms produced as a result of the unfolding
@@
   t\typ{\delta} \in \tEc
-  \quad
+  \betweenantes
   \Ctr^\delta = \{ \const{C}_1, \ldots, \const{C}_m \}
@@
   t, u\typ{\delta} \in \tEc
-  \quad
+  \betweenantes
   \delta \in \Codata
-  \quad
+  \betweenantes
   \left|\delta\right| = 1
@@
   {\Varec{x}}\typ{\tau} \in \FV( \Val )
-  \quad
+  \betweenantes
   \MU {\Varec{x}}.\; t \aequiv v
-  \quad
+  \betweenantes
   v \in \M( \tau )
-  \quad
+  \betweenantes
   v \not\in \mathcal{T}^\M_{\Varec{x}}(\Val)

### 2015-02-02 tex/body.tex
more

@@
 $\thD$-satisfiability of sets of ground constraints over a fixed
-signature~$\Sig$. It is formulated as a calculus, whose rules are applied
-until saturation or contradiction.
+signature~$\Sig$. It is formulated as a calculus, whose derivation rules are
+applied until saturation or contradiction.
 %
@@
 \inferrule{
-  t\typ{\delta}
-  \betweenantes
+%  t\typ{\delta}
+%  \betweenantes
   \delta \in \Data
@@
   %\ttpath{\Ec}{\ec{t}}{\ec{t}} \neq \emptyset
-  \Val \ec{t} = \MU x.\; u
+  \Val \ec{t\typ{\delta}} = \MU x.\; u
   \betweenantes
@@
 \inferrule{
- t, u\typ{\delta}
- \betweenantes
+% t, u\typ{\delta}
+% \betweenantes
  \delta \in \Codata
@@
  %\tpath{\Ec}{\ec{t}} = \tpath{\Ec}{\ec{u}}\sigma \neq \emptyset
- \Val \ec{t} \aequiv \Val \ec{u}
+ \Val \ec{t\typ{\delta}} \aequiv \Val \ec{u}
 }{
@@
 
-Given the mapping $\Val$, the \rn{Acyclic} and \rn{Unique} rules can be stated as follows.
-For acyclicity, if $\ec{t\typ{\delta}}$ is an equivalence class such that $\delta$ is a datatype and
-whose %class of  -- "whose" already refers to "equivalence class"
-values $\Val \ec{t} = \MU x.\; u$ are cyclic,
-as indicated by the assumption $x \in \FV( u )$,
+Given the mapping $\Val$, the \rn{Acyclic} and \rn{Unique} rules operate as follows.
+For acyclicity, if $\ec{t\typ{\delta}}$ is a datatype equivalence class
+whose values $\Val \ec{t} = \MU x.\; u$ are cyclic
+(as indicated by the precondition $x \in \FV( u )$),
 then $\Ec$ is unsatisfiable.
-For uniqueness, if $\ec{t\typ{\delta}}$ and $\ec{u\typ{\delta}}$ are two equivalence classes such
-that $\delta$ is a codatatype and
-whose %classes of  -- "whose" already refers to "equivalence classes"
-values $\Val \ec{t}$ and $\Val \ec{u}$ are $\alpha$-equivalent,
+For uniqueness, if $\ec{t}$ and $\ec{u}$ are two codatatype equivalence classes
+whose values $\Val \ec{t}$ and $\Val \ec{u}$ are $\alpha$-equivalent,
 %for a capture-avoiding renaming of $\mu$-bound variables,
-then $t \teq u$. This rule may seem too restrictive, since
-$\MU x.\; \const{S}(x)$ and $\MU y.\; \const{S}(\const{S}(y))$ denote the same
-value despite being $\alpha$-disequivalent, but the construction also works in
-such cases, as illustrated by the following example.
+then $t \teq u$. This rule may seem overly restrictive, since
+$x \teq \const{S}(x)$ and $y \teq \const{S}(\const{S}(y))$ specify the same
+value despite having $\alpha$-disequivalent right-hand sides, but the
+construction also works in such cases, by first deriving the equality
+$y \teq \const{S}(y)$.
 
@@
 
-\begin{figure}[tbh]
+\begin{figure}[b!]
 \normalsize

### 2015-02-02 tex/body.tex
more

@@
 %for a capture-avoiding renaming of $\mu$-bound variables,
-then $t \teq u$. This rule may seem overly restrictive, since
-$x \teq \const{S}(x)$ and $y \teq \const{S}(\const{S}(y))$ specify the same
-value despite having $\alpha$-disequivalent right-hand sides, but the
-construction also works in such cases, by first deriving the equality
-$y \teq \const{S}(y)$.
+then $t \teq u$. This rule may seem too restrictive, since
+$\MU x.\; \const{S}(x)$ and $\MU y.\; \const{S}(\const{S}(y))$ specify the same
+value despite being $\alpha$-disequivalent, but even in such cases the rule will
+make some progress by discovering that the subterm $\const{S}(y)$ of $\MU y.\;
+\const{S}(\const{S}(y))$ must be equal to the entire term.
 
@@
 \begin{examplex}
-Let $\const{C} : $ $\tau \times \delta \rightarrow \delta$ be a constructor,
-and let $\Ec = \{ x \teq \const{C}(y,\const{C}(y,x)) \}$.
+Let $\Ec = \{ y \teq \const{S}(\const{S}(y)) \}$.
 After phase~1, the equivalence classes %of $\Ec$
 are
-$\{ y \}$, $\{ \const{C}(y,x) \}$, and $\{ x,\, \const{C}(y,\const{C}(y,x)) \}$.
+$\{ \const{S}(y) \}$ and $\{ y,\, \const{S}(\const{S}(y)) \}$.
 Constructing $\Val$ yields
-\[\begin{array}{rcrlrl}
-\Val \ec{x} & \,=\, &
-\MU {\Varec{x}}. & \const{C}( {\Varec{y}},\, & \smash{\MU {\VAREC{\const{C}(y, x)}}}. & \const{C}( {\Varec{y}},\, {\Varec{x}} ) ) \\
-\Val \ec{\const{C}(y,x)} & = &
-\MU {\VAREC{\const{C}(y,x)}}. & \const{C}( {\Varec{y}},\, & \MU {\Varec{x}}. & \const{C}( {\Varec{y}},\, {\VAREC{\const{C}(y,x)}} ) )
+\[\begin{array}{rcrl@{}rl}
+\Val \ec{y} & \,=\, &
+\MU {\Varec{y}}. & \const{S}( & \smash{\MU {\VAREC{\const{S}(y)}}}. & \const{S}( {\Varec{y}} ) ) \\
+\Val \ec{\const{S}(y)} & = &
+\MU {\VAREC{\const{S}(y)}}. & \const{S}( & \MU {\Varec{y}}. & \const{S}( {\VAREC{\const{S}(y)}} ) )
 \end{array}\]
-Since $\Val \ec{x} \aequiv \Val \ec{\const{C}(y,x)}$,
+Since $\Val \ec{y} \aequiv \Val \ec{\const{S}(y)}$,
 %for the renaming
-%$\{ {\Varec{x}} \mapsto {\VAREC{\const{C}(y,x)}}, {\VAREC{\const{C}(y,x)}} \mapsto {\Varec{x}} \}$,
-the \rn{Unique} rule applies to derive $x \teq \const{C}(y,x)$.
-%Intuitively, this equality holds, since the values of $x$ and $\const{C}( y, x )$
-%are equivalent for any value of ${\Varec{y}}$.
+%$\{ {\Varec{y}} \mapsto {\VAREC{\const{S}(y)}}, {\VAREC{\const{S}(y)}} \mapsto {\Varec{y}} \}$,
+the \rn{Unique} rule applies to derive $y \teq \const{S}(y)$.
+%Intuitively, this equality holds, since the values of $y$ and $\const{S}(y)$
+%are equivalent.
 \xend

### 2015-02-02 tex/body.tex
more
finished playing with Example 2

@@
 \begin{examplex}
-Let $\Ec = \{ y \teq \const{S}(\const{S}(y)) \}$.
+Let $\Ec = \{ x \teq \const{S}(x),\; y \teq \const{S}(\const{S}(y)) \}$.
 After phase~1, the equivalence classes %of $\Ec$
 are
-$\{ \const{S}(y) \}$ and $\{ y,\, \const{S}(\const{S}(y)) \}$.
+$\{ x,\, \const{S}(x) \}$,
+$\{ y,\, \const{S}(\const{S}(y)) \}$,
+and $\{ \const{S}(y) \}$.
 Constructing $\Val$ yields
-\[\begin{array}{rcrl@{}rl}
-\Val \ec{y} & \,=\, &
-\MU {\Varec{y}}. & \const{S}( & \smash{\MU {\VAREC{\const{S}(y)}}}. & \const{S}( {\Varec{y}} ) ) \\
-\Val \ec{\const{S}(y)} & = &
-\MU {\VAREC{\const{S}(y)}}. & \const{S}( & \MU {\Varec{y}}. & \const{S}( {\VAREC{\const{S}(y)}} ) )
-\end{array}\]
+\begin{align*}
+\Val \ec{x} & \,=\, \MU {\Varec{x}}.\; \const{S}(\Varec{x}) &
+\enskip %% TYPESETTING
+\Val \ec{y} & \,=\,
+\MU {\Varec{y}}.\; \const{S}(\smash{\MU {\VAREC{\const{S}(y)}}}.\; \const{S}( {\Varec{y}} ) )
+\enskip %% TYPESETTING
+ &
+\Val \ec{\const{S}(y)} & =
+\MU {\VAREC{\const{S}(y)}}.\; \const{S}(\MU {\Varec{y}}.\; \const{S}( {\VAREC{\const{S}(y)}} ) )
+\end{align*}
 Since $\Val \ec{y} \aequiv \Val \ec{\const{S}(y)}$,
@@
 %are equivalent.
+%
+Next, phase~1 is activated again to produce the equivalence classes
+$\{ x,\, \const{S}(x) \}$ and
+$\{ y,\, \const{S}(y),\, \const{S}(\const{S}(y)) \}$.
+The mapping $\Val$ is updated accordingly:
+\begin{align*}
+\Val \ec{x} & \,=\, \MU {\Varec{x}}.\; \const{S}(\Varec{x})
+&
+\Val \ec{y} & \,=\, \MU {\Varec{y}}.\; \const{S}({\Varec{y}})
+\end{align*}
+Finally, since $\Val \ec{x} \aequiv \Val \ec{y}$,
+%for the renaming
+%$\{ {\Varec{y}} \mapsto {\VAREC{\const{S}(y)}}, {\VAREC{\const{S}(y)}} \mapsto {\Varec{y}} \}$,
+the \rn{Unique} rule produces $x \teq y$.
 \xend

### 2015-02-02 tex/body.tex
more

@@
 \newcommand{\is}[1]{\const{is#1}}
-\newcommand{\ror}{\kern.45em \parallel \kern.45em}
+\newcommand{\ror}{\parallel}
+\newcommand{\ROR}{\kern.45em \ror \kern.45em}
 \newcommand{\tpath}[2]{\mathcal{P}_{ #2 \rightarrow \_ }( #1 )}
@@
 %\newcommand{\Varec}[1]{\Var{\ec{#1}}}
-\newcommand{\Varec}[1]{\widetilde{#1}}
+\newcommand{\Varec}[1]{\vvthinspace\widetilde{#1}\vvthinspace}
 \newcommand{\VAREC}[1]{\widetilde{\vphantom{\scriptstyle x^i}\smash{#1}}}
@@
 A rule may have multiple conclusions separated by $\parallel$,
-denoting nondeterministic branching. The rules belonging to a phase have
+denoting branching.
+
+The rules belonging to a phase have
 priority over those of the subsequent phases. The rules are applied until $\bot$
@@
 The mapping $\Val$ is updated accordingly:
+\begin{conf}%
+$\Val \ec{y} \,=\, \MU {\Varec{y}}.\; \const{S}({\Varec{y}})$.
+\end{conf}%
+\begin{rep}
 \begin{align*}
@@
 \end{align*}
+\end{rep}
 Finally, since $\Val \ec{x} \aequiv \Val \ec{y}$,
@@
 %$\{ {\Varec{y}} \mapsto {\VAREC{\const{S}(y)}}, {\VAREC{\const{S}(y)}} \mapsto {\Varec{y}} \}$,
-the \rn{Unique} rule produces $x \teq y$.
+the \rn{Unique} rule derives $x \teq y$.
 \xend
@@
 }{
-  \Ec := \Ec,\: t \teq \const{C}_1\bigl(\const s^1_1( t ), \ldots,\const s^{n_1}_1( t ) \bigr) \ror \cdots \ror \Ec := \Ec,\: t \teq \const{C}_m\bigl(\const s^1_m( t ), \ldots,\const s^{n_m}_m( t ) \bigr)
+  \Ec := \Ec,\: t \teq \const{C}_1\bigl(\const s^1_1( t ), \ldots,\const s^{n_1}_1( t ) \bigr) \ROR \cdots \ROR \Ec := \Ec,\: t \teq \const{C}_m\bigl(\const s^1_m( t ), \ldots,\const s^{n_m}_m( t ) \bigr)
 }
@@
 If a selector is applied to a term $t$, or if $t$'s type is a finite datatype,
-the equivalence class of $t$ must contain an application of one of
-$\delta$'s constructors.
+the equivalence class of $t$ must contain an application of a
+$\delta$ constructor.
 This is enforced in the third phase by the \rn{Split} rule.
@@
 terms from $\tEc$ have a corecursive singleton type
-(Section~\ref{sec:the-theory-of-co-datatypes}), in which case they must be
-equivalent. Both the finiteness assumption on the datatype in \rn{Split} and
-the singleton constraint on the codatatype in \rn{Single} can be 
-computed statically
+(Section~\ref{sec:the-theory-of-co-datatypes}) and must therefore be equal. Both
+the finiteness assumption on the datatype in \rn{Split} and the singleton
+constraint on the codatatype in \rn{Single} can be computed statically
 based on a recursive computation of the cardinalities of the

### 2015-02-02 tex/body.tex
more

@@
 \newcommand{\is}[1]{\const{is#1}}
-\newcommand{\ror}{\parallel}
+%\newcommand{\ror}{\parallel}
+\newcommand{\ror}{{\&}}
 \newcommand{\ROR}{\kern.45em \ror \kern.45em}
@@
 % (in which case we call an application of it \emph{terminal}).
-A rule may have multiple conclusions separated by $\parallel$,
+A rule may have multiple conclusions separated by $\ror$,
 denoting branching.
 
-The rules belonging to a phase have
-priority over those of the subsequent phases. The rules are applied until $\bot$
-is derived or until no rule can be applied to extend~$\Ec$ further.
+An application of a rule is \emph{redundant} if
+at least one branch in the conclusion is an assignment to $\Ec$ that leaves it
+unchanged.
+A \emph{derivation tree} is a tree whose nodes are finite sets of equalities and
+where each non-root node is obtained by
+a nonredundant application of a derivation rule to its parent node.
+A derivation tree is \emph{closed} if all of its leaf nodes are $\bot$.
+A node is \emph{saturated} if no nonredundant instance of a rule can be applied to it.
+
+The rules belonging to a phase have priority over those of the subsequent
+phases. The rules are applied until the derivation tree is closed or until no
+rule can be applied to extend~$\Ec$ further.
 
@@
 %where a rule can be applied to $\Ec$ if it meets all of the specified preconditions for $\Ec$.
-%The conclusion of a rule either describes equalities to be added to $\Ec$ (in which case we call an application of it \emph{nonterminal}),
-%or is $\bot$ (in which case we call an application of it \emph{terminal}).
-%A rule may have multiple conclusions separated by $\parallel$, which denotes nondeterministic branching.
 
@@
 \paragraph{Correctness.}
-An application of a rule is \emph{redundant} if
-at least one branch in the conclusion is an assignment to $\Ec$ that leaves it
-unchanged.
-A \emph{derivation tree} is a tree whose nodes are finite sets of equalities and
-where each non-root node is obtained by
-a nonredundant application of a derivation rule to its parent node.
-A derivation tree is \emph{closed} if all of its leaf nodes are $\bot$.
-A node is \emph{saturated} if no nonredundant instance of a rule can be applied to it.
-
 Correctness means that if there exists a closed derivation tree with root node

### 2015-02-02 tex/body.tex
more

@@
 
-\newcommand\betweenantes{\kern1.5em}
+\newcommand\betweenantes{\kern1.25em}
 
@@
 %\newcommand{\ror}{\parallel}
-\newcommand{\ror}{{\&}}
-\newcommand{\ROR}{\kern.45em \ror \kern.45em}
+%\newcommand{\ror}{{\&}}
+%\newcommand{\ROR}{\kern.45em \ror \kern.45em}
+\newcommand\ROR{\betweenantes}
 \newcommand{\tpath}[2]{\mathcal{P}_{ #2 \rightarrow \_ }( #1 )}
@@
 $\thD$-satisfiability of sets of ground constraints over a fixed
-signature~$\Sig$. It is formulated as a calculus, whose derivation rules are
-applied until saturation or contradiction.
-%
-The rules operate on a finite set $\Ec$ of equalities and disequalities
-between $\Sig$-terms, where $\Ec$ is taken as a conjunction.
-We write $\tEc$ to denote the set of $\Sig$-terms occurring in $\Ec$.
+signature~$\Sig$. It is formulated as a calculus.
+% whose derivation rules are applied until saturation or contradiction.
+%
+The derivation rules operate on a finite set $\Ec$ of equalities and
+disequalities between $\Sig$-terms, where $\Ec$ is taken as a conjunction. We
+write $\tEc$ to denote the set of $\Sig$-terms occurring in $\Ec$.
+
 %We will commonly denote tuples of terms $( t_1, \ldots, t_n )$ in bold font, as $\vec t$.
@@
 
-To simplify the presentation, we make the following assumptions about $\Sig$.
+A rule can be applied to $\Ec$ if all the specified preconditions are met.
+The conclusion of a rule either describes equalities to be added to $\Ec$
+or is $\bot$ (contradiction).
+% (in which case we call an application of it \emph{nonterminal})
+% (in which case we call an application of it \emph{terminal}).
+A rule may have multiple conclusions, % separated by $\ror$,
+denoting branching.
+%
+An application of a rule is \emph{redundant} if at least one branch in the
+conclusion is an assignment to $\Ec$ that leaves it unchanged. A
+\emph{derivation tree} is a finite tree whose nodes are finite sets of
+equalities and where child nodes are obtained by a nonredundant application of a
+derivation rule to the parent node. A derivation tree is \emph{closed} if all of
+its leaf nodes are $\bot$. A node is \emph{saturated} if no nonredundant
+instance of a rule can be applied to it.
+
+The calculus consists of three sets of rules, presented in
+\figuresname~\ref{fig:cc-rules} to \ref{fig:split-rule}, corresponding to three
+phases. The first phase computes the bidirectional closure of $\Ec$. The second
+phase makes inferences based on acyclicity (for datatypes) and uniqueness (for
+codatatypes). The third phase performs case splitting on constructors for
+various terms in $\tEc$.
+%Following the conventions from~\cite{},
+%The derivation rules are given in \relax{guarded assignment form},
+%where
+%
+The rules belonging to a phase have priority over those of the subsequent
+phases. Rules are applied until the derivation tree is closed or all leaf nodes
+are saturated.
+
+To simplify the presentation, we make the following assumptions about
+the signature $\Sig$.
 First, all codatatypes in $\Codata$ are corecursive. This is not restrictive
@@
 treat all types in $\Nondata$ as infinite.
-
-The calculus consists of three sets of rules, given in \figuresname~\ref{fig:cc-rules} to
-\ref{fig:split-rule}, corresponding to three phases. The first
-phase computes the bidirectional closure of $\Ec$. The second phase makes
-inferences based on acyclicity (for datatypes) and uniqueness (for
-codatatypes). The third phase performs case splitting on constructors for
-various terms in $\tEc$.
-%Following the conventions from~\cite{},
-%The derivation rules are given in \relax{guarded assignment form},
-%where
-A rule can be applied to $\Ec$ if all the specified preconditions are met.
-The conclusion of a rule either describes equalities to be added to $\Ec$
-or is $\bot$ (contradiction).
-% (in which case we call an application of it \emph{nonterminal})
-% (in which case we call an application of it \emph{terminal}).
-A rule may have multiple conclusions separated by $\ror$,
-denoting branching.
-
-An application of a rule is \emph{redundant} if
-at least one branch in the conclusion is an assignment to $\Ec$ that leaves it
-unchanged.
-A \emph{derivation tree} is a tree whose nodes are finite sets of equalities and
-where each non-root node is obtained by
-a nonredundant application of a derivation rule to its parent node.
-A derivation tree is \emph{closed} if all of its leaf nodes are $\bot$.
-A node is \emph{saturated} if no nonredundant instance of a rule can be applied to it.
-
-The rules belonging to a phase have priority over those of the subsequent
-phases. The rules are applied until the derivation tree is closed or until no
-rule can be applied to extend~$\Ec$ further.
 
@@
 }{
-  \Ec := \Ec,\: t \teq \const{C}_1\bigl(\const s^1_1( t ), \ldots,\const s^{n_1}_1( t ) \bigr) \ROR \cdots \ROR \Ec := \Ec,\: t \teq \const{C}_m\bigl(\const s^1_m( t ), \ldots,\const s^{n_m}_m( t ) \bigr)
+  \Ec := \Ec,\: t \teq \const{C}_1\bigl(\const s^1_1( t ), \ldots,\const s^{n_1}_1( t ) \bigr) %\ROR
+  \ROR \cdots \ROR \Ec := \Ec,\: t \teq \const{C}_m\bigl(\const s^1_m( t ), \ldots,\const s^{n_m}_m( t ) \bigr)
 }

### 2015-02-02 tex/body.tex
more

@@
 conclusion is an assignment to $\Ec$ that leaves it unchanged. A
-\emph{derivation tree} is a finite tree whose nodes are finite sets of
+\emph{derivation tree} is a %finite
+tree whose nodes are finite sets of
 equalities and where child nodes are obtained by a nonredundant application of a
@@
 then $\Ec$ is unsatisfiable.
-For uniqueness, if $\ec{t}$ and $\ec{u}$ are two codatatype equivalence classes
-whose values $\Val \ec{t}$ and $\Val \ec{u}$ are $\alpha$-equivalent,
+For uniqueness, if $\ec{t}$, $\ec{u}$ are two codatatype equivalence classes
+whose values $\Val \ec{t}$, $\Val \ec{u}$ are $\alpha$-equivalent,
 %for a capture-avoiding renaming of $\mu$-bound variables,
@@
 $\MU x.\; \const{S}(x)$ and $\MU y.\; \const{S}(\const{S}(y))$ specify the same
-value despite being $\alpha$-disequivalent, but even in such cases the rule will
+value despite being $\alpha$-disequivalent, but the rule will
 make some progress by discovering that the subterm $\const{S}(y)$ of $\MU y.\;
@@
 If a selector is applied to a term $t$, or if $t$'s type is a finite datatype,
-the equivalence class of $t$ must contain an application of a
-$\delta$ constructor.
+the equivalence class of $t$ must contain a
+$\delta$ constructor term.
 This is enforced in the third phase by the \rn{Split} rule.
@@
 based on a recursive computation of the cardinalities of the
-constructors' argument types.
+constructors' argument types (and assuming all types in $\Nondata$ are
+infinite).
 

### 2015-02-10 tex/body.tex
more
minor

@@
 
+\vskip3\smallskipamount
+
+\noindent\hfill\emph{In memoriam Morgan Deters 1979--2015}
+
 \begin{abstract}
@@
 both methods naturally depend on a notion of datatype.
+% \textbf{TODO:} http://frocos2013.loria.fr/slides/Peltier.pdf
+% \textbf{TODO:} Others (or at least add "e.g.")
+
+%\textbf{TODO}:
+%  * mu-calculus
+%  * Klop et al. (streams) Andrew Lice % http://joerg.endrullis.de/
+%  * Ro\c{s}u --- equality of streams is...
+%     http://dl.acm.org/citation.cfm?doid=1159803.1159827
 
@@
 Cesare Tinelli, and Christoph Weidenbach, have either encouraged the research on
-codatatype or at least endured
-%tolerated?
-it, both of which we are thankful for.
+codatatype or at least benevolently tolerated it, both of which we are thankful
+for.
 %

### 2015-02-13 tex/body.tex
testing hg-git
more
Minor fixes, work on proof.  Rearrange implementation section.

@@
 \usepackage{amssymb}
-%\usepackage{arydshln}
 \usepackage[scaled=.82]{beramono}
@@
 constructors and selectors. All properties below are implicitly universally
-quantified and range over all $i$, $j$, $\jPrime$, and $k$ within bounds:
+quantified and range over all $i$, $j$, $\jPrime$, and $k$ within bounds:\rem{Introduce that $\is{C}$ is notation for discriminators.}
 % and to all possible splits of the
@@
 applied selector terms but may differ on wrongly applied
-selector terms. \rem{and variables?}
+selector terms and variables.
 %%% @ANDY: Variables are normally not part of the interpretation, but given by
@@
 can be replaced by an enumeration datatype with $n$~nullary constructors.
-Moreover, since the constraints are ground, they cannot entail an upper bound
+Moreover, since the constraints in $\Ec$ are ground, they cannot entail an upper bound
 on the cardinality of any uninterpreted type $\tau \in \Nondata$; thus,
@@
 }{
-  \Val := \nf{\Val \{ {\Varec{x}} \mapsto \MU {\Varec{x}}.\; t \}}
+  \Val := \nf{\Val [ {\Varec{x}} \mapsto \MU {\Varec{x}}.\; t ]}
 }
@@
 
+Intuitively, this lemma states three properties of $\ValC$ which ensure a normal model $\M$ can be trivially constructed for $\Ec$.
+Part~\PART{1} states that the values in the range of $\ValC$ are $\alpha$-equivalent to a value in normal model.
+In particular, this means they are closed, normal, and acyclic when required.
+Part~\PART{2} states that the interpretation of all subterms in the range of $\ValC$ depends on its associated variable only.
+In other words, the interpretation of a subterm $t$ where $\muvar(t) = {\Varec{x}}$ is equal to $\ValC \ec{x}$, independent of the context in which it is interpreted.
+Part~\PART{3} states that $\ValC$ is injective, which in the following ensures that a distinct values can be associated with each equivalence class in a corresponding model.
+
 \begin{theorem}[Solution Soundness]%
@@
 and interprets the variables of $\Fc$ based on the completion $\ValC$ of the mapping $\Val$ from phase~2.
-For the latter, let $\M( x )$ be the value in $\M$ that is $\alpha$-equivalent with $\ValC \ec{x}$ for each variable $x \in \mathcal{T}(\Fc)$.
+For the latter, let $\M( x )$ be the value in $\M$ that is $\alpha$-equivalent with $\ValC \ec{x}$ for each variable $x \in \mathcal{T}(\Fc)$,
+which by Lemma~\ref{lem:model-completion}\PART{1} is guaranteed to exist.
 
@@
 where for each $i \in [n]$,
-$\muvar(w_i) = {\Varec{u_i}}$ and thus by Lemma~\ref{lem:model-completion}\PART{2}, 
+$\muvar(w_i) = {\Varec{u_i}}$.  Thus by Lemma~\ref{lem:model-completion}\PART{2},
 $\interp{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$.
@@
 which is $\ValC \ec{t}$.
-If $j \neq \jPrime$, we are free to interpret $\smash{\const s^k_{\negvthinspace j}( \M( u ) )}$ as the value in $\M( \tau )$ that is $\alpha$-equivalent with $\ValC \ec{ t }$.
-Since \rn{Cong} does not apply and due to Lemma~\ref{lem:model-completion}\PART{3},
-pairs of wrongly applied selectors are assigned the same value if and only if they reside in the same equivalence class,
-and thus this definition is well-defined.
+If $j \neq \jPrime$, since \rn{Cong} does not apply,
+$\smash{\const s^k_{\negvthinspace j}( u )}$ and $\smash{\const s^k_{\negvthinspace j}( u' )}$
+are in distinct equivalence classes only if $u$ and $u'$ are in distinct equivalence classes.
+Due to the induction hypothesis and Lemma~\ref{lem:model-completion}\PART{3}, $\M( u ) \neq \M( u' )$ for all such $u, u'$.
+Thus, we may interpret $\smash{\M(\const s^k_{\negvthinspace j})( \M( u ) )}$ as the value in $\M( \tau )$ that is $\alpha$-equivalent with $\ValC \ec{ t }$.
 
@@
 This section describes how the calculus in Section~\ref{sec:a-ground-decision-procedure-for-co-datatypes} is implemented within the SMT solver CVC4.
-
-\paragraph{Optimizations.}
-We first discuss several optimizations that are not reflected in the presentation of the calculus, which closely follow the approach outlined by Barrett et al.\ \cite{barrett-et-al-2007}.
-We only briefly mention these optimizations, since each applies to datatypes as well as codatatypes, and most have been presented thoroughly in Barrett et al.
-
-Discriminators are treated as predicate symbols, so a discriminator $\const{d}_{\negvthinspace j}( t )$ is such that $\const{d}_{\negvthinspace j}$
-is a predicate symbol, instead of reducing this constraint to $t \teq \const{C}_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^n_{\negvthinspace j}( t ) \bigr)$.
-We have found that this leads to better performance since this reduction introduces terms more eagerly to $\tEc$.
-Handling discriminators requires extending the decision procedure with several
-additional rules (see~Barrett et al.\ \cite{barrett-et-al-2007} for more
-details), which apply uniformly to both datatypes and codatatypes.
-Also, selectors are collapsed eagerly:
-if $\const s^k_{\negvthinspace j}( t )$ is a term in $\tEc$ and $t$ contains the constructor
-$\const{C}_{\negvthinspace j}( \vec u )$, we immediately infer $\const s^k_{\negvthinspace j}( t ) \teq u_k$, whereas the calculus we presented would apply \rn{Split} and \rn{Inject} before inferring this equality.
-For the sake of reducing the number of unique constraints considered by the calculus, we compute a normal form for constraints as a preprocessing step.
-In particular, we
-replace $t \teq s$ by $s \teq t$ if $s$ is less than $t$ based on a term ordering,
-replace $\const{C}_{\negvthinspace j}( \vec t ) \teq \const{C}_{\negvthinspace\jPrime}( \vec u )$ with $\bot$ when $j \neq \jPrime$,
-replace all selector terms of the form $\const{s}^{\,k}_{\negvthinspace j}( \const{C}_{\negvthinspace j}(t_1,\ldots,t_n) )$ by $t_k$,
-and replace all occurrences of discriminators $\const{d}_{\negvthinspace j}( \const{C}_{\negvthinspace\jPrime}( \vec t ) )$ by $\top$ if $j = \jPrime$ and $\bot$ otherwise.
 
@@
     * theory combination more complicated (and Theorem Completeness too weak)
+
+\paragraph{Optimizations.}
+Our implementation uses several optimizations not reflected in the presentation of the calculus, which closely follow the approach outlined by Barrett et al.\ \cite{barrett-et-al-2007}.
+We only briefly mention these optimizations, since each applies to datatypes as well as codatatypes, and most have been presented thoroughly in Barrett et al.
+
+Discriminators are treated as predicate symbols, so a discriminator $\const{d}_{\negvthinspace j}( t )$ is such that $\const{d}_{\negvthinspace j}$
+is a predicate symbol, instead of reducing this constraint to $t \teq \const{C}_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^n_{\negvthinspace j}( t ) \bigr)$.
+We have found that this leads to better performance since this reduction introduces terms more eagerly to $\tEc$.
+Handling discriminators requires extending the decision procedure with several
+additional rules (see~Barrett et al.\ \cite{barrett-et-al-2007} for more
+details), which apply uniformly to both datatypes and codatatypes.
+Also, selectors are collapsed eagerly:
+if $\const s^k_{\negvthinspace j}( t )$ is a term in $\tEc$ and $t$ contains the constructor
+$\const{C}_{\negvthinspace j}( \vec u )$, we immediately infer $\const s^k_{\negvthinspace j}( t ) \teq u_k$, whereas the calculus we presented would apply \rn{Split} and \rn{Inject} before inferring this equality.
+For the sake of reducing the number of unique constraints considered by the calculus, we compute a normal form for constraints as a preprocessing step.
+In particular, we
+replace $t \teq s$ by $s \teq t$ if $s$ is less than $t$ based on a term ordering,
+replace $\const{C}_{\negvthinspace j}( \vec t ) \teq \const{C}_{\negvthinspace\jPrime}( \vec u )$ with $\bot$ when $j \neq \jPrime$,
+replace all selector terms of the form $\const{s}^{\,k}_{\negvthinspace j}( \const{C}_{\negvthinspace j}(t_1,\ldots,t_n) )$ by $t_k$,
+and replace all occurrences of discriminators $\const{d}_{\negvthinspace j}( \const{C}_{\negvthinspace\jPrime}( \vec t ) )$ by $\top$ if $j = \jPrime$ and $\bot$ otherwise.
 

### 2015-02-16 tex/body.tex
some edits

@@
 (Section~\ref{sec:the-theory-solver}).
-\rem{CHECK WITH IMPLEMENTATION, but round to nearest 50 or 1000.}
+\rem{CHECK WITH IMPLEMENTATION, but round to nearest 50 or 100 LOC.}
 It consists of about 2000 lines of \cpp{} code, among which 1600 are
@@
 can be replaced by an enumeration datatype with $n$~nullary constructors.
-Moreover, since the constraints in $\Ec$ are ground, they cannot entail an upper bound
-on the cardinality of any uninterpreted type $\tau \in \Nondata$; thus,
+Finally, since the constraints in $\Ec$ are ground, they cannot entail an upper
+bound on the cardinality of any uninterpreted type $\tau \in \Nondata$; thus,
 we may without loss of generality %or completeness
@@
 %not occurring in $\tEc$,
-and set $\Val\ec{u} = {\Varec{u}}$.
-Thus, there are initially no constraints on the values for any equivalence class $\ec{u}$.
-The mapping $\Val$ is refined by applying the following unfolding rule exhaustively:\strut
+and let $\Val\ec{u} = {\Varec{u}}$. Therefore, there are initially no
+constraints on the values for any equivalence class $\ec{u}$. The mapping $\Val$
+is refined by applying the following unfolding rule exhaustively:\strut
 \[
@@
 Given the mapping $\Val$, the \rn{Acyclic} and \rn{Unique} rules operate as follows.
-For acyclicity, if $\ec{t\typ{\delta}}$ is a datatype equivalence class
+For acyclicity, if $\ec{t}$ is a datatype equivalence class
 whose values $\Val \ec{t} = \MU x.\; u$ are cyclic
-(as indicated by the precondition $x \in \FV( u )$),
+(as required by the precondition $x \in \FV( u )$),
 then $\Ec$ is unsatisfiable.
@@
 %for a capture-avoiding renaming of $\mu$-bound variables,
-then $t \teq u$. This rule may seem too restrictive, since
+then $t \teq u$. This may seem too restrictive, since
 $\MU x.\; \const{S}(x)$ and $\MU y.\; \const{S}(\const{S}(y))$ specify the same
@@
 %
-Next, phase~1 is activated again to produce the equivalence classes
+At this point, phase~1 is activated again to refine the equivalence classes,
+yielding
 $\{ x,\, \const{S}(x) \}$ and
-$\{ y,\, \const{S}(y),\, \const{S}(\const{S}(y)) \}$.
+$\{ y,\, \const{S}(y),\allowbreak\, \const{S}(\const{S}(y)) \}$.
 The mapping $\Val$ is updated accordingly:
@@
 \end{rep}
-Finally, since $\Val \ec{x} \aequiv \Val \ec{y}$,
+Since $\Val \ec{x} \aequiv \Val \ec{y}$,
 %for the renaming
 %$\{ {\Varec{y}} \mapsto {\VAREC{\const{S}(y)}}, {\VAREC{\const{S}(y)}} \mapsto {\Varec{y}} \}$,
-the \rn{Unique} rule derives $x \teq y$.
+the \rn{Unique} rule can finally derive $x \teq y$.
 \xend
@@
 based on a recursive computation of the cardinalities of the
-constructors' argument types (and assuming all types in $\Nondata$ are
+constructors' argument types (assuming all types in $\Nondata$ are
 infinite).

### 2015-02-16 tex/body.tex
systematically write thD-(un)sat

@@
 \newcommand{\rem}[1]{\textcolor{red}{[#1]}}
-%\newcommand{\is}[1]{is\text{-} #1}
-\newcommand{\is}[1]{\const{is#1}}
 %\newcommand{\ror}{\parallel}
@@
 %%% before by other authors.
-\newcommand{\thD}{\mathcal{CD}}
+\newcommand{\thD}{\mathcal{DC}}
 
@@
 \begin{conf}%
-\vthinspace$\keyw{datatype}~\vthinspace\ty{option}_{\,\tau} = \const{None} \mid \const{Some}(\tau)$.%
+\vthinspace$\keyw{datatype}~\vthinspace\ty{option}_{\,\tau} \,=\, \const{None} \,\mid\, \const{Some}(\tau)$.%
 \end{conf}%
@@
 \[\begin{aligned}[t]
-      \keyw{datatype}\;\, \ty{option}_{\,\tau} & \,=\, \const{None} \mid \const{Some}(\tau) \\[-.5\jot]
-      \keyw{codatatype}\; \,\ty{complex} & \,=\, \const{Complex}(\const{re}{:}\; \ty{real},\; \const{im}{:}\;\ty{real})
+      \keyw{datatype}~\vthinspace\ty{option}_{\,\tau\!} & \,=\, \const{None} \,\mid\, \const{Some}(\tau) \\[-.5\jot]
+      \keyw{codatatype}~\vthinspace\ty{complex} & \,=\, \const{Complex}(\const{re}{:}\; \ty{real},\; \const{im}{:}\;\ty{real})
 \end{aligned}
@@
 this gives rise to spurious equalities between unrelated terms---e.g.,
-$\const{s}_{11}(\const{C}_2) \teq \const{s}_{11}(\const{C}_3)$. This flaw
+$\const{s}_1^1(\const{C}_2) \teq \const{s}_1^1(\const{C}_3)$. This flaw
 could be corrected, but the added complexity seems to suggest that selectors
@@
 constructors and selectors. All properties below are implicitly universally
-quantified and range over all $i$, $j$, $\jPrime$, and $k$ within bounds:\rem{Introduce that $\is{C}$ is notation for discriminators.}
+quantified and range over all $i$, $j$, $\jPrime$, and $k$ within bounds:
 % and to all possible splits of the
@@
 \text{Exhaustiveness:}\quad
-  & \smash{\is{C}_{i1}(x) \mathrel\lor \cdots \mathrel\lor \is{C}_{im_i}(x)}
+  & \smash{\const{d}_{i1}(x) \mathrel\lor \cdots \mathrel\lor \const{d}_{im_i}(x)}
   \\[-.5\jot]
@@
 \begin{aligned}[t]
-\text{Coinduction:}\kern.8em %%% TYPESETTING: should be \quad
-\AXC{\strut$\begin{gathered}\textstyle \bigwedge_{\,i}\, R_i[v_i, w_i] \\[-\jot]\textstyle
+\text{Coinduction:}\quad
+\AXC{\strut$\begin{gathered}\textstyle \smash{\bigwedge_{\,i}\, R_i[v_i, w_i]} \\[-\jot]\textstyle
 \bigwedge_{\,i}\, \forall v\; w.\;\, R_i[v, w] \longrightarrow
@@
 
-Given a signature $\Sig$, we write $\thD$ to refer to the \emph{theory of}
-(\emph{co})\emph{datatypes},
+Given a signature $\Sig$, we write $\thD$ to refer to the \emph{theory of
+datatypes and codatatypes},
 which in addition to $\Sig$ defines a class of $\Sig$-interpretations $\mathbf{I}$,
@@
 (as required by the precondition $x \in \FV( u )$),
-then $\Ec$ is unsatisfiable.
+then $\Ec$ is $\thD$-unsatisfiable.
 For uniqueness, if $\ec{t}$, $\ec{u}$ are two codatatype equivalence classes
@@
 
-\begin{figure}[b!]
+\begin{figure}[t!]
 \normalsize
@@
 \end{tabular}
+%%% TODO: Clarify s's?
 \caption{\,Derivation rules for branching%.
@@
 the finiteness assumption on the datatype in \rn{Split} and the singleton
-constraint on the codatatype in \rn{Single} can be computed statically
+constraint on the codatatype in \rn{Single} can be evaluated statically
 based on a recursive computation of the cardinalities of the
-constructors' argument types (assuming all types in $\Nondata$ are
-infinite).
+constructors' argument types. This step relies on the assumption that all types
+in $\Nondata$ are infinite.
 
@@
 \begin{proof}
-Consider a derivation tree with root node $\Ec$.
-Let $\SSS \subseteq \tEc$ be the set of terms occurring as arguments of selectors.
-Let $T \subseteq \tEc$ be the set of terms whose types are finite datatypes.
-For each term $t \in T$\negvthinspace,
-let
+Consider a derivation tree with root node $\Ec$. Let $D \subseteq \tEc$ be the
+set of terms whose types are finite datatypes, and let $\SSS \subseteq \tEc$ be
+the set of terms occurring as arguments to selectors. For each term $t \in
+D$\negvthinspace, let
 \begin{align*}
 \smash{\SSSS 0_t} & = \{ t \}
-& \smash{\SSSS {i+1}_t} & = \smash{\SSSS i_t \mathrel\cup \{ \const s( u ) \mid u\typ{\delta} \in \SSSS i{,}~ \delta \in \Data{,}~ \left|\delta\right| \text{ is finite}{,}~ \const{s} \in \Sel^{\delta}}  \}
+& \smash{\SSSS {i+1}_t} & = \smash{\SSSS i_t \mathrel\cup \{ \const s( u ) \mid u\typ{\delta} \in \SSSS i_t{,}~ \delta \in \Data{,}~ \left|\delta\right| \text{ is finite}{,}~ \const{s} \in \Sel^{\delta}}  \}
 \end{align*}
@@
 and let $\SSSS \AST_t$ be the limit of this sequence.
-This is a finite set for each $t$ because the values of the type of $t$ are of finite size.
-Let $\SSSS \AST$ be the union of all sets $\SSSS \AST_t$ where $t \in T$,
-and let $\tcEc$ be the set of subterms of $\Ec \mathrel\cup \{ \const C_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\smash{\,n_{\negvthinspace j}}}_{\negvthinspace j}( t ) \bigr) \mid t\typ{\delta} \in \SSS \mathrel\cup \SSSS \AST{,}~ \const C_{\negvthinspace j} \in \Ctr^\delta \}$.
+This is a finite set for each $t$ because the values of the type of $t$ are of finite size\rem{Clarify}.
+Let $\SSSS \AST$ be the union of all sets $\SSSS \AST_t$ where $t \in D$,
+and let $\tcEc$ be the set of subterms of
+%
+$$\Ec \mathrel\cup \{ \const C_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\smash{\,n_{\negvthinspace j}}}_{\negvthinspace j}( t ) \bigr) \mid t\typ{\delta} \in \SSS \mathrel\cup \SSSS \AST{,}~ \const C_{\negvthinspace j} \in \Ctr^\delta \}$$
+%
 In a derivation tree with root node~$\Ec$,
-it can be shown by case analysis on the rules of the calculus that each non-root node $\Fc$ is such that
+it can be shown by %case analysis
+induction on the rules of the calculus that each non-root node $\Fc$ is such that
 $\mathcal{T}(\Fc) \subseteq \tcEc$, and hence contains an equality between two terms from $\tcEc$ not occurring in its parent node.
@@
 For \rn{Clash}, this is a consequence of distinctness.
-For \rn{Acyclic}, the construction of $\Val$ indicates that the class of values that term~$t$ can take in models of $\Ec$ is infeasible,
-and thus $\Ec$ is unsatisfiable.
+For \rn{Acyclic}, the construction of $\Val$ indicates that the class of values that term~$t$ can take in models of $\Ec$ is infeasible.
 If the child nodes of $\Ec$ are closed derivation trees
-whose root nodes are the result of applying the rule \rn{Split} on $t\typ{\delta}$,
-by the induction hypothesis $\Ec \mathrel\cup t \teq \smash{\const C_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\,\smash{n_{\negvthinspace j}}}_{\negvthinspace j}( t ) \bigr)}$ is unsatisfiable
+whose root nodes are the result of applying \rn{Split} on $t\typ{\delta}$,
+by the induction hypothesis $\Ec \mathrel\cup t \teq \smash{\const C_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\,\smash{n_{\negvthinspace j}}}_{\negvthinspace j}( t ) \bigr)}$ is
+$\thD$-unsatisfiable
 for each $\const C_{\negvthinspace j} \in \smash{\Ctr^\delta}$.

### 2015-02-16 tex/body.tex
more edits

@@
 %\newcommand{\interp}[2]{\lceil #2 \rceil_{#1}}
-\newcommand{\interp}[2]{\left< \smash{#2} \right>_{#1}}
+\newcommand{\interp}[2]{\langle \smash{#2} \rangle_{#1}}
 \newcommand{\dpath}[3]{\delta^{#1}_{#2}( #3 )}
@@
 Otherwise, the child node of $\Ec$ is a closed derivation tree
-whose root node is $\Ec \mathrel\cup t \teq u$ obtained by applying one of the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, \rn{Inject}, \rn{Unique}, or \rn{Single}.
-In all cases, we have that $\Ec \models_{\thD} t \teq u$.
+whose root node $\Ec \mathrel\cup t \teq u$ is obtained by applying one of the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, \rn{Inject}, \rn{Unique}, or \rn{Single}.
+In all these cases, $\Ec \models_{\thD} t \teq u$.
 For \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, this is a consequence of equality reasoning.
@@
 
-It remains to show the converse of the previous theorem: When a derivation tree
+It remains to show the converse of the previous theorem: If a derivation tree
 with root node $\Ec$ contains a saturated node, then $\Ec$ is
 $\thD$-satisfiable.
-To do so, we exhibit a specific model $\M$ of $\thD$ that satisfies $\Ec$.
+The proof relies on a specific model $\M$ of $\thD$ that satisfies $\Ec$.
 
@@
 which requires additional terminology concerning $\mu$-terms.
-For a $\mu$-term $t$ with subterm $u$,
+Given a $\mu$-term $t$ with subterm $u$,
 %%% @ANDY: I'm taking the liberty of renaming "interpretation" to "expansion". Interpretation
@@
 \interp{t}{x}^B & {} \,=\, {} &
-\textstyle\smash{\begin{cases}
+\textstyle\relax{\begin{cases}
     \MU x.\; \const{C}( \interp{t}{ \vec u}^{\smash{B \mathrel\cup \{ x \}}}) & \text{if this $x \notin B$ is
@@
     x & \text{otherwise}
-\end{cases}} ~ \\[1.5ex] %% TYPESETTING: trick with \smash
+\end{cases}} ~ \\
 \interp{t}{\MU x.\; \const{C}( \vec u )}^B & = &

### 2015-02-16 tex/body.tex
more edits

@@
 \textstyle\relax{\begin{cases}
-    \MU x.\; \const{C}( \interp{t}{ \vec u}^{\smash{B \mathrel\cup \{ x \}}}) & \text{if this $x \notin B$ is
+    \MU x.\; \const{C}\smash{\bigl( \interp{t}{ \vec u}^{\smash{B \,\cup\, \{ x \}}}\bigr)} & \text{if this $x \notin B$ is
     bound by $\MU x.\; \const{C}( \vec u )$ in $t$}  \\[-\jot]
@@
 \textstyle\begin{cases}
-    \MU x.\; \const{C}( \interp{t}{ \vec u }^{\smash{B \mathrel\cup \{ x \}}} ) & \text{if }x \not\in B \\[-\jot]
+    \MU x.\; \const{C}\smash{\bigl( \interp{t}{ \vec u }^{\smash{B \,\cup\, \{ x \}}}\bigr)} & \text{if }x \not\in B \\[-\jot]
     x & \text{otherwise}
@@
 because $\MU y.\; \const{C}( y )$ is a self-similar subterm of $t$.
-In particular, note that their arguments have the same expansion with respect to $t$,
-that is,
-$\interp{t}{\MU y.\; \const{C}( y )} =$
-$\MU y.\; \const{C}( \interp{t}{y}^{\smash{\{y\}}} ) =$
-$\MU y.\; \const{C}( y )$
+Their arguments have the same expansion with respect to
+$t$:
+$\interp{t}{\MU y.\; \const{C}( y )} =
+\MU y.\; \const{C}\bigl( \interp{t}{y}^{\smash{\{y\}}} \bigr) =
+\MU y.\; \const{C}( y )$
 is $\alpha$-equivalent to
-$\interp{t}{y} =$
-$\MU y.\; \const{C}( \interp{t}{y}^{\smash{\{y\}}} ) =$
-$\MU y.\; \const{C}( y )$.
+$\interp{t}{y} =
+\MU y.\; \const{C}\bigl( \interp{t}{y}^{\smash{\{y\}}} \bigr) =
+\MU y.\; \const{C}( y )$.
 The term $t = \MU x.\; \const{C}( \MU y.\; \const{C}( x ) )$ is also abnormal,
@@
 \interp{t}{\MU y.\; \const{C}( x )}
-& = \MU y.\; \const{C}( \interp{t}{x}^{\smash{\{y\}}} ) \\
-& = \MU y.\; \const{C}( \interp{t}{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) ) }^{\smash{\{y\}}} ) \\
-& = \MU y.\; \const{C}( \MU x.\; \const{C}( \interp{t}{\MU y.\; \const{C}( x )}^{\smash{\{x,y\}}} ) ) \\
+& = \MU y.\; \const{C}\bigl( \interp{t}{x}^{\smash{\{y\}}} \bigr) \\
+& = \MU y.\; \const{C}\bigl( \interp{t}{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) ) }^{\smash{\{y\}}} \bigr) \\
+& = \MU y.\; \const{C}\bigl( \MU x.\; \const{C}( \interp{t}{\MU y.\; \const{C}( x )}^{\smash{\{x,y\}}} ) \bigr) \\
 & = \MU y.\allowbreak\; \const{C}( \MU x.\allowbreak\; \const{C}( y ) )

### 2015-02-16 tex/body.tex
removed comments to self

@@
 For any $\mu$-term $t$ of the form $\MU x.\; \const C( \vec u )$, we can
-recursively construct its \emph{normal form} $\nf{t}$
+construct its \emph{normal form} $\nf{t}$
 by replacing all of the self-similar subterms of $t$ with $x$
 %and by replacing all of the abnormal subterms of $t$ with their corresponding normal forms.
-and by normalizing the subterms.
+and by recursively normalizing $t$'s subterms.
 For example, $\nf{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) )} = \MU x.\; \const{C}( x )$.
@@
 We now define the class of interpretations for $\thD$.
-$\M( \tau )$ denotes the interpretation type $\tau$ in $\M$,
-that is, a set of domain elements for that type.
+$\M( \tau )$ denotes the interpretation type $\tau$ in $\M$---that is, a set of
+domain elements for that type.
 For a term $t\typ{\tau}$, $\M( t ) \in \M( \tau )$ denotes the interpretation of $t$ in $\M$.
-To simplify the presentation, all types are interpreted as sets of $\mu$-terms; but
+%To simplify the presentation,
+All types are interpreted as sets of $\mu$-terms, but
 only values of types in $\Codata$ may contain cycles.
@@
 \item
-%%% TODO: What does maximal set mean? Does it make sense for \Nondata?
 For each type $\tau$,
 $\M( \tau )$ is a maximal set of closed normal $\mu$-terms of that type that are
-unique up to $\alpha$-equivalence and acyclic when $\tau \not\in \Codata$.
+unique up to $\alpha$-equivalence and acyclic if $\tau \not\in \Codata$.
 \item
@@
 $\nf{\MU x.\; \const{C}( \M( \vec t \,) )}$, where $x$ is fresh.
-%%% TODO: I'm a bit worried to see a semantic value \M under a syntactic
-%%% constructor. I know our models consist of \mu-terms, but is this intended?
 \item
-For all selector terms $\const s^k_{\negvthinspace j}( t )$ of type $\tau$, % for some $\vec u$,
+For each selector term $\const s^k_{\negvthinspace j}( t )$ of type $\tau$, % for some $\vec u$,
 $\M\bigl( \const s^k_{\negvthinspace j}( t ) \bigr)$ is the value

### 2015-02-16 tex/body.tex
Minor fix

--- 2015-02-16 tex/body.tex
@@
 \newcommand{\rem}[1]{\textcolor{red}{[#1]}}
+%\newcommand{\is}[1]{is\text{-} #1}
+\newcommand{\is}[1]{\const{is#1}}
 %\newcommand{\ror}{\parallel}
@@
 %\newcommand{\interp}[2]{\lceil #2 \rceil_{#1}}
-\newcommand{\interp}[2]{\langle \smash{#2} \rangle_{#1}}
+\newcommand{\interp}[2]{\left< \smash{#2} \right>_{#1}}
 \newcommand{\dpath}[3]{\delta^{#1}_{#2}( #3 )}
@@
 %%% before by other authors.
-\newcommand{\thD}{\mathcal{DC}}
+\newcommand{\thD}{\mathcal{CD}}
 
@@
 (Section~\ref{sec:the-theory-solver}).
-\rem{CHECK WITH IMPLEMENTATION, but round to nearest 50 or 100 LOC.}
+\rem{CHECK WITH IMPLEMENTATION, but round to nearest 50 or 1000.}
 It consists of about 2000 lines of \cpp{} code, among which 1600 are
@@
 \begin{conf}%
-\vthinspace$\keyw{datatype}~\vthinspace\ty{option}_{\,\tau} \,=\, \const{None} \,\mid\, \const{Some}(\tau)$.%
+\vthinspace$\keyw{datatype}~\vthinspace\ty{option}_{\,\tau} = \const{None} \mid \const{Some}(\tau)$.%
 \end{conf}%
@@
 \[\begin{aligned}[t]
-      \keyw{datatype}~\vthinspace\ty{option}_{\,\tau\!} & \,=\, \const{None} \,\mid\, \const{Some}(\tau) \\[-.5\jot]
-      \keyw{codatatype}~\vthinspace\ty{complex} & \,=\, \const{Complex}(\const{re}{:}\; \ty{real},\; \const{im}{:}\;\ty{real})
+      \keyw{datatype}\;\, \ty{option}_{\,\tau} & \,=\, \const{None} \mid \const{Some}(\tau) \\[-.5\jot]
+      \keyw{codatatype}\; \,\ty{complex} & \,=\, \const{Complex}(\const{re}{:}\; \ty{real},\; \const{im}{:}\;\ty{real})
 \end{aligned}
@@
 this gives rise to spurious equalities between unrelated terms---e.g.,
-$\const{s}_1^1(\const{C}_2) \teq \const{s}_1^1(\const{C}_3)$. This flaw
+$\const{s}_{11}(\const{C}_2) \teq \const{s}_{11}(\const{C}_3)$. This flaw
 could be corrected, but the added complexity seems to suggest that selectors
@@
 \text{Exhaustiveness:}\quad
-  & \smash{\const{d}_{i1}(x) \mathrel\lor \cdots \mathrel\lor \const{d}_{im_i}(x)}
+  & \smash{\is{C}_{i1}(x) \mathrel\lor \cdots \mathrel\lor \is{C}_{im_i}(x)}
   \\[-.5\jot]
@@
 \begin{aligned}[t]
-\text{Coinduction:}\quad
-\AXC{\strut$\begin{gathered}\textstyle \smash{\bigwedge_{\,i}\, R_i[v_i, w_i]} \\[-\jot]\textstyle
+\text{Coinduction:}\kern.8em %%% TYPESETTING: should be \quad
+\AXC{\strut$\begin{gathered}\textstyle \bigwedge_{\,i}\, R_i[v_i, w_i] \\[-\jot]\textstyle
 \bigwedge_{\,i}\, \forall v\; w.\;\, R_i[v, w] \longrightarrow
@@
 
-Given a signature $\Sig$, we write $\thD$ to refer to the \emph{theory of
-datatypes and codatatypes},
+Given a signature $\Sig$, we write $\thD$ to refer to the \emph{theory of}
+(\emph{co})\emph{datatypes},
 which in addition to $\Sig$ defines a class of $\Sig$-interpretations $\mathbf{I}$,
@@
 can be replaced by an enumeration datatype with $n$~nullary constructors.
-Finally, since the constraints in $\Ec$ are ground, they cannot entail an upper
-bound on the cardinality of any uninterpreted type $\tau \in \Nondata$; thus,
+Moreover, since the constraints in $\Ec$ are ground, they cannot entail an upper bound
+on the cardinality of any uninterpreted type $\tau \in \Nondata$; thus,
 we may without loss of generality %or completeness
@@
 %not occurring in $\tEc$,
-and let $\Val\ec{u} = {\Varec{u}}$. Therefore, there are initially no
-constraints on the values for any equivalence class $\ec{u}$. The mapping $\Val$
-is refined by applying the following unfolding rule exhaustively:\strut
+and set $\Val\ec{u} = {\Varec{u}}$.
+Thus, there are initially no constraints on the values for any equivalence class $\ec{u}$.
+The mapping $\Val$ is refined by applying the following unfolding rule exhaustively:\strut
 \[
@@
 Given the mapping $\Val$, the \rn{Acyclic} and \rn{Unique} rules operate as follows.
-For acyclicity, if $\ec{t}$ is a datatype equivalence class
+For acyclicity, if $\ec{t\typ{\delta}}$ is a datatype equivalence class
 whose values $\Val \ec{t} = \MU x.\; u$ are cyclic
-(as required by the precondition $x \in \FV( u )$),
-then $\Ec$ is $\thD$-unsatisfiable.
+(as indicated by the precondition $x \in \FV( u )$),
+then $\Ec$ is unsatisfiable.
 For uniqueness, if $\ec{t}$, $\ec{u}$ are two codatatype equivalence classes
@@
 %for a capture-avoiding renaming of $\mu$-bound variables,
-then $t \teq u$. This may seem too restrictive, since
+then $t \teq u$. This rule may seem too restrictive, since
 $\MU x.\; \const{S}(x)$ and $\MU y.\; \const{S}(\const{S}(y))$ specify the same
@@
 %
-At this point, phase~1 is activated again to refine the equivalence classes,
-yielding
+Next, phase~1 is activated again to produce the equivalence classes
 $\{ x,\, \const{S}(x) \}$ and
-$\{ y,\, \const{S}(y),\allowbreak\, \const{S}(\const{S}(y)) \}$.
+$\{ y,\, \const{S}(y),\, \const{S}(\const{S}(y)) \}$.
 The mapping $\Val$ is updated accordingly:
@@
 \end{rep}
-Since $\Val \ec{x} \aequiv \Val \ec{y}$,
+Finally, since $\Val \ec{x} \aequiv \Val \ec{y}$,
 %for the renaming
 %$\{ {\Varec{y}} \mapsto {\VAREC{\const{S}(y)}}, {\VAREC{\const{S}(y)}} \mapsto {\Varec{y}} \}$,
-the \rn{Unique} rule can finally derive $x \teq y$.
+the \rn{Unique} rule derives $x \teq y$.
 \xend
@@
 
-\begin{figure}[t!]
+\begin{figure}[b!]
 \normalsize
@@
 \inferrule{
-  t, u\typ{\delta} \in \tEc
+  t\typ{\delta}, u\typ{\delta} \in \tEc
   \betweenantes
@@
 \end{tabular}
-%%% TODO: Clarify s's?
 \caption{\,Derivation rules for branching%.
@@
 the finiteness assumption on the datatype in \rn{Split} and the singleton
-constraint on the codatatype in \rn{Single} can be evaluated statically
+constraint on the codatatype in \rn{Single} can be computed statically
 based on a recursive computation of the cardinalities of the
-constructors' argument types. This step relies on the assumption that all types
-in $\Nondata$ are infinite.
+constructors' argument types (and assuming all types in $\Nondata$ are
+infinite).
 
@@
 \begin{proof}
-Consider a derivation tree with root node $\Ec$. Let $D \subseteq \tEc$ be the
-set of terms whose types are finite datatypes, and let $\SSS \subseteq \tEc$ be
-the set of terms occurring as arguments to selectors. For each term $t \in
-D$\negvthinspace, let
+Consider a derivation tree with root node $\Ec$.
+Let $\SSS \subseteq \tEc$ be the set of terms occurring as arguments of selectors.
+Let $T \subseteq \tEc$ be the set of terms whose types are finite datatypes.
+For each term $t \in T$\negvthinspace,
+let
 \begin{align*}
 \smash{\SSSS 0_t} & = \{ t \}
-& \smash{\SSSS {i+1}_t} & = \smash{\SSSS i_t \mathrel\cup \{ \const s( u ) \mid u\typ{\delta} \in \SSSS i_t{,}~ \delta \in \Data{,}~ \left|\delta\right| \text{ is finite}{,}~ \const{s} \in \Sel^{\delta}}  \}
+& \smash{\SSSS {i+1}_t} & = \smash{\SSSS i_t \mathrel\cup \{ \const s( u ) \mid u\typ{\delta} \in \SSSS i{,}~ \delta \in \Data{,}~ \left|\delta\right| \text{ is finite}{,}~ \const{s} \in \Sel^{\delta}}  \}
 \end{align*}
@@
 and let $\SSSS \AST_t$ be the limit of this sequence.
-This is a finite set for each $t$ because the values of the type of $t$ are of finite size\rem{Clarify}.
-Let $\SSSS \AST$ be the union of all sets $\SSSS \AST_t$ where $t \in D$,
-and let $\tcEc$ be the set of subterms of
-%
-$$\Ec \mathrel\cup \{ \const C_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\smash{\,n_{\negvthinspace j}}}_{\negvthinspace j}( t ) \bigr) \mid t\typ{\delta} \in \SSS \mathrel\cup \SSSS \AST{,}~ \const C_{\negvthinspace j} \in \Ctr^\delta \}$$
-%
+This is a finite set for each $t$ because the values of the type of $t$ are of finite size.
+Let $\SSSS \AST$ be the union of all sets $\SSSS \AST_t$ where $t \in T$,
+and let $\tcEc$ be the set of subterms of $\Ec \mathrel\cup \{ \const C_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\smash{\,n_{\negvthinspace j}}}_{\negvthinspace j}( t ) \bigr) \mid t\typ{\delta} \in \SSS \mathrel\cup \SSSS \AST{,}~ \const C_{\negvthinspace j} \in \Ctr^\delta \}$.
 In a derivation tree with root node~$\Ec$,
-it can be shown by %case analysis
-induction on the rules of the calculus that each non-root node $\Fc$ is such that
+it can be shown by case analysis on the rules of the calculus that each non-root node $\Fc$ is such that
 $\mathcal{T}(\Fc) \subseteq \tcEc$, and hence contains an equality between two terms from $\tcEc$ not occurring in its parent node.
@@
 For \rn{Clash}, this is a consequence of distinctness.
-For \rn{Acyclic}, the construction of $\Val$ indicates that the class of values that term~$t$ can take in models of $\Ec$ is infeasible.
+For \rn{Acyclic}, the construction of $\Val$ indicates that the class of values that term~$t$ can take in models of $\Ec$ is infeasible,
+and thus $\Ec$ is unsatisfiable.
 If the child nodes of $\Ec$ are closed derivation trees
-whose root nodes are the result of applying \rn{Split} on $t\typ{\delta}$,
-by the induction hypothesis $\Ec \mathrel\cup t \teq \smash{\const C_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\,\smash{n_{\negvthinspace j}}}_{\negvthinspace j}( t ) \bigr)}$ is
-$\thD$-unsatisfiable
+whose root nodes are the result of applying the rule \rn{Split} on $t\typ{\delta}$,
+by the induction hypothesis $\Ec \mathrel\cup t \teq \smash{\const C_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\,\smash{n_{\negvthinspace j}}}_{\negvthinspace j}( t ) \bigr)}$ is unsatisfiable
 for each $\const C_{\negvthinspace j} \in \smash{\Ctr^\delta}$.
@@
 Otherwise, the child node of $\Ec$ is a closed derivation tree
-whose root node $\Ec \mathrel\cup t \teq u$ is obtained by applying one of the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, \rn{Inject}, \rn{Unique}, or \rn{Single}.
-In all these cases, $\Ec \models_{\thD} t \teq u$.
+whose root node is $\Ec \mathrel\cup t \teq u$ obtained by applying one of the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, \rn{Inject}, \rn{Unique}, or \rn{Single}.
+In all cases, we have that $\Ec \models_{\thD} t \teq u$.
 For \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, this is a consequence of equality reasoning.
@@
 
-It remains to show the converse of the previous theorem: If a derivation tree
+It remains to show the converse of the previous theorem: When a derivation tree
 with root node $\Ec$ contains a saturated node, then $\Ec$ is
 $\thD$-satisfiable.
-The proof relies on a specific model $\M$ of $\thD$ that satisfies $\Ec$.
+To do so, we exhibit a specific model $\M$ of $\thD$ that satisfies $\Ec$.
 
@@
 which requires additional terminology concerning $\mu$-terms.
-Given a $\mu$-term $t$ with subterm $u$,
+For a $\mu$-term $t$ with subterm $u$,
 %%% @ANDY: I'm taking the liberty of renaming "interpretation" to "expansion". Interpretation
@@
 \interp{t}{x}^B & {} \,=\, {} &
-\textstyle\relax{\begin{cases}
+\textstyle\smash{\begin{cases}
     \MU x.\; \const{C}( \interp{t}{ \vec u}^{\smash{B \mathrel\cup \{ x \}}}) & \text{if this $x \notin B$ is
@@
     x & \text{otherwise}
-\end{cases}} ~ \\
+\end{cases}} ~ \\[1.5ex] %% TYPESETTING: trick with \smash
 \interp{t}{\MU x.\; \const{C}( \vec u )}^B & = &

### 2015-02-16 tex/body.tex
more

--- 2015-02-16 tex/body.tex
@@
 For each selector term $\const s^k_{\negvthinspace j}( t )$ of type $\tau$, % for some $\vec u$,
-$\M\bigl( \const s^k_{\negvthinspace j}( t ) \bigr)$ is the value
+$\M\negvthinspace\bigl( \const s^k_{\negvthinspace j}( t ) \bigr)$ is the value
 in $\M( \tau )$ that is $\alpha$-equivalent to
@@
 it remains only to define how $\M$ interprets wrongly applied selector terms and variables.
-For the latter, this will be based on the mapping $\Val$ constructed in phase~2 of the calculus.
-We compute the \emph{completion} $\ValC$ of $\Val$ for normal model $\M$
+For the latter, this will be based on the mapping $\Val$ computed in phase~2 of the calculus.
+We obtain the \emph{completion} $\ValC$ of $\Val$ for normal model $\M$
 by assigning values from $\M$ to unassigned variables in the domain of $\Val$.
@@
 We write $t \vsimv{x} u$ if $\mu$-terms $t$ and $u$ are syntactically equivalent
-for some renaming that avoids capturing any variable other than $x$.
+for some renaming that avoids capturing any variable other than $x$.%
+\rem{Clarify previous definition}
 For example,

### 2015-02-16 tex/body.tex
started updating stats
updated

@@
 %is simply $\const{Z}$ after dropping the $\mu$-binding of $x$.
+
+\rem{TODO: Normal model vs. actual model.}
 
@@
 No (co)datatypes
-  & 221 & 760 & \phantom{0}55 & 1036 \\[\jot]
+  & 229 & 765 & \phantom{0}61 & 1055 \\[\jot]
 %Datatype simplication only
@@
 Datatypes without acyclicity
-  & 227 & 765 & \phantom{0}55 & 1047 \\
+  & 235 & 770 & \phantom{0}61 & 1066 \\
 Datatypes with acyclicity
-  & 227 & 771 & \phantom{0}55 & 1053 \\[\jot]
+  & 235 & 773 & \phantom{0}61 & 1069 \\[\jot]
 %Codatatype simplication only
@@
 Codatatypes without uniqueness
-  & 222 & 789 & \phantom{0}56 & 1067 \\
+  & 230 & 794 & \phantom{0}65 & 1089 \\
 Codatatypes with uniqueness
-  & 223 & 789 & \bfseries \phantom{0}59 & 1071 \\[\jot]
+  & 231 & 794 & \bfseries \phantom{0}68 & 1093 \\[\jot]
 Full (co)datatypes
-  & \bfseries 229 & \bfseries 800 & \bfseries \phantom{0}59 & \bfseries 1088 \\[\jot]
+  & \bfseries 237 & \bfseries 802 & \bfseries \phantom{0}68 & \bfseries 1107 \\[\jot]
 Total number of goals

### 2015-02-16 tex/body.tex
more

@@
 \makeatother
+
+\newcommand\NOK[1]{\textcolor{red}{\textbf{#1}}}
+\newcommand\OK[1]{#1}
 
@@
 %  &  \\
-Datatypes without acyclicity
+Datatypes without \rn{Acyclic}
   & 235 & 770 & \phantom{0}61 & 1066 \\
-Datatypes with acyclicity
+Full datatypes
   & 235 & 773 & \phantom{0}61 & 1069 \\[\jot]
@@
 %  &  \\
-Codatatypes without uniqueness
+Codatatypes without \rn{Unique}
   & 230 & 794 & \phantom{0}65 & 1089 \\
-Codatatypes with uniqueness
+Full codatatypes
   & 231 & 794 & \bfseries \phantom{0}68 & 1093 \\[\jot]
@@
 Codatatypes being a recent addition to Isabelle
-\cite{blanchette-et-al-2014-impl}, benchmarks are somewhat hard to come by. We
-included the theory files from the Isabelle distribution (Distro) and the \emph{Archive
+\cite{blanchette-et-al-2014-impl}, benchmarks are not abundant. We
+included all the theory files from the Isabelle distribution (Distro) and the \emph{Archive
 of Formal Proofs} (AFP) \cite{klein-et-al-afp} that define codatatypes falling
@@
 To also exercise the the support
-for datatypes, the benchmarks are complemented by theories about various list
-and tree data structures. The theory files were selected before running the
-experiments. The experimental data is publicly available \cite{our-eval-data}.
+for datatypes, the benchmarks are complemented by theories about lists
+and tree data structures. The theories were selected before running the
+experiments. The experimental data is publicly available.%
+\footnote{\url{http://TODO/FIXME.}}
 
@@
 
-\newcommand\BAD[1]{\textcolor{red}{\textbf{#1}}}
-
 %  * experience with such features is not extremely high -- e.g. arithmetic,
@@
 The results are summarized in Table~\ref{tab:bench}. The decision procedure
-makes a difference across all theories, and an overall difference of
-\BAD{4.7\%}. Moreover, every aspect of the procedure, including the more
-expensive rules, make a contribution. \textbf{NOT TRUE}: What is not visible in the table, but
-readily available when looking at the raw data \cite{our-eval-data}, is that
-the stronger decision procedures subsume the weaker ones. In the context of
-Sledgehammer, the power of interpreted (co)datatypes is roughly
-comparable to that of arithmetic \cite{blanchette-et-al-2013-smt}.
-
+makes a difference across all theories. Overall, it accounts for a success rate
+increase of \OK{4.9\%}. Moreover, every aspect of the procedure, including the
+more expensive rules, make a contribution.
 %* and finally, we look in more detail at one or two such proofs, by presenting
 %  it and explaining it (and, before that, understanding it)
-
-Among the four uniqueness proofs, three corresponded to one-line
-proofs in Isabelle, of the form \keyw{by}~\textit{coinduction}~\textit{auto}
-\cite{blanchette-et-al-2014-impl}. The fourth proof was somewhat more elaborate:
+%
+Three proofs were found thanks to
+the \rn{Acyclic} rule and four required the \rn{Unique} rule. Among the latter,
+three correspond to one-line proofs in Isabelle, of the form
+\keyw{by}~\textit{coinduction}~\textit{auto} \cite{blanchette-et-al-2014-impl}.
+The fourth proof, by Gammie and Lochbihler, is more elaborate:
 %
@@
 \const{Node}\;0\;\const{num}\;\const{X0}$.
+
+The raw data also reveals that the
+stronger decision procedures almost completely subsume the weaker ones. For
+Sledgehammer, the power of interpreted (co)datatypes is roughly comparable to
+that of arithmetic \cite{blanchette-et-al-2013-smt}.
 
@@
 with high-level features and at reducing the gap between automatic and
-interactive theorem proving. As future work, we want to implement proof
-reconstruction for CVC4's (co)datatype inferences in Isabelle. We also want to
-try out CVC4 for higher-order model finding in Isabelle, as an alternative to
-the counterexample generator Nitpick \cite{blanchette-nipkow-2010}. We
-see some opportunities to enrich SMT solvers with recursive and corecursive
-functions. Finally, it might be possible to go further in terms of supporting
-nested and mixed (co)recursion and quantified formulas over (co)datatypes in
-solvers.
+interactive theorem proving. As future work, it would be useful to implement
+proof reconstruction for CVC4's (co)datatype inferences in Isabelle. We also
+want to try out CVC4 for higher-order model finding in Isabelle, as an
+alternative to the counterexample generator Nitpick
+\cite{blanchette-nipkow-2010}. Finally, it might be worthwhile to enrich SMT
+solvers with explicit support for recursive and corecursive functions.
+
+%Finally, it might be possible to go
+%further in terms of supporting nested and mixed (co)recursion and quantified
+%ormulas over (co)datatypes in solvers.
 

### 2015-02-16 tex/body.tex
more

@@
 When constructing a model $\M$ for $\Ec$,
-it remains only to define how $\M$ interprets wrongly applied selector terms and variables.
+it remains only to specify how $\M$ interprets wrongly applied selector terms and variables.
 For the latter, this will be based on the mapping $\Val$ computed in phase~2 of the calculus.
-We obtain the \emph{completion} $\ValC$ of $\Val$ for normal model $\M$
-by assigning values from $\M$ to unassigned variables in the domain of $\Val$.
 
@@
 We write $t \vsimv{x} u$ if $\mu$-terms $t$ and $u$ are syntactically equivalent
-for some renaming that avoids capturing any variable other than $x$.%
-\rem{Clarify previous definition}
+for some renaming that avoids capturing any variable other than $x$.
 For example,
-$\MU x.\; \const{D}( x ) \vsimv{y} \MU x.\; \const{D}( y )$,
-$\MU x.\; \const{C}( x, x ) \vsimv{x} \MU y.\; \const{C}( x, y )$ and
+$\MU x.\; \const{D}( x ) \vsimv{y} \MU x.\; \const{D}( y )$
+(by renaming $y$ to $x$),
+$\MU x.\; \const{C}( x, x ) \vsimv{x} \MU y.\; \const{C}( x, y )$, and
 $\MU x.\; \const{C}( z, x ) \vsimv{z} \MU y.\; \const{C}( z, y )$,
@@
 For a variable $x\typ{\tau}$ and a fixed normal model $\M$,
-we write $\mathcal{T}^\M_x( \Val )$ to denote the set consisting of all values $v$ from $\M( \tau )$
+we write $\mathcal{T}^{\smash{\M}}_x( \Val )$ to denote the set consisting of all values
+$v \in \M( \tau )$
 such that $v \vsimv{x} \interp{t}{u}$ for some term $t$ with subterm $u$ occurring in the range of $\Val$.
-
-We construct $\ValC$ by exhaustively applying the following rule to $\nf{\Val}$:
+\rem{Intuition?}
+
+The \emph{completion} $\ValC$ of $\Val$ for a normal model $\M$
+assigns values from $\M$ to unassigned variables in the domain of $\Val$.
+We construct $\ValC$ by exhaustively applying the following rule to $\nf{\Val}$:%
+\rem{A bit clumsy}
 \[
@@
   \betweenantes
-  v \not\in \mathcal{T}^\M_{\Varec{x}}(\Val)
+  v \not\in \mathcal{T}^{\smash{\M}}_{\Varec{x}}(\Val)
 }{
@@
 %
-The rule chooses an unassigned variable in $\Val$
-and assigns it a value that is not $\alpha$-equivalent to one occurring in $\mathcal{T}^\M_x(\Val)$.
-Since this update removes one variable from the set $\FV( \Val )$ and does not add any variables to $\FV( \Val )$,
-it can only be applied a finite number of times.
-After assigning this value, we normalize all terms in the range of $\Val$.
+The rule chooses an unassigned variable in $\Val$ and assigns it a value that is
+not $\alpha$-equivalent to one occurring in
+$\mathcal{T}^{\smash{\M}}_{\Varec{x}}(\Val)$.
+\rem{What guarantees that such a value exists? (Answer: infinitely many values
+of codatatypes, of which only finitely many are in
+$\mathcal{T}^{\smash{\M}}_{\Varec{x}}(\Val)$?)}
+Since this update removes one
+variable from the set $\FV( \Val )$ and does not add any variables to $\FV( \Val
+)$, it can only be applied a finite number of times. We normalize all terms in
+the range of $\Val$ after each rule application.\rem{clumsy}
 
@@
 \}$.
-After running the calculus to saturation on $\Ec$, the mapping $\Val$ contains
+After applying the calculus to saturation on $\Ec$, the mapping $\Val$ is as follows:
 \[\begin{array}{r@{}c@{}l@{\;}l@{\qquad}r@{}c@{}r@{\;}l@{}l@{\;}l}
@@
 %
-To construct a completion of $\Val$, we must choose values for the free variables of $\Val$,
-namely ${\Varec{w_3}}$ and ${\Varec{x}}$.
-The content of the set \smash{$\mathcal{T}^\M_{{\Varec{x}}}( \Val )$} is $\alpha$-equivalent to
+To construct a completion $\ValC$, we must choose values for the free
+variables of $\Val$, namely ${\Varec{w_3}}$ and ${\Varec{x}}$.
+Modulo $\alpha$-equivalence, the set \smash{$\mathcal{T}^{\smash{\M}}_{{\Varec{x}}}( \Val )$} is
 \[%\begin{array}{c}
@@
 $x$ is interpreted as $\Val \ec{x}$, $y_1$ as $\Val \ec{y_1}$, and so on.
-Assigning a value for ${\Varec{x}}$ that is $\alpha$-equivalent to a value in $\mathcal{T}^\M_{{\Varec{x}}}( \Val )$
+Assigning a value for ${\Varec{x}}$ that is $\alpha$-equivalent to a value in $\mathcal{T}^{\smash{\M}}_{{\Varec{x}}}( \Val )$
 may cause values in the range of $\Val$ to become $\alpha$-equivalent,
-which in turn may cause $\M$ to be inconsistent.
+which in turn may cause $\M$ to
+lose its modelhood.%
+\rem{Improve terminology.}
+% "be inconsistent." Surely, models can't be "inconsistent".
 For example, assign $\MU {\Varec{x}}.\; \const{D}( {\Varec{x}} )$ for ${\Varec{x}}$.
@@
 However, this contradicts the disequality $x \not\teq y_2$ in~$\Ec$.
-On the other hand, if the value assigned to ${\Varec{x}}$ is not $\alpha$-equivalent to any term in $\mathcal{T}^\M_{{\Varec{x}}}( \Val )$,
+On the other hand, if the value assigned to ${\Varec{x}}$ is not $\alpha$-equivalent to any term in $\mathcal{T}^{\smash{\M}}_{{\Varec{x}}}( \Val )$,
 the values in the range of $\Val$ remain $\alpha$-disequivalent.
@@
 Legal substitutions for ${\Varec{x}}$ may cause the range of $\Val$ to contain abnormal terms.
-After applying the substitution for the third term mentioned above,
+After applying the substitution for the third term mentioned above\rem{which third term?},
 $\Val \ec{w_1}$ is $\MU {\Varec{w_1}}.\; \const{C}( \MU {\Varec{x}}.\; \const{D}( \const{C}( {\Varec{x}} ) ) )$,
@@
 $\vphantom{\Bigl(_j}\muvar( t ) = \smash{\begin{cases}
-  x & \text{if $t$ is of the form $\MU {x}.\; u$} \\[-\jot]
-  t & \text{otherwise.}
+  t & \text{if $t$ is a variable} \\[-\jot]
+  x & \text{if $t$ is of the form $\MU {x}.\; u$.} 
 \end{cases}}$
@@
 there is only one equivalence class of type $\tau$ in $\Ec$,
-and thus there are no terms in $\mathcal{T}^\M_{\Varec{y}}(\ValC)$ of type $\tau$.
+and thus there are no terms in $\mathcal{T}^{\smash{\M}}_{\Varec{y}}(\ValC)$ of type $\tau$.
 This is a contradiction, since our model completion may assign the value in the domain of $\tau$ to ${\Varec{y}}$.
@@
 This is also a contradiction,
-since there are only a finite number of closed terms in $\mathcal{T}^\M_{\Varec{y}}(\ValC)$,
-and thus our model completion can assign a value not occurring in $\mathcal{T}^\M_{\Varec{y}}(\ValC)$ to ${\Varec{y}}$.
+since there are only a finite number of closed terms in $\mathcal{T}^{\smash{\M}}_{\Varec{y}}(\ValC)$,
+and thus our model completion can assign a value not occurring in $\mathcal{T}^{\smash{\M}}_{\Varec{y}}(\ValC)$ to ${\Varec{y}}$.
 By construction, $\ValC \ec{x}$ is normal.
@@
 ${\Varec{x}} \in FV( \Val_1 )$,
-and $\MU {\Varec{x}}.\; t$ is not $\alpha$-equivalent to a term in $\mathcal{T}^\M_{\Varec{x}}(\Val_1)$.
+and $\MU {\Varec{x}}.\; t$ is not $\alpha$-equivalent to a term in $\mathcal{T}^{\smash{\M}}_{\Varec{x}}(\Val_1)$.
 To show 2,
@@
 Since $\interp{\Val_1 \ec{y}}{t_1}$ is normal,
-there is a closed $\mu$-term $v \in \mathcal{T}^\M_{\Varec{x}}( \Val_1 )$
+there is a closed $\mu$-term $v \in \mathcal{T}^{\smash{\M}}_{\Varec{x}}( \Val_1 )$
 such that $v \vsimv{\Varec{x}} \interp{\Val_1 \ec{y}}{t_1}$.
@@
 \normalsize
-\begin{center}\begin{tabular}{l@{\kern1.5em}c@{\kern.5em}c@{\kern.5em}c@{\kern1.5em}c}
-  & \HD{Distro} & \HD{AFP} & \HD{\gandl} & All
+\begin{center}\begin{tabular}{@{\;}l@{\kern1.5em}c@{\kern.5em}c@{\kern.5em}c@{\kern1.3em}c@{\;}}
+  & \HD{Distro} & \HD{AFP} & \HD{\gandl} & \relax{All}
 \MIDRULE

### 2015-02-16 tex/body.tex
minor

@@
+%%% TODO: Clarify that \mu-terms appear in generated models
+%%% TODO: In report, give small example where Single is needed
+
 \usepackage[T1]{fontenc}
@@
 \makeatother
+
+%\newcommand\XXXL{DPLL}
+\newcommand\XXXL{CDCL} %% TODO: Convince Andy this is right
 
@@
 \rem{TODO}
+\qed
 \end{proof}
@@
 %\rem{TODO}
+%\qed
 %\end{proof}
@@
 \rem{TODO: Normal model vs. actual model.}
+\rem{TODO: Model or interpretation? Grep elsewhere.}
 
@@
 %
-The rule chooses an unassigned variable in $\Val$ and assigns it a value that is
-not $\alpha$-equivalent to one occurring in
-$\mathcal{T}^{\smash{\M}}_{\Varec{x}}(\Val)$.
+The rule chooses an unassigned variable in $\Val$ and assigns it a fresh
+value---one that does not occur in $\mathcal{T}^{\smash{\M}}_{\Varec{x}}(\Val)$
+modulo $\alpha$-equivalence.
+\rem{The ``not alpha equivalent'' bit is expressed differently in the rule,
+with $\notin$. This looks fishy.}
 \rem{What guarantees that such a value exists? (Answer: infinitely many values
@@
 which in turn may cause $\M$ to
-lose its modelhood.%
-\rem{Improve terminology.}
-% "be inconsistent." Surely, models can't be "inconsistent".
+become a non-model.
+% "be inconsistent." Models can't be "inconsistent".
 For example, assign $\MU {\Varec{x}}.\; \const{D}( {\Varec{x}} )$ for ${\Varec{x}}$.
@@
 However, this contradicts the disequality $x \not\teq y_2$ in~$\Ec$.
-On the other hand, if the value assigned to ${\Varec{x}}$ is not $\alpha$-equivalent to any term in $\mathcal{T}^{\smash{\M}}_{{\Varec{x}}}( \Val )$,
+On the other hand, if the value assigned to ${\Varec{x}}$ is fresh, %not $\alpha$-equivalent to any term in $\mathcal{T}^{\smash{\M}}_{{\Varec{x}}}( \Val )$,
 the values in the range of $\Val$ remain $\alpha$-disequivalent.
@@
 \begin{enumerate}
-\item[\rm (1)] $\ValC \ec{x}$ is $\alpha$-equivalent to a value in $\M( \tau )$, where $\ec{x}$ has type $\tau$.
+\item[\rm (1)] $\ValC \ec{x\typ{\tau}}$ is $\alpha$-equivalent to a value in $\M( \tau )$.
 \item[\rm (2)]
@@
 
-Intuitively, this lemma states three properties of $\ValC$ which ensure a normal model $\M$ can be trivially constructed for $\Ec$.
-Part~\PART{1} states that the values in the range of $\ValC$ are $\alpha$-equivalent to a value in normal model.
-In particular, this means they are closed, normal, and acyclic when required.
-Part~\PART{2} states that the interpretation of all subterms in the range of $\ValC$ depends on its associated variable only.
-In other words, the interpretation of a subterm $t$ where $\muvar(t) = {\Varec{x}}$ is equal to $\ValC \ec{x}$, independent of the context in which it is interpreted.
-Part~\PART{3} states that $\ValC$ is injective, which in the following ensures that a distinct values can be associated with each equivalence class in a corresponding model.
+Intuitively, this lemma states three properties of $\ValC$ that ensure a normal
+model $\M$ can be trivially constructed for $\Ec$. Part~\PART{1} states that the
+values in the range of $\ValC$ are $\alpha$-equivalent to a value in normal
+model. In particular, this means they are closed, normal, and acyclic when
+required. Part~\PART{2} states that the interpretation of all subterms in the
+range of $\ValC$ depends on its associated variable only. In other words, the
+interpretation of a subterm $t$ where $\muvar(t) = {\Varec{x}}$ is equal to
+$\ValC \ec{x}$, independently of the context in which it is interpreted.
+Part~\PART{3} states that $\ValC$ is injective, which in the following ensures
+that a distinct values can be associated with each equivalence class in a
+corresponding model.
 
@@
 We consider a normal model $\M$
-that interprets wrongly applied selectors based on equality information in $\Fc$,
-and interprets the variables of $\Fc$ based on the completion $\ValC$ of the mapping $\Val$ from phase~2.
+that interprets wrongly applied selectors based on equality information in $\Fc$
+and that interprets the variables of $\Fc$ based on the completion $\ValC$ of the mapping $\Val$ from phase~2.
 For the latter, let $\M( x )$ be the value in $\M$ that is $\alpha$-equivalent with $\ValC \ec{x}$ for each variable $x \in \mathcal{T}(\Fc)$,
@@
 then $\M( t )$ is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \M( u_1 ), \ldots, \M( u_n ) ) }$ for some fresh $x$,
-which by the induction hypothesis is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \ValC \ec{ u_1 }, \ldots, \ValC \ec{ u_n } )}$, call this term $t_u$.
+which by the induction hypothesis is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \ValC \ec{ u_1 }, \ldots, \ValC \ec{ u_n } )}$.
+Call this term $t_u$.
 %By Lemma~\ref{lem:model-completion}\PART{2} and due to our construction of $\ValC$, we have $\ValC \ec{t}$ is a term
@@
 are in distinct equivalence classes only if $u$ and $u'$ are in distinct equivalence classes.
-Due to the induction hypothesis and Lemma~\ref{lem:model-completion}\PART{3}, $\M( u ) \neq \M( u' )$ for all such $u, u'$.
+Due to the induction hypothesis and Lemma~\ref{lem:model-completion}\PART{3}, $\M( u ) \neq \M( u' )$ for all such $u$, $u'$.
 Thus, we may interpret $\smash{\M(\const s^k_{\negvthinspace j})( \M( u ) )}$ as the value in $\M( \tau )$ that is $\alpha$-equivalent with $\ValC \ec{ t }$.
@@
 
-This section describes how the calculus in Section~\ref{sec:a-ground-decision-procedure-for-co-datatypes} is implemented within the SMT solver CVC4.
-
-\paragraph{Integration into DPLL(T).}
-The calculus is implemented as \emph{theory solver} of CVC4,
-that is, a specialized solver for determining the satisfiability of conjunctions of constraints for its theory.
-Given a theory $T = T_1 \mathrel\cup \ldots T_n$ and a input set of clauses $F$ in CNF,
-the DPLL($T$) \rem{CDCL($T$)?} procedure
-(incrementally) builds partial assignments from the atomic formulas of $F$ to truth values such that no clause in $F$ is falsified.
-We may interpret this partial assignment as a set of literals where $a \in M$ for all atoms $a$ that it assigned the value $\top$, and $\neg a \in M$ for all atoms that it assigns the value $\bot$.
-Then, at a high level, by the Nelson--Oppen combination framework,
-each $T_i$-solver for $i = 1, \ldots, l$ takes as input a combination of (1) the purified form of $T_i$-constraints occurring in $M$ where terms containing symbols not belonging to $T_i$ are replaced by fresh variables of the same type,
-(2) additional equalities and disequalities between variables of types not belonging to $T_i$.
-Let us call this set $M_i$.
-Each $T_i$ solver either
-reports that a subset $C$ of $M_i$ is $T_i$-unsatisfiable, in which case $\neg C$ is added to $F$,
-adds a clause $C$ to $F$,
-or does nothing.
-When $M$ is a complete assignment for $F$, a theory solver can only choose to do nothing when $M_i$ is indeed $T_i$-satisfiable.
-
-Assume $\Ec$ is initially the set $M_i$ described above.
-For each equality added to $\Ec$, we associate a set of equalities from $M_i$ that together entail $t \teq s$,
-which we call its \emph{explanation}.
-Similarly, each $\Val \ec{x}$ is also associated an explanation, that is,
-a set of equalities from $M_i$ that entail that the values of $\ec{x}$ in models of $\Ec$ are of the form $\Val \ec{x}$.
-For example, if $x \teq \const{C}( x ) \in M_i$, then $x \teq \const{C}( x )$ is a (possible) explanation for $\Val \ec{x} = \MU {\Varec{x}}.\; \const{C}( {\Varec{x}} )$.
-Thus, the rules of the calculus are implemented as follows.
-For all rules with conclusion $\bot$,
-we report the union of the explanations for all premises is $\thD$-unsatisfiable.
-For the \rn{Split} rule, we add a clause of the form
-$\const{C}_1\bigl( \const s^1_1( t ), \ldots, \const s^{n_1}_1( t ) \bigr) \mathrel\lor \cdots \mathrel\lor t \teq \const{C}_m\bigl( \const s^1_m( t ), \ldots, \const s^{n_m}_m( t ) \bigr)$
-to $F$.
-Decisions on which branch to take are thus performed externally by the SAT solver, and not by the theory solver.
-All other rules add equalities to the internal state of the theory solver.
-The rules in phase~1 (along with the optimization for collapsing selectors) and the \rn{Single} rule are performed eagerly,
-that is for partial satisfying assignments $M$, while the rules in phase~2 and the \rn{Split} rule are performed only for complete satisfying assignments $M$,
-and in that order.
-
-    * sharing is caring (let's cite it for a change)
-    * theory combination more complicated (and Theorem Completeness too weak)
+The decision procedure for (co)datatypes and its extension to quantified
+formulas were described at a high level of abstraction. This section describes
+the implementation within the SMT solver CVC4.
+
+\paragraph{Integration into {\relax\XXXL($T$)}.}
+The calculus is implemented as a \emph{theory solver} of CVC4, that is, a
+specialized solver for determining the satisfiability of conjunctions of
+constraints for its theory.
+Given a theory $T = T_1 \mathrel\cup \cdots \mathrel\cup T_n$ and a input set of
+clauses $F$ in CNF, the \XXXL($T$) procedure
+incrementally builds partial assignments from the atomic formulas of $F$ to
+truth values such that no clause in $F$ is falsified. We may interpret such a
+partial assignment as a set of literals where $a \in M$ for all atoms $a$ that
+it assigned the value $\top$, and $\neg\; a \in M$ for all atoms that it assigns
+the value $\bot$. Then, at a high level, by the Nelson--Oppen combination
+framework, each $T_i$-solver for $i = 1, \ldots, l$ takes as input a combination
+of (1) the purified form of $T_i$-constraints occurring in $M$, where terms
+containing symbols not belonging to $T_i$ are replaced by fresh variables of the
+same type; (2) additional equalities and disequalities between variables of
+types not belonging to $T_i$. Let us call this set $M_i$. Each $T_i$ solver
+either reports that a subset $C$ of $M_i$ is $T_i$-unsatisfiable, in which case
+$\neg\; C$ is added to $F$, adds a clause $C$ to $F$, or does nothing.
+When $M$ is a complete assignment for $F$, a theory solver can only choose to do
+nothing when $M_i$ is indeed $T_i$-satisfiable.
+
+Assume $\Ec$ is initially the set $M_i$ described above. For each equality added
+to $\Ec$, we associate a set of equalities from $M_i$ that together entail $t
+\teq s$, which we call its \emph{explanation}.
+Similarly, each $\Val \ec{x}$ is also associated an explanation---that is, a set
+of equalities from $M_i$ that entail that the values of $\ec{x}$ in models of
+$\Ec$ are of the form $\Val \ec{x}$. For example, if $x \teq \const{C}( x ) \in
+M_i$, then $x \teq \const{C}( x )$ is an %(possible)
+explanation for $\Val \ec{x} = \MU {\Varec{x}}.\; \const{C}( {\Varec{x}} )$.
+Thus, the rules of the calculus are implemented as follows. For all rules with
+conclusion $\bot$, we report the union of the explanations for all premises is
+$\thD$-unsatisfiable. For the \rn{Split} rule, we add a clause of the form
+$\const{C}_1\bigl( \const s^1_1( t ), \ldots, \const s^{n_1}_1( t ) \bigr)
+\mathrel\lor \cdots \mathrel\lor t \teq \const{C}_m\bigl( \const s^1_m( t ),
+\ldots, \const s^{n_m}_m( t ) \bigr)$ to $F$.
+Decisions on which branch to take are thus performed externally by the SAT
+solver and not by the theory solver. All other rules add equalities to the
+internal state of the theory solver. The rules in phase~1 (along with the
+optimization for collapsing selectors) and the \rn{Single} rule are performed
+eagerly---that is, for partial satisfying assignments $M$---while the rules in
+phase~2 and the \rn{Split} rule are performed only for complete satisfying
+assignments $M$, and in that order.\rem{in what order?}
+
+\rem{TODO: * sharing is caring (let's cite it for a change)}
+\rem{TODO: ``theory combination more complicated (and Theorem Completeness too weak)''?}
 
 \paragraph{Optimizations.}
-Our implementation uses several optimizations not reflected in the presentation of the calculus, which closely follow the approach outlined by Barrett et al.\ \cite{barrett-et-al-2007}.
-We only briefly mention these optimizations, since each applies to datatypes as well as codatatypes, and most have been presented thoroughly in Barrett et al.
+Our implementation uses several optimizations not reflected in the presentation
+of the calculus, which closely follow the approach outlined by Barrett et al.\
+\cite{barrett-et-al-2007}.%
+\rem{What closely follows the approach by Barrett et al.? The optimizations or
+the calculus?}
+We only briefly mention these optimizations, since
+each applies to datatypes as well as codatatypes, and most have been presented
+thoroughly in Barrett et al.
 
@@
 Handling discriminators requires extending the decision procedure with several
-additional rules (see~Barrett et al.\ \cite{barrett-et-al-2007} for more
-details), which apply uniformly to both datatypes and codatatypes.
+additional rules \cite{barrett-et-al-2007},
+which apply uniformly to both datatypes and codatatypes.
 Also, selectors are collapsed eagerly:
-if $\const s^k_{\negvthinspace j}( t )$ is a term in $\tEc$ and $t$ contains the constructor
-$\const{C}_{\negvthinspace j}( \vec u )$, we immediately infer $\const s^k_{\negvthinspace j}( t ) \teq u_k$, whereas the calculus we presented would apply \rn{Split} and \rn{Inject} before inferring this equality.
-For the sake of reducing the number of unique constraints considered by the calculus, we compute a normal form for constraints as a preprocessing step.
-In particular, we
-replace $t \teq s$ by $s \teq t$ if $s$ is less than $t$ based on a term ordering,
-replace $\const{C}_{\negvthinspace j}( \vec t ) \teq \const{C}_{\negvthinspace\jPrime}( \vec u )$ with $\bot$ when $j \neq \jPrime$,
-replace all selector terms of the form $\const{s}^{\,k}_{\negvthinspace j}( \const{C}_{\negvthinspace j}(t_1,\ldots,t_n) )$ by $t_k$,
-and replace all occurrences of discriminators $\const{d}_{\negvthinspace j}( \const{C}_{\negvthinspace\jPrime}( \vec t ) )$ by $\top$ if $j = \jPrime$ and $\bot$ otherwise.
+If $\const s^k_{\negvthinspace j}( t )$ is a term in $\tEc$ and $t$ contains the
+constructor $\const{C}_{\negvthinspace j}( \vec u )$, we immediately infer
+$\const s^k_{\negvthinspace j}( t ) \teq u_k$, whereas the calculus we presented
+would apply \rn{Split} and \rn{Inject} before inferring this equality. For the
+sake of reducing the number of unique constraints considered by the calculus, we
+compute a normal form for constraints as a preprocessing step. In particular, we
+replace $t \teq s$ by $s \teq t$ if $s$ is less than $t$ based on a term
+ordering, replace $\const{C}_{\negvthinspace j}( \vec t ) \teq
+\const{C}_{\negvthinspace\jPrime}( \vec u )$ with $\bot$ when $j \neq \jPrime$,
+replace all selector terms of the form $\const{s}^{\,k}_{\negvthinspace j}(
+\const{C}_{\negvthinspace j}(t_1,\ldots,t_n) )$ by $t_k$, and replace
+occurrences of discriminators $\const{d}_{\negvthinspace j}(
+\const{C}_{\negvthinspace\jPrime}( \vec t ) )$ by $\top$ or $\bot$
+based on whether $j = \jPrime$.
 
@@
 \newcommand\HD[1]{\hbox to2.5em{\hfill#1\hfill}}
+
+To evaluate the decision procedure, we generated problems from existing
+Isabelle formalizations using Sledgehammer \cite{paulson-blanchette-2010}.
+Codatatypes being a recent addition to Isabelle
+\cite{blanchette-et-al-2014-impl}, benchmarks are not abundant. We
+included all the theory files from the Isabelle distribution (Distro) and the \emph{Archive
+of Formal Proofs} (AFP) \cite{klein-et-al-afp} that define codatatypes falling
+within the supported fragment. We also included two unpublished theories by
+Peter Gammie and Andreas Lochbihler (\gandl), about Bird and Stern--Brocot trees.
+To also exercise the support
+for datatypes, the benchmarks are complemented by theories about lists
+and tree data structures. The theories were selected before running the
+experiments. The experimental data is publicly available.%
+\footnote{\url{http://TODO/FIXME.}}
+
+For each goal in each theory file, Sledgehammer was invoked to select
+256~lemmas as axioms, which were then monomorphized and translated to SMT-LIB
+along with the goal \cite{boehme-2012-phd}. The resulting problem was given to
+CVC4, running for up to \textbf{TODO: XXX} seconds on the StarExec cluster
+\cite{stump-et-al-2014-starexec}. Problems involving no (co)datatypes were
+filtered out. Due to the lack of machinery for reconstructing inferences about
+(co)datatypes in Isabelle, CVC4 is trusted as an oracle in these experiments.
+
+CVC4 was run on each problem several times, with the support for datatypes and
+codatatypes either enabled or disabled. The contributions of the acyclicity and
+uniqueness rules were also measured, so find out whether these somewhat
+expensive rules are useful in practice. Even when the decision procedure is
+disabled, some of the generated axioms typically include basic properties of
+constructors and selectors, which the decision procedure would recognize as
+tautologies.
+%This allows us to answer the question, \relax{What are the
+%benefits of activating the decision procedure as opposed to letting
+%Sledgehammer do what it would normally do?}
+
+%  * experience with such features is not extremely high -- e.g. arithmetic,
+%    a most useful theory on lots of benchmarks, increases Sledgehammer's
+%    success rate by 2 percentage points, or 4\%, with Z3 in earlier work \cite{xxx}
+
+%  * so when analyzing statistics of this kind, with highly varied problems,
+%    we must not overestimate the importance of a single trick
+%  * on the other hand: every percentage point counts! cite Tom Hales
 
@@
 
-To evaluate the decision procedure, we generated problems from existing
-Isabelle formalizations using Sledgehammer \cite{paulson-blanchette-2010}.
-Codatatypes being a recent addition to Isabelle
-\cite{blanchette-et-al-2014-impl}, benchmarks are not abundant. We
-included all the theory files from the Isabelle distribution (Distro) and the \emph{Archive
-of Formal Proofs} (AFP) \cite{klein-et-al-afp} that define codatatypes falling
-within the supported fragment. We also included two unpublished theories by
-Peter Gammie and Andreas Lochbihler (\gandl), about Bird and Stern--Brocot trees.
-To also exercise the the support
-for datatypes, the benchmarks are complemented by theories about lists
-and tree data structures. The theories were selected before running the
-experiments. The experimental data is publicly available.%
-\footnote{\url{http://TODO/FIXME.}}
-
-For each goal in each theory file, Sledgehammer was invoked to select
-256~lemmas as axioms, which were then monomorphized and translated to SMT-LIB
-along with the goal \cite{boehme-2012-phd}. The resulting problem was given to
-CVC4, running for up to \textbf{TODO: XXX} seconds on the StarExec cluster
-\cite{stump-et-al-2014-starexec}. Problems involving no (co)datatypes were
-filtered out. Due to the lack of machinery for reconstructing inferences about
-(co)datatypes in Isabelle, CVC4 is trusted as an oracle in these experiments.
-
-CVC4 was run on each problem several times, with the support for datatypes and
-codatatypes either enabled or disabled. The contributions of the acyclicity and
-uniqueness rules were also measured, so find out whether these somewhat
-expensive rules are useful in practice. Even when the decision procedure is
-disabled, some of the generated axioms typically include basic properties of
-constructors and selectors, which the decision procedure would recognize as
-tautologies.
-%This allows us to answer the question, \relax{What are the
-%benefits of activating the decision procedure as opposed to letting
-%Sledgehammer do what it would normally do?}
-
-%  * experience with such features is not extremely high -- e.g. arithmetic,
-%    a most useful theory on lots of benchmarks, increases Sledgehammer's
-%    success rate by 2 percentage points, or 4\%, with Z3 in earlier work \cite{xxx}
-
-%  * so when analyzing statistics of this kind, with highly varied problems,
-%    we must not overestimate the importance of a single trick
-%  * on the other hand: every percentage point counts! cite Tom Hales
-
 The results are summarized in Table~\ref{tab:bench}. The decision procedure

### 2015-02-17 tex/body.tex
more

@@
 %not occurring in $\tEc$,
-and let $\Val\ec{u} = {\Varec{u}}$. Therefore, there are initially no
+and set $\Val\ec{u} := {\Varec{u}}$. Therefore, there are initially no
 constraints on the values for any equivalence class $\ec{u}$. The mapping $\Val$
@@
 \end{array}\]
+%
+\rem{TODO: add small examples}
 
@@
 is a \emph{self-similar subterm} of~$t$ if
-$u$ is a strict subterm of $t$,
+$u$ is a proper subterm of $t$,
 $t$ is of the form $\MU x.\; \const{C}( t_1, \ldots, t_n )$,
@@
 The term $t$ is \emph{normal} if it does not contain a self-similar subterm
-and all of its strict subterms are also normal.
+and all of its proper subterms are also normal.
 For example, $t = \MU x.\; \const{C}( \MU y.\; \const{C}( y ) )$ is abnormal
 because $\MU y.\; \const{C}( y )$ is a self-similar subterm of $t$.
-Their arguments have the same expansion with respect to
-$t$:
+Their arguments have the same expansion with respect to~$t$:
 $\interp{t}{\MU y.\; \const{C}( y )} =
@@
 
-For any $\mu$-term $t$ of the form $\MU x.\; \const C( \vec u )$, we can
-construct its \emph{normal form} $\nf{t}$
+For any $\mu$-term $t$ of the form $\MU x.\; \const C( \vec u )$,
+its \emph{normal form} $\nf{t}$ is obtained
 by replacing all of the self-similar subterms of $t$ with $x$
 %and by replacing all of the abnormal subterms of $t$ with their corresponding normal forms.
-and by recursively normalizing $t$'s subterms.
-For example, $\nf{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) )} = \MU x.\; \const{C}( x )$.
+and by recursively normalizing the subterms.
+For variables, $\nf{x} = x$.
+Thus, $\nf{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) )} = \MU x.\; \const{C}( x )$.
 

### 2015-02-17 tex/body.tex
more

@@
 When constructing a model $\M$ for $\Ec$,
-it remains only to specify how $\M$ interprets wrongly applied selector terms and variables.
+it remains only to specify how $\M$ interprets wrongly applied selector terms and variables.%
+\rem{Variables are really part of the model then?}
 For the latter, this will be based on the mapping $\Val$ computed in phase~2 of the calculus.
@@
 Intuitively, this lemma states three properties of $\ValC$ that ensure a normal
-model $\M$ can be trivially constructed for $\Ec$. Part~\PART{1} states that the
+model $\M$ can be trivially constructed for $\Ec$. Property~\PART{1} states that the
 values in the range of $\ValC$ are $\alpha$-equivalent to a value in normal
 model. In particular, this means they are closed, normal, and acyclic when
-required. Part~\PART{2} states that the interpretation of all subterms in the
+required. Property~\PART{2} states that the interpretation of all subterms in the
 range of $\ValC$ depends on its associated variable only. In other words, the
@@
 $\ValC \ec{x}$, independently of the context in which it is interpreted.
-Part~\PART{3} states that $\ValC$ is injective, which in the following ensures
-that a distinct values can be associated with each equivalence class in a
-corresponding model.
+Property~\PART{3} states that $\ValC$ is injective (modulo
+$\alpha$-equivalence), which in the following ensures that a distinct value can
+be associated with each equivalence class in a corresponding model.
 

### 2015-02-17 tex/body.tex
more

@@
 
-%\paragraph{Related Work.}
+\paragraph{Related Work.}
 Although it was written several years ago,
@@
 % \textbf{TODO:} Others (or at least add "e.g.")
+% \textbf{TODO:} Nitpick
 
@@
 since \rn{Split} does not apply to $\Fc$,
-$\ec{u}$ must contain a term of the form $\const C_{\negvthinspace\jPrime}\bigl( \const s^1_{\negvthinspace\jPrime}( u ), \ldots, \const s^{n}_{\negvthinspace\jPrime}( u ) \bigr)$ for some $\jPrime$.
+$\ec{u}$ must contain a term of the form $\smash{\const C_{\negvthinspace\jPrime}\bigl( \const s^1_{\negvthinspace\jPrime}( u ), \ldots, \const s^{n}_{\negvthinspace\jPrime}( u ) \bigr)}$ for some $\jPrime$.
 %By Lemma~\ref{lem:model-completion}\PART{2} and due to our construction of $\ValC$, we have $\ValC \ec{u}$ is of the form
@@
 \rem{TODO: ``theory combination more complicated (and Theorem Completeness too weak)''?}
+\rem{TODO: representation of $\mu$-terms?}
 

### 2015-02-17 tex/body.tex
cleaner handling of memoriam
more

@@
 
-\noindent\hfill\emph{In memoriam Morgan Deters 1979--2015}
+\begin{nomemoriam}
+%%% TODO: Unphantom
+\noindent\hfill\emph{\phantom{In memoriam Morgan Deters 1979--2015}}
+
+\vspace*{-6\smallskipamount}
+\end{nomemoriam}
+\begin{memoriam}
+\noindent\hfill\emph{\relax{In memoriam Morgan Deters 1979--2015}}
+\end{memoriam}
 
@@
 \end{abstract}
+
+\begin{nomemoriam}
+%%% TODO: Unphantom
+\vskip3\smallskipamount
+\end{nomemoriam}
 
@@
 Closely related is the automatic structural
-induction in CVC4 by Reynolds and Kuncak \cite{reynolds-kuncak-2015} and in
-the SPASS-like prototype by Wand and Weidenbach \cite{wand-weidenbach-201x};
-both methods naturally depend on a notion of datatype.
-% \textbf{TODO:} http://frocos2013.loria.fr/slides/Peltier.pdf
-% \textbf{TODO:} Others (or at least add "e.g.")
+induction in superposition-based provers \cite{kersani-peltier-2013,wand-weidenbach-201x}
+and SMT solvers \cite{reynolds-kuncak-2015}.
+%
+Datatypes are explicitly supported by the higher-order model finder Refute
+\cite{weber-2008}. Its successor, Nitpick \cite{blanchette-nipkow-2010}, can
+also find models involving cyclic codatatype values. The $\mu$-notation is
+inspired by the $\mu$-calculus \cite[etc.]{kozen-1983,endrullis-et-al-2011}.
+
 % \textbf{TODO:} Nitpick
@@
 %\textbf{TODO}:
-%  * mu-calculus
 %  * Klop et al. (streams) Andrew Lice % http://joerg.endrullis.de/

### 2015-02-17 tex/body.tex
more

@@
 
+\looseness=-1
 Third, \emph{it must be possible to express cyclic }(\emph{$\omega$-regular}) \emph{values as closed terms and
@@
 and for theory combinations. The $\mu$-binder notation associates a name with
-a subterm; it is used to represent cyclic values in the generated models and
-in the metatheory. For example,
+a subterm; it is used to represent cyclic values in the generated
+models\begin{rep} and in the metatheory\end{rep}. For example,
 the $\mu$-term $\const{SCons}(1,\: \MU s.\; \const{SCons}(0,\: \const{SCons}(9,\: s)))$
-stands for the lasso-shaped sequence $1, 0, 9, 0, 9, 0, 9, \ldots$\afterLdots.
+stands for the lasso\begin{rep}-shaped\end{rep} sequence $1, 0, 9, 0, 9, 0, 9,
+\ldots$\afterLdots.
 
@@
 (Section~\ref{sec:the-theory-solver}).
-\rem{CHECK WITH IMPLEMENTATION, but round to nearest 50 or 100 LOC.}
-It consists of about 2000 lines of \cpp{} code, among which 1600 are
+It consists of about 2000 lines of \cpp{} code, among which \NOK{1600} are
 shared between datatypes and codatatypes. The code is integrated in the
@@
 
-\paragraph{Related Work.}
+\begin{rep}\paragraph{Related Work.} \end{rep}%
 Although it was written several years ago,
@@
 related work. Since then, datatypes have been added to the SMT solver
-Z3 in unpublished work by Leonardo de Moura and to a SPASS-like prototypical
-superposition prover called Pirate by Daniel Wand.
+Z3 in unpublished work by Leonardo de Moura and to a SPASS-like %prototypical
+superposition prover by Daniel Wand \cite{wand-2014}.
 Closely related is the automatic structural
-induction in superposition-based provers \cite{kersani-peltier-2013,wand-weidenbach-201x}
+induction in superposition provers \cite{kersani-peltier-2013,wand-weidenbach-201x}
 and SMT solvers \cite{reynolds-kuncak-2015}.

### 2015-02-17 tex/body.tex
more edits

@@
 For example, the type
-%
-\[\keyw{codatatype}~\;\ty{stream}_{\,\tau} \,=\, \const{SCons}(\tau,\: \ty{stream}_{\,\tau})\]
-%
+
+\begin{conf}
+\kern-2pt %%% TYPESETTING: HACK
+\end{conf}
+
+\vskip\abovedisplayskip
+\noindent\centerline{$\keyw{codatatype}~\;\ty{stream}_{\,\tau} \,=\, \const{SCons}(\tau,\: \ty{stream}_{\,\tau})$}
+\vskip\belowdisplayskip
+
+\begin{conf}
+\kern-2pt %%% TYPESETTING: HACK
+\end{conf}
+
+\noindent
 of infinite sequences, or streams, over an element type $\tau$ is allowed; the
@@
 
-\looseness=-1
 Third, \emph{it must be possible to express cyclic }(\emph{$\omega$-regular}) \emph{values as closed terms and
@@
 \begin{rep}\paragraph{Related Work.} \end{rep}%
-Although it was written several years ago,
+Although it was written a few years ago,
 Barrett et al.\ \cite{barrett-et-al-2007} still provides a good account of
 related work. Since then, datatypes have been added to the SMT solver
-Z3 in unpublished work by Leonardo de Moura and to a SPASS-like %prototypical
-superposition prover by Daniel Wand \cite{wand-2014}.
+Z3 in unpublished work \begin{rep}by Leonardo de Moura \end{rep}and to a SPASS-like %prototypical
+superposition prover \begin{rep}by Daniel Wand\end{rep} \cite{wand-2014}.
 Closely related is the automatic structural
@@
 
+\begin{conf}
+\kern-2.9pt %%% TYPESETTING: HACK
+\end{conf}
+
 \paragraph{Conventions.}
@@
 A signature $\Sig = (\Types, \Funcs)$ consists of a set of types $\Types$ and a
-set of function symbols $\Funcs$. Types are simply atomic sorts, with no
-structure, and interpreted by nonempty domains. The set~$\Types$ must contain a
+set of function symbols $\Funcs$. Types are atomic sorts and interpreted by
+nonempty domains. The set~$\Types$ must contain a
 distinguished type \ty{bool}\begin{rep} interpreted as the set of truth
@@
 
+\pagebreak
+
 \begin{rep}
@@
 $\const{S}$, $\const{Infty}$) are reserved for constructors. With each function symbol \const{f}
-is associated a list of argument types $\tau_1,\ldots,\tau_n$ (with $n \ge 0$)
-and a return type $\tau$. This connection can be expressed %compactly
-as $\const{f} : \tau_1 \times \cdots \times \tau_n \to \tau$\begin{rep},
+is associated a list of argument types $\tau_1,\ldots,\tau_n$ ($n \ge 0$)
+and a return type $\tau$. This connection is written %compactly
+$\const{f} : \tau_1 \times \cdots \times \tau_n \to \tau$\begin{rep},
 which collapses to $\const{f} : \tau$ if $n = 0$\end{rep}.
-For a term $t$, the notation $t\typ{\tau}$ indicates that it has type $\tau$.
-Functions are invoked in the standard way, with $\const{f}(t_1,\ldots,t_n)$
-applying the $n$-ary function symbol
+The notation $t\typ{\tau}$ indicates that term $t$ has type $\tau$.
+Functions invocations $\const{f}(t_1,\ldots,t_n)$
+apply the $n$-ary function symbol
 \const{f} to $n$ well-typed arguments $t_1 :\nobreak \tau_1$, \ldots, $t_n :

### 2015-02-17 tex/body.tex
more

@@
 Compared
-with the conventional inductive definition
+with the conventional
 \[\keyw{datatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat}) \,\mid\, \const{Infty}\]
@@
 modulo theories) solvers, includes a theory of datatypes.
-
-\pagebreak[3]
 
@@
 
-\begin{rep}\paragraph{Related Work.} \end{rep}%
-Although it was written a few years ago,
+\begin{rep}\paragraph{Related Work.} Although\end{rep}%
+\begin{conf}Although\end{conf} it was written a few years ago,
 Barrett et al.\ \cite{barrett-et-al-2007} still provides a good account of
@@
 % at the metalevel
-through subscripts---for example, $\ty{stream}_{\,\tau}$ denotes a
+through subscripts%
+\begin{conf}
+(e.g., $\ty{stream}_{\,\tau}$)%
+\end{conf}%
+\begin{rep}---for example, $\ty{stream}_{\,\tau}$ denotes a
 family of ground types including
 $\ty{stream}_{\,\ty{int}}$, $\ty{stream}_{\,\ty{bool}}$,
-and \smash{$\ty{stream}_{\,\ty{stream}_{\,\ty{real}}}$}.
+and \smash{$\ty{stream}_{\,\ty{stream}_{\,\ty{real}}}$}\end{rep}.
 

### 2015-02-17 tex/body.tex
updated stats

@@
 No (co)datatypes
-  & 229 & 765 & \phantom{0}61 & 1055 \\[\jot]
+  & 221 & 760 & \phantom{0}55 & 1036 \\[\jot]
 %Datatype simplication only
@@
 Datatypes without \rn{Acyclic}
-  & 235 & 770 & \phantom{0}61 & 1066 \\
+  & 227 & 765 & \phantom{0}55 & 1047 \\
 Full datatypes
-  & 235 & 773 & \phantom{0}61 & 1069 \\[\jot]
+  & 227 & 771 & \phantom{0}55 & 1053 \\[\jot]
 %Codatatype simplication only
@@
 Codatatypes without \rn{Unique}
-  & 230 & 794 & \phantom{0}65 & 1089 \\
+  & 222 & 789 & \phantom{0}56 & 1067 \\
 Full codatatypes
-  & 231 & 794 & \bfseries \phantom{0}68 & 1093 \\[\jot]
+  & 223 & 789 & \bfseries \phantom{0}59 & 1071 \\[\jot]
 Full (co)datatypes
-  & \bfseries 237 & \bfseries 802 & \bfseries \phantom{0}68 & \bfseries 1107 \\[\jot]
+  & \bfseries 229 & \bfseries 800 & \bfseries \phantom{0}59 & \bfseries 1088 \\[\jot]
 Total number of goals
@@
 makes a difference across all theories. Overall, it accounts for a success rate
-increase of \OK{4.9\%}. Moreover, every aspect of the procedure, including the
+increase of \OK{5.0\%}. Moreover, every aspect of the procedure, including the
 more expensive rules, make a contribution.

### 2015-02-17 tex/body.tex
more

--- 2015-02-17 tex/body.tex
@@
 \newcommand\gandl{G\&L}
-\newcommand\HD[1]{\hbox to2.5em{\hfill#1\hfill}}
+\newcommand\HD[1]{\hbox to2.3em{\hfill\!\!\!\!#1\!\!\!\!\hfill}}
 
@@
 
+%  & 879 & 2974\phantom{0} & 317 & 4170
+
 For each goal in each theory file, Sledgehammer was invoked to select
@@
 
+%%% TODO: Fix data
+
 \begin{table*}[tbh!]
 \normalsize
-\begin{center}\begin{tabular}{@{\;}l@{\kern1.5em}c@{\kern.5em}c@{\kern.5em}c@{\kern1.3em}c@{\;}}
-  & \HD{Distro} & \HD{AFP} & \HD{\gandl} & \relax{All}
+\begin{center}\begin{tabular}{@{\;}l@{\kern1.5em}cc@{\kern.5em}cc@{\kern.5em}cc@{\kern1.3em}cc@{\;}}
+  & \multicolumn{2}{c}{Distro} & \multicolumn{2}{c}{AFP} & \multicolumn{2}{c}{\gandl} & \multicolumn{2}{c}{All} \\
+  & \HD{Z3} & \HD{CVC4} & \HD{Z3} & \HD{CVC4} & \HD{Z3} & \HD{CVC4} & \HD{Z3} & \HD{CVC4}
 \MIDRULE
 No (co)datatypes
-  & 229 & 765 & \phantom{0}61 & 1055 \\[\jot]
+  & 209 & 221 & 777 & 785 & \phantom{0}51 & \phantom{0}52 & 1037 & 1048 \\[\jot]
 %Datatype simplication only
 %  &  \\
-Datatypes without \rn{Acyclic}
-  & 235 & 770 & \phantom{0}61 & 1066 \\
+Datatypes w/o \rn{Acyclic}
+  & & 235 & & 770 & & \phantom{0}61 & & 1066 \\
 Full datatypes
-  & 235 & 773 & \phantom{0}61 & 1069 \\[\jot]
+  & 213 & 227 & 791 & 786 & \phantom{0}51 & \phantom{0}52 & 1055 & 1065 \\[\jot]
 %Codatatype simplication only
 %  &  \\
-Codatatypes without \rn{Unique}
-  & 230 & 794 & \phantom{0}65 & 1089 \\
+Codatatypes w/o \rn{Unique}
+  & & 230 & & 794 & & \phantom{0}65 & & 1089 \\
 Full codatatypes
-  & 231 & 794 & \bfseries \phantom{0}68 & 1093 \\[\jot]
+  & & 231 & & 794 & & \bfseries \phantom{0}68 & & 1093 \\[\jot]
 Full (co)datatypes
-  & \bfseries 237 & \bfseries 802 & \bfseries \phantom{0}68 & \bfseries 1107 \\[\jot]
-Total number of goals
-  & 879 & 2974\phantom{0} & 317 & 4170
+  & & \bfseries 237 & & \bfseries 802 & & \bfseries \phantom{0}68 & & \bfseries 1107% \\[\jot]
 \end{tabular}\end{center}

### 2015-02-17 tex/body.tex
Fix minor things, add paragraph on implicitly computing mu-terms.  Add citation.

@@
 Legal substitutions for ${\Varec{x}}$ may cause the range of $\Val$ to contain abnormal terms.
-After applying the substitution for the third term mentioned above\rem{which third term?},
+After applying the substitution for $\MU {\Varec{x}}.\; \const{D}( \const{C}( {\Varec{x}} ) )$, we have that
 $\Val \ec{w_1}$ is $\MU {\Varec{w_1}}.\; \const{C}( \MU {\Varec{x}}.\; \const{D}( \const{C}( {\Varec{x}} ) ) )$,
@@
 the value $\bot$. Then, at a high level, by the Nelson--Oppen combination
-framework, each $T_i$-solver for $i = 1, \ldots, l$ takes as input a combination
+framework (its variant used in CVC4 is described in~\cite{jovanovic2011sharing}), 
+each $T_i$-solver for $i = 1, \ldots, l$ takes as input a combination
 of (1) the purified form of $T_i$-constraints occurring in $M$, where terms
@@
 solver and not by the theory solver. All other rules add equalities to the
-internal state of the theory solver. The rules in phase~1 (along with the
-optimization for collapsing selectors) and the \rn{Single} rule are performed
+internal state of the theory solver. The rules in phase~1 along with the
+optimization for collapsing selectors are performed
 eagerly---that is, for partial satisfying assignments $M$---while the rules in
-phase~2 and the \rn{Split} rule are performed only for complete satisfying
-assignments $M$, and in that order.\rem{in what order?}
-
-\rem{TODO: * sharing is caring (let's cite it for a change)}
+phases~2 and~3 are performed only for complete satisfying
+assignments $M$.
+
+Prior to constructing a model for $F$, the theory solver does not explicitly construct $\mu$-terms nor the map $\Val$.
+Instead, $\Val$ is computed implicitly by traversing the equivalence classes of $\Ec$ during phase~2.
+To recognize when the precondition of \rn{Acyclic} holds,
+the procedure considers each datatype equivalence class $\ec{t}$ containing a constructor $\const{C}( t_1, \ldots, t_n )$.
+It visits $\ec{t_1}, \ldots, \ec{t_n}$ and all arguments of constructors in these equivalence classes in the same way.
+If while doing so it returns to $\ec{t}$, then it deduces that \rn{Acyclic} applies to $t$.
+To recognize when the precondition of \rn{Unique} holds,
+the procedure considers the set $S$ of all codatatype equivalence classes.
+It simultaneously visits the equivalence classes of arguments of constructor terms in each equivalence class in $S$,
+while partitioning $S$ into $S_1, \ldots, S_n$,
+where all terms in $S_i$ have the same constructor type for all their subterms of type $\Data \cup \Codata$,
+and are equivalent for all subterms of type $\Nondata$.
+If after doing so, some $S_i$ contains two terms $u$ and $v$, then it deduces that \rn{Unique} applies to $u$ and $v$.
+
+%\rem{TODO: * sharing is caring (let's cite it for a change)}
 \rem{TODO: ``theory combination more complicated (and Theorem Completeness too weak)''?}
-\rem{TODO: representation of $\mu$-terms?}
+%\rem{TODO: representation of $\mu$-terms?}
 
 \paragraph{Optimizations.}
-Our implementation uses several optimizations not reflected in the presentation
-of the calculus, which closely follow the approach outlined by Barrett et al.\
-\cite{barrett-et-al-2007}.%
-\rem{What closely follows the approach by Barrett et al.? The optimizations or
-the calculus?}
+Our implementation uses several optimizations which
+follow the approach outlined by Barrett et al.\cite{barrett-et-al-2007}
+and are not reflected in the presentation of the calculus.
 We only briefly mention these optimizations, since
@@
 along with the goal \cite{boehme-2012-phd}. The resulting problem was given to
-CVC4, running for up to \textbf{TODO: XXX} seconds on the StarExec cluster
+CVC4, running for up to 60 seconds on the StarExec cluster
 \cite{stump-et-al-2014-starexec}. Problems involving no (co)datatypes were
@@
 %% alliteration
-On the practical side, we obtained respectable results on Isabelle benchmarks.
+On the practical side, we obtained respectable results on Isabelle benchmarks. \rem{emphasize new capabilities allows us to improve state of the art.}
 
@@
 \paragraph{\ackname.}
-We owe a great debt to Clark Barrett and Cesare Tinelli for datatype case blah
-blah blah. \textbf{TODO FIXME}
-Morgan Deters?
+We owe a great debt to the development team of CVC4, 
+including Clark Barrett and Cesare Tinelli,
+and in particular Morgan Deters,
+who jointly the first author developed the theory solver for datatypes in CVC4.
 %

### 2015-02-17 tex/body.tex
more

--- 2015-02-17 tex/body.tex
@@
 \end{conf}%
-The procedure is implemented in the SMT solver CVC4 as a combination
+The procedure is implemented in the SMT solver CVC4 \cite{barrett-et-al-2011} as a combination
 of rewriting and a theory solver
@@
 related work. Since then, datatypes have been added to the SMT solver
-Z3 in unpublished work \begin{rep}by Leonardo de Moura \end{rep}and to a SPASS-like %prototypical
+Z3 \cite{de-moura-bjoerner-2008} in unpublished work \begin{rep}by Leonardo de Moura \end{rep}and to a SPASS-like %prototypical
 superposition prover \begin{rep}by Daniel Wand\end{rep} \cite{wand-2014}.
@@
 \newcommand\gandl{G\&L}
-\newcommand\HD[1]{\hbox to2.3em{\hfill\!\!\!\!#1\!\!\!\!\hfill}}
+\newcommand\HD[1]{\hbox to2.25em{\footnotesize\hfill\!\!\!\!#1\!\!\!\!\hfill}}
 
@@
 constructors and selectors, which the decision procedure would recognize as
-tautologies.
+tautologies. Z3 is included for comparison when this is possible.
 %This allows us to answer the question, \relax{What are the
@@
 
-\begin{table*}[tbh!]
+\begin{table*}[t!]
 \normalsize

### 2015-02-17 tex/body.tex
more

@@
 (Section~\ref{sec:a-ground-decision-procedure-for-co-datatypes}).
-We also discuss extensions of the procedure for handling quantifiers
-(Section~\ref{sec:extension-to-quantified-formulas}).
 The procedure is described abstractly as a calculus and is
@@
 \begin{conf}
-Detailed proofs of soundness and completeness are included in the technical report
-associated with this paper \cite{our-report}.
+Detailed proofs of soundness and completeness are included in a technical report
+\cite{our-report}.
 \end{conf}%
 The procedure is implemented in the SMT solver CVC4 \cite{barrett-et-al-2011} as a combination
-of rewriting and a theory solver
-(Section~\ref{sec:the-theory-solver}).
+of rewriting and a theory solver (Section~\ref{sec:implementation-as-a-theory-solver-in-cvc4}).
 It consists of about 2000 lines of \cpp{} code, among which \NOK{1600} are
@@
 
-\pagebreak
+\pagebreak[2]
 
@@
 
-\section{Extension to Quantified Formulas}
-\label{sec:extension-to-quantified-formulas}
-
-  * universal conjecture is no problem:
-    * falls into the ground fragment via negation
-      and skolemization
-  * nor is existential axiom a problem
-  * what's interesting: universal axioms and existential conjectures
-
-Blah.
-
 \section{Implementation as a Theory Solver in CVC4}
-\label{sec:the-theory-solver}
+\label{sec:implementation-as-a-theory-solver-in-cvc4}
 
@@
 \rem{TODO: representation of $\mu$-terms?}
+
+\paragraph{Extension to Quantified Formulas.}
+Two paras. One general one that talks about:
+\begin{verbatim}
+  * universal conjecture is no problem:
+    * falls into the ground fragment via negation
+      and skolemization
+  * nor is existential axiom a problem
+  * what's interesting: universal axioms and existential conjectures
+\end{verbatim}
+One more specific that talks about the cardinality calculus.
 
@@
 We owe a great debt to Clark Barrett and Cesare Tinelli for datatype case blah
-blah blah. \textbf{TODO FIXME}
+blah blah. \rem{TODO FIXME}
 Morgan Deters?

### 2015-02-17 tex/body.tex

--- 2015-02-17 tex/body.tex
@@
 \begin{conf}
-Detailed proofs of soundness and completeness are included in the technical report
-associated with this paper \cite{our-report}.
+Detailed proofs of soundness and completeness are included in a technical report
+\cite{our-report}.
 \end{conf}%
-The procedure is implemented in the SMT solver CVC4 as a combination
-of rewriting and a theory solver
-(Section~\ref{sec:the-theory-solver}).
+The procedure is implemented in the SMT solver CVC4 \cite{barrett-et-al-2011} as a combination
+of rewriting and a theory solver (Section~\ref{sec:implementation-as-a-theory-solver-in-cvc4}).
 It consists of about 2000 lines of \cpp{} code, among which \NOK{1600} are
@@
 
+\paragraph{Extension to Quantified Formulas.}
+Two paras. One general one that talks about:
+\begin{verbatim}
+  * universal conjecture is no problem:
+    * falls into the ground fragment via negation
+      and skolemization
+  * nor is existential axiom a problem
+  * what's interesting: universal axioms and existential conjectures
+\end{verbatim}
+One more specific that talks about the cardinality calculus.
+
 \paragraph{Optimizations.}
@@
 and in particular Morgan Deters,
-who jointly the first author developed the theory solver for datatypes in CVC4.
+who jointly with the first author developed the theory solver for datatypes in
+CVC4.
 %
@@
 %
-Blanchette's research was partially supported by the Deutsche
+The second author's research was partially supported by the Deutsche
 Forschungs\-gemein\-schaft (DFG) project

### 2015-02-17 tex/body.tex
more

@@
 %% alliteration
-On the practical side, we obtained respectable results on Isabelle benchmarks. \rem{emphasize new capabilities allows us to improve state of the art.}
-
-This work is part of a larger program that aims at enriching automatic provers
+On the practical side, the empirical results on Isabelle benchmarks
+confirm that CVC4's new capabilities improve the state of the art.
+
+This work is part of a program that aims at enriching automatic provers
 with high-level features and at reducing the gap between automatic and
@@
 proof reconstruction for CVC4's (co)datatype inferences in Isabelle. We also
-want to try out CVC4 for higher-order model finding in Isabelle, as an
-alternative to the counterexample generator Nitpick
-\cite{blanchette-nipkow-2010}. Finally, it might be worthwhile to enrich SMT
-solvers with explicit support for recursive and corecursive functions.
+want to try out the solver for model finding in Isabelle, as an alternative to
+Nitpick. Finally, it might be worthwhile to extend SMT solvers with dedicated
+support for (co)recursive functions.
 

### 2015-02-17 tex/body.tex
more

@@
 We introduced a decision procedure for the ground theory of datatypes and
-codatatypes and implemented it in the SMT solver CVC4. Our main
+codatatypes. Our main
 contribution has been the support for codatatypes. Both the metatheory and
-the implementation rely heavily on $\mu$-terms to represent cyclic values,
+the implementation in CVC4 rely heavily on $\mu$-terms to represent cyclic values,
 a complication specific to codatatypes.
@@
 %% alliteration
-On the practical side, the empirical results on Isabelle benchmarks
+The empirical results on Isabelle benchmarks
 confirm that CVC4's new capabilities improve the state of the art.
@@
 interactive theorem proving. As future work, it would be useful to implement
-proof reconstruction for CVC4's (co)datatype inferences in Isabelle. We also
-want to try out the solver for model finding in Isabelle, as an alternative to
+proof reconstruction for (co)data\-type inferences in Isabelle. We also
+want to try out CVC4 for model finding in Isabelle, as an alternative to
 Nitpick. Finally, it might be worthwhile to extend SMT solvers with dedicated

### 2015-02-17 tex/body.tex
more

@@
 value~$\infty$.
-Compared
-with the conventional
+Compared with the conventional inductive definition
 \[\keyw{datatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat}) \,\mid\, \const{Infty}\]
@@
 of infinite sequences, or streams, over an element type $\tau$ is allowed; the
-corresponding datatype would be rejected as non-well-founded \cite{blanchette-et-al-2015-esop}.
+corresponding datatype would be rejected as non-well-founded%
+\begin{rep} \cite{blanchette-et-al-2015-esop}\end{rep}.
 
@@
 \begin{center}\begin{tabular}{@{\;}l@{\kern1.5em}cc@{\kern.5em}cc@{\kern.5em}cc@{\kern1.3em}cc@{\;}}
-  & \multicolumn{2}{c}{Distro} & \multicolumn{2}{c}{AFP} & \multicolumn{2}{c}{\gandl} & \multicolumn{2}{c}{All} \\
+  & \multicolumn{2}{c}{Distro} & \multicolumn{2}{c}{AFP} & \multicolumn{2}{c}{\gandl} & \multicolumn{2}{c}{Overall} \\
   & \HD{Z3} & \HD{CVC4} & \HD{Z3} & \HD{CVC4} & \HD{Z3} & \HD{CVC4} & \HD{Z3} & \HD{CVC4}
@@
 No (co)datatypes
-  & 209 & 221 & 777 & 785 & \phantom{0}51 & \phantom{0}52 & 1037 & 1048 \\[\jot]
+  & 209 & 221 & 777 & 785 & \phantom{0}51 & \phantom{0}52 & 1037 & 1048 \\%[\jot]
 %Datatype simplication only
@@
 Full datatypes
-  & 213 & 227 & 791 & 786 & \phantom{0}51 & \phantom{0}52 & 1055 & 1065 \\[\jot]
+  & 213 & 227 & 791 & 786 & \phantom{0}51 & \phantom{0}52 & 1055 & 1065 \\%[\jot]
 %Codatatype simplication only
@@
 Full codatatypes
-  & & 231 & & 794 & & \bfseries \phantom{0}68 & & 1093 \\[\jot]
+  & & 231 & & 794 & & \bfseries \phantom{0}68 & & 1093 \\%[\jot]
 Full (co)datatypes

### 2015-02-17 tex/body.tex
squeeze

@@
 Compared with the conventional inductive definition
+\begin{rep}
 \[\keyw{datatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat}) \,\mid\, \const{Infty}\]
+\end{rep}
 with a dedicated constructor for representing infinity, the codatatype avoids
@@
 modulo theories) solvers, includes a theory of datatypes.
+
+\pagebreak[2]
 
@@
 \begin{conf}
-\kern-2pt %%% TYPESETTING: HACK
+\kern-2pt %%% TYPESETTING: Crude hack
 \end{conf}
@@
 
-\begin{rep}\paragraph{Related Work.} Although\end{rep}%
-\begin{conf}Although\end{conf} it was written a few years ago,
+\begin{rep}\paragraph{Related Work.} Despite\end{rep}%
+\begin{conf}Despite\end{conf} its age,
 Barrett et al.\ \cite{barrett-et-al-2007} still provides a good account of
 related work. Since then, datatypes have been added to the SMT solver
-Z3 \cite{de-moura-bjoerner-2008} in unpublished work \begin{rep}by Leonardo de Moura \end{rep}and to a SPASS-like %prototypical
+Z3 \cite{de-moura-bjoerner-2008} \begin{rep}in unpublished work by Leonardo de Moura \end{rep}and to a SPASS-like %prototypical
 superposition prover \begin{rep}by Daniel Wand\end{rep} \cite{wand-2014}.
 Closely related is the automatic structural
-induction in superposition provers \cite{kersani-peltier-2013,wand-weidenbach-201x}
-and SMT solvers \cite{reynolds-kuncak-2015}.
+induction in both types of provers \cite{kersani-peltier-2013,wand-weidenbach-201x,reynolds-kuncak-2015}.
 %
@@
 \begin{conf}
-\kern-2.9pt %%% TYPESETTING: HACK
+\kern-2.9pt %%% TYPESETTING: Crude hack
 \end{conf}
@@
 
-\pagebreak[2]
-
 \begin{rep}
@@
 \end{rep}
-Symbols starting with an uppercase letter (e.g.,
-$\const{S}$, $\const{Infty}$) are reserved for constructors. With each function symbol \const{f}
-is associated a list of argument types $\tau_1,\ldots,\tau_n$ ($n \ge 0$)
-and a return type $\tau$. This connection is written %compactly
+Names starting with an uppercase letter (e.g.,
+$\const{S}$) are reserved for constructors. With each function symbol \const{f}
+is associated a list of argument types $\tau_1,\ldots,\tau_n$ (for $n \ge 0$)
+and a return type $\tau$, written
 $\const{f} : \tau_1 \times \cdots \times \tau_n \to \tau$\begin{rep},
 which collapses to $\const{f} : \tau$ if $n = 0$\end{rep}.
-The notation $t\typ{\tau}$ indicates that term $t$ has type $\tau$.
+The notation $t\typ{\tau}$ indicates that term $t$ has type~$\tau$.
+\begin{rep}%
 Functions invocations $\const{f}(t_1,\ldots,t_n)$
@@
 parentheses in terms.
-\begin{rep}%
 The operator $\bigwedge_{\,i}\, \varphi_i$ abbeviates a conjunction
@@
 $\bar x$ abbreviates a list or tuple $x_1,\ldots,x_n$.
+
+\pagebreak
 
@@
 because they can be abstracted away as variables when combining theories.
-Exceptionally, it is convenient to use numeric constants ($0$, $1$, \ldots)\ in
-examples.
+%Exceptionally, it is convenient to use numeric constants ($0$, $1$, \ldots)\ in
+%examples.
 
@@
 together in a mutual fashion even if they are not actually (co)recursive.
+\begin{rep}%
 The semantic notion is more precise and is the one that interests us.
+\end{rep}
 %

### 2015-02-17 tex/body.tex
using J instead of I -- I looks too naked

@@
 datatypes and codatatypes},
-which in addition to $\Sig$ defines a class of $\Sig$-interpretations $\mathbf{I}$,
+which in addition to $\Sig$ defines a class of $\Sig$-interpretations $\mathbf{J}$,
 namely the ones that satisfy the %axioms --- existence of all infinite codatatype values cannot be axiomatized
 properties mentioned in this section. In particular, the interpretations
-in~$\mathbf{I}$ share the same fixed interpretation for constructor terms and correctly
+in~$\mathbf{J}$ share the same fixed interpretation for constructor terms and correctly
 applied selector terms but may differ on wrongly applied
@@
 A formula $\varphi$ is $\thD$-\emph{satisfiable} if there
-exists an interpretation in $\mathbf{I}$ that satisfies $\varphi$.
+exists an interpretation in $\mathbf{J}$ that satisfies $\varphi$.
 % otherwise, it is $\thD$-unsatisfiable.  (I trust the readers' intelligence when

### 2015-02-18 tex/body.tex
Shrink example 3.  Minor changes regarding model/interpretation.

@@
 %for a capture-avoiding renaming of $\mu$-bound variables,
-then $t \teq u$. This may seem too restrictive, since
+then $t \teq u$. This precondition may seem too restrictive, since
 $\MU x.\; \const{S}(x)$ and $\MU y.\; \const{S}(\const{S}(y))$ specify the same
@@
 make some progress by discovering that the subterm $\const{S}(y)$ of $\MU y.\;
-\const{S}(\const{S}(y))$ must be equal to the entire term.
+\const{S}(\const{S}(y))$ must be equal to the entire term,
+as demonstrated in the following example.
 
@@
 $\{
-w_1 \teq \const{C}( x ),\;
-w_2 \teq \const{E}( w_3 ),\;
-x \not\teq y_2,\;
-y_1 \teq \const{C}( y_2 ),\;
-y_2 \teq \const{D}( x ),\;
-z_1 \teq \const{E}( z_2 ),\;
-z_2 \teq \const{C}( z_1 )
+w \teq \const{C}( z ),\;
+z \teq \const{D}( x ),\;
+u \teq \const{C}( x ),\;
+v \teq \const{E}( y ),\;
+x \not\teq z
 \}$.
 After applying the calculus to saturation on $\Ec$, the mapping $\Val$ is as follows:
-\[\begin{array}{r@{}c@{}l@{\;}l@{\qquad}r@{}c@{}r@{\;}l@{}l@{\;}l}
-\Val \ec{w_1} & {}\,=\,{} & \MU {\Varec{w_1}}. & \const{C}( {\Varec{x}} ) &
-\Val \ec{y_1} & {}\,=\,{} & \MU {\Varec{y_1}}. & \const{C}( & \MU {\Varec{y_2}}. & \const{D}( {\Varec{x}} ) ) \\
-\Val \ec{w_2} & = & \MU {\Varec{w_2}}. & \const{E}( {\Varec{w_3}} ) &
-\Val \ec{y_2} & = & \MU {\Varec{y_2}}. & \multicolumn{2}{@{}l@{}}{\const{D}( {\Varec{x}} )} \\
-\Val \ec{w_3} & = & {\Varec{w_3}} &&
-\Val \ec{z_1} & = & \MU {\Varec{z_1}}. & \const{E}( & \MU {\Varec{z_2}}. & \const{C}( {\Varec{z_1}} ) ) \\
-\Val \ec{x} & = & {\Varec{x}} &&
-\Val \ec{z_2} & = & \MU {\Varec{z_2}}. & \const{C}( & \MU {\Varec{z_1}}. & \const{E}( {\Varec{z_2}} ) )
+%\[\begin{array}{r@{}c@{}l@{\;}l@{\qquad}r@{}c@{}r@{\;}l@{}l@{\;}l}
+%\[\begin{array}{r@{}c@{}l@{\;}l@{\qquad}r@{}c@{}r@{\;}l@{}l@{\;}l@{\qquad}r@{}c@{}r@{\;}l}
+\[\begin{array}{r@{}c@{}r@{\qquad}r@{}c@{}r@{\;}l@{}l@{\;}l@{\qquad}r@{}c@{}l@{\;}l}
+\Val \ec{x} & = & {\Varec{x}} &
+\Val \ec{z} & = & \MU {\Varec{z}}. & \multicolumn{2}{@{}l@{}}{\const{D}( {\Varec{x}} )} & &
+\Val \ec{u} & {}\,=\,{} & \MU {\Varec{u}}. & \const{C}( {\Varec{x}} ) \\
+\Val \ec{y} & = & {\Varec{y}} &
+\Val \ec{w} & {}\,=\,{} & \MU {\Varec{w}}. & \const{C}( & \MU {\Varec{z}}. & \const{D}( {\Varec{x}} ) ) &
+\Val \ec{v} & = & \MU {\Varec{v}}. & \const{E}( {\Varec{y}} )
 \end{array}\]
@@
 To construct a completion $\ValC$, we must choose values for the free
-variables of $\Val$, namely ${\Varec{w_3}}$ and ${\Varec{x}}$.
+variables of $\Val$, namely ${\Varec{x}}$ and ${\Varec{y}}$.
 Modulo $\alpha$-equivalence, the set \smash{$\mathcal{T}^{\smash{\M}}_{{\Varec{x}}}( \Val )$} is
@@
 \{
-\MU x.\; \const{C}( x ),\;
+\MU x.\; \const{D}( x ),\;
 \MU x.\; \const{C}( \const{D}( x )),\;
 \const{C}( \MU x.\; \const{D}( x )),\;
-\MU x.\; \const{C}( \const{E}( x )),\;
-\MU x.\; \const{D}( x ),\;
-\MU x.\; \const{E}( \const{C}( x ))
+\MU x.\; \const{C}( x )
 \}
@@
 \]
-Consider a model $\M$ that interprets variables in $\Ec$ based on the mapping $\Val$:
-$x$ is interpreted as $\Val \ec{x}$, $y_1$ as $\Val \ec{y_1}$, and so on.
-Assigning a value for ${\Varec{x}}$ that is $\alpha$-equivalent to a value in $\mathcal{T}^{\smash{\M}}_{{\Varec{x}}}( \Val )$
+Consider a normal interpretation $\mathbf{J}$ \rem{Appropriate symbol?} that interprets variables in $\Ec$ based on the mapping $\Val$:
+$x$ is interpreted as $\Val \ec{x}$, $y$ as $\Val \ec{y}$, and so on.
+Assigning a value for $\Val \ec{x}$ that is $\alpha$-equivalent to a value in $\mathcal{T}^{\smash{\M}}_{{\Varec{x}}}( \Val )$
 may cause values in the range of $\Val$ to become $\alpha$-equivalent,
-which in turn may cause $\M$ to
-become a non-model.
-% "be inconsistent." Models can't be "inconsistent".
+which in turn may cause $\Ec$ to be falsified by $\mathbf{J}$.
 For example, assign $\MU {\Varec{x}}.\; \const{D}( {\Varec{x}} )$ for ${\Varec{x}}$.
-After this substitution, $\Val \ec{y_2}$ is $\MU {\Varec{y_2}}.\; \const{D}( \MU {\Varec{x}}.\; \const{D}( {\Varec{x}} ) )$,
-which has normal form $\MU {\Varec{y_2}}.\; \const{D}( {\Varec{y_2}} )$,
+After this substitution, $\Val \ec{z}$ is $\MU {\Varec{z}}.\; \const{D}( \MU {\Varec{x}}.\; \const{D}( {\Varec{x}} ) )$,
+which has normal form $\MU {\Varec{z}}.\; \const{D}( {\Varec{z}} )$,
 which is $\alpha$-equivalent to $\MU {\Varec{x}}.\; \const{D}( {\Varec{x}} )$.
-However, this contradicts the disequality $x \not\teq y_2$ in~$\Ec$.
+However, this contradicts the disequality $x \not\teq z$ in~$\Ec$.
 On the other hand, if the value assigned to ${\Varec{x}}$ is fresh, %not $\alpha$-equivalent to any term in $\mathcal{T}^{\smash{\M}}_{{\Varec{x}}}( \Val )$,
 the values in the range of $\Val$ remain $\alpha$-disequivalent.
-For example, we may assign values such as
-$\MU {\Varec{x}}.\; \const{E}( {\Varec{x}} )$,
-$\MU {\Varec{x}}.\; \const{E}( \MU x_1.\; \const{C}( x_1) )$, or
+For example, we may assign a value such as
+$\MU {\Varec{x}}.\; \const{E}( {\Varec{x}} )$ or
 $\MU {\Varec{x}}.\; \const{D}( \const{C}( {\Varec{x}} ) )$
@@
 Legal substitutions for ${\Varec{x}}$ may cause the range of $\Val$ to contain abnormal terms.
-After applying the substitution for $\MU {\Varec{x}}.\; \const{D}( \const{C}( {\Varec{x}} ) )$, we have that
-$\Val \ec{w_1}$ is $\MU {\Varec{w_1}}.\; \const{C}( \MU {\Varec{x}}.\; \const{D}( \const{C}( {\Varec{x}} ) ) )$,
+For example, after assigning $\MU {\Varec{x}}.\; \const{D}( \const{C}( {\Varec{x}} ) )$ to $\Varec{x}$, we have that
+$\Val \ec{u}$ is $\MU {\Varec{u}}.\; \const{C}( \MU {\Varec{x}}.\; \const{D}( \const{C}( {\Varec{x}} ) ) )$,
 which has normal form
-$\MU {\Varec{w_1}}.\; \const{C}( \MU {\Varec{x}}.\; \const{D}( {\Varec{w_1}} ) )$.
+$\MU {\Varec{u}}.\; \const{C}( \MU {\Varec{x}}.\; \const{D}( {\Varec{u}} ) )$.
 \xend
@@
 Intuitively, this lemma states three properties of $\ValC$ that ensure a normal
-model $\M$ can be trivially constructed for $\Ec$. Property~\PART{1} states that the
+interpretation $\mathbf{J}$ can be constructed that satisfies $\Ec$. Property~\PART{1} states that the
 values in the range of $\ValC$ are $\alpha$-equivalent to a value in normal
-model. In particular, this means they are closed, normal, and acyclic when
+interpretation. In particular, this means they are closed, normal, and acyclic when
 required. Property~\PART{2} states that the interpretation of all subterms in the
@@
 $\alpha$-equivalence), which in the following ensures that a distinct value can
-be associated with each equivalence class in a corresponding model.
+be associated with each equivalence class.
 
@@
 %The latter also comes with no loss of generality since nested applications of selectors in $\Fc_0$ can be replaced by fresh variables while adding equalities to $\Fc$.
-We consider a normal model $\M$
+We consider a normal model $\M$ \rem{Change to interpretation, with proper symbol.}
 that interprets wrongly applied selectors based on equality information in $\Fc$
@@
 solver and not by the theory solver. All other rules add equalities to the
-internal state of the theory solver. The rules in phase~1 along with the
-optimization for collapsing selectors are performed
+internal state of the theory solver. The rules in phase~1 are performed
 eagerly---that is, for partial satisfying assignments $M$---while the rules in
@@
 and are equivalent for all subterms of type $\Nondata$.
-If after doing so, some $S_i$ contains two terms $u$ and $v$, then it deduces that \rn{Unique} applies to $u$ and $v$.
+If after doing so, some $S_i$ in the resulting partition contains two terms $u$ and $v$, then it deduces that \rn{Unique} applies to $u$ and $v$.
 
@@
 If $\const s^k_{\negvthinspace j}( t )$ is a term in $\tEc$ and $t$ contains the
-constructor $\const{C}_{\negvthinspace j}( \vec u )$, we immediately infer
-$\const s^k_{\negvthinspace j}( t ) \teq u_k$, whereas the calculus we presented
-would apply \rn{Split} and \rn{Inject} before inferring this equality. For the
-sake of reducing the number of unique constraints considered by the calculus, we
+constructor $\const{C}_{\negvthinspace j}( u_1, \ldots, u_n )$, we immediately add
+$\const s^k_{\negvthinspace j}( t ) \teq u_k$ to $\Ec$, whereas the presented calculus
+would apply \rn{Split} and \rn{Inject} before adding this equality. 
+\begin{rep}  % lower priority to mention this
+For the sake of reducing the number of unique constraints considered by the calculus, we
 compute a normal form for constraints as a preprocessing step. In particular, we
@@
 based on whether $j = \jPrime$.
+\end{rep}
 

### 2015-02-18 tex/body.tex
edits

@@
 
-\pagebreak
-
 %* although nothing prevents composing the decision procedure with theories
@@
 %\paragraph{Specifications.}
-In an SMT problem, the signature is normally given by specifying first the
+In an SMT problem, the signature is typically given by specifying first the
 uninterpreted types in any order, then the (co)datatypes with their constructors
@@
 \!(\keyw{co})\keyw{datatype}\;\,
-  \delta_1 & {}= \smash{\const{C}_{11\!}(\bigl[\const{s}_{11\!}^1{:}\bigr]\vthinspace \tau_{11\!}^1, \ldots, \bigl[\const{s}_{11\!}^{n_{11\!}}{:}\bigr]\vthinspace \tau_{11\!}^{n_{11\!}})} \mid \cdots \mid \smash{\const{C}_{1m_1\!}(\ldots)} \\
-   \smash{\vdots\,\,\,} \\[-1\jot]
+  \delta_1 & {}= \smash{\const{C}_{11\!}(\bigl[\const{s}_{11\!}^1{:}\bigr]\vthinspace \tau_{11\!}^1, \ldots, \bigl[\const{s}_{11\!}^{n_{11\!}}{:}\bigr]\vthinspace \tau_{11\!}^{n_{11\!}})} \mid \cdots \mid \smash{\const{C}_{1m_1\!}(\ldots)} \\[-.5\jot]
+   \smash{\vdots\,\,\,} \\[-1.5\jot]
   \keyw{and}\; \,\delta_{\elllx} & =\, \smash{\const{C}_{\elll 1\!}(\bigl[\const{s}_{\elll 1\!}^1{:}\bigr]\vthinspace \tau_{\elll 1\!}^1, \ldots, \bigl[\const{s}_{\elll 1\!}^{n_{\elll 1\!}}{:}\bigr]\vthinspace \tau_{\elll 1\!}^{n_{\elll 1\!}})} \mid \cdots \mid \smash{\const{C}_{\elll m_{\elllx}\!}(\ldots)}
@@
 types is \emph{mutually} (\emph{co})\emph{recursive} if and only if the
-associated dependency graph is strongly connected. Types can be declared
-together in a mutual fashion even if they are not actually (co)recursive.
+associated dependency graph is strongly connected.
 \begin{rep}%
-The semantic notion is more precise and is the one that interests us.
-\end{rep}
+Types can be declared together in a mutual fashion even if they are not actually
+(co)recursive. The semantic notion is more precise and is the one that interests
+us.
+\end{rep}%
 %
@@
 
-Datatypes and codatatypes share many basic properties pertaining to
-constructors and selectors. All properties below are implicitly universally
-quantified and range over all $i$, $j$, $\jPrime$, and $k$ within bounds:
+Datatypes and codatatypes share many basic properties\begin{rep}. All properties
+below are implicitly universally quantified and range over all $i$, $j$,
+$\jPrime$, and $k$ within bounds\end{rep}:
 % and to all possible splits of the
@@
     \smash{\const{C}_{i\negvthinspace j}(\bar x) \tneq \const{C}_{i\negvthinspace \jPrime}(\bar y) \quad\text{if $j \not= \jPrime$}}
-  \\[-.5\jot]
+  \\[-1\jot]
 \text{Injectivity:}\quad
@@
     \smash{\const{C}_{i\negvthinspace j}(x_1,\ldots,x_{n_{i\negvthinspace j}}) \teq \const{C}_{i\negvthinspace j}(x_1,\ldots,x_{k-1},y,x_{k+1},\ldots,x_{n_{i\negvthinspace j}}) \longrightarrow x_k \teq y}
-  \\[-.5\jot]
+  \\[-1\jot]
 \text{Exhaustiveness:}\quad
   & \smash{\const{d}_{i1}(x) \mathrel\lor \cdots \mathrel\lor \const{d}_{im_i}(x)}
-  \\[-.5\jot]
+  \\[-1\jot]
 \text{Selection:}\quad
@@
 \end{rep}
-
+%
 %  * how to deal with "wrong" selectors, e.g.
@@
 %        and not in other models
-
+%
+\noindent\kern\parindent
 Datatypes are additionally characterized by an induction axiom schema\begin{rep}:
@@
 $\const{S}(\const{S}(\ldots))$.
-%
-\begin{rep}\par\end{rep}
-%
+
+\pagebreak
+
 For codatatypes, the dual notion is called coinduction. \begin{rep}It depends on
@@
 The derivation rules operate on a finite set $\Ec$ of equalities and
-disequalities between $\Sig$-terms, where $\Ec$ is taken as a conjunction. We
+disequalities between $\Sig$-terms, taken as a conjunction. We
 write $\tEc$ to denote the set of $\Sig$-terms occurring in $\Ec$.
@@
 Isabelle formalizations using Sledgehammer \cite{paulson-blanchette-2010}.
-Codatatypes being a recent addition to Isabelle
-\cite{blanchette-et-al-2014-impl}, benchmarks are not abundant. We
+Codatatypes being a recent addition to Isabelle\begin{rep}
+\cite{blanchette-et-al-2014-impl}\end{rep}, benchmarks are not abundant. We
 included all the theory files from the Isabelle distribution (Distro) and the \emph{Archive
@@
 256~lemmas as axioms, which were then monomorphized and translated to SMT-LIB
-along with the goal \cite{boehme-2012-phd}. The resulting problem was given to
-CVC4, running for up to 60 seconds on the StarExec cluster
+along with the goal \cite{blanchette-et-al-2013-smt}. The resulting problem was
+given to CVC4, running for up to 60 seconds on the StarExec cluster
 \cite{stump-et-al-2014-starexec}. Problems involving no (co)datatypes were
@@
 three correspond to one-line proofs in Isabelle, of the form
-\keyw{by}~\textit{coinduction}~\textit{auto} \cite{blanchette-et-al-2014-impl}.
+\keyw{by}~\textit{coinduction}~\textit{auto}\begin{rep} \cite{blanchette-et-al-2014-impl}\end{rep}.
 The fourth proof, by Gammie and Lochbihler, is more elaborate:

### 2015-02-18 tex/body.tex
updated numbers

@@
 No (co)datatypes
-  & 209 & 221 & 777 & 785 & \phantom{0}51 & \phantom{0}52 & 1037 & 1048 \\%[\jot]
+  & 209 & 221 & 777 & 775 & \phantom{0}51 & \phantom{0}52 & 1037 & 1048 \\%[\jot]
 %Datatype simplication only
@@
 Datatypes w/o \rn{Acyclic}
-  & & 235 & & 770 & & \phantom{0}61 & & 1066 \\
+  & & 227 & & 780 & & \phantom{0}52 & & 1059 \\
 Full datatypes
@@
 Codatatypes w/o \rn{Unique}
-  & & 230 & & 794 & & \phantom{0}65 & & 1089 \\
+  & & 222 & & 804 & & \phantom{0}56 & & 1082 \\
 Full codatatypes
-  & & 231 & & 794 & & \bfseries \phantom{0}68 & & 1093 \\%[\jot]
+  & & 223 & & 804 & & \bfseries \phantom{0}59 & & 1086 \\%[\jot]
 Full (co)datatypes
-  & & \bfseries 237 & & \bfseries 802 & & \bfseries \phantom{0}68 & & \bfseries 1107% \\[\jot]
+  & & \bfseries 229 & & \bfseries 815 & & \bfseries \phantom{0}59 & & \bfseries 1103% \\[\jot]
 \end{tabular}\end{center}

### 2015-02-18 tex/body.tex
up
Add text about quantified formulas, minor.

@@
  %\tpath{\Ec}{\ec{t}} = \tpath{\Ec}{\ec{u}}\sigma \neq \emptyset
- \Val \ec{t\typ{\delta}} \aequiv \Val \ec{u}
+ \Val \ec{t\typ{\delta}} \aequiv \Val \ec{u\typ{\delta}}
 }{
@@
 For \rn{Inject}, this is a consequence of injectivity.
-For \rn{Unique}, our construction of $\Val$ indicates that the value of $t$ and $u$ are equivalent in all models of $\Ec$.
+For \rn{Unique}, our construction of $\Val$ indicates that the values of $t$ and $u$ are equivalent in all models of $\Ec$.
 For \rn{Single}, $t$~and~$u$ must have the same value since the cardinality of their type is $1$.
@@
 For each selector term $\const s^k_{\negvthinspace j}( t )$ of type $\tau$, % for some $\vec u$,
-$\M\negvthinspace\bigl( \const s^k_{\negvthinspace j}( t ) \bigr)$ is the value
+if $\M( t )$ is of the form $\MU x.\; \const{C}_{\negvthinspace j}( \vec u )$, 
+then $\M\negvthinspace\bigl( \const s^k_{\negvthinspace j}( t ) \bigr)$ is the value
 in $\M( \tau )$ that is $\alpha$-equivalent to
-$\interp{\M( t )}{u_k}$,
-assuming $\M( t )$ is of the form $\MU x.\; \const{C}_{\negvthinspace j}( \vec u )$.
+$\interp{\M( t )}{u_k}$.
 \end{enumerate}
@@
 the value $\bot$. Then, at a high level, by the Nelson--Oppen combination
-framework (its variant used in CVC4 is described in~\cite{jovanovic2011sharing}), 
+framework (its variant in CVC4 is described in~\cite{jovanovic2011sharing}), 
 each $T_i$-solver for $i = 1, \ldots, l$ takes as input a combination
@@
 either reports that a subset $C$ of $M_i$ is $T_i$-unsatisfiable, in which case
-$\neg\; C$ is added to $F$, adds a clause $C$ to $F$, or does nothing.
+$\neg\; C$ is added to $F$, adds a clause to $F$, or does nothing.
 When $M$ is a complete assignment for $F$, a theory solver can only choose to do
@@
 
-Assume $\Ec$ is initially the set $M_i$ described above. For each equality added
+Assume $\Ec$ is initially the set $M_i$ described above. For each equality $t \teq s$ added
 to $\Ec$, we associate a set of equalities from $M_i$ that together entail $t
@@
 It simultaneously visits the equivalence classes of arguments of constructor terms in each equivalence class in $S$,
-while partitioning $S$ into $S_1, \ldots, S_n$,
-where all terms in $S_i$ have the same constructor type for all their subterms of type $\Data \cup \Codata$,
-and are equivalent for all subterms of type $\Nondata$.
-If after doing so, some $S_i$ in the resulting partition contains two terms $u$ and $v$, then it deduces that \rn{Unique} applies to $u$ and $v$.
+while partitioning $S$ into $S_1, \ldots, S_n$ based on the top-most symbol of constructor terms in these equivalence classes
+and the equivalence of their arguments of type $\Nondata$.
+It then partitions each set in this partition in the same manner recursively.
+If after doing so, the resulting partition contains a set $S_i$ containing distinct terms $u$ and $v$, then it deduces that \rn{Unique} applies to $u$ and $v$.
 
@@
 \paragraph{Extension to Quantified Formulas.}
-Two paras. One general one that talks about:
-\begin{verbatim}
-  * universal conjecture is no problem:
-    * falls into the ground fragment via negation
-      and skolemization
-  * nor is existential axiom a problem
-  * what's interesting: universal axioms and existential conjectures
-\end{verbatim}
-One more specific that talks about the cardinality calculus.
+While our calculus is a decision procedure for ground constraints in the theory of datatypes and codatatypes,
+in practice we are often interested in problems that combine reasoning about universal axioms and existential conjectures,
+where checking satisfiability requires reasoning about such ground constraints in the presence of universally quantified formulas.
+Many SMT solvers, such as CVC4, support reasoning about quantified formulas using incomplete instantiation-based methods~\cite{MouraBjoerner07,ReynoldsTinelliMoura14}.
+These methods extend naturally to quantified formulas over the signature of the theory of datatypes and codatatypes.
+
+However, the presence of quantified formulas poses an additional challenge in the context of (co)datatypes.
+Namely, since quantified formulas may entail cardinality constraints on uninterpreted types $U$,
+we may no longer assume that $U$ is of infinite cardinality.
+When assuming that a finite type has infinite cardinality, the calculus presented in this paper is \emph{incomplete},
+since the preconditions on \rn{Split} and \rn{Single} may not apply when they indeed are applicable.
+This does not necessarily impact the correctness of our procedure in this setting,
+since by assumption the solver is already incomplete when quantified formulas are present.
+Regardless, we may be interested in increasing the precision of the solver by doing the following.
+First, we may apply \rn{Split} to terms of datatype types whose cardinality depends on the finiteness of uninterpreted types.
+Second, we may conditionally apply \rn{Single} to terms of codatatype types that may have cardinality one.
+For instance, consider the codatatype $\keyw{codatatype}~\ty{c} = \const{C}(\ty{U},\: \ty{c})$, where $U$ is an uninterpreted type.
+This codatatype has cardinality one when $\ty{U}$ has cardinality one.
+If there exists two equivalence classes $\ec{t\typ{c}}$ and $\ec{u\typ{c}}$,
+our implementation adds a clause of the form $( \exists x\,y.\, x \not\teq y \vee t \teq u )$ to $F$,
+where $x$ and $y$ are variables of type $U$.
+In other words, either the cardinality of $U$ is greater than one, or $t$ must be equal to $u$.
 
@@
 makes a difference across all theories. Overall, it accounts for a success rate
-increase of \OK{4.9\%}. Moreover, every aspect of the procedure, including the
+increase of \OK{5.2\%}. Moreover, every aspect of the procedure, including the
 more expensive rules, make a contribution.
@@
 and in particular Morgan Deters,
-who jointly with the first author developed the theory solver for datatypes in
-CVC4.
+who jointly with the first author developed the initial version 
+of the theory solver for datatypes in CVC4.
 %

### 2015-02-18 tex/body.tex
more

--- 2015-02-18 tex/body.tex
@@
 \noindent
-of infinite sequences, or streams, over an element type $\tau$ is allowed; the
-corresponding datatype would be rejected as non-well-founded%
-\begin{rep} \cite{blanchette-et-al-2015-esop}\end{rep}.
+of infinite sequences, or streams, over an element type $\tau$ is allowed\begin{rep}; the
+corresponding datatype would be rejected as non-well-founded
+\cite{blanchette-et-al-2015-esop}\end{rep}.
 
@@
 of rewriting and a theory solver (Section~\ref{sec:implementation-as-a-theory-solver-in-cvc4}).
-It consists of about 2000 lines of \cpp{} code, among which \NOK{1600} are
-shared between datatypes and codatatypes. The code is integrated in the
-development version of the solver and is expected to be part of the CVC4~1.5 release.
+It consists of about 2000 lines of \cpp{} code, %among which \NOK{1600} are
+most of which are shared between datatypes and codatatypes. The code is
+integrated in the development version of the solver and is expected to be part
+of the CVC4~1.5 release.
 %
@@
 nonempty domains. The set~$\Types$ must contain a
-distinguished type \ty{bool}\begin{rep} interpreted as the set of truth
-values $\{\bot, \top\}$\end{rep}. %, and may contain other interpreted types (e.g., \ty{int}, \ty{real}).
-The only predicate is equality ($\teq$)\begin{rep} and belongs to the logical symbols\end{rep}.
-Other predicates can be represented as functions to $\ty{bool}$,
-with $\const{p}(\ldots)$ abbreviating $\const{p}(\ldots) \teq \const{true}$.
+distinguished type \ty{bool} interpreted as the set of truth
+values. %, and may contain other interpreted types (e.g., \ty{int}, \ty{real}).
 \begin{rep}
@@
 whereas $\tau,\:\upsilon$ range over arbitrary types.
-When applied to terms, the symbol $=$ denotes syntactic equality.\end{rep}
-
-\begin{rep}
+When applied to terms, the symbol $=$ denotes syntactic equality.
+
 Function symbols are written in a sans-serif font (e.g., $\const{f}$, $\const{g}$) to
@@
 \end{rep}
-Names starting with an uppercase letter (e.g.,
-$\const{S}$) are reserved for constructors. With each function symbol \const{f}
+Names starting with an uppercase letter \begin{rep}(e.g.,
+$\const{S}$) \end{rep}are reserved for constructors. With each function symbol \const{f}
 is associated a list of argument types $\tau_1,\ldots,\tau_n$ (for $n \ge 0$)
@@
 \const{f} to $n$ well-typed arguments $t_1 :\nobreak \tau_1$, \ldots, $t_n :
-\tau_n$. Nullary function symbols are called constants and can appear without
+\tau_n$. Nullary function symbols, called constants, can appear without
 parentheses in terms.
-The operator $\bigwedge_{\,i}\, \varphi_i$ abbeviates a conjunction
+\end{rep}
+The set $\Funcs$ must at least contain
+$\const{true},\, \const{false} : \ty{bool}$, interpreted as truth values.
+The only predicate is equality ($\teq$)\begin{rep} and belongs to the logical symbols\end{rep}.
+Other predicates can be represented as functions to $\ty{bool}$,
+with $\const{p}(\ldots)$ abbreviating $\const{p}(\ldots) \teq \const{true}$.
+\begin{rep}The operator $\bigwedge_{\,i}\, \varphi_i$ abbeviates a conjunction
 $\varphi_1 \mathrel\land \cdots \mathrel\land \varphi_n$. \end{rep}%
-Finally, $[n]$ denotes the set $\{1,\ldots, n\}$, and
+Finally, $[n] = \{1,\ldots, n\}$ and
 $\bar x$ abbreviates a list or tuple $x_1,\ldots,x_n$.
@@
 
+\newcommand\nummodden{num\hbox{\rm\_}mod\hbox{\rm\_}den}
+
 The results are summarized in Table~\ref{tab:bench}. The decision procedure
 makes a difference across all theories. Overall, it accounts for a success rate
-increase of \OK{4.9\%}. Moreover, every aspect of the procedure, including the
+increase of \OK{5.2\%}. Moreover, every aspect of the procedure, including the
 more expensive rules, make a contribution.
@@
 \begin{quote}
-\keyw{lemma} \,\textit{X0\_unique}: \,$x = \const{Node}\;0\;\const{num}\; x \Longrightarrow x = \const{X0}$ \\
+\keyw{lemma} \,\textit{\nummodden\_unique}: \,$x = \const{Node}\;0\;\const{num}\; x \Longrightarrow x = \const{\nummodden}$ \\
 \keyw{proof} \,(\textit{coinduction arbitrary}: $x$ \textit{rule}: \textit{tree.coinduct\_strong}) \\
@@
 \noindent
-where \const{X0} is defined as $\const{X0} =
-\const{Node}\;0\;\const{num}\;\const{X0}$.
+where \const{\nummodden} is defined as $\const{\nummodden} =
+\const{Node}\;0\;\const{num}\;\const{\nummodden}$.
 

### 2015-02-18 tex/body.tex
ack

--- 2015-02-18 tex/body.tex
@@
 
+\newcommand\nummodden{num\hbox{\rm\_}mod\hbox{\rm\_}den}
+
 The results are summarized in Table~\ref{tab:bench}. The decision procedure
@@
 %
+Andreas Lochbihler suggested several textual improvements.
 The second author's research was partially supported by the Deutsche

### 2015-02-18 tex/body.tex
more comments by Andreas

@@
 $\varphi_1 \mathrel\land \cdots \mathrel\land \varphi_n$. \end{rep}%
-Finally, $[n] = \{1,\ldots, n\}$ and
-$\bar x$ abbreviates a list or tuple $x_1,\ldots,x_n$.
+Finally, $\bar x$ abbreviates a list or tuple $x_1,\ldots,x_n$.
 
@@
 
-Each datatype $\delta$ is equipped with
+Each (co)datatype $\delta$ is equipped with
 $m \ge 1$ constructors, and each constructor for $\delta$ takes zero or more
@@
 namely the ones that satisfy the %axioms --- existence of all infinite codatatype values cannot be axiomatized
-properties mentioned in this section. In particular, the interpretations
+properties mentioned in this section, including (co)induction. In particular, the interpretations
 in~$\mathbf{J}$ share the same fixed interpretation for constructor terms and correctly
@@
 $\FV( \Val )$ denotes the union of the free variables occurring in $\Val$'s range,
-and $\Val[x \mapsto t]$ denotes the substitution of $t$ for $x$ in $\Val$'s range.
-It is easy to see that the height of terms produced as a result of the unfolding
-is bounded by the number of equivalence classes of $\Ec$,
-and thus the construction of $\Val$ will terminate.
+and $\Val[x \mapsto t]$ denotes the variable-capturing substitution of $t$ for
+$x$ in $\Val$'s range. It is easy to see that the height of terms produced as a
+result of the unfolding is bounded by the number of equivalence classes of
+$\Ec$, and thus the construction of $\Val$ will terminate.
 
@@
 \rem{TODO: add small examples}
+\rem{TODO: explain why terminates}
 
@@
 $t$ is of the form $\MU x.\; \const{C}( t_1, \ldots, t_n )$,
-$u$ it is of the form $\MU y.\; \const{C}( u_1, \ldots, u_n )$,
-and $\interp{t}{ t_k } \aequiv \interp{t}{ u_k }$ for all $k \in [n]$.
+$u$ is of the form $\MU y.\; \const{C}( u_1, \ldots, u_n )$,
+and $\interp{t}{ t_k } \aequiv \interp{t}{ u_k }$ for all $k$. % \in [n]
 The term $t$ is \emph{normal} if it does not contain a self-similar subterm
@@
 Due to our construction of $\Val$,
-we know $\ec{x}$ contains a constructor $\const{C}( z_1, \ldots, z_n )$ and $\muvar( t_i ) = z_i$ for $i \in [n]$.
+we know $\ec{x}$ contains a constructor $\const{C}( z_1, \ldots, z_n )$ and $\muvar( t_i ) = z_i$ for some $i$ % \in [n]$.
 Since $\Val \ec{x}$ is a minimal, it
 contains a subterm of the form $\MU {\Varec{y}}.\; \const{C}( u_1, \ldots, u_n )$
-where $\interp{\Val \ec{x}}{ t_i } \aequiv \interp{\Val \ec{x} }{ u_i }$ for $i \in [n]$.
-Due to our construction of $\Val$, $\ec{y}$ contains a constructor $\const{C}( w_1, \ldots, w_n )$ and $\muvar( u_i ) = w_i$ for $i \in [n]$.
+where $\interp{\Val \ec{x}}{ t_i } \aequiv \interp{\Val \ec{x} }{ u_i }$ for some $i$. % \in [n]$.
+Due to our construction of $\Val$, $\ec{y}$ contains a constructor $\const{C}( w_1, \ldots, w_n )$ and $\muvar( u_i ) = w_i$ for some $i$. % \in [n]$.
 Since \rn{Cong} does not apply to $\Ec$
@@
 of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$
-where for each $i \in [n]$,
+where for each $i \in \{1,\ldots, n\}$, % \in [n]$,
 $\muvar(w_i) = {\Varec{u_i}}$.  Thus by Lemma~\ref{lem:model-completion}\PART{2},
 $\interp{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$.
-For each $i \in [n]$, let $u_\iPrime$ be the $i$\vvthinspace th argument of $t_u$.
+For each $i \in \{1,\ldots, n\}$, let $u_\iPrime$ be the $i$\vvthinspace th argument of $t_u$.
 By Lemma~\ref{lem:mu-norm-arg}, $\interp{t_u}{u_\iPrime} \vsim \ValC \ec{u_i}$.
@@
 $\MU {\Varec{u}}.\; \const C_{\negvthinspace\jPrime}( w_1, \ldots, w_n )$,
-where for each $i \in [n]$,
+where for each $i \in \{1,\ldots, n\}$,
 $\muvar(w_i) = \smash{\Varec{\vphantom{I}\smash{\const s^i_{\negvthinspace\jPrime}( u )}}}$ and thus by Lemma~\ref{lem:model-completion}\PART{2},
@@
 constraints for its theory.
-Given a theory $T = T_1 \mathrel\cup \cdots \mathrel\cup T_n$ and a input set of
+Given a theory $T = T_1 \mathrel\cup \cdots \mathrel\cup T_n$ and an input set of
 clauses $F$ in CNF, the \XXXL($T$) procedure
@@
 conclusion $\bot$, we report the union of the explanations for all premises is
-$\thD$-unsatisfiable. For the \rn{Split} rule, we add a clause of the form
-$\const{C}_1\bigl( \const s^1_1( t ), \ldots, \const s^{n_1}_1( t ) \bigr)
+$\thD$-unsatisfiable. For \rn{Split}, we add a clause %of the form
+$t \teq \const{C}_1\bigl( \const s^1_1( t ), \ldots, \const s^{n_1}_1( t ) \bigr)
 \mathrel\lor \cdots \mathrel\lor t \teq \const{C}_m\bigl( \const s^1_m( t ),
@@
 codatatypes either enabled or disabled. The contributions of the acyclicity and
-uniqueness rules were also measured, so find out whether these somewhat
+uniqueness rules were also measured, to find out whether these somewhat
 expensive rules are useful in practice. Even when the decision procedure is

### 2015-02-18 tex/body.tex
more

@@
 \newcommand{\ec}[1]{[#1]}
+\newcommand{\J}{\mathcal{J}}
 \newcommand{\M}{\mathcal{M}}
@@
 \newcommand\negvthinspace{\kern-0.083333em}
+\newcommand\negvvthinspace{\kern-0.041667em}
 
@@
 problems involving datatypes and codatatypes in combination
-(Section~\ref{sec:a-ground-decision-procedure-for-co-datatypes}).
+(Section~\ref{sec:the-ground-decision-procedure}).
 The procedure is described abstractly as a calculus and is
@@
 constructor is left completely unspecified.
-\end{rep}
+\end{rep}%
 %
@@
 %
-\noindent\kern\parindent
 Datatypes are additionally characterized by an induction axiom schema\begin{rep}:
@@
 
-Given a signature $\Sig$, we write $\thD$ to refer to the \emph{theory of
+Given a signature $\Sig$, $\thD$ refers to the \emph{theory of
 datatypes and codatatypes},
@@
 % it comes to English prefixes like non-, in-, un-, and a-.)
-\rem{more?}
+%\rem{more?}
 
@@
 A type may also be a corecursive singleton only in some models. If the example
-above is altered to leave \ty{unit} uninterpreted, \ty{b} and \ty{c} will be
+above is altered to make \ty{unit} an ordinary type, \ty{b} and \ty{c} will be
 singletons precisely when \ty{unit} is interpreted as a singleton.
@@
 \section{The Ground Decision Procedure} % for (Co)datatypes}
-\label{sec:a-ground-decision-procedure-for-co-datatypes}
+\label{sec:sec:the-ground-decision-procedure}
 
@@
 The derivation rules operate on a finite set $\Ec$ of equalities and
-disequalities between $\Sig$-terms, taken as a conjunction. We
-write $\tEc$ to denote the set of $\Sig$-terms occurring in $\Ec$.
+disequalities between $\Sig$-terms. %, taken as a conjunction.
+We let $\tEc$ denote the set of $\Sig$-terms occurring in $\Ec$.
 
@@
 
-The calculus consists of three sets of rules, presented in
+The calculus consists of three sets of rules, given in
 \figuresname~\ref{fig:cc-rules} to \ref{fig:split-rule}, corresponding to three
 phases. The first phase computes the bidirectional closure of $\Ec$. The second
-phase makes inferences based on acyclicity (for datatypes) and uniqueness (for
-codatatypes). The third phase performs case splitting on constructors for
+phase
+makes \hfill inferences \hfill based \hfill on \hfill acyclicity \hfill (for
+\hfill datatypes) \hfill and \hfill uniqueness \hfill (for \hfill codatatypes).
+
+%%% TYPESETTING: horrible hack
+\pagebreak[2]
+\noindent
+The third phase performs case splitting on constructors for
 various terms in $\tEc$.
@@
 
-\rem{TODO: Normal model vs. actual model.}
-\rem{TODO: Model or interpretation? Grep elsewhere.}
-
-\begin{definition}[Normal Model]
+\rem{TODO: Normal interpretation vs. actual interpretation.}
+
+\begin{definition}[Normal Interpretation]
 \afterDot%
-\label{def:norm-model}%
+\label{def:norm-interpretation}%
 \rm
-A model $\M$ is \emph{normal} if these conditions are met:
+An interpretation $\J$ is \emph{normal} if these conditions are met:
 \begin{enumerate}
@@
 For each type $\tau$,
-$\M( \tau )$ is a maximal set of closed normal $\mu$-terms of that type that are
+$\J( \tau )$ is a maximal set of closed normal $\mu$-terms of that type that are
 unique up to $\alpha$-equivalence and acyclic if $\tau \not\in \Codata$.
@@
 For each constructor term $\const{C}( \vec t \,)$ of type $\tau$,
-$\M( \const{C}( \vec t \,) )$ is the value
-in $\M( \tau )$ that is $\alpha$-equivalent to
-$\nf{\MU x.\; \const{C}( \M( \vec t \,) )}$, where $x$ is fresh.
+$\J( \const{C}( \vec t \,) )$ is the value
+in $\J( \tau )$ that is $\alpha$-equivalent to
+$\nf{\MU x.\; \const{C}( \J( \vec t \,) )}$, where $x$ is fresh.
 \item
 For each selector term $\const s^k_{\negvthinspace j}( t )$ of type $\tau$, % for some $\vec u$,
-if $\M( t )$ is of the form $\MU x.\; \const{C}_{\negvthinspace j}( \vec u )$, 
-then $\M\negvthinspace\bigl( \const s^k_{\negvthinspace j}( t ) \bigr)$ is the value
-in $\M( \tau )$ that is $\alpha$-equivalent to
-$\interp{\M( t )}{u_k}$.
+if $\J( t )$ is of the form $\MU x.\; \const{C}_{\negvthinspace j}( \vec u )$, 
+then $\J\negvvthinspace\bigl( \const s^k_{\negvthinspace j}( t ) \bigr)$ is the value
+in $\J( \tau )$ that is $\alpha$-equivalent to
+$\interp{\J( t )}{u_k}$.
 \end{enumerate}
@@
 
-In the following, we consider only normal models.
+In the following, we consider only normal interpretations.
 When constructing a model $\M$ for $\Ec$,
@@
 $\MU x.\; \const{C}( x, x ) \not\vsimv{y} \MU y.\; \const{C}( x, y )$.
-For a variable $x\typ{\tau}$ and a fixed normal model $\M$,
-we write $\mathcal{T}^{\smash{\M}}_x( \Val )$ to denote the set consisting of all values
-$v \in \M( \tau )$
+For a variable $x\typ{\tau}$ and a fixed normal interpretation $\J$,
+we write $\mathcal{T}^{\smash{\J}}_x( \Val )$ to denote the set consisting of all values
+$v \in \J( \tau )$
 such that $v \vsimv{x} \interp{t}{u}$ for some term $t$ with subterm $u$ occurring in the range of $\Val$.
@@
 
-The \emph{completion} $\ValC$ of $\Val$ for a normal model $\M$
-assigns values from $\M$ to unassigned variables in the domain of $\Val$.
+The \emph{completion} $\ValC$ of $\Val$ for a normal interpretation $\J$
+assigns values from $\J$ to unassigned variables in the domain of $\Val$.
 We construct $\ValC$ by exhaustively applying the following rule to $\nf{\Val}$:%
@@
   \betweenantes
-  v \in \M( \tau )
+  v \in \J( \tau )
   \betweenantes
-  v \not\in \mathcal{T}^{\smash{\M}}_{\Varec{x}}(\Val)
+  v \not\in \mathcal{T}^{\smash{\J}}_{\Varec{x}}(\Val)
 }{
@@
 The rule chooses an unassigned variable in $\Val$ and assigns it a fresh
-value---one that does not occur in $\mathcal{T}^{\smash{\M}}_{\Varec{x}}(\Val)$
+value---one that does not occur in $\mathcal{T}^{\smash{\J}}_{\Varec{x}}(\Val)$
 modulo $\alpha$-equivalence.
@@
 of codatatypes, of which only finitely many are in
-$\mathcal{T}^{\smash{\M}}_{\Varec{x}}(\Val)$?)}
+$\mathcal{T}^{\smash{\J}}_{\Varec{x}}(\Val)$?)}
 Since this update removes one
@@
 variables of $\Val$, namely ${\Varec{x}}$ and ${\Varec{y}}$.
-Modulo $\alpha$-equivalence, the set \smash{$\mathcal{T}^{\smash{\M}}_{{\Varec{x}}}( \Val )$} is
+Modulo $\alpha$-equivalence, the set \smash{$\mathcal{T}^{\smash{\J}}_{{\Varec{x}}}( \Val )$} is
 \[%\begin{array}{c}
@@
 \]
-Consider a normal interpretation $\mathbf{J}$ \rem{Appropriate symbol?} that interprets variables in $\Ec$ based on the mapping $\Val$:
+Consider a normal interpretation $\J$ that interprets variables in $\Ec$ based on the mapping $\Val$:
 $x$ is interpreted as $\Val \ec{x}$, $y$ as $\Val \ec{y}$, and so on.
-Assigning a value for $\Val \ec{x}$ that is $\alpha$-equivalent to a value in $\mathcal{T}^{\smash{\M}}_{{\Varec{x}}}( \Val )$
+Assigning a value for $\Val \ec{x}$ that is $\alpha$-equivalent to a value in $\mathcal{T}^{\smash{\J}}_{{\Varec{x}}}( \Val )$
 may cause values in the range of $\Val$ to become $\alpha$-equivalent,
-which in turn may cause $\Ec$ to be falsified by $\mathbf{J}$.
+which in turn may cause $\Ec$ to be falsified by $\J$.
 For example, assign $\MU {\Varec{x}}.\; \const{D}( {\Varec{x}} )$ for ${\Varec{x}}$.
@@
 \begin{lemma}
-\label{lem:model-completion}
+\label{lem:interpretation-completion}
 If $\Val$ is constructed for a saturated set $\Ec$
-and $\ValC$ is a completion of $\Val$ for a normal model $\M$, the following properties hold:
+and $\ValC$ is a completion of $\Val$ for a normal interpretation $\J$, the following properties hold:
 \begin{enumerate}
-\item[\rm (1)] $\ValC \ec{x\typ{\tau}}$ is $\alpha$-equivalent to a value in $\M( \tau )$.
+\item[\rm (1)] $\ValC \ec{x\typ{\tau}}$ is $\alpha$-equivalent to a value in $\J( \tau )$.
 \item[\rm (2)]
@@
 there is only one equivalence class of type $\tau$ in $\Ec$,
-and thus there are no terms in $\mathcal{T}^{\smash{\M}}_{\Varec{y}}(\ValC)$ of type $\tau$.
-This is a contradiction, since our model completion may assign the value in the domain of $\tau$ to ${\Varec{y}}$.
+and thus there are no terms in $\mathcal{T}^{\smash{\J}}_{\Varec{y}}(\ValC)$ of type $\tau$.
+This is a contradiction, since %our %model
+completion can assign the value in the domain of $\tau$ to ${\Varec{y}}$.
 Now, consider the case when $\tau$ is infinite.
 This is also a contradiction,
-since there are only a finite number of closed terms in $\mathcal{T}^{\smash{\M}}_{\Varec{y}}(\ValC)$,
-and thus our model completion can assign a value not occurring in $\mathcal{T}^{\smash{\M}}_{\Varec{y}}(\ValC)$ to ${\Varec{y}}$.
+since there are only a finite number of closed terms in $\mathcal{T}^{\smash{\J}}_{\Varec{y}}(\ValC)$,
+and thus %our %model
+completion can assign a value not occurring in $\mathcal{T}^{\smash{\J}}_{\Varec{y}}(\ValC)$ to ${\Varec{y}}$.
 By construction, $\ValC \ec{x}$ is normal.
@@
 Thus, $\ValC \ec{x}$ is acyclic when $\tau \not\in \Codata$.
-Therefore, by definition, $\ValC \ec{x}$ is $\alpha$-equivalent to a value in $\M( \tau )$.
+Therefore, by definition, $\ValC \ec{x}$ is $\alpha$-equivalent to a value in $\J( \tau )$.
 
@@
 ${\Varec{x}} \in FV( \Val_1 )$,
-and $\MU {\Varec{x}}.\; t$ is not $\alpha$-equivalent to a term in $\mathcal{T}^{\smash{\M}}_{\Varec{x}}(\Val_1)$.
+and $\MU {\Varec{x}}.\; t$ is not $\alpha$-equivalent to a term in $\mathcal{T}^{\smash{\J}}_{\Varec{x}}(\Val_1)$.
 To show 2,
@@
 Since $\interp{\Val_1 \ec{y}}{t_1}$ is normal,
-there is a closed $\mu$-term $v \in \mathcal{T}^{\smash{\M}}_{\Varec{x}}( \Val_1 )$
+there is a closed $\mu$-term $v \in \mathcal{T}^{\smash{\J}}_{\Varec{x}}( \Val_1 )$
 such that $v \vsimv{\Varec{x}} \interp{\Val_1 \ec{y}}{t_1}$.
@@
 Intuitively, this lemma states three properties of $\ValC$ that ensure a normal
-interpretation $\mathbf{J}$ can be constructed that satisfies $\Ec$. Property~\PART{1} states that the
+interpretation $\J$ can be constructed that satisfies $\Ec$. Property~\PART{1} states that the
 values in the range of $\ValC$ are $\alpha$-equivalent to a value in normal
@@
 %The latter also comes with no loss of generality since nested applications of selectors in $\Fc_0$ can be replaced by fresh variables while adding equalities to $\Fc$.
-We consider a normal model $\M$ \rem{Change to interpretation, with proper symbol.}
+We consider a normal model $\M$
 that interprets wrongly applied selectors based on equality information in $\Fc$
@@
 For the latter, let $\M( x )$ be the value in $\M$ that is $\alpha$-equivalent with $\ValC \ec{x}$ for each variable $x \in \mathcal{T}(\Fc)$,
-which by Lemma~\ref{lem:model-completion}\PART{1} is guaranteed to exist.
+which by Lemma~\ref{lem:interpretation-completion}\PART{1} is guaranteed to exist.
 
@@
 Call this term $t_u$.
-%By Lemma~\ref{lem:model-completion}\PART{2} and due to our construction of $\ValC$, we have $\ValC \ec{t}$ is a term
+%By Lemma~\ref{lem:interpretation-completion}\PART{2} and due to our construction of $\ValC$, we have $\ValC \ec{t}$ is a term
 %of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$
@@
 where for each $i \in \{1,\ldots, n\}$, % \in [n]$,
-$\muvar(w_i) = {\Varec{u_i}}$.  Thus by Lemma~\ref{lem:model-completion}\PART{2},
+$\muvar(w_i) = {\Varec{u_i}}$.  Thus by Lemma~\ref{lem:interpretation-completion}\PART{2},
 $\interp{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$.
@@
 $\ec{u}$ must contain a term of the form $\smash{\const C_{\negvthinspace\jPrime}\bigl( \const s^1_{\negvthinspace\jPrime}( u ), \ldots, \const s^{n}_{\negvthinspace\jPrime}( u ) \bigr)}$ for some $\jPrime$.
-%By Lemma~\ref{lem:model-completion}\PART{2} and due to our construction of $\ValC$, we have $\ValC \ec{u}$ is of the form
+%By Lemma~\ref{lem:interpretation-completion}\PART{2} and due to our construction of $\ValC$, we have $\ValC \ec{u}$ is of the form
 %$\MU {\Varec{u}}.\; \const C_{\negvthinspace\jPrime}( w_1, \ldots, w_n )$,
@@
 where for each $i \in \{1,\ldots, n\}$,
-$\muvar(w_i) = \smash{\Varec{\vphantom{I}\smash{\const s^i_{\negvthinspace\jPrime}( u )}}}$ and thus by Lemma~\ref{lem:model-completion}\PART{2},
+$\muvar(w_i) = \smash{\Varec{\vphantom{I}\smash{\const s^i_{\negvthinspace\jPrime}( u )}}}$ and thus by Lemma~\ref{lem:interpretation-completion}\PART{2},
 $\interp{\ValC \ec{u}}{ w_i } = \ValC \ec{ \smash{\const s^i_{\negvthinspace\jPrime}( u )}}$.
@@
 are in distinct equivalence classes only if $u$ and $u'$ are in distinct equivalence classes.
-Due to the induction hypothesis and Lemma~\ref{lem:model-completion}\PART{3}, $\M( u ) \neq \M( u' )$ for all such $u$, $u'$.
+Due to the induction hypothesis and Lemma~\ref{lem:interpretation-completion}\PART{3}, $\M( u ) \neq \M( u' )$ for all such $u$, $u'$.
 Thus, we may interpret $\smash{\M(\const s^k_{\negvthinspace j})( \M( u ) )}$ as the value in $\M( \tau )$ that is $\alpha$-equivalent with $\ValC \ec{ t }$.
@@
 Since $\M( t_1 ) \vsim \ValC \ec{t_1}$ and $\M( t_2 ) \vsim \ValC \ec{t_2}$,
-by Lemma~\ref{lem:model-completion}\PART{3}, $\M( t_1 ) \neq \M( t_2 )$, and thus $\M \models t_1 \tneq t_2$.
+by Lemma~\ref{lem:interpretation-completion}\PART{3}, $\M( t_1 ) \neq \M( t_2 )$, and thus $\M \models t_1 \tneq t_2$.
 

### 2015-02-18 tex/body.tex
more

@@
 %
-\begin{align*}
-\interp{t}{\MU y.\; \const{C}( x )}
-& = \MU y.\; \const{C}\bigl( \interp{t}{x}^{\smash{\{y\}}} \bigr) \\
-& = \MU y.\; \const{C}\bigl( \interp{t}{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) ) }^{\smash{\{y\}}} \bigr) \\
-& = \MU y.\; \const{C}\bigl( \MU x.\; \const{C}( \interp{t}{\MU y.\; \const{C}( x )}^{\smash{\{x,y\}}} ) \bigr) \\
-& = \MU y.\allowbreak\; \const{C}( \MU x.\allowbreak\; \const{C}( y ) )
-\end{align*}
+$\interp{t}{\MU y.\; \const{C}( x )}
+ = \MU y.\; \const{C}\bigl( \interp{t}{x}^{\smash{\{y\}}} \bigr)
+ = \MU y.\; \const{C}\bigl( \interp{t}{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) ) }^{\smash{\{y\}}} \bigr)
+ = \MU y.\; \const{C}\bigl( \MU x.\; \const{C}( \interp{t}{\MU y.\; \const{C}( x )}^{\smash{\{x,y\}}} ) \bigr)
+ = \MU y.\allowbreak\; \const{C}( \MU x.\allowbreak\; \const{C}( y ) )$
 is $\alpha$-equivalent to $\interp{t}{x} = t$.
@@
 it remains only to specify how $\M$ interprets wrongly applied selector terms and variables.%
-\rem{Variables are really part of the model then?}
 For the latter, this will be based on the mapping $\Val$ computed in phase~2 of the calculus.
@@
 modulo $\alpha$-equivalence.
-\rem{The ``not alpha equivalent'' bit is expressed differently in the rule,
-with $\notin$. This looks fishy.}
 \rem{What guarantees that such a value exists? (Answer: infinitely many values

### 2015-02-18 tex/body.tex
more

@@
 An evaluation on %hand-crafted examples and on
-problems generated from Isabelle theories using the Sledgehammer tool
+problems generated from Isabelle theories using the Sledgehammer tool \cite{paulson-blanchette-2010}
 demonstrates the usefulness of the approach (Section~\ref{sec:experimental-results}).
@@
 \section{The Ground Decision Procedure} % for (Co)datatypes}
-\label{sec:sec:the-ground-decision-procedure}
+\label{sec:the-ground-decision-procedure}
 
@@
 \end{tabular}
+\vspace*{-3pt} %% TYPESETTING HACK
 \caption{\,Derivation rules for bidirectional closure%.
@@
   %\ttpath{\Ec}{\ec{t}}{\ec{t}} \neq \emptyset
-  \Val \ec{t\typ{\delta}} = \MU x.\; u
+  \Val \smash{\ec{t\typ{\delta}}} = \MU x.\; u
   \betweenantes
@@
  %\tpath{\Ec}{\ec{t}} = \tpath{\Ec}{\ec{u}}\sigma \neq \emptyset
- \Val \ec{t\typ{\delta}} \aequiv \Val \ec{u\typ{\delta}}
+ \Val \smash{\ec{t\typ{\delta}}} \aequiv \Val \smash{\ec{u\typ{\delta}}}
 }{
@@
 \end{tabular}
+\vspace*{-3pt} %% TYPESETTING HACK
 \caption{\,Derivation rules for acyclicity and uniqueness%.
@@
 
-\paragraph{Integration into {\relax\XXXL($T$)}.}
+\paragraph{Integration into {\relax\XXXL($T\negvthinspace$)}.}
 The calculus is implemented as a \emph{theory solver} of CVC4, that is, a
@@
 Decisions on which branch to take are thus performed externally by the SAT
-solver and not by the theory solver. All other rules add equalities to the
+solver. % and not by the theory solver.
+All other rules add equalities to the
 internal state of the theory solver. The rules in phase~1 are performed
@@
 %\rem{TODO: * sharing is caring (let's cite it for a change)}
-\rem{TODO: ``theory combination more complicated (and Theorem Completeness too weak)''?}
+%\rem{TODO: ``theory combination more complicated (and Theorem Completeness too weak)''?}
 %\rem{TODO: representation of $\mu$-terms?}
@@
 To evaluate the decision procedure, we generated problems from existing
-Isabelle formalizations using Sledgehammer \cite{paulson-blanchette-2010}.
+Isabelle formalizations using Sledgehammer.
 Codatatypes being a recent addition to Isabelle\begin{rep}
 \cite{blanchette-et-al-2014-impl}\end{rep}, benchmarks are not abundant. We
-included all the theory files from the Isabelle distribution (Distro) and the \emph{Archive
-of Formal Proofs} (AFP) \cite{klein-et-al-afp} that define codatatypes falling
-within the supported fragment. We also included two unpublished theories by
-Peter Gammie and Andreas Lochbihler (\gandl), about Bird and Stern--Brocot trees.
-To also exercise the support
+included all the theory files from the Isabelle distribution (Distro, 879 goals)
+and the \emph{Archive of Formal Proofs} (AFP, 2974 goals) \cite{klein-et-al-afp}
+that define codatatypes falling within the supported fragment. We also included
+two unpublished theories by Peter Gammie and Andreas Lochbihler (\gandl, 317
+goals), about Bird and Stern--Brocot trees. To also exercise the support
 for datatypes, the benchmarks are complemented by theories about lists
 and tree data structures. The theories were selected before running the
-experiments. The experimental data is publicly available.%
-\footnote{\url{http://TODO/FIXME.}}
+experiments. The experimental data are publicly available.%
+\footnote{\url{http://www21.in.tum.de/~blanchet/codp_data.tgz}}
 
@@
 filtered out. Due to the lack of machinery for reconstructing inferences about
-(co)datatypes in Isabelle, CVC4 is trusted as an oracle in these experiments.
+(co)datatypes in Isabelle, the external solver is trusted as an oracle in these
+experiments.
 
@@
 constructors and selectors, which the decision procedure would recognize as
-tautologies. Z3 is included for comparison when this is possible.
+tautologies. Z3 is included for comparison. % when this is possible.
 %This allows us to answer the question, \relax{What are the

### 2015-02-18 tex/body.tex
shorten

@@
 Correctness means that if there exists a closed derivation tree with root node
-$\Ec$, we can conclude that $\Ec$ is $\thD$-unsatisfiable; and if there exists
-a derivation tree with root node $\Ec$ that contains a saturated node, we can
-conclude that $\Ec$ is $\thD$-satisfiable.
+$\Ec$, then $\Ec$ is $\thD$-unsatisfiable; and if there exists
+a derivation tree with root node $\Ec$ that contains a saturated node, then
+$\Ec$ is $\thD$-satisfiable.
 

### 2015-02-18 tex/body.tex
shorten
more

@@
 D$\negvthinspace, let
+\begin{conf}
+$\smash{\SSSS 0_t} = \{ t \}$ and
+$\smash{\SSSS {i+1}_t} = \smash{\SSSS i_t \mathrel\cup \{ \const s( u ) \mid u\typ{\delta} \in \SSSS i_t{,}}~\allowbreak\smash{ \delta \in \Data{,}~ \left|\delta\right| \text{ is finite}{,}~ \const{s} \in \Sel^{\delta}}  \}$,
+\end{conf}%
+\begin{rep}
 \begin{align*}
@@
 \end{align*}
+\end{rep}%
 %for $i \geq 0$, let ;
 and let $\SSSS \AST_t$ be the limit of this sequence.
-This is a finite set for each $t$ because the values of the type of $t$ are of finite size\rem{Clarify}.
+This is a finite set for each $t$ because there can only be finitely many
+values of the same type as $t$.
+%%% TODO: Clarify above step in report.
 Let $\SSSS \AST$ be the union of all sets $\SSSS \AST_t$ where $t \in D$,
@@
 %
-$$\Ec \mathrel\cup \{ \const C_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\smash{\,n_{\negvthinspace j}}}_{\negvthinspace j}( t ) \bigr) \mid t\typ{\delta} \in \SSS \mathrel\cup \SSSS \AST{,}~ \const C_{\negvthinspace j} \in \Ctr^\delta \}$$
+$\Ec \mathrel\cup \{ \const C_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\smash{\,n_{\negvthinspace j}}}_{\negvthinspace j}( t ) \bigr) \mid t\typ{\delta} \in \SSS \mathrel\cup \SSSS \AST{,}~ \const C_{\negvthinspace j} \in \Ctr^\delta \}$.
 %
@@
 constructor $\const{C}_{\negvthinspace j}( u_1, \ldots, u_n )$, we immediately add
-$\const s^k_{\negvthinspace j}( t ) \teq u_k$ to $\Ec$, whereas the presented calculus
+$\smash{\const s^k_{\negvthinspace j}( t ) \teq u_k}$ to $\Ec$, whereas the presented calculus
 would apply \rn{Split} and \rn{Inject} before adding this equality. 

### 2015-02-18 tex/body.tex
shorten

@@
 
-% \textbf{TODO:} Nitpick
+% TODO: Leon, RADA -- http://www-users.cs.umn.edu/~hung/papers/fse13.pdf
 
@@
 constructor $\const{C}_{\negvthinspace j}( u_1, \ldots, u_n )$, we immediately add
-$\smash{\const s^k_{\negvthinspace j}( t ) \teq u_k}$ to $\Ec$, whereas the presented calculus
-would apply \rn{Split} and \rn{Inject} before adding this equality. 
+$\smash{\const s^k_{\negvthinspace j}( t ) \teq u_k}$ to $\Ec$\begin{rep}, whereas the presented calculus
+would apply \rn{Split} and \rn{Inject} before adding this equality\end{rep}. 
 \begin{rep}  % lower priority to mention this
-For the sake of reducing the number of unique constraints considered by the calculus, we
+To reduce the number of unique constraints considered by the calculus, we
 compute a normal form for constraints as a preprocessing step. In particular, we
@@
 
-To evaluate the decision procedure, we generated problems from existing
+To evaluate the decision procedure, we generated benchmark problems from existing
 Isabelle formalizations using Sledgehammer.
@@
 the \rn{Acyclic} rule and four required the \rn{Unique} rule. Among the latter,
-three correspond to one-line proofs in Isabelle, of the form
+three are one-line proofs in Isabelle, of the form
 \keyw{by}~\textit{coinduction}~\textit{auto}\begin{rep} \cite{blanchette-et-al-2014-impl}\end{rep}.

### 2015-02-18 tex/body.tex
more

@@
 including Agda, Coq, Isabelle, Matita, and PVS. In recent years, datatypes
-have emerged in a few automatic theorem provers as well. The SMT-LIB~2
-\cite{barrett-et-al-2010} syntax, implemented by most SMT (satisfiability
-modulo theories) solvers, includes a theory of datatypes.
+have emerged in a few automatic theorem provers as well. The SMT-LIB
+\cite{barrett-et-al-2010} format, implemented by most SMT %(satisfiability
+%modulo theories)
+solvers, has been extended with a syntax for datatypes.
 
@@
 as long as the argument types coincide. % \cite[Section~3]{blanchette-et-al-2014-codata}.
-However, this is disallowed by SMT-LIB 2, so we do not consider it here.}
+However, this is disallowed by CVC4 and Z3, so we do not consider it here.}
 Schematically:
@@
 For each goal in each theory file, Sledgehammer was invoked to select
-256~lemmas as axioms, which were then monomorphized and translated to SMT-LIB
+256~lemmas as axioms, which were then monomorphized and translated to SMT-LIB~2
 along with the goal \cite{blanchette-et-al-2013-smt}. The resulting problem was

### 2015-02-18 tex/body.tex
more

@@
 
-The decision procedure for the ground theory of (co)datatypes $\thD$ determines the
-$\thD$-satisfiability of sets of ground constraints over a fixed
-signature~$\Sig$. It is formulated as a calculus.
+Given a fixed signature~$\Sig$,
+the decision procedure for the ground theory of (co)datatypes $\thD$ determines the
+$\thD$-satisfiability of finite sets $\Ec$ of constraints:\ equalities
+and disequalities between ground $\Sig$-terms. The procedure is formulated as a
+calculus.
 % whose derivation rules are applied until saturation or contradiction.
-%
-The derivation rules operate on a finite set $\Ec$ of equalities and
-disequalities between $\Sig$-terms. %, taken as a conjunction.
-We let $\tEc$ denote the set of $\Sig$-terms occurring in $\Ec$.
+
+To simplify the presentation, we make a few assumptions about
+$\Sig$.
+First, all codatatypes in $\Codata$ are corecursive. This is not restrictive
+because noncorecursive codatatypes can be seen as nonrecursive
+datatypes.
+Second, all types $\tau \in \Nondata$ have infinite cardinality.
+This is not overly restrictive, because each interpreted type in $\Nondata$
+having finite cardinality~$n$
+can be replaced by an enumeration datatype with $n$~nullary constructors.
+As for uninterpreted types, since the constraints in $\Ec$ are ground, they cannot entail an upper
+bound on the cardinality of any such type, so it is safe to consider them infinite.
 
@@
 An application of a rule is \emph{redundant} if at least one branch in the
-conclusion is an assignment to $\Ec$ that leaves it unchanged. A
-\emph{derivation tree} is a %finite
+conclusion is an assignment to $\Ec$ that leaves it unchanged.
+A \emph{derivation tree} is a %finite
 tree whose nodes are finite sets of
@@
 phases. The first phase computes the bidirectional closure of $\Ec$. The second
-phase
-makes \hfill inferences \hfill based \hfill on \hfill acyclicity \hfill (for
-\hfill datatypes) \hfill and \hfill uniqueness \hfill (for \hfill codatatypes).
-
-%%% TYPESETTING: horrible hack
-\pagebreak[2]
-\noindent
+phase makes inferences based on acyclicity (for \hfill datatypes) and uniqueness
+(for codatatypes).
 The third phase performs case splitting on constructors for
-various terms in $\tEc$.
+various terms in $\tEc$, the set of $\Sig$-terms occurring in $\Ec$.
 %Following the conventions from~\cite{},
@@
 are saturated.
-
-To simplify the presentation, we make the following assumptions about
-the signature $\Sig$.
-First, all codatatypes in $\Codata$ are corecursive. This is not restrictive
-because noncorecursive codatatypes can be seen instead as nonrecursive
-datatypes.
-Second, all types $\tau \in \Nondata$ have infinite cardinality.
-This is not overly restrictive, because each interpreted type $\tau \in \Nondata$
-having finite cardinality~$n$
-can be replaced by an enumeration datatype with $n$~nullary constructors.
-Finally, since the constraints in $\Ec$ are ground, they cannot entail an upper
-bound on the cardinality of any uninterpreted type $\tau \in \Nondata$; thus,
-we may without loss of generality %or completeness
-treat all types in $\Nondata$ as infinite.
 
@@
 \begin{figure}[t!]
+\vspace*{+3pt} %% TYPESETTING HACK
 \normalsize

### 2015-02-18 tex/body.tex
more squeezing

@@
 
-When looking at the ground theory, induction can be replaced by the acyclicity
-axiom schema, which states that any ground constructor term cannot be equal to
-a variable it contains \cite{barrett-et-al-2007}. Dually, coinduction can be
-replaced by the uniqueness axiom schema, which asserts that (co)recursive
-equations giving rise to the same infinite expansion---e.g., $x \teq
-\const{C}(x)$ and $y \teq \const{C}(\const{C}(y))$---must be equal.
+For the ground theory, induction is equivalent to the acyclicity
+axiom schema, which states that ground constructor terms cannot be equal to
+any of their proper subterms \cite{barrett-et-al-2007}. Dually, coinduction
+is equivalent to the uniqueness axiom schema, which asserts that codatatype
+values are fully characterized by their possibly infinite expansion.
 
@@
 A rule can be applied to $\Ec$ if all the specified preconditions are met.
-The conclusion of a rule either describes equalities to be added to $\Ec$
+The conclusion either describes equalities to be added to $\Ec$
 or is $\bot$ (contradiction).
@@
 % (in which case we call an application of it \emph{terminal}).
-A rule may have multiple conclusions, % separated by $\ror$,
+One rule has multiple conclusions, % separated by $\ror$,
 denoting branching.
@@
 equalities and where child nodes are obtained by a nonredundant application of a
-derivation rule to the parent node. A derivation tree is \emph{closed} if all of
+derivation rule to the parent node. \hfill A
+derivation tree
+\goodbreak\noindent %%% TYPESETTING: horrible HACKS
+is \emph{closed} if all of
 its leaf nodes are $\bot$. A node is \emph{saturated} if no nonredundant

### 2015-02-18 tex/body.tex
more

@@
 The third phase performs case splitting on constructors for
-various terms in $\tEc$, the set of $\Sig$-terms occurring in $\Ec$.
+various terms occurring in $\Ec$.
 %Following the conventions from~\cite{},
@@
 At the end of this phase, $\Ec$ induces an equivalence
-relation over $\tEc$ such that two terms $t$ and $u$ are equivalent if and
+relation over the set of $\Sig$-terms occurring in $\Ec$,
+denoted by $\tEc$, such that two terms $t$ and $u$ are equivalent if and
 only if $t \teq u \in \Ec$.

### 2015-02-18 tex/body.tex
shorten

@@
 have emerged in a few automatic theorem provers as well. The SMT-LIB
-\cite{barrett-et-al-2010} format, implemented by most SMT %(satisfiability
-%modulo theories)
+%\cite{barrett-et-al-2010}
+format, implemented by most SMT
+%(satisfiability modulo theories)
 solvers, has been extended with a syntax for datatypes.
@@
 One way to define datatypes semantically is as the initial model of the
-selector--constructor equations \cite{barrett-et-al-2010}.
+selector--constructor equations \cite{barrett-et-al-2007}.
 \begin{rep}

### 2015-02-18 tex/body.tex
more

@@
 \begin{enumerate}
-\item \noindent\rlap{Initially:}\phantom{Unfold $\Varec{x}$:\enskip}$\{  \}$
-\item \noindent\rlap{Unfold $\Varec{x}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) \vthinspace\}$
-\item \noindent\rlap{Unfold $\Varec{y}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\;
+%\item \noindent\rlap{Initially:}\phantom{Unfold $\Varec{x}$:\enskip}$\{  \}$
+\item \noindent\rlap{Unfold $\Varec{x}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\Val = \{\vthinspace  \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) \vthinspace\}$
+\item \noindent\rlap{Unfold $\vthinspace\Varec{y}\negvthinspace$}\phantom{Unfold $\Varec{x}$}:\enskip$\Val = \{\vthinspace  \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\;
   \ec{y} \mapsto \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) \vthinspace\}$
-\item \noindent\rlap{Unfold $\Varec{x}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\{\vthinspace  \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\;
+\item \noindent\rlap{Unfold $\Varec{x}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\Val = \{\vthinspace  \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\;
   \ec{y} \mapsto \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\; \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) ) \vthinspace\}$
@@
 %
-The result indicates that the values for $x$ and $y$ in models of $\Ec$
-are of the forms $\const{C}( {\Varec{w}},\allowbreak \const{C}( {\Varec{z}},\allowbreak \const{C}( {\Varec{w}},\allowbreak \const{C}( {\Varec{z}},\ldots ))))$
+The resulting $\Val$ indicates that the values for $x$ and $y$ in models of $\Ec$
+must be of the forms $\const{C}( {\Varec{w}}, \const{C}( {\Varec{z}},\allowbreak \const{C}( {\Varec{w}},\allowbreak \const{C}( {\Varec{z}},\ldots ))))$
 and $\const{C}( {\Varec{z}}, \const{C}( {\Varec{w}}, \const{C}( {\Varec{z}}, \const{C}( {\Varec{w}}, \ldots ))))$,

### 2015-02-18 tex/body.tex
more

@@
 whose values $\Val \ec{t} = \MU x.\; u$ are cyclic
-(as required by the precondition $x \in \FV( u )$),
+(as expressed by $x \in \FV( u )$),
 then $\Ec$ is $\thD$-unsatisfiable.
@@
 %for a capture-avoiding renaming of $\mu$-bound variables,
-then $t \teq u$. This precondition may seem too restrictive, since
+then $t \teq u$. This requirement may seem too restrictive, since
 $\MU x.\; \const{S}(x)$ and $\MU y.\; \const{S}(\const{S}(y))$ specify the same
@@
 \const{S}(\const{S}(y))$ must be equal to the entire term,
-as demonstrated in the following example.
+as demonstrated next.
 

### 2015-02-18 tex/body.tex
less

@@
 %\paragraph{Special Cases.}
+{\looseness=-1
 Some codatatypes are so degenerate as to have infinite values
@@
 singletons precisely when \ty{unit} is interpreted as a singleton.
-Fortunately, given cardinalities for the ordinary types, it is easy to
-characterize this degenerate case.
-
+Fortunately, \begin{rep}given cardinalities for the ordinary types, \end{rep}%
+it is easy to characterize this degenerate case.
+
+}
 % In other words, all finite corecursive codatatypes are corecursive singletons.
@@
 either infinite or a singleton. In the latter case, $\delta$ necessarily has a
-single constructor, whose arguments have types that are interpreted as
+single constructor, whose arguments have types that are interpreted by
 singletons.
@@
 
-A rule can be applied to $\Ec$ if all the specified preconditions are met.
+{\looseness=-1
+A rule can be applied to $\Ec$ if %all
+the %specified
+preconditions are met.
 The conclusion either describes equalities to be added to $\Ec$
-or is $\bot$ (contradiction).
+or is $\bot$. % (contradiction).
 % (in which case we call an application of it \emph{nonterminal})
@@
 equalities and where child nodes are obtained by a nonredundant application of a
-derivation rule to the parent node. \hfill A
-derivation tree
-\goodbreak\noindent %%% TYPESETTING: horrible HACKS
-is \emph{closed} if all of
+derivation rule to the parent. A derivation tree is \emph{closed} if all of
 its leaf nodes are $\bot$. A node is \emph{saturated} if no nonredundant
 instance of a rule can be applied to it.
+
+}
 

### 2015-02-18 tex/body.tex
fifteen pages!

@@
 also find models involving cyclic codatatype values. The $\mu$-notation is
-inspired by the $\mu$-calculus \cite[etc.]{kozen-1983,endrullis-et-al-2011}.
+inspired by the $\mu$-calculus
+\begin{conf}\cite{kozen-1983}\end{conf}%
+\begin{rep}\cite[etc.]{kozen-1983,endrullis-et-al-2011}\end{rep}%
+.
 
@@
 
+\kern-1.5pt %% TYPESETTING hack
+
 \paragraph{Phase 1: Computing the Bidirectional Closure.}
@@
 
+\kern-1.5pt %% TYPESETTING hack
+
 \paragraph{Phase 2: Applying Acyclicity and Uniqueness.}
@@
 $\FV( u )$ denotes the variables occurring freely in~$u$.
-
 %%% But there are no constants of type $\tau \in \Nondata$ -- at most variables!
@@
 %\const{c}$ to be a $\mu$-term if $\const{c}$ is a constant of type $\tau \in \Nondata$.
-
+%
 %%% I think this should be clear enough from the "Conventions" paragraph at the
@@
 %then $\MU x.\; \const{C}( \vec v )$ is a well-typed term of type $\upsilon$.
-
+%
 A $\mu$-term is \emph{closed} if %and only if --- by convention, "and only if" is not necessary for *definitions* (but add it back if you disagree)
-it contains no free variables. It is \emph{cyclic} if %and only if --- ditto
-it contains an occurrence of a bound variable.
-The notation $t \aequiv u$, expressing $\alpha$-\emph{equivalence},
+it contains no free variables. \hfill It is \emph{cyclic} if %and only if --- ditto
+it contains a bound variable. \hfill The
+\goodbreak %%% TYPESETTING: horrible hacks (including two hfills)
+\noindent
+notation $t \aequiv u$, for $\alpha$-\emph{equivalence},
 indicates that the $\mu$-terms $t$ and $u$
-are syntactically equivalent for some capture-avoiding renaming of $\mu$-bound variables---e.g.,
+are syntactically equivalent for some capture-avoiding renaming of $\mu$-bound variables\begin{rep}---e.g.,
 %$\MU x.\; \const{C}( x ) \aequiv \MU y.\; \const{C}( y )$ and
-$\MU x.\; \const{D}( y, x ) \aequiv \MU z.\; \const{D}( y, z )$\begin{rep},
+$\MU x.\; \const{D}( y, x ) \aequiv \MU z.\; \const{D}( y, z )$,
 but
@@
 \not\aequiv \MU y.\; \const{D}( y, x )$\end{rep}.
-Two $\mu$-terms can represent the same value while not being $\alpha$-equivalent---e.g.,
+Two $\mu$-terms can denote the same value despite being $\alpha$-disequivalent---e.g.,
 $\MU x.\; \const{C}( x ) \not\aequiv \MU y.\; \const{C}( \const{C}( y ) )$.
@@
 %
-At this point, phase~1 is activated again to refine the equivalence classes,
+At this point, phase~1 is activated again, % to XXX the equivalence classes,
 yielding
@@
 %$\{ {\Varec{y}} \mapsto {\VAREC{\const{S}(y)}}, {\VAREC{\const{S}(y)}} \mapsto {\Varec{y}} \}$,
-the \rn{Unique} rule can finally derive $x \teq y$.
+\rn{Unique} can finally derive $x \teq y$.
 \xend
@@
 terms from $\tEc$ have a corecursive singleton type
-(Section~\ref{sec:the-theory-of-co-datatypes}) and must therefore be equal. Both
-the finiteness assumption on the datatype in \rn{Split} and the singleton
-constraint on the codatatype in \rn{Single} can be evaluated statically
+%(Section~\ref{sec:the-theory-of-co-datatypes})
+and must therefore be equal. Both
+\rn{Split}'s finiteness assumption %on the datatype
+and \rn{Single}'s singleton
+constraint %on the codatatype
+can be evaluated statically
 based on a recursive computation of the cardinalities of the

### 2015-02-18 tex/body.tex
more

@@
 \newcommand{\tpos}[2]{#1\!\mid_{#2}}
-\newcommand{\muvar}{\mathcal{Var}}
+%\newcommand{\muvar}{\mathcal{Var}}
+\newcommand{\muvar}{\mathrm{Var}} %% cf. FV
 
@@
 Modulo $\alpha$-equivalence, the set \smash{$\mathcal{T}^{\smash{\J}}_{{\Varec{x}}}( \Val )$} is
-\[%\begin{array}{c}
+$%\begin{array}{c}
 \{
 \MU x.\; \const{D}( x ),\;
-\MU x.\; \const{C}( \const{D}( x )),\;
+\MU x.\; \const{C}( \const{D}( x )),\allowbreak\;
 \const{C}( \MU x.\; \const{D}( x )),\;
@@
 %\end{array}
-\]
+$.
 Consider a normal interpretation $\J$ that interprets variables in $\Ec$ based on the mapping $\Val$:
@@
 
+\pagebreak[2]
+
 Intuitively, this lemma states three properties of $\ValC$ that ensure a normal
@@
 
-%%% TODO: Fix data
+\newcommand\nummodden{num\hbox{\rm\_}mod\hbox{\rm\_}den}
 
@@
 
-\newcommand\nummodden{num\hbox{\rm\_}mod\hbox{\rm\_}den}
-
 The results are summarized in Table~\ref{tab:bench}. The decision procedure

### 2015-02-18 tex/body.tex
more

@@
 }{
-  \Ec := \Ec,\: t \teq \const{C}_1\bigl(\const s^1_1( t ), \ldots,\const s^{n_1}_1( t ) \bigr) %\ROR
+  \Ec := \Ec,\: t \teq \const{C}_1\bigl(\const s^1_1( t ), \ldots,\const s^{n_{1\!}}_1( t ) \bigr) %\ROR
   \ROR \cdots \ROR \Ec := \Ec,\: t \teq \const{C}_m\bigl(\const s^1_m( t ), \ldots,\const s^{n_m}_m( t ) \bigr)
@@
 $\thD$-unsatisfiable. For \rn{Split}, we add a clause %of the form
-$t \teq \const{C}_1\bigl( \const s^1_1( t ), \ldots, \const s^{n_1}_1( t ) \bigr)
+$t \teq \const{C}_1\bigl( \const s^1_1( t ), \ldots, \const s^{n_{1\!}}_1( t ) \bigr)
 \mathrel\lor \cdots \mathrel\lor t \teq \const{C}_m\bigl( \const s^1_m( t ),
@@
 However, the presence of quantified formulas poses an additional challenge in the context of (co)datatypes.
-Namely, since quantified formulas may entail cardinality constraints on uninterpreted types $U$,
-we may no longer assume that $U$ is of infinite cardinality.
+Namely, since quantified formulas may entail cardinality constraints on uninterpreted types $\ty{u}$,
+we may no longer assume that $\ty{u}$ is of infinite cardinality.
 When assuming that a finite type has infinite cardinality, the calculus presented in this paper is \emph{incomplete},
@@
 Second, we may conditionally apply \rn{Single} to terms of codatatype types that may have cardinality one.
-For instance, consider the codatatype $\keyw{codatatype}~\ty{c} = \const{C}(\ty{U},\: \ty{c})$, where $U$ is an uninterpreted type.
-This codatatype has cardinality one when $\ty{U}$ has cardinality one.
-If there exists two equivalence classes $\ec{t\typ{c}}$ and $\ec{u\typ{c}}$,
-our implementation adds a clause of the form $( \exists x\,y.\, x \not\teq y \vee t \teq u )$ to $F$,
-where $x$ and $y$ are variables of type $U$.
-In other words, either the cardinality of $U$ is greater than one, or $t$ must be equal to $u$.
+For example, consider the codatatype $\keyw{codatatype}~\ty{c} = \const{C}(\ty{u},\: \ty{c})$, where $\ty{u}$ is an uninterpreted type.
+This codatatype has cardinality one when $\ty{u}$ has cardinality one.
+If there exist two equivalence classes $\ec{s\typ{c}}$ and $\ec{t\typ{c}}$,
+our implementation adds the clause %of the form
+$\exists x\;y\typ{\ty{u}}.\;\, x \not\teq y \mathrel\vee s \teq t$ to $F$.
+In other words, either the cardinality of $\ty{u}$ is greater than one or $s$ is equal to $t$.
 

### 2015-02-18 tex/body.tex
more

@@
 formulas were described at a high level of abstraction. This section describes
-the implementation within the SMT solver CVC4.
-
-\paragraph{Integration into {\relax\XXXL($T\negvthinspace$)}.}
+the main aspects of the implementation within the SMT solver CVC4:\ the
+integration into CDCL($T$), the extension to quantified formulas, and
+some optimizations.
+
+%%% @ANDY: I'm trying without \paragraph's, to save space but also to make the
+%%% section look more like a whole than a random list of stuff. Revert if you
+%%% disagree.
+
+%\paragraph{Integration into {\relax\XXXL($T\negvthinspace$)}.}
 The calculus is implemented as a \emph{theory solver} of CVC4, that is, a
@@
 
-\paragraph{Extension to Quantified Formulas.}
+%\paragraph{Extension to Quantified Formulas.}
 While our calculus is a decision procedure for ground constraints in the theory of datatypes and codatatypes,
@@
 
-\paragraph{Optimizations.}
+%\paragraph{Optimizations.}
 Our implementation uses several optimizations which

### 2015-02-18 tex/body.tex
some edits

@@
 
-\newcommand\SSS{\mathit{S\vthinspace}}
+\newcommand\SSS{\mathit{S}}
 \newcommand\SSSS[1]{\mathit{S}^{\,#1}}
@@
 The decision procedure for (co)datatypes and its extension to quantified
-formulas were described at a high level of abstraction. This section describes
+formulas were presented at a high level of abstraction. This section describes
 the main aspects of the implementation within the SMT solver CVC4:\ the
@@
 %\paragraph{Integration into {\relax\XXXL($T\negvthinspace$)}.}
-The calculus is implemented as a \emph{theory solver} of CVC4, that is, a
+The calculus is implemented as a \relax{theory solver} of CVC4, that is, a
 specialized solver for determining the satisfiability of conjunctions of
 constraints for its theory.
-Given a theory $T = T_1 \mathrel\cup \cdots \mathrel\cup T_n$ and an input set of
-clauses $F$ in CNF, the \XXXL($T$) procedure
+Given a theory $T = T_1 \mathrel\cup \cdots \mathrel\cup T_n$ and a set of
+input clauses $F$ in CNF, the \XXXL($T$) procedure
 incrementally builds partial assignments from the atomic formulas of $F$ to
-truth values such that no clause in $F$ is falsified. We may interpret such a
-partial assignment as a set of literals where $a \in M$ for all atoms $a$ that
-it assigned the value $\top$, and $\neg\; a \in M$ for all atoms that it assigns
-the value $\bot$. Then, at a high level, by the Nelson--Oppen combination
-framework (its variant in CVC4 is described in~\cite{jovanovic2011sharing}), 
+truth values such that no clause in $F$ is falsified. We can interpret such a
+partial assignment as a set $M$ of literals where $a \in M$ for all atoms $a$ that
+it assigned the value $\top$ and $\neg\; a \in M$ for all atoms that it assigns
+the value $\bot$. Then, %at a high level,
+by a variant
+\cite{jovanovic2011sharing} of the Nelson--Oppen method,
 each $T_i$-solver for $i = 1, \ldots, l$ takes as input a combination
 of (1) the purified form of $T_i$-constraints occurring in $M$, where terms
-containing symbols not belonging to $T_i$ are replaced by fresh variables of the
-same type; (2) additional equalities and disequalities between variables of
+containing symbols not belonging to $T_i$ are replaced by fresh variables \begin{rep}of the
+same type\end{rep}; (2) additional equalities and disequalities between variables of
 types not belonging to $T_i$. Let us call this set $M_i$. Each $T_i$ solver
 either reports that a subset $C$ of $M_i$ is $T_i$-unsatisfiable, in which case
-$\neg\; C$ is added to $F$, adds a clause to $F$, or does nothing.
-When $M$ is a complete assignment for $F$, a theory solver can only choose to do
+$\neg\; C$ is added to $F\!$, adds a clause to $F\!$, or does nothing.
+When $M$ is a complete assignment for $F\!$, a theory solver can only choose to do
 nothing when $M_i$ is indeed $T_i$-satisfiable.
 
-Assume $\Ec$ is initially the set $M_i$ described above. For each equality $t \teq s$ added
+Assume $\Ec$ is initially the set $M_i$ described above. With each equality $t \teq s$ added
 to $\Ec$, we associate a set of equalities from $M_i$ that together entail $t
 \teq s$, which we call its \emph{explanation}.
-Similarly, each $\Val \ec{x}$ is also associated an explanation---that is, a set
+Similarly, each $\Val \ec{x}$ is assigned an explanation---that is, a set
 of equalities from $M_i$ that entail that the values of $\ec{x}$ in models of
@@
 
-Prior to constructing a model for $F$, the theory solver does not explicitly construct $\mu$-terms nor the map $\Val$.
+Prior to constructing a model for $F\!$, the theory solver does not explicitly construct $\mu$-terms nor the mapping $\Val$.
 Instead, $\Val$ is computed implicitly by traversing the equivalence classes of $\Ec$ during phase~2.
 To recognize when the precondition of \rn{Acyclic} holds,
-the procedure considers each datatype equivalence class $\ec{t}$ containing a constructor $\const{C}( t_1, \ldots, t_n )$.
-It visits $\ec{t_1}, \ldots, \ec{t_n}$ and all arguments of constructors in these equivalence classes in the same way.
+the procedure considers each equivalence class $\ec{t}$ containing a datatype constructor $\const{C}( t_1, \ldots, t_n )$.
+It visits $\ec{t_1}, \ldots, \ec{t_n}$ and all arguments of constructors in these equivalence classes recursively.
 If while doing so it returns to $\ec{t}$, then it deduces that \rn{Acyclic} applies to $t$.
 To recognize when the precondition of \rn{Unique} holds,
-the procedure considers the set $S$ of all codatatype equivalence classes.
-It simultaneously visits the equivalence classes of arguments of constructor terms in each equivalence class in $S$,
-while partitioning $S$ into $S_1, \ldots, S_n$ based on the top-most symbol of constructor terms in these equivalence classes
+the procedure considers the set $\SSS$ of all codatatype equivalence classes.
+It simultaneously visits the equivalence classes of arguments of constructor terms in each equivalence class in $\SSS$,
+while partitioning $\SSS$ into $\SSS_1, \ldots, \SSS_n$ based on the top-most symbol of constructor terms in these equivalence classes
 and the equivalence of their arguments of type $\Nondata$.
 It then partitions each set in this partition in the same manner recursively.
-If after doing so, the resulting partition contains a set $S_i$ containing distinct terms $u$ and $v$, then it deduces that \rn{Unique} applies to $u$ and $v$.
+If %after doing so,
+the resulting partition contains a set $\SSS_i$ containing distinct terms $u$
+and $v$, it deduces that \rn{Unique} applies to $u$ and $v$.
 

### 2015-02-18 tex/body.tex
more

@@
 %\paragraph{Extension to Quantified Formulas.}
-While our calculus is a decision procedure for ground constraints in the theory of datatypes and codatatypes,
+While the decision procedure is restricted to ground constraints,
 in practice we are often interested in problems that combine reasoning about universal axioms and existential conjectures,
 where checking satisfiability requires reasoning about such ground constraints in the presence of universally quantified formulas.
-Many SMT solvers, such as CVC4, support reasoning about quantified formulas using incomplete instantiation-based methods~\cite{MouraBjoerner07,ReynoldsTinelliMoura14}.
+Many SMT solvers, including CVC4, support reasoning about quantified formulas using incomplete instantiation-based methods~\cite{MouraBjoerner07,ReynoldsTinelliMoura14}.
 These methods extend naturally to quantified formulas over the signature of the theory of datatypes and codatatypes.
@@
 Namely, since quantified formulas may entail cardinality constraints on uninterpreted types $\ty{u}$,
-we may no longer assume that $\ty{u}$ is of infinite cardinality.
-When assuming that a finite type has infinite cardinality, the calculus presented in this paper is \emph{incomplete},
-since the preconditions on \rn{Split} and \rn{Single} may not apply when they indeed are applicable.
+we may no longer assume that they are of infinite cardinality.
+When assuming that a finite type has infinite cardinality, the calculus presented in this paper is \relax{incomplete},
+due to the overly restrictive preconditions on \rn{Split} and \rn{Single}.
 This does not necessarily impact the correctness of our procedure in this setting,
 since by assumption the solver is already incomplete when quantified formulas are present.
-Regardless, we may be interested in increasing the precision of the solver by doing the following.
-First, we may apply \rn{Split} to terms of datatype types whose cardinality depends on the finiteness of uninterpreted types.
-Second, we may conditionally apply \rn{Single} to terms of codatatype types that may have cardinality one.
-For example, consider the codatatype $\keyw{codatatype}~\ty{c} = \const{C}(\ty{u},\: \ty{c})$, where $\ty{u}$ is an uninterpreted type.
-This codatatype has cardinality one when $\ty{u}$ has cardinality one.
-If there exist two equivalence classes $\ec{s\typ{c}}$ and $\ec{t\typ{c}}$,
+Regardless, we can increase the precision of the solver by doing the following.
+First, we can apply \rn{Split} to datatype terms whose cardinality depends on the finiteness of uninterpreted types.
+Second, we can conditionally apply \rn{Single} to codatatype terms that may have cardinality one.
+For example, consider the $\ty{stream}_{\,\ty{u}}$ codatatype, where $\ty{u}$ is uninterpreted.
+This type has cardinality precisely when $\ty{u}$ has cardinality one.
+If there exist two equivalence classes $\ec{s}$ and $\ec{t}$ for this type,
 our implementation adds the clause %of the form
 $\exists x\;y\typ{\ty{u}}.\;\, x \not\teq y \mathrel\vee s \teq t$ to $F$.
-In other words, either the cardinality of $\ty{u}$ is greater than one or $s$ is equal to $t$.
+\begin{rep}In other words, either the cardinality of $\ty{u}$ is greater than one or $s$ is equal to $t$.\end{rep}
 
 %\paragraph{Optimizations.}
-Our implementation uses several optimizations which
+The implementation of the decision procedure uses several optimizations which
 follow the approach outlined by Barrett et al.\cite{barrett-et-al-2007}
 and are not reflected in the presentation of the calculus.
-We only briefly mention these optimizations, since
-each applies to datatypes as well as codatatypes, and most have been presented
-thoroughly in Barrett et al.
-
+We briefly mention the main ones.
+%
 Discriminators are treated as predicate symbols, so a discriminator $\const{d}_{\negvthinspace j}( t )$ is such that $\const{d}_{\negvthinspace j}$
 is a predicate symbol, instead of reducing this constraint to $t \teq \const{C}_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^n_{\negvthinspace j}( t ) \bigr)$.
-We have found that this leads to better performance since this reduction introduces terms more eagerly to $\tEc$.
+We have found that this leads to better performance because this reduction introduces terms more eagerly to $\tEc$.
 Handling discriminators requires extending the decision procedure with several
-additional rules \cite{barrett-et-al-2007},
+rules \cite{barrett-et-al-2007},
 which apply uniformly to both datatypes and codatatypes.
 Also, selectors are collapsed eagerly:
-If $\const s^k_{\negvthinspace j}( t )$ is a term in $\tEc$ and $t$ contains the
-constructor $\const{C}_{\negvthinspace j}( u_1, \ldots, u_n )$, we immediately add
+If $\const s^k_{\negvthinspace j}( t ) \in \tEc$ and $t =
+\const{C}_{\negvthinspace j}( u_1, \ldots, u_n )$, we immediately add
 $\smash{\const s^k_{\negvthinspace j}( t ) \teq u_k}$ to $\Ec$\begin{rep}, whereas the presented calculus

### 2015-02-18 tex/body.tex
more

@@
 
-Datatypes and codatatypes are an integral part of most modern proof assistants,
+Datatypes and codatatypes are an integral part of many proof assistants,
 including Agda, Coq, Isabelle, Matita, and PVS. In recent years, datatypes
@@
 solvers, has been extended with a syntax for datatypes.
-
-\pagebreak[2]
-
 In this \thewordpaper, we introduce a unified decision procedure for ground
-problems involving datatypes and codatatypes in combination
-(Section~\ref{sec:the-ground-decision-procedure}).
+problems involving datatypes and codatatypes in
+\begin{conf}\goodbreak\noindent\end{conf}%  %%% TYPESETTING: terrible hack
+combination (Section~\ref{sec:the-ground-decision-procedure}).
 The procedure is described abstractly as a calculus and is
@@
 Closely related is the automatic structural
-induction in both types of provers \cite{kersani-peltier-2013,wand-weidenbach-201x,reynolds-kuncak-2015}.
-%
-Datatypes are explicitly supported by the higher-order model finder Refute
+induction in both types of provers \cite{kersani-peltier-2013,cruanes-201x,wand-weidenbach-201x,reynolds-kuncak-2015},
+the (co)datatype and (co)induction support in Dafny \cite{leino-moskal-2014},
+and the (semi-)decision procedures implemented in Leon \cite{suter-et-al-2011} and RADA \cite{pham-et-al-2013}
+for datatypes.
+%
+Datatypes are supported by the higher-order model finder Refute
 \cite{weber-2008}. Its successor, Nitpick \cite{blanchette-nipkow-2010}, can
@@
 .
-
-% TODO: Leon, RADA -- http://www-users.cs.umn.edu/~hung/papers/fse13.pdf
 
@@
 nonempty domains. The set~$\Types$ must contain a
-distinguished type \ty{bool} interpreted as the set of truth
+\begin{rep}distinguished \end{rep}type \ty{bool} interpreted as the set of truth
 values. %, and may contain other interpreted types (e.g., \ty{int}, \ty{real}).
@@
 The only predicate is equality ($\teq$)\begin{rep} and belongs to the logical symbols\end{rep}.
-Other predicates can be represented as functions to $\ty{bool}$,
-with $\const{p}(\ldots)$ abbreviating $\const{p}(\ldots) \teq \const{true}$.
+Other predicates can be represented as functions to $\ty{bool}$\begin{rep},
+with $\const{p}(\ldots)$ abbreviating $\const{p}(\ldots) \teq \const{true}$\end{rep}.
 \begin{rep}The operator $\bigwedge_{\,i}\, \varphi_i$ abbeviates a conjunction
@@
 contribution has been the support for codatatypes. Both the metatheory and
-the implementation in CVC4 rely heavily on $\mu$-terms to represent cyclic values,
+the implementation in CVC4 rely on $\mu$-terms to represent cyclic values,
 a complication specific to codatatypes.

### 2015-02-18 tex/body.tex
more

@@
 % it comes to English prefixes like non-, in-, un-, and a-.)
-%\rem{more?}
 
@@
 %
-\rem{TODO: add small examples}
-\rem{TODO: explain why terminates}
+%\rem{TODO: add small examples}
+%\rem{TODO: explain why terminates}
 
@@
 \end{lemma}
-\begin{proof}
-\rem{TODO}
-\qed
-\end{proof}
+% TODO:
+%\begin{proof}
+%\qed
+%\end{proof}
 
@@
 
-\rem{TODO: Normal interpretation vs. actual interpretation.}
+%\rem{TODO: Normal interpretation vs. actual interpretation.}
 
@@
 such that $v \vsimv{x} \interp{t}{u}$ for some term $t$ with subterm $u$ occurring in the range of $\Val$.
-\rem{Intuition?}
+%\rem{TODO: Intuition?}
 
@@
 We construct $\ValC$ by exhaustively applying the following rule to $\nf{\Val}$:%
-\rem{A bit clumsy}
+%\rem{TODO: A bit clumsy}
 \[
@@
 modulo $\alpha$-equivalence.
-\rem{What guarantees that such a value exists? (Answer: infinitely many values
-of codatatypes, of which only finitely many are in
-$\mathcal{T}^{\smash{\J}}_{\Varec{x}}(\Val)$?)}
+%\rem{TODO: What guarantees that such a value exists? (Answer: infinitely many values
+%of codatatypes, of which only finitely many are in
+%$\mathcal{T}^{\smash{\J}}_{\Varec{x}}(\Val)$?)}
 Since this update removes one
@@
 )$, it can only be applied a finite number of times. We normalize all terms in
-the range of $\Val$ after each rule application.\rem{clumsy}
+the range of $\Val$ after each rule application.%\rem{TODO: clumsy}
 
@@
 the expansion of $\interp{\Val \ec{y}}{ t }$ as a term
-constructed by a sequence of substitutions of the form mentioned above to ${\Varec{z}}$. \rem{more detail}
+constructed by a sequence of substitutions of the form mentioned above to ${\Varec{z}}$. %\rem{TODO: more detail}
 Thus, we have
@@
 and $t_2$ is of the form $\MU w_2.\; \const{D}( \vec u )$,
-then the expansion \rem{define} of $\nf{\Val_1 \sigma} \ec{y}$ and $\nf{\Val_1 \sigma} \ec{z}$
+then the expansion %\rem{TODO: define}
+of $\nf{\Val_1 \sigma} \ec{y}$ and $\nf{\Val_1 \sigma} \ec{z}$
 are $\alpha$-disequivalent at position $p$.
@@
 which implies that the expansion of $\nf{\Val_1 \sigma} \ec{y}$ and $\nf{\Val_1 \sigma} \ec{z}$
-are $\alpha$-disequivalent at position $p$. \rem{more detail}
+are $\alpha$-disequivalent at position $p$. %\rem{TODO: more detail}
 
@@
 and $v$, it deduces that \rn{Unique} applies to $u$ and $v$.
-
-%\rem{TODO: * sharing is caring (let's cite it for a change)}
-%\rem{TODO: ``theory combination more complicated (and Theorem Completeness too weak)''?}
-%\rem{TODO: representation of $\mu$-terms?}
 
@@
 Isabelle formalizations using Sledgehammer.
-Codatatypes being a recent addition to Isabelle\begin{rep}
-\cite{blanchette-et-al-2014-impl}\end{rep}, benchmarks are not abundant. We
+Codatatypes being a recent addition to Isabelle
+\cite{blanchette-et-al-2014-impl}, benchmarks are not abundant. We
 included all the theory files from the Isabelle distribution (Distro, 879 goals)

### 2015-02-18 tex/body.tex
cleaner
more

@@
 For each $i \in \{1,\ldots, n\}$, let $u_\iPrime$ be the $i$\vvthinspace th argument of $t_u$.
-By Lemma~\ref{lem:mu-norm-arg}, $\interp{t_u}{u_\iPrime} \vsim \ValC \ec{u_i}$.
+\begin{rep}By Lemma~\ref{lem:mu-norm-arg},\end{rep}%
+\begin{conf}Clearly,\end{conf} %% TODO: Fix! ;)
+$\interp{t_u}{u_\iPrime} \vsim \ValC \ec{u_i}$.
 Thus $\interp{t_u}{u_\iPrime} \vsim \interp{\ValC \ec{t}}{ w_i }$. % for each $i \in [n]$.
@@
 two unpublished theories by Peter Gammie and Andreas Lochbihler (\gandl, 317
-goals), about Bird and Stern--Brocot trees. To also exercise the support
+goals), about Bird and Stern--Brocot trees. To exercise the support
 for datatypes, the benchmarks are complemented by theories about lists
-and tree data structures. The theories were selected before running the
+and trees. The theories were selected before running the
 experiments. The experimental data are publicly available.%
@@
 stronger decision procedures almost completely subsume the weaker ones. For
-Sledgehammer, the power of interpreted (co)datatypes is roughly comparable to
-that of arithmetic \cite{blanchette-et-al-2013-smt}.
+Sledgehammer, the practical usefulness of interpreted (co)datatypes appears
+roughly comparable to that of arithmetic \cite{blanchette-et-al-2013-smt}.
 

### 2015-02-18 tex/body.tex
minor

@@
 The results are summarized in Table~\ref{tab:bench}. The decision procedure
-makes a difference across all theories. Overall, it accounts for a success rate
-increase of \OK{5.2\%}. Moreover, every aspect of the procedure, including the
+makes a difference across all theories. It accounts for an overall success rate
+increase of over \OK{5\%}. Moreover, every aspect of the procedure, including the
 more expensive rules, make a contribution.
@@
 the \rn{Acyclic} rule and four required the \rn{Unique} rule. Among the latter,
-three are one-line proofs in Isabelle, of the form
-\keyw{by}~\textit{coinduction}~\textit{auto}\begin{rep} \cite{blanchette-et-al-2014-impl}\end{rep}.
+three are simple proofs of the form
+\keyw{by}~\textit{coinduction}~\textit{auto} %\cite{blanchette-et-al-2014-impl}
+in Isabelle.
 The fourth proof, by Gammie and Lochbihler, is more elaborate:
@@
 \begin{quote}
-\keyw{lemma} \,\textit{\nummodden\_unique}: \,$x = \const{Node}\;0\;\const{num}\; x \Longrightarrow x = \const{\nummodden}$ \\
+\keyw{lemma} \,\textit{\nummodden\_unique}: \,$x = \const{Node}\;0\;\const{num}\; x \vthinspace\Longrightarrow\vthinspace x = \const{\nummodden}$ \\
 \keyw{proof} \,(\textit{coinduction arbitrary}: $x$ \textit{rule}: \textit{tree.coinduct\_strong}) \\

### 2015-02-18 tex/body.tex
more

@@
 \DeclareFontShape{OT1}{pzc}{m}{it}{<-> s * [1.10] pzcmi7t}{}
-\DeclareMathAlphabet{\mathcal}{OT1}{pzc}{m}{it}
+\DeclareMathAlphabet{\mathcalx}{OT1}{pzc}{m}{it}
 
@@
 \newcommand{\Ec}{E}
-%\newcommand{\Ec}{\mathcal{E}}
+%\newcommand{\Ec}{\mathcalx{E}}
 %\newcommand{\Fc}{\mathsf{F}}
 \newcommand{\Fc}{F}
-%\newcommand{\Fc}{\mathcal{F}}
+%\newcommand{\Fc}{\mathcalx{F}}
 %\newcommand{\Gc}{\mathsf{G}}
@@
 \newcommand{\vdom}{\mathsf{dom}}
-\newcommand{\tEc}{\mathcal{T}(\Ec)}
-\newcommand{\tcEc}{\mathcal{T}^\AST(\Ec)}
+\newcommand{\tEc}{\mathcalx{T}(\Ec)}
+\newcommand{\tcEc}{\mathcalx{T}^\AST(\Ec)}
 \newcommand{\rn}[1]{\textsf{\small #1}}
@@
 \newcommand\ROR{\betweenantes}
-\newcommand{\tpath}[2]{\mathcal{P}_{ #2 \rightarrow \_ }( #1 )}
-\newcommand{\ttpath}[3]{\mathcal{P}_{ #2 \rightarrow #3 }( #1 )}
+\newcommand{\tpath}[2]{\mathcalx{P}_{ #2 \rightarrow \_ }( #1 )}
+\newcommand{\ttpath}[3]{\mathcalx{P}_{ #2 \rightarrow #3 }( #1 )}
 %\newcommand{\interp}[2]{#1\llbracket #2 \rrbracket}
@@
 \newcommand{\ec}[1]{[#1]}
-\newcommand{\J}{\mathcal{J}}
-\newcommand{\M}{\mathcal{M}}
-\newcommand{\Val}{\mathcal{A}}
-\newcommand{\ValC}{\mathcal{A}^\star}
-\newcommand{\Var}{\mathcal{V}}
+\newcommand{\JJJJ}{\mathcal{\!J\!}}
+\newcommand{\J}{\mathcalx{J}}
+\newcommand{\M}{\mathcalx{M}}
+\newcommand{\Val}{\mathcalx{A}}
+\newcommand{\ValC}{\mathcalx{A}^\star}
+\newcommand{\Var}{\mathcalx{V}}
 %\newcommand{\Varec}[1]{\Var{\ec{#1}}}
@@
 \newcommand{\tpos}[2]{#1\!\mid_{#2}}
-%\newcommand{\muvar}{\mathcal{Var}}
+%\newcommand{\muvar}{\mathcalx{Var}}
 \newcommand{\muvar}{\mathrm{Var}} %% cf. FV
@@
 %%% like you did in the string paper).
-%%% How about the following? I'm pretty sure \mathcal has been used for theories
+%%% How about the following? I'm pretty sure \mathcalx has been used for theories
 %%% before by other authors.
-\newcommand{\thD}{\mathcal{DC}}
+\newcommand{\thD}{\mathcalx{DC}}
 
 %\newcommand\Types{\mathcal{T}}
-\newcommand\Types{\mathcal{Y}}
-\newcommand\Funcs{\mathcal{F}}
+\newcommand\Types{\mathcalx{Y}}
+\newcommand\Funcs{\mathcalx{F}}
 
@@
 problems and is composable via the Nelson--Oppen method. It has been
-implemented in the latest version of CVC4, a state-of-the-art SMT solver. An
-evaluation based on problems generated from Isabelle theories demonstrates the
+implemented in CVC4, a state-of-the-art SMT solver. An evaluation based on
+problems generated from theories developed with Isabelle demonstrates the
 potential of the procedure.
@@
 models the natural numbers $\const{Z}$, $\const{S}(\const{Z})$, $\const{S}(\const{S}(\const{Z}))$, $\ldots$\afterLdots{},
-using Peano notation but extended with an
+using Peano notation, but extended with an
 infinite value $\infty = \const{S}(\const{S}(\const{S}(\ldots)))$.
@@
 codatatypes.
+\begin{conf}
+Detailed proofs of soundness and completeness are included in a technical report
+\cite{our-report}.
+\end{conf}%
 
@@
 
-\begin{conf}
-Detailed proofs of soundness and completeness are included in a technical report
-\cite{our-report}.
-\end{conf}%
 The procedure is implemented in the SMT solver CVC4 \cite{barrett-et-al-2011} as a combination
@@
 superposition prover \begin{rep}by Daniel Wand\end{rep} \cite{wand-2014}.
-Closely related is the automatic structural
+Closely related are the automatic structural
 induction in both types of provers \cite{kersani-peltier-2013,cruanes-201x,wand-weidenbach-201x,reynolds-kuncak-2015},
 the (co)datatype and (co)induction support in Dafny \cite{leino-moskal-2014},
-and the (semi-)decision procedures implemented in Leon \cite{suter-et-al-2011} and RADA \cite{pham-et-al-2013}
-for datatypes.
+and the (semi-)decision procedures for datatypes
+implemented in Leon \cite{suter-et-al-2011} and RADA \cite{pham-et-al-2013}.
 %
@@
 \cite{weber-2008}. Its successor, Nitpick \cite{blanchette-nipkow-2010}, can
-also find models involving cyclic codatatype values. The $\mu$-notation is
+also generate models involving cyclic codatatype values. The $\mu$-notation is
 inspired by the $\mu$-calculus
@@
 In an SMT problem, the signature is typically given by specifying first the
-uninterpreted types in any order, then the (co)datatypes with their constructors
+uninterpreted ordinary types in any order, then the (co)datatypes with their constructors
 and selectors in groups of mutually (co)recursive groups of (co)datatypes, and
@@
 %
-Each (co)datatype specification consists of $l$~mutually recursive types which are
+Each (co)datatype specification consists of $l$~mutually recursive types that are
 either all datatypes or all codatatypes. Polymorphic types, nested
@@
 \cite{blanchette-et-al-2014-codata}).}
-In the presentation, we allow ourselves some metalevel parameterization
+We allow ourselves some notational parameterization
 % at the metalevel
@@
 it is not necessary to extend the signature:
-The discriminator term $\const{d}_{i\negvthinspace j}(x)$ can be seen as an abbreviation for
+The discriminator atom $\const{d}_{i\negvthinspace j}(x) \begin{conf}\teq
+\const{true}\end{conf}$ can be seen as an abbreviation for
 $x = \const{C}_{i\negvthinspace j}\bigl(\const{s}_{i\negvthinspace j}^1(x), \ldots, \const{s}_{i\negvthinspace j}^{\,n_{\smash{{i\negvthinspace j}}}}(x)\bigr)$.
@@
 datatypes and codatatypes},
-which in addition to $\Sig$ defines a class of $\Sig$-interpretations $\mathbf{J}$,
+which in addition to $\Sig$ defines a class of $\Sig$-interpretations $\JJJJ$,
 namely the ones that satisfy the %axioms --- existence of all infinite codatatype values cannot be axiomatized
 properties mentioned in this section, including (co)induction. In particular, the interpretations
-in~$\mathbf{J}$ share the same fixed interpretation for constructor terms and correctly
+in~$\JJJJ$ share the same fixed interpretation for constructor terms and correctly
 applied selector terms but may differ on wrongly applied
@@
 A formula $\varphi$ is $\thD$-\emph{satisfiable} if there
-exists an interpretation in $\mathbf{J}$ that satisfies $\varphi$.
+exists an interpretation in $\JJJJ$ that satisfies $\varphi$.
 % otherwise, it is $\thD$-unsatisfiable.  (I trust the readers' intelligence when
@@
 induction on the rules of the calculus that each non-root node $\Fc$ is such that
-$\mathcal{T}(\Fc) \subseteq \tcEc$, and hence contains an equality between two terms from $\tcEc$ not occurring in its parent node.
+$\mathcalx{T}(\Fc) \subseteq \tcEc$, and hence contains an equality between two terms from $\tcEc$ not occurring in its parent node.
 Thus, the depth of a branch in a derivation tree with root node $\Ec$ is at most $\left| \smash{\tcEc}\vphantom{X} \right|{\!}^2$,
@@
 For a variable $x\typ{\tau}$ and a fixed normal interpretation $\J$,
-we write $\mathcal{T}^{\smash{\J}}_x( \Val )$ to denote the set consisting of all values
+we write $\mathcalx{T}^{\smash{\J}}_x( \Val )$ to denote the set consisting of all values
 $v \in \J( \tau )$
@@
   \betweenantes
-  v \not\in \mathcal{T}^{\smash{\J}}_{\Varec{x}}(\Val)
+  v \not\in \mathcalx{T}^{\smash{\J}}_{\Varec{x}}(\Val)
 }{
@@
 The rule chooses an unassigned variable in $\Val$ and assigns it a fresh
-value---one that does not occur in $\mathcal{T}^{\smash{\J}}_{\Varec{x}}(\Val)$
+value---one that does not occur in $\mathcalx{T}^{\smash{\J}}_{\Varec{x}}(\Val)$
 modulo $\alpha$-equivalence.
@@
 %of codatatypes, of which only finitely many are in
-%$\mathcal{T}^{\smash{\J}}_{\Varec{x}}(\Val)$?)}
+%$\mathcalx{T}^{\smash{\J}}_{\Varec{x}}(\Val)$?)}
 Since this update removes one
@@
 $x$ is interpreted as $\Val \ec{x}$, $y$ as $\Val \ec{y}$, and so on.
-Assigning a value for $\Val \ec{x}$ that is $\alpha$-equivalent to a value in $\mathcal{T}^{\smash{\J}}_{{\Varec{x}}}( \Val )$
+Assigning a value for $\Val \ec{x}$ that is $\alpha$-equivalent to a value in $\mathcalx{T}^{\smash{\J}}_{{\Varec{x}}}( \Val )$
 may cause values in the range of $\Val$ to become $\alpha$-equivalent,
@@
 However, this contradicts the disequality $x \not\teq z$ in~$\Ec$.
-On the other hand, if the value assigned to ${\Varec{x}}$ is fresh, %not $\alpha$-equivalent to any term in $\mathcal{T}^{\smash{\M}}_{{\Varec{x}}}( \Val )$,
+On the other hand, if the value assigned to ${\Varec{x}}$ is fresh, %not $\alpha$-equivalent to any term in $\mathcalx{T}^{\smash{\M}}_{{\Varec{x}}}( \Val )$,
 the values in the range of $\Val$ remain $\alpha$-disequivalent.
@@
 there is only one equivalence class of type $\tau$ in $\Ec$,
-and thus there are no terms in $\mathcal{T}^{\smash{\J}}_{\Varec{y}}(\ValC)$ of type $\tau$.
+and thus there are no terms in $\mathcalx{T}^{\smash{\J}}_{\Varec{y}}(\ValC)$ of type $\tau$.
 This is a contradiction, since %our %model
@@
 This is also a contradiction,
-since there are only a finite number of closed terms in $\mathcal{T}^{\smash{\J}}_{\Varec{y}}(\ValC)$,
+since there are only a finite number of closed terms in $\mathcalx{T}^{\smash{\J}}_{\Varec{y}}(\ValC)$,
 and thus %our %model
-completion can assign a value not occurring in $\mathcal{T}^{\smash{\J}}_{\Varec{y}}(\ValC)$ to ${\Varec{y}}$.
+completion can assign a value not occurring in $\mathcalx{T}^{\smash{\J}}_{\Varec{y}}(\ValC)$ to ${\Varec{y}}$.
 By construction, $\ValC \ec{x}$ is normal.
@@
 ${\Varec{x}} \in FV( \Val_1 )$,
-and $\MU {\Varec{x}}.\; t$ is not $\alpha$-equivalent to a term in $\mathcal{T}^{\smash{\J}}_{\Varec{x}}(\Val_1)$.
+and $\MU {\Varec{x}}.\; t$ is not $\alpha$-equivalent to a term in $\mathcalx{T}^{\smash{\J}}_{\Varec{x}}(\Val_1)$.
 To show 2,
@@
 Since $\interp{\Val_1 \ec{y}}{t_1}$ is normal,
-there is a closed $\mu$-term $v \in \mathcal{T}^{\smash{\J}}_{\Varec{x}}( \Val_1 )$
+there is a closed $\mu$-term $v \in \mathcalx{T}^{\smash{\J}}_{\Varec{x}}( \Val_1 )$
 such that $v \vsimv{\Varec{x}} \interp{\Val_1 \ec{y}}{t_1}$.
@@
 and that interprets the variables of $\Fc$ based on the completion $\ValC$ of the mapping $\Val$ from phase~2.
-For the latter, let $\M( x )$ be the value in $\M$ that is $\alpha$-equivalent with $\ValC \ec{x}$ for each variable $x \in \mathcal{T}(\Fc)$,
+For the latter, let $\M( x )$ be the value in $\M$ that is $\alpha$-equivalent with $\ValC \ec{x}$ for each variable $x \in \mathcalx{T}(\Fc)$,
 which by Lemma~\ref{lem:interpretation-completion}\PART{1} is guaranteed to exist.
@@
 by structural induction on $t$,
-that $\M( t ) \aequiv \ValC \ec{t}$ for all terms $t \in \mathcal{T}( \Fc )$,
+that $\M( t ) \aequiv \ValC \ec{t}$ for all terms $t \in \mathcalx{T}( \Fc )$,
 which since $\M$ is normal, implies $\M \models t_1 \teq t_2$.

### 2015-02-18 tex/body.tex
more

@@
 For the ground theory, induction is equivalent to the acyclicity
-axiom schema, which states that ground constructor terms cannot be equal to
+axiom schema, which states that constructor terms cannot be equal to
 any of their proper subterms \cite{barrett-et-al-2007}. Dually, coinduction
@@
 \vthinspace$\keyw{codatatype}~\ty{a} = \const{A}(\ty{a})$, whose
-only value is $\MU a.\; A(a)$. Other specimens are
+unique value is $\MU a.\; A(a)$. Other specimens are
 $\ty{stream}_{\,\ty{unit}\,}$

### 2015-02-18 tex/body.tex
more

@@
 Given a fixed signature~$\Sig$,
-the decision procedure for the ground theory of (co)datatypes $\thD$ determines the
+the decision procedure for the ground theory of (co)datatypes %$\thD$
+determines the
 $\thD$-satisfiability of finite sets $\Ec$ of constraints:\ equalities
@@
 Second, all types $\tau \in \Nondata$ have infinite cardinality.
-This is not overly restrictive, because each interpreted type in $\Nondata$
+This is not very limitative, because each interpreted type in $\Nondata$
 having finite cardinality~$n$
@@
 {\looseness=-1
-A rule can be applied to $\Ec$ if %all
+A derivation rule can be applied to $\Ec$ if %all
 the %specified
@@
 %
-An application of a rule is \emph{redundant} if at least one branch in the
+An application of a rule is \emph{redundant} if one branch in the
 conclusion is an assignment to $\Ec$ that leaves it unchanged.
@@
 phases. The first phase computes the bidirectional closure of $\Ec$. The second
-phase makes inferences based on acyclicity (for \hfill datatypes) and uniqueness
+phase makes inferences based on acyclicity (for datatypes) and uniqueness
 (for codatatypes).
-The third phase performs case splitting on constructors for
+The third phase performs case distinctions on constructors for
 various terms occurring in $\Ec$.

### 2015-02-18 tex/body.tex
reorder

@@
 The premises of the rules that make up the second phase refer to a mapping $\Val$
-that associates a $\mu$-term as representative with each equivalence class.
+that associates a $\mu$-term with each equivalence class as its representative.
 %Recall that $\mu$-bindings can be used for representing possibly cyclic terms and values.
@@
 % for example, $\MU x.\; \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
-A $\mu$-term $t$ of type $\tau$ describes a class of values of
-$\tau$ that terms in that equivalence class can take in the models of $\Ec$.
+A $\mu$-term $t\typ{\tau}$ describes a class of $\tau$ values
+that terms in that equivalence class can take in the models of $\Ec$.
 %either in the case when $\tau$ is a codatatype or datatype type,
 When $\tau$ is a datatype, % \in \Data$,
-a cyclic $t$ describes an infeasible class of values.
+a cyclic $\mu$-term describes an infeasible class of values.
 %$\tau \not\in \Codata$.
@@
 $\mu$-terms $\vec t$ of the expected types.
-The variable $x$ need not occur freely in the binder's body, in which case
+The variable $x$ need not occur freely in the $\mu$-binder's body, in which case
 the binder can be omitted.
@@
 A $\mu$-term is \emph{closed} if %and only if --- by convention, "and only if" is not necessary for *definitions* (but add it back if you disagree)
-it contains no free variables. \hfill It is \emph{cyclic} if %and only if --- ditto
-it contains a bound variable. \hfill The
-\goodbreak %%% TYPESETTING: horrible hacks (including two hfills)
+it contains no free variables. \,It is \emph{cyclic} if %and only if --- ditto
+it contains a bound variable. \,The
+\goodbreak %%% TYPESETTING: horrible hacks (including two \,'s)
 \noindent
-notation $t \aequiv u$, for $\alpha$-\emph{equivalence},
+$\alpha$-\emph{equivalence} relation $t \aequiv u$
 indicates that the $\mu$-terms $t$ and $u$
@@
 
-The mapping $\Val$ is computed as follows.
-%
-With each equivalence class $\ec{u}$, we associate a fresh variable $\Varec{u}$
-of the same type as $u$
+The mapping $\Val$, on which the rules depends, is computed as follows.
+%
+With each equivalence class $\ec{u\typ{\tau}}$, we associate a fresh variable $\Varec{u}\typ{\tau}$
+%of the same type as $u$
 %not occurring in $\tEc$,
@@
 $\FV( \Val )$ denotes the union of the free variables occurring in $\Val$'s range,
-and $\Val[x \mapsto t]$ denotes the variable-capturing substitution of $t$ for
+and $\Val[x \mapsto t]$ denotes the \emph{variable-capturing} substitution of $t$ for
 $x$ in $\Val$'s range. It is easy to see that the height of terms produced as a
@@
 \begin{examplex}
-Let $\delta$ be a codatatype with the constructors $\const{C}, \const{D}, \const{E} :\delta \rightarrow \delta$.
+Let $\delta$ be a codatatype with the constructors $\const{C},\, \const{D},\, \const{E} :\delta \rightarrow \delta$.
 Let $\Ec$ be the set
@@
 %\[\begin{array}{r@{}c@{}l@{\;}l@{\qquad}r@{}c@{}r@{\;}l@{}l@{\;}l@{\qquad}r@{}c@{}r@{\;}l}
-\[\begin{array}{r@{}c@{}r@{\qquad}r@{}c@{}r@{\;}l@{}l@{\;}l@{\qquad}r@{}c@{}l@{\;}l}
-\Val \ec{x} & = & {\Varec{x}} &
-\Val \ec{z} & = & \MU {\Varec{z}}. & \multicolumn{2}{@{}l@{}}{\const{D}( {\Varec{x}} )} & &
-\Val \ec{u} & {}\,=\,{} & \MU {\Varec{u}}. & \const{C}( {\Varec{x}} ) \\
-\Val \ec{y} & = & {\Varec{y}} &
-\Val \ec{w} & {}\,=\,{} & \MU {\Varec{w}}. & \const{C}( & \MU {\Varec{z}}. & \const{D}( {\Varec{x}} ) ) &
-\Val \ec{v} & = & \MU {\Varec{v}}. & \const{E}( {\Varec{y}} )
+\[\begin{array}{r@{}c@{}r@{}l@{\qquad}r@{}c@{}r@{}l@{\qquad}r@{}c@{}l}
+\Val \ec{u} & {}\,=\,{} & \MU {\Varec{u}}.\; & \const{C}( {\Varec{x}} ) &
+\Val \ec{v} & = & \MU {\Varec{v}}.\; & \const{E}( {\Varec{y}} ) &
+\Val \ec{y} & {}\,=\,{} & {\Varec{y}} \\
+\Val \ec{z} & = & \MU {\Varec{z}}.\; & \const{D}( {\Varec{x}} ) &
+\Val \ec{w} & {}\,=\,{} & \MU {\Varec{w}}.\; & \const{C}( \MU {\Varec{z}}.\; \const{D}( {\Varec{x}} ) ) &
+\Val \ec{x} & = & {\Varec{x}}
 \end{array}\]

### 2015-02-18 tex/body.tex
renamed vars

@@
 $\{
-w \teq \const{C}( z ),\;
-z \teq \const{D}( x ),\;
-u \teq \const{C}( x ),\;
-v \teq \const{E}( y ),\;
-x \not\teq z
+u \teq \const{C}( z ),\;
+v \teq \const{D}( z ),\;
+w \teq \const{E}( y ),\;
+x \teq \const{C}( v ),\;
+z \not\teq v
 \}$.
@@
 \[\begin{array}{r@{}c@{}r@{}l@{\qquad}r@{}c@{}r@{}l@{\qquad}r@{}c@{}l}
-\Val \ec{u} & {}\,=\,{} & \MU {\Varec{u}}.\; & \const{C}( {\Varec{x}} ) &
-\Val \ec{v} & = & \MU {\Varec{v}}.\; & \const{E}( {\Varec{y}} ) &
+\Val \ec{u} & {}\,=\,{} & \MU {\Varec{u}}.\; & \const{C}( {\Varec{z}} ) &
+\Val \ec{w} & = & \MU {\Varec{w}}.\; & \const{E}( {\Varec{y}} ) &
 \Val \ec{y} & {}\,=\,{} & {\Varec{y}} \\
-\Val \ec{z} & = & \MU {\Varec{z}}.\; & \const{D}( {\Varec{x}} ) &
-\Val \ec{w} & {}\,=\,{} & \MU {\Varec{w}}.\; & \const{C}( \MU {\Varec{z}}.\; \const{D}( {\Varec{x}} ) ) &
-\Val \ec{x} & = & {\Varec{x}}
+\Val \ec{v} & = & \MU {\Varec{v}}.\; & \const{D}( {\Varec{z}} ) &
+\Val \ec{x} & {}\,=\,{} & \MU {\Varec{x}}.\; & \const{C}( \MU {\Varec{v}}.\; \const{D}( {\Varec{z}} ) ) &
+\Val \ec{z} & = & {\Varec{z}}
 \end{array}\]
@@
 To construct a completion $\ValC$, we must choose values for the free
-variables of $\Val$, namely ${\Varec{x}}$ and ${\Varec{y}}$.
-Modulo $\alpha$-equivalence, the set \smash{$\mathcal{T}^{\smash{\J}}_{{\Varec{x}}}( \Val )$} is
+variables of $\Val$, namely ${\Varec{z}}$ and ${\Varec{y}}$.
+Modulo $\alpha$-equivalence, the set \smash{$\mathcal{T}^{\smash{\J}}_{{\Varec{z}}}( \Val )$} is
 $%\begin{array}{c}
 \{
-\MU x.\; \const{D}( x ),\;
-\MU x.\; \const{C}( \const{D}( x )),\allowbreak\;
-\const{C}( \MU x.\; \const{D}( x )),\;
-\MU x.\; \const{C}( x )
+\MU z.\; \const{D}( z ),\;
+\MU z.\; \const{C}( \const{D}( z )),\allowbreak\;
+\const{C}( \MU z.\; \const{D}( z )),\;
+\MU z.\; \const{C}( z )
 \}
@@
 Consider a normal interpretation $\J$ that interprets variables in $\Ec$ based on the mapping $\Val$:
-$x$ is interpreted as $\Val \ec{x}$, $y$ as $\Val \ec{y}$, and so on.
-Assigning a value for $\Val \ec{x}$ that is $\alpha$-equivalent to a value in $\mathcalx{T}^{\smash{\J}}_{{\Varec{x}}}( \Val )$
+$z$ is interpreted as $\Val \ec{z}$, $y$ as $\Val \ec{y}$, and so on.
+Assigning a value for $\Val \ec{z}$ that is $\alpha$-equivalent to a value in $\mathcalx{T}^{\smash{\J}}_{{\Varec{z}}}( \Val )$
 may cause values in the range of $\Val$ to become $\alpha$-equivalent,
 which in turn may cause $\Ec$ to be falsified by $\J$.
-For example, assign $\MU {\Varec{x}}.\; \const{D}( {\Varec{x}} )$ for ${\Varec{x}}$.
-After this substitution, $\Val \ec{z}$ is $\MU {\Varec{z}}.\; \const{D}( \MU {\Varec{x}}.\; \const{D}( {\Varec{x}} ) )$,
-which has normal form $\MU {\Varec{z}}.\; \const{D}( {\Varec{z}} )$,
-which is $\alpha$-equivalent to $\MU {\Varec{x}}.\; \const{D}( {\Varec{x}} )$.
-However, this contradicts the disequality $x \not\teq z$ in~$\Ec$.
-On the other hand, if the value assigned to ${\Varec{x}}$ is fresh, %not $\alpha$-equivalent to any term in $\mathcalx{T}^{\smash{\M}}_{{\Varec{x}}}( \Val )$,
+For example, assign $\MU {\Varec{z}}.\; \const{D}( {\Varec{z}} )$ for ${\Varec{z}}$.
+After this substitution, $\Val \ec{v}$ is $\MU {\Varec{v}}.\; \const{D}( \MU {\Varec{z}}.\; \const{D}( {\Varec{z}} ) )$,
+which has normal form $\MU {\Varec{v}}.\; \const{D}( {\Varec{v}} )$,
+which is $\alpha$-equivalent to $\MU {\Varec{z}}.\; \const{D}( {\Varec{z}} )$.
+However, this contradicts the disequality $z \not\teq v$ in~$\Ec$.
+On the other hand, if the value assigned to ${\Varec{z}}$ is fresh, %not $\alpha$-equivalent to any term in $\mathcalx{T}^{\smash{\M}}_{{\Varec{z}}}( \Val )$,
 the values in the range of $\Val$ remain $\alpha$-disequivalent.
 For example, we may assign a value such as
-$\MU {\Varec{x}}.\; \const{E}( {\Varec{x}} )$ or
-$\MU {\Varec{x}}.\; \const{D}( \const{C}( {\Varec{x}} ) )$
-to ${\Varec{x}}$.
-Legal substitutions for ${\Varec{x}}$ may cause the range of $\Val$ to contain abnormal terms.
-For example, after assigning $\MU {\Varec{x}}.\; \const{D}( \const{C}( {\Varec{x}} ) )$ to $\Varec{x}$, we have that
-$\Val \ec{u}$ is $\MU {\Varec{u}}.\; \const{C}( \MU {\Varec{x}}.\; \const{D}( \const{C}( {\Varec{x}} ) ) )$,
+$\MU {\Varec{z}}.\; \const{E}( {\Varec{z}} )$ or
+$\MU {\Varec{z}}.\; \const{D}( \const{C}( {\Varec{z}} ) )$
+to ${\Varec{z}}$.
+Legal substitutions for ${\Varec{z}}$ may cause the range of $\Val$ to contain abnormal terms.
+For example, after assigning $\MU {\Varec{z}}.\; \const{D}( \const{C}( {\Varec{z}} ) )$ to $\Varec{z}$, we have that
+$\Val \ec{u}$ is $\MU {\Varec{u}}.\; \const{C}( \MU {\Varec{z}}.\; \const{D}( \const{C}( {\Varec{z}} ) ) )$,
 which has normal form
-$\MU {\Varec{u}}.\; \const{C}( \MU {\Varec{x}}.\; \const{D}( {\Varec{u}} ) )$.
+$\MU {\Varec{u}}.\; \const{C}( \MU {\Varec{z}}.\; \const{D}( {\Varec{u}} ) )$.
 \xend

### 2015-02-18 tex/body.tex
more

@@
 \newcommand{\vdom}{\mathsf{dom}}
-\newcommand{\tEc}{\mathcalx{T}(\Ec)}
-\newcommand{\tcEc}{\mathcalx{T}^\AST(\Ec)}
+\newcommand{\tEc}{\Terms(\Ec)}
+\newcommand{\tcEc}{\Terms^\AST(\Ec)}
 \newcommand{\rn}[1]{\textsf{\small #1}}
@@
 
-%\newcommand\Types{\mathcal{T}}
+\newcommand\Terms{\mathcalx{T}}
 \newcommand\Types{\mathcalx{Y}}
@@
 induction on the rules of the calculus that each non-root node $\Fc$ is such that
-$\mathcalx{T}(\Fc) \subseteq \tcEc$, and hence contains an equality between two terms from $\tcEc$ not occurring in its parent node.
+$\Terms(\Fc) \subseteq \tcEc$, and hence contains an equality between two terms from $\tcEc$ not occurring in its parent node.
 Thus, the depth of a branch in a derivation tree with root node $\Ec$ is at most $\left| \smash{\tcEc}\vphantom{X} \right|{\!}^2$,
@@
 For a variable $x\typ{\tau}$ and a fixed normal interpretation $\J$,
-we write $\mathcalx{T}^{\smash{\J}}_x( \Val )$ to denote the set consisting of all values
+we write $\Terms^{\smash{\J}}_x( \Val )$ to denote the set consisting of all values
 $v \in \J( \tau )$
@@
   \betweenantes
-  v \not\in \mathcalx{T}^{\smash{\J}}_{\Varec{x}}(\Val)
+  v \not\in \Terms^{\smash{\J}}_{\Varec{x}}(\Val)
 }{
@@
 The rule chooses an unassigned variable in $\Val$ and assigns it a fresh
-value---one that does not occur in $\mathcalx{T}^{\smash{\J}}_{\Varec{x}}(\Val)$
+value---one that does not occur in $\Terms^{\smash{\J}}_{\Varec{x}}(\Val)$
 modulo $\alpha$-equivalence.
@@
 %of codatatypes, of which only finitely many are in
-%$\mathcalx{T}^{\smash{\J}}_{\Varec{x}}(\Val)$?)}
+%$\Terms^{\smash{\J}}_{\Varec{x}}(\Val)$?)}
 Since this update removes one
@@
 variables of $\Val$, namely ${\Varec{z}}$ and ${\Varec{y}}$.
-Modulo $\alpha$-equivalence, the set \smash{$\mathcal{T}^{\smash{\J}}_{{\Varec{z}}}( \Val )$} is
+Modulo $\alpha$-equivalence, the set \smash{$\Terms^{\smash{\J}}_{{\Varec{z}}}( \Val )$} is
 $%\begin{array}{c}
@@
 $z$ is interpreted as $\Val \ec{z}$, $y$ as $\Val \ec{y}$, and so on.
-Assigning a value for $\Val \ec{z}$ that is $\alpha$-equivalent to a value in $\mathcalx{T}^{\smash{\J}}_{{\Varec{z}}}( \Val )$
+Assigning a value for $\Val \ec{z}$ that is $\alpha$-equivalent to a value in $\Terms^{\smash{\J}}_{{\Varec{z}}}( \Val )$
 may cause values in the range of $\Val$ to become $\alpha$-equivalent,
@@
 However, this contradicts the disequality $z \not\teq v$ in~$\Ec$.
-On the other hand, if the value assigned to ${\Varec{z}}$ is fresh, %not $\alpha$-equivalent to any term in $\mathcalx{T}^{\smash{\M}}_{{\Varec{z}}}( \Val )$,
+On the other hand, if the value assigned to ${\Varec{z}}$ is fresh, %not $\alpha$-equivalent to any term in $\Terms^{\smash{\M}}_{{\Varec{z}}}( \Val )$,
 the values in the range of $\Val$ remain $\alpha$-disequivalent.
@@
 there is only one equivalence class of type $\tau$ in $\Ec$,
-and thus there are no terms in $\mathcalx{T}^{\smash{\J}}_{\Varec{y}}(\ValC)$ of type $\tau$.
+and thus there are no terms in $\Terms^{\smash{\J}}_{\Varec{y}}(\ValC)$ of type $\tau$.
 This is a contradiction, since %our %model
@@
 This is also a contradiction,
-since there are only a finite number of closed terms in $\mathcalx{T}^{\smash{\J}}_{\Varec{y}}(\ValC)$,
+since there are only a finite number of closed terms in $\Terms^{\smash{\J}}_{\Varec{y}}(\ValC)$,
 and thus %our %model
-completion can assign a value not occurring in $\mathcalx{T}^{\smash{\J}}_{\Varec{y}}(\ValC)$ to ${\Varec{y}}$.
+completion can assign a value not occurring in $\Terms^{\smash{\J}}_{\Varec{y}}(\ValC)$ to ${\Varec{y}}$.
 By construction, $\ValC \ec{x}$ is normal.
@@
 ${\Varec{x}} \in FV( \Val_1 )$,
-and $\MU {\Varec{x}}.\; t$ is not $\alpha$-equivalent to a term in $\mathcalx{T}^{\smash{\J}}_{\Varec{x}}(\Val_1)$.
+and $\MU {\Varec{x}}.\; t$ is not $\alpha$-equivalent to a term in $\Terms^{\smash{\J}}_{\Varec{x}}(\Val_1)$.
 To show 2,
@@
 Since $\interp{\Val_1 \ec{y}}{t_1}$ is normal,
-there is a closed $\mu$-term $v \in \mathcalx{T}^{\smash{\J}}_{\Varec{x}}( \Val_1 )$
+there is a closed $\mu$-term $v \in \Terms^{\smash{\J}}_{\Varec{x}}( \Val_1 )$
 such that $v \vsimv{\Varec{x}} \interp{\Val_1 \ec{y}}{t_1}$.
@@
 and that interprets the variables of $\Fc$ based on the completion $\ValC$ of the mapping $\Val$ from phase~2.
-For the latter, let $\M( x )$ be the value in $\M$ that is $\alpha$-equivalent with $\ValC \ec{x}$ for each variable $x \in \mathcalx{T}(\Fc)$,
+For the latter, let $\M( x )$ be the value in $\M$ that is $\alpha$-equivalent with $\ValC \ec{x}$ for each variable $x \in \Terms(\Fc)$,
 which by Lemma~\ref{lem:interpretation-completion}\PART{1} is guaranteed to exist.
@@
 by structural induction on $t$,
-that $\M( t ) \aequiv \ValC \ec{t}$ for all terms $t \in \mathcalx{T}( \Fc )$,
+that $\M( t ) \aequiv \ValC \ec{t}$ for all terms $t \in \Terms( \Fc )$,
 which since $\M$ is normal, implies $\M \models t_1 \teq t_2$.

### 2015-02-18 tex/body.tex
renamed vars

@@
 \newcommand{\M}{\mathcalx{M}}
-\newcommand{\Val}{\mathcalx{A}}
+\newcommand{\Val}{\mathcalx{A}\vvthinspace}
 \newcommand{\ValC}{\mathcalx{A}^\star}
@@
 To construct a completion $\ValC$, we must choose values for the free
-variables of $\Val$, namely ${\Varec{z}}$ and ${\Varec{y}}$.
-Modulo $\alpha$-equivalence, the set \smash{$\Terms^{\smash{\J}}_{{\Varec{z}}}( \Val )$} is
-$%\begin{array}{c}
+variables of $\Val$, namely ${\Varec{y}}$ and ${\Varec{z}}$.
+Modulo $\alpha$-equivalence, $\smash{\Terms^{\smash{\J}}_{{\Varec{z}}}( \Val )} =
+%\begin{array}{c}
 \{
-\MU z.\; \const{D}( z ),\;
-\MU z.\; \const{C}( \const{D}( z )),\allowbreak\;
-\const{C}( \MU z.\; \const{D}( z )),\;
-\MU z.\; \const{C}( z )
+\MU a.\: \const{D}( a ),\:
+\MU a.\: \const{C}( \const{D}( a )),\allowbreak\:
+\const{C}( \MU a.\: \const{D}( a )),\allowbreak\:
+\MU a.\: \const{C}( a )
 \}
@@
 $.
-Consider a normal interpretation $\J$ that interprets variables in $\Ec$ based on the mapping $\Val$:
+Consider a normal interpretation $\J$ that evaluates variables in $\Ec$ based on the mapping $\Val$:
 $z$ is interpreted as $\Val \ec{z}$, $y$ as $\Val \ec{y}$, and so on.
@@
 For example, assign $\MU {\Varec{z}}.\; \const{D}( {\Varec{z}} )$ for ${\Varec{z}}$.
-After this substitution, $\Val \ec{v}$ is $\MU {\Varec{v}}.\; \const{D}( \MU {\Varec{z}}.\; \const{D}( {\Varec{z}} ) )$,
+After the substitution, $\Val \ec{v} = \MU {\Varec{v}}.\; \const{D}( \MU {\Varec{z}}.\; \const{D}( {\Varec{z}} ) )$,
 which has normal form $\MU {\Varec{v}}.\; \const{D}( {\Varec{v}} )$,

### 2015-02-18 tex/body.tex
minor

@@
 Legal substitutions for ${\Varec{z}}$ may cause the range of $\Val$ to contain abnormal terms.
-For example, after assigning $\MU {\Varec{z}}.\; \const{D}( \const{C}( {\Varec{z}} ) )$ to $\Varec{z}$, we have that
-$\Val \ec{u}$ is $\MU {\Varec{u}}.\; \const{C}( \MU {\Varec{z}}.\; \const{D}( \const{C}( {\Varec{z}} ) ) )$,
-which has normal form
+For example, after assigning $\MU {\Varec{z}}.\; \const{D}( \const{C}( {\Varec{z}} ) )$ to $\Varec{z}$, we have
+$\Val \ec{u} = \MU {\Varec{u}}.\; \const{C}( \MU {\Varec{z}}.\; \const{D}( \const{C}( {\Varec{z}} ) ) )$,
+with normal form
 $\MU {\Varec{u}}.\; \const{C}( \MU {\Varec{z}}.\; \const{D}( {\Varec{u}} ) )$.

### 2015-02-18 tex/body.tex
uniformized style

@@
 \begin{proof}
-To show 1, we first show that $\ValC$ contains no free variables.
+To show (1), we first show that $\ValC$ contains no free variables.
 Assume by contradiction that $\ValC$ contained a free variable ${\Varec{y}}$ for some $\ec{y}$ of type $\tau$.
@@
 If $\tau \in \Codata$, then $\tau$ is corecursive by assumption, and by Lemma~\ref{lem:corecursive-singletons},
-the cardinality of $\tau$ must be $1$.
+the cardinality of $\tau$ must be one.
 Since \rn{Single} does not apply to $\Ec$,
@@
 
-We first show that 2 and 3 hold initially for $\Val$.
+We first show that (2) and (3) hold initially for $\Val$.
 For all equivalence classes $\ec{z}$,
@@
 Thus, we have
-$\interp{\Val \ec{y}}{ t } = \interp{\Val \ec{x}}{ \Val \ec{x} } = \Val \ec{x}$, and thus 2 holds for $\Val$.
-Statement 3 holds for $\Val$ since \rn{Unique} does not apply.
+$\interp{\Val \ec{y}}{ t } = \interp{\Val \ec{x}}{ \Val \ec{x} } = \Val \ec{x}$, and thus (2) holds for $\Val$.
+Property (3) holds for $\Val$ since \rn{Unique} does not apply.
 
@@
 and $\MU {\Varec{x}}.\; t$ is not $\alpha$-equivalent to a term in $\Terms^{\smash{\J}}_{\Varec{x}}(\Val_1)$.
-To show 2,
-by assumption of 2 on $\Val_1$, we have $\interp{\Val_1 \ec{y}}{ u } = \Val_1 \ec{x}$ for
+To show (2),
+by assumption of (2) on $\Val_1$, we have $\interp{\Val_1 \ec{y}}{ u } = \Val_1 \ec{x}$ for
 all subterms $u$ of $\Val_1 \ec{y}$ where $\muvar(t) = {\Varec{x}}$.
@@
 $\interp{\nf{\Val_1 \sigma} \ec{y}}{ u } = \nf{\Val_1 \sigma} \ec{x}$.
-To show 3,
+To show (3),
 consider two distinct equivalence classes $\ec{y}$ and $\ec{z}$,

### 2015-02-18 tex/body.tex
minor

@@
 of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$
-where for each $i \in \{1,\ldots, n\}$, % \in [n]$,
-$\muvar(w_i) = {\Varec{u_i}}$.  Thus by Lemma~\ref{lem:interpretation-completion}\PART{2},
+where
+$\muvar(w_i) = {\Varec{u_i}}$ for each $i$.  Thus by Lemma~\ref{lem:interpretation-completion}\PART{2},
 $\interp{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$.
-For each $i \in \{1,\ldots, n\}$, let $u_\iPrime$ be the $i$\vvthinspace th argument of $t_u$.
+For each $i$, let $u_\iPrime$ be the $i$\vvthinspace th argument of $t_u$.
 \begin{rep}By Lemma~\ref{lem:mu-norm-arg},\end{rep}%
@@
 $\MU {\Varec{u}}.\; \const C_{\negvthinspace\jPrime}( w_1, \ldots, w_n )$,
-where for each $i \in \{1,\ldots, n\}$,
-$\muvar(w_i) = \smash{\Varec{\vphantom{I}\smash{\const s^i_{\negvthinspace\jPrime}( u )}}}$ and thus by Lemma~\ref{lem:interpretation-completion}\PART{2},
+where
+$\muvar(w_i) = \smash{\Varec{\vphantom{I}\smash{\const s^i_{\negvthinspace\jPrime}( u )}}}$
+for each $i$ and thus by Lemma~\ref{lem:interpretation-completion}\PART{2},
 $\interp{\ValC \ec{u}}{ w_i } = \ValC \ec{ \smash{\const s^i_{\negvthinspace\jPrime}( u )}}$.

### 2015-02-19 tex/body.tex
minor edits

@@
 
-The mapping $\Val$, on which the rules depends, is computed as follows.
+The mapping $\Val$, on which the rules depend, is computed as follows.
 %
@@
 \]
-$\FV( \Val )$ denotes the union of the free variables occurring in $\Val$'s range,
+$\FV( \Val )$ denotes the set of free variables occurring in $\Val$'s range,
 and $\Val[x \mapsto t]$ denotes the \emph{variable-capturing} substitution of $t$ for
@@
 
-Given the mapping $\Val$, the \rn{Acyclic} and \rn{Unique} rules operate as follows.
+Given the mapping $\Val$, the \rn{Acyclic} and \rn{Unique} rules work as follows.
 For acyclicity, if $\ec{t}$ is a datatype equivalence class
@@
 %for a capture-avoiding renaming of $\mu$-bound variables,
-then $t \teq u$. This requirement may seem too restrictive, since
+then $t \teq u$. Comparison for $\alpha$-equivalence may seem too restrictive, since
 $\MU x.\; \const{S}(x)$ and $\MU y.\; \const{S}(\const{S}(y))$ specify the same
 value despite being $\alpha$-disequivalent, but the rule will
-make some progress by discovering that the subterm $\const{S}(y)$ of $\MU y.\;
-\const{S}(\const{S}(y))$ must be equal to the entire term,
-as demonstrated next.
+make progress by discovering that the subterm $\const{S}(y)$ of $\MU y.\;
+\const{S}(\const{S}(y))$ must be equal to the entire term\begin{rep}
+as demonstrated next\end{rep}.
 

### 2015-02-19 tex/body.tex
more

@@
 incrementally builds partial assignments from the atomic formulas of $F$ to
-truth values such that no clause in $F$ is falsified. We can interpret such a
+truth values such that no clause in $F$ is falsified. We can regard such a
 partial assignment as a set $M$ of literals where $a \in M$ for all atoms $a$ that
@@
 \cite{jovanovic2011sharing} of the Nelson--Oppen method,
-each $T_i$-solver for $i = 1, \ldots, l$ takes as input a combination
+each $T_i$-solver %for $i = 1, \ldots, l$  n?
+takes as input a combination
 of (1) the purified form of $T_i$-constraints occurring in $M$, where terms
-containing symbols not belonging to $T_i$ are replaced by fresh variables \begin{rep}of the
+containing symbols not belonging to $T_i$ are replaced by fresh variables\begin{rep} of the
 same type\end{rep}; (2) additional equalities and disequalities between variables of
@@
 $\neg\; C$ is added to $F\!$, adds a clause to $F\!$, or does nothing.
-When $M$ is a complete assignment for $F\!$, a theory solver can only choose to do
-nothing when $M_i$ is indeed $T_i$-satisfiable.
+When $M$ is a complete assignment for $F\!$, a theory solver can choose to do
+nothing only if $M_i$ is indeed $T_i$-satisfiable.
 

### 2015-02-19 tex/body.tex
more

@@
 truth values such that no clause in $F$ is falsified. We can regard such a
-partial assignment as a set $M$ of literals where $a \in M$ for all atoms $a$ that
-it assigned the value $\top$ and $\neg\; a \in M$ for all atoms that it assigns
-the value $\bot$. Then, %at a high level,
+partial assignment as a set $M$ of true literals.
+% where $a \in M$ for all atoms $a$ that
+%it assigned the value $\top$ and $\neg\; a \in M$ for all atoms that it assigns
+%the value $\bot$.
+Then, %at a high level,
 by a variant
@@
 takes as input a combination
+%%% TODO, @ANDY: "a combination" -> "the union"?
 of (1) the purified form of $T_i$-constraints occurring in $M$, where terms
@@
 same type\end{rep}; (2) additional equalities and disequalities between variables of
-types not belonging to $T_i$. Let us call this set $M_i$. Each $T_i$ solver
+types not belonging to $T_i$. Let us call this set $M_i$.
+%%% TODO, @ANDY: which set M_i?
+Each $T_i$ solver
 either reports that a subset $C$ of $M_i$ is $T_i$-unsatisfiable, in which case
@@
 
-Assume $\Ec$ is initially the set $M_i$ described above. With each equality $t \teq s$ added
+Assume $\Ec$ is initially the set $M_i$ described above. With each equality $t \teq u$ added
 to $\Ec$, we associate a set of equalities from $M_i$ that together entail $t
-\teq s$, which we call its \emph{explanation}.
+\teq u$, which we call its \emph{explanation}.
 Similarly, each $\Val \ec{x}$ is assigned an explanation---that is, a set
@@
 explanation for $\Val \ec{x} = \MU {\Varec{x}}.\; \const{C}( {\Varec{x}} )$.
-Thus, the rules of the calculus are implemented as follows. For all rules with
+The rules of the calculus are implemented as follows. For all rules with
 conclusion $\bot$, we report the union of the explanations for all premises is
@@
 It visits $\ec{t_1}, \ldots, \ec{t_n}$ and all arguments of constructors in these equivalence classes recursively.
-If while doing so it returns to $\ec{t}$, then it deduces that \rn{Acyclic} applies to $t$.
+If while doing so it returns to $\ec{t}$, it deduces that \rn{Acyclic} applies. % to $t$.
 To recognize when the precondition of \rn{Unique} holds,
@@
 While the decision procedure is restricted to ground constraints,
-in practice we are often interested in problems that combine reasoning about universal axioms and existential conjectures,
+in practice we often want to solve problems that combine reasoning about universal axioms and existential conjectures,
 where checking satisfiability requires reasoning about such ground constraints in the presence of universally quantified formulas.
@@
 
-However, the presence of quantified formulas poses an additional challenge in the context of (co)datatypes.
-Namely, since quantified formulas may entail cardinality constraints on uninterpreted types $\ty{u}$,
+However, the presence of quantified formulas poses an additional challenge in the context of (co)datatypes:
+Since quantified formulas may entail cardinality constraints on uninterpreted types $\ty{u}$,
 we may no longer assume that they are of infinite cardinality.
@@
 due to the overly restrictive preconditions on \rn{Split} and \rn{Single}.
-This does not necessarily impact the correctness of our procedure in this setting,
-since by assumption the solver is already incomplete when quantified formulas are present.
-Regardless, we can increase the precision of the solver by doing the following.
+This does not %necessarily
+impact the correctness of our procedure in this setting,
+since %by assumption
+the solver is already incomplete when quantified formulas are present.
+Nonetheless, two techniques help increase the precision of the solver.
 First, we can apply \rn{Split} to datatype terms whose cardinality depends on the finiteness of uninterpreted types.
@@
 The implementation of the decision procedure uses several optimizations which
-follow the approach outlined by Barrett et al.\cite{barrett-et-al-2007}
-and are not reflected in the presentation of the calculus.
+follow the approach outlined by Barrett et al. \cite{barrett-et-al-2007}.
+%and are not reflected in the presentation of the calculus.
 We briefly mention the main ones.
 %
-Discriminators are treated as predicate symbols, so a discriminator $\const{d}_{\negvthinspace j}( t )$ is such that $\const{d}_{\negvthinspace j}$
-is a predicate symbol, instead of reducing this constraint to $t \teq \const{C}_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^n_{\negvthinspace j}( t ) \bigr)$.
-We have found that this leads to better performance because this reduction introduces terms more eagerly to $\tEc$.
+Discriminators are part of the signature instead
+of being abbreviations.
+We have found that this leads to better performance because the alternative introduces terms more eagerly to $\tEc$.
 Handling discriminators requires extending the decision procedure with several
@@
 \const{C}_{\negvthinspace j}( u_1, \ldots, u_n )$, we immediately add
-$\smash{\const s^k_{\negvthinspace j}( t ) \teq u_k}$ to $\Ec$\begin{rep}, whereas the presented calculus
-would apply \rn{Split} and \rn{Inject} before adding this equality\end{rep}. 
+$\smash{\const s^k_{\negvthinspace j}( t ) \teq u_k}$ to $\Ec$, whereas the presented calculus
+would apply \rn{Split} and \rn{Inject} before adding this equality. 
 \begin{rep}  % lower priority to mention this

### 2015-02-19 tex/body.tex
more

@@
 goals), about Bird and Stern--Brocot trees. To exercise the support
-for datatypes, the benchmarks are complemented by theories about lists
+for datatypes, the benchmarks were complemented by theories about lists
 and trees. The theories were selected before running the
@@
 No (co)datatypes
-  & 209 & 221 & 777 & 775 & \phantom{0}51 & \phantom{0}52 & 1037 & 1048 \\%[\jot]
+  & 209 & 221 & 777 & 775 & 51 & 52 & 1037 & 1048 \\%[\jot]
 %Datatype simplication only
@@
 Datatypes w/o \rn{Acyclic}
-  & & 227 & & 780 & & \phantom{0}52 & & 1059 \\
+  & -- & 227 & -- & 780 & -- & 52 & -- & 1059 \\
 Full datatypes
-  & 213 & 227 & 791 & 786 & \phantom{0}51 & \phantom{0}52 & 1055 & 1065 \\%[\jot]
+  & 213 & 227 & 791 & 786 & 51 & 52 & 1055 & 1065 \\%[\jot]
 %Codatatype simplication only
@@
 Codatatypes w/o \rn{Unique}
-  & & 222 & & 804 & & \phantom{0}56 & & 1082 \\
+  & -- & 222 & -- & 804 & -- & 56 & -- & 1082 \\
 Full codatatypes
-  & & 223 & & 804 & & \bfseries \phantom{0}59 & & 1086 \\%[\jot]
+  & -- & 223 & -- & 804 & -- & \bfseries 59 & -- & 1086 \\%[\jot]
 Full (co)datatypes
-  & & \bfseries 229 & & \bfseries 815 & & \bfseries \phantom{0}59 & & \bfseries 1103% \\[\jot]
+  & -- & \bfseries 229 & -- & \bfseries 815 & -- & \bfseries 59 & -- & \bfseries 1103% \\[\jot]
 \end{tabular}\end{center}

### 2015-02-19 tex/body.tex
more

@@
 problems involving datatypes and codatatypes in
-\begin{conf}\goodbreak\noindent\end{conf}%  %%% TYPESETTING: terrible hack
+\begin{conf}\goodbreak\noindent\end{conf}%  %%% TYPESETTING: terrible HACK
 combination (Section~\ref{sec:the-ground-decision-procedure}).
@@
 %
-In addition, nothing forbids non(co)recursive specifications
+In addition, non(co)recursive specifications
 such~as
 \begin{conf}%
-\vthinspace$\keyw{datatype}~\vthinspace\ty{option}_{\,\tau} \,=\, \const{None} \,\mid\, \const{Some}(\tau)$.%
+\vthinspace$\keyw{datatype}~\vthinspace\ty{option}_{\,\tau} \,=\, \const{None} \,\mid\, \const{Some}(\tau)$
 \end{conf}%
@@
 \]
-At the semantic level, it makes no difference whether such types are
-introduced as datatypes or as codatatypes.
+\end{rep}%
+are permitted.
+\begin{rep}At the semantic level, it makes no difference whether
+such types are introduced as datatypes or as codatatypes.
 \end{rep}%
@@
 
-\kern-1.5pt %% TYPESETTING hack
-
-\paragraph{Phase 1: Computing the Bidirectional Closure.}
+\paragraph{Phase 1: Computing the Bidirectional Closure \rm(\figurename~\ref{fig:cc-rules}).}
 In conjunction with \rn{Refl}, \rn{Sym}, and \rn{Trans}, the \rn{Cong} rule computes the (upward) congruence closure,
@@
 
+\paragraph{Phase 2: Applying Acyclicity and Uniqueness \rm(\figurename~\ref{fig:ab-rules}).}
+%For presentation of the rules in this phase, we rely on a representation of terms in the $\mu$-notation.
+The premises of the rules that make up the second phase refer to a mapping $\Val$
+that associates a $\mu$-term with each equivalence class as its representative.
+%Recall that $\mu$-bindings can be used for representing possibly cyclic terms and values.
+%(Section~\ref{sec:introduction}).
+% for example, $\MU x.\; \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
+A $\mu$-term $t\typ{\tau}$ describes a class of $\tau$ values
+that terms in that equivalence class can take in the models of $\Ec$.
+%either in the case when $\tau$ is a codatatype or datatype type,
+When $\tau$ is a datatype, % \in \Data$,
+a cyclic $\mu$-term describes an infeasible class of values.
+%$\tau \not\in \Codata$.
+
+Formally, $\mu$-\emph{terms} are defined recursively as being either a variable $x$
+or a node of the form
+$\MU x.\: \const{C}( \vec t\, )$ for some constructor $\const{C} \in \Ctr$ and
+$\mu$-terms $\vec t$ of the expected types.
+The variable $x$ need not occur freely in the $\mu$-binder's body, in which case
+the binder can be omitted.
+In the \rn{Acyclic} rule,
+$\FV( u )$ denotes the variables occurring freely in~$u$.
+%%% But there are no constants of type $\tau \in \Nondata$ -- at most variables!
+%
+%For uniformity, we also consider $%\MU x.\;
+%\const{c}$ to be a $\mu$-term if $\const{c}$ is a constant of type $\tau \in \Nondata$.
+%
+%%% I think this should be clear enough from the "Conventions" paragraph at the
+%%% end of the introduction.
+%%
+% which we define recursively:
+%if $\const{C} : \vec \tau \rightarrow \upsilon$
+%and $\vec v$ are well-typed $\mu$-terms of type $\vec \tau$ under the assumption that $x$ has type $\upsilon$,
+%then $\MU x.\; \const{C}( \vec v )$ is a well-typed term of type $\upsilon$.
+%
+A $\mu$-term is \emph{closed} if %and only if --- by convention, "and only if" is not necessary for *definitions* (but add it back if you disagree)
+it contains no free variables. \,It is \emph{cyclic} if %and only if --- ditto
+it contains a bound variable. \,The
+\goodbreak %%% TYPESETTING: horrible hacks (including two \,'s)
+\noindent
+$\alpha$-\emph{equivalence} relation $t \aequiv u$
+indicates that the $\mu$-terms $t$ and $u$
+are syntactically equivalent for some capture-avoiding renaming of $\mu$-bound variables\begin{rep}---e.g.,
+%$\MU x.\; \const{C}( x ) \aequiv \MU y.\; \const{C}( y )$ and
+$\MU x.\; \const{D}( y, x ) \aequiv \MU z.\; \const{D}( y, z )$,
+but
+$\MU x.\; \const{C}( x ) \not\aequiv \MU x.\; \const{D}( y, x ) \not\aequiv \MU x.\; \const{D}( z, x )
+\not\aequiv \MU y.\; \const{D}( y, x )$\end{rep}.
+Two $\mu$-terms can denote the same value despite being $\alpha$-disequivalent---e.g.,
+$\MU x.\; \const{C}( x ) \not\aequiv \MU y.\; \const{C}( \const{C}( y ) )$.
+%, or informally are \emph{observationally equivalent}.
+%For convienience, we use $\mu$-terms to refer to (classes of) values for both codatatype and datatype terms,
+%where in the latter case, a $\mu$-term with a bound variable denotes an infeasible term.
+%where the latter case adds the restriction on $\mu$-terms $t$ that no subterms of $t$ contain are bound variables.
+
 \begin{figure}[t!]
-\vspace*{+3pt} %% TYPESETTING HACK
+%\vspace*{+3pt}
 \normalsize
@@
 \end{figure}
-
-\kern-1.5pt %% TYPESETTING hack
-
-\paragraph{Phase 2: Applying Acyclicity and Uniqueness.}
-%For presentation of the rules in this phase, we rely on a representation of terms in the $\mu$-notation.
-The premises of the rules that make up the second phase refer to a mapping $\Val$
-that associates a $\mu$-term with each equivalence class as its representative.
-%Recall that $\mu$-bindings can be used for representing possibly cyclic terms and values.
-%(Section~\ref{sec:introduction}).
-% for example, $\MU x.\; \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
-A $\mu$-term $t\typ{\tau}$ describes a class of $\tau$ values
-that terms in that equivalence class can take in the models of $\Ec$.
-%either in the case when $\tau$ is a codatatype or datatype type,
-When $\tau$ is a datatype, % \in \Data$,
-a cyclic $\mu$-term describes an infeasible class of values.
-%$\tau \not\in \Codata$.
-
-Formally, $\mu$-\emph{terms} are defined recursively as being either a variable $x$
-or a node of the form
-$\MU x.\: \const{C}( \vec t\, )$ for some constructor $\const{C} \in \Ctr$ and
-$\mu$-terms $\vec t$ of the expected types.
-The variable $x$ need not occur freely in the $\mu$-binder's body, in which case
-the binder can be omitted.
-In the \rn{Acyclic} rule,
-$\FV( u )$ denotes the variables occurring freely in~$u$.
-%%% But there are no constants of type $\tau \in \Nondata$ -- at most variables!
-%
-%For uniformity, we also consider $%\MU x.\;
-%\const{c}$ to be a $\mu$-term if $\const{c}$ is a constant of type $\tau \in \Nondata$.
-%
-%%% I think this should be clear enough from the "Conventions" paragraph at the
-%%% end of the introduction.
-%%
-% which we define recursively:
-%if $\const{C} : \vec \tau \rightarrow \upsilon$
-%and $\vec v$ are well-typed $\mu$-terms of type $\vec \tau$ under the assumption that $x$ has type $\upsilon$,
-%then $\MU x.\; \const{C}( \vec v )$ is a well-typed term of type $\upsilon$.
-%
-A $\mu$-term is \emph{closed} if %and only if --- by convention, "and only if" is not necessary for *definitions* (but add it back if you disagree)
-it contains no free variables. \,It is \emph{cyclic} if %and only if --- ditto
-it contains a bound variable. \,The
-\goodbreak %%% TYPESETTING: horrible hacks (including two \,'s)
-\noindent
-$\alpha$-\emph{equivalence} relation $t \aequiv u$
-indicates that the $\mu$-terms $t$ and $u$
-are syntactically equivalent for some capture-avoiding renaming of $\mu$-bound variables\begin{rep}---e.g.,
-%$\MU x.\; \const{C}( x ) \aequiv \MU y.\; \const{C}( y )$ and
-$\MU x.\; \const{D}( y, x ) \aequiv \MU z.\; \const{D}( y, z )$,
-but
-$\MU x.\; \const{C}( x ) \not\aequiv \MU x.\; \const{D}( y, x ) \not\aequiv \MU x.\; \const{D}( z, x )
-\not\aequiv \MU y.\; \const{D}( y, x )$\end{rep}.
-Two $\mu$-terms can denote the same value despite being $\alpha$-disequivalent---e.g.,
-$\MU x.\; \const{C}( x ) \not\aequiv \MU y.\; \const{C}( \const{C}( y ) )$.
-%, or informally are \emph{observationally equivalent}.
-%For convienience, we use $\mu$-terms to refer to (classes of) values for both codatatype and datatype terms,
-%where in the latter case, a $\mu$-term with a bound variable denotes an infeasible term.
-%where the latter case adds the restriction on $\mu$-terms $t$ that no subterms of $t$ contain are bound variables.
 
@@
 
-\paragraph{Phase 3: Branching.}
+\paragraph{Phase 3: Branching \rm(\figurename~\ref{fig:split-rule}).}
 If a selector is applied to a term $t$, or if $t$'s type is a finite datatype,

### 2015-02-19 tex/body.tex
more

@@
 % for example, $\MU x.\; \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
-A $\mu$-term $t\typ{\tau}$ describes a class of $\tau$ values
-that terms in that equivalence class can take in the models of $\Ec$.
+The $\mu$-term $\Val[t\typ{\tau}]$ describes a class of $\tau$ values
+that $t$ and other members of $t$'s equivalence class can take in models of $\Ec$.
 %either in the case when $\tau$ is a codatatype or datatype type,
@@
 A companion rule, \rn{Single}, focuses on the degenerate case where two
-terms from $\tEc$ have a corecursive singleton type
+terms %from $\tEc$
+have a corecursive singleton type
 %(Section~\ref{sec:the-theory-of-co-datatypes})
@@
 based on a recursive computation of the cardinalities of the
-constructors' argument types. This step relies on the assumption that all types
-in $\Nondata$ are infinite.
+constructors' argument types. This step relies on the assumption that
+ordinary types are infinite.
 
@@
 When constructing a model $\M$ for $\Ec$,
-it remains only to specify how $\M$ interprets wrongly applied selector terms and variables.%
+it remains only to specify how $\M$ interprets wrongly applied selector terms and variables.
 For the latter, this will be based on the mapping $\Val$ computed in phase~2 of the calculus.
@@
 To construct a completion $\ValC$, we must choose values for the free
-variables of $\Val$, namely ${\Varec{y}}$ and ${\Varec{z}}$.
+variables of $\Val$, specifically ${\Varec{y}}$ and ${\Varec{z}}$.
 Modulo $\alpha$-equivalence, $\smash{\Terms^{\smash{\J}}_{{\Varec{z}}}( \Val )} =
@@
 values in the range of $\ValC$ are $\alpha$-equivalent to a value in normal
-interpretation. In particular, this means they are closed, normal, and acyclic when
+interpretation. This means they are closed, normal, and acyclic when
 required. Property~\PART{2} states that the interpretation of all subterms in the
@@
 interpretation of a subterm $t$ where $\muvar(t) = {\Varec{x}}$ is equal to
-$\ValC \ec{x}$, independently of the context in which it is interpreted.
+$\ValC \ec{x}$, independently of the context. % in which it is interpreted.
 Property~\PART{3} states that $\ValC$ is injective (modulo
-$\alpha$-equivalence), which in the following ensures that a distinct value can
-be associated with each equivalence class.
+$\alpha$-equivalence), which ensures %in the following
+that distinct values are
+assigned to distinct equivalence classes.
 
@@
 While the decision procedure is restricted to ground constraints,
-in practice we often want to solve problems that combine reasoning about universal axioms and existential conjectures,
+in practice we often want to solve problems that feature universal axioms and existential conjectures,
 where checking satisfiability requires reasoning about such ground constraints in the presence of universally quantified formulas.
-Many SMT solvers, including CVC4, support reasoning about quantified formulas using incomplete instantiation-based methods~\cite{MouraBjoerner07,ReynoldsTinelliMoura14}.
-These methods extend naturally to quantified formulas over the signature of the theory of datatypes and codatatypes.
+Many SMT solvers, including CVC4, can reason about quantified formulas using incomplete instantiation-based methods~\cite{MouraBjoerner07,ReynoldsTinelliMoura14}.
+These methods extend naturally to quantified formulas involving datatypes and codatatypes.
 
@@
 Isabelle formalizations using Sledgehammer.
-Codatatypes being a recent addition to Isabelle
-\cite{blanchette-et-al-2014-impl}, benchmarks are not abundant. We
+There are not yet many benchmarks available, since codatatypes are a
+recent addition to Isabelle \cite{blanchette-et-al-2014-impl}. We
 included all the theory files from the Isabelle distribution (Distro, 879 goals)
 and the \emph{Archive of Formal Proofs} (AFP, 2974 goals) \cite{klein-et-al-afp}
-that define codatatypes falling within the supported fragment. We also included
+that define codatatypes falling within the supported fragment. We added
 two unpublished theories by Peter Gammie and Andreas Lochbihler (\gandl, 317
@@
 given to CVC4, running for up to 60 seconds on the StarExec cluster
-\cite{stump-et-al-2014-starexec}. Problems involving no (co)datatypes were
-filtered out. Due to the lack of machinery for reconstructing inferences about
-(co)datatypes in Isabelle, the external solver is trusted as an oracle in these
-experiments.
+\cite{stump-et-al-2014-starexec}.
+%TODO: Keep or not? Problems not involving (co)datatypes were filtered out.
+Due to the lack of machinery for reconstructing inferences about (co)datatypes
+in Isabelle, the solver is trusted as an oracle in these experiments.
 
@@
 Cesare Tinelli, and Christoph Weidenbach, have either encouraged the research on
-codatatype or at least benevolently tolerated it, both of which we are thankful
+codatatypes or at least benevolently tolerated it, both of which we are thankful
 for.
@@
 %
-Andreas Lochbihler suggested several textual improvements.
+Andreas Lochbihler and Mark Summerfield suggested several textual improvements.
 The second author's research was partially supported by the Deutsche

### 2015-02-19 tex/body.tex
more

@@
 each $T_i$-solver %for $i = 1, \ldots, l$  n?
-takes as input a combination
-%%% TODO, @ANDY: "a combination" -> "the union"?
+takes as input the union
 of (1) the purified form of $T_i$-constraints occurring in $M$, where terms

### 2015-02-19 tex/body.tex
dropped a few "our"s

@@
 For \rn{Inject}, this is a consequence of injectivity.
-For \rn{Unique}, our construction of $\Val$ indicates that the values of $t$ and $u$ are equivalent in all models of $\Ec$.
+For \rn{Unique}, the construction of $\Val$ indicates that the values of $t$ and $u$ are equivalent in all models of $\Ec$.
 For \rn{Single}, $t$~and~$u$ must have the same value since the cardinality of their type is $1$.
@@
 Since \rn{Acyclic} does not apply, $\Val \ec{x}$ is acyclic when $\tau \in \Data$.
-Moreover, our construction of $\ValC$ applies substitutions of the form
+Moreover, the construction of $\ValC$ applies substitutions of the form
 $\{ {\Varec{y}} \mapsto t \}$, where $t$ is acyclic when the type of ${\Varec{y}}$ is not a codatatype.
@@
 We have that $\Val \ec{x}$ is of the form $\MU {\Varec{x}}.\; \const{C}( t_1, \ldots, t_n )$.
-Due to our construction of $\Val$,
+Due to the construction of $\Val$,
 we know $\ec{x}$ contains a constructor $\const{C}( z_1, \ldots, z_n )$ and $\muvar( t_i ) = z_i$ for some $i$ % \in [n]$.
@@
 where $\interp{\Val \ec{x}}{ t_i } \aequiv \interp{\Val \ec{x} }{ u_i }$ for some $i$. % \in [n]$.
-Due to our construction of $\Val$, $\ec{y}$ contains a constructor $\const{C}( w_1, \ldots, w_n )$ and $\muvar( u_i ) = w_i$ for some $i$. % \in [n]$.
+Due to the construction of $\Val$, $\ec{y}$ contains a constructor $\const{C}( w_1, \ldots, w_n )$ and $\muvar( u_i ) = w_i$ for some $i$. % \in [n]$.
 Since \rn{Cong} does not apply to $\Ec$
@@
 such that $v \vsimv{\Varec{x}} \interp{\Val_1 \ec{y}}{t_1}$.
-Thus, by assumption on our selection of $\MU {\Varec{x}}.\; t$,
+Thus, by assumption on the selection of $\MU {\Varec{x}}.\; t$,
 we have $\MU {\Varec{x}}.\; t \not\vsimv{\Varec{x}} \interp{\Val_1 \ec{y}}{t_1}$,
@@
 Since \rn{Inject} and \rn{Clash} do not apply to $\Fc$,
-by our construction of $\ValC$ we have $\ValC \ec{t}$ is a term
+by the construction of $\ValC$ we have $\ValC \ec{t}$ is a term
 of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$
@@
 each $T_i$-solver %for $i = 1, \ldots, l$  n?
-takes as input the union
+takes as input the union $M_i$
 of (1) the purified form of $T_i$-constraints occurring in $M$, where terms
@@
 same type\end{rep}; (2) additional equalities and disequalities between variables of
-types not belonging to $T_i$. Let us call this set $M_i$.
-%%% TODO, @ANDY: which set M_i?
+types not belonging to $T_i$.
 Each $T_i$ solver
@@
 This does not %necessarily
-impact the correctness of our procedure in this setting,
+impact the correctness of the procedure in this setting,
 since %by assumption
@@
 If there exist two equivalence classes $\ec{s}$ and $\ec{t}$ for this type,
-our implementation adds the clause %of the form
+the implementation adds the clause %of the form
 $\exists x\;y\typ{\ty{u}}.\;\, x \not\teq y \mathrel\vee s \teq t$ to $F$.

### 2015-02-19 tex/body.tex
more

@@
 takes as input the union $M_i$
-of (1) the purified form of $T_i$-constraints occurring in $M$, where terms
+of (1)~the purified form of $T_i$-constraints occurring in $M$, where terms
 containing symbols not belonging to $T_i$ are replaced by fresh variables\begin{rep} of the
-same type\end{rep}; (2) additional equalities and disequalities between variables of
+same type\end{rep}; (2)~additional equalities and disequalities between variables of
 types not belonging to $T_i$.

### 2015-02-19 tex/body.tex
more

@@
 Instead, $\Val$ is computed implicitly by traversing the equivalence classes of $\Ec$ during phase~2.
-To recognize when the precondition of \rn{Acyclic} holds,
+To detect whether \rn{Acyclic} applies,
 the procedure considers each equivalence class $\ec{t}$ containing a datatype constructor $\const{C}( t_1, \ldots, t_n )$.
@@
 %\paragraph{Optimizations.}
-The implementation of the decision procedure uses several optimizations which
-follow the approach outlined by Barrett et al. \cite{barrett-et-al-2007}.
+The implementation of the decision procedure uses several optimizations
+following the lines of Barrett et al. \cite{barrett-et-al-2007}.
 %and are not reflected in the presentation of the calculus.

### 2015-02-19 tex/body.tex
more

@@
 An evaluation on %hand-crafted examples and on
-problems generated from Isabelle theories using the Sledgehammer tool \cite{paulson-blanchette-2010}
+problems generated from Isabelle \cite{nipkow-et-al-2002} theories using the Sledgehammer tool \cite{paulson-blanchette-2010}
 demonstrates the usefulness of the approach (Section~\ref{sec:experimental-results}).
@@
 inspired by the $\mu$-calculus
-\begin{conf}\cite{kozen-1983}\end{conf}%
-\begin{rep}\cite[etc.]{kozen-1983,endrullis-et-al-2011}\end{rep}%
-.
+%\begin{conf}\cite{kozen-1983}\end{conf}%
+\cite[etc.]{kozen-1983,endrullis-et-al-2011}.
 

### 2015-02-19 tex/body.tex
more

@@
 %\newcommand\XXXL{DPLL}
-\newcommand\XXXL{CDCL} %% TODO: Convince Andy this is right
+\newcommand\XXXL{CDCL}
 
@@
 \begin{nomemoriam}
-%%% TODO: Unphantom
 \noindent\hfill\emph{\phantom{In memoriam Morgan Deters 1979--2015}}
@@
 \begin{nomemoriam}
-%%% TODO: Unphantom
 \vskip3\smallskipamount
@@
 
-%\textbf{TODO}:
-%  * Klop et al. (streams) Andrew Lice % http://joerg.endrullis.de/
 %  * Ro\c{s}u --- equality of streams is...
@@
 
-%TODO: Andy, introduce theory here. There were some ideas below in the comments
-%you might want to mention either here or in Section 3. I'm not sure what's the
-%best way to justify it though, but for the conference at least we can just claim it. (we
-%won't have the place for a proof anyway).
-
 Given a signature $\Sig$, $\thD$ refers to the \emph{theory of
@@
 applied selector terms but may differ on wrongly applied
-selector terms and variables.
-%%% @ANDY: Variables are normally not part of the interpretation, but given by
-%%% an additional environment. Or is this done differently in the SMT community?
+selector terms and on variables.
 A formula $\varphi$ is $\thD$-\emph{satisfiable} if there
@@
 
-\begin{rep}
-TODO: State more precisely and proof?
-\end{rep}
+%\begin{rep}
+% TODO: State more precisely and prove?
+%\end{rep}
 
@@
 
-If $t$ is a selector term of the form $\const s^k_{\negvthinspace j}( u )$,
+If $t$ is a selector term %of the form
+$\const s^k_{\negvthinspace j}( u )$,
 since \rn{Split} does not apply to $\Fc$,
@@
 Many SMT solvers, including CVC4, can reason about quantified formulas using incomplete instantiation-based methods~\cite{MouraBjoerner07,ReynoldsTinelliMoura14}.
-These methods extend naturally to quantified formulas involving datatypes and codatatypes.
+These methods extend naturally to %quantified
+formulas involving datatypes and codatatypes.
 

### 2015-02-19 tex/body.tex
more

@@
 \makeatother
+
+%\newcommand\DISC[1]{t \teq \const{C}_{#1}\bigl(\const s^1_{#1}( t ), \ldots,\const s^{n_{#1}}_{#1}( t )\bigr)}
+\newcommand\DISC[1]{\const{d}_{#1}(t)}
 
@@
 A signature $\Sig = (\Types, \Funcs)$ consists of a set of types $\Types$ and a
-set of function symbols $\Funcs$. Types are atomic sorts and interpreted by
+set of function symbols $\Funcs$. Types are atomic sorts and interpreted as
 nonempty domains. The set~$\Types$ must contain a
@@
 it is not necessary to extend the signature:
-The discriminator atom $\const{d}_{i\negvthinspace j}(x) \begin{conf}\teq
-\const{true}\end{conf}$ can be seen as an abbreviation for
-$x = \const{C}_{i\negvthinspace j}\bigl(\const{s}_{i\negvthinspace j}^1(x), \ldots, \const{s}_{i\negvthinspace j}^{\,n_{\smash{{i\negvthinspace j}}}}(x)\bigr)$.
+We let $\const{d}_{i\negvthinspace j}(t)$
+abbreviate
+$t \teq \const{C}_{i\negvthinspace j}\bigl(\const{s}_{i\negvthinspace j}^1(x), \ldots, \const{s}_{i\negvthinspace j}^{\,n_{\smash{{i\negvthinspace j}}}}(x)\bigr)$.
 %This will simplify the presentation. % of the decision procedure.
@@
 either infinite or a singleton. In the latter case, $\delta$ necessarily has a
-single constructor, whose arguments have types that are interpreted by
-singletons.
+single constructor, whose arguments have types that are interpreted as
+singleton domains.
 \end{lemma}
@@
 }{
-  \Ec := \Ec,\: t \teq \const{C}_1\bigl(\const s^1_1( t ), \ldots,\const s^{n_{1\!}}_1( t ) \bigr) %\ROR
-  \ROR \cdots \ROR \Ec := \Ec,\: t \teq \const{C}_m\bigl(\const s^1_m( t ), \ldots,\const s^{n_m}_m( t ) \bigr)
+  \Ec := \Ec,\: \DISC{1} %\ROR
+  \ROR \cdots \ROR \Ec := \Ec,\: \DISC{m}
 }
@@
 for each $\const C_{\negvthinspace j} \in \smash{\Ctr^\delta}$.
-Since by exhaustiveness, all models of $\thD$ entail (exactly) one equality $t \teq \const C_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\,\smash{n_{\negvthinspace j}}}_{\negvthinspace j}( t ) \bigr)$,
-this means $\Ec$ is $\thD$-unsatisfiable.
+Since by exhaustiveness, all models of $\thD$ entail (exactly) one
+$\DISC{\negvthinspace j}$,
+$\Ec$ is $\thD$-unsatisfiable.
 Otherwise, the child node of $\Ec$ is a closed derivation tree
@@
 conclusion $\bot$, we report the union of the explanations for all premises is
-$\thD$-unsatisfiable. For \rn{Split}, we add a clause %of the form
-$t \teq \const{C}_1\bigl( \const s^1_1( t ), \ldots, \const s^{n_{1\!}}_1( t ) \bigr)
-\mathrel\lor \cdots \mathrel\lor t \teq \const{C}_m\bigl( \const s^1_m( t ),
-\ldots, \const s^{n_m}_m( t ) \bigr)$ to $F$.
+$\thD$-unsatisfiable. For \rn{Split}, we add the exhaustiveness clause %of the form
+$\DISC{1} \mathrel\lor \cdots \mathrel\lor \DISC{m}$ to $F$.
 Decisions on which branch to take are thus performed externally by the SAT

### 2015-02-19 tex/body.tex
more

@@
 \end{array}\]
+The recursion will eventually terminate because each recursion adds one bound
+variable to $B$ and there are finitely many distinct bound variables in a $\mu$-term.
 %
@@
 %is simply $\const{Z}$ after dropping the $\mu$-binding of $x$.
-
-%\rem{TODO: Normal interpretation vs. actual interpretation.}
 
@@
 
-In the following, we consider only normal interpretations.
+Strictly speaking, normal interpretations are not models of codatatypes,
+because they are restricted to an $\omega$-regular values, whereas models must
+also cater for acyclic infinite terms. However, by monotonicity of ground
+constraints, it is trivial to extend a normal interpretation into a model.
+In the following, we focus on normal interpretations.
+
 When constructing a model $\M$ for $\Ec$,

### 2015-02-19 tex/body.tex
cleaner

@@
 assigns values from $\J$ to unassigned variables in the domain of $\Val$.
-We construct $\ValC$ by exhaustively applying the following rule to $\nf{\Val}$:%
-%\rem{TODO: A bit clumsy}
+We construct $\ValC$ by initially taking $\ValC := \nf{\Val}$
+and by exhaustively applying the following rule:%
 \[
@@
 }{
-  \Val := \nf{\Val [ {\Varec{x}} \mapsto \MU {\Varec{x}}.\; t ]}
+  \ValC := \nf{\ValC [ {\Varec{x}} \mapsto \MU {\Varec{x}}.\; t ]}
 }
@@
 %
-The rule chooses an unassigned variable in $\Val$ and assigns it a fresh
-value---one that does not occur in $\Terms^{\smash{\J}}_{\Varec{x}}(\Val)$
+The rule chooses an unassigned variable in $\ValC$ and assigns it a fresh
+value---one that does not occur in $\Terms^{\smash{\J}}_{\Varec{x}}(\ValC)$
 modulo $\alpha$-equivalence.
@@
 %of codatatypes, of which only finitely many are in
-%$\Terms^{\smash{\J}}_{\Varec{x}}(\Val)$?)}
+%$\Terms^{\smash{\J}}_{\Varec{x}}(\ValC)$?)}
 Since this update removes one
-variable from the set $\FV( \Val )$ and does not add any variables to $\FV( \Val
-)$, it can only be applied a finite number of times. We normalize all terms in
-the range of $\Val$ after each rule application.%\rem{TODO: clumsy}
+variable from $\FV( \ValC )$ and does not add any variables to $\FV( \ValC
+)$, the process eventually terminates. We normalize all terms in
+the range of $\ValC$ at each step.
 

### 2015-02-19 tex/body.tex
more

@@
 assigns values from $\J$ to unassigned variables in the domain of $\Val$.
-We construct $\ValC$ by initially taking $\ValC := \nf{\Val}$
+We construct $\ValC$ by initially setting $\ValC := \nf{\Val}$
 and by exhaustively applying the following rule:%
@@
 \inferrule{
-  {\Varec{x}}\typ{\tau} \in \FV( \Val )
+  {\Varec{x}}\typ{\tau} \in \FV( \ValC )
   \betweenantes
@@
   \betweenantes
-  v \not\in \Terms^{\smash{\J}}_{\Varec{x}}(\Val)
+  v \not\in \Terms^{\smash{\J}}_{\Varec{x}}(\ValC)
 }{
@@
 %
-The rule chooses an unassigned variable in $\ValC$ and assigns it a fresh
+Given an unassigned variable in $\ValC$, this rule assigns it a fresh
 value---one that does not occur in $\Terms^{\smash{\J}}_{\Varec{x}}(\ValC)$

### 2015-02-19 tex/body.tex
Minor.

@@
 If $j \neq \jPrime$, since \rn{Cong} does not apply,
-$\smash{\const s^k_{\negvthinspace j}( u )}$ and $\smash{\const s^k_{\negvthinspace j}( u' )}$
-are in distinct equivalence classes only if $u$ and $u'$ are in distinct equivalence classes.
+any term of the form $\smash{\const s^k_{\negvthinspace j}( u' )}$ not occurring in $\ec{t}$ 
+is such that $u$ and $u'$ are in distinct equivalence classes.
 Due to the induction hypothesis and Lemma~\ref{lem:interpretation-completion}\PART{3}, $\M( u ) \neq \M( u' )$ for all such $u$, $u'$.
@@
 we may no longer assume that they are of infinite cardinality.
-When assuming that a finite type has infinite cardinality, the calculus presented in this paper is \relax{incomplete},
-due to the overly restrictive preconditions on \rn{Split} and \rn{Single}.
+When assuming that a finite type has infinite cardinality, the calculus presented in this paper is \relax{incomplete}
+since it may fail to recognize cases when the preconditions of \rn{Split} and \rn{Single} hold.
 This does not %necessarily
@@
 the implementation adds the clause %of the form
-$\exists x\;y\typ{\ty{u}}.\;\, x \not\teq y \mathrel\vee s \teq t$ to $F$.
+$(\exists x\;y\typ{\ty{u}}.\;\, x \not\teq y ) \mathrel\vee s \teq t$ to $F$.
 \begin{rep}In other words, either the cardinality of $\ty{u}$ is greater than one or $s$ is equal to $t$.\end{rep}
@@
 of being abbreviations.
-We have found that this leads to better performance because the alternative introduces terms more eagerly to $\tEc$.
+We have found that this leads to better performance since it introduces terms less eagerly to $\tEc$.
 Handling discriminators requires extending the decision procedure with several
@@
 There are not yet many benchmarks available, since codatatypes are a
-recent addition to Isabelle \cite{blanchette-et-al-2014-impl}. We
+recent addition to Isabelle \cite{blanchette-et-al-2014-impl}. \rem{Necessary to mention this? Seems to detract from significance of contribution.} We
 included all the theory files from the Isabelle distribution (Distro, 879 goals)

### 2015-02-19 tex/body.tex
more

--- 2015-02-19 tex/body.tex
@@
 Strictly speaking, normal interpretations are not models of codatatypes,
-because they are restricted to an $\omega$-regular values, whereas models must
+because they are restricted to an $\omega$-regular values, whereas actual models must
 also cater for acyclic infinite terms. However, by monotonicity of ground
-constraints, it is trivial to extend a normal interpretation into a model.
-In the following, we focus on normal interpretations.
-
-When constructing a model $\M$ for $\Ec$,
+constraints, it is trivial to extend a normal interpretation to an actual model.
+In the following, we focus on the former and unapologetically call them
+models.
+
+When constructing a model $\M$ of $\Ec$,
 it remains only to specify how $\M$ interprets wrongly applied selector terms and variables.

### 2015-02-19 tex/body.tex
more

@@
 Isabelle formalizations using Sledgehammer.
-There are not yet many benchmarks available, since codatatypes are a
-recent addition to Isabelle \cite{blanchette-et-al-2014-impl}. \rem{Necessary to mention this? Seems to detract from significance of contribution.} We
+%There are not yet many benchmarks available, since codatatypes are a
+%recent addition to Isabelle \cite{blanchette-et-al-2014-impl}. \rem{Necessary to mention this? Seems to detract from significance of contribution.}
+We
 included all the theory files from the Isabelle distribution (Distro, 879 goals)

### 2015-02-19 tex/body.tex
as discussed on Skype

@@
 
-%\newcommand\DISC[1]{t \teq \const{C}_{#1}\bigl(\const s^1_{#1}( t ), \ldots,\const s^{n_{#1}}_{#1}( t )\bigr)}
-\newcommand\DISC[1]{\const{d}_{#1}(t)}
+\newcommand\DISC[1]{t \teq \const{C}_{#1}\bigl(\const s^1_{#1}( t ), \ldots,\const s^{n_{#1}}_{#1}( t )\bigr)}
+%\newcommand\DISC[1]{\const{d}_{#1}(t)}
 
@@
 \newcommand{\J}{\mathcalx{J}}
-\newcommand{\M}{\mathcalx{M}}
+%\newcommand{\M}{\mathcalx{M}}
+\newcommand{\M}{\mathcalx{J}}
 \newcommand{\Val}{\mathcalx{A}\vvthinspace}
@@
 $\Sig$.
-First, all codatatypes in $\Codata$ are corecursive. This is not restrictive
+First, all codatatypes in $\Codata$ are corecursive. This is reasonable
 because noncorecursive codatatypes can be seen as nonrecursive
@@
 Second, all types $\tau \in \Nondata$ have infinite cardinality.
-This is not very limitative, because each interpreted type in $\Nondata$
-having finite cardinality~$n$
-can be replaced by an enumeration datatype with $n$~nullary constructors.
-As for uninterpreted types, since the constraints in $\Ec$ are ground, they cannot entail an upper
+This is not very restrictive, because in principle each interpreted type in
+$\Nondata$ having finite cardinality~$n$
+can be replaced by a %n enumeration
+datatype with $n$~nullary constructors
+\cite{barrett-et-al-2007}.
+As for uninterpreted types, since the constraints in $\Ec$
+are ground, they cannot entail an upper
 bound on the cardinality of any such type, so it is safe to consider them infinite.
@@
 based on a recursive computation of the cardinalities of the
-constructors' argument types. This step relies on the assumption that
-ordinary types are infinite.
+constructors' argument types.
+%This step relies on the assumption that ordinary types are infinite.
 
@@
 and let $\SSSS \AST_t$ be the limit of this sequence.
-This is a finite set for each $t$ because there can only be finitely many
-values of the same type as $t$.
+This is a finite set for each $t$, because in the absence of (co)recursion
+one eventually reaches a point where no selector can be applied.
 %%% TODO: Clarify above step in report.
@@
 we may no longer assume that they are of infinite cardinality.
-When assuming that a finite type has infinite cardinality, the calculus presented in this paper is \relax{incomplete}
-since it may fail to recognize cases when the preconditions of \rn{Split} and \rn{Single} hold.
+When assuming that a finite type has infinite cardinality, the calculus
+presented in this paper is \relax{incomplete}
+since it may fail to recognize cases where \rn{Split} and \rn{Single} should be
+applied.
 This does not %necessarily
@@
 The implementation of the decision procedure uses several optimizations
-following the lines of Barrett et al. \cite{barrett-et-al-2007}.
+following the lines of Barrett et al.\ \cite{barrett-et-al-2007}.
 %and are not reflected in the presentation of the calculus.

### 2015-02-19 tex/body.tex
removed | | around \delta in "\delta is finite" -- better stay syntactic

@@
 
-%%% @ANDY: I don't like using \qed or anything too similar (e.g. \Box) for
+%%% I don't like using \qed or anything too similar (e.g. \Box) for
 %%% anything else than "quid erat demonstrandum". I'm open to other symbols,
@@
 %\newcommand{\interp}[2]{#1\llbracket #2 \rrbracket}
-%%% @ANDY: I prefer a notation that puts more emphasis on #2 (which is, in essence,
+%%% I prefer a notation that puts more emphasis on #2 (which is, in essence,
 %%% the object that is returned) and less on #1
@@
 %\newcommand{\thD}{T_{\!\mathrm{cd}}}
-%%% @ANDY: the italicized "T" makes it look as though it were a variable;
+%%% the italicized "T" makes it look as though it were a variable;
 %%% and the subscript means we have two layers of subscripts in things like
@@
 
-%% @ANDY: Minor terminology point:
+%% Minor terminology point:
 %% A function symbol can have a "type signature" (but we're trying to avoid that terminology);
@@
 
-%%% @ANDY: I explicitly write "|\delta| is finite" instead of "\delta is finite"
-%%% to emphasize the dependency on the model. In principle, some types are
-%%% interpreted finitely in some models and infinitely in others. In Isabelle,
-%%% those wouldn't be considered "finite" types, but they wouldn't be "infinite"
-%%% either. (We have Haskell-style type classes for both of these concepts.)
-%%% Naturally, thanks to our assumptions at the beginning of Sect. 3, our types
-%%% are either finite or infinite.
-
 \begin{figure}[t!]
@@
   \text{ or }
-  \bigl( \delta \in \Data \text{ and } \left|\delta\right| \text{ is finite} \bigr)
+  \bigl( \delta \in \Data \text{ and } %\left|
+  \delta
+  %\right|
+  \text{ is finite} \bigr)
 }{
@@
 For \rn{Unique}, the construction of $\Val$ indicates that the values of $t$ and $u$ are equivalent in all models of $\Ec$.
-For \rn{Single}, $t$~and~$u$ must have the same value since the cardinality of their type is $1$.
+For \rn{Single}, $t$~and~$u$ must have the same value since the cardinality of their type is one.
 By the induction hypothesis, $\Ec \mathrel\cup t \teq u$ is $\thD$-unsatisfiable
@@
 Given a $\mu$-term $t$ with subterm $u$,
-%%% @ANDY: I'm taking the liberty of renaming "interpretation" to "expansion". Interpretation
-%%% sounds semantic to me (although it is sometimes used for syntactic transformations).
 the \emph{expansion of $u$ with respect to $t$} is the $\mu$-term $\interp{t}{u}^\emptyset$,
@@
 
-%%% @ANDY: I think it's enough to point this out clearly in the introduction.
+%%% I think it's enough to point this out clearly in the introduction.
 %%% At least, that always worked fine for my other papers. :)

### 2015-02-19 tex/body.tex
more

@@
 inspired by the $\mu$-calculus
-%\begin{conf}\cite{kozen-1983}\end{conf}%
-\cite[etc.]{kozen-1983,endrullis-et-al-2011}.
+\begin{conf}\cite{kozen-1983}\end{conf}%
+\begin{rep}\cite[etc.]{kozen-1983,endrullis-et-al-2011}\end{rep}.
 
@@
   \betweenantes
-  \left|\delta\right| = 1
+  \delta \text{ is a singleton}
 }{
@@
 variable to $B$ and there are finitely many distinct bound variables in a $\mu$-term.
+Intuitively, the expansion of a subterm is a stand-alone
+$\mu$-term that denotes the same value as the original subterm---e.g.,
+$\interp{\MU x.\: \const{C}(\MU y.\: \const{D}(x))}{\MU y.\: \const{D}(x)} =
+ \MU y.\: \const{D}(\MU x.\: \const{C}(y))$.
 %
@@
 
-%First, we need the following definition.
+First, we need the following definitions.
 We write $t \vsimv{x} u$ if $\mu$-terms $t$ and $u$ are syntactically equivalent
@@
 $\MU x.\; \const{C}( x, x ) \not\vsimv{y} \MU y.\; \const{C}( x, y )$.
-For a variable $x\typ{\tau}$ and a fixed normal interpretation $\J$,
+For a variable $x\typ{\tau}$ and a normal interpretation $\J$,
 we write $\Terms^{\smash{\J}}_x( \Val )$ to denote the set consisting of all values
@@
 such that $v \vsimv{x} \interp{t}{u}$ for some term $t$ with subterm $u$ occurring in the range of $\Val$.
-%\rem{TODO: Intuition?}
 
@@
 the range of $\ValC$ at each step.
+
+The last precondition is admittedly rather technical. Its role is to ensure that
+$\ValC$ remains injective modulo $\alpha$-equivalence. A simpler, less efficient
+way to express the rule would be to check that property explicitly before
+performing the assignment.
 

### 2015-02-19 tex/body.tex
more insights

@@
 If a selector is applied to a term $t$, or if $t$'s type is a finite datatype,
-the equivalence class of $t$ must contain a
+$t$'s equivalence class must contain a
 $\delta$ constructor term.
 This is enforced in the third phase by the \rn{Split} rule.
-A companion rule, \rn{Single}, focuses on the degenerate case where two
+Another rule, \rn{Single}, focuses on the degenerate case where two
 terms %from $\tEc$
@@
 value---one that does not occur in $\Terms^{\smash{\J}}_{\Varec{x}}(\ValC)$
-modulo $\alpha$-equivalence.
+modulo $\alpha$-equivalence---excluding
+not only existing terms in the range of $\ValC$ but also
+terms that could emerge as a result of the update.
 %\rem{TODO: What guarantees that such a value exists? (Answer: infinitely many values
@@
 $\ValC$ remains injective modulo $\alpha$-equivalence. A simpler, less efficient
-way to express the rule would be to check that property explicitly before
-performing the assignment.
+way to proceed would be to compute $\nf{\ValC [ {\Varec{x}} \mapsto \MU {\Varec{x}}.\; t ]}$ optimistically and check
+injectivity afterward.
 

### 2015-02-19 tex/body.tex
Make normal interpretation more explicit.  Reword for contains maximal set of values.

@@
 domain elements for that type.
-For a term $t\typ{\tau}$, $\M( t ) \in \M( \tau )$ denotes the interpretation of $t$ in $\M$.
+%For a term $t\typ{\tau}$, $\M( t ) \in \M( \tau )$ denotes the interpretation of $t$ in $\M$.
+$\M( f )$ denotes the interpretation of a function $f$ in $\M$.
+If $f$ is of type $ \tau_1 \times \cdots \times \tau_n \to \tau$,
+then $\M( f )$ is a total function of type $ \M( \tau_1 ) \times \cdots \times \M( \tau_n ) \to \M( \tau )$.
 %To simplify the presentation,
@@
 For each type $\tau$,
-$\J( \tau )$ is a maximal set of closed normal $\mu$-terms of that type that are
+$\J( \tau )$ contains a maximal set of closed normal $\mu$-terms of that type that are
 unique up to $\alpha$-equivalence and acyclic if $\tau \not\in \Codata$.
@@
 For each constructor term $\const{C}( \vec t \,)$ of type $\tau$,
-$\J( \const{C}( \vec t \,) )$ is the value
+$\J( \const{C} )( \J( \vec t ) \,)$ is the value
 in $\J( \tau )$ that is $\alpha$-equivalent to
@@
 For each selector term $\const s^k_{\negvthinspace j}( t )$ of type $\tau$, % for some $\vec u$,
-if $\J( t )$ is of the form $\MU x.\; \const{C}_{\negvthinspace j}( \vec u )$, 
-then $\J\negvvthinspace\bigl( \const s^k_{\negvthinspace j}( t ) \bigr)$ is the value
+if $\J( t )$ is $\MU x.\; \const{C}_{\negvthinspace j}( \vec u )$, 
+then $\J\negvvthinspace( \const s^k_{\negvthinspace j} )\bigl( \J( t ) \bigr)$ is the value
 in $\J( \tau )$ that is $\alpha$-equivalent to
@@
 
-Strictly speaking, normal interpretations are not models of codatatypes,
-because they are restricted to an $\omega$-regular values, whereas actual models must
-also cater for acyclic infinite terms. However, by monotonicity of ground
-constraints, it is trivial to extend a normal interpretation to an actual model.
-In the following, we focus on the former and unapologetically call them
-models.
+% AJR : rewording above "contains" makes this paragraph not necessary?  this reworded definition is more accurate and doesn't break the proofs.
+%  maybe comment that acyclic infinite values are included but are irrelevant for our purposes?
+
+%Strictly speaking, normal interpretations are not models of codatatypes,
+%because they are restricted to an $\omega$-regular values, whereas actual models must
+%also cater for acyclic infinite terms. However, by monotonicity of ground
+%constraints, it is trivial to extend a normal interpretation to an actual model.
+%In the following, we focus on the former and unapologetically call them
+%models.
 
@@
 %The latter also comes with no loss of generality since nested applications of selectors in $\Fc_0$ can be replaced by fresh variables while adding equalities to $\Fc$.
-We consider a normal model $\M$
+We consider a normal interpretation $\M$
 that interprets wrongly applied selectors based on equality information in $\Fc$

### 2015-02-19 tex/body.tex
more

@@
 For a variable $x\typ{\tau}$ and a normal interpretation $\J$,
-we write $\Terms^{\smash{\J}}_x( \Val )$ to denote the set consisting of all values
+we let $\Terms^{\smash{\J}}_x( \Val )$ denote the set consisting of all values
 $v \in \J( \tau )$
-such that $v \vsimv{x} \interp{t}{u}$ for some term $t$ with subterm $u$ occurring in the range of $\Val$.
+such that $v \vsimv{x} \interp{t}{u}$ for some subterm $u$ of a term $t$
+occurring in the range of $\Val$. This set describes shapes of terms to avoid.
 
@@
 \{
-\MU a.\: \const{D}( a ),\:
+\MU a.\: \const{C}( a ),\:
+\MU a.\: \const{D}( a ),\allowbreak\:
 \MU a.\: \const{C}( \const{D}( a )),\allowbreak\:
-\const{C}( \MU a.\: \const{D}( a )),\allowbreak\:
-\MU a.\: \const{C}( a )
+\const{C}( \MU a.\: \const{D}( a ))
 \}

### 2015-02-19 tex/body.tex
more

@@
 \newcommand{\J}{\mathcalx{J}}
-%\newcommand{\M}{\mathcalx{M}}
-\newcommand{\M}{\mathcalx{J}}
 \newcommand{\Val}{\mathcalx{A}\vvthinspace}
@@
 $\thD$-satisfiable.
-The proof relies on a specific model $\M$ of $\thD$ that satisfies $\Ec$.
+The proof relies on a specific model $\J$ of $\thD$ that satisfies $\Ec$.
 
@@
 We now define the class of interpretations for $\thD$.
-$\M( \tau )$ denotes the interpretation type $\tau$ in $\M$---that is, a set of
+$\J( \tau )$ denotes the interpretation type $\tau$ in $\J$---that is, a set of
 domain elements for that type.
-%For a term $t\typ{\tau}$, $\M( t ) \in \M( \tau )$ denotes the interpretation of $t$ in $\M$.
-$\M( f )$ denotes the interpretation of a function $f$ in $\M$.
+%For a term $t\typ{\tau}$, $\J( t ) \in \J( \tau )$ denotes the interpretation of $t$ in $\J$.
+$\J( f )$ denotes the interpretation of a function $f$ in $\J$.
 If $f$ is of type $ \tau_1 \times \cdots \times \tau_n \to \tau$,
-then $\M( f )$ is a total function of type $ \M( \tau_1 ) \times \cdots \times \M( \tau_n ) \to \M( \tau )$.
+then $\J( f )$ is a total function of type $ \J( \tau_1 ) \times \cdots \times \J( \tau_n ) \to \J( \tau )$.
 %To simplify the presentation,
@@
 For each constructor term $\const{C}( \vec t \,)$ of type $\tau$,
-$\J( \const{C} )( \J( \vec t ) \,)$ is the value
+$\J( \const{C} )( \J( \vec t \,))$ is the value
 in $\J( \tau )$ that is $\alpha$-equivalent to
@@
 
-% AJR : rewording above "contains" makes this paragraph not necessary?  this reworded definition is more accurate and doesn't break the proofs.
-%  maybe comment that acyclic infinite values are included but are irrelevant for our purposes?
-
-%Strictly speaking, normal interpretations are not models of codatatypes,
-%because they are restricted to an $\omega$-regular values, whereas actual models must
-%also cater for acyclic infinite terms. However, by monotonicity of ground
-%constraints, it is trivial to extend a normal interpretation to an actual model.
-%In the following, we focus on the former and unapologetically call them
-%models.
-
-When constructing a model $\M$ of $\Ec$,
-it remains only to specify how $\M$ interprets wrongly applied selector terms and variables.
+Not all normal interpretations are actually models of codatatypes, because
+models must contain all possible infinite terms, not only cyclic ones. However,
+only cyclic terms are interesting to us, and by monotonicity of ground
+constraints it is trivial to extend a normal interpretation to an actual model.
+Hence, we focus on normal interpretations.
+
+When constructing a model $\J$ of $\Ec$,
+it remains only to specify how $\J$ interprets wrongly applied selector terms and variables.
 For the latter, this will be based on the mapping $\Val$ computed in phase~2 of the calculus.
@@
 However, this contradicts the disequality $z \not\teq v$ in~$\Ec$.
-On the other hand, if the value assigned to ${\Varec{z}}$ is fresh, %not $\alpha$-equivalent to any term in $\Terms^{\smash{\M}}_{{\Varec{z}}}( \Val )$,
+On the other hand, if the value assigned to ${\Varec{z}}$ is fresh, %not $\alpha$-equivalent to any term in $\Terms^{\smash{\J}}_{{\Varec{z}}}( \Val )$,
 the values in the range of $\Val$ remain $\alpha$-disequivalent.
@@
 Let $\Fc$ be a saturated node in a derivation tree with root node $\Ec$.
-%We will construct a model $\M$ for a set of equalities $\Fc$ that is equivalent to $\Fc_0$, and where
+%We will construct a model $\J$ for a set of equalities $\Fc$ that is equivalent to $\Fc_0$, and where
 %all equivalence classes of $\Fc$ contain at least one variable, and all selectors in $\Fc$ are applied to variables only.
@@
 %The latter also comes with no loss of generality since nested applications of selectors in $\Fc_0$ can be replaced by fresh variables while adding equalities to $\Fc$.
-We consider a normal interpretation $\M$
+We consider a normal interpretation $\J$
 that interprets wrongly applied selectors based on equality information in $\Fc$
 and that interprets the variables of $\Fc$ based on the completion $\ValC$ of the mapping $\Val$ from phase~2.
-For the latter, let $\M( x )$ be the value in $\M$ that is $\alpha$-equivalent with $\ValC \ec{x}$ for each variable $x \in \Terms(\Fc)$,
+For the latter, let $\J( x )$ be the value in $\J$ that is $\alpha$-equivalent with $\ValC \ec{x}$ for each variable $x \in \Terms(\Fc)$,
 which by Lemma~\ref{lem:interpretation-completion}\PART{1} is guaranteed to exist.
 
-We first show $\M$ satisfies all equalities in $t_1 \teq t_2 \in \Fc$, where $t_1$ and $t_2$ have type $\tau$.
+We first show $\J$ satisfies all equalities in $t_1 \teq t_2 \in \Fc$, where $t_1$ and $t_2$ have type $\tau$.
 We show,
 by structural induction on $t$,
-that $\M( t ) \aequiv \ValC \ec{t}$ for all terms $t \in \Terms( \Fc )$,
-which since $\M$ is normal, implies $\M \models t_1 \teq t_2$.
-
-If $t$ is a variable, then $\M( t ) \vsim \ValC \ec{ t }$ by construction.
+that $\J( t ) \aequiv \ValC \ec{t}$ for all terms $t \in \Terms( \Fc )$,
+which since $\J$ is normal, implies $\J \models t_1 \teq t_2$.
+
+If $t$ is a variable, then $\J( t ) \vsim \ValC \ec{ t }$ by construction.
 
 If $t$ is a constructor term of the form $\const{C}( u_1, \ldots, u_n )$,
-then $\M( t )$ is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \M( u_1 ), \ldots, \M( u_n ) ) }$ for some fresh $x$,
+then $\J( t )$ is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \J( u_1 ), \ldots, \J( u_n ) ) }$ for some fresh $x$,
 which by the induction hypothesis is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \ValC \ec{ u_1 }, \ldots, \ValC \ec{ u_n } )}$.
@@
 %By Lemma~\ref{lem:mu-cong}, we have $t_u \vsim t_w$.
-Thus, $\M( t ) \vsim t_u \vsim \ValC \ec{t}$, and
-we have $\M( t ) \vsim \ValC \ec{t}$.
+Thus, $\J( t ) \vsim t_u \vsim \ValC \ec{t}$, and
+we have $\J( t ) \vsim \ValC \ec{t}$.
 
@@
 $\interp{\ValC \ec{u}}{ w_i } = \ValC \ec{ \smash{\const s^i_{\negvthinspace\jPrime}( u )}}$.
-If $j = \jPrime$, then $\M( t )$ is $\alpha$-equivalent with $\interp{\ValC \ec{u}}{w_k}$, which is equal to $\ValC \ec{ \smash{ \const s^k_{\negvthinspace j}( u )}}$,
+If $j = \jPrime$, then $\J( t )$ is $\alpha$-equivalent with $\interp{\ValC \ec{u}}{w_k}$, which is equal to $\ValC \ec{ \smash{ \const s^k_{\negvthinspace j}( u )}}$,
 which is $\ValC \ec{t}$.
@@
 is such that $u$ and $u'$ are in distinct equivalence classes.
-Due to the induction hypothesis and Lemma~\ref{lem:interpretation-completion}\PART{3}, $\M( u ) \neq \M( u' )$ for all such $u$, $u'$.
-Thus, we may interpret $\smash{\M(\const s^k_{\negvthinspace j})( \M( u ) )}$ as the value in $\M( \tau )$ that is $\alpha$-equivalent with $\ValC \ec{ t }$.
-
-We now show that all disequalities in $\Fc$ are satisfied by $\M$.
+Due to the induction hypothesis and Lemma~\ref{lem:interpretation-completion}\PART{3}, $\J( u ) \neq \J( u' )$ for all such $u$, $u'$.
+Thus, we may interpret $\smash{\J(\const s^k_{\negvthinspace j})( \J( u ) )}$ as the value in $\J( \tau )$ that is $\alpha$-equivalent with $\ValC \ec{ t }$.
+
+We now show that all disequalities in $\Fc$ are satisfied by $\J$.
 Say $t_1 \tneq t_2 \in \Fc$.
 Since \rn{Conflict} does not apply, $t_1 \teq t_2 \not\in \Fc$ and thus $\ec{t_1}$ and $\ec{t_2}$ are distinct.
-Since $\M( t_1 ) \vsim \ValC \ec{t_1}$ and $\M( t_2 ) \vsim \ValC \ec{t_2}$,
-by Lemma~\ref{lem:interpretation-completion}\PART{3}, $\M( t_1 ) \neq \M( t_2 )$, and thus $\M \models t_1 \tneq t_2$.
-
-Since by assumption $\Fc$ contains only equalities and disequalities, we have $\M \models \Fc$,
-and since $\Fc$ is a superset of $\Ec$, we have that $\Ec$ is satisfied by $\M$.
+Since $\J( t_1 ) \vsim \ValC \ec{t_1}$ and $\J( t_2 ) \vsim \ValC \ec{t_2}$,
+by Lemma~\ref{lem:interpretation-completion}\PART{3}, $\J( t_1 ) \neq \J( t_2 )$, and thus $\J \models t_1 \tneq t_2$.
+
+Since by assumption $\Fc$ contains only equalities and disequalities, we have $\J \models \Fc$,
+and since $\Fc$ is a superset of $\Ec$, we have that $\Ec$ is satisfied by $\J$.
 \qed

### 2015-02-19 tex/body.tex
more

@@
 
+{\looseness=-1
 First, we need the following definitions.
@@
 such that $v \vsimv{x} \interp{t}{u}$ for some subterm $u$ of a term $t$
-occurring in the range of $\Val$. This set describes shapes of terms to avoid.
+occurring in the range of $\Val$. This set describes shapes of terms to avoid
+when assigning a $\mu$-term to $x$.
+
+}
 
@@
 way to proceed would be to compute $\nf{\ValC [ {\Varec{x}} \mapsto \MU {\Varec{x}}.\; t ]}$ optimistically and check
-injectivity afterward.
+injectivity afterward. Injectivity is important to satisfy disequality constraints.
+For infinite codatatypes, it is always possible to find suitable values of $v$.
 
@@
 \}$.
-After applying the calculus to saturation on $\Ec$, the mapping $\Val$ is as follows:
+After applying the calculus to saturation, % on $\Ec$,
+the mapping $\Val$ is as follows:
 %\[\begin{array}{r@{}c@{}l@{\;}l@{\qquad}r@{}c@{}r@{\;}l@{}l@{\;}l}

### 2015-02-19 tex/body.tex
macro

@@
 \newcommand{\ttpath}[3]{\mathcalx{P}_{ #2 \rightarrow #3 }( #1 )}
-%\newcommand{\interp}[2]{#1\llbracket #2 \rrbracket}
+
+%\newcommand{\expand}[2]{#1\llbracket #2 \rrbracket}
 %%% I prefer a notation that puts more emphasis on #2 (which is, in essence,
 %%% the object that is returned) and less on #1
-%\newcommand{\interp}[2]{\llbracket #2 \rrbracket_{#1}}
+%\newcommand{\expand}[2]{\llbracket #2 \rrbracket_{#1}}
 %%% And one that doesn't look so semantic
-%\newcommand{\interp}[2]{\lceil #2 \rceil_{#1}}
-\newcommand{\interp}[2]{\langle \smash{#2} \rangle_{#1}}
+%\newcommand{\expand}[2]{\lceil #2 \rceil_{#1}}
+\newcommand{\expand}[2]{\langle \smash{#2} \rangle_{#1}}
+
+\newcommand{\interp}[2]{#1(#2)}
+
 \newcommand{\dpath}[3]{\delta^{#1}_{#2}( #3 )}
@@
 Given a $\mu$-term $t$ with subterm $u$,
-the \emph{expansion of $u$ with respect to $t$} is the $\mu$-term $\interp{t}{u}^\emptyset$,
+the \emph{expansion of $u$ with respect to $t$} is the $\mu$-term $\expand{t}{u}^\emptyset$,
 abbreviated to
-$\interp{t}{u}$, as returned by the following recursive function:
+$\expand{t}{u}$, as returned by the following recursive function:
 %where $x \leadsto \MU x.\; \const{C}( \vec u ) \in t$ indicates that $\MU x.\;
@@
 \[\begin{array}{r@{}c@{}l}
-\interp{t}{x}^B & {} \,=\, {} &
+\expand{t}{x}^B & {} \,=\, {} &
 \textstyle\relax{\begin{cases}
-    \MU x.\; \const{C}\smash{\bigl( \interp{t}{ \vec u}^{\smash{B \,\cup\, \{ x \}}}\bigr)} & \text{if this $x \notin B$ is
+    \MU x.\; \const{C}\smash{\bigl( \expand{t}{ \vec u}^{\smash{B \,\cup\, \{ x \}}}\bigr)} & \text{if this $x \notin B$ is
     bound by $\MU x.\; \const{C}( \vec u )$ in $t$}  \\[-\jot]
@@
 \end{cases}} ~ \\
-\interp{t}{\MU x.\; \const{C}( \vec u )}^B & = &
+\expand{t}{\MU x.\; \const{C}( \vec u )}^B & = &
 \textstyle\begin{cases}
-    \MU x.\; \const{C}\smash{\bigl( \interp{t}{ \vec u }^{\smash{B \,\cup\, \{ x \}}}\bigr)} & \text{if }x \not\in B \\[-\jot]
+    \MU x.\; \const{C}\smash{\bigl( \expand{t}{ \vec u }^{\smash{B \,\cup\, \{ x \}}}\bigr)} & \text{if }x \not\in B \\[-\jot]
     x & \text{otherwise}
@@
 $\mu$-term that denotes the same value as the original subterm---e.g.,
-$\interp{\MU x.\: \const{C}(\MU y.\: \const{D}(x))}{\MU y.\: \const{D}(x)} =
+$\expand{\MU x.\: \const{C}(\MU y.\: \const{D}(x))}{\MU y.\: \const{D}(x)} =
  \MU y.\: \const{D}(\MU x.\: \const{C}(y))$.
@@
 $u$ is of the form $\MU y.\; \const{C}( u_1, \ldots, u_n )$,
-and $\interp{t}{ t_k } \aequiv \interp{t}{ u_k }$ for all $k$. % \in [n]
+and $\expand{t}{ t_k } \aequiv \expand{t}{ u_k }$ for all $k$. % \in [n]
 The term $t$ is \emph{normal} if it does not contain a self-similar subterm
@@
 Their arguments have the same expansion with respect to~$t$:
-$\interp{t}{\MU y.\; \const{C}( y )} =
-\MU y.\; \const{C}\bigl( \interp{t}{y}^{\smash{\{y\}}} \bigr) =
+$\expand{t}{\MU y.\; \const{C}( y )} =
+\MU y.\; \const{C}\bigl( \expand{t}{y}^{\smash{\{y\}}} \bigr) =
 \MU y.\; \const{C}( y )$
 is $\alpha$-equivalent to
-$\interp{t}{y} =
-\MU y.\; \const{C}\bigl( \interp{t}{y}^{\smash{\{y\}}} \bigr) =
+$\expand{t}{y} =
+\MU y.\; \const{C}\bigl( \expand{t}{y}^{\smash{\{y\}}} \bigr) =
 \MU y.\; \const{C}( y )$.
@@
 %
-$\interp{t}{\MU y.\; \const{C}( x )}
- = \MU y.\; \const{C}\bigl( \interp{t}{x}^{\smash{\{y\}}} \bigr)
- = \MU y.\; \const{C}\bigl( \interp{t}{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) ) }^{\smash{\{y\}}} \bigr)
- = \MU y.\; \const{C}\bigl( \MU x.\; \const{C}( \interp{t}{\MU y.\; \const{C}( x )}^{\smash{\{x,y\}}} ) \bigr)
+$\expand{t}{\MU y.\; \const{C}( x )}
+ = \MU y.\; \const{C}\bigl( \expand{t}{x}^{\smash{\{y\}}} \bigr)
+ = \MU y.\; \const{C}\bigl( \expand{t}{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) ) }^{\smash{\{y\}}} \bigr)
+ = \MU y.\; \const{C}\bigl( \MU x.\; \const{C}( \expand{t}{\MU y.\; \const{C}( x )}^{\smash{\{x,y\}}} ) \bigr)
  = \MU y.\allowbreak\; \const{C}( \MU x.\allowbreak\; \const{C}( y ) )$
-is $\alpha$-equivalent to $\interp{t}{x} = t$.
+is $\alpha$-equivalent to $\expand{t}{x} = t$.
 % for the renaming $\{ x \mapsto y, y \mapsto x \}$.
@@
 If the $\mu$-terms $\vec u$ are in normal form and $t = \nf{\MU x.\; \const{C}( \vec u )} = \MU x.\; \const{C}( \vec w )$,
-then $\vec u \vsim \interp{t}{\vec w}$\vthinspace.
+then $\vec u \vsim \expand{t}{\vec w}$\vthinspace.
 \end{lemma}
@@
 \label{lem:mu-norm-interp}
-If the $\mu$-term $t$ is normal and $u$ is a subterm of $t$, then $\interp{t}{u}$ is normal.
+If the $\mu$-term $t$ is normal and $u$ is a subterm of $t$, then $\expand{t}{u}$ is normal.
 \end{lemma}
@@
 %If $t_u = \MU x.\; \const{C}( \vec u )$ and $t_w = \MU y.\; \const{C}( \vec w )$ are in normal form,
-%then $\interp{t_u}{ \vec u } \vsim \interp{t_w}{ \vec w }$ if and only if $t_u \vsim t_w$.
+%then $\expand{t_u}{ \vec u } \vsim \expand{t_w}{ \vec w }$ if and only if $t_u \vsim t_w$.
 %\end{lemma}
@@
 We now define the class of interpretations for $\thD$.
-$\J( \tau )$ denotes the interpretation type $\tau$ in $\J$---that is, a set of
+$\interp{\J}{\tau}$ denotes the interpretation type $\tau$ in $\J$---that is, a set of
 domain elements for that type.
-%For a term $t\typ{\tau}$, $\J( t ) \in \J( \tau )$ denotes the interpretation of $t$ in $\J$.
-$\J( f )$ denotes the interpretation of a function $f$ in $\J$.
+$\interp{\J}{f}$ denotes the interpretation of a function $f$ in $\J$.
 If $f$ is of type $ \tau_1 \times \cdots \times \tau_n \to \tau$,
-then $\J( f )$ is a total function of type $ \J( \tau_1 ) \times \cdots \times \J( \tau_n ) \to \J( \tau )$.
+then $\interp{\J}{f}$ is a total function of type $ \interp{\J}{\tau_1} \times \cdots \times \interp{\J}{\tau_n} \to \interp{\J}{\tau}$.
 %To simplify the presentation,
@@
 For each type $\tau$,
-$\J( \tau )$ contains a maximal set of closed normal $\mu$-terms of that type that are
+$\interp{\J}{\tau}$ contains a maximal set of closed normal $\mu$-terms of that type that are
 unique up to $\alpha$-equivalence and acyclic if $\tau \not\in \Codata$.
@@
 For each constructor term $\const{C}( \vec t \,)$ of type $\tau$,
-$\J( \const{C} )( \J( \vec t \,))$ is the value
-in $\J( \tau )$ that is $\alpha$-equivalent to
-$\nf{\MU x.\; \const{C}( \J( \vec t \,) )}$, where $x$ is fresh.
+$\interp{\J}{\const{C}}( \interp{\J}{\vec t} )$ is the value
+in $\interp{\J}{\tau}$ that is $\alpha$-equivalent to
+$\nf{\MU x.\; \const{C}( \interp{\J}{\vec t} )}$, where $x$ is fresh.
 \item
 For each selector term $\const s^k_{\negvthinspace j}( t )$ of type $\tau$, % for some $\vec u$,
-if $\J( t )$ is $\MU x.\; \const{C}_{\negvthinspace j}( \vec u )$, 
-then $\J\negvvthinspace( \const s^k_{\negvthinspace j} )\bigl( \J( t ) \bigr)$ is the value
-in $\J( \tau )$ that is $\alpha$-equivalent to
-$\interp{\J( t )}{u_k}$.
+if $\interp{\J}{t}$ is $\MU x.\; \const{C}_{\negvthinspace j}( \vec u )$, 
+then $\J\negvvthinspace( \const s^k_{\negvthinspace j} )\bigl( \interp{\J}{t} \bigr)$ is the value
+in $\interp{\J}{\tau}$ that is $\alpha$-equivalent to
+$\expand{\interp{\J}{t}}{u_k}$.
 \end{enumerate}
@@
 we let $\Terms^{\smash{\J}}_x( \Val )$ denote the set consisting of all values
-$v \in \J( \tau )$
-such that $v \vsimv{x} \interp{t}{u}$ for some subterm $u$ of a term $t$
+$v \in \interp{\J}{\tau}$
+such that $v \vsimv{x} \expand{t}{u}$ for some subterm $u$ of a term $t$
 occurring in the range of $\Val$. This set describes shapes of terms to avoid
@@
   \betweenantes
-  v \in \J( \tau )
+  v \in \interp{\J}{\tau}
   \betweenantes
@@
 \}$.
-After applying the calculus to saturation, % on $\Ec$,
+After applying the calculus to saturation on $\Ec$,
 the mapping $\Val$ is as follows:
@@
 \begin{enumerate}
-\item[\rm (1)] $\ValC \ec{x\typ{\tau}}$ is $\alpha$-equivalent to a value in $\J( \tau )$.
+\item[\rm (1)] $\ValC \ec{x\typ{\tau}}$ is $\alpha$-equivalent to a value in $\interp{\J}{\tau}$.
 \item[\rm (2)]
-$\ValC \ec{x} = \interp{\ValC \ec{y}}{ t }$ for
+$\ValC \ec{x} = \expand{\ValC \ec{y}}{ t }$ for
 all subterms $t$ of $\ValC \ec{y}$ with $\muvar(t) = {\Varec{x}}$.
@@
 Thus, $\ValC \ec{x}$ is acyclic when $\tau \not\in \Codata$.
-Therefore, by definition, $\ValC \ec{x}$ is $\alpha$-equivalent to a value in $\J( \tau )$.
+Therefore, by definition, $\ValC \ec{x}$ is $\alpha$-equivalent to a value in $\interp{\J}{\tau}$.
 
@@
 Moreover, free variables ${\Varec{z}}$ in $t$ that are bound in $\Val \ec{y}$ are interpreted in
-the expansion of $\interp{\Val \ec{y}}{ t }$ as a term
+the expansion of $\expand{\Val \ec{y}}{ t }$ as a term
 constructed by a sequence of substitutions of the form mentioned above to ${\Varec{z}}$. %\rem{TODO: more detail}
 Thus, we have
-$\interp{\Val \ec{y}}{ t } = \interp{\Val \ec{x}}{ \Val \ec{x} } = \Val \ec{x}$, and thus (2) holds for $\Val$.
+$\expand{\Val \ec{y}}{ t } = \expand{\Val \ec{x}}{ \Val \ec{x} } = \Val \ec{x}$, and thus (2) holds for $\Val$.
 Property (3) holds for $\Val$ since \rn{Unique} does not apply.
@@
 contains a subterm of the form $\MU {\Varec{y}}.\; \const{C}( u_1, \ldots, u_n )$
-where $\interp{\Val \ec{x}}{ t_i } \aequiv \interp{\Val \ec{x} }{ u_i }$ for some $i$. % \in [n]$.
+where $\expand{\Val \ec{x}}{ t_i } \aequiv \expand{\Val \ec{x} }{ u_i }$ for some $i$. % \in [n]$.
 Due to the construction of $\Val$, $\ec{y}$ contains a constructor $\const{C}( w_1, \ldots, w_n )$ and $\muvar( u_i ) = w_i$ for some $i$. % \in [n]$.
@@
 we have $\ec{w_{\negvthinspace j}}$ and $\ec{z_{\negvthinspace j}}$ are distinct for some $j$.
-By (2), $\interp{\Val \ec{x}}{ t_{\negvthinspace j} } = \Val \ec{z_{\negvthinspace j}}$
-and $\interp{\Val \ec{x}}{ u_{\negvthinspace j} } = \Val \ec{w_{\negvthinspace j}}$,
+By (2), $\expand{\Val \ec{x}}{ t_{\negvthinspace j} } = \Val \ec{z_{\negvthinspace j}}$
+and $\expand{\Val \ec{x}}{ u_{\negvthinspace j} } = \Val \ec{w_{\negvthinspace j}}$,
 which are not $\alpha$-equivalent by (3), contradicting the fact that $\MU {\Varec{y}}.\; \const{C}( u_1, \ldots, u_n )$ is a self-similar subterm of $\Val \ec{x}$.
@@
 To show (2),
-by assumption of (2) on $\Val_1$, we have $\interp{\Val_1 \ec{y}}{ u } = \Val_1 \ec{x}$ for
+by assumption of (2) on $\Val_1$, we have $\expand{\Val_1 \ec{y}}{ u } = \Val_1 \ec{x}$ for
 all subterms $u$ of $\Val_1 \ec{y}$ where $\muvar(t) = {\Varec{x}}$.
-Thus, $\interp{\Val_1 \sigma \ec{y}}{ u } = \Val_1 \sigma \ec{x}$ and
-$\interp{\nf{\Val_1 \sigma} \ec{y}}{ u } = \nf{\Val_1 \sigma} \ec{x}$.
+Thus, $\expand{\Val_1 \sigma \ec{y}}{ u } = \Val_1 \sigma \ec{x}$ and
+$\expand{\nf{\Val_1 \sigma} \ec{y}}{ u } = \nf{\Val_1 \sigma} \ec{x}$.
 To show (3),
@@
 are $\alpha$-disequivalent at position $p$ since they do not contain the same free variables.
-Since $\interp{\Val_1 \ec{y}}{t_1}$ is normal,
+Since $\expand{\Val_1 \ec{y}}{t_1}$ is normal,
 there is a closed $\mu$-term $v \in \Terms^{\smash{\J}}_{\Varec{x}}( \Val_1 )$
-such that $v \vsimv{\Varec{x}} \interp{\Val_1 \ec{y}}{t_1}$.
+such that $v \vsimv{\Varec{x}} \expand{\Val_1 \ec{y}}{t_1}$.
 Thus, by assumption on the selection of $\MU {\Varec{x}}.\; t$,
-we have $\MU {\Varec{x}}.\; t \not\vsimv{\Varec{x}} \interp{\Val_1 \ec{y}}{t_1}$,
+we have $\MU {\Varec{x}}.\; t \not\vsimv{\Varec{x}} \expand{\Val_1 \ec{y}}{t_1}$,
 which implies that the expansion of $\nf{\Val_1 \sigma} \ec{y}$ and $\nf{\Val_1 \sigma} \ec{z}$
@@
 and that interprets the variables of $\Fc$ based on the completion $\ValC$ of the mapping $\Val$ from phase~2.
-For the latter, let $\J( x )$ be the value in $\J$ that is $\alpha$-equivalent with $\ValC \ec{x}$ for each variable $x \in \Terms(\Fc)$,
+For the latter, let $\interp{\J}{x}$ be the value in $\J$ that is $\alpha$-equivalent with $\ValC \ec{x}$ for each variable $x \in \Terms(\Fc)$,
 which by Lemma~\ref{lem:interpretation-completion}\PART{1} is guaranteed to exist.
@@
 by structural induction on $t$,
-that $\J( t ) \aequiv \ValC \ec{t}$ for all terms $t \in \Terms( \Fc )$,
+that $\interp{\J}{t} \aequiv \ValC \ec{t}$ for all terms $t \in \Terms( \Fc )$,
 which since $\J$ is normal, implies $\J \models t_1 \teq t_2$.
 
-If $t$ is a variable, then $\J( t ) \vsim \ValC \ec{ t }$ by construction.
+If $t$ is a variable, then $\interp{\J}{t} \vsim \ValC \ec{ t }$ by construction.
 
 If $t$ is a constructor term of the form $\const{C}( u_1, \ldots, u_n )$,
-then $\J( t )$ is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \J( u_1 ), \ldots, \J( u_n ) ) }$ for some fresh $x$,
+then $\interp{\J}{t}$ is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \interp{\J}{u_1}, \ldots, \interp{\J}{u_n} ) }$ for some fresh $x$,
 which by the induction hypothesis is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \ValC \ec{ u_1 }, \ldots, \ValC \ec{ u_n } )}$.
@@
 %of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$
-%where $\interp{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$ for $i \in [n]$.
+%where $\expand{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$ for $i \in [n]$.
 Since \rn{Inject} and \rn{Clash} do not apply to $\Fc$,
@@
 $\muvar(w_i) = {\Varec{u_i}}$ for each $i$.  Thus by Lemma~\ref{lem:interpretation-completion}\PART{2},
-$\interp{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$.
+$\expand{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$.
 For each $i$, let $u_\iPrime$ be the $i$\vvthinspace th argument of $t_u$.
@@
 \begin{conf}Clearly,\end{conf} %% TODO: Fix! ;)
-$\interp{t_u}{u_\iPrime} \vsim \ValC \ec{u_i}$.
-Thus $\interp{t_u}{u_\iPrime} \vsim \interp{\ValC \ec{t}}{ w_i }$. % for each $i \in [n]$.
+$\expand{t_u}{u_\iPrime} \vsim \ValC \ec{u_i}$.
+Thus $\expand{t_u}{u_\iPrime} \vsim \expand{\ValC \ec{t}}{ w_i }$. % for each $i \in [n]$.
 %By Lemma~\ref{lem:mu-cong}, we have $t_u \vsim t_w$.
-Thus, $\J( t ) \vsim t_u \vsim \ValC \ec{t}$, and
-we have $\J( t ) \vsim \ValC \ec{t}$.
+Thus, $\interp{\J}{t} \vsim t_u \vsim \ValC \ec{t}$, and
+we have $\interp{\J}{t} \vsim \ValC \ec{t}$.
 
@@
 %$\MU {\Varec{u}}.\; \const C_{\negvthinspace\jPrime}( w_1, \ldots, w_n )$,
-%where $\interp{\ValC \ec{u} }{ w_i } = \ValC \ec{ \const s^i_{\negvthinspace\jPrime}( u )}$ for $i \in [n]$.
+%where $\expand{\ValC \ec{u} }{ w_i } = \ValC \ec{ \const s^i_{\negvthinspace\jPrime}( u )}$ for $i \in [n]$.
 Since \rn{Inject} and \rn{Clash} do not apply to $\Fc$,
@@
 for each $i$ and thus by Lemma~\ref{lem:interpretation-completion}\PART{2},
-$\interp{\ValC \ec{u}}{ w_i } = \ValC \ec{ \smash{\const s^i_{\negvthinspace\jPrime}( u )}}$.
-If $j = \jPrime$, then $\J( t )$ is $\alpha$-equivalent with $\interp{\ValC \ec{u}}{w_k}$, which is equal to $\ValC \ec{ \smash{ \const s^k_{\negvthinspace j}( u )}}$,
+$\expand{\ValC \ec{u}}{ w_i } = \ValC \ec{ \smash{\const s^i_{\negvthinspace\jPrime}( u )}}$.
+If $j = \jPrime$, then $\interp{\J}{t}$ is $\alpha$-equivalent with $\expand{\ValC \ec{u}}{w_k}$, which is equal to $\ValC \ec{ \smash{ \const s^k_{\negvthinspace j}( u )}}$,
 which is $\ValC \ec{t}$.
@@
 is such that $u$ and $u'$ are in distinct equivalence classes.
-Due to the induction hypothesis and Lemma~\ref{lem:interpretation-completion}\PART{3}, $\J( u ) \neq \J( u' )$ for all such $u$, $u'$.
-Thus, we may interpret $\smash{\J(\const s^k_{\negvthinspace j})( \J( u ) )}$ as the value in $\J( \tau )$ that is $\alpha$-equivalent with $\ValC \ec{ t }$.
+Due to the induction hypothesis and Lemma~\ref{lem:interpretation-completion}\PART{3}, $\interp{\J}{u} \neq \interp{\J}{u'}$ for all such $u$, $u'$.
+Thus, we may interpret $\smash{\interp{\J}{\const s^k_{\negvthinspace j}}( \interp{\J}{u} )}$ as the value in $\interp{\J}{\tau}$ that is $\alpha$-equivalent with $\ValC \ec{ t }$.
 
@@
 Since \rn{Conflict} does not apply, $t_1 \teq t_2 \not\in \Fc$ and thus $\ec{t_1}$ and $\ec{t_2}$ are distinct.
-Since $\J( t_1 ) \vsim \ValC \ec{t_1}$ and $\J( t_2 ) \vsim \ValC \ec{t_2}$,
-by Lemma~\ref{lem:interpretation-completion}\PART{3}, $\J( t_1 ) \neq \J( t_2 )$, and thus $\J \models t_1 \tneq t_2$.
+Since $\interp{\J}{t_1} \vsim \ValC \ec{t_1}$ and $\interp{\J}{t_2} \vsim \ValC \ec{t_2}$,
+by Lemma~\ref{lem:interpretation-completion}\PART{3}, $\interp{\J}{t_1} \neq \interp{\J}{t_2}$, and thus $\J \models t_1 \tneq t_2$.
 

### 2015-02-19 tex/body.tex
more

@@
 
-\newcommand{\interp}[2]{#1(#2)}
+%\newcommand{\interp}[2]{#1(#2)}
+\newcommand{\interp}[2]{{#2}^{#1}}
 
@@
 domain elements for that type.
-$\interp{\J}{f}$ denotes the interpretation of a function $f$ in $\J$.
-If $f$ is of type $ \tau_1 \times \cdots \times \tau_n \to \tau$,
-then $\interp{\J}{f}$ is a total function of type $ \interp{\J}{\tau_1} \times \cdots \times \interp{\J}{\tau_n} \to \interp{\J}{\tau}$.
+$\interp{\J}{\const{f}}$ denotes the interpretation of a function $\const{f}$ in $\J$.
+If $\const{f}$ is of type $ \tau_1 \times \cdots \times \tau_n \to \tau$,
+then $\interp{\J}{\const{f}}$ is a total function from $\interp{\J}{\tau_1} \times \cdots \times \interp{\J}{\tau_n}$ to $\interp{\J}{\tau}$.
 %To simplify the presentation,

### 2015-02-19 tex/body.tex
going back to avoid too many subscripts/superscripts
terminology

@@
 
-%\newcommand{\interp}[2]{#1(#2)}
-\newcommand{\interp}[2]{{#2}^{#1}}
+\newcommand{\interp}[2]{#1(#2)}
+%\newcommand{\interp}[2]{{#2}^{#1}}
 
@@
 For each type $\tau$,
-$\interp{\J}{\tau}$ contains a maximal set of closed normal $\mu$-terms of that type that are
+%%% sets *contain* elements, they *include* other sets
+$\interp{\J}{\tau}$ includes a maximal set of closed normal $\mu$-terms of that type that are
 unique up to $\alpha$-equivalence and acyclic if $\tau \not\in \Codata$.

### 2015-02-19 tex/body.tex
more

@@
 
-Not all normal interpretations are actually models of codatatypes, because
+Not all normal interpretations are models of codatatypes, because
 models must contain all possible infinite terms, not only cyclic ones. However,
 only cyclic terms are interesting to us, and by monotonicity of ground
-constraints it is trivial to extend a normal interpretation to an actual model.
-Hence, we focus on normal interpretations.
+constraints it is trivial to extend any normal interpretation with additional
+values. Hence, we focus on normal interpretations.
 

### 2015-02-19 tex/body.tex
looks better

@@
 to ${\Varec{z}}$.
-Legal substitutions for ${\Varec{z}}$ may cause the range of $\Val$ to contain abnormal terms.
+\begin{rep}Legal substitutions for ${\Varec{z}}$ may cause the range of $\Val$ to contain abnormal terms.
 For example, after assigning $\MU {\Varec{z}}.\; \const{D}( \const{C}( {\Varec{z}} ) )$ to $\Varec{z}$, we have
@@
 with normal form
-$\MU {\Varec{u}}.\; \const{C}( \MU {\Varec{z}}.\; \const{D}( {\Varec{u}} ) )$.
+$\MU {\Varec{u}}.\; \const{C}( \MU {\Varec{z}}.\; \const{D}( {\Varec{u}} ) )$.\end{rep}
 \xend
@@
 the cardinality of $\tau$ must be one.
-Since \rn{Single} does not apply to $\Ec$,
+Since \rn{Single} does not apply, % to $\Ec$,
 there is only one equivalence class of type $\tau$ in $\Ec$,
@@
 each pair of constructor terms $\const{C}_{\negvthinspace j}( \vec t )$ and $\const{C}_{\negvthinspace\jPrime}( \vec u )$ in $\ec{z}$
-are such that $j = \jPrime$ since \rn{Clash} does not apply to $\Ec$,
-and are such that $\ec{ \vec t } = \ec{ \vec u }$ since \rn{Inject} does not apply to $\Ec$.
+are such that $j = \jPrime$ since \rn{Clash} does not apply, % to $\Ec$,
+and are such that $\ec{ \vec t } = \ec{ \vec u }$ since \rn{Inject} does not apply. % to $\Ec$.
 Thus,
@@
 Due to the construction of $\Val$, $\ec{y}$ contains a constructor $\const{C}( w_1, \ldots, w_n )$ and $\muvar( u_i ) = w_i$ for some $i$. % \in [n]$.
-Since \rn{Cong} does not apply to $\Ec$
+Since \rn{Cong} does not apply, % to $\Ec$
 we have $\ec{w_{\negvthinspace j}}$ and $\ec{z_{\negvthinspace j}}$ are distinct for some $j$.
@@
 %where $\expand{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$ for $i \in [n]$.
-Since \rn{Inject} and \rn{Clash} do not apply to $\Fc$,
+Since \rn{Inject} and \rn{Clash} do not apply, % to $\Fc$,
 by the construction of $\ValC$ we have $\ValC \ec{t}$ is a term
@@
 %where $\expand{\ValC \ec{u} }{ w_i } = \ValC \ec{ \const s^i_{\negvthinspace\jPrime}( u )}$ for $i \in [n]$.
-Since \rn{Inject} and \rn{Clash} do not apply to $\Fc$,
+Since \rn{Inject} and \rn{Clash} do not apply, % to $\Fc$,
 by construction $\ValC$ must be of the form
@@
 $\expand{\ValC \ec{u}}{ w_i } = \ValC \ec{ \smash{\const s^i_{\negvthinspace\jPrime}( u )}}$.
-If $j = \jPrime$, then $\interp{\J}{t}$ is $\alpha$-equivalent with $\expand{\ValC \ec{u}}{w_k}$, which is equal to $\ValC \ec{ \smash{ \const s^k_{\negvthinspace j}( u )}}$,
-which is $\ValC \ec{t}$.
+If $j = \jPrime$, then $\interp{\J}{t}$ is $\alpha$-equivalent with $\expand{\ValC \ec{u}}{w_k}$, which is equal to $\ValC \ec{ \smash{ \const s^k_{\negvthinspace j}( u )}}
+= \ValC \ec{t}$.
 If $j \neq \jPrime$, since \rn{Cong} does not apply,
 any term of the form $\smash{\const s^k_{\negvthinspace j}( u' )}$ not occurring in $\ec{t}$ 
-is such that $u$ and $u'$ are in distinct equivalence classes.
-Due to the induction hypothesis and Lemma~\ref{lem:interpretation-completion}\PART{3}, $\interp{\J}{u} \neq \interp{\J}{u'}$ for all such $u$, $u'$.
+is such that $[u] \not= [u']$.
+By the induction hypothesis and Lemma~\ref{lem:interpretation-completion}\PART{3}, $\interp{\J}{u} \neq \interp{\J}{u'}$ for all such $u$, $u'$.
 Thus, we may interpret $\smash{\interp{\J}{\const s^k_{\negvthinspace j}}( \interp{\J}{u} )}$ as the value in $\interp{\J}{\tau}$ that is $\alpha$-equivalent with $\ValC \ec{ t }$.
@@
 We now show that all disequalities in $\Fc$ are satisfied by $\J$.
-Say $t_1 \tneq t_2 \in \Fc$.
-Since \rn{Conflict} does not apply, $t_1 \teq t_2 \not\in \Fc$ and thus $\ec{t_1}$ and $\ec{t_2}$ are distinct.
-Since $\interp{\J}{t_1} \vsim \ValC \ec{t_1}$ and $\interp{\J}{t_2} \vsim \ValC \ec{t_2}$,
-by Lemma~\ref{lem:interpretation-completion}\PART{3}, $\interp{\J}{t_1} \neq \interp{\J}{t_2}$, and thus $\J \models t_1 \tneq t_2$.
+Assume $t \tneq u \in \Fc$.
+Since \rn{Conflict} does not apply, $t \teq u \not\in \Fc$ and thus $\ec{t}$ and $\ec{u}$ are distinct.
+Since $\interp{\J}{t} \vsim \ValC \ec{t}$ and $\interp{\J}{u} \vsim \ValC \ec{u}$,
+by Lemma~\ref{lem:interpretation-completion}\PART{3}, $\interp{\J}{t} \neq \interp{\J}{u}$, and thus $\J \models t \tneq u$.
 
@@
 %and are not reflected in the presentation of the calculus.
-We briefly mention the main ones.
-%
-Discriminators are part of the signature instead
-of being abbreviations.
-We have found that this leads to better performance since it introduces terms less eagerly to $\tEc$.
-Handling discriminators requires extending the decision procedure with several
+\begin{rep}We briefly mention the main ones. \end{rep}%
+Discriminators are part of the signature and not abbreviations.
+This requires extending the decision procedure with several
 rules \cite{barrett-et-al-2007},
-which apply uniformly to both datatypes and codatatypes.
-Also, selectors are collapsed eagerly:
+which apply uniformly to datatypes and codatatypes.
+This approach often leads to better performance because it introduces terms
+less eagerly to $\tEc$.
+Selectors are collapsed eagerly:
 If $\const s^k_{\negvthinspace j}( t ) \in \tEc$ and $t =
-\const{C}_{\negvthinspace j}( u_1, \ldots, u_n )$, we immediately add
+\const{C}_{\negvthinspace j}( u_1, \ldots, u_n )$, the solver directly adds
 $\smash{\const s^k_{\negvthinspace j}( t ) \teq u_k}$ to $\Ec$, whereas the presented calculus

### 2015-02-19 tex/body.tex
edits

@@
 designed to be composable via the Nelson--Oppen method \cite{nelson-oppen-1979}.
-For the datatype case, it follows the lines of Barrett et al.\ \cite{barrett-et-al-2007}.
+It generalizes the procedure by Barrett et al.\ \cite{barrett-et-al-2007}, which only covers
+datatypes.
 To our knowledge, our procedure is the first of its kind for the theory of
@@
 A formula $\varphi$ is $\thD$-\emph{satisfiable} if there
-exists an interpretation in $\JJJJ$ that satisfies $\varphi$.
+exists an interpretation in $\JJJJ$ that satisfies~$\varphi$.
 % otherwise, it is $\thD$-unsatisfiable. (I trust the readers' intelligence when
@@
 models must contain all possible infinite terms, not only cyclic ones. However,
-only cyclic terms are interesting to us, and by monotonicity of ground
-constraints it is trivial to extend any normal interpretation with additional
-values. Hence, we focus on normal interpretations.
+acyclic infinite values are uninteresting to us, and by monotonicity of ground
+constraints it is trivial to extend any normal interpretation with extra
+domain elements to obtain a genuine model if desired. %Hence, we focus on normal interpretations.
 

### 2015-02-19 tex/body.tex
more edits

@@
 \end{tabular}
-\vspace*{-3pt} %% TYPESETTING HACK
+%\vspace*{-3pt} %% TYPESETTING HACK
 \caption{\,Derivation rules for bidirectional closure%.
@@
 \end{tabular}
-\vspace*{-3pt} %% TYPESETTING HACK
+%\vspace*{-3pt} %% TYPESETTING HACK
 \caption{\,Derivation rules for acyclicity and uniqueness%.
@@
 for each $\const C_{\negvthinspace j} \in \smash{\Ctr^\delta}$.
-Since by exhaustiveness, all models of $\thD$ entail (exactly) one
+Since by exhaustiveness, all models of $\thD$ entail exactly one
 $\DISC{\negvthinspace j}$,
@@
 Otherwise, the child node of $\Ec$ is a closed derivation tree
-whose root node $\Ec \mathrel\cup t \teq u$ is obtained by applying one of the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, \rn{Inject}, \rn{Unique}, or \rn{Single}.
+whose root
+node $\Ec \mathrel\cup t \teq u$ is obtained by applying one of the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, \rn{Inject}, \rn{Unique}, or \rn{Single}.
 In all these cases, $\Ec \models_{\thD} t \teq u$.
@@
 \MU y.\; \const{C}( y )$.
-The term $t = \MU x.\; \const{C}( \MU y.\; \const{C}( x ) )$ is also abnormal,
-since $\MU y.\; \const{C}( x )$ is a self-similar subterm of $t$,
+The term $u = \MU x.\; \const{C}( \MU y.\; \const{C}( x ) )$ is also abnormal,
+since $\MU y.\; \const{C}( x )$ is a self-similar subterm of $u$,
 noting that
 %
-$\expand{t}{\MU y.\; \const{C}( x )}
- = \MU y.\; \const{C}\bigl( \expand{t}{x}^{\smash{\{y\}}} \bigr)
- = \MU y.\; \const{C}\bigl( \expand{t}{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) ) }^{\smash{\{y\}}} \bigr)
- = \MU y.\; \const{C}\bigl( \MU x.\; \const{C}( \expand{t}{\MU y.\; \const{C}( x )}^{\smash{\{x,y\}}} ) \bigr)
+$\expand{u}{\MU y.\; \const{C}( x )}
+ = \MU y.\; \const{C}\bigl( \expand{u}{x}^{\smash{\{y\}}} \bigr)
+ = \MU y.\; \const{C}\bigl( \expand{u}{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) ) }^{\smash{\{y\}}} \bigr)
+ = \MU y.\; \const{C}\bigl( \MU x.\; \const{C}( \expand{u}{\MU y.\; \const{C}( x )}^{\smash{\{x,y\}}} ) \bigr)
  = \MU y.\allowbreak\; \const{C}( \MU x.\allowbreak\; \const{C}( y ) )$
-is $\alpha$-equivalent to $\expand{t}{x} = t$.
+is $\alpha$-equivalent to $\expand{u}{x} = u$.
 % for the renaming $\{ x \mapsto y, y \mapsto x \}$.
@@
 %and by replacing all of the abnormal subterms of $t$ with their corresponding normal forms.
-and by recursively normalizing the subterms.
+and by recursively normalizing the other subterms.
 For variables, $\nf{x} = x$.
@@
 We now define the class of interpretations for $\thD$.
-$\interp{\J}{\tau}$ denotes the interpretation type $\tau$ in $\J$---that is, a set of
-domain elements for that type.
+$\interp{\J}{\tau}$ denotes the interpretation type $\tau$ in $\J$---that is, a
+nonempty set of domain elements for that type.
 $\interp{\J}{\const{f}}$ denotes the interpretation of a function $\const{f}$ in $\J$.
@@
 unique up to $\alpha$-equivalence and acyclic if $\tau \not\in \Codata$.
+\pagebreak[2] %% TYPESETTING
 \item
@@
 that interprets wrongly applied selectors based on equality information in $\Fc$
-and that interprets the variables of $\Fc$ based on the completion $\ValC$ of the mapping $\Val$ from phase~2.
-For the latter, let $\interp{\J}{x}$ be the value in $\J$ that is $\alpha$-equivalent with $\ValC \ec{x}$ for each variable $x \in \Terms(\Fc)$,
+and that interprets the variables of $\Fc$ based on the completion $\ValC$.
+% of %the mapping $\Val$. % from phase~2.
+For the variables, let $\interp{\J}{x\typ{\tau}}$ be the value in $\J(\tau)$ that is $\alpha$-equivalent with $\ValC \ec{x}$ for each variable $x \in \Terms(\Fc)$,
 which by Lemma~\ref{lem:interpretation-completion}\PART{1} is guaranteed to exist.
 
-We first show $\J$ satisfies all equalities in $t_1 \teq t_2 \in \Fc$, where $t_1$ and $t_2$ have type $\tau$.
-We show,
-by structural induction on $t$,
+We first show that $\J$ satisfies all equalities $t_1 \teq t_2 \in \Fc$. % where $t_1$ and $t_2$ have type $\tau$.
+To achieve this, we show
+by structural induction on $t\typ\tau$
 that $\interp{\J}{t} \aequiv \ValC \ec{t}$ for all terms $t \in \Terms( \Fc )$,
-which since $\J$ is normal, implies $\J \models t_1 \teq t_2$.
+which implies $\J \models t_1 \teq t_2$ since $\J$ is normal.
 
@@
 which by the induction hypothesis is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \ValC \ec{ u_1 }, \ldots, \ValC \ec{ u_n } )}$.
-Call this term $t_u$.
+Call this term $t'$.
 %By Lemma~\ref{lem:interpretation-completion}\PART{2} and due to our construction of $\ValC$, we have $\ValC \ec{t}$ is a term
@@
 %where $\expand{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$ for $i \in [n]$.
-Since \rn{Inject} and \rn{Clash} do not apply, % to $\Fc$,
-by the construction of $\ValC$ we have $\ValC \ec{t}$ is a term
-of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$
+Since \rn{Inject} and \rn{Clash} do not apply to $\Fc$,
+by the construction of $\ValC$ we have that $\ValC \ec{t}$ is a term
+of the form $\MU {\Varec{t}}.\; \const{C}( w_1,\allowbreak \ldots,\allowbreak w_n )$
 where
@@
 $\expand{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$.
-For each $i$, let $u_\iPrime$ be the $i$\vvthinspace th argument of $t_u$.
+For each $i$, let $u_\iPrime$ be the $i$\vvthinspace th argument of $t'$.
 \begin{rep}By Lemma~\ref{lem:mu-norm-arg},\end{rep}%
 \begin{conf}Clearly,\end{conf} %% TODO: Fix! ;)
-$\expand{t_u}{u_\iPrime} \vsim \ValC \ec{u_i}$.
-Thus $\expand{t_u}{u_\iPrime} \vsim \expand{\ValC \ec{t}}{ w_i }$. % for each $i \in [n]$.
-%By Lemma~\ref{lem:mu-cong}, we have $t_u \vsim t_w$.
-Thus, $\interp{\J}{t} \vsim t_u \vsim \ValC \ec{t}$, and
+$\expand{t'}{u_\iPrime} \vsim \ValC \ec{u_i}$.
+Thus $\expand{t'}{u_\iPrime} \vsim \expand{\ValC \ec{t}}{ w_i }$. % for each $i \in [n]$.
+%By Lemma~\ref{lem:mu-cong}, we have $t' \vsim t_w$.
+Thus, $\interp{\J}{t} \vsim t' \vsim \ValC \ec{t}$, and
 we have $\interp{\J}{t} \vsim \ValC \ec{t}$.

### 2015-02-19 tex/body.tex
minor

@@
 Since by assumption $\Fc$ contains only equalities and disequalities, we have $\J \models \Fc$,
-and since $\Fc$ is a superset of $\Ec$, we have that $\Ec$ is satisfied by $\J$.
+and since $\Ec \subseteq \Fc$,
+% is a superset of $\Ec$,
+we have that $\Ec$ is satisfied by $\J$.
 \qed

### 2015-02-19 tex/body.tex
more

@@
 inspired by the $\mu$-calculus
-\begin{conf}\cite{kozen-1983}\end{conf}%
-\begin{rep}\cite[etc.]{kozen-1983,endrullis-et-al-2011}\end{rep}.
+%\begin{conf}\cite{kozen-1983}.\end{conf}%
+%\begin{rep}\cite[etc.]{kozen-1983,endrullis-et-al-2011}.\end{rep}%
+\cite[etc.]{kozen-1983,endrullis-et-al-2011}.
 
@@
 %\paragraph{Integration into {\relax\XXXL($T\negvthinspace$)}.}
+{\looseness=-1
 The calculus is implemented as a \relax{theory solver} of CVC4, that is, a
@@
 Given a theory $T = T_1 \mathrel\cup \cdots \mathrel\cup T_n$ and a set of
-input clauses $F$ in CNF, the \XXXL($T$) procedure
+input clauses $F$ in conjunctive normal form, the \XXXL($T$) procedure
 incrementally builds partial assignments from the atomic formulas of $F$ to
@@
 %the value $\bot$.
-Then, %at a high level,
-by a variant
+%at a high level,
+By a variant
 \cite{jovanovic2011sharing} of the Nelson--Oppen method,
@@
 containing symbols not belonging to $T_i$ are replaced by fresh variables\begin{rep} of the
-same type\end{rep}; (2)~additional equalities and disequalities between variables of
+same type\end{rep}; (2)~additional (dis)equalities between variables of
 types not belonging to $T_i$.
-Each $T_i$ solver
+Each $T_i$-solver
 either reports that a subset $C$ of $M_i$ is $T_i$-unsatisfiable, in which case
@@
 nothing only if $M_i$ is indeed $T_i$-satisfiable.
+
+}
 
@@
 
+\kern-0.7pt %% TYPESETTING: hack
+
 \section{Evaluation on Isabelle Problems}
 \label{sec:experimental-results}
+
+\kern-0.3pt %% TYPESETTING: hack
 
@@
 and the \emph{Archive of Formal Proofs} (AFP, 2974 goals) \cite{klein-et-al-afp}
-that define codatatypes falling within the supported fragment. We added
-two unpublished theories by Peter Gammie and Andreas Lochbihler (\gandl, 317
+that define codatatypes falling within the supported fragment. \hfill We added
+two
+
+\pagebreak %%% TYPESETTING: hack (including \hfill)
+
+\noindent
+unpublished theories by Peter Gammie and Andreas Lochbihler (\gandl, 317
 goals), about Bird and Stern--Brocot trees. To exercise the support

### 2015-02-19 tex/body.tex
too much space
more

@@
 
-\newcommand\betweenantes{\kern1.25em}
+\newcommand\betweenantes{\kern1.125em}
 
@@
 
-\nopagebreak
-
 Datatypes and codatatypes are an integral part of many proof assistants,
@@
 Closely related are the automatic structural
-induction in both types of provers \cite{kersani-peltier-2013,cruanes-201x,wand-weidenbach-201x,reynolds-kuncak-2015},
+induction in both kinds of provers \cite{kersani-peltier-2013,cruanes-201x,wand-weidenbach-201x,reynolds-kuncak-2015},
 the (co)datatype and (co)induction support in Dafny \cite{leino-moskal-2014},
@@
 it is not necessary to extend the signature:
-We let $\const{d}_{i\negvthinspace j}(t)$
+We can simply let $\const{d}_{i\negvthinspace j}(t)$
 abbreviate
@@
 \text{Selection:}\quad
-  & \smash{\const{s}_{i\negvthinspace j}^{\,k}(\const{C}_{i\negvthinspace j}(x_1,\ldots,x_{n_{i\negvthinspace j}})) = x_k}
+  & \smash{\const{s}_{i\negvthinspace j}^{\,k}(\const{C}_{i\negvthinspace j}(x_1,\ldots,x_{n_{i\negvthinspace j}})) \teq x_k}
 \end{aligned}
@@
 \centering
-\begin{tabular}{c}
+\begin{tabular}{@{}c@{}}
 \(
@@
 \rn{Acyclic}
-\\[4\jot]
-\(
+\kern1.5em\( %%% TYPESETTING: should be \quad
 \inferrule{
@@
 unique up to $\alpha$-equivalence and acyclic if $\tau \not\in \Codata$.
-\pagebreak[2] %% TYPESETTING
 \item
@@
 
-%%% TYPESETTING: Evil hack!!
-%We first state the following properties of $\ValC$,
 In the following lemma about $\ValC$,
-$\vphantom{\Bigl(_j}\muvar( t ) = \smash{\begin{cases}
+$\vphantom{\Bigl(_j}\muvar( t ) = \relax{\begin{cases}
   t & \text{if $t$ is a variable} \\[-\jot]
@@
 
-\pagebreak[2]
-
 Intuitively, this lemma states three properties of $\ValC$ that ensure a normal
@@
 and the \emph{Archive of Formal Proofs} (AFP, 2974 goals) \cite{klein-et-al-afp}
-that define codatatypes falling within the supported fragment. \hfill We added
-two
-
-\pagebreak %%% TYPESETTING: hack (including \hfill)
-
-\noindent
-unpublished theories by Peter Gammie and Andreas Lochbihler (\gandl, 317
+that define codatatypes falling within the supported fragment. We added
+two unpublished theories by Peter Gammie and Andreas Lochbihler (\gandl, 317
 goals), about Bird and Stern--Brocot trees. To exercise the support
@@
 three are simple proofs of the form
-\keyw{by}~\textit{coinduction}~\textit{auto} %\cite{blanchette-et-al-2014-impl}
-in Isabelle.
+\keyw{by}~\textit{coinduction}~\textit{auto}
+in Isabelle \cite{blanchette-et-al-2014-impl}.
 The fourth proof, by Gammie and Lochbihler, is more elaborate:

### 2015-02-19 tex/body.tex
reintroduced para

@@
 would apply \rn{Split} and \rn{Inject} before adding this equality.
-\begin{rep}  % lower priority to mention this
+
 To reduce the number of unique constraints considered by the calculus, we
@@
 based on whether $j = \jPrime$.
-\end{rep}
 
@@
 
-\kern-0.7pt %% TYPESETTING: hack
-
 \section{Evaluation on Isabelle Problems}
 \label{sec:experimental-results}
-
-\kern-0.3pt %% TYPESETTING: hack
 

### 2015-02-19 tex/body.tex
spell-checker found two real mistakes

@@
 with $\const{p}(\ldots)$ abbreviating $\const{p}(\ldots) \teq \const{true}$\end{rep}.
-\begin{rep}The operator $\bigwedge_{\,i}\, \varphi_i$ abbeviates a conjunction
+\begin{rep}The operator $\bigwedge_{\,i}\, \varphi_i$ abbreviates a conjunction
 $\varphi_1 \mathrel\land \cdots \mathrel\land \varphi_n$. \end{rep}%
@@
 %In the first step, we compute the bidirectional closure of $\Ec$;
-%in the second step, we make inferences based on cyclicity and uniqueness;
+%in the second step, we make inferences based on acyclicity and uniqueness;
 %and in the third step, %(when necessary),
@@
 %, or informally are \emph{observationally equivalent}.
-%For convienience, we use $\mu$-terms to refer to (classes of) values for both codatatype and datatype terms,
+%For convenience, we use $\mu$-terms to refer to (classes of) values for both codatatype and datatype terms,
 %where in the latter case, a $\mu$-term with a bound variable denotes an infeasible term.
@@
 such that $\const{C}(w,y) \in \ec{x}$ and $\const{C}(z,x) \in \ec{y}$ for some $\const{C} \in \Ctr$.
-A possible sequence of unfoldings is given below, omitting
+A possible sequence of unfolding steps is given below, omitting
 trivial entries such as $\ec{w} \mapsto {\Varec{w}}$.
@@
 We now show that $\Val = \nf{\Val}$.
-Assume by contradition $\Val \ec{x} \neq \nf{\Val \ec{x}}$ for some $\Val \ec{x}$ of minimal size.
+Assume by contradiction $\Val \ec{x} \neq \nf{\Val \ec{x}}$ for some $\Val \ec{x}$ of minimal size.
 We have that $\Val \ec{x}$ is of the form $\MU {\Varec{x}}.\; \const{C}( t_1, \ldots, t_n )$.
@@
   & 209 & 221 & 777 & 775 & 51 & 52 & 1037 & 1048 \\%[\jot]
-%Datatype simplication only
+%Datatype simplification only
 %  &  \\
@@
   & 213 & 227 & 791 & 786 & 51 & 52 & 1055 & 1065 \\%[\jot]
-%Codatatype simplication only
+%Codatatype simplification only
 %  &  \\

### 2015-02-19 tex/body.tex
minor

@@
 \noindent
-of infinite sequences, or streams, over an element type $\tau$ is allowed\begin{rep}; the
+of infinite sequences or streams over an element type $\tau$ is
+allowed\begin{conf}, with no base case\end{conf}\begin{rep}; the
 corresponding datatype would be rejected as non-well-founded
@@
 An evaluation on %hand-crafted examples and on
-problems generated from Isabelle \cite{nipkow-et-al-2002} theories using the Sledgehammer tool \cite{paulson-blanchette-2010}
+problems generated from Isabelle theories using the Sledgehammer tool \cite{paulson-blanchette-2010}
 demonstrates the usefulness of the approach (Section~\ref{sec:experimental-results}).
@@
 which collapses to $\const{f} : \tau$ if $n = 0$\end{rep}.
-The notation $t\typ{\tau}$ indicates that term $t$ has type~$\tau$.
 \begin{rep}%
@@
 $\const{true},\, \const{false} : \ty{bool}$, interpreted as truth values.
-The only predicate is equality ($\teq$)\begin{rep} and belongs to the logical symbols\end{rep}.
-Other predicates can be represented as functions to $\ty{bool}$\begin{rep},
+The only predicate is equality ($\teq$)\begin{rep} and belongs to the logical symbols\end{rep};
+other predicates can be represented as functions to $\ty{bool}$\begin{rep},
 with $\const{p}(\ldots)$ abbreviating $\const{p}(\ldots) \teq \const{true}$\end{rep}.
+The notation $t\typ{\tau}$ indicates that term $t$ has type~$\tau$.
 \begin{rep}The operator $\bigwedge_{\,i}\, \varphi_i$ abbreviates a conjunction

### 2015-02-19 tex/body.tex
more

@@
 $\Sel^\delta$.
-
+%
 For types with several constructors, it is customary to provide discriminators
-$\const{d}_{i\negvthinspace j} : \delta_i \to \ty{bool}$. However,
-it is not necessary to extend the signature:
-We can simply let $\const{d}_{i\negvthinspace j}(t)$
+$\const{d}_{i\negvthinspace j} : \delta_i \to \ty{bool}$. Instead of extending $\Funcs$,
+we let $\const{d}_{i\negvthinspace j}(t)$
 abbreviate
@@
 associated dependency graph is strongly connected.
+A type is (\emph{co})\emph{recursive} if it belongs to such a set of types.
 \begin{rep}%
@@
 %
-In addition, non(co)recursive specifications
-such~as
+Non(co)recursive type specifications such~as
 \begin{conf}%
-\vthinspace$\keyw{datatype}~\vthinspace\ty{option}_{\,\tau} \,=\, \const{None} \,\mid\, \const{Some}(\tau)$
+\vthinspace $\keyw{datatype}~\vthinspace\ty{option}_{\,\tau} \,=\, \const{None} \,\mid\, \const{Some}(\tau)$
 \end{conf}%
@@
 %
-The axiom schema ensures that the interpretation of datatypes
+The schema ensures that the interpretation of datatypes
 is standard.
@@
 \emph{corecursive singletons}. For the decision procedure, it will be
-crucial to detect these types. %, even if they rarely arise in practice.
+crucial to detect these. %, even if they rarely arise in practice.
 A type may also be a corecursive singleton only in some models. If the example

### 2015-02-19 tex/body.tex
more

@@
 $\thD$-satisfiability of finite sets $\Ec$ of constraints:\ equalities
-and disequalities between ground $\Sig$-terms. The procedure is formulated as a
-calculus.
+and disequalities between ground $\Sig$-terms. The
+procedure is formulated as a calculus.
 % whose derivation rules are applied until saturation or contradiction.
@@
 are ground, they cannot entail an upper
-bound on the cardinality of any such type, so it is safe to consider them infinite.
+bound on the cardinality of any such type, so it is safe to consider these types infinite.
 
@@
 preconditions are met.
-The conclusion either describes equalities to be added to $\Ec$
+The conclusion either specifies equalities to be added to $\Ec$
 or is $\bot$. % (contradiction).

### 2015-02-19 tex/body.tex
more

@@
 \usepackage[scaled=.82]{beramono}
+\usepackage[scaled=.9]{helvet}
 \usepackage{booktabs}
@@
 \vthinspace$\keyw{codatatype}~\ty{a} = \const{A}(\ty{a})$, whose
-unique value is $\MU a.\; A(a)$. Other specimens are
+unique value is $\MU a.\; \const{A}(a)$. Other specimens are
 $\ty{stream}_{\,\ty{unit}\,}$
@@
 Formally, $\mu$-\emph{terms} are defined recursively as being either a variable $x$
-or a node of the form
-$\MU x.\: \const{C}( \vec t\, )$ for some constructor $\const{C} \in \Ctr$ and
+or an applied constructor
+$\MU x.\: \const{C}( \vec t\, )$ for some $\const{C} \in \Ctr$ and
 $\mu$-terms $\vec t$ of the expected types.

### 2015-02-19 tex/body.tex
fonts

@@
 \makeatother
+
+%\newcommand\ourparagraph[1]{\paragraph{\upshape\textbf{#1}}}
+%\newcommand\beforefig{}
+
+\newcommand\ourparagraph[1]{\paragraph{{#1}}}
+\newcommand\beforefig{\rm}
 
@@
 
-\begin{rep}\paragraph{Related Work.} Despite\end{rep}%
+\begin{rep}\ourparagraph{Related Work.} Despite\end{rep}%
 \begin{conf}Despite\end{conf} its age,
@@
 
-\paragraph{Conventions.}
+\ourparagraph{Conventions.}
 The setting is a monomorphic (or many-sorted) first-order logic.
@@
 
-%\paragraph{Specifications.}
+%\ourparagraph{Specifications.}
 In an SMT problem, the signature is typically given by specifying first the
@@
 
-%\paragraph{Characterization.}
+%\ourparagraph{Characterization.}
 One way to define datatypes semantically is as the initial model of the
@@
 
-%\paragraph{Special Cases.}
+%\ourparagraph{Special Cases.}
 {\looseness=-1
@@
 
-\paragraph{Phase 1: Computing the Bidirectional Closure \rm(\figurename~\ref{fig:cc-rules}).}
+\ourparagraph{Phase 1: Computing the Bidirectional Closure \beforefig(\figurename~\ref{fig:cc-rules}).}
 In conjunction with \rn{Refl}, \rn{Sym}, and \rn{Trans}, the \rn{Cong} rule computes the (upward) congruence closure,
@@
 
-\paragraph{Phase 2: Applying Acyclicity and Uniqueness \rm(\figurename~\ref{fig:ab-rules}).}
+\ourparagraph{Phase 2: Applying Acyclicity and Uniqueness \beforefig(\figurename~\ref{fig:ab-rules}).}
 %For presentation of the rules in this phase, we rely on a representation of terms in the $\mu$-notation.
@@
 
-\paragraph{Phase 3: Branching \rm(\figurename~\ref{fig:split-rule}).}
+\ourparagraph{Phase 3: Branching \beforefig(\figurename~\ref{fig:split-rule}).}
 If a selector is applied to a term $t$, or if $t$'s type is a finite datatype,
@@
 
-\paragraph{Correctness.}
+\ourparagraph{Correctness.}
 Correctness means that if there exists a closed derivation tree with root node
@@
 
-%%% @ANDY: I'm trying without \paragraph's, to save space but also to make the
+%%% @ANDY: I'm trying without \ourparagraph's, to save space but also to make the
 %%% section look more like a whole than a random list of stuff. Revert if you
@@
 
-%\paragraph{Integration into {\relax\XXXL($T\negvthinspace$)}.}
+%\ourparagraph{Integration into {\relax\XXXL($T\negvthinspace$)}.}
 {\looseness=-1
@@
 
-%\paragraph{Extension to Quantified Formulas.}
+%\ourparagraph{Extension to Quantified Formulas.}
 While the decision procedure is restricted to ground constraints,
@@
 
-%\paragraph{Optimizations.}
+%\ourparagraph{Optimizations.}
 The implementation of the decision procedure uses several optimizations
@@
 \def\ackname{Acknowledgment}
-\paragraph{\ackname.}
+\ourparagraph{\ackname.}
 We owe a great debt to the development team of CVC4, 

### 2015-02-19 tex/body.tex
more

@@
 models the natural numbers $\const{Z}$, $\const{S}(\const{Z})$, $\const{S}(\const{S}(\const{Z}))$, $\ldots$\afterLdots{},
-using Peano notation, but extended with an
+using Peano notation but extended with an
 infinite value $\infty = \const{S}(\const{S}(\const{S}(\ldots)))$.
@@
 combination (Section~\ref{sec:the-ground-decision-procedure}).
-The procedure is described abstractly as a calculus and is
-designed to be composable via the Nelson--Oppen method \cite{nelson-oppen-1979}.
+The procedure is described abstractly as a calculus and can be
+composed via the Nelson--Oppen method \cite{nelson-oppen-1979}.
 It generalizes the procedure by Barrett et al.\ \cite{barrett-et-al-2007}, which only covers
@@
 
-Second, \emph{a uniqueness rule replaces the acyclicity rule of datatypes.}
+Second, \emph{a uniqueness rule takes the place of the acyclicity rule of datatypes.}
 Cyclic constraints such as
@@
 injectivity afterward. Injectivity is important to satisfy disequality constraints.
-For infinite codatatypes, it is always possible to find suitable values of $v$.
+For infinite codatatypes, it is always possible to find suitable values $v$. For example,
+if the longest cycle in $\Terms^{\smash{\J}}_{\Varec{x}}(\ValC)$ is of length $n$, a fresh value
+can be constructed with a cycle of length $n + 1$.
 
@@
 %
-To construct a completion $\ValC$, we must choose values for the free
-variables of $\Val$, specifically ${\Varec{y}}$ and ${\Varec{z}}$.
+To construct a completion $\ValC$, we must choose values for ${\Varec{y}}$ and ${\Varec{z}}$,
+which are free in~$\Val$.
 Modulo $\alpha$-equivalence, $\smash{\Terms^{\smash{\J}}_{{\Varec{z}}}( \Val )} =
@@
 $.
-Consider a normal interpretation $\J$ that evaluates variables in $\Ec$ based on the mapping $\Val$:
-$z$ is interpreted as $\Val \ec{z}$, $y$ as $\Val \ec{y}$, and so on.
+Consider a normal interpretation $\J$ that evaluates variables in $\Ec$ based on~$\Val$:
+$\interp{\J}{u} = \Val \ec{u}$, $\interp{\J}{v} = \Val \ec{v}$, and so on.
 Assigning a value for $\Val \ec{z}$ that is $\alpha$-equivalent to a value in $\Terms^{\smash{\J}}_{{\Varec{z}}}( \Val )$
@@
 the values in the range of $\Val$ remain $\alpha$-disequivalent.
-For example, we may assign a value such as
-$\MU {\Varec{z}}.\; \const{E}( {\Varec{z}} )$ or
-$\MU {\Varec{z}}.\; \const{D}( \const{C}( {\Varec{z}} ) )$
+We can assign a value such as
+$\MU {\Varec{z}}.\; \const{E}( {\Varec{z}} )$,
+$\MU {\Varec{z}}.\; \const{D}( \const{C}( {\Varec{z}} ) )$, or
+$\MU {\Varec{z}}.\; \const{C}( \const{C}( \const{D}( {\Varec{z}} ) ) )$
 to ${\Varec{z}}$.

### 2015-02-19 tex/body.tex
swapped subscript and superscript (cf. =_\alpha^x)

@@
 For a variable $x\typ{\tau}$ and a normal interpretation $\J$,
-we let $\Terms^{\smash{\J}}_x( \Val )$ denote the set consisting of all values
+we let $\Terms_{\smash{\J}}^x( \Val )$ denote the set consisting of all values
 $v \in \interp{\J}{\tau}$
@@
   \betweenantes
-  v \not\in \Terms^{\smash{\J}}_{\Varec{x}}(\ValC)
+  v \not\in \Terms_{\smash{\J}}^{\Varec{x}}(\ValC)
 }{
@@
 Given an unassigned variable in $\ValC$, this rule assigns it a fresh
-value---one that does not occur in $\Terms^{\smash{\J}}_{\Varec{x}}(\ValC)$
+value---one that does not occur in $\Terms_{\smash{\J}}^{\Varec{x}}(\ValC)$
 modulo $\alpha$-equivalence---excluding
@@
 %of codatatypes, of which only finitely many are in
-%$\Terms^{\smash{\J}}_{\Varec{x}}(\ValC)$?)}
+%$\Terms_{\smash{\J}}^{\Varec{x}}(\ValC)$?)}
 Since this update removes one
@@
 For infinite codatatypes, it is always possible to find suitable values $v$. For example,
-if the longest cycle in $\Terms^{\smash{\J}}_{\Varec{x}}(\ValC)$ is of length $n$, a fresh value
+if the longest cycle in $\Terms_{\smash{\J}}^{\Varec{x}}(\ValC)$ is of length $n$, a fresh value
 can be constructed with a cycle of length $n + 1$.
@@
 which are free in~$\Val$.
-Modulo $\alpha$-equivalence, $\smash{\Terms^{\smash{\J}}_{{\Varec{z}}}( \Val )} =
+Modulo $\alpha$-equivalence, $\smash{\Terms_{\smash{\J}}^{{\Varec{z}}}( \Val )} =
 %\begin{array}{c}
@@
 $\interp{\J}{u} = \Val \ec{u}$, $\interp{\J}{v} = \Val \ec{v}$, and so on.
-Assigning a value for $\Val \ec{z}$ that is $\alpha$-equivalent to a value in $\Terms^{\smash{\J}}_{{\Varec{z}}}( \Val )$
+Assigning a value for $\Val \ec{z}$ that is $\alpha$-equivalent to a value in $\Terms_{\smash{\J}}^{{\Varec{z}}}( \Val )$
 may cause values in the range of $\Val$ to become $\alpha$-equivalent,
@@
 However, this contradicts the disequality $z \not\teq v$ in~$\Ec$.
-On the other hand, if the value assigned to ${\Varec{z}}$ is fresh, %not $\alpha$-equivalent to any term in $\Terms^{\smash{\J}}_{{\Varec{z}}}( \Val )$,
+On the other hand, if the value assigned to ${\Varec{z}}$ is fresh, %not $\alpha$-equivalent to any term in $\Terms_{\smash{\J}}^{{\Varec{z}}}( \Val )$,
 the values in the range of $\Val$ remain $\alpha$-disequivalent.
@@
 there is only one equivalence class of type $\tau$ in $\Ec$,
-and thus there are no terms in $\Terms^{\smash{\J}}_{\Varec{y}}(\ValC)$ of type $\tau$.
+and thus there are no terms in $\Terms_{\smash{\J}}^{\Varec{y}}(\ValC)$ of type $\tau$.
 This is a contradiction, since %our %model
@@
 This is also a contradiction,
-since there are only a finite number of closed terms in $\Terms^{\smash{\J}}_{\Varec{y}}(\ValC)$,
+since there are only a finite number of closed terms in $\Terms_{\smash{\J}}^{\Varec{y}}(\ValC)$,
 and thus %our %model
-completion can assign a value not occurring in $\Terms^{\smash{\J}}_{\Varec{y}}(\ValC)$ to ${\Varec{y}}$.
+completion can assign a value not occurring in $\Terms_{\smash{\J}}^{\Varec{y}}(\ValC)$ to ${\Varec{y}}$.
 By construction, $\ValC \ec{x}$ is normal.
@@
 ${\Varec{x}} \in FV( \Val_1 )$,
-and $\MU {\Varec{x}}.\; t$ is not $\alpha$-equivalent to a term in $\Terms^{\smash{\J}}_{\Varec{x}}(\Val_1)$.
+and $\MU {\Varec{x}}.\; t$ is not $\alpha$-equivalent to a term in $\Terms_{\smash{\J}}^{\Varec{x}}(\Val_1)$.
 To show (2),
@@
 Since $\expand{\Val_1 \ec{y}}{t_1}$ is normal,
-there is a closed $\mu$-term $v \in \Terms^{\smash{\J}}_{\Varec{x}}( \Val_1 )$
+there is a closed $\mu$-term $v \in \Terms_{\smash{\J}}^{\Varec{x}}( \Val_1 )$
 such that $v \vsimv{\Varec{x}} \expand{\Val_1 \ec{y}}{t_1}$.

### 2015-02-19 tex/body.tex
more
whitespace
avoid ugly abbrev

@@
 %  &  \\
-Datatypes w/o \rn{Acyclic}
+Datatypes without \rn{Acyclic}
   & -- & 227 & -- & 780 & -- & 52 & -- & 1059 \\
@@
 %  &  \\
-Codatatypes w/o \rn{Unique}
+Codatatypes without \rn{Unique}\!\!
   & -- & 222 & -- & 804 & -- & 56 & -- & 1082 \\
@@
 interactive theorem proving. As future work, it would be useful to implement
-proof reconstruction for (co)data\-type inferences in Isabelle. We also
-want to try out CVC4 for model finding in Isabelle, as an alternative to
-Nitpick. Finally, it might be worthwhile to extend SMT solvers with dedicated
-support for (co)recursive functions.
+proof reconstruction for (co)data\-type inferences in Isabelle.
+CVC4 cound be interfaced for generating counterexamples in proof assistants, as
+an alternative to Nitpick. Finally, it might be worthwhile to extend SMT solvers
+with dedicated support for (co)recursive functions.
 

### 2015-02-20 tex/body.tex
todo
update 1st para of chap 4

@@
-%%% TODO: Clarify that \mu-terms appear in generated models
+%%% TODO: Archive!!
 %%% TODO: In report, give small example where Single is needed
@@
 
-The decision procedure for (co)datatypes and its extension to quantified
-formulas were presented at a high level of abstraction. This section describes
+The ground decision procedure for (co)datatypes
+was presented at a high level of abstraction. This section describes
 the main aspects of the implementation within the SMT solver CVC4:\ the
 integration into CDCL($T$), the extension to quantified formulas, and
-some optimizations.
+some of the optimizations.
 

### 2015-02-20 tex/body.tex
more

@@
 {\looseness=-1
-The calculus is implemented as a \relax{theory solver} of CVC4, that is, a
+The decision procedure is implemented as a \relax{theory solver} of CVC4, that is, a
 specialized solver for determining the satisfiability of conjunctions of
@@
 input clauses $F$ in conjunctive normal form, the \XXXL($T$) procedure
-incrementally builds partial assignments from the atomic formulas of $F$ to
-truth values such that no clause in $F$ is falsified. We can regard such a
+incrementally builds partial assignments of
+truth values to the atoms of $F$ such that no clause in $F$ is falsified. We can regard such a
 partial assignment as a set $M$ of true literals.
@@
 %at a high level,
-By a variant
-\cite{jovanovic2011sharing} of the Nelson--Oppen method,
+By a variant of the Nelson--Oppen method \cite{jovanovic2011sharing,nelson-oppen-1979},
 each $T_i$-solver %for $i = 1, \ldots, l$  n?
 takes as input the union $M_i$
-of (1)~the purified form of $T_i$-constraints occurring in $M$, where terms
-containing symbols not belonging to $T_i$ are replaced by fresh variables\begin{rep} of the
-same type\end{rep}; (2)~additional (dis)equalities between variables of
-types not belonging to $T_i$.
-Each $T_i$-solver
+of (1)~the purified form of $T_i$-constraints occurring in $M$, where fresh
+variables replace terms containing symbols not belonging to~$T_i$;
+(2)~additional (dis)equalities between variables of types not belonging to
+$T_i$. Each $T_i$-solver
 either reports that a subset $C$ of $M_i$ is $T_i$-unsatisfiable, in which case

### 2015-02-20 tex/body.tex
more edits

@@
 For all equivalence classes $\ec{z}$,
-each pair of constructor terms $\const{C}_{\negvthinspace j}( \vec t )$ and $\const{C}_{\negvthinspace\jPrime}( \vec u )$ in $\ec{z}$
+each pair of constructor terms $\const{C}_{\negvthinspace j}( \vec t \,)$ and $\const{C}_{\negvthinspace\jPrime}( \vec u )$ in $\ec{z}$
 are such that $j = \jPrime$ since \rn{Clash} does not apply, % to $\Ec$,
@@
 and $\nf{\Val_1 \sigma} \ec{z}$ (resp. $\nf{\Val_1 \sigma} \ec{y}$) does not.
-If $t_1$ is of the form $\MU w_1.\; \const{C}( \vec t )$
+If $t_1$ is of the form $\MU w_1.\; \const{C}( \vec t \,)$
 and $t_2$ is of the form $\MU w_2.\; \const{D}( \vec u )$,
@@
 Since $t_1$ and $t_2$ are minimal,
-say $t_1$ is of the form $\MU w.\; \const{C}( \vec t )$,
+say $t_1$ is of the form $\MU w.\; \const{C}( \vec t \,)$,
 and $t_2$ is ${\Varec{x}}$.
@@
 
-However, the presence of quantified formulas poses an additional challenge in the context of (co)datatypes:
-Since quantified formulas may entail cardinality constraints on uninterpreted types $\ty{u}$,
-we may no longer assume that they are of infinite cardinality.
-When assuming that a finite type has infinite cardinality, the calculus
+However, the presence of quantified formulas poses an additional challenge in
+the context of (co)datatypes. Quantified formulas may entail an upper bound on
+the cardinality of an uninterpreted type $\ty{u}$.
+When assuming that $\ty{u}$ has infinite cardinality, the calculus
 presented in this paper is \relax{incomplete}
@@
 since %by assumption
-the solver is already incomplete when quantified formulas are present.
+the solver is already incomplete in the presence of quantified formulas.
 Nonetheless, two techniques help increase the precision of the solver.
@@
 Second, we can conditionally apply \rn{Single} to codatatype terms that may have cardinality one.
-For example, consider the $\ty{stream}_{\,\ty{u}}$ codatatype, where $\ty{u}$ is uninterpreted.
-This type has cardinality precisely when $\ty{u}$ has cardinality one.
+For example, the $\ty{stream}_{\,\ty{u}}$ codatatype %, where $\ty{u}$ is uninterpreted.
+has cardinality precisely when $\ty{u}$ has cardinality one.
 If there exist two equivalence classes $\ec{s}$ and $\ec{t}$ for this type,
@@
 replace $t \teq s$ by $s \teq t$ if $s$ is less than $t$ based on a term
-ordering, replace $\const{C}_{\negvthinspace j}( \vec t ) \teq
-\const{C}_{\negvthinspace\jPrime}( \vec u )$ with $\bot$ when $j \neq \jPrime$,
-replace all selector terms of the form $\const{s}^{\,k}_{\negvthinspace j}(
-\const{C}_{\negvthinspace j}(t_1,\ldots,t_n) )$ by $t_k$, and replace
-occurrences of discriminators $\const{d}_{\negvthinspace j}(
-\const{C}_{\negvthinspace\jPrime}( \vec t ) )$ by $\top$ or $\bot$
+ordering, replace $\smash{\const{C}_{\negvthinspace j}( \vec t \,) \teq
+\const{C}_{\negvthinspace\jPrime}( \vec u )}$ with $\bot$ when $j \neq \jPrime$,
+replace all selector terms of the form $\smash{\const{s}^{\,k}_{\negvthinspace j}(
+\const{C}_{\negvthinspace j}(t_1,\ldots,t_n) )}$ by $t_k$, and replace
+occurrences of discriminators $\smash{\const{d}_{\negvthinspace j}(
+\const{C}_{\negvthinspace\jPrime}( \vec t \,) )}$ by $\top$ or $\bot$
 based on whether $j = \jPrime$.
@@
 two unpublished theories by Peter Gammie and Andreas Lochbihler (\gandl, 317
-goals), about Bird and Stern--Brocot trees. To exercise the support
-for datatypes, the benchmarks were complemented by theories about lists
-and trees. The theories were selected before running the
+goals), about Bird and Stern--Brocot trees. To exercise the datatype-specific
+part of the decision procedure, the benchmarks were complemented by theories
+about lists and trees. The theories were selected before running the
 experiments. The experimental data are publicly available.%

### 2015-02-20 tex/body.tex
more

@@
 
-By Theorems~\ref{thm:t},~\ref{thm:rs}, and~\ref{thm:ss}, the calculus is sound and complete for $\thD$.
+By Theorems~\ref{thm:t},~\ref{thm:rs}, and~\ref{thm:ss}, the
+calculus is sound and complete for the ground $\thD$ theory, and we can
+rightly call it a decision procedure.
 
@@
 We
-included all the theory files from the Isabelle distribution (Distro, 879 goals)
-and the \emph{Archive of Formal Proofs} (AFP, 2974 goals) \cite{klein-et-al-afp}
+included all the theory files from the Isabelle distribution (Distro, 879~goals)
+and the \emph{Archive of Formal Proofs} (AFP, 2974~goals) \cite{klein-et-al-afp}
 that define codatatypes falling within the supported fragment. We added
@@
 
-For each goal in each theory file, Sledgehammer was invoked to select
-256~lemmas as axioms, which were then monomorphized and translated to SMT-LIB~2
+For each goal in each theory, Sledgehammer was invoked to select up to
+256~lemmas, which were monomorphized and translated to SMT-LIB~2
 along with the goal \cite{blanchette-et-al-2013-smt}. The resulting problem was
@@
 constructors and selectors, which the decision procedure would recognize as
-tautologies. Z3 is included for comparison. % when this is possible.
+tautologies. Z3 is included for comparison when this makes sense.
 %This allows us to answer the question, \relax{What are the
@@
 \begin{center}\begin{tabular}{@{\;}l@{\kern1.5em}cc@{\kern.5em}cc@{\kern.5em}cc@{\kern1.3em}cc@{\;}}
-  & \multicolumn{2}{c}{Distro} & \multicolumn{2}{c}{AFP} & \multicolumn{2}{c}{\gandl} & \multicolumn{2}{c}{Overall} \\
+  & \multicolumn{2}{c}{Distro} & \multicolumn{2}{c}{AFP} & \multicolumn{2}{c}{\gandl} & \multicolumn{2}{c}{Overall} \\[-0.5pt]
   & \HD{Z3} & \HD{CVC4} & \HD{Z3} & \HD{CVC4} & \HD{Z3} & \HD{CVC4} & \HD{Z3} & \HD{CVC4}
@@
 The results are summarized in Table~\ref{tab:bench}. The decision procedure
-makes a difference across all theories. It accounts for an overall success rate
+makes a difference across all three benchmark suites. It accounts for an overall success rate
 increase of over \OK{5\%}. Moreover, every aspect of the procedure, including the

### 2015-02-20 tex/body.tex
more

@@
 with
-$\smash{\const{C}_{i\negvthinspace j} : \tau_{i\negvthinspace j}^1\times\cdots\times\tau_{i\negvthinspace j}^{\,k_{\smash{i\negvthinspace j}}} \to \delta_i}$
+$\smash{\const{C}_{i\negvthinspace j} : \tau_{i\negvthinspace j}^1\times\cdots\times\tau_{i\negvthinspace j}^{\,n_{\smash{i\negvthinspace j}}} \to \delta_i}$
 and $\smash{\const{s}_{i\negvthinspace j}^{\,k} : \delta_i \to \tau_{i\negvthinspace j}^{\,k}}$.
@@
 abbreviate
-$t \teq \const{C}_{i\negvthinspace j}\bigl(\const{s}_{i\negvthinspace j}^1(x), \ldots, \const{s}_{i\negvthinspace j}^{\,n_{\smash{{i\negvthinspace j}}}}(x)\bigr)$.
+$t \teq \const{C}_{i\negvthinspace j}\bigl(\const{s}_{i\negvthinspace j}^1(t), \ldots, \const{s}_{i\negvthinspace j}^{\,n_{\smash{{i\negvthinspace j}}}}(t)\bigr)$.
 %This will simplify the presentation. % of the decision procedure.
@@
 datatypes and codatatypes},
-which in addition to $\Sig$ defines a class of $\Sig$-interpretations $\JJJJ$,
+which %in addition to $\Sig$
+defines a class of $\Sig$-interpretations $\JJJJ$,
 namely the ones that satisfy the %axioms --- existence of all infinite codatatype values cannot be axiomatized
-properties mentioned in this section, including (co)induction. In particular, the interpretations
-in~$\JJJJ$ share the same fixed interpretation for constructor terms and correctly
-applied selector terms but may differ on wrongly applied
-selector terms and on variables.
-A formula $\varphi$ is $\thD$-\emph{satisfiable} if there
-exists an interpretation in $\JJJJ$ that satisfies~$\varphi$.
+properties mentioned in this section, including (co)induction. The interpretations
+in~$\JJJJ$ share the same interpretation for constructor terms and correctly
+applied selector terms (modulo renaming of domain elements) but may differ on
+variables and wrongly applied selector terms.
+A formula is $\thD$-\emph{satisfiable} if there
+exists an interpretation in $\JJJJ$ that satisfies it.
 % otherwise, it is $\thD$-unsatisfiable. (I trust the readers' intelligence when
@@
 {\looseness=-1
-Some codatatypes are so degenerate as to have infinite values
-even though they are finite. A simple example is
+Some codatatypes are so degenerate as be finite even though they
+have infinite values. A simple example is
 \vthinspace$\keyw{codatatype}~\ty{a} = \const{A}(\ty{a})$, whose

### 2015-02-20 tex/body.tex
more

@@
 in~$\JJJJ$ share the same interpretation for constructor terms and correctly
-applied selector terms (modulo renaming of domain elements) but may differ on
+applied selector terms (up to renaming of domain elements) but may differ on
 variables and wrongly applied selector terms.
@@
 Andrei Popescu helped clarify our thoughts regarding the axiomatization of
-codatatypes. Dmitriy Traytel took part in discussions about degenerate
-codatatypes.
-%
-Andreas Lochbihler and Mark Summerfield suggested several textual improvements.
-The second author's research was partially supported by the Deutsche
-Forschungs\-gemein\-schaft (DFG) project
-\relax{Hardening the Hammer} (grant Ni\,491\slash 14-1).
+codatatypes and indicated related work. Dmitriy Traytel took part in discussions
+about degenerate codatatypes.
+%
+Andreas Lochbihler, Mark Summerfield, and Dmitriy Traytel suggested several
+textual improvements. The second author's research was partially supported by
+the Deutsche Forschungs\-gemein\-schaft (DFG) project \relax{Hardening the
+Hammer} (grant Ni\,491\slash 14-1).
 

### 2015-02-20 tex/body.tex
more

@@
 \usepackage[scaled=.82]{beramono}
-\usepackage[scaled=.9]{helvet}
+\usepackage[scaled=.95]{helvet}
 \usepackage{booktabs}
@@
 Three proofs were found thanks to
-the \rn{Acyclic} rule and four required the \rn{Unique} rule. Among the latter,
+the acyclicity rule and four required uniqueness. Among the latter,
 three are simple proofs of the form
@@
 CVC4 cound be interfaced for generating counterexamples in proof assistants, as
-an alternative to Nitpick. Finally, it might be worthwhile to extend SMT solvers
-with dedicated support for (co)recursive functions.
+an alternative to Nitpick; in this context, the acyclicity and uniqueness
+rules are essential to exclude spurious models. Finally, it might be
+worthwhile to extend SMT solvers with dedicated support for (co)recursive
+functions.
 

### 2015-02-20 tex/body.tex
more

@@
 uniqueness rule that identifies observationally equal values, also in the presence of cyclic
-($\omega$-regular) data. The procedure decides ground and universal
+(regular) data. The procedure decides ground and universal
 problems and is composable via the Nelson--Oppen method. It has been
@@
 models the natural numbers $\const{Z}$, $\const{S}(\const{Z})$, $\const{S}(\const{S}(\const{Z}))$, $\ldots$\afterLdots{},
-using Peano notation but extended with an
+in Peano notation but extended with an
 infinite value $\infty = \const{S}(\const{S}(\const{S}(\ldots)))$.
@@
 
-Third, \emph{it must be possible to express cyclic }(\emph{$\omega$-regular}) \emph{values as closed terms and
+Third, \emph{it must be possible to express cyclic }(\emph{regular}) \emph{values as closed terms and
 to enumerate them.} This is necessary both for finite model finding (modulo theories)
@@
 \Val \ec{x} & \,=\, \MU {\Varec{x}}.\; \const{S}(\Varec{x}) &
-\enskip %% TYPESETTING
+\kern0.25em %% TYPESETTING
 \Val \ec{y} & \,=\,
 \MU {\Varec{y}}.\; \const{S}(\smash{\MU {\VAREC{\const{S}(y)}}}.\; \const{S}( {\Varec{y}} ) )
-\enskip %% TYPESETTING
+\kern0.25em %% TYPESETTING
  &
@@
 $.
-Consider a normal interpretation $\J$ that evaluates variables in $\Ec$ based on~$\Val$:
+Now consider a normal interpretation $\J$ that evaluates variables in $\Ec$ based on~$\Val$:
 $\interp{\J}{u} = \Val \ec{u}$, $\interp{\J}{v} = \Val \ec{v}$, and so on.

### 2015-02-20 tex/body.tex
more

@@
 an alternative to Nitpick; in this context, the acyclicity and uniqueness
-rules are essential to exclude spurious models. Finally, it might be
+rules are crucial to exclude spurious models. Finally, it might be
 worthwhile to extend SMT solvers with dedicated support for (co)recursive

### 2015-02-20 tex/body.tex
more

@@
 For example, the $\ty{stream}_{\,\ty{u}}$ codatatype %, where $\ty{u}$ is uninterpreted.
-has cardinality precisely when $\ty{u}$ has cardinality one.
+has cardinality one precisely when $\ty{u}$ has cardinality one.
 If there exist two equivalence classes $\ec{s}$ and $\ec{t}$ for this type,
@@
 compute a normal form for constraints as a preprocessing step. In particular, we
-replace $t \teq s$ by $s \teq t$ if $s$ is less than $t$ based on a term
+replace $u \teq t$ by $t \teq u$ if $t$ is less than $u$ based on a term
 ordering, replace $\smash{\const{C}_{\negvthinspace j}( \vec t \,) \teq
@@
 The results are summarized in Table~\ref{tab:bench}. The decision procedure
-makes a difference across all three benchmark suites. It accounts for an overall success rate
-increase of over \OK{5\%}. Moreover, every aspect of the procedure, including the
+makes a difference across all three benchmark suites. It accounts for an overall
+success rate increase of over \OK{5\%}, which is quite significant. The
+practical usefulness of interpreted (co)datatypes appears roughly comparable to
+that of arithmetic \cite{blanchette-et-al-2013-smt}.
+The raw data also reveals that the
+stronger decision procedures almost completely subsume the weaker ones.
+
+Moreover, every aspect of the procedure, including the
 more expensive rules, make a contribution.
@@
 \const{Node}\;0\;\const{num}\;\const{\nummodden}$.
-
-The raw data also reveals that the
-stronger decision procedures almost completely subsume the weaker ones. For
-Sledgehammer, the practical usefulness of interpreted (co)datatypes appears
-roughly comparable to that of arithmetic \cite{blanchette-et-al-2013-smt}.
 
@@
 %
-Andreas Lochbihler, Mark Summerfield, and Dmitriy Traytel suggested several
-textual improvements. The second author's research was partially supported by
-the Deutsche Forschungs\-gemein\-schaft (DFG) project \relax{Hardening the
-Hammer} (grant Ni\,491\slash 14-1).
+Pascal Fontaine, Andreas Lochbihler, Mark Summerfield, and Dmitriy Traytel
+suggested several textual improvements. The second author's research was
+partially supported by the Deutsche Forschungs\-gemein\-schaft (DFG) project
+\relax{Hardening the Hammer} (grant Ni\,491\slash 14-1).
 

### 2015-02-20 tex/body.tex
more

@@
   & \multicolumn{2}{c}{Distro} & \multicolumn{2}{c}{AFP} & \multicolumn{2}{c}{\gandl} & \multicolumn{2}{c}{Overall} \\[-0.5pt]
-  & \HD{Z3} & \HD{CVC4} & \HD{Z3} & \HD{CVC4} & \HD{Z3} & \HD{CVC4} & \HD{Z3} & \HD{CVC4}
+  & \HD{CVC4} & \HD{Z3} & \HD{CVC4} & \HD{Z3} & \HD{CVC4} & \HD{Z3} & \HD{CVC4}& \HD{Z3} 
 \MIDRULE
 No (co)datatypes
-  & 209 & 221 & 777 & 775 & 51 & 52 & 1037 & 1048 \\%[\jot]
+  & 221 & 209 & 775 & 777 & 52 & 51 & 1048 & 1037 \\%[\jot]
 %Datatype simplification only
@@
 Datatypes without \rn{Acyclic}
-  & -- & 227 & -- & 780 & -- & 52 & -- & 1059 \\
+  & 227 & -- & 780 & -- & 52 & -- & 1059 & -- \\
 Full datatypes
-  & 213 & 227 & 791 & 786 & 51 & 52 & 1055 & 1065 \\%[\jot]
+  & 227 & 213 & 786 & 791 & 52 & 51 & 1065 & 1055 \\%[\jot]
 %Codatatype simplification only
@@
 Codatatypes without \rn{Unique}\!\!
-  & -- & 222 & -- & 804 & -- & 56 & -- & 1082 \\
+  & 222 & -- & 804 & -- & 56 & -- & 1082 & -- \\
 Full codatatypes
-  & -- & 223 & -- & 804 & -- & \bfseries 59 & -- & 1086 \\%[\jot]
+  & 223 & -- & 804 & -- & \bfseries 59 & -- & 1086 & -- \\%[\jot]
 Full (co)datatypes
-  & -- & \bfseries 229 & -- & \bfseries 815 & -- & \bfseries 59 & -- & \bfseries 1103% \\[\jot]
+  & \bfseries 229 & -- & \bfseries 815 & -- & \bfseries 59 & -- & \bfseries 1103 & -- %\\[\jot]
 \end{tabular}\end{center}

### 2015-02-20 tex/body.tex
more

@@
 
-CVC4 was run on each problem several times, with the support for datatypes and
-codatatypes either enabled or disabled. The contributions of the acyclicity and
-uniqueness rules were also measured, to find out whether these somewhat
-expensive rules are useful in practice. Even when the decision procedure is
-disabled, some of the generated axioms typically include basic properties of
-constructors and selectors, which the decision procedure would recognize as
-tautologies. Z3 is included for comparison when this makes sense.
+The development version of CVC4 was run on each problem several times, with the
+support for datatypes and codatatypes either enabled or disabled. The
+contributions of the acyclicity and uniqueness rules were also measured, to find
+out whether these somewhat expensive rules are useful in practice. Even when the
+decision procedure is disabled, some of the generated axioms typically include
+basic properties of constructors and selectors, which the decision procedure
+would recognize as tautologies. Z3 version 4.3.2 is included for comparison.
+% when this makes sense.
 %This allows us to answer the question, \relax{What are the
@@
 makes a difference across all three benchmark suites. It accounts for an overall
-success rate increase of over \OK{5\%}, which is quite significant. The
-practical usefulness of interpreted (co)datatypes appears roughly comparable to
-that of arithmetic \cite{blanchette-et-al-2013-smt}.
-The raw data also reveals that the
-stronger decision procedures almost completely subsume the weaker ones.
+success rate increase of over \OK{5\%}, which is quite significant. For Isabelle
+users, the practical usefulness of interpreted (co)datatypes appears roughly
+comparable to that of arithmetic \cite{blanchette-et-al-2013-smt}. The raw data
+also reveals that the stronger decision procedures almost fully subsume the weaker
+ones.
 

### 2015-02-20 tex/body.tex
otherk ind of l
more

@@
 %
-Each (co)datatype specification consists of $l$~mutually recursive types that are
+Each (co)datatype specification consists of $\ell$~mutually recursive types that are
 either all datatypes or all codatatypes. Polymorphic types, nested
@@
 
-\newcommand\elll{\kern.18ex l\kern.11ex}
-\newcommand\elllx{\kern.11ex l\kern.18ex}
+\newcommand\elll{\kern.05ex \ell\kern.05ex}
+\newcommand\elllx{\kern.05ex \ell\kern.05ex}
 
@@
 \!(\keyw{co})\keyw{datatype}\;\,
-  \delta_1 & {}= \smash{\const{C}_{11\!}(\bigl[\const{s}_{11\!}^1{:}\bigr]\vthinspace \tau_{11\!}^1, \ldots, \bigl[\const{s}_{11\!}^{n_{11\!}}{:}\bigr]\vthinspace \tau_{11\!}^{n_{11\!}})} \mid \cdots \mid \smash{\const{C}_{1m_1\!}(\ldots)} \\[-.5\jot]
+  \delta_1 & {}=\, \smash{\const{C}_{11\!}(\bigl[\const{s}_{11\!}^1{:}\bigr]\vthinspace \tau_{11\!}^1, \ldots, \bigl[\const{s}_{11\!}^{n_{11\!}}{:}\bigr]\vthinspace \tau_{11\!}^{n_{11\!}})} \mid \cdots \mid \smash{\const{C}_{1m_1\!}(\ldots)} \\[-.5\jot]
    \smash{\vdots\,\,\,} \\[-1.5\jot]
@@
 an alternative to Nitpick; in this context, the acyclicity and uniqueness
-rules are crucial to exclude spurious models. Finally, it might be
+rules are crucial to exclude spurious countermodels. Finally, it might be
 worthwhile to extend SMT solvers with dedicated support for (co)recursive

### 2015-02-20 tex/body.tex
more

@@
 differences.
-
+%
 First, \emph{codatatypes need not be well-founded.}
@@
 \cite{blanchette-et-al-2015-esop}\end{rep}.
-
+%
 Second, \emph{a uniqueness rule takes the place of the acyclicity rule of datatypes.}
@@
 $\const{y} \teq \const{S}(\const{y})$, it deduces $\const{x} \teq \const{y}$.
-The acyclicity and uniqueness rules are needed to ensure completeness on
-ground problems. %and the absence of spurious models.
+The acyclicity and uniqueness rules are needed to ensure completeness
+(model soundness) on ground problems. %and the absence of spurious models.
 They cannot be expressed as finite axiomatizations, so they naturally belong in
@@
 %axiom.
-
+%
 Third, \emph{it must be possible to express cyclic }(\emph{regular}) \emph{values as closed terms and
-to enumerate them.} This is necessary both for finite model finding (modulo theories)
+to enumerate them.} This is necessary both for finite model finding
+%modulo theories
 and for theory combinations. The $\mu$-binder notation associates a name with

### 2015-02-20 tex/body.tex
shaped is back

@@
 the $\mu$-term $\const{SCons}(1,\: \MU s.\; \const{SCons}(0,\: \const{SCons}(9,\: s)))$
-stands for the lasso\begin{rep}-shaped\end{rep} sequence $1, 0, 9, 0, 9, 0, 9,
+stands for the lasso-shaped sequence $1, 0, 9, 0, 9, 0, 9,
 \ldots$\afterLdots.

### 2015-02-20 tex/body.tex
more

@@
 $\thD$-satisfiable.
-The proof relies on a specific model $\J$ of $\thD$ that satisfies $\Ec$.
+The proof relies on a specific interpretation $\J$ that satisfies $\Ec$.
 
@@
 $\interp{\J}{\const{f}}$ denotes the interpretation of a function $\const{f}$ in $\J$.
-If $\const{f}$ is of type $ \tau_1 \times \cdots \times \tau_n \to \tau$,
+If $\const{f} : \tau_1 \times \cdots \times \tau_n \to \tau$,
 then $\interp{\J}{\const{f}}$ is a total function from $\interp{\J}{\tau_1} \times \cdots \times \interp{\J}{\tau_n}$ to $\interp{\J}{\tau}$.

### 2015-02-20 tex/body.tex
commented out junk
added some stuff

@@
 \usepackage[scaled=.82]{beramono}
-\usepackage[scaled=.95]{helvet}
+\usepackage[scaled=.88]{helvet}
 \usepackage{booktabs}
@@
 \newcommand{\tcEc}{\Terms^\AST(\Ec)}
-\newcommand{\rn}[1]{\textsf{\small #1}}
-\newcommand{\cvc}{\textsc{cvc}{\small 4}\xspace}
+\newcommand{\rn}[1]{\textsf{#1}}
 \newcommand{\teq}{\approx}
@@
 injectivity afterward. Injectivity is important to satisfy disequality constraints.
-For infinite codatatypes, it is always possible to find suitable values $v$. For example,
-if the longest cycle in $\Terms_{\smash{\J}}^{\Varec{x}}(\ValC)$ is of length $n$, a fresh value
-can be constructed with a cycle of length $n + 1$.
+For infinite codatatypes, it is always possible to find suitable values $v$.
+%For example,
+%if the longest cycle in $\Terms_{\smash{\J}}^{\Varec{x}}(\ValC)$ is of length $n$, a fresh value
+%can be constructed with a cycle of length $n + 1$.
 
@@
 
-Prior to constructing a model for $F\!$, the theory solver does not explicitly construct $\mu$-terms nor the mapping $\Val$.
+Before constructing a model for $F\!$, the theory solver constructs neither $\mu$-terms nor the mapping $\Val$.
 Instead, $\Val$ is computed implicitly by traversing the equivalence classes of $\Ec$ during phase~2.
@@
 the procedure considers each equivalence class $\ec{t}$ containing a datatype constructor $\const{C}( t_1, \ldots, t_n )$.
-It visits $\ec{t_1}, \ldots, \ec{t_n}$ and all arguments of constructors in these equivalence classes recursively.
+It visits $\ec{t_1}, \ldots, \ec{t_n}$ and all constructor arguments in these equivalence classes recursively.
 If while doing so it returns to $\ec{t}$, it deduces that \rn{Acyclic} applies. % to $t$.
@@
 the procedure considers the set $\SSS$ of all codatatype equivalence classes.
-It simultaneously visits the equivalence classes of arguments of constructor terms in each equivalence class in $\SSS$,
+It\begin{conf}%
+\pagebreak
+\noindent
+\end{conf}
+simultaneously visits the equivalence classes of arguments of constructor terms in each equivalence class in $\SSS$,
 while partitioning $\SSS$ into $\SSS_1, \ldots, \SSS_n$ based on the top-most symbol of constructor terms in these equivalence classes
@@
 When assuming that $\ty{u}$ has infinite cardinality, the calculus
-presented in this paper is \relax{incomplete}
+presented in Section~\ref{sec:the-ground-decision-procedure} is \relax{incomplete}
 since it may fail to recognize cases where \rn{Split} and \rn{Single} should be
@@
 the implementation adds the clause %of the form
-$(\exists x\;y\typ{\ty{u}}.\;\, x \not\teq y ) \mathrel\vee s \teq t$ to $F$.
+$(\exists x\;y\typ{\ty{u}}.\;\, x \not\teq y ) \mathrel\vee s \teq t$ to $F\!$.
 \begin{rep}In other words, either the cardinality of $\ty{u}$ is greater than one or $s$ is equal to $t$.\end{rep}
@@
 
+The efficiency properties of the procedure are mostly as described in Barrett et
+al.\ \cite{barrett-et-al-2007}. In particular, it is theoretically and
+empirically advantageous to delay applications of \rn{Split} as long as
+possible. Similarly, \rn{Acyclic} and \rn{Unique} are fairly expensive because
+they require traversing the equivalence classes, which is why they are part of
+phase~2.
+
 %\section{Examples}
@@
 two unpublished theories by Peter Gammie and Andreas Lochbihler (\gandl, 317
-goals), about Bird and Stern--Brocot trees. To exercise the datatype-specific
-part of the decision procedure, the benchmarks were complemented by theories
+goals), about Bird and Stern--Brocot trees. To exercise the datatype
+support, the benchmarks were complemented by theories
 about lists and trees. The theories were selected before running the
@@
 support for datatypes and codatatypes either enabled or disabled. The
-contributions of the acyclicity and uniqueness rules were also measured, to find
-out whether these somewhat expensive rules are useful in practice. Even when the
-decision procedure is disabled, some of the generated axioms typically include
-basic properties of constructors and selectors, which the decision procedure
-would recognize as tautologies. Z3 version 4.3.2 is included for comparison.
+contributions of the acyclicity and uniqueness rules were also measured.
+% to find
+% out whether these somewhat expensive rules are useful in practice.
+Even when the
+decision procedure is disabled, the problems may contain
+basic lemmas about constructors and selectors, allowing some (co)datatype
+reasoning. Z3 4.3.2 is included for comparison.
 % when this makes sense.

### 2015-02-20 tex/body.tex
more stuff

@@
   & %\forall x_1,\ldots,x_{n_{i\negvthinspace j}} y.\;\,
-    \smash{\const{C}_{i\negvthinspace j}(x_1,\ldots,x_{n_{i\negvthinspace j}}) \teq \const{C}_{i\negvthinspace j}(x_1,\ldots,x_{k-1},y,x_{k+1},\ldots,x_{n_{i\negvthinspace j}}) \longrightarrow x_k \teq y}
+    \smash{\const{C}_{i\negvthinspace j}(x_1,\ldots,x_{n_{i\negvthinspace j}}) \teq \const{C}_{i\negvthinspace j}(y_1,\ldots,y_{n_{i\negvthinspace j}}) \longrightarrow x_k \teq y_k}
   \\[-1\jot]
@@
 the procedure considers the set $\SSS$ of all codatatype equivalence classes.
-It\begin{conf}%
-\pagebreak
-\noindent
-\end{conf}
+It
+%
+\begin{conf}\goodbreak\noindent\end{conf}%
 simultaneously visits the equivalence classes of arguments of constructor terms in each equivalence class in $\SSS$,
 while partitioning $\SSS$ into $\SSS_1, \ldots, \SSS_n$ based on the top-most symbol of constructor terms in these equivalence classes
-and the equivalence of their arguments of type $\Nondata$.
-It then partitions each set in this partition in the same manner recursively.
+and the equivalence of their arguments of ordinary types.
+It then partitions each set %in this partition %in the same manner
+recursively.
 If %after doing so,
 the resulting partition contains a set $\SSS_i$ containing distinct terms $u$
-and $v$, it deduces that \rn{Unique} applies to $u$ and $v$.
+and $v$, it deduces that \rn{Unique} applies to $u$ and~$v$.
 

### 2015-02-20 tex/body.tex
Revise paragraph on intuition behind bad set.

@@
 %not occurring in $\tEc$,
-and set $\Val\ec{u} := {\Varec{u}}$. Therefore, there are initially no
+and set $\Val\ec{u} := {\Varec{u}}$, that is to say there are initially no
 constraints on the values for any equivalence class $\ec{u}$. The mapping $\Val$
@@
 
-The last precondition is admittedly rather technical. Its role is to ensure that
-$\ValC$ remains injective modulo $\alpha$-equivalence. A simpler, less efficient
-way to proceed would be to compute $\nf{\ValC [ {\Varec{x}} \mapsto \MU {\Varec{x}}.\; t ]}$ optimistically and check
-injectivity afterward. Injectivity is important to satisfy disequality constraints.
-For infinite codatatypes, it is always possible to find suitable values $v$.
+%The last precondition is admittedly rather technical.
+Ensuring disequality constraints are satisfied by an interpretation based on $\ValC$
+relies on showing that $\ValC$ is injective modulo $\alpha$-equivalence.
+The last precondition in the above rule ensures that $\ValC$ remains injective after each substitution.
+More precisely, the set $\Terms_{\smash{\J}}^{\Varec{x}}(\ValC)$ is an overapproximation of the values
+that, when assigned to ${\Varec{x}}$, cause values in the range of $\ValC$ to become $\alpha$-equivalent.
+This precondition is not strictly necessary in practice, 
+as one could compute $\nf{\ValC [ {\Varec{x}} \mapsto \MU {\Varec{x}}.\; t ]}$ optimistically and check
+injectivity afterward, repeating until a suitable value is found.
+For infinite codatatypes, it is always possible to find suitable values $v$,
+due to the fact that $\Terms_{\smash{\J}}^{\Varec{x}}(\ValC)$ is a finite set.
+%Injectivity is important to satisfy disequality constraints.
+%For infinite codatatypes, it is always possible to find suitable values $v$.
 %For example,

### 2015-02-20 tex/body.tex
more

--- 2015-02-20 tex/body.tex
@@
 
-\newcommand\DISC[1]{t \teq \const{C}_{#1}\bigl(\const s^1_{#1}( t ), \ldots,\const s^{n_{#1}}_{#1}( t )\bigr)}
+\newcommand\DISC[1]{t \teq \const{C}_{#1}\negvthinspace\bigl(\const s^1_{#1}( t ), \ldots,\const s^{n_{#1}}_{#1}( t )\bigr)}
 %\newcommand\DISC[1]{\const{d}_{#1}(t)}
@@
 \hyphenation{data-type data-types co-data-type co-data-types isa-belle sledge-hammer
-non-redun-dant}
+non-redun-dant counter-examples counter-example}
 
@@
 
-\begin{rep}\ourparagraph{Related Work.} Despite\end{rep}%
-\begin{conf}Despite\end{conf} its age,
-Barrett et al.\ \cite{barrett-et-al-2007} still provides a good account of
+\begin{rep}\ourparagraph{Related Work.} Barrett\end{rep}%
+\begin{conf}Barrett\end{conf}
+et al.\ \cite{barrett-et-al-2007} provide a good account of
 related work. Since then, datatypes have been added to the SMT solver
@@
 %\ourparagraph{Integration into {\relax\XXXL($T\negvthinspace$)}.}
-{\looseness=-1
 The decision procedure is implemented as a \relax{theory solver} of CVC4, that is, a
@@
 % where $a \in M$ for all atoms $a$ that
-%it assigned the value $\top$ and $\neg\; a \in M$ for all atoms that it assigns
+%it assigned the value $\top$ and $\neg\, a \in M$ for all atoms that it assigns
 %the value $\bot$.
@@
 either reports that a subset $C$ of $M_i$ is $T_i$-unsatisfiable, in which case
-$\neg\; C$ is added to $F\!$, adds a clause to $F\!$, or does nothing.
+$\neg\, C$ is added to $F\!$, adds a clause to $F\!$, or does nothing.
 When $M$ is a complete assignment for $F\!$, a theory solver can choose to do
 nothing only if $M_i$ is indeed $T_i$-satisfiable.
-
-}
 
@@
 experiments. The experimental data are publicly available.%
-\footnote{\url{http://www21.in.tum.de/~blanchet/codp_data.tgz}}
+\footnote{\url{http://lara.epfl.ch/~reynolds/CADE2015-cdt/}}
+%http://www21.in.tum.de/~blanchet/codp_data.tgz}}
 
@@
 contribution has been the support for codatatypes. Both the metatheory and
-the implementation in CVC4 rely on $\mu$-terms to represent cyclic values,
-a complication specific to codatatypes.
+the implementation in CVC4 rely on $\mu$-terms to represent cyclic values.
+Although this aspect is primarily motivated by codatatypes, it enables
+a uniform account of datatypes and codatatypes---in particular,
+the acyclicity rule for datatypes exploits $\mu$-terms to detect cycles.
 % are not very
@@
 proof reconstruction for (co)data\-type inferences in Isabelle.
-CVC4 cound be interfaced for generating counterexamples in proof assistants, as
+CVC4 could also be interfaced for generating counterexamples in proof assistants, as
 an alternative to Nitpick; in this context, the acyclicity and uniqueness

### 2015-02-20 tex/body.tex
due to the fact that -> because (or since)
fit again

@@
 that, when assigned to ${\Varec{x}}$, cause values in the range of $\ValC$ to become $\alpha$-equivalent.
-This precondition is not strictly necessary in practice, 
-as one could compute $\nf{\ValC [ {\Varec{x}} \mapsto \MU {\Varec{x}}.\; t ]}$ optimistically and check
-injectivity afterward, repeating until a suitable value is found.
+%This precondition is not strictly necessary in practice, 
+%as one could compute $\nf{\ValC [ {\Varec{x}} \mapsto \MU {\Varec{x}}.\; t ]}$ optimistically and check
+%injectivity afterward, repeating until a suitable value is found.
 For infinite codatatypes, it is always possible to find suitable values $v$,
-due to the fact that $\Terms_{\smash{\J}}^{\Varec{x}}(\ValC)$ is a finite set.
+because $\Terms_{\smash{\J}}^{\Varec{x}}(\ValC)$ is a finite set.
 %Injectivity is important to satisfy disequality constraints.

### 2015-02-20 tex/body.tex
more

@@
 
+\newcommand\BAD{\mathcalx{V}}
 \newcommand\Terms{\mathcalx{T}}
@@
 For a variable $x\typ{\tau}$ and a normal interpretation $\J$,
-we let $\Terms_{\smash{\J}}^x( \Val )$ denote the set consisting of all values
+we let $\BAD_{\smash{\J}}^x( \Val )$ denote the set consisting of all values
 $v \in \interp{\J}{\tau}$
@@
   \betweenantes
-  v \not\in \Terms_{\smash{\J}}^{\Varec{x}}(\ValC)
+  v \not\in \BAD_{\smash{\J}}^{\Varec{x}}(\ValC)
 }{
@@
 Given an unassigned variable in $\ValC$, this rule assigns it a fresh
-value---one that does not occur in $\Terms_{\smash{\J}}^{\Varec{x}}(\ValC)$
+value---one that does not occur in $\BAD_{\smash{\J}}^{\Varec{x}}(\ValC)$
 modulo $\alpha$-equivalence---excluding
@@
 %of codatatypes, of which only finitely many are in
-%$\Terms_{\smash{\J}}^{\Varec{x}}(\ValC)$?)}
+%$\BAD_{\smash{\J}}^{\Varec{x}}(\ValC)$?)}
 Since this update removes one
@@
 The last precondition in the above rule ensures that $\ValC$ remains injective after each substitution.
-More precisely, the set $\Terms_{\smash{\J}}^{\Varec{x}}(\ValC)$ is an overapproximation of the values
+More precisely, the set $\BAD_{\smash{\J}}^{\Varec{x}}(\ValC)$ is an overapproximation of the values
 that, when assigned to ${\Varec{x}}$, cause values in the range of $\ValC$ to become $\alpha$-equivalent.
@@
 For infinite codatatypes, it is always possible to find suitable values $v$,
-because $\Terms_{\smash{\J}}^{\Varec{x}}(\ValC)$ is a finite set.
+because $\BAD_{\smash{\J}}^{\Varec{x}}(\ValC)$ is a finite set.
 %Injectivity is important to satisfy disequality constraints.
@@
 %For example,
-%if the longest cycle in $\Terms_{\smash{\J}}^{\Varec{x}}(\ValC)$ is of length $n$, a fresh value
+%if the longest cycle in $\BAD_{\smash{\J}}^{\Varec{x}}(\ValC)$ is of length $n$, a fresh value
 %can be constructed with a cycle of length $n + 1$.
@@
 which are free in~$\Val$.
-Modulo $\alpha$-equivalence, $\smash{\Terms_{\smash{\J}}^{{\Varec{z}}}( \Val )} =
+Modulo $\alpha$-equivalence, $\smash{\BAD_{\smash{\J}}^{{\Varec{z}}}( \Val )} =
 %\begin{array}{c}
@@
 $\interp{\J}{u} = \Val \ec{u}$, $\interp{\J}{v} = \Val \ec{v}$, and so on.
-Assigning a value for $\Val \ec{z}$ that is $\alpha$-equivalent to a value in $\Terms_{\smash{\J}}^{{\Varec{z}}}( \Val )$
+Assigning a value for $\Val \ec{z}$ that is $\alpha$-equivalent to a value in $\BAD_{\smash{\J}}^{{\Varec{z}}}( \Val )$
 may cause values in the range of $\Val$ to become $\alpha$-equivalent,
@@
 However, this contradicts the disequality $z \not\teq v$ in~$\Ec$.
-On the other hand, if the value assigned to ${\Varec{z}}$ is fresh, %not $\alpha$-equivalent to any term in $\Terms_{\smash{\J}}^{{\Varec{z}}}( \Val )$,
+On the other hand, if the value assigned to ${\Varec{z}}$ is fresh, %not $\alpha$-equivalent to any term in $\BAD_{\smash{\J}}^{{\Varec{z}}}( \Val )$,
 the values in the range of $\Val$ remain $\alpha$-disequivalent.
@@
 there is only one equivalence class of type $\tau$ in $\Ec$,
-and thus there are no terms in $\Terms_{\smash{\J}}^{\Varec{y}}(\ValC)$ of type $\tau$.
+and thus there are no terms in $\BAD_{\smash{\J}}^{\Varec{y}}(\ValC)$ of type $\tau$.
 This is a contradiction, since %our %model
@@
 This is also a contradiction,
-since there are only a finite number of closed terms in $\Terms_{\smash{\J}}^{\Varec{y}}(\ValC)$,
+since there are only a finite number of closed terms in $\BAD_{\smash{\J}}^{\Varec{y}}(\ValC)$,
 and thus %our %model
-completion can assign a value not occurring in $\Terms_{\smash{\J}}^{\Varec{y}}(\ValC)$ to ${\Varec{y}}$.
+completion can assign a value not occurring in $\BAD_{\smash{\J}}^{\Varec{y}}(\ValC)$ to ${\Varec{y}}$.
 By construction, $\ValC \ec{x}$ is normal.
@@
 ${\Varec{x}} \in FV( \Val_1 )$,
-and $\MU {\Varec{x}}.\; t$ is not $\alpha$-equivalent to a term in $\Terms_{\smash{\J}}^{\Varec{x}}(\Val_1)$.
+and $\MU {\Varec{x}}.\; t$ is not $\alpha$-equivalent to a term in $\BAD_{\smash{\J}}^{\Varec{x}}(\Val_1)$.
 To show (2),
@@
 Since $\expand{\Val_1 \ec{y}}{t_1}$ is normal,
-there is a closed $\mu$-term $v \in \Terms_{\smash{\J}}^{\Varec{x}}( \Val_1 )$
+there is a closed $\mu$-term $v \in \BAD_{\smash{\J}}^{\Varec{x}}( \Val_1 )$
 such that $v \vsimv{\Varec{x}} \expand{\Val_1 \ec{y}}{t_1}$.

### 2015-02-20 tex/body.tex
more

@@
 %The last precondition is admittedly rather technical.
-Ensuring disequality constraints are satisfied by an interpretation based on $\ValC$
-relies on showing that $\ValC$ is injective modulo $\alpha$-equivalence.
-The last precondition in the above rule ensures that $\ValC$ remains injective after each substitution.
-More precisely, the set $\BAD_{\smash{\J}}^{\Varec{x}}(\ValC)$ is an overapproximation of the values
-that, when assigned to ${\Varec{x}}$, cause values in the range of $\ValC$ to become $\alpha$-equivalent.
+To ensure disequality constraints are satisfied by an interpretation based on $\ValC$,
+it suffices that $\ValC$ is injective modulo $\alpha$-equivalence.
+This invariant holds initially, and
+the last precondition in the above rule ensures that it is maintained.
+%More precisely,
+The set $\BAD_{\smash{\J}}^{\Varec{x}}(\ValC)$ is an overapproximation of the values
+that, when assigned to ${\Varec{x}}$, will cause values in the range of $\ValC$ to become $\alpha$-equivalent.
 %This precondition is not strictly necessary in practice, 
@@
 %injectivity afterward, repeating until a suitable value is found.
-For infinite codatatypes, it is always possible to find suitable values $v$,
+For infinite codatatypes, it is always possible to find fresh values $v$
 because $\BAD_{\smash{\J}}^{\Varec{x}}(\ValC)$ is a finite set.

### 2015-02-20 tex/body.tex
some fixes

@@
 A $\mu$-term is \emph{closed} if %and only if --- by convention, "and only if" is not necessary for *definitions* (but add it back if you disagree)
-it contains no free variables. \,It is \emph{cyclic} if %and only if --- ditto
-it contains a bound variable. \,The
-\goodbreak %%% TYPESETTING: horrible hacks (including two \,'s)
-\noindent
-$\alpha$-\emph{equivalence} relation $t \aequiv u$
+it contains no free variables. It is \emph{cyclic} if %and only if --- ditto
+it contains a bound variable. The $\alpha$-\emph{equivalence} relation $t \aequiv u$
 indicates that the $\mu$-terms $t$ and $u$
@@
 To recognize when the precondition of \rn{Unique} holds,
-the procedure considers the set $\SSS$ of all codatatype equivalence classes.
-It
-%
+the procedure considers the set $\SSS$ of all codatatype equivalence classes. It
 \begin{conf}\goodbreak\noindent\end{conf}%

### 2015-02-22 tex/body.tex
reintroduced spaces

@@
 a natural choice is to turn to coinductive datatypes, or \emph{codatatypes},
-the non-well-founded dual of inductive \emph{datatypes}.
+the non-well-founded dual of inductive \emph{datatypes.}
 %
@@
     \smash{\const{C}_{i\negvthinspace j}(\bar x) \tneq \const{C}_{i\negvthinspace \jPrime}(\bar y) \quad\text{if $j \not= \jPrime$}}
-  \\[-1\jot]
+  \\[-0.5\jot]
 \text{Injectivity:}\quad
@@
     \smash{\const{C}_{i\negvthinspace j}(x_1,\ldots,x_{n_{i\negvthinspace j}}) \teq \const{C}_{i\negvthinspace j}(y_1,\ldots,y_{n_{i\negvthinspace j}}) \longrightarrow x_k \teq y_k}
-  \\[-1\jot]
+  \\[-0.5\jot]
 \text{Exhaustiveness:}\quad
   & \smash{\const{d}_{i1}(x) \mathrel\lor \cdots \mathrel\lor \const{d}_{im_i}(x)}
-  \\[-1\jot]
+  \\[-0.5\jot]
 \text{Selection:}\quad

### 2015-02-22 tex/body.tex
more

@@
 composed via the Nelson--Oppen method \cite{nelson-oppen-1979}.
-It generalizes the procedure by Barrett et al.\ \cite{barrett-et-al-2007}, which only covers
-datatypes.
+It generalizes the procedure by Barrett et al.\ \cite{barrett-et-al-2007}, which covers
+only datatypes.
 To our knowledge, our procedure is the first of its kind for the theory of
@@
 et al.\ \cite{barrett-et-al-2007} provide a good account of
-related work. Since then, datatypes have been added to the SMT solver
-Z3 \cite{de-moura-bjoerner-2008} \begin{rep}in unpublished work by Leonardo de Moura \end{rep}and to a SPASS-like %prototypical
+related work as of 2007. Since then, datatypes have been added not only to CVC4
+but also to the SMT solver
+Z3 \cite{de-moura-bjoerner-2008} \begin{rep}in unpublished work by Leonardo de Moura \end{rep}and \begin{rep}to \end{rep}a SPASS-like %prototypical
 superposition prover \begin{rep}by Daniel Wand\end{rep} \cite{wand-2014}.

### 2015-02-22 tex/body.tex
minor

@@
 By Theorems~\ref{thm:t},~\ref{thm:rs}, and~\ref{thm:ss}, the
-calculus is sound and complete for the ground $\thD$ theory, and we can
+calculus is sound and complete for the ground $\thD$ theory. We can
 rightly call it a decision procedure.
@@
 The ground decision procedure for (co)datatypes
-was presented at a high level of abstraction. This section describes
-the main aspects of the implementation within the SMT solver CVC4:\ the
-integration into CDCL($T$), the extension to quantified formulas, and
+was presented at a high level of abstraction, omitting quite a few details. This
+section describes the main aspects of the implementation within the SMT solver
+CVC4:\ the integration into CDCL($T$), the extension to quantified formulas, and
 some of the optimizations.
@@
 
-The efficiency properties of the procedure are mostly as described in Barrett et
-al.\ \cite{barrett-et-al-2007}. In particular, it is theoretically and
-empirically advantageous to delay applications of \rn{Split} as long as
+The efficiency properties of the procedure are mostly as described by Barrett et
+al.\ \cite{barrett-et-al-2007}. In particular, it is both theoretically and
+empirically beneficial to delay applications of \rn{Split} as long as
 possible. Similarly, \rn{Acyclic} and \rn{Unique} are fairly expensive because
@@
 that define codatatypes falling within the supported fragment. We added
-two unpublished theories by Peter Gammie and Andreas Lochbihler (\gandl, 317
-goals), about Bird and Stern--Brocot trees. To exercise the datatype
+two unpublished theories about Bird and Stern--Brocot trees
+by Peter Gammie and Andreas Lochbihler (\gandl,
+317~goals). To exercise the datatype
 support, the benchmarks were complemented by theories
@@
 success rate increase of over \OK{5\%}, which is quite significant. For Isabelle
-users, the practical usefulness of interpreted (co)datatypes appears roughly
-comparable to that of arithmetic \cite{blanchette-et-al-2013-smt}. The raw data
-also reveals that the stronger decision procedures almost fully subsume the weaker
-ones.
+users, the practical usefulness of interpreted (co)datatypes is roughly
+comparable to that of arithmetic \cite{blanchette-et-al-2013-smt}, although this
+obviously depends a lot on the application. The raw evaluation data also
+reveals that the stronger decision procedures almost completely subsume the
+weaker ones in practice.
 

### 2015-02-22 tex/body.tex
added tableau-like
more

@@
 datatypes and codatatypes. The dual of the acyclicity rule for datatypes is a
-uniqueness rule that identifies observationally equal values, also in the presence of cyclic
-(regular) data. The procedure decides ground and universal
+uniqueness rule that identifies observationally equal codatatype values, also in the presence of
+cyclic data. The procedure decides ground and universal
 problems and is composable via the Nelson--Oppen method. It has been
@@
 \begin{conf}
-Detailed proofs of soundness and completeness are included in a technical report
+Detailed proofs of completeness are included in a technical report
 \cite{our-report}.
@@
 
-\newcommand\elll{\kern.05ex \ell\kern.05ex}
-\newcommand\elllx{\kern.05ex \ell\kern.05ex}
+\newcommand\elll{\kern.0515ex \ell\kern.0515ex}
+\newcommand\elllx{\kern.0515ex \ell\kern.0515ex}
 
@@
 %
-For types with several constructors, it is customary to provide discriminators
+For types with several constructors, it is useful to provide discriminators
 $\const{d}_{i\negvthinspace j} : \delta_i \to \ty{bool}$. Instead of extending $\Funcs$,
 we let $\const{d}_{i\negvthinspace j}(t)$
-abbreviate
+be an abbreviation for
 $t \teq \const{C}_{i\negvthinspace j}\bigl(\const{s}_{i\negvthinspace j}^1(t), \ldots, \const{s}_{i\negvthinspace j}^{\,n_{\smash{{i\negvthinspace j}}}}(t)\bigr)$.
@@
 \afterDot
-Let $\delta$ be a corecursive codatatype. The domain interpreting $\delta$ is
-either infinite or a singleton. In the latter case, $\delta$ necessarily has a
-single constructor, whose arguments have types that are interpreted as
-singleton domains.
+Let $\delta$ be a corecursive codatatype. For any interpretation in $\JJJJ$,
+the domain interpreting $\delta$ is either infinite or a singleton. In the
+latter case, $\delta$ necessarily has a single constructor, whose arguments have
+types that are interpreted as singleton domains.
 \end{lemma}
@@
 and disequalities between ground $\Sig$-terms. The
-procedure is formulated as a calculus.
+procedure is formulated as a tableau-like calculus.
 % whose derivation rules are applied until saturation or contradiction.
@@
 %The former comes with no loss of generality since new equalities of the form $y \teq t$ for fresh variable $y$ can be added to $\Fc$ without affecting its satisfiability,
-%The latter also comes with no loss of generality since nested applications of selectors in $\Fc_0$ can be replaced by fresh variables while adding equalities to $\Fc$.
+%The latter also comes with no loss of generality since nested applications of selectors in $\Fc_0$ can be replaced by fresh variables while adding equalities to $\Fc\!$.
 We consider a normal interpretation $\J$
@@
 
-We first show that $\J$ satisfies all equalities $t_1 \teq t_2 \in \Fc$. % where $t_1$ and $t_2$ have type $\tau$.
+We first show that $\J$ satisfies all equalities $t_1 \teq t_2 \in \Fc\!$. % where $t_1$ and $t_2$ have type $\tau$.
 To achieve this, we show
@@
 %where $\expand{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$ for $i \in [n]$.
-Since \rn{Inject} and \rn{Clash} do not apply to $\Fc$,
+Since \rn{Inject} and \rn{Clash} do not apply to $\Fc\!$,
 by the construction of $\ValC$ we have that $\ValC \ec{t}$ is a term
@@
 $\const s^k_{\negvthinspace j}( u )$,
-since \rn{Split} does not apply to $\Fc$,
+since \rn{Split} does not apply to $\Fc\!$,
 $\ec{u}$ must contain a term of the form $\smash{\const C_{\negvthinspace\jPrime}\bigl( \const s^1_{\negvthinspace\jPrime}( u ), \ldots, \const s^{n}_{\negvthinspace\jPrime}( u ) \bigr)}$ for some $\jPrime$.
@@
 %where $\expand{\ValC \ec{u} }{ w_i } = \ValC \ec{ \const s^i_{\negvthinspace\jPrime}( u )}$ for $i \in [n]$.
-Since \rn{Inject} and \rn{Clash} do not apply, % to $\Fc$,
+Since \rn{Inject} and \rn{Clash} do not apply, % to $\Fc\!$,
 by construction $\ValC$ must be of the form
@@
 We now show that all disequalities in $\Fc$ are satisfied by $\J$.
-Assume $t \tneq u \in \Fc$.
+Assume $t \tneq u \in \Fc\!$.
 Since \rn{Conflict} does not apply, $t \teq u \not\in \Fc$ and thus $\ec{t}$ and $\ec{u}$ are distinct.
@@
 
-Since by assumption $\Fc$ contains only equalities and disequalities, we have $\J \models \Fc$,
-and since $\Ec \subseteq \Fc$,
+Since by assumption $\Fc$ contains only equalities and disequalities, we have $\J \models \Fc\!$,
+and since $\Ec \subseteq \Fc\!$,
 % is a superset of $\Ec$,
-we have that $\Ec$ is satisfied by $\J$.
+we conclude that $\J \models \Ec$.
 \qed

### 2015-02-22 tex/body.tex
some edits

@@
 %
-The rules belonging to a phase have priority over those of the subsequent
-phases. Rules are applied until the derivation tree is closed or all leaf nodes
+The rules belonging to a phase have priority over those of %the
+subsequent
+phases. The rules are applied until the derivation tree is closed or all leaf nodes
 are saturated.
@@
 \ourparagraph{Phase 1: Computing the Bidirectional Closure \beforefig(\figurename~\ref{fig:cc-rules}).}
-In conjunction with \rn{Refl}, \rn{Sym}, and \rn{Trans}, the \rn{Cong} rule computes the (upward) congruence closure,
-whereas the \rn{Inject} and \rn{Clash} rules together compute (downward) unification.
+In conjunction with \rn{Refl}, \rn{Sym}, and \rn{Trans}, the \rn{Cong} rule computes the congruence (upward) closure,
+whereas the \rn{Inject} and \rn{Clash} rules %together
+compute the unification (downward) closure.
 For unification, additional equalities are inferred based on the injectivity of constructors by \rn{Inject},

### 2015-02-22 tex/body.tex
more

@@
 \usepackage[scaled=.82]{beramono}
-\usepackage[scaled=.88]{helvet}
+\usepackage[scaled=.86]{helvet}
 \usepackage{booktabs}
@@
 Cyclic constraints such as
-$\const{x} \teq \const{S}(\const{x})$ %, where $\const{C}$ is a constructor,
+$x \teq \const{S}(x)$ %, where $\const{C}$ is a constructor,
 are unsatisfiable for datatypes but satisfiable for codatatypes.
 However, the uniqueness principle states that two bisimilar values---i.e., two values
-having the same possibly infinite expansion---must be equal; from $\const{x}
-\teq \const{S}(\const{x})$ and
-$\const{y} \teq \const{S}(\const{y})$, it deduces $\const{x} \teq \const{y}$.
+having the same possibly infinite expansion---must be equal; from $x
+\teq \const{S}(x)$ and
+$y \teq \const{S}(y)$, it deduces $x \teq y$.
 The acyclicity and uniqueness rules are needed to ensure completeness
@@
 compute the unification (downward) closure.
-For unification, additional equalities are inferred based on the injectivity of constructors by \rn{Inject},
+For unification, %additional
+equalities are inferred based on the injectivity of constructors by \rn{Inject},
 and failures to unify equated terms are recognized by \rn{Clash}.
@@
 The premises of the rules that make up the second phase refer to a mapping $\Val$
-that associates a $\mu$-term with each equivalence class as its representative.
+that associates with each equivalence class a $\mu$-term as its representative.
 %Recall that $\mu$-bindings can be used for representing possibly cyclic terms and values.
@@
 $\mu$-terms $\vec t$ of the expected types.
-The variable $x$ need not occur freely in the $\mu$-binder's body, in which case
+The variable~$x$ need not occur freely in the $\mu$-binder's body, in which case
 the binder can be omitted.

### 2015-02-22 tex/body.tex
more

@@
 Third, \emph{it must be possible to express cyclic }(\emph{regular}) \emph{values as closed terms and
-to enumerate them.} This is necessary both for finite model finding
+to enumerate them.} This is necessary both for finite model finding \cite{reynolds-et-al-2013}
 %modulo theories
@@
 terms %from $\tEc$
-have a corecursive singleton type
+have the same corecursive singleton type
 %(Section~\ref{sec:the-theory-of-co-datatypes})
-and must therefore be equal. Both
+and are therefore equal. Both
 \rn{Split}'s finiteness assumption %on the datatype
@@
 \keyw{by}~\textit{coinduction}~\textit{auto}
-in Isabelle \cite{blanchette-et-al-2014-impl}.
+in Isabelle\begin{rep} \cite{blanchette-et-al-2014-impl}\end{rep}.
 The fourth proof, by Gammie and Lochbihler, is more elaborate:
@@
 proof reconstruction for (co)data\-type inferences in Isabelle.
-CVC4 could also be interfaced for generating counterexamples in proof assistants, as
-an alternative to Nitpick; in this context, the acyclicity and uniqueness
-rules are crucial to exclude spurious countermodels. Finally, it might be
-worthwhile to extend SMT solvers with dedicated support for (co)recursive
-functions.
+CVC4's finite model finding capabilities \cite{reynolds-et-al-2013} could also
+be interfaced for generating counterexamples in proof assistants, as an
+alternative to Nitpick; in this context, the acyclicity and uniqueness rules are
+crucial to exclude spurious countermodels. Finally, it might be worthwhile to
+extend SMT solvers with dedicated support for (co)recursive functions.
 

### 2015-02-23 tex/body.tex
minor

@@
 
-This work is part of a program that aims at enriching automatic provers
+This work is part of a wider program that aims at enriching automatic provers
 with high-level features and at reducing the gap between automatic and
@@
 CVC4's finite model finding capabilities \cite{reynolds-et-al-2013} could also
-be interfaced for generating counterexamples in proof assistants, as an
-alternative to Nitpick; in this context, the acyclicity and uniqueness rules are
+be interfaced for generating counterexamples in proof assistants,
+as an alternative to Nitpick;
+in this context, the acyclicity and uniqueness rules are
 crucial to exclude spurious countermodels. Finally, it might be worthwhile to

### 2015-02-23 tex/body.tex
clearer

@@
 Compared with the conventional inductive definition
-\begin{rep}
-\[\keyw{datatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat}) \,\mid\, \const{Infty}\]
-\end{rep}
-with a dedicated constructor for representing infinity, the codatatype avoids
+\begin{conf}$\vthinspace \keyw{datatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat}) \,\mid\, \const{Infty}$, \end{conf}%
+\begin{rep}\[\vthinspace \keyw{datatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat}) \,\mid\, \const{Infty}\]\end{rep}%
+%with a dedicated constructor for representing infinity, the
+the codatatype avoids
 one case by unifying the finite and infinite nonzero cases.

### 2015-02-23 tex/body.tex
more

@@
 inspired by the $\mu$-calculus
-%\begin{conf}\cite{kozen-1983}.\end{conf}%
-%\begin{rep}\cite[etc.]{kozen-1983,endrullis-et-al-2011}.\end{rep}%
-\cite[etc.]{kozen-1983,endrullis-et-al-2011}.
+\begin{conf}\cite{kozen-1983}.\end{conf}
+\begin{rep}\cite[etc.]{kozen-1983,endrullis-et-al-2011}.\end{rep}
 
@@
    \smash{\vdots\,\,\,} \\[-1.5\jot]
-  \keyw{and}\; \,\delta_{\elllx} & =\, \smash{\const{C}_{\elll 1\!}(\bigl[\const{s}_{\elll 1\!}^1{:}\bigr]\vthinspace \tau_{\elll 1\!}^1, \ldots, \bigl[\const{s}_{\elll 1\!}^{n_{\elll 1\!}}{:}\bigr]\vthinspace \tau_{\elll 1\!}^{n_{\elll 1\!}})} \mid \cdots \mid \smash{\const{C}_{\elll m_{\elllx}\!}(\ldots)}
+  \keyw{and}\; \,\delta_{\elllx} & =\, \smash{\const{C}_{\elll 1\!}(%\bigl[\const{s}_{\elll 1\!}^1{:}\bigr]\vthinspace \tau_{\elll 1\!}^1, \ldots, \bigl[\const{s}_{\elll 1\!}^{n_{\elll 1\!}}{:}\bigr]\vthinspace \tau_{\elll 1\!}^{n_{\elll 1\!}}
+  \ldots)} \mid \cdots \mid \smash{\const{C}_{\elll m_{\elllx}\!}(\ldots)}
 \end{aligned}
@@
 determines the
-$\thD$-satisfiability of finite sets $\Ec$ of constraints:\ equalities
+$\thD$-satisfiability of finite sets $\Ec$ of literals:\ equalities
 and disequalities between ground $\Sig$-terms. The
@@
 \cite{barrett-et-al-2007}.
-As for uninterpreted types, since the constraints in $\Ec$
+As for uninterpreted types, since the literals in $\Ec$
 are ground, they cannot entail an upper
@@
 %We assume that all $\Sig$-terms in $\Ec$ are \emph{normalized}, meaning that all subterms of the form $\const{s}^{\,k}_{\negvthinspace j}( \const{C}_{\negvthinspace j}(t_1,\ldots,t_n) )$
-%are simplified to $t_k$. All constraints added to $\Ec$ are also implicitly
+%are simplified to $t_k$. All literals added to $\Ec$ are also implicitly
 %normalized in this manner.
@@
 %are simplified to $t_i$, and moreover assume that all %additional
-%constraints added to $\Ec$ are normalized in this manner.
+%literals added to $\Ec$ are normalized in this manner.
 %We present the calculus in three steps.
@@
 acyclic infinite values are uninteresting to us, and by monotonicity of ground
-constraints it is trivial to extend any normal interpretation with extra
+literals it is trivial to extend any normal interpretation with extra
 domain elements to obtain a genuine model if desired. %Hence, we focus on normal interpretations.
@@
 %The last precondition is admittedly rather technical.
-To ensure disequality constraints are satisfied by an interpretation based on $\ValC$,
+To ensure disequality literals are satisfied by an interpretation based on $\ValC$,
 it suffices that $\ValC$ is injective modulo $\alpha$-equivalence.
@@
 because $\BAD_{\smash{\J}}^{\Varec{x}}(\ValC)$ is a finite set.
-%Injectivity is important to satisfy disequality constraints.
+%Injectivity is important to satisfy disequality literals.
 %For infinite codatatypes, it is always possible to find suitable values $v$.
@@
 specialized solver for determining the satisfiability of conjunctions of
-constraints for its theory.
+literals for its theory.
 Given a theory $T = T_1 \mathrel\cup \cdots \mathrel\cup T_n$ and a set of
@@
 takes as input the union $M_i$
-of (1)~the purified form of $T_i$-constraints occurring in $M$, where fresh
+of (1)~the purified form of $T_i$-literals occurring in $M$, where fresh
 variables replace terms containing symbols not belonging to~$T_i$;
@@
 %\ourparagraph{Extension to Quantified Formulas.}
-While the decision procedure is restricted to ground constraints,
+While the decision procedure is restricted to ground literals,
 in practice we often want to solve problems that feature universal axioms and existential conjectures,
-where checking satisfiability requires reasoning about such ground constraints in the presence of universally quantified formulas.
+where checking satisfiability requires reasoning about such ground literals in the presence of universally quantified formulas.
 Many SMT solvers, including CVC4, can reason about quantified formulas using incomplete instantiation-based methods~\cite{MouraBjoerner07,ReynoldsTinelliMoura14}.
@@
 
-To reduce the number of unique constraints considered by the calculus, we
-compute a normal form for constraints as a preprocessing step. In particular, we
+To reduce the number of unique literals considered by the calculus, we
+compute a normal form for literals as a preprocessing step. In particular, we
 replace $u \teq t$ by $t \teq u$ if $t$ is less than $u$ based on a term

### 2015-02-23 tex/body.tex
reorder paras

@@
 
-{\looseness=-1
 A derivation rule can be applied to $\Ec$ if %all
@@
 %
-An application of a rule is \emph{redundant} if one branch in the
-conclusion is an assignment to $\Ec$ that leaves it unchanged.
+An application of a rule is \emph{redundant} if one of its non-$\bot$
+conclusions leaves $\Ec$ unchanged.
 A \emph{derivation tree} is a %finite
@@
 instance of a rule can be applied to it.
-
-}
 
@@
 that associates with each equivalence class a $\mu$-term as its representative.
-%Recall that $\mu$-bindings can be used for representing possibly cyclic terms and values.
-%(Section~\ref{sec:introduction}).
-% for example, $\MU x.\; \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
-The $\mu$-term $\Val[t\typ{\tau}]$ describes a class of $\tau$ values
-that $t$ and other members of $t$'s equivalence class can take in models of $\Ec$.
-%either in the case when $\tau$ is a codatatype or datatype type,
-When $\tau$ is a datatype, % \in \Data$,
-a cyclic $\mu$-term describes an infeasible class of values.
-%$\tau \not\in \Codata$.
-
+%
 Formally, $\mu$-\emph{terms} are defined recursively as being either a variable $x$
@@
 the binder can be omitted.
-In the \rn{Acyclic} rule,
-$\FV( u )$ denotes the variables occurring freely in~$u$.
-%%% But there are no constants of type $\tau \in \Nondata$ -- at most variables!
-%
-%For uniformity, we also consider $%\MU x.\;
-%\const{c}$ to be a $\mu$-term if $\const{c}$ is a constant of type $\tau \in \Nondata$.
-%
-%%% I think this should be clear enough from the "Conventions" paragraph at the
-%%% end of the introduction.
-%%
-% which we define recursively:
-%if $\const{C} : \vec \tau \rightarrow \upsilon$
-%and $\vec v$ are well-typed $\mu$-terms of type $\vec \tau$ under the assumption that $x$ has type $\upsilon$,
-%then $\MU x.\; \const{C}( \vec v )$ is a well-typed term of type $\upsilon$.
-%
+
+$\FV( t )$ denotes the set of free variables occurring in the $\mu$-term~$t$.
 A $\mu$-term is \emph{closed} if %and only if --- by convention, "and only if" is not necessary for *definitions* (but add it back if you disagree)
 it contains no free variables. It is \emph{cyclic} if %and only if --- ditto
-it contains a bound variable. The $\alpha$-\emph{equivalence} relation $t \aequiv u$
+it contains a bound variable. The $\alpha$\emph{-equiv\-alence} relation $t \aequiv u$
 indicates that the $\mu$-terms $t$ and $u$
@@
 $\MU x.\; \const{C}( x ) \not\aequiv \MU y.\; \const{C}( \const{C}( y ) )$.
+%In the \rn{Acyclic} rule,
+
+%Recall that $\mu$-bindings can be used for representing possibly cyclic terms and values.
+%(Section~\ref{sec:introduction}).
+% for example, $\MU x.\; \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
+%either in the case when $\tau$ is a codatatype or datatype type,
+%$\tau \not\in \Codata$.
+
+%%% But there are no constants of type $\tau \in \Nondata$ -- at most variables!
+%
+%For uniformity, we also consider $%\MU x.\;
+%\const{c}$ to be a $\mu$-term if $\const{c}$ is a constant of type $\tau \in \Nondata$.
+%
+%%% I think this should be clear enough from the "Conventions" paragraph at the
+%%% end of the introduction.
+%%
+% which we define recursively:
+%if $\const{C} : \vec \tau \rightarrow \upsilon$
+%and $\vec v$ are well-typed $\mu$-terms of type $\vec \tau$ under the assumption that $x$ has type $\upsilon$,
+%then $\MU x.\; \const{C}( \vec v )$ is a well-typed term of type $\upsilon$.
+%
 %, or informally are \emph{observationally equivalent}.
@@
 
-The mapping $\Val$, on which the rules depend, is computed as follows.
+The $\mu$-term $\Val[t\typ{\tau}]$ describes a class of $\tau$ values
+that $t$ and other members of $t$'s equivalence class can take in models of $\Ec$.
+When $\tau$ is a datatype, % \in \Data$,
+a cyclic $\mu$-term describes an infeasible class of values.
+
+The mapping $\Val$ is computed as follows.
 %

### 2015-02-23 tex/body.tex
more

@@
 
-At the end of this phase, $\Ec$ induces an equivalence
-relation over the set of $\Sig$-terms occurring in $\Ec$,
-denoted by $\tEc$, such that two terms $t$ and $u$ are equivalent if and
+Let $\tEc$ denote the set of $\Sig$-terms occurring in $\Ec$.
+At the end of the first phase, $\Ec$ induces an equivalence
+relation over $\tEc$ such that two terms $t$ and $u$ are equivalent if and
 only if $t \teq u \in \Ec$.
@@
 equivalence classes of terms. For a term $t \in \tEc$, we write $\ec{t}$ to
-denote the equivalence class in $\Ec$ that contains $t$.
+denote the equivalence class of $t$ in $\Ec$.
 
@@
 $\mu$-terms $\vec t$ of the expected types.
-The variable~$x$ need not occur freely in the $\mu$-binder's body, in which case
+The variable~$x$ need not occur free in the $\mu$-binder's body, in which case
 the binder can be omitted.
@@
 Two $\mu$-terms can denote the same value despite being $\alpha$-disequivalent---e.g.,
-$\MU x.\; \const{C}( x ) \not\aequiv \MU y.\; \const{C}( \const{C}( y ) )$.
+$\MU x.\; \const{S}( x ) \not\aequiv \MU y.\; \const{S}( \const{S}( y ) )$.
 %In the \rn{Acyclic} rule,
@@
 and all of its proper subterms are also normal.
-For example, $t = \MU x.\; \const{C}( \MU y.\; \const{C}( y ) )$ is abnormal
+For example, $t = \MU x.\; \const{C}( \MU y.\; \const{C}( y ) )$ is not normal
 because $\MU y.\; \const{C}( y )$ is a self-similar subterm of $t$.

### 2015-02-23 tex/body.tex
more

@@
 %
-Pascal Fontaine, Andreas Lochbihler, Mark Summerfield, and Dmitriy Traytel
-suggested several textual improvements. The second author's research was
-partially supported by the Deutsche Forschungs\-gemein\-schaft (DFG) project
-\relax{Hardening the Hammer} (grant Ni\,491\slash 14-1).
+Pascal Fontaine, Andreas Lochbihler, Christophe Ringeissen, Mark Summerfield,
+and Dmitriy Traytel suggested several textual improvements. The second author's
+research was partially supported by the Deutsche Forschungs\-gemein\-schaft
+(DFG) project \relax{Hardening the Hammer} (grant Ni\,491\slash 14-1).
 

### 2015-02-23 tex/body.tex
more refs

@@
 only datatypes.
-To our knowledge, our procedure is the first of its kind for the theory of
-codatatypes.
+%To our knowledge, our procedure is the first of its kind for the theory of
+%codatatypes.
 \begin{conf}
-Detailed proofs of completeness are included in a technical report
+Detailed proofs %of completeness
+are included in a %technical
+report
 \cite{our-report}.
@@
 et al.\ \cite{barrett-et-al-2007} provide a good account of
-related work as of 2007. Since then, datatypes have been added not only to CVC4
+related work on datatypes as of 2007. Since then, datatypes have been added not only to CVC4
 but also to the SMT solver
@@
 \cite{weber-2008}. Its successor, Nitpick \cite{blanchette-nipkow-2010}, can
-also generate models involving cyclic codatatype values. The $\mu$-notation is
-inspired by the $\mu$-calculus
+also generate models involving cyclic codatatype values.
+Cyclic values have been studied extensively under the headings of
+regular or rational trees---see Carayol and Morvan
+\cite{carayol-morvan-2006} and Djellou et al.\ \cite{djelloul-et-al-2008}
+for recent work.
+The $\mu$-notation is inspired by the $\mu$-calculus
 \begin{conf}\cite{kozen-1983}.\end{conf}
 \begin{rep}\cite[etc.]{kozen-1983,endrullis-et-al-2011}.\end{rep}
+
+%, notably in the context of constraint logic programming
 
@@
 
+{\footnotesize
 \def\ackname{Acknowledgment}
-\ourparagraph{\ackname.}
+\ourparagraph{\footnotesize\ackname.}
 We owe a great debt to the development team of CVC4,
@@
 Pascal Fontaine, Andreas Lochbihler, Christophe Ringeissen, Mark Summerfield,
-and Dmitriy Traytel suggested several textual improvements. The second author's
+and Dmitriy Traytel suggested many textual improvements. The second author's
 research was partially supported by the Deutsche Forschungs\-gemein\-schaft
@@
 
+}
+
 \bibliographystyle{splncs03}

### 2015-02-23 tex/body.tex
more

@@
 of infinite sequences or streams over an element type $\tau$ is
-allowed\begin{conf}, with no base case\end{conf}\begin{rep}; the
+allowed\begin{conf}, even though it has no base case\end{conf}\begin{rep}; the
 corresponding datatype would be rejected as non-well-founded
@@
 \emph{selectors}. There is no need to consider further function symbols
-because they can be abstracted away as variables when combining theories.
+because they can be abstracted away as variables when combining theories
+(Section~\ref{sec:implementation-as-a-theory-solver-in-cvc4}).
 %Exceptionally, it is convenient to use numeric constants ($0$, $1$, \ldots)\ in
@@
 \end{rep}%
-This axiom schema guarantees that two values that yield the same
-observations must be equal, where the observations are made by using the
+This schema encodes a form of extensionality:\ Two values that yield the same
+observations must be equal, where the observations are made through
 selectors and discriminators.
@@
 axiom schema, which states that constructor terms cannot be equal to
-any of their proper subterms \cite{barrett-et-al-2007}. Dually, coinduction
+any of their proper subterms. %\cite{barrett-et-al-2007}.
+Dually, coinduction
 is equivalent to the uniqueness axiom schema, which asserts that codatatype

### 2015-02-23 tex/body.tex
more

@@
 %variable $x$:
-\[\begin{array}{r@{}c@{}l}
+\[\begin{array}{@{}r@{}c@{}l@{}}
 \expand{t}{x}^B & {} \,=\, {} &
 \textstyle\relax{\begin{cases}
-    \MU x.\; \const{C}\smash{\bigl( \expand{t}{ \vec u}^{\smash{B \,\cup\, \{ x \}}}\bigr)} & \text{if this $x \notin B$ is
-    bound by $\MU x.\; \const{C}( \vec u )$ in $t$}  \\[-\jot]
-    x & \text{otherwise}
+    x & \text{if $x \in B$}   \\[-\jot]
+    \MU x.\; \const{C}\smash{\bigl( \expand{t}{ \vec u}^{\smash{B \,\uplus\, \{ x \}}}\bigr)} & \text{if this occurrence of $x$ is
+    bound by $\MU x.\; \const{C}( \vec u )$ in $t$}
 \end{cases}} ~ \\
@@
 \textstyle\begin{cases}
-    \MU x.\; \const{C}\smash{\bigl( \expand{t}{ \vec u }^{\smash{B \,\cup\, \{ x \}}}\bigr)} & \text{if }x \not\in B \\[-\jot]
-    x & \text{otherwise}
+    x & \text{if $x \in B$} \\[-\jot]
+    \MU x.\; \const{C}\smash{\bigl( \expand{t}{ \vec u }^{\smash{B \,\uplus\, \{ x \}}}\bigr)} & \text{otherwise}
 \end{cases}

### 2015-02-23 tex/body.tex
more

@@
 First, we define the set of interpretations of the theory $\thD$,
-which requires additional terminology concerning $\mu$-terms.
+which requires custom terminology concerning $\mu$-terms.
 Given a $\mu$-term $t$ with subterm $u$,
@@
     x & \text{if $x \in B$}   \\[-\jot]
-    \MU x.\; \const{C}\smash{\bigl( \expand{t}{ \vec u}^{\smash{B \,\uplus\, \{ x \}}}\bigr)} & \text{if this occurrence of $x$ is
-    bound by $\MU x.\; \const{C}( \vec u )$ in $t$}
+    \MU x.\; \const{C}\smash{\bigl( \expand{t}{ \vec u}^{\smash{B \,\uplus\, \{ x \}}}\bigr)} & \text{if
+    $\MU x.\; \const{C}( \vec u )$ binds this occurrence of $x \notin B$ in $t$}
 \end{cases}} ~ \\

### 2015-02-23 tex/body.tex
more edits

@@
 compute a normal form for literals as a preprocessing step. In particular, we
-replace $u \teq t$ by $t \teq u$ if $t$ is less than $u$ based on a term
+replace $u \teq t$ by $t \teq u$ if $t$ is smaller than $u$ with respect to a term
 ordering, replace $\smash{\const{C}_{\negvthinspace j}( \vec t \,) \teq
@@
 
+\kern-1pt
+
 %\section{Examples}
@@
 and the \emph{Archive of Formal Proofs} (AFP, 2974~goals) \cite{klein-et-al-afp}
-that define codatatypes falling within the supported fragment. We added
+that define codatatypes falling within the supported fragment. We also
+
+%%% TYPESETTING: Terrible hack
+\pagebreak
+\noindent
+added
 two unpublished theories about Bird and Stern--Brocot trees
@@
 support, the benchmarks were complemented by theories
-about lists and trees. The theories were selected before running the
+about lists and trees. The theories were selected before conducting the
 experiments. The experimental data are publicly available.%
@@
 along with the goal \cite{blanchette-et-al-2013-smt}. The resulting problem was
-given to CVC4, running for up to 60 seconds on the StarExec cluster
-\cite{stump-et-al-2014-starexec}.
+given to the development version of CVC4 and to Z3 4.3.2, each running for up to 60
+seconds on the StarExec cluster \cite{stump-et-al-2014-starexec}.
 %TODO: Keep or not? Problems not involving (co)datatypes were filtered out.
 Due to the lack of machinery for reconstructing inferences about (co)datatypes
-in Isabelle, the solver is trusted as an oracle in these experiments.
+in Isabelle, the solvers are trusted as oracles in these experiments. Z3
+is included for comparison.
 
@@
 basic lemmas about constructors and selectors, allowing some (co)datatype
-reasoning. Z3 4.3.2 is included for comparison.
+reasoning.
 % when this makes sense.

### 2015-02-23 tex/body.tex
more

@@
 compute a normal form for literals as a preprocessing step. In particular, we
-replace $u \teq t$ by $t \teq u$ if $t$ is smaller than $u$ with respect to a term
+replace $u \teq t$ by $t \teq u$ if $t$ is smaller than $u$ with respect to some term
 ordering, replace $\smash{\const{C}_{\negvthinspace j}( \vec t \,) \teq

### 2015-02-23 tex/body.tex
more
tried to make intro slightly less technical

@@
 $x \teq \const{S}(x)$ %, where $\const{C}$ is a constructor,
-are unsatisfiable for datatypes but satisfiable for codatatypes.
-However, the uniqueness principle states that two bisimilar values---i.e., two values
-having the same possibly infinite expansion---must be equal; from $x
+are unsatisfiable for datatypes,
+thanks to an acyclicity rule,
+but satisfiable for codatatypes. 
+For the latter, a uniqueness rule ensures that two values
+having the same infinite expansion must be equal; from $x
 \teq \const{S}(x)$ and
 $y \teq \const{S}(y)$, it deduces $x \teq y$.
-The acyclicity and uniqueness rules are needed to ensure completeness
+These two rules are needed to ensure completeness
 (model soundness) on ground problems. %and the absence of spurious models.
@@
 and the \emph{Archive of Formal Proofs} (AFP, 2974~goals) \cite{klein-et-al-afp}
-that define codatatypes falling within the supported fragment. We also
-
+that define codatatypes falling within the supported fragment. \hfill We~also%
+\begin{conf}%
+\par
 %%% TYPESETTING: Terrible hack
@@
 \noindent
-added
-two unpublished theories about Bird and Stern--Brocot trees
+\end{conf}%
+added two unpublished theories about Bird and Stern--Brocot trees
 by Peter Gammie and Andreas Lochbihler (\gandl,

### 2015-02-23 tex/body.tex
clarify FO

@@
 
-We fix a signature $\Sig = (\Types, \Funcs)$. The types are partitioned into
+{\looseness=-1
+We fix a first-order signature $\Sig = (\Types, \Funcs)$. The types are partitioned into
 $\Types = \Data \mathrel{\uplus} \Codata \mathrel{\uplus} \Nondata$, where $\Data$ are the
@@
 \emph{selectors}. There is no need to consider further function symbols
-because they can be abstracted away as variables when combining theories
-(Section~\ref{sec:implementation-as-a-theory-solver-in-cvc4}).
+because they can be abstracted away as variables when combining theories.
+%(Section~\ref{sec:implementation-as-a-theory-solver-in-cvc4}).
 %Exceptionally, it is convenient to use numeric constants ($0$, $1$, \ldots)\ in
 %examples.
+
+}
 

### 2015-02-23 tex/body.tex
more

@@
 \newcommand\Codata{\Types_{\mathrm{codt}}}
-\newcommand\Nondata{\Types_{\mathrm{ord}}}
+\newcommand\Nondata{\Types_{\mathrm{un}}}
 
@@
 
-{\looseness=-1
-We fix a first-order signature $\Sig = (\Types, \Funcs)$. The types are partitioned into
+We fix a signature $\Sig = (\Types, \Funcs)$. The types are partitioned into
 $\Types = \Data \mathrel{\uplus} \Codata \mathrel{\uplus} \Nondata$, where $\Data$ are the
 \emph{datatypes}, $\Codata$ are the \emph{codatatypes}, and $\Nondata$ are the %remaining
-\emph{ordinary types}. The function symbols are partitioned into $\Funcs = \Ctr
+\emph{uninterpreted types}. The function symbols are partitioned into $\Funcs = \Ctr
 \mathrel{\uplus} \Sel$, where $\Ctr$ are the \emph{constructors} and $\Sel$ are the
@@
 
-}
-
 %\ourparagraph{Specifications.}
 In an SMT problem, the signature is typically given by specifying first the
-uninterpreted ordinary types in any order, then the (co)datatypes with their constructors
+uninterpreted types in any order, then the (co)datatypes with their constructors
 and selectors in groups of mutually (co)recursive groups of (co)datatypes, and
@@
 arguments and returns a $\delta$ value. The argument types must be either
-ordinary, among the already known (co)datatypes, or among the (co)datatypes
+uninterpreted, among the already known (co)datatypes, or among the (co)datatypes
 being introduced.
@@
 A type may also be a corecursive singleton only in some models. If the example
-above is altered to make \ty{unit} an ordinary type, \ty{b} and \ty{c} will be
+above is altered to make \ty{unit} an uninterpreted type, \ty{b} and \ty{c} will be
 singletons precisely when \ty{unit} is interpreted as a singleton.
-Fortunately, \begin{rep}given cardinalities for the ordinary types, \end{rep}%
+Fortunately, \begin{rep}given cardinalities for the uninterpreted types, \end{rep}%
 it is easy to characterize this degenerate case.
@@
 %  * leaves us with the case of a single constructor
-% * since we are looking at a specific model, we can assume all ordinary
+% * since we are looking at a specific model, we can assume all uninterpreted
 %   types are finite datatypes with nullary constructors corresponding to their
@@
 $\Sig$.
-First, all codatatypes in $\Codata$ are corecursive. This is reasonable
+First, all codatatypes %in $\Codata$
+are corecursive. This is reasonable
 because noncorecursive codatatypes can be seen as nonrecursive
 datatypes.
-Second, all types $\tau \in \Nondata$ have infinite cardinality.
-This is not very restrictive, because in principle each interpreted type in
-$\Nondata$ having finite cardinality~$n$
-can be replaced by a %n enumeration
-datatype with $n$~nullary constructors
-\cite{barrett-et-al-2007}.
-As for uninterpreted types, since the literals in $\Ec$
-are ground, they cannot entail an upper
-bound on the cardinality of any such type, so it is safe to consider these types infinite.
+Second, all uninterpreted types have infinite cardinality.
+Since the literals in $\Ec$ are ground, they cannot entail an upper
+bound on the cardinality of any such type, so it is safe to consider these types infinite.%
+%\footnote{%
+%The theory could be extended This is not very restrictive, because in principle each interpreted type in
+%$\Nondata$ having finite cardinality~$n$
+%can be replaced by a %n enumeration
+%datatype with $n$~nullary constructors
+%\cite{barrett-et-al-2007}.
+%}
 
@@
 %either in the case when $\tau$ is a codatatype or datatype type,
-%$\tau \not\in \Codata$.
+%$\tau \notin \Codata$.
 
@@
 constructors' argument types.
-%This step relies on the assumption that ordinary types are infinite.
+%This step relies on the assumption that uninterpreted types are infinite.
 
@@
 $\interp{\J}{\tau}$ includes a maximal set of closed normal $\mu$-terms of that type that are
-unique up to $\alpha$-equivalence and acyclic if $\tau \not\in \Codata$.
+unique up to $\alpha$-equivalence and acyclic if $\tau \notin \Codata$.
 \item
@@
   \betweenantes
-  v \not\in \BAD_{\smash{\J}}^{\Varec{x}}(\ValC)
+  v \notin \BAD_{\smash{\J}}^{\Varec{x}}(\ValC)
 }{
@@
 Consider the case when $\tau$ is finite.
-By assumption, $\tau \not\in \Nondata$.
-Since \rn{Split} does not apply to $\Ec$, we have $\tau \not\in \Data$.
+By assumption, $\tau \notin \Nondata$.
+Since \rn{Split} does not apply to $\Ec$, we have $\tau \notin \Data$.
 If $\tau \in \Codata$, then $\tau$ is corecursive by assumption, and by Lemma~\ref{lem:corecursive-singletons},
@@
 $\{ {\Varec{y}} \mapsto t \}$, where $t$ is acyclic when the type of ${\Varec{y}}$ is not a codatatype.
-Thus, $\ValC \ec{x}$ is acyclic when $\tau \not\in \Codata$.
+Thus, $\ValC \ec{x}$ is acyclic when $\tau \notin \Codata$.
 Therefore, by definition, $\ValC \ec{x}$ is $\alpha$-equivalent to a value in $\interp{\J}{\tau}$.
@@
 Assume $t \tneq u \in \Fc\!$.
-Since \rn{Conflict} does not apply, $t \teq u \not\in \Fc$ and thus $\ec{t}$ and $\ec{u}$ are distinct.
+Since \rn{Conflict} does not apply, $t \teq u \notin \Fc$ and thus $\ec{t}$ and $\ec{u}$ are distinct.
 Since $\interp{\J}{t} \vsim \ValC \ec{t}$ and $\interp{\J}{u} \vsim \ValC \ec{u}$,
@@
 while partitioning $\SSS$ into $\SSS_1, \ldots, \SSS_n$ based on the top-most symbol of constructor terms in these equivalence classes
-and the equivalence of their arguments of ordinary types.
+and the equivalence of their arguments of uninterpreted types.
 It then partitions each set %in this partition %in the same manner

### 2015-02-23 tex/body.tex
more

@@
 uniqueness rule that identifies observationally equal codatatype values, also in the presence of
-cyclic data. The procedure decides ground and universal
+cyclic data. The procedure decides universal
 problems and is composable via the Nelson--Oppen method. It has been
@@
 infinit\-e constructor terms are also legitimate values for codatatypes
-(Section~\ref{sec:the-theory-of-co-datatypes}). Intuitively, the
+(Section~\ref{sec:co-datatypes}). Intuitively, the
 values of a codatatype consist of all well-typed finite and infinite ground
@@
 
+{\looseness=-1
 Datatypes and codatatypes are an integral part of many proof assistants,
@@
 solvers, has been extended with a syntax for datatypes.
-In this \thewordpaper, we introduce a unified decision procedure for ground
-problems involving datatypes and codatatypes in
+In this \thewordpaper, we introduce a
+unified \hfill decision \hfill procedure \hfill for \hfill universal \hfill problems \hfill involving \hfill datatypes \hfill and \hfill codatatypes
 \begin{conf}\goodbreak\noindent\end{conf}%  %%% TYPESETTING: terrible HACK
-combination (Section~\ref{sec:the-ground-decision-procedure}).
+in combination (Section~\ref{sec:the-decision-procedure}).
 The procedure is described abstractly as a calculus and can be
@@
 
+}
+
 Datatypes and codatatypes share many properties, so it makes sense
@@
 These two rules are needed to ensure completeness
-(model soundness) on ground problems. %and the absence of spurious models.
+(model soundness) on universal problems. %and the absence of spurious models.
 They cannot be expressed as finite axiomatizations, so they naturally belong in
@@
 (Co)datatypes}
-\label{sec:the-theory-of-co-datatypes}
+\label{sec:co-datatypes}
 
@@
 
-For the ground theory, induction is equivalent to the acyclicity
+For the universal theory, induction is equivalent to the acyclicity
 axiom schema, which states that constructor terms cannot be equal to
@@
 %    * codatatypes:
-%      * when looking at the ground theory, enough to
+%      * when looking at the universal theory, enough to
 %        consider uniqueness -- no way to express that certain infinite
@@
 
-\section{The Ground Decision Procedure} % for (Co)datatypes}
-\label{sec:the-ground-decision-procedure}
+\section{The Decision Procedure} % for (Co)datatypes}
+\label{sec:the-decision-procedure}
 
 Given a fixed signature~$\Sig$,
-the decision procedure for the ground theory of (co)datatypes %$\thD$
+the decision procedure for the universal theory of (co)data\-types %$\thD$
 determines the
 $\thD$-satisfiability of finite sets $\Ec$ of literals:\ equalities
-and disequalities between ground $\Sig$-terms. The
+and disequalities between $\Sig$-terms. The
 procedure is formulated as a tableau-like calculus.
@@
 Second, all uninterpreted types have infinite cardinality.
-Since the literals in $\Ec$ are ground, they cannot entail an upper
-bound on the cardinality of any such type, so it is safe to consider these types infinite.%
+Without quantifiers, the constraints $\Ec$ cannot entail an upper
+bound on the cardinality of any such type, so it is safe to consider these types
+infinite.%
 %\footnote{%
@@
 have the same corecursive singleton type
-%(Section~\ref{sec:the-theory-of-co-datatypes})
+%(Section~\ref{sec:co-datatypes})
 and are therefore equal. Both
@@
 models must contain all possible infinite terms, not only cyclic ones. However,
-acyclic infinite values are uninteresting to us, and by monotonicity of ground
-literals it is trivial to extend any normal interpretation with extra
+acyclic infinite values are uninteresting to us, and for quantifier-free
+formulas it is trivial to extend any normal interpretation with extra
 domain elements to obtain a genuine model if desired. %Hence, we focus on normal interpretations.
@@
 By Theorems~\ref{thm:t},~\ref{thm:rs}, and~\ref{thm:ss}, the
-calculus is sound and complete for the ground $\thD$ theory. We can
+calculus is sound and complete for the universal $\thD$ theory. We can
 rightly call it a decision procedure.
@@
 
-The ground decision procedure for (co)datatypes
-was presented at a high level of abstraction, omitting quite a few details. This
-section describes the main aspects of the implementation within the SMT solver
-CVC4:\ the integration into CDCL($T$), the extension to quantified formulas, and
-some of the optimizations.
+The decision procedure was presented at a high level of abstraction, omitting
+quite a few details. This section describes the main aspects of the
+implementation within the SMT solver CVC4:\ the integration into CDCL($T$), the
+extension to quantified formulas, and some of the optimizations.
 
@@
 %\ourparagraph{Extension to Quantified Formulas.}
-While the decision procedure is restricted to ground literals,
-in practice we often want to solve problems that feature universal axioms and existential conjectures,
-where checking satisfiability requires reasoning about such ground literals in the presence of universally quantified formulas.
+While the decision procedure is restricted to universal conjectures, in practice
+we often want to solve problems that feature universal axioms and existential
+conjectures.
+%%% @ANDY: not sure how to phrase this now
+%where checking satisfiability requires reasoning about ground literals in the
+%presence of universally quantified formulas.
 Many SMT solvers, including CVC4, can reason about quantified formulas using incomplete instantiation-based methods~\cite{MouraBjoerner07,ReynoldsTinelliMoura14}.
@@
 When assuming that $\ty{u}$ has infinite cardinality, the calculus
-presented in Section~\ref{sec:the-ground-decision-procedure} is \relax{incomplete}
+presented in Section~\ref{sec:the-decision-procedure} is \relax{incomplete}
 since it may fail to recognize cases where \rn{Split} and \rn{Single} should be
@@
 
-We introduced a decision procedure for the ground theory of datatypes and
+We introduced a decision procedure for the universal theory of datatypes and
 codatatypes. Our main

### 2015-02-23 tex/body.tex
more

@@
 Compared with the conventional inductive definition
-\begin{conf}$\vthinspace \keyw{datatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat}) \,\mid\, \const{Infty}$, \end{conf}%
+\begin{conf}$\vthinspace \keyw{datatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat}) \,\mid\, \const{Infty}$\vvthinspace, \end{conf}%
 \begin{rep}\[\vthinspace \keyw{datatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat}) \,\mid\, \const{Infty}\]\end{rep}%
@@
 In this \thewordpaper, we introduce a
-unified \hfill decision \hfill procedure \hfill for \hfill universal \hfill problems \hfill involving \hfill datatypes \hfill and \hfill codatatypes
+unified \hfill decision \hfill procedure \hfill for \hfill universal \hfill problems \hfill involving \hfill datatypes \hfill and \hfill codatatypes%
 \begin{conf}\goodbreak\noindent\end{conf}%  %%% TYPESETTING: terrible HACK
@@
 
-For the universal theory, induction is equivalent to the acyclicity
+For deciding universal formulas, induction can be replaced by the acyclicity
 axiom schema, which states that constructor terms cannot be equal to
-any of their proper subterms. %\cite{barrett-et-al-2007}.
-Dually, coinduction
-is equivalent to the uniqueness axiom schema, which asserts that codatatype
-values are fully characterized by their possibly infinite expansion.
+any of their proper subterms \cite{barrett-et-al-2007}.
+Dually, coinduction can be replaced by the uniqueness axiom schema, which
+asserts that codatatype values are fully characterized by their possibly
+infinite expansion \cite[Theorem 8.1, 2$\,{\Leftrightarrow}\,$5]{rutten-2000}.
 

### 2015-02-23 tex/body.tex
more

@@
 % it comes to English prefixes like non-, in-, un-, and a-.)
-
+%
 For deciding universal formulas, induction can be replaced by the acyclicity
@@
 any of their proper subterms \cite{barrett-et-al-2007}.
-Dually, coinduction can be replaced by the uniqueness axiom schema, which
-asserts that codatatype values are fully characterized by their possibly
-infinite expansion \cite[Theorem 8.1, 2$\,{\Leftrightarrow}\,$5]{rutten-2000}.
+Dually, coinduction can be replaced by the uniqueness schema, which
+asserts that codatatype values are fully characterized by their %possibly
+%infinite
+expansion \cite[Theorem 8.1, 2$\,{\Leftrightarrow}\,$5]{rutten-2000}.
 
@@
 {\looseness=-1
-Some codatatypes are so degenerate as be finite even though they
+Some codatatypes are so degenerate as to be finite even though they
 have infinite values. A simple example is
@@
 \vthinspace$\keyw{codatatype}~\ty{b} = \const{B}(\ty{b},\: \ty{c},\: \ty{b},\: \ty{unit})
-~\keyw{and}~ \ty{c} = \const{C}(\ty{a},\: \ty{unit},\: \ty{b},\: \ty{c})$,
+~\keyw{and}~ \ty{c} = \const{C}(\ty{a},\: \ty{b},\: \ty{c})$,
 \end{conf}%
@@
       \keyw{codatatype}\;\, \ty{b} & \,=\, \const{B}(\ty{b},\: \ty{c},\: \ty{b},\: \ty{unit}) \\[-1\jot]
-      \keyw{and}\;\, \ty{c} & \,=\, \const{C}(\ty{a},\: \ty{unit},\: \ty{b},\: \ty{c})
+      \keyw{and}\;\, \ty{c} & \,=\, \const{C}(\ty{a},\: \ty{b},\: \ty{c})
 \end{aligned}
@@
 $\thD$-satisfiability of finite sets $\Ec$ of literals:\ equalities
-and disequalities between $\Sig$-terms. The
-procedure is formulated as a tableau-like calculus.
+and disequalities between $\Sig$-terms,
+whose variables are interpreted existentially.
+The procedure is formulated as a tableau-like calculus.
+%Following the tradition of refutational theorem proving,
+Proving a universal quantifier-free conjecture is reduced to showing that
+its negation is unsatisfiable.
 % whose derivation rules are applied until saturation or contradiction.

### 2015-02-23 tex/body.tex
15 pages again

@@
 \newcommand\Codata{\Types_{\mathrm{codt}}}
-\newcommand\Nondata{\Types_{\mathrm{un}}}
+\newcommand\Nondata{\Types_{\mathrm{ord}}}
 
@@
 \emph{datatypes}, $\Codata$ are the \emph{codatatypes}, and $\Nondata$ are the %remaining
-\emph{uninterpreted types}. The function symbols are partitioned into $\Funcs = \Ctr
-\mathrel{\uplus} \Sel$, where $\Ctr$ are the \emph{constructors} and $\Sel$ are the
-\emph{selectors}. There is no need to consider further function symbols
-because they can be abstracted away as variables when combining theories.
+\emph{ordinary types} (which can be interpreted or not). The function symbols are
+partitioned into $\Funcs = \Ctr \mathrel{\uplus} \Sel$, where $\Ctr$ are the
+\emph{constructors} and $\Sel$ are the \emph{selectors}. There is no need to
+consider further function symbols because they can be abstracted away as
+variables when combining theories.
 %(Section~\ref{sec:implementation-as-a-theory-solver-in-cvc4}).
@@
 Each (co)datatype specification consists of $\ell$~mutually recursive types that are
-either all datatypes or all codatatypes. Polymorphic types, nested
-(co)recursion, and datatype--codatatype mixtures fall outside this fragment%
+either all datatypes or all codatatypes. \begin{rep}Polymorphic types, nested
+(co)recursion,\end{rep}\begin{conf}Nested (co)recursion\end{conf} and datatype--codatatype mixtures fall outside this fragment%
 \dotReportFootnote{%
@@
 \cite{blanchette-et-al-2014-codata}).}
-We allow ourselves some notational parameterization
+\begin{rep}We allow ourselves some notational parameterization
 % at the metalevel
-through subscripts%
-\begin{conf}
-(e.g., $\ty{stream}_{\,\tau}$)%
-\end{conf}%
-\begin{rep}---for example, $\ty{stream}_{\,\tau}$ denotes a
+through subscripts---for example, $\ty{stream}_{\,\tau}$ denotes a
 family of ground types including
 $\ty{stream}_{\,\ty{int}}$, $\ty{stream}_{\,\ty{bool}}$,
-and \smash{$\ty{stream}_{\,\ty{stream}_{\,\ty{real}}}$}\end{rep}.
+and \smash{$\ty{stream}_{\,\ty{stream}_{\,\ty{real}}}$}.\end{rep}
 
@@
 arguments and returns a $\delta$ value. The argument types must be either
-uninterpreted, among the already known (co)datatypes, or among the (co)datatypes
+ordinary, among the already known (co)datatypes, or among the (co)datatypes
 being introduced.
@@
     \smash{\const{C}_{i\negvthinspace j}(\bar x) \tneq \const{C}_{i\negvthinspace \jPrime}(\bar y) \quad\text{if $j \not= \jPrime$}}
-  \\[-0.5\jot]
+  \\[-1\jot]
 \text{Injectivity:}\quad
@@
     \smash{\const{C}_{i\negvthinspace j}(x_1,\ldots,x_{n_{i\negvthinspace j}}) \teq \const{C}_{i\negvthinspace j}(y_1,\ldots,y_{n_{i\negvthinspace j}}) \longrightarrow x_k \teq y_k}
-  \\[-0.5\jot]
+  \\[-1\jot]
 \text{Exhaustiveness:}\quad
   & \smash{\const{d}_{i1}(x) \mathrel\lor \cdots \mathrel\lor \const{d}_{im_i}(x)}
-  \\[-0.5\jot]
+  \\[-1\jot]
 \text{Selection:}\quad
@@
 datatypes.
-Second, all uninterpreted types have infinite cardinality.
+Second, all ordinary types have infinite cardinality.
 Without quantifiers, the constraints $\Ec$ cannot entail an upper
-bound on the cardinality of any such type, so it is safe to consider these types
-infinite.%
-%\footnote{%
-%The theory could be extended This is not very restrictive, because in principle each interpreted type in
-%$\Nondata$ having finite cardinality~$n$
-%can be replaced by a %n enumeration
-%datatype with $n$~nullary constructors
-%\cite{barrett-et-al-2007}.
-%}
+bound on the cardinality of any uninterpreted type, so it is safe to consider these types
+infinite. As for ordinary types interpreted finitely by other theories (e.g.,
+bit vectors), each interpreted type having finite cardinality~$n$
+can be replaced by a %n enumeration
+datatype with $n$~nullary constructors
+\cite{barrett-et-al-2007}.
 
@@
 The conclusion either specifies equalities to be added to $\Ec$
-or is $\bot$. % (contradiction).
+or is $\bot$ (contradiction).
 % (in which case we call an application of it \emph{nonterminal})
@@
 tree whose nodes are finite sets of
-equalities and where child nodes are obtained by a nonredundant application of a
+equalities, such that child nodes are obtained by a nonredundant application of~a%
+\begin{conf}
+\par
+\pagebreak %% TYPESETTING: crude hack
+\noindent
+\end{conf}\begin{rep} \end{rep}%
 derivation rule to the parent. A derivation tree is \emph{closed} if all of
@@
 %where a rule can be applied to $\Ec$ if it meets all of the specified preconditions for $\Ec$.
+
+\ourparagraph{Phase 1: Computing the Bidirectional Closure \beforefig(\figurename~\ref{fig:cc-rules}).}
+In conjunction with \rn{Refl}, \rn{Sym}, and \rn{Trans}, the \rn{Cong} rule computes the congruence (upward) closure,
+whereas the \rn{Inject} and \rn{Clash} rules %together
+compute the unification (downward) closure.
+For unification, %additional
+equalities are inferred based on the injectivity of constructors by \rn{Inject},
+and failures to unify equated terms are recognized by \rn{Clash}.
+The \rn{Conflict} rule recognizes when an equality and its negation both occur in $\Ec$, in which case $\Ec$ has no model.
+
+Let $\tEc$ denote the set of $\Sig$-terms occurring in $\Ec$.
+At the end of the first phase, $\Ec$ induces an equivalence
+relation over $\tEc$ such that two terms $t$ and $u$ are equivalent if and
+only if $t \teq u \in \Ec$.
+Thus, we can regard $\Ec$ as a set of
+equivalence classes of terms. For a term $t \in \tEc$, we write $\ec{t}$ to
+denote the equivalence class of $t$ in $\Ec$.
+
+\ourparagraph{Phase 2: Applying Acyclicity and Uniqueness \beforefig(\figurename~\ref{fig:ab-rules}).}
+%For presentation of the rules in this phase, we rely on a representation of terms in the $\mu$-notation.
+The premises of the rules that make up the second phase refer to a mapping $\Val$
+that associates with each equivalence class a $\mu$-term as its representative.
+%
+Formally, $\mu$-\emph{terms} are defined recursively as being either a variable $x$
+or an applied constructor
+$\MU x.\: \const{C}( \vec t\, )$ for some $\const{C} \in \Ctr$ and
+$\mu$-terms $\vec t$ of the expected types.
+The variable~$x$ need not occur free in the $\mu$-binder's body, in which case
+the binder can be omitted.
+
+$\FV( t )$ denotes the set of free variables occurring in the $\mu$-term~$t$.
+A $\mu$-term is \emph{closed} if %and only if --- by convention, "and only if" is not necessary for *definitions* (but add it back if you disagree)
+it contains no free variables. It is \emph{cyclic} if %and only if --- ditto
+it contains a bound variable. The $\alpha$\emph{-equiv\-alence} relation $t \aequiv u$
+indicates that the $\mu$-terms $t$ and $u$
+are syntactically equivalent for some capture-avoiding renaming of $\mu$-bound variables\begin{rep}---e.g.,
+%$\MU x.\; \const{C}( x ) \aequiv \MU y.\; \const{C}( y )$ and
+$\MU x.\; \const{D}( y, x ) \aequiv \MU z.\; \const{D}( y, z )$,
+but
+$\MU x.\; \const{C}( x ) \not\aequiv \MU x.\; \const{D}( y, x ) \not\aequiv \MU x.\; \const{D}( z, x )
+\not\aequiv \MU y.\; \const{D}( y, x )$\end{rep}.
+Two $\mu$-terms can denote the same value despite being $\alpha$-disequivalent---e.g.,
+$\MU x.\; \const{S}( x ) \not\aequiv \MU y.\; \const{S}( \const{S}( y ) )$.
+%In the \rn{Acyclic} rule,
+
+%Recall that $\mu$-bindings can be used for representing possibly cyclic terms and values.
+%(Section~\ref{sec:introduction}).
+% for example, $\MU x.\; \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
+%either in the case when $\tau$ is a codatatype or datatype type,
+%$\tau \notin \Codata$.
+
+%%% But there are no constants of type $\tau \in \Nondata$ -- at most variables!
+%
+%For uniformity, we also consider $%\MU x.\;
+%\const{c}$ to be a $\mu$-term if $\const{c}$ is a constant of type $\tau \in \Nondata$.
+%
+%%% I think this should be clear enough from the "Conventions" paragraph at the
+%%% end of the introduction.
+%%
+% which we define recursively:
+%if $\const{C} : \vec \tau \rightarrow \upsilon$
+%and $\vec v$ are well-typed $\mu$-terms of type $\vec \tau$ under the assumption that $x$ has type $\upsilon$,
+%then $\MU x.\; \const{C}( \vec v )$ is a well-typed term of type $\upsilon$.
+%
+%, or informally are \emph{observationally equivalent}.
+%For convenience, we use $\mu$-terms to refer to (classes of) values for both codatatype and datatype terms,
+%where in the latter case, a $\mu$-term with a bound variable denotes an infeasible term.
+%where the latter case adds the restriction on $\mu$-terms $t$ that no subterms of $t$ contain are bound variables.
+
+The $\mu$-term $\Val[t\typ{\tau}]$ describes a class of $\tau$ values
+that $t$ and other members of $t$'s equivalence class can take in models of $\Ec$.
+When $\tau$ is a datatype, % \in \Data$,
+a cyclic $\mu$-term describes an infeasible class of values.
+
+The mapping $\Val$ is computed as follows.
+%
+With each equivalence class $\ec{u\typ{\tau}}$, we associate a fresh variable $\Varec{u}\typ{\tau}$
+%of the same type as $u$
+%not occurring in $\tEc$,
+and set $\Val\ec{u} := {\Varec{u}}$, that is to say there are initially no
+constraints on the values for any equivalence class $\ec{u}$. The mapping $\Val$
+is refined by applying the following unfolding rule exhaustively:\strut
+\[
+\hbox{\(
+\inferrule{
+  {\Varec{u}} \in \FV( \Val )
+  \betweenantes
+  \const{C}( t_1, \ldots, t_n ) \in \ec{u}
+  \betweenantes
+  \const{C} \in \Ctr
+}{
+  \Val := \Val [{\Varec{u}} \mapsto \MU {\Varec{u}}.\; \const{C}( {\Varec{t_1}}, \ldots, {\Varec{t_n}} )]
+}
+\)
+%\rn{Unfold}
+}
+\]
+$\FV( \Val )$ denotes the set of free variables occurring in $\Val$'s range,
+and $\Val[x \mapsto t]$ denotes the \emph{variable-capturing} substitution of $t$ for
+$x$ in $\Val$'s range. It is easy to see that the height of terms produced as a
+result of the unfolding is bounded by the number of equivalence classes of
+$\Ec$, and thus the construction of $\Val$ will terminate.
 
@@
 \end{tabular}
-%\vspace*{-3pt} %% TYPESETTING HACK
+\vspace*{-2pt} %% TYPESETTING HACK
 \caption{\,Derivation rules for bidirectional closure%.
@@
 
-\ourparagraph{Phase 1: Computing the Bidirectional Closure \beforefig(\figurename~\ref{fig:cc-rules}).}
-In conjunction with \rn{Refl}, \rn{Sym}, and \rn{Trans}, the \rn{Cong} rule computes the congruence (upward) closure,
-whereas the \rn{Inject} and \rn{Clash} rules %together
-compute the unification (downward) closure.
-For unification, %additional
-equalities are inferred based on the injectivity of constructors by \rn{Inject},
-and failures to unify equated terms are recognized by \rn{Clash}.
-The \rn{Conflict} rule recognizes when an equality and its negation both occur in $\Ec$, in which case $\Ec$ has no model.
-
-Let $\tEc$ denote the set of $\Sig$-terms occurring in $\Ec$.
-At the end of the first phase, $\Ec$ induces an equivalence
-relation over $\tEc$ such that two terms $t$ and $u$ are equivalent if and
-only if $t \teq u \in \Ec$.
-Thus, we can regard $\Ec$ as a set of
-equivalence classes of terms. For a term $t \in \tEc$, we write $\ec{t}$ to
-denote the equivalence class of $t$ in $\Ec$.
-
-\ourparagraph{Phase 2: Applying Acyclicity and Uniqueness \beforefig(\figurename~\ref{fig:ab-rules}).}
-%For presentation of the rules in this phase, we rely on a representation of terms in the $\mu$-notation.
-The premises of the rules that make up the second phase refer to a mapping $\Val$
-that associates with each equivalence class a $\mu$-term as its representative.
-%
-Formally, $\mu$-\emph{terms} are defined recursively as being either a variable $x$
-or an applied constructor
-$\MU x.\: \const{C}( \vec t\, )$ for some $\const{C} \in \Ctr$ and
-$\mu$-terms $\vec t$ of the expected types.
-The variable~$x$ need not occur free in the $\mu$-binder's body, in which case
-the binder can be omitted.
-
-$\FV( t )$ denotes the set of free variables occurring in the $\mu$-term~$t$.
-A $\mu$-term is \emph{closed} if %and only if --- by convention, "and only if" is not necessary for *definitions* (but add it back if you disagree)
-it contains no free variables. It is \emph{cyclic} if %and only if --- ditto
-it contains a bound variable. The $\alpha$\emph{-equiv\-alence} relation $t \aequiv u$
-indicates that the $\mu$-terms $t$ and $u$
-are syntactically equivalent for some capture-avoiding renaming of $\mu$-bound variables\begin{rep}---e.g.,
-%$\MU x.\; \const{C}( x ) \aequiv \MU y.\; \const{C}( y )$ and
-$\MU x.\; \const{D}( y, x ) \aequiv \MU z.\; \const{D}( y, z )$,
-but
-$\MU x.\; \const{C}( x ) \not\aequiv \MU x.\; \const{D}( y, x ) \not\aequiv \MU x.\; \const{D}( z, x )
-\not\aequiv \MU y.\; \const{D}( y, x )$\end{rep}.
-Two $\mu$-terms can denote the same value despite being $\alpha$-disequivalent---e.g.,
-$\MU x.\; \const{S}( x ) \not\aequiv \MU y.\; \const{S}( \const{S}( y ) )$.
-%In the \rn{Acyclic} rule,
-
-%Recall that $\mu$-bindings can be used for representing possibly cyclic terms and values.
-%(Section~\ref{sec:introduction}).
-% for example, $\MU x.\; \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
-%either in the case when $\tau$ is a codatatype or datatype type,
-%$\tau \notin \Codata$.
-
-%%% But there are no constants of type $\tau \in \Nondata$ -- at most variables!
-%
-%For uniformity, we also consider $%\MU x.\;
-%\const{c}$ to be a $\mu$-term if $\const{c}$ is a constant of type $\tau \in \Nondata$.
-%
-%%% I think this should be clear enough from the "Conventions" paragraph at the
-%%% end of the introduction.
-%%
-% which we define recursively:
-%if $\const{C} : \vec \tau \rightarrow \upsilon$
-%and $\vec v$ are well-typed $\mu$-terms of type $\vec \tau$ under the assumption that $x$ has type $\upsilon$,
-%then $\MU x.\; \const{C}( \vec v )$ is a well-typed term of type $\upsilon$.
-%
-%, or informally are \emph{observationally equivalent}.
-%For convenience, we use $\mu$-terms to refer to (classes of) values for both codatatype and datatype terms,
-%where in the latter case, a $\mu$-term with a bound variable denotes an infeasible term.
-%where the latter case adds the restriction on $\mu$-terms $t$ that no subterms of $t$ contain are bound variables.
-
 \begin{figure}[t!]
-%\vspace*{+3pt}
+\vspace*{+6pt} %% TYPESETTING HACK
 \normalsize
@@
 \end{tabular}
-%\vspace*{-3pt} %% TYPESETTING HACK
+\vspace*{-2pt} %% TYPESETTING HACK
 \caption{\,Derivation rules for acyclicity and uniqueness%.
@@
 
-The $\mu$-term $\Val[t\typ{\tau}]$ describes a class of $\tau$ values
-that $t$ and other members of $t$'s equivalence class can take in models of $\Ec$.
-When $\tau$ is a datatype, % \in \Data$,
-a cyclic $\mu$-term describes an infeasible class of values.
-
-The mapping $\Val$ is computed as follows.
-%
-With each equivalence class $\ec{u\typ{\tau}}$, we associate a fresh variable $\Varec{u}\typ{\tau}$
-%of the same type as $u$
-%not occurring in $\tEc$,
-and set $\Val\ec{u} := {\Varec{u}}$, that is to say there are initially no
-constraints on the values for any equivalence class $\ec{u}$. The mapping $\Val$
-is refined by applying the following unfolding rule exhaustively:\strut
-\[
-\hbox{\(
-\inferrule{
-  {\Varec{u}} \in \FV( \Val )
-  \betweenantes
-  \const{C}( t_1, \ldots, t_n ) \in \ec{u}
-  \betweenantes
-  \const{C} \in \Ctr
-}{
-  \Val := \Val [{\Varec{u}} \mapsto \MU {\Varec{u}}.\; \const{C}( {\Varec{t_1}}, \ldots, {\Varec{t_n}} )]
-}
-\)
-%\rn{Unfold}
-}
-\]
-$\FV( \Val )$ denotes the set of free variables occurring in $\Val$'s range,
-and $\Val[x \mapsto t]$ denotes the \emph{variable-capturing} substitution of $t$ for
-$x$ in $\Val$'s range. It is easy to see that the height of terms produced as a
-result of the unfolding is bounded by the number of equivalence classes of
-$\Ec$, and thus the construction of $\Val$ will terminate.
-
-\begin{examplex}
-Suppose that $\Ec$ contains four distinct equivalence classes $\ec{w}$, $\ec{x}$, $\ec{y}$, and~$\ec{z}$
-such that $\const{C}(w,y) \in \ec{x}$ and $\const{C}(z,x) \in \ec{y}$ for some $\const{C} \in \Ctr$.
-A possible sequence of unfolding steps is given below, omitting
-trivial entries such as $\ec{w} \mapsto {\Varec{w}}$.
-%
-\begin{enumerate}
-%\item \noindent\rlap{Initially:}\phantom{Unfold $\Varec{x}$:\enskip}$\{  \}$
-\item \noindent\rlap{Unfold $\Varec{x}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\Val = \{\vthinspace  \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) \vthinspace\}$
-\item \noindent\rlap{Unfold $\vthinspace\Varec{y}\negvthinspace$}\phantom{Unfold $\Varec{x}$}:\enskip$\Val = \{\vthinspace  \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\;
-  \ec{y} \mapsto \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) \vthinspace\}$
-\item \noindent\rlap{Unfold $\Varec{x}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\Val = \{\vthinspace  \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\;
-  \ec{y} \mapsto \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\; \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) ) \vthinspace\}$
-\end{enumerate}
-%
-The resulting $\Val$ indicates that the values for $x$ and $y$ in models of $\Ec$
-must be of the forms $\const{C}( {\Varec{w}}, \const{C}( {\Varec{z}},\allowbreak \const{C}( {\Varec{w}},\allowbreak \const{C}( {\Varec{z}},\ldots ))))$
-and $\const{C}( {\Varec{z}}, \const{C}( {\Varec{w}}, \const{C}( {\Varec{z}}, \const{C}( {\Varec{w}}, \ldots ))))$,
-respectively. %, for some values of ${\Varec{w}}$ and ${\Varec{z}}$.
-\xend
-\end{examplex}
-
-Given the mapping $\Val$, the \rn{Acyclic} and \rn{Unique} rules work as follows.
-For acyclicity, if $\ec{t}$ is a datatype equivalence class
-whose values $\Val \ec{t} = \MU x.\; u$ are cyclic
-(as expressed by $x \in \FV( u )$),
-then $\Ec$ is $\thD$-unsatisfiable.
-For uniqueness, if $\ec{t}$, $\ec{u}$ are two codatatype equivalence classes
-whose values $\Val \ec{t}$, $\Val \ec{u}$ are $\alpha$-equivalent,
-%for a capture-avoiding renaming of $\mu$-bound variables,
-then $t \teq u$. Comparison for $\alpha$-equivalence may seem too restrictive, since
-$\MU x.\; \const{S}(x)$ and $\MU y.\; \const{S}(\const{S}(y))$ specify the same
-value despite being $\alpha$-disequivalent, but the rule will
-make progress by discovering that the subterm $\const{S}(y)$ of $\MU y.\;
-\const{S}(\const{S}(y))$ must be equal to the entire term\begin{rep}
-as demonstrated next\end{rep}.
-
-%% Minor terminology point:
-%% A function symbol can have a "type signature" (but we're trying to avoid that terminology);
-%% unless it's a constant, it doesn't have a (first-order) type. So I'm rephrasing.
-
-\begin{examplex}
-Let $\Ec = \{ x \teq \const{S}(x),\; y \teq \const{S}(\const{S}(y)) \}$.
-After phase~1, the equivalence classes %of $\Ec$
-are
-$\{ x,\, \const{S}(x) \}$,
-$\{ y,\, \const{S}(\const{S}(y)) \}$,
-and $\{ \const{S}(y) \}$.
-Constructing $\Val$ yields
-\begin{align*}
-\Val \ec{x} & \,=\, \MU {\Varec{x}}.\; \const{S}(\Varec{x}) &
-\kern0.25em %% TYPESETTING
-\Val \ec{y} & \,=\,
-\MU {\Varec{y}}.\; \const{S}(\smash{\MU {\VAREC{\const{S}(y)}}}.\; \const{S}( {\Varec{y}} ) )
-\kern0.25em %% TYPESETTING
- &
-\Val \ec{\const{S}(y)} & =
-\MU {\VAREC{\const{S}(y)}}.\; \const{S}(\MU {\Varec{y}}.\; \const{S}( {\VAREC{\const{S}(y)}} ) )
-\end{align*}
-Since $\Val \ec{y} \aequiv \Val \ec{\const{S}(y)}$,
-%for the renaming
-%$\{ {\Varec{y}} \mapsto {\VAREC{\const{S}(y)}}, {\VAREC{\const{S}(y)}} \mapsto {\Varec{y}} \}$,
-the \rn{Unique} rule applies to derive $y \teq \const{S}(y)$.
-%Intuitively, this equality holds, since the values of $y$ and $\const{S}(y)$
-%are equivalent.
-%
-At this point, phase~1 is activated again, % to XXX the equivalence classes,
-yielding
-$\{ x,\, \const{S}(x) \}$ and
-$\{ y,\, \const{S}(y),\allowbreak\, \const{S}(\const{S}(y)) \}$.
-The mapping $\Val$ is updated accordingly:
-\begin{conf}%
-$\Val \ec{y} \,=\, \MU {\Varec{y}}.\; \const{S}({\Varec{y}})$.
-\end{conf}%
-\begin{rep}
-\begin{align*}
-\Val \ec{x} & \,=\, \MU {\Varec{x}}.\; \const{S}(\Varec{x})
-&
-\Val \ec{y} & \,=\, \MU {\Varec{y}}.\; \const{S}({\Varec{y}})
-\end{align*}
-\end{rep}
-Since $\Val \ec{x} \aequiv \Val \ec{y}$,
-%for the renaming
-%$\{ {\Varec{y}} \mapsto {\VAREC{\const{S}(y)}}, {\VAREC{\const{S}(y)}} \mapsto {\Varec{y}} \}$,
-\rn{Unique} can finally derive $x \teq y$.
-\xend
-\end{examplex}
-
 \begin{figure}[t!]
+\vspace*{+6pt} %% TYPESETTING HACK
 \normalsize
@@
 \end{tabular}
+\vspace*{-2pt} %% TYPESETTING HACK
 \caption{\,Derivation rules for branching%.
@@
 \end{figure}
+
+\begin{examplex}
+Suppose that $\Ec$ contains four distinct equivalence classes $\ec{w}$, $\ec{x}$, $\ec{y}$, and~$\ec{z}$
+such that $\const{C}(w,y) \in \ec{x}$ and $\const{C}(z,x) \in \ec{y}$ for some $\const{C} \in \Ctr$.
+A possible sequence of unfolding steps is given below, omitting
+trivial entries such as $\ec{w} \mapsto {\Varec{w}}$.
+%
+\begin{enumerate}
+%\item \noindent\rlap{Initially:}\phantom{Unfold $\Varec{x}$:\enskip}$\{  \}$
+\item \noindent\rlap{Unfold $\Varec{x}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\Val = \{\vthinspace  \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) \vthinspace\}$
+\item \noindent\rlap{Unfold $\vthinspace\Varec{y}\negvthinspace$}\phantom{Unfold $\Varec{x}$}:\enskip$\Val = \{\vthinspace  \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\;
+  \ec{y} \mapsto \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) \vthinspace\}$
+\item \noindent\rlap{Unfold $\Varec{x}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\Val = \{\vthinspace  \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\;
+  \ec{y} \mapsto \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\; \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) ) \vthinspace\}$
+\end{enumerate}
+%
+The resulting $\Val$ indicates that the values for $x$ and $y$ in models of $\Ec$
+must be of the forms $\const{C}( {\Varec{w}}, \const{C}( {\Varec{z}},\allowbreak \const{C}( {\Varec{w}},\allowbreak \const{C}( {\Varec{z}},\ldots ))))$
+and $\const{C}( {\Varec{z}}, \const{C}( {\Varec{w}}, \const{C}( {\Varec{z}}, \const{C}( {\Varec{w}}, \ldots ))))$,
+respectively. %, for some values of ${\Varec{w}}$ and ${\Varec{z}}$.
+\xend
+\end{examplex}
+
+Given the mapping $\Val$, the \rn{Acyclic} and \rn{Unique} rules work as follows.
+For acyclicity, if $\ec{t}$ is a datatype equivalence class
+whose values $\Val \ec{t} = \MU x.\; u$ are cyclic
+(as expressed by $x \in \FV( u )$),
+then $\Ec$ is $\thD$-unsatisfiable.
+For uniqueness, if $\ec{t}$, $\ec{u}$ are two codatatype equivalence classes
+whose values $\Val \ec{t}$, $\Val \ec{u}$ are $\alpha$-equivalent,
+%for a capture-avoiding renaming of $\mu$-bound variables,
+then $t \teq u$. Comparison for $\alpha$-equivalence may seem too restrictive, since
+$\MU x.\; \const{S}(x)$ and $\MU y.\; \const{S}(\const{S}(y))$ specify the same
+value despite being $\alpha$-disequivalent, but the rule will
+make progress by discovering that the subterm $\const{S}(y)$ of $\MU y.\;
+\const{S}(\const{S}(y))$ must be equal to the entire term\begin{rep}
+as demonstrated next\end{rep}.
+
+%% Minor terminology point:
+%% A function symbol can have a "type signature" (but we're trying to avoid that terminology);
+%% unless it's a constant, it doesn't have a (first-order) type. So I'm rephrasing.
+
+\begin{examplex}
+Let $\Ec = \{ x \teq \const{S}(x),\; y \teq \const{S}(\const{S}(y)) \}$.
+After phase~1, the equivalence classes %of $\Ec$
+are
+$\{ x,\, \const{S}(x) \}$,
+$\{ y,\, \const{S}(\const{S}(y)) \}$,
+and $\{ \const{S}(y) \}$.
+Constructing $\Val$ yields
+\begin{align*}
+\Val \ec{x} & \,=\, \MU {\Varec{x}}.\; \const{S}(\Varec{x}) &
+\kern0.25em %% TYPESETTING
+\Val \ec{y} & \,=\,
+\MU {\Varec{y}}.\; \const{S}(\smash{\MU {\VAREC{\const{S}(y)}}}.\; \const{S}( {\Varec{y}} ) )
+\kern0.25em %% TYPESETTING
+ &
+\Val \ec{\const{S}(y)} & =
+\MU {\VAREC{\const{S}(y)}}.\; \const{S}(\MU {\Varec{y}}.\; \const{S}( {\VAREC{\const{S}(y)}} ) )
+\end{align*}
+Since $\Val \ec{y} \aequiv \Val \ec{\const{S}(y)}$,
+%for the renaming
+%$\{ {\Varec{y}} \mapsto {\VAREC{\const{S}(y)}}, {\VAREC{\const{S}(y)}} \mapsto {\Varec{y}} \}$,
+the \rn{Unique} rule applies to derive $y \teq \const{S}(y)$.
+%Intuitively, this equality holds, since the values of $y$ and $\const{S}(y)$
+%are equivalent.
+%
+At this point, phase~1 is activated again, % to XXX the equivalence classes,
+yielding
+$\{ x,\, \const{S}(x) \}$ and
+$\{ y,\, \const{S}(y),\allowbreak\, \const{S}(\const{S}(y)) \}$.
+The mapping $\Val$ is updated accordingly:
+\begin{conf}%
+$\Val \ec{y} \,=\, \MU {\Varec{y}}.\; \const{S}({\Varec{y}})$.
+\end{conf}%
+\begin{rep}
+\begin{align*}
+\Val \ec{x} & \,=\, \MU {\Varec{x}}.\; \const{S}(\Varec{x})
+&
+\Val \ec{y} & \,=\, \MU {\Varec{y}}.\; \const{S}({\Varec{y}})
+\end{align*}
+\end{rep}
+Since $\Val \ec{x} \aequiv \Val \ec{y}$,
+%for the renaming
+%$\{ {\Varec{y}} \mapsto {\VAREC{\const{S}(y)}}, {\VAREC{\const{S}(y)}} \mapsto {\Varec{y}} \}$,
+\rn{Unique} can finally derive $x \teq y$.
+\xend
+\end{examplex}
 

### 2015-02-23 tex/body.tex
clarification
uninterpreted vs. ordinary

@@
-%%% TODO: Archive!!
 %%% TODO: In report, give small example where Single is needed
@@
 consider further function symbols because they can be abstracted away as
-variables when combining theories.
+variables when combining theories. $\mathrm\Sigma$-terms are standard first-order terms,
+without $\mu$-binders.
 %(Section~\ref{sec:implementation-as-a-theory-solver-in-cvc4}).
@@
 singletons precisely when \ty{unit} is interpreted as a singleton.
-Fortunately, \begin{rep}given cardinalities for the uninterpreted types, \end{rep}%
+Fortunately, \begin{rep}given cardinalities for the ordinary types, \end{rep}%
 it is easy to characterize this degenerate case.
@@
 bit vectors), each interpreted type having finite cardinality~$n$
-can be replaced by a %n enumeration
+can be viewed as a %n enumeration
 datatype with $n$~nullary constructors
@@
 tree whose nodes are finite sets of
-equalities, such that child nodes are obtained by a nonredundant application of~a%
-\begin{conf}
-\par
-\pagebreak %% TYPESETTING: crude hack
-\noindent
-\end{conf}\begin{rep} \end{rep}%
+equalities, such that child nodes are obtained by a nonredundant application of a
 derivation rule to the parent. A derivation tree is \emph{closed} if all of
@@
 while partitioning $\SSS$ into $\SSS_1, \ldots, \SSS_n$ based on the top-most symbol of constructor terms in these equivalence classes
-and the equivalence of their arguments of uninterpreted types.
+and the equivalence of their arguments of ordinary types.
 It then partitions each set %in this partition %in the same manner
@@
 Nonetheless, two techniques help increase the precision of the solver.
-First, we can apply \rn{Split} to datatype terms whose cardinality depends on the finiteness of uninterpreted types.
+First, we can apply \rn{Split} to datatype terms whose cardinality depends on the finiteness of
+ordinary types.
 Second, we can conditionally apply \rn{Single} to codatatype terms that may have cardinality one.

### 2015-02-23 tex/body.tex
more

@@
 useful to represent finite data structures in computer science applications but
-also arise when formalizing mathematics.
+also arise in formalized mathematics.
 They can be implemented efficiently and enjoy
@@
 in combination (Section~\ref{sec:the-decision-procedure}).
-The procedure is described abstractly as a calculus and can be
+The procedure is described abstractly as a %tableau-like
+calculus and can be
 composed via the Nelson--Oppen method \cite{nelson-oppen-1979}.
@@
 \ourparagraph{Conventions.}
-The setting is a monomorphic (or many-sorted) first-order logic.
+Our setting is a monomorphic (or many-sorted) first-order logic.
 A signature $\Sig = (\Types, \Funcs)$ consists of a set of types $\Types$ and a
@@
 with $\const{p}(\ldots)$ abbreviating $\const{p}(\ldots) \teq \const{true}$\end{rep}.
-The notation $t\typ{\tau}$ indicates that term $t$ has type~$\tau$.
+The notation $t\typ{\tau}$ stands for a term $t$ of type~$\tau$.
 \begin{rep}The operator $\bigwedge_{\,i}\, \varphi_i$ abbreviates a conjunction

### 2015-02-23 tex/body.tex
Minor fixes

@@
 \MU y.\; \const{C}( y )$.
-The term $u = \MU x.\; \const{C}( \MU y.\; \const{C}( x ) )$ is also abnormal,
+The term $u = \MU x.\; \const{C}( \MU y.\; \const{C}( x ) )$ is also not normal,
 since $\MU y.\; \const{C}( x )$ is a self-similar subterm of $u$,
@@
 First, we can apply \rn{Split} to datatype terms whose cardinality depends on the finiteness of
-ordinary types.
+uninterpreted types.
 Second, we can conditionally apply \rn{Single} to codatatype terms that may have cardinality one.

### 2015-02-23 tex/body.tex
more
killed one 'ordinary'

@@
 
+{\looseness=-1
 We fix a signature $\Sig = (\Types, \Funcs)$. The types are partitioned into
@@
 consider further function symbols because they can be abstracted away as
-variables when combining theories. $\mathrm\Sigma$-terms are standard first-order terms,
+variables when combining theories. $\mathrm\Sigma$-terms are standard first-order terms
+over $\mathrm\Sigma$,
 without $\mu$-binders.
@@
 
+}
+
 %\ourparagraph{Specifications.}
 In an SMT problem, the signature is typically given by specifying first the
-uninterpreted ordinary types in any order, then the (co)datatypes with their constructors
+uninterpreted %ordinary
+types in any order, then the (co)datatypes with their constructors
 and selectors in groups of mutually (co)recursive groups of (co)datatypes, and

### 2015-02-23 tex/body.tex
more
minor

@@
 unified \hfill decision \hfill procedure \hfill for \hfill universal \hfill problems \hfill involving \hfill datatypes \hfill and \hfill codatatypes%
-\begin{conf}\goodbreak\noindent\end{conf}%  %%% TYPESETTING: terrible HACK
+\begin{conf}\goodbreak\noindent\end{conf}%  %%% TYPESETTING BREAK
+\begin{rep} \end{rep}%
 in combination (Section~\ref{sec:the-decision-procedure}).
@@
 the procedure considers the set $\SSS$ of all codatatype equivalence classes. It
-\begin{conf}\goodbreak\noindent\end{conf}%
+\begin{conf}\goodbreak\noindent\end{conf}%  %%% TYPESETTING BREAK
 simultaneously visits the equivalence classes of arguments of constructor terms in each equivalence class in $\SSS$,
@@
 and the \emph{Archive of Formal Proofs} (AFP, 2974~goals) \cite{klein-et-al-afp}
-that define codatatypes falling within the supported fragment. \hfill We~also%
-\begin{conf}%
-\par
-%%% TYPESETTING: Terrible hack
-\pagebreak
-\noindent
-\end{conf}%
+that define codatatypes falling within the supported fragment. We
 added two unpublished theories about Bird and Stern--Brocot trees

### 2015-02-23 tex/body.tex
added AP again

@@
 theories with us so we could include them in the benchmarks.
-Andrei Popescu helped clarify our thoughts regarding the axiomatization of
-codatatypes and indicated related work. Dmitriy Traytel took part in discussions
-about degenerate codatatypes.
-%
-Pascal Fontaine, Andreas Lochbihler, Christophe Ringeissen, Mark Summerfield,
-and Dmitriy Traytel suggested many textual improvements. The second author's
-research was partially supported by the Deutsche Forschungs\-gemein\-schaft
-(DFG) project \relax{Hardening the Hammer} (grant Ni\,491\slash 14-1).
+Andrei Popescu helped clarify our thoughts regarding codatatypes and indicated
+related work. Dmitriy Traytel took part in discussions about degenerate
+codatatypes.
+%
+Pascal Fontaine, Andreas Lochbihler, Andrei Popescu, Christophe Ringeissen, Mark
+Summerfield, and Dmitriy Traytel suggested many textual improvements. The second
+author's research was partially supported by the Deutsche
+Forschungs\-gemein\-schaft (DFG) project \relax{Hardening the Hammer} (grant
+Ni\,491\slash 14-1).
 

### 2015-02-23 tex/body.tex
changes suggested by Andrei

@@
 Freely generated inductive datatypes are ubiquitous in functional programs and
-logical specifications. They are especially
+formal specifications. They are especially
 useful to represent finite data structures in computer science applications but
@@
 constructor
-terms, and only those. For example, the coinductive specification
+terms, and only those. As a simple example, the coinductive specification
 %
@@
 The procedure is described abstractly as a %tableau-like
-calculus and can be
-composed via the Nelson--Oppen method \cite{nelson-oppen-1979}.
+calculus and is composable via the Nelson--Oppen method \cite{nelson-oppen-1979}.
 It generalizes the procedure by Barrett et al.\ \cite{barrett-et-al-2007}, which covers
@@
 
-The procedure is implemented in the SMT solver CVC4 \cite{barrett-et-al-2011} as a combination
+Our procedure is implemented in the SMT solver CVC4 \cite{barrett-et-al-2011} as a combination
 of rewriting and a theory solver (Section~\ref{sec:implementation-as-a-theory-solver-in-cvc4}).
@@
 also generate models involving cyclic codatatype values.
-Cyclic values have been studied extensively under the headings of
+Cyclic values have been studied extensively under the heading of
 regular or rational trees---see Carayol and Morvan
@@
 %
-The schema ensures that the interpretation of datatypes
-is standard.
+%The schema ensures that the interpretation of datatypes
+%is standard.
 For the natural numbers constructed from $\const{Z}$ and $\const{S}$,
-induction prohibits nonstandard models, which could contain cyclic values---e.g.,
+induction prohibits models that contain cyclic values---e.g.,
 an $n$ such that $n \teq \const{S}(n)$---or even infinite acyclic values

### 2015-02-23 tex/body.tex
rephased

@@
 %
-Datatypes are additionally characterized by an induction axiom schema\begin{rep}:
+Datatypes are additionally characterized by an induction principle\begin{rep}:
 %
@@
 the formula is type-correct or else $\top$\end{rep}.
-%
-%The schema ensures that the interpretation of datatypes
-%is standard.
+The principle ensures that the interpretation of datatypes is standard.
 For the natural numbers constructed from $\const{Z}$ and $\const{S}$,
@@
 \end{rep}%
-This schema encodes a form of extensionality:\ Two values that yield the same
+This axiom encodes a form of extensionality:\ Two values that yield the same
 observations must be equal, where the observations are made through

### 2015-02-24 tex/body.tex
more

@@
 An evaluation on %hand-crafted examples and on
-problems generated from Isabelle theories using the Sledgehammer tool \cite{paulson-blanchette-2010}
+problems generated from Isabelle theories using the Sledgehammer tool \cite{blanchette-et-al-2013-smt}
 demonstrates the usefulness of the approach (Section~\ref{sec:experimental-results}).
@@
 quite a few details. This section describes the main aspects of the
-implementation within the SMT solver CVC4:\ the integration into CDCL($T$), the
-extension to quantified formulas, and some of the optimizations.
+implementation within the SMT solver CVC4:\ the integration into \XXXL($T$)
+\cite{ganzinger-et-al-2004}, the extension to quantified formulas, and some of
+the optimizations.
 
@@
 CVC4's finite model finding capabilities \cite{reynolds-et-al-2013} could also
-be interfaced for generating counterexamples in proof assistants,
-as an alternative to Nitpick;
+be interfaced for generating counterexamples in proof assistants;
+%as an alternative to Nitpick;
 in this context, the acyclicity and uniqueness rules are
 crucial to exclude spurious countermodels. Finally, it might be worthwhile to
-extend SMT solvers with dedicated support for (co)recursive functions.
+extend SMT solvers with dedicated support for (co)recursion.
 
@@
 \ourparagraph{\footnotesize\ackname.}
+\looseness=-1
 We owe a great debt to the development team of CVC4,
@@
 and in particular Morgan Deters,
-who jointly with the first author developed the initial version
+who jointly with the first author
+developed the initial version
 of the theory solver for datatypes in CVC4.
@@
 %
-Peter Gammie and Andreas Lochbihler shared their private
-theories with us so we could include them in the benchmarks.
+Peter Gammie and Andreas Lochbihler provided useful benchmarks.
 Andrei Popescu helped clarify our thoughts regarding codatatypes and indicated
@@
 Pascal Fontaine, Andreas Lochbihler, Andrei Popescu, Christophe Ringeissen, Mark
-Summerfield, and Dmitriy Traytel suggested many textual improvements. The second
-author's research was partially supported by the Deutsche
-Forschungs\-gemein\-schaft (DFG) project \relax{Hardening the Hammer} (grant
-Ni\,491\slash 14-1).
+Summerfield, and Dmitriy Traytel suggested many textual improvements.
+The second author's
+research was partially supported by the Deutsche
+Forschungs\-gemein\-schaft %(DFG)
+project \relax{Hardening the Hammer} (grant Ni\,491\slash 14-1).
 

### 2015-02-24 tex/body.tex
Minor improvements.

@@
 and let $\SSSS \AST_t$ be the limit of this sequence.
-This is a finite set for each $t$, because in the absence of (co)recursion
-one eventually reaches a point where no selector can be applied.
+This is a finite set for each $t$, 
+because all chains of selectors applied to $t$ are finite.
+%because in the absence of (co)recursion
+%one eventually reaches a point where no selector can be applied.
 %%% TODO: Clarify above step in report.
@@
 In a derivation tree with root node~$\Ec$,
-it can be shown by %case analysis
-induction on the rules of the calculus that each non-root node $\Fc$ is such that
+it can be shown by case analysis
+on the rules of the calculus and by induction that each non-root node $\Fc$ is such that
 $\Terms(\Fc) \subseteq \tcEc$, and hence contains an equality between two terms from $\tcEc$ not occurring in its parent node.
@@
 would apply \rn{Split} and \rn{Inject} before adding this equality.
-
 To reduce the number of unique literals considered by the calculus, we
@@
 
-The efficiency properties of the procedure are mostly as described by Barrett et
+An evaluation of the efficiency of the procedure is described by Barrett et
 al.\ \cite{barrett-et-al-2007}. In particular, it is both theoretically and

### 2015-02-24 tex/body.tex
rule induction

@@
 In a derivation tree with root node~$\Ec$,
-it can be shown by case analysis
-on the rules of the calculus and by induction that each non-root node $\Fc$ is such that
+it can be shown by induction
+on the rules of the calculus that each non-root node $\Fc$ is such that
 $\Terms(\Fc) \subseteq \tcEc$, and hence contains an equality between two terms from $\tcEc$ not occurring in its parent node.

### 2015-02-24 tex/body.tex
minor

@@
 By Theorems~\ref{thm:t},~\ref{thm:rs}, and~\ref{thm:ss}, the
-calculus is sound and complete for the universal $\thD$ theory. We can
-rightly call it a decision procedure.
+calculus is sound and complete for the universal theory of (co)datatypes. We can
+rightly call it a decision procedure for that theory.
 

### 2015-02-24 tex/body.tex
more

@@
 
+{\looseness=-1
 For each goal in each theory, Sledgehammer was invoked to select up to
@@
 along with the goal \cite{blanchette-et-al-2013-smt}. The resulting problem was
-given to the development version of CVC4 and to Z3 4.3.2, each running for up to 60
-seconds on the StarExec cluster \cite{stump-et-al-2014-starexec}.
-%TODO: Keep or not? Problems not involving (co)datatypes were filtered out.
+given to the development version of CVC4 and to Z3 4.3.2 for comparison, each
+running for up to 60 seconds on the StarExec cluster
+\cite{stump-et-al-2014-starexec}.
+%Problems not involving (co)datatypes were left out.
 Due to the lack of machinery for reconstructing inferences about (co)datatypes
-in Isabelle, the solvers are trusted as oracles in these experiments. Z3
-is included for comparison.
+in Isabelle, the solvers are trusted as oracles. % in these experiments.
+%Z3 is included for comparison.
+
+}
 

### 2015-02-24 tex/body.tex
tweaked

@@
 given to the development version of CVC4 and to Z3 4.3.2 for comparison, each
-running for up to 60 seconds on the StarExec cluster
+running for up to 60~s on the StarExec cluster
 \cite{stump-et-al-2014-starexec}.
-%Problems not involving (co)datatypes were left out.
+%Problems not involving any (co)datatypes were left out.
 Due to the lack of machinery for reconstructing inferences about (co)datatypes
@@
 comparable to that of arithmetic \cite{blanchette-et-al-2013-smt}, although this
-obviously depends a lot on the application. The raw evaluation data also
-reveals that the stronger decision procedures almost completely subsume the
-weaker ones in practice.
+obviously depends on the application. The raw evaluation data also
+suggest that the theoretically stronger decision procedures almost completely
+subsume the weaker ones in practice.
 

### 2015-02-26 tex/body.tex
more

@@
 We
-included all the theory files from the Isabelle distribution (Distro, 879~goals)
-and the \emph{Archive of Formal Proofs} (AFP, 2974~goals) \cite{klein-et-al-afp}
+included all the theory files from the Isabelle distribution (Distro, \NOK{1123}~goals)
+and the \emph{Archive of Formal Proofs} (AFP, \OK{2995}~goals) \cite{klein-et-al-afp}
 that define codatatypes falling within the supported fragment. We
@@
 by Peter Gammie and Andreas Lochbihler (\gandl,
-317~goals). To exercise the datatype
+\OK{316}~goals). To exercise the datatype
 support, the benchmarks were complemented by theories
@@
 given to the development version of CVC4 and to Z3 4.3.2 for comparison, each
-running for up to 60~s on the StarExec cluster
+running for up to \OK{60}~s on the StarExec cluster
 \cite{stump-et-al-2014-starexec}.
-%Problems not involving any (co)datatypes were left out.
+Problems not involving any (co)datatypes were left out.
 Due to the lack of machinery for reconstructing inferences about (co)datatypes
@@
 No (co)datatypes
-  & 221 & 209 & 775 & 777 & 52 & 51 & 1048 & 1037 \\%[\jot]
+  & 584 & 522 & 1476 & 1261 & 92 & 87 & 2152 & 1870 \\%[\jot]
 %Datatype simplification only
@@
 Datatypes without \rn{Acyclic}
-  & 227 & -- & 780 & -- & 52 & -- & 1059 & -- \\
+  & 582 & -- & \bfseries 1481 & -- & 92 & -- & 2155 & -- \\
 Full datatypes
-  & 227 & 213 & 786 & 791 & 52 & 51 & 1065 & 1055 \\%[\jot]
+  & 582 & 522 & \bfseries 1481 & 1256 & 92 & 85 & 2155 & 1863 \\%[\jot]
 %Codatatype simplification only
@@
 Codatatypes without \rn{Unique}\!\!
-  & 222 & -- & 804 & -- & 56 & -- & 1082 & -- \\
+  & 583 & -- & 1474 & -- & 89 & -- & 2146 & -- \\
 Full codatatypes
-  & 223 & -- & 804 & -- & \bfseries 59 & -- & 1086 & -- \\%[\jot]
+  & \bfseries 586 & -- & 1475 & -- & \bfseries 99 & -- & 2160 & -- \\%[\jot]
 Full (co)datatypes
-  & \bfseries 229 & -- & \bfseries 815 & -- & \bfseries 59 & -- & \bfseries 1103 & -- %\\[\jot]
+  & 584 & -- & 1480 & -- & \bfseries 99 & -- & \bfseries 2163 & -- %\\[\jot]
 \end{tabular}\end{center}

### 2015-02-26 tex/body.tex
slight rephrasing

--- 2015-02-24 tex/body.tex
@@
 
-{\looseness=-1
-For each goal in each theory, Sledgehammer was invoked to select up to
+For each goal in each theory, Sledgehammer selected between 16 and
 256~lemmas, which were monomorphized and translated to SMT-LIB~2
@@
 \cite{stump-et-al-2014-starexec}.
-%Problems not involving any (co)datatypes were left out.
+Problems not involving any (co)datatypes were left out.
 Due to the lack of machinery for reconstructing inferences about (co)datatypes
@@
 %Z3 is included for comparison.
-
-}
 

### 2015-02-27 tex/body.tex
simpler

@@
 in~$\JJJJ$ share the same interpretation for constructor terms and correctly
-applied selector terms (up to renaming of domain elements) but may differ on
-variables and wrongly applied selector terms.
+applied selector terms (up to isomorphism) %renaming of domain elements
+but may differ on variables and wrongly applied selector terms.
 A formula is $\thD$-\emph{satisfiable} if there

### 2015-03-12 tex/body.tex
memoriam

@@
 datatypes and codatatypes. The dual of the acyclicity rule for datatypes is a
-uniqueness rule that identifies observationally equal codatatype values, also in the presence of
-cyclic data. The procedure decides universal
+uniqueness rule that identifies observationally equal codatatype values,
+including cyclic values. The procedure decides universal
 problems and is composable via the Nelson--Oppen method. It has been
@@
 %
-(using an ML-like syntax) introduces a type that
+%(using an ML-like syntax)
+{\looseness=-1
+introduces a type that
 models the natural numbers $\const{Z}$, $\const{S}(\const{Z})$, $\const{S}(\const{S}(\const{Z}))$, $\ldots$\afterLdots{},
@@
 because both sides expand to the infinite term
-$\const{S}(\const{S}(\const{S}(\ldots)))$, which uniquely identifies the
-value~$\infty$.
-Compared with the conventional inductive definition
+$\const{S}(\const{S}(\const{S}(\ldots)))$, which uniquely identifies
+%the value~
+$\infty$.
+Compared with the conventional %inductive
+definition
 \begin{conf}$\vthinspace \keyw{datatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat}) \,\mid\, \const{Infty}$\vvthinspace, \end{conf}%
@@
 the codatatype avoids
-one case by unifying the finite and infinite nonzero cases.
+one case by unifying the infinite and finite nonzero cases.
+
+}
 
@@
 %codatatypes.
-\begin{conf}
+\begin{conf}%
 Detailed proofs %of completeness

### 2015-05-08 tex/body.tex
avoid so many "semantics"
more

@@
 
-Freely generated inductive datatypes are ubiquitous in functional programs and
+Freely generated algebraic datatypes are ubiquitous in functional programs and
 formal specifications. They are especially
@@
 To represent infinite objects, % such as $\const{S}(\const{S}(\const{S}(\ldots))$,
-a natural choice is to turn to coinductive datatypes, or \emph{codatatypes},
-the non-well-founded dual of inductive \emph{datatypes.}
+a natural choice is to turn to coalgebraic datatypes, or \emph{codatatypes},
+the non-well-founded dual of algebraic \emph{datatypes.}
 %
@@
 constructor
-terms, and only those. As a simple example, the coinductive specification
+terms, and only those. As a simple example, the coalgebraic specification
 %
@@
 $\infty$.
-Compared with the conventional %inductive
+Compared with the conventional %algebraic
 definition
@@
 %\ourparagraph{Characterization.}
-One way to define datatypes semantically is as the initial model of the
+One way to characterize datatypes %semantically
+is as the initial model of the
 selector--constructor equations \cite{barrett-et-al-2007}.

### 2015-05-08 tex/body.tex
fixed grant

@@
 Closely related are the automatic structural
-induction in both kinds of provers \cite{kersani-peltier-2013,cruanes-201x,wand-weidenbach-201x,reynolds-kuncak-2015},
+% cruanes-201x,wand-weidenbach-201x,  -- both were rejected
+induction in both kinds of provers \cite{kersani-peltier-2013,reynolds-kuncak-2015},
 the (co)datatype and (co)induction support in Dafny \cite{leino-moskal-2014},
@@
 either all datatypes or all codatatypes. \begin{rep}Polymorphic types, nested
-(co)recursion,\end{rep}\begin{conf}Nested (co)recursion\end{conf} and datatype--codatatype mixtures fall outside this fragment%
+(co)recursion,\end{rep}\begin{conf}Nested (co)recursion\end{conf} and datatype--codatatype mixtures
+fall outside this fragment%
 \dotReportFootnote{%
@@
 Forschungs\-gemein\-schaft %(DFG)
-project \relax{Hardening the Hammer} (grant Ni\,491\slash 14-1).
+project \relax{Hardening the Hammer} (grant NI\,491\slash 14-1).
 

### 2015-05-11 tex/REBUTTAL.txt
moved with the others

Reviewer 1:

Regarding the evaluation, analysis of raw data http://lara.epfl.ch/~reynolds/CADE2015-cdt/ shows that the overhead of using the more expensive rules (Acyclic and Unique) is negligible for this set. The 1059 benchmarks solved by both CVC4 (full datatypes) and CVC4 (datatypes without Acyclic) are solved collectively in 48.7 and 48.4 seconds respectively, and the 1082 benchmarks solved by both CVC4 (full codatatypes) and CVC4 (codatatypes without Unique) are solved in 54.0 and 54.1 seconds respectively. Notice that these rules may increase runtime due to the time spent on computing when they are applicable, but may also decrease runtime due to the solver's ability to find a smaller proof. We will comment on this point in the evaluation.

The statement "the stronger decision procedures almost completely subsume the weaker ones in practice" references the fact that in only 1 case (out of 4170) did a configuration with a weaker decision procedure solve a benchmark that was unsolved by a stronger one. Such outliers are common for benchmarks containing universally quantified formulas, since as we mention in Section 4, the SMT solver is generally incomplete when such formulas are present. We will make this statement explicit.

Reviewer 2 and 3:

We should clarify the relationship between this paper and Barrett et al [2], which discussed an implementation of the datatypes theory solver in CVC3. Regarding implementation, CVC4 was a from-scratch rewrite of CVC3, and does not share any code used in [2]. These two implementations share the same ideas, but differ significantly in practice, even when restricted to datatypes only. For instance, the implementation in CVC4 is built on top of a general-purpose data structure for efficient congruence closure that represents equivalence classes as circular lists, whereas the one in CVC3 uses a union-find data structure. Some of these differences are implicitly reflected in the calculus, in that [2] describes the use of oriented equations, whereas this paper does not. The datatypes theory solver in CVC4 is (as of yet) unpublished work. We will elaborate on this relationship in the paper.

Reviewer 2:

While the addition of codatatypes adds only two rules to the calculus, in our opinion, in terms of metatheory, the difference is significant. Most notably, the proof of completeness in this paper is not a straightforward lifting of the proof completeness in [2]. Since the proof from Section 4 of [2] does not include codatatypes, the proof of completeness in this paper required both a different formalization of values (as closed normal mu-terms), and new reasoning to recognize the set of (possibly cyclic) values V^x_I that may lead us to non-models when constructing A*.

We will add more details regarding the role of A and A* (notice they are required not only to argue for correctness, but also for the SMT solver to be model-producing). We will also comment more on how they are implemented, which we will reference earlier in the paper.

### 2015-05-11 tex/body.tex
Address reviewer comments.

@@
 %For presentation of the rules in this phase, we rely on a representation of terms in the $\mu$-notation.
-The premises of the rules that make up the second phase refer to a mapping $\Val$
+%The premises of the rules that make up the second phase refer to
+We describe the rules in this phase in terms of a mapping $\Val$
 that associates with each equivalence class a $\mu$-term as its representative.
-%
+\rem{Not explicitly constructed, add here?}
+
 Formally, $\mu$-\emph{terms} are defined recursively as being either a variable $x$
@@
 the binder can be omitted.
-
 $\FV( t )$ denotes the set of free variables occurring in the $\mu$-term~$t$.
@@
 
-The mapping $\Val$ is computed as follows.
+The mapping $\Val$ is defined as follows.
 %
@@
 rightly call it a decision procedure for that theory.
+We remark that the proof of solution soundness is constructive,
+meaning that it provides a method for constructing a model for a saturated configuration, by means of the mapping $\ValC$.
 
@@
 
-An evaluation of the efficiency of the procedure is described by Barrett et
-al.\ \cite{barrett-et-al-2007}. In particular, it is both theoretically and
-empirically beneficial to delay applications of \rn{Split} as long as
-possible. Similarly, \rn{Acyclic} and \rn{Unique} are fairly expensive because
-they require traversing the equivalence classes, which is why they are part of
-phase~2.
+% AJR : this paragraph may have been what confused Reviewer #2. It does suggest that we are copying verbatim what was in CVC3 (which is what Reviewer #2 thought).
+% Since we aren't saying anything new beyond what is said in Barrett et al (and honestly, we didn't critically evaluate the performance of the updated implementation), 
+% maybe we should just remove this paragraph.
+
+%An evaluation of the efficiency of the procedure is described by Barrett et
+%al.\ \cite{barrett-et-al-2007}. In particular, it is both theoretically and
+%empirically beneficial to delay applications of \rn{Split} as long as
+%possible. Similarly, \rn{Acyclic} and \rn{Unique} are fairly expensive because
+%they require traversing the equivalence classes, which is why they are part of
+%phase~2.
 
@@
 subsume the weaker ones in practice.
+In particular, we encountered only one benchmark over the entire set of 4170 benchmarks
+that was solved by a configuration of CVC4 and unsolved by a configuration of CVC4 with more features enabled.
 

### 2015-05-11 tex/body.tex
some work on eval

--- 2015-05-08 tex/body.tex
@@
 \hyphenation{data-type data-types co-data-type co-data-types isa-belle sledge-hammer
-non-redun-dant counter-examples counter-example}
+non-redun-dant counter-examples counter-example mono-morph-ize mono-morph-ized}
 
@@
 
-For each goal in each theory, Sledgehammer selected between 16 and
-256~lemmas, which were monomorphized and translated to SMT-LIB~2
-along with the goal \cite{blanchette-et-al-2013-smt}. The resulting problem was
-given to the development version of CVC4 and to Z3 4.3.2 for comparison, each
-running for up to 60~s on the StarExec cluster
+For each goal in each theory, Sledgehammer selected
+\begin{conf}16~lemmas\end{conf}%
+\begin{rep}a fixed number of lemmas\end{rep},
+which were monomorphized and translated to SMT-LIB~2 along with the goal
+\cite{blanchette-et-al-2013-smt}. The resulting problem was given to the
+development version of CVC4 and to Z3 4.3.2 for comparison, each running for
+up to 60~s on the StarExec cluster
 \cite{stump-et-al-2014-starexec}.
@@
 \newcommand\nummodden{num\hbox{\rm\_}mod\hbox{\rm\_}den}
+
+\newcommand\captioni{\,Number of solved goals for the three benchmark suites}
+\begin{rep}
+\renewcommand\captioni{\,Number of solved goals with 16~lemmas per goal}
+\end{rep}
 
@@
 \end{tabular}\end{center}
-\caption{\,Number of solved goals for the three benchmark suites}
-\label{tab:bench}
+\caption{\captioni}
+\label{tab:benchsixteen}
 \end{table*}
 
-The results are summarized in Table~\ref{tab:bench}. The decision procedure
+\begin{rep}
+\begin{table*}[t!]
+\normalsize
+\begin{center}\begin{tabular}{@{\;}l@{\kern1.5em}cc@{\kern.5em}cc@{\kern.5em}cc@{\kern1.3em}cc@{\;}}
+  & \multicolumn{2}{c}{Distro} & \multicolumn{2}{c}{AFP} & \multicolumn{2}{c}{\gandl} & \multicolumn{2}{c}{Overall} \\[-0.5pt]
+  & \HD{CVC4} & \HD{Z3} & \HD{CVC4} & \HD{Z3} & \HD{CVC4} & \HD{Z3} & \HD{CVC4}& \HD{Z3} 
+\MIDRULE
+No (co)datatypes
+  & 221 & 209 & 775 & 777 & 52 & 51 & 1048 & 1037 \\%[\jot]
+%Datatype simplification only
+%  &  \\
+Datatypes without \rn{Acyclic}
+  & 227 & -- & 780 & -- & 52 & -- & 1059 & -- \\
+Full datatypes
+  & 227 & 213 & 786 & 791 & 52 & 51 & 1065 & 1055 \\%[\jot]
+%Codatatype simplification only
+%  &  \\
+Codatatypes without \rn{Unique}\!\!
+  & 222 & -- & 804 & -- & 56 & -- & 1082 & -- \\
+Full codatatypes
+  & 223 & -- & 804 & -- & \bfseries 59 & -- & 1086 & -- \\%[\jot]
+Full (co)datatypes
+  & \bfseries 229 & -- & \bfseries 815 & -- & \bfseries 59 & -- & \bfseries 1103 & -- %\\[\jot]
+\end{tabular}\end{center}
+\caption{\,Number of solved goals with 256~lemmas per goal}
+\label{tab:benchtwofiftysix}
+\end{table*}
+\end{rep}
+
+The results are summarized in
+\begin{conf}Table~\ref{tab:benchsixteen}\end{conf}%
+\begin{rep}Tables \ref{tab:benchsixteen} and \ref{tab:benchtwofiftysix}\end{rep}.
+\begin{rep}%
+The goals in Table \ref{tab:benchsixteen} include 16
+heuristically selected lemmas (before monomorphization);
+%\cite{meng-paulson-2009-relev};
+that number is increased to 256 for Table \ref{tab:benchtwofiftysix}.
+\par
+\end{rep}%
+The decision procedure
 makes a difference across all three benchmark suites. It accounts for an overall
-success rate increase of over \OK{5\%}, which is quite significant. For Isabelle
-users, the practical usefulness of interpreted (co)datatypes is roughly
-comparable to that of arithmetic \cite{blanchette-et-al-2013-smt}, although this
-obviously depends on the application. The raw evaluation data also
+success rate increase of over \OK{5\%}, which is quite significant.
+%For Isabelle
+%users, the practical usefulness of interpreted (co)datatypes is roughly
+%comparable to that of arithmetic \cite{blanchette-et-al-2013-smt}, although this
+%obviously depends on the application.
+The raw evaluation data also
 suggest that the theoretically stronger decision procedures almost completely

### 2015-05-11 tex/body.tex
more

@@
 of rewriting and a theory solver (Section~\ref{sec:implementation-as-a-theory-solver-in-cvc4}).
-It consists of about 2000 lines of \cpp{} code, %among which \NOK{1600} are
+It consists of about 2000 lines of \cpp{} code, %among which \OK{1600} are
 most of which are shared between datatypes and codatatypes. The code is
@@
 
-For each goal in each theory, Sledgehammer selected
-\begin{conf}16~lemmas\end{conf}%
-\begin{rep}a fixed number of lemmas\end{rep},
+For each goal in each theory, Sledgehammer selected about 16~lemmas,
+%\end{conf}%
+%\begin{rep}a fixed number of lemmas\end{rep},
 which were monomorphized and translated to SMT-LIB~2 along with the goal
@@
 \newcommand\captioni{\,Number of solved goals for the three benchmark suites}
-\begin{rep}
-\renewcommand\captioni{\,Number of solved goals with 16~lemmas per goal}
-\end{rep}
+%\begin{rep}
+%\renewcommand\captioni{\,Number of solved goals with 16~lemmas per goal}
+%\end{rep}
 
@@
 
-\begin{rep}
-\begin{table*}[t!]
-\normalsize
-\begin{center}\begin{tabular}{@{\;}l@{\kern1.5em}cc@{\kern.5em}cc@{\kern.5em}cc@{\kern1.3em}cc@{\;}}
-  & \multicolumn{2}{c}{Distro} & \multicolumn{2}{c}{AFP} & \multicolumn{2}{c}{\gandl} & \multicolumn{2}{c}{Overall} \\[-0.5pt]
-  & \HD{CVC4} & \HD{Z3} & \HD{CVC4} & \HD{Z3} & \HD{CVC4} & \HD{Z3} & \HD{CVC4}& \HD{Z3} 
-\MIDRULE
-No (co)datatypes
-  & 221 & 209 & 775 & 777 & 52 & 51 & 1048 & 1037 \\%[\jot]
-%Datatype simplification only
-%  &  \\
-Datatypes without \rn{Acyclic}
-  & 227 & -- & 780 & -- & 52 & -- & 1059 & -- \\
-Full datatypes
-  & 227 & 213 & 786 & 791 & 52 & 51 & 1065 & 1055 \\%[\jot]
-%Codatatype simplification only
-%  &  \\
-Codatatypes without \rn{Unique}\!\!
-  & 222 & -- & 804 & -- & 56 & -- & 1082 & -- \\
-Full codatatypes
-  & 223 & -- & 804 & -- & \bfseries 59 & -- & 1086 & -- \\%[\jot]
-Full (co)datatypes
-  & \bfseries 229 & -- & \bfseries 815 & -- & \bfseries 59 & -- & \bfseries 1103 & -- %\\[\jot]
-\end{tabular}\end{center}
-\caption{\,Number of solved goals with 256~lemmas per goal}
-\label{tab:benchtwofiftysix}
-\end{table*}
-\end{rep}
-
-The results are summarized in
-\begin{conf}Table~\ref{tab:benchsixteen}\end{conf}%
-\begin{rep}Tables \ref{tab:benchsixteen} and \ref{tab:benchtwofiftysix}\end{rep}.
-\begin{rep}%
-The goals in Table \ref{tab:benchsixteen} include 16
-heuristically selected lemmas (before monomorphization);
+The results are summarized in Table~\ref{tab:benchsixteen}.
+%The goals in Table \ref{tab:benchsixteen} include 16
+%heuristically selected lemmas (before monomorphization);
 %\cite{meng-paulson-2009-relev};
-that number is increased to 256 for Table \ref{tab:benchtwofiftysix}.
-\par
-\end{rep}%
+%that number is increased to 256 for Table \ref{tab:benchtwofiftysix}.
 The decision procedure
@@
 subsume the weaker ones in practice.
-In particular, we encountered only one benchmark over the entire set of 4170 benchmarks
+In particular, we encountered only one benchmark over the entire set of \OK{4170} benchmarks
 that was solved by a configuration of CVC4 and unsolved by a configuration of CVC4 with more features enabled.
@@
 Pascal Fontaine, Andreas Lochbihler, Andrei Popescu, Christophe Ringeissen, Mark
-Summerfield, and Dmitriy Traytel suggested many textual improvements.
-The second author's
-research was partially supported by the Deutsche
+Summerfield, Dmitriy Traytel, and the anonymous reviewers suggested many
+textual improvements. The second author's research was partially supported by
+the Deutsche
 Forschungs\-gemein\-schaft %(DFG)
-project \relax{Hardening the Hammer} (grant NI\,491\slash 14-1).
+project ``Den Hammer h\"arten'' (grant NI\,491\slash 14-1)
+and the Inria technological development
+action ``Contre-exemples Utilisables par Isabelle et Coq'' (CUIC).
 

### 2015-05-11 tex/body.tex
commented out remark for now
reintroduced para in modified form

@@
 that associates with each equivalence class a $\mu$-term as its representative.
-\rem{Not explicitly constructed, add here?}
+%\rem{Not explicitly constructed, add here?}
 
@@
 rightly call it a decision procedure for that theory.
-We remark that the proof of solution soundness is constructive,
-meaning that it provides a method for constructing a model for a saturated configuration, by means of the mapping $\ValC$.
+The proof of solution soundness is constructive in that it provides a
+method for constructing a model for a saturated configuration, by means of the
+mapping $\ValC$.
 
@@
 
-%An evaluation of the efficiency of the procedure is described by Barrett et
-%al.\ \cite{barrett-et-al-2007}. In particular, it is both theoretically and
-%empirically beneficial to delay applications of \rn{Split} as long as
-%possible. Similarly, \rn{Acyclic} and \rn{Unique} are fairly expensive because
-%they require traversing the equivalence classes, which is why they are part of
-%phase~2.
+As Barrett et al.\ \cite{barrett-et-al-2007} observed for their calculus,
+it is both theoretically and
+empirically beneficial to delay applications of \rn{Split} as long as
+possible. Similarly, \rn{Acyclic} and \rn{Unique} are fairly expensive because
+they require traversing the equivalence classes, which is why they are part of
+phase~2.
 

### 2015-05-11 tex/body.tex
some edits

@@
 abbreviated to
-$\expand{t}{u}$, as returned by the following recursive function:
+$\expand{t}{u}$, as returned by the function
 %where $x \leadsto \MU x.\; \const{C}( \vec u ) \in t$ indicates that $\MU x.\;
@@
 domain elements to obtain a genuine model if desired. %Hence, we focus on normal interpretations.
+
+\pagebreak[2]
 
@@
 \end{cases}}$
+
+\pagebreak[2]
 
@@
 by Peter Gammie and Andreas Lochbihler (\gandl,
-317~goals). To exercise the datatype
-support, the benchmarks were complemented by theories
-about lists and trees. The theories were selected before conducting the
-experiments. The experimental data are publicly available.%
+317~goals). To exercise the datatype support, theories about lists and trees
+were added to the first two benchmark sets. The theories were selected before
+conducting any experiments. The experimental data are available
+online.%
 \footnote{\url{http://lara.epfl.ch/~reynolds/CADE2015-cdt/}}
+
 %http://www21.in.tum.de/~blanchet/codp_data.tgz}}
@@
 suggest that the theoretically stronger decision procedures almost completely
-subsume the weaker ones in practice.
-In particular, we encountered only one benchmark over the entire set of \OK{4170} benchmarks
-that was solved by a configuration of CVC4 and unsolved by a configuration of CVC4 with more features enabled.
+subsume the weaker ones in practice: We encountered only one goal (out of \OK{4170})
+that was solved by a configuration of CVC4 and unsolved by a configuration of
+CVC4 with more features enabled.
 

### 2015-05-11 tex/body.tex
comparison with Barrett et al.

@@
 calculus and is composable via the Nelson--Oppen method \cite{nelson-oppen-1979}.
-It generalizes the procedure by Barrett et al.\ \cite{barrett-et-al-2007}, which covers
-only datatypes.
+It generalizes the procedure by Barrett et al.\ \cite{barrett-et-al-2007},
+which covers only datatypes.
 %To our knowledge, our procedure is the first of its kind for the theory of
@@
 These two rules are needed to ensure completeness
-(model soundness) on universal problems. %and the absence of spurious models.
+(solution soundness) on universal problems. %and the absence of spurious models.
 They cannot be expressed as finite axiomatizations, so they naturally belong in
@@
 et al.\ \cite{barrett-et-al-2007} provide a good account of
-related work on datatypes as of 2007. Since then, datatypes have been added not only to CVC4
+related work on datatypes as of 2007, in addition to describing their
+implementation in CVC3.
+Since then, datatypes have been added not only to CVC4 (a complete rewrite of CVC3)
 but also to the SMT solver
@@
 \begin{conf}
-\kern-2.9pt %%% TYPESETTING: Crude hack
+\kern-3pt %%% TYPESETTING: Crude hack
 \end{conf}
@@
 $\const{true},\, \const{false} : \ty{bool}$, interpreted as truth values.
-The only predicate is equality ($\teq$)\begin{rep} and belongs to the logical symbols\end{rep};
-other predicates can be represented as functions to $\ty{bool}$\begin{rep},
+The only predicate is equality ($\teq$)\begin{rep} and belongs to the logical symbols;
+other predicates can be represented as functions to $\ty{bool}$,
 with $\const{p}(\ldots)$ abbreviating $\const{p}(\ldots) \teq \const{true}$\end{rep}.
@@
 \begin{rep}The operator $\bigwedge_{\,i}\, \varphi_i$ abbreviates a conjunction
-$\varphi_1 \mathrel\land \cdots \mathrel\land \varphi_n$. \end{rep}%
-Finally, $\bar x$ abbreviates a list or tuple $x_1,\ldots,x_n$.
+$\varphi_1 \mathrel\land \cdots \mathrel\land \varphi_n$.
+Finally, $\bar x$ abbreviates a list or tuple $x_1,\ldots,x_n$.\end{rep}
 
@@
 Proving a universal quantifier-free conjecture is reduced to showing that
-its negation is unsatisfiable.
+its negation is unsatisfiable. The presentation is
+inspired by Barrett et al.\ \cite{barrett-et-al-2007} but
+higher-level, using unoriented equations instead of oriented ones.
+
 % whose derivation rules are applied until saturation or contradiction.
@@
 
-As Barrett et al.\ \cite{barrett-et-al-2007} observed for their calculus,
+As Barrett et al.\ \cite{barrett-et-al-2007} observed for their procedure,
 it is both theoretically and

### 2015-05-11 tex/body.tex
minor

@@
 %
-$\Ec \mathrel\cup \{ \const C_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\smash{\,n_{\negvthinspace j}}}_{\negvthinspace j}( t ) \bigr) \mid t\typ{\delta} \in \SSS \mathrel\cup \SSSS \AST{,}~ \const C_{\negvthinspace j} \in \Ctr^\delta \}$.
+$\Ec \mathrel\cup \{ \smash{\const C_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\smash{\,n_{\negvthinspace j}}}_{\negvthinspace j}( t ) \bigr)} \mid \smash{t\typ{\delta} \in \SSS \mathrel\cup \SSSS \AST{,}~ \const C_{\negvthinspace j} \in \smash{\Ctr^\delta}} \}$.
 %
@@
 This requires extending the decision procedure with several
-rules \cite{barrett-et-al-2007},
+rules, % \cite{barrett-et-al-2007},
 which apply uniformly to datatypes and codatatypes.
@@
 To evaluate the decision procedure, we generated benchmark problems from existing
-Isabelle formalizations using Sledgehammer.
+Isabelle formalizations using Sledgehammer \cite{blanchette-et-al-2013-smt}.
 %There are not yet many benchmarks available, since codatatypes are a
@@
 %\begin{rep}a fixed number of lemmas\end{rep},
-which were monomorphized and translated to SMT-LIB~2 along with the goal
-\cite{blanchette-et-al-2013-smt}. The resulting problem was given to the
+which were monomorphized and translated to SMT-LIB~2 along with the goal.
+The resulting problem was given to the
 development version of CVC4 and to Z3 4.3.2 for comparison, each running for

### 2015-05-12 tex/body.tex
v minor edits

@@
 \newcommand{\Val}{\mathcalx{A}\vvthinspace}
-\newcommand{\ValC}{\mathcalx{A}^\star}
+\newcommand{\ValC}{\smash{\mathcalx{A}^\star}}
 \newcommand{\Var}{\mathcalx{V}}
@@
 $u$ is a proper subterm of $t$,
-$t$ is of the form $\MU x.\; \const{C}( t_1, \ldots, t_n )$,
-$u$ is of the form $\MU y.\; \const{C}( u_1, \ldots, u_n )$,
+$t$ and $u$ are of the forms $\MU x.\; \const{C}( t_1, \ldots, t_n )$
+and $\MU y.\; \const{C}( u_1, \ldots, u_n )$,
 and $\expand{t}{ t_k } \aequiv \expand{t}{ u_k }$ for all $k$. % \in [n]
-The term $t$ is \emph{normal} if it does not contain a self-similar subterm
+The $\mu$-term $t$ is \emph{normal} if it does not contain %any
+self-similar subterms
 and all of its proper subterms are also normal.
-For example, $t = \MU x.\; \const{C}( \MU y.\; \const{C}( y ) )$ is not normal
+Thus, $t = \MU x.\; \const{C}( \MU y.\; \const{C}( y ) )$ is not normal
 because $\MU y.\; \const{C}( y )$ is a self-similar subterm of $t$.
@@
  = \MU y.\; \const{C}\bigl( \expand{u}{x}^{\smash{\{y\}}} \bigr)
- = \MU y.\; \const{C}\bigl( \expand{u}{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) ) }^{\smash{\{y\}}} \bigr)
- = \MU y.\; \const{C}\bigl( \MU x.\; \const{C}( \expand{u}{\MU y.\; \const{C}( x )}^{\smash{\{x,y\}}} ) \bigr)
+ = \MU y.\; \const{C}\bigl( \expand{u}{\MU x.\allowbreak\; \const{C}( \MU y.\allowbreak\; \const{C}( x ) ) }^{\smash{\{y\}}} \bigr)
+ = \MU y.\; \const{C}\bigl( \MU x.\allowbreak\; \const{C}( \expand{u}{\MU y.\allowbreak\; \const{C}( x )}^{\smash{\{x,y\}}} ) \bigr)
  = \MU y.\allowbreak\; \const{C}( \MU x.\allowbreak\; \const{C}( y ) )$
@@
 $\interp{\J}{\tau}$ includes a maximal set of closed normal $\mu$-terms of that type that are
-unique up to $\alpha$-equivalence and acyclic if $\tau \notin \Codata$.
+unique up to $\alpha$-equivalence and acyclic if $\tau \notin \Codata$.\strut
+
 \item
-For each constructor term $\const{C}( \vec t \,)$ of type $\tau$,
+For each constructor term $\vphantom{s^k}\const{C}( \vec t \,)$ of type $\tau$,
 $\interp{\J}{\const{C}}( \interp{\J}{\vec t} )$ is the value
 in $\interp{\J}{\tau}$ that is $\alpha$-equivalent to
-$\nf{\MU x.\; \const{C}( \interp{\J}{\vec t} )}$, where $x$ is fresh.
+$\nf{\MU x.\; \const{C}( \interp{\J}{\vec t} )}$, where $x$ is fresh.\strut
+
 \item
@@
 We write $t \vsimv{x} u$ if $\mu$-terms $t$ and $u$ are syntactically equivalent
-for some renaming that avoids capturing any variable other than $x$.
+for some renaming that avoids capturing any variable other than~$x$.
 For example,
@@
 %where $\expand{\ValC \ec{u} }{ w_i } = \ValC \ec{ \const s^i_{\negvthinspace\jPrime}( u )}$ for $i \in [n]$.
-Since \rn{Inject} and \rn{Clash} do not apply, % to $\Fc\!$,
-by construction $\ValC$ must be of the form
+Since \rn{Inject} and \rn{Clash} are not applicable, % to $\Fc\!$,
+by construction $\ValC\ec{u}$ must be of the form
 $\MU {\Varec{u}}.\; \const C_{\negvthinspace\jPrime}( w_1, \ldots, w_n )$,
@@
 $\muvar(w_i) = \smash{\Varec{\vphantom{I}\smash{\const s^i_{\negvthinspace\jPrime}( u )}}}$
-for each $i$ and thus by Lemma~\ref{lem:interpretation-completion}\PART{2},
+for each $i$, and thus by Lemma~\ref{lem:interpretation-completion}\PART{2},
 $\expand{\ValC \ec{u}}{ w_i } = \ValC \ec{ \smash{\const s^i_{\negvthinspace\jPrime}( u )}}$.
@@
 % where $a \in M$ for all atoms $a$ that
-%it assigned the value $\top$ and $\neg\, a \in M$ for all atoms that it assigns
+%it assigned the value $\top$ and $\neg\: a \in M$ for all atoms that it assigns
 %the value $\bot$.
@@
 either reports that a subset $C$ of $M_i$ is $T_i$-unsatisfiable, in which case
-$\neg\, C$ is added to $F\!$, adds a clause to $F\!$, or does nothing.
+$\neg\: C$ is added to $F\!$, adds a clause to $F\!$, or does nothing.
 When $M$ is a complete assignment for $F\!$, a theory solver can choose to do

### 2015-05-19 tex/body.tex
more

@@
 \usepackage{bussproofs}
-\usepackage{calc}
 \usepackage{cite}
 \usepackage{mathptmx}
-%\usepackage{txfonts}
-%\usepackage{mathrsfs}
-%\usepackage{pifont}
-%\usepackage{smallcap}
 \usepackage{mathpartir}
@@
 \usepackage{subfigure}
-\usepackage[usenames]{color}
-%\usepackage{graphicx}
-%\usepackage{newcent}
 \usepackage{textcomp}
-%\usepackage{tipa}
-\usepackage{units}
 \usepackage{url}
 \usepackage[all]{xy}
+
+\newcommand\beforefig{\rm}
+
+\newcommand\DISC[1]{t \teq \const{C}_{#1}\negvthinspace\bigl(\const s^1_{#1}( t ), \ldots,\const s^{n_{#1}}_{#1}( t )\bigr)}
+%\newcommand\DISC[1]{\const{d}_{#1}(t)}
+
+%\newcommand\XXXL{DPLL}
+\newcommand\XXXL{CDCL}
+
+\newcommand\NOK[1]{\textcolor{red}{\textbf{#1}}}
+\newcommand\OK[1]{#1}
+
+%\newcommand\typ[1]{: #1}
+\newcommand\typ[1]{^{\vthinspace #1}}
+
+%\newcommand\PART[1]{.#1}
+\newcommand\PART[1]{(#1)}
+
+%\newcommand\AST{\ast}
+\newcommand\AST{\infty} %%% matter of taste -- revert if you don't like
+
+\newcommand\MU{\vvthinspace\mu\vvthinspace}
+
+%%% not part of shared counter
 
@@
 
-%\newcommand\ourparagraph[1]{\paragraph{\upshape\textbf{#1}}}
-%\newcommand\beforefig{}
-
-\newcommand\ourparagraph[1]{\paragraph{{#1}}}
-\newcommand\beforefig{\rm}
-
-\newcommand\DISC[1]{t \teq \const{C}_{#1}\negvthinspace\bigl(\const s^1_{#1}( t ), \ldots,\const s^{n_{#1}}_{#1}( t )\bigr)}
-%\newcommand\DISC[1]{\const{d}_{#1}(t)}
-
-%\newcommand\XXXL{DPLL}
-\newcommand\XXXL{CDCL}
-
-\newcommand\NOK[1]{\textcolor{red}{\textbf{#1}}}
-\newcommand\OK[1]{#1}
-
-%\newcommand\typ[1]{: #1}
-\newcommand\typ[1]{^{\vthinspace #1}}
-
-%\newcommand\PART[1]{.#1}
-\newcommand\PART[1]{(#1)}
-
-%\newcommand\AST{\ast}
-\newcommand\AST{\infty} %%% matter of taste -- revert if you don't like
-
-\newcommand\MU{\vvthinspace\mu\vvthinspace}
-
-%%% not part of shared counter
 \spnXwtheorem{examplex}{Example}{\itshape}{\rmfamily}
@@
 
-\begin{rep}\ourparagraph{Related Work.} Barrett\end{rep}%
+\begin{rep}\paragraph{Related Work.} Barrett\end{rep}%
 \begin{conf}Barrett\end{conf}
@@
 
-\ourparagraph{Conventions.}
+\paragraph{Conventions.}
 Our setting is a monomorphic (or many-sorted) first-order logic.
@@
 
-%\ourparagraph{Specifications.}
+%\paragraph{Specifications.}
 In an SMT problem, the signature is typically given by specifying first the
@@
 
-%\ourparagraph{Characterization.}
 One way to characterize datatypes %semantically
@@
 
-%\ourparagraph{Special Cases.}
 {\looseness=-1
@@
 
-\ourparagraph{Phase 1: Computing the Bidirectional Closure \beforefig(\figurename~\ref{fig:cc-rules}).}
+\paragraph{Phase 1: Computing the Bidirectional Closure \beforefig(\figurename~\ref{fig:cc-rules}).}
 In conjunction with \rn{Refl}, \rn{Sym}, and \rn{Trans}, the \rn{Cong} rule computes the congruence (upward) closure,
@@
 
-\ourparagraph{Phase 2: Applying Acyclicity and Uniqueness \beforefig(\figurename~\ref{fig:ab-rules}).}
+\paragraph{Phase 2: Applying Acyclicity and Uniqueness \beforefig(\figurename~\ref{fig:ab-rules}).}
 %For presentation of the rules in this phase, we rely on a representation of terms in the $\mu$-notation.
@@
 
-\ourparagraph{Phase 3: Branching \beforefig(\figurename~\ref{fig:split-rule}).}
+\paragraph{Phase 3: Branching \beforefig(\figurename~\ref{fig:split-rule}).}
 If a selector is applied to a term $t$, or if $t$'s type is a finite datatype,
@@
 
-\ourparagraph{Correctness.}
+\paragraph{Correctness.}
 Correctness means that if there exists a closed derivation tree with root node
@@
 
-%%% @ANDY: I'm trying without \ourparagraph's, to save space but also to make the
+%%% @ANDY: I'm trying without \paragraph's, to save space but also to make the
 %%% section look more like a whole than a random list of stuff. Revert if you
@@
 
-%\ourparagraph{Integration into {\relax\XXXL($T\negvthinspace$)}.}
 The decision procedure is implemented as a \relax{theory solver} of CVC4, that is, a
@@
 
-%\ourparagraph{Extension to Quantified Formulas.}
 While the decision procedure is restricted to universal conjectures, in practice
@@
 
-%\ourparagraph{Optimizations.}
 The implementation of the decision procedure uses several optimizations

### 2015-05-19 tex/body.tex
more
chased macros

@@
 
-\newcommand\beforefig{\rm}
-
 \newcommand\DISC[1]{t \teq \const{C}_{#1}\negvthinspace\bigl(\const s^1_{#1}( t ), \ldots,\const s^{n_{#1}}_{#1}( t )\bigr)}
-%\newcommand\DISC[1]{\const{d}_{#1}(t)}
-
-%\newcommand\XXXL{DPLL}
-\newcommand\XXXL{CDCL}
-
-\newcommand\NOK[1]{\textcolor{red}{\textbf{#1}}}
-\newcommand\OK[1]{#1}
-
-%\newcommand\typ[1]{: #1}
+
 \newcommand\typ[1]{^{\vthinspace #1}}
 
-%\newcommand\PART[1]{.#1}
-\newcommand\PART[1]{(#1)}
-
-%\newcommand\AST{\ast}
-\newcommand\AST{\infty} %%% matter of taste -- revert if you don't like
-
 \newcommand\MU{\vvthinspace\mu\vvthinspace}
-
-%%% not part of shared counter
 
@@
 \makeatother
-
 \spnXwtheorem{examplex}{Example}{\itshape}{\rmfamily}
@@
 
-\renewcommand\models{\vDash} %%% matter of taste
-
 \def\thewordpaper{paper}
@@
 
-%%% I don't like using \qed or anything too similar (e.g. \Box) for
-%%% anything else than "quid erat demonstrandum". I'm open to other symbols,
-%%% e.g. a filled black square. But if the example are short enough, which
-%%% they currently are, we could live with nothing.
-%%%
 \newcommand\xend{{\hfill$\scriptstyle\blacksquare$}}
-%\newcommand\xend{{\qed}}
-%\newcommand\xend{{}}
-
-%\renewcommand{\vec}[1]{\mathbf{#1}}
+
 \renewcommand{\vec}[1]{\bar #1}
-%\newcommand{\Ec}{\mathsf{E}}
 \newcommand{\Ec}{E}
-%\newcommand{\Ec}{\mathcalx{E}}
-%\newcommand{\Fc}{\mathsf{F}}
 \newcommand{\Fc}{F}
-%\newcommand{\Fc}{\mathcalx{F}}
-%\newcommand{\Gc}{\mathsf{G}}
-%\newcommand{\Ac}{\mathsf{A}}
-%\newcommand{\Dc}{\mathsf{D}}
-%\newcommand{\Rc}{\mathsf{R}}
-\newcommand{\vrange}{\mathsf{range}}
-\newcommand{\vdom}{\mathsf{dom}}
 \newcommand{\tEc}{\Terms(\Ec)}
-\newcommand{\tcEc}{\Terms^\AST(\Ec)}
+\newcommand{\tcEc}{\Terms^\infty(\Ec)}
 \newcommand{\rn}[1]{\textsf{#1}}
@@
 \newcommand{\tneq}{\not\teq}
-\newcommand{\rem}[1]{\textcolor{red}{[#1]}}
-%\newcommand{\ror}{\parallel}
-%\newcommand{\ror}{{\&}}
-%\newcommand{\ROR}{\kern.45em \ror \kern.45em}
 \newcommand\ROR{\betweenantes}
-\newcommand{\tpath}[2]{\mathcalx{P}_{ #2 \rightarrow \_ }( #1 )}
-\newcommand{\ttpath}[3]{\mathcalx{P}_{ #2 \rightarrow #3 }( #1 )}
-
-%\newcommand{\expand}[2]{#1\llbracket #2 \rrbracket}
-%%% I prefer a notation that puts more emphasis on #2 (which is, in essence,
-%%% the object that is returned) and less on #1
-%\newcommand{\expand}[2]{\llbracket #2 \rrbracket_{#1}}
-%%% And one that doesn't look so semantic
-%\newcommand{\expand}[2]{\lceil #2 \rceil_{#1}}
+
 \newcommand{\expand}[2]{\langle \smash{#2} \rangle_{#1}}
@@
 \newcommand{\interp}[2]{#1(#2)}
-%\newcommand{\interp}[2]{{#2}^{#1}}
-
-\newcommand{\dpath}[3]{\delta^{#1}_{#2}( #3 )}
+
 \newcommand{\ec}[1]{[#1]}
@@
 \newcommand{\ValC}{\smash{\mathcalx{A}^\star}}
-\newcommand{\Var}{\mathcalx{V}}
-%\newcommand{\Varec}[1]{\Var{\ec{#1}}}
 \newcommand{\Varec}[1]{\vvthinspace\widetilde{#1}\vvthinspace}
@@
 \newcommand{\vsimv}[1]{\mathrel{=^{#1}_\alpha}}
-\newcommand{\tpos}[2]{#1\!\mid_{#2}}
-%\newcommand{\muvar}{\mathcalx{Var}}
 \newcommand{\muvar}{\mathrm{Var}} %% cf. FV
@@
 
-% change?
-\newcommand{\thO}{T_{\mathrm{o}}}
-%\newcommand{\thD}{T_{\!\mathrm{cd}}}
-%%% the italicized "T" makes it look as though it were a variable;
-%%% and the subscript means we have two layers of subscripts in things like
-%%% \model_\thD (although we could drop the "T" and keep only the subscript
-%%% like you did in the string paper).
-%%% How about the following? I'm pretty sure \mathcalx has been used for theories
-%%% before by other authors.
 \newcommand{\thD}{\mathcalx{DC}}
@@
 
-%%% REMOVE BEFORE SUBMITTING ABSOLUTELY FINAL VERSION
 \makeatletter
@@
 of rewriting and a theory solver (Section~\ref{sec:implementation-as-a-theory-solver-in-cvc4}).
-It consists of about 2000 lines of \cpp{} code, %among which \OK{1600} are
+It consists of about 2000 lines of \cpp{} code,
 most of which are shared between datatypes and codatatypes. The code is
@@
 
-\paragraph{Phase 1: Computing the Bidirectional Closure \beforefig(\figurename~\ref{fig:cc-rules}).}
+\paragraph{Phase 1: Computing the Bidirectional Closure \rm(\figurename~\ref{fig:cc-rules}).}
 In conjunction with \rn{Refl}, \rn{Sym}, and \rn{Trans}, the \rn{Cong} rule computes the congruence (upward) closure,
@@
 
-\paragraph{Phase 2: Applying Acyclicity and Uniqueness \beforefig(\figurename~\ref{fig:ab-rules}).}
+\paragraph{Phase 2: Applying Acyclicity and Uniqueness \rm(\figurename~\ref{fig:ab-rules}).}
 %For presentation of the rules in this phase, we rely on a representation of terms in the $\mu$-notation.
@@
   \betweenantes
-  %\ttpath{\Ec}{\ec{t}}{\ec{t}} \neq \emptyset
   \Val \smash{\ec{t\typ{\delta}}} = \MU x.\; u
@@
  \betweenantes
- %\tpath{\Ec}{\ec{t}} = \tpath{\Ec}{\ec{u}}\sigma \neq \emptyset
  \Val \smash{\ec{t\typ{\delta}}} \aequiv \Val \smash{\ec{u\typ{\delta}}}
@@
 
-\paragraph{Phase 3: Branching \beforefig(\figurename~\ref{fig:split-rule}).}
+\paragraph{Phase 3: Branching \rm(\figurename~\ref{fig:split-rule}).}
 If a selector is applied to a term $t$, or if $t$'s type is a finite datatype,
@@
 %for $i \geq 0$, let ;
-and let $\SSSS \AST_t$ be the limit of this sequence.
+and let $\SSSS \infty_t$ be the limit of this sequence.
 This is a finite set for each $t$, 
@@
 %%% TODO: Clarify above step in report.
-Let $\SSSS \AST$ be the union of all sets $\SSSS \AST_t$ where $t \in D$,
+Let $\SSSS \infty$ be the union of all sets $\SSSS \infty_t$ where $t \in D$,
 and let $\tcEc$ be the set of subterms of
 %
-$\Ec \mathrel\cup \{ \smash{\const C_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\smash{\,n_{\negvthinspace j}}}_{\negvthinspace j}( t ) \bigr)} \mid \smash{t\typ{\delta} \in \SSS \mathrel\cup \SSSS \AST{,}~ \const C_{\negvthinspace j} \in \smash{\Ctr^\delta}} \}$.
+$\Ec \mathrel\cup \{ \smash{\const C_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\smash{\,n_{\negvthinspace j}}}_{\negvthinspace j}( t ) \bigr)} \mid \smash{t\typ{\delta} \in \SSS \mathrel\cup \SSSS \infty{,}~ \const C_{\negvthinspace j} \in \smash{\Ctr^\delta}} \}$.
 %
@@
 node $\Ec \mathrel\cup t \teq u$ is obtained by applying one of the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, \rn{Inject}, \rn{Unique}, or \rn{Single}.
-In all these cases, $\Ec \models_{\thD} t \teq u$.
+In all these cases, $\Ec \vDash_{\thD} t \teq u$.
 For \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, this is a consequence of equality reasoning.
@@
 Intuitively, this lemma states three properties of $\ValC$ that ensure a normal
-interpretation $\J$ can be constructed that satisfies $\Ec$. Property~\PART{1} states that the
+interpretation $\J$ can be constructed that satisfies $\Ec$. Property~(1) states that the
 values in the range of $\ValC$ are $\alpha$-equivalent to a value in normal
 interpretation. This means they are closed, normal, and acyclic when
-required. Property~\PART{2} states that the interpretation of all subterms in the
+required. Property~(2) states that the interpretation of all subterms in the
 range of $\ValC$ depends on its associated variable only. In other words, the
@@
 $\ValC \ec{x}$, independently of the context. % in which it is interpreted.
-Property~\PART{3} states that $\ValC$ is injective (modulo
+Property~(3) states that $\ValC$ is injective (modulo
 $\alpha$-equivalence), which ensures %in the following
@@
 For the variables, let $\interp{\J}{x\typ{\tau}}$ be the value in $\J(\tau)$ that is $\alpha$-equivalent with $\ValC \ec{x}$ for each variable $x \in \Terms(\Fc)$,
-which by Lemma~\ref{lem:interpretation-completion}\PART{1} is guaranteed to exist.
+which by Lemma~\ref{lem:interpretation-completion}(1) is guaranteed to exist.
 
@@
 that $\interp{\J}{t} \aequiv \ValC \ec{t}$ for all terms $t \in \Terms( \Fc )$,
-which implies $\J \models t_1 \teq t_2$ since $\J$ is normal.
+which implies $\J \vDash t_1 \teq t_2$ since $\J$ is normal.
 
@@
 Call this term $t'$.
-%By Lemma~\ref{lem:interpretation-completion}\PART{2} and due to our construction of $\ValC$, we have $\ValC \ec{t}$ is a term
+%By Lemma~\ref{lem:interpretation-completion}(2) and due to our construction of $\ValC$, we have $\ValC \ec{t}$ is a term
 %of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$
@@
 where
-$\muvar(w_i) = {\Varec{u_i}}$ for each $i$.  Thus by Lemma~\ref{lem:interpretation-completion}\PART{2},
+$\muvar(w_i) = {\Varec{u_i}}$ for each $i$.  Thus by Lemma~\ref{lem:interpretation-completion}(2),
 $\expand{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$.
@@
 $\ec{u}$ must contain a term of the form $\smash{\const C_{\negvthinspace\jPrime}\bigl( \const s^1_{\negvthinspace\jPrime}( u ), \ldots, \const s^{n}_{\negvthinspace\jPrime}( u ) \bigr)}$ for some $\jPrime$.
-%By Lemma~\ref{lem:interpretation-completion}\PART{2} and due to our construction of $\ValC$, we have $\ValC \ec{u}$ is of the form
+%By Lemma~\ref{lem:interpretation-completion}(2) and due to our construction of $\ValC$, we have $\ValC \ec{u}$ is of the form
 %$\MU {\Varec{u}}.\; \const C_{\negvthinspace\jPrime}( w_1, \ldots, w_n )$,
@@
 $\muvar(w_i) = \smash{\Varec{\vphantom{I}\smash{\const s^i_{\negvthinspace\jPrime}( u )}}}$
-for each $i$, and thus by Lemma~\ref{lem:interpretation-completion}\PART{2},
+for each $i$, and thus by Lemma~\ref{lem:interpretation-completion}(2),
 $\expand{\ValC \ec{u}}{ w_i } = \ValC \ec{ \smash{\const s^i_{\negvthinspace\jPrime}( u )}}$.
@@
 is such that $[u] \not= [u']$.
-By the induction hypothesis and Lemma~\ref{lem:interpretation-completion}\PART{3}, $\interp{\J}{u} \neq \interp{\J}{u'}$ for all such $u$, $u'$.
+By the induction hypothesis and Lemma~\ref{lem:interpretation-completion}(3), $\interp{\J}{u} \neq \interp{\J}{u'}$ for all such $u$, $u'$.
 Thus, we may interpret $\smash{\interp{\J}{\const s^k_{\negvthinspace j}}( \interp{\J}{u} )}$ as the value in $\interp{\J}{\tau}$ that is $\alpha$-equivalent with $\ValC \ec{ t }$.
@@
 Since $\interp{\J}{t} \vsim \ValC \ec{t}$ and $\interp{\J}{u} \vsim \ValC \ec{u}$,
-by Lemma~\ref{lem:interpretation-completion}\PART{3}, $\interp{\J}{t} \neq \interp{\J}{u}$, and thus $\J \models t \tneq u$.
-
-Since by assumption $\Fc$ contains only equalities and disequalities, we have $\J \models \Fc\!$,
+by Lemma~\ref{lem:interpretation-completion}(3), $\interp{\J}{t} \neq \interp{\J}{u}$, and thus $\J \vDash t \tneq u$.
+
+Since by assumption $\Fc$ contains only equalities and disequalities, we have $\J \vDash \Fc\!$,
 and since $\Ec \subseteq \Fc\!$,
 % is a superset of $\Ec$,
-we conclude that $\J \models \Ec$.
+we conclude that $\J \vDash \Ec$.
 \qed
@@
 quite a few details. This section describes the main aspects of the
-implementation within the SMT solver CVC4:\ the integration into \XXXL($T$)
+implementation within the SMT solver CVC4:\ the integration into CDCL($T$)
 \cite{ganzinger-et-al-2004}, the extension to quantified formulas, and some of
@@
 Given a theory $T = T_1 \mathrel\cup \cdots \mathrel\cup T_n$ and a set of
-input clauses $F$ in conjunctive normal form, the \XXXL($T$) procedure
+input clauses $F$ in conjunctive normal form, the CDCL($T$) procedure
 incrementally builds partial assignments of
@@
 makes a difference across all three benchmark suites. It accounts for an overall
-success rate increase of over \OK{5\%}, which is quite significant.
+success rate increase of over 5\%, which is quite significant.
 %For Isabelle
@@
 suggest that the theoretically stronger decision procedures almost completely
-subsume the weaker ones in practice: We encountered only one goal (out of \OK{4170})
+subsume the weaker ones in practice: We encountered only one goal (out of 4170)
 that was solved by a configuration of CVC4 and unsolved by a configuration of
@@
 \def\ackname{Acknowledgment}
-\ourparagraph{\footnotesize\ackname.}
+\paragraph{\footnotesize\ackname.}
 \looseness=-1

### 2015-05-19 tex/cade_final/main.tex
more

--- 2015-05-19 tex/body.tex
@@
-%%% TODO: In report, give small example where Single is needed
+%envcountsect,
+\documentclass[a4paper,oribibl,envcountsame,draft]{llncs}
+\usepackage{version}
+
+\excludeversion{rep}
+\includeversion{conf}
+\excludeversion{nomemoriam}
+\includeversion{memoriam}
 
@@
 For example, the type
-
-\begin{conf}
-\kern-2pt %%% TYPESETTING: Crude hack
-\end{conf}
-
-\vskip\abovedisplayskip
-\noindent\centerline{$\keyw{codatatype}~\;\ty{stream}_{\,\tau} \,=\, \const{SCons}(\tau,\: \ty{stream}_{\,\tau})$}
-\vskip\belowdisplayskip
-
-\begin{conf}
-\kern-2pt %%% TYPESETTING: Crude hack
-\end{conf}
-
-\noindent
+$\vthinspace\keyw{codatatype}~\;\ty{stream}_{\,\tau} \,=\allowbreak\, \const{SCons}(\tau,\allowbreak\: \ty{stream}_{\,\tau})\vthinspace$
 of infinite sequences or streams over an element type $\tau$ is
@@
 
-\begin{conf}
-\kern-3pt %%% TYPESETTING: Crude hack
-\end{conf}
-
 \paragraph{Conventions.}
@@
 nonempty domains. The set~$\Types$ must contain a
-\begin{rep}distinguished \end{rep}type \ty{bool} interpreted as the set of truth
+distinguished type \ty{bool} interpreted as the set of truth
 values. %, and may contain other interpreted types (e.g., \ty{int}, \ty{real}).
@@
 whereas $\tau,\:\upsilon$ range over arbitrary types.
-When applied to terms, the symbol $=$ denotes syntactic equality.
 
@@
 The notation $t\typ{\tau}$ stands for a term $t$ of type~$\tau$.
+When applied to terms, the symbol $=$ denotes syntactic equality.
 \begin{rep}The operator $\bigwedge_{\,i}\, \varphi_i$ abbreviates a conjunction
@@
 
-\kern-1pt
-
 %\section{Examples}

### 2015-09-10 conf_and_rep/cade_final/main.tex
renamed

--- 2015-05-19 tex/cade_final/main.tex
@@
-%envcountsect,
 \documentclass[a4paper,oribibl,envcountsame,draft]{llncs}
@@
 \makeatletter
-\def\spnXwtheorem#1#2#3#4{\@spynthm{#1}{#2}{#3}{#4}%
-                         \@addtoreset{#1}{chapter}}%
-\makeatother
+\def\spnXwtheorem#1#2#3#4{\@spynthm{#1}{#2}{#3}{#4}\@addtoreset{#1}{chapter}}\makeatother
 \spnXwtheorem{examplex}{Example}{\itshape}{\rmfamily}
@@
 \newcommand\MIDRULE{
-\\[-1pt] %%% TYPESETTING HACK
-\midrule
-%$\enatT$
-\\[-11pt] %%% TYPESETTING HACK
-}
+\\[-1pt] \midrule
+\\[-11pt] }
 
@@
 
-\newcommand\afterDot{\;} %%% Too little space by default after "Lemma Foo."
-
+\newcommand\afterDot{\;} 
 \begin{rep}
@@
 
-\newcommand\afterLdots{\kern.1em} %% TYPESETTING
-
-% for "bussproofs" package
+\newcommand\afterLdots{\kern.1em} 
 \EnableBpAbbreviations
@@
 
-\let\labelitemi=\labelitemii %% CHEAT!
-
+\let\labelitemi=\labelitemii 
 \newcommand\cpp{C\nobreak\raisebox{.05ex}{+}\nobreak\raisebox{.05ex}{+}}
@@
 \newcommand{\VAREC}[1]{\widetilde{\vphantom{\scriptstyle x^i}\smash{#1}}}
-%\newcommand{\nf}[1]{{{#1}{\downarrow}}}
 \newcommand{\nf}[1]{\lfloor#1\rfloor}
@@
 \newcommand{\vsimv}[1]{\mathrel{=^{#1}_\alpha}}
-\newcommand{\muvar}{\mathrm{Var}} %% cf. FV
-
+\newcommand{\muvar}{\mathrm{Var}} 
 \newcommand\SSS{\mathit{S}}
@@
 \newcommand\Sel{\Funcs_{\smash{\mathrm{sel}}}}
-%\newcommand\Plainfuncs{\Funcs_{\mathrm{other}}}
 
@@
 
-%%% For final version as well?
 \usepackage[
@@
 
-\include{defs}
-
 \hyphenation{data-type data-types co-data-type co-data-types isa-belle sledge-hammer
@@
 \begin{abstract}
-%Codatatypes naturally capture potentially infinite data structures and
-%processes.
 We present a decision procedure that combines reasoning about
@@
 
-%% The institutions above shouldn't count as footnotes
 \setcounter{footnote}{0}
@@
 properties that can be exploited in automated reasoners.
-%
-%However, because datatype values correspond to finite ground terms, they
-%are generally not adequate to represent infinite objects.
-%For example, the datatype of natural numbers
-%constructed by $\const{Z} : \ty{nat}$ and $\const{S} : \ty{nat} \to \ty{nat}$
-%only allow values of the form $\const{S}(\ldots(\const{S}(\const{Z}))\ldots)$.
-
-To represent infinite objects, % such as $\const{S}(\const{S}(\const{S}(\ldots))$,
-a natural choice is to turn to coalgebraic datatypes, or \emph{codatatypes},
+
+To represent infinite objects, a natural choice is to turn to coalgebraic datatypes, or \emph{codatatypes},
 the non-well-founded dual of algebraic \emph{datatypes.}
-%
 Despite their reputation for being esoteric, codatatypes have a
@@
 terms, and only those. As a simple example, the coalgebraic specification
-%
 \[\keyw{codatatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat})\]
-%
-%(using an ML-like syntax)
 {\looseness=-1
@@
 $\const{S}(\const{S}(\const{S}(\ldots)))$, which uniquely identifies
-%the value~
 $\infty$.
-Compared with the conventional %algebraic
-definition
-\begin{conf}$\vthinspace \keyw{datatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat}) \,\mid\, \const{Infty}$\vvthinspace, \end{conf}%
-\begin{rep}\[\vthinspace \keyw{datatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat}) \,\mid\, \const{Infty}\]\end{rep}%
-%with a dedicated constructor for representing infinity, the
-the codatatype avoids
+Compared with the conventional definition
+\begin{conf}$\vthinspace \keyw{datatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat}) \,\mid\, \const{Infty}$\vvthinspace, \end{conf}\begin{rep}\[\vthinspace \keyw{datatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat}) \,\mid\, \const{Infty}\]\end{rep}the codatatype avoids
 one case by unifying the infinite and finite nonzero cases.
@@
 have emerged in a few automatic theorem provers as well. The SMT-LIB
-%\cite{barrett-et-al-2010}
 format, implemented by most SMT
-%(satisfiability modulo theories)
 solvers, has been extended with a syntax for datatypes.
 In this \thewordpaper, we introduce a
-unified \hfill decision \hfill procedure \hfill for \hfill universal \hfill problems \hfill involving \hfill datatypes \hfill and \hfill codatatypes%
-\begin{conf}\goodbreak\noindent\end{conf}%  %%% TYPESETTING BREAK
-\begin{rep} \end{rep}%
-in combination (Section~\ref{sec:the-decision-procedure}).
-The procedure is described abstractly as a %tableau-like
-calculus and is composable via the Nelson--Oppen method \cite{nelson-oppen-1979}.
+unified \hfill decision \hfill procedure \hfill for \hfill universal \hfill problems \hfill involving \hfill datatypes \hfill and \hfill codatatypes\begin{conf}\goodbreak\noindent\end{conf}\begin{rep} \end{rep}in combination (Section~\ref{sec:the-decision-procedure}).
+The procedure is described abstractly as a calculus and is composable via the Nelson--Oppen method \cite{nelson-oppen-1979}.
 It generalizes the procedure by Barrett et al.\ \cite{barrett-et-al-2007},
 which covers only datatypes.
-%To our knowledge, our procedure is the first of its kind for the theory of
-%codatatypes.
-\begin{conf}%
-Detailed proofs %of completeness
-are included in a %technical
-report
+\begin{conf}Detailed proofs are included in a report
 \cite{our-report}.
-\end{conf}%
-
+\end{conf}
 }
@@
 differences.
-%
 First, \emph{codatatypes need not be well-founded.}
@@
 \cite{blanchette-et-al-2015-esop}\end{rep}.
-%
 Second, \emph{a uniqueness rule takes the place of the acyclicity rule of datatypes.}
 Cyclic constraints such as
-$x \teq \const{S}(x)$ %, where $\const{C}$ is a constructor,
-are unsatisfiable for datatypes,
+$x \teq \const{S}(x)$ are unsatisfiable for datatypes,
 thanks to an acyclicity rule,
@@
 These two rules are needed to ensure completeness
-(solution soundness) on universal problems. %and the absence of spurious models.
-They cannot be expressed as finite axiomatizations, so they naturally belong in
+(solution soundness) on universal problems. They cannot be expressed as finite axiomatizations, so they naturally belong in
 a decision procedure.
-%    * in particular, acyclicity and uniqueness are necessary for some proofs
-%%      (and cannot be axiomatized finitely)
-%    * and for model finding, without them we quickly get spurious models
-%    * explain how finite model finding works
-%(For the other (co)datatype properties---the injectivity, distinctness, and
-%exhaustiveness of constructors and the selector laws---it is
-%widely recognized that decision procedures can be more efficient than
-%axiomatizations.)
-%axiom.
-%
 Third, \emph{it must be possible to express cyclic }(\emph{regular}) \emph{values as closed terms and
 to enumerate them.} This is necessary both for finite model finding \cite{reynolds-et-al-2013}
-%modulo theories
 and for theory combinations. The $\mu$-binder notation associates a name with
@@
 
-%%% 1090909 is a prime number
-
 Our procedure is implemented in the SMT solver CVC4 \cite{barrett-et-al-2011} as a combination
@@
 of the CVC4~1.5 release.
-%
-An evaluation on %hand-crafted examples and on
-problems generated from Isabelle theories using the Sledgehammer tool \cite{blanchette-et-al-2013-smt}
+An evaluation on problems generated from Isabelle theories using the Sledgehammer tool \cite{blanchette-et-al-2013-smt}
 demonstrates the usefulness of the approach (Section~\ref{sec:experimental-results}).
 
-%  * useful both for proving and for model finding
-
-%\ref{sec:examples}
-
-%  * benchmarking is often an issue -- esp. codatatypes
-%
-%Polymorphic types, nested (co)recursion, and datatype--codatatype mixtures fall
-
-%  * setting: FOL
-%    * core procedure is restricted to ground
-%      * but theory solver cooperates
-
-%  * codatatypes were added later, motivated by the use of SMT solvers as
-%    backends to proof assistants (more specifically, CVC4 to Isabelle/HOL)
-
-%  * setting :
-%    * universal formulas
-%    * many-sorted logic
-%   * mutually (co)recursive types with constructors, selectors, and
-%      discriminators
-
-%  * codatatypes: from a theoretical and implementational point of view, like
-%      datatypes but:
-%    * infinite values (infinitely many nested constructors)
-%    * codatatypes are never empty (e.g. finite streams are rejected)
-
-%\[
-%      \keyw{codatatype}\; \,\ty{llist} \,=\, \const{LNil} \,\mid\, \const{LCons}(\ty{int},\: \ty{llist})
-%\]
-
-%  * consider a natural fragment---datatypes as supported in modern SMT solvers
-%    and the SMT-LIB 2 standard, and codatatypes as their duals
-%    * mutual recursion, but no polymorphism, nested recursion
-%  * integrated with Nelson--Oppen
-
-% * SMT-Lib
-
-%  * one implication is that if $m$ equals $ES(n)$ and $n$ equals $ES(m)$, necessarily
-%    $m$ and $n$ must be equal.
-
-%  * perhaps the most commonly used codatatype is that of lazy lists or sequences.
-%    using a syntax similar to Standard ML, Haskell, or SMT-LIB
-
-%  * less briefly: codatatypes and why they are useful
-%    * in Agda, Coq, Matita -- recently also in Isabelle/HOL \cite{nipkow-et-al-2002}
-%    * but not in SMT-LIB 2 \cite{barrett-et-al-2010}
-
-\begin{rep}\paragraph{Related Work.} Barrett\end{rep}%
-\begin{conf}Barrett\end{conf}
+\begin{rep}\paragraph{Related Work.} Barrett\end{rep}\begin{conf}Barrett\end{conf}
 et al.\ \cite{barrett-et-al-2007} provide a good account of
@@
 but also to the SMT solver
-Z3 \cite{de-moura-bjoerner-2008} \begin{rep}in unpublished work by Leonardo de Moura \end{rep}and \begin{rep}to \end{rep}a SPASS-like %prototypical
-superposition prover \begin{rep}by Daniel Wand\end{rep} \cite{wand-2014}.
+Z3 \cite{de-moura-bjoerner-2008} \begin{rep}in unpublished work by Leonardo de Moura \end{rep}and \begin{rep}to \end{rep}a SPASS-like superposition prover \begin{rep}by Daniel Wand\end{rep} \cite{wand-2014}.
 Closely related are the automatic structural
-% cruanes-201x,wand-weidenbach-201x,  -- both were rejected
 induction in both kinds of provers \cite{kersani-peltier-2013,reynolds-kuncak-2015},
@@
 implemented in Leon \cite{suter-et-al-2011} and RADA \cite{pham-et-al-2013}.
-%
 Datatypes are supported by the higher-order model finder Refute
@@
 
-%, notably in the context of constraint logic programming
-
-%  * Ro\c{s}u --- equality of streams is...
-%     http://dl.acm.org/citation.cfm?doid=1159803.1159827
-
-%     * point to Barrett et al. for SMT datatypes
-%       * about their own work, they say: "our focus is on generality and
-%         efficiency rather than immediacy of implementation"
-%       * (deal more directly with finite sorts than Barrett et al., Section 6.1)
-%     * additional ones since then (e.g. strings?)
-%     * anything about codatatypes?
-%       * proof assistants like Agda, Coq, etc. have them
-%       * Dafny, CoALP
-%       * also a lot of theoretical research, some of which is loosely connected,
-%         e.g. decision procedure for corecursive functions (Henning in Nijmegen)
-%
-%     * ODDITY: Oppen 1980: single-constructor, recursive -- infinite values?
-
 \paragraph{Conventions.}
@@
 distinguished type \ty{bool} interpreted as the set of truth
-values. %, and may contain other interpreted types (e.g., \ty{int}, \ty{real}).
-\begin{rep}
+values. \begin{rep}
 The metavariables $\delta,\:\varepsilon$ range over (co)datatypes,
@@
 which collapses to $\const{f} : \tau$ if $n = 0$\end{rep}.
-\begin{rep}%
-Functions invocations $\const{f}(t_1,\ldots,t_n)$
+\begin{rep}Functions invocations $\const{f}(t_1,\ldots,t_n)$
 apply the $n$-ary function symbol
@@
 
-%* although nothing prevents composing the decision procedure with theories
-%  providing polymorphic types (parametric sorts), such as for arrays (e.g., $\ty{array}(\alpha,\beta)$)
-
-\section{%The Theory of
-(Co)datatypes}
+\section{(Co)datatypes}
 \label{sec:co-datatypes}
-
-%    (terminology: freely-generated, inductive, algebraic, ..., sometimes
-%    with different meanings; we'll clarify below what we use)
 
@@
 $\Types = \Data \mathrel{\uplus} \Codata \mathrel{\uplus} \Nondata$, where $\Data$ are the
-\emph{datatypes}, $\Codata$ are the \emph{codatatypes}, and $\Nondata$ are the %remaining
-\emph{ordinary types} (which can be interpreted or not). The function symbols are
+\emph{datatypes}, $\Codata$ are the \emph{codatatypes}, and $\Nondata$ are the \emph{ordinary types} (which can be interpreted or not). The function symbols are
 partitioned into $\Funcs = \Ctr \mathrel{\uplus} \Sel$, where $\Ctr$ are the
@@
 without $\mu$-binders.
-%(Section~\ref{sec:implementation-as-a-theory-solver-in-cvc4}).
-%Exceptionally, it is convenient to use numeric constants ($0$, $1$, \ldots)\ in
-%examples.
-
-}
-
-%\paragraph{Specifications.}
+
+}
+
 In an SMT problem, the signature is typically given by specifying first the
-uninterpreted %ordinary
-types in any order, then the (co)datatypes with their constructors
+uninterpreted types in any order, then the (co)datatypes with their constructors
 and selectors in groups of mutually (co)recursive groups of (co)datatypes, and
 finally any other function symbols.
-%
 Each (co)datatype specification consists of $\ell$~mutually recursive types that are
@@
 (co)recursion,\end{rep}\begin{conf}Nested (co)recursion\end{conf} and datatype--codatatype mixtures
-fall outside this fragment%
-\dotReportFootnote{%
-In principle, rank-1 (top-level) polymorphism \cite{blanchette-paskevich-2013}
+fall outside this fragment\dotReportFootnote{In principle, rank-1 (top-level) polymorphism \cite{blanchette-paskevich-2013}
 should not raise any special difficulties. Nesting datatypes inside datatypes,
@@
 \begin{rep}We allow ourselves some notational parameterization
-% at the metalevel
 through subscripts---for example, $\ty{stream}_{\,\tau}$ denotes a
@@
 being introduced.
-%
 To every argument corresponds a selector. The names for the (co)data\-types, the
@@
 \begin{conf}fresh\end{conf}\begin{rep}distinct and different from
-existing names\end{rep}%
-\dotReportFootnote{It can be convenient to specify the same selector
+existing names\end{rep}\dotReportFootnote{It can be convenient to specify the same selector
 for several constructors associated with the same (co)data\-type,
-as long as the argument types coincide. % \cite[Section~3]{blanchette-et-al-2014-codata}.
-However, this is disallowed by CVC4 and Z3, so we do not consider it here.}
+as long as the argument types coincide. However, this is disallowed by CVC4 and Z3, so we do not consider it here.}
 Schematically:
-%
 \[
@@
    \smash{\vdots\,\,\,} \\[-1.5\jot]
-  \keyw{and}\; \,\delta_{\elllx} & =\, \smash{\const{C}_{\elll 1\!}(%\bigl[\const{s}_{\elll 1\!}^1{:}\bigr]\vthinspace \tau_{\elll 1\!}^1, \ldots, \bigl[\const{s}_{\elll 1\!}^{n_{\elll 1\!}}{:}\bigr]\vthinspace \tau_{\elll 1\!}^{n_{\elll 1\!}}
-  \ldots)} \mid \cdots \mid \smash{\const{C}_{\elll m_{\elllx}\!}(\ldots)}
+  \keyw{and}\; \,\delta_{\elllx} & =\, \smash{\const{C}_{\elll 1\!}(\ldots)} \mid \cdots \mid \smash{\const{C}_{\elll m_{\elllx}\!}(\ldots)}
 \end{aligned}
 \]
-%
 with
@@
 $\Sel^\delta$.
-%
 For types with several constructors, it is useful to provide discriminators
@@
 $t \teq \const{C}_{i\negvthinspace j}\bigl(\const{s}_{i\negvthinspace j}^1(t), \ldots, \const{s}_{i\negvthinspace j}^{\,n_{\smash{{i\negvthinspace j}}}}(t)\bigr)$.
-%This will simplify the presentation. % of the decision procedure.
-
-\begin{rep}%
-Here are a few examples of legal specifications of (co)datatype families:
+
+\begin{rep}Here are a few examples of legal specifications of (co)datatype families:
 \[\begin{aligned}[t]
-      \keyw{codatatype}\; \,\ty{llist}_{\,\tau} & \,=\, \const{LNil} \,\mid\, \const{LCons}(%\const{lhead}{:}\;
-      \tau,\: %\const{ltail}{:}\;
-      \ty{llist}_{\,\tau}) \\%[-.5\jot]
+      \keyw{codatatype}\; \,\ty{llist}_{\,\tau} & \,=\, \const{LNil} \,\mid\, \const{LCons}(\tau,\: \ty{llist}_{\,\tau}) \\%[-.5\jot]
       \keyw{datatype}\;\, \ty{tree}_{\,\tau} & \,=\, \const{Node}(\tau,\:\, \ty{forest}_{\,\tau}) \\[-\jot]
@@
 A type is (\emph{co})\emph{recursive} if it belongs to such a set of types.
-\begin{rep}%
-Types can be declared together in a mutual fashion even if they are not actually
+\begin{rep}Types can be declared together in a mutual fashion even if they are not actually
 (co)recursive. The semantic notion is more precise and is the one that interests
 us.
-\end{rep}%
-%
-Non(co)recursive type specifications such~as
-\begin{conf}%
-\vthinspace $\keyw{datatype}~\vthinspace\ty{option}_{\,\tau} \,=\, \const{None} \,\mid\, \const{Some}(\tau)$
-\end{conf}%
-\begin{rep}%
-\[\begin{aligned}[t]
+\end{rep}Non(co)recursive type specifications such~as
+\begin{conf}\vthinspace $\keyw{datatype}~\vthinspace\ty{option}_{\,\tau} \,=\, \const{None} \,\mid\, \const{Some}(\tau)$
+\end{conf}\begin{rep}\[\begin{aligned}[t]
       \keyw{datatype}~\vthinspace\ty{option}_{\,\tau\!} & \,=\, \const{None} \,\mid\, \const{Some}(\tau) \\[-.5\jot]
@@
 \]
-\end{rep}%
-are permitted.
+\end{rep}are permitted.
 \begin{rep}At the semantic level, it makes no difference whether
 such types are introduced as datatypes or as codatatypes.
-\end{rep}%
-%Without loss of generality, we consider that these types are datatypes
-%instead of codatatypes.
-
-%Given the specification
-%
-%\[\keyw{codatatype}\;\, \ty{stream} \,=\, \const{SCons}(\ty{int},\:\ty{stream})\]
-%
-%the infinite value $\const{SCons}(0, \const{SCons}(0, \ldots))$ witnesses
-%nonemptiness.
-
-One way to characterize datatypes %semantically
-is as the initial model of the
+\end{rep}
+
+One way to characterize datatypes is as the initial model of the
 selector--constructor equations \cite{barrett-et-al-2007}.
@@
 their selectors.
-%By uniformly focusing on the constructors, the
-%axiomatic approach emphasizes the commonality between datatypes and
-%codatatypes, while sacrificing a theoretically fruitful notion of duality.
 
@@
 $\jPrime$, and $k$ within bounds\end{rep}:
-% and to all possible splits of the
-%$n$-ary constructor $\const{C}_{i\negvthinspace j}$'s argument list into $\bar x,
-%y, \bar z$:
-%
 \[
@@
 \text{Distinctness:}\quad
-  & %\forall \bar x, \bar y.\;\,
-    \smash{\const{C}_{i\negvthinspace j}(\bar x) \tneq \const{C}_{i\negvthinspace \jPrime}(\bar y) \quad\text{if $j \not= \jPrime$}}
+  & \smash{\const{C}_{i\negvthinspace j}(\bar x) \tneq \const{C}_{i\negvthinspace \jPrime}(\bar y) \quad\text{if $j \not= \jPrime$}}
   \\[-0.5\jot]
 \text{Injectivity:}\quad
-  & %\forall x_1,\ldots,x_{n_{i\negvthinspace j}} y.\;\,
-    \smash{\const{C}_{i\negvthinspace j}(x_1,\ldots,x_{n_{i\negvthinspace j}}) \teq \const{C}_{i\negvthinspace j}(y_1,\ldots,y_{n_{i\negvthinspace j}}) \longrightarrow x_k \teq y_k}
+  & \smash{\const{C}_{i\negvthinspace j}(x_1,\ldots,x_{n_{i\negvthinspace j}}) \teq \const{C}_{i\negvthinspace j}(y_1,\ldots,y_{n_{i\negvthinspace j}}) \longrightarrow x_k \teq y_k}
   \\[-0.5\jot]
@@
 \]
-%
-\begin{rep}%
-Expressed in the algebraic jargon, exhaustiveness helps ensure that ``no
+\begin{rep}Expressed in the algebraic jargon, exhaustiveness helps ensure that ``no
 junk'' exists, whereas distinctness and injectivity guarantee that ``no
@@
 constructor is left completely unspecified.
-\end{rep}%
-%
-%  * how to deal with "wrong" selectors, e.g.
-%        hd(nil1) = hd(nil2)?
-%      * leave them unspecified; hence hd(nil1) = hd(nil2) in some models,
-%        and not in other models
-%
-Datatypes are additionally characterized by an induction principle\begin{rep}:
-%
+\end{rep}Datatypes are additionally characterized by an induction principle\begin{rep}:
 \[
@@
 witnesses $R_i$ that are required to be bisimulations:
-%
 \[
@@
 $\iPrime$ such that the formula is type-correct or $x \teq y$ otherwise.
-\end{rep}%
-This axiom encodes a form of extensionality:\ Two values that yield the same
+\end{rep}This axiom encodes a form of extensionality:\ Two values that yield the same
 observations must be equal, where the observations are made through
 selectors and discriminators.
-%
 \begin{rep}\par\end{rep}
-  %
   In addition, codatatypes are guaranteed to contain all values corresponding to
@@
 datatypes and codatatypes},
-which %in addition to $\Sig$
-defines a class of $\Sig$-interpretations $\JJJJ$,
-namely the ones that satisfy the %axioms --- existence of all infinite codatatype values cannot be axiomatized
-properties mentioned in this section, including (co)induction. The interpretations
+which defines a class of $\Sig$-interpretations $\JJJJ$,
+namely the ones that satisfy the properties mentioned in this section, including (co)induction. The interpretations
 in~$\JJJJ$ share the same interpretation for constructor terms and correctly
-applied selector terms (up to isomorphism) %renaming of domain elements
-but may differ on variables and wrongly applied selector terms.
+applied selector terms (up to isomorphism) but may differ on variables and wrongly applied selector terms.
 A formula is $\thD$-\emph{satisfiable} if there
 exists an interpretation in $\JJJJ$ that satisfies it.
-% otherwise, it is $\thD$-unsatisfiable. (I trust the readers' intelligence when
-% it comes to English prefixes like non-, in-, un-, and a-.)
-%
 For deciding universal formulas, induction can be replaced by the acyclicity
@@
 Dually, coinduction can be replaced by the uniqueness schema, which
-asserts that codatatype values are fully characterized by their %possibly
-%infinite
-expansion \cite[Theorem 8.1, 2$\,{\Leftrightarrow}\,$5]{rutten-2000}.
-
-%\begin{rep}
-% TODO: State more precisely and prove?
-%\end{rep}
-
-%    * :
-%        enough to consider acyclicity -- no way to specify infinite objects
-%        otherwise
-
-%    * codatatypes:
-%      * when looking at the universal theory, enough to
-%        consider uniqueness -- no way to express that certain infinite
-%        acyclic objects do not exist
+asserts that codatatype values are fully characterized by their expansion \cite[Theorem 8.1, 2$\,{\Leftrightarrow}\,$5]{rutten-2000}.
 
@@
 and
-\begin{conf}%
-\vthinspace$\keyw{codatatype}~\ty{b} = \const{B}(\ty{b},\: \ty{c},\: \ty{b},\: \ty{unit})
+\begin{conf}\vthinspace$\keyw{codatatype}~\ty{b} = \const{B}(\ty{b},\: \ty{c},\: \ty{b},\: \ty{unit})
 ~\keyw{and}~ \ty{c} = \const{C}(\ty{a},\: \ty{b},\: \ty{c})$,
-\end{conf}%
-\begin{rep}%
-\[\begin{aligned}[t]
+\end{conf}\begin{rep}\[\begin{aligned}[t]
       \keyw{codatatype}\;\, \ty{b} & \,=\, \const{B}(\ty{b},\: \ty{c},\: \ty{b},\: \ty{unit}) \\[-1\jot]
@@
 \]
-\end{rep}%
-where \ty{unit} is a datatype with the single constructor $\const{Unity} :
+\end{rep}where \ty{unit} is a datatype with the single constructor $\const{Unity} :
 \ty{unit}$. We call such types
 \emph{corecursive singletons}. For the decision procedure, it will be
-crucial to detect these. %, even if they rarely arise in practice.
-A type may also be a corecursive singleton only in some models. If the example
+crucial to detect these. A type may also be a corecursive singleton only in some models. If the example
 above is altered to make \ty{unit} an uninterpreted type, \ty{b} and \ty{c} will be
 singletons precisely when \ty{unit} is interpreted as a singleton.
-Fortunately, \begin{rep}given cardinalities for the ordinary types, \end{rep}%
-it is easy to characterize this degenerate case.
-
-}
-% In other words, all finite corecursive codatatypes are corecursive singletons.
-
-\begin{lemma}%[Corecursive Singletons]%
-\label{lem:corecursive-singletons}%
-\afterDot
+Fortunately, \begin{rep}given cardinalities for the ordinary types, \end{rep}it is easy to characterize this degenerate case.
+
+}
+
+\begin{lemma}\label{lem:corecursive-singletons}\afterDot
 Let $\delta$ be a corecursive codatatype. For any interpretation in $\JJJJ$,
@@
 
-%  * assume for simplicity no indirect recursion, but this does not radically
-%    change the argument
-%  * assume there are at least two values built with C.
-%    they must be different at some point, e.g.
-%    C(0, C(1, C(0, ...))) and C(0, C(1, C(1, ...)))
-%    can use that to create infinitely alternation patterns
-%  * leaves us with the case of a single constructor
-% * since we are looking at a specific model, we can assume all uninterpreted
-%   types are finite datatypes with nullary constructors corresponding to their
-%   elements
-% * if two
-% * either there is only one ``path'' from ctr to itself
-%   or at least two;
-
-\section{The Decision Procedure} % for (Co)datatypes}
-\label{sec:the-decision-procedure}
+\section{The Decision Procedure} \label{sec:the-decision-procedure}
 
 Given a fixed signature~$\Sig$,
-the decision procedure for the universal theory of (co)data\-types %$\thD$
-determines the
+the decision procedure for the universal theory of (co)data\-types determines the
 $\thD$-satisfiability of finite sets $\Ec$ of literals:\ equalities
@@
 The procedure is formulated as a tableau-like calculus.
-%Following the tradition of refutational theorem proving,
 Proving a universal quantifier-free conjecture is reduced to showing that
@@
 
-% whose derivation rules are applied until saturation or contradiction.
-
 To simplify the presentation, we make a few assumptions about
 $\Sig$.
-First, all codatatypes %in $\Codata$
-are corecursive. This is reasonable
+First, all codatatypes are corecursive. This is reasonable
 because noncorecursive codatatypes can be seen as nonrecursive
@@
 bit vectors), each interpreted type having finite cardinality~$n$
-can be viewed as a %n enumeration
-datatype with $n$~nullary constructors
+can be viewed as a datatype with $n$~nullary constructors
 \cite{barrett-et-al-2007}.
 
-%We will commonly denote tuples of terms $( t_1, \ldots, t_n )$ in bold font, as $\vec t$.
-%We assume that all $\Sig$-terms in $\Ec$ are \emph{normalized}, meaning that all subterms of the form $\const{s}^{\,k}_{\negvthinspace j}( \const{C}_{\negvthinspace j}(t_1,\ldots,t_n) )$
-%are simplified to $t_k$. All literals added to $\Ec$ are also implicitly
-%normalized in this manner.
-
-A derivation rule can be applied to $\Ec$ if %all
-the %specified
-preconditions are met.
+A derivation rule can be applied to $\Ec$ if the preconditions are met.
 The conclusion either specifies equalities to be added to $\Ec$
 or is $\bot$ (contradiction).
-% (in which case we call an application of it \emph{nonterminal})
-% (in which case we call an application of it \emph{terminal}).
-One rule has multiple conclusions, % separated by $\ror$,
-denoting branching.
-%
+One rule has multiple conclusions, denoting branching.
 An application of a rule is \emph{redundant} if one of its non-$\bot$
 conclusions leaves $\Ec$ unchanged.
-A \emph{derivation tree} is a %finite
-tree whose nodes are finite sets of
+A \emph{derivation tree} is a tree whose nodes are finite sets of
 equalities, such that child nodes are obtained by a nonredundant application of a
@@
 various terms occurring in $\Ec$.
-%Following the conventions from~\cite{},
-%The derivation rules are given in \relax{guarded assignment form},
-%where
-%
-The rules belonging to a phase have priority over those of %the
-subsequent
+The rules belonging to a phase have priority over those of subsequent
 phases. The rules are applied until the derivation tree is closed or all leaf nodes
@@
 
-%We assume that all $\Sig$-terms in $\Ec$ are \emph{normalized}, meaning that all subterms of the form $\const{s}^{\,k}_{\negvthinspace j}( \const{C}_{\negvthinspace j}(t_1,\ldots,t_n) )$
-%are simplified to $t_i$, and moreover assume that all %additional
-%literals added to $\Ec$ are normalized in this manner.
-%We present the calculus in three steps.
-%In the first step, we compute the bidirectional closure of $\Ec$;
-%in the second step, we make inferences based on acyclicity and uniqueness;
-%and in the third step, %(when necessary),
-%we branch on constructor types for various terms in $\tEc$.
-%Following the conventions from~\cite{},
-%the derivation rules of our calculus are given in \emph{guarded assignment form},
-%where a rule can be applied to $\Ec$ if it meets all of the specified preconditions for $\Ec$.
-
 \paragraph{Phase 1: Computing the Bidirectional Closure \rm(\figurename~\ref{fig:cc-rules}).}
 In conjunction with \rn{Refl}, \rn{Sym}, and \rn{Trans}, the \rn{Cong} rule computes the congruence (upward) closure,
-whereas the \rn{Inject} and \rn{Clash} rules %together
-compute the unification (downward) closure.
-For unification, %additional
-equalities are inferred based on the injectivity of constructors by \rn{Inject},
+whereas the \rn{Inject} and \rn{Clash} rules compute the unification (downward) closure.
+For unification, equalities are inferred based on the injectivity of constructors by \rn{Inject},
 and failures to unify equated terms are recognized by \rn{Clash}.
@@
 \paragraph{Phase 2: Applying Acyclicity and Uniqueness \rm(\figurename~\ref{fig:ab-rules}).}
-%For presentation of the rules in this phase, we rely on a representation of terms in the $\mu$-notation.
-%The premises of the rules that make up the second phase refer to
 We describe the rules in this phase in terms of a mapping $\Val$
 that associates with each equivalence class a $\mu$-term as its representative.
-%\rem{Not explicitly constructed, add here?}
 
@@
 $\FV( t )$ denotes the set of free variables occurring in the $\mu$-term~$t$.
-A $\mu$-term is \emph{closed} if %and only if --- by convention, "and only if" is not necessary for *definitions* (but add it back if you disagree)
-it contains no free variables. It is \emph{cyclic} if %and only if --- ditto
-it contains a bound variable. The $\alpha$\emph{-equiv\-alence} relation $t \aequiv u$
+A $\mu$-term is \emph{closed} if it contains no free variables. It is \emph{cyclic} if it contains a bound variable. The $\alpha$\emph{-equiv\-alence} relation $t \aequiv u$
 indicates that the $\mu$-terms $t$ and $u$
 are syntactically equivalent for some capture-avoiding renaming of $\mu$-bound variables\begin{rep}---e.g.,
-%$\MU x.\; \const{C}( x ) \aequiv \MU y.\; \const{C}( y )$ and
 $\MU x.\; \const{D}( y, x ) \aequiv \MU z.\; \const{D}( y, z )$,
@@
 $\MU x.\; \const{S}( x ) \not\aequiv \MU y.\; \const{S}( \const{S}( y ) )$.
-%In the \rn{Acyclic} rule,
-
-%Recall that $\mu$-bindings can be used for representing possibly cyclic terms and values.
-%(Section~\ref{sec:introduction}).
-% for example, $\MU x.\; \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
-%either in the case when $\tau$ is a codatatype or datatype type,
-%$\tau \notin \Codata$.
-
-%%% But there are no constants of type $\tau \in \Nondata$ -- at most variables!
-%
-%For uniformity, we also consider $%\MU x.\;
-%\const{c}$ to be a $\mu$-term if $\const{c}$ is a constant of type $\tau \in \Nondata$.
-%
-%%% I think this should be clear enough from the "Conventions" paragraph at the
-%%% end of the introduction.
-%%
-% which we define recursively:
-%if $\const{C} : \vec \tau \rightarrow \upsilon$
-%and $\vec v$ are well-typed $\mu$-terms of type $\vec \tau$ under the assumption that $x$ has type $\upsilon$,
-%then $\MU x.\; \const{C}( \vec v )$ is a well-typed term of type $\upsilon$.
-%
-%, or informally are \emph{observationally equivalent}.
-%For convenience, we use $\mu$-terms to refer to (classes of) values for both codatatype and datatype terms,
-%where in the latter case, a $\mu$-term with a bound variable denotes an infeasible term.
-%where the latter case adds the restriction on $\mu$-terms $t$ that no subterms of $t$ contain are bound variables.
 
@@
 that $t$ and other members of $t$'s equivalence class can take in models of $\Ec$.
-When $\tau$ is a datatype, % \in \Data$,
-a cyclic $\mu$-term describes an infeasible class of values.
+When $\tau$ is a datatype, a cyclic $\mu$-term describes an infeasible class of values.
 
 The mapping $\Val$ is defined as follows.
-%
 With each equivalence class $\ec{u\typ{\tau}}$, we associate a fresh variable $\Varec{u}\typ{\tau}$
-%of the same type as $u$
-%not occurring in $\tEc$,
 and set $\Val\ec{u} := {\Varec{u}}$, that is to say there are initially no
@@
 \)
-%\rn{Unfold}
 }
@@
 \end{tabular}
-\vspace*{-2pt} %% TYPESETTING HACK
-\caption{\,Derivation rules for bidirectional closure%.
-}
+\vspace*{-2pt} \caption{\,Derivation rules for bidirectional closure}
 \label{fig:cc-rules}
@@
 \begin{figure}[t!]
-\vspace*{+6pt} %% TYPESETTING HACK
-\normalsize
+\vspace*{+6pt} \normalsize
 \centering
@@
 \inferrule{
-%  t\typ{\delta}
-%  \betweenantes
 \delta \in \Data
@@
 \rn{Acyclic}
-\kern1.5em\( %%% TYPESETTING: should be \quad
-\inferrule{
-% t, u\typ{\delta}
-% \betweenantes
+\kern1.5em\( \inferrule{
 \delta \in \Codata
@@
 \end{tabular}
-\vspace*{-2pt} %% TYPESETTING HACK
-\caption{\,Derivation rules for acyclicity and uniqueness%.
-}
+\vspace*{-2pt} \caption{\,Derivation rules for acyclicity and uniqueness}
 \label{fig:ab-rules}
@@
   \text{ or }
-  \bigl( \delta \in \Data \text{ and } %\left|
-  \delta
-  %\right|
+  \bigl( \delta \in \Data \text{ and } \delta
   \text{ is finite} \bigr)
 }{
-  \Ec := \Ec,\: \DISC{1} %\ROR
-  \ROR \cdots \ROR \Ec := \Ec,\: \DISC{m}
+  \Ec := \Ec,\: \DISC{1} \ROR \cdots \ROR \Ec := \Ec,\: \DISC{m}
 }
@@
 \end{tabular}
-\vspace*{-2pt} %% TYPESETTING HACK
-\caption{\,Derivation rules for branching%.
-%All pairs of terms whose type has cardinality 1 are entailed to be equal (the degenerate case).
-%Constructors must be assigned for all terms $t$ if has a selector is applied to it, or if $t$ has finite datatype type.
-}
+\vspace*{-2pt} \caption{\,Derivation rules for branching}
 \label{fig:split-rule}
@@
 trivial entries such as $\ec{w} \mapsto {\Varec{w}}$.
-%
 \begin{enumerate}
-%\item \noindent\rlap{Initially:}\phantom{Unfold $\Varec{x}$:\enskip}$\{  \}$
 \item \noindent\rlap{Unfold $\Varec{x}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\Val = \{\vthinspace  \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) \vthinspace\}$
@@
 \end{enumerate}
-%
 The resulting $\Val$ indicates that the values for $x$ and $y$ in models of $\Ec$
@@
 and $\const{C}( {\Varec{z}}, \const{C}( {\Varec{w}}, \const{C}( {\Varec{z}}, \const{C}( {\Varec{w}}, \ldots ))))$,
-respectively. %, for some values of ${\Varec{w}}$ and ${\Varec{z}}$.
-\xend
+respectively. \xend
 \end{examplex}
@@
 whose values $\Val \ec{t}$, $\Val \ec{u}$ are $\alpha$-equivalent,
-%for a capture-avoiding renaming of $\mu$-bound variables,
 then $t \teq u$. Comparison for $\alpha$-equivalence may seem too restrictive, since
@@
 
-%% Minor terminology point:
-%% A function symbol can have a "type signature" (but we're trying to avoid that terminology);
-%% unless it's a constant, it doesn't have a (first-order) type. So I'm rephrasing.
-
 \begin{examplex}
 Let $\Ec = \{ x \teq \const{S}(x),\; y \teq \const{S}(\const{S}(y)) \}$.
-After phase~1, the equivalence classes %of $\Ec$
-are
+After phase~1, the equivalence classes are
 $\{ x,\, \const{S}(x) \}$,
@@
 \Val \ec{x} & \,=\, \MU {\Varec{x}}.\; \const{S}(\Varec{x}) &
-\kern0.25em %% TYPESETTING
-\Val \ec{y} & \,=\,
+\kern0.25em \Val \ec{y} & \,=\,
 \MU {\Varec{y}}.\; \const{S}(\smash{\MU {\VAREC{\const{S}(y)}}}.\; \const{S}( {\Varec{y}} ) )
-\kern0.25em %% TYPESETTING
- &
+\kern0.25em &
 \Val \ec{\const{S}(y)} & =
@@
 Since $\Val \ec{y} \aequiv \Val \ec{\const{S}(y)}$,
-%for the renaming
-%$\{ {\Varec{y}} \mapsto {\VAREC{\const{S}(y)}}, {\VAREC{\const{S}(y)}} \mapsto {\Varec{y}} \}$,
 the \rn{Unique} rule applies to derive $y \teq \const{S}(y)$.
-%Intuitively, this equality holds, since the values of $y$ and $\const{S}(y)$
-%are equivalent.
-%
-At this point, phase~1 is activated again, % to XXX the equivalence classes,
-yielding
+At this point, phase~1 is activated again, yielding
 $\{ x,\, \const{S}(x) \}$ and
@@
 The mapping $\Val$ is updated accordingly:
-\begin{conf}%
-$\Val \ec{y} \,=\, \MU {\Varec{y}}.\; \const{S}({\Varec{y}})$.
-\end{conf}%
-\begin{rep}
+\begin{conf}$\Val \ec{y} \,=\, \MU {\Varec{y}}.\; \const{S}({\Varec{y}})$.
+\end{conf}\begin{rep}
 \begin{align*}
@@
 Since $\Val \ec{x} \aequiv \Val \ec{y}$,
-%for the renaming
-%$\{ {\Varec{y}} \mapsto {\VAREC{\const{S}(y)}}, {\VAREC{\const{S}(y)}} \mapsto {\Varec{y}} \}$,
 \rn{Unique} can finally derive $x \teq y$.
@@
 Another rule, \rn{Single}, focuses on the degenerate case where two
-terms %from $\tEc$
-have the same corecursive singleton type
-%(Section~\ref{sec:co-datatypes})
+terms have the same corecursive singleton type
 and are therefore equal. Both
-\rn{Split}'s finiteness assumption %on the datatype
-and \rn{Single}'s singleton
-constraint %on the codatatype
-can be evaluated statically
+\rn{Split}'s finiteness assumption and \rn{Single}'s singleton
+constraint can be evaluated statically
 based on a recursive computation of the cardinalities of the
 constructors' argument types.
-%This step relies on the assumption that uninterpreted types are infinite.
 
@@
 
-\begin{theorem}[Termination]%
-\label{thm:t}%
-\afterDot
+\begin{theorem}[Termination]\label{thm:t}\afterDot
 All derivation trees are finite.
@@
 $\smash{\SSSS {i+1}_t} = \smash{\SSSS i_t \mathrel\cup \{ \const s( u ) \mid u\typ{\delta} \in \SSSS i_t{,}}~\allowbreak\smash{ \delta \in \Data{,}~ \left|\delta\right| \text{ is finite}{,}~ \const{s} \in \Sel^{\delta}}  \}$,
-\end{conf}%
-\begin{rep}
+\end{conf}\begin{rep}
 \begin{align*}
@@
 \end{align*}
-\end{rep}%
-%for $i \geq 0$, let ;
-and let $\SSSS \infty_t$ be the limit of this sequence.
+\end{rep}and let $\SSSS \infty_t$ be the limit of this sequence.
 This is a finite set for each $t$, 
 because all chains of selectors applied to $t$ are finite.
-%because in the absence of (co)recursion
-%one eventually reaches a point where no selector can be applied.
-%%% TODO: Clarify above step in report.
 Let $\SSSS \infty$ be the union of all sets $\SSSS \infty_t$ where $t \in D$,
 and let $\tcEc$ be the set of subterms of
-%
 $\Ec \mathrel\cup \{ \smash{\const C_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\smash{\,n_{\negvthinspace j}}}_{\negvthinspace j}( t ) \bigr)} \mid \smash{t\typ{\delta} \in \SSS \mathrel\cup \SSSS \infty{,}~ \const C_{\negvthinspace j} \in \smash{\Ctr^\delta}} \}$.
-%
 In a derivation tree with root node~$\Ec$,
@@
 
-\begin{theorem}[Refutation Soundness]%
-\label{thm:rs}%
-\afterDot
+\begin{theorem}[Refutation Soundness]\label{thm:rs}\afterDot
 If there exists a closed derivation tree with root node\/ $\Ec$, then\/ $\Ec$ is $\thD$-unsatisfiable.
@@
 $\expand{t}{u}$, as returned by the function
-%where $x \leadsto \MU x.\; \const{C}( \vec u ) \in t$ indicates that $\MU x.\;
-%\const{C}( \vec u )$ is the subterm of $t$ that binds this occurrence of
-%variable $x$:
 \[\begin{array}{@{}r@{}c@{}l@{}}
@@
  \MU y.\: \const{D}(\MU x.\: \const{C}(y))$.
-%
-%\rem{TODO: add small examples}
-%\rem{TODO: explain why terminates}
 
@@
 and $\MU y.\; \const{C}( u_1, \ldots, u_n )$,
-and $\expand{t}{ t_k } \aequiv \expand{t}{ u_k }$ for all $k$. % \in [n]
-The $\mu$-term $t$ is \emph{normal} if it does not contain %any
-self-similar subterms
+and $\expand{t}{ t_k } \aequiv \expand{t}{ u_k }$ for all $k$. The $\mu$-term $t$ is \emph{normal} if it does not contain self-similar subterms
 and all of its proper subterms are also normal.
@@
 noting that
-%
 $\expand{u}{\MU y.\; \const{C}( x )}
@@
 is $\alpha$-equivalent to $\expand{u}{x} = u$.
-% for the renaming $\{ x \mapsto y, y \mapsto x \}$.
 
@@
 by replacing all of the self-similar subterms of $t$ with $x$
-%and by replacing all of the abnormal subterms of $t$ with their corresponding normal forms.
 and by recursively normalizing the other subterms.
@@
 \end{lemma}
-% TODO:
-%\begin{proof}
-%\qed
-%\end{proof}
 
@@
 \end{lemma}
-% TODO:
-%\begin{proof}
-%\qed
-%\end{proof}
-
-%\begin{lemma}
-%\label{lem:mu-cong}
-%If $t_u = \MU x.\; \const{C}( \vec u )$ and $t_w = \MU y.\; \const{C}( \vec w )$ are in normal form,
-%then $\expand{t_u}{ \vec u } \vsim \expand{t_w}{ \vec w }$ if and only if $t_u \vsim t_w$.
-%\end{lemma}
-%\begin{proof}
-%\rem{TODO}
-%\qed
-%\end{proof}
+
 \end{rep}
@@
 then $\interp{\J}{\const{f}}$ is a total function from $\interp{\J}{\tau_1} \times \cdots \times \interp{\J}{\tau_n}$ to $\interp{\J}{\tau}$.
-%To simplify the presentation,
 All types are interpreted as sets of $\mu$-terms, but
 only values of types in $\Codata$ may contain cycles.
-%Since all values of types in $\Data$ and $\Nondata$ are acyclic,
-%a typical representation for these values can be obtained by
-%dropping each of its $\mu$-bindings.
-%For example, the $\mu$-term $\MU x.\; \const{Z}$ representing the integer constant zero
-%is simply $\const{Z}$ after dropping the $\mu$-binding of $x$.
 
 \begin{definition}[Normal Interpretation]
-\afterDot%
-\label{def:norm-interpretation}%
-\rm
+\afterDot\label{def:norm-interpretation}\rm
 An interpretation $\J$ is \emph{normal} if these conditions are met:
@@
 For each type $\tau$,
-%%% sets *contain* elements, they *include* other sets
 $\interp{\J}{\tau}$ includes a maximal set of closed normal $\mu$-terms of that type that are
@@
 \item
-For each selector term $\const s^k_{\negvthinspace j}( t )$ of type $\tau$, % for some $\vec u$,
-if $\interp{\J}{t}$ is $\MU x.\; \const{C}_{\negvthinspace j}( \vec u )$,
+For each selector term $\const s^k_{\negvthinspace j}( t )$ of type $\tau$, if $\interp{\J}{t}$ is $\MU x.\; \const{C}_{\negvthinspace j}( \vec u )$,
 then $\J\negvvthinspace( \const s^k_{\negvthinspace j} )\bigl( \interp{\J}{t} \bigr)$ is the value
@@
 formulas it is trivial to extend any normal interpretation with extra
-domain elements to obtain a genuine model if desired. %Hence, we focus on normal interpretations.
-
+domain elements to obtain a genuine model if desired. 
 \pagebreak[2]
@@
 We construct $\ValC$ by initially setting $\ValC := \nf{\Val}$
-and by exhaustively applying the following rule:%
-\[
+and by exhaustively applying the following rule:\[
 \hbox{\(
@@
 \)
-%\rn{Assign}
 }
 \]
-%
 Given an unassigned variable in $\ValC$, this rule assigns it a fresh
@@
 terms that could emerge as a result of the update.
-%\rem{TODO: What guarantees that such a value exists? (Answer: infinitely many values
-%of codatatypes, of which only finitely many are in
-%$\BAD_{\smash{\J}}^{\Varec{x}}(\ValC)$?)}
 Since this update removes one
@@
 
-%The last precondition is admittedly rather technical.
 To ensure disequality literals are satisfied by an interpretation based on $\ValC$,
@@
 the last precondition in the above rule ensures that it is maintained.
-%More precisely,
 The set $\BAD_{\smash{\J}}^{\Varec{x}}(\ValC)$ is an overapproximation of the values
 that, when assigned to ${\Varec{x}}$, will cause values in the range of $\ValC$ to become $\alpha$-equivalent.
-%This precondition is not strictly necessary in practice, 
-%as one could compute $\nf{\ValC [ {\Varec{x}} \mapsto \MU {\Varec{x}}.\; t ]}$ optimistically and check
-%injectivity afterward, repeating until a suitable value is found.
 For infinite codatatypes, it is always possible to find fresh values $v$
 because $\BAD_{\smash{\J}}^{\Varec{x}}(\ValC)$ is a finite set.
-%Injectivity is important to satisfy disequality literals.
-%For infinite codatatypes, it is always possible to find suitable values $v$.
-%For example,
-%if the longest cycle in $\BAD_{\smash{\J}}^{\Varec{x}}(\ValC)$ is of length $n$, a fresh value
-%can be constructed with a cycle of length $n + 1$.
 
@@
 the mapping $\Val$ is as follows:
-%\[\begin{array}{r@{}c@{}l@{\;}l@{\qquad}r@{}c@{}r@{\;}l@{}l@{\;}l}
-%\[\begin{array}{r@{}c@{}l@{\;}l@{\qquad}r@{}c@{}r@{\;}l@{}l@{\;}l@{\qquad}r@{}c@{}r@{\;}l}
 \[\begin{array}{r@{}c@{}r@{}l@{\qquad}r@{}c@{}r@{}l@{\qquad}r@{}c@{}l}
@@
 \end{array}\]
-%
 To construct a completion $\ValC$, we must choose values for ${\Varec{y}}$ and ${\Varec{z}}$,
@@
 Modulo $\alpha$-equivalence, $\smash{\BAD_{\smash{\J}}^{{\Varec{z}}}( \Val )} =
-%\begin{array}{c}
 \{
@@
 \}
-%\end{array}
 $.
@@
 However, this contradicts the disequality $z \not\teq v$ in~$\Ec$.
-On the other hand, if the value assigned to ${\Varec{z}}$ is fresh, %not $\alpha$-equivalent to any term in $\BAD_{\smash{\J}}^{{\Varec{z}}}( \Val )$,
-the values in the range of $\Val$ remain $\alpha$-disequivalent.
+On the other hand, if the value assigned to ${\Varec{z}}$ is fresh, the values in the range of $\Val$ remain $\alpha$-disequivalent.
 We can assign a value such as
@@
 \pagebreak[2]
-
-%%% I think it's enough to point this out clearly in the introduction.
-%%% At least, that always worked fine for my other papers. :)
-%
-%\begin{conf}
-%The proof of this lemma can be found in the extended version of this paper.
-%\end{conf}
 
@@
 the cardinality of $\tau$ must be one.
-Since \rn{Single} does not apply, % to $\Ec$,
-there is only one equivalence class of type $\tau$ in $\Ec$,
+Since \rn{Single} does not apply, there is only one equivalence class of type $\tau$ in $\Ec$,
 and thus there are no terms in $\BAD_{\smash{\J}}^{\Varec{y}}(\ValC)$ of type $\tau$.
-This is a contradiction, since %our %model
-completion can assign the value in the domain of $\tau$ to ${\Varec{y}}$.
+This is a contradiction, since completion can assign the value in the domain of $\tau$ to ${\Varec{y}}$.
 Now, consider the case when $\tau$ is infinite.
@@
 since there are only a finite number of closed terms in $\BAD_{\smash{\J}}^{\Varec{y}}(\ValC)$,
-and thus %our %model
-completion can assign a value not occurring in $\BAD_{\smash{\J}}^{\Varec{y}}(\ValC)$ to ${\Varec{y}}$.
+and thus completion can assign a value not occurring in $\BAD_{\smash{\J}}^{\Varec{y}}(\ValC)$ to ${\Varec{y}}$.
 By construction, $\ValC \ec{x}$ is normal.
@@
 each pair of constructor terms $\const{C}_{\negvthinspace j}( \vec t \,)$ and $\const{C}_{\negvthinspace\jPrime}( \vec u )$ in $\ec{z}$
-are such that $j = \jPrime$ since \rn{Clash} does not apply, % to $\Ec$,
-and are such that $\ec{ \vec t } = \ec{ \vec u }$ since \rn{Inject} does not apply. % to $\Ec$.
-Thus,
+are such that $j = \jPrime$ since \rn{Clash} does not apply, and are such that $\ec{ \vec t } = \ec{ \vec u }$ since \rn{Inject} does not apply. Thus,
 $\Val$ was constructed by applying a sequence of substitutions
@@
 the expansion of $\expand{\Val \ec{y}}{ t }$ as a term
-constructed by a sequence of substitutions of the form mentioned above to ${\Varec{z}}$. %\rem{TODO: more detail}
-Thus, we have
+constructed by a sequence of substitutions of the form mentioned above to ${\Varec{z}}$. Thus, we have
 $\expand{\Val \ec{y}}{ t } = \expand{\Val \ec{x}}{ \Val \ec{x} } = \Val \ec{x}$, and thus (2) holds for $\Val$.
@@
 Due to the construction of $\Val$,
-we know $\ec{x}$ contains a constructor $\const{C}( z_1, \ldots, z_n )$ and $\muvar( t_i ) = z_i$ for some $i$ % \in [n]$.
-Since $\Val \ec{x}$ is a minimal, it
+we know $\ec{x}$ contains a constructor $\const{C}( z_1, \ldots, z_n )$ and $\muvar( t_i ) = z_i$ for some $i$ Since $\Val \ec{x}$ is a minimal, it
 contains a subterm of the form $\MU {\Varec{y}}.\; \const{C}( u_1, \ldots, u_n )$
-where $\expand{\Val \ec{x}}{ t_i } \aequiv \expand{\Val \ec{x} }{ u_i }$ for some $i$. % \in [n]$.
-Due to the construction of $\Val$, $\ec{y}$ contains a constructor $\const{C}( w_1, \ldots, w_n )$ and $\muvar( u_i ) = w_i$ for some $i$. % \in [n]$.
-Since \rn{Cong} does not apply, % to $\Ec$
-we have $\ec{w_{\negvthinspace j}}$ and $\ec{z_{\negvthinspace j}}$ are distinct for some $j$.
+where $\expand{\Val \ec{x}}{ t_i } \aequiv \expand{\Val \ec{x} }{ u_i }$ for some $i$.
+Due to the construction of $\Val$, $\ec{y}$ contains a constructor $\const{C}( w_1, \ldots, w_n )$ and $\muvar( u_i ) = w_i$ for some $i$.
+Since \rn{Cong} does not apply, we have $\ec{w_{\negvthinspace j}}$ and $\ec{z_{\negvthinspace j}}$ are distinct for some $j$.
 By (2), $\expand{\Val \ec{x}}{ t_{\negvthinspace j} } = \Val \ec{z_{\negvthinspace j}}$
@@
 and $t_1 \not\aequiv t_2$.
-%We first show that $\nf{\Val_1 \sigma} \ec{y}$ (and for the same reasons, $\nf{\Val_1 \sigma} \ec{z}$)
-%also contains position $p$.
-%If this were not the case,
-%then without loss of generality we may assume
-%$\Val_1 \sigma \ec{y}$ contains a self-similar subterm
-%$\MU {\Varec{w}}.\; u$ that strictly contains $t_1 \sigma$.
 If $t_1$ (resp. $t_2$) is a free variable that is not ${\Varec{x}}$,
@@
 and $t_2$ is of the form $\MU w_2.\; \const{D}( \vec u )$,
-then the expansion %\rem{TODO: define}
-of $\nf{\Val_1 \sigma} \ec{y}$ and $\nf{\Val_1 \sigma} \ec{z}$
+then the expansion of $\nf{\Val_1 \sigma} \ec{y}$ and $\nf{\Val_1 \sigma} \ec{z}$
 are $\alpha$-disequivalent at position $p$.
@@
 which implies that the expansion of $\nf{\Val_1 \sigma} \ec{y}$ and $\nf{\Val_1 \sigma} \ec{z}$
-are $\alpha$-disequivalent at position $p$. %\rem{TODO: more detail}
-
+are $\alpha$-disequivalent at position $p$. 
 Thus, by induction on the number of applications of the above rule used to obtain $\ValC$,
@@
 interpretation of a subterm $t$ where $\muvar(t) = {\Varec{x}}$ is equal to
-$\ValC \ec{x}$, independently of the context. % in which it is interpreted.
-Property~(3) states that $\ValC$ is injective (modulo
-$\alpha$-equivalence), which ensures %in the following
-that distinct values are
+$\ValC \ec{x}$, independently of the context. Property~(3) states that $\ValC$ is injective (modulo
+$\alpha$-equivalence), which ensures that distinct values are
 assigned to distinct equivalence classes.
 
-\begin{theorem}[Solution Soundness]%
-\label{thm:ss}%
-\afterDot
+\begin{theorem}[Solution Soundness]\label{thm:ss}\afterDot
 If there exists a derivation tree with root node $\Ec$ containing a saturated node, then $\Ec$ is $\thD$-satisfiable.
@@
 Let $\Fc$ be a saturated node in a derivation tree with root node $\Ec$.
-%We will construct a model $\J$ for a set of equalities $\Fc$ that is equivalent to $\Fc_0$, and where
-%all equivalence classes of $\Fc$ contain at least one variable, and all selectors in $\Fc$ are applied to variables only.
-%The former comes with no loss of generality since new equalities of the form $y \teq t$ for fresh variable $y$ can be added to $\Fc$ without affecting its satisfiability,
-%The latter also comes with no loss of generality since nested applications of selectors in $\Fc_0$ can be replaced by fresh variables while adding equalities to $\Fc\!$.
 We consider a normal interpretation $\J$
@@
 and that interprets the variables of $\Fc$ based on the completion $\ValC$.
-% of %the mapping $\Val$. % from phase~2.
 For the variables, let $\interp{\J}{x\typ{\tau}}$ be the value in $\J(\tau)$ that is $\alpha$-equivalent with $\ValC \ec{x}$ for each variable $x \in \Terms(\Fc)$,
@@
 
-We first show that $\J$ satisfies all equalities $t_1 \teq t_2 \in \Fc\!$. % where $t_1$ and $t_2$ have type $\tau$.
-To achieve this, we show
+We first show that $\J$ satisfies all equalities $t_1 \teq t_2 \in \Fc\!$. To achieve this, we show
 by structural induction on $t\typ\tau$
@@
 Call this term $t'$.
-%By Lemma~\ref{lem:interpretation-completion}(2) and due to our construction of $\ValC$, we have $\ValC \ec{t}$ is a term
-%of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$
-%where $\expand{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$ for $i \in [n]$.
 Since \rn{Inject} and \rn{Clash} do not apply to $\Fc\!$,
@@
 For each $i$, let $u_\iPrime$ be the $i$\vvthinspace th argument of $t'$.
-\begin{rep}By Lemma~\ref{lem:mu-norm-arg},\end{rep}%
-\begin{conf}Clearly,\end{conf} %% TODO: Fix! ;)
-$\expand{t'}{u_\iPrime} \vsim \ValC \ec{u_i}$.
-Thus $\expand{t'}{u_\iPrime} \vsim \expand{\ValC \ec{t}}{ w_i }$. % for each $i \in [n]$.
-%By Lemma~\ref{lem:mu-cong}, we have $t' \vsim t_w$.
-Thus, $\interp{\J}{t} \vsim t' \vsim \ValC \ec{t}$, and
+\begin{rep}By Lemma~\ref{lem:mu-norm-arg},\end{rep}\begin{conf}Clearly,\end{conf} $\expand{t'}{u_\iPrime} \vsim \ValC \ec{u_i}$.
+Thus $\expand{t'}{u_\iPrime} \vsim \expand{\ValC \ec{t}}{ w_i }$. Thus, $\interp{\J}{t} \vsim t' \vsim \ValC \ec{t}$, and
 we have $\interp{\J}{t} \vsim \ValC \ec{t}$.
 
-If $t$ is a selector term %of the form
-$\const s^k_{\negvthinspace j}( u )$,
+If $t$ is a selector term $\const s^k_{\negvthinspace j}( u )$,
 since \rn{Split} does not apply to $\Fc\!$,
 $\ec{u}$ must contain a term of the form $\smash{\const C_{\negvthinspace\jPrime}\bigl( \const s^1_{\negvthinspace\jPrime}( u ), \ldots, \const s^{n}_{\negvthinspace\jPrime}( u ) \bigr)}$ for some $\jPrime$.
-%By Lemma~\ref{lem:interpretation-completion}(2) and due to our construction of $\ValC$, we have $\ValC \ec{u}$ is of the form
-%$\MU {\Varec{u}}.\; \const C_{\negvthinspace\jPrime}( w_1, \ldots, w_n )$,
-%where $\expand{\ValC \ec{u} }{ w_i } = \ValC \ec{ \const s^i_{\negvthinspace\jPrime}( u )}$ for $i \in [n]$.
-Since \rn{Inject} and \rn{Clash} are not applicable, % to $\Fc\!$,
-by construction $\ValC\ec{u}$ must be of the form
+Since \rn{Inject} and \rn{Clash} are not applicable, by construction $\ValC\ec{u}$ must be of the form
 $\MU {\Varec{u}}.\; \const C_{\negvthinspace\jPrime}( w_1, \ldots, w_n )$,
@@
 and since $\Ec \subseteq \Fc\!$,
-% is a superset of $\Ec$,
 we conclude that $\J \vDash \Ec$.
@@
 the optimizations.
-
-%%% @ANDY: I'm trying without \paragraph's, to save space but also to make the
-%%% section look more like a whole than a random list of stuff. Revert if you
-%%% disagree.
 
@@
 partial assignment as a set $M$ of true literals.
-% where $a \in M$ for all atoms $a$ that
-%it assigned the value $\top$ and $\neg\: a \in M$ for all atoms that it assigns
-%the value $\bot$.
-%at a high level,
 By a variant of the Nelson--Oppen method \cite{jovanovic2011sharing,nelson-oppen-1979},
-each $T_i$-solver %for $i = 1, \ldots, l$  n?
-takes as input the union $M_i$
+each $T_i$-solver takes as input the union $M_i$
 of (1)~the purified form of $T_i$-literals occurring in $M$, where fresh
@@
 $\Ec$ are of the form $\Val \ec{x}$. For example, if $x \teq \const{C}( x ) \in
-M_i$, then $x \teq \const{C}( x )$ is an %(possible)
-explanation for $\Val \ec{x} = \MU {\Varec{x}}.\; \const{C}( {\Varec{x}} )$.
+M_i$, then $x \teq \const{C}( x )$ is an explanation for $\Val \ec{x} = \MU {\Varec{x}}.\; \const{C}( {\Varec{x}} )$.
 The rules of the calculus are implemented as follows. For all rules with
 conclusion $\bot$, we report the union of the explanations for all premises is
-$\thD$-unsatisfiable. For \rn{Split}, we add the exhaustiveness clause %of the form
-$\DISC{1} \mathrel\lor \cdots \mathrel\lor \DISC{m}$ to $F$.
+$\thD$-unsatisfiable. For \rn{Split}, we add the exhaustiveness clause $\DISC{1} \mathrel\lor \cdots \mathrel\lor \DISC{m}$ to $F$.
 Decisions on which branch to take are thus performed externally by the SAT
-solver. % and not by the theory solver.
-All other rules add equalities to the
+solver. All other rules add equalities to the
 internal state of the theory solver. The rules in phase~1 are performed
@@
 It visits $\ec{t_1}, \ldots, \ec{t_n}$ and all constructor arguments in these equivalence classes recursively.
-If while doing so it returns to $\ec{t}$, it deduces that \rn{Acyclic} applies. % to $t$.
-To recognize when the precondition of \rn{Unique} holds,
+If while doing so it returns to $\ec{t}$, it deduces that \rn{Acyclic} applies. To recognize when the precondition of \rn{Unique} holds,
 the procedure considers the set $\SSS$ of all codatatype equivalence classes. It
-\begin{conf}\goodbreak\noindent\end{conf}%  %%% TYPESETTING BREAK
-simultaneously visits the equivalence classes of arguments of constructor terms in each equivalence class in $\SSS$,
+\begin{conf}\goodbreak\noindent\end{conf}simultaneously visits the equivalence classes of arguments of constructor terms in each equivalence class in $\SSS$,
 while partitioning $\SSS$ into $\SSS_1, \ldots, \SSS_n$ based on the top-most symbol of constructor terms in these equivalence classes
 and the equivalence of their arguments of ordinary types.
-It then partitions each set %in this partition %in the same manner
-recursively.
-If %after doing so,
-the resulting partition contains a set $\SSS_i$ containing distinct terms $u$
+It then partitions each set recursively.
+If the resulting partition contains a set $\SSS_i$ containing distinct terms $u$
 and $v$, it deduces that \rn{Unique} applies to $u$ and~$v$.
@@
 conjectures.
-%%% @ANDY: not sure how to phrase this now
-%where checking satisfiability requires reasoning about ground literals in the
-%presence of universally quantified formulas.
 Many SMT solvers, including CVC4, can reason about quantified formulas using incomplete instantiation-based methods~\cite{MouraBjoerner07,ReynoldsTinelliMoura14}.
-These methods extend naturally to %quantified
-formulas involving datatypes and codatatypes.
+These methods extend naturally to formulas involving datatypes and codatatypes.
 
@@
 applied.
-This does not %necessarily
-impact the correctness of the procedure in this setting,
-since %by assumption
-the solver is already incomplete in the presence of quantified formulas.
+This does not impact the correctness of the procedure in this setting,
+since the solver is already incomplete in the presence of quantified formulas.
 Nonetheless, two techniques help increase the precision of the solver.
@@
 Second, we can conditionally apply \rn{Single} to codatatype terms that may have cardinality one.
-For example, the $\ty{stream}_{\,\ty{u}}$ codatatype %, where $\ty{u}$ is uninterpreted.
-has cardinality one precisely when $\ty{u}$ has cardinality one.
+For example, the $\ty{stream}_{\,\ty{u}}$ codatatype has cardinality one precisely when $\ty{u}$ has cardinality one.
 If there exist two equivalence classes $\ec{s}$ and $\ec{t}$ for this type,
-the implementation adds the clause %of the form
-$(\exists x\;y\typ{\ty{u}}.\;\, x \not\teq y ) \mathrel\vee s \teq t$ to $F\!$.
+the implementation adds the clause $(\exists x\;y\typ{\ty{u}}.\;\, x \not\teq y ) \mathrel\vee s \teq t$ to $F\!$.
 \begin{rep}In other words, either the cardinality of $\ty{u}$ is greater than one or $s$ is equal to $t$.\end{rep}
@@
 following the lines of Barrett et al.\ \cite{barrett-et-al-2007}.
-%and are not reflected in the presentation of the calculus.
-\begin{rep}We briefly mention the main ones. \end{rep}%
-Discriminators are part of the signature and not abbreviations.
+\begin{rep}We briefly mention the main ones. \end{rep}Discriminators are part of the signature and not abbreviations.
 This requires extending the decision procedure with several
-rules, % \cite{barrett-et-al-2007},
-which apply uniformly to datatypes and codatatypes.
+rules, which apply uniformly to datatypes and codatatypes.
 This approach often leads to better performance because it introduces terms
@@
 
-% AJR : this paragraph may have been what confused Reviewer #2. It does suggest that we are copying verbatim what was in CVC3 (which is what Reviewer #2 thought).
-% Since we aren't saying anything new beyond what is said in Barrett et al (and honestly, we didn't critically evaluate the performance of the updated implementation), 
-% maybe we should just remove this paragraph.
-
 As Barrett et al.\ \cite{barrett-et-al-2007} observed for their procedure,
@@
 
-%\section{Examples}
-%\label{sec:examples}
-
 \section{Evaluation on Isabelle Problems}
@@
 Isabelle formalizations using Sledgehammer \cite{blanchette-et-al-2013-smt}.
-%There are not yet many benchmarks available, since codatatypes are a
-%recent addition to Isabelle \cite{blanchette-et-al-2014-impl}. \rem{Necessary to mention this? Seems to detract from significance of contribution.}
 We
@@
 conducting any experiments. The experimental data are available
-online.%
-\footnote{\url{http://lara.epfl.ch/~reynolds/CADE2015-cdt/}}
-
-%http://www21.in.tum.de/~blanchet/codp_data.tgz}}
-
-%  & 879 & 2974\phantom{0} & 317 & 4170
+online.\footnote{\url{http://lara.epfl.ch/~reynolds/CADE2015-cdt/}}
 
 For each goal in each theory, Sledgehammer selected about 16~lemmas,
-%\end{conf}%
-%\begin{rep}a fixed number of lemmas\end{rep},
 which were monomorphized and translated to SMT-LIB~2 along with the goal.
@@
 Due to the lack of machinery for reconstructing inferences about (co)datatypes
-in Isabelle, the solvers are trusted as oracles. % in these experiments.
-%Z3 is included for comparison.
-
+in Isabelle, the solvers are trusted as oracles. 
 The development version of CVC4 was run on each problem several times, with the
@@
 contributions of the acyclicity and uniqueness rules were also measured.
-% to find
-% out whether these somewhat expensive rules are useful in practice.
 Even when the
@@
 reasoning.
-% when this makes sense.
-%This allows us to answer the question, \relax{What are the
-%benefits of activating the decision procedure as opposed to letting
-%Sledgehammer do what it would normally do?}
-
-%  * experience with such features is not extremely high -- e.g. arithmetic,
-%    a most useful theory on lots of benchmarks, increases Sledgehammer's
-%    success rate by 2 percentage points, or 4\%, with Z3 in earlier work \cite{xxx}
-
-%  * so when analyzing statistics of this kind, with highly varied problems,
-%    we must not overestimate the importance of a single trick
-%  * on the other hand: every percentage point counts! cite Tom Hales
 
@@
 \newcommand\captioni{\,Number of solved goals for the three benchmark suites}
-%\begin{rep}
-%\renewcommand\captioni{\,Number of solved goals with 16~lemmas per goal}
-%\end{rep}
 
@@
   & 221 & 209 & 775 & 777 & 52 & 51 & 1048 & 1037 \\%[\jot]
-%Datatype simplification only
-%  &  \\
 Datatypes without \rn{Acyclic}
@@
   & 227 & 213 & 786 & 791 & 52 & 51 & 1065 & 1055 \\%[\jot]
-%Codatatype simplification only
-%  &  \\
 Codatatypes without \rn{Unique}\!\!
@@
 Full (co)datatypes
-  & \bfseries 229 & -- & \bfseries 815 & -- & \bfseries 59 & -- & \bfseries 1103 & -- %\\[\jot]
-\end{tabular}\end{center}
+  & \bfseries 229 & -- & \bfseries 815 & -- & \bfseries 59 & -- & \bfseries 1103 & -- \end{tabular}\end{center}
 \caption{\captioni}
@@
 The results are summarized in Table~\ref{tab:benchsixteen}.
-%The goals in Table \ref{tab:benchsixteen} include 16
-%heuristically selected lemmas (before monomorphization);
-%\cite{meng-paulson-2009-relev};
-%that number is increased to 256 for Table \ref{tab:benchtwofiftysix}.
 The decision procedure
@@
 success rate increase of over 5\%, which is quite significant.
-%For Isabelle
-%users, the practical usefulness of interpreted (co)datatypes is roughly
-%comparable to that of arithmetic \cite{blanchette-et-al-2013-smt}, although this
-%obviously depends on the application.
 The raw evaluation data also
@@
 more expensive rules, make a contribution.
-%* and finally, we look in more detail at one or two such proofs, by presenting
-%  it and explaining it (and, before that, understanding it)
-%
 Three proofs were found thanks to
@@
 The fourth proof, by Gammie and Lochbihler, is more elaborate:
-%
 \begin{quote}
@@
 \end{quote}
-%
 \noindent
@@
 the acyclicity rule for datatypes exploits $\mu$-terms to detect cycles.
-% are not very
-% difficult per se, but there are tricky corner cases to take care of if we care
-% about completeness (which we do).
-%% alliteration
 The empirical results on Isabelle benchmarks
@@
 be interfaced for generating counterexamples in proof assistants;
-%as an alternative to Nitpick;
 in this context, the acyclicity and uniqueness rules are
@@
 
-%Finally, it might be possible to go
-%further in terms of supporting nested and mixed (co)recursion and quantified
-%ormulas over (co)datatypes in solvers.
-
-{\footnotesize
+{%\footnotesize
 \def\ackname{Acknowledgment}
-\paragraph{\footnotesize\ackname.}
+\paragraph{%\footnotesize
+\ackname.}
 \looseness=-1
@@
 of the theory solver for datatypes in CVC4.
-%
 Our present and former bosses, Viktor Kuncak, Stephan Merz, Tobias Nipkow,
@@
 for.
-%
 Peter Gammie and Andreas Lochbihler provided useful benchmarks.
@@
 codatatypes.
-%
 Pascal Fontaine, Andreas Lochbihler, Andrei Popescu, Christophe Ringeissen, Mark
@@
 the Deutsche
-Forschungs\-gemein\-schaft %(DFG)
-project ``Den Hammer h\"arten'' (grant NI\,491\slash 14-1)
+Forschungs\-gemein\-schaft project ``Den Hammer h\"arten'' (grant NI\,491\slash 14-1)
 and the Inria technological development

### 2015-09-10 jour/jour.tex
starting to port

--- 2015-05-19 tex/cade_final/main.tex
@@
-%envcountsect,
-\documentclass[a4paper,oribibl,envcountsame,draft]{llncs}
-\usepackage{version}
-
-\excludeversion{rep}
-\includeversion{conf}
-\excludeversion{nomemoriam}
-\includeversion{memoriam}
+%%% TODO: Give small example where Single is needed
+
+\documentclass[smallcondensed,draft]{svjour3}
 
@@
 
-\def\thewordpaper{paper}
-\newcommand\dotReportFootnote[1]{.}
-
 \newcommand\FV{\mathrm{FV}}
@@
 \newcommand\afterDot{\;} %%% Too little space by default after "Lemma Foo."
-
-\begin{rep}
-\def\thewordpaper{report}
-\renewcommand\dotReportFootnote[1]{.\footnote{#1}}
-\end{rep}
 
@@
 
-\begin{nomemoriam}
-\noindent\hfill\emph{\phantom{In memoriam Morgan Deters 1979--2015}}
-
-\vspace*{-6\smallskipamount}
-\end{nomemoriam}
-\begin{memoriam}
 \noindent\hfill\emph{\relax{In memoriam Morgan Deters 1979--2015}}
-\end{memoriam}
 
@@
 
-\begin{nomemoriam}
-\vskip3\smallskipamount
-\end{nomemoriam}
-
 %% The institutions above shouldn't count as footnotes
@@
 %(using an ML-like syntax)
-{\looseness=-1
 introduces a type that
@@
 definition
-\begin{conf}$\vthinspace \keyw{datatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat}) \,\mid\, \const{Infty}$\vvthinspace, \end{conf}%
-\begin{rep}\[\vthinspace \keyw{datatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat}) \,\mid\, \const{Infty}\]\end{rep}%
+\[\vthinspace \keyw{datatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat}) \,\mid\, \const{Infty}\]
 %with a dedicated constructor for representing infinity, the
@@
 
-}
-
-{\looseness=-1
 Datatypes and codatatypes are an integral part of many proof assistants,
@@
 solvers, has been extended with a syntax for datatypes.
-In this \thewordpaper, we introduce a
-unified \hfill decision \hfill procedure \hfill for \hfill universal \hfill problems \hfill involving \hfill datatypes \hfill and \hfill codatatypes%
-\begin{conf}\goodbreak\noindent\end{conf}%  %%% TYPESETTING BREAK
-\begin{rep} \end{rep}%
+In this article, we introduce a
+unified decision procedure for universal problems involving datatypes and codatatypes
 in combination (Section~\ref{sec:the-decision-procedure}).
@@
 %codatatypes.
-\begin{conf}%
-Detailed proofs %of completeness
-are included in a %technical
-report
-\cite{our-report}.
-\end{conf}%
-
-}
 
@@
 of infinite sequences or streams over an element type $\tau$ is
-allowed\begin{conf}, even though it has no base case\end{conf}\begin{rep}; the
+allowed, even though it has no base case; the
 corresponding datatype would be rejected as non-well-founded
-\cite{blanchette-et-al-2015-esop}\end{rep}.
+\cite{blanchette-et-al-2015-esop}.
 %
@@
 a subterm; it is used to represent cyclic values in the generated
-models\begin{rep} and in the metatheory\end{rep}. For example,
+models and in the metatheory. For example,
 the $\mu$-term $\const{SCons}(1,\: \MU s.\; \const{SCons}(0,\: \const{SCons}(9,\: s)))$
@@
 
-\begin{rep}\paragraph{Related Work.} Barrett\end{rep}%
-\begin{conf}Barrett\end{conf}
+\paragraph{Related Work.} Barrett
 et al.\ \cite{barrett-et-al-2007} provide a good account of
@@
 but also to the SMT solver
-Z3 \cite{de-moura-bjoerner-2008} \begin{rep}in unpublished work by Leonardo de Moura \end{rep}and \begin{rep}to \end{rep}a SPASS-like %prototypical
-superposition prover \begin{rep}by Daniel Wand\end{rep} \cite{wand-2014}.
+Z3 \cite{de-moura-bjoerner-2008} in unpublished work by Leonardo de Moura and to a SPASS-like %prototypical
+superposition prover by Daniel Wand \cite{wand-2014}.
 Closely related are the automatic structural
@@
 The $\mu$-notation is inspired by the $\mu$-calculus
-\begin{conf}\cite{kozen-1983}.\end{conf}
-\begin{rep}\cite[etc.]{kozen-1983,endrullis-et-al-2011}.\end{rep}
+\cite[etc.]{kozen-1983,endrullis-et-al-2011}.
 
@@
 values. %, and may contain other interpreted types (e.g., \ty{int}, \ty{real}).
-\begin{rep}
 The metavariables $\delta,\:\varepsilon$ range over (co)datatypes,
@@
 distinguish them from variables (e.g., $x$, $y$).
-\end{rep}
-Names starting with an uppercase letter \begin{rep}(e.g.,
-$\const{S}$) \end{rep}are reserved for constructors. With each function symbol \const{f}
+Names starting with an uppercase letter (e.g.,
+$\const{S}$) are reserved for constructors. With each function symbol \const{f}
 is associated a list of argument types $\tau_1,\ldots,\tau_n$ (for $n \ge 0$)
 and a return type $\tau$, written
-$\const{f} : \tau_1 \times \cdots \times \tau_n \to \tau$\begin{rep},
-which collapses to $\const{f} : \tau$ if $n = 0$\end{rep}.
-\begin{rep}%
+$\const{f} : \tau_1 \times \cdots \times \tau_n \to \tau$,
+which collapses to $\const{f} : \tau$ if $n = 0$.
+%
 Functions invocations $\const{f}(t_1,\ldots,t_n)$
@@
 parentheses in terms.
-\end{rep}
 The set $\Funcs$ must at least contain
 $\const{true},\, \const{false} : \ty{bool}$, interpreted as truth values.
-The only predicate is equality ($\teq$)\begin{rep} and belongs to the logical symbols;
+The only predicate is equality ($\teq$) and belongs to the logical symbols;
 other predicates can be represented as functions to $\ty{bool}$,
-with $\const{p}(\ldots)$ abbreviating $\const{p}(\ldots) \teq \const{true}$\end{rep}.
+with $\const{p}(\ldots)$ abbreviating $\const{p}(\ldots) \teq \const{true}$.
 The notation $t\typ{\tau}$ stands for a term $t$ of type~$\tau$.
 When applied to terms, the symbol $=$ denotes syntactic equality.
-\begin{rep}The operator $\bigwedge_{\,i}\, \varphi_i$ abbreviates a conjunction
+The operator $\bigwedge_{\,i}\, \varphi_i$ abbreviates a conjunction
 $\varphi_1 \mathrel\land \cdots \mathrel\land \varphi_n$.
-Finally, $\bar x$ abbreviates a list or tuple $x_1,\ldots,x_n$.\end{rep}
+Finally, $\bar x$ abbreviates a list or tuple $x_1,\ldots,x_n$.
 
@@
 Each (co)datatype specification consists of $\ell$~mutually recursive types that are
-either all datatypes or all codatatypes. \begin{rep}Polymorphic types, nested
-(co)recursion,\end{rep}\begin{conf}Nested (co)recursion\end{conf} and datatype--codatatype mixtures
-fall outside this fragment%
-\dotReportFootnote{%
-In principle, rank-1 (top-level) polymorphism \cite{blanchette-paskevich-2013}
+either all datatypes or all codatatypes. Polymorphic types, nested
+(co)recursion, and datatype--codatatype mixtures
+fall outside this fragment.\footnote{In principle, rank-1 (top-level) polymorphism \cite{blanchette-paskevich-2013}
 should not raise any special difficulties. Nesting datatypes inside datatypes,
@@
 \cite{blanchette-et-al-2014-codata}).}
-\begin{rep}We allow ourselves some notational parameterization
+We allow ourselves some notational parameterization
 % at the metalevel
@@
 $\ty{stream}_{\,\ty{int}}$, $\ty{stream}_{\,\ty{bool}}$,
-and \smash{$\ty{stream}_{\,\ty{stream}_{\,\ty{real}}}$}.\end{rep}
+and \smash{$\ty{stream}_{\,\ty{stream}_{\,\ty{real}}}$}.
 
@@
 constructors, and the selectors must be
-\begin{conf}fresh\end{conf}\begin{rep}distinct and different from
-existing names\end{rep}%
-\dotReportFootnote{It can be convenient to specify the same selector
+distinct and different from existing names.%
+\footnote{It can be convenient to specify the same selector
 for several constructors associated with the same (co)data\-type,
@@
 
-\begin{rep}%
 Here are a few examples of legal specifications of (co)datatype families:
@@
 finite or infinite, that witnesses nonemptiness \cite{blanchette-et-al-2015-esop}.
-\end{rep}
 
@@
 A type is (\emph{co})\emph{recursive} if it belongs to such a set of types.
-\begin{rep}%
 Types can be declared together in a mutual fashion even if they are not actually
 (co)recursive. The semantic notion is more precise and is the one that interests
-us.
-\end{rep}%
-%
-Non(co)recursive type specifications such~as
-\begin{conf}%
-\vthinspace $\keyw{datatype}~\vthinspace\ty{option}_{\,\tau} \,=\, \const{None} \,\mid\, \const{Some}(\tau)$
-\end{conf}%
-\begin{rep}%
+us. Non(co)recursive type specifications such~as
 \[\begin{aligned}[t]
@@
 \]
-\end{rep}%
 are permitted.
-\begin{rep}At the semantic level, it makes no difference whether
+At the semantic level, it makes no difference whether
 such types are introduced as datatypes or as codatatypes.
-\end{rep}%
 %Without loss of generality, we consider that these types are datatypes
@@
 selector--constructor equations \cite{barrett-et-al-2007}.
-\begin{rep}
 A drawback of this approach is that it does
@@
 
-\end{rep}
 A related semantic view of datatypes is as initial algebras. Codatatypes are
@@
 
-Datatypes and codatatypes share many basic properties\begin{rep}. All properties
+Datatypes and codatatypes share many basic properties. All properties
 below are implicitly universally quantified and range over all $i$, $j$,
-$\jPrime$, and $k$ within bounds\end{rep}:
+$\jPrime$, and $k$ within bounds:
 % and to all possible splits of the
@@
 %
-\begin{rep}%
 Expressed in the algebraic jargon, exhaustiveness helps ensure that ``no
@@
 constructor is left completely unspecified.
-\end{rep}%
+%
 %
@@
 %
-Datatypes are additionally characterized by an induction principle\begin{rep}:
+Datatypes are additionally characterized by an induction principle:
 %
@@
 denotes either $P_{\iPrime}(x)$ if there exists some $\iPrime$ such that
-the formula is type-correct or else $\top$\end{rep}.
+the formula is type-correct or else $\top$.
 The principle ensures that the interpretation of datatypes is standard.
@@
 
-For codatatypes, the dual notion is called coinduction. \begin{rep}It depends on
+For codatatypes, the dual notion is called coinduction. It depends on
 witnesses $R_i$ that are required to be bisimulations:
@@
 $\iPrime$ such that the formula is type-correct or $x \teq y$ otherwise.
-\end{rep}%
 This axiom encodes a form of extensionality:\ Two values that yield the same
@@
 selectors and discriminators.
-%
-\begin{rep}\par\end{rep}
-  %
+
 In addition, codatatypes are guaranteed to contain all values corresponding to
 infinite ground constructor terms.
-\begin{rep}
 In general, this cannot be captured by a
@@
 function space $\ty{nat} \to \ty{int}$.
-\end{rep}
 
@@
 
-%\begin{rep}
 % TODO: State more precisely and prove?
-%\end{rep}
 
@@
 and
-\begin{conf}%
-\vthinspace$\keyw{codatatype}~\ty{b} = \const{B}(\ty{b},\: \ty{c},\: \ty{b},\: \ty{unit})
-~\keyw{and}~ \ty{c} = \const{C}(\ty{a},\: \ty{b},\: \ty{c})$,
-\end{conf}%
-\begin{rep}%
 \[\begin{aligned}[t]
@@
 \]
-\end{rep}%
 where \ty{unit} is a datatype with the single constructor $\const{Unity} :
@@
 singletons precisely when \ty{unit} is interpreted as a singleton.
-Fortunately, \begin{rep}given cardinalities for the ordinary types, \end{rep}%
+Fortunately, given cardinalities for the ordinary types,
 it is easy to characterize this degenerate case.
@@
 
-\begin{rep}
 \begin{proof}
@@
 \end{proof}
-\end{rep}
 
@@
 indicates that the $\mu$-terms $t$ and $u$
-are syntactically equivalent for some capture-avoiding renaming of $\mu$-bound variables\begin{rep}---e.g.,
+are syntactically equivalent for some capture-avoiding renaming of $\mu$-bound variables---e.g.,
 %$\MU x.\; \const{C}( x ) \aequiv \MU y.\; \const{C}( y )$ and
@@
 $\MU x.\; \const{C}( x ) \not\aequiv \MU x.\; \const{D}( y, x ) \not\aequiv \MU x.\; \const{D}( z, x )
-\not\aequiv \MU y.\; \const{D}( y, x )$\end{rep}.
+\not\aequiv \MU y.\; \const{D}( y, x )$.
 Two $\mu$-terms can denote the same value despite being $\alpha$-disequivalent---e.g.,
@@
 make progress by discovering that the subterm $\const{S}(y)$ of $\MU y.\;
-\const{S}(\const{S}(y))$ must be equal to the entire term\begin{rep}
-as demonstrated next\end{rep}.
+\const{S}(\const{S}(y))$ must be equal to the entire term,
+as demonstrated next.
 
@@
 The mapping $\Val$ is updated accordingly:
-\begin{conf}%
-$\Val \ec{y} \,=\, \MU {\Varec{y}}.\; \const{S}({\Varec{y}})$.
-\end{conf}%
-\begin{rep}
 \begin{align*}
@@
 \end{align*}
-\end{rep}
 Since $\Val \ec{x} \aequiv \Val \ec{y}$,
@@
 D$\negvthinspace, let
-\begin{conf}
-$\smash{\SSSS 0_t} = \{ t \}$ and
-$\smash{\SSSS {i+1}_t} = \smash{\SSSS i_t \mathrel\cup \{ \const s( u ) \mid u\typ{\delta} \in \SSSS i_t{,}}~\allowbreak\smash{ \delta \in \Data{,}~ \left|\delta\right| \text{ is finite}{,}~ \const{s} \in \Sel^{\delta}}  \}$,
-\end{conf}%
-\begin{rep}
 \begin{align*}
@@
 \end{align*}
-\end{rep}%
 %for $i \geq 0$, let ;
@@
 
-\begin{rep}
 \begin{lemma}
@@
 %\end{proof}
-\end{rep}
 
@@
 to ${\Varec{z}}$.
-\begin{rep}Legal substitutions for ${\Varec{z}}$ may cause the range of $\Val$ to contain abnormal terms.
+Legal substitutions for ${\Varec{z}}$ may cause the range of $\Val$ to contain abnormal terms.
 For example, after assigning $\MU {\Varec{z}}.\; \const{D}( \const{C}( {\Varec{z}} ) )$ to $\Varec{z}$, we have
@@
 with normal form
-$\MU {\Varec{u}}.\; \const{C}( \MU {\Varec{z}}.\; \const{D}( {\Varec{u}} ) )$.\end{rep}
+$\MU {\Varec{u}}.\; \const{C}( \MU {\Varec{z}}.\; \const{D}( {\Varec{u}} ) )$.
 \xend
@@
 \end{lemma}
-\begin{rep}
+
 \begin{proof}
@@
 For each $i$, let $u_\iPrime$ be the $i$\vvthinspace th argument of $t'$.
-\begin{rep}By Lemma~\ref{lem:mu-norm-arg},\end{rep}%
-\begin{conf}Clearly,\end{conf} %% TODO: Fix! ;)
+By Lemma~\ref{lem:mu-norm-arg},
 $\expand{t'}{u_\iPrime} \vsim \ValC \ec{u_i}$.
@@
 the procedure considers the set $\SSS$ of all codatatype equivalence classes. It
-\begin{conf}\goodbreak\noindent\end{conf}%  %%% TYPESETTING BREAK
 simultaneously visits the equivalence classes of arguments of constructor terms in each equivalence class in $\SSS$,
@@
 $(\exists x\;y\typ{\ty{u}}.\;\, x \not\teq y ) \mathrel\vee s \teq t$ to $F\!$.
-\begin{rep}In other words, either the cardinality of $\ty{u}$ is greater than one or $s$ is equal to $t$.\end{rep}
+In other words, either the cardinality of $\ty{u}$ is greater than one or $s$ is equal to $t$.
 
@@
 %and are not reflected in the presentation of the calculus.
-\begin{rep}We briefly mention the main ones. \end{rep}%
+We briefly mention the main ones.
 Discriminators are part of the signature and not abbreviations.
@@
 For each goal in each theory, Sledgehammer selected about 16~lemmas,
-%\end{conf}%
-%\begin{rep}a fixed number of lemmas\end{rep},
 which were monomorphized and translated to SMT-LIB~2 along with the goal.
@@
 \newcommand\captioni{\,Number of solved goals for the three benchmark suites}
-%\begin{rep}
 %\renewcommand\captioni{\,Number of solved goals with 16~lemmas per goal}
-%\end{rep}
 
@@
 \keyw{by}~\textit{coinduction}~\textit{auto}
-in Isabelle\begin{rep} \cite{blanchette-et-al-2014-impl}\end{rep}.
+in Isabelle \cite{blanchette-et-al-2014-impl}.
 The fourth proof, by Gammie and Lochbihler, is more elaborate:

### 2015-09-11 jour/jour.tex
more

@@
 %%% TODO: Give small example where Single is needed
+
+%%% GOOD IDEAS in impl of codata models:
+%%% 1. de Bruijn
+%%% 2. enum: "X nat"
+%%% 3. norm by DFA minim
+%%% 4. check membership in bad set by matching
 
@@
 
-\makeatletter
-\def\spnXwtheorem#1#2#3#4{\@spynthm{#1}{#2}{#3}{#4}%
-                         \@addtoreset{#1}{chapter}}%
-\makeatother
-\spnXwtheorem{examplex}{Example}{\itshape}{\rmfamily}
+\spnewtheorem{examplex}{Example}{\itshape}{\rmfamily}
 
@@
 
-\makeatletter
-\ps@myheadings
-\makeatother
-
-\include{defs}
-
 \hyphenation{data-type data-types co-data-type co-data-types isa-belle sledge-hammer
@@
 
-\author {Andrew Reynolds\inst{1} \and Jasmin Christian Blanchette\inst{2,3}}
+\author {Andrew Reynolds \and Jasmin Christian Blanchette}
 \authorrunning {A. Reynolds \and J. C. Blanchette}
+
 \institute{
-\'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL), Switzerland
+Andrew Reynolds
+\at \'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL), Switzerland
 \and
-Inria Nancy \& LORIA, Villers-l\`es-Nancy, France
-\and
-Max-Planck-Institut f\"ur Informatik, Saarbr\"ucken, Germany
-}
+Jasmin Christian Blanchette%, funded by DFG grant \textit{Hardening the Hammer} (Ni\,491\slash 14-1)
+\at Inria \& LORIA, Nancy, France \\
+Max-Planck-Institut f\"ur Informatik, Saarbr\"ucken, Germany}
 
@@
 \noindent\hfill\emph{\relax{In memoriam Morgan Deters 1979--2015}}
+
+\bigskip
 
@@
 $\infty$.
-Compared with the conventional %algebraic
-definition
-\[\vthinspace \keyw{datatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat}) \,\mid\, \const{Infty}\]
-%with a dedicated constructor for representing infinity, the
-the codatatype avoids
-one case by unifying the infinite and finite nonzero cases.
 
@@
 and the Inria technological development
-action ``Contre-exemples Utilisables par Isabelle et Coq'' (CUIC).
-
-}
-
-\bibliographystyle{splncs03}
+action ``Contre-exemples utilisables par Isabelle et Coq'' (CUIC).
+
+}
+
+\bibliographystyle{spmpsci}
 \bibliography{bib}{}

### 2015-10-27 jour/jour.tex
more

@@
 have emerged in a few automatic theorem provers as well. The SMT-LIB
-%\cite{barrett-et-al-2010}
+\cite{xxxxxxx}
 format, implemented by most SMT
@@
 A signature $\Sig = (\Types, \Funcs)$ consists of a set of types $\Types$ and a
-set of function symbols $\Funcs$. Types are atomic sorts and interpreted as
+set of function symbols $\Funcs$\negvthinspace. Types are atomic sorts and interpreted as
 nonempty domains. The set~$\Types$ must contain a
@@
 $\const{S}(\const{S}(\ldots))$.
-
-\pagebreak[2]
 
@@
 \newcommand\gandl{G\&L}
-\newcommand\HD[1]{\hbox to2.25em{\footnotesize\hfill\!\!\!\!#1\!\!\!\!\hfill}}
+\newcommand\HD[1]{\hbox to2em{\footnotesize\hfill\!\!\!\!#1\!\!\!\!\hfill}}
 
@@
 
-\newcommand\captioni{\,Number of solved goals for the three benchmark suites}
-%\renewcommand\captioni{\,Number of solved goals with 16~lemmas per goal}
+\newcommand\captioni{\,Number of solved goals with 16~lemmas per goal}
+\newcommand\captionii{\,Number of solved goals with 256~lemmas per goal}
 
@@
 \normalsize
-\begin{center}\begin{tabular}{@{\;}l@{\kern1.5em}cc@{\kern.5em}cc@{\kern.5em}cc@{\kern1.3em}cc@{\;}}
-  & \multicolumn{2}{c}{Distro} & \multicolumn{2}{c}{AFP} & \multicolumn{2}{c}{\gandl} & \multicolumn{2}{c}{Overall} \\[-0.5pt]
+\begin{center}\begin{tabular}{@{\;}l@{\kern1.5em}c@{\kern.5em}c@{\kern1em}c@{\kern.5em}c@{\kern1em}c@{\kern.5em}c@{\kern1.5em}c@{\kern.5em}c@{\;}}
+  & \multicolumn{2}{@{}c@{\kern1em}}{Distro} & \multicolumn{2}{@{}c@{\kern1em}}{AFP} & \multicolumn{2}{@{}c@{\kern1.5em}}{\gandl} & \multicolumn{2}{@{}c@{\;}}{Overall} \\[-0.5pt]
   & \HD{CVC4} & \HD{Z3} & \HD{CVC4} & \HD{Z3} & \HD{CVC4} & \HD{Z3} & \HD{CVC4}& \HD{Z3} 
@@
 
+\begin{table*}[t!]
+\normalsize
+\begin{center}\begin{tabular}{@{\;}l@{\kern1.5em}c@{\kern.5em}c@{\kern1em}c@{\kern.5em}c@{\kern1em}c@{\kern.5em}c@{\kern1.5em}c@{\kern.5em}c@{\;}}
+  & \multicolumn{2}{@{}c@{\kern1em}}{Distro} & \multicolumn{2}{@{}c@{\kern1em}}{AFP} & \multicolumn{2}{@{}c@{\kern1.5em}}{\gandl} & \multicolumn{2}{@{}c@{\;}}{Overall} \\[-0.5pt]
+  & \HD{CVC4} & \HD{Z3} & \HD{CVC4} & \HD{Z3} & \HD{CVC4} & \HD{Z3} & \HD{CVC4}& \HD{Z3} 
+\MIDRULE
+No (co)datatypes
+  & 221 & 209 & 775 & 777 & 52 & 51 & 1048 & 1037 \\%[\jot]
+%Datatype simplification only
+%  &  \\
+Datatypes without \rn{Acyclic}
+  & 227 & -- & 780 & -- & 52 & -- & 1059 & -- \\
+Full datatypes
+  & 227 & 213 & 786 & 791 & 52 & 51 & 1065 & 1055 \\%[\jot]
+%Codatatype simplification only
+%  &  \\
+Codatatypes without \rn{Unique}\!\!
+  & 222 & -- & 804 & -- & 56 & -- & 1082 & -- \\
+Full codatatypes
+  & 223 & -- & 804 & -- & \bfseries 59 & -- & 1086 & -- \\%[\jot]
+Full (co)datatypes
+  & \bfseries 229 & -- & \bfseries 815 & -- & \bfseries 59 & -- & \bfseries 1103 & -- %\\[\jot]
+\end{tabular}\end{center}
+\caption{\captionii}
+\label{tab:benchtwofittysix}
+\end{table*}
+
+\newpage
+
 The results are summarized in Table~\ref{tab:benchsixteen}.

### 2015-10-27 jour/jour.tex
more

@@
 that define codatatypes falling within the supported fragment. We
-added two unpublished theories about Bird and Stern--Brocot trees
+added unpublished theories about Bird and Stern--Brocot trees
 by Peter Gammie and Andreas Lochbihler (\gandl,
@@
 
-For each goal in each theory, Sledgehammer selected about 16~lemmas,
-which were monomorphized and translated to SMT-LIB~2 along with the goal.
-The resulting problem was given to the
+For each goal in each theory, we used Sledgehammer to select either 16 or
+256~lemmas, which were monomorphized and translated to SMT-LIB~2 along with the
+goal. The resulting problem was given to the
 development version of CVC4 and to Z3 4.3.2 for comparison, each running for
@@
 basic lemmas about constructors and selectors, allowing some (co)datatype
-reasoning.
+reasoning. This is especially the case for problems generated using 256~lemmas.
 % when this makes sense.
@@
 No (co)datatypes
-  & 221 & 209 & 775 & 777 & 52 & 51 & 1048 & 1037 \\%[\jot]
+  & 287 & 282 & 765 & 771 & 47 & 44 & 1099 & 1097 \\%[\jot]
 %Datatype simplification only
@@
 Datatypes without \rn{Acyclic}
-  & 227 & -- & 780 & -- & 52 & -- & 1059 & -- \\
+  & 298 & -- & 771 & -- & 47 & -- & 1116 & -- \\
 Full datatypes
-  & 227 & 213 & 786 & 791 & 52 & 51 & 1065 & 1055 \\%[\jot]
+  & 298 & 294 & 775 & 783 & 47 & 44 & 1120 & 1121 \\%[\jot]
 %Codatatype simplification only
@@
 Codatatypes without \rn{Unique}\!\!
-  & 222 & -- & 804 & -- & 56 & -- & 1082 & -- \\
+  & 288 & -- & 797 & -- & 47 & -- & 1132 & -- \\
 Full codatatypes
-  & 223 & -- & 804 & -- & \bfseries 59 & -- & 1086 & -- \\%[\jot]
+  & 288 & -- & 797 & -- & \bfseries 52 & -- & 1137 & -- \\%[\jot]
 Full (co)datatypes
-  & \bfseries 229 & -- & \bfseries 815 & -- & \bfseries 59 & -- & \bfseries 1103 & -- %\\[\jot]
+  & \bfseries 299 & -- & \bfseries 806 & -- & \bfseries 52 & -- & \bfseries 1157 & -- %\\[\jot]
 \end{tabular}\end{center}
@@
 No (co)datatypes
-  & 221 & 209 & 775 & 777 & 52 & 51 & 1048 & 1037 \\%[\jot]
+  & 617 & 560 & 1503 & 1271 & 89 & 80 & 2209 & 1911 \\%[\jot]
 %Datatype simplification only
@@
 Datatypes without \rn{Acyclic}
-  & 227 & -- & 780 & -- & 52 & -- & 1059 & -- \\
+  & 617 & -- & \bfseries 1504 & -- & 90 & -- & 2211 & -- \\
 Full datatypes
-  & 227 & 213 & 786 & 791 & 52 & 51 & 1065 & 1055 \\%[\jot]
+  & 617 & 560 & \bfseries 1504 & 1263 & 90 & 78 & 2211 & 1901 \\%[\jot]
 %Codatatype simplification only
@@
 Codatatypes without \rn{Unique}\!\!
-  & 222 & -- & 804 & -- & 56 & -- & 1082 & -- \\
+  & 617 & -- & 1501 & -- & 90 & -- & 2208 & -- \\
 Full codatatypes
-  & 223 & -- & 804 & -- & \bfseries 59 & -- & 1086 & -- \\%[\jot]
+  & \bfseries 620 & -- & 1502 & -- & 98 & -- & \bfseries 2220 & -- \\%[\jot]
 Full (co)datatypes
-  & \bfseries 229 & -- & \bfseries 815 & -- & \bfseries 59 & -- & \bfseries 1103 & -- %\\[\jot]
+  & 619 & -- & 1501 & -- & \bfseries 99 & -- & 2219 & -- %\\[\jot]
 \end{tabular}\end{center}

### 2015-11-01 jour/jour.tex
more

@@
+%%% TODO: Mealy machine minimization
 %%% TODO: Give small example where Single is needed
@@
 
+  * decision procedure useful both for proving and for model finding
+  * model finding: critical, without it unsound models
+  * contribution to proving is less clear
+    * some parts can be axiomatized
+    * only acyclicity and uniqueness, but how often do they pop up in practice?
+
 To evaluate the decision procedure, we generated benchmark problems from existing
@@
 online.%
-\footnote{\url{http://lara.epfl.ch/~reynolds/CADE2015-cdt/}}
+\footnote{\url{http://lara.epfl.ch/~reynolds/JAR-CADE2015-cdt/}}
 
@@
 goal. The resulting problem was given to the
-development version of CVC4 and to Z3 4.3.2 for comparison, each running for
+development version of CVC4 (\textbf{which revision?}) and to
+Z3 4.3.2 (\textbf{really?}) for comparison, each running for
 up to 60~s on the StarExec cluster
@@
 reasoning. This is especially the case for problems generated using 256~lemmas.
+The problems with 16~lemmas put more stress on the decision procedure but are
+less typical of Sledgehammer.
+
 % when this makes sense.

### 2015-11-09 jour/jour.tex
edits on Sects. 1 and 2, mostly

@@
 \usepackage{cite}
+\usepackage{color}
 \usepackage{mathptmx}
@@
 \newcommand\ty[1]{\textit{#1}}
+
+\newcommand\medrightarrow{\mathrel{{{\color{black}\relbar}\kern-0.9ex\rlap{\color{white}\ensuremath{\blacksquare}}\kern-0.9ex}\joinrel{\color{black}\rightarrow}}}
 
@@
 \newcommand{\ec}[1]{[#1]}
-\newcommand{\JJJJ}{\mathcal{\!J\!}}
+\newcommand{\JJJJ}{\mathcal{\!I\!}}
 \newcommand{\J}{\mathcalx{J}}
@@
 
-\author {Andrew Reynolds \and Jasmin Christian Blanchette}
+\author {Andrew Reynolds \and Jasmin Christian Blanchette\thanks{In memoriam Morgan Deters 1979--2015}}
 \authorrunning {A. Reynolds \and J. C. Blanchette}
@@
 \maketitle
-
-\vskip3\smallskipamount
-
-\noindent\hfill\emph{\relax{In memoriam Morgan Deters 1979--2015}}
-
-\bigskip
 
@@
 (Section~\ref{sec:co-datatypes}). Intuitively, the
-values of a codatatype consist of all well-typed finite and infinite ground
+values of a codatatype consist of all well-typed finite \emph{and infinite} ground
 constructor
@@
 have emerged in a few automatic theorem provers as well. The SMT-LIB
-\cite{xxxxxxx}
-format, implemented by most SMT
-%(satisfiability modulo theories)
+\cite{barrett-et-al-2015}
+format, supported by most SMT
+(satisfiability modulo theories)
 solvers, has been extended with a syntax for datatypes.
@@
 For example, the type
-$\vthinspace\keyw{codatatype}~\;\ty{stream}_{\,\tau} \,=\allowbreak\, \const{SCons}(\tau,\allowbreak\: \ty{stream}_{\,\tau})\vthinspace$
+$$\vthinspace\keyw{codatatype}~\;\ty{stream}_{\,\tau} \,=\allowbreak\, \const{SCons}(\tau,\allowbreak\: \ty{stream}_{\,\tau})\vthinspace$$
 of infinite sequences or streams over an element type $\tau$ is
@@
 (solution soundness) on universal problems. %and the absence of spurious models.
-They cannot be expressed as finite axiomatizations, so they naturally belong in
+They cannot be finitely axiomatized, so they naturally belong in
 a decision procedure.
@@
 and for theory combinations. The $\mu$-binder notation associates a name with
-a subterm; it is used to represent cyclic values in the generated
-models and in the metatheory. For example,
-the $\mu$-term $\const{SCons}(1,\: \MU s.\; \const{SCons}(0,\: \const{SCons}(9,\: s)))$
+a subterm; it is used to represent cyclic values in the metatheory and
+in the generated models. For example,
+the $\mu$-term $\const{SCons}(1,\: \MU z.\; \const{SCons}(0,\: \const{SCons}(9,\: z)))$
 stands for the lasso-shaped sequence $1, 0, 9, 0, 9, 0, 9,
@@
 problems generated from Isabelle theories using the Sledgehammer tool \cite{blanchette-et-al-2013-smt}
-demonstrates the usefulness of the approach (Section~\ref{sec:experimental-results}).
+demonstrates the usefulness of the approach (Section~\ref{sec:evaluation-on-isabelle-problems}).
+
+An earlier version of this article was presented at the CADE-25 conference in
+Berlin, Germany. The article extends the conference paper with additional
+background material on (co)datatypes (Section~\ref{sec:co-datatypes}), more
+elaborated metatheoretical proofs (Section~\ref{sec:the-decision-procedure}),
+a description of the generation of models with $\mu$-binders in CVC4
+(Section~\ref{sec:implementation-as-a-theory-solver-in-cvc4}), and a more
+comprehensive evaluation (Section~\ref{sec:evaluation-on-isabelle-problems}).
 
@@
 Z3 \cite{de-moura-bjoerner-2008} in unpublished work by Leonardo de Moura and to a SPASS-like %prototypical
-superposition prover by Daniel Wand \cite{wand-2014}.
+superposition prover by Wand \cite{wand-2014}.
 Closely related are the automatic structural
 % cruanes-201x,wand-weidenbach-201x,  -- both were rejected
-induction in both kinds of provers \cite{kersani-peltier-2013,reynolds-kuncak-2015},
+induction in SMT solvers
+\cite{reynolds-kuncak-2014}
+and superposition provers\cite{kersani-peltier-2013,cruanes-2015},
 the (co)datatype and (co)induction support in Dafny \cite{leino-moskal-2014},
-and the (semi-)decision procedures for datatypes
+and the (\hbox{semi-})decision procedures for datatypes
 implemented in Leon \cite{suter-et-al-2011} and RADA \cite{pham-et-al-2013}.
@@
 Datatypes are supported by the higher-order model finder Refute
-\cite{weber-2008}. Its successor, Nitpick \cite{blanchette-nipkow-2010}, can
-also generate models involving cyclic codatatype values.
-Cyclic values have been studied extensively under the heading of
-regular or rational trees---see Carayol and Morvan
+\cite{weber-2008} for Isabelle. Its successor, Nitpick
+\cite{blanchette-nipkow-2010}, can also generate models involving cyclic
+codatatype values. Cyclic values have been studied extensively under the heading
+of regular or rational trees---see Carayol and Morvan
 \cite{carayol-morvan-2006} and Djellou et al.\ \cite{djelloul-et-al-2008}
@@
 
-{\looseness=-1
-We fix a signature $\Sig = (\Types, \Funcs)$. The types are partitioned into
+We fix a signature $\Sig = (\Types, \Funcs)$. The types are partitioned as
 $\Types = \Data \mathrel{\uplus} \Codata \mathrel{\uplus} \Nondata$, where $\Data$ are the
-\emph{datatypes}, $\Codata$ are the \emph{codatatypes}, and $\Nondata$ are the %remaining
+\emph{datatypes}, $\Codata$ are the \emph{codatatypes}, and $\Nondata$ are the remaining
 \emph{ordinary types} (which can be interpreted or not). The function symbols are
@@
 consider further function symbols because they can be abstracted away as
-variables when combining theories. $\mathrm\Sigma$-terms are standard first-order terms
+variables when combining theories.
+Exceptionally, it is convenient to use numeric constants ($0$, $1$, \ldots)\ in
+examples.
+$\mathrm\Sigma$-terms are standard first-order terms
 over $\mathrm\Sigma$,
@@
 %(Section~\ref{sec:implementation-as-a-theory-solver-in-cvc4}).
-%Exceptionally, it is convenient to use numeric constants ($0$, $1$, \ldots)\ in
-%examples.
-
-}
 
@@
 Types can be declared together in a mutual fashion even if they are not actually
-(co)recursive. The semantic notion is more precise and is the one that interests
-us. Non(co)recursive type specifications such~as
+mutually (co)recursive. The semantic notion is more precise and is the one that interests
+us.
+
+Non(co)recursive type specifications such~as
 \[\begin{aligned}[t]
@@
 
-A related semantic view of datatypes is as initial algebras. Codatatypes are
+A related semantic view of datatypes is as initial algebras of suitable
+functors. Codatatypes are
 then defined dually as final coalgebras \cite{rutten-2000}. The datatypes are
@@
   & %\forall x_1,\ldots,x_{n_{i\negvthinspace j}} y.\;\,
-    \smash{\const{C}_{i\negvthinspace j}(x_1,\ldots,x_{n_{i\negvthinspace j}}) \teq \const{C}_{i\negvthinspace j}(y_1,\ldots,y_{n_{i\negvthinspace j}}) \longrightarrow x_k \teq y_k}
+    \smash{\const{C}_{i\negvthinspace j}(x_1,\ldots,x_{n_{i\negvthinspace j}}) \teq \const{C}_{i\negvthinspace j}(y_1,\ldots,y_{n_{i\negvthinspace j}}) \medrightarrow x_k \teq y_k}
   \\[-0.5\jot]
@@
 %
-Datatypes are additionally characterized by an induction principle:
+Datatypes are additionally characterized by an induction principle for proving
+a conjunction of properties $P_1,\ldots,P_{\ell}$ over arbitrary
+values $v_1 : \delta_1, \ldots, v_\ell : \delta_\ell$:
 %
@@
 \text{Induction:}\quad
-\AXC{\strut$\bigwedge_{\,i,j}\, \forall x_1 \ldots \vthinspace x_{n_{i\negvthinspace j}}.\; \bigl(\bigwedge_{\,k}\, \mathit{IH}_{i\negvthinspace j}^k[x_k]\bigr) \longrightarrow P_i[\const C_{i\negvthinspace j}(x_1,\ldots,x_{n_{ij}})]$}
+\AXC{\strut$\bigwedge_{\,i,j}\, \forall x_1 \ldots \vthinspace x_{n_{i\negvthinspace j}}.\; \bigl(\bigwedge_{\,k}\, \mathcalx{IH}[x_k]\bigr) \medrightarrow P_i[\const C_{i\negvthinspace j}(x_1,\ldots,x_{n_{ij}})]$}
 \UIC{\strut$\bigwedge_{\,i}\, P_i[v_i]$}
@@
 \]
-where the induction hypothesis $\mathit{IH}_{i\negvthinspace j}^k(x)$
-denotes either $P_{\iPrime}(x)$ if there exists some $\iPrime$ such that
-the formula is type-correct or else $\top$.
-The principle ensures that the interpretation of datatypes is standard.
+The notation $\mathcalx{IH}[x]$
+denotes either $P_{\iPrime}[x]$ if there exists some $\iPrime$ such that
+the formula is type-correct or else $\top$ (truth).
+The induction principle ensures that the interpretation of datatypes is standard.
 For the natural numbers constructed from $\const{Z}$ and $\const{S}$,
@@
 
-For codatatypes, the dual notion is called coinduction. It depends on
-witnesses $R_i$ that are required to be bisimulations:
+For codatatypes, the dual notion is called coinduction. It makes it possible to
+derive the equality of pairs of codatatype values
+$v_1, w_1 : \delta_1,\,\ldots,\,v_\ell, w_\ell : \delta_\ell$, based on
+coinduction witnesses $R_1,\ldots,R_\ell$ that must be supplied upon use:
 %
@@
 \AXC{\strut$\begin{gathered}\textstyle \smash{\bigwedge_{\,i}\, R_i[v_i, w_i]} \\[-\jot]\textstyle
-\bigwedge_{\,i}\, \forall v\; w.\;\, R_i[v, w] \longrightarrow
-  \bigwedge_{\,j}\, \const{d}_{\negvthinspace j}(v) \teq \const{d}_{\negvthinspace j}(w)
-  \mathrel\land
-  \const{d}_{\negvthinspace j}(v) \longrightarrow \bigwedge_{\,k}\, \const{s}_{i\negvthinspace j}^{\,k}(v) \sim \const{s}_{i\negvthinspace j}^{\,k}(w)
+\bigwedge_{\,i}\, \forall v\; w.\;\, R_i[v, w] \medrightarrow
+  \bigwedge_{\,j}\, \const{d}_{\negvthinspace j}(v) \mathrel\land \const{d}_{\negvthinspace j}(w)
+  \medrightarrow \bigwedge_{\,k}\, \const{s}_{i\negvthinspace j}^{\,k}(v) \sim \const{s}_{i\negvthinspace j}^{\,k}(w)
 \end{gathered}$}
@@
 \]
-where $x \sim y$ denotes either $R_{\iPrime}[x, y]$ if there exists some
+The notation $x \sim y$ stands for $R_{\iPrime}[x, y]$ if there exists some
 $\iPrime$ such that the formula is type-correct or $x \teq y$ otherwise.
-This axiom encodes a form of extensionality:\ Two values that yield the same
-observations must be equal, where the observations are made through
-selectors and discriminators.
+The second premise ensures that the coinduction witnesses are
+bisimulations. The first premise and the conclusion capture the notion that
+equality is the largest bisimulation on codatatypes. Thus, the coinduction
+principle encodes a form of extensionality:\ Two values that yield the same
+observations must be equal, where the observations are made through selectors
+and discriminators.
 
@@
 
-{\looseness=-1
+For datatypes, any recursive specification gives rise to an infinite datatype.
+Paradoxically, this does not extend to codatatypes:
 Some codatatypes are so degenerate as to be finite even though they
 have infinite values. A simple example is
-\vthinspace$\keyw{codatatype}~\ty{a} = \const{A}(\ty{a})$, whose
-unique value is $\MU a.\; \const{A}(a)$. Other specimens are
+\vthinspace$\keyw{codatatype}~\ty{a} = \const{A}(\ty{a})$, which has a cardinality
+of one; its unique value is $\MU a.\; \const{A}(a)$. Other specimens are
 $\ty{stream}_{\,\ty{unit}\,}$
-and
+and both \ty{b} and \ty{c} in the specification
 \[\begin{aligned}[t]
@@
 \]
-where \ty{unit} is a datatype with the single constructor $\const{Unity} :
-\ty{unit}$. We call such types
+assuming \ty{unit} is a datatype with the single constructor $\const{Unity} :
+\ty{unit}$. We call such codatatypes
 \emph{corecursive singletons}. For the decision procedure, it will be
@@
 
-}
 % In other words, all finite corecursive codatatypes are corecursive singletons.
@@
 noncorecursive constructor $\const{E}$, it is possible to create terms of
-arbitrary depths---e.g., $\const{C}(\ldots(\const{C}(\const{E}))\ldots)$. In
-both cases, there can be no finite models.
+arbitrary depth---e.g., $\const{C}(\ldots(\const{C}(\const{E}))\ldots)$. In
+either case, there can be no finite models.
 
@@
 \section{Evaluation on Isabelle Problems}
-\label{sec:experimental-results}
+\label{sec:evaluation-on-isabelle-problems}
 

### 2015-11-09 jour/jour.tex
more edits

@@
+%%% TODO: Put http://lara.epfl.ch/~reynolds/JAR-CADE2015-cdt/ online
 %%% TODO: Mealy machine minimization
@@
 \newcommand{\ec}[1]{[#1]}
-\newcommand{\JJJJ}{\mathcal{\!I\!}}
+\newcommand{\JJJJ}{\mathcal{\!J\!}}
 \newcommand{\J}{\mathcalx{J}}
@@
 
-\global\def\figurename{Figure}
-\global\def\figuresname{Figures}
-
 \DeclareSymbolFont{letters}{OML}{txmi}{m}{it}
@@
 implemented in CVC4, a state-of-the-art SMT solver. An evaluation based on
-problems generated from theories developed with Isabelle demonstrates the
+problems generated from formalizations developed with Isabelle demonstrates the
 potential of the procedure.
@@
 Third, \emph{it must be possible to express cyclic }(\emph{regular}) \emph{values as closed terms and
-to enumerate them.} This is necessary both for finite model finding \cite{reynolds-et-al-2013}
-%modulo theories
+to enumerate them.} This is necessary both for finite model finding
+modulo theories \cite{reynolds-et-al-2013,ge-de-moura-2009}
 and for theory combinations. The $\mu$-binder notation associates a name with
@@
 An evaluation on %hand-crafted examples and on
-problems generated from Isabelle theories using the Sledgehammer tool \cite{blanchette-et-al-2013-smt}
+problems generated from Isabelle/HOL \cite{nipkow-et-al-2002}
+formalizations using the Sledgehammer tool \cite{blanchette-et-al-2013-smt}
 demonstrates the usefulness of the approach (Section~\ref{sec:evaluation-on-isabelle-problems}).
@@
 The calculus consists of three sets of rules, given in
-\figuresname~\ref{fig:cc-rules} to \ref{fig:split-rule}, corresponding to three
+Figs.\ \ref{fig:cc-rules}~to~\ref{fig:split-rule}, corresponding to three
 phases. The first phase computes the bidirectional closure of $\Ec$. The second
@@
 
-\paragraph{Phase 1: Computing the Bidirectional Closure \rm(\figurename~\ref{fig:cc-rules}).}
+\paragraph{Phase 1: Computing the Bidirectional Closure \rm(Fig.~\ref{fig:cc-rules}).}
 In conjunction with \rn{Refl}, \rn{Sym}, and \rn{Trans}, the \rn{Cong} rule computes the congruence (upward) closure,
@@
 
-\paragraph{Phase 2: Applying Acyclicity and Uniqueness \rm(\figurename~\ref{fig:ab-rules}).}
+\paragraph{Phase 2: Applying Acyclicity and Uniqueness \rm(Fig.~\ref{fig:ab-rules}).}
 %For presentation of the rules in this phase, we rely on a representation of terms in the $\mu$-notation.
@@
 %
-With each equivalence class $\ec{u\typ{\tau}}$, we associate a fresh variable $\Varec{u}\typ{\tau}$
-%of the same type as $u$
+With each equivalence class $\ec{u}$, we associate a fresh variable $\Varec{u}$
+of the same type as $u$
 %not occurring in $\tEc$,
-and set $\Val\ec{u} := {\Varec{u}}$, that is to say there are initially no
+and set $\Val\ec{u} := {\Varec{u}}$. Because $\Varec{u}$ is unspecified,
+this indicates that there are initially no
 constraints on the values for any equivalence class $\ec{u}$. The mapping $\Val$
@@
 whose values $\Val \ec{t} = \MU x.\; u$ are cyclic
-(as expressed by $x \in \FV( u )$),
+(expressed by $x \in \FV( u )$),
 then $\Ec$ is $\thD$-unsatisfiable.
@@
 
-\paragraph{Phase 3: Branching \rm(\figurename~\ref{fig:split-rule}).}
+\paragraph{Phase 3: Branching \rm(Fig.~\ref{fig:split-rule}).}
 If a selector is applied to a term $t$, or if $t$'s type is a finite datatype,
@@
 \label{lem:mu-norm-arg}
-If the $\mu$-terms $\vec u$ are in normal form and $t = \nf{\MU x.\; \const{C}( \vec u )} = \MU x.\; \const{C}( \vec w )$,
+If the $\mu$-terms $\vec u$ are in normal form and $t = \nf{\MU x.\; \const{C}( \vec u )} = \MU x.\; \const{C}( \vec w )$
+for some $\mu$-terms $\vec w$,
 then $\vec u \vsim \expand{t}{\vec w}$\vthinspace.
@@
 \rm
-An interpretation $\J$ is \emph{normal} if these conditions are met:
+An interpretation $\J$ is \emph{normal} if the following conditions are met:
 \begin{enumerate}
@@
 models must contain all possible infinite terms, not only cyclic ones. However,
-acyclic infinite values are uninteresting to us, and for quantifier-free
-formulas it is trivial to extend any normal interpretation with extra
+acyclic infinite values are not interesting for deciding universal formulas:
+For such formulas, it is trivial to extend any normal interpretation with extra
 domain elements to obtain a genuine model if desired. %Hence, we focus on normal interpretations.
-
-\pagebreak[2]
 
@@
 of $\nf{\Val_1 \sigma} \ec{y}$ and $\nf{\Val_1 \sigma} \ec{z}$
-are $\alpha$-disequivalent at position $p$.
+are $\alpha$-disequivalent at position~$p$.
 Since $t_1$ and $t_2$ are minimal,
@@
 or else the expansion of $\nf{\Val_1 \sigma} \ec{y}$ and $\nf{\Val_1 \sigma} \ec{z}$
-are $\alpha$-disequivalent at position $p$ since they do not contain the same free variables.
+are $\alpha$-disequivalent at position~$p$ since they do not contain the same free variables.
 Since $\expand{\Val_1 \ec{y}}{t_1}$ is normal,
@@
 which implies that the expansion of $\nf{\Val_1 \sigma} \ec{y}$ and $\nf{\Val_1 \sigma} \ec{z}$
-are $\alpha$-disequivalent at position $p$. %\rem{TODO: more detail}
+are $\alpha$-disequivalent at position~$p$. %\rem{TODO: more detail}
 
@@
 
-Intuitively, this lemma states three properties of $\ValC$ that ensure a normal
-interpretation $\J$ can be constructed that satisfies $\Ec$. Property~(1) states that the
-values in the range of $\ValC$ are $\alpha$-equivalent to a value in normal
-interpretation. This means they are closed, normal, and acyclic when
-required. Property~(2) states that the interpretation of all subterms in the
-range of $\ValC$ depends on its associated variable only. In other words, the
-interpretation of a subterm $t$ where $\muvar(t) = {\Varec{x}}$ is equal to
-$\ValC \ec{x}$, independently of the context. % in which it is interpreted.
+Intuitively, this lemma states three properties of $\ValC$ that together ensure that
+a normal interpretation~$\J$ can be constructed that satisfies $\Ec$.
+Property~(1) states that the values in the range of $\ValC$ are
+$\alpha$-equivalent to a value in normal interpretation. This means they are
+closed, normal, and acyclic when required. Property~(2) states that the
+interpretation of all subterms in the range of $\ValC$ depends on its associated
+variable only. In other words, the interpretation of a subterm $t$ where
+$\muvar(t) = {\Varec{x}}$ is equal to
+$\ValC \ec{x}$, independently of the context.
 Property~(3) states that $\ValC$ is injective (modulo
@@
 % of %the mapping $\Val$. % from phase~2.
-For the variables, let $\interp{\J}{x\typ{\tau}}$ be the value in $\J(\tau)$ that is $\alpha$-equivalent with $\ValC \ec{x}$ for each variable $x \in \Terms(\Fc)$,
+For the variables, let $\interp{\J}{x\typ{\tau}}$ be the value in $\J(\tau)$ that is $\alpha$-equivalent
+with $\ValC \ec{x}$ for each variable $x \in \Terms(\Fc)$,
 which by Lemma~\ref{lem:interpretation-completion}(1) is guaranteed to exist.
@@
 If $t$ is a constructor term of the form $\const{C}( u_1, \ldots, u_n )$,
-then $\interp{\J}{t}$ is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \interp{\J}{u_1}, \ldots, \interp{\J}{u_n} ) }$ for some fresh $x$,
+then $\interp{\J}{t}$ is $\alpha$-equivalent with $\nf{\MU x.\allowbreak\; \const{C}( \interp{\J}{u_1},\allowbreak \ldots,\allowbreak \interp{\J}{u_n} ) }$ for some fresh $x$,
 which by the induction hypothesis is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \ValC \ec{ u_1 }, \ldots, \ValC \ec{ u_n } )}$.
@@
 where
-$\muvar(w_i) = {\Varec{u_i}}$ for each $i$.  Thus by Lemma~\ref{lem:interpretation-completion}(2),
+$\muvar(w_i) = {\Varec{u_i}}$ for each~$i$.  Thus by Lemma~\ref{lem:interpretation-completion}(2),
 $\expand{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$.
@@
 $\muvar(w_i) = \smash{\Varec{\vphantom{I}\smash{\const s^i_{\negvthinspace\jPrime}( u )}}}$
-for each $i$, and thus by Lemma~\ref{lem:interpretation-completion}(2),
+for each~$i$, and thus by Lemma~\ref{lem:interpretation-completion}(2),
 $\expand{\ValC \ec{u}}{ w_i } = \ValC \ec{ \smash{\const s^i_{\negvthinspace\jPrime}( u )}}$.
@@
 By Theorems~\ref{thm:t},~\ref{thm:rs}, and~\ref{thm:ss}, the
-calculus is sound and complete for the universal theory of (co)datatypes. We can
+calculus is sound and complete for the universal theory of (co)datatypes. We may
 rightly call it a decision procedure for that theory.
@@
 
-While the decision procedure is restricted to universal conjectures, in practice
-we often want to solve problems that feature universal axioms and existential
+While the decision procedure is restricted to universal conjectures, users often
+want to solve problems that feature universal axioms and existential
 conjectures.
@@
 %presence of universally quantified formulas.
-Many SMT solvers, including CVC4, can reason about quantified formulas using incomplete instantiation-based methods~\cite{MouraBjoerner07,ReynoldsTinelliMoura14}.
-These methods extend naturally to %quantified
+Many SMT solvers, including CVC4, can reason about quantified formulas using
+incomplete instantiation-based
+techniques~\cite{MouraBjoerner07,ReynoldsTinelliMoura14}.
+These techniques extend naturally to %quantified
 formulas involving datatypes and codatatypes.
 
-However, the presence of quantified formulas poses an additional challenge in
+However, the presence of quantifiers poses an additional challenge in
 the context of (co)datatypes. Quantified formulas may entail an upper bound on
@@
 since %by assumption
-the solver is already incomplete in the presence of quantified formulas.
+the solver is already incomplete for quantified formulas.
 Nonetheless, two techniques help increase the precision of the solver.
@@
 
-As Barrett et al.\ \cite{barrett-et-al-2007} observed for their procedure,
+% JB to AJR: I'd rather keep the para, based on feedback by Pascal.
+
+As Barrett et al.\ %\cite{barrett-et-al-2007}
+observed for their procedure,
 it is both theoretically and
@@
 
-  * decision procedure useful both for proving and for model finding
-  * model finding: critical, without it unsound models
-  * contribution to proving is less clear
-    * some parts can be axiomatized
-    * only acyclicity and uniqueness, but how often do they pop up in practice?
-
-To evaluate the decision procedure, we generated benchmark problems from existing
-Isabelle formalizations using Sledgehammer \cite{blanchette-et-al-2013-smt}.
+The decision procedure for (co)datatypes is useful both for proving (via
+negation, in the refutational style) and finite model finding.
+It is in fact vital for finite model finding,
+because the acyclicity and uniqueness rules are necessary for solution
+soundness, without which the generated models would often be spurious. The
+contributions to proving are less clear; they depend on how often the need for
+acyclicity or uniqueness arises in practice.
+
+To evaluate the decision procedure for proving conjectures,
+we generated benchmark problems from existing interactive
+proof goals arising in existing
+Isabelle formalizations, using Sledgehammer \cite{blanchette-et-al-2013-smt}
+as translator from Isabelle to SMT-LIB.
 %There are not yet many benchmarks available, since codatatypes are a
@@
 We
-included all the theory files from the Isabelle distribution (Distro, 879~goals)
-and the \emph{Archive of Formal Proofs} (AFP, 2974~goals) \cite{klein-et-al-afp}
+included all the formalizations from the Isabelle distribution (Distro, 1179~goals)
+and the \emph{Archive of Formal Proofs} (AFP, 3014~goals) \cite{klein-et-al-afp}
 that define codatatypes falling within the supported fragment. We
-added unpublished theories about Bird and Stern--Brocot trees
+added unpublished formalizations about Bird and Stern--Brocot trees
 by Peter Gammie and Andreas Lochbihler (\gandl,
-317~goals). To exercise the datatype support, theories about lists and trees
-were added to the first two benchmark sets. The theories were selected before
+265~goals). To exercise the datatype support, formalizations about lists and trees
+were added to the first two benchmark sets. The formalizations were selected before
 conducting any experiments. The experimental data are available
@@
 
-For each goal in each theory, we used Sledgehammer to select either 16 or
+For each goal in each formalization, we used Sledgehammer to select either 16 or
 256~lemmas, which were monomorphized and translated to SMT-LIB~2 along with the
 goal. The resulting problem was given to the
-development version of CVC4 (\textbf{which revision?}) and to
-Z3 4.3.2 (\textbf{really?}) for comparison, each running for
+development version of CVC4 (\textbf{TODO: which revision?}) and to
+Z3 4.3.2 (\textbf{TODO: really?}) for comparison, each running for
 up to 60~s on the StarExec cluster
@@
 Problems not involving any (co)datatypes were left out.
-Due to the lack of machinery for reconstructing inferences about (co)datatypes
-in Isabelle, the solvers are trusted as oracles. % in these experiments.
+Due to the lack of machinery in Isabelle for parsing CVC4 proofs and
+reconstructing inferences about (co)datatypes, the solvers are trusted
+as oracles. % in these experiments.
 %Z3 is included for comparison.
 
-The development version of CVC4 was run on each problem several times, with the
+CVC4 was run on each problem several times, with the
 support for datatypes and codatatypes either enabled or disabled. The
@@
 basic lemmas about constructors and selectors, allowing some (co)datatype
-reasoning. This is especially the case for problems generated using 256~lemmas.
+reasoning. This is especially true for problems generated using 256~lemmas.
 The problems with 16~lemmas put more stress on the decision procedure but are
-less typical of Sledgehammer.
+less typical of Sledgehammer-generated problems.
 
@@
 \caption{\captionii}
-\label{tab:benchtwofittysix}
+\label{tab:benchtwofiftysix}
 \end{table*}
@@
 
-The results are summarized in Table~\ref{tab:benchsixteen}.
+The results are summarized in Tables \ref{tab:benchsixteen} and
+\ref{tab:benchtwofiftysix}.
 %The goals in Table \ref{tab:benchsixteen} include 16
@@
 suggest that the theoretically stronger decision procedures almost completely
-subsume the weaker ones in practice: We encountered only one goal (out of 4170)
+subsume the weaker ones in practice. We encountered only one
+(\textbf{TODO: FIXME}) goal (out of $2 \times 4458$)
 that was solved by a configuration of CVC4 and unsolved by a configuration of
 CVC4 with more features enabled.
-
+%
 Moreover, every aspect of the procedure, including the
@@
 %
-Three proofs were found thanks to
-the acyclicity rule and four required uniqueness. Among the latter,
-three are simple proofs of the form
+Four proofs were found thanks to
+the acyclicity rule and 17~required uniqueness. Among the latter,
+some were simple proofs of the form
 \keyw{by}~\textit{coinduction}~\textit{auto}
-in Isabelle \cite{blanchette-et-al-2014-impl}.
-The fourth proof, by Gammie and Lochbihler, is more elaborate:
+in Isabelle \cite{blanchette-et-al-2014-impl}, while others
+required more elaborate arguments---for example:
 %
@@
 We introduced a decision procedure for the universal theory of datatypes and
-codatatypes. Our main
+codatatypes. Our principal
 contribution has been the support for codatatypes. Both the metatheory and
@@
 with high-level features and at reducing the gap between automatic and
-interactive theorem proving. As future work, it would be useful to implement
-proof reconstruction for (co)data\-type inferences in Isabelle.
-CVC4's finite model finding capabilities \cite{reynolds-et-al-2013} could also
-be interfaced for generating counterexamples in proof assistants;
+interactive theorem proving.
+We are currently interfacing CVC4's finite model finding capabilities
+for generating counterexamples in proof assistants
+\cite{reynolds-et-al-2015-smt};
 %as an alternative to Nitpick;
 in this context, the acyclicity and uniqueness rules are
-crucial to exclude spurious countermodels. Finally, it might be worthwhile to
+crucial to exclude spurious countermodels.
+As future work, it would be useful to implement
+proof reconstruction for (co)data\-type inferences in Isabelle.
+Finally, it might be worthwhile to
 extend SMT solvers with dedicated support for (co)recursion.
@@
 Summerfield, Dmitriy Traytel, and the anonymous reviewers suggested many
-textual improvements. The second author's research was partially supported by
+textual improvements. The second author's work was partially supported by
 the Deutsche

### 2015-11-09 jour/jour.tex
Starting work on extending sec 4.

@@
 
-\section{Implementation as a Theory Solver in CVC4}
+\section{Implementation in CVC4}
 \label{sec:implementation-as-a-theory-solver-in-cvc4}
@@
 %%% disagree.
-
+\subsection{Theory Solver for $\thD$}
 The decision procedure is implemented as a \relax{theory solver} of CVC4, that is, a
@@
 and $v$, it deduces that \rn{Unique} applies to $u$ and~$v$.
-
-While the decision procedure is restricted to universal conjectures, users often
-want to solve problems that feature universal axioms and existential
-conjectures.
-%%% @ANDY: not sure how to phrase this now
-%where checking satisfiability requires reasoning about ground literals in the
-%presence of universally quantified formulas.
-Many SMT solvers, including CVC4, can reason about quantified formulas using
-incomplete instantiation-based
-techniques~\cite{MouraBjoerner07,ReynoldsTinelliMoura14}.
-These techniques extend naturally to %quantified
-formulas involving datatypes and codatatypes.
-
-However, the presence of quantifiers poses an additional challenge in
-the context of (co)datatypes. Quantified formulas may entail an upper bound on
-the cardinality of an uninterpreted type $\ty{u}$.
-When assuming that $\ty{u}$ has infinite cardinality, the calculus
-presented in Section~\ref{sec:the-decision-procedure} is \relax{incomplete}
-since it may fail to recognize cases where \rn{Split} and \rn{Single} should be
-applied.
-This does not %necessarily
-impact the correctness of the procedure in this setting,
-since %by assumption
-the solver is already incomplete for quantified formulas.
-Nonetheless, two techniques help increase the precision of the solver.
-First, we can apply \rn{Split} to datatype terms whose cardinality depends on the finiteness of
-uninterpreted types.
-Second, we can conditionally apply \rn{Single} to codatatype terms that may have cardinality one.
-For example, the $\ty{stream}_{\,\ty{u}}$ codatatype %, where $\ty{u}$ is uninterpreted.
-has cardinality one precisely when $\ty{u}$ has cardinality one.
-If there exist two equivalence classes $\ec{s}$ and $\ec{t}$ for this type,
-the implementation adds the clause %of the form
-$(\exists x\;y\typ{\ty{u}}.\;\, x \not\teq y ) \mathrel\vee s \teq t$ to $F\!$.
-In other words, either the cardinality of $\ty{u}$ is greater than one or $s$ is equal to $t$.
 
@@
 phase~2.
+
+\begin{figure}[t!]
+\normalsize
+\centering
+\begin{enumerate}
+\item Let $\mathcal{U}$ be a partition of the equivalence classes of $\Ec$ such that $\ec{x} =_{\mathcal{U}} \ec{y}$ if and only if:
+\begin{itemize}
+\item $\const{C}( x_1, \ldots, x_k ) \in \ec{x}$ and $\const{C}( y_1, \ldots, y_k ) \in \ec{y}$.
+\end{itemize}
+\item Repeat until $\mathcal{U}$ is unchanged
+\begin{enumerate}
+\item $\mathcal{U}' := \emptyset$
+\item for each $U_i \in \mathcal{U}$:
+\begin{itemize}
+\item[\ ] Let $\mathcal{U}_i$ be a partition of $U_i$ such that $\ec{x} =_{\mathcal{U}_i} \ec{y}$ for distinct $\ec{x}, \ec{y}$ if and only if:
+\begin{itemize}
+\item $\const{C}( x_1, \ldots, x_k ) \in \ec{x}$ and $\const{C}( y_1, \ldots, y_k ) \in \ec{y}$, and
+\item $\ec{x_j} =_{\mathcal{U}} \ec{y_j}$ for each $j = 1, \ldots, k$.
+\end{itemize}
+\item[\ ] $\mathcal{U}' := \mathcal{U}' \cup \mathcal{U}_i$.
+\end{itemize}
+\item $\mathcal{U} := \mathcal{U}'$
+\end{enumerate}
+\item If distinct $t\typ{\delta}, s\typ{\delta} \in U \in \mathcal{U}$ where $\delta \in \Codata$, apply \rn{Unique} to $t,s$.
+\end{enumerate}
+\vspace*{-2pt} %% TYPESETTING HACK
+\caption{\,Algorithm for checking Uniqueness.
+}
+\label{fig:cc-rules}
+\end{figure}
+
+\subsection{Model Construction}
+
+[TODO]
+
+\subsection{Extension to Quantified Formulas}
+While the decision procedure is restricted to universal conjectures, users often
+want to solve problems that feature universal axioms and existential
+conjectures.
+%%% @ANDY: not sure how to phrase this now
+%where checking satisfiability requires reasoning about ground literals in the
+%presence of universally quantified formulas.
+Many SMT solvers, including CVC4, can reason about quantified formulas using incomplete 
+instantiation-based 
+techniques~\cite{MouraBjoerner07,ReynoldsTinelliMoura14}.
+These techniques extend naturally to %quantified
+formulas involving datatypes and codatatypes.
+
+However, the presence of quantifiers poses an additional challenge in
+the context of (co)datatypes. Quantified formulas may entail an upper bound on
+the cardinality of an uninterpreted type $\ty{u}$.
+When assuming that $\ty{u}$ has infinite cardinality, the calculus
+presented in Section~\ref{sec:the-decision-procedure} is \relax{incomplete}
+since it may fail to recognize cases where \rn{Split} and \rn{Single} should be
+applied.
+This does not %necessarily
+impact the correctness of the procedure in this setting,
+since %by assumption
+the solver is already incomplete for quantified formulas.
+Nonetheless, two techniques help increase the precision of the solver.
+First, we can apply \rn{Split} to datatype terms whose cardinality depends on the finiteness of
+uninterpreted types.
+Second, we can conditionally apply \rn{Single} to codatatype terms that may have cardinality one.
+For example, the $\ty{stream}_{\,\ty{u}}$ codatatype %, where $\ty{u}$ is uninterpreted.
+has cardinality one precisely when $\ty{u}$ has cardinality one.
+If there exist two equivalence classes $\ec{s}$ and $\ec{t}$ for this type,
+the implementation adds the clause %of the form
+$(\exists x\;y\typ{\ty{u}}.\;\, x \not\teq y ) \mathrel\vee s \teq t$ to $F\!$.
+In other words, either the cardinality of $\ty{u}$ is greater than one or $s$ is equal to $t$.
 

### 2015-11-09 jour/jour.tex
More updates to algo

@@
 \begin{enumerate}
-\item Let $\mathcal{U}$ be a partition of the equivalence classes of $\Ec$ such that $\ec{x} =_{\mathcal{U}} \ec{y}$ if and only if:
+\item Let $\mathcal{U}$ be a partition of the codatatype equivalence classes of $\Ec$ such that $\ec{t} =_{\mathcal{U}} \ec{u}$ for distinct $\ec{t}, \ec{u}$ if and only if:
 \begin{itemize}
-\item $\const{C}( x_1, \ldots, x_k ) \in \ec{x}$ and $\const{C}( y_1, \ldots, y_k ) \in \ec{y}$.
+\item $\const{C}( t_1, \ldots, t_k ) \in \ec{t}$ and $\const{C}( u_1, \ldots, u_k ) \in \ec{u}$.
 \end{itemize}
@@
 \begin{itemize}
-\item[\ ] Let $\mathcal{U}_i$ be a partition of $U_i$ such that $\ec{x} =_{\mathcal{U}_i} \ec{y}$ for distinct $\ec{x}, \ec{y}$ if and only if:
+\item[\ ] Let $\mathcal{U}_i$ be a partition of $U_i$ such that $\ec{t} =_{\mathcal{U}_i} \ec{u}$ for distinct $\ec{t}, \ec{u}$ if and only if:
 \begin{itemize}
-\item $\const{C}( x_1, \ldots, x_k ) \in \ec{x}$ and $\const{C}( y_1, \ldots, y_k ) \in \ec{y}$, and
-\item $\ec{x_j} =_{\mathcal{U}} \ec{y_j}$ for each $j = 1, \ldots, k$.
+\item $\const{C}( t_1, \ldots, t_k ) \in \ec{t}$ and $\const{C}( u_1, \ldots, u_k ) \in \ec{u}$, and
+\item $\ec{t_j} =_{\mathcal{U}} \ec{u_j}$ for each $j = 1, \ldots, k$.
 \end{itemize}
@@
 \end{enumerate}
-\item If distinct $t\typ{\delta}, s\typ{\delta} \in U \in \mathcal{U}$ where $\delta \in \Codata$, apply \rn{Unique} to $t,s$.
+\item If distinct $\ec{t}, \ec{u} \in U$ for some $U \in \mathcal{U}$, apply \rn{Unique} to $t,u$.
 \end{enumerate}
@@
 }
-\label{fig:cc-rules}
+\label{fig:uniqueness-algo}
+\end{figure}
+
+\begin{figure}[t!]
+\normalsize
+\centering
+\begin{enumerate}
+\item Let $U$ be the datatype equivalence classes of $\Ec$.
+\item Repeat until $U \neq \emptyset$.
+\begin{itemize}
+\item[\ ] For some $\ec{t} \in U$, $\mathrm{traverse}( \ec{t}, U, \emptyset )$
+\end{itemize}
+\end{enumerate}
+
+\flushleft $\mathrm{traverse}( \ec{t}, U, P )$:
+\begin{enumerate}
+\item If $\ec{t} \in P$, apply \rn{Acyclic} to $t$.
+\item If $\ec{t} \in U$
+\begin{enumerate}
+\item If $\const{C}( t_1, \ldots, t_k ) \in \ec{t}$, then for each $j = 1, \ldots k$,
+\begin{itemize}
+\item[\ ] $\mathrm{traverse}( \ec{t_j}, U, P \cup \{ \ec{t} \} )$
+\end{itemize}
+\item $U := U \textbackslash \{ \ec{t} \}$.
+\end{enumerate}
+\end{enumerate}
+\vspace*{-2pt} %% TYPESETTING HACK
+\caption{\,Algorithm for checking Acyclic.
+}
+\label{fig:acyclic-algo}
 \end{figure}

### 2015-11-09 jour/jour.tex
Minor

@@
 \begin{enumerate}
-\item Let $\mathcal{U}$ be a partition of the codatatype equivalence classes of $\Ec$ such that $\ec{t} =_{\mathcal{U}} \ec{u}$ for distinct $\ec{t}, \ec{u}$ if and only if:
+\item Let $\mathcal{U}$ be a partition of the equivalence classes of $\Ec$ such that $\ec{t} =_{\mathcal{U}} \ec{u}$ for distinct $\ec{t}, \ec{u}$ if and only if:
 \begin{itemize}
-\item $\const{C}( t_1, \ldots, t_k ) \in \ec{t}$ and $\const{C}( u_1, \ldots, u_k ) \in \ec{u}$.
+\item $t\typ{\delta}$, $u\typ{\delta}$ where $\delta \in \Codata$, and $\const{C}( t_1, \ldots, t_k ) \in \ec{t}$ and $\const{C}( u_1, \ldots, u_k ) \in \ec{u}$.
 \end{itemize}
@@
 \item If $\ec{t} \in P$, apply \rn{Acyclic} to $t$.
-\item If $\ec{t} \in U$
+\item If $\ec{t} \in U$,
 \begin{enumerate}

### 2015-11-09 jour/jour.tex
minor
tuning
more
Rearrange, describe figures.

@@
 
+\begin{figure}[t!]
+\normalsize
+\centering
+\begin{enumerate}
+\item Let $U$ be the datatype equivalence classes of $\Ec$.
+\item Repeat until $U \neq \emptyset$.
+\begin{itemize}
+\item[\ ] For some $\ec{t} \in U$, $\mathrm{traverse}( \ec{t}, U, \emptyset )$
+\end{itemize}
+\end{enumerate}
+
+\flushleft $\mathrm{traverse}( \ec{t}, U, P )$:
+\begin{enumerate}
+\item If $\ec{t} \in P$, apply \rn{Acyclic} to $t$.
+\item If $\ec{t} \in U$,
+\begin{enumerate}
+\item If $\const{C}( t_1, \ldots, t_k ) \in \ec{t}$, then for each $j = 1, \ldots k$,
+\begin{itemize}
+\item[\ ] $\mathrm{traverse}( \ec{t_j}, U, P \cup \{ \ec{t} \} )$
+\end{itemize}
+\item $U := U \textbackslash \{ \ec{t} \}$.
+\end{enumerate}
+\end{enumerate}
+\vspace*{-2pt} %% TYPESETTING HACK
+\caption{\,Algorithm for checking \rn{Acyclic}}
+\label{fig:acyclic-algo}
+\end{figure}
+
+\begin{figure}[t!]
+\normalsize
+\centering
+\begin{enumerate}
+\item Let $\mathcal{U}$ be a partition of the equivalence classes of $\Ec$ such that $\ec{t} =_{\mathcal{U}} \ec{u}$ for distinct $\ec{t}, \ec{u}$ if and only if:
+\begin{itemize}
+\item $t\typ{\delta}$, $u\typ{\delta}$ where $\delta \in \Codata$, and $\const{C}( t_1, \ldots, t_k ) \in \ec{t}$ and $\const{C}( u_1, \ldots, u_k ) \in \ec{u}$.
+\end{itemize}
+\item Repeat until $\mathcal{U}$ is unchanged
+\begin{enumerate}
+\item $\mathcal{U}' := \emptyset$
+\item for each $U_i \in \mathcal{U}$:
+\begin{itemize}
+\item[\ ] Let $\mathcal{U}_i$ be a partition of $U_i$ such that $\ec{t} =_{\mathcal{U}_i} \ec{u}$ for distinct $\ec{t}, \ec{u}$ if and only if:
+\begin{itemize}
+\item $\const{C}( t_1, \ldots, t_k ) \in \ec{t}$ and $\const{C}( u_1, \ldots, u_k ) \in \ec{u}$, and
+\item $\ec{t_j} =_{\mathcal{U}} \ec{u_j}$ for each $j = 1, \ldots, k$.
+\end{itemize}
+\item[\ ] $\mathcal{U}' := \mathcal{U}' \cup \mathcal{U}_i$.
+\end{itemize}
+\item $\mathcal{U} := \mathcal{U}'$
+\end{enumerate}
+\item If distinct $\ec{t}, \ec{u} \in U$ for some $U \in \mathcal{U}$, apply \rn{Unique} to $t,u$.
+\end{enumerate}
+\vspace*{-2pt} %% TYPESETTING HACK
+\caption{\,Algorithm for checking \rn{Unique}}
+\label{fig:uniqueness-algo}
+\end{figure}
+
 Before constructing a model for $F\!$, the theory solver constructs neither $\mu$-terms nor the mapping $\Val$.
-Instead, $\Val$ is computed implicitly by traversing the equivalence classes of $\Ec$ during phase~2.
-To detect whether \rn{Acyclic} applies,
-the procedure considers each equivalence class $\ec{t}$ containing a datatype constructor $\const{C}( t_1, \ldots, t_n )$.
-It visits $\ec{t_1}, \ldots, \ec{t_n}$ and all constructor arguments in these equivalence classes recursively.
-If while doing so it returns to $\ec{t}$, it deduces that \rn{Acyclic} applies. % to $t$.
-To recognize when the precondition of \rn{Unique} holds,
-the procedure considers the set $\SSS$ of all codatatype equivalence classes. It
-simultaneously visits the equivalence classes of arguments of constructor terms in each equivalence class in $\SSS$,
-while partitioning $\SSS$ into $\SSS_1, \ldots, \SSS_n$ based on the top-most symbol of constructor terms in these equivalence classes
-and the equivalence of their arguments of ordinary types.
-It then partitions each set %in this partition %in the same manner
-recursively.
-If %after doing so,
-the resulting partition contains a set $\SSS_i$ containing distinct terms $u$
-and $v$, it deduces that \rn{Unique} applies to $u$ and~$v$.
+Instead, we use the algorithms in Figures~\ref{fig:acyclic-algo} and~\ref{fig:uniqueness-algo}
+for determining whether the rules \rn{Acyclic} and \rn{Unique} apply to our set $\Ec$.
+
+For \rn{Acyclic}, Figure~\ref{fig:acyclic-algo} considers the set $U$ of all datatype equivalence classes of $\Ec$.
+We choose an arbitrary $\ec{t}$ in $U$ and call the recursive subprocedure $\mathrm{traverse}$,
+which takes as input the current equivalence class we are processing, the set of equivalence classes $U$ we have yet to process,
+and the set of equivalence classes $P$ we are currently processing.
+If a call to $\mathrm{traverse}( \ec{t}, U, P )$ is such that $\ec{t} \in P$,
+we know that the rule \rn{Acyclic} applies to $t$.
+Otherwise, if we have yet to process $\ec{t}$, as indicated by the condition $\ec{t} \in U$,
+then if $\ec{t}$ contains a constructor term $\const{C}( t_1, \ldots, t_k )$, 
+then we recursively call $\mathrm{traverse}$ on each of $\ec{t_1}, \ldots, \ec{t_n}$.
+If this succeeds, we remove $\ec{t}$ from $U$.
+
+For \rn{Unique}, Figure~\ref{fig:uniqueness-algo} considers a partition $\mathcal{U}$ of the equivalence classes of our set $\Ec$
+such that equivalence classes having a constructor term with top symbol $\const{C}$ are placed in the same subset, for each constructor $\const{C}$.
+We write $\ec{t} =_{\mathcal{U}} \ec{u}$ to denote that $\ec{t}$ and $\ec{u}$ reside in the same subset within $\mathcal{U}$.
+We then refine $\mathcal{U}$ by 
+constructing a partition $\mathcal{U}_i$ of each $U_i \in \mathcal{U}$ such that distinct $\ec{t}$ and $\ec{u}$ reside in the same subset of $\mathcal{U}_i$ 
+if and only if they contain constructor terms $\const{C}( t_1, \ldots, t_k )$ and $\const{C}( u_1, \ldots, u_k )$
+where each $t_j, u_j$ are such that $\ec{t_j}$ and $\ec{u_j}$ reside in the same subset of $\mathcal{U}$ for each $j=1, \ldots, k$.
+Let $\mathcal{U'}$ be the union of each of these partitions.
+We update $\mathcal{U}$ to $\mathcal{U'}$.
+This refinement is repeated until $\mathcal{U}$ is left unchanged.
+If the resulting partition $\mathcal{U}$ contains a subset $U$ having two distinct equivalence classes $\ec{t}$ and $\ec{u}$, then \rn{Unique} applies to $t$ and $u$.
+
+%Instead, $\Val$ is computed implicitly by traversing the equivalence classes of $\Ec$ during phase~2.
+%To detect whether \rn{Acyclic} applies,
+%the procedure considers each equivalence class $\ec{t}$ containing a datatype constructor $\const{C}( t_1, \ldots, t_n )$.
+%It visits $\ec{t_1}, \ldots, \ec{t_n}$ and all constructor arguments in these equivalence classes recursively.
+%If while doing so it returns to $\ec{t}$, it deduces that \rn{Acyclic} applies. % to $t$.
+%To recognize when the precondition of \rn{Unique} holds,
+%the procedure considers the set $\SSS$ of all codatatype equivalence classes. It
+%simultaneously visits the equivalence classes of arguments of constructor terms in each equivalence class in $\SSS$,
+%while partitioning $\SSS$ into $\SSS_1, \ldots, \SSS_n$ based on the top-most symbol of constructor terms in these equivalence classes
+%and the equivalence of their arguments of ordinary types.
+%It then partitions each set %in this partition %in the same manner
+%recursively.
+%If %after doing so,
+%the resulting partition contains a set $\SSS_i$ containing distinct terms $u$
+%and $v$, it deduces that \rn{Unique} applies to $u$ and~$v$.
 
@@
 phase~2.
-
-\begin{figure}[t!]
-\normalsize
-\centering
-\begin{enumerate}
-\item Let $\mathcal{U}$ be a partition of the equivalence classes of $\Ec$ such that $\ec{t} =_{\mathcal{U}} \ec{u}$ for distinct $\ec{t}, \ec{u}$ if and only if:
-\begin{itemize}
-\item $t\typ{\delta}$, $u\typ{\delta}$ where $\delta \in \Codata$, and $\const{C}( t_1, \ldots, t_k ) \in \ec{t}$ and $\const{C}( u_1, \ldots, u_k ) \in \ec{u}$.
-\end{itemize}
-\item Repeat until $\mathcal{U}$ is unchanged
-\begin{enumerate}
-\item $\mathcal{U}' := \emptyset$
-\item for each $U_i \in \mathcal{U}$:
-\begin{itemize}
-\item[\ ] Let $\mathcal{U}_i$ be a partition of $U_i$ such that $\ec{t} =_{\mathcal{U}_i} \ec{u}$ for distinct $\ec{t}, \ec{u}$ if and only if:
-\begin{itemize}
-\item $\const{C}( t_1, \ldots, t_k ) \in \ec{t}$ and $\const{C}( u_1, \ldots, u_k ) \in \ec{u}$, and
-\item $\ec{t_j} =_{\mathcal{U}} \ec{u_j}$ for each $j = 1, \ldots, k$.
-\end{itemize}
-\item[\ ] $\mathcal{U}' := \mathcal{U}' \cup \mathcal{U}_i$.
-\end{itemize}
-\item $\mathcal{U} := \mathcal{U}'$
-\end{enumerate}
-\item If distinct $\ec{t}, \ec{u} \in U$ for some $U \in \mathcal{U}$, apply \rn{Unique} to $t,u$.
-\end{enumerate}
-\vspace*{-2pt} %% TYPESETTING HACK
-\caption{\,Algorithm for checking Uniqueness.
-}
-\label{fig:uniqueness-algo}
-\end{figure}
-
-\begin{figure}[t!]
-\normalsize
-\centering
-\begin{enumerate}
-\item Let $U$ be the datatype equivalence classes of $\Ec$.
-\item Repeat until $U \neq \emptyset$.
-\begin{itemize}
-\item[\ ] For some $\ec{t} \in U$, $\mathrm{traverse}( \ec{t}, U, \emptyset )$
-\end{itemize}
-\end{enumerate}
-
-\flushleft $\mathrm{traverse}( \ec{t}, U, P )$:
-\begin{enumerate}
-\item If $\ec{t} \in P$, apply \rn{Acyclic} to $t$.
-\item If $\ec{t} \in U$,
-\begin{enumerate}
-\item If $\const{C}( t_1, \ldots, t_k ) \in \ec{t}$, then for each $j = 1, \ldots k$,
-\begin{itemize}
-\item[\ ] $\mathrm{traverse}( \ec{t_j}, U, P \cup \{ \ec{t} \} )$
-\end{itemize}
-\item $U := U \textbackslash \{ \ec{t} \}$.
-\end{enumerate}
-\end{enumerate}
-\vspace*{-2pt} %% TYPESETTING HACK
-\caption{\,Algorithm for checking Acyclic.
-}
-\label{fig:acyclic-algo}
-\end{figure}
 
@@
 
-{\footnotesize
+{
 \def\ackname{Acknowledgment}
-\paragraph{\footnotesize\ackname.}
+\paragraph{\ackname.}
 \looseness=-1

### 2015-11-09 jour/jour.tex
Minor

@@
 For \rn{Unique}, Figure~\ref{fig:uniqueness-algo} considers a partition $\mathcal{U}$ of the equivalence classes of our set $\Ec$
-such that equivalence classes having a constructor term with top symbol $\const{C}$ are placed in the same subset, for each constructor $\const{C}$.
+such that codatatype equivalence classes having a constructor term with top symbol $\const{C}$ are placed in the same subset, 
+for each constructor $\const{C}$.
 We write $\ec{t} =_{\mathcal{U}} \ec{u}$ to denote that $\ec{t}$ and $\ec{u}$ reside in the same subset within $\mathcal{U}$.

### 2015-11-10 jour/jour.tex
Splitting section 3 into 2 subsections, minor.

@@
 whose variables are interpreted existentially.
-The procedure is formulated as a tableau-like calculus.
+The decision procedure is formulated as a tableau-like calculus.
 %Following the tradition of refutational theorem proving,
@@
 higher-level, using unoriented equations instead of oriented ones.
-
-% whose derivation rules are applied until saturation or contradiction.
 
@@
 
+%AJR: would prefer splitting this section into two parts (calculus, correctness), let me know if you agree.
+\subsection{Calculus for $\thD$}
+\label{sec:calculus-for-dc}
+
+% whose derivation rules are applied until saturation or contradiction.
+
 %We will commonly denote tuples of terms $( t_1, \ldots, t_n )$ in bold font, as $\vec t$.
@@
 
+In this section, we describe the derivation rules of the calculus for $\thD$.
 A derivation rule can be applied to $\Ec$ if %all
@@
 
-\paragraph{Correctness.}
+\subsection{Termination and Correctness}
+\label{sec:termination-and-correctness}
+
+We now show the termination and correctness of the calculus.
 Correctness means that if there exists a closed derivation tree with root node
@@
 
-\begin{lemma}
-\label{lem:mu-norm-interp}
-If the $\mu$-term $t$ is normal and $u$ is a subterm of $t$, then $\expand{t}{u}$ is normal.
-\end{lemma}
+%\begin{lemma}
+%\label{lem:mu-norm-interp}
+%If the $\mu$-term $t$ is normal and $u$ is a subterm of $t$, then $\expand{t}{u}$ is normal.
+%\end{lemma}
 % TODO:
@@
 If the resulting partition $\mathcal{U}$ contains a subset $U$ having two distinct equivalence classes $\ec{t}$ and $\ec{u}$, then \rn{Unique} applies to $t$ and $u$.
+\textbf{Mention this coincides with DFA minimization algorithm.}
 
@@
 
+The implementation is able to produce models for satisfiable inputs.
+
 [TODO]
@@
 goal. The resulting problem was given to the
-development version of CVC4 (\textbf{TODO: which revision?}) and to
-Z3 4.3.2 (\textbf{TODO: really?}) for comparison, each running for
+development version of CVC4 (\textbf{TODO: which revision?}) %AJR : version is from September 15 2015, git doesn't give nice revision numbers unfortunately
+and to
+Z3 4.3.2 %(\textbf{TODO: really?}) AJR : yes
+for comparison, each running for
 up to 60~s on the StarExec cluster

### 2015-11-10 jour/jour.tex
Some clarifications

@@
 denote the equivalence class of $t$ in $\Ec$.
+Moreover, at the end of this phase, each equivalence class $\ec{t}$ contains (at most) 
+one constructor term that is unique up to congruence.
+Thus, in the subsequent phases, 
+when considering the case that $\ec{t}$ contains a constructor term,
+it is enough for our purposes to select an arbitrary constructor term $\const{C}( \vec t\, ) \in \ec{t}$ among these.
 
@@
 With each equivalence class $\ec{u}$, we associate a fresh variable $\Varec{u}$
-of the same type as $u$
-%not occurring in $\tEc$,
-and set $\Val\ec{u} := {\Varec{u}}$. Because $\Varec{u}$ is unspecified,
+of the same type as $u$. %not occurring in $\tEc$,
+%AJR : this sentence is necessary, since we are abusing the \Varec notation a bit
+For a term $t \in \tEc$, 
+we write $\Varec{t}$ to denote the variable associated with equivalence class $\ec{t}$.
+Initially, we set $\Val\ec{u} := {\Varec{u}}$ for each equivalence class $\ec{u}$. Because $\Varec{u}$ is unspecified,
 this indicates that there are initially no
-constraints on the values for any equivalence class $\ec{u}$. The mapping $\Val$
+constraints on the values for any equivalence class $\ec{u}$. 
+The mapping $\Val$
 is refined by applying the following unfolding rule exhaustively:\strut
@@
 yielding
-$\{ x,\, \const{S}(x) \}$ and
+$\{ x,\, \const{S}(x) \}$ andwill
 $\{ y,\, \const{S}(y),\allowbreak\, \const{S}(\const{S}(y)) \}$.
@@
 The implementation is able to produce models for satisfiable inputs.
-
-[TODO]
+As described in Section~\ref{sec:termination-and-correctness},
+given a saturated set $\Ec$, we may construct a map $\ValC$ from variables (corresponding to the equivalence classes of $\Ec$) to closed normal $\MU$-terms.
+%In practice, this requires a fair enumerator for each (co)datatype $\tau$.
+
+%Recall that to assign a value $\MU {\Varec{x}}.\; t$ to variable
+
+%We do not explicitly construct the set $\BAD_{\smash{\J}}^x( \Val )$.
+%Instead,
 

### 2015-11-10 jour/jour.tex
Minor fix

@@
 yielding
-$\{ x,\, \const{S}(x) \}$ andwill
+$\{ x,\, \const{S}(x) \}$ and
 $\{ y,\, \const{S}(y),\allowbreak\, \const{S}(\const{S}(y)) \}$.

### 2015-11-11 jour/jour.tex
Paragraph on model construction.

@@
 If the resulting partition $\mathcal{U}$ contains a subset $U$ having two distinct equivalence classes $\ec{t}$ and $\ec{u}$, then \rn{Unique} applies to $t$ and $u$.
-\textbf{Mention this coincides with DFA minimization algorithm.}
+This algorithm is similar to the classic algorithm for minimizing deterministic finite automata (DFA). \textbf{maybe cite?}
 
@@
 
-The implementation is able to produce models for satisfiable inputs.
+When instructed, the implementation produces models for satisfiable inputs.
 As described in Section~\ref{sec:termination-and-correctness},
-given a saturated set $\Ec$, we may construct a map $\ValC$ from variables (corresponding to the equivalence classes of $\Ec$) to closed normal $\MU$-terms.
-%In practice, this requires a fair enumerator for each (co)datatype $\tau$.
-
-%Recall that to assign a value $\MU {\Varec{x}}.\; t$ to variable
-
-%We do not explicitly construct the set $\BAD_{\smash{\J}}^x( \Val )$.
-%Instead,
+given a saturated set $\Ec$, we may construct a map $\ValC$ from 
+the equivalence classes of $\Ec$ to closed normal $\MU$-terms.
+Recall our construction of $\ValC$ requires picking values $\MU {\Varec{x}}.\; t$ for each $\Varec{x} \in \FV( \ValC )$
+such that $\MU {\Varec{x}}.\; t$ does not occur in the set $\BAD_{\smash{\J}}^{\Varec{x}}( \ValC )$.
+To pick values $\MU {\Varec{x}}.\; t$, we use a fair enumerator for each (co)datatype $\tau$ which enumerates the values
+of type $\tau$ in a normal interpretation, which are closed normal $\MU$-terms.
+Concretely, we enumerate a stream of \emph{every} $\MU$-term of our signature $\Sigma$ in a fair manner,
+discarding whose that have free variables or are not normal.
+For those that are closed and normal, we then check whether they occur in the set $\BAD_{\smash{\J}}^{\Varec{x}}( \ValC )$,
+which we do not explicitly construct.
+%We do not explicitly construct the set $\BAD_{\smash{\J}}^{\Varec{x}}( \ValC )$.
+Instead, to check if a term $\MU {\Varec{x}}.\; t$ occurs in the set $\BAD_{\smash{\J}}^{\Varec{x}}( \ValC )$, 
+we check if it is alpha-equivalent to a closed-term $(\MU {\Varec{y}}.\; u) \{ \Varec{x} \mapsto z \}$ for
+some $\MU {\Varec{y}}.\; u$ in the range of $\ValC$ and variable $z$.
+This can be efficiently checked by matching the term $\MU {\Varec{x}}.\; t$ with the term $\MU {\Varec{y}}.\; u$.
+
+%Once the mapping $\ValC$ is constructed, our model $\J$ interprets
+%$\J( x ) = \ValC \ec{x}$ for each variable $x$ occurring free in $\Ec$.
 

### 2015-11-11 jour/jour.tex
Minor

@@
 goal. The resulting problem was given to the
-development version of CVC4 (\textbf{TODO: which revision?}) %AJR : version is from September 15 2015, git doesn't give nice revision numbers unfortunately
+development version (from 15 September 2015) of CVC4 %AJR : version is from September 15 2015, git doesn't give nice revision numbers unfortunately
 and to

### 2015-11-12 jour/jour.tex
minor edits

@@
 %AJR: would prefer splitting this section into two parts (calculus, correctness), let me know if you agree.
-\subsection{Calculus for $\thD$}
+%JB: I fully agree, but I tuned the titles a little bit. Also, I tried
+%%% to avoid such "meta" (or "road-sign") wording as
+%%% "This section will describe such-and-such".
+
+\subsection{A Calculus for $\thD$}
 \label{sec:calculus-for-dc}
@@
 
-In this section, we describe the derivation rules of the calculus for $\thD$.
+Our calculus for $\thD$ consists of derivation rules.
 A derivation rule can be applied to $\Ec$ if %all
-the %specified
+the specified
 preconditions are met.
@@
 
-The calculus consists of three sets of rules, given in
+The derivation rules are partitioned into three sets of rules, given in
 Figs.\ \ref{fig:cc-rules}~to~\ref{fig:split-rule}, corresponding to three
-phases. The first phase computes the bidirectional closure of $\Ec$. The second
+phases of the calculus. The first phase computes the bidirectional closure of $\Ec$. The second
 phase makes inferences based on acyclicity (for datatypes) and uniqueness
@@
 denote the equivalence class of $t$ in $\Ec$.
-Moreover, at the end of this phase, each equivalence class $\ec{t}$ contains (at most) 
+Moreover, at the end of this phase, each equivalence class $\ec{t}$ contains
+at most %%% JB: dropped ()'s
 one constructor term that is unique up to congruence.
@@
 when considering the case that $\ec{t}$ contains a constructor term,
-it is enough for our purposes to select an arbitrary constructor term $\const{C}( \vec t\, ) \in \ec{t}$ among these.
+it is enough %for our purposes
+to select an arbitrary constructor term $\const{C}( \vec t\, ) \in \ec{t}$ among these.
 
@@
 of the same type as $u$. %not occurring in $\tEc$,
-%AJR : this sentence is necessary, since we are abusing the \Varec notation a bit
 For a term $t \in \tEc$, 
-we write $\Varec{t}$ to denote the variable associated with equivalence class $\ec{t}$.
+we write $\Varec{t}$ to denote the variable associated with
+the equivalence class $\ec{t}$.
 Initially, we set $\Val\ec{u} := {\Varec{u}}$ for each equivalence class $\ec{u}$. Because $\Varec{u}$ is unspecified,
@@
 
-\subsection{Termination and Correctness}
+\subsection{Termination and Correctness of the Calculus}
 \label{sec:termination-and-correctness}

### 2015-11-12 jour/jour.tex
more

@@
 %\end{lemma}
-% TODO:
 %\begin{proof}
@@
 quite a few details. This section describes the main aspects of the
-implementation within the SMT solver CVC4:\ the integration into CDCL($T$)
-\cite{ganzinger-et-al-2004}, the extension to quantified formulas, and some of
-the optimizations.
-
-%%% @ANDY: I'm trying without \paragraph's, to save space but also to make the
-%%% section look more like a whole than a random list of stuff. Revert if you
-%%% disagree.
-\subsection{Theory Solver for $\thD$}
-The decision procedure is implemented as a \relax{theory solver} of CVC4, that is, a
+implementation within the SMT solver CVC4:\ the integration of the
+procedure into CDCL($T$)
+\cite{ganzinger-et-al-2004}, the construction of models with $\mu$-terms,
+and the extension of the procedure to quantified formulas.
+%, and some of the optimizations.
+
+\subsection{A Theory Solver for $\thD$}
+
+The decision procedure is implemented as a \relax{theory solver} of CVC4---that is, a
 specialized solver for determining the satisfiability of conjunctions of
@@
 takes as input the union $M_i$
-of (1)~the purified form of $T_i$-literals occurring in $M$, where fresh
+of
+\begin{itemize}
+\item the purified form of $T_i$-literals occurring in $M$, where fresh
 variables replace terms containing symbols not belonging to~$T_i$;
-(2)~additional (dis)equalities between variables of types not belonging to
-$T_i$. Each $T_i$-solver
+\item additional (dis)equalities between variables of types not belonging to
+$T_i$.
+\end{itemize}
+Each $T_i$-solver
 either reports that a subset $C$ of $M_i$ is $T_i$-unsatisfiable, in which case
@@
 \centering
+\flushleft $\rn{Acyclic}(t)$:
+
+\vspace*{-\topsep}
+
 \begin{enumerate}
 \item Let $U$ be the datatype equivalence classes of $\Ec$.
-\item Repeat until $U \neq \emptyset$.
+\item Repeat until $U \neq \emptyset$:
 \begin{itemize}
-\item[\ ] For some $\ec{t} \in U$, $\mathrm{traverse}( \ec{t}, U, \emptyset )$
+\item[\ ] For some $\ec{t} \in U$, $\rn{Traverse}( \ec{t}, U, \emptyset )$.
 \end{itemize}
@@
 
-\flushleft $\mathrm{traverse}( \ec{t}, U, P )$:
+\flushleft $\rn{Traverse}( \ec{t}, U, P )$:
+
+\vspace*{-\topsep}
+
 \begin{enumerate}
-\item If $\ec{t} \in P$, apply \rn{Acyclic} to $t$.
-\item If $\ec{t} \in U$,
+\item If $\ec{t} \in P$, $\rn{Acyclic}(t)$.
+\item If $\ec{t} \in U$:
 \begin{enumerate}
-\item If $\const{C}( t_1, \ldots, t_k ) \in \ec{t}$, then for each $j = 1, \ldots k$,
+\item[2.1.] If $\const{C}( t_1, \ldots, t_k ) \in \ec{t}$, then for each $j = 1, \ldots k$:
 \begin{itemize}
-\item[\ ] $\mathrm{traverse}( \ec{t_j}, U, P \cup \{ \ec{t} \} )$
+\item[2.1.1.] $\rn{Traverse}( \ec{t_j}, U, P \cup \{ \ec{t} \} )$.
 \end{itemize}
-\item $U := U \textbackslash \{ \ec{t} \}$.
+\item[2.2.] $U := U \textbackslash \{ \ec{t} \}$.
 \end{enumerate}
 \end{enumerate}
-\vspace*{-2pt} %% TYPESETTING HACK
+\vspace*{-\topsep}
 \caption{\,Algorithm for checking \rn{Acyclic}}
@@
 \centering
+
+\flushleft $\rn{Unique}(t, u)$:
+
+\vspace*{-\topsep}
+
 \begin{enumerate}
-\item Let $\mathcal{U}$ be a partition of the equivalence classes of $\Ec$ such that $\ec{t} =_{\mathcal{U}} \ec{u}$ for distinct $\ec{t}, \ec{u}$ if and only if:
+\item Let $\mathcalx{U}$ be a partition of the equivalence classes of $\Ec$ such that $\ec{t} =_{\mathcalx{U}} \ec{u}$ for distinct $\ec{t}, \ec{u}$ if and only if
+  $t$, $u$ have type $\delta \in \Codata$,
+  $\const{C}( t_1, \ldots, t_k ) \in \ec{t}$,
+  and $\const{C}( u_1, \ldots, u_k ) \in \ec{u}$.
+
+\item Repeat until $\mathcalx{U}$ is unchanged:
+\begin{enumerate}
+\item[2.1.] $\mathcalx{U}' := \emptyset$.
+\item[2.2.] For each $U_i \in \mathcalx{U}$:
 \begin{itemize}
-\item $t\typ{\delta}$, $u\typ{\delta}$ where $\delta \in \Codata$, and $\const{C}( t_1, \ldots, t_k ) \in \ec{t}$ and $\const{C}( u_1, \ldots, u_k ) \in \ec{u}$.
-\end{itemize}
-\item Repeat until $\mathcal{U}$ is unchanged
-\begin{enumerate}
-\item $\mathcal{U}' := \emptyset$
-\item for each $U_i \in \mathcal{U}$:
-\begin{itemize}
-\item[\ ] Let $\mathcal{U}_i$ be a partition of $U_i$ such that $\ec{t} =_{\mathcal{U}_i} \ec{u}$ for distinct $\ec{t}, \ec{u}$ if and only if:
+\item[2.2.1.] Let $\mathcalx{U}_i$ be a partition of $U_i$ such that $\ec{t} =_{\mathcalx{U}_i} \ec{u}$ for distinct $\ec{t}, \ec{u}$ if and only if
 \begin{itemize}
 \item $\const{C}( t_1, \ldots, t_k ) \in \ec{t}$ and $\const{C}( u_1, \ldots, u_k ) \in \ec{u}$, and
-\item $\ec{t_j} =_{\mathcal{U}} \ec{u_j}$ for each $j = 1, \ldots, k$.
+\item $\ec{t_j} =_{\mathcalx{U}} \ec{u_j}$ for each $j = 1, \ldots, k$.
 \end{itemize}
-\item[\ ] $\mathcal{U}' := \mathcal{U}' \cup \mathcal{U}_i$.
+\item[2.2.2.] $\mathcalx{U}' := \mathcalx{U}' \cup \mathcalx{U}_i$.
 \end{itemize}
-\item $\mathcal{U} := \mathcal{U}'$
+\item[2.3.] $\mathcalx{U} := \mathcalx{U}'$.
 \end{enumerate}
-\item If distinct $\ec{t}, \ec{u} \in U$ for some $U \in \mathcal{U}$, apply \rn{Unique} to $t,u$.
+\item If distinct $\ec{t}, \ec{u} \in U$ for some $U \in \mathcalx{U}$, $\rn{Unique}(t,u)$.
 \end{enumerate}
-\vspace*{-2pt} %% TYPESETTING HACK
+\vspace*{-\topsep}
 \caption{\,Algorithm for checking \rn{Unique}}
@@
 Before constructing a model for $F\!$, the theory solver constructs neither $\mu$-terms nor the mapping $\Val$.
-Instead, we use the algorithms in Figures~\ref{fig:acyclic-algo} and~\ref{fig:uniqueness-algo}
-for determining whether the rules \rn{Acyclic} and \rn{Unique} apply to our set $\Ec$.
-
-For \rn{Acyclic}, Figure~\ref{fig:acyclic-algo} considers the set $U$ of all datatype equivalence classes of $\Ec$.
-We choose an arbitrary $\ec{t}$ in $U$ and call the recursive subprocedure $\mathrm{traverse}$,
+Instead, it relies on the algorithms in Figs.~\ref{fig:acyclic-algo} and~\ref{fig:uniqueness-algo}
+for determining whether the rules \rn{Acyclic} and \rn{Unique} apply to the set $\Ec$.
+
+For \rn{Acyclic}, Fig.~\ref{fig:acyclic-algo} considers the set $U$ of all datatype equivalence classes of $\Ec$.
+We choose an arbitrary $\ec{t}$ in $U$ and call the recursive subprocedure $\rn{Traverse}$,
 which takes as input the current equivalence class we are processing, the set of equivalence classes $U$ we have yet to process,
 and the set of equivalence classes $P$ we are currently processing.
-If a call to $\mathrm{traverse}( \ec{t}, U, P )$ is such that $\ec{t} \in P$,
+If a call to $\rn{Traverse}( \ec{t}, U, P )$ is such that $\ec{t} \in P$,
 we know that the rule \rn{Acyclic} applies to $t$.
@@
 then if $\ec{t}$ contains a constructor term $\const{C}( t_1, \ldots, t_k )$, 
-then we recursively call $\mathrm{traverse}$ on each of $\ec{t_1}, \ldots, \ec{t_n}$.
+then we recursively call $\rn{Traverse}$ on each of $\ec{t_1}, \ldots, \ec{t_n}$.
 If this succeeds, we remove $\ec{t}$ from $U$.
 
-For \rn{Unique}, Figure~\ref{fig:uniqueness-algo} considers a partition $\mathcal{U}$ of the equivalence classes of our set $\Ec$
+For \rn{Unique}, Fig.~\ref{fig:uniqueness-algo} considers a partition $\mathcalx{U}$ of the equivalence classes of our set $\Ec$
 such that codatatype equivalence classes having a constructor term with top symbol $\const{C}$ are placed in the same subset, 
 for each constructor $\const{C}$.
-We write $\ec{t} =_{\mathcal{U}} \ec{u}$ to denote that $\ec{t}$ and $\ec{u}$ reside in the same subset within $\mathcal{U}$.
-We then refine $\mathcal{U}$ by 
-constructing a partition $\mathcal{U}_i$ of each $U_i \in \mathcal{U}$ such that distinct $\ec{t}$ and $\ec{u}$ reside in the same subset of $\mathcal{U}_i$ 
+We write $\ec{t} =_{\mathcalx{U}} \ec{u}$ to denote that $\ec{t}$ and $\ec{u}$ reside in the same subset within $\mathcalx{U}$.
+We then refine $\mathcalx{U}$ by 
+constructing a partition $\mathcalx{U}_i$ of each $U_i \in \mathcalx{U}$ such that distinct $\ec{t}$ and $\ec{u}$ reside in the same subset of $\mathcalx{U}_i$ 
 if and only if they contain constructor terms $\const{C}( t_1, \ldots, t_k )$ and $\const{C}( u_1, \ldots, u_k )$
-where each $t_j, u_j$ are such that $\ec{t_j}$ and $\ec{u_j}$ reside in the same subset of $\mathcal{U}$ for each $j=1, \ldots, k$.
-Let $\mathcal{U'}$ be the union of each of these partitions.
-We update $\mathcal{U}$ to $\mathcal{U'}$.
-This refinement is repeated until $\mathcal{U}$ is left unchanged.
-If the resulting partition $\mathcal{U}$ contains a subset $U$ having two distinct equivalence classes $\ec{t}$ and $\ec{u}$, then \rn{Unique} applies to $t$ and $u$.
-This algorithm is similar to the classic algorithm for minimizing deterministic finite automata (DFA). \textbf{maybe cite?}
+where each $t_j, u_j$ are such that $\ec{t_j}$ and $\ec{u_j}$ reside in the same subset of $\mathcalx{U}$ for each $j=1, \ldots, k$.
+Let $\mathcalx{U'}$ be the union of each of these partitions.
+We update $\mathcalx{U}$ to $\mathcalx{U'}$.
+This refinement is repeated until $\mathcalx{U}$ is left unchanged.
+If the resulting partition $\mathcalx{U}$ contains a subset $U$ having two distinct equivalence classes $\ec{t}$ and $\ec{u}$, then \rn{Unique} applies to $t$ and $u$.
+This algorithm is analogous to Hopcroft's algorithm for minimizing
+deterministic finite automata \cite{hopcroft-1971}---terms correspond to
+states, argument positions $1,\ldots,k$ correspond to input symbols, and the top constructor symbol
+of a term generalizes the accepting/rejecting status of a state.
 
@@
 \subsection{Extension to Quantified Formulas}
+
 While the decision procedure is restricted to universal conjectures, users often

### 2015-11-12 jour/jour.tex
more

@@
 Instead, to check if a term $\MU {\Varec{x}}.\; t$ occurs in the set $\BAD_{\smash{\J}}^{\Varec{x}}( \ValC )$, 
-we check if it is alpha-equivalent to a closed-term $(\MU {\Varec{y}}.\; u) \{ \Varec{x} \mapsto z \}$ for
+we check if it is $\alpha$-equivalent to a closed term $(\MU {\Varec{y}}.\; u) \{ \Varec{x} \mapsto z \}$ for
 some $\MU {\Varec{y}}.\; u$ in the range of $\ValC$ and variable $z$.

### 2015-11-13 jour/jour.tex
more minor edits

--- 2015-11-12 jour/jour.tex
@@
 
-When instructed, the implementation produces models for satisfiable inputs.
+When instructed to do so, the implementation produces models for satisfiable inputs.
 As described in Section~\ref{sec:termination-and-correctness},
-given a saturated set $\Ec$, we may construct a map $\ValC$ from 
+given a saturated set $\Ec$, we %may
+construct a map $\ValC$ from 
 the equivalence classes of $\Ec$ to closed normal $\MU$-terms.
-Recall our construction of $\ValC$ requires picking values $\MU {\Varec{x}}.\; t$ for each $\Varec{x} \in \FV( \ValC )$
+Recall that our construction of $\ValC$ requires picking values $\MU {\Varec{x}}.\; t$ for each $\Varec{x} \in \FV( \ValC )$
 such that $\MU {\Varec{x}}.\; t$ does not occur in the set $\BAD_{\smash{\J}}^{\Varec{x}}( \ValC )$.
-To pick values $\MU {\Varec{x}}.\; t$, we use a fair enumerator for each (co)datatype $\tau$ which enumerates the values
-of type $\tau$ in a normal interpretation, which are closed normal $\MU$-terms.
-Concretely, we enumerate a stream of \emph{every} $\MU$-term of our signature $\Sigma$ in a fair manner,
-discarding whose that have free variables or are not normal.
-For those that are closed and normal, we then check whether they occur in the set $\BAD_{\smash{\J}}^{\Varec{x}}( \ValC )$,
-which we do not explicitly construct.
+To pick values $\MU {\Varec{x}}.\; t$, we use a fair enumerator for each (co)datatype $\tau$,
+which lists all the values
+of type $\tau$ in a normal interpretation. %, which are closed normal $\MU$-terms.
+
+%Concretely,
+For codatatypes, we enumerate a stream of \relax{every} $\MU$-term of our signature $\Sigma$ in a fair manner,
+discarding those that have free variables or are not normal.
+For those that are closed and normal, we then check whether they occur in the set $\BAD_{\smash{\J}}^{\Varec{x}}( \ValC )$.
+This set does not need to be explicitly constructed.
 %We do not explicitly construct the set $\BAD_{\smash{\J}}^{\Varec{x}}( \ValC )$.
-Instead, to check if a term $\MU {\Varec{x}}.\; t$ occurs in the set $\BAD_{\smash{\J}}^{\Varec{x}}( \ValC )$, 
-we check if it is alpha-equivalent to a closed-term $(\MU {\Varec{y}}.\; u) \{ \Varec{x} \mapsto z \}$ for
+To determine if a term $\MU {\Varec{x}}.\; t$ occurs in the set $\BAD_{\smash{\J}}^{\Varec{x}}( \ValC )$, 
+we check if it is $\alpha$-equivalent to a closed term $(\MU {\Varec{y}}.\; u) \{ \Varec{x} \mapsto z \}$ for
 some $\MU {\Varec{y}}.\; u$ in the range of $\ValC$ and variable $z$.
-This can be efficiently checked by matching the term $\MU {\Varec{x}}.\; t$ with the term $\MU {\Varec{y}}.\; u$.
+This can be efficiently achieved by matching the term $\MU {\Varec{x}}.\; t$ with %the term
+$\MU {\Varec{y}}.\; u$.
 
@@
 the implementation adds the clause %of the form
-$(\exists x\;y\typ{\ty{u}}.\;\, x \not\teq y ) \mathrel\vee s \teq t$ to $F\!$.
-In other words, either the cardinality of $\ty{u}$ is greater than one or $s$ is equal to $t$.
+$(\exists x\;y\typ{\ty{u}}.\;\, x \not\teq y ) \mathrel\vee s \teq t$ to $F\!$,
+which states that either the cardinality of $\ty{u}$ is greater than one or
+$s$ must be equal to $t$.
 

### 2015-11-13 jour/jour.tex
added data

@@
 soundness, without which the generated models would often be spurious. The
-contributions to proving are less clear; they depend on how often the need for
-acyclicity or uniqueness arises in practice.
+contributions to proving are less clear; they depend on how often
+acyclicity and uniqueness are necessary for a proof.
 
@@
 that define codatatypes falling within the supported fragment. We
-added unpublished formalizations about Bird and Stern--Brocot trees
+also included unpublished formalizations about Bird and Stern--Brocot trees
 by Peter Gammie and Andreas Lochbihler (\gandl,
-265~goals). To exercise the datatype support, formalizations about lists and trees
+265~goals). To exercise the datatype support, formalizations about finite
+lists and trees
 were added to the first two benchmark sets. The formalizations were selected before
@@
 goal. The resulting problem was given to the
-development version (from 15 September 2015) of CVC4 %AJR : version is from September 15 2015, git doesn't give nice revision numbers unfortunately
-and to
-Z3 4.3.2 %(\textbf{TODO: really?}) AJR : yes
-for comparison, each running for
+development version of CVC4
+(from 15 September 2015) and to
+Z3 4.3.2 for comparison, each running for
 up to 60~s on the StarExec cluster
@@
 support for datatypes and codatatypes either enabled or disabled. The
-contributions of the acyclicity and uniqueness rules were also measured.
+contributions of the acyclicity and uniqueness rules were also measured,
+by selectively enabling or disabling the rules.
 % to find
@@
 
-\newpage
-
 The results are summarized in Tables \ref{tab:benchsixteen} and

### 2015-11-13 jour/jour.tex
more

@@
 because the acyclicity and uniqueness rules are necessary for solution
-soundness, without which the generated models would often be spurious. The
-contributions to proving are less clear; they depend on how often
-acyclicity and uniqueness are necessary for a proof.
-
-To evaluate the decision procedure for proving conjectures,
+soundness, without which the generated models would often be
+spurious. For example, given the axioms
+\begin{align*}
+\const{zeros} & = \const{SCons}(0,\:\const{zeros})
+& \const{sconst}~n & = \const{SCons}(n,\:\const{sconst}~n)
+\end{align*}
+the conjecture
+$\const{zeros} = \const{sconst}~0$
+would be ``refuted'' by a spurious countermodel where
+\const{zero} and $\const{sconst}~0$ are interpreted by two distinct
+values that both correspond to the \hbox{$\mu$-term}
+$\MU s.\; \const{SCons}(0, s)$,
+violating uniqueness.
+
+By contrast, the contributions of the decision procedure
+to proving are less clear; they depend on how
+often acyclicity and uniqueness are necessary for a proof.
+To evaluate this,
 we generated benchmark problems from existing interactive
@@
 The decision procedure
-makes a difference across all three benchmark suites. It accounts for an overall
-success rate increase of over 5\%, which is quite significant.
+makes a difference across all three benchmark suites. For the 16-lemma
+problems, it accounts for an overall
+success rate increase of over 5\%.
+Moreover, every aspect of the procedure, including the
+more expensive rules, make a contribution.
+
+For the 256-lemma problems, the difference
+is much smaller, at 0.5\%.
 %For Isabelle
@@
 %obviously depends on the application.
-The raw evaluation data also
-suggest that the theoretically stronger decision procedures almost completely
-subsume the weaker ones in practice. We encountered only one
-(\textbf{TODO: FIXME}) goal (out of $2 \times 4458$)
-that was solved by a configuration of CVC4 and unsolved by a configuration of
-CVC4 with more features enabled.
-%
-Moreover, every aspect of the procedure, including the
-more expensive rules, make a contribution.
+Table~\ref{tab:benchtwofiftysix} clearly indicates that
+the theoretically stronger decision
+procedures do not always subsume the weaker ones in practice.
+The raw data reveal that the decision procedure proved 27~goals that
+could not be proved without it, but failed for 17~goals that could be proved
+without it. This potentially points to poor interactions between the decision
+procedure and the quantifier instantiation module
+\cite{ReynoldsTinelliMoura14}.
+
 %* and finally, we look in more detail at one or two such proofs, by presenting
@@
 %
-Four proofs were found thanks to
-the acyclicity rule and 17~required uniqueness. Among the latter,
-some were simple proofs of the form
+Overall, four proofs were found thanks to
+the acyclicity rule and 17~required uniqueness. Interestingly,
+no proofs were lost by enabling these rules. Among 17 proofs requiring uniqueness,
+some were simple arguments of the form
 \keyw{by}~\textit{coinduction}~\textit{auto}
 in Isabelle \cite{blanchette-et-al-2014-impl}, while others
-required more elaborate arguments---for example:
+involved more elaborate reasoning, including the following example:
 %

### 2015-11-13 jour/jour.tex
more

@@
 models the natural numbers $\const{Z}$, $\const{S}(\const{Z})$, $\const{S}(\const{S}(\const{Z}))$, $\ldots$\afterLdots{},
-in Peano notation but extended with an
+in Peano notation, extended with an
 infinite value $\infty = \const{S}(\const{S}(\const{S}(\ldots)))$.
@@
 of infinite sequences or streams over an element type $\tau$ is
-allowed, even though it has no base case; the
-corresponding datatype would be rejected as non-well-founded
+allowed, even though it has no base case. By contrast, the
+corresponding datatype specification
+$$\vthinspace\keyw{datatype}~\;\ty{fstream}_{\,\tau} \,=\allowbreak\, \const{FSCons}(\tau,\allowbreak\: \ty{fstream}_{\,\tau})\vthinspace$$
+would be rejected as non-well-founded
 \cite{blanchette-et-al-2015-esop}.
@@
 An earlier version of this article was presented at the CADE-25 conference in
-Berlin, Germany. The article extends the conference paper with additional
+Berlin, Germany. This article extends the conference paper with additional
 background material on (co)datatypes (Section~\ref{sec:co-datatypes}), more
-elaborated metatheoretical proofs (Section~\ref{sec:the-decision-procedure}),
+detailed metatheoretical proofs (Section~\ref{sec:the-decision-procedure}),
 a description of the generation of models with $\mu$-binders in CVC4
@@
 distinguish them from variables (e.g., $x$, $y$).
-Names starting with an uppercase letter (e.g.,
+Symbol names starting with an uppercase letter (e.g.,
 $\const{S}$) are reserved for constructors. With each function symbol \const{f}
@@
 apply the $n$-ary function symbol
-\const{f} to $n$ well-typed arguments $t_1 :\nobreak \tau_1$, \ldots, $t_n :
-\tau_n$. Nullary function symbols, called constants, can appear without
+\const{f} to $n$ well-typed arguments $t_1 \mathbin{:} \tau_1$, \ldots, $t_n \mathbin{:}
+\tau_n$. Nullary function symbols, also called constants, can appear without
 parentheses in terms.
@@
 with $\const{p}(\ldots)$ abbreviating $\const{p}(\ldots) \teq \const{true}$.
-The notation $t\typ{\tau}$ stands for a term $t$ of type~$\tau$.
+The notation $t\typ{\tau}$ stands for a term $t$ of type~$\tau$;
+it should not be confused with the notation~$t :\nobreak \tau$,
+which is a statement expressing that $t$ has type $\tau$.
 When applied to terms, the symbol $=$ denotes syntactic equality.
@@
 proof reconstruction for (co)data\-type inferences in Isabelle.
-Finally, it might be worthwhile to
+In addition, it might be worthwhile to
 extend SMT solvers with dedicated support for (co)recursion.
@@
 %further in terms of supporting nested and mixed (co)recursion and quantified
-%ormulas over (co)datatypes in solvers.
+%formulas over (co)datatypes in solvers.
 

### 2015-11-13 jour/jour.tex
more

@@
 
-We fix a signature $\Sig = (\Types, \Funcs)$. The types are partitioned as
+We fix a signature $\Sig = (\Types, \Funcs)$. The types are partitioned into
 $\Types = \Data \mathrel{\uplus} \Codata \mathrel{\uplus} \Nondata$, where $\Data$ are the
@@
 mixed nested (co)recursion as well as
-(co)recursion through a non-(co)datatype (both of which make sense
-\cite{blanchette-et-al-2014-codata}).}
+(co)recursion through a non-(co)datatype (both of which make sense,
+at least in a higher-order setting \cite{blanchette-et-al-2014-codata}).}
 We allow ourselves some notational parameterization

### 2015-11-13 jour/jour.tex
more

@@
 For example, the type
-$$\vthinspace\keyw{codatatype}~\;\ty{stream}_{\,\tau} \,=\allowbreak\, \const{SCons}(\tau,\allowbreak\: \ty{stream}_{\,\tau})\vthinspace$$
+$$\vthinspace\keyw{codatatype}~\;\ty{stream}_{\vthinspace\tau} \,=\allowbreak\, \const{SCons}(\tau,\allowbreak\: \ty{stream}_{\vthinspace\tau})\vthinspace$$
 of infinite sequences or streams over an element type $\tau$ is
@@
 corresponding datatype specification
-$$\vthinspace\keyw{datatype}~\;\ty{fstream}_{\,\tau} \,=\allowbreak\, \const{FSCons}(\tau,\allowbreak\: \ty{fstream}_{\,\tau})\vthinspace$$
+$$\vthinspace\keyw{datatype}~\;\ty{fstream}_{\vthinspace\tau} \,=\allowbreak\, \const{FSCons}(\tau,\allowbreak\: \ty{fstream}_{\vthinspace\tau})\vthinspace$$
 would be rejected as non-well-founded
@@
 When applied to terms, the symbol $=$ denotes syntactic equality.
-The operator $\bigwedge_{\,i}\, \varphi_i$ abbreviates a conjunction
+The operator $\bigwedge_{\vthinspace i}\, \varphi_i$ abbreviates a conjunction
 $\varphi_1 \mathrel\land \cdots \mathrel\land \varphi_n$.
@@
 % at the metalevel
-through subscripts---for example, $\ty{stream}_{\,\tau}$ denotes a
+through subscripts---for example, $\ty{stream}_{\vthinspace\tau}$ denotes a
 family of ground types including
-$\ty{stream}_{\,\ty{int}}$, $\ty{stream}_{\,\ty{bool}}$,
-and \smash{$\ty{stream}_{\,\ty{stream}_{\,\ty{real}}}$}.
+$\ty{stream}_{\vthinspace\ty{int}}$, $\ty{stream}_{\vthinspace\ty{bool}}$,
+and \smash{$\ty{stream}_{\vthinspace\ty{stream}_{\vthinspace\ty{real}}}$}.
 
@@
 as long as the argument types coincide. % \cite[Section~3]{blanchette-et-al-2014-codata}.
-However, this is disallowed by CVC4 and Z3, so we do not consider it here.}
+However, this is disallowed by SMT-LIB, so we do not consider it here.}
 Schematically:
@@
 \[\begin{aligned}[t]
-      \keyw{codatatype}\; \,\ty{llist}_{\,\tau} & \,=\, \const{LNil} \,\mid\, \const{LCons}(%\const{lhead}{:}\;
+      \keyw{codatatype}\; \,\ty{llist}_{\vthinspace\tau} & \,=\, \const{LNil} \,\mid\, \const{LCons}(%\const{lhead}{:}\;
       \tau,\: %\const{ltail}{:}\;
-      \ty{llist}_{\,\tau}) \\%[-.5\jot]
-      \keyw{datatype}\;\, \ty{tree}_{\,\tau} & \,=\, \const{Node}(\tau,\:\, \ty{forest}_{\,\tau}) \\[-\jot]
-      \keyw{and}\;\,\, \ty{forest}_{\,\tau} & \,=\, \const{FNil} \,\mid\, \const{FCons}(\ty{tree}_{\,\tau},\:\, \ty{forest}_{\,\tau})
+      \ty{llist}_{\vthinspace\tau}) \\%[-.5\jot]
+      \keyw{datatype}\;\, \ty{tree}_{\vthinspace\tau} & \,=\, \const{Node}(\tau,\:\, \ty{forest}_{\vthinspace\tau}) \\[-\jot]
+      \keyw{and}\;\,\, \ty{forest}_{\vthinspace\tau} & \,=\, \const{FNil} \,\mid\, \const{FCons}(\ty{tree}_{\vthinspace\tau},\:\, \ty{forest}_{\vthinspace\tau})
 \end{aligned}
@@
 This rules out non-well-founded specifications such as
-\[\keyw{datatype}\;\, \ty{fstream}_{\,\tau} \,=\, \const{FSCons}(\tau,\:\, \ty{fstream}_{\,\tau})\]
+that of \ty{fstream} in Section~\ref{sec:introduction}.
 For codatatypes, no admissibility check is necessary because there is always a term,
@@
 
-Non(co)recursive type specifications such~as
+Non(co)recursive type specifications such~as either of
 \[\begin{aligned}[t]
-      \keyw{datatype}~\vthinspace\ty{option}_{\,\tau\!} & \,=\, \const{None} \,\mid\, \const{Some}(\tau) \\[-.5\jot]
+      \keyw{datatype}~\vthinspace\ty{option}_{\vthinspace\tau\!} & \,=\, \const{None} \,\mid\, \const{Some}(\tau) \\[-.5\jot]
       \keyw{codatatype}~\vthinspace\ty{complex} & \,=\, \const{Complex}(\const{re}{:}\; \ty{real},\; \const{im}{:}\;\ty{real})
@@
 \text{Induction:}\quad
-\AXC{\strut$\bigwedge_{\,i,j}\, \forall x_1 \ldots \vthinspace x_{n_{i\negvthinspace j}}.\; \bigl(\bigwedge_{\,k}\, \mathcalx{IH}[x_k]\bigr) \medrightarrow P_i[\const C_{i\negvthinspace j}(x_1,\ldots,x_{n_{ij}})]$}
-\UIC{\strut$\bigwedge_{\,i}\, P_i[v_i]$}
+\AXC{\strut$\bigwedge_{\vthinspace i,j}\, \forall x_1 \ldots \vthinspace x_{n_{i\negvthinspace j}}.\; \bigl(\bigwedge_{\vthinspace k}\, \mathcalx{IH}[x_k]\bigr) \medrightarrow P_i[\const C_{i\negvthinspace j}(x_1,\ldots,x_{n_{ij}})]$}
+\UIC{\strut$\bigwedge_{\vthinspace i}\, P_i[v_i]$}
 \DP
@@
 derive the equality of pairs of codatatype values
-$v_1, w_1 : \delta_1,\,\ldots,\,v_\ell, w_\ell : \delta_\ell$, based on
-coinduction witnesses $R_1,\ldots,R_\ell$ that must be supplied upon use:
+$v_1, w_1 \mathbin: \delta_1,\,\ldots,\,v_\ell, w_\ell \mathbin: \delta_\ell$, based on
+suitable coinduction witnesses $R_1,\ldots,R_\ell$:
 %
@@
 \text{Coinduction:}\quad
-\AXC{\strut$\begin{gathered}\textstyle \smash{\bigwedge_{\,i}\, R_i[v_i, w_i]} \\[-\jot]\textstyle
-\bigwedge_{\,i}\, \forall v\; w.\;\, R_i[v, w] \medrightarrow
-  \bigwedge_{\,j}\, \const{d}_{\negvthinspace j}(v) \mathrel\land \const{d}_{\negvthinspace j}(w)
-  \medrightarrow \bigwedge_{\,k}\, \const{s}_{i\negvthinspace j}^{\,k}(v) \sim \const{s}_{i\negvthinspace j}^{\,k}(w)
+\AXC{\strut$\begin{gathered}\textstyle \smash{\bigwedge_{\vthinspace i}\, R_i[v_i, w_i]} \\[-\jot]\textstyle
+\bigwedge_{\vthinspace i}\, \forall v\; w.\;\, R_i[v, w] \medrightarrow
+  \bigwedge_{\vthinspace j}\, \const{d}_{\negvthinspace j}(v) \mathrel\land \const{d}_{\negvthinspace j}(w)
+  \medrightarrow \bigwedge_{\vthinspace k}\, \const{s}_{i\negvthinspace j}^{\,k}(v) \sim \const{s}_{i\negvthinspace j}^{\,k}(w)
 \end{gathered}$}
-\UIC{\strut$\bigwedge_{\,i}\, v_i \teq w_i$}
+\UIC{\strut$\bigwedge_{\vthinspace i}\, v_i \teq w_i$}
 \DP
@@
 first-order axiomatization, since there may be uncountably many of them.
-For example, $\ty{stream}_{\,\ty{int}}$ is isomorphic to the uncountable
+For example, $\ty{stream}_{\vthinspace\ty{int}}$ is isomorphic to the uncountable
 function space $\ty{nat} \to \ty{int}$.
@@
 defines a class of $\Sig$-interpretations $\JJJJ$,
-namely the ones that satisfy the %axioms --- existence of all infinite codatatype values cannot be axiomatized
+namely those that satisfy the %axioms --- existence of all infinite codatatype values cannot be axiomatized
 properties mentioned in this section, including (co)induction. The interpretations
@@
 have infinite values. A simple example is
-\vthinspace$\keyw{codatatype}~\ty{a} = \const{A}(\ty{a})$, which has a cardinality
-of one; its unique value is $\MU a.\; \const{A}(a)$. Other specimens are
-$\ty{stream}_{\,\ty{unit}\,}$
+\vthinspace$\keyw{codatatype}~\ty{a} = \const{A}(\ty{a})$, which is corecursive
+and yet has a cardinality of one;
+its unique value is $\MU a.\; \const{A}(a)$. Other specimens are
+$\ty{stream}_{\vthinspace\ty{unit}\,}$
 and both \ty{b} and \ty{c} in the specification
@@
 \ty{unit}$. We call such codatatypes
-\emph{corecursive singletons}. For the decision procedure, it will be
+\emph{corecursive singletons}, or simply singletons.
+For the decision procedure, it will be
 crucial to detect these. %, even if they rarely arise in practice.
@@
 Second, we can conditionally apply \rn{Single} to codatatype terms that may have cardinality one.
-For example, the $\ty{stream}_{\,\ty{u}}$ codatatype %, where $\ty{u}$ is uninterpreted.
+For example, the $\ty{stream}_{\vthinspace\ty{u}}$ codatatype %, where $\ty{u}$ is uninterpreted.
 has cardinality one precisely when $\ty{u}$ has cardinality one.

### 2015-11-13 jour/jour.tex
more

@@
 its negation is unsatisfiable. The presentation is
-inspired by Barrett et al.\ \cite{barrett-et-al-2007} but
-higher-level, using unoriented equations instead of oriented ones.
+inspired by Barrett et al.\ \cite{barrett-et-al-2007} but at
+a higher level, using unoriented equations instead of oriented ones.
 
@@
 % (in which case we call an application of it \emph{terminal}).
-One rule has multiple conclusions, % separated by $\ror$,
+One of the rules has multiple conclusions, % separated by $\ror$,
 denoting branching.
@@
 %The premises of the rules that make up the second phase refer to
-We describe the rules in this phase in terms of a mapping $\Val$
-that associates with each equivalence class a $\mu$-term as its representative.
+The rules in this phase are described in terms of a mapping $\Val$
+that assigns to each equivalence class a $\mu$-term as its representative.
 %\rem{Not explicitly constructed, add here?}
@@
 $\FV( \Val )$ denotes the set of free variables occurring in $\Val$'s range,
-and $\Val[x \mapsto t]$ denotes the \emph{variable-capturing} substitution of $t$ for
+and $\Val[x \mapsto t]$ denotes the variable-\emph{capturing} substitution of $t$ for
 $x$ in $\Val$'s range. It is easy to see that the height of terms produced as a

### 2015-11-13 jour/jour.tex
more

@@
 At this point, phase~1 is activated again, % to XXX the equivalence classes,
-yielding
+yielding the equivalence classes
 $\{ x,\, \const{S}(x) \}$ and
@@
 %$\{ {\Varec{y}} \mapsto {\VAREC{\const{S}(y)}}, {\VAREC{\const{S}(y)}} \mapsto {\Varec{y}} \}$,
-\rn{Unique} can finally derive $x \teq y$.
+\rn{Unique} can finally be applied to derive $x \teq y$.
 \xend

### 2015-11-13 jour/jour.tex
more

@@
 the specified
-preconditions are met.
+premises are met.
 The conclusion either specifies equalities to be added to $\Ec$
@@
 %the derivation rules of our calculus are given in \emph{guarded assignment form},
-%where a rule can be applied to $\Ec$ if it meets all of the specified preconditions for $\Ec$.
+%where a rule can be applied to $\Ec$ if it meets all of the specified premises for $\Ec$.
 
@@
 We now show the termination and correctness of the calculus.
-Correctness means that if there exists a closed derivation tree with root node
+Correctness means the following:
+If there exists a closed derivation tree with root node
 $\Ec$, then $\Ec$ is $\thD$-unsatisfiable; and if there exists
@@
 \end{array}\]
-The recursion will eventually terminate because each recursion adds one bound
+The recursion will eventually terminate because each recursive call adds one bound
 variable to $B$ and there are finitely many distinct bound variables in a $\mu$-term.
-Intuitively, the expansion of a subterm is a stand-alone
-$\mu$-term that denotes the same value as the original subterm---e.g.,
+Intuitively, the expansion of a subterm is a self-contained
+$\mu$-term that denotes the same value as the original subterm---\allowbreak e.g.,
 $\expand{\MU x.\: \const{C}(\MU y.\: \const{D}(x))}{\MU y.\: \const{D}(x)} =
@@
 is a \emph{self-similar subterm} of~$t$ if
-$u$ is a proper subterm of $t$,
+$u$ is a proper subterm of~$t$,
 $t$ and $u$ are of the forms $\MU x.\; \const{C}( t_1, \ldots, t_n )$
@@
 Thus, $t = \MU x.\; \const{C}( \MU y.\; \const{C}( y ) )$ is not normal
-because $\MU y.\; \const{C}( y )$ is a self-similar subterm of $t$.
+because $\MU y.\; \const{C}( y )$ is a self-similar subterm of~$t$.
 Their arguments have the same expansion with respect to~$t$:
@@
 its \emph{normal form} $\nf{t}$ is obtained
-by replacing all of the self-similar subterms of $t$ with $x$
+by replacing all of the self-similar subterms of~$t$ with $x$
 %and by replacing all of the abnormal subterms of $t$ with their corresponding normal forms.
@@
 $\interp{\J}{\tau}$ includes a maximal set of closed normal $\mu$-terms of that type that are
-unique up to $\alpha$-equivalence and acyclic if $\tau \notin \Codata$.\strut
+unique up to $\alpha$-equivalence and acyclic unless $\tau \in \Codata$.\strut
 
@@
 For each constructor term $\vphantom{s^k}\const{C}( \vec t \,)$ of type $\tau$,
-$\interp{\J}{\const{C}}( \interp{\J}{\vec t} )$ is the value
+$\interp{\J}{\const{C}}( \interp{\J}{\vec t\,} )$ is the value
 in $\interp{\J}{\tau}$ that is $\alpha$-equivalent to
-$\nf{\MU x.\; \const{C}( \interp{\J}{\vec t} )}$, where $x$ is fresh.\strut
+$\nf{\MU x.\; \const{C}( \interp{\J}{\vec t\,} )}$, where $x$ is fresh.\strut
 
@@
 
-%The last precondition is admittedly rather technical.
+%The last premise is admittedly rather technical.
 To ensure disequality literals are satisfied by an interpretation based on $\ValC$,
@@
 This invariant holds initially, and
-the last precondition in the above rule ensures that it is maintained.
+the last premise in the above rule ensures that it is maintained.
 %More precisely,
@@
 that, when assigned to ${\Varec{x}}$, will cause values in the range of $\ValC$ to become $\alpha$-equivalent.
-%This precondition is not strictly necessary in practice, 
+%This premise is not strictly necessary in practice, 
 %as one could compute $\nf{\ValC [ {\Varec{x}} \mapsto \MU {\Varec{x}}.\; t ]}$ optimistically and check
@@
 %If while doing so it returns to $\ec{t}$, it deduces that \rn{Acyclic} applies. % to $t$.
-%To recognize when the precondition of \rn{Unique} holds,
+%To recognize when the premise of \rn{Unique} holds,
 %the procedure considers the set $\SSS$ of all codatatype equivalence classes. It

### 2015-11-13 jour/jour.tex
more

@@
 \newcommand\xend{{\hfill$\scriptstyle\blacksquare$}}
+\newcommand\xqed{{\hfill$\scriptstyle\squareforqed$}}
 
@@
 that the type has at most one value.
-\qed
+\xqed
 \end{proof}
@@
 which is finite since $\tcEc$ is finite.
-\qed
+\xqed
 \end{proof}
@@
 and thus $\Ec$ is $\thD$-unsatisfiable.
-\qed
+\xqed
 \end{proof}
@@
 may cause values in the range of $\Val$ to become $\alpha$-equivalent,
-which in turn may cause $\Ec$ to be falsified by $\J$.
+which in turn may cause $\Ec$ to be falsified by~$\J$.
 For example, assign $\MU {\Varec{z}}.\; \const{D}( {\Varec{z}} )$ for ${\Varec{z}}$.
@@
 to ${\Varec{z}}$.
+
 Legal substitutions for ${\Varec{z}}$ may cause the range of $\Val$ to contain abnormal terms.
@@
 $\MU {\Varec{u}}.\; \const{C}( \MU {\Varec{z}}.\; \const{D}( {\Varec{u}} ) )$.
+This explains why the term is normalized in the rule's conclusion.
 \xend
@@
 To show (1), we first show that $\ValC$ contains no free variables.
-Assume by contradiction that $\ValC$ contained a free variable ${\Varec{y}}$ for some $\ec{y}$ of type $\tau$.
+Assume by contradiction that $\ValC$ contains a free variable ${\Varec{y}}$ for some $\ec{y}$ of type $\tau$.
 Then it must be the case that $\ec{y}$ does not contain a constructor term,
-or else ${\Varec{y}}$ would not occur as a free variable in $\Val$.
+or else ${\Varec{y}}$ would not occur as a free variable in~$\Val$.
 Consider the case when $\tau$ is finite.
@@
 each pair of constructor terms $\const{C}_{\negvthinspace j}( \vec t \,)$ and $\const{C}_{\negvthinspace\jPrime}( \vec u )$ in $\ec{z}$
-are such that $j = \jPrime$ since \rn{Clash} does not apply, % to $\Ec$,
-and are such that $\ec{ \vec t } = \ec{ \vec u }$ since \rn{Inject} does not apply. % to $\Ec$.
+are such that $j = \jPrime$, since \rn{Clash} does not apply, % to $\Ec$,
+and are such that $\ec{ \vec t \,} = \ec{ \vec u }$, since \rn{Inject} does not apply. % to $\Ec$.
 Thus,
@@
 when $\ec{z}$ contains a constructor $\const{C}_{\negvthinspace j}( t_1, \ldots, t_n )$.
-Say $\Val \ec{y}$ has a subterm $t$ where $\muvar(t) = {\Varec{x}}$.
+Suppose $\Val \ec{y}$ has a subterm $t$ such that $\muvar(t) = {\Varec{x}}$.
 Both $\Val \ec{x}$ and the subterm $t$ of $\Val \ec{y}$ were
@@
 Thus, we have
-$\expand{\Val \ec{y}}{ t } = \expand{\Val \ec{x}}{ \Val \ec{x} } = \Val \ec{x}$, and thus (2) holds for $\Val$.
+$\expand{\Val \ec{y}}{ t } = \expand{\Val \ec{x}}{ \Val \ec{x} } = \Val \ec{x}$, and %thus
+(2) holds for $\Val$.
 Property (3) holds for $\Val$ since \rn{Unique} does not apply.
@@
 We have that $\Val \ec{x}$ is of the form $\MU {\Varec{x}}.\; \const{C}( t_1, \ldots, t_n )$.
-Due to the construction of $\Val$,
-we know $\ec{x}$ contains a constructor $\const{C}( z_1, \ldots, z_n )$ and $\muvar( t_i ) = z_i$ for some $i$ % \in [n]$.
+Due to the construction of~$\Val$,
+we know $\ec{x}$ contains a constructor $\const{C}( z_1, \ldots, z_n )$ and $\muvar( t_i ) = z_i$ for some $i$. % \in [n]$.
 Since $\Val \ec{x}$ is a minimal, it
@@
 which are not $\alpha$-equivalent by (3), contradicting the fact that $\MU {\Varec{y}}.\; \const{C}( u_1, \ldots, u_n )$ is a self-similar subterm of $\Val \ec{x}$.
-Thus, $\Val = \nf{\Val}$, and (2) and (3) hold for $\nf{\Val}$.
+Thus, $\Val = \nf{\Val}$, and (2) and~(3) hold for $\nf{\Val}$.
 
@@
 $\ec{y}$ and $\ec{z}$ must have (minimal) subterms
-where $t_1$ occurs in $\Val_1 \ec{y}$ the same position (call it $p$) as $t_2$ occurs in $\Val_1 \ec{z}$,
+where $t_1$ occurs in $\Val_1 \ec{y}$ the same position~$p$ as $t_2$ occurs in $\Val_1 \ec{z}$,
 and $t_1 \not\aequiv t_2$.
@@
 then $\nf{\Val_1 \sigma} \ec{y}$ (resp. $\nf{\Val_1 \sigma} \ec{z}$)
-contains $t_1$ (resp. $t_2$) at position $p$,
+contains $t_1$ (resp. $t_2$) at position~$p$,
 and $\nf{\Val_1 \sigma} \ec{z}$ (resp. $\nf{\Val_1 \sigma} \ec{y}$) does not.
@@
 Thus, by assumption on the selection of $\MU {\Varec{x}}.\; t$,
-we have $\MU {\Varec{x}}.\; t \not\vsimv{\Varec{x}} \expand{\Val_1 \ec{y}}{t_1}$,
+we have $\smash{\MU {\Varec{x}}.\; t \not\vsimv{\Varec{x}} \expand{\Val_1 \ec{y}}{t_1}}$,
 which implies that the expansion of $\nf{\Val_1 \sigma} \ec{y}$ and $\nf{\Val_1 \sigma} \ec{z}$
@@
 we have that $\ValC$ satisfies (2) and (3).
-\qed
+\xqed
 \end{proof}
@@
 We consider a normal interpretation $\J$
-that interprets wrongly applied selectors based on equality information in $\Fc$
+that interprets wrongly applied selectors based on equality information in~$\Fc$
 and that interprets the variables of $\Fc$ based on the completion $\ValC$.
@@
 %where $\expand{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$ for $i \in [n]$.
-Since \rn{Inject} and \rn{Clash} do not apply to $\Fc\!$,
+Since \rn{Inject} and \rn{Clash} do not apply to~$\Fc\!$,
 by the construction of $\ValC$ we have that $\ValC \ec{t}$ is a term
@@
 we conclude that $\J \vDash \Ec$.
-\qed
+\xqed
 \end{proof}

### 2015-11-13 jour/jour.tex
more

@@
 examples.
-$\mathrm\Sigma$-terms are standard first-order terms
-over $\mathrm\Sigma$,
+$\Sig$-terms are standard first-order terms
+over $\Sig$,
 without $\mu$-binders.
@@
 The decision procedure is implemented as a \relax{theory solver} of CVC4---that is, a
-specialized solver for determining the satisfiability of conjunctions of
+specialized procedure for determining the satisfiability of conjunctions of
 literals for its theory.
@@
 incrementally builds partial assignments of
-truth values to the atoms of $F$ such that no clause in $F$ is falsified. We can regard such a
+truth values to the atoms of~$F$ such that no clause in $F$ is falsified. We can regard such a
 partial assignment as a set $M$ of true literals.
@@
 %at a high level,
-By a variant of the Nelson--Oppen method \cite{jovanovic2011sharing,nelson-oppen-1979},
+By a variant \cite{jovanovic2011sharing}
+of the Nelson--Oppen method \cite{nelson-oppen-1979},
 each $T_i$-solver %for $i = 1, \ldots, l$  n?
@@
 \begin{itemize}
-\item[2.2.1.] Let $\mathcalx{U}_i$ be a partition of $U_i$ such that $\ec{t} =_{\mathcalx{U}_i} \ec{u}$ for distinct $\ec{t}, \ec{u}$ if and only if
+\item[2.2.1.] Let $\mathcalx{U}_{\vthinspace i}$ be a partition of $U_i$ such that $\ec{t} =_{\mathcalx{U}_{\vthinspace i}} \ec{u}$ for distinct $\ec{t}, \ec{u}$ if and only if
 \begin{itemize}
@@
 \end{itemize}
-\item[2.2.2.] $\mathcalx{U}' := \mathcalx{U}' \cup \mathcalx{U}_i$.
+\item[2.2.2.] $\mathcalx{U}' := \mathcalx{U}' \cup \mathcalx{U}_{\vthinspace i}$.
 \end{itemize}
@@
 We then refine $\mathcalx{U}$ by 
-constructing a partition $\mathcalx{U}_i$ of each $U_i \in \mathcalx{U}$ such that distinct $\ec{t}$ and $\ec{u}$ reside in the same subset of $\mathcalx{U}_i$ 
+constructing a partition $\mathcalx{U}_{\vthinspace i}$ of each $U_i \in \mathcalx{U}$ such that distinct $\ec{t}$ and $\ec{u}$ reside in the same subset of $\mathcalx{U}_{\vthinspace i}$ 
 if and only if they contain constructor terms $\const{C}( t_1, \ldots, t_k )$ and $\const{C}( u_1, \ldots, u_k )$
@@
 the equivalence classes of $\Ec$ to closed normal $\MU$-terms.
-Recall that our construction of $\ValC$ requires picking values $\MU {\Varec{x}}.\; t$ for each $\Varec{x} \in \FV( \ValC )$
+Recall that our construction of $\ValC$ requires choosing values $\MU {\Varec{x}}.\; t$ for each $\Varec{x} \in \FV( \ValC )$
 such that $\MU {\Varec{x}}.\; t$ does not occur in the set $\BAD_{\smash{\J}}^{\Varec{x}}( \ValC )$.
-To pick values $\MU {\Varec{x}}.\; t$, we use a fair enumerator for each (co)datatype $\tau$,
+To choose values, % $\MU {\Varec{x}}.\; t$,
+we use a fair enumerator for each (co)datatype $\tau$,
 which lists all the values
@@
 %Concretely,
-For codatatypes, we enumerate a stream of \relax{every} $\MU$-term of our signature $\Sigma$ in a fair manner,
+For codatatypes, we enumerate a stream of \relax{every} $\MU$-term of our signature $\Sig$ in a fair manner,
 discarding those that have free variables or are not normal.
@@
 techniques~\cite{MouraBjoerner07,ReynoldsTinelliMoura14}.
-These techniques extend naturally to %quantified
+These techniques extend naturally to quantified
 formulas involving datatypes and codatatypes.
@@
 However, the presence of quantifiers poses an additional challenge in
-the context of (co)datatypes. Quantified formulas may entail an upper bound on
+the context of (co)datatypes. Quantified formulas can entail an upper bound on
 the cardinality of an uninterpreted type $\ty{u}$.
-When assuming that $\ty{u}$ has infinite cardinality, the calculus
+By assuming that uninterpreted types have infinite cardinality, the calculus
 presented in Section~\ref{sec:the-decision-procedure} is \relax{incomplete}
@@
 the solver is already incomplete for quantified formulas.
+
 Nonetheless, two techniques help increase the precision of the solver.
@@
 Second, we can conditionally apply \rn{Single} to codatatype terms that may have cardinality one.
-For example, the $\ty{stream}_{\vthinspace\ty{u}}$ codatatype %, where $\ty{u}$ is uninterpreted.
+For example, the codatatype $\ty{stream}_{\vthinspace\ty{u}}$ %, where $\ty{u}$ is uninterpreted.
 has cardinality one precisely when $\ty{u}$ has cardinality one.
@@
 The decision procedure for (co)datatypes is useful both for proving (via
-negation, in the refutational style) and finite model finding.
+negation, in the refutational style) and for finite model finding.
 It is in fact vital for finite model finding,

### 2015-11-13 jour/jour.tex
more

@@
 %  * consider a natural fragment---datatypes as supported in modern SMT solvers
-%    and the SMT-LIB 2 standard, and codatatypes as their duals
+%    and the SMT-LIB standard, and codatatypes as their duals
 %    * mutual recursion, but no polymorphism, nested recursion
@@
 (co)recursion through a non-(co)datatype (both of which make sense,
-at least in a higher-order setting \cite{blanchette-et-al-2014-codata}).}
+at least in a higher-order setting \cite{blanchette-et-al-2014-impl}).}
 We allow ourselves some notational parameterization
@@
 for several constructors associated with the same (co)data\-type,
-as long as the argument types coincide. % \cite[Section~3]{blanchette-et-al-2014-codata}.
+as long as the argument types coincide. % \cite[Section~3]{blanchette-et-al-2014-impl}.
 However, this is disallowed by SMT-LIB, so we do not consider it here.}
@@
 For each goal in each formalization, we used Sledgehammer to select either 16 or
-256~lemmas, which were monomorphized and translated to SMT-LIB~2 along with the
+256~lemmas, which were monomorphized and translated to SMT-LIB along with the
 goal. The resulting problem was given to the
@@
 more expensive rules, make a contribution.
-
+%
 For the 256-lemma problems, the difference
@@
 %obviously depends on the application.
-Table~\ref{tab:benchtwofiftysix} clearly indicates that
+Table~\ref{tab:benchtwofiftysix} indicates that
 the theoretically stronger decision

### 2015-11-16 jour/jour.tex
minor edits

@@
 in the generated models. For example,
-the $\mu$-term $\const{SCons}(1,\: \MU z.\; \const{SCons}(0,\: \const{SCons}(9,\: z)))$
+the $\mu$-term $\const{SCons}(1,\: \MU s.\; \const{SCons}(0,\: \const{SCons}(9,\: s)))$
 stands for the lasso-shaped sequence $1, 0, 9, 0, 9, 0, 9,
@@
 An earlier version of this article was presented at the CADE-25 conference in
-Berlin, Germany. This article extends the conference paper with additional
+Berlin, Germany \cite{reynolds-blanchette-2015-codata}.
+This article extends the conference paper with additional
 background material on (co)datatypes (Section~\ref{sec:co-datatypes}), more
@@
 and a return type $\tau$, written
-$\const{f} : \tau_1 \times \cdots \times \tau_n \to \tau$,
-which collapses to $\const{f} : \tau$ if $n = 0$.
+$\const{f} : \tau_1 \times \cdots \times \tau_n \to \tau$;
+this notation collapses to $\const{f} : \tau$ if $n = 0$.
 %
@@
 apply the $n$-ary function symbol
-\const{f} to $n$ well-typed arguments $t_1 \mathbin{:} \tau_1$, \ldots, $t_n \mathbin{:}
-\tau_n$. Nullary function symbols, also called constants, can appear without
+\const{f} to $n$ %well-typed
+arguments $t_1 \mathbin{:} \tau_1$, \ldots, $t_n \mathbin{:}
+\tau_n$ of the right types.
+Nullary function symbols, also called constants, can appear without
 parentheses in terms.
@@
 $\const{true},\, \const{false} : \ty{bool}$, interpreted as truth values.
-The only predicate is equality ($\teq$) and belongs to the logical symbols;
-other predicates can be represented as functions to $\ty{bool}$,
+The only predicate is equality, written $\teq$;
+it belongs to the logical symbols.
+Other predicates can be represented as functions to $\ty{bool}$,
 with $\const{p}(\ldots)$ abbreviating $\const{p}(\ldots) \teq \const{true}$.
 The notation $t\typ{\tau}$ stands for a term $t$ of type~$\tau$;
-it should not be confused with the notation~$t :\nobreak \tau$,
+it should not be confused with $t :\nobreak \tau$,
 which is a statement expressing that $t$ has type $\tau$.
-When applied to terms, the symbol $=$ denotes syntactic equality.
+When applied to terms, the standard equality symbol~$=$ denotes syntactic equality.
 The operator $\bigwedge_{\vthinspace i}\, \varphi_i$ abbreviates a conjunction
@@
 fall outside this fragment.\footnote{In principle, rank-1 (top-level) polymorphism \cite{blanchette-paskevich-2013}
-should not raise any special difficulties. Nesting datatypes inside datatypes,
+does not raise any special difficulties. Nesting datatypes inside datatypes,
 and likewise for codatatypes, can be reduced to the mutual case
@@
 \end{align*}
-the conjecture
+on streams, the conjecture
 $\const{zeros} = \const{sconst}~0$
-would be ``refuted'' by a spurious countermodel where
-\const{zero} and $\const{sconst}~0$ are interpreted by two distinct
+would be ``refuted'' by a spurious countermodel that interprets
+\const{zero} and $\const{sconst}~0$ by two distinct
 values that both correspond to the \hbox{$\mu$-term}
@@
 
-For each goal in each formalization, we used Sledgehammer to select either 16 or
+For each proof goal in each formalization, we used Sledgehammer to select either 16 or
 256~lemmas, which were monomorphized and translated to SMT-LIB along with the
@@
 Z3 4.3.2 for comparison, each running for
-up to 60~s on the StarExec cluster
-\cite{stump-et-al-2014-starexec}.
+up to 60~s on StarExec \cite{stump-et-al-2014-starexec}.
 Problems not involving any (co)datatypes were left out.
@@
 Moreover, every aspect of the procedure, including the
-more expensive rules, make a contribution.
+more expensive rules, makes a contribution.
 %
@@
 Table~\ref{tab:benchtwofiftysix} indicates that
-the theoretically stronger decision
-procedures do not always subsume the weaker ones in practice.
-The raw data reveal that the decision procedure proved 27~goals that
+the theoretically stronger instances of the decision
+procedure do not always subsume the weaker ones in practice.
+The raw data reveal that the full procedure proved 27~goals that
 could not be proved without it, but failed for 17~goals that could be proved
@@
 the acyclicity rule and 17~required uniqueness. Interestingly,
-no proofs were lost by enabling these rules. Among 17 proofs requiring uniqueness,
+no proofs were lost by enabling these rules. Among the 17 proofs requiring uniqueness,
 some were simple arguments of the form
@@
 in Isabelle \cite{blanchette-et-al-2014-impl}, while others
-involved more elaborate reasoning, including the following example:
+involved more elaborate reasoning, including the following example
+about Stern--Brocot trees:
 %
@@
 \noindent
-where \const{\nummodden} is defined as $\const{\nummodden} =
+The tree \const{\nummodden} is defined as $\const{\nummodden} =
 \const{Node}\;0\;\const{num}\;\const{\nummodden}$.
@@
 We introduced a decision procedure for the universal theory of datatypes and
-codatatypes. Our principal
+codatatypes. Our main
 contribution has been the support for codatatypes. Both the metatheory and
 the implementation in CVC4 rely on $\mu$-terms to represent cyclic values.
-Although this aspect is primarily motivated by codatatypes, it enables
-a uniform account of datatypes and codatatypes---in particular,
+Although this aspect is primarily motivated by codatatypes, it makes
+a uniform account of datatypes and codatatypes possible---in particular,
 the acyclicity rule for datatypes exploits $\mu$-terms to detect cycles.

### 2015-11-25 jour/jour.tex
more

@@
 (satisfiability modulo theories)
-solvers, has been extended with a syntax for datatypes.
+solvers, is being extended with a syntax for datatypes.
 In this article, we introduce a
@@
 Third, \emph{it must be possible to express cyclic }(\emph{regular}) \emph{values as closed terms and
-to enumerate them.} This is necessary both for finite model finding
-modulo theories \cite{reynolds-et-al-2013,ge-de-moura-2009}
-and for theory combinations. The $\mu$-binder notation associates a name with
+to enumerate them.} This is necessary when generating models.
+The $\mu$-binder notation associates a name with
 a subterm; it is used to represent cyclic values in the metatheory and
@@
 the $\mu$-term $\const{SCons}(1,\: \MU s.\; \const{SCons}(0,\: \const{SCons}(9,\: s)))$
-stands for the lasso-shaped sequence $1, 0, 9, 0, 9, 0, 9,
+stands for the lasso-shaped sequence $1, 0, 9, 0, 9, %0, 9,
 \ldots$\afterLdots.
@@
 The decision procedure for (co)datatypes is useful both for proving (via
-negation, in the refutational style) and for finite model finding.
+negation, in the refutational style) and for model finding
+\cite{reynolds-et-al-2013,ge-de-moura-2009}.
 It is in fact vital for finite model finding,

### 2015-11-25 jour/jour.tex
Andy's suggestions

@@
 superposition prover by Wand \cite{wand-2014}.
+In his Ph.D.\ thesis \cite{bjoerner-1998}, Bj\o rner introduced a decision procedure for
+(co)datatypes in STeP, the Stanford Temporal Prover.
 Closely related are the automatic structural
@@
 al.\ address this by parameterizing the construction by default values, but
-this gives rise to spurious equalities between unrelated terms---e.g.,
-$\const{s}_1^1(\const{C}_2) \teq \const{s}_1^1(\const{C}_3)$. This flaw
+this gives rise to spurious equalities between unrelated terms. For example,
+given
+\[\keyw{datatype}\;\, \ty{x} \,=\, \const{C}(\const{s}{:}\; \ty{int}) \mid \const{D} \mid \const{E}\]
+we would have the spurious equality
+$\const{s}(\const{D}) \teq \const{s}(\const{E})$. This flaw
 could be corrected, but the added complexity seems to suggest that selectors
@@
 For example, $\ty{stream}_{\vthinspace\ty{int}}$ is isomorphic to the uncountable
-function space $\ty{nat} \to \ty{int}$.
+function space $\ty{nat} \to \ty{int}$. Bj\o rner gives a rigorous treatment of
+this aspect \cite{bjoerner-1998}.
 
@@
 Thus, in the subsequent phases, 
-when considering the case that $\ec{t}$ contains a constructor term,
+when considering the case that $\ec{t}$ contains constructor terms,
 it is enough %for our purposes
@@
 the equivalence class $\ec{t}$.
-Initially, we set $\Val\ec{u} := {\Varec{u}}$ for each equivalence class $\ec{u}$. Because $\Varec{u}$ is unspecified,
+Initially, we set $\Val\ec{u} := {\Varec{u}}$ for each equivalence class $\ec{u}$. Because $\Varec{u}$ is
+unconstrained,
 this indicates that there are initially no
@@
 node $\Ec \mathrel\cup t \teq u$ is obtained by applying one of the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, \rn{Inject}, \rn{Unique}, or \rn{Single}.
-In all these cases, $\Ec \vDash_{\thD} t \teq u$.
+In each of these cases, $\Ec \vDash_{\thD} t \teq u$.
 For \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, this is a consequence of equality reasoning.
@@
 
-We first show that (2) and (3) hold initially for $\Val$.
+We now show that (2) and (3) hold initially for $\Val$.
 For all equivalence classes $\ec{z}$,
@@
 
-Since by assumption $\Fc$ contains only equalities and disequalities, we have $\J \vDash \Fc\!$,
+Since $\Fc$ contains only equalities and disequalities, we have $\J \vDash \Fc\!$,
 and since $\Ec \subseteq \Fc\!$,

### 2015-11-25 jour/jour.tex
Inline lemma.

@@
 
-\begin{lemma}
-\label{lem:mu-norm-arg}
-If the $\mu$-terms $\vec u$ are in normal form and $t = \nf{\MU x.\; \const{C}( \vec u )} = \MU x.\; \const{C}( \vec w )$
-for some $\mu$-terms $\vec w$,
-then $\vec u \vsim \expand{t}{\vec w}$\vthinspace.
-\end{lemma}
+%\begin{lemma}
+%\label{lem:mu-norm-arg}
+%If the $\mu$-terms $\vec u$ are in normal form and $t = \nf{\MU x.\; \const{C}( \vec u )} = \MU x.\; \const{C}( \vec w )$
+%for some $\mu$-terms $\vec w$,
+%then $\vec u \vsim \expand{t}{\vec w}$\vthinspace.
+%\end{lemma}
 % TODO:
@@
 For each $i$, let $u_\iPrime$ be the $i$\vvthinspace th argument of $t'$.
-By Lemma~\ref{lem:mu-norm-arg},
-$\expand{t'}{u_\iPrime} \vsim \ValC \ec{u_i}$.
-Thus $\expand{t'}{u_\iPrime} \vsim \expand{\ValC \ec{t}}{ w_i }$. % for each $i \in [n]$.
+%By Lemma~\ref{lem:mu-norm-arg},
+We have that
+$\expand{t'}{u_\iPrime} \vsim \ValC \ec{u_i}$, and
+thus $\expand{t'}{u_\iPrime} \vsim \expand{\ValC \ec{t}}{ w_i }$. % for each $i \in [n]$.
 %By Lemma~\ref{lem:mu-cong}, we have $t' \vsim t_w$.

### 2016-02-06 sister/sister.tex
added sister conf

--- 2015-11-25 jour/jour.tex
@@
 
-\newcommand\gandl{G\&L}
+\newcommand\gandl{SBT}
 \newcommand\HD[1]{\hbox to2em{\footnotesize\hfill\!\!\!\!#1\!\!\!\!\hfill}}
@@
 that define codatatypes falling within the supported fragment. We
-also included unpublished formalizations about Bird and Stern--Brocot trees
-by Peter Gammie and Andreas Lochbihler (\gandl,
-265~goals). To exercise the datatype support, formalizations about finite
+also included formalizations about Bird and Stern--Brocot trees (\gandl,
+265~goals) \cite{gammie-lochbihler-2016}. To exercise the datatype support, formalizations about finite
 lists and trees
@@
 {
+\footnotesize
 \def\ackname{Acknowledgment}
-\paragraph{\ackname.}
+\paragraph{\footnotesize\ackname.}
 \looseness=-1
@@
 %
-Peter Gammie and Andreas Lochbihler provided useful benchmarks.
+Peter Gammie and Andreas Lochbihler pointed us to useful benchmarks.
 Andrei Popescu helped clarify our thoughts regarding codatatypes and indicated

### 2016-02-08 sister/sister.tex
submitted first version
more

@@
-%%% TODO: Put http://lara.epfl.ch/~reynolds/JAR-CADE2015-cdt/ online
-%%% TODO: Mealy machine minimization
-%%% TODO: Give small example where Single is needed
-
-%%% GOOD IDEAS in impl of codata models:
-%%% 1. de Bruijn
-%%% 2. enum: "X nat"
-%%% 3. norm by DFA minim
-%%% 4. check membership in bad set by matching
-
-\documentclass[smallcondensed,draft]{svjour3}
+\documentclass[letter]{article}
+% The file ijcai16.sty is the style file for IJCAI-16 (same as ijcai07.sty).
+\usepackage{ijcai16}
 
@@
 \usepackage{amssymb}
+\usepackage{amsthm}
 \usepackage[scaled=.82]{beramono}
@@
 \usepackage{bussproofs}
-\usepackage{cite}
+%\usepackage{cite}
 \usepackage{color}
@@
 
-\spnewtheorem{examplex}{Example}{\itshape}{\rmfamily}
+\newtheorem{examplex}{Example}%{\itshape}{\rmfamily}
+\newtheorem{lemma}{Lemma}
+\newtheorem{theorem}{Theorem}
+\newtheorem{definition}{Definition}
 
@@
 \newcommand\xend{{\hfill$\scriptstyle\blacksquare$}}
-\newcommand\xqed{{\hfill$\scriptstyle\squareforqed$}}
+\newcommand\xqed{{\hfill$\scriptstyle\square$}}
 
@@
 
-\author {Andrew Reynolds \and Jasmin Christian Blanchette\thanks{In memoriam Morgan Deters 1979--2015}}
-\authorrunning {A. Reynolds \and J. C. Blanchette}
-
-\institute{
-Andrew Reynolds
-\at \'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL), Switzerland
-\and
-Jasmin Christian Blanchette%, funded by DFG grant \textit{Hardening the Hammer} (Ni\,491\slash 14-1)
-\at Inria \& LORIA, Nancy, France \\
-Max-Planck-Institut f\"ur Informatik, Saarbr\"ucken, Germany}
+\author {Andrew Reynolds \and Jasmin Christian Blanchette\thanks{In memoriam Morgan Deters 1979--2015} \\
+Department of Computer Science, The University of Iowa, USA \and \\
+Inria Nancy -- Grand Est \& LORIA Villers-l\`es-Nancy, France
+}
+
+%\authorrunning {A. Reynolds \and J. C. Blanchette}
+
+%\institute{
+%Andrew Reynolds
+%\at \'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL), Switzerland
+%\and
+%Jasmin Christian Blanchette%, funded by DFG grant \textit{Hardening the Hammer} (Ni\,491\slash 14-1)
+%\at Inria \& LORIA, Nancy, France \\
+%Max-Planck-Institut f\"ur Informatik, Saarbr\"ucken, Germany}
 
 \maketitle
+
+\begin{abstract}
+New abstract.
+\end{abstract}
+
+\section{Introduction}
+\label{sec:introduction}
+
+  * a general introduction to automatic theorem proving \\
+
+  * relation to AI
+   * among the earliest AI goals
+   * yet somewhat disconnected from mainstream AI
+     (with some exceptions, e.g. Urban ..., MaSh)
+
+  * followed by an intro to SMT and what it's used for \\
+  * followed to why theories are important (more efficient reasoning) \\
+  * followed to why (co)datatypes are important
+
+  * both for proving
+   * not finitely axiomatizable
+  * and for model finding
+   * Nunchaku
+
+\section{%The Theory of
+(Co)datatypes}
+\label{sec:co-datatypes}
+
+xxx
+
+\section{The Decision Procedure} % for (Co)datatypes}
+\label{sec:the-decision-procedure}
+
+\section{Evaluation} % on Isabelle Problems}
+\label{sec:evaluation}
+
+\section{Conclusion}
+\label{sec:conclusion}
+
+\newpage
+
+X
+
+\newpage
 
@@
 \end{abstract}
-
-%% The institutions above shouldn't count as footnotes
-\setcounter{footnote}{0}
 
@@
 
-\bibliographystyle{spmpsci}
+\bibliographystyle{named}
 \bibliography{bib}{}

### 2016-02-08 sister/sister.tex
push
put text in

@@
 \begin{abstract}
-New abstract.
-\end{abstract}
-
-\section{Introduction}
-\label{sec:introduction}
-
-  * a general introduction to automatic theorem proving \\
-
-  * relation to AI
-   * among the earliest AI goals
-   * yet somewhat disconnected from mainstream AI
-     (with some exceptions, e.g. Urban ..., MaSh)
-
-  * followed by an intro to SMT and what it's used for \\
-  * followed to why theories are important (more efficient reasoning) \\
-  * followed to why (co)datatypes are important
-
-  * both for proving
-   * not finitely axiomatizable
-  * and for model finding
-   * Nunchaku
-
-\section{%The Theory of
-(Co)datatypes}
-\label{sec:co-datatypes}
-
-xxx
-
-\section{The Decision Procedure} % for (Co)datatypes}
-\label{sec:the-decision-procedure}
-
-\section{Evaluation} % on Isabelle Problems}
-\label{sec:evaluation}
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-\newpage
-
-X
-
-\newpage
-
-\begin{abstract}
 %Codatatypes naturally capture potentially infinite data structures and
@@
 
-\section{Introduction}
+\section{Introduction -- JB 2 columns}
 \label{sec:introduction}
+
+%  * a general introduction to automatic theorem proving \\
+
+%  * relation to AI
+%   * among the earliest AI goals
+%   * yet somewhat disconnected from mainstream AI
+%     (with some exceptions, e.g. Urban ..., MaSh)
+%
+%  * followed by an intro to SMT and what it's used for \\
+%  * followed to why theories are important (more efficient reasoning) \\
+%  * followed to why (co)datatypes are important
+%
+%  * both for proving
+%   * not finitely axiomatizable
+%  * and for model finding
+%   * Nunchaku
 
@@
 
-%  * useful both for proving and for model finding
-
-%\ref{sec:examples}
-
-%  * benchmarking is often an issue -- esp. codatatypes
-%
-%Polymorphic types, nested (co)recursion, and datatype--codatatype mixtures fall
-
-%  * setting: FOL
-%    * core procedure is restricted to ground
-%      * but theory solver cooperates
-
-%  * codatatypes were added later, motivated by the use of SMT solvers as
-%    backends to proof assistants (more specifically, CVC4 to Isabelle/HOL)
-
-%  * setting :
-%    * universal formulas
-%    * many-sorted logic
-%   * mutually (co)recursive types with constructors, selectors, and
-%      discriminators
-
-%  * codatatypes: from a theoretical and implementational point of view, like
-%      datatypes but:
-%    * infinite values (infinitely many nested constructors)
-%    * codatatypes are never empty (e.g. finite streams are rejected)
-
-%\[
-%      \keyw{codatatype}\; \,\ty{llist} \,=\, \const{LNil} \,\mid\, \const{LCons}(\ty{int},\: \ty{llist})
-%\]
-
-%  * consider a natural fragment---datatypes as supported in modern SMT solvers
-%    and the SMT-LIB standard, and codatatypes as their duals
-%    * mutual recursion, but no polymorphism, nested recursion
-%  * integrated with Nelson--Oppen
-
-% * SMT-Lib
-
-%  * one implication is that if $m$ equals $ES(n)$ and $n$ equals $ES(m)$, necessarily
-%    $m$ and $n$ must be equal.
-
-%  * perhaps the most commonly used codatatype is that of lazy lists or sequences.
-%    using a syntax similar to Standard ML, Haskell, or SMT-LIB
-
-%  * less briefly: codatatypes and why they are useful
-%    * in Agda, Coq, Matita -- recently also in Isabelle/HOL \cite{nipkow-et-al-2002}
-%    * but not in SMT-LIB 2 \cite{barrett-et-al-2010}
-
-\paragraph{Related Work.} Barrett
-et al.\ \cite{barrett-et-al-2007} provide a good account of
-related work on datatypes as of 2007, in addition to describing their
-implementation in CVC3.
-Since then, datatypes have been added not only to CVC4 (a complete rewrite of CVC3)
-but also to the SMT solver
-Z3 \cite{de-moura-bjoerner-2008} in unpublished work by Leonardo de Moura and to a SPASS-like %prototypical
-superposition prover by Wand \cite{wand-2014}.
-In his Ph.D.\ thesis \cite{bjoerner-1998}, Bj\o rner introduced a decision procedure for
-(co)datatypes in STeP, the Stanford Temporal Prover.
-Closely related are the automatic structural
-% cruanes-201x,wand-weidenbach-201x,  -- both were rejected
-induction in SMT solvers
-\cite{reynolds-kuncak-2014}
-and superposition provers\cite{kersani-peltier-2013,cruanes-2015},
-the (co)datatype and (co)induction support in Dafny \cite{leino-moskal-2014},
-and the (\hbox{semi-})decision procedures for datatypes
-implemented in Leon \cite{suter-et-al-2011} and RADA \cite{pham-et-al-2013}.
-%
-Datatypes are supported by the higher-order model finder Refute
-\cite{weber-2008} for Isabelle. Its successor, Nitpick
-\cite{blanchette-nipkow-2010}, can also generate models involving cyclic
-codatatype values. Cyclic values have been studied extensively under the heading
-of regular or rational trees---see Carayol and Morvan
-\cite{carayol-morvan-2006} and Djellou et al.\ \cite{djelloul-et-al-2008}
-for recent work.
-The $\mu$-notation is inspired by the $\mu$-calculus
-\cite[etc.]{kozen-1983,endrullis-et-al-2011}.
-
-%, notably in the context of constraint logic programming
-
-%  * Ro\c{s}u --- equality of streams is...
-%     http://dl.acm.org/citation.cfm?doid=1159803.1159827
-
-%     * point to Barrett et al. for SMT datatypes
-%       * about their own work, they say: "our focus is on generality and
-%         efficiency rather than immediacy of implementation"
-%       * (deal more directly with finite sorts than Barrett et al., Section 6.1)
-%     * additional ones since then (e.g. strings?)
-%     * anything about codatatypes?
-%       * proof assistants like Agda, Coq, etc. have them
-%       * Dafny, CoALP
-%       * also a lot of theoretical research, some of which is loosely connected,
-%         e.g. decision procedure for corecursive functions (Henning in Nijmegen)
-%
-%     * ODDITY: Oppen 1980: single-constructor, recursive -- infinite values?
-
 \paragraph{Conventions.}
@@
 
-%* although nothing prevents composing the decision procedure with theories
-%  providing polymorphic types (parametric sorts), such as for arrays (e.g., $\ty{array}(\alpha,\beta)$)
-
 \section{%The Theory of
-(Co)datatypes}
+(Co)datatypes -- AJR 0.75 column}
 \label{sec:co-datatypes}
-
-%    (terminology: freely-generated, inductive, algebraic, ..., sometimes
-%    with different meanings; we'll clarify below what we use)
 
@@
 
-%  * assume for simplicity no indirect recursion, but this does not radically
-%    change the argument
-%  * assume there are at least two values built with C.
-%    they must be different at some point, e.g.
-%    C(0, C(1, C(0, ...))) and C(0, C(1, C(1, ...)))
-%    can use that to create infinitely alternation patterns
-%  * leaves us with the case of a single constructor
-% * since we are looking at a specific model, we can assume all uninterpreted
-%   types are finite datatypes with nullary constructors corresponding to their
-%   elements
-% * if two
-% * either there is only one ``path'' from ctr to itself
-%   or at least two;
-
-\section{The Decision Procedure} % for (Co)datatypes}
+\section{The Decision Procedure -- AJR 4 columns} % for (Co)datatypes}
 \label{sec:the-decision-procedure}
@@
 
-\section{Implementation in CVC4}
-\label{sec:implementation-as-a-theory-solver-in-cvc4}
-
-The decision procedure was presented at a high level of abstraction, omitting
-quite a few details. This section describes the main aspects of the
-implementation within the SMT solver CVC4:\ the integration of the
-procedure into CDCL($T$)
-\cite{ganzinger-et-al-2004}, the construction of models with $\mu$-terms,
-and the extension of the procedure to quantified formulas.
-%, and some of the optimizations.
-
-\subsection{A Theory Solver for $\thD$}
-
-The decision procedure is implemented as a \relax{theory solver} of CVC4---that is, a
-specialized procedure for determining the satisfiability of conjunctions of
-literals for its theory.
-Given a theory $T = T_1 \mathrel\cup \cdots \mathrel\cup T_n$ and a set of
-input clauses $F$ in conjunctive normal form, the CDCL($T$) procedure
-incrementally builds partial assignments of
-truth values to the atoms of~$F$ such that no clause in $F$ is falsified. We can regard such a
-partial assignment as a set $M$ of true literals.
-% where $a \in M$ for all atoms $a$ that
-%it assigned the value $\top$ and $\neg\: a \in M$ for all atoms that it assigns
-%the value $\bot$.
-%at a high level,
-By a variant \cite{jovanovic2011sharing}
-of the Nelson--Oppen method \cite{nelson-oppen-1979},
-each $T_i$-solver %for $i = 1, \ldots, l$  n?
-takes as input the union $M_i$
-of
-\begin{itemize}
-\item the purified form of $T_i$-literals occurring in $M$, where fresh
-variables replace terms containing symbols not belonging to~$T_i$;
-\item additional (dis)equalities between variables of types not belonging to
-$T_i$.
-\end{itemize}
-Each $T_i$-solver
-either reports that a subset $C$ of $M_i$ is $T_i$-unsatisfiable, in which case
-$\neg\: C$ is added to $F\!$, adds a clause to $F\!$, or does nothing.
-When $M$ is a complete assignment for $F\!$, a theory solver can choose to do
-nothing only if $M_i$ is indeed $T_i$-satisfiable.
-
-Assume $\Ec$ is initially the set $M_i$ described above. With each equality $t \teq u$ added
-to $\Ec$, we associate a set of equalities from $M_i$ that together entail $t
-\teq u$, which we call its \emph{explanation}.
-Similarly, each $\Val \ec{x}$ is assigned an explanation---that is, a set
-of equalities from $M_i$ that entail that the values of $\ec{x}$ in models of
-$\Ec$ are of the form $\Val \ec{x}$. For example, if $x \teq \const{C}( x ) \in
-M_i$, then $x \teq \const{C}( x )$ is an %(possible)
-explanation for $\Val \ec{x} = \MU {\Varec{x}}.\; \const{C}( {\Varec{x}} )$.
-The rules of the calculus are implemented as follows. For all rules with
-conclusion $\bot$, we report the union of the explanations for all premises is
-$\thD$-unsatisfiable. For \rn{Split}, we add the exhaustiveness clause %of the form
-$\DISC{1} \mathrel\lor \cdots \mathrel\lor \DISC{m}$ to $F$.
-Decisions on which branch to take are thus performed externally by the SAT
-solver. % and not by the theory solver.
-All other rules add equalities to the
-internal state of the theory solver. The rules in phase~1 are performed
-eagerly---that is, for partial satisfying assignments $M$---while the rules in
-phases~2 and~3 are performed only for complete satisfying
-assignments $M$.
-
-\begin{figure}[t!]
-\normalsize
-\centering
-\flushleft $\rn{Acyclic}(t)$:
-
-\vspace*{-\topsep}
-
-\begin{enumerate}
-\item Let $U$ be the datatype equivalence classes of $\Ec$.
-\item Repeat until $U \neq \emptyset$:
-\begin{itemize}
-\item[\ ] For some $\ec{t} \in U$, $\rn{Traverse}( \ec{t}, U, \emptyset )$.
-\end{itemize}
-\end{enumerate}
-
-\flushleft $\rn{Traverse}( \ec{t}, U, P )$:
-
-\vspace*{-\topsep}
-
-\begin{enumerate}
-\item If $\ec{t} \in P$, $\rn{Acyclic}(t)$.
-\item If $\ec{t} \in U$:
-\begin{enumerate}
-\item[2.1.] If $\const{C}( t_1, \ldots, t_k ) \in \ec{t}$, then for each $j = 1, \ldots k$:
-\begin{itemize}
-\item[2.1.1.] $\rn{Traverse}( \ec{t_j}, U, P \cup \{ \ec{t} \} )$.
-\end{itemize}
-\item[2.2.] $U := U \textbackslash \{ \ec{t} \}$.
-\end{enumerate}
-\end{enumerate}
-\vspace*{-\topsep}
-\caption{\,Algorithm for checking \rn{Acyclic}}
-\label{fig:acyclic-algo}
-\end{figure}
-
-\begin{figure}[t!]
-\normalsize
-\centering
-
-\flushleft $\rn{Unique}(t, u)$:
-
-\vspace*{-\topsep}
-
-\begin{enumerate}
-\item Let $\mathcalx{U}$ be a partition of the equivalence classes of $\Ec$ such that $\ec{t} =_{\mathcalx{U}} \ec{u}$ for distinct $\ec{t}, \ec{u}$ if and only if
-  $t$, $u$ have type $\delta \in \Codata$,
-  $\const{C}( t_1, \ldots, t_k ) \in \ec{t}$,
-  and $\const{C}( u_1, \ldots, u_k ) \in \ec{u}$.
-
-\item Repeat until $\mathcalx{U}$ is unchanged:
-\begin{enumerate}
-\item[2.1.] $\mathcalx{U}' := \emptyset$.
-\item[2.2.] For each $U_i \in \mathcalx{U}$:
-\begin{itemize}
-\item[2.2.1.] Let $\mathcalx{U}_{\vthinspace i}$ be a partition of $U_i$ such that $\ec{t} =_{\mathcalx{U}_{\vthinspace i}} \ec{u}$ for distinct $\ec{t}, \ec{u}$ if and only if
-\begin{itemize}
-\item $\const{C}( t_1, \ldots, t_k ) \in \ec{t}$ and $\const{C}( u_1, \ldots, u_k ) \in \ec{u}$, and
-\item $\ec{t_j} =_{\mathcalx{U}} \ec{u_j}$ for each $j = 1, \ldots, k$.
-\end{itemize}
-\item[2.2.2.] $\mathcalx{U}' := \mathcalx{U}' \cup \mathcalx{U}_{\vthinspace i}$.
-\end{itemize}
-\item[2.3.] $\mathcalx{U} := \mathcalx{U}'$.
-\end{enumerate}
-\item If distinct $\ec{t}, \ec{u} \in U$ for some $U \in \mathcalx{U}$, $\rn{Unique}(t,u)$.
-\end{enumerate}
-\vspace*{-\topsep}
-\caption{\,Algorithm for checking \rn{Unique}}
-\label{fig:uniqueness-algo}
-\end{figure}
-
-Before constructing a model for $F\!$, the theory solver constructs neither $\mu$-terms nor the mapping $\Val$.
-Instead, it relies on the algorithms in Figs.~\ref{fig:acyclic-algo} and~\ref{fig:uniqueness-algo}
-for determining whether the rules \rn{Acyclic} and \rn{Unique} apply to the set $\Ec$.
-
-For \rn{Acyclic}, Fig.~\ref{fig:acyclic-algo} considers the set $U$ of all datatype equivalence classes of $\Ec$.
-We choose an arbitrary $\ec{t}$ in $U$ and call the recursive subprocedure $\rn{Traverse}$,
-which takes as input the current equivalence class we are processing, the set of equivalence classes $U$ we have yet to process,
-and the set of equivalence classes $P$ we are currently processing.
-If a call to $\rn{Traverse}( \ec{t}, U, P )$ is such that $\ec{t} \in P$,
-we know that the rule \rn{Acyclic} applies to $t$.
-Otherwise, if we have yet to process $\ec{t}$, as indicated by the condition $\ec{t} \in U$,
-then if $\ec{t}$ contains a constructor term $\const{C}( t_1, \ldots, t_k )$, 
-then we recursively call $\rn{Traverse}$ on each of $\ec{t_1}, \ldots, \ec{t_n}$.
-If this succeeds, we remove $\ec{t}$ from $U$.
-
-For \rn{Unique}, Fig.~\ref{fig:uniqueness-algo} considers a partition $\mathcalx{U}$ of the equivalence classes of our set $\Ec$
-such that codatatype equivalence classes having a constructor term with top symbol $\const{C}$ are placed in the same subset, 
-for each constructor $\const{C}$.
-We write $\ec{t} =_{\mathcalx{U}} \ec{u}$ to denote that $\ec{t}$ and $\ec{u}$ reside in the same subset within $\mathcalx{U}$.
-We then refine $\mathcalx{U}$ by 
-constructing a partition $\mathcalx{U}_{\vthinspace i}$ of each $U_i \in \mathcalx{U}$ such that distinct $\ec{t}$ and $\ec{u}$ reside in the same subset of $\mathcalx{U}_{\vthinspace i}$ 
-if and only if they contain constructor terms $\const{C}( t_1, \ldots, t_k )$ and $\const{C}( u_1, \ldots, u_k )$
-where each $t_j, u_j$ are such that $\ec{t_j}$ and $\ec{u_j}$ reside in the same subset of $\mathcalx{U}$ for each $j=1, \ldots, k$.
-Let $\mathcalx{U'}$ be the union of each of these partitions.
-We update $\mathcalx{U}$ to $\mathcalx{U'}$.
-This refinement is repeated until $\mathcalx{U}$ is left unchanged.
-If the resulting partition $\mathcalx{U}$ contains a subset $U$ having two distinct equivalence classes $\ec{t}$ and $\ec{u}$, then \rn{Unique} applies to $t$ and $u$.
-This algorithm is analogous to Hopcroft's algorithm for minimizing
-deterministic finite automata \cite{hopcroft-1971}---terms correspond to
-states, argument positions $1,\ldots,k$ correspond to input symbols, and the top constructor symbol
-of a term generalizes the accepting/rejecting status of a state.
-
-%Instead, $\Val$ is computed implicitly by traversing the equivalence classes of $\Ec$ during phase~2.
-%To detect whether \rn{Acyclic} applies,
-%the procedure considers each equivalence class $\ec{t}$ containing a datatype constructor $\const{C}( t_1, \ldots, t_n )$.
-%It visits $\ec{t_1}, \ldots, \ec{t_n}$ and all constructor arguments in these equivalence classes recursively.
-%If while doing so it returns to $\ec{t}$, it deduces that \rn{Acyclic} applies. % to $t$.
-%To recognize when the premise of \rn{Unique} holds,
-%the procedure considers the set $\SSS$ of all codatatype equivalence classes. It
-%simultaneously visits the equivalence classes of arguments of constructor terms in each equivalence class in $\SSS$,
-%while partitioning $\SSS$ into $\SSS_1, \ldots, \SSS_n$ based on the top-most symbol of constructor terms in these equivalence classes
-%and the equivalence of their arguments of ordinary types.
-%It then partitions each set %in this partition %in the same manner
-%recursively.
-%If %after doing so,
-%the resulting partition contains a set $\SSS_i$ containing distinct terms $u$
-%and $v$, it deduces that \rn{Unique} applies to $u$ and~$v$.
-
-The implementation of the decision procedure uses several optimizations
-following the lines of Barrett et al.\ \cite{barrett-et-al-2007}.
-%and are not reflected in the presentation of the calculus.
-We briefly mention the main ones.
-Discriminators are part of the signature and not abbreviations.
-This requires extending the decision procedure with several
-rules, % \cite{barrett-et-al-2007},
-which apply uniformly to datatypes and codatatypes.
-This approach often leads to better performance because it introduces terms
-less eagerly to $\tEc$.
-Selectors are collapsed eagerly:
-If $\const s^k_{\negvthinspace j}( t ) \in \tEc$ and $t =
-\const{C}_{\negvthinspace j}( u_1, \ldots, u_n )$, the solver directly adds
-$\smash{\const s^k_{\negvthinspace j}( t ) \teq u_k}$ to $\Ec$, whereas the presented calculus
-would apply \rn{Split} and \rn{Inject} before adding this equality.
-To reduce the number of unique literals considered by the calculus, we
-compute a normal form for literals as a preprocessing step. In particular, we
-replace $u \teq t$ by $t \teq u$ if $t$ is smaller than $u$ with respect to some term
-ordering, replace $\smash{\const{C}_{\negvthinspace j}( \vec t \,) \teq
-\const{C}_{\negvthinspace\jPrime}( \vec u )}$ with $\bot$ when $j \neq \jPrime$,
-replace all selector terms of the form $\smash{\const{s}^{\,k}_{\negvthinspace j}(
-\const{C}_{\negvthinspace j}(t_1,\ldots,t_n) )}$ by $t_k$, and replace
-occurrences of discriminators $\smash{\const{d}_{\negvthinspace j}(
-\const{C}_{\negvthinspace\jPrime}( \vec t \,) )}$ by $\top$ or $\bot$
-based on whether $j = \jPrime$.
-
-% AJR : this paragraph may have been what confused Reviewer #2. It does suggest that we are copying verbatim what was in CVC3 (which is what Reviewer #2 thought).
-% Since we aren't saying anything new beyond what is said in Barrett et al (and honestly, we didn't critically evaluate the performance of the updated implementation), 
-% maybe we should just remove this paragraph.
-
-% JB to AJR: I'd rather keep the para, based on feedback by Pascal.
-
-As Barrett et al.\ %\cite{barrett-et-al-2007}
-observed for their procedure,
-it is both theoretically and
-empirically beneficial to delay applications of \rn{Split} as long as
-possible. Similarly, \rn{Acyclic} and \rn{Unique} are fairly expensive because
-they require traversing the equivalence classes, which is why they are part of
-phase~2.
-
-\subsection{Model Construction}
-
-When instructed to do so, the implementation produces models for satisfiable inputs.
-As described in Section~\ref{sec:termination-and-correctness},
-given a saturated set $\Ec$, we %may
-construct a map $\ValC$ from 
-the equivalence classes of $\Ec$ to closed normal $\MU$-terms.
-Recall that our construction of $\ValC$ requires choosing values $\MU {\Varec{x}}.\; t$ for each $\Varec{x} \in \FV( \ValC )$
-such that $\MU {\Varec{x}}.\; t$ does not occur in the set $\BAD_{\smash{\J}}^{\Varec{x}}( \ValC )$.
-To choose values, % $\MU {\Varec{x}}.\; t$,
-we use a fair enumerator for each (co)datatype $\tau$,
-which lists all the values
-of type $\tau$ in a normal interpretation. %, which are closed normal $\MU$-terms.
-
-%Concretely,
-For codatatypes, we enumerate a stream of \relax{every} $\MU$-term of our signature $\Sig$ in a fair manner,
-discarding those that have free variables or are not normal.
-For those that are closed and normal, we then check whether they occur in the set $\BAD_{\smash{\J}}^{\Varec{x}}( \ValC )$.
-This set does not need to be explicitly constructed.
-%We do not explicitly construct the set $\BAD_{\smash{\J}}^{\Varec{x}}( \ValC )$.
-To determine if a term $\MU {\Varec{x}}.\; t$ occurs in the set $\BAD_{\smash{\J}}^{\Varec{x}}( \ValC )$, 
-we check if it is $\alpha$-equivalent to a closed term $(\MU {\Varec{y}}.\; u) \{ \Varec{x} \mapsto z \}$ for
-some $\MU {\Varec{y}}.\; u$ in the range of $\ValC$ and variable $z$.
-This can be efficiently achieved by matching the term $\MU {\Varec{x}}.\; t$ with %the term
-$\MU {\Varec{y}}.\; u$.
-
-%Once the mapping $\ValC$ is constructed, our model $\J$ interprets
-%$\J( x ) = \ValC \ec{x}$ for each variable $x$ occurring free in $\Ec$.
-
-\subsection{Extension to Quantified Formulas}
-
-While the decision procedure is restricted to universal conjectures, users often
-want to solve problems that feature universal axioms and existential
-conjectures.
-%%% @ANDY: not sure how to phrase this now
-%where checking satisfiability requires reasoning about ground literals in the
-%presence of universally quantified formulas.
-Many SMT solvers, including CVC4, can reason about quantified formulas using incomplete 
-instantiation-based 
-techniques~\cite{MouraBjoerner07,ReynoldsTinelliMoura14}.
-These techniques extend naturally to quantified
-formulas involving datatypes and codatatypes.
-
-However, the presence of quantifiers poses an additional challenge in
-the context of (co)datatypes. Quantified formulas can entail an upper bound on
-the cardinality of an uninterpreted type $\ty{u}$.
-By assuming that uninterpreted types have infinite cardinality, the calculus
-presented in Section~\ref{sec:the-decision-procedure} is \relax{incomplete}
-since it may fail to recognize cases where \rn{Split} and \rn{Single} should be
-applied.
-This does not %necessarily
-impact the correctness of the procedure in this setting,
-since %by assumption
-the solver is already incomplete for quantified formulas.
-
-Nonetheless, two techniques help increase the precision of the solver.
-First, we can apply \rn{Split} to datatype terms whose cardinality depends on the finiteness of
-uninterpreted types.
-Second, we can conditionally apply \rn{Single} to codatatype terms that may have cardinality one.
-For example, the codatatype $\ty{stream}_{\vthinspace\ty{u}}$ %, where $\ty{u}$ is uninterpreted.
-has cardinality one precisely when $\ty{u}$ has cardinality one.
-If there exist two equivalence classes $\ec{s}$ and $\ec{t}$ for this type,
-the implementation adds the clause %of the form
-$(\exists x\;y\typ{\ty{u}}.\;\, x \not\teq y ) \mathrel\vee s \teq t$ to $F\!$,
-which states that either the cardinality of $\ty{u}$ is greater than one or
-$s$ must be equal to $t$.
-
-%\section{Examples}
-%\label{sec:examples}
-
-\section{Evaluation on Isabelle Problems}
-\label{sec:evaluation-on-isabelle-problems}
+\section{Evaluation -- JB 1 column} % on Isabelle Problems}
+\label{sec:evaluation}
 
@@
 
-\section{Conclusion}
+\section{Conclusion -- JB 0.25 column}
 \label{sec:conclusion}
@@
 
-%Finally, it might be possible to go
-%further in terms of supporting nested and mixed (co)recursion and quantified
-%formulas over (co)datatypes in solvers.
-
 {

### 2016-02-08 sister/sister.tex
tuning

@@
 Department of Computer Science, The University of Iowa, USA \and \\
-Inria Nancy -- Grand Est \& LORIA Villers-l\`es-Nancy, France
+Inria Nancy -- Grand Est \& LORIA, Villers-l\`es-Nancy, France
 }
@@
 %
-\[\keyw{codatatype}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat})\]
+\[\keyw{codata}~\,\ty{enat} \,=\, \const{Z} \,\mid\, \const{S}(\ty{enat})\]
 %
@@
 For example, the type
-$$\vthinspace\keyw{codatatype}~\;\ty{stream}_{\vthinspace\tau} \,=\allowbreak\, \const{SCons}(\tau,\allowbreak\: \ty{stream}_{\vthinspace\tau})\vthinspace$$
+$$\vthinspace\keyw{codata}~\;\ty{stream}_{\vthinspace\tau} \,=\allowbreak\, \const{SCons}(\tau,\allowbreak\: \ty{stream}_{\vthinspace\tau})\vthinspace$$
 of infinite sequences or streams over an element type $\tau$ is
@@
 corresponding datatype specification
-$$\vthinspace\keyw{datatype}~\;\ty{fstream}_{\vthinspace\tau} \,=\allowbreak\, \const{FSCons}(\tau,\allowbreak\: \ty{fstream}_{\vthinspace\tau})\vthinspace$$
+$$\vthinspace\keyw{data}~\;\ty{fstream}_{\vthinspace\tau} \,=\allowbreak\, \const{FSCons}(\tau,\allowbreak\: \ty{fstream}_{\vthinspace\tau})\vthinspace$$
 would be rejected as non-well-founded
@@
 (co)recursion, and datatype--codatatype mixtures
-fall outside this fragment.\footnote{In principle, rank-1 (top-level) polymorphism \cite{blanchette-paskevich-2013}
-does not raise any special difficulties. Nesting datatypes inside datatypes,
-and likewise for codatatypes, can be reduced to the mutual case
-\cite{gunter-1993-not}. So the only genuinely interesting cases missing are
-mixed nested (co)recursion as well as
-(co)recursion through a non-(co)datatype (both of which make sense,
-at least in a higher-order setting \cite{blanchette-et-al-2014-impl}).}
-We allow ourselves some notational parameterization
-% at the metalevel
-through subscripts---for example, $\ty{stream}_{\vthinspace\tau}$ denotes a
-family of ground types including
-$\ty{stream}_{\vthinspace\ty{int}}$, $\ty{stream}_{\vthinspace\ty{bool}}$,
-and \smash{$\ty{stream}_{\vthinspace\ty{stream}_{\vthinspace\ty{real}}}$}.
+fall outside this fragment.
+
+%We allow ourselves some notational parameterization
+%% at the metalevel
+%through subscripts---for example, $\ty{stream}_{\vthinspace\tau}$ denotes a
+%family of ground types including
+%$\ty{stream}_{\vthinspace\ty{int}}$, $\ty{stream}_{\vthinspace\ty{bool}}$,
+%and \smash{$\ty{stream}_{\vthinspace\ty{stream}_{\vthinspace\ty{real}}}$}.
 
@@
 constructors, and the selectors must be
-distinct and different from existing names.%
-\footnote{It can be convenient to specify the same selector
-for several constructors associated with the same (co)data\-type,
-as long as the argument types coincide. % \cite[Section~3]{blanchette-et-al-2014-impl}.
-However, this is disallowed by SMT-LIB, so we do not consider it here.}
+distinct and different from existing names.
 Schematically:
@@
 \begin{aligned}[t]
-\!(\keyw{co})\keyw{datatype}\;\,
+\!(\keyw{co})\keyw{data}\;\,
   \delta_1 & {}=\, \smash{\const{C}_{11\!}(\bigl[\const{s}_{11\!}^1{:}\bigr]\vthinspace \tau_{11\!}^1, \ldots, \bigl[\const{s}_{11\!}^{n_{11\!}}{:}\bigr]\vthinspace \tau_{11\!}^{n_{11\!}})} \mid \cdots \mid \smash{\const{C}_{1m_1\!}(\ldots)} \\[-.5\jot]
@@
 
-Here are a few examples of legal specifications of (co)datatype families:
-\[\begin{aligned}[t]
-      \keyw{codatatype}\; \,\ty{llist}_{\vthinspace\tau} & \,=\, \const{LNil} \,\mid\, \const{LCons}(%\const{lhead}{:}\;
-      \tau,\: %\const{ltail}{:}\;
-      \ty{llist}_{\vthinspace\tau}) \\%[-.5\jot]
-      \keyw{datatype}\;\, \ty{tree}_{\vthinspace\tau} & \,=\, \const{Node}(\tau,\:\, \ty{forest}_{\vthinspace\tau}) \\[-\jot]
-      \keyw{and}\;\,\, \ty{forest}_{\vthinspace\tau} & \,=\, \const{FNil} \,\mid\, \const{FCons}(\ty{tree}_{\vthinspace\tau},\:\, \ty{forest}_{\vthinspace\tau})
-\end{aligned}
-\]
-
-Because all types must be inhabited (nonempty), a datatype specification is
-admissible only if a ground constructor term can be exhibited.
-This rules out non-well-founded specifications such as
-that of \ty{fstream} in Section~\ref{sec:introduction}.
-For codatatypes, no admissibility check is necessary because there is always a term,
-finite or infinite, that witnesses nonemptiness \cite{blanchette-et-al-2015-esop}.
-
-A type $\delta$ depends on another type $\varepsilon$ if $\varepsilon$ is the
-type of an argument to one of $\delta$'s constructors. Semantically, a set of
-types is \emph{mutually} (\emph{co})\emph{recursive} if and only if the
-associated dependency graph is strongly connected.
-A type is (\emph{co})\emph{recursive} if it belongs to such a set of types.
-Types can be declared together in a mutual fashion even if they are not actually
-mutually (co)recursive. The semantic notion is more precise and is the one that interests
-us.
-
-Non(co)recursive type specifications such~as either of
-\[\begin{aligned}[t]
-      \keyw{datatype}~\vthinspace\ty{option}_{\vthinspace\tau\!} & \,=\, \const{None} \,\mid\, \const{Some}(\tau) \\[-.5\jot]
-      \keyw{codatatype}~\vthinspace\ty{complex} & \,=\, \const{Complex}(\const{re}{:}\; \ty{real},\; \const{im}{:}\;\ty{real})
-\end{aligned}
-\]
-are permitted.
-At the semantic level, it makes no difference whether
-such types are introduced as datatypes or as codatatypes.
+%A type $\delta$ depends on another type $\varepsilon$ if $\varepsilon$ is the
+%type of an argument to one of $\delta$'s constructors. Semantically, a set of
+%types is \emph{mutually} (\emph{co})\emph{recursive} if and only if the
+%associated dependency graph is strongly connected.
+%A type is (\emph{co})\emph{recursive} if it belongs to such a set of types.
+%Types can be declared together in a mutual fashion even if they are not %actually
+%mutually (co)recursive. The semantic notion is more precise and is the one %that interests
+%us.
+
+%Non(co)recursive type specifications such~as either of
+%\[\begin{aligned}[t]
+%      \keyw{data}~\vthinspace\ty{option}_{\vthinspace\tau\!} & \,=\, %\const{None} \,\mid\, \const{Some}(\tau) \\[-.5\jot]
+%      \keyw{codata}~\vthinspace\ty{complex} & \,=\, %\const{Complex}(\const{re}{:}\; \ty{real},\; \const{im}{:}\;\ty{real})
+%\end{aligned}
+%\]
+%are permitted.
+%At the semantic level, it makes no difference whether
+%such types are introduced as datatypes or as codatatypes.
 %Without loss of generality, we consider that these types are datatypes
@@
 %
-%\[\keyw{codatatype}\;\, \ty{stream} \,=\, \const{SCons}(\ty{int},\:\ty{stream})\]
+%\[\keyw{codata}\;\, \ty{stream} \,=\, \const{SCons}(\ty{int},\:\ty{stream})\]
 %
@@
 
-One way to characterize datatypes %semantically
-is as the initial model of the
-selector--constructor equations \cite{barrett-et-al-2007}.
-A drawback of this approach is that it does
-not naturally account for selectors applied to wrong constructors. Barrett et
-al.\ address this by parameterizing the construction by default values, but
-this gives rise to spurious equalities between unrelated terms. For example,
-given
-\[\keyw{datatype}\;\, \ty{x} \,=\, \const{C}(\const{s}{:}\; \ty{int}) \mid \const{D} \mid \const{E}\]
-we would have the spurious equality
-$\const{s}(\const{D}) \teq \const{s}(\const{E})$. This flaw
-could be corrected, but the added complexity seems to suggest that selectors
-are better characterized axiomatically.
-
-A related semantic view of datatypes is as initial algebras of suitable
-functors. Codatatypes are
-then defined dually as final coalgebras \cite{rutten-2000}. The datatypes are
-generated by their constructors, whereas the codatatypes are viewed through
-their selectors.
+%One way to characterize datatypes %semantically
+%is as the initial model of the
+%selector--constructor equations \cite{barrett-et-al-2007}.
+%A drawback of this approach is that it does
+%not naturally account for selectors applied to wrong constructors. Barrett et
+%al.\ address this by parameterizing the construction by default values, but
+%this gives rise to spurious equalities between unrelated terms. For example,
+%given
+%\[\keyw{data}\;\, \ty{x} \,=\, \const{C}(\const{s}{:}\; \ty{int}) \mid %\const{D} \mid \const{E}\]
+%we would have the spurious equality
+%$\const{s}(\const{D}) \teq \const{s}(\const{E})$. This flaw
+%could be corrected, but the added complexity seems to suggest that selectors
+%are better characterized axiomatically.
+%
+%A related semantic view of datatypes is as initial algebras of suitable
+%functors. Codatatypes are
+%then defined dually as final coalgebras \cite{rutten-2000}. The datatypes are
+%generated by their constructors, whereas the codatatypes are viewed through
+%their selectors.
 %By uniformly focusing on the constructors, the
@@
 \begin{aligned}[t]
-\text{Distinctness:}\quad
+\text{Distinctness:}\enskip
   & %\forall \bar x, \bar y.\;\,
@@
   \\[-0.5\jot]
-\text{Injectivity:}\quad
+\text{Injectivity:}\enskip
   & %\forall x_1,\ldots,x_{n_{i\negvthinspace j}} y.\;\,
@@
   \\[-0.5\jot]
-\text{Exhaustiveness:}\quad
+\text{Exhaustiveness:}\enskip
   & \smash{\const{d}_{i1}(x) \mathrel\lor \cdots \mathrel\lor \const{d}_{im_i}(x)}
   \\[-0.5\jot]
-\text{Selection:}\quad
+\text{Selection:}\enskip
   & \smash{\const{s}_{i\negvthinspace j}^{\,k}(\const{C}_{i\negvthinspace j}(x_1,\ldots,x_{n_{i\negvthinspace j}})) \teq x_k}
@@
 %
-Datatypes are additionally characterized by an induction principle for proving
-a conjunction of properties $P_1,\ldots,P_{\ell}$ over arbitrary
-values $v_1 : \delta_1, \ldots, v_\ell : \delta_\ell$:
-%
-\[
-\begin{aligned}[t]
-\text{Induction:}\quad
-\AXC{\strut$\bigwedge_{\vthinspace i,j}\, \forall x_1 \ldots \vthinspace x_{n_{i\negvthinspace j}}.\; \bigl(\bigwedge_{\vthinspace k}\, \mathcalx{IH}[x_k]\bigr) \medrightarrow P_i[\const C_{i\negvthinspace j}(x_1,\ldots,x_{n_{ij}})]$}
-\UIC{\strut$\bigwedge_{\vthinspace i}\, P_i[v_i]$}
-\DP
-\end{aligned}
-\]
-The notation $\mathcalx{IH}[x]$
-denotes either $P_{\iPrime}[x]$ if there exists some $\iPrime$ such that
-the formula is type-correct or else $\top$ (truth).
-The induction principle ensures that the interpretation of datatypes is standard.
+Datatypes are additionally characterized by an induction principle.
 For the natural numbers constructed from $\const{Z}$ and $\const{S}$,
@@
 
-For codatatypes, the dual notion is called coinduction. It makes it possible to
-derive the equality of pairs of codatatype values
-$v_1, w_1 \mathbin: \delta_1,\,\ldots,\,v_\ell, w_\ell \mathbin: \delta_\ell$, based on
-suitable coinduction witnesses $R_1,\ldots,R_\ell$:
-%
-\[
-\begin{aligned}[t]
-\text{Coinduction:}\quad
-\AXC{\strut$\begin{gathered}\textstyle \smash{\bigwedge_{\vthinspace i}\, R_i[v_i, w_i]} \\[-\jot]\textstyle
-\bigwedge_{\vthinspace i}\, \forall v\; w.\;\, R_i[v, w] \medrightarrow
-  \bigwedge_{\vthinspace j}\, \const{d}_{\negvthinspace j}(v) \mathrel\land \const{d}_{\negvthinspace j}(w)
-  \medrightarrow \bigwedge_{\vthinspace k}\, \const{s}_{i\negvthinspace j}^{\,k}(v) \sim \const{s}_{i\negvthinspace j}^{\,k}(w)
-\end{gathered}$}
-\UIC{\strut$\bigwedge_{\vthinspace i}\, v_i \teq w_i$}
-\DP
-\end{aligned}
-\]
-The notation $x \sim y$ stands for $R_{\iPrime}[x, y]$ if there exists some
-$\iPrime$ such that the formula is type-correct or $x \teq y$ otherwise.
-The second premise ensures that the coinduction witnesses are
-bisimulations. The first premise and the conclusion capture the notion that
-equality is the largest bisimulation on codatatypes. Thus, the coinduction
-principle encodes a form of extensionality:\ Two values that yield the same
-observations must be equal, where the observations are made through selectors
-and discriminators.
-
-In addition, codatatypes are guaranteed to contain all values corresponding to
+For codatatypes, the dual notion is called coinduction.
+By this principle, two values that yield the same observations must be equal, 
+where the observations are made through selectors and discriminators. In 
+addition, codatatypes are guaranteed to contain all values corresponding to 
 infinite ground constructor terms.
-In general, this cannot be captured by a
-first-order axiomatization, since there may be uncountably many of them.
-For example, $\ty{stream}_{\vthinspace\ty{int}}$ is isomorphic to the uncountable
-function space $\ty{nat} \to \ty{int}$. Bj\o rner gives a rigorous treatment of
-this aspect \cite{bjoerner-1998}.
 
@@
 
-For datatypes, any recursive specification gives rise to an infinite datatype.
-Paradoxically, this does not extend to codatatypes:
-Some codatatypes are so degenerate as to be finite even though they
-have infinite values. A simple example is
-\vthinspace$\keyw{codatatype}~\ty{a} = \const{A}(\ty{a})$, which is corecursive
-and yet has a cardinality of one;
-its unique value is $\MU a.\; \const{A}(a)$. Other specimens are
-$\ty{stream}_{\vthinspace\ty{unit}\,}$
-and both \ty{b} and \ty{c} in the specification
-\[\begin{aligned}[t]
-      \keyw{codatatype}\;\, \ty{b} & \,=\, \const{B}(\ty{b},\: \ty{c},\: \ty{b},\: \ty{unit}) \\[-1\jot]
-      \keyw{and}\;\, \ty{c} & \,=\, \const{C}(\ty{a},\: \ty{b},\: \ty{c})
-\end{aligned}
-\]
-assuming \ty{unit} is a datatype with the single constructor $\const{Unity} :
-\ty{unit}$. We call such codatatypes
-\emph{corecursive singletons}, or simply singletons.
-For the decision procedure, it will be
-crucial to detect these. %, even if they rarely arise in practice.
-A type may also be a corecursive singleton only in some models. If the example
-above is altered to make \ty{unit} an uninterpreted type, \ty{b} and \ty{c} will be
-singletons precisely when \ty{unit} is interpreted as a singleton.
-Fortunately, given cardinalities for the ordinary types,
-it is easy to characterize this degenerate case.
-
-% In other words, all finite corecursive codatatypes are corecursive singletons.
-
-\begin{lemma}%[Corecursive Singletons]%
-\label{lem:corecursive-singletons}%
-\afterDot
-Let $\delta$ be a corecursive codatatype. For any interpretation in $\JJJJ$,
-the domain interpreting $\delta$ is either infinite or a singleton. In the
-latter case, $\delta$ necessarily has a single constructor, whose arguments have
-types that are interpreted as singleton domains.
-\end{lemma}
-
-\begin{proof}
-By definition, the type is equipped with at least one (directly or indirectly)
-corecursive constructor $\const{C}$. If it additional has second
-corecursive constructor $\const{D}$, it is possible to encode infinitely many
-alternation patterns---e.g.,
-$\const{C}(\const{D}(\const{C}(\const{C}(\ldots))))$---all of which correspond
-to distinct values (by distinctness and injectivity). If the type has a
-noncorecursive constructor $\const{E}$, it is possible to create terms of
-arbitrary depth---e.g., $\const{C}(\ldots(\const{C}(\const{E}))\ldots)$. In
-either case, there can be no finite models.
-
-Therefore, $\const{C}$ must be the only constructor.
-If any of its noncorecursive arguments has a cardinality greater than 1,
-it is possible to encode alternation patterns using it---e.g.,
-$\const{C}(0,\: \const{C}(1,\: \const{C}(0,\: \const{C}(0,\: \ldots))))$---which
-again excludes finite models. Otherwise, the coinduction principle ensures
-that the type has at most one value.
-\xqed
-\end{proof}
+%For datatypes, any recursive specification gives rise to an infinite %datatype.
+%Paradoxically, this does not extend to codatatypes:
+%Some codatatypes are so degenerate as to be finite even though they
+%have infinite values. A simple example is
+%\vthinspace$\keyw{codata}~\ty{a} = \const{A}(\ty{a})$, which is corecursive
+%and yet has a cardinality of one;
+%its unique value is $\MU a.\; \const{A}(a)$. Other specimens are
+%$\ty{stream}_{\vthinspace\ty{unit}\,}$
+%and both \ty{b} and \ty{c} in the specification
+%\[\begin{aligned}[t]
+%      \keyw{codata}\;\, \ty{b} & \,=\, \const{B}(\ty{b},\: \ty{c},\: %\ty{b},\: \ty{unit}) \\[-1\jot]
+%      \keyw{and}\;\, \ty{c} & \,=\, \const{C}(\ty{a},\: \ty{b},\: \ty{c})
+%\end{aligned}
+%\]
+%assuming \ty{unit} is a datatype with the single constructor $\const{Unity} :
+%\ty{unit}$. We call such codatatypes
+%\emph{corecursive singletons}, or simply singletons.
+%For the decision procedure, it will be
+%crucial to detect these. %, even if they rarely arise in practice.
+%A type may also be a corecursive singleton only in some models. If the %example
+%above is altered to make \ty{unit} an uninterpreted type, \ty{b} and \ty{c} %will be
+%singletons precisely when \ty{unit} is interpreted as a singleton.
+%Fortunately, given cardinalities for the ordinary types,
+%it is easy to characterize this degenerate case.
+%
+%% In other words, all finite corecursive codatatypes are corecursive %singletons.
+%
+%\begin{lemma}%[Corecursive Singletons]%
+%\label{lem:corecursive-singletons}%
+%\afterDot
+%Let $\delta$ be a corecursive codatatype. For any interpretation in $\JJJJ$,
+%the domain interpreting $\delta$ is either infinite or a singleton. In the
+%latter case, $\delta$ necessarily has a single constructor, whose arguments %have
+%types that are interpreted as singleton domains.
+%\end{lemma}
 
@@
 %%% "This section will describe such-and-such".
-
-\subsection{A Calculus for $\thD$}
-\label{sec:calculus-for-dc}
 
@@
 
-\subsection{Termination and Correctness of the Calculus}
+\paragraph{Termination and Correctness.}
 \label{sec:termination-and-correctness}
@@
 \end{theorem}
-\begin{proof}
-Consider a derivation tree with root node $\Ec$. Let $D \subseteq \tEc$ be the
-set of terms whose types are finite datatypes, and let $\SSS \subseteq \tEc$ be
-the set of terms occurring as arguments to selectors. For each term $t \in
-D$\negvthinspace, let
-\begin{align*}
-\smash{\SSSS 0_t} & = \{ t \}
-& \smash{\SSSS {i+1}_t} & = \smash{\SSSS i_t \mathrel\cup \{ \const s( u ) \mid u\typ{\delta} \in \SSSS i_t{,}~ \delta \in \Data{,}~ \left|\delta\right| \text{ is finite}{,}~ \const{s} \in \Sel^{\delta}}  \}
-\end{align*}
-%for $i \geq 0$, let ;
-and let $\SSSS \infty_t$ be the limit of this sequence.
-This is a finite set for each $t$, 
-because all chains of selectors applied to $t$ are finite.
-%because in the absence of (co)recursion
-%one eventually reaches a point where no selector can be applied.
-%%% TODO: Clarify above step in report.
-Let $\SSSS \infty$ be the union of all sets $\SSSS \infty_t$ where $t \in D$,
-and let $\tcEc$ be the set of subterms of
-%
-$\Ec \mathrel\cup \{ \smash{\const C_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\smash{\,n_{\negvthinspace j}}}_{\negvthinspace j}( t ) \bigr)} \mid \smash{t\typ{\delta} \in \SSS \mathrel\cup \SSSS \infty{,}~ \const C_{\negvthinspace j} \in \smash{\Ctr^\delta}} \}$.
-%
-In a derivation tree with root node~$\Ec$,
-it can be shown by induction
-on the rules of the calculus that each non-root node $\Fc$ is such that
-$\Terms(\Fc) \subseteq \tcEc$, and hence contains an equality between two terms from $\tcEc$ not occurring in its parent node.
-Thus, the depth of a branch in a derivation tree with root node $\Ec$ is at most $\left| \smash{\tcEc}\vphantom{X} \right|{\!}^2$,
-which is finite since $\tcEc$ is finite.
-\xqed
-\end{proof}
 
@@
 \end{theorem}
-\begin{proof}
-The proof is by structural induction on the derivation tree with root node $\Ec$.
-If the tree is an application of \rn{Conflict}, \rn{Clash}, or \rn{Acyclic},
-then $\Ec$ is $\thD$-unsatisfiable.
-For \rn{Conflict}, this is a consequence of equality reasoning.
-For \rn{Clash}, this is a consequence of distinctness.
-For \rn{Acyclic}, the construction of $\Val$ indicates that the class of values that term~$t$ can take in models of $\Ec$ is infeasible.
-If the child nodes of $\Ec$ are closed derivation trees
-whose root nodes are the result of applying \rn{Split} on $t\typ{\delta}$,
-by the induction hypothesis $\Ec \mathrel\cup t \teq \smash{\const C_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\,\smash{n_{\negvthinspace j}}}_{\negvthinspace j}( t ) \bigr)}$ is
-$\thD$-unsatisfiable
-for each $\const C_{\negvthinspace j} \in \smash{\Ctr^\delta}$.
-Since by exhaustiveness, all models of $\thD$ entail exactly one
-$\DISC{\negvthinspace j}$,
-$\Ec$ is $\thD$-unsatisfiable.
-Otherwise, the child node of $\Ec$ is a closed derivation tree
-whose root
-node $\Ec \mathrel\cup t \teq u$ is obtained by applying one of the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, \rn{Inject}, \rn{Unique}, or \rn{Single}.
-In each of these cases, $\Ec \vDash_{\thD} t \teq u$.
-For \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, this is a consequence of equality reasoning.
-For \rn{Inject}, this is a consequence of injectivity.
-For \rn{Unique}, the construction of $\Val$ indicates that the values of $t$ and $u$ are equivalent in all models of $\Ec$.
-For \rn{Single}, $t$~and~$u$ must have the same value since the cardinality of their type is one.
-By the induction hypothesis, $\Ec \mathrel\cup t \teq u$ is $\thD$-unsatisfiable
-and thus $\Ec$ is $\thD$-unsatisfiable.
-\xqed
-\end{proof}
 
@@
 The proof relies on a specific interpretation $\J$ that satisfies $\Ec$.
-
-First, we define the set of interpretations of the theory $\thD$,
-which requires custom terminology concerning $\mu$-terms.
-Given a $\mu$-term $t$ with subterm $u$,
-the \emph{expansion of $u$ with respect to $t$} is the $\mu$-term $\expand{t}{u}^\emptyset$,
-abbreviated to
-$\expand{t}{u}$, as returned by the function
-%where $x \leadsto \MU x.\; \const{C}( \vec u ) \in t$ indicates that $\MU x.\;
-%\const{C}( \vec u )$ is the subterm of $t$ that binds this occurrence of
-%variable $x$:
-\[\begin{array}{@{}r@{}c@{}l@{}}
-\expand{t}{x}^B & {} \,=\, {} &
-\textstyle\relax{\begin{cases}
-    x & \text{if $x \in B$}   \\[-\jot]
-    \MU x.\; \const{C}\smash{\bigl( \expand{t}{ \vec u}^{\smash{B \,\uplus\, \{ x \}}}\bigr)} & \text{if
-    $\MU x.\; \const{C}( \vec u )$ binds this occurrence of $x \notin B$ in $t$}
-\end{cases}} ~ \\
-\expand{t}{\MU x.\; \const{C}( \vec u )}^B & = &
-\textstyle\begin{cases}
-    x & \text{if $x \in B$} \\[-\jot]
-    \MU x.\; \const{C}\smash{\bigl( \expand{t}{ \vec u }^{\smash{B \,\uplus\, \{ x \}}}\bigr)} & \text{otherwise}
-\end{cases}
-\end{array}\]
-The recursion will eventually terminate because each recursive call adds one bound
-variable to $B$ and there are finitely many distinct bound variables in a $\mu$-term.
-Intuitively, the expansion of a subterm is a self-contained
-$\mu$-term that denotes the same value as the original subterm---\allowbreak e.g.,
-$\expand{\MU x.\: \const{C}(\MU y.\: \const{D}(x))}{\MU y.\: \const{D}(x)} =
- \MU y.\: \const{D}(\MU x.\: \const{C}(y))$.
-%
-%\rem{TODO: add small examples}
-%\rem{TODO: explain why terminates}
-
-The $\mu$-term $u$
-is a \emph{self-similar subterm} of~$t$ if
-$u$ is a proper subterm of~$t$,
-$t$ and $u$ are of the forms $\MU x.\; \const{C}( t_1, \ldots, t_n )$
-and $\MU y.\; \const{C}( u_1, \ldots, u_n )$,
-and $\expand{t}{ t_k } \aequiv \expand{t}{ u_k }$ for all $k$. % \in [n]
-The $\mu$-term $t$ is \emph{normal} if it does not contain %any
-self-similar subterms
-and all of its proper subterms are also normal.
-Thus, $t = \MU x.\; \const{C}( \MU y.\; \const{C}( y ) )$ is not normal
-because $\MU y.\; \const{C}( y )$ is a self-similar subterm of~$t$.
-Their arguments have the same expansion with respect to~$t$:
-$\expand{t}{\MU y.\; \const{C}( y )} =
-\MU y.\; \const{C}\bigl( \expand{t}{y}^{\smash{\{y\}}} \bigr) =
-\MU y.\; \const{C}( y )$
-is $\alpha$-equivalent to
-$\expand{t}{y} =
-\MU y.\; \const{C}\bigl( \expand{t}{y}^{\smash{\{y\}}} \bigr) =
-\MU y.\; \const{C}( y )$.
-The term $u = \MU x.\; \const{C}( \MU y.\; \const{C}( x ) )$ is also not normal,
-since $\MU y.\; \const{C}( x )$ is a self-similar subterm of $u$,
-noting that
-%
-$\expand{u}{\MU y.\; \const{C}( x )}
- = \MU y.\; \const{C}\bigl( \expand{u}{x}^{\smash{\{y\}}} \bigr)
- = \MU y.\; \const{C}\bigl( \expand{u}{\MU x.\allowbreak\; \const{C}( \MU y.\allowbreak\; \const{C}( x ) ) }^{\smash{\{y\}}} \bigr)
- = \MU y.\; \const{C}\bigl( \MU x.\allowbreak\; \const{C}( \expand{u}{\MU y.\allowbreak\; \const{C}( x )}^{\smash{\{x,y\}}} ) \bigr)
- = \MU y.\allowbreak\; \const{C}( \MU x.\allowbreak\; \const{C}( y ) )$
-is $\alpha$-equivalent to $\expand{u}{x} = u$.
-% for the renaming $\{ x \mapsto y, y \mapsto x \}$.
-
-For any $\mu$-term $t$ of the form $\MU x.\; \const C( \vec u )$,
-its \emph{normal form} $\nf{t}$ is obtained
-by replacing all of the self-similar subterms of~$t$ with $x$
-%and by replacing all of the abnormal subterms of $t$ with their corresponding normal forms.
-and by recursively normalizing the other subterms.
-For variables, $\nf{x} = x$.
-Thus, $\nf{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) )} = \MU x.\; \const{C}( x )$.
-
-%\begin{lemma}
-%\label{lem:mu-norm-arg}
-%If the $\mu$-terms $\vec u$ are in normal form and $t = \nf{\MU x.\; \const{C}( \vec u )} = \MU x.\; \const{C}( \vec w )$
-%for some $\mu$-terms $\vec w$,
-%then $\vec u \vsim \expand{t}{\vec w}$\vthinspace.
-%\end{lemma}
-% TODO:
-%\begin{proof}
-%\qed
-%\end{proof}
-
-%\begin{lemma}
-%\label{lem:mu-norm-interp}
-%If the $\mu$-term $t$ is normal and $u$ is a subterm of $t$, then $\expand{t}{u}$ is normal.
-%\end{lemma}
-%\begin{proof}
-%\qed
-%\end{proof}
-
-%\begin{lemma}
-%\label{lem:mu-cong}
-%If $t_u = \MU x.\; \const{C}( \vec u )$ and $t_w = \MU y.\; \const{C}( \vec w )$ are in normal form,
-%then $\expand{t_u}{ \vec u } \vsim \expand{t_w}{ \vec w }$ if and only if $t_u \vsim t_w$.
-%\end{lemma}
-%\begin{proof}
-%\rem{TODO}
-%\qed
-%\end{proof}
-
-We now define the class of interpretations for $\thD$.
-$\interp{\J}{\tau}$ denotes the interpretation type $\tau$ in $\J$---that is, a
-nonempty set of domain elements for that type.
-$\interp{\J}{\const{f}}$ denotes the interpretation of a function $\const{f}$ in $\J$.
-If $\const{f} : \tau_1 \times \cdots \times \tau_n \to \tau$,
-then $\interp{\J}{\const{f}}$ is a total function from $\interp{\J}{\tau_1} \times \cdots \times \interp{\J}{\tau_n}$ to $\interp{\J}{\tau}$.
-%To simplify the presentation,
-All types are interpreted as sets of $\mu$-terms, but
-only values of types in $\Codata$ may contain cycles.
-%Since all values of types in $\Data$ and $\Nondata$ are acyclic,
-%a typical representation for these values can be obtained by
-%dropping each of its $\mu$-bindings.
-%For example, the $\mu$-term $\MU x.\; \const{Z}$ representing the integer constant zero
-%is simply $\const{Z}$ after dropping the $\mu$-binding of $x$.
-
-\begin{definition}[Normal Interpretation]
-\afterDot%
-\label{def:norm-interpretation}%
-\rm
-An interpretation $\J$ is \emph{normal} if the following conditions are met:
-\begin{enumerate}
-\item
-For each type $\tau$,
-%%% sets *contain* elements, they *include* other sets
-$\interp{\J}{\tau}$ includes a maximal set of closed normal $\mu$-terms of that type that are
-unique up to $\alpha$-equivalence and acyclic unless $\tau \in \Codata$.\strut
-
-\item
-For each constructor term $\vphantom{s^k}\const{C}( \vec t \,)$ of type $\tau$,
-$\interp{\J}{\const{C}}( \interp{\J}{\vec t\,} )$ is the value
-in $\interp{\J}{\tau}$ that is $\alpha$-equivalent to
-$\nf{\MU x.\; \const{C}( \interp{\J}{\vec t\,} )}$, where $x$ is fresh.\strut
-
-\item
-For each selector term $\const s^k_{\negvthinspace j}( t )$ of type $\tau$, % for some $\vec u$,
-if $\interp{\J}{t}$ is $\MU x.\; \const{C}_{\negvthinspace j}( \vec u )$,
-then $\J\negvvthinspace( \const s^k_{\negvthinspace j} )\bigl( \interp{\J}{t} \bigr)$ is the value
-in $\interp{\J}{\tau}$ that is $\alpha$-equivalent to
-$\expand{\interp{\J}{t}}{u_k}$.
-\end{enumerate}
-\end{definition}
-
-Not all normal interpretations are models of codatatypes, because
-models must contain all possible infinite terms, not only cyclic ones. However,
-acyclic infinite values are not interesting for deciding universal formulas:
-For such formulas, it is trivial to extend any normal interpretation with extra
-domain elements to obtain a genuine model if desired. %Hence, we focus on normal interpretations.
-
-When constructing a model $\J$ of $\Ec$,
-it remains only to specify how $\J$ interprets wrongly applied selector terms and variables.
-For the latter, this will be based on the mapping $\Val$ computed in phase~2 of the calculus.
-
-{\looseness=-1
-First, we need the following definitions.
-We write $t \vsimv{x} u$ if $\mu$-terms $t$ and $u$ are syntactically equivalent
-for some renaming that avoids capturing any variable other than~$x$.
-For example,
-$\MU x.\; \const{D}( x ) \vsimv{y} \MU x.\; \const{D}( y )$
-(by renaming $y$ to $x$),
-$\MU x.\; \const{C}( x, x ) \vsimv{x} \MU y.\; \const{C}( x, y )$, and
-$\MU x.\; \const{C}( z, x ) \vsimv{z} \MU y.\; \const{C}( z, y )$,
-but
-$\MU x.\; \const{D}( x ) \not\vsimv{x} \MU x.\; \const{D}( y )$ and
-$\MU x.\; \const{C}( x, x ) \not\vsimv{y} \MU y.\; \const{C}( x, y )$.
-For a variable $x\typ{\tau}$ and a normal interpretation $\J$,
-we let $\BAD_{\smash{\J}}^x( \Val )$ denote the set consisting of all values
-$v \in \interp{\J}{\tau}$
-such that $v \vsimv{x} \expand{t}{u}$ for some subterm $u$ of a term $t$
-occurring in the range of $\Val$. This set describes shapes of terms to avoid
-when assigning a $\mu$-term to $x$.
-
-}
-
-The \emph{completion} $\ValC$ of $\Val$ for a normal interpretation $\J$
-assigns values from $\J$ to unassigned variables in the domain of $\Val$.
-We construct $\ValC$ by initially setting $\ValC := \nf{\Val}$
-and by exhaustively applying the following rule:%
-\[
-\hbox{\(
-\inferrule{
-  {\Varec{x}}\typ{\tau} \in \FV( \ValC )
-  \betweenantes
-  \MU {\Varec{x}}.\; t \aequiv v
-  \betweenantes
-  v \in \interp{\J}{\tau}
-  \betweenantes
-  v \notin \BAD_{\smash{\J}}^{\Varec{x}}(\ValC)
-}{
-  \ValC := \nf{\ValC [ {\Varec{x}} \mapsto \MU {\Varec{x}}.\; t ]}
-}
-\)
-%\rn{Assign}
-}
-\]
-%
-Given an unassigned variable in $\ValC$, this rule assigns it a fresh
-value---one that does not occur in $\BAD_{\smash{\J}}^{\Varec{x}}(\ValC)$
-modulo $\alpha$-equivalence---excluding
-not only existing terms in the range of $\ValC$ but also
-terms that could emerge as a result of the update.
-%\rem{TODO: What guarantees that such a value exists? (Answer: infinitely many values
-%of codatatypes, of which only finitely many are in
-%$\BAD_{\smash{\J}}^{\Varec{x}}(\ValC)$?)}
-Since this update removes one
-variable from $\FV( \ValC )$ and does not add any variables to $\FV( \ValC
-)$, the process eventually terminates. We normalize all terms in
-the range of $\ValC$ at each step.
-
-%The last premise is admittedly rather technical.
-To ensure disequality literals are satisfied by an interpretation based on $\ValC$,
-it suffices that $\ValC$ is injective modulo $\alpha$-equivalence.
-This invariant holds initially, and
-the last premise in the above rule ensures that it is maintained.
-%More precisely,
-The set $\BAD_{\smash{\J}}^{\Varec{x}}(\ValC)$ is an overapproximation of the values
-that, when assigned to ${\Varec{x}}$, will cause values in the range of $\ValC$ to become $\alpha$-equivalent.
-%This premise is not strictly necessary in practice, 
-%as one could compute $\nf{\ValC [ {\Varec{x}} \mapsto \MU {\Varec{x}}.\; t ]}$ optimistically and check
-%injectivity afterward, repeating until a suitable value is found.
-For infinite codatatypes, it is always possible to find fresh values $v$
-because $\BAD_{\smash{\J}}^{\Varec{x}}(\ValC)$ is a finite set.
-%Injectivity is important to satisfy disequality literals.
-%For infinite codatatypes, it is always possible to find suitable values $v$.
-%For example,
-%if the longest cycle in $\BAD_{\smash{\J}}^{\Varec{x}}(\ValC)$ is of length $n$, a fresh value
-%can be constructed with a cycle of length $n + 1$.
-
-\begin{examplex}
-Let $\delta$ be a codatatype with the constructors $\const{C},\, \const{D},\, \const{E} :\delta \rightarrow \delta$.
-Let $\Ec$ be the set
-$\{
-u \teq \const{C}( z ),\;
-v \teq \const{D}( z ),\;
-w \teq \const{E}( y ),\;
-x \teq \const{C}( v ),\;
-z \not\teq v
-\}$.
-After applying the calculus to saturation on $\Ec$,
-the mapping $\Val$ is as follows:
-%\[\begin{array}{r@{}c@{}l@{\;}l@{\qquad}r@{}c@{}r@{\;}l@{}l@{\;}l}
-%\[\begin{array}{r@{}c@{}l@{\;}l@{\qquad}r@{}c@{}r@{\;}l@{}l@{\;}l@{\qquad}r@{}c@{}r@{\;}l}
-\[\begin{array}{r@{}c@{}r@{}l@{\qquad}r@{}c@{}r@{}l@{\qquad}r@{}c@{}l}
-\Val \ec{u} & {}\,=\,{} & \MU {\Varec{u}}.\; & \const{C}( {\Varec{z}} ) &
-\Val \ec{w} & = & \MU {\Varec{w}}.\; & \const{E}( {\Varec{y}} ) &
-\Val \ec{y} & {}\,=\,{} & {\Varec{y}} \\
-\Val \ec{v} & = & \MU {\Varec{v}}.\; & \const{D}( {\Varec{z}} ) &
-\Val \ec{x} & {}\,=\,{} & \MU {\Varec{x}}.\; & \const{C}( \MU {\Varec{v}}.\; \const{D}( {\Varec{z}} ) ) &
-\Val \ec{z} & = & {\Varec{z}}
-\end{array}\]
-%
-To construct a completion $\ValC$, we must choose values for ${\Varec{y}}$ and ${\Varec{z}}$,
-which are free in~$\Val$.
-Modulo $\alpha$-equivalence, $\smash{\BAD_{\smash{\J}}^{{\Varec{z}}}( \Val )} =
-%\begin{array}{c}
-\{
-\MU a.\: \const{C}( a ),\:
-\MU a.\: \const{D}( a ),\allowbreak\:
-\MU a.\: \const{C}( \const{D}( a )),\allowbreak\:
-\const{C}( \MU a.\: \const{D}( a ))
-\}
-%\end{array}
-$.
-Now consider a normal interpretation $\J$ that evaluates variables in $\Ec$ based on~$\Val$:
-$\interp{\J}{u} = \Val \ec{u}$, $\interp{\J}{v} = \Val \ec{v}$, and so on.
-Assigning a value for $\Val \ec{z}$ that is $\alpha$-equivalent to a value in $\BAD_{\smash{\J}}^{{\Varec{z}}}( \Val )$
-may cause values in the range of $\Val$ to become $\alpha$-equivalent,
-which in turn may cause $\Ec$ to be falsified by~$\J$.
-For example, assign $\MU {\Varec{z}}.\; \const{D}( {\Varec{z}} )$ for ${\Varec{z}}$.
-After the substitution, $\Val \ec{v} = \MU {\Varec{v}}.\; \const{D}( \MU {\Varec{z}}.\; \const{D}( {\Varec{z}} ) )$,
-which has normal form $\MU {\Varec{v}}.\; \const{D}( {\Varec{v}} )$,
-which is $\alpha$-equivalent to $\MU {\Varec{z}}.\; \const{D}( {\Varec{z}} )$.
-However, this contradicts the disequality $z \not\teq v$ in~$\Ec$.
-On the other hand, if the value assigned to ${\Varec{z}}$ is fresh, %not $\alpha$-equivalent to any term in $\BAD_{\smash{\J}}^{{\Varec{z}}}( \Val )$,
-the values in the range of $\Val$ remain $\alpha$-disequivalent.
-We can assign a value such as
-$\MU {\Varec{z}}.\; \const{E}( {\Varec{z}} )$,
-$\MU {\Varec{z}}.\; \const{D}( \const{C}( {\Varec{z}} ) )$, or
-$\MU {\Varec{z}}.\; \const{C}( \const{C}( \const{D}( {\Varec{z}} ) ) )$
-to ${\Varec{z}}$.
-
-Legal substitutions for ${\Varec{z}}$ may cause the range of $\Val$ to contain abnormal terms.
-For example, after assigning $\MU {\Varec{z}}.\; \const{D}( \const{C}( {\Varec{z}} ) )$ to $\Varec{z}$, we have
-$\Val \ec{u} = \MU {\Varec{u}}.\; \const{C}( \MU {\Varec{z}}.\; \const{D}( \const{C}( {\Varec{z}} ) ) )$,
-with normal form
-$\MU {\Varec{u}}.\; \const{C}( \MU {\Varec{z}}.\; \const{D}( {\Varec{u}} ) )$.
-This explains why the term is normalized in the rule's conclusion.
-\xend
-\end{examplex}
-
-In the following lemma about $\ValC$,
-$\vphantom{\Bigl(_j}\muvar( t ) = \relax{\begin{cases}
-  t & \text{if $t$ is a variable} \\[-\jot]
-  x & \text{if $t$ is of the form $\MU {x}.\; u$.}
-\end{cases}}$
-
-\pagebreak[2]
-
-\begin{lemma}
-\label{lem:interpretation-completion}
-If $\Val$ is constructed for a saturated set $\Ec$
-and $\ValC$ is a completion of $\Val$ for a normal interpretation $\J$, the following properties hold:
-\begin{enumerate}
-\item[\rm (1)] $\ValC \ec{x\typ{\tau}}$ is $\alpha$-equivalent to a value in $\interp{\J}{\tau}$.
-\item[\rm (2)]
-$\ValC \ec{x} = \expand{\ValC \ec{y}}{ t }$ for
-all subterms $t$ of $\ValC \ec{y}$ with $\muvar(t) = {\Varec{x}}$.
-\item[\rm (3)]
-$\ValC \ec{x} \vsim \ValC \ec{y}$ if and only if $\ec{x} = \ec{y}$.
-\end{enumerate}
-\end{lemma}
-
-\begin{proof}
-To show (1), we first show that $\ValC$ contains no free variables.
-Assume by contradiction that $\ValC$ contains a free variable ${\Varec{y}}$ for some $\ec{y}$ of type $\tau$.
-Then it must be the case that $\ec{y}$ does not contain a constructor term,
-or else ${\Varec{y}}$ would not occur as a free variable in~$\Val$.
-Consider the case when $\tau$ is finite.
-By assumption, $\tau \notin \Nondata$.
-Since \rn{Split} does not apply to $\Ec$, we have $\tau \notin \Data$.
-If $\tau \in \Codata$, then $\tau$ is corecursive by assumption, and by Lemma~\ref{lem:corecursive-singletons},
-the cardinality of $\tau$ must be one.
-Since \rn{Single} does not apply, % to $\Ec$,
-there is only one equivalence class of type $\tau$ in $\Ec$,
-and thus there are no terms in $\BAD_{\smash{\J}}^{\Varec{y}}(\ValC)$ of type $\tau$.
-This is a contradiction, since %our %model
-completion can assign the value in the domain of $\tau$ to ${\Varec{y}}$.
-Now, consider the case when $\tau$ is infinite.
-This is also a contradiction,
-since there are only a finite number of closed terms in $\BAD_{\smash{\J}}^{\Varec{y}}(\ValC)$,
-and thus %our %model
-completion can assign a value not occurring in $\BAD_{\smash{\J}}^{\Varec{y}}(\ValC)$ to ${\Varec{y}}$.
-By construction, $\ValC \ec{x}$ is normal.
-Since \rn{Acyclic} does not apply, $\Val \ec{x}$ is acyclic when $\tau \in \Data$.
-Moreover, the construction of $\ValC$ applies substitutions of the form
-$\{ {\Varec{y}} \mapsto t \}$, where $t$ is acyclic when the type of ${\Varec{y}}$ is not a codatatype.
-Thus, $\ValC \ec{x}$ is acyclic when $\tau \notin \Codata$.
-Therefore, by definition, $\ValC \ec{x}$ is $\alpha$-equivalent to a value in $\interp{\J}{\tau}$.
-
-We now show that (2) and (3) hold initially for $\Val$.
-For all equivalence classes $\ec{z}$,
-each pair of constructor terms $\const{C}_{\negvthinspace j}( \vec t \,)$ and $\const{C}_{\negvthinspace\jPrime}( \vec u )$ in $\ec{z}$
-are such that $j = \jPrime$, since \rn{Clash} does not apply, % to $\Ec$,
-and are such that $\ec{ \vec t \,} = \ec{ \vec u }$, since \rn{Inject} does not apply. % to $\Ec$.
-Thus,
-$\Val$ was constructed by applying a sequence of substitutions
-where all substitutions for variables ${\Varec{z}}$
-were uniquely of the form $\{ {\Varec{z}} \mapsto \const{C}_{\negvthinspace j}( {\Varec{t_1}}, \ldots, {\Varec{t_n}} ) \}$
-when $\ec{z}$ contains a constructor $\const{C}_{\negvthinspace j}( t_1, \ldots, t_n )$.
-Suppose $\Val \ec{y}$ has a subterm $t$ such that $\muvar(t) = {\Varec{x}}$.
-Both $\Val \ec{x}$ and the subterm $t$ of $\Val \ec{y}$ were
-constructed by applying a sequence of substitutions of the form mentioned above to ${\Varec{x}}$.
-Moreover, free variables ${\Varec{z}}$ in $t$ that are bound in $\Val \ec{y}$ are interpreted in
-the expansion of $\expand{\Val \ec{y}}{ t }$ as a term
-constructed by a sequence of substitutions of the form mentioned above to ${\Varec{z}}$. %\rem{TODO: more detail}
-Thus, we have
-$\expand{\Val \ec{y}}{ t } = \expand{\Val \ec{x}}{ \Val \ec{x} } = \Val \ec{x}$, and %thus
-(2) holds for $\Val$.
-Property (3) holds for $\Val$ since \rn{Unique} does not apply.
-
-We now show that $\Val = \nf{\Val}$.
-Assume by contradiction $\Val \ec{x} \neq \nf{\Val \ec{x}}$ for some $\Val \ec{x}$ of minimal size.
-We have that $\Val \ec{x}$ is of the form $\MU {\Varec{x}}.\; \const{C}( t_1, \ldots, t_n )$.
-Due to the construction of~$\Val$,
-we know $\ec{x}$ contains a constructor $\const{C}( z_1, \ldots, z_n )$ and $\muvar( t_i ) = z_i$ for some $i$. % \in [n]$.
-Since $\Val \ec{x}$ is a minimal, it
-contains a subterm of the form $\MU {\Varec{y}}.\; \const{C}( u_1, \ldots, u_n )$
-where $\expand{\Val \ec{x}}{ t_i } \aequiv \expand{\Val \ec{x} }{ u_i }$ for some $i$. % \in [n]$.
-Due to the construction of $\Val$, $\ec{y}$ contains a constructor $\const{C}( w_1, \ldots, w_n )$ and $\muvar( u_i ) = w_i$ for some $i$. % \in [n]$.
-Since \rn{Cong} does not apply, % to $\Ec$
-we have $\ec{w_{\negvthinspace j}}$ and $\ec{z_{\negvthinspace j}}$ are distinct for some $j$.
-By (2), $\expand{\Val \ec{x}}{ t_{\negvthinspace j} } = \Val \ec{z_{\negvthinspace j}}$
-and $\expand{\Val \ec{x}}{ u_{\negvthinspace j} } = \Val \ec{w_{\negvthinspace j}}$,
-which are not $\alpha$-equivalent by (3), contradicting the fact that $\MU {\Varec{y}}.\; \const{C}( u_1, \ldots, u_n )$ is a self-similar subterm of $\Val \ec{x}$.
-Thus, $\Val = \nf{\Val}$, and (2) and~(3) hold for $\nf{\Val}$.
-
-We now show that if (2) and (3) hold for some $\Val_1$,
-they also hold for $\nf{\Val_1 \sigma}$,
-where $\sigma$ is a substitution of the form $\{ {\Varec{x}} \mapsto \MU{\Varec{x}}.\; t \}$,
-${\Varec{x}} \in FV( \Val_1 )$,
-and $\MU {\Varec{x}}.\; t$ is not $\alpha$-equivalent to a term in $\BAD_{\smash{\J}}^{\Varec{x}}(\Val_1)$.
-To show (2),
-by assumption of (2) on $\Val_1$, we have $\expand{\Val_1 \ec{y}}{ u } = \Val_1 \ec{x}$ for
-all subterms $u$ of $\Val_1 \ec{y}$ where $\muvar(t) = {\Varec{x}}$.
-Thus, $\expand{\Val_1 \sigma \ec{y}}{ u } = \Val_1 \sigma \ec{x}$ and
-$\expand{\nf{\Val_1 \sigma} \ec{y}}{ u } = \nf{\Val_1 \sigma} \ec{x}$.
-To show (3),
-consider two distinct equivalence classes $\ec{y}$ and $\ec{z}$,
-and assume by contradiction that $\nf{\Val_1 \sigma} \ec{y} \aequiv \nf{\Val_1 \sigma} \ec{z}$.
-Due to (3) for $\Val_1$,
-$\ec{y}$ and $\ec{z}$ must have (minimal) subterms
-where $t_1$ occurs in $\Val_1 \ec{y}$ the same position~$p$ as $t_2$ occurs in $\Val_1 \ec{z}$,
-and $t_1 \not\aequiv t_2$.
-%We first show that $\nf{\Val_1 \sigma} \ec{y}$ (and for the same reasons, $\nf{\Val_1 \sigma} \ec{z}$)
-%also contains position $p$.
-%If this were not the case,
-%then without loss of generality we may assume
-%$\Val_1 \sigma \ec{y}$ contains a self-similar subterm
-%$\MU {\Varec{w}}.\; u$ that strictly contains $t_1 \sigma$.
-If $t_1$ (resp. $t_2$) is a free variable that is not ${\Varec{x}}$,
-then $\nf{\Val_1 \sigma} \ec{y}$ (resp. $\nf{\Val_1 \sigma} \ec{z}$)
-contains $t_1$ (resp. $t_2$) at position~$p$,
-and $\nf{\Val_1 \sigma} \ec{z}$ (resp. $\nf{\Val_1 \sigma} \ec{y}$) does not.
-If $t_1$ is of the form $\MU w_1.\; \const{C}( \vec t \,)$
-and $t_2$ is of the form $\MU w_2.\; \const{D}( \vec u )$,
-then the expansion %\rem{TODO: define}
-of $\nf{\Val_1 \sigma} \ec{y}$ and $\nf{\Val_1 \sigma} \ec{z}$
-are $\alpha$-disequivalent at position~$p$.
-Since $t_1$ and $t_2$ are minimal,
-say $t_1$ is of the form $\MU w.\; \const{C}( \vec t \,)$,
-and $t_2$ is ${\Varec{x}}$.
-Since $\sigma$ maps ${\Varec{x}}$ to a closed $\mu$-term $\MU {\Varec{x}}.\; t$,
-we have that $FV( t_1 ) \subseteq \{ {\Varec{x}} \}$,
-or else the expansion of $\nf{\Val_1 \sigma} \ec{y}$ and $\nf{\Val_1 \sigma} \ec{z}$
-are $\alpha$-disequivalent at position~$p$ since they do not contain the same free variables.
-Since $\expand{\Val_1 \ec{y}}{t_1}$ is normal,
-there is a closed $\mu$-term $v \in \BAD_{\smash{\J}}^{\Varec{x}}( \Val_1 )$
-such that $v \vsimv{\Varec{x}} \expand{\Val_1 \ec{y}}{t_1}$.
-Thus, by assumption on the selection of $\MU {\Varec{x}}.\; t$,
-we have $\smash{\MU {\Varec{x}}.\; t \not\vsimv{\Varec{x}} \expand{\Val_1 \ec{y}}{t_1}}$,
-which implies that the expansion of $\nf{\Val_1 \sigma} \ec{y}$ and $\nf{\Val_1 \sigma} \ec{z}$
-are $\alpha$-disequivalent at position~$p$. %\rem{TODO: more detail}
-
-Thus, by induction on the number of applications of the above rule used to obtain $\ValC$,
-we have that $\ValC$ satisfies (2) and (3).
-\xqed
-\end{proof}
-
-Intuitively, this lemma states three properties of $\ValC$ that together ensure that
-a normal interpretation~$\J$ can be constructed that satisfies $\Ec$.
-Property~(1) states that the values in the range of $\ValC$ are
-$\alpha$-equivalent to a value in normal interpretation. This means they are
-closed, normal, and acyclic when required. Property~(2) states that the
-interpretation of all subterms in the range of $\ValC$ depends on its associated
-variable only. In other words, the interpretation of a subterm $t$ where
-$\muvar(t) = {\Varec{x}}$ is equal to
-$\ValC \ec{x}$, independently of the context.
-Property~(3) states that $\ValC$ is injective (modulo
-$\alpha$-equivalence), which ensures %in the following
-that distinct values are
-assigned to distinct equivalence classes.
 
@@
 \end{theorem}
-\begin{proof}
-Let $\Fc$ be a saturated node in a derivation tree with root node $\Ec$.
-%We will construct a model $\J$ for a set of equalities $\Fc$ that is equivalent to $\Fc_0$, and where
-%all equivalence classes of $\Fc$ contain at least one variable, and all selectors in $\Fc$ are applied to variables only.
-%The former comes with no loss of generality since new equalities of the form $y \teq t$ for fresh variable $y$ can be added to $\Fc$ without affecting its satisfiability,
-%The latter also comes with no loss of generality since nested applications of selectors in $\Fc_0$ can be replaced by fresh variables while adding equalities to $\Fc\!$.
-We consider a normal interpretation $\J$
-that interprets wrongly applied selectors based on equality information in~$\Fc$
-and that interprets the variables of $\Fc$ based on the completion $\ValC$.
-% of %the mapping $\Val$. % from phase~2.
-For the variables, let $\interp{\J}{x\typ{\tau}}$ be the value in $\J(\tau)$ that is $\alpha$-equivalent
-with $\ValC \ec{x}$ for each variable $x \in \Terms(\Fc)$,
-which by Lemma~\ref{lem:interpretation-completion}(1) is guaranteed to exist.
-
-We first show that $\J$ satisfies all equalities $t_1 \teq t_2 \in \Fc\!$. % where $t_1$ and $t_2$ have type $\tau$.
-To achieve this, we show
-by structural induction on $t\typ\tau$
-that $\interp{\J}{t} \aequiv \ValC \ec{t}$ for all terms $t \in \Terms( \Fc )$,
-which implies $\J \vDash t_1 \teq t_2$ since $\J$ is normal.
-
-If $t$ is a variable, then $\interp{\J}{t} \vsim \ValC \ec{ t }$ by construction.
-
-If $t$ is a constructor term of the form $\const{C}( u_1, \ldots, u_n )$,
-then $\interp{\J}{t}$ is $\alpha$-equivalent with $\nf{\MU x.\allowbreak\; \const{C}( \interp{\J}{u_1},\allowbreak \ldots,\allowbreak \interp{\J}{u_n} ) }$ for some fresh $x$,
-which by the induction hypothesis is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \ValC \ec{ u_1 }, \ldots, \ValC \ec{ u_n } )}$.
-Call this term $t'$.
-%By Lemma~\ref{lem:interpretation-completion}(2) and due to our construction of $\ValC$, we have $\ValC \ec{t}$ is a term
-%of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$
-%where $\expand{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$ for $i \in [n]$.
-Since \rn{Inject} and \rn{Clash} do not apply to~$\Fc\!$,
-by the construction of $\ValC$ we have that $\ValC \ec{t}$ is a term
-of the form $\MU {\Varec{t}}.\; \const{C}( w_1,\allowbreak \ldots,\allowbreak w_n )$
-where
-$\muvar(w_i) = {\Varec{u_i}}$ for each~$i$.  Thus by Lemma~\ref{lem:interpretation-completion}(2),
-$\expand{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$.
-For each $i$, let $u_\iPrime$ be the $i$\vvthinspace th argument of $t'$.
-%By Lemma~\ref{lem:mu-norm-arg},
-We have that
-$\expand{t'}{u_\iPrime} \vsim \ValC \ec{u_i}$, and
-thus $\expand{t'}{u_\iPrime} \vsim \expand{\ValC \ec{t}}{ w_i }$. % for each $i \in [n]$.
-%By Lemma~\ref{lem:mu-cong}, we have $t' \vsim t_w$.
-Thus, $\interp{\J}{t} \vsim t' \vsim \ValC \ec{t}$, and
-we have $\interp{\J}{t} \vsim \ValC \ec{t}$.
-
-If $t$ is a selector term %of the form
-$\const s^k_{\negvthinspace j}( u )$,
-since \rn{Split} does not apply to $\Fc\!$,
-$\ec{u}$ must contain a term of the form $\smash{\const C_{\negvthinspace\jPrime}\bigl( \const s^1_{\negvthinspace\jPrime}( u ), \ldots, \const s^{n}_{\negvthinspace\jPrime}( u ) \bigr)}$ for some $\jPrime$.
-%By Lemma~\ref{lem:interpretation-completion}(2) and due to our construction of $\ValC$, we have $\ValC \ec{u}$ is of the form
-%$\MU {\Varec{u}}.\; \const C_{\negvthinspace\jPrime}( w_1, \ldots, w_n )$,
-%where $\expand{\ValC \ec{u} }{ w_i } = \ValC \ec{ \const s^i_{\negvthinspace\jPrime}( u )}$ for $i \in [n]$.
-Since \rn{Inject} and \rn{Clash} are not applicable, % to $\Fc\!$,
-by construction $\ValC\ec{u}$ must be of the form
-$\MU {\Varec{u}}.\; \const C_{\negvthinspace\jPrime}( w_1, \ldots, w_n )$,
-where
-$\muvar(w_i) = \smash{\Varec{\vphantom{I}\smash{\const s^i_{\negvthinspace\jPrime}( u )}}}$
-for each~$i$, and thus by Lemma~\ref{lem:interpretation-completion}(2),
-$\expand{\ValC \ec{u}}{ w_i } = \ValC \ec{ \smash{\const s^i_{\negvthinspace\jPrime}( u )}}$.
-If $j = \jPrime$, then $\interp{\J}{t}$ is $\alpha$-equivalent with $\expand{\ValC \ec{u}}{w_k}$, which is equal to $\ValC \ec{ \smash{ \const s^k_{\negvthinspace j}( u )}}
-= \ValC \ec{t}$.
-If $j \neq \jPrime$, since \rn{Cong} does not apply,
-any term of the form $\smash{\const s^k_{\negvthinspace j}( u' )}$ not occurring in $\ec{t}$
-is such that $[u] \not= [u']$.
-By the induction hypothesis and Lemma~\ref{lem:interpretation-completion}(3), $\interp{\J}{u} \neq \interp{\J}{u'}$ for all such $u$, $u'$.
-Thus, we may interpret $\smash{\interp{\J}{\const s^k_{\negvthinspace j}}( \interp{\J}{u} )}$ as the value in $\interp{\J}{\tau}$ that is $\alpha$-equivalent with $\ValC \ec{ t }$.
-
-We now show that all disequalities in $\Fc$ are satisfied by $\J$.
-Assume $t \tneq u \in \Fc\!$.
-Since \rn{Conflict} does not apply, $t \teq u \notin \Fc$ and thus $\ec{t}$ and $\ec{u}$ are distinct.
-Since $\interp{\J}{t} \vsim \ValC \ec{t}$ and $\interp{\J}{u} \vsim \ValC \ec{u}$,
-by Lemma~\ref{lem:interpretation-completion}(3), $\interp{\J}{t} \neq \interp{\J}{u}$, and thus $\J \vDash t \tneq u$.
-
-Since $\Fc$ contains only equalities and disequalities, we have $\J \vDash \Fc\!$,
-and since $\Ec \subseteq \Fc\!$,
-% is a superset of $\Ec$,
-we conclude that $\J \vDash \Ec$.
-\xqed
-\end{proof}
 
@@
 
-%* and finally, we look in more detail at one or two such proofs, by presenting
-%  it and explaining it (and, before that, understanding it)
-%
-Overall, four proofs were found thanks to
-the acyclicity rule and 17~required uniqueness. Interestingly,
-no proofs were lost by enabling these rules. Among the 17 proofs requiring uniqueness,
-some were simple arguments of the form
-\keyw{by}~\textit{coinduction}~\textit{auto}
-in Isabelle \cite{blanchette-et-al-2014-impl}, while others
-involved more elaborate reasoning, including the following example
-about Stern--Brocot trees:
-%
-\begin{quote}
-\keyw{lemma} \,\textit{\nummodden\_unique}: \,$x = \const{Node}\;0\;\const{num}\; x \vthinspace\Longrightarrow\vthinspace x = \const{\nummodden}$ \\
-\keyw{proof} \,(\textit{coinduction arbitrary}: $x$ \textit{rule}: \textit{tree.coinduct\_strong}) \\
-\noindent\hbox{}\quad  \keyw{case} (\textit{Eq\_tree} $x$) \keyw{show} \textit{?case} \\
-\noindent\hbox{}\qquad  \keyw{by} (\textit{subst} (1 2 3 4) \textit{Eq\_tree}) (\textit{simp add}: \textit{eqTrueI}[OF \textit{Eq\_tree}]) \\
-\keyw{qed}
-\end{quote}
-%
-\noindent
-The tree \const{\nummodden} is defined as $\const{\nummodden} =
-\const{Node}\;0\;\const{num}\;\const{\nummodden}$.
-
 \section{Conclusion -- JB 0.25 column}

### 2016-02-09 sister/sister.tex
tweaking

@@
 \begin{abstract}
-%Codatatypes naturally capture potentially infinite data structures and
-%processes.
-We present a decision procedure that combines reasoning about
-datatypes and codatatypes. The dual of the acyclicity rule for datatypes is a
-uniqueness rule that identifies observationally equal codatatype values,
-including cyclic values. The procedure decides universal
-problems and is composable via the Nelson--Oppen method. It has been
-implemented in CVC4, a state-of-the-art SMT solver. An evaluation based on
-problems generated from formalizations developed with Isabelle demonstrates the
-potential of the procedure.
+Datatypes and codatatypes are useful to represent infinite and potentially
+infinite data structures and processes. We describe a decision procedure that
+combines reasoning about datatypes and codatatypes. The procedure has been
+integrated in CVC4, a state-of-the-art SMT (satisfiability modulo theories)
+solver, which can be used both as a general constraint solver and as an
+automatic theorem prover. An evaluation based on problems generated from
+interactive formal proofs developed using the Isabelle proof assistant
+demonstrates the potential of the decision procedure.
 \end{abstract}
@@
 have emerged in a few automatic theorem provers as well. The SMT-LIB
-\cite{barrett-et-al-2015}
-format, supported by most SMT
+format \cite{barrett-et-al-2015}, supported by most SMT
 (satisfiability modulo theories)
@@
 The procedure is described abstractly as a %tableau-like
-calculus and is composable via the Nelson--Oppen method \cite{nelson-oppen-1979}.
-It generalizes the procedure by Barrett et al.\ \cite{barrett-et-al-2007},
+calculus and is composable via the Nelson--Oppen \shortcite{nelson-oppen-1979}
+method.
+It generalizes the procedure by Barrett et al.\
+\shortcite{barrett-et-al-2007},
 which covers only datatypes.
@@
 $$\vthinspace\keyw{data}~\;\ty{fstream}_{\vthinspace\tau} \,=\allowbreak\, \const{FSCons}(\tau,\allowbreak\: \ty{fstream}_{\vthinspace\tau})\vthinspace$$
-would be rejected as non-well-founded
-\cite{blanchette-et-al-2015-esop}.
+would be rejected as non-well-founded.
 %
@@
 
-An earlier version of this article was presented at the CADE-25 conference in
-Berlin, Germany \cite{reynolds-blanchette-2015-codata}.
-This article extends the conference paper with additional
-background material on (co)datatypes (Section~\ref{sec:co-datatypes}), more
-detailed metatheoretical proofs (Section~\ref{sec:the-decision-procedure}),
-a description of the generation of models with $\mu$-binders in CVC4
-(Section~\ref{sec:implementation-as-a-theory-solver-in-cvc4}), and a more
-comprehensive evaluation (Section~\ref{sec:evaluation-on-isabelle-problems}).
+The original version of this paper was presented at the CADE-25 conference in
+Berlin, Germany \cite{reynolds-blanchette-2015-codata}. An more comprehensive
+version, including detailed proofs, has been accepted in a special issue of
+the \emph{Journal of Automated Reasoning}
+\cite{reynolds-blanchette-201x-codata}.
 
@@
 its negation is unsatisfiable. The presentation is
-inspired by Barrett et al.\ \cite{barrett-et-al-2007} but at
+inspired by Barrett et al.\ \shortcite{barrett-et-al-2007} but at
 a higher level, using unoriented equations instead of oriented ones.
@@
 included all the formalizations from the Isabelle distribution (Distro, 1179~goals)
-and the \emph{Archive of Formal Proofs} (AFP, 3014~goals) \cite{klein-et-al-afp}
-that define codatatypes falling within the supported fragment. We
+and the \emph{Archive of Formal Proofs} (AFP, 3014~goals) that define codatatypes falling within the supported fragment. We
 also included formalizations about Bird and Stern--Brocot trees (\gandl,
@@
 We introduced a decision procedure for the universal theory of datatypes and
-codatatypes. Our main
-contribution has been the support for codatatypes. Both the metatheory and
-the implementation in CVC4 rely on $\mu$-terms to represent cyclic values.
+codatatypes. Our approach relies on $\mu$-terms to represent cyclic values.
 Although this aspect is primarily motivated by codatatypes, it makes
@@
 crucial to exclude spurious countermodels.
-As future work, it would be useful to implement
-proof reconstruction for (co)data\-type inferences in Isabelle.
-In addition, it might be worthwhile to
-extend SMT solvers with dedicated support for (co)recursion.
-
-{
-\footnotesize
-\def\ackname{Acknowledgment}
-\paragraph{\footnotesize\ackname.}
-\looseness=-1
-We owe a great debt to the development team of CVC4,
-including Clark Barrett and Cesare Tinelli,
-and in particular Morgan Deters,
-who jointly with the first author
-developed the initial version
-of the theory solver for datatypes in CVC4.
-%
-Our present and former bosses, Viktor Kuncak, Stephan Merz, Tobias Nipkow,
-Cesare Tinelli, and Christoph Weidenbach, have either encouraged the research on
-codatatypes or at least benevolently tolerated it, both of which we are thankful
-for.
-%
-Peter Gammie and Andreas Lochbihler pointed us to useful benchmarks.
-Andrei Popescu helped clarify our thoughts regarding codatatypes and indicated
-related work. Dmitriy Traytel took part in discussions about degenerate
-codatatypes.
-%
-Pascal Fontaine, Andreas Lochbihler, Andrei Popescu, Christophe Ringeissen, Mark
-Summerfield, Dmitriy Traytel, and the anonymous reviewers suggested many
-textual improvements. The second author's work was partially supported by
+%As future work, it would be useful to implement
+%proof reconstruction for (co)data\-type inferences in Isabelle.
+%In addition, it might be worthwhile to
+%extend SMT solvers with dedicated support for (co)recursion.
+
+\section*{Acknowledgment}
+
+We owe a great debt to the development team of CVC4, including Clark Barrett
+and Cesare Tinelli, and in particular Morgan Deters, who jointly with the
+first author developed the initial version of the decision procedure for
+datatypes in CVC4.
+We want to thank everyone who helped us with the longer versions of
+this paper.
+The second author's work was partially supported by
 the Deutsche
@@
 
-}
-
 \bibliographystyle{named}

### 2016-02-09 sister/sister.tex
some work on intro

@@
 interactive formal proofs developed using the Isabelle proof assistant
-demonstrates the potential of the decision procedure.
+demonstrates the potential of our decision procedure.
 \end{abstract}
@@
 
+Automatic theorem proving
+* much focus on FOL
+* two main approaches:
+  superposition---e.g., E, SPASS, Vampire
+  SMT (satisfiability modulo theory) solvers---e.g., CVC4, veriT, Yices, Z3
+* SMT is also useful for model finding, or constraint solving: given some   
+  constraints, find an assignment of its variables that satisfy them
+
+* strength of the SMT approach is that it can compose different efficient
+  decision procedures and other solvers for various theories---e.g., equality,
+  linear arithmetic, and first-order quantifiers.
+* the main scientific question is: which theories would be useful to
+  applications, and which ones can be supported efficiently in SMT solvers?
+
 %  * a general introduction to automatic theorem proving \\
 
-%  * relation to AI
-%   * among the earliest AI goals
-%   * yet somewhat disconnected from mainstream AI
-%     (with some exceptions, e.g. Urban ..., MaSh)
-%
 %  * followed by an intro to SMT and what it's used for \\

### 2016-02-09 sister/sister.tex
got rid of one convention

@@
 
-\newcommand\typ[1]{^{\vthinspace #1}}
-
 \newcommand\MU{\vvthinspace\mu\vvthinspace}
@@
 Datatypes and codatatypes are an integral part of many proof assistants,
-including Agda, Coq, Isabelle, Matita, and PVS. In recent years, datatypes
-have emerged in a few automatic theorem provers as well. The SMT-LIB
-format \cite{barrett-et-al-2015}, supported by most SMT
-(satisfiability modulo theories)
-solvers, is being extended with a syntax for datatypes.
+including Coq, Isabelle, and PVS. In recent years, datatypes
+have emerged in a few automatic theorem provers as well.
 In this article, we introduce a
@@
 the $\mu$-term $\const{SCons}(1,\: \MU s.\; \const{SCons}(0,\: \const{SCons}(9,\: s)))$
-stands for the lasso-shaped sequence $1, 0, 9, 0, 9, %0, 9,
-\ldots$\afterLdots.
+stands for the lasso-shaped sequence $1, 0, 9, 0, 9, 0, 9,
+\dotsc$.
 
@@
 An evaluation on %hand-crafted examples and on
-problems generated from Isabelle/HOL \cite{nipkow-et-al-2002}
-formalizations using the Sledgehammer tool \cite{blanchette-et-al-2013-smt}
-demonstrates the usefulness of the approach (Section~\ref{sec:evaluation-on-isabelle-problems}).
+problems generated from Isabelle\slash HOL
+formalizations demonstrates the usefulness of the approach (Section~\ref{sec:evaluation-on-isabelle-problems}).
 
@@
 Berlin, Germany \cite{reynolds-blanchette-2015-codata}. An more comprehensive
-version, including detailed proofs, has been accepted in a special issue of
-the \emph{Journal of Automated Reasoning}
-\cite{reynolds-blanchette-201x-codata}.
-
-\paragraph{Conventions.}
-Our setting is a monomorphic (or many-sorted) first-order logic.
-A signature $\Sig = (\Types, \Funcs)$ consists of a set of types $\Types$ and a
-set of function symbols $\Funcs$\negvthinspace. Types are atomic sorts and interpreted as
-nonempty domains. The set~$\Types$ must contain a
-distinguished type \ty{bool} interpreted as the set of truth
-values. %, and may contain other interpreted types (e.g., \ty{int}, \ty{real}).
-The metavariables $\delta,\:\varepsilon$ range over (co)datatypes,
-whereas $\tau,\:\upsilon$ range over arbitrary types.
-
-Function symbols are written in a sans-serif font (e.g., $\const{f}$, $\const{g}$) to
-distinguish them from variables (e.g., $x$, $y$).
-Symbol names starting with an uppercase letter (e.g.,
-$\const{S}$) are reserved for constructors. With each function symbol \const{f}
-is associated a list of argument types $\tau_1,\ldots,\tau_n$ (for $n \ge 0$)
-and a return type $\tau$, written
-$\const{f} : \tau_1 \times \cdots \times \tau_n \to \tau$;
-this notation collapses to $\const{f} : \tau$ if $n = 0$.
-%
-Functions invocations $\const{f}(t_1,\ldots,t_n)$
-apply the $n$-ary function symbol
-\const{f} to $n$ %well-typed
-arguments $t_1 \mathbin{:} \tau_1$, \ldots, $t_n \mathbin{:}
-\tau_n$ of the right types.
-Nullary function symbols, also called constants, can appear without
-parentheses in terms.
-The set $\Funcs$ must at least contain
-$\const{true},\, \const{false} : \ty{bool}$, interpreted as truth values.
-The only predicate is equality, written $\teq$;
-it belongs to the logical symbols.
-Other predicates can be represented as functions to $\ty{bool}$,
-with $\const{p}(\ldots)$ abbreviating $\const{p}(\ldots) \teq \const{true}$.
-The notation $t\typ{\tau}$ stands for a term $t$ of type~$\tau$;
-it should not be confused with $t :\nobreak \tau$,
-which is a statement expressing that $t$ has type $\tau$.
-When applied to terms, the standard equality symbol~$=$ denotes syntactic equality.
-The operator $\bigwedge_{\vthinspace i}\, \varphi_i$ abbreviates a conjunction
-$\varphi_1 \mathrel\land \cdots \mathrel\land \varphi_n$.
-Finally, $\bar x$ abbreviates a list or tuple $x_1,\ldots,x_n$.
+version, including detailed proofs, will appear in a special issue of the
+\emph{Journal of Automated Reasoning}.
+%\cite{reynolds-blanchette-201x-codata2}.
 
@@
 
-We fix a signature $\Sig = (\Types, \Funcs)$. The types are partitioned into
+Our setting is a monomorphic (or many-sorted) first-order logic.
+We fix a signature $\Sig = (\Types, \Funcs)$ consisting of a set of types $\Types$ and a
+set of function symbols $\Funcs$\negvthinspace.
+The types are partitioned into
 $\Types = \Data \mathrel{\uplus} \Codata \mathrel{\uplus} \Nondata$, where $\Data$ are the
@@
 
-%\paragraph{Specifications.}
+%\paragraph{Specifications}
 In an SMT problem, the signature is typically given by specifying first the
@@
 
-\paragraph{Phase 1: Computing the Bidirectional Closure \rm(Fig.~\ref{fig:cc-rules}).}
+\paragraph{Phase 1: Computing the Bidirectional Closure \rm(Fig.~\ref{fig:cc-rules})}
 In conjunction with \rn{Refl}, \rn{Sym}, and \rn{Trans}, the \rn{Cong} rule computes the congruence (upward) closure,
@@
 
-\paragraph{Phase 2: Applying Acyclicity and Uniqueness \rm(Fig.~\ref{fig:ab-rules}).}
+\paragraph{Phase 2: Applying Acyclicity and Uniqueness \rm(Fig.~\ref{fig:ab-rules})}
 %For presentation of the rules in this phase, we rely on a representation of terms in the $\mu$-notation.
@@
 
-The $\mu$-term $\Val[t\typ{\tau}]$ describes a class of $\tau$ values
+The $\mu$-term $\Val[t]$ describes a class of values
 that $t$ and other members of $t$'s equivalence class can take in models of $\Ec$.
@@
 \inferrule{
-%  t\typ{\delta}
-%  \betweenantes
-  \delta \in \Data
+  t : \delta \in \Data
   \betweenantes
-  \Val \smash{\ec{t\typ{\delta}}} = \MU x.\; u
+  \Val \smash{\ec{t}} = \MU x.\; u
   \betweenantes
@@
 \rn{Acyclic}
-\kern1.5em\( %%% TYPESETTING: should be \quad
+\\[5\jot]
+\(
 \inferrule{
-% t, u\typ{\delta}
-% \betweenantes
- \delta \in \Codata
+ t, u : \delta \in \Codata
  \betweenantes
- \Val \smash{\ec{t\typ{\delta}}} \aequiv \Val \smash{\ec{u\typ{\delta}}}
+ \Val \smash{\ec{t}} \aequiv \Val \smash{\ec{u}}
 }{
@@
 \inferrule{
-  t\typ{\delta} \in \tEc
+  t : \delta
+  \betweenantes
+  t \in \tEc
   \betweenantes
@@
 \inferrule{
-  t\typ{\delta}, u\typ{\delta} \in \tEc
+  t, u : \delta \in \Codata
   \betweenantes
-  \delta \in \Codata
+  t, u \in \tEc
   \betweenantes
@@
 
-\paragraph{Phase 3: Branching \rm(Fig.~\ref{fig:split-rule}).}
+\paragraph{Phase 3: Branching \rm(Fig.~\ref{fig:split-rule})}
 If a selector is applied to a term $t$, or if $t$'s type is a finite datatype,
@@
 
-\paragraph{Termination and Correctness.}
-\label{sec:termination-and-correctness}
-
+\paragraph{Termination and Correctness}
 We now show the termination and correctness of the calculus.
@@
 Isabelle formalizations, using Sledgehammer \cite{blanchette-et-al-2013-smt}
-as translator from Isabelle to SMT-LIB.
+as translator.
 %There are not yet many benchmarks available, since codatatypes are a

### 2016-02-09 jour/jour.tex
tuning

@@
 \const{zero} and $\const{sconst}~0$ by two distinct
-values that both correspond to the \hbox{$\mu$-term}
+values that both correspond to the term
 $\MU s.\; \const{SCons}(0, s)$,
@@
 By contrast, the contributions of the decision procedure
-to proving are less clear; they depend on how
+to proving are less obvious; they depend on how
 often acyclicity and uniqueness are necessary for a proof.

### 2016-02-09 jour/jour.tex
ooch

@@
 types in any order, then the (co)datatypes with their constructors
-and selectors in groups of mutually (co)recursive groups of (co)datatypes, and
+and selectors in groups of mutually (co)recursive (co)datatypes, and
 finally any other function symbols.

### 2016-02-09 sister/sister.tex
tuned sect 2

@@
 Datatypes and codatatypes are useful to represent infinite and potentially
-infinite data structures and processes. We describe a decision procedure that
-combines reasoning about datatypes and codatatypes. The procedure has been
-integrated in CVC4, a state-of-the-art SMT (satisfiability modulo theories)
-solver, which can be used both as a general constraint solver and as an
-automatic theorem prover. An evaluation based on problems generated from
-interactive formal proofs developed using the Isabelle proof assistant
-demonstrates the potential of our decision procedure.
+infinite data structures and processes. We describe a decision procedure
+to reason about these types. The procedure has been
+integrated in CVC4, a modern SMT (satisfiability modulo theories)
+solver, which can be used both as a constraint solver and as an
+automatic theorem prover. An evaluation based on formalizations developed
+in the Isabelle proof assistant demonstrates the potential of our
+procedure.
 \end{abstract}
 
-\section{Introduction -- JB 2 columns}
+\section{Introduction}
 \label{sec:introduction}
 
-Automatic theorem proving
-* much focus on FOL
-* two main approaches:
-  superposition---e.g., E, SPASS, Vampire
-  SMT (satisfiability modulo theory) solvers---e.g., CVC4, veriT, Yices, Z3
-* SMT is also useful for model finding, or constraint solving: given some   
-  constraints, find an assignment of its variables that satisfy them
-
-* strength of the SMT approach is that it can compose different efficient
-  decision procedures and other solvers for various theories---e.g., equality,
-  linear arithmetic, and first-order quantifiers.
-* the main scientific question is: which theories would be useful to
-  applications, and which ones can be supported efficiently in SMT solvers?
-
-%  * a general introduction to automatic theorem proving \\
-
-%  * followed by an intro to SMT and what it's used for \\
-%  * followed to why theories are important (more efficient reasoning) \\
-%  * followed to why (co)datatypes are important
-%
-%  * both for proving
-%   * not finitely axiomatizable
-%  * and for model finding
-%   * Nunchaku
-
-Freely generated algebraic datatypes are ubiquitous in functional programs and
-formal specifications. They are especially
-useful to represent finite data structures in computer science applications but
-also arise in formalized mathematics.
-They can be implemented efficiently and enjoy
-properties that can be exploited in automated reasoners.
-%
-%However, because datatype values correspond to finite ground terms, they
-%are generally not adequate to represent infinite objects.
-%For example, the datatype of natural numbers
-%constructed by $\const{Z} : \ty{nat}$ and $\const{S} : \ty{nat} \to \ty{nat}$
-%only allow values of the form $\const{S}(\ldots(\const{S}(\const{Z}))\ldots)$.
-
-To represent infinite objects, % such as $\const{S}(\const{S}(\const{S}(\ldots))$,
-a natural choice is to turn to coalgebraic datatypes, or \emph{codatatypes},
-the non-well-founded dual of algebraic \emph{datatypes.}
-%
-Despite their reputation for being esoteric, codatatypes have a
+In the past decade, satisfiability modulo theories (SMT) solvers
+\cite{nieuwenhuis-et-al-2006}
+% \cite{ganzinger-et-al-2004}
+have emerged as one of the most powerful ways to
+prove theorems in first-order logic. These solvers are also useful for model
+finding, or constraint solving: given some constraints, they can often find a
+variable assignment that satisfy them.
+
+A great benefit of the SMT approach is that it provides a general framework
+for composing optimized decision procedures and other solvers for various
+theories, including equality, linear arithmetic on $\mathbb{Z}$ and
+$\mathbb{R}$, and first-order quantifiers ($\forall$,~$\exists$). In this
+context, two recurrent scientific questions are:
+\begin{itemize}
+\item Which theories would be useful to applications?
+\item Which theories can be coded efficiently in SMT solvers?
+\end{itemize}
+
+This paper suggests the following partial answer:\ freely generated algebraic
+and coalgebraic datatypes---also called \emph{datatypes} and
+\emph{codatatypes}, respectively.
+Datatypes are ubiquitous in functional programs and formal specifications.
+They are useful to represent finite data structures in computer
+science but also arise in formalized mathematics.
+To represent infinite objects,
+a natural choice is to turn to codatatypes,
+the non-well-founded dual of datatypes.
+%
+Despite their reputation for esotericism, codatatypes have a
 role to play in computer science. The verified C compiler CompCert
@@
 infinit\-e constructor terms are also legitimate values for codatatypes
-(Section~\ref{sec:co-datatypes}). Intuitively, the
-values of a codatatype consist of all well-typed finite \emph{and infinite} ground
+(Section~\ref{sec:co-datatypes}). The
+values of a codatatype consist of all well-typed finite \emph{and infinite} variable-free
 constructor
@@
 introduces a type that
-models the natural numbers $\const{Z}$, $\const{S}(\const{Z})$, $\const{S}(\const{S}(\const{Z}))$, $\ldots$\afterLdots{},
-in Peano notation, extended with an
+captures the natural numbers $\const{Z}$, $\const{S}(\const{Z})$, $\const{S}(\const{S}(\const{Z}))$, $\ldots$\afterLdots{},
+in Peano (unary) notation, extended with an
 infinite value $\infty = \const{S}(\const{S}(\const{S}(\ldots)))$.
@@
 have emerged in a few automatic theorem provers as well.
-In this article, we introduce a
+In this paper, we describe a
 unified decision procedure for universal problems involving datatypes and codatatypes
@@
 The procedure is described abstractly as a %tableau-like
-calculus and is composable via the Nelson--Oppen \shortcite{nelson-oppen-1979}
-method.
+calculus and can be composed with other theories in an SMT solver.
 It generalizes the procedure by Barrett et al.\
@@
 of infinite sequences or streams over an element type $\tau$ is
-allowed, even though it has no base case. By contrast, the
-corresponding datatype specification
-$$\vthinspace\keyw{data}~\;\ty{fstream}_{\vthinspace\tau} \,=\allowbreak\, \const{FSCons}(\tau,\allowbreak\: \ty{fstream}_{\vthinspace\tau})\vthinspace$$
-would be rejected as non-well-founded.
-%
+allowed, even though it has no base case.
 Second, \emph{a uniqueness rule takes the place of the acyclicity rule of datatypes.}
@@
 $y \teq \const{S}(y)$, it deduces $x \teq y$.
-These two rules are needed to ensure completeness
-(solution soundness) on universal problems. %and the absence of spurious models.
-They cannot be finitely axiomatized, so they naturally belong in
+These two rules cannot be finitely axiomatized, so they naturally belong in
 a decision procedure.
@@
 %
-Third, \emph{it must be possible to express cyclic }(\emph{regular}) \emph{values as closed terms and
+Third, \emph{it must be possible to express cyclic values as closed terms and
 to enumerate them.} This is necessary when generating models.
 The $\mu$-binder notation associates a name with
-a subterm; it is used to represent cyclic values in the metatheory and
-in the generated models. For example,
+a subterm, making it possible to represent cyclic values.
+For example,
 the $\mu$-term $\const{SCons}(1,\: \MU s.\; \const{SCons}(0,\: \const{SCons}(9,\: s)))$
@@
 
-Our procedure is implemented in the SMT solver CVC4 \cite{barrett-et-al-2011} as a combination
-of rewriting and a theory solver (Section~\ref{sec:implementation-as-a-theory-solver-in-cvc4}).
-It consists of about 2000 lines of \cpp{} code,
-most of which are shared between datatypes and codatatypes. The code is
-integrated in the development version of the solver and is expected to be part
-of the CVC4~1.5 release.
+Our procedure is implemented in the SMT solver CVC4 \cite{barrett-et-al-2011}. 
+It consists of about 2000 lines of \cpp{}. % code.
+%most of which are shared between datatypes and codatatypes.
 %
@@
 problems generated from Isabelle\slash HOL
-formalizations demonstrates the usefulness of the approach (Section~\ref{sec:evaluation-on-isabelle-problems}).
+formalizations demonstrates its usefulness (Section~\ref{sec:evaluation}).
 
@@
 \section{%The Theory of
-(Co)datatypes -- AJR 0.75 column}
+(Co)datatypes}
 \label{sec:co-datatypes}
 
-Our setting is a monomorphic (or many-sorted) first-order logic.
-We fix a signature $\Sig = (\Types, \Funcs)$ consisting of a set of types $\Types$ and a
+Our setting is a monomorphic (many-sorted) first-order logic.
+We fix a signature %$\Sig = (\Types, \Funcs)$
+consisting of a set of types $\Types$ and a
 set of function symbols $\Funcs$\negvthinspace.
@@
 \emph{datatypes}, $\Codata$ are the \emph{codatatypes}, and $\Nondata$ are the remaining
-\emph{ordinary types} (which can be interpreted or not). The function symbols are
+\emph{ordinary types}.
+% (which can be interpreted or not).
+The function symbols are
 partitioned into $\Funcs = \Ctr \mathrel{\uplus} \Sel$, where $\Ctr$ are the
@@
 variables when combining theories.
-Exceptionally, it is convenient to use numeric constants ($0$, $1$, \ldots)\ in
-examples.
-$\Sig$-terms are standard first-order terms
-over $\Sig$,
-without $\mu$-binders.
+%Exceptionally, it is convenient to use numeric constants ($0$, $1$, \ldots)\ 
+%in
+%examples.
+%$\Sig$-terms are standard first-order terms
+%over $\Sig$,
+%without $\mu$-binders.
 %(Section~\ref{sec:implementation-as-a-theory-solver-in-cvc4}).
@@
 types in any order, then the (co)datatypes with their constructors
-and selectors in groups of mutually (co)recursive groups of (co)datatypes, and
+and selectors in groups of $\ell$ mutually recursive datatypes or corecursive
+codatatypes, and
 finally any other function symbols.
-%
-Each (co)datatype specification consists of $\ell$~mutually recursive types that are
-either all datatypes or all codatatypes. Polymorphic types, nested
-(co)recursion, and datatype--codatatype mixtures
-fall outside this fragment.
 
@@
 Each (co)datatype $\delta$ is equipped with
-$m \ge 1$ constructors, and each constructor for $\delta$ takes zero or more
-arguments and returns a $\delta$ value. The argument types must be either
-ordinary, among the already known (co)datatypes, or among the (co)datatypes
-being introduced.
+$m \ge 1$ constructors, and each constructor %for $\delta$
+takes zero or more arguments and returns a $\delta$ value.
+%The argument types must be either
+%ordinary, among the already known (co)datatypes, or among the (co)datatypes
+%being introduced.
 %
 To every argument corresponds a selector. The names for the (co)data\-types, the
-constructors, and the selectors must be
-distinct and different from existing names.
+constructors, and the selectors must be fresh.
+%distinct and different from existing names.
 Schematically:
@@
 \!(\keyw{co})\keyw{data}\;\,
-  \delta_1 & {}=\, \smash{\const{C}_{11\!}(\bigl[\const{s}_{11\!}^1{:}\bigr]\vthinspace \tau_{11\!}^1, \ldots, \bigl[\const{s}_{11\!}^{n_{11\!}}{:}\bigr]\vthinspace \tau_{11\!}^{n_{11\!}})} \mid \cdots \mid \smash{\const{C}_{1m_1\!}(\ldots)} \\[-.5\jot]
-   \smash{\vdots\,\,\,} \\[-1.5\jot]
+  \delta_1 & {}=\, \smash{\const{C}_{11\!}(\bigl[\const{s}_{11\!}^1{:}\bigr]\vthinspace \tau_{11\!}^1, \ldots, \bigl[\const{s}_{11\!}^{n_{11\!}}{:}\bigr]\vthinspace \tau_{11\!}^{n_{11\!}})} \mid \cdots \mid \smash{\const{C}_{1m_1\!}(\ldots)} \\[+0.5\jot]
+   \smash{\vdots\,\,\,} \\[-0.5\jot]
   \keyw{and}\; \,\delta_{\elllx} & =\, \smash{\const{C}_{\elll 1\!}(%\bigl[\const{s}_{\elll 1\!}^1{:}\bigr]\vthinspace \tau_{\elll 1\!}^1, \ldots, \bigl[\const{s}_{\elll 1\!}^{n_{\elll 1\!}}{:}\bigr]\vthinspace \tau_{\elll 1\!}^{n_{\elll 1\!}}
@@
 and $\smash{\const{s}_{i\negvthinspace j}^{\,k} : \delta_i \to \tau_{i\negvthinspace j}^{\,k}}$.
-Defaults are assumed for the selector names if they are omitted.
+%Defaults are assumed for the selector names if they are omitted.
 The $\delta$ constructors and selectors are denoted by $\Ctr^\delta$ and
@@
 For types with several constructors, it is useful to provide discriminators
-$\const{d}_{i\negvthinspace j} : \delta_i \to \ty{bool}$. Instead of extending $\Funcs$,
-we let $\const{d}_{i\negvthinspace j}(t)$
+$\const{d}_{i\negvthinspace j} : \delta_i \to \ty{bool}$.
+We let $\const{d}_{i\negvthinspace j}(t)$
 be an abbreviation for
-$t \teq \const{C}_{i\negvthinspace j}\bigl(\const{s}_{i\negvthinspace j}^1(t), \ldots, \const{s}_{i\negvthinspace j}^{\,n_{\smash{{i\negvthinspace j}}}}(t)\bigr)$.
+$t \teq \const{C}_{i\negvthinspace j}\relax(\const{s}_{i\negvthinspace j}^1(t), \ldots, \const{s}_{i\negvthinspace j}^{\,n_{\smash{{i\negvthinspace j}}}}(t)\relax)$.
 %This will simplify the presentation. % of the decision procedure.
@@
 %
-Expressed in the algebraic jargon, exhaustiveness helps ensure that ``no
-junk'' exists, whereas distinctness and injectivity guarantee that ``no
-confusion'' can arise.
-The result of selectors applied to the wrong
-constructor is left completely unspecified.
+%In the algebraic jargon, exhaustiveness helps ensure that ``no
+%junk'' exists, whereas distinctness and injectivity guarantee that ``no
+%confusion'' can arise.
+%The result of selectors applied to the wrong
+%constructor is left completely unspecified.
 %
@@
 %
+%
 Datatypes are additionally characterized by an induction principle.
@@
 induction prohibits models that contain cyclic values---e.g.,
-an $n$ such that $n \teq \const{S}(n)$---or even infinite acyclic values
-$\const{S}(\const{S}(\ldots))$.
-
-For codatatypes, the dual notion is called coinduction.
-By this principle, two values that yield the same observations must be equal, 
-where the observations are made through selectors and discriminators. In 
-addition, codatatypes are guaranteed to contain all values corresponding to 
-infinite ground constructor terms.
-
-Given a signature $\Sig$, $\thD$ refers to the \emph{theory of
-datatypes and codatatypes},
-which %in addition to $\Sig$
-defines a class of $\Sig$-interpretations $\JJJJ$,
-namely those that satisfy the %axioms --- existence of all infinite codatatype values cannot be axiomatized
-properties mentioned in this section, including (co)induction. The interpretations
-in~$\JJJJ$ share the same interpretation for constructor terms and correctly
-applied selector terms (up to isomorphism) %renaming of domain elements
-but may differ on variables and wrongly applied selector terms.
-A formula is $\thD$-\emph{satisfiable} if there
-exists an interpretation in $\JJJJ$ that satisfies it.
-% otherwise, it is $\thD$-unsatisfiable. (I trust the readers' intelligence when
-% it comes to English prefixes like non-, in-, un-, and a-.)
-%
-For deciding universal formulas, induction can be replaced by the acyclicity
-axiom schema, which states that constructor terms cannot be equal to
-any of their proper subterms \cite{barrett-et-al-2007}.
-Dually, coinduction can be replaced by the uniqueness schema, which
-asserts that codatatype values are fully characterized by their %possibly
-%infinite
-expansion \cite[Theorem 8.1, 2$\,{\Leftrightarrow}\,$5]{rutten-2000}.
+an $n$ such that $n \teq \const{S}(n)$---or even infinite acyclic values.
+% $\const{S}(\const{S}(\ldots))$.
+%
+For codatatypes, the dual notion is called coinduction:
+Two values that yield the same observations must be equal, 
+where the observations are made through selectors and discriminators.
+Codatatypes are also guaranteed to contain all values corresponding to 
+infinite variable-free constructor terms.
+
+%Given a signature $\Sig$, $\thD$ refers to the \emph{theory of
+%datatypes and codatatypes},
+%which %in addition to $\Sig$
+%defines a class of $\Sig$-interpretations $\JJJJ$,
+%namely those that satisfy the %axioms --- existence of all infinite %codatatype values cannot be axiomatized
+%properties mentioned in this section, including (co)induction. The %interpretations
+%in~$\JJJJ$ share the same interpretation for constructor terms and correctly
+%applied selector terms (up to isomorphism) %renaming of domain elements
+%but may differ on variables and wrongly applied selector terms.
+%A formula is $\thD$-\emph{satisfiable} if there
+%exists an interpretation in $\JJJJ$ that satisfies it.
+%% otherwise, it is $\thD$-unsatisfiable. (I trust the readers' intelligence %when
+%% it comes to English prefixes like non-, in-, un-, and a-.)
+%%
+%For deciding universal formulas, induction can be replaced by the acyclicity
+%axiom schema, which states that constructor terms cannot be equal to
+%any of their proper subterms \cite{barrett-et-al-2007}.
+%Dually, coinduction can be replaced by the uniqueness schema, which
+%asserts that codatatype values are fully characterized by their %possibly
+%%infinite
+%expansion \cite[Theorem 8.1, 2$\,{\Leftrightarrow}\,$5]{rutten-2000}.
 
@@
 
-\section{The Decision Procedure -- AJR 4 columns} % for (Co)datatypes}
+\section{The Decision Procedure} % for (Co)datatypes}
 \label{sec:the-decision-procedure}
 
-Given a fixed signature~$\Sig$,
+Given a fixed signature, %~$\Sig$,
 the decision procedure for the universal theory of (co)data\-types %$\thD$
 determines the
-$\thD$-satisfiability of finite sets $\Ec$ of literals:\ equalities
-and disequalities between $\Sig$-terms,
+%$\thD$-
+satisfiability of finite sets $\Ec$ of literals:\ equalities
+and disequalities between first-order terms, %$\Sig$-terms,
 whose variables are interpreted existentially.
@@
 To simplify the presentation, we make a few assumptions about
-$\Sig$.
+the signature. %$\Sig$.
 First, all codatatypes %in $\Codata$
@@
 
-Our calculus for $\thD$ consists of derivation rules.
+Our calculus for the theory of (co)datatypes
+consists of derivation rules.
 A derivation rule can be applied to $\Ec$ if %all
@@
 
-Let $\tEc$ denote the set of $\Sig$-terms occurring in $\Ec$.
+Let $\tEc$ denote the set of %$\Sig$-
+terms occurring in $\Ec$.
 At the end of the first phase, $\Ec$ induces an equivalence
@@
 (expressed by $x \in \FV( u )$),
-then $\Ec$ is $\thD$-unsatisfiable.
+then $\Ec$ is %$\thD$-
+unsatisfiable.
 For uniqueness, if $\ec{t}$, $\ec{u}$ are two codatatype equivalence classes
@@
 If there exists a closed derivation tree with root node
-$\Ec$, then $\Ec$ is $\thD$-unsatisfiable; and if there exists
+$\Ec$, then $\Ec$ is %$\thD$-
+unsatisfiable; and if there exists
 a derivation tree with root node $\Ec$ that contains a saturated node, then
-$\Ec$ is $\thD$-satisfiable.
+$\Ec$ is %$\thD$-
+satisfiable.
 
@@
 \afterDot
-If there exists a closed derivation tree with root node\/ $\Ec$, then\/ $\Ec$ is $\thD$-unsatisfiable.
+If there exists a closed derivation tree with root node\/ $\Ec$, then\/ $\Ec$ is %$\thD$-
+unsatisfiable.
 \end{theorem}
-
-It remains to show the converse of the previous theorem: If a derivation tree
-with root node $\Ec$ contains a saturated node, then $\Ec$ is
-$\thD$-satisfiable.
-The proof relies on a specific interpretation $\J$ that satisfies $\Ec$.
 
@@
 \afterDot
-If there exists a derivation tree with root node $\Ec$ containing a saturated node, then $\Ec$ is $\thD$-satisfiable.
+If there exists a derivation tree with root node $\Ec$ containing a saturated node, then $\Ec$ is %$\thD$-
+satisfiable.
 \end{theorem}
@@
 
-\section{Evaluation -- JB 1 column} % on Isabelle Problems}
+\section{Implementation and Evaluation} % on Isabelle Problems}
 \label{sec:evaluation}
@@
 negation, in the refutational style) and for model finding
-\cite{reynolds-et-al-2013,ge-de-moura-2009}.
+\cite{ge-de-moura-2009,reynolds-et-al-2013}.
 It is in fact vital for finite model finding,
@@
 soundness, without which the generated models would often be
-spurious. For example, given the axioms
+spurious. For example, given the constraints
 \begin{align*}
@@
 \end{align*}
-on streams, the conjecture
+the conjecture
 $\const{zeros} = \const{sconst}~0$
@@
 \const{zero} and $\const{sconst}~0$ by two distinct
-values that both correspond to the \hbox{$\mu$-term}
-$\MU s.\; \const{SCons}(0, s)$,
+values $\MU s.\; \const{SCons}(0, s)$,
 violating uniqueness.
@@
 also included formalizations about Bird and Stern--Brocot trees (\gandl,
-265~goals) \cite{gammie-lochbihler-2016}. To exercise the datatype support, formalizations about finite
+265~goals). % \cite{gammie-lochbihler-2016}.
+To exercise the datatype support, formalizations about finite
 lists and trees
 were added to the first two benchmark sets. The formalizations were selected before
-conducting any experiments. The experimental data are available
-online.%
-\footnote{\url{http://lara.epfl.ch/~reynolds/JAR-CADE2015-cdt/}}
+conducting any experiments.
+%The experimental data are available
+%online.%
+%\footnote{\url{http://lara.epfl.ch/~reynolds/JAR-CADE2015-cdt/}}
 
@@
 
-\section{Conclusion -- JB 0.25 column}
+\section{Conclusion}
 \label{sec:conclusion}
@@
 for generating counterexamples in proof assistants
-\cite{reynolds-et-al-2015-smt};
+\cite{reynolds-et-al-2015-smt}.
 %as an alternative to Nitpick;
-in this context, the acyclicity and uniqueness rules are
-crucial to exclude spurious countermodels.
+The acyclicity and uniqueness rules are
+crucial to exclude spurious counterexamples.
 %As future work, it would be useful to implement
@@
 datatypes in CVC4.
-We want to thank everyone who helped us with the longer versions of
+We also want to thank everyone who helped us with the longer versions of
 this paper.

### 2016-02-09 sister/sister.tex
more

@@
 
-\section{Implementation and Evaluation} % on Isabelle Problems}
+\section{Evaluation} % on Isabelle Problems}
 \label{sec:evaluation}
@@
 \newcommand\gandl{SBT}
-\newcommand\HD[1]{\hbox to2em{\footnotesize\hfill\!\!\!\!#1\!\!\!\!\hfill}}
+\newcommand\HD[1]{\hbox to2em{\relax\hfill\!\!\!\!#1\!\!\!\!\hfill}}
 
@@
 lists and trees
-were added to the first two benchmark sets. The formalizations were selected before
-conducting any experiments.
+were added to the first two benchmark sets.
+%The formalizations were selected before
+%conducting any experiments.
 %The experimental data are available
@@
 Z3 4.3.2 for comparison, each running for
-up to 60~s on StarExec \cite{stump-et-al-2014-starexec}.
+up to 60~s. % on StarExec \cite{stump-et-al-2014-starexec}.
 Problems not involving any (co)datatypes were left out.
-Due to the lack of machinery in Isabelle for parsing CVC4 proofs and
-reconstructing inferences about (co)datatypes, the solvers are trusted
-as oracles. % in these experiments.
-%Z3 is included for comparison.
 
@@
 The problems with 16~lemmas put more stress on the decision procedure but are
-less typical of Sledgehammer-generated problems.
+less typical of Sledgehammer problems.
 
@@
 
-\begin{table*}[t!]
+%%% TYPESETTING: hack
+\setcounter{table}{3}
+
+\begin{table}[t!]
 \normalsize
-\begin{center}\begin{tabular}{@{\;}l@{\kern1.5em}c@{\kern.5em}c@{\kern1em}c@{\kern.5em}c@{\kern1em}c@{\kern.5em}c@{\kern1.5em}c@{\kern.5em}c@{\;}}
-  & \multicolumn{2}{@{}c@{\kern1em}}{Distro} & \multicolumn{2}{@{}c@{\kern1em}}{AFP} & \multicolumn{2}{@{}c@{\kern1.5em}}{\gandl} & \multicolumn{2}{@{}c@{\;}}{Overall} \\[-0.5pt]
-  & \HD{CVC4} & \HD{Z3} & \HD{CVC4} & \HD{Z3} & \HD{CVC4} & \HD{Z3} & \HD{CVC4}& \HD{Z3} 
+\begin{center}\begin{tabular}{@{\;}l@{\kern1.5em}c@{\kern.5em}c@{\kern1em}c@{\kern.5em}c@{\;}}
+  & \multicolumn{2}{@{}c@{\kern1em}}{16 lemmas\;} & \multicolumn{2}{@{}c@{\;}}{256 lemmas} \\%[-0.5pt]
+  & \HD{CVC4} & \HD{Z3} & \HD{CVC4} & \HD{\,Z3}
 \MIDRULE
 No (co)datatypes
-  & 287 & 282 & 765 & 771 & 47 & 44 & 1099 & 1097 \\%[\jot]
-%Datatype simplification only
-%  &  \\
+ &  1099  &  1097  &  2209  &  1911 \\%[\jot]
 Datatypes without \rn{Acyclic}
-  & 298 & -- & 771 & -- & 47 & -- & 1116 & -- \\
+ &  1116  &  -- &  2211  &  -- \\
 Full datatypes
-  & 298 & 294 & 775 & 783 & 47 & 44 & 1120 & 1121 \\%[\jot]
-%Codatatype simplification only
-%  &  \\
+ &  1120  &  1121  &  2211  &  1901 \\%[\jot]
 Codatatypes without \rn{Unique}\!\!
-  & 288 & -- & 797 & -- & 47 & -- & 1132 & -- \\
+ &  1132  &  --  &  2208  &  -- \\
 Full codatatypes
-  & 288 & -- & 797 & -- & \bfseries 52 & -- & 1137 & -- \\%[\jot]
+ &  1137  &  -- &  \bfseries 2220  &  -- \\%[\jot]
 Full (co)datatypes
-  & \bfseries 299 & -- & \bfseries 806 & -- & \bfseries 52 & -- & \bfseries 1157 & -- %\\[\jot]
+ &  \bfseries 1157  &  -- &  2219  &  -- %\\[\jot]
 \end{tabular}\end{center}
 \caption{\captioni}
-\label{tab:benchsixteen}
-\end{table*}
-
-\begin{table*}[t!]
-\normalsize
-\begin{center}\begin{tabular}{@{\;}l@{\kern1.5em}c@{\kern.5em}c@{\kern1em}c@{\kern.5em}c@{\kern1em}c@{\kern.5em}c@{\kern1.5em}c@{\kern.5em}c@{\;}}
-  & \multicolumn{2}{@{}c@{\kern1em}}{Distro} & \multicolumn{2}{@{}c@{\kern1em}}{AFP} & \multicolumn{2}{@{}c@{\kern1.5em}}{\gandl} & \multicolumn{2}{@{}c@{\;}}{Overall} \\[-0.5pt]
-  & \HD{CVC4} & \HD{Z3} & \HD{CVC4} & \HD{Z3} & \HD{CVC4} & \HD{Z3} & \HD{CVC4}& \HD{Z3} 
-\MIDRULE
-No (co)datatypes
-  & 617 & 560 & 1503 & 1271 & 89 & 80 & 2209 & 1911 \\%[\jot]
-%Datatype simplification only
-%  &  \\
-Datatypes without \rn{Acyclic}
-  & 617 & -- & \bfseries 1504 & -- & 90 & -- & 2211 & -- \\
-Full datatypes
-  & 617 & 560 & \bfseries 1504 & 1263 & 90 & 78 & 2211 & 1901 \\%[\jot]
-%Codatatype simplification only
-%  &  \\
-Codatatypes without \rn{Unique}\!\!
-  & 617 & -- & 1501 & -- & 90 & -- & 2208 & -- \\
-Full codatatypes
-  & \bfseries 620 & -- & 1502 & -- & 98 & -- & \bfseries 2220 & -- \\%[\jot]
-Full (co)datatypes
-  & 619 & -- & 1501 & -- & \bfseries 99 & -- & 2219 & -- %\\[\jot]
-\end{tabular}\end{center}
-\caption{\captionii}
-\label{tab:benchtwofiftysix}
-\end{table*}
-
-The results are summarized in Tables \ref{tab:benchsixteen} and
-\ref{tab:benchtwofiftysix}.
-%The goals in Table \ref{tab:benchsixteen} include 16
-%heuristically selected lemmas (before monomorphization);
-%\cite{meng-paulson-2009-relev};
-%that number is increased to 256 for Table \ref{tab:benchtwofiftysix}.
-The decision procedure
-makes a difference across all three benchmark suites. For the 16-lemma
-problems, it accounts for an overall
+\label{tab:bench}
+\end{table}
+
+The results are summarized in Table \ref{tab:bench}.
+For the 16-lemma problems, it accounts for an overall
 success rate increase of over 5\%.
@@
 could not be proved without it, but failed for 17~goals that could be proved
-without it. This potentially points to poor interactions between the decision
-procedure and the quantifier instantiation module
-\cite{ReynoldsTinelliMoura14}.
+without it.
+%This potentially points to poor interactions between the decision
+%procedure and the quantifier instantiation module
+%\cite{ReynoldsTinelliMoura14}.
 

### 2016-02-09 sister/sister.tex
more

@@
 
-\newtheorem{examplex}{Example}%{\itshape}{\rmfamily}
 \newtheorem{lemma}{Lemma}
 \newtheorem{theorem}{Theorem}
+
+\theoremstyle{definition}
+\newtheorem{examplex}{Example}%{\itshape}{\rmfamily}
 \newtheorem{definition}{Definition}
@@
 Despite their reputation for esotericism, codatatypes have a
-role to play in computer science. The verified C compiler CompCert
-\cite{leroy-2009}, the verified Java compiler Jinja\-Threads
-\cite{lochbihler-2010-jinja}, and the formalized Java memory model
-\cite{lochbihler-2014-jmm} all depend on codatatypes to capture infinite
+role to play in computer science. For example,
+Leroy's verified C compiler % Comp\-Cert
+\shortcite{leroy-2009} %, the verified Java compiler Jinja\-Threads
+%\cite{lochbihler-2010-jinja},
+and Lochbihler's formalized Java memory model
+\shortcite{lochbihler-2014-jmm} both depend on codatatypes to capture infinite
 processes.
@@
 have emerged in a few automatic theorem provers as well.
-In this paper, we describe a
+In this paper, we present a
 unified decision procedure for universal problems involving datatypes and codatatypes
@@
 Third, \emph{it must be possible to express cyclic values as closed terms and
-to enumerate them.} This is necessary when generating models.
+to enumerate them.} %This is necessary when generating models.
 The $\mu$-binder notation associates a name with
@@
 the $\mu$-term $\const{SCons}(1,\: \MU s.\; \const{SCons}(0,\: \const{SCons}(9,\: s)))$
-stands for the lasso-shaped sequence $1, 0, 9, 0, 9, 0, 9,
-\dotsc$.
+stands for the %lasso-shaped
+sequence $1, 0, 9, 0, 9, 0, 9, \dotsc$.
 
@@
 version, including detailed proofs, will appear in a special issue of the
-\emph{Journal of Automated Reasoning}.
+\emph{Journal of Automated Reasoning}. We refer to these for a description of
+related work.
 %\cite{reynolds-blanchette-201x-codata2}.
@@
 The derivation rules are partitioned into three sets of rules, given in
-Figs.\ \ref{fig:cc-rules}~to~\ref{fig:split-rule}, corresponding to three
+Figures \ref{fig:cc-rules}, \ref{fig:ab-rules}, and \ref{fig:split-rule},
+corresponding to three
 phases of the calculus. The first phase computes the bidirectional closure of $\Ec$. The second
@@
 
-\paragraph{Phase 1: Computing the Bidirectional Closure \rm(Fig.~\ref{fig:cc-rules})}
+\subsection{Phase 1: Computing the Bidirectional Closure}
+
 In conjunction with \rn{Refl}, \rn{Sym}, and \rn{Trans}, the \rn{Cong} rule computes the congruence (upward) closure,
@@
 
-\paragraph{Phase 2: Applying Acyclicity and Uniqueness \rm(Fig.~\ref{fig:ab-rules})}
+\subsection{Phase 2: Applying Acyclicity and Uniqueness}
+
 %For presentation of the rules in this phase, we rely on a representation of terms in the $\mu$-notation.
@@
 
-\paragraph{Phase 3: Branching \rm(Fig.~\ref{fig:split-rule})}
+\subsection{Phase 3: Branching}
+
 If a selector is applied to a term $t$, or if $t$'s type is a finite datatype,
@@
 
-\paragraph{Termination and Correctness}
-We now show the termination and correctness of the calculus.
+\subsection{Termination and Correctness}
+
+We now show (FIXME) the termination and correctness of the calculus.
 Correctness means the following:
@@
 %obviously depends on the application.
-Table~\ref{tab:benchtwofiftysix} indicates that
+Table~\ref{tab:bench} indicates that
 the theoretically stronger instances of the decision
@@
 
-We introduced a decision procedure for the universal theory of datatypes and
+We presented a decision procedure for the universal theory of datatypes and
 codatatypes. Our approach relies on $\mu$-terms to represent cyclic values.
@@
 
+\raggedright
+
 \bibliographystyle{named}

### 2016-02-09 sister/sister.tex
more

@@
 \begin{abstract}
-Datatypes and codatatypes are useful to represent infinite and potentially
+Datatypes and codatatypes are useful to represent finite and potentially
 infinite data structures and processes. We describe a decision procedure
-to reason about these types. The procedure has been
+to reason about such types. The procedure has been
 integrated in CVC4, a modern SMT (satisfiability modulo theories)
@@
 prove theorems in first-order logic. These solvers are also useful for model
-finding, or constraint solving: given some constraints, they can often find a
-variable assignment that satisfy them.
+finding, or constraint solving: given a set of constraints, they can be used
+to find a variable assignment that satisfy them.
 
@@
 science but also arise in formalized mathematics.
-To represent infinite objects,
+And to represent infinite objects,
 a natural choice is to turn to codatatypes,
-the non-well-founded dual of datatypes.
+their non-well-founded dual.
 %
@@
 and Lochbihler's formalized Java memory model
-\shortcite{lochbihler-2014-jmm} both depend on codatatypes to capture infinite
+\shortcite{lochbihler-2014-jmm} both depend on codatatypes to express infinite
 processes.
@@
 constructor
-terms, and only those. As a simple example, the coalgebraic specification
+terms, and only those. As a simple example, the codatatype specification
 %
@@
 to enumerate them.} %This is necessary when generating models.
-The $\mu$-binder notation associates a name with
+The \hbox{$\mu$-binder} notation associates a name with
 a subterm, making it possible to represent cyclic values.
@@
 An evaluation on %hand-crafted examples and on
-problems generated from Isabelle\slash HOL
+problems generated from Isabelle
 formalizations demonstrates its usefulness (Section~\ref{sec:evaluation}).
@@
 %(Section~\ref{sec:co-datatypes})
-and are therefore equal. Both
+and are therefore equal.
+(FIXME)
+
+Some codatatypes are so degenerate as to be finite even though they
+have infinite values. A simple example is
+\vthinspace$\keyw{codata}~\ty{a} = \const{A}(\ty{a})$, which is corecursive
+and yet has a cardinality of one;
+its unique value is $\MU a.\; \const{A}(a)$.
+
+Both
 \rn{Split}'s finiteness assumption %on the datatype

### 2016-02-09 sister/sister.tex
Minor

@@
 Proving a universal quantifier-free conjecture is reduced to showing that
-its negation is unsatisfiable. The presentation is
-inspired by Barrett et al.\ \shortcite{barrett-et-al-2007} but at
-a higher level, using unoriented equations instead of oriented ones.
+its negation is unsatisfiable. 
+%The presentation is
+%inspired by Barrett et al.\ \shortcite{barrett-et-al-2007} but at
+%a higher level, using unoriented equations instead of oriented ones.
 
@@
 it is enough %for our purposes
-to select an arbitrary constructor term $\const{C}( \vec t\, ) \in \ec{t}$ among these.
+to select an arbitrary constructor term from $\ec{t}$ among these.
 
@@
 }{
-  \Ec := \Ec,\: \DISC{1} %\ROR
-  \ROR \cdots \ROR \Ec := \Ec,\: \DISC{m}
+  \parallel_{i=1 \ldots n} \Ec := \Ec,\: \DISC{i} %\ROR
+ % \ROR \cdots \ROR \Ec := \Ec,\: \DISC{m}
 }

### 2016-02-09 sister/sister.tex
Address remaining issues in sec 3.

@@
 The rules in this phase are described in terms of a mapping $\Val$
-that assigns to each equivalence class a $\mu$-term as its representative.
+that assigns to each equivalence class a $\mu$-term. % as its representative.
 %\rem{Not explicitly constructed, add here?}
@@
 
-The $\mu$-term $\Val[t]$ describes a class of values
-that $t$ and other members of $t$'s equivalence class can take in models of $\Ec$.
-When $\tau$ is a datatype, % \in \Data$,
-a cyclic $\mu$-term describes an infeasible class of values.
-
-The mapping $\Val$ is defined as follows.
+The mapping $\Val$ is constructed as follows.
 %
@@
 
+%AJR: moved this paragraph after the construction to avoid mistaking this paragraph as part of the formal definition.
+The $\mu$-term $\Val[t]$ describes a class of values
+that $t$ and other members of $t$'s equivalence class can take in models of $\Ec$.
+When $\tau$ is a datatype, % \in \Data$,
+a cyclic $\mu$-term describes an infeasible class of values.
+
 \begin{figure}[t!]
-\normalsize
+\small
 \centering
@@
 \rn{Refl}
-\qquad
+\quad
 \(
@@
 \rn{Sym}
-\qquad
+\quad
 \(
@@
 \rn{Cong}
-\qquad
+\quad
 \(
@@
 \rn{Inject}
-\qquad
+\quad
 \(
@@
 \vspace*{+6pt} %% TYPESETTING HACK
-\normalsize
+\small
 \centering
@@
 %for a capture-avoiding renaming of $\mu$-bound variables,
-then $t \teq u$. Comparison for $\alpha$-equivalence may seem too restrictive, since
+then $t$ is equal to $u$. Comparison for $\alpha$-equivalence may seem too restrictive, since
 $\MU x.\; \const{S}(x)$ and $\MU y.\; \const{S}(\const{S}(y))$ specify the same
@@
 \begin{align*}
-\Val \ec{x} & \,=\, \MU {\Varec{x}}.\; \const{S}(\Varec{x}) &
-\kern0.25em %% TYPESETTING
-\Val \ec{y} & \,=\,
+\Val \ec{x} & \,=\, \MU {\Varec{x}}.\; \const{S}(\Varec{x}) 
+\kern2.0em %% TYPESETTING
+\Val \ec{y}  \,=\,
 \MU {\Varec{y}}.\; \const{S}(\smash{\MU {\VAREC{\const{S}(y)}}}.\; \const{S}( {\Varec{y}} ) )
-\kern0.25em %% TYPESETTING
- &
-\Val \ec{\const{S}(y)} & =
+%\kern0.25em %% TYPESETTING
+% &
+\\
+\Val \ec{\const{S}(y)} & \,=\, 
 \MU {\VAREC{\const{S}(y)}}.\; \const{S}(\MU {\Varec{y}}.\; \const{S}( {\VAREC{\const{S}(y)}} ) )
@@
 terms %from $\tEc$
-have the same corecursive singleton type
-%(Section~\ref{sec:co-datatypes})
+are of a \emph{singleton} type (one for which there exists only one value),
 and are therefore equal.
-(FIXME)
-
-Some codatatypes are so degenerate as to be finite even though they
-have infinite values. A simple example is
+Notice that corecursive singleton types may have infinite values. A simple example is
 \vthinspace$\keyw{codata}~\ty{a} = \const{A}(\ty{a})$, which is corecursive
@@
 
-We now show (FIXME) the termination and correctness of the calculus.
-Correctness means the following:
-If there exists a closed derivation tree with root node
-$\Ec$, then $\Ec$ is %$\thD$-
-unsatisfiable; and if there exists
-a derivation tree with root node $\Ec$ that contains a saturated node, then
-$\Ec$ is %$\thD$-
-satisfiable.
+%We now show (FIXME) the termination and correctness of the calculus.
+%Correctness means the following:
+%If there exists a closed derivation tree with root node
+%$\Ec$, then $\Ec$ is %$\thD$-
+%unsatisfiable; and if there exists
+%a derivation tree with root node $\Ec$ that contains a saturated node, then
+%$\Ec$ is %$\thD$-
+%satisfiable.
+We state the following properties of derivation trees generated by the calculus.
 
@@
 The proof of solution soundness is constructive in that it provides a
-method for constructing a model for a saturated configuration, by means of the
-mapping $\ValC$.
+method for constructing a model for a saturated configuration.
+%AJR : if necessary...
+%To give an intuition how this model is constructed,
+%we consider the mapping $\Val$ from Step 2 of the procedure,
+%and apply additional substitutions over the free variables in its range
+%to obtain a mapping, call it $\ValC$, 
+%from equivalence classes to closed $\MU$-terms.
+%Then, for each equivalence class $\ec{t}$ in the domain of $\ValC$,
+%we interpret each variable in $\ec{t}$ as $\ValC \ec{t}$.
+%The properties of a saturated configuration ensure
+%that $\ValC$ can be constructed such that all constraints in our input are satisfied.
 

### 2016-02-09 sister/sister.tex
tuning

@@
 can be viewed as a %n enumeration
-datatype with $n$~nullary constructors
-\cite{barrett-et-al-2007}.
+datatype with $n$~nullary constructors.
+%\cite{barrett-et-al-2007}.
 
@@
 }{
-  \parallel_{i=1 \ldots n} \Ec := \Ec,\: \DISC{i} %\ROR
+  {\mathop{\big|\kern-.25ex\big|}\vthinspace}_{i=1}^n\, \Ec := \Ec,\: \DISC{i} %\ROR
  % \ROR \cdots \ROR \Ec := \Ec,\: \DISC{m}
@@
 \begin{enumerate}
-%\item \noindent\rlap{Initially:}\phantom{Unfold $\Varec{x}$:\enskip}$\{  \}$
-\item \noindent\rlap{Unfold $\Varec{x}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\Val = \{\vthinspace  \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) \vthinspace\}$
-\item \noindent\rlap{Unfold $\vthinspace\Varec{y}\negvthinspace$}\phantom{Unfold $\Varec{x}$}:\enskip$\Val = \{\vthinspace  \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\;
-  \ec{y} \mapsto \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) \vthinspace\}$
-\item \noindent\rlap{Unfold $\Varec{x}$:}\phantom{Unfold $\Varec{x}$:\enskip}$\Val = \{\vthinspace  \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\;
-  \ec{y} \mapsto \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\; \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) ) \vthinspace\}$
+\item Unfold $\Varec{x}$:\enskip $\Val = \{\vthinspace  \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) \vthinspace\}$
+\item Unfold $\Varec{y}$:\enskip $\Val = \{\vthinspace  \!\begin{aligned}[t]
+  & \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\; \\[-.5\jot]
+  & \ec{y} \mapsto \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) \vthinspace\}\end{aligned}$
+\item Unfold $\Varec{x}$:\enskip $\Val = \{\vthinspace \!\begin{aligned}[t]
+  & \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\; \\[-.5\jot]
+  & \ec{y} \mapsto \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\; \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) ) \vthinspace\}\end{aligned}$
 \end{enumerate}

### 2016-02-09 sister/sister.tex
tune

@@
 
-\newcommand\captioni{\,Number of solved goals with 16~lemmas per goal}
+\newcommand\captioni{\,Number of solved goals with $n$~lemmas per goal}
 \newcommand\captionii{\,Number of solved goals with 256~lemmas per goal}
@@
 \begin{center}\begin{tabular}{@{\;}l@{\kern1.5em}c@{\kern.5em}c@{\kern1em}c@{\kern.5em}c@{\;}}
-  & \multicolumn{2}{@{}c@{\kern1em}}{16 lemmas\;} & \multicolumn{2}{@{}c@{\;}}{256 lemmas} \\%[-0.5pt]
+  & \multicolumn{2}{@{}c@{\kern1em}}{$n = 16$\,} & \multicolumn{2}{@{}c@{\;}}{$n = 256$\;} \\%[-0.5pt]
   & \HD{CVC4} & \HD{Z3} & \HD{CVC4} & \HD{\,Z3}

### 2016-02-09 jour/jour.tex
more

@@
 \text{Induction:}\quad
-\AXC{\strut$\bigwedge_{\vthinspace i,j}\, \forall x_1 \ldots \vthinspace x_{n_{i\negvthinspace j}}.\; \bigl(\bigwedge_{\vthinspace k}\, \mathcalx{IH}[x_k]\bigr) \medrightarrow P_i[\const C_{i\negvthinspace j}(x_1,\ldots,x_{n_{ij}})]$}
-\UIC{\strut$\bigwedge_{\vthinspace i}\, P_i[v_i]$}
+\AXC{\strut$\bigwedge_{\vthinspace i=1}^\ell \bigwedge_{\vthinspace j=1}^{m_i}\, \forall x_1 \ldots \vthinspace x_{n_{i\negvthinspace j}}.\; \bigl(\bigwedge_{\vthinspace k=1}^{ n_{i\negvthinspace j}}\, \mathcalx{IH}[x_k]\bigr) \medrightarrow P_i[\const C_{i\negvthinspace j}(x_1,\ldots,x_{n_{ij}})]$}
+\UIC{\strut$\bigwedge_{\vthinspace i=1}^{\ell}\, P_i[v_i]$}
 \DP
@@
 \text{Coinduction:}\quad
-\AXC{\strut$\begin{gathered}\textstyle \smash{\bigwedge_{\vthinspace i}\, R_i[v_i, w_i]} \\[-\jot]\textstyle
-\bigwedge_{\vthinspace i}\, \forall v\; w.\;\, R_i[v, w] \medrightarrow
-  \bigwedge_{\vthinspace j}\, \const{d}_{\negvthinspace j}(v) \mathrel\land \const{d}_{\negvthinspace j}(w)
-  \medrightarrow \bigwedge_{\vthinspace k}\, \const{s}_{i\negvthinspace j}^{\,k}(v) \sim \const{s}_{i\negvthinspace j}^{\,k}(w)
+\AXC{\strut$\begin{gathered}\textstyle \smash{\bigwedge_{\vthinspace i=1}^\ell \, R_i[v_i, w_i]} \\[-\jot]\textstyle
+\bigwedge_{\vthinspace i=1}^\ell\, \bigl(\forall v\; w.\;\, R_i[v, w] \medrightarrow
+  \bigwedge_{\vthinspace j=1}^{m_i}\, \bigl(\const{d}_{\negvthinspace j}(v) \mathrel\land \const{d}_{\negvthinspace j}(w)
+  \medrightarrow \bigwedge_{\vthinspace k=1}^{n_{i\negvthinspace j}}\, \const{s}_{i\negvthinspace j}^{\,k}(v) \sim \const{s}_{i\negvthinspace j}^{\,k}(w)\bigr)\bigr)
 \end{gathered}$}
-\UIC{\strut$\bigwedge_{\vthinspace i}\, v_i \teq w_i$}
+\UIC{\strut$\bigwedge_{\vthinspace i=1}^\ell \, v_i \teq w_i$}
 \DP
@@
 but
-$\MU x.\; \const{C}( x ) \not\aequiv \MU x.\; \const{D}( y, x ) \not\aequiv \MU x.\; \const{D}( z, x )
-\not\aequiv \MU y.\; \const{D}( y, x )$.
+$\MU x.\; \const{C}( x )$, $\MU x.\; \const{D}( y, x )$, $\MU x.\; \const{D}( z, x )$, and
+$\MU y.\; \const{D}( y, x )$ are all $\alpha$-disequivalent.
 Two $\mu$-terms can denote the same value despite being $\alpha$-disequivalent---e.g.,
@@
 the cardinality of an uninterpreted type $\ty{u}$.
-By assuming that uninterpreted types have infinite cardinality, the calculus
+Since it assumes that uninterpreted types have infinite cardinality, the calculus
 presented in Section~\ref{sec:the-decision-procedure} is \relax{incomplete}
@@
 uninterpreted types.
-Second, we can conditionally apply \rn{Single} to codatatype terms that may have cardinality one.
+Second, we can conditionally apply \rn{Single} to codatatype terms whose type
+potentially has cardinality one.
 For example, the codatatype $\ty{stream}_{\vthinspace\ty{u}}$ %, where $\ty{u}$ is uninterpreted.
@@
 soundness, without which the generated models would often be
-spurious. For example, given the axioms
+spurious. For example, given the constraints
 \begin{align*}
-\const{zeros} & = \const{SCons}(0,\:\const{zeros})
-& \const{sconst}~n & = \const{SCons}(n,\:\const{sconst}~n)
+\const{zeros} & \teq \const{SCons}(0,\:\const{zeros})
+& \const{sconst}~n & \teq \const{SCons}(n,\:\const{sconst}~n)
 \end{align*}
 on streams, the conjecture
-$\const{zeros} = \const{sconst}~0$
+$\const{zeros} \teq \const{sconst}~0$
 would be ``refuted'' by a spurious countermodel that interprets

### 2016-02-09 sister/sister.tex
avoid "tableau" terminology at IJCAI
more
more typesetting

@@
 \usepackage[all]{xy}
+
+\newcommand\yasx{\mskip.5mu y\mskip.5mu} %%% TYPESETTING: hack so that y is exactly as wide as x
 
@@
 whose variables are interpreted existentially.
-The decision procedure is formulated as a tableau-like calculus.
+The decision procedure is formulated as an %tableau-like
+abstract calculus.
 %Following the tradition of refutational theorem proving,
@@
 \subsection{Phase 1: Computing the Bidirectional Closure}
+
+\begin{figure}[b!]
+%\small
+\centering
+\begin{tabular}{@{}c@{}}
+\(
+\inferrule{
+  t \in \tEc
+}{
+  \Ec := \Ec,\: t \teq t
+}
+\)
+\rn{Refl}
+\quad
+\(
+\inferrule{
+ t \teq u \in \Ec
+}{
+ \Ec := \Ec,\: u \teq t
+}
+\)
+\rn{Sym}
+\\[5\jot]
+\(
+\inferrule{
+  s \teq t,\; t \teq u \in \Ec
+}{
+  \Ec := \Ec,\: s \teq u
+}
+\)
+\rn{Trans}
+\quad
+\(
+\inferrule{
+  \vec t \teq \vec u \in \Ec
+  \betweenantes
+  \const f( \vec t \,),\, \const f( \vec u ) \in \tEc
+}{
+  \Ec := \Ec,\: \const f( \vec t \,) \teq \const f( \vec u )
+}
+\)
+\rn{Cong}\\[5\jot]
+\(
+\inferrule{
+  t \teq u,\; t \tneq u \in \Ec
+}{
+  \bot
+}
+\)
+\rn{Conflict}
+\quad
+\(
+\inferrule{
+  \const{C}( \vec t \,) \teq \const{C}( \vec u ) \in \Ec
+}{
+  \Ec := \Ec,\: \vec t \teq \vec u
+}
+\)
+\rn{Inject}
+\\[5\jot]
+\(
+\inferrule{
+  \const{C}( \vec t \,) \teq \const{D}( \vec u ) \in \Ec
+  \betweenantes
+  \const{C} \not= \const{D}
+}{
+  \bot
+}
+\)
+\rn{Clash}
+\end{tabular}
+\vspace*{-2pt} %% TYPESETTING HACK
+\caption{\,Derivation rules for bidirectional closure%.
+}
+\label{fig:cc-rules}
+\end{figure}
 
@@
 %\rem{Not explicitly constructed, add here?}
+
+\begin{figure}[b!]
+%\vspace*{+6pt} %% TYPESETTING HACK
+%\small
+\centering
+\begin{tabular}{@{}c@{}}
+\(
+\inferrule{
+  t : \delta \in \Data
+  \betweenantes
+  \Val \smash{\ec{t}} = \MU x.\; u
+  \betweenantes
+  x \in \FV( u )
+}{
+  \bot
+}
+\)
+\rn{Acyclic}
+\\[5\jot]
+\(
+\inferrule{
+ t, u : \delta \in \Codata
+ \betweenantes
+ \Val \smash{\ec{t}} \aequiv \Val \smash{\ec{u}}
+}{
+ \Ec := \Ec,\: t \teq u
+}
+\)
+\rn{Unique}
+\end{tabular}
+\vspace*{-2pt} %% TYPESETTING HACK
+\caption{\,Derivation rules for acyclicity and uniqueness%.
+}
+\label{fig:ab-rules}
+\end{figure}
 
@@
 
+\begin{examplex}
+Suppose that $\Ec$ contains four distinct equivalence classes $\ec{w}$, $\ec{x}$, $\ec{\yasx}$, and~$\ec{z}$
+such that $\const{C}(w,y) \in \ec{x}$ and $\const{C}(z,x) \in \ec{\yasx}$ for some $\const{C} \in \Ctr$.
+A possible sequence of unfolding steps is given below, omitting
+trivial entries such as $\ec{w} \mapsto {\Varec{w}}$.
+%
+\begin{enumerate}
+\item Unfold $\Varec{x}$:\enskip $\Val = \{\vthinspace  \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{\yasx}} ) \vthinspace\}$
+\item Unfold $\Varec{\yasx}$:\enskip $\Val = \{\vthinspace  \!\begin{aligned}[t]
+  & \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \MU {\Varec{\yasx}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\; \\[-.5\jot]
+  & \ec{\yasx} \mapsto \MU {\Varec{\yasx}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) \vthinspace\}\end{aligned}$
+\item Unfold $\Varec{x}$:\enskip $\Val = \{\vthinspace \!\begin{aligned}[t]
+  & \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \MU {\Varec{\yasx}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\; \\[-.5\jot]
+  & \ec{\yasx} \mapsto \MU {\Varec{\yasx}}.\; \const{C}( {\Varec{z}},\; \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{\yasx}} ) ) \vthinspace\}\end{aligned}$
+\end{enumerate}
+%
+The resulting $\Val$ indicates that the values for $x$ and $y$ in models of $\Ec$
+must be of the forms $\const{C}( {\Varec{w}}, \const{C}( {\Varec{z}},\allowbreak \const{C}( {\Varec{w}},\allowbreak \const{C}( {\Varec{z}},\ldots ))))$
+and $\const{C}( {\Varec{z}}, \const{C}( {\Varec{w}}, \const{C}( {\Varec{z}}, \const{C}( {\Varec{w}}, \ldots ))))$,
+respectively. %, for some values of ${\Varec{w}}$ and ${\Varec{z}}$.
+\xend
+\end{examplex}
+
+Given the mapping $\Val$, the \rn{Acyclic} and \rn{Unique} rules work as follows.
+For acyclicity, if $\ec{t}$ is a datatype equivalence class
+whose values $\Val \ec{t} = \MU x.\; u$ are cyclic
+(expressed by $x \in \FV( u )$),
+then $\Ec$ is %$\thD$-
+unsatisfiable.
+For uniqueness, if $\ec{t}$, $\ec{u}$ are two codatatype equivalence classes
+whose values $\Val \ec{t}$, $\Val \ec{u}$ are $\alpha$-equivalent,
+%for a capture-avoiding renaming of $\mu$-bound variables,
+then $t$ is equal to $u$. Comparison for $\alpha$-equivalence may seem too restrictive, since
+$\MU x.\; \const{S}(x)$ and $\MU y.\; \const{S}(\const{S}(y))$ specify the same
+value despite being $\alpha$-disequivalent, but the rule will
+make progress by discovering that the subterm $\const{S}(y)$ of $\MU y.\;
+\const{S}(\const{S}(y))$ must be equal to the entire term,
+as demonstrated next.
+
+%% Minor terminology point:
+%% A function symbol can have a "type signature" (but we're trying to avoid that terminology);
+%% unless it's a constant, it doesn't have a (first-order) type. So I'm rephrasing.
+
+\begin{examplex}
+Let $\Ec = \{ x \teq \const{S}(x),\; y \teq \const{S}(\const{S}(y)) \}$.
+After phase~1, the equivalence classes %of $\Ec$
+are
+$\{ x,\, \const{S}(x) \}$,
+$\{ y,\, \const{S}(\const{S}(y)) \}$,
+and $\{ \const{S}(y) \}$.
+Constructing $\Val$ yields
+\begin{align*}
+\Val \ec{x} & \,=\, \MU {\Varec{x}}.\; \const{S}(\Varec{x}) 
+\kern2.0em %% TYPESETTING
+\Val \ec{\yasx}  \,=\,
+\MU {\Varec{\yasx}}.\; \const{S}(\smash{\MU {\VAREC{\const{S}(y)}}}.\; \const{S}( {\Varec{\yasx}} ) )
+%\kern0.25em %% TYPESETTING
+% &
+\\
+\Val \ec{\const{S}(y)} & \,=\, 
+\MU {\VAREC{\const{S}(y)}}.\; \const{S}(\MU {\Varec{\yasx}}.\; \const{S}( {\VAREC{\const{S}(y)}} ) )
+\end{align*}
+Since $\Val \ec{\yasx} \aequiv \Val \ec{\const{S}(y)}$,
+%for the renaming
+%$\{ {\Varec{\yasx}} \mapsto {\VAREC{\const{S}(y)}}, {\VAREC{\const{S}(y)}} \mapsto {\Varec{\yasx}} \}$,
+the \rn{Unique} rule applies to derive $y \teq \const{S}(y)$.
+%Intuitively, this equality holds, since the values of $y$ and $\const{S}(y)$
+%are equivalent.
+%
+At this point, phase~1 is activated again, % to XXX the equivalence classes,
+yielding the equivalence classes
+$\{ x,\, \const{S}(x) \}$ and
+$\{ y,\, \const{S}(y),\allowbreak\, \const{S}(\const{S}(y)) \}$.
+The mapping $\Val$ is updated accordingly:
+\begin{align*}
+\Val \ec{x} & \,=\, \MU {\Varec{x}}.\; \const{S}(\Varec{x})
+&
+\Val \ec{\yasx} & \,=\, \MU {\Varec{\yasx}}.\; \const{S}({\Varec{\yasx}})
+\end{align*}
+Since $\Val \ec{x} \aequiv \Val \ec{\yasx}$,
+%for the renaming
+%$\{ {\Varec{\yasx}} \mapsto {\VAREC{\const{S}(y)}}, {\VAREC{\const{S}(y)}} \mapsto {\Varec{\yasx}} \}$,
+\rn{Unique} can finally be applied to derive $x \teq y$.
+\xend
+\end{examplex}
+
+\subsection{Phase 3: Branching}
+
+If a selector is applied to a term $t$, or if $t$'s type is a finite datatype,
+$t$'s equivalence class must contain a
+$\delta$ constructor term.
+This is enforced in the third phase by the \rn{Split} rule.
+Another rule, \rn{Single}, focuses on the degenerate case where two
+terms %from $\tEc$
+are of a \emph{singleton} type (one for which there exists only one value),
+and are therefore equal.
+Notice that corecursive singleton types may have infinite values. A simple example is
+\vthinspace$\keyw{codata}~\ty{a} = \const{A}(\ty{a})$, which is corecursive
+and yet has a cardinality of one;
+its unique value is $\MU a.\; \const{A}(a)$.
+
 \begin{figure}[t!]
-\small
-\centering
-\begin{tabular}{c}
-\(
-\inferrule{
-  t \in \tEc
-}{
-  \Ec := \Ec,\: t \teq t
-}
-\)
-\rn{Refl}
-\quad
-\(
-\inferrule{
- t \teq u \in \Ec
-}{
- \Ec := \Ec,\: u \teq t
-}
-\)
-\rn{Sym}
-\quad
-\(
-\inferrule{
-  s \teq t,\; t \teq u \in \Ec
-}{
-  \Ec := \Ec,\: s \teq u
-}
-\)
-\rn{Trans}
-\\[5\jot]
-\(
-\inferrule{
-  \vec t \teq \vec u \in \Ec
-  \betweenantes
-  \const f( \vec t \,),\, \const f( \vec u ) \in \tEc
-}{
-  \Ec := \Ec,\: \const f( \vec t \,) \teq \const f( \vec u )
-}
-\)
-\rn{Cong}
-\quad
-\(
-\inferrule{
-  t \teq u,\; t \tneq u \in \Ec
-}{
-  \bot
-}
-\)
-\rn{Conflict}
-\\[5\jot]
-\(
-\inferrule{
-  \const{C}( \vec t \,) \teq \const{C}( \vec u ) \in \Ec
-}{
-  \Ec := \Ec,\: \vec t \teq \vec u
-}
-\)
-\rn{Inject}
-\quad
-\(
-\inferrule{
-  \const{C}( \vec t \,) \teq \const{D}( \vec u ) \in \Ec
-  \betweenantes
-  \const{C} \not= \const{D}
-}{
-  \bot
-}
-\)
-\rn{Clash}
-\end{tabular}
-\vspace*{-2pt} %% TYPESETTING HACK
-\caption{\,Derivation rules for bidirectional closure%.
-}
-\label{fig:cc-rules}
-\end{figure}
-
-\begin{figure}[t!]
-\vspace*{+6pt} %% TYPESETTING HACK
-\small
+%\vspace*{+6pt} %% TYPESETTING HACK
+%\small
 \centering
 \begin{tabular}{@{}c@{}}
-\(
-\inferrule{
-  t : \delta \in \Data
-  \betweenantes
-  \Val \smash{\ec{t}} = \MU x.\; u
-  \betweenantes
-  x \in \FV( u )
-}{
-  \bot
-}
-\)
-\rn{Acyclic}
-\\[5\jot]
-\(
-\inferrule{
- t, u : \delta \in \Codata
- \betweenantes
- \Val \smash{\ec{t}} \aequiv \Val \smash{\ec{u}}
-}{
- \Ec := \Ec,\: t \teq u
-}
-\)
-\rn{Unique}
-\end{tabular}
-\vspace*{-2pt} %% TYPESETTING HACK
-\caption{\,Derivation rules for acyclicity and uniqueness%.
-}
-\label{fig:ab-rules}
-\end{figure}
-
-\begin{figure}[t!]
-\vspace*{+6pt} %% TYPESETTING HACK
-\small
-\centering
-\begin{tabular}{c}
 \(
@@
 \end{figure}
-
-\begin{examplex}
-Suppose that $\Ec$ contains four distinct equivalence classes $\ec{w}$, $\ec{x}$, $\ec{y}$, and~$\ec{z}$
-such that $\const{C}(w,y) \in \ec{x}$ and $\const{C}(z,x) \in \ec{y}$ for some $\const{C} \in \Ctr$.
-A possible sequence of unfolding steps is given below, omitting
-trivial entries such as $\ec{w} \mapsto {\Varec{w}}$.
-%
-\begin{enumerate}
-\item Unfold $\Varec{x}$:\enskip $\Val = \{\vthinspace  \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) \vthinspace\}$
-\item Unfold $\Varec{y}$:\enskip $\Val = \{\vthinspace  \!\begin{aligned}[t]
-  & \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\; \\[-.5\jot]
-  & \ec{y} \mapsto \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) \vthinspace\}\end{aligned}$
-\item Unfold $\Varec{x}$:\enskip $\Val = \{\vthinspace \!\begin{aligned}[t]
-  & \ec{x} \mapsto \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\; \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\, {\Varec{x}} ) ),\; \\[-.5\jot]
-  & \ec{y} \mapsto \MU {\Varec{y}}.\; \const{C}( {\Varec{z}},\; \MU {\Varec{x}}.\; \const{C}( {\Varec{w}},\, {\Varec{y}} ) ) \vthinspace\}\end{aligned}$
-\end{enumerate}
-%
-The resulting $\Val$ indicates that the values for $x$ and $y$ in models of $\Ec$
-must be of the forms $\const{C}( {\Varec{w}}, \const{C}( {\Varec{z}},\allowbreak \const{C}( {\Varec{w}},\allowbreak \const{C}( {\Varec{z}},\ldots ))))$
-and $\const{C}( {\Varec{z}}, \const{C}( {\Varec{w}}, \const{C}( {\Varec{z}}, \const{C}( {\Varec{w}}, \ldots ))))$,
-respectively. %, for some values of ${\Varec{w}}$ and ${\Varec{z}}$.
-\xend
-\end{examplex}
-
-Given the mapping $\Val$, the \rn{Acyclic} and \rn{Unique} rules work as follows.
-For acyclicity, if $\ec{t}$ is a datatype equivalence class
-whose values $\Val \ec{t} = \MU x.\; u$ are cyclic
-(expressed by $x \in \FV( u )$),
-then $\Ec$ is %$\thD$-
-unsatisfiable.
-For uniqueness, if $\ec{t}$, $\ec{u}$ are two codatatype equivalence classes
-whose values $\Val \ec{t}$, $\Val \ec{u}$ are $\alpha$-equivalent,
-%for a capture-avoiding renaming of $\mu$-bound variables,
-then $t$ is equal to $u$. Comparison for $\alpha$-equivalence may seem too restrictive, since
-$\MU x.\; \const{S}(x)$ and $\MU y.\; \const{S}(\const{S}(y))$ specify the same
-value despite being $\alpha$-disequivalent, but the rule will
-make progress by discovering that the subterm $\const{S}(y)$ of $\MU y.\;
-\const{S}(\const{S}(y))$ must be equal to the entire term,
-as demonstrated next.
-
-%% Minor terminology point:
-%% A function symbol can have a "type signature" (but we're trying to avoid that terminology);
-%% unless it's a constant, it doesn't have a (first-order) type. So I'm rephrasing.
-
-\begin{examplex}
-Let $\Ec = \{ x \teq \const{S}(x),\; y \teq \const{S}(\const{S}(y)) \}$.
-After phase~1, the equivalence classes %of $\Ec$
-are
-$\{ x,\, \const{S}(x) \}$,
-$\{ y,\, \const{S}(\const{S}(y)) \}$,
-and $\{ \const{S}(y) \}$.
-Constructing $\Val$ yields
-\begin{align*}
-\Val \ec{x} & \,=\, \MU {\Varec{x}}.\; \const{S}(\Varec{x}) 
-\kern2.0em %% TYPESETTING
-\Val \ec{y}  \,=\,
-\MU {\Varec{y}}.\; \const{S}(\smash{\MU {\VAREC{\const{S}(y)}}}.\; \const{S}( {\Varec{y}} ) )
-%\kern0.25em %% TYPESETTING
-% &
-\\
-\Val \ec{\const{S}(y)} & \,=\, 
-\MU {\VAREC{\const{S}(y)}}.\; \const{S}(\MU {\Varec{y}}.\; \const{S}( {\VAREC{\const{S}(y)}} ) )
-\end{align*}
-Since $\Val \ec{y} \aequiv \Val \ec{\const{S}(y)}$,
-%for the renaming
-%$\{ {\Varec{y}} \mapsto {\VAREC{\const{S}(y)}}, {\VAREC{\const{S}(y)}} \mapsto {\Varec{y}} \}$,
-the \rn{Unique} rule applies to derive $y \teq \const{S}(y)$.
-%Intuitively, this equality holds, since the values of $y$ and $\const{S}(y)$
-%are equivalent.
-%
-At this point, phase~1 is activated again, % to XXX the equivalence classes,
-yielding the equivalence classes
-$\{ x,\, \const{S}(x) \}$ and
-$\{ y,\, \const{S}(y),\allowbreak\, \const{S}(\const{S}(y)) \}$.
-The mapping $\Val$ is updated accordingly:
-\begin{align*}
-\Val \ec{x} & \,=\, \MU {\Varec{x}}.\; \const{S}(\Varec{x})
-&
-\Val \ec{y} & \,=\, \MU {\Varec{y}}.\; \const{S}({\Varec{y}})
-\end{align*}
-Since $\Val \ec{x} \aequiv \Val \ec{y}$,
-%for the renaming
-%$\{ {\Varec{y}} \mapsto {\VAREC{\const{S}(y)}}, {\VAREC{\const{S}(y)}} \mapsto {\Varec{y}} \}$,
-\rn{Unique} can finally be applied to derive $x \teq y$.
-\xend
-\end{examplex}
-
-\subsection{Phase 3: Branching}
-
-If a selector is applied to a term $t$, or if $t$'s type is a finite datatype,
-$t$'s equivalence class must contain a
-$\delta$ constructor term.
-This is enforced in the third phase by the \rn{Split} rule.
-Another rule, \rn{Single}, focuses on the degenerate case where two
-terms %from $\tEc$
-are of a \emph{singleton} type (one for which there exists only one value),
-and are therefore equal.
-Notice that corecursive singleton types may have infinite values. A simple example is
-\vthinspace$\keyw{codata}~\ty{a} = \const{A}(\ty{a})$, which is corecursive
-and yet has a cardinality of one;
-its unique value is $\MU a.\; \const{A}(a)$.
 

### 2016-02-09 sister/sister.tex
more tuning

@@
 
-\newtheorem{lemma}{Lemma}
-\newtheorem{theorem}{Theorem}
+\theoremstyle{plain}
+\newtheorem*{theoremi}{Theorem 1 (Termination)}
+\newtheorem*{theoremii}{Theorem 2 (Refutation Soundness)}
+\newtheorem*{theoremiii}{Theorem 3 (Solution Soundness)}
 
@@
 % &
-\\
+\\[-.5\jot]
 \Val \ec{\const{S}(y)} & \,=\, 
@@
 %$\{ {\Varec{\yasx}} \mapsto {\VAREC{\const{S}(y)}}, {\VAREC{\const{S}(y)}} \mapsto {\Varec{\yasx}} \}$,
-\rn{Unique} can finally be applied to derive $x \teq y$.
+\rn{Unique} can finally derive $x \teq y$.
 \xend
@@
 its unique value is $\MU a.\; \const{A}(a)$.
+Both
+\rn{Split}'s finiteness assumption %on the datatype
+and \rn{Single}'s singleton
+constraint %on the codatatype
+can be evaluated statically
+based on a recursive computation of the cardinalities of the
+constructors' argument types.
+%This step relies on the assumption that uninterpreted types are infinite.
 
@@
 
-Both
-\rn{Split}'s finiteness assumption %on the datatype
-and \rn{Single}'s singleton
-constraint %on the codatatype
-can be evaluated statically
-based on a recursive computation of the cardinalities of the
-constructors' argument types.
-%This step relies on the assumption that uninterpreted types are infinite.
-
 \subsection{Termination and Correctness}
@@
 
-\begin{theorem}[Termination]%
+\begin{theoremi}
 \label{thm:t}%
@@
 All derivation trees are finite.
-\end{theorem}
-
-\begin{theorem}[Refutation Soundness]%
+\end{theoremi}
+
+\begin{theoremii}
 \label{thm:rs}%
@@
 unsatisfiable.
-\end{theorem}
-
-\begin{theorem}[Solution Soundness]%
+\end{theoremii}
+
+\begin{theoremiii}
 \label{thm:ss}%
@@
 satisfiable.
-\end{theorem}
-
-By Theorems~\ref{thm:t},~\ref{thm:rs}, and~\ref{thm:ss}, the
+\end{theoremiii}
+
+\looseness=-1
+By Theorems~1, 2, and 3, the
 calculus is sound and complete for the universal theory of (co)datatypes. We may
@@
 method for constructing a model for a saturated configuration.
+
 %AJR : if necessary...
@@
 \newcommand\captioni{\,Number of solved goals with $n$~lemmas per goal}
-\newcommand\captionii{\,Number of solved goals with 256~lemmas per goal}
 
@@
 
-\begin{table}[t!]
+\begin{table}[h!]
 \normalsize

### 2016-02-09 sister/sister.tex
fixed

@@
 \begin{align*}
-\const{zeros} & = \const{SCons}(0,\:\const{zeros})
-& \const{sconst}~n & = \const{SCons}(n,\:\const{sconst}~n)
+\const{zeros} & \teq \const{SCons}(0,\:\const{zeros})
+& \const{repeat}(n) & \teq \const{SCons}(n,\:\const{repeat}(n))
 \end{align*}
 the conjecture
-$\const{zeros} = \const{sconst}~0$
+$\const{zeros} = \const{repeat}(0)$
 would be ``refuted'' by a spurious countermodel that interprets
-\const{zero} and $\const{sconst}~0$ by two distinct
+\const{zero} and $\const{repeat}(0)$ by two distinct
 values $\MU s.\; \const{SCons}(0, s)$,

### 2016-02-09 jour/jour.tex
more
more fixes

@@
 \const{zeros} & \teq \const{SCons}(0,\:\const{zeros})
-& \const{sconst}~n & \teq \const{SCons}(n,\:\const{sconst}~n)
+& \const{repeat}(n) & \teq \const{SCons}(n,\:\const{repeat}(n))
 \end{align*}
 on streams, the conjecture
-$\const{zeros} \teq \const{sconst}~0$
+$\const{zeros} \teq \const{repeat}(0)$
 would be ``refuted'' by a spurious countermodel that interprets
-\const{zero} and $\const{sconst}~0$ by two distinct
+\const{zero} and $\const{repeat}(0)$ by two distinct
 values that both correspond to the term

### 2016-02-09 sister/sister.tex
tuned typesetting

@@
 Berlin, Germany \cite{reynolds-blanchette-2015-codata}. An more comprehensive
-version, including detailed proofs, will appear in a special issue of the
+version %, including detailed proofs,
+will appear in a special issue of the
 \emph{Journal of Automated Reasoning}. We refer to these for a description of
@@
 For the natural numbers constructed from $\const{Z}$ and $\const{S}$,
-induction prohibits models that contain cyclic values---e.g.,
-an $n$ such that $n \teq \const{S}(n)$---or even infinite acyclic values.
+induction prohibits models that contain infinite values.
 % $\const{S}(\const{S}(\ldots))$.
@@
 
-The derivation rules are partitioned into three sets of rules, given in
-Figures \ref{fig:cc-rules}, \ref{fig:ab-rules}, and \ref{fig:split-rule},
-corresponding to three
-phases of the calculus. The first phase computes the bidirectional closure of $\Ec$. The second
+The derivation rules are partitioned into three phases, given in
+Figures \ref{fig:cc-rules}, \ref{fig:ab-rules}, and \ref{fig:split-rule}.
+The first phase computes the bidirectional closure of $\Ec$. The second
 phase makes inferences based on acyclicity (for datatypes) and uniqueness
@@
 and failures to unify equated terms are recognized by \rn{Clash}.
-The \rn{Conflict} rule recognizes when an equality and its negation both occur in $\Ec$, in which case $\Ec$ has no model.
+\rn{Conflict} recognizes when an equality and its negation both occur in $\Ec$, in which case $\Ec$ has no model.
 
@@
 
-If a selector is applied to a term $t$, or if $t$'s type is a finite datatype,
-$t$'s equivalence class must contain a
-$\delta$ constructor term.
-This is enforced in the third phase by the \rn{Split} rule.
-Another rule, \rn{Single}, focuses on the degenerate case where two
-terms %from $\tEc$
-are of a \emph{singleton} type (one for which there exists only one value),
-and are therefore equal.
-Notice that corecursive singleton types may have infinite values. A simple example is
-\vthinspace$\keyw{codata}~\ty{a} = \const{A}(\ty{a})$, which is corecursive
-and yet has a cardinality of one;
-its unique value is $\MU a.\; \const{A}(a)$.
-Both
-\rn{Split}'s finiteness assumption %on the datatype
-and \rn{Single}'s singleton
-constraint %on the codatatype
-can be evaluated statically
-based on a recursive computation of the cardinalities of the
-constructors' argument types.
-%This step relies on the assumption that uninterpreted types are infinite.
-
-\begin{figure}[t!]
+\begin{figure}[b!]
 %\vspace*{+6pt} %% TYPESETTING HACK
@@
 \end{figure}
+
+If a selector is applied to a term $t$, or if $t$'s type is a finite datatype,
+$t$'s equivalence class must contain a
+$\delta$ constructor term.
+This is enforced in the third phase by the \rn{Split} rule.
+Another rule, \rn{Single}, focuses on the degenerate case where two
+terms %from $\tEc$
+are of a \emph{singleton} type (one for which there exists only one value),
+and are therefore equal.
+Notice that corecursive singleton types may have infinite values. A simple example is
+\vthinspace$\keyw{codata}~\ty{a} = \const{A}(\ty{a})$, which is corecursive
+and yet has a cardinality of one;
+its unique value is $\MU a.\; \const{A}(a)$.
+Both
+\rn{Split}'s finiteness assumption %on the datatype
+and \rn{Single}'s singleton
+constraint %on the codatatype
+can be evaluated statically
+based on a recursive computation of the cardinalities of the
+constructors' argument types.
+%This step relies on the assumption that uninterpreted types are infinite.
 

### 2016-02-09 sister/sister.tex
tuning
more

@@
 \usepackage[T1]{fontenc}
+\usepackage{flushend}
 \usepackage{amsmath}
@@
 
+%\let\emph=\textit
+%\let\em=\relax %%% avoid emphasized "et al."
+
 \newcommand\yasx{\mskip.5mu y\mskip.5mu} %%% TYPESETTING: hack so that y is exactly as wide as x
@@
 
-\title{A Decision Procedure for (Co)datatypes in SMT Solvers}
-
-\author {Andrew Reynolds \and Jasmin Christian Blanchette\thanks{In memoriam Morgan Deters 1979--2015} \\
-Department of Computer Science, The University of Iowa, USA \and \\
-Inria Nancy -- Grand Est \& LORIA, Villers-l\`es-Nancy, France
+\title{A Decision Procedure for (Co)datatypes in SMT Solvers\thanks{In memoriam Morgan Deters 1979--2015}}
+
+\author {Andrew Reynolds \\
+Department of Computer Science \\
+The University of Iowa, USA
+\And
+Jasmin Christian Blanchette \\
+Inria Nancy -- Grand Est \& LORIA \\
+Villers-l\`es-Nancy, France
 }
@@
 Datatypes and codatatypes are useful to represent finite and potentially
-infinite data structures and processes. We describe a decision procedure
+infinite objects. We describe a decision procedure
 to reason about such types. The procedure has been
@@
 %
-%(using an ML-like syntax)
+(using an ML- or Haskell-like syntax)
 introduces a type that
 captures the natural numbers $\const{Z}$, $\const{S}(\const{Z})$, $\const{S}(\const{S}(\const{Z}))$, $\ldots$\afterLdots{},
-in Peano (unary) notation, extended with an
+in unary notation, extended with an
 infinite value $\infty = \const{S}(\const{S}(\const{S}(\ldots)))$.
@@
 because both sides expand to the infinite term
-$\const{S}(\const{S}(\const{S}(\ldots)))$, which uniquely identifies
-%the value~
-$\infty$.
+$\const{S}(\const{S}(\const{S}(\ldots)))$, i.e., $\infty$.
 
@@
 calculus and can be composed with other theories in an SMT solver.
-It generalizes the procedure by Barrett et al.\
+It generalizes the procedure by Barrett {\em et al.}\
 \shortcite{barrett-et-al-2007},
@@
 \[
-\begin{aligned}[t]
+\!\begin{aligned}[t]
 \text{Distinctness:}\enskip
@@
 %The presentation is
-%inspired by Barrett et al.\ \shortcite{barrett-et-al-2007} but at
+%inspired by Barrett {\em et al.} \shortcite{barrett-et-al-2007} but at
 %a higher level, using unoriented equations instead of oriented ones.
@@
 
+%%% TYPESETTING: hacks
+
+\vfill
+
+~
+
 \end{document}

### 2016-02-09 sister/sister.tex
more tuning

@@
 
-\newcommand\DISC[1]{t \teq \const{C}_{#1}\negvthinspace\bigl(\const s^1_{#1}( t ), \ldots,\const s^{n_{#1}}_{#1}( t )\bigr)}
+\newcommand\DISC[1]{t \teq \const{C}_{#1}\negvthinspace\bigl(\const s^1_{#1}( t ), \dotsc,\const s^{n_{#1}}_{#1}( t )\bigr)}
 
@@
 automatic theorem prover. An evaluation based on formalizations developed
-in the Isabelle proof assistant demonstrates the potential of our
+in the Isabelle proof assistant demonstrates the potential of the
 procedure.
@@
 prove theorems in first-order logic. These solvers are also useful for model
-finding, or constraint solving: given a set of constraints, they can be used
+finding, or constraint solving: Given a set of constraints, they can be used
 to find a variable assignment that satisfy them.
 
-A great benefit of the SMT approach is that it provides a general framework
+A great benefit of the SMT approach is that it provides a flexible framework
 for composing optimized decision procedures and other solvers for various
@@
 \item Which theories would be useful to applications?
-\item Which theories can be coded efficiently in SMT solvers?
+\item Which theories could be implemented efficiently in \\ %%% TYPESETTING: hack
+SMT solvers?
 \end{itemize}
@@
 introduces a type that
-captures the natural numbers $\const{Z}$, $\const{S}(\const{Z})$, $\const{S}(\const{S}(\const{Z}))$, $\ldots$\afterLdots{},
+captures the natural numbers $\const{Z}$, $\const{S}(\const{Z})$, $\const{S}(\const{S}(\const{Z}))$, $\dotsc$,
 in unary notation, extended with an
@@
 Berlin, Germany \cite{reynolds-blanchette-2015-codata}. An more comprehensive
-version %, including detailed proofs,
+article %, including detailed proofs,
 will appear in a special issue of the
@@
 \!(\keyw{co})\keyw{data}\;\,
-  \delta_1 & {}=\, \smash{\const{C}_{11\!}(\bigl[\const{s}_{11\!}^1{:}\bigr]\vthinspace \tau_{11\!}^1, \ldots, \bigl[\const{s}_{11\!}^{n_{11\!}}{:}\bigr]\vthinspace \tau_{11\!}^{n_{11\!}})} \mid \cdots \mid \smash{\const{C}_{1m_1\!}(\ldots)} \\[+0.5\jot]
+  \delta_1 & {}=\, \smash{\const{C}_{11\!}(\bigl[\const{s}_{11\!}^1{:}\bigr]\vthinspace \tau_{11\!}^1, \dotsc, \bigl[\const{s}_{11\!}^{n_{11\!}}{:}\bigr]\vthinspace \tau_{11\!}^{n_{11\!}})} \mid \cdots \mid \smash{\const{C}_{1m_1\!}(\ldots)} \\[+0.5\jot]
    \smash{\vdots\,\,\,} \\[-0.5\jot]
-  \keyw{and}\; \,\delta_{\elllx} & =\, \smash{\const{C}_{\elll 1\!}(%\bigl[\const{s}_{\elll 1\!}^1{:}\bigr]\vthinspace \tau_{\elll 1\!}^1, \ldots, \bigl[\const{s}_{\elll 1\!}^{n_{\elll 1\!}}{:}\bigr]\vthinspace \tau_{\elll 1\!}^{n_{\elll 1\!}}
+  \keyw{and}\; \,\delta_{\elllx} & =\, \smash{\const{C}_{\elll 1\!}(%\bigl[\const{s}_{\elll 1\!}^1{:}\bigr]\vthinspace \tau_{\elll 1\!}^1, \dotsc, \bigl[\const{s}_{\elll 1\!}^{n_{\elll 1\!}}{:}\bigr]\vthinspace \tau_{\elll 1\!}^{n_{\elll 1\!}}
   \ldots)} \mid \cdots \mid \smash{\const{C}_{\elll m_{\elllx}\!}(\ldots)}
@@
 be an abbreviation for
-$t \teq \const{C}_{i\negvthinspace j}\relax(\const{s}_{i\negvthinspace j}^1(t), \ldots, \const{s}_{i\negvthinspace j}^{\,n_{\smash{{i\negvthinspace j}}}}(t)\relax)$.
+$t \teq \const{C}_{i\negvthinspace j}\relax(\const{s}_{i\negvthinspace j}^1(t), \dotsc, \const{s}_{i\negvthinspace j}^{\,n_{\smash{{i\negvthinspace j}}}}(t)\relax)$.
 %This will simplify the presentation. % of the decision procedure.
@@
 \text{Injectivity:}\enskip
-  & %\forall x_1,\ldots,x_{n_{i\negvthinspace j}} y.\;\,
-    \smash{\const{C}_{i\negvthinspace j}(x_1,\ldots,x_{n_{i\negvthinspace j}}) \teq \const{C}_{i\negvthinspace j}(y_1,\ldots,y_{n_{i\negvthinspace j}}) \medrightarrow x_k \teq y_k}
+  & %\forall x_1,\dotsc,x_{n_{i\negvthinspace j}} y.\;\,
+    \smash{\const{C}_{i\negvthinspace j}(x_1,\dotsc,x_{n_{i\negvthinspace j}}) \teq \const{C}_{i\negvthinspace j}(y_1,\dotsc,y_{n_{i\negvthinspace j}}) \medrightarrow x_k \teq y_k}
   \\[-0.5\jot]
@@
 \text{Selection:}\enskip
-  & \smash{\const{s}_{i\negvthinspace j}^{\,k}(\const{C}_{i\negvthinspace j}(x_1,\ldots,x_{n_{i\negvthinspace j}})) \teq x_k}
+  & \smash{\const{s}_{i\negvthinspace j}^{\,k}(\const{C}_{i\negvthinspace j}(x_1,\dotsc,x_{n_{i\negvthinspace j}})) \teq x_k}
 \end{aligned}
@@
 
-%We will commonly denote tuples of terms $( t_1, \ldots, t_n )$ in bold font, as $\vec t$.
-%We assume that all $\Sig$-terms in $\Ec$ are \emph{normalized}, meaning that all subterms of the form $\const{s}^{\,k}_{\negvthinspace j}( \const{C}_{\negvthinspace j}(t_1,\ldots,t_n) )$
+%We will commonly denote tuples of terms $( t_1, \dotsc, t_n )$ in bold font, as $\vec t$.
+%We assume that all $\Sig$-terms in $\Ec$ are \emph{normalized}, meaning that all subterms of the form $\const{s}^{\,k}_{\negvthinspace j}( \const{C}_{\negvthinspace j}(t_1,\dotsc,t_n) )$
 %are simplified to $t_k$. All literals added to $\Ec$ are also implicitly
@@
 
-%We assume that all $\Sig$-terms in $\Ec$ are \emph{normalized}, meaning that all subterms of the form $\const{s}^{\,k}_{\negvthinspace j}( \const{C}_{\negvthinspace j}(t_1,\ldots,t_n) )$
+%We assume that all $\Sig$-terms in $\Ec$ are \emph{normalized}, meaning that all subterms of the form $\const{s}^{\,k}_{\negvthinspace j}( \const{C}_{\negvthinspace j}(t_1,\dotsc,t_n) )$
 %are simplified to $t_i$, and moreover assume that all %additional
@@
 %(Section~\ref{sec:introduction}).
-% for example, $\MU x.\; \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
+% for example, $\MU x.\; \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots))$.
 %either in the case when $\tau$ is a codatatype or datatype type,
@@
   \betweenantes
-  \const{C}( t_1, \ldots, t_n ) \in \ec{u}
+  \const{C}( t_1, \dotsc, t_n ) \in \ec{u}
   \betweenantes
@@
 }{
-  \Val := \Val [{\Varec{u}} \mapsto \MU {\Varec{u}}.\; \const{C}( {\Varec{t_1}}, \ldots, {\Varec{t_n}} )]
+  \Val := \Val [{\Varec{u}} \mapsto \MU {\Varec{u}}.\; \const{C}( {\Varec{t_1}}, \dotsc, {\Varec{t_n}} )]
 }
@@
 The resulting $\Val$ indicates that the values for $x$ and $y$ in models of $\Ec$
-must be of the forms $\const{C}( {\Varec{w}}, \const{C}( {\Varec{z}},\allowbreak \const{C}( {\Varec{w}},\allowbreak \const{C}( {\Varec{z}},\ldots ))))$
-and $\const{C}( {\Varec{z}}, \const{C}( {\Varec{w}}, \const{C}( {\Varec{z}}, \const{C}( {\Varec{w}}, \ldots ))))$,
+must be of the forms $\const{C}( {\Varec{w}}, \const{C}( {\Varec{z}},\allowbreak \const{C}( {\Varec{w}},\allowbreak \const{C}( {\Varec{z}},\ldots))))$
+and $\const{C}( {\Varec{z}}, \const{C}( {\Varec{w}}, \const{C}( {\Varec{z}}, \const{C}( {\Varec{w}}, \ldots))))$,
 respectively. %, for some values of ${\Varec{w}}$ and ${\Varec{z}}$.
@@
   \betweenantes
-  \Ctr^\delta = \{ \const{C}_1, \ldots, \const{C}_m \}
+  \Ctr^\delta = \{ \const{C}_1, \dotsc, \const{C}_m \}
 \\

### 2016-02-10 sister/sister.tex
more
clearer
more polish

@@
 to reason about such types. The procedure has been
-integrated in CVC4, a modern SMT (satisfiability modulo theories)
+integrated into CVC4, a modern SMT (satisfiability modulo theories)
 solver, which can be used both as a constraint solver and as an
 automatic theorem prover. An evaluation based on formalizations developed
-in the Isabelle proof assistant demonstrates the potential of the
+in the Isabelle proof assistant shows the potential of the
 procedure.
@@
 prove theorems in first-order logic. These solvers are also useful for model
-finding, or constraint solving: Given a set of constraints, they can be used
+finding, or constraint solving: Given some constraints, they can be used
 to find a variable assignment that satisfy them.
@@
 A great benefit of the SMT approach is that it provides a flexible framework
-for composing optimized decision procedures and other solvers for various
+for composing dedicated decision procedures and other solvers for various
 theories, including equality, linear arithmetic on $\mathbb{Z}$ and
@@
 \emph{codatatypes}, respectively.
-Datatypes are ubiquitous in functional programs and formal specifications.
+Datatypes are ubiquitous in functional programming and formal specifications.
 They are useful to represent finite data structures in computer
@@
 For the latter, a uniqueness rule ensures that two values
-having the same infinite expansion must be equal; from $x
+having the same infinite expansion are equal; from $x
 \teq \const{S}(x)$ and
@@
 We fix a signature %$\Sig = (\Types, \Funcs)$
-consisting of a set of types $\Types$ and a
-set of function symbols $\Funcs$\negvthinspace.
+consisting of a set of types and a
+set of function symbols.
 The types are partitioned into
-$\Types = \Data \mathrel{\uplus} \Codata \mathrel{\uplus} \Nondata$, where $\Data$ are the
-\emph{datatypes}, $\Codata$ are the \emph{codatatypes}, and $\Nondata$ are the remaining
-\emph{ordinary types}.
+the \emph{datatypes} $\Data$,
+the \emph{codatatypes} $\Codata$,
+and the are the remaining \emph{ordinary types} $\Nondata$.
 % (which can be interpreted or not).
 The function symbols are
-partitioned into $\Funcs = \Ctr \mathrel{\uplus} \Sel$, where $\Ctr$ are the
-\emph{constructors} and $\Sel$ are the \emph{selectors}. There is no need to
-consider further function symbols because they can be abstracted away as
-variables when combining theories.
+partitioned into the
+\emph{constructors} $\Ctr$ and the \emph{selectors} $\Sel$.
+There is no need to consider further function symbols because they can be
+abstracted away as variables when combining theories.
 %Exceptionally, it is convenient to use numeric constants ($0$, $1$, \ldots)\ 
@@
 For the natural numbers constructed from $\const{Z}$ and $\const{S}$,
-induction prohibits models that contain infinite values.
-% $\const{S}(\const{S}(\ldots))$.
+induction prohibits infinite values $\const{S}(\const{S}(\ldots))$.
 %
@@
 %$\thD$-
-satisfiability of finite sets $\Ec$ of literals:\ equalities
-and disequalities between first-order terms, %$\Sig$-terms,
+satisfiability of finite sets $\Ec$ of constraints:\ equalities ($\teq$)
+and disequalities ($\not\teq$) between first-order terms, %$\Sig$-terms,
 whose variables are interpreted existentially.
@@
 % (in which case we call an application of it \emph{terminal}).
-One of the rules has multiple conclusions, % separated by $\ror$,
+One of the rules has multiple conclusions separated by $\left|\kern-.1ex\right|$,
 denoting branching.
@@
 and failures to unify equated terms are recognized by \rn{Clash}.
-\rn{Conflict} recognizes when an equality and its negation both occur in $\Ec$, in which case $\Ec$ has no model.
+\rn{Conflict} recognizes when an equality and its negation both occur in $\Ec$, in which case $\Ec$ is
+unsatisfiable.
 
@@
 \rn{Split}
-\\[5\jot]
+\\[6\jot]
 \(

### 2016-02-10 jour/jour.tex
more polish
more

@@
 A derivation rule can be applied to $\Ec$ if %all
-the specified
-premises are met.
+its premises are met.
 The conclusion either specifies equalities to be added to $\Ec$

### 2016-02-10 sister/sister.tex
more

@@
 % \cite{ganzinger-et-al-2004}
-have emerged as one of the most powerful ways to
-prove theorems in first-order logic. These solvers are also useful for model
-finding, or constraint solving: Given some constraints, they can be used
-to find a variable assignment that satisfy them.
+have emerged as one of the most powerful ways to prove theorems in first-order
+logic. These solvers are also useful for constraint solving, or model finding:
+Given some constraints, they can be used to find an assignment, or model, that
+satisfies them.
 
@@
 theories, including equality, linear arithmetic on $\mathbb{Z}$ and
-$\mathbb{R}$, and first-order quantifiers ($\forall$,~$\exists$). In this
+$\mathbb{R}$, bit vectors (for representing $n$-bit machine words),
+and first-order quantifiers ($\forall$,~$\exists$). In this
 context, two recurrent scientific questions are:
@@
 
-This paper suggests the following partial answer:\ freely generated algebraic
+This paper offers the following partial answer:\ freely generated algebraic
 and coalgebraic datatypes---also called \emph{datatypes} and
@@
 %
-Third, \emph{it must be possible to express cyclic values as closed terms and
-to enumerate them.} %This is necessary when generating models.
-The \hbox{$\mu$-binder} notation associates a name with
-a subterm, making it possible to represent cyclic values.
-For example,
-the $\mu$-term $\const{SCons}(1,\: \MU s.\; \const{SCons}(0,\: \const{SCons}(9,\: s)))$
+Third, \emph{it must be possible to express cyclic values as closed terms.} %This is necessary when generating models.
+For this, we use the \hbox{$\mu$-binder} notation to give a name to
+a repeated subterm. Thus,
+the $\mu$-term $\const{SCons}(1,\:\allowbreak \MU s.\; \const{SCons}(0,\:\allowbreak \const{SCons}(9,\: s)))$
 stands for the %lasso-shaped

### 2016-02-10 sister/sister.tex
used lighter notations
tuning
more edits

@@
 }{
-  \Ec := \Ec,\: t \teq t
+  t \teq t \in \Ec
 }
@@
 }{
- \Ec := \Ec,\: u \teq t
+ u \teq t \in \Ec
 }
@@
 \rn{Sym}
-\\[5\jot]
+\quad
 \(
@@
 }{
-  \Ec := \Ec,\: s \teq u
+  s \teq u \in \Ec
 }
@@
 \rn{Trans}
-\quad
+\\[5\jot]
 \(
@@
 }{
-  \Ec := \Ec,\: \const f( \vec t \,) \teq \const f( \vec u )
+  \const f( \vec t \,) \teq \const f( \vec u ) \in \Ec
 }
@@
 }{
-  \Ec := \Ec,\: \vec t \teq \vec u
+  \vec t \teq \vec u \in \Ec
 }
@@
 }{
- \Ec := \Ec,\: t \teq u
+ t \teq u \in \Ec
 }
@@
 but
-$\MU x.\; \const{C}( x ) \not\aequiv \MU x.\; \const{D}( y, x ) \not\aequiv \MU x.\; \const{D}( z, x )
-\not\aequiv \MU y.\; \const{D}( y, x )$.
+$\MU x.\; \const{C}( x )$, $\MU x.\; \const{D}( y, x )$, $\MU x.\; \const{D}( z, x )$,
+and $\MU y.\; \const{D}( y, x )$ are all $\alpha$-disequivalent.
 Two $\mu$-terms can denote the same value despite being $\alpha$-disequivalent---e.g.,
@@
 A possible sequence of unfolding steps is given below, omitting
-trivial entries such as $\ec{w} \mapsto {\Varec{w}}$.
+trivial entries $\ec{t} \mapsto {\Varec{t}}$.
 %
@@
 }{
-  {\mathop{\big|\kern-.25ex\big|}\vthinspace}_{i=1}^n\, \Ec := \Ec,\: \DISC{i} %\ROR
- % \ROR \cdots \ROR \Ec := \Ec,\: \DISC{m}
+  {\mathop{\big|\kern-.25ex\big|}\vthinspace}_{i=1}^n\, \DISC{i} \in \Ec %\ROR
+ % \ROR \cdots \ROR \DISC{m} \in \Ec
 }
@@
 }{
-  \Ec := \Ec,\: t \teq u
+  t \teq u \in \Ec
 }
@@
 $t$'s equivalence class must contain a
-$\delta$ constructor term.
+%$\delta$
+constructor term.
 This is enforced in the third phase by the \rn{Split} rule.
@@
 the conjecture
-$\const{zeros} = \const{repeat}(0)$
+$\const{zeros} \teq \const{repeat}(0)$
 would be ``refuted'' by a spurious countermodel that interprets
@@
 
-\begin{table}[h!]
+\begin{table}[t!]
 \normalsize
@@
 %obviously depends on the application.
-Table~\ref{tab:bench} indicates that
+The table indicates that
 the theoretically stronger instances of the decision

### 2016-02-10 jour/jour.tex
more

@@
 \[\begin{aligned}[t]
-      \keyw{codatatype}\; \,\ty{llist}_{\vthinspace\tau} & \,=\, \const{LNil} \,\mid\, \const{LCons}(%\const{lhead}{:}\;
-      \tau,\: %\const{ltail}{:}\;
+      \keyw{datatype}\; \,\ty{list}_{\vthinspace\tau} & \,=\, \const{Nil} \,\mid\, \const{Cons}(\const{hd}{:}\;
+      \tau,\: \const{tl}{:}\; \ty{list}_{\vthinspace\tau}) \\%[-.5\jot]
+      \keyw{codatatype}\; \,\ty{llist}_{\vthinspace\tau} & \,=\, \const{LNil} \,\mid\, \const{LCons}(\const{lhd}{:}\;
+      \tau,\: \const{ltl}{:}\;
       \ty{llist}_{\vthinspace\tau}) \\%[-.5\jot]
@@
 For the natural numbers constructed from $\const{Z}$ and $\const{S}$,
-induction prohibits models that contain cyclic values---e.g.,
-an $n$ such that $n \teq \const{S}(n)$---or even infinite acyclic values
+induction prohibits models that contain infinite values
 $\const{S}(\const{S}(\ldots))$.
@@
 \AXC{\strut$\begin{gathered}\textstyle \smash{\bigwedge_{\vthinspace i=1}^\ell \, R_i[v_i, w_i]} \\[-\jot]\textstyle
-\bigwedge_{\vthinspace i=1}^\ell\, \bigl(\forall v\; w.\;\, R_i[v, w] \medrightarrow
-  \bigwedge_{\vthinspace j=1}^{m_i}\, \bigl(\const{d}_{\negvthinspace j}(v) \mathrel\land \const{d}_{\negvthinspace j}(w)
-  \medrightarrow \bigwedge_{\vthinspace k=1}^{n_{i\negvthinspace j}}\, \const{s}_{i\negvthinspace j}^{\,k}(v) \sim \const{s}_{i\negvthinspace j}^{\,k}(w)\bigr)\bigr)
+\bigwedge_{\vthinspace i=1}^\ell\, \bigl(\forall x\; y.\;\, R_i[x, y] \medrightarrow
+  \bigwedge_{\vthinspace j=1}^{m_i}\, \bigl(\const{d}_{\negvthinspace j}(x) \medrightarrow \const{d}_{\negvthinspace j}(y)
+  \mathrel\land \bigwedge_{\vthinspace k=1}^{n_{i\negvthinspace j}}\, \const{s}_{i\negvthinspace j}^{\,k}(x) \sim \const{s}_{i\negvthinspace j}^{\,k}(y)\bigr)\bigr)
 \end{gathered}$}
@@
 
-In addition, codatatypes are guaranteed to contain all values corresponding to
+\begin{examplex}
+The induction principle for $\ty{list}_{\vthinspace\tau}$ is given below:
+\[
+\AXC{\strut$P[\const{Nil}]$}
+  \AXC{\strut$\forall x,\, \mathit{xs}.\;\vthinspace P[\mathit{xs}] \medrightarrow P[\const{Cons}(x, \mathit{xs})]$}
+\BIC{\strut$P[vs]$}
+\DP
+\]
+Assuming that $\const{lnull}$ and its negation are the discriminators
+associated with \const{LNil} and \const{LCons},
+the coinduction principle for $\ty{llist}_{\vthinspace\tau}$ is
+\[
+\AXC{\strut$R[\mathit{vs},\, \mathit{ws}]$}
+  \AXC{$\phantom{\forall \mathit{xs}\; \mathit{ys}.\;\,} \!\begin{aligned}[b]
+      \llap{\ensuremath{\forall \mathit{xs}\; \mathit{ys}.\;\,}}
+    & R[\mathit{xs}, \mathit{ys}]
+      \medrightarrow
+      \bigl(\const{lnull}(\mathit{xs}) \medrightarrow \const{lnull}(\mathit{ys})\bigr)
+\\[-\jot]      
+    & \kern2em{\mathrel\land}\;
+      \bigl(\lnot\;\const{lnull}(\mathit{xs}) \medrightarrow \lnot\;\const{lnull}(\mathit{ys})
+        \mathrel\land \const{hd}(\mathit{xs}) \teq \const{hd}(\mathit{ys})
+        \mathrel\land R[\const{tl}(\mathit{xs}), \const{tl}(\mathit{ys})]\bigr)
+\end{aligned}$}
+\BIC{\strut$\mathit{vs} \teq \mathit{ws}$}
+\DP
+\]
+\xend
+\end{examplex}
+
+Codatatypes are guaranteed to contain all values corresponding to
 infinite ground constructor terms.
@@
 }{
-  \Ec := \Ec,\: \DISC{1} %\ROR
-  \ROR \cdots \ROR \Ec := \Ec,\: \DISC{m}
+  \DISC{1} \in \Ec \ROR \cdots \ROR \DISC{m} \in \Ec
 }

### 2016-02-20 sister/sister.tex
tuned
Minor
minor

@@
 % \cite{ganzinger-et-al-2004}
-have emerged as one of the most powerful ways to prove theorems in first-order
-logic. These solvers are also useful for constraint solving, or model finding:
+have emerged as one of the most powerful ways to prove theorems in classical
+first-order logic. These solvers are also useful for constraint solving, or model finding:
 Given some constraints, they can be used to find an assignment, or model, that
@@
 and first-order quantifiers ($\forall$,~$\exists$). In this
-context, two recurrent scientific questions are:
+context, recurrent scientific questions include:
 \begin{itemize}
@@
 
-This paper offers the following partial answer:\ freely generated algebraic
+This paper offers the following partial answer:\ 
+freely generated algebraic
 and coalgebraic datatypes---also called \emph{datatypes} and
-\emph{codatatypes}, respectively.
+\emph{codatatypes}.
 Datatypes are ubiquitous in functional programming and formal specifications.
@@
 its premises are met.
-The conclusion either specifies equalities to be added to $\Ec$
+The conclusion either specifies an equality to be added to $\Ec$
 or is $\bot$ (contradiction).

### 2016-02-23 jour/jour.tex
Address minor points from reviewer.

@@
 If $\Val$ is constructed for a saturated set $\Ec$
-and $\ValC$ is a completion of $\Val$ for a normal interpretation $\J$, the following properties hold:
+and $\ValC$ is a completion of $\Val$ for a normal interpretation $\J$, the following hold for all $\ec{x}$ and $\ec{y}$ in the domain of $\ValC$:
 \begin{enumerate}
@@
 explanation for $\Val \ec{x} = \MU {\Varec{x}}.\; \const{C}( {\Varec{x}} )$.
+In the case that multiple rules could derive the same conclusion,
+the implementation simply keeps the first explanation it encountered.
 The rules of the calculus are implemented as follows. For all rules with
 conclusion $\bot$, we report the union of the explanations for all premises is
-$\thD$-unsatisfiable. For \rn{Split}, we add the exhaustiveness clause %of the form
+$\thD$-unsatisfiable. The implementation does not 
+use any techniques to minimize the size of this set, which in some cases may be non-minimal.
+For \rn{Split}, we add the exhaustiveness clause %of the form
 $\DISC{1} \mathrel\lor \cdots \mathrel\lor \DISC{m}$ to $F$.
@@
 based on whether $j = \jPrime$.
+Finally, finite ordinary types are not converted to finite datatypes.
+We instead rely on standard extensions of Nelson-Oppen theory combination~\cite{tinelli2005combining}, which
+impose requirements regarding terms having finite sorts are shared
+between multiple theories.  In particular, given an ordinary type $\tau$ of finite cardinality $n$
+belonging to theory $T_i$, the theory solver for $T_i$ must 
+communicate enough equalities over shared terms to
+ensure all other theory solvers, including the one for $\thD$, 
+interpret $\tau$ as a set having cardinality at most $n$.
 

### 2016-02-24 jour/jour.tex
Extend example to completition -> interpretation.

@@
 \begin{examplex}
+\label{ex:assign}
 Let $\delta$ be a codatatype with the constructors $\const{C},\, \const{D},\, \const{E} :\delta \rightarrow \delta$.
@@
 x \teq \const{C}( v ),\;
+v \teq s,\;
 z \not\teq v
@@
 %
+Here, $\ec{u}$ denotes the equivalence class $\{ u, \const{C}( z ) \}$,
+$\ec{v}$ denotes $\{ v, \const{C}( z ), s \}$, and so on.
 To construct a completion $\ValC$, we must choose values for ${\Varec{y}}$ and ${\Varec{z}}$,
@@
 
+\begin{examplex}
+\label{ex:assign-complete}
+Continuing with Example~\ref{ex:assign}, while extending $\Val$ to its corresponding completion $\ValC$,
+we may, for instance, assign $\MU {\Varec{z}}.\; \const{E}( {\Varec{z}} )$ to $\Varec{z}$,
+and subsequently assign $\MU {\Varec{y}}.\; \const{D}( {\Varec{y}} )$ to $\Varec{y}$.
+We obtain the following mapping $\ValC$ from equivalence classes to $\MU$-terms:
+\[\begin{array}{r@{}c@{}r@{}l@{\qquad}r@{}c@{}r@{}l@{\qquad}r@{}c@{}l}
+\ValC \ec{u} & {}\,=\,{} & \MU {\Varec{u}}.\; & \const{C}( \MU {\Varec{z}}.\; \const{E}( {\Varec{z}} ) ) &
+\ValC \ec{w} & = & \MU {\Varec{w}}.\; & \const{E}( \MU {\Varec{y}}.\; \const{D}( {\Varec{y}} ) ) &
+\ValC \ec{y} & {}\,=\,{} & \MU {\Varec{y}}.\; \const{D}( {\Varec{y}} ) \\
+\ValC \ec{v} & = & \MU {\Varec{v}}.\; & \const{D}( \MU {\Varec{z}}.\; \const{E}( {\Varec{z}} ) ) &
+\ValC \ec{x} & {}\,=\,{} & \MU {\Varec{x}}.\; & \const{C}( \MU {\Varec{v}}.\; \const{D}( \MU {\Varec{z}}.\; \const{E}( {\Varec{z}} ) ) ) &
+\ValC \ec{z} & = & \MU {\Varec{z}}.\; \const{E}( {\Varec{z}} )
+\end{array}\]
+Notice that $\ValC$ satisfies the properties from Lemma~\ref{lem:interpretation-completion} for a normal interpretation $\J$:
+all terms in the range of $\ValC$ are closed and normal, 
+$\ValC \ec{s} = \expand{\ValC \ec{y}}{ t }$ for all subterms $t$ where $\muvar(t) = {\Varec{s}}$ for all $s$,
+and all terms in the range of $\ValC$ are pairwise alpha-disequivalent.
+We may construct an interpretation $\J$ where variables are assigned the value
+that is alpha-equivalent to one from $\interp{\J}{\delta}$, and is uniquely associated to its equivalence class by $\ValC$.
+For instance, up to renaming,
+$\interp{\J}{u}$ is $\const{C}( \MU {\Varec{z}}.\; \const{E}( {\Varec{z}} ) )$,
+$\interp{\J}{v}$ and $\interp{\J}{s}$ are $\const{D}( \MU {\Varec{z}}.\; \const{E}( {\Varec{z}} ) )$, and so on.
+\xend
+\end{examplex}
+
 \begin{theorem}[Solution Soundness]%

### 2016-02-24 jour/jour.tex
Finish examples to supplment solution soundness.

@@
 \label{ex:assign-complete}
-Continuing with Example~\ref{ex:assign}, while extending $\Val$ to its corresponding completion $\ValC$,
+Consider $\Val$ from Example~\ref{ex:assign}.
+While extending $\Val$ to its corresponding completion $\ValC$,
 we may, for instance, assign $\MU {\Varec{z}}.\; \const{E}( {\Varec{z}} )$ to $\Varec{z}$,
@@
 and all terms in the range of $\ValC$ are pairwise alpha-disequivalent.
-We may construct an interpretation $\J$ where variables are assigned the value
+We may construct an interpretation $\J$ where variables are interpreted as the value
 that is alpha-equivalent to one from $\interp{\J}{\delta}$, and is uniquely associated to its equivalence class by $\ValC$.
-For instance, up to renaming,
-$\interp{\J}{u}$ is $\const{C}( \MU {\Varec{z}}.\; \const{E}( {\Varec{z}} ) )$,
-$\interp{\J}{v}$ and $\interp{\J}{s}$ are $\const{D}( \MU {\Varec{z}}.\; \const{E}( {\Varec{z}} ) )$, and so on.
+%For instance, up to renaming,
+%$\interp{\J}{u}$ is $\const{C}( \MU {\Varec{z}}.\; \const{E}( {\Varec{z}} ) )$,
+%$\interp{\J}{v}$ and $\interp{\J}{s}$ are $\const{D}( \MU {\Varec{z}}.\; \const{E}( {\Varec{z}} ) )$, and so on.
 \xend
@@
 \end{proof}
+
+\begin{examplex}
+\label{ex:model}
+Continuing Example~\ref{ex:assign-complete}, let $\J$ be a normal interpretation, where
+up to renaming of $\MU$-bound variables, we have:
+\[\begin{array}{r@{}c@{}r@{}c@{}l@{\qquad}r@{}c@{}l@{\qquad}r@{}c@{}l}
+& &\interp{\J}{u} & {}\,=\,{} &  \const{C}( \MU {\Varec{z}}.\; \const{E}( {\Varec{z}} ) ) &
+\interp{\J}{w} & = &  \const{E}( \MU {\Varec{y}}.\; \const{D}( {\Varec{y}} ) ) &
+\interp{\J}{y} & {}\,=\,{} & \MU {\Varec{y}}.\; \const{D}( {\Varec{y}} ) \\
+\interp{\J}{v} & = & \interp{\J}{s} & = & \const{D}( \MU {\Varec{z}}.\; \const{E}( {\Varec{z}} ) ) &
+\interp{\J}{x} & {}\,=\,{} & \const{C}( \const{D}( \MU {\Varec{z}}.\; \const{E}( {\Varec{z}} ) ) ) &
+\interp{\J}{z} & = & \MU {\Varec{z}}.\; \const{E}( {\Varec{z}} )
+\end{array}\]
+It can be shown $\J$ satisfies $\Ec =
+\{
+u \teq \const{C}( z ),\;
+v \teq \const{D}( z ),\;
+w \teq \const{E}( y ),\;
+x \teq \const{C}( v ),\;
+v \teq s,\;
+z \not\teq v
+\}$.
+\xend
+\end{examplex}
 

### 2016-02-26 jour/jour.tex
tuning

@@
 If $\Val$ is constructed for a saturated set $\Ec$
-and $\ValC$ is a completion of $\Val$ for a normal interpretation $\J$, the following hold for all $\ec{x}$ and $\ec{y}$ in the domain of $\ValC$:
+and $\ValC$ is a completion of $\Val$ for a normal interpretation $\J$, the following
+properties hold for all $\ec{x}$ and $\ec{y}$ in the domain of $\ValC$:
 \begin{enumerate}
@@
 \end{array}\]
-Notice that $\ValC$ satisfies the properties from Lemma~\ref{lem:interpretation-completion} for a normal interpretation $\J$:
-all terms in the range of $\ValC$ are closed and normal, 
+Notice that $\ValC$ satisfies the properties from Lemma~\ref{lem:interpretation-completion}
+for a normal interpretation $\J$:
+All terms in the range of $\ValC$ are closed and normal,
 $\ValC \ec{s} = \expand{\ValC \ec{y}}{ t }$ for all subterms $t$ where $\muvar(t) = {\Varec{s}}$ for all $s$,
-and all terms in the range of $\ValC$ are pairwise alpha-disequivalent.
+and all terms in the range of $\ValC$ are pairwise $\alpha$-disequivalent.
 We may construct an interpretation $\J$ where variables are interpreted as the value
-that is alpha-equivalent to one from $\interp{\J}{\delta}$, and is uniquely associated to its equivalence class by $\ValC$.
+that is $\alpha$-equivalent to one from $\interp{\J}{\delta}$ and that is uniquely associated
+with its equivalence class by $\ValC$.
 %For instance, up to renaming,
@@
 Continuing Example~\ref{ex:assign-complete}, let $\J$ be a normal interpretation, where
-up to renaming of $\MU$-bound variables, we have:
+up to renaming of $\MU$-bound variables, we have
 \[\begin{array}{r@{}c@{}r@{}c@{}l@{\qquad}r@{}c@{}l@{\qquad}r@{}c@{}l}
@@
 \end{array}\]
-It can be shown $\J$ satisfies $\Ec =
+It can be shown that $\J$ satisfies $\Ec =
 \{
@@
 explanation for $\Val \ec{x} = \MU {\Varec{x}}.\; \const{C}( {\Varec{x}} )$.
-In the case that multiple rules could derive the same conclusion,
-the implementation simply keeps the first explanation it encountered.
+If multiple rules can derive the same conclusion,
+the implementation simply keeps the first explanation it encounters.
+
 The rules of the calculus are implemented as follows. For all rules with
@@
 Finally, finite ordinary types are not converted to finite datatypes.
-We instead rely on standard extensions of Nelson-Oppen theory combination~\cite{tinelli2005combining}, which
+We instead rely on standard extensions of the Nelson--Oppen method~\cite{tinelli2005combining}, which
 impose requirements regarding terms having finite sorts are shared
@@
 communicate enough equalities over shared terms to
-ensure all other theory solvers, including the one for $\thD$, 
+ensure that all other theory solvers, including the one for $\thD$,
 interpret $\tau$ as a set having cardinality at most $n$.

### 2016-02-26 jour/jour.tex
tune
tuning spaces before punctuation

@@
 %constructed by $\const{Z} : \ty{nat}$ and $\const{S} : \ty{nat} \to \ty{nat}$
-%only allow values of the form $\const{S}(\ldots(\const{S}(\const{Z}))\ldots)$.
-
-To represent infinite objects, % such as $\const{S}(\const{S}(\const{S}(\ldots))$,
+%only allow values of the form $\const{S}(\ldots(\const{S}(\const{Z}))\ldots).$
+
+To represent infinite objects, % such as $\const{S}(\const{S}(\const{S}(\ldots)),$
 a natural choice is to turn to coalgebraic datatypes, or \emph{codatatypes},
@@
 introduces a type that
-models the natural numbers $\const{Z}$, $\const{S}(\const{Z})$, $\const{S}(\const{S}(\const{Z}))$, $\ldots$\afterLdots{},
+models the natural numbers $\const{Z},$ $\const{S}(\const{Z}),$ $\const{S}(\const{S}(\const{Z})),$ $\ldots$\afterLdots{},
 in Peano notation, extended with an
-infinite value $\infty = \const{S}(\const{S}(\const{S}(\ldots)))$.
+infinite value $\infty = \const{S}(\const{S}(\const{S}(\ldots))).$
 The equation $\const{S}(\infty) \teq \infty$ holds as expected,
 because both sides expand to the infinite term
-$\const{S}(\const{S}(\const{S}(\ldots)))$, which uniquely identifies
+$\const{S}(\const{S}(\const{S}(\ldots))),$ which uniquely identifies
 %the value~
-$\infty$.
+$\infty.$
 
@@
 \teq \const{S}(x)$ and
-$y \teq \const{S}(y)$, it deduces $x \teq y$.
+$y \teq \const{S}(y),$ it deduces $x \teq y.$
 These two rules are needed to ensure completeness
@@
 
-%  * one implication is that if $m$ equals $ES(n)$ and $n$ equals $ES(m)$, necessarily
+%  * one implication is that if $m$ equals $ES(n)$ and $n$ equals $ES(m),$ necessarily
 %    $m$ and $n$ must be equal.
@@
 
-Function symbols are written in a sans-serif font (e.g., $\const{f}$, $\const{g}$) to
-distinguish them from variables (e.g., $x$, $y$).
+Function symbols are written in a sans-serif font (e.g., $\const{f},$ $\const{g}$) to
+distinguish them from variables (e.g., $x,$ $y$).
 Symbol names starting with an uppercase letter (e.g.,
@@
 is associated a list of argument types $\tau_1,\ldots,\tau_n$ (for $n \ge 0$)
-and a return type $\tau$, written
+and a return type $\tau,$ written
 $\const{f} : \tau_1 \times \cdots \times \tau_n \to \tau$;
-this notation collapses to $\const{f} : \tau$ if $n = 0$.
+this notation collapses to $\const{f} : \tau$ if $n = 0.$
 %
@@
 \const{f} to $n$ %well-typed
-arguments $t_1 \mathbin{:} \tau_1$, \ldots, $t_n \mathbin{:}
+arguments $t_1 \mathbin{:} \tau_1,$ \ldots, $t_n \mathbin{:}
 \tau_n$ of the right types.
@@
 The set $\Funcs$ must at least contain
-$\const{true},\, \const{false} : \ty{bool}$, interpreted as truth values.
+$\const{true},\, \const{false} : \ty{bool},$ interpreted as truth values.
 The only predicate is equality, written $\teq$;
 it belongs to the logical symbols.
-Other predicates can be represented as functions to $\ty{bool}$,
-with $\const{p}(\ldots)$ abbreviating $\const{p}(\ldots) \teq \const{true}$.
+Other predicates can be represented as functions to $\ty{bool},$
+with $\const{p}(\ldots)$ abbreviating $\const{p}(\ldots) \teq \const{true}.$
 The notation $t\typ{\tau}$ stands for a term $t$ of type~$\tau$;
-it should not be confused with $t :\nobreak \tau$,
-which is a statement expressing that $t$ has type $\tau$.
+it should not be confused with $t :\nobreak \tau,$
+which is a statement expressing that $t$ has type $\tau.$
 When applied to terms, the standard equality symbol~$=$ denotes syntactic equality.
 The operator $\bigwedge_{\vthinspace i}\, \varphi_i$ abbreviates a conjunction
-$\varphi_1 \mathrel\land \cdots \mathrel\land \varphi_n$.
-Finally, $\bar x$ abbreviates a list or tuple $x_1,\ldots,x_n$.
+$\varphi_1 \mathrel\land \cdots \mathrel\land \varphi_n.$
+Finally, $\bar x$ abbreviates a list or tuple $x_1,\ldots,x_n.$
 
@@
 
-We fix a signature $\Sig = (\Types, \Funcs)$. The types are partitioned into
-$\Types = \Data \mathrel{\uplus} \Codata \mathrel{\uplus} \Nondata$, where $\Data$ are the
+We fix a signature $\Sig = (\Types, \Funcs).$ The types are partitioned into
+$\Types = \Data \mathrel{\uplus} \Codata \mathrel{\uplus} \Nondata,$ where $\Data$ are the
 \emph{datatypes}, $\Codata$ are the \emph{codatatypes}, and $\Nondata$ are the remaining
 \emph{ordinary types} (which can be interpreted or not). The function symbols are
-partitioned into $\Funcs = \Ctr \mathrel{\uplus} \Sel$, where $\Ctr$ are the
+partitioned into $\Funcs = \Ctr \mathrel{\uplus} \Sel,$ where $\Ctr$ are the
 \emph{constructors} and $\Sel$ are the \emph{selectors}. There is no need to
@@
 variables when combining theories.
-Exceptionally, it is convenient to use numeric constants ($0$, $1$, \ldots)\ in
+Exceptionally, it is convenient to use numeric constants ($0,$ $1,$ \ldots)\ in
 examples.
 $\Sig$-terms are standard first-order terms
-over $\Sig$,
+over $\Sig,$
 without $\mu$-binders.
@@
 family of ground types including
-$\ty{stream}_{\vthinspace\ty{int}}$, $\ty{stream}_{\vthinspace\ty{bool}}$,
+$\ty{stream}_{\vthinspace\ty{int}},$ $\ty{stream}_{\vthinspace\ty{bool}},$
 and \smash{$\ty{stream}_{\vthinspace\ty{stream}_{\vthinspace\ty{real}}}$}.
@@
 $\smash{\const{C}_{i\negvthinspace j} : \tau_{i\negvthinspace j}^1\times\cdots\times\tau_{i\negvthinspace j}^{\,n_{\smash{i\negvthinspace j}}} \to \delta_i}$
-and $\smash{\const{s}_{i\negvthinspace j}^{\,k} : \delta_i \to \tau_{i\negvthinspace j}^{\,k}}$.
+and $\smash{\const{s}_{i\negvthinspace j}^{\,k} : \delta_i \to \tau_{i\negvthinspace j}^{\,k}}.$
 Defaults are assumed for the selector names if they are omitted.
 The $\delta$ constructors and selectors are denoted by $\Ctr^\delta$ and
-$\Sel^\delta$.
+$\Sel^\delta.$
 %
 For types with several constructors, it is useful to provide discriminators
-$\const{d}_{i\negvthinspace j} : \delta_i \to \ty{bool}$. Instead of extending $\Funcs$,
+$\const{d}_{i\negvthinspace j} : \delta_i \to \ty{bool}.$ Instead of extending $\Funcs,$
 we let $\const{d}_{i\negvthinspace j}(t)$
 be an abbreviation for
-$t \teq \const{C}_{i\negvthinspace j}\bigl(\const{s}_{i\negvthinspace j}^1(t), \ldots, \const{s}_{i\negvthinspace j}^{\,n_{\smash{{i\negvthinspace j}}}}(t)\bigr)$.
+$t \teq \const{C}_{i\negvthinspace j}\bigl(\const{s}_{i\negvthinspace j}^1(t), \ldots, \const{s}_{i\negvthinspace j}^{\,n_{\smash{{i\negvthinspace j}}}}(t)\bigr).$
 %This will simplify the presentation. % of the decision procedure.
@@
 we would have the spurious equality
-$\const{s}(\const{D}) \teq \const{s}(\const{E})$. This flaw
+$\const{s}(\const{D}) \teq \const{s}(\const{E}).$ This flaw
 could be corrected, but the added complexity seems to suggest that selectors
@@
 Datatypes and codatatypes share many basic properties. All properties
-below are implicitly universally quantified and range over all $i$, $j$,
-$\jPrime$, and $k$ within bounds:
+below are implicitly universally quantified and range over all $i,$ $j,$
+$\jPrime,$ and $k$ within bounds:
 % and to all possible splits of the
@@
 The induction principle ensures that the interpretation of datatypes is standard.
-For the natural numbers constructed from $\const{Z}$ and $\const{S}$,
+For the natural numbers constructed from $\const{Z}$ and $\const{S},$
 induction prohibits models that contain infinite values
-$\const{S}(\const{S}(\ldots))$.
+$\const{S}(\const{S}(\ldots)).$
 
@@
 derive the equality of pairs of codatatype values
-$v_1, w_1 \mathbin: \delta_1,\,\ldots,\,v_\ell, w_\ell \mathbin: \delta_\ell$, based on
+$v_1, w_1 \mathbin: \delta_1,\,\ldots,\,v_\ell, w_\ell \mathbin: \delta_\ell,$ based on
 suitable coinduction witnesses $R_1,\ldots,R_\ell$:
@@
 For example, $\ty{stream}_{\vthinspace\ty{int}}$ is isomorphic to the uncountable
-function space $\ty{nat} \to \ty{int}$. Bj\o rner gives a rigorous treatment of
+function space $\ty{nat} \to \ty{int}.$ Bj\o rner gives a rigorous treatment of
 this aspect \cite{bjoerner-1998}.
 
-Given a signature $\Sig$, $\thD$ refers to the \emph{theory of
+Given a signature $\Sig,$ $\thD$ refers to the \emph{theory of
 datatypes and codatatypes},
 which %in addition to $\Sig$
-defines a class of $\Sig$-interpretations $\JJJJ$,
+defines a class of $\Sig$-interpretations $\JJJJ,$
 namely those that satisfy the %axioms --- existence of all infinite codatatype values cannot be axiomatized
@@
 have infinite values. A simple example is
-\vthinspace$\keyw{codatatype}~\ty{a} = \const{A}(\ty{a})$, which is corecursive
+\vthinspace$\keyw{codatatype}~\ty{a} = \const{A}(\ty{a}),$ which is corecursive
 and yet has a cardinality of one;
-its unique value is $\MU a.\; \const{A}(a)$. Other specimens are
+its unique value is $\MU a.\; \const{A}(a).$ Other specimens are
 $\ty{stream}_{\vthinspace\ty{unit}\,}$
@@
 assuming \ty{unit} is a datatype with the single constructor $\const{Unity} :
-\ty{unit}$. We call such codatatypes
+\ty{unit}.$ We call such codatatypes
 \emph{corecursive singletons}, or simply singletons.
@@
 \afterDot
-Let $\delta$ be a corecursive codatatype. For any interpretation in $\JJJJ$,
+Let $\delta$ be a corecursive codatatype. For any interpretation in $\JJJJ,$
 the domain interpreting $\delta$ is either infinite or a singleton. In the
@@
 By definition, the type is equipped with at least one (directly or indirectly)
-corecursive constructor $\const{C}$. If it additional has second
-corecursive constructor $\const{D}$, it is possible to encode infinitely many
+corecursive constructor $\const{C}.$ If it additional has second
+corecursive constructor $\const{D},$ it is possible to encode infinitely many
 alternation patterns---e.g.,
@@
 to distinct values (by distinctness and injectivity). If the type has a
-noncorecursive constructor $\const{E}$, it is possible to create terms of
-arbitrary depth---e.g., $\const{C}(\ldots(\const{C}(\const{E}))\ldots)$. In
+noncorecursive constructor $\const{E},$ it is possible to create terms of
+arbitrary depth---e.g., $\const{C}(\ldots(\const{C}(\const{E}))\ldots).$ In
 either case, there can be no finite models.
@@
 
-Given a fixed signature~$\Sig$,
+Given a fixed signature~$\Sig,$
 the decision procedure for the universal theory of (co)data\-types %$\thD$
@@
 To simplify the presentation, we make a few assumptions about
-$\Sig$.
+$\Sig.$
 First, all codatatypes %in $\Codata$
@@
 
-%We will commonly denote tuples of terms $( t_1, \ldots, t_n )$ in bold font, as $\vec t$.
+%We will commonly denote tuples of terms $( t_1, \ldots, t_n )$ in bold font, as $\vec t.$
 %We assume that all $\Sig$-terms in $\Ec$ are \emph{normalized}, meaning that all subterms of the form $\const{s}^{\,k}_{\negvthinspace j}( \const{C}_{\negvthinspace j}(t_1,\ldots,t_n) )$
-%are simplified to $t_k$. All literals added to $\Ec$ are also implicitly
+%are simplified to $t_k.$ All literals added to $\Ec$ are also implicitly
 %normalized in this manner.
@@
 % (in which case we call an application of it \emph{terminal}).
-One of the rules has multiple conclusions, % separated by $\ror$,
+One of the rules has multiple conclusions, % separated by $\ror,$
 denoting branching.
@@
 derivation rule to the parent. A derivation tree is \emph{closed} if all of
-its leaf nodes are $\bot$. A node is \emph{saturated} if no nonredundant
+its leaf nodes are $\bot.$ A node is \emph{saturated} if no nonredundant
 instance of a rule can be applied to it.
@@
 Figs.\ \ref{fig:cc-rules}~to~\ref{fig:split-rule}, corresponding to three
-phases of the calculus. The first phase computes the bidirectional closure of $\Ec$. The second
+phases of the calculus. The first phase computes the bidirectional closure of $\Ec.$ The second
 phase makes inferences based on acyclicity (for datatypes) and uniqueness
@@
 The third phase performs case distinctions on constructors for
-various terms occurring in $\Ec$.
+various terms occurring in $\Ec.$
 %Following the conventions from~\cite{},
@@
 %We assume that all $\Sig$-terms in $\Ec$ are \emph{normalized}, meaning that all subterms of the form $\const{s}^{\,k}_{\negvthinspace j}( \const{C}_{\negvthinspace j}(t_1,\ldots,t_n) )$
-%are simplified to $t_i$, and moreover assume that all %additional
+%are simplified to $t_i,$ and moreover assume that all %additional
 %literals added to $\Ec$ are normalized in this manner.
@@
 %and in the third step, %(when necessary),
-%we branch on constructor types for various terms in $\tEc$.
+%we branch on constructor types for various terms in $\tEc.$
 %Following the conventions from~\cite{},
 %the derivation rules of our calculus are given in \emph{guarded assignment form},
-%where a rule can be applied to $\Ec$ if it meets all of the specified premises for $\Ec$.
+%where a rule can be applied to $\Ec$ if it meets all of the specified premises for $\Ec.$
 
@@
 and failures to unify equated terms are recognized by \rn{Clash}.
-The \rn{Conflict} rule recognizes when an equality and its negation both occur in $\Ec$, in which case $\Ec$ has no model.
-
-Let $\tEc$ denote the set of $\Sig$-terms occurring in $\Ec$.
+The \rn{Conflict} rule recognizes when an equality and its negation both occur in $\Ec,$ in which case $\Ec$ has no model.
+
+Let $\tEc$ denote the set of $\Sig$-terms occurring in $\Ec.$
 At the end of the first phase, $\Ec$ induces an equivalence
 relation over $\tEc$ such that two terms $t$ and $u$ are equivalent if and
-only if $t \teq u \in \Ec$.
+only if $t \teq u \in \Ec.$
 Thus, we can regard $\Ec$ as a set of
-equivalence classes of terms. For a term $t \in \tEc$, we write $\ec{t}$ to
-denote the equivalence class of $t$ in $\Ec$.
+equivalence classes of terms. For a term $t \in \tEc,$ we write $\ec{t}$ to
+denote the equivalence class of $t$ in $\Ec.$
 Moreover, at the end of this phase, each equivalence class $\ec{t}$ contains
@@
 the binder can be omitted.
-$\FV( t )$ denotes the set of free variables occurring in the $\mu$-term~$t$.
+$\FV( t )$ denotes the set of free variables occurring in the $\mu$-term~$t.$
 A $\mu$-term is \emph{closed} if %and only if --- by convention, "and only if" is not necessary for *definitions* (but add it back if you disagree)
@@
 %$\MU x.\; \const{C}( x ) \aequiv \MU y.\; \const{C}( y )$ and
-$\MU x.\; \const{D}( y, x ) \aequiv \MU z.\; \const{D}( y, z )$,
+$\MU x.\; \const{D}( y, x ) \aequiv \MU z.\; \const{D}( y, z ),$
 but
-$\MU x.\; \const{C}( x )$, $\MU x.\; \const{D}( y, x )$, $\MU x.\; \const{D}( z, x )$, and
+$\MU x.\; \const{C}( x ),$ $\MU x.\; \const{D}( y, x ),$ $\MU x.\; \const{D}( z, x ),$ and
 $\MU y.\; \const{D}( y, x )$ are all $\alpha$-disequivalent.
 Two $\mu$-terms can denote the same value despite being $\alpha$-disequivalent---e.g.,
-$\MU x.\; \const{S}( x ) \not\aequiv \MU y.\; \const{S}( \const{S}( y ) )$.
+$\MU x.\; \const{S}( x ) \not\aequiv \MU y.\; \const{S}( \const{S}( y ) ).$
 %In the \rn{Acyclic} rule,
@@
 %(Section~\ref{sec:introduction}).
-% for example, $\MU x.\; \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
+% for example, $\MU x.\; \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots )).$
 %either in the case when $\tau$ is a codatatype or datatype type,
-%$\tau \notin \Codata$.
+%$\tau \notin \Codata.$
 
@@
 %For uniformity, we also consider $%\MU x.\;
-%\const{c}$ to be a $\mu$-term if $\const{c}$ is a constant of type $\tau \in \Nondata$.
+%\const{c}$ to be a $\mu$-term if $\const{c}$ is a constant of type $\tau \in \Nondata.$
 %
@@
 %if $\const{C} : \vec \tau \rightarrow \upsilon$
-%and $\vec v$ are well-typed $\mu$-terms of type $\vec \tau$ under the assumption that $x$ has type $\upsilon$,
-%then $\MU x.\; \const{C}( \vec v )$ is a well-typed term of type $\upsilon$.
+%and $\vec v$ are well-typed $\mu$-terms of type $\vec \tau$ under the assumption that $x$ has type $\upsilon,$
+%then $\MU x.\; \const{C}( \vec v )$ is a well-typed term of type $\upsilon.$
 %
@@
 The $\mu$-term $\Val[t\typ{\tau}]$ describes a class of $\tau$ values
-that $t$ and other members of $t$'s equivalence class can take in models of $\Ec$.
-When $\tau$ is a datatype, % \in \Data$,
+that $t$ and other members of $t$'s equivalence class can take in models of $\Ec.$
+When $\tau$ is a datatype, % \in \Data,$
 a cyclic $\mu$-term describes an infeasible class of values.
@@
 %
-With each equivalence class $\ec{u}$, we associate a fresh variable $\Varec{u}$
-of the same type as $u$. %not occurring in $\tEc$,
-For a term $t \in \tEc$,
+With each equivalence class $\ec{u},$ we associate a fresh variable $\Varec{u}$
+of the same type as $u.$ %not occurring in $\tEc,$
+For a term $t \in \tEc,$
 we write $\Varec{t}$ to denote the variable associated with
-the equivalence class $\ec{t}$.
-Initially, we set $\Val\ec{u} := {\Varec{u}}$ for each equivalence class $\ec{u}$. Because $\Varec{u}$ is
+the equivalence class $\ec{t}.$
+Initially, we set $\Val\ec{u} := {\Varec{u}}$ for each equivalence class $\ec{u}.$ Because $\Varec{u}$ is
 unconstrained,
 this indicates that there are initially no
-constraints on the values for any equivalence class $\ec{u}$.
+constraints on the values for any equivalence class $\ec{u}.$
 The mapping $\Val$
@@
 result of the unfolding is bounded by the number of equivalence classes of
-$\Ec$, and thus the construction of $\Val$ will terminate.
+$\Ec,$ and thus the construction of $\Val$ will terminate.
 
@@
 \begin{examplex}
-Suppose that $\Ec$ contains four distinct equivalence classes $\ec{w}$, $\ec{x}$, $\ec{y}$, and~$\ec{z}$
-such that $\const{C}(w,y) \in \ec{x}$ and $\const{C}(z,x) \in \ec{y}$ for some $\const{C} \in \Ctr$.
+Suppose that $\Ec$ contains four distinct equivalence classes $\ec{w},$ $\ec{x},$ $\ec{y},$ and~$\ec{z}$
+such that $\const{C}(w,y) \in \ec{x}$ and $\const{C}(z,x) \in \ec{y}$ for some $\const{C} \in \Ctr.$
 A possible sequence of unfolding steps is given below, omitting
-trivial entries such as $\ec{w} \mapsto {\Varec{w}}$.
+trivial entries such as $\ec{w} \mapsto {\Varec{w}}.$
 %
@@
 must be of the forms $\const{C}( {\Varec{w}}, \const{C}( {\Varec{z}},\allowbreak \const{C}( {\Varec{w}},\allowbreak \const{C}( {\Varec{z}},\ldots ))))$
-and $\const{C}( {\Varec{z}}, \const{C}( {\Varec{w}}, \const{C}( {\Varec{z}}, \const{C}( {\Varec{w}}, \ldots ))))$,
-respectively. %, for some values of ${\Varec{w}}$ and ${\Varec{z}}$.
+and $\const{C}( {\Varec{z}}, \const{C}( {\Varec{w}}, \const{C}( {\Varec{z}}, \const{C}( {\Varec{w}}, \ldots )))),$
+respectively. %, for some values of ${\Varec{w}}$ and ${\Varec{z}}.$
 \xend
@@
 
-Given the mapping $\Val$, the \rn{Acyclic} and \rn{Unique} rules work as follows.
+Given the mapping $\Val,$ the \rn{Acyclic} and \rn{Unique} rules work as follows.
 For acyclicity, if $\ec{t}$ is a datatype equivalence class
@@
 then $\Ec$ is $\thD$-unsatisfiable.
-For uniqueness, if $\ec{t}$, $\ec{u}$ are two codatatype equivalence classes
-whose values $\Val \ec{t}$, $\Val \ec{u}$ are $\alpha$-equivalent,
+For uniqueness, if $\ec{t},$ $\ec{u}$ are two codatatype equivalence classes
+whose values $\Val \ec{t},$ $\Val \ec{u}$ are $\alpha$-equivalent,
 %for a capture-avoiding renaming of $\mu$-bound variables,
-then $t \teq u$. Comparison for $\alpha$-equivalence may seem too restrictive, since
+then $t \teq u.$ Comparison for $\alpha$-equivalence may seem too restrictive, since
 $\MU x.\; \const{S}(x)$ and $\MU y.\; \const{S}(\const{S}(y))$ specify the same
@@
 \begin{examplex}
-Let $\Ec = \{ x \teq \const{S}(x),\; y \teq \const{S}(\const{S}(y)) \}$.
+Let $\Ec = \{ x \teq \const{S}(x),\; y \teq \const{S}(\const{S}(y)) \}.$
 After phase~1, the equivalence classes %of $\Ec$
 are
-$\{ x,\, \const{S}(x) \}$,
-$\{ y,\, \const{S}(\const{S}(y)) \}$,
-and $\{ \const{S}(y) \}$.
+$\{ x,\, \const{S}(x) \},$
+$\{ y,\, \const{S}(\const{S}(y)) \},$
+and $\{ \const{S}(y) \}.$
 Constructing $\Val$ yields
@@
 \end{align*}
-Since $\Val \ec{y} \aequiv \Val \ec{\const{S}(y)}$,
+Since $\Val \ec{y} \aequiv \Val \ec{\const{S}(y)},$
 %for the renaming
-%$\{ {\Varec{y}} \mapsto {\VAREC{\const{S}(y)}}, {\VAREC{\const{S}(y)}} \mapsto {\Varec{y}} \}$,
-the \rn{Unique} rule applies to derive $y \teq \const{S}(y)$.
+%$\{ {\Varec{y}} \mapsto {\VAREC{\const{S}(y)}}, {\VAREC{\const{S}(y)}} \mapsto {\Varec{y}} \},$
+the \rn{Unique} rule applies to derive $y \teq \const{S}(y).$
 %Intuitively, this equality holds, since the values of $y$ and $\const{S}(y)$
@@
 $\{ x,\, \const{S}(x) \}$ and
-$\{ y,\, \const{S}(y),\allowbreak\, \const{S}(\const{S}(y)) \}$.
+$\{ y,\, \const{S}(y),\allowbreak\, \const{S}(\const{S}(y)) \}.$
 The mapping $\Val$ is updated accordingly:
@@
 \end{align*}
-Since $\Val \ec{x} \aequiv \Val \ec{y}$,
+Since $\Val \ec{x} \aequiv \Val \ec{y},$
 %for the renaming
-%$\{ {\Varec{y}} \mapsto {\VAREC{\const{S}(y)}}, {\VAREC{\const{S}(y)}} \mapsto {\Varec{y}} \}$,
-\rn{Unique} can finally be applied to derive $x \teq y$.
+%$\{ {\Varec{y}} \mapsto {\VAREC{\const{S}(y)}}, {\VAREC{\const{S}(y)}} \mapsto {\Varec{y}} \},$
+\rn{Unique} can finally be applied to derive $x \teq y.$
 \xend
@@
 \paragraph{Phase 3: Branching \rm(Fig.~\ref{fig:split-rule}).}
-If a selector is applied to a term $t$, or if $t$'s type is a finite datatype,
+If a selector is applied to a term $t,$ or if $t$'s type is a finite datatype,
 $t$'s equivalence class must contain a
@@
 If there exists a closed derivation tree with root node
-$\Ec$, then $\Ec$ is $\thD$-unsatisfiable; and if there exists
+$\Ec,$ then $\Ec$ is $\thD$-unsatisfiable; and if there exists
 a derivation tree with root node $\Ec$ that contains a saturated node, then
@@
 \begin{proof}
-Consider a derivation tree with root node $\Ec$. Let $D \subseteq \tEc$ be the
+Consider a derivation tree with root node $\Ec.$ Let $D \subseteq \tEc$ be the
 set of terms whose types are finite datatypes, and let $\SSS \subseteq \tEc$ be
@@
 \end{align*}
-%for $i \geq 0$, let ;
+%for $i \geq 0,$ let ;
 and let $\SSSS \infty_t$ be the limit of this sequence.
-This is a finite set for each $t$,
+This is a finite set for each $t,$
 because all chains of selectors applied to $t$ are finite.
@@
 %%% TODO: Clarify above step in report.
-Let $\SSSS \infty$ be the union of all sets $\SSSS \infty_t$ where $t \in D$,
+Let $\SSSS \infty$ be the union of all sets $\SSSS \infty_t$ where $t \in D,$
 and let $\tcEc$ be the set of subterms of
 %
-$\Ec \mathrel\cup \{ \smash{\const C_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\smash{\,n_{\negvthinspace j}}}_{\negvthinspace j}( t ) \bigr)} \mid \smash{t\typ{\delta} \in \SSS \mathrel\cup \SSSS \infty{,}~ \const C_{\negvthinspace j} \in \smash{\Ctr^\delta}} \}$.
-%
-In a derivation tree with root node~$\Ec$,
+$\Ec \mathrel\cup \{ \smash{\const C_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\smash{\,n_{\negvthinspace j}}}_{\negvthinspace j}( t ) \bigr)} \mid \smash{t\typ{\delta} \in \SSS \mathrel\cup \SSSS \infty{,}~ \const C_{\negvthinspace j} \in \smash{\Ctr^\delta}} \}.$
+%
+In a derivation tree with root node~$\Ec,$
 it can be shown by induction
 on the rules of the calculus that each non-root node $\Fc$ is such that
-$\Terms(\Fc) \subseteq \tcEc$, and hence contains an equality between two terms from $\tcEc$ not occurring in its parent node.
-Thus, the depth of a branch in a derivation tree with root node $\Ec$ is at most $\left| \smash{\tcEc}\vphantom{X} \right|{\!}^2$,
+$\Terms(\Fc) \subseteq \tcEc,$ and hence contains an equality between two terms from $\tcEc$ not occurring in its parent node.
+Thus, the depth of a branch in a derivation tree with root node $\Ec$ is at most $\left| \smash{\tcEc}\vphantom{X} \right|{\!}^2,$
 which is finite since $\tcEc$ is finite.
@@
 \afterDot
-If there exists a closed derivation tree with root node\/ $\Ec$, then\/ $\Ec$ is $\thD$-unsatisfiable.
+If there exists a closed derivation tree with root node\/ $\Ec,$ then\/ $\Ec$ is $\thD$-unsatisfiable.
 \end{theorem}
 \begin{proof}
-The proof is by structural induction on the derivation tree with root node $\Ec$.
+The proof is by structural induction on the derivation tree with root node $\Ec.$
 If the tree is an application of \rn{Conflict}, \rn{Clash}, or \rn{Acyclic},
@@
 If the child nodes of $\Ec$ are closed derivation trees
-whose root nodes are the result of applying \rn{Split} on $t\typ{\delta}$,
+whose root nodes are the result of applying \rn{Split} on $t\typ{\delta},$
 by the induction hypothesis $\Ec \mathrel\cup t \teq \smash{\const C_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\,\smash{n_{\negvthinspace j}}}_{\negvthinspace j}( t ) \bigr)}$ is
 $\thD$-unsatisfiable
-for each $\const C_{\negvthinspace j} \in \smash{\Ctr^\delta}$.
+for each $\const C_{\negvthinspace j} \in \smash{\Ctr^\delta}.$
 Since by exhaustiveness, all models of $\thD$ entail exactly one
-$\DISC{\negvthinspace j}$,
+$\DISC{\negvthinspace j},$
 $\Ec$ is $\thD$-unsatisfiable.
@@
 node $\Ec \mathrel\cup t \teq u$ is obtained by applying one of the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, \rn{Inject}, \rn{Unique}, or \rn{Single}.
-In each of these cases, $\Ec \vDash_{\thD} t \teq u$.
+In each of these cases, $\Ec \vDash_{\thD} t \teq u.$
 For \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, this is a consequence of equality reasoning.
 For \rn{Inject}, this is a consequence of injectivity.
-For \rn{Unique}, the construction of $\Val$ indicates that the values of $t$ and $u$ are equivalent in all models of $\Ec$.
+For \rn{Unique}, the construction of $\Val$ indicates that the values of $t$ and $u$ are equivalent in all models of $\Ec.$
 For \rn{Single}, $t$~and~$u$ must have the same value since the cardinality of their type is one.
@@
 $\thD$-satisfiable.
-The proof relies on a specific interpretation $\J$ that satisfies $\Ec$.
-
-First, we define the set of interpretations of the theory $\thD$,
+The proof relies on a specific interpretation $\J$ that satisfies $\Ec.$
+
+First, we define the set of interpretations of the theory $\thD,$
 which requires custom terminology concerning $\mu$-terms.
-Given a $\mu$-term $t$ with subterm $u$,
-the \emph{expansion of $u$ with respect to $t$} is the $\mu$-term $\expand{t}{u}^\emptyset$,
+Given a $\mu$-term $t$ with subterm $u,$
+the \emph{expansion of $u$ with respect to $t$} is the $\mu$-term $\expand{t}{u}^\emptyset,$
 abbreviated to
-$\expand{t}{u}$, as returned by the function
+$\expand{t}{u},$ as returned by the function
 %where $x \leadsto \MU x.\; \const{C}( \vec u ) \in t$ indicates that $\MU x.\;
@@
 $\expand{\MU x.\: \const{C}(\MU y.\: \const{D}(x))}{\MU y.\: \const{D}(x)} =
- \MU y.\: \const{D}(\MU x.\: \const{C}(y))$.
+ \MU y.\: \const{D}(\MU x.\: \const{C}(y)).$
 %
@@
 is a \emph{self-similar subterm} of~$t$ if
-$u$ is a proper subterm of~$t$,
+$u$ is a proper subterm of~$t,$
 $t$ and $u$ are of the forms $\MU x.\; \const{C}( t_1, \ldots, t_n )$
-and $\MU y.\; \const{C}( u_1, \ldots, u_n )$,
-and $\expand{t}{ t_k } \aequiv \expand{t}{ u_k }$ for all $k$. % \in [n]
+and $\MU y.\; \const{C}( u_1, \ldots, u_n ),$
+and $\expand{t}{ t_k } \aequiv \expand{t}{ u_k }$ for all $k.$ % \in [n]
 The $\mu$-term $t$ is \emph{normal} if it does not contain %any
@@
 Thus, $t = \MU x.\; \const{C}( \MU y.\; \const{C}( y ) )$ is not normal
-because $\MU y.\; \const{C}( y )$ is a self-similar subterm of~$t$.
+because $\MU y.\; \const{C}( y )$ is a self-similar subterm of~$t.$
 Their arguments have the same expansion with respect to~$t$:
@@
 \MU y.\; \const{C}\bigl( \expand{t}{y}^{\smash{\{y\}}} \bigr) =
-\MU y.\; \const{C}( y )$.
+\MU y.\; \const{C}( y ).$
 The term $u = \MU x.\; \const{C}( \MU y.\; \const{C}( x ) )$ is also not normal,
-since $\MU y.\; \const{C}( x )$ is a self-similar subterm of $u$,
+since $\MU y.\; \const{C}( x )$ is a self-similar subterm of $u,$
 noting that
@@
  = \MU y.\allowbreak\; \const{C}( \MU x.\allowbreak\; \const{C}( y ) )$
-is $\alpha$-equivalent to $\expand{u}{x} = u$.
-% for the renaming $\{ x \mapsto y, y \mapsto x \}$.
-
-For any $\mu$-term $t$ of the form $\MU x.\; \const C( \vec u )$,
+is $\alpha$-equivalent to $\expand{u}{x} = u.$
+% for the renaming $\{ x \mapsto y, y \mapsto x \}.$
+
+For any $\mu$-term $t$ of the form $\MU x.\; \const C( \vec u ),$
 its \emph{normal form} $\nf{t}$ is obtained
@@
 and by recursively normalizing the other subterms.
-For variables, $\nf{x} = x$.
-Thus, $\nf{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) )} = \MU x.\; \const{C}( x )$.
+For variables, $\nf{x} = x.$
+Thus, $\nf{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) )} = \MU x.\; \const{C}( x ).$
 
@@
 %If the $\mu$-terms $\vec u$ are in normal form and $t = \nf{\MU x.\; \const{C}( \vec u )} = \MU x.\; \const{C}( \vec w )$
-%for some $\mu$-terms $\vec w$,
+%for some $\mu$-terms $\vec w,$
 %then $\vec u \vsim \expand{t}{\vec w}$\vthinspace.
@@
 %\label{lem:mu-norm-interp}
-%If the $\mu$-term $t$ is normal and $u$ is a subterm of $t$, then $\expand{t}{u}$ is normal.
+%If the $\mu$-term $t$ is normal and $u$ is a subterm of $t,$ then $\expand{t}{u}$ is normal.
 %\end{lemma}
@@
 %If $t_u = \MU x.\; \const{C}( \vec u )$ and $t_w = \MU y.\; \const{C}( \vec w )$ are in normal form,
-%then $\expand{t_u}{ \vec u } \vsim \expand{t_w}{ \vec w }$ if and only if $t_u \vsim t_w$.
+%then $\expand{t_u}{ \vec u } \vsim \expand{t_w}{ \vec w }$ if and only if $t_u \vsim t_w.$
 %\end{lemma}
@@
 
-We now define the class of interpretations for $\thD$.
+We now define the class of interpretations for $\thD.$
 $\interp{\J}{\tau}$ denotes the interpretation type $\tau$ in $\J$---that is, a
 nonempty set of domain elements for that type.
-$\interp{\J}{\const{f}}$ denotes the interpretation of a function $\const{f}$ in $\J$.
-If $\const{f} : \tau_1 \times \cdots \times \tau_n \to \tau$,
-then $\interp{\J}{\const{f}}$ is a total function from $\interp{\J}{\tau_1} \times \cdots \times \interp{\J}{\tau_n}$ to $\interp{\J}{\tau}$.
+$\interp{\J}{\const{f}}$ denotes the interpretation of a function $\const{f}$ in $\J.$
+If $\const{f} : \tau_1 \times \cdots \times \tau_n \to \tau,$
+then $\interp{\J}{\const{f}}$ is a total function from $\interp{\J}{\tau_1} \times \cdots \times \interp{\J}{\tau_n}$ to $\interp{\J}{\tau}.$
 %To simplify the presentation,
@@
 %For example, the $\mu$-term $\MU x.\; \const{Z}$ representing the integer constant zero
-%is simply $\const{Z}$ after dropping the $\mu$-binding of $x$.
+%is simply $\const{Z}$ after dropping the $\mu$-binding of $x.$
 
@@
 \item
-For each type $\tau$,
+For each type $\tau,$
 %%% sets *contain* elements, they *include* other sets
 $\interp{\J}{\tau}$ includes a maximal set of closed normal $\mu$-terms of that type that are
-unique up to $\alpha$-equivalence and acyclic unless $\tau \in \Codata$.\strut
+unique up to $\alpha$-equivalence and acyclic unless $\tau \in \Codata.$\strut
 
 \item
-For each constructor term $\vphantom{s^k}\const{C}( \vec t \,)$ of type $\tau$,
+For each constructor term $\vphantom{s^k}\const{C}( \vec t \,)$ of type $\tau,$
 $\interp{\J}{\const{C}}( \interp{\J}{\vec t\,} )$ is the value
 in $\interp{\J}{\tau}$ that is $\alpha$-equivalent to
-$\nf{\MU x.\; \const{C}( \interp{\J}{\vec t\,} )}$, where $x$ is fresh.\strut
+$\nf{\MU x.\; \const{C}( \interp{\J}{\vec t\,} )},$ where $x$ is fresh.\strut
 
 \item
-For each selector term $\const s^k_{\negvthinspace j}( t )$ of type $\tau$, % for some $\vec u$,
-if $\interp{\J}{t}$ is $\MU x.\; \const{C}_{\negvthinspace j}( \vec u )$,
+For each selector term $\const s^k_{\negvthinspace j}( t )$ of type $\tau,$ % for some $\vec u,$
+if $\interp{\J}{t}$ is $\MU x.\; \const{C}_{\negvthinspace j}( \vec u ),$
 then $\J\negvvthinspace( \const s^k_{\negvthinspace j} )\bigl( \interp{\J}{t} \bigr)$ is the value
 in $\interp{\J}{\tau}$ that is $\alpha$-equivalent to
-$\expand{\interp{\J}{t}}{u_k}$.
+$\expand{\interp{\J}{t}}{u_k}.$
 \end{enumerate}
@@
 
-When constructing a model $\J$ of $\Ec$,
+When constructing a model $\J$ of $\Ec,$
 it remains only to specify how $\J$ interprets wrongly applied selector terms and variables.
@@
 We write $t \vsimv{x} u$ if $\mu$-terms $t$ and $u$ are syntactically equivalent
-for some renaming that avoids capturing any variable other than~$x$.
+for some renaming that avoids capturing any variable other than~$x.$
 For example,
@@
 (by renaming $y$ to $x$),
-$\MU x.\; \const{C}( x, x ) \vsimv{x} \MU y.\; \const{C}( x, y )$, and
-$\MU x.\; \const{C}( z, x ) \vsimv{z} \MU y.\; \const{C}( z, y )$,
+$\MU x.\; \const{C}( x, x ) \vsimv{x} \MU y.\; \const{C}( x, y ),$ and
+$\MU x.\; \const{C}( z, x ) \vsimv{z} \MU y.\; \const{C}( z, y ),$
 but
 $\MU x.\; \const{D}( x ) \not\vsimv{x} \MU x.\; \const{D}( y )$ and
-$\MU x.\; \const{C}( x, x ) \not\vsimv{y} \MU y.\; \const{C}( x, y )$.
-For a variable $x\typ{\tau}$ and a normal interpretation $\J$,
+$\MU x.\; \const{C}( x, x ) \not\vsimv{y} \MU y.\; \const{C}( x, y ).$
+For a variable $x\typ{\tau}$ and a normal interpretation $\J,$
 we let $\BAD_{\smash{\J}}^x( \Val )$ denote the set consisting of all values
@@
 such that $v \vsimv{x} \expand{t}{u}$ for some subterm $u$ of a term $t$
-occurring in the range of $\Val$. This set describes shapes of terms to avoid
-when assigning a $\mu$-term to $x$.
+occurring in the range of $\Val.$ This set describes shapes of terms to avoid
+when assigning a $\mu$-term to $x.$
 
@@
 The \emph{completion} $\ValC$ of $\Val$ for a normal interpretation $\J$
-assigns values from $\J$ to unassigned variables in the domain of $\Val$.
+assigns values from $\J$ to unassigned variables in the domain of $\Val.$
 We construct $\ValC$ by initially setting $\ValC := \nf{\Val}$
@@
 %
-Given an unassigned variable in $\ValC$, this rule assigns it a fresh
+Given an unassigned variable in $\ValC,$ this rule assigns it a fresh
 value---one that does not occur in $\BAD_{\smash{\J}}^{\Varec{x}}(\ValC)$
@@
 variable from $\FV( \ValC )$ and does not add any variables to $\FV( \ValC
-)$, the process eventually terminates. We normalize all terms in
+),$ the process eventually terminates. We normalize all terms in
 the range of $\ValC$ at each step.
@@
 %The last premise is admittedly rather technical.
-To ensure disequality literals are satisfied by an interpretation based on $\ValC$,
+To ensure disequality literals are satisfied by an interpretation based on $\ValC,$
 it suffices that $\ValC$ is injective modulo $\alpha$-equivalence.
@@
 The set $\BAD_{\smash{\J}}^{\Varec{x}}(\ValC)$ is an overapproximation of the values
-that, when assigned to ${\Varec{x}}$, will cause values in the range of $\ValC$ to become $\alpha$-equivalent.
+that, when assigned to ${\Varec{x}},$ will cause values in the range of $\ValC$ to become $\alpha$-equivalent.
 %This premise is not strictly necessary in practice,
@@
 %Injectivity is important to satisfy disequality literals.
-%For infinite codatatypes, it is always possible to find suitable values $v$.
+%For infinite codatatypes, it is always possible to find suitable values $v.$
 %For example,
-%if the longest cycle in $\BAD_{\smash{\J}}^{\Varec{x}}(\ValC)$ is of length $n$, a fresh value
-%can be constructed with a cycle of length $n + 1$.
+%if the longest cycle in $\BAD_{\smash{\J}}^{\Varec{x}}(\ValC)$ is of length $n,$ a fresh value
+%can be constructed with a cycle of length $n + 1.$
 
@@
 \label{ex:assign}
-Let $\delta$ be a codatatype with the constructors $\const{C},\, \const{D},\, \const{E} :\delta \rightarrow \delta$.
+Let $\delta$ be a codatatype with the constructors $\const{C},\, \const{D},\, \const{E} :\delta \rightarrow \delta.$
 Let $\Ec$ be the set
@@
 z \not\teq v
-\}$.
-After applying the calculus to saturation on $\Ec$,
+\}.$
+After applying the calculus to saturation on $\Ec,$
 the mapping $\Val$ is as follows:
@@
 %
-Here, $\ec{u}$ denotes the equivalence class $\{ u, \const{C}( z ) \}$,
-$\ec{v}$ denotes $\{ v, \const{C}( z ), s \}$, and so on.
-To construct a completion $\ValC$, we must choose values for ${\Varec{y}}$ and ${\Varec{z}}$,
-which are free in~$\Val$.
+Here, $\ec{u}$ denotes the equivalence class $\{ u, \const{C}( z ) \},$
+$\ec{v}$ denotes $\{ v, \const{C}( z ), s \},$ and so on.
+To construct a completion $\ValC,$ we must choose values for ${\Varec{y}}$ and ${\Varec{z}},$
+which are free in~$\Val.$
 Modulo $\alpha$-equivalence, $\smash{\BAD_{\smash{\J}}^{{\Varec{z}}}( \Val )} =
@@
 %\end{array}
-$.
+.$
 Now consider a normal interpretation $\J$ that evaluates variables in $\Ec$ based on~$\Val$:
-$\interp{\J}{u} = \Val \ec{u}$, $\interp{\J}{v} = \Val \ec{v}$, and so on.
+$\interp{\J}{u} = \Val \ec{u},$ $\interp{\J}{v} = \Val \ec{v},$ and so on.
 Assigning a value for $\Val \ec{z}$ that is $\alpha$-equivalent to a value in $\BAD_{\smash{\J}}^{{\Varec{z}}}( \Val )$
 may cause values in the range of $\Val$ to become $\alpha$-equivalent,
-which in turn may cause $\Ec$ to be falsified by~$\J$.
-For example, assign $\MU {\Varec{z}}.\; \const{D}( {\Varec{z}} )$ for ${\Varec{z}}$.
-After the substitution, $\Val \ec{v} = \MU {\Varec{v}}.\; \const{D}( \MU {\Varec{z}}.\; \const{D}( {\Varec{z}} ) )$,
-which has normal form $\MU {\Varec{v}}.\; \const{D}( {\Varec{v}} )$,
-which is $\alpha$-equivalent to $\MU {\Varec{z}}.\; \const{D}( {\Varec{z}} )$.
-However, this contradicts the disequality $z \not\teq v$ in~$\Ec$.
-On the other hand, if the value assigned to ${\Varec{z}}$ is fresh, %not $\alpha$-equivalent to any term in $\BAD_{\smash{\J}}^{{\Varec{z}}}( \Val )$,
+which in turn may cause $\Ec$ to be falsified by~$\J.$
+For example, assign $\MU {\Varec{z}}.\; \const{D}( {\Varec{z}} )$ for ${\Varec{z}}.$
+After the substitution, $\Val \ec{v} = \MU {\Varec{v}}.\; \const{D}( \MU {\Varec{z}}.\; \const{D}( {\Varec{z}} ) ),$
+which has normal form $\MU {\Varec{v}}.\; \const{D}( {\Varec{v}} ),$
+which is $\alpha$-equivalent to $\MU {\Varec{z}}.\; \const{D}( {\Varec{z}} ).$
+However, this contradicts the disequality $z \not\teq v$ in~$\Ec.$
+On the other hand, if the value assigned to ${\Varec{z}}$ is fresh, %not $\alpha$-equivalent to any term in $\BAD_{\smash{\J}}^{{\Varec{z}}}( \Val ),$
 the values in the range of $\Val$ remain $\alpha$-disequivalent.
 We can assign a value such as
-$\MU {\Varec{z}}.\; \const{E}( {\Varec{z}} )$,
-$\MU {\Varec{z}}.\; \const{D}( \const{C}( {\Varec{z}} ) )$, or
+$\MU {\Varec{z}}.\; \const{E}( {\Varec{z}} ),$
+$\MU {\Varec{z}}.\; \const{D}( \const{C}( {\Varec{z}} ) ),$ or
 $\MU {\Varec{z}}.\; \const{C}( \const{C}( \const{D}( {\Varec{z}} ) ) )$
-to ${\Varec{z}}$.
+to ${\Varec{z}}.$
 
 Legal substitutions for ${\Varec{z}}$ may cause the range of $\Val$ to contain abnormal terms.
-For example, after assigning $\MU {\Varec{z}}.\; \const{D}( \const{C}( {\Varec{z}} ) )$ to $\Varec{z}$, we have
-$\Val \ec{u} = \MU {\Varec{u}}.\; \const{C}( \MU {\Varec{z}}.\; \const{D}( \const{C}( {\Varec{z}} ) ) )$,
+For example, after assigning $\MU {\Varec{z}}.\; \const{D}( \const{C}( {\Varec{z}} ) )$ to $\Varec{z},$ we have
+$\Val \ec{u} = \MU {\Varec{u}}.\; \const{C}( \MU {\Varec{z}}.\; \const{D}( \const{C}( {\Varec{z}} ) ) ),$
 with normal form
-$\MU {\Varec{u}}.\; \const{C}( \MU {\Varec{z}}.\; \const{D}( {\Varec{u}} ) )$.
+$\MU {\Varec{u}}.\; \const{C}( \MU {\Varec{z}}.\; \const{D}( {\Varec{u}} ) ).$
 This explains why the term is normalized in the rule's conclusion.
@@
 
-In the following lemma about $\ValC$,
+In the following lemma about $\ValC,$
 $\vphantom{\Bigl(_j}\muvar( t ) = \relax{\begin{cases}
   t & \text{if $t$ is a variable} \\[-\jot]
-  x & \text{if $t$ is of the form $\MU {x}.\; u$.}
+  x & \text{if $t$ is of the form $\MU {x}.\; u.$}
 \end{cases}}$
@@
 If $\Val$ is constructed for a saturated set $\Ec$
-and $\ValC$ is a completion of $\Val$ for a normal interpretation $\J$, the following
+and $\ValC$ is a completion of $\Val$ for a normal interpretation $\J,$ the following
 properties hold for all $\ec{x}$ and $\ec{y}$ in the domain of $\ValC$:
 \begin{enumerate}
-\item[\rm (1)] $\ValC \ec{x\typ{\tau}}$ is $\alpha$-equivalent to a value in $\interp{\J}{\tau}$.
+\item[\rm (1)] $\ValC \ec{x\typ{\tau}}$ is $\alpha$-equivalent to a value in $\interp{\J}{\tau}.$
 \item[\rm (2)]
 $\ValC \ec{x} = \expand{\ValC \ec{y}}{ t }$ for
-all subterms $t$ of $\ValC \ec{y}$ with $\muvar(t) = {\Varec{x}}$.
+all subterms $t$ of $\ValC \ec{y}$ with $\muvar(t) = {\Varec{x}}.$
 \item[\rm (3)]
-$\ValC \ec{x} \vsim \ValC \ec{y}$ if and only if $\ec{x} = \ec{y}$.
+$\ValC \ec{x} \vsim \ValC \ec{y}$ if and only if $\ec{x} = \ec{y}.$
 \end{enumerate}
@@
 To show (1), we first show that $\ValC$ contains no free variables.
-Assume by contradiction that $\ValC$ contains a free variable ${\Varec{y}}$ for some $\ec{y}$ of type $\tau$.
+Assume by contradiction that $\ValC$ contains a free variable ${\Varec{y}}$ for some $\ec{y}$ of type $\tau.$
 Then it must be the case that $\ec{y}$ does not contain a constructor term,
-or else ${\Varec{y}}$ would not occur as a free variable in~$\Val$.
+or else ${\Varec{y}}$ would not occur as a free variable in~$\Val.$
 Consider the case when $\tau$ is finite.
-By assumption, $\tau \notin \Nondata$.
-Since \rn{Split} does not apply to $\Ec$, we have $\tau \notin \Data$.
-If $\tau \in \Codata$, then $\tau$ is corecursive by assumption, and by Lemma~\ref{lem:corecursive-singletons},
+By assumption, $\tau \notin \Nondata.$
+Since \rn{Split} does not apply to $\Ec,$ we have $\tau \notin \Data.$
+If $\tau \in \Codata,$ then $\tau$ is corecursive by assumption, and by Lemma~\ref{lem:corecursive-singletons},
 the cardinality of $\tau$ must be one.
-Since \rn{Single} does not apply, % to $\Ec$,
-there is only one equivalence class of type $\tau$ in $\Ec$,
-and thus there are no terms in $\BAD_{\smash{\J}}^{\Varec{y}}(\ValC)$ of type $\tau$.
+Since \rn{Single} does not apply, % to $\Ec,$
+there is only one equivalence class of type $\tau$ in $\Ec,$
+and thus there are no terms in $\BAD_{\smash{\J}}^{\Varec{y}}(\ValC)$ of type $\tau.$
 This is a contradiction, since %our %model
-completion can assign the value in the domain of $\tau$ to ${\Varec{y}}$.
+completion can assign the value in the domain of $\tau$ to ${\Varec{y}}.$
 Now, consider the case when $\tau$ is infinite.
 This is also a contradiction,
-since there are only a finite number of closed terms in $\BAD_{\smash{\J}}^{\Varec{y}}(\ValC)$,
+since there are only a finite number of closed terms in $\BAD_{\smash{\J}}^{\Varec{y}}(\ValC),$
 and thus %our %model
-completion can assign a value not occurring in $\BAD_{\smash{\J}}^{\Varec{y}}(\ValC)$ to ${\Varec{y}}$.
+completion can assign a value not occurring in $\BAD_{\smash{\J}}^{\Varec{y}}(\ValC)$ to ${\Varec{y}}.$
 By construction, $\ValC \ec{x}$ is normal.
-Since \rn{Acyclic} does not apply, $\Val \ec{x}$ is acyclic when $\tau \in \Data$.
+Since \rn{Acyclic} does not apply, $\Val \ec{x}$ is acyclic when $\tau \in \Data.$
 Moreover, the construction of $\ValC$ applies substitutions of the form
-$\{ {\Varec{y}} \mapsto t \}$, where $t$ is acyclic when the type of ${\Varec{y}}$ is not a codatatype.
-Thus, $\ValC \ec{x}$ is acyclic when $\tau \notin \Codata$.
-Therefore, by definition, $\ValC \ec{x}$ is $\alpha$-equivalent to a value in $\interp{\J}{\tau}$.
-
-We now show that (2) and (3) hold initially for $\Val$.
-For all equivalence classes $\ec{z}$,
+$\{ {\Varec{y}} \mapsto t \},$ where $t$ is acyclic when the type of ${\Varec{y}}$ is not a codatatype.
+Thus, $\ValC \ec{x}$ is acyclic when $\tau \notin \Codata.$
+Therefore, by definition, $\ValC \ec{x}$ is $\alpha$-equivalent to a value in $\interp{\J}{\tau}.$
+
+We now show that (2) and (3) hold initially for $\Val.$
+For all equivalence classes $\ec{z},$
 each pair of constructor terms $\const{C}_{\negvthinspace j}( \vec t \,)$ and $\const{C}_{\negvthinspace\jPrime}( \vec u )$ in $\ec{z}$
-are such that $j = \jPrime$, since \rn{Clash} does not apply, % to $\Ec$,
-and are such that $\ec{ \vec t \,} = \ec{ \vec u }$, since \rn{Inject} does not apply. % to $\Ec$.
+are such that $j = \jPrime,$ since \rn{Clash} does not apply, % to $\Ec,$
+and are such that $\ec{ \vec t \,} = \ec{ \vec u },$ since \rn{Inject} does not apply. % to $\Ec.$
 Thus,
@@
 were uniquely of the form $\{ {\Varec{z}} \mapsto \const{C}_{\negvthinspace j}( {\Varec{t_1}}, \ldots, {\Varec{t_n}} ) \}$
-when $\ec{z}$ contains a constructor $\const{C}_{\negvthinspace j}( t_1, \ldots, t_n )$.
-Suppose $\Val \ec{y}$ has a subterm $t$ such that $\muvar(t) = {\Varec{x}}$.
+when $\ec{z}$ contains a constructor $\const{C}_{\negvthinspace j}( t_1, \ldots, t_n ).$
+Suppose $\Val \ec{y}$ has a subterm $t$ such that $\muvar(t) = {\Varec{x}}.$
 Both $\Val \ec{x}$ and the subterm $t$ of $\Val \ec{y}$ were
-constructed by applying a sequence of substitutions of the form mentioned above to ${\Varec{x}}$.
+constructed by applying a sequence of substitutions of the form mentioned above to ${\Varec{x}}.$
 Moreover, free variables ${\Varec{z}}$ in $t$ that are bound in $\Val \ec{y}$ are interpreted in
 the expansion of $\expand{\Val \ec{y}}{ t }$ as a term
-constructed by a sequence of substitutions of the form mentioned above to ${\Varec{z}}$. %\rem{TODO: more detail}
+constructed by a sequence of substitutions of the form mentioned above to ${\Varec{z}}.$ %\rem{TODO: more detail}
 Thus, we have
-$\expand{\Val \ec{y}}{ t } = \expand{\Val \ec{x}}{ \Val \ec{x} } = \Val \ec{x}$, and %thus
-(2) holds for $\Val$.
+$\expand{\Val \ec{y}}{ t } = \expand{\Val \ec{x}}{ \Val \ec{x} } = \Val \ec{x},$ and %thus
+(2) holds for $\Val.$
 Property (3) holds for $\Val$ since \rn{Unique} does not apply.
 
-We now show that $\Val = \nf{\Val}$.
+We now show that $\Val = \nf{\Val}.$
 Assume by contradiction $\Val \ec{x} \neq \nf{\Val \ec{x}}$ for some $\Val \ec{x}$ of minimal size.
-We have that $\Val \ec{x}$ is of the form $\MU {\Varec{x}}.\; \const{C}( t_1, \ldots, t_n )$.
-Due to the construction of~$\Val$,
-we know $\ec{x}$ contains a constructor $\const{C}( z_1, \ldots, z_n )$ and $\muvar( t_i ) = z_i$ for some $i$. % \in [n]$.
+We have that $\Val \ec{x}$ is of the form $\MU {\Varec{x}}.\; \const{C}( t_1, \ldots, t_n ).$
+Due to the construction of~$\Val,$
+we know $\ec{x}$ contains a constructor $\const{C}( z_1, \ldots, z_n )$ and $\muvar( t_i ) = z_i$ for some $i.$ % \in [n].$
 Since $\Val \ec{x}$ is a minimal, it
 contains a subterm of the form $\MU {\Varec{y}}.\; \const{C}( u_1, \ldots, u_n )$
-where $\expand{\Val \ec{x}}{ t_i } \aequiv \expand{\Val \ec{x} }{ u_i }$ for some $i$. % \in [n]$.
-Due to the construction of $\Val$, $\ec{y}$ contains a constructor $\const{C}( w_1, \ldots, w_n )$ and $\muvar( u_i ) = w_i$ for some $i$. % \in [n]$.
+where $\expand{\Val \ec{x}}{ t_i } \aequiv \expand{\Val \ec{x} }{ u_i }$ for some $i.$ % \in [n].$
+Due to the construction of $\Val,$ $\ec{y}$ contains a constructor $\const{C}( w_1, \ldots, w_n )$ and $\muvar( u_i ) = w_i$ for some $i.$ % \in [n].$
 Since \rn{Cong} does not apply, % to $\Ec$
-we have $\ec{w_{\negvthinspace j}}$ and $\ec{z_{\negvthinspace j}}$ are distinct for some $j$.
+we have $\ec{w_{\negvthinspace j}}$ and $\ec{z_{\negvthinspace j}}$ are distinct for some $j.$
 By (2), $\expand{\Val \ec{x}}{ t_{\negvthinspace j} } = \Val \ec{z_{\negvthinspace j}}$
-and $\expand{\Val \ec{x}}{ u_{\negvthinspace j} } = \Val \ec{w_{\negvthinspace j}}$,
-which are not $\alpha$-equivalent by (3), contradicting the fact that $\MU {\Varec{y}}.\; \const{C}( u_1, \ldots, u_n )$ is a self-similar subterm of $\Val \ec{x}$.
-Thus, $\Val = \nf{\Val}$, and (2) and~(3) hold for $\nf{\Val}$.
-
-We now show that if (2) and (3) hold for some $\Val_1$,
-they also hold for $\nf{\Val_1 \sigma}$,
-where $\sigma$ is a substitution of the form $\{ {\Varec{x}} \mapsto \MU{\Varec{x}}.\; t \}$,
-${\Varec{x}} \in FV( \Val_1 )$,
-and $\MU {\Varec{x}}.\; t$ is not $\alpha$-equivalent to a term in $\BAD_{\smash{\J}}^{\Varec{x}}(\Val_1)$.
+and $\expand{\Val \ec{x}}{ u_{\negvthinspace j} } = \Val \ec{w_{\negvthinspace j}},$
+which are not $\alpha$-equivalent by (3), contradicting the fact that $\MU {\Varec{y}}.\; \const{C}( u_1, \ldots, u_n )$ is a self-similar subterm of $\Val \ec{x}.$
+Thus, $\Val = \nf{\Val},$ and (2) and~(3) hold for $\nf{\Val}.$
+
+We now show that if (2) and (3) hold for some $\Val_1,$
+they also hold for $\nf{\Val_1 \sigma},$
+where $\sigma$ is a substitution of the form $\{ {\Varec{x}} \mapsto \MU{\Varec{x}}.\; t \},$
+${\Varec{x}} \in FV( \Val_1 ),$
+and $\MU {\Varec{x}}.\; t$ is not $\alpha$-equivalent to a term in $\BAD_{\smash{\J}}^{\Varec{x}}(\Val_1).$
 To show (2),
-by assumption of (2) on $\Val_1$, we have $\expand{\Val_1 \ec{y}}{ u } = \Val_1 \ec{x}$ for
-all subterms $u$ of $\Val_1 \ec{y}$ where $\muvar(t) = {\Varec{x}}$.
+by assumption of (2) on $\Val_1,$ we have $\expand{\Val_1 \ec{y}}{ u } = \Val_1 \ec{x}$ for
+all subterms $u$ of $\Val_1 \ec{y}$ where $\muvar(t) = {\Varec{x}}.$
 Thus, $\expand{\Val_1 \sigma \ec{y}}{ u } = \Val_1 \sigma \ec{x}$ and
-$\expand{\nf{\Val_1 \sigma} \ec{y}}{ u } = \nf{\Val_1 \sigma} \ec{x}$.
+$\expand{\nf{\Val_1 \sigma} \ec{y}}{ u } = \nf{\Val_1 \sigma} \ec{x}.$
 To show (3),
-consider two distinct equivalence classes $\ec{y}$ and $\ec{z}$,
-and assume by contradiction that $\nf{\Val_1 \sigma} \ec{y} \aequiv \nf{\Val_1 \sigma} \ec{z}$.
-Due to (3) for $\Val_1$,
+consider two distinct equivalence classes $\ec{y}$ and $\ec{z},$
+and assume by contradiction that $\nf{\Val_1 \sigma} \ec{y} \aequiv \nf{\Val_1 \sigma} \ec{z}.$
+Due to (3) for $\Val_1,$
 $\ec{y}$ and $\ec{z}$ must have (minimal) subterms
-where $t_1$ occurs in $\Val_1 \ec{y}$ the same position~$p$ as $t_2$ occurs in $\Val_1 \ec{z}$,
-and $t_1 \not\aequiv t_2$.
+where $t_1$ occurs in $\Val_1 \ec{y}$ the same position~$p$ as $t_2$ occurs in $\Val_1 \ec{z},$
+and $t_1 \not\aequiv t_2.$
 %We first show that $\nf{\Val_1 \sigma} \ec{y}$ (and for the same reasons, $\nf{\Val_1 \sigma} \ec{z}$)
-%also contains position $p$.
+%also contains position $p.$
 %If this were not the case,
@@
 %$\Val_1 \sigma \ec{y}$ contains a self-similar subterm
-%$\MU {\Varec{w}}.\; u$ that strictly contains $t_1 \sigma$.
-If $t_1$ (resp. $t_2$) is a free variable that is not ${\Varec{x}}$,
+%$\MU {\Varec{w}}.\; u$ that strictly contains $t_1 \sigma.$
+If $t_1$ (resp. $t_2$) is a free variable that is not ${\Varec{x}},$
 then $\nf{\Val_1 \sigma} \ec{y}$ (resp. $\nf{\Val_1 \sigma} \ec{z}$)
-contains $t_1$ (resp. $t_2$) at position~$p$,
+contains $t_1$ (resp. $t_2$) at position~$p,$
 and $\nf{\Val_1 \sigma} \ec{z}$ (resp. $\nf{\Val_1 \sigma} \ec{y}$) does not.
 If $t_1$ is of the form $\MU w_1.\; \const{C}( \vec t \,)$
-and $t_2$ is of the form $\MU w_2.\; \const{D}( \vec u )$,
+and $t_2$ is of the form $\MU w_2.\; \const{D}( \vec u ),$
 then the expansion %\rem{TODO: define}
 of $\nf{\Val_1 \sigma} \ec{y}$ and $\nf{\Val_1 \sigma} \ec{z}$
-are $\alpha$-disequivalent at position~$p$.
+are $\alpha$-disequivalent at position~$p.$
 Since $t_1$ and $t_2$ are minimal,
-say $t_1$ is of the form $\MU w.\; \const{C}( \vec t \,)$,
-and $t_2$ is ${\Varec{x}}$.
-Since $\sigma$ maps ${\Varec{x}}$ to a closed $\mu$-term $\MU {\Varec{x}}.\; t$,
-we have that $FV( t_1 ) \subseteq \{ {\Varec{x}} \}$,
+say $t_1$ is of the form $\MU w.\; \const{C}( \vec t \,),$
+and $t_2$ is ${\Varec{x}}.$
+Since $\sigma$ maps ${\Varec{x}}$ to a closed $\mu$-term $\MU {\Varec{x}}.\; t,$
+we have that $FV( t_1 ) \subseteq \{ {\Varec{x}} \},$
 or else the expansion of $\nf{\Val_1 \sigma} \ec{y}$ and $\nf{\Val_1 \sigma} \ec{z}$
@@
 there is a closed $\mu$-term $v \in \BAD_{\smash{\J}}^{\Varec{x}}( \Val_1 )$
-such that $v \vsimv{\Varec{x}} \expand{\Val_1 \ec{y}}{t_1}$.
-Thus, by assumption on the selection of $\MU {\Varec{x}}.\; t$,
-we have $\smash{\MU {\Varec{x}}.\; t \not\vsimv{\Varec{x}} \expand{\Val_1 \ec{y}}{t_1}}$,
+such that $v \vsimv{\Varec{x}} \expand{\Val_1 \ec{y}}{t_1}.$
+Thus, by assumption on the selection of $\MU {\Varec{x}}.\; t,$
+we have $\smash{\MU {\Varec{x}}.\; t \not\vsimv{\Varec{x}} \expand{\Val_1 \ec{y}}{t_1}},$
 which implies that the expansion of $\nf{\Val_1 \sigma} \ec{y}$ and $\nf{\Val_1 \sigma} \ec{z}$
-are $\alpha$-disequivalent at position~$p$. %\rem{TODO: more detail}
-
-Thus, by induction on the number of applications of the above rule used to obtain $\ValC$,
+are $\alpha$-disequivalent at position~$p.$ %\rem{TODO: more detail}
+
+Thus, by induction on the number of applications of the above rule used to obtain $\ValC,$
 we have that $\ValC$ satisfies (2) and (3).
@@
 Intuitively, this lemma states three properties of $\ValC$ that together ensure that
-a normal interpretation~$\J$ can be constructed that satisfies $\Ec$.
+a normal interpretation~$\J$ can be constructed that satisfies $\Ec.$
 Property~(1) states that the values in the range of $\ValC$ are
@@
 $\muvar(t) = {\Varec{x}}$ is equal to
-$\ValC \ec{x}$, independently of the context.
+$\ValC \ec{x},$ independently of the context.
 Property~(3) states that $\ValC$ is injective (modulo
@@
 Consider $\Val$ from Example~\ref{ex:assign}.
-While extending $\Val$ to its corresponding completion $\ValC$,
-we may, for instance, assign $\MU {\Varec{z}}.\; \const{E}( {\Varec{z}} )$ to $\Varec{z}$,
-and subsequently assign $\MU {\Varec{y}}.\; \const{D}( {\Varec{y}} )$ to $\Varec{y}$.
+While extending $\Val$ to its corresponding completion $\ValC,$
+we may, for instance, assign $\MU {\Varec{z}}.\; \const{E}( {\Varec{z}} )$ to $\Varec{z},$
+and subsequently assign $\MU {\Varec{y}}.\; \const{D}( {\Varec{y}} )$ to $\Varec{y}.$
 We obtain the following mapping $\ValC$ from equivalence classes to $\MU$-terms:
@@
 All terms in the range of $\ValC$ are closed and normal,
-$\ValC \ec{s} = \expand{\ValC \ec{y}}{ t }$ for all subterms $t$ where $\muvar(t) = {\Varec{s}}$ for all $s$,
+$\ValC \ec{s} = \expand{\ValC \ec{y}}{ t }$ for all subterms $t$ where $\muvar(t) = {\Varec{s}}$ for all $s,$
 and all terms in the range of $\ValC$ are pairwise $\alpha$-disequivalent.
@@
 that is $\alpha$-equivalent to one from $\interp{\J}{\delta}$ and that is uniquely associated
-with its equivalence class by $\ValC$.
+with its equivalence class by $\ValC.$
 %For instance, up to renaming,
-%$\interp{\J}{u}$ is $\const{C}( \MU {\Varec{z}}.\; \const{E}( {\Varec{z}} ) )$,
-%$\interp{\J}{v}$ and $\interp{\J}{s}$ are $\const{D}( \MU {\Varec{z}}.\; \const{E}( {\Varec{z}} ) )$, and so on.
+%$\interp{\J}{u}$ is $\const{C}( \MU {\Varec{z}}.\; \const{E}( {\Varec{z}} ) ),$
+%$\interp{\J}{v}$ and $\interp{\J}{s}$ are $\const{D}( \MU {\Varec{z}}.\; \const{E}( {\Varec{z}} ) ),$ and so on.
 \xend
@@
 \begin{proof}
-Let $\Fc$ be a saturated node in a derivation tree with root node $\Ec$.
-%We will construct a model $\J$ for a set of equalities $\Fc$ that is equivalent to $\Fc_0$, and where
+Let $\Fc$ be a saturated node in a derivation tree with root node $\Ec.$
+%We will construct a model $\J$ for a set of equalities $\Fc$ that is equivalent to $\Fc_0,$ and where
 %all equivalence classes of $\Fc$ contain at least one variable, and all selectors in $\Fc$ are applied to variables only.
 %The former comes with no loss of generality since new equalities of the form $y \teq t$ for fresh variable $y$ can be added to $\Fc$ without affecting its satisfiability,
-%The latter also comes with no loss of generality since nested applications of selectors in $\Fc_0$ can be replaced by fresh variables while adding equalities to $\Fc\!$.
+%The latter also comes with no loss of generality since nested applications of selectors in $\Fc_0$ can be replaced by fresh variables while adding equalities to $\Fc\!.$
 We consider a normal interpretation $\J$
 that interprets wrongly applied selectors based on equality information in~$\Fc$
-and that interprets the variables of $\Fc$ based on the completion $\ValC$.
-% of %the mapping $\Val$. % from phase~2.
+and that interprets the variables of $\Fc$ based on the completion $\ValC.$
+% of %the mapping $\Val.$ % from phase~2.
 For the variables, let $\interp{\J}{x\typ{\tau}}$ be the value in $\J(\tau)$ that is $\alpha$-equivalent
-with $\ValC \ec{x}$ for each variable $x \in \Terms(\Fc)$,
+with $\ValC \ec{x}$ for each variable $x \in \Terms(\Fc),$
 which by Lemma~\ref{lem:interpretation-completion}(1) is guaranteed to exist.
 
-We first show that $\J$ satisfies all equalities $t_1 \teq t_2 \in \Fc\!$. % where $t_1$ and $t_2$ have type $\tau$.
+We first show that $\J$ satisfies all equalities $t_1 \teq t_2 \in \Fc\!.$ % where $t_1$ and $t_2$ have type $\tau.$
 To achieve this, we show
 by structural induction on $t\typ\tau$
-that $\interp{\J}{t} \aequiv \ValC \ec{t}$ for all terms $t \in \Terms( \Fc )$,
+that $\interp{\J}{t} \aequiv \ValC \ec{t}$ for all terms $t \in \Terms( \Fc ),$
 which implies $\J \vDash t_1 \teq t_2$ since $\J$ is normal.
@@
 
-If $t$ is a constructor term of the form $\const{C}( u_1, \ldots, u_n )$,
-then $\interp{\J}{t}$ is $\alpha$-equivalent with $\nf{\MU x.\allowbreak\; \const{C}( \interp{\J}{u_1},\allowbreak \ldots,\allowbreak \interp{\J}{u_n} ) }$ for some fresh $x$,
-which by the induction hypothesis is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \ValC \ec{ u_1 }, \ldots, \ValC \ec{ u_n } )}$.
-Call this term $t'$.
-%By Lemma~\ref{lem:interpretation-completion}(2) and due to our construction of $\ValC$, we have $\ValC \ec{t}$ is a term
+If $t$ is a constructor term of the form $\const{C}( u_1, \ldots, u_n ),$
+then $\interp{\J}{t}$ is $\alpha$-equivalent with $\nf{\MU x.\allowbreak\; \const{C}( \interp{\J}{u_1},\allowbreak \ldots,\allowbreak \interp{\J}{u_n} ) }$ for some fresh $x,$
+which by the induction hypothesis is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \ValC \ec{ u_1 }, \ldots, \ValC \ec{ u_n } )}.$
+Call this term $t'.$
+%By Lemma~\ref{lem:interpretation-completion}(2) and due to our construction of $\ValC,$ we have $\ValC \ec{t}$ is a term
 %of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$
-%where $\expand{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$ for $i \in [n]$.
-Since \rn{Inject} and \rn{Clash} do not apply to~$\Fc\!$,
+%where $\expand{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$ for $i \in [n].$
+Since \rn{Inject} and \rn{Clash} do not apply to~$\Fc\!,$
 by the construction of $\ValC$ we have that $\ValC \ec{t}$ is a term
@@
 where
-$\muvar(w_i) = {\Varec{u_i}}$ for each~$i$.  Thus by Lemma~\ref{lem:interpretation-completion}(2),
-$\expand{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$.
-For each $i$, let $u_\iPrime$ be the $i$\vvthinspace th argument of $t'$.
+$\muvar(w_i) = {\Varec{u_i}}$ for each~$i.$  Thus by Lemma~\ref{lem:interpretation-completion}(2),
+$\expand{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}.$
+For each $i,$ let $u_\iPrime$ be the $i$\vvthinspace th argument of $t'.$
 %By Lemma~\ref{lem:mu-norm-arg},
 We have that
-$\expand{t'}{u_\iPrime} \vsim \ValC \ec{u_i}$, and
-thus $\expand{t'}{u_\iPrime} \vsim \expand{\ValC \ec{t}}{ w_i }$. % for each $i \in [n]$.
-%By Lemma~\ref{lem:mu-cong}, we have $t' \vsim t_w$.
-Thus, $\interp{\J}{t} \vsim t' \vsim \ValC \ec{t}$, and
-we have $\interp{\J}{t} \vsim \ValC \ec{t}$.
+$\expand{t'}{u_\iPrime} \vsim \ValC \ec{u_i},$ and
+thus $\expand{t'}{u_\iPrime} \vsim \expand{\ValC \ec{t}}{ w_i }.$ % for each $i \in [n].$
+%By Lemma~\ref{lem:mu-cong}, we have $t' \vsim t_w.$
+Thus, $\interp{\J}{t} \vsim t' \vsim \ValC \ec{t},$ and
+we have $\interp{\J}{t} \vsim \ValC \ec{t}.$
 
 If $t$ is a selector term %of the form
-$\const s^k_{\negvthinspace j}( u )$,
-since \rn{Split} does not apply to $\Fc\!$,
-$\ec{u}$ must contain a term of the form $\smash{\const C_{\negvthinspace\jPrime}\bigl( \const s^1_{\negvthinspace\jPrime}( u ), \ldots, \const s^{n}_{\negvthinspace\jPrime}( u ) \bigr)}$ for some $\jPrime$.
-%By Lemma~\ref{lem:interpretation-completion}(2) and due to our construction of $\ValC$, we have $\ValC \ec{u}$ is of the form
-%$\MU {\Varec{u}}.\; \const C_{\negvthinspace\jPrime}( w_1, \ldots, w_n )$,
-%where $\expand{\ValC \ec{u} }{ w_i } = \ValC \ec{ \const s^i_{\negvthinspace\jPrime}( u )}$ for $i \in [n]$.
-Since \rn{Inject} and \rn{Clash} are not applicable, % to $\Fc\!$,
+$\const s^k_{\negvthinspace j}( u ),$
+since \rn{Split} does not apply to $\Fc\!,$
+$\ec{u}$ must contain a term of the form $\smash{\const C_{\negvthinspace\jPrime}\bigl( \const s^1_{\negvthinspace\jPrime}( u ), \ldots, \const s^{n}_{\negvthinspace\jPrime}( u ) \bigr)}$ for some $\jPrime.$
+%By Lemma~\ref{lem:interpretation-completion}(2) and due to our construction of $\ValC,$ we have $\ValC \ec{u}$ is of the form
+%$\MU {\Varec{u}}.\; \const C_{\negvthinspace\jPrime}( w_1, \ldots, w_n ),$
+%where $\expand{\ValC \ec{u} }{ w_i } = \ValC \ec{ \const s^i_{\negvthinspace\jPrime}( u )}$ for $i \in [n].$
+Since \rn{Inject} and \rn{Clash} are not applicable, % to $\Fc\!,$
 by construction $\ValC\ec{u}$ must be of the form
-$\MU {\Varec{u}}.\; \const C_{\negvthinspace\jPrime}( w_1, \ldots, w_n )$,
+$\MU {\Varec{u}}.\; \const C_{\negvthinspace\jPrime}( w_1, \ldots, w_n ),$
 where
 $\muvar(w_i) = \smash{\Varec{\vphantom{I}\smash{\const s^i_{\negvthinspace\jPrime}( u )}}}$
-for each~$i$, and thus by Lemma~\ref{lem:interpretation-completion}(2),
-$\expand{\ValC \ec{u}}{ w_i } = \ValC \ec{ \smash{\const s^i_{\negvthinspace\jPrime}( u )}}$.
-If $j = \jPrime$, then $\interp{\J}{t}$ is $\alpha$-equivalent with $\expand{\ValC \ec{u}}{w_k}$, which is equal to $\ValC \ec{ \smash{ \const s^k_{\negvthinspace j}( u )}}
-= \ValC \ec{t}$.
-If $j \neq \jPrime$, since \rn{Cong} does not apply,
+for each~$i,$ and thus by Lemma~\ref{lem:interpretation-completion}(2),
+$\expand{\ValC \ec{u}}{ w_i } = \ValC \ec{ \smash{\const s^i_{\negvthinspace\jPrime}( u )}}.$
+If $j = \jPrime,$ then $\interp{\J}{t}$ is $\alpha$-equivalent with $\expand{\ValC \ec{u}}{w_k},$ which is equal to $\ValC \ec{ \smash{ \const s^k_{\negvthinspace j}( u )}}
+= \ValC \ec{t}.$
+If $j \neq \jPrime,$ since \rn{Cong} does not apply,
 any term of the form $\smash{\const s^k_{\negvthinspace j}( u' )}$ not occurring in $\ec{t}$
-is such that $[u] \not= [u']$.
-By the induction hypothesis and Lemma~\ref{lem:interpretation-completion}(3), $\interp{\J}{u} \neq \interp{\J}{u'}$ for all such $u$, $u'$.
-Thus, we may interpret $\smash{\interp{\J}{\const s^k_{\negvthinspace j}}( \interp{\J}{u} )}$ as the value in $\interp{\J}{\tau}$ that is $\alpha$-equivalent with $\ValC \ec{ t }$.
-
-We now show that all disequalities in $\Fc$ are satisfied by $\J$.
-Assume $t \tneq u \in \Fc\!$.
+is such that $[u] \not= [u'].$
+By the induction hypothesis and Lemma~\ref{lem:interpretation-completion}(3), $\interp{\J}{u} \neq \interp{\J}{u'}$ for all such $u,$ $u'.$
+Thus, we may interpret $\smash{\interp{\J}{\const s^k_{\negvthinspace j}}( \interp{\J}{u} )}$ as the value in $\interp{\J}{\tau}$ that is $\alpha$-equivalent with $\ValC \ec{ t }.$
+
+We now show that all disequalities in $\Fc$ are satisfied by $\J.$
+Assume $t \tneq u \in \Fc\!.$
 Since \rn{Conflict} does not apply, $t \teq u \notin \Fc$ and thus $\ec{t}$ and $\ec{u}$ are distinct.
-Since $\interp{\J}{t} \vsim \ValC \ec{t}$ and $\interp{\J}{u} \vsim \ValC \ec{u}$,
-by Lemma~\ref{lem:interpretation-completion}(3), $\interp{\J}{t} \neq \interp{\J}{u}$, and thus $\J \vDash t \tneq u$.
-
-Since $\Fc$ contains only equalities and disequalities, we have $\J \vDash \Fc\!$,
-and since $\Ec \subseteq \Fc\!$,
-% is a superset of $\Ec$,
-we conclude that $\J \vDash \Ec$.
+Since $\interp{\J}{t} \vsim \ValC \ec{t}$ and $\interp{\J}{u} \vsim \ValC \ec{u},$
+by Lemma~\ref{lem:interpretation-completion}(3), $\interp{\J}{t} \neq \interp{\J}{u},$ and thus $\J \vDash t \tneq u.$
+
+Since $\Fc$ contains only equalities and disequalities, we have $\J \vDash \Fc\!,$
+and since $\Ec \subseteq \Fc\!,$
+% is a superset of $\Ec,$
+we conclude that $\J \vDash \Ec.$
 \xqed
@@
 z \not\teq v
-\}$.
+\}.$
 \xend
@@
 method for constructing a model for a saturated configuration, by means of the
-mapping $\ValC$.
+mapping $\ValC.$
 
@@
 %it assigned the value $\top$ and $\neg\: a \in M$ for all atoms that it assigns
-%the value $\bot$.
+%the value $\bot.$
 %at a high level,
@@
 \begin{itemize}
-\item the purified form of $T_i$-literals occurring in $M$, where fresh
+\item the purified form of $T_i$-literals occurring in $M,$ where fresh
 variables replace terms containing symbols not belonging to~$T_i$;
 \item additional (dis)equalities between variables of types not belonging to
-$T_i$.
+$T_i.$
 \end{itemize}
@@
 either reports that a subset $C$ of $M_i$ is $T_i$-unsatisfiable, in which case
-$\neg\: C$ is added to $F\!$, adds a clause to $F\!$, or does nothing.
-When $M$ is a complete assignment for $F\!$, a theory solver can choose to do
+$\neg\: C$ is added to $F\!,$ adds a clause to $F\!,$ or does nothing.
+When $M$ is a complete assignment for $F\!,$ a theory solver can choose to do
 nothing only if $M_i$ is indeed $T_i$-satisfiable.
@@
 Assume $\Ec$ is initially the set $M_i$ described above. With each equality $t \teq u$ added
-to $\Ec$, we associate a set of equalities from $M_i$ that together entail $t
-\teq u$, which we call its \emph{explanation}.
+to $\Ec,$ we associate a set of equalities from $M_i$ that together entail $t
+\teq u,$ which we call its \emph{explanation}.
 Similarly, each $\Val \ec{x}$ is assigned an explanation---that is, a set
 of equalities from $M_i$ that entail that the values of $\ec{x}$ in models of
-$\Ec$ are of the form $\Val \ec{x}$. For example, if $x \teq \const{C}( x ) \in
-M_i$, then $x \teq \const{C}( x )$ is an %(possible)
-explanation for $\Val \ec{x} = \MU {\Varec{x}}.\; \const{C}( {\Varec{x}} )$.
+$\Ec$ are of the form $\Val \ec{x}.$ For example, if $x \teq \const{C}( x ) \in
+M_i,$ then $x \teq \const{C}( x )$ is an %(possible)
+explanation for $\Val \ec{x} = \MU {\Varec{x}}.\; \const{C}( {\Varec{x}} ).$
 If multiple rules can derive the same conclusion,
-the implementation simply keeps the first explanation it encounters.
+the solver simply keeps the first explanation it encounters.
 
 The rules of the calculus are implemented as follows. For all rules with
-conclusion $\bot$, we report the union of the explanations for all premises is
+conclusion $\bot,$ we report the union of the explanations for all premises is
 $\thD$-unsatisfiable. The implementation does not
@@
 For \rn{Split}, we add the exhaustiveness clause %of the form
-$\DISC{1} \mathrel\lor \cdots \mathrel\lor \DISC{m}$ to $F$.
+$\DISC{1} \mathrel\lor \cdots \mathrel\lor \DISC{m}$ to $F.$
 Decisions on which branch to take are thus performed externally by the SAT
@@
 phases~2 and~3 are performed only for complete satisfying
-assignments $M$.
+assignments $M.$
 
@@
 \begin{enumerate}
-\item Let $U$ be the datatype equivalence classes of $\Ec$.
+\item Let $U$ be the datatype equivalence classes of $\Ec.$
 \item Repeat until $U \neq \emptyset$:
 \begin{itemize}
-\item[\ ] For some $\ec{t} \in U$, $\rn{Traverse}( \ec{t}, U, \emptyset )$.
+\item[\ ] For some $\ec{t} \in U,$ $\rn{Traverse}( \ec{t}, U, \emptyset ).$
 \end{itemize}
@@
 \begin{enumerate}
-\item If $\ec{t} \in P$, $\rn{Acyclic}(t)$.
+\item If $\ec{t} \in P,$ $\rn{Acyclic}(t).$
 \item If $\ec{t} \in U$:
 \begin{enumerate}
-\item[2.1.] If $\const{C}( t_1, \ldots, t_k ) \in \ec{t}$, then for each $j = 1, \ldots k$:
+\item[2.1.] If $\const{C}( t_1, \ldots, t_k ) \in \ec{t},$ then for each $j = 1, \ldots k$:
 \begin{itemize}
-\item[2.1.1.] $\rn{Traverse}( \ec{t_j}, U, P \cup \{ \ec{t} \} )$.
+\item[2.1.1.] $\rn{Traverse}( \ec{t_j}, U, P \cup \{ \ec{t} \} ).$
 \end{itemize}
-\item[2.2.] $U := U \textbackslash \{ \ec{t} \}$.
+\item[2.2.] $U := U \textbackslash \{ \ec{t} \}.$
 \end{enumerate}
@@
 \item Let $\mathcalx{U}$ be a partition of the equivalence classes of $\Ec$ such that $\ec{t} =_{\mathcalx{U}} \ec{u}$ for distinct $\ec{t}, \ec{u}$ if and only if
-  $t$, $u$ have type $\delta \in \Codata$,
-  $\const{C}( t_1, \ldots, t_k ) \in \ec{t}$,
-  and $\const{C}( u_1, \ldots, u_k ) \in \ec{u}$.
+  $t,$ $u$ have type $\delta \in \Codata,$
+  $\const{C}( t_1, \ldots, t_k ) \in \ec{t},$
+  and $\const{C}( u_1, \ldots, u_k ) \in \ec{u}.$
 
@@
 \begin{enumerate}
-\item[2.1.] $\mathcalx{U}' := \emptyset$.
+\item[2.1.] $\mathcalx{U}' := \emptyset.$
 \item[2.2.] For each $U_i \in \mathcalx{U}$:
@@
 \begin{itemize}
-\item $\const{C}( t_1, \ldots, t_k ) \in \ec{t}$ and $\const{C}( u_1, \ldots, u_k ) \in \ec{u}$, and
-\item $\ec{t_j} =_{\mathcalx{U}} \ec{u_j}$ for each $j = 1, \ldots, k$.
+\item $\const{C}( t_1, \ldots, t_k ) \in \ec{t}$ and $\const{C}( u_1, \ldots, u_k ) \in \ec{u},$ and
+\item $\ec{t_j} =_{\mathcalx{U}} \ec{u_j}$ for each $j = 1, \ldots, k.$
 \end{itemize}
-\item[2.2.2.] $\mathcalx{U}' := \mathcalx{U}' \cup \mathcalx{U}_{\vthinspace i}$.
+\item[2.2.2.] $\mathcalx{U}' := \mathcalx{U}' \cup \mathcalx{U}_{\vthinspace i}.$
 \end{itemize}
-\item[2.3.] $\mathcalx{U} := \mathcalx{U}'$.
+\item[2.3.] $\mathcalx{U} := \mathcalx{U}'.$
 \end{enumerate}
-\item If distinct $\ec{t}, \ec{u} \in U$ for some $U \in \mathcalx{U}$, $\rn{Unique}(t,u)$.
+\item If distinct $\ec{t}, \ec{u} \in U$ for some $U \in \mathcalx{U},$ $\rn{Unique}(t,u).$
 \end{enumerate}
@@
 
-Before constructing a model for $F\!$, the theory solver constructs neither $\mu$-terms nor the mapping $\Val$.
+Before constructing a model for $F\!,$ the theory solver constructs neither $\mu$-terms nor the mapping $\Val.$
 Instead, it relies on the algorithms in Figs.~\ref{fig:acyclic-algo} and~\ref{fig:uniqueness-algo}
-for determining whether the rules \rn{Acyclic} and \rn{Unique} apply to the set $\Ec$.
-
-For \rn{Acyclic}, Fig.~\ref{fig:acyclic-algo} considers the set $U$ of all datatype equivalence classes of $\Ec$.
-We choose an arbitrary $\ec{t}$ in $U$ and call the recursive subprocedure $\rn{Traverse}$,
+for determining whether the rules \rn{Acyclic} and \rn{Unique} apply to the set $\Ec.$
+
+For \rn{Acyclic}, Fig.~\ref{fig:acyclic-algo} considers the set $U$ of all datatype equivalence classes of $\Ec.$
+We choose an arbitrary $\ec{t}$ in $U$ and call the recursive subprocedure $\rn{Traverse},$
 which takes as input the current equivalence class we are processing, the set of equivalence classes $U$ we have yet to process,
 and the set of equivalence classes $P$ we are currently processing.
-If a call to $\rn{Traverse}( \ec{t}, U, P )$ is such that $\ec{t} \in P$,
-we know that the rule \rn{Acyclic} applies to $t$.
-Otherwise, if we have yet to process $\ec{t}$, as indicated by the condition $\ec{t} \in U$,
-then if $\ec{t}$ contains a constructor term $\const{C}( t_1, \ldots, t_k )$,
-then we recursively call $\rn{Traverse}$ on each of $\ec{t_1}, \ldots, \ec{t_n}$.
-If this succeeds, we remove $\ec{t}$ from $U$.
+If a call to $\rn{Traverse}( \ec{t}, U, P )$ is such that $\ec{t} \in P,$
+we know that the rule \rn{Acyclic} applies to $t.$
+Otherwise, if we have yet to process $\ec{t},$ as indicated by the condition $\ec{t} \in U,$
+then if $\ec{t}$ contains a constructor term $\const{C}( t_1, \ldots, t_k ),$
+then we recursively call $\rn{Traverse}$ on each of $\ec{t_1}, \ldots, \ec{t_n}.$
+If this succeeds, we remove $\ec{t}$ from $U.$
 
@@
 such that codatatype equivalence classes having a constructor term with top symbol $\const{C}$ are placed in the same subset,
-for each constructor $\const{C}$.
-We write $\ec{t} =_{\mathcalx{U}} \ec{u}$ to denote that $\ec{t}$ and $\ec{u}$ reside in the same subset within $\mathcalx{U}$.
+for each constructor $\const{C}.$
+We write $\ec{t} =_{\mathcalx{U}} \ec{u}$ to denote that $\ec{t}$ and $\ec{u}$ reside in the same subset within $\mathcalx{U}.$
 We then refine $\mathcalx{U}$ by
@@
 if and only if they contain constructor terms $\const{C}( t_1, \ldots, t_k )$ and $\const{C}( u_1, \ldots, u_k )$
-where each $t_j, u_j$ are such that $\ec{t_j}$ and $\ec{u_j}$ reside in the same subset of $\mathcalx{U}$ for each $j=1, \ldots, k$.
+where each $t_j, u_j$ are such that $\ec{t_j}$ and $\ec{u_j}$ reside in the same subset of $\mathcalx{U}$ for each $j=1, \ldots, k.$
 Let $\mathcalx{U'}$ be the union of each of these partitions.
-We update $\mathcalx{U}$ to $\mathcalx{U'}$.
+We update $\mathcalx{U}$ to $\mathcalx{U'}.$
 This refinement is repeated until $\mathcalx{U}$ is left unchanged.
-If the resulting partition $\mathcalx{U}$ contains a subset $U$ having two distinct equivalence classes $\ec{t}$ and $\ec{u}$, then \rn{Unique} applies to $t$ and $u$.
+If the resulting partition $\mathcalx{U}$ contains a subset $U$ having two distinct equivalence classes $\ec{t}$ and $\ec{u},$ then \rn{Unique} applies to $t$ and $u.$
 This algorithm is analogous to Hopcroft's algorithm for minimizing
@@
 %To detect whether \rn{Acyclic} applies,
-%the procedure considers each equivalence class $\ec{t}$ containing a datatype constructor $\const{C}( t_1, \ldots, t_n )$.
+%the procedure considers each equivalence class $\ec{t}$ containing a datatype constructor $\const{C}( t_1, \ldots, t_n ).$
 %It visits $\ec{t_1}, \ldots, \ec{t_n}$ and all constructor arguments in these equivalence classes recursively.
-%If while doing so it returns to $\ec{t}$, it deduces that \rn{Acyclic} applies. % to $t$.
+%If while doing so it returns to $\ec{t},$ it deduces that \rn{Acyclic} applies. % to $t.$
 %To recognize when the premise of \rn{Unique} holds,
 %the procedure considers the set $\SSS$ of all codatatype equivalence classes. It
-%simultaneously visits the equivalence classes of arguments of constructor terms in each equivalence class in $\SSS$,
+%simultaneously visits the equivalence classes of arguments of constructor terms in each equivalence class in $\SSS,$
 %while partitioning $\SSS$ into $\SSS_1, \ldots, \SSS_n$ based on the top-most symbol of constructor terms in these equivalence classes
@@
 %the resulting partition contains a set $\SSS_i$ containing distinct terms $u$
-%and $v$, it deduces that \rn{Unique} applies to $u$ and~$v$.
+%and $v,$ it deduces that \rn{Unique} applies to $u$ and~$v.$
 
@@
 This approach often leads to better performance because it introduces terms
-less eagerly to $\tEc$.
+less eagerly to $\tEc.$
 Selectors are collapsed eagerly:
 If $\const s^k_{\negvthinspace j}( t ) \in \tEc$ and $t =
-\const{C}_{\negvthinspace j}( u_1, \ldots, u_n )$, the solver directly adds
-$\smash{\const s^k_{\negvthinspace j}( t ) \teq u_k}$ to $\Ec$, whereas the presented calculus
+\const{C}_{\negvthinspace j}( u_1, \ldots, u_n ),$ the solver directly adds
+$\smash{\const s^k_{\negvthinspace j}( t ) \teq u_k}$ to $\Ec,$ whereas the presented calculus
 would apply \rn{Split} and \rn{Inject} before adding this equality.
@@
 ordering, replace $\smash{\const{C}_{\negvthinspace j}( \vec t \,) \teq
-\const{C}_{\negvthinspace\jPrime}( \vec u )}$ with $\bot$ when $j \neq \jPrime$,
+\const{C}_{\negvthinspace\jPrime}( \vec u )}$ with $\bot$ when $j \neq \jPrime,$
 replace all selector terms of the form $\smash{\const{s}^{\,k}_{\negvthinspace j}(
-\const{C}_{\negvthinspace j}(t_1,\ldots,t_n) )}$ by $t_k$, and replace
+\const{C}_{\negvthinspace j}(t_1,\ldots,t_n) )}$ by $t_k,$ and replace
 occurrences of discriminators $\smash{\const{d}_{\negvthinspace j}(
 \const{C}_{\negvthinspace\jPrime}( \vec t \,) )}$ by $\top$ or $\bot$
-based on whether $j = \jPrime$.
+based on whether $j = \jPrime.$
 Finally, finite ordinary types are not converted to finite datatypes.
@@
 between multiple theories.  In particular, given an ordinary type $\tau$ of finite cardinality $n$
-belonging to theory $T_i$, the theory solver for $T_i$ must
+belonging to theory $T_i,$ the theory solver for $T_i$ must
 communicate enough equalities over shared terms to
-ensure that all other theory solvers, including the one for $\thD$,
-interpret $\tau$ as a set having cardinality at most $n$.
+ensure that all other theory solvers, including the one for $\thD,$
+interpret $\tau$ as a set having cardinality at most $n.$
 
@@
 As described in Section~\ref{sec:termination-and-correctness},
-given a saturated set $\Ec$, we %may
+given a saturated set $\Ec,$ we %may
 construct a map $\ValC$ from
@@
 Recall that our construction of $\ValC$ requires choosing values $\MU {\Varec{x}}.\; t$ for each $\Varec{x} \in \FV( \ValC )$
-such that $\MU {\Varec{x}}.\; t$ does not occur in the set $\BAD_{\smash{\J}}^{\Varec{x}}( \ValC )$.
-To choose values, % $\MU {\Varec{x}}.\; t$,
-we use a fair enumerator for each (co)datatype $\tau$,
+such that $\MU {\Varec{x}}.\; t$ does not occur in the set $\BAD_{\smash{\J}}^{\Varec{x}}( \ValC ).$
+To choose values, % $\MU {\Varec{x}}.\; t,$
+we use a fair enumerator for each (co)datatype $\tau,$
 which lists all the values
@@
 discarding those that have free variables or are not normal.
-For those that are closed and normal, we then check whether they occur in the set $\BAD_{\smash{\J}}^{\Varec{x}}( \ValC )$.
+For those that are closed and normal, we then check whether they occur in the set $\BAD_{\smash{\J}}^{\Varec{x}}( \ValC ).$
 This set does not need to be explicitly constructed.
-%We do not explicitly construct the set $\BAD_{\smash{\J}}^{\Varec{x}}( \ValC )$.
-To determine if a term $\MU {\Varec{x}}.\; t$ occurs in the set $\BAD_{\smash{\J}}^{\Varec{x}}( \ValC )$,
+%We do not explicitly construct the set $\BAD_{\smash{\J}}^{\Varec{x}}( \ValC ).$
+To determine if a term $\MU {\Varec{x}}.\; t$ occurs in the set $\BAD_{\smash{\J}}^{\Varec{x}}( \ValC ),$
 we check if it is $\alpha$-equivalent to a closed term $(\MU {\Varec{y}}.\; u) \{ \Varec{x} \mapsto z \}$ for
-some $\MU {\Varec{y}}.\; u$ in the range of $\ValC$ and variable $z$.
+some $\MU {\Varec{y}}.\; u$ in the range of $\ValC$ and variable $z.$
 This can be efficiently achieved by matching the term $\MU {\Varec{x}}.\; t$ with %the term
-$\MU {\Varec{y}}.\; u$.
+$\MU {\Varec{y}}.\; u.$
 
 %Once the mapping $\ValC$ is constructed, our model $\J$ interprets
-%$\J( x ) = \ValC \ec{x}$ for each variable $x$ occurring free in $\Ec$.
+%$\J( x ) = \ValC \ec{x}$ for each variable $x$ occurring free in $\Ec.$
 
@@
 the context of (co)datatypes. Quantified formulas can entail an upper bound on
-the cardinality of an uninterpreted type $\ty{u}$.
+the cardinality of an uninterpreted type $\ty{u}.$
 Since it assumes that uninterpreted types have infinite cardinality, the calculus
@@
 the implementation adds the clause %of the form
-$(\exists x\;y\typ{\ty{u}}.\;\, x \not\teq y ) \mathrel\vee s \teq t$ to $F\!$,
+$(\exists x\;y\typ{\ty{u}}.\;\, x \not\teq y ) \mathrel\vee s \teq t$ to $F\!,$
 which states that either the cardinality of $\ty{u}$ is greater than one or
-$s$ must be equal to $t$.
+$s$ must be equal to $t.$
 
@@
 values that both correspond to the term
-$\MU s.\; \const{SCons}(0, s)$,
+$\MU s.\; \const{SCons}(0, s),$
 violating uniqueness.
@@
 The tree \const{\nummodden} is defined as $\const{\nummodden} =
-\const{Node}\;0\;\const{num}\;\const{\nummodden}$.
+\const{Node}\;0\;\const{num}\;\const{\nummodden}.$
 

### 2016-03-02 jour/jour.tex
minor changes
added URL

@@
 
-\newcommand\xend{{\hfill$\scriptstyle\blacksquare$}}
+\newcommand\xend{{\hfill\llap{$\scriptstyle\blacksquare$}}}
 \newcommand\xqed{{\hfill$\scriptstyle\squareforqed$}}
@@
 The $\mu$-notation is inspired by the $\mu$-calculus
-\cite[etc.]{kozen-1983,endrullis-et-al-2011}.
+\cite{kozen-1983,endrullis-et-al-2011}.
+% [etc.]
 
@@
 up to renaming of $\MU$-bound variables, we have
-\[\begin{array}{r@{}c@{}r@{}c@{}l@{\qquad}r@{}c@{}l@{\qquad}r@{}c@{}l}
-& &\interp{\J}{u} & {}\,=\,{} &  \const{C}( \MU {\Varec{z}}.\; \const{E}( {\Varec{z}} ) ) &
+\[\begin{array}{r@{}c@{}l@{\qquad}r@{}c@{}l@{\qquad}r@{}c@{}l}
+\interp{\J}{u} & {}\,=\,{} &  \const{C}( \MU {\Varec{z}}.\; \const{E}( {\Varec{z}} ) ) &
 \interp{\J}{w} & = &  \const{E}( \MU {\Varec{y}}.\; \const{D}( {\Varec{y}} ) ) &
 \interp{\J}{y} & {}\,=\,{} & \MU {\Varec{y}}.\; \const{D}( {\Varec{y}} ) \\
-\interp{\J}{v} & = & \interp{\J}{s} & = & \const{D}( \MU {\Varec{z}}.\; \const{E}( {\Varec{z}} ) ) &
+\interp{\J}{s} \,=\, \interp{\J}{v} & = & \const{D}( \MU {\Varec{z}}.\; \const{E}( {\Varec{z}} ) ) &
 \interp{\J}{x} & {}\,=\,{} & \const{C}( \const{D}( \MU {\Varec{z}}.\; \const{E}( {\Varec{z}} ) ) ) &
@@
 \end{array}\]
-It can be shown that $\J$ satisfies $\Ec =
-\{
-u \teq \const{C}( z ),\;
-v \teq \const{D}( z ),\;
-w \teq \const{E}( y ),\;
-x \teq \const{C}( v ),\;
-v \teq s,\;
-z \not\teq v
-\}.$
-\xend
+%It can be shown that
+It is easy to see that $\J$ satisfies the constraints
+
+\vskip\abovedisplayskip
+
+\noindent
+\hbox{}{\color{white}\xend}$u \teq \const{C}( z )$\hfill
+$v \teq \const{D}( z )$\hfill
+$w \teq \const{E}( y )$\hfill
+$x \teq \const{C}( v )$\hfill
+$v \teq\nobreak s$\hfill
+$z \not\teq v$\xend
 \end{examplex}
@@
 \begin{itemize}
-\item[\ ] For some $\ec{t} \in U,$ $\rn{Traverse}( \ec{t}, U, \emptyset ).$
+\item[2.1.] For some $\ec{t} \in U,$ $\rn{Traverse}( \ec{t}, U, \emptyset ).$
 \end{itemize}
@@
 \vspace*{-\topsep}
-\caption{\,Algorithm for checking \rn{Acyclic}}
+\caption{\,Algorithm for applying the \rn{Acyclic} rule}
 \label{fig:acyclic-algo}
@@
 \vspace*{-\topsep}
-\caption{\,Algorithm for checking \rn{Unique}}
+\caption{\,Algorithm for applying the \rn{Unique} rule}
 \label{fig:uniqueness-algo}
@@
 online.%
-\footnote{\url{http://lara.epfl.ch/~reynolds/JAR-CADE2015-cdt/}}
+\footnote{\url{http://lara.epfl.ch/~reynolds/JAR-CADE2015-cdt/} or
+\url{http://www21.in.tum.de/~blanchet/JAR-CADE2015-cdt/}}
 

### 2016-04-11 sister/sister.tex
updated ref
more

@@
 
-The original version of this paper was presented at the CADE-25 conference in
+The original version of this paper was presented at the 25th
+Conference on Automated Deduction (CADE-25) in
 Berlin, Germany \cite{reynolds-blanchette-2015-codata}. An more comprehensive
@@
 will appear in a special issue of the
-\emph{Journal of Automated Reasoning}. We refer to these for a description of
-related work.
+\emph{J.\ Autom.\ Reasoning}. % \cite{reynolds-blanchette-201x-codata3}.
+We refer to these for a description of related work.
 %\cite{reynolds-blanchette-201x-codata2}.
@@
 %satisfiable.
-We state the following properties of derivation trees generated by the calculus.
+We proved the following properties of derivation trees generated by the calculus
+\cite{reynolds-blanchette-2015-codata}.
 
@@
 rightly call it a decision procedure for that theory.
-The proof of solution soundness is constructive in that it provides a
-method for constructing a model for a saturated configuration.
+The proof of solution soundness %is constructive in that it
+provides a method for constructing a model for a saturated configuration.
 
@@
 for generating counterexamples in proof assistants
-\cite{reynolds-et-al-2015-smt}.
+\cite{reynolds-et-al-2016-smt}.
 %as an alternative to Nitpick;

### 2016-04-15 sister/sister.tex
Add reference to application paper.

@@
 % \cite{ganzinger-et-al-2004}
-have emerged as one of the most powerful ways to prove theorems in classical
-first-order logic. These solvers are also useful for constraint solving, or model finding:
+have emerged as useful tools in a number of applications~\cite{DeMoura2011SMT}.
+These solvers are one of the most powerful ways to prove theorems in classical
+first-order logic, and are also useful for constraint solving, or model finding:
 Given some constraints, they can be used to find an assignment, or model, that

### 2016-04-16 sister/sister.tex
Fixes for submission.

@@
 \usepackage{ijcai16}
+\usepackage{times}
 
 \usepackage[T1]{fontenc}
-\usepackage{flushend}
+%\usepackage{flushend}
 \usepackage{amsmath}
@@
 %%% For final version as well?
-\usepackage[
-   a4paper,
-   pdftex,
-   pdftitle={A Decision Procedure for (Co)datatypes in SMT Solvers},
-   pdfauthor={Andrew Reynolds and Jasmin Christian Blanchette},
-   pdfkeywords={},
-   pdfborder={0 0 0},
-   draft=false,
-   bookmarksnumbered,
-   bookmarks,
-   bookmarksdepth=2,
-   bookmarksopenlevel=2,
-   bookmarksopen]{hyperref}
+%\usepackage[
+%   a4paper,
+%   pdftex,
+%   pdftitle={A Decision Procedure for (Co)datatypes in SMT Solvers},
+%   pdfauthor={Andrew Reynolds, Jasmin Christian Blanchette},
+%   pdfkeywords={Decision Procedures, Automated Theorem Proving, Satisfiability Modulo Theories},
+%   pdfborder={0 0 0},
+%   draft=false,
+%   bookmarksnumbered,
+%   bookmarks,
+%   bookmarksdepth=2,
+%   bookmarksopenlevel=2,
+%   bookmarksopen]{hyperref}
+
+\pdfinfo{
+  /Title (A Decision Procedure for (Co)datatypes in SMT Solvers)
+  /Author (Andrew Reynolds, Jasmin Christian Blanchette)
+  /Keywords (Decision Procedures, Automated Theorem Proving, Satisfiability Modulo Theories)
+}
 
